### 2020-05-04 tour-vis.Rmd
init a distill article

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Sherry Zhang 
    url: https://github.com/huizezhang-sherry
    affiliation: Monash University
    affiliation_url: https://www.monash.edu/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

### 2020-05-04 tour-vis.Rmd
lit review and global structure

@@
 output: distill::distill_article
+bibliography: biblio.bib
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
+library(tidyverse)
+load(here::here("data/temp.rda"))
 ```
 
-Distill is a publication format for scientific and technical writing, native to the web.
+# Introduction
 
-Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
+## Tour
+
+Tour provides a way to explore multivariate data interactively via an established tour path.
+A tour path is formed by interpolating between randomly generated plane.
+Different types of tours are available depends on the purpose of the exploration i.e.
+a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data.
+a manual tour allows user to manually control the projection @cook1997manual.
+
+## Guided tour
+
+Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data.
+A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine.
+The projection pursuit index measures the "interestingness" of data defined as its departure from normality.
+Numerous indices have been proposed in the literature, including lengendre index @friedman1974projection, hermite index @hall1989polynomial, natural hermite index @cook1993projection, chi-square index @posse1995projection, LDA index @lee2005projection and PDA index @lee2010projection.
+An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index.
+The discussion of existing optimisation procedure will be discussed in the next section.
+
+Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path.
+An illustration from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.
+
+![](figures/tour-path.png)
+
+## Optimisation methods existing in tourr
+
+Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure.
+Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data.
+Simulated annealing, proposed by [@posse1995projection], is widely used to optimise the projection pursuit index.
+Geodesic searching method is an analogue to gradient acsent with the evaluation of derivative being replaced by randomly evaluation five direction in a tiny nearby neighbourhood and then find the projection basis along the geodesic direction that maximise the projection pursuit index.
+[need refine and more details here!]
+
+## problems experienced with existing versions
+
+# Recording the guided tour
+
+## Tour components
+
+Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) @wickham2011tourrpackage.
+A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.
+
+The psudo-code below illustrates the implementation of guided tour in the tourr package.
+Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
+This information is then used to compute a series of interpolating bases inside the `tour()` function.
+All the basis will be sent to create animation for visualising the tour in the `animate()` function.
+
+```{r eval = FALSE, echo = TRUE}
+animation <- function(){
+
+  # compute projection basis 
+  tour <- function(){
+
+    # construct bases on the tour path
+    new_geodesic_path <- function(){
+      tour_path <- function(){
+
+        # GENERATOR: generate projection basis via projection pursuit
+        guided_tour <- function(){
+          generator <- function(){
+
+            # define projection pursuit index
+            # generate the target basis from the current basis via optimisation
+          }
+        }
+
+        # prepare geodesic information needed for interpolating along the tour path
+      }
+    }
+
+    # INTERPOLATOR: interpolate between the current and target basis 
+    function(){
+      # generate interpolating bases on the geodesic path
+    }
+  }
+
+  # animate according to different display methods
+}
+```
+
+## Data Structure
+
+In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randonly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario [needa rephrase this part].
+
+The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process.
+The first five rows of the record of a five-variable 1d projection is printed below.
+
+```{r eval = FALSE}
+devtools::load_all()
+set.seed(123456)
+temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+                                           search_f =  search_geodesic),
+                  rescale = FALSE, verbose = TRUE)
+
+#save(temp, file = here::here("vignettes", "data", "temp.rda"))
+```
+
+```{r echo = TRUE}
+temp %>% head(5)
+```
+
+`tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in.
+This would include the `interpolation` stage and the detailed stage in the optimisation i.e.
+`direction_search`, `best_direction_search`, `line_search`and `best_line_search` for geodesic searching (`search_geodesic`); `random_search` and `new_basis` for simulating annealing (`search_better`).
+`loop` is the counter used for the optimisation procedure and thus will be `NA` for interpolation steps.
+`id` creates a sequential order of the basis.
+This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`.
+Additional messages during the optimisation can be displayed via `verbose = TRUE`.
+
+# Visual methods
+## index vs time plots
+## target basis vs interpolation steps, index vs time
+## low-d representation, PCA
+
+# Animating plots
+
+# Finding errors and developing improvements
+## interrupt
+## polish
+
+# Vis package

### 2020-05-04 tour-vis.Rmd
lit rev for existing optim method

@@
 Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data.
-Simulated annealing, proposed by [@posse1995projection], is widely used to optimise the projection pursuit index.
-Geodesic searching method is an analogue to gradient acsent with the evaluation of derivative being replaced by randomly evaluation five direction in a tiny nearby neighbourhood and then find the projection basis along the geodesic direction that maximise the projection pursuit index.
-[need refine and more details here!]
+Posse @posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis.
+The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$.
+The new basis is taken as the target basis if it has higher index value, or the sampling continues.
+If no basis is found to have higher index value after a certain humber of tries $n$, the radius $c$ is halved.
+The algorithm stops when the maximum numbere of iteration is attained or the radius $c$ is less than a pre-determined number.
+[Pursuit package uses this method and it works greate!
+But I don't think we implement this - although don't think it is too hard to do it].
+
+Cook @cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
+Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
+Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
+
+Simulated annealing @lee2005projection @klein1989experiments is a non-derivative procedure based on a cooling *temperature* parametere $T_i$.
+Given an initial $T_0$, the temperature at iteration $i$ is defined as $T_i = \frac{T_0}{log(i + 1)}$.
+The simulated annealing algorithm works as follows.
+Given a proportion $p$ and a randomly generated orthonormal basis $B'$, a candidate basis is constructed as $(1- p)B + p B'$ where $B$ is the current basis.
+If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
+If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(\frac{I(B) - I(B')}{T_i}), 1 \right)$ where $I(.)$ is the index function.
+
+    [The mathmetical expression is from Lee 2005 - without a negative in the $\exp$ but in the initial paper where SA is proposed 
+    @article{bertsimas1993simulated,
+      title={Simulated annealing},
+      author={Bertsimas, Dimitris and Tsitsiklis, John and others},
+      journal={Statistical science},
+      volume={8},
+      number={1},
+      pages={10--15},
+      year={1993},
+      publisher={Institute of Mathematical Statistics}
+    }
+    ,
+
+    @article{kirkpatrick1983optimization,
+      title={Optimization by simulated annealing},
+      author={Kirkpatrick, Scott and Gelatt, C Daniel and Vecchi, Mario P},
+      journal={science},
+      volume={220},
+      number={4598},
+      pages={671--680},
+      year={1983},
+      publisher={American association for the advancement of science}
+    }
+
+    and
+
+    @article{klein1989experiments,
+      title={Experiments in projection and clustering by simulated annealing},
+      author={Klein, Raymond W and Dubes, Richard C},
+      journal={Pattern Recognition},
+      volume={22},
+      number={2},
+      pages={213--220},
+      year={1989},
+      publisher={Elsevier}
+    }
+
+    The definition of $T_i$ is $T_i = \exp(-\frac{\Delta I}{k T})$ - with negative in the $\exp$. It is the gibb/boltzman distribution!]
 
 ## problems experienced with existing versions
+
+- the simulated annealing algorithm doesn't include the acceptance probability
+
+- is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995). Any tools available to  better understand its searching process?
+
+- it is not understood how the index value changes during interpolation - is any interpolating basis better candidate for the next-round current basis than the target basis?
+
+- How does the projected basis found by the optim look like? Are they close to each other/ far away? Any high-d or low-d vis of the projection basis?
+
+- Do I want to introduce search_polish in this article?
 
@@
 In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randonly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario [needa rephrase this part].
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[needa rephrase this part].**
 

### 2020-05-11 tour-vis.Rmd
more vis method and development

@@
 bibliography: biblio.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 library(tidyverse)
+library(ferrn)
+library(patchwork)
+
 load(here::here("data/temp.rda"))
+load(here::here("data/data_mult.rda"))
+load(here::here("data/holes_1d_geo.rda"))
+load(here::here("data/holes_2d_geo.rda"))
+load(here::here("data/holes_2d_better.rda"))
+
+load(here::here("data/holes_2d_geo_polish.rda"))
+load(here::here("data/holes_2d_better_polish.rda"))
+
+load(here::here("data/interrupt_yes.rda"))
+load(here::here("data/interrupt_no.rda"))
 ```
@@
 
-Simulated annealing @lee2005projection @klein1989experiments is a non-derivative procedure based on a cooling *temperature* parametere $T_i$.
-Given an initial $T_0$, the temperature at iteration $i$ is defined as $T_i = \frac{T_0}{log(i + 1)}$.
+Simulated annealing @bertsimas1993simulated @kirkpatrick1983optimization is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
+Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$.
 The simulated annealing algorithm works as follows.
-Given a proportion $p$ and a randomly generated orthonormal basis $B'$, a candidate basis is constructed as $(1- p)B + p B'$ where $B$ is the current basis.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(\frac{I(B) - I(B')}{T_i}), 1 \right)$ where $I(.)$ is the index function.
-
-    [The mathmetical expression is from Lee 2005 - without a negative in the $\exp$ but in the initial paper where SA is proposed 
-    @article{bertsimas1993simulated,
-      title={Simulated annealing},
-      author={Bertsimas, Dimitris and Tsitsiklis, John and others},
-      journal={Statistical science},
-      volume={8},
-      number={1},
-      pages={10--15},
-      year={1993},
-      publisher={Institute of Mathematical Statistics}
-    }
-    ,
-
-    @article{kirkpatrick1983optimization,
-      title={Optimization by simulated annealing},
-      author={Kirkpatrick, Scott and Gelatt, C Daniel and Vecchi, Mario P},
-      journal={science},
-      volume={220},
-      number={4598},
-      pages={671--680},
-      year={1983},
-      publisher={American association for the advancement of science}
-    }
-
-    and
-
-    @article{klein1989experiments,
-      title={Experiments in projection and clustering by simulated annealing},
-      author={Klein, Raymond W and Dubes, Richard C},
-      journal={Pattern Recognition},
-      volume={22},
-      number={2},
-      pages={213--220},
-      year={1989},
-      publisher={Elsevier}
-    }
-
-    The definition of $T_i$ is $T_i = \exp(-\frac{\Delta I}{k T})$ - with negative in the $\exp$. It is the gibb/boltzman distribution!]
+If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.
 
@@
 
+Two set of simulated data are used in the demonstration of the visualiation and diagnostics of the tour optimisation.
+A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`).
+`x2` is the informative variable simulated from two bi-modal normal distribution centered at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`.
+The data has been scaled to ensure `x2` has variance of 1.
+
+A larger dataset contains more informative variables (`x3` to `x7`) of different types.
+`x3` takes 500 positive one and 500 negative one.
+The distribution of all the variables except `x3` is plotted below.
+*[should I introduce the dist for each var?]*
+
+```{r}
+origin_dt_bi <- data_mult %>% 
+  dplyr::select(-x3) %>% 
+  gather(names, values) %>%
+  mutate(names = as_factor(names))
+
+origin_dt_bi %>%
+  ggplot(aes(x = values)) +
+  geom_histogram(binwidth = 0.3) +
+  geom_density(aes(y = 0.3 * ..count..)) +
+  facet_wrap(vars(names), ncol = 3)
+```
+
+Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure.
 The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process.
-The first five rows of the record of a five-variable 1d projection is printed below.
+The example below presents the global object of a 1D projection of the small dataset with geodesic seraching method.
 
 ```{r eval = FALSE}
-devtools::load_all()
 set.seed(123456)
@@
 Additional messages during the optimisation can be displayed via `verbose = TRUE`.
+Another examples is a 2D projection of the larger dataset with two informative variable (`x2` and `x7`) using search_better method.
+Notice in this example, the dimension of the bases becomes 6 by 2.
+
+```{r echo = TRUE}
+holes_2d_better %>% head(5)
+```
 
 # Visual methods
+
+The creation of the global object facilitates the diagnostics of the optimisation in the projection pursuit guided tourr.
+Different diagnostic plots could be made to understand the optimisation **[pretty cheesy here]**
+
+*The examples and results below still need to be modified as the development / modification of both ferrn and tourr package*.
+
 ## index vs time plots
+
+```{r echo = TRUE}
+holes_2d_geo %>% explore_trace_all()
+```
+
 ## target basis vs interpolation steps, index vs time
+
+```{r echo = TRUE}
+holes_2d_geo %>% explore_trace_interp()
+holes_2d_better %>% explore_trace_interp()
+
+interrupt_yes %>% explore_trace_interp()
+```
+
 ## low-d representation, PCA
+
+```{r echo = TRUE}
+holes_1d_geo %>% explore_proj_pca()
+```
 
@@
 
+```{r echo = TRUE}
+holes_1d_geo %>% explore_proj_pca(animate = TRUE)
+```
+
 # Finding errors and developing improvements
+
 ## interrupt
-## polish
+
+```{r}
+p1 <- interrupt_no %>% explore_trace_interp() + ggtitle("without interruption")
+
+p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption")
+
+p1 | p2
+```
+
+Further the case when the index function is not smooth, the interpolation may not attain -> modification
+
+## Polish
+
+In principle, all the optimisation routines should present the same output on the same problem.
+However, as we observed, the results from using `search_better` and `search_geodesic` on our 2D projection problem are different.
+Although both methods have detected `x2` as an informative variable on the second column, the weights for each variable on the first column vary, which causes difference on the index value.
+This motivates us to create another searching function that polishes the ending basis of the previous search and achieve unity on different methods.
+
+```{r echo = TRUE}
+holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+
+holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
+holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
+```
+
+**[Talk about how search_polish works]**
+
+The usage of search_polish is as follows.
+After the first tour, the final basis from the interpolation is extracted and supplied intoa new tour with the `start` argument and `search_polish` as teh searching function in the guided_tour.
+All the other arguments should remain the same.
+
+```{r eval = FALSE, echo = TRUE}
+set.seed(123456)
+holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],
+                           tour_path = guided_tour(holes(), d = 2, 
+                                                   search_f = tourr:::search_geodesic),
+                           rescale = FALSE, verbose = TRUE)
+
+last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+
+set.seed(123456)
+holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)],
+                                     tour_path = guided_tour(holes(), d = 2, 
+                                                             search_f = tourr:::search_polish),
+                                     rescale = FALSE, verbose = TRUE, start = last_basis)
+```
+
+**[This example is not ideal because currently the geodesic is not doing a good job in the first place]**
+
+```{r echo = TRUE}
+holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)
+holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)
+
+holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]
+holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]
+
+```
+
+## posse?
+
+search_better annd search_geodesic methods are not doing good on 2d projection -> see the simple case above: abs best is 0.96 while they only get to 0.88-ish.
+Would be worse for higher dimension.
+thus search_posse.
+get 0.95 and then polish to 0.96 :)
 

### 2020-05-11 vis_method.R

@@ -1,0 +1,81 @@
+library(tidyverse)
+library(tourr)
+
+set.seed(1234)
+x1 <- rnorm(1000, 0, 1)
+x2 <- c(rnorm(500, -3, 1), rnorm(500, 3, 1))
+x8 <- rnorm(1000, 0, 1)
+x9 <- rnorm(1000, 0, 1)
+x10 <- rnorm(1000, 0, 1)
+
+
+data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
+                       x9 = x9, x10 = x10) %>% purrr::map_df(scale) %>% as.matrix()
+########################
+
+set.seed(1234)
+x1 <- rnorm(1000, 0, 1)
+x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
+x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
+x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
+x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
+x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
+x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
+x8 <- rnorm(1000, 0, 1)
+x9 <- rnorm(1000, 0, 1)
+x10 <- rnorm(1000, 0, 1)
+
+data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
+                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
+                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
+
+########################
+set.seed(123456)
+holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],
+                           tour_path = guided_tour(holes(), d = 2, search_f = tourr:::search_geodesic),
+                           rescale = FALSE, verbose = TRUE)
+
+last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+
+set.seed(123456)
+holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)],
[…]

### 2020-05-12 tour-vis.Rmd
improvement

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, cache= TRUE)
 library(tidyverse)
@@
 library(patchwork)
+library(ggrepel)
 
@@
 load(here::here("data/holes_2d_better.rda"))
-
+load(here::here("data/holes_2d_better_random.rda"))
 load(here::here("data/holes_2d_geo_polish.rda"))
 load(here::here("data/holes_2d_better_polish.rda"))
+load(here::here("data/holes_2d_better_random_polish.rda"))
 
@@
 holes_2d_better %>% explore_trace_interp()
-
-interrupt_yes %>% explore_trace_interp()
 ```
@@
 
-## interrupt
+The visualisation methods introduced above allow us to assess the performance of each optimisation routine and thus provide improvments.
 
-```{r}
-p1 <- interrupt_no %>% explore_trace_interp() + ggtitle("without interruption")
+## Interrupt
 
-p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption")
+The left panel of Figure \@ref(fig:interruption) presents the trace plot of the interpolated basis using `search_better` to optimise the holes index (2D projection using the larger dataset).
+Tour interpolates from the current basis to the target basis, which has been found to have a higher index value by projection pursuit.
+The target basis will then be sent to the projection pursuit as the current basis to find the next target basis.
+We can observe from the plot that there are basis with even higher index value on the interpolation path.
+These bases could be used to search for new basis in the next round.
+Therefore, an interruption is constructed to only take interpolated basis up to the point where the index value stops increasing and the last basis is taken as the current basis for the next round of search.
+After implementing the interruption, the tracing plot with the same configuration is shown on the right panel.
+Rather than interpolate to the target basis, the interruption interrupt the process after id = 60 and proceeds the next round of search from the interpolated basis.
+Taking advantage of the interpolated basis results in a higher index value in the end with fewer steps.
+
+```{r interruption, fig.cap = "\\label{interruption} Trace plots of the interpolated basis with and without the interruption.
+The interruption stops the interpolation when the index value starts to decrease, see the difference at id being around 60.
+The implementation of the interuption finds an ending basis with higher index value using fewer steps.
+"}
+p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>% 
+  mutate(anno = c("current basis", "interpolated basis", "target basis"))
+
+p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") + 
+  geom_point(data = p1_anno) + 
+  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80)
+
+p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80)
 
@@
 In principle, all the optimisation routines should present the same output on the same problem.
-However, as we observed, the results from using `search_better` and `search_geodesic` on our 2D projection problem are different.
-Although both methods have detected `x2` as an informative variable on the second column, the weights for each variable on the first column vary, which causes difference on the index value.
-This motivates us to create another searching function that polishes the ending basis of the previous search and achieve unity on different methods.
+In Figure \@ref(fig:trace-compare), red dots shows the trace of the interpolated basis using three different methods on the 2D projection problem.
+We can observe that different optimisation methods attains different ending index value.
+This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
 
-```{r echo = TRUE}
-holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
-holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+`search_polish` takes the ending basis of a given search as the starting basis and uses a brutal-force approach to sample a larger number of basis (`n_sample`) in the neightbourhood, controlled by `polish_alpha`.
+Among the `n_sample` basis, the one with the largest index value becomes the candidate.
+If its index value is larger than that of the starting basis, it becomes the center of the searching neightbourhood in the next round.
+If no basis is found to have larger index value than the starting basis, the searching neighbourhood will shrink and the search continues.
+The polishing search ends when one of the four stopping criteria is satisfied:
 
-holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
-holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) 
-```
-
-**[Talk about how search_polish works]**
+1) the two basis can't be too close 
+2) the improvement can't be too small 
+3) the searching neighbourhood can't be too small
+4) the number of iteration can't exceed the max.tries
 
 The usage of search_polish is as follows.
-After the first tour, the final basis from the interpolation is extracted and supplied intoa new tour with the `start` argument and `search_polish` as teh searching function in the guided_tour.
+After the first tour, the final basis from the interpolation is extracted and supplied into a new tour with the `start` argument and `search_polish` as the searching function in the guided_tour.
 All the other arguments should remain the same.
@@
 set.seed(123456)
-holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],
-                           tour_path = guided_tour(holes(), d = 2, 
+holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],tour_path = 
+                             guided_tour(holes(), d = 2, 
                                          search_f = tourr:::search_geodesic),
@@
 
-last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% 
+  tail(1) %>% pull(basis) %>% .[[1]]
 
 set.seed(123456)
-holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)],
-                                     tour_path = guided_tour(holes(), d = 2, 
+holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)], tour_path = 
+                                    guided_tour(holes(), d = 2, 
                                                 search_f = tourr:::search_polish),
-                                     rescale = FALSE, verbose = TRUE, start = last_basis)
+                                  rescale = FALSE, verbose = TRUE, 
+                                  start = last_basis)
 ```
 
-**[This example is not ideal because currently the geodesic is not doing a good job in the first place]**
+After polishing, all three methods attain the same index value as shown in Figure \@ref(fig:trace-compare).
+This allows us to evaluate the performance of each searching method.
+Search_better_random finds an ending basis with a decent high index value while search_geodesic and search_better don't and make search polish do a large proportion of the work.
 
-```{r echo = TRUE}
-holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)
-holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(index_val)
+```{r trace-compare, fig.cap="Breakdown of index value when using different searching methods in conjunction with search polish.
+All three methods attain the same index value after the polishing.
+Search better random finds a decent ending basis while search better and search geodesic don't and search polish has done a large proportion of polishing."}
+p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("search_geodesic")
 
-holes_2d_better_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]
-holes_2d_geo_polish %>% filter(info == "polish_best") %>% tail(1) %>% pull(basis) %>% .[[1]]
+p2 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("search_better")
 
+p3 <- bind_rows(holes_2d_better_random, holes_2d_better_random_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("search_better_random")
+
+(p1 | p2 | p3) &
+  theme(legend.position = "none") 
 ```

### 2020-05-13 tour-vis.Rmd
modify polish; new tour path plot; new data for geo search

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache= TRUE)
+knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
 library(tidyverse)
@@
 load(here::here("data/holes_2d_better.rda"))
+load(here::here("data/holes_2d_better_max_tries.rda"))
 load(here::here("data/holes_2d_better_random.rda"))
@@
 load(here::here("data/holes_2d_better_polish.rda"))
+load(here::here("data/holes_2d_better_max_tries_polish.rda"))
 load(here::here("data/holes_2d_better_random_polish.rda"))
-
 load(here::here("data/interrupt_yes.rda"))
@@
 Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path.
-An illustration from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.
+An illustration modified from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.
 
-![](figures/tour-path.png)
+![](figures/tour_path.svg)
 
@@
 set.seed(123456)
-temp <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+holes_1d_geo <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
                                            search_f =  search_geodesic),
@@
 ```{r echo = TRUE}
-temp %>% head(5)
+holes_1d_geo %>% head(5)
 ```
@@
 
+The tracing plot is useful to explore the index value. Given a global object created, `explore_trace_all()` plots the index value of all the bases.
+
 ```{r echo = TRUE}
-holes_2d_geo %>% explore_trace_all()
+holes_2d_geo %>% explore_trace_all(magnify = TRUE)
+holes_2d_geo %>% explore_trace_interp()
 ```
@@
 ## target basis vs interpolation steps, index vs time
-
-```{r echo = TRUE}
-holes_2d_geo %>% explore_trace_interp()
-holes_2d_better %>% explore_trace_interp()
-```
 
@@
 In principle, all the optimisation routines should present the same output on the same problem.
-In Figure \@ref(fig:trace-compare), red dots shows the trace of the interpolated basis using three different methods on the 2D projection problem.
-We can observe that different optimisation methods attains different ending index value.
+In Figure \@ref(fig:trace-compare), red dots shows the trace of the interpolated basis using `search_geodesic` and `search_better` (with max.tries = 100), respectively, on the 2D projection problem.
+We can observe that they attain slightly different ending index value, which is not ideal.
 This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
+
+```{r}
+p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("search_geodesic")
+
+p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("search_better")
+
+(p1 | p2) &
+  theme(legend.position = "none") 
+```
 
@@
 
-After polishing, all three methods attain the same index value as shown in Figure \@ref(fig:trace-compare).
-This allows us to evaluate the performance of each searching method.
-Search_better_random finds an ending basis with a decent high index value while search_geodesic and search_better don't and make search polish do a large proportion of the work.
+The following example conducted a 2D projection on the larger dataset using search better with different configurations.
+`max.tries` is a hyperparameter that controls the maximum number of try without improvement and its default value is 25.
+As shown in Figure \@ref(fig:trace-compare), after polishing, both trials attain the same index value.
+However, a small `max.tries` of 25 is not sufficient for the algorithm to find the true maximum.
+This is because 25 tries is not sufficient for the 2D searching space.
 
-```{r trace-compare, fig.cap="Breakdown of index value when using different searching methods in conjunction with search polish.
-All three methods attain the same index value after the polishing.
-Search better random finds a decent ending basis while search better and search geodesic don't and search polish has done a large proportion of polishing."}
-p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("search_geodesic")
+```{r trace-compare, fig.cap="Breakdown of index value when using different max.tries in search better in conjunction with search polish.
+Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}
 
-p2 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("search_better")
+p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("max.tries = 25")
 
-p3 <- bind_rows(holes_2d_better_random, holes_2d_better_random_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("search_better_random")
+p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
+  explore_trace_interp(point = TRUE) + ggtitle("max.tries = 500")
 
-(p1 | p2 | p3) &
+(p1 | p2) &
   theme(legend.position = "none") 

### 2020-05-13 source/cheatsheet.Rmd
supplementary materials

---
title: "Performance of different optimisation routines on projection pursuit index"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ferrn)
library(patchwork)

load(here::here("source/test/holes_1d_geo.rda"))
load(here::here("source/test/holes_1d_better.rda"))
load(here::here("source/test/holes_2d_geo.rda"))
load(here::here("source/test/holes_2d_geo_polish.rda"))
load(here::here("source/test/holes_2d_better.rda"))
load(here::here("source/test/holes_2d_better_polish.rda"))
load(here::here("source/test/holes_2d_better_max_tries.rda"))
load(here::here("source/test/holes_2d_better_max_tries_polish.rda"))
load(here::here("source/test/holes_2d_better_random.rda"))
load(here::here("source/test/holes_2d_better_random_polish.rda"))
load(here::here("source/test/holes_2d_pos.rda"))
load(here::here("source/test/holes_2d_pos_polish.rda"))

```

# Holes Index

## 1D projection

```{r}
p1 <- holes_1d_geo %>% explore_trace_interp()
p2 <- holes_1d_better %>% explore_trace_interp()
(p1 | p2 ) & ylim(0.74, 0.95) 
```

## 2D projection

```{r}
p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("geodesic")
p2 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("better, max.tries = 25")
p3 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>%
  explore_trace_interp(point = TRUE) + ggtitle("better, max.tries = 500")
p4 <- bind_rows(holes_2d_better_random, holes_2d_better_random_polish) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("better_random")
p5 <- bind_rows(holes_2d_pos, holes_2d_pos_polish) %>% 
  mutate(method = fct_relevel(method, c("search_posse", "search_polish"))) %>% 
  explore_trace_interp(point = TRUE) + ggtitle("posse")

((p1 | p4 | p5 ) / (p2 | p3 )) & 
  theme(legend.position = "none") & 
  ylim(0.8, 0.99)

```

### 2020-05-13 source/my_test_file.R

@@ -1,0 +1,122 @@
+library(tidyverse)
+library(tourr)
+set.seed(1234)
+x1 <- rnorm(1000, 0, 1)
+x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
+x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
+x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
+x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
+x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
+x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
+x8 <- rnorm(1000, 0, 1)
+x9 <- rnorm(1000, 0, 1)
+x10 <- rnorm(1000, 0, 1)
+
+data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
+                       x9 = x9, x10 = x10) %>% purrr::map_df(scale)
+
+data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
+                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
+                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
+
+###############################
+devtools::load_all()
+set.seed(123456)
+holes_1d_geo <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+                                                   search_f =  tourr:::search_geodesic),
+                     rescale = FALSE, verbose = TRUE)
+
+devtools::load_all()
+set.seed(123456)
+holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+                                                           search_f =  tourr:::search_better),
+                             rescale = FALSE, verbose = TRUE)
+
+############
+devtools::load_all()
+set.seed(123456)
+holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)], tour_path = guided_tour(holes(), d = 2,
+                                                                             search_f =  tourr:::search_geodesic),
+                           rescale = FALSE, verbose = TRUE)
+
+last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
+
[…]

### 2020-05-17 paper/paper.Rmd
JCGS format and add some general optim difficulties

--- 2020-05-13 tour-vis.Rmd
@@
 ---
-title: "Untitled"
-description: |
-  A new article created using the Distill format.
-author:
-  - name: Sherry Zhang 
-    url: https://github.com/huizezhang-sherry
-    affiliation: Monash University
-    affiliation_url: https://www.monash.edu/
-date: "`r Sys.Date()`"
-output: distill::distill_article
+title: Title here
+
+# to produce blinded version set to 1
+blinded: 0
+
+authors: 
+- name: Author 1
+  thanks: The authors gratefully acknowledge ...
+  affiliation: Department of YYY, University of XXX
+
+- name: Author 2
+  affiliation: Department of ZZZ, University of WWW
+
+keywords:
+- 3 to 6 keywords
+- that do not appear in the title
+
+abstract: |
+  1. check consistency of using PP for projection pursuit and PPI for projection pursuit index
+
 bibliography: biblio.bib
-editor_options: 
-  chunk_output_type: console
+output: rticles::asa_article
 ---
@@
 knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+
 library(tidyverse)
@@
 load(here::here("data/interrupt_no.rda"))
+
 ```
@@
 a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data.
-a manual tour allows user to manually control the projection @cook1997manual.
+a manual tour allows user to manually control the projection [@cook1997manual].
 
@@
 The projection pursuit index measures the "interestingness" of data defined as its departure from normality.
-Numerous indices have been proposed in the literature, including lengendre index @friedman1974projection, hermite index @hall1989polynomial, natural hermite index @cook1993projection, chi-square index @posse1995projection, LDA index @lee2005projection and PDA index @lee2010projection.
+Numerous indices have been proposed in the literature, including lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index.
@@
 Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path.
-An illustration modified from @buja2005computational entails the geodesic interpolation bewteen the plane generated by projection pursuit.
+An illustration modified from [@buja2005computational] entails the geodesic interpolation bewteen the plane generated by projection pursuit.
 
-![](figures/tour_path.svg)
+```{r}
+knitr::include_graphics(path = here::here("figures/tour_path_new.png"))
+```
 
-## Optimisation methods existing in tourr
+## Optimisation methods in projection pursuit literature
 
-Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure.
-Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data.
-Posse @posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis.
+As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the com- putational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe.
+An ideal optimisation procedure needs to have following characteristics:
+
+- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentible
+
+- *Being able to optimise with constraints*: the projection matrix is restricted to an orthornormal matrix.
+
+- *Being able to reveal both local and global maximum*: Although the primary interest is to find the global maximum, distinct local structures are also of our interest.
+
+Below present three existing methods in tour.
+
+@posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis.
 The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$.
@@
 
-Cook @cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
+@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
 Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
@@
 
-Simulated annealing @bertsimas1993simulated @kirkpatrick1983optimization is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
 Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$.
@@
 
-## problems experienced with existing versions
+<!-- Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure.
+Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data.
+-->
 
-- the simulated annealing algorithm doesn't include the acceptance probability
+## problems and difficulties in PP optimisation
 
-- is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995). Any tools available to  better understand its searching process?
+Below listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific for PP optimisation.
 
-- it is not understood how the index value changes during interpolation - is any interpolating basis better candidate for the next-round current basis than the target basis?
+- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
+Also, there are discussions on how to avoid getting trapped in a local optimal in the literature.
 
-- How does the projected basis found by the optim look like? Are they close to each other/ far away? Any high-d or low-d vis of the projection basis?
+- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
+Stochastic optimisation method could be an alternative to solve these problems.
 
-- Do I want to introduce search_polish in this article?
+- *computation speed*: The optimisation procedure needs to be fast to compute since tours produces real-time animation of the projected data.
+
+- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm producing the same result for every run.
+This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
+
+- *high-dimensional decision variable*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
+Researcher would be better off if they can understand the relative position of different projection matrix in the high-dimensional space.
+
+- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximises the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
+It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
+
+<!-- the interpolation can be seen as giving information on the direction of the derivative. -->
+
+<!-- - is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995).
+Any tools available to  better understand its searching process?
+-->
+
+*Think about how does your pacakge help people to understand optimisation*
+
+ - diagnostic on stochastic optim
+ - vis the prograssion of multi-parameter decision variable 
+ - understanding learning rate - neightbourhood parameter
+ - understand where the local & global maximum is found - trace plot - see if noisy function
 
@@
 
-Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) @wickham2011tourrpackage.
+Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.
@@
 
+## add something about general vis
+
 The creation of the global object facilitates the diagnostics of the optimisation in the projection pursuit guided tourr.
@@
 
-```{r echo = TRUE}
+```{r echo = TRUE, eval = FALSE}
 holes_1d_geo %>% explore_proj_pca(animate = TRUE)

### 2020-05-17 paper/paper.tex
JCGS format and add some general optim difficulties

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Title here}

  \author{
        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of YYY, University of XXX\\
     and \\     Author 2 \\
    Department of ZZZ, University of WWW\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Title here}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  check consistency of using PP for projection pursuit and PPI for
  projection pursuit index
\end{enumerate}
\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{tour}{%
\subsection{Tour}\label{tour}}

Tour provides a way to explore multivariate data interactively via an
established tour path. A tour path is formed by interpolating between
randomly generated plane. Different types of tours are available depends
on the purpose of the exploration i.e.~a grand tour is suitable for
randomly exploring the data from different angles; a guided tour detects
a particular structure in the data. a manual tour allows user to
manually control the projection \citep{cook1997manual}.

\hypertarget{guided-tour}{%
\subsection{Guided tour}\label{guided-tour}}

Guided tour is usually used in conjunction projection pursuit, a method
coined by \citet{friedman1974projection} to detect ``interesting''
low-diemsion projection of multivariate data. A projection pursuit
requires the definition of a projection pursuit index function and an
optimisation routine. The projection pursuit index measures the
``interestingness'' of data defined as its departure from normality.
Numerous indices have been proposed in the literature, including
lengendre index \citep{friedman1974projection}, hermite index
\citep{hall1989polynomial}, natural hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}. An optimisation routine is required to
find the projection basis (thus projection) that maximises the
projection pursuit index. The discussion of existing optimisation
procedure will be discussed in the next section.

Guided tour creates visualisation for the projections found by
projection pursuit by constructing a tour path. An illustration modified
from \citep{buja2005computational} entails the geodesic interpolation
bewteen the plane generated by projection pursuit.

\includegraphics[width=7.75in]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_new}

\hypertarget{optimisation-methods-in-projection-pursuit-literature}{%
\subsection{Optimisation methods in projection pursuit
literature}\label{optimisation-methods-in-projection-pursuit-literature}}

As \citet{friedman1974projection} said ``\ldots{}, the technique use for
maximizing the (one- and two-dimensional) projection index strongly
influences both the statistical and the com- putational aspects of the
procedure.'' The quality of the optimisation procedure largely affect
the tour view and thus, the interesting projection one could possibly
observe. An ideal optimisation procedure needs to have following
characteristics:

\begin{itemize}
\item
  \emph{Being able to handle non-differentiable index function}: the
  index function could be noisy and non-differentible
\item
  \emph{Being able to optimise with constraints}: the projection matrix
  is restricted to an orthornormal matrix.
\item
  \emph{Being able to reveal both local and global maximum}: Although
  the primary interest is to find the global maximum, distinct local
  structures are also of our interest.
\end{itemize}

Below present three existing methods in tour.

\citet{posse1995projection} presented a random search algorithm that
samples new basis in the neighbourhood of the current basis. The
neighbourhood is defined via the radius of the p-dimenionsal sphere,
\(c\). The new basis is taken as the target basis if it has higher index
value, or the sampling continues. If no basis is found to have higher
index value after a certain humber of tries \(n\), the radius \(c\) is
halved. The algorithm stops when the maximum numbere of iteration is
attained or the radius \(c\) is less than a pre-determined number.
{[}Pursuit package uses this method and it works greate! But I don't
think we implement this - although don't think it is too hard to do
it{]}.

\citet{cook1995grand} explained the use of a gradient ascent
optimisation with the assumption that the index function is continuous
and differentiable. Since some indices could be non-differentiable, the
computation of derivative is replaced by a psudo-derivative of
evaluating five randomly generated directions in a tiny nearby
neighbourhood. Taking a step on the straight derivative direction has
been modified to maximise the projection pursuit index along the
geodesic direction.

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(B\), a candidate basis is
constructed as \(B_j = (1- \alpha)B_i + \alpha B\) where \(B_i\) is the
current basis. If the index value of the candidate basis is larger than
the one of the current basis, the candidate basis becomes the target
basis. If it is smaller, the candidate is accepted with probability
\(A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)\) where
\(I(.)\) is the index function.

\hypertarget{problems-and-difficulties-in-pp-optimisation}{%
\subsection{problems and difficulties in PP
optimisation}\label{problems-and-difficulties-in-pp-optimisation}}

Below listed several issues in projection pursuit optimisation. Some of
them are general problems in optimisation literature, while others are
more specific for PP optimisation.

\begin{itemize}
\item
  \emph{Finding global maximum}: Although finding local maximum is
  relatively easy with developed algorithms, it is generally hard to
  guarantee global maximum in a problem where the objective function is
  complex or the number of decision variables is large. Also, there are
  discussions on how to avoid getting trapped in a local optimal in the
  literature.
\item
  \emph{optimising on non-smooth function}: When the objective function
  is non-differentiable, derivative information can not be obtained.
  This means traditional gradient- or Hessian- based methods are not
  feasible, stochastic optimisation method could be an alternative to
  solve these problems.
\item
  \emph{computation speed}: The optimisation procedure needs to be fast
  to compute because tours produces real-time animation of the projected
  data.
\item
  \emph{consistency result in stochastic optimisation}: In stochastic
  algorithm, researchers usually set a seed to ensure the algorithm
  producing the same result for every run. While this practice supports
  reproducibility, less efforts has been made to guarantee different
  seeds will provide the same result.
\item
  \emph{high-dimensional decision variable}: In projection pursuit, the
  decision variable is the entry in the projection matrix, which is
  usually high-dimensional. Researcher would be better off if they can
  understand the relative position of different projection matrix in the
  high-dimensional space.
\item
  \emph{role of interpolation in PP optimisation}: An optimisation
  procedure usually involves iteratively finding projection bases that
  maximise the index function, while tour requires geodesic
  interpolation between these bases to produce a continuous view for the
  users. It would be interesting to see if the interpolated bases could,
  in reverse, help the optimisation reach faster convergence.
\end{itemize}

\emph{Think about how does your pacakge help people to understand
optimisation}

\begin{itemize}
\tightlist
\item
  diagnostic on stochastic optim
\item
  vis the prograssion of multi-parameter decision variable
\item
  understanding learning rate - neightbourhood parameter
\item
  understand where the local \& global maximum is found - trace plot -
  see if noisy function
\end{itemize}

\hypertarget{recording-the-guided-tour}{%
\section{Recording the guided tour}\label{recording-the-guided-tour}}

\hypertarget{tour-components}{%
\subsection{Tour components}\label{tour-components}}

Guided tour, along with other types of tour, has been implemented in the
\emph{tourr} package in R, available on the Comprehensive R Archive
Network at \url{https://cran.r-project.org/web/packages/tourr/}
\citep{wickham2011tourrpackage}. A tour includes two major components: a
\emph{generator} that generating the projection basis according to
projection pursuit and an \emph{interpolator} that performing geodesic
interpolation between the projection basis.

The psudo-code below illustrates the implementation of guided tour in
the tourr package. Given an projection pursuit index function and a
randomly generated projection basis (current basis), the optimisation
procedure produces a target basis inside \texttt{generator()} Both the
current basis and the target basis will be supplied to
\texttt{tour\_path()} to prepare information needed for constructing a
geodesic path. This information is then used to compute a series of
interpolating bases inside the \texttt{tour()} function. All the basis
will be sent to create animation for visualising the tour in the
\texttt{animate()} function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{animation <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}

  \CommentTok{# compute projection basis }
\NormalTok{  tour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}

    \CommentTok{# construct bases on the tour path}
\NormalTok{    new_geodesic_path <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
\NormalTok{      tour_path <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}

        \CommentTok{# GENERATOR: generate projection basis via projection pursuit}
\NormalTok{        guided_tour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
\NormalTok{          generator <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}

            \CommentTok{# define projection pursuit index}
            \CommentTok{# generate the target basis from the current basis via optimisation}
\NormalTok{          \}}
\NormalTok{        \}}

        \CommentTok{# prepare geodesic information needed for interpolating along the tour path}
\NormalTok{      \}}
\NormalTok{    \}}

    \CommentTok{# INTERPOLATOR: interpolate between the current and target basis }
    \ControlFlowTok{function}\NormalTok{()\{}
      \CommentTok{# generate interpolating bases on the geodesic path}
\NormalTok{    \}}
\NormalTok{  \}}

  \CommentTok{# animate according to different display methods}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{data-structure}{%
\subsection{Data Structure}\label{data-structure}}

In the current implementation of the \texttt{tourr} package, while the
target basis generated by the projection pursuit can be accessed later
via \texttt{save\_history()}, interpolating bases and those randonly
nearby bases generated in the optimisation are not stored. This creates
difficulties for fully understand the behaviour of the optimisation and
interpolation of tour in complex scenario \textbf{{[}needa rephrase this
part{]}.}

Two set of simulated data are used in the demonstration of the
visualiation and diagnostics of the tour optimisation. A small dataset
consists of 1000 randomly simulated observations of five variables
(\texttt{x1}, \texttt{x2}, \texttt{x8}, \texttt{x9}, \texttt{x10}).
\texttt{x2} is the informative variable simulated from two bi-modal
normal distribution centered at -3 and 3 with variance being 1 and the
other four are simulated from \texttt{N(0,\ 1)}. The data has been
scaled to ensure \texttt{x2} has variance of 1.

A larger dataset contains more informative variables (\texttt{x3} to
\texttt{x7}) of different types. \texttt{x3} takes 500 positive one and
500 negative one. The distribution of all the variables except
\texttt{x3} is plotted below. \emph{{[}should I introduce the dist for
each var?{]}}

\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}

Once the dataset is sent to the tourr package, all the information
generated will be stored in a global structure. The global structure
consists of six columnes: \texttt{basis}, \texttt{index\_val},
\texttt{tries}, \texttt{info}, \texttt{loop}, \texttt{id} and captured
all the basis generated during whole tour process. The example below
presents the global object of a 1D projection of the small dataset with
geodesic seraching method.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 5 x 7
##   basis             index_val tries info              loop method             id
##   <list>                <dbl> <dbl> <chr>            <dbl> <chr>           <int>
## 1 <dbl[,1] [5 x 1]>     0.749     1 start               NA <NA>                1
## 2 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     2
## 3 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     3
## 4 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     4
## 5 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     5
\end{verbatim}

\texttt{tries} has an increment of one once the generator is called
(equivalently a new target basis is generated); \texttt{info} records
the stage the basis is in. This would include the \texttt{interpolation}
stage and the detailed stage in the optimisation i.e.
\texttt{direction\_search}, \texttt{best\_direction\_search},
\texttt{line\_search}and \texttt{best\_line\_search} for geodesic
searching (\texttt{search\_geodesic}); \texttt{random\_search} and
\texttt{new\_basis} for simulating annealing (\texttt{search\_better}).
\texttt{loop} is the counter used for the optimisation procedure and
thus will be \texttt{NA} for interpolation steps. \texttt{id} creates a
sequential order of the basis. This information will be stored and
printed when the optimisation ends and can be turned off via
\texttt{print\ =\ FALSE}. Additional messages during the optimisation
can be displayed via \texttt{verbose\ =\ TRUE}. Another examples is a 2D
projection of the larger dataset with two informative variable
(\texttt{x2} and \texttt{x7}) using search\_better method. Notice in
this example, the dimension of the bases becomes 6 by 2.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_2d_better }\OperatorTok{%>%}\StringTok{ }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 5 x 7
##   basis             index_val tries info           loop method           id
##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>         <int>
## 1 <dbl[,2] [6 x 2]>     0.804     1 start            NA <NA>              1
## 2 <dbl[,2] [6 x 2]>     0.793     1 random_search     1 search_better     2
## 3 <dbl[,2] [6 x 2]>     0.784     1 random_search     2 search_better     3
## 4 <dbl[,2] [6 x 2]>     0.773     1 random_search     3 search_better     4
## 5 <dbl[,2] [6 x 2]>     0.795     1 random_search     4 search_better     5
\end{verbatim}

\hypertarget{visual-methods}{%
\section{Visual methods}\label{visual-methods}}

\hypertarget{add-something-about-general-vis}{%
\subsection{add something about general
vis}\label{add-something-about-general-vis}}

The creation of the global object facilitates the diagnostics of the
optimisation in the projection pursuit guided tourr. Different
diagnostic plots could be made to understand the optimisation
\textbf{{[}pretty cheesy here{]}}

\emph{The examples and results below still need to be modified as the
development / modification of both ferrn and tourr package}.

\hypertarget{index-vs-time-plots}{%
\subsection{index vs time plots}\label{index-vs-time-plots}}

The tracing plot is useful to explore the index value. Given a global
object created, \texttt{explore\_trace\_all()} plots the index value of
all the bases.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_trace_all}\NormalTok{(}\DataTypeTok{magnify =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_trace_interp}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-2.pdf}

\hypertarget{target-basis-vs-interpolation-steps-index-vs-time}{%
\subsection{target basis vs interpolation steps, index vs
time}\label{target-basis-vs-interpolation-steps-index-vs-time}}

\hypertarget{low-d-representation-pca}{%
\subsection{low-d representation, PCA}\label{low-d-representation-pca}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: The `x` argument of `as_tibble.matrix()` must have column names if `.name_repair` is omitted as of tibble 2.0.0.
## Using compatibility `.name_repair`.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_warnings()` to see where this warning was generated.
\end{verbatim}

\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}

\hypertarget{animating-plots}{%
\section{Animating plots}\label{animating-plots}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{finding-errors-and-developing-improvements}{%
\section{Finding errors and developing
improvements}\label{finding-errors-and-developing-improvements}}

The visualisation methods introduced above allow us to assess the
performance of each optimisation routine and thus provide improvments.

\hypertarget{interrupt}{%
\subsection{Interrupt}\label{interrupt}}

The left panel of Figure @ref(fig:interruption) presents the trace plot
of the interpolated basis using \texttt{search\_better} to optimise the
holes index (2D projection using the larger dataset). Tour interpolates
from the current basis to the target basis, which has been found to have
a higher index value by projection pursuit. The target basis will then
be sent to the projection pursuit as the current basis to find the next
target basis. We can observe from the plot that there are basis with
even higher index value on the interpolation path. These bases could be
used to search for new basis in the next round. Therefore, an
interruption is constructed to only take interpolated basis up to the
point where the index value stops increasing and the last basis is taken
as the current basis for the next round of search. After implementing
the interruption, the tracing plot with the same configuration is shown
on the right panel. Rather than interpolate to the target basis, the
interruption interrupt the process after id = 60 and proceeds the next
round of search from the interpolated basis. Taking advantage of the
interpolated basis results in a higher index value in the end with fewer
steps.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/interruption-1.pdf}
\caption{\label{interruption} Trace plots of the interpolated basis with
and without the interruption. The interruption stops the interpolation
when the index value starts to decrease, see the difference at id being
around 60. The implementation of the interuption finds an ending basis
with higher index value using fewer steps.}
\end{figure}

Further the case when the index function is not smooth, the
interpolation may not attain -\textgreater{} modification

\hypertarget{polish}{%
\subsection{Polish}\label{polish}}

In principle, all the optimisation routines should present the same
output on the same problem. In Figure @ref(fig:trace-compare), red dots
shows the trace of the interpolated basis using
\texttt{search\_geodesic} and \texttt{search\_better} (with max.tries =
100), respectively, on the 2D projection problem. We can observe that
they attain slightly different ending index value, which is not ideal.
This motivates the creation of a polishing search that polishes the
ending basis and achieves unity on different methods.

\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}

\texttt{search\_polish} takes the ending basis of a given search as the
starting basis and uses a brutal-force approach to sample a larger
number of basis (\texttt{n\_sample}) in the neightbourhood, controlled
by \texttt{polish\_alpha}. Among the \texttt{n\_sample} basis, the one
with the largest index value becomes the candidate. If its index value
is larger than that of the starting basis, it becomes the center of the
searching neightbourhood in the next round. If no basis is found to have
larger index value than the starting basis, the searching neighbourhood
will shrink and the search continues. The polishing search ends when one
of the four stopping criteria is satisfied:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the two basis can't be too close
\item
  the improvement can't be too small
\item
  the searching neighbourhood can't be too small
\item
  the number of iteration can't exceed the max.tries
\end{enumerate}

The usage of search\_polish is as follows. After the first tour, the
final basis from the interpolation is extracted and supplied into a new
tour with the \texttt{start} argument and \texttt{search\_polish} as the
searching function in the guided\_tour. All the other arguments should
remain the same.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)],}\DataTypeTok{tour_path =} 
                             \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                         \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_geodesic),}
                           \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}

\NormalTok{last_basis <-}\StringTok{ }\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(info }\OperatorTok{==}\StringTok{ "interpolation"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{tail}\NormalTok{(}\DecValTok{1}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}

\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
\NormalTok{holes_2d_geo_polish <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)], }\DataTypeTok{tour_path =} 
                                    \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
                                                \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_polish),}
                                  \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }
                                  \DataTypeTok{start =}\NormalTok{ last_basis)}
\end{Highlighting}
\end{Shaded}

The following example conducted a 2D projection on the larger dataset
using search better with different configurations. \texttt{max.tries} is
a hyperparameter that controls the maximum number of try without
improvement and its default value is 25. As shown in Figure
@ref(fig:trace-compare), after polishing, both trials attain the same
index value. However, a small \texttt{max.tries} of 25 is not sufficient
for the algorithm to find the true maximum. This is because 25 tries is
not sufficient for the 2D searching space.

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/trace-compare-1.pdf}
\caption{Breakdown of index value when using different max.tries in
search better in conjunction with search polish. Both attain the same
final index value after the polishing while using a max.tries 25 is not
sufficient to find the ture maximum.}
\end{figure}

\hypertarget{posse}{%
\subsection{posse?}\label{posse}}

search\_better annd search\_geodesic methods are not doing good on 2d
projection -\textgreater{} see the simple case above: abs best is 0.96
while they only get to 0.88-ish. Would be worse for higher dimension.
thus search\_posse. get 0.95 and then polish to 0.96 :)

\bibliographystyle{agsm}
\bibliography{biblio.bib}

\end{document}

### 2020-05-19 paper/paper.Rmd
write GOG; explore post-run diagnostics using vis

@@
 title: Title here
+blinded: 0
+authors:
 
-# to produce blinded version set to 1
-blinded: 0
-
-authors: 
 - name: Author 1
@@
   affiliation: Department of YYY, University of XXX
-
 - name: Author 2
   affiliation: Department of ZZZ, University of WWW
-
 keywords:
@@
 - that do not appear in the title
-
 abstract: |
   1. check consistency of using PP for projection pursuit and PPI for projection pursuit index 
-
 bibliography: biblio.bib
 output: rticles::asa_article
+editor_options: 
+  chunk_output_type: console
 ---
@@
 load(here::here("data/holes_2d_better_random_polish.rda"))
+load(here::here("data/holes_2d_pos.rda"))
 load(here::here("data/interrupt_yes.rda"))
@@
 
-# Recording the guided tour
+# Iterative algorithm and its diagnostics
 
@@
 ```
+
+## Diagnostics
+
+Visualisatioin has been widely used for exploring and understanding data.
+Visualisation presents information in a graphicial manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation.
+Visual diagnostics can be real-time or post-run.
+Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
+This section focuses on the definition and production of post-run diagnoistics and the next section discusses real-time diagnostics.
+
+Post-run diagnostics requires the data produced during the algorithm to be stored/ saved in order to produce plot diagnostics.
+
+The graphical system in R is established based on the conception gramma of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.
+A layer includes 1) the dataset that powers the plot; 2) a geometric object that represents the visual shape of the plot and 3) relevant statistical transformation if needed.
+An important concept in the gramma of graphic is *aesthetic mapping*.
+Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object.
+For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot.
+To create a boxplot, we first map one variable on the x-axis and then map the five point summary of another variable on the y-axis.
+This computation of five point summary from the origin variable is statistical transformation.
+This definition of graphic through layers provides advantages to produce complex plots since all the plots, however complex, can be decomposed into basic geometric objects shaped by the variable supplied.
+
+This idea of decompositing a graph to basic elements inspires me to characterise the diagnostics of iterative algorithm in a similar fashion.
+The optimisation in many machine learning algorithms these days is iterative in nature, while remains as a black box.
+Being able to diagnose it visually allows researchers to have the tool to unfold the myth and thus provides more understand to the algorithm.
+
+The concept of gramma of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
+The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
+The next section shows how a global object is created in projection pursuit guided tour.
 
@@
 
-## add something about general vis
+Given all the information of interest in a global object, the next thing is to design its visual representation that can diagnose our algorithm.
+In an iterative algorithm, one would be interested to know how one parameter changes as the algorithm iterates.
+The parameter could be could be the decision variable, the value of the optimisation function or other parameter that facilitate the algorithm.
+In projectio pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`).
+These are the variables that will be mapped into the diagnostic plot.
+The global object also has its time series nature since the observations are recorded as the algorithm progresses and thus can't be switch in rows.
+This time series structure is always hierachical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm.
+Take our example, in the global object, each row has an `id` label which is the smallest unit the observation is ordered.
+A larger ordering unit is `tries` since it is updated everytime a new target basis is found.
+A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`.
+These features explain the definition of the global object for projection pursuit guided tour in Figure \@ref(fig:glb-obj).
 
-The creation of the global object facilitates the diagnostics of the optimisation in the projection pursuit guided tourr.
-Different diagnostic plots could be made to understand the optimisation **[pretty cheesy here]**
+```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="this is xxxx"}
+knitr::include_graphics(path = here::here("figures/global_obj.png"))
+```
 
@@
 
-## index vs time plots
+## Explore scalar parameters
 
-The tracing plot is useful to explore the index value. Given a global object created, `explore_trace_all()` plots the index value of all the bases.
+The most interesting parameter to explore is the value of objective function.
+Given the structure of projection pursuit guided tour algorithm, whatever searching method we use, we could roughly divide the observation into three category: searching, updating and interpolating points.
 
-```{r echo = TRUE}
-holes_2d_geo %>% explore_trace_all(magnify = TRUE)
-holes_2d_geo %>% explore_trace_interp()
+- *Searching points* include the observations that ones recorded during the trials and errors in the searching algorithm to find the target basis.
+In one `tries`, there could be only a few searching points while towards the end of the algorithm, they can easily go to hundreds.
+
+- *Updating points* are the observations for the target points and there are only one points per `tries`.
+We use "all searching points" to refer to the collection of both searching and updating points.
+
+- *interpolation points* exist specifically in the guided tour in order to provide continuous animated view from one target basis to another, thus they are not included in the searching procedure.
+There are multiple points in one tries but it doesn't have `loop` value.
+
+### Explore all searching points
+
+As mentioned previously, the largest feature of exploring searching points is its uncertainty in the number per `tries`.
+This makes it difficult to map `id` on the x-axis since the first few tries that can be informative to the evolving of the algorithm occupies a small number of points on the graph while the large less informative search points towards the end occupies the vast majority of the plot space.
+
+This motivates the use of summarised statistics.
+At each iteration, rather than knowing the value of *every* points, we are more interested to know a general summary of all the points and more importantly, the point with the largest `index_val` since it is the one that prescribe the interpolation and future search.
+Therefore, boxplot is a suitable candidate.
+
+However, boxplot doesn't tells us the number of point in each box.
+By showing boxplot for all `tries`, we lose information on how long it takes for the search to find the target basis.
+We may still want to show the exact points of every point if there are only a few points.
+Therefore a cutoff point is placed as an argument to decise when to switch from point geometry to boxplot geometry.
+
+This design is easy to implement with the concept of gramma of graphic.
+With the idea of layed gramma, we only need to subset the data and map it onto the either a point geometry or boxplot geometry.
+We can also add a layer with line geometry that connect the points with the largest `index_val` in each tries.
+This would show us which `tries` has the largest improvement.
+
+A comparison of using scatter plot (all point geometry) and my new design is shown in Figure \@ref(fig:explore-all-search).
+As we can see the points for the first few tries on hte scatter plot are all squeezed in a tiny region, being nearly unseeable.
+While the plot on the right balances out the points in each `tries`.
+By linking the updating points for each `tries`, we observe that the largest improvement occurs in `tries = 5` where the number of searching points dramastically increased from 11 points in `tries = 4` to 86 points in `tries = 5`.
+
+```{r explore-all-search, fig.cap="This is xxx"}
+dt <- holes_2d_better_max_tries %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+p1 <- dt %>% 
+  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
+  geom_point()  +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1)) 
+  #theme(legend.position = "none")
+
+dt_count <- dt %>%
+  group_by(as.factor(tries)) %>%
+  count() %>%
+  rename(tries = `as.factor(tries)`)
+
+p2 <- dt %>%
+  ggplot(aes(x = as.factor(tries), y = index_val, col = as.factor(tries))) +
+  geom_boxplot(data = dt %>% filter(tries %in% which(dt_count$n > 15))) +
+  geom_point(data = dt%>% filter(tries %in% which(dt_count$n < 15))) +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1)) 
+  #theme(legend.position = "none")
+
+(p1 | p2) & theme(legend.position = "bottom")
+
 ```
 
-## target basis vs interpolation steps, index vs time
+### Explore only interpolation points
 
-## low-d representation, PCA
+While mapping `tries` to the x axis allows for viewing the magnitude of the interpolation and also its density, we can't really see the curvature of the interpolated path.
+Think this would be important when the index is no longer smooth and the interpolation will jump up and down.
+Thus the geom_point + geom_line specification is still preferred.
+
+EXAMPLE: prevention on the decrese of interpolated value - interruption
+
+EXAMPLE: comparison of interpolation in noisy index function
+
+<!-- The tracing plot is useful to explore the index value. Given a global object created, `explore_trace_all()` plots the index value of all the bases.  -->
+
+```{r}
+# points is always included in the newest version
+p1 <- holes_2d_better %>% 
+  filter(info == "interpolation") %>% 
+  ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
+  geom_point() + 
+  geom_line(aes(group = 1))
+
+p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
+
+(p1 | p2)  &
+  theme(legend.position = "bottom")
+
+```
+
+### Adding more variables to the mapping
+
+For now we only include the iterator and `index_val` in the mapping of the plot, while more paramters of interest could be added to explore the effect their effect.
+
+EXAMPLE: polishing - add method : compare with max.tries and without for search_better.
+
+EXAMPLE: add alpha: compare the cooling system of the neighbourhood parameter in serach_better & search_posse.
+
+```{r}
+dt <- holes_2d_better_max_tries %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+p1 <- dt %>%
+  ggplot(aes(x = as.factor(tries), y = alpha, col = as.factor(alpha))) +
+  geom_jitter() +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1)) + 
+  theme(legend.position = "none")
+
+dt <- holes_2d_pos %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+p2 <- dt %>%
+  ggplot(aes(x = as.factor(tries), y = alpha, col = as.factor(alpha))) +
+  geom_jitter() +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1))  + 
+  theme(legend.position = "none")
+
+p1 | p2
+```
+
+## Explore matrix parameter
 

### 2020-05-19 paper/paper.tex
write GOG; explore post-run diagnostics using vis

@@
 
-Below listed several issues in projection pursuit optimisation. Some of
-them are general problems in optimisation literature, while others are
-more specific for PP optimisation.
+Below listed several issues in projection pursuit optimisation. Some are
+general optimisation problems, while others are more specific for PP
+optimisation.
 
@@
 \item
-  \emph{optimising on non-smooth function}: When the objective function
-  is non-differentiable, derivative information can not be obtained.
-  This means traditional gradient- or Hessian- based methods are not
-  feasible, stochastic optimisation method could be an alternative to
+  \emph{optimising non-smooth function}: When the objective function is
+  non-differentiable, derivative information can not be obtained, which
+  means traditional gradient- or Hessian- based methods are not
+  feasible. Stochastic optimisation method could be an alternative to
   solve these problems.
@@
   \emph{computation speed}: The optimisation procedure needs to be fast
-  to compute because tours produces real-time animation of the projected
+  to compute since tours produces real-time animation of the projected
   data.
@@
   algorithm, researchers usually set a seed to ensure the algorithm
-  producing the same result for every run. While this practice supports
-  reproducibility, less efforts has been made to guarantee different
-  seeds will provide the same result.
+  producing the same result for every run. This practice supports
+  reproducibility, while less efforts has been made to guarantee
+  different seeds will provide the same result.
 \item
   \emph{high-dimensional decision variable}: In projection pursuit, the
-  decision variable is the entry in the projection matrix, which is
-  usually high-dimensional. Researcher would be better off if they can
+  decision variable includes all the entries in the projection matrix,
+  which is high-dimensional. Researcher would be better off if they can
   understand the relative position of different projection matrix in the
@@
   procedure usually involves iteratively finding projection bases that
-  maximise the index function, while tour requires geodesic
+  maximises the index function, while tour requires geodesic
   interpolation between these bases to produce a continuous view for the
@@
 
-\hypertarget{recording-the-guided-tour}{%
-\section{Recording the guided tour}\label{recording-the-guided-tour}}
+\hypertarget{iterative-algorithm-and-its-diagnostics}{%
+\section{Iterative algorithm and its
+diagnostics}\label{iterative-algorithm-and-its-diagnostics}}
 
@@
 \end{Shaded}
+
+\hypertarget{diagnostics}{%
+\subsection{Diagnostics}\label{diagnostics}}
+
+Visualisatioin has been widely used for exploring and understanding
+data. Visualisation presents information in a graphicial manner and
+often allows people to see information they would otherwise not seen in
+the reporting of numerical summarisation. Visual diagnostics can be
+real-time or post-run. Real time diagnostic directly uses the data
+produced in the algorithm to produce visual representation and thus
+doesn't need to store the data. This section focuses on the definition
+and production of post-run diagnoistics and the next section discusses
+real-time diagnostics.
+
+Post-run diagnostics requires the data produced during the algorithm to
+be stored/ saved in order to produce plot diagnostics.
+
+The graphical system in R is established based on the conception gramma
+of graphic \citep{wickham2010layered}, where a graphic is defined using
+stacked layers in a coordination system. A layer includes 1) the dataset
+that powers the plot; 2) a geometric object that represents the visual
+shape of the plot and 3) relevant statistical transformation if needed.
+An important concept in the gramma of graphic is \emph{aesthetic
+mapping}. Aesthetic mapping links the variable in a dataset to
+information needed to produce a geometric object. For example, we map
+one variable on the x-axis and another on the y-axis to create a
+scatterplot. To create a boxplot, we first map one variable on the
+x-axis and then map the five point summary of another variable on the
+y-axis. This computation of five point summary from the origin variable
+is statistical transformation. This definition of graphic through layers
+provides advantages to produce complex plots since all the plots,
+however complex, can be decomposed into basic geometric objects shaped
+by the variable supplied.
+
+This idea of decompositing a graph to basic elements inspires me to
+characterise the diagnostics of iterative algorithm in a similar
+fashion. The optimisation in many machine learning algorithms these days
+is iterative in nature, while remains as a black box. Being able to
+diagnose it visually allows researchers to have the tool to unfold the
+myth and thus provides more understand to the algorithm.
+
+The concept of gramma of graphic requires a dataset to be supplied in
+making a plot, thus a global object needs to be created in the iterative
+algorithm. The variables will be mapped into element of a graphic to
+explore and thus should contain all the parameters of interest. The next
+section shows how a global object is created in projection pursuit
+guided tour.
 
@@
 
-\hypertarget{add-something-about-general-vis}{%
-\subsection{add something about general
-vis}\label{add-something-about-general-vis}}
+Given all the information of interest in a global object, the next thing
+is to design its visual representation that can diagnose our algorithm.
+In an iterative algorithm, one would be interested to know how one
+parameter changes as the algorithm iterates. The parameter could be
+could be the decision variable, the value of the optimisation function
+or other parameter that facilitate the algorithm. In projectio pursuit
+guided tour, apart from the decision variable projection basis, the
+value of index function (\texttt{index\_val}), we are also interested to
+explore the effect the different searching methods, the neighbourhood
+parameter alpha, and the stage a particular observation is in
+(\texttt{info}). These are the variables that will be mapped into the
+diagnostic plot. The global object also has its time series nature since
+the observations are recorded as the algorithm progresses and thus can't
+be switch in rows. This time series structure is always hierachical and
+nested since each row can be labelled with the smallest unit \texttt{id}
+and up by the iterative structure of the algorithm. Take our example, in
+the global object, each row has an \texttt{id} label which is the
+smallest unit the observation is ordered. A larger ordering unit is
+\texttt{tries} since it is updated everytime a new target basis is
+found. A nested ordering unit is \texttt{loop}, which increases by one
+as searching method iterates and starts over at a new \texttt{tries}.
+These features explain the definition of the global object for
+projection pursuit guided tour in Figure @ref(fig:glb-obj).
 
-The creation of the global object facilitates the diagnostics of the
-optimisation in the projection pursuit guided tourr. Different
-diagnostic plots could be made to understand the optimisation
-\textbf{{[}pretty cheesy here{]}}
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{this is xxxx}\label{fig:glb-obj}
+\end{figure}
 
@@
 
-\hypertarget{index-vs-time-plots}{%
-\subsection{index vs time plots}\label{index-vs-time-plots}}
+\hypertarget{explore-scalar-parameters}{%
+\subsection{Explore scalar parameters}\label{explore-scalar-parameters}}
 
-The tracing plot is useful to explore the index value. Given a global
-object created, \texttt{explore\_trace\_all()} plots the index value of
-all the bases.
+The most interesting parameter to explore is the value of objective
+function. Given the structure of projection pursuit guided tour
+algorithm, whatever searching method we use, we could roughly divide the
+observation into three category: searching, updating and interpolating
+points.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_trace_all}\NormalTok{(}\DataTypeTok{magnify =} \OtherTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\begin{itemize}
+\item
+  \emph{Searching points} include the observations that ones recorded
+  during the trials and errors in the searching algorithm to find the
+  target basis. In one \texttt{tries}, there could be only a few
+  searching points while towards the end of the algorithm, they can
+  easily go to hundreds.
+\item
+  \emph{Updating points} are the observations for the target points and
+  there are only one points per \texttt{tries}. We use ``all searching
+  points'' to refer to the collection of both searching and updating
+  points.
+\item
+  \emph{interpolation points} exist specifically in the guided tour in
+  order to provide continuous animated view from one target basis to
+  another, thus they are not included in the searching procedure. There
+  are multiple points in one tries but it doesn't have \texttt{loop}
+  value.
+\end{itemize}
+
+\hypertarget{explore-all-searching-points}{%
+\subsubsection{Explore all searching
+points}\label{explore-all-searching-points}}
+
+As mentioned previously, the largest feature of exploring searching
+points is its uncertainty in the number per \texttt{tries}. This makes
+it difficult to map \texttt{id} on the x-axis since the first few tries
+that can be informative to the evolving of the algorithm occupies a
+small number of points on the graph while the large less informative
+search points towards the end occupies the vast majority of the plot
+space.
+
+This motivates the use of summarised statistics. At each iteration,
+rather than knowing the value of \emph{every} points, we are more
+interested to know a general summary of all the points and more
+importantly, the point with the largest \texttt{index\_val} since it is
+the one that prescribe the interpolation and future search. Therefore,
+boxplot is a suitable candidate.
+
+However, boxplot doesn't tells us the number of point in each box. By
+showing boxplot for all \texttt{tries}, we lose information on how long
+it takes for the search to find the target basis. We may still want to
+show the exact points of every point if there are only a few points.
+Therefore a cutoff point is placed as an argument to decise when to
+switch from point geometry to boxplot geometry.
+
+This design is easy to implement with the concept of gramma of graphic.
+With the idea of layed gramma, we only need to subset the data and map
+it onto the either a point geometry or boxplot geometry. We can also add
+a layer with line geometry that connect the points with the largest
+\texttt{index\_val} in each tries. This would show us which
+\texttt{tries} has the largest improvement.
+
+A comparison of using scatter plot (all point geometry) and my new
+design is shown in Figure @ref(fig:explore-all-search). As we can see
+the points for the first few tries on hte scatter plot are all squeezed
+in a tiny region, being nearly unseeable. While the plot on the right
+balances out the points in each \texttt{tries}. By linking the updating
+points for each \texttt{tries}, we observe that the largest improvement
+occurs in \texttt{tries\ =\ 5} where the number of searching points
+dramastically increased from 11 points in \texttt{tries\ =\ 4} to 86
+points in \texttt{tries\ =\ 5}.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/explore-all-search-1.pdf}
+\caption{This is xxx}
+\end{figure}
+
+\hypertarget{explore-only-interpolation-points}{%
+\subsubsection{Explore only interpolation
+points}\label{explore-only-interpolation-points}}
+
+While mapping \texttt{tries} to the x axis allows for viewing the
+magnitude of the interpolation and also its density, we can't really see
+the curvature of the interpolated path. Think this would be important
+when the index is no longer smooth and the interpolation will jump up
+and down. Thus the geom\_point + geom\_line specification is still
+preferred.
+
+EXAMPLE: prevention on the decrese of interpolated value - interruption
+
+EXAMPLE: comparison of interpolation in noisy index function
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_trace_interp}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{adding-more-variables-to-the-mapping}{%
+\subsubsection{Adding more variables to the
+mapping}\label{adding-more-variables-to-the-mapping}}
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-2.pdf}
+For now we only include the iterator and \texttt{index\_val} in the
+mapping of the plot, while more paramters of interest could be added to
+explore the effect their effect.
 
-\hypertarget{target-basis-vs-interpolation-steps-index-vs-time}{%
-\subsection{target basis vs interpolation steps, index vs
-time}\label{target-basis-vs-interpolation-steps-index-vs-time}}
+EXAMPLE: polishing - add method : compare with max.tries and without for
+search\_better.
 
-\hypertarget{low-d-representation-pca}{%
-\subsection{low-d representation, PCA}\label{low-d-representation-pca}}
+EXAMPLE: add alpha: compare the cooling system of the neighbourhood
+parameter in serach\_better \& search\_posse.
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+\hypertarget{explore-matrix-parameter}{%
+\subsection{Explore matrix parameter}\label{explore-matrix-parameter}}
 
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
 

### 2020-05-19 paper/paper.Rmd
knit again
bit on explore basis

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 
@@
 load(here::here("data/holes_1d_geo.rda"))
+load(here::here("data/holes_1d_geo_polish.rda"))
 load(here::here("data/holes_2d_geo.rda"))
@@
 
+```{r echo = TRUE}
+nrow(holes_2d_better_max_tries)
+nrow(holes_2d_pos)
+```
+
 ```{r}
@@
 ## Explore matrix parameter
+
+The variable or object to explore could sometimes be a more than a scale, for example, in a tour, we would also be interested to explore the relative position of the projection basis in the vector space.
+This imposes difficulties in visualisation since as human, we are bounded to preceive at most three dimensions.
+Dimension reduction is reuqired to summarise the projection matrix to a managable number of variable to map to a plot.
+Principal copmonent analysis is used to extract the direction that reflects most variation in the data.
 

### 2020-05-19 paper/paper.tex
bit on explore basis

@@
 \begin{verbatim}
-## # A tibble: 5 x 7
-##   basis             index_val tries info              loop method             id
-##   <list>                <dbl> <dbl> <chr>            <dbl> <chr>           <int>
-## 1 <dbl[,1] [5 x 1]>     0.749     1 start               NA <NA>                1
-## 2 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     2
-## 3 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     3
-## 4 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     4
-## 5 <dbl[,1] [5 x 1]>     0.749     1 direction_search     1 search_geodesic     5
+## # A tibble: 5 x 8
+##   basis           index_val tries info            loop method        alpha    id
+##   <list>              <dbl> <dbl> <chr>          <dbl> <chr>         <dbl> <int>
+## 1 <dbl[,1] [5 x ~     0.749     1 start             NA <NA>            0.5     1
+## 2 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       2
+## 3 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       3
+## 4 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       4
+## 5 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       5
 \end{verbatim}
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{nrow}\NormalTok{(holes_2d_better_max_tries)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 1531
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{nrow}\NormalTok{(holes_2d_pos)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 959
+\end{verbatim}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
+The variable or object to explore could sometimes be a more than a
+scale, for example, in a tour, we would also be interested to explore
+the relative position of the projection basis in the vector space. This
+imposes difficulties in visualisation since as human, we are bounded to
+preceive at most three dimensions. Dimension reduction is reuqired to
+summarise the projection matrix to a managable number of variable to map
+to a plot. Principal copmonent analysis is used to extract the direction
+that reflects most variation in the data.
+
 \begin{Shaded}
 \begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{()}
+\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{() }
 \end{Highlighting}
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
 
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1.pdf}
 

### 2020-05-19 source/explore_basis.Rmd
a document for illustrating the exploration of the basis

---
title: "explore_basis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ferrn)

load(here::here("data/holes_1d_geo.rda"))
load(here::here("data/holes_1d_geo_polish.rda"))
load(here::here("data/holes_2d_geo.rda"))
load(here::here("data/holes_2d_geo_polish.rda"))
```

*Explore the first two principle component along with info*

This helps to graphically illustrate the geodesic searching algorithm

```{r}
holes_1d_geo %>% explore_proj_pca() 
```

*Explore the first two principle component along with method*

It doesn't seem that the polishing seraching is very efficient - the space is too large. We could color by alpha in the polish to see it better.

```{r}
glb_obj <- bind_rows(holes_1d_geo, holes_1d_geo_polish)

num_col <- ncol(glb_obj$basis[[1]])
num_row <- nrow(glb_obj$basis[[1]])

basis <- purrr::flatten_dbl(glb_obj$basis)  %>% matrix(ncol = num_row * num_col, byrow = TRUE)

loadings <- basis %>%
  stats::prcomp() %>%
  stats::predict() %>%
  tibble::as_tibble()

combined <- tibble::as_tibble(basis) %>%
  dplyr::bind_cols(loadings) %>%
  dplyr::bind_cols(glb_obj %>% dplyr::select(-basis))

combined %>% 
  ggplot(aes(x = PC1, y = PC2, col = method)) +
  geom_point() +
  geom_point(data = combined %>% filter(method == "search_geodesic")) +
  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
  theme(aspect.ratio = 1)
```

*Explore the first two principle component along with alpha*

We can indeed see that an init of 0.5 for alpha is too big and unnecessary. What about the projection distant between the last two target basis in the search geodesic? Their distance is only 0.02! Filter for alpha < 0.02

```{r}
combined %>% 
  ggplot(aes(x = PC1, y = PC2, col = as.factor(alpha))) +
  geom_point() +
  geom_point(data = combined %>% filter(method == "search_geodesic")) +
  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
  theme(aspect.ratio = 1)
```

```{r echo = TRUE}
last_two <- holes_1d_geo %>% filter(info == "best_line_search") %>% tail(2) %>% pull(basis)
last_two_dist <- tourr::proj_dist(last_two[[1]], last_two[[2]])
last_two_dist
```

```{r}
combined %>% filter(alpha < 0.03) %>% 
  ggplot(aes(x = PC1, y = PC2, col = as.factor(alpha))) +
  geom_point() +
  geom_point(data = combined %>% filter(method == "search_geodesic")) +
  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
  theme(aspect.ratio = 1)
```

```{r}
combined %>% filter(info == "polish_best") %>% dplyr::select(index_val, alpha, loop)
```

*Conclusion*

This flow of analysis shows that an init value of 0.5 for alpha is not ideal. A value based on the distance of the last two target basis would be a better option.

*Explore the first two principle component along with index_val*

The problem of mapping index_val to color is that human eyes can only preceive the cahnge of index_val, but not its magnitude. We can't see how fast it increases! That's why I think it is indeed necessary to create a 3D plot, but just to make the plot fancy but because both the principal compoennt and index_val need to be mapped to axis for us to preceive the changes.

```{r}
glb_obj <- holes_1d_geo #%>% filter(info == "interpolation")

num_col <- ncol(glb_obj$basis[[1]])
num_row <- nrow(glb_obj$basis[[1]])

basis <- purrr::flatten_dbl(glb_obj$basis)  %>% matrix(ncol = num_row * num_col, byrow = TRUE)

loadings <- basis %>%
  stats::prcomp() %>%
  stats::predict() %>%
  tibble::as_tibble()

combined <- tibble::as_tibble(basis) %>%
  dplyr::bind_cols(loadings) %>%
  dplyr::bind_cols(glb_obj %>% dplyr::select(-basis))

combined %>% 
  ggplot(aes(x = PC1, y = PC2, col = index_val)) +
  geom_point() +
  geom_point(data = combined %>% filter(method == "search_geodesic")) +
  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
  theme(aspect.ratio = 1)
```

Haven't seen an option to do 3D graph in ggplot - the best I can find is plotly. Now it feels much like a hill climbing. tries 12, the last tries is not doing that good tho. The largest improvement happens around tries = 7/8.

```{r}
library(plotly)
plot_ly(combined, x = ~PC1, y = ~PC2, z = ~index_val) %>%
  add_markers(color = ~tries) %>% 
  layout(scene = list(aspectratio = list(x = 1,y = 1, z= 1)))

```

*A 2D example*

We did the PCA for both columns and take the first PC of both. Again, mapping index_val to color doesn't show much info.

```{r}
glb_obj <- holes_2d_geo
num_col <- ncol(glb_obj$basis[[1]])
num_row <- nrow(glb_obj$basis[[1]])

basis <- purrr::flatten_dbl(glb_obj$basis)  %>% matrix(ncol = num_row * num_col, byrow = TRUE)

v1 <- stats::prcomp(basis[,1:num_row]) %>%
  stats::predict() %>%
  tibble::as_tibble() %>%
  dplyr::select(PC1)

v2 <- stats::prcomp(basis[,(num_row + 1):(2*num_row)]) %>%
  stats::predict() %>%
  tibble::as_tibble() %>%
  dplyr::select(PC1) %>% rename(PC2 = PC1)

combined <- tibble::as_tibble(basis) %>%
  dplyr::bind_cols(v1) %>%
  dplyr::bind_cols(v2) %>%
  dplyr::bind_cols(glb_obj %>% dplyr::select(-basis))

combined %>% 
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(col = index_val)) +
  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
  theme(aspect.ratio = 1)

```

Using plotly, we can preceive the changes better - there seems to be a local optimum somewhere and an only way to get to the higher hill.

```{r}
library(plotly)
plot_ly(combined, x = ~PC1, y = ~PC2, z = ~index_val, hoverlabel = ~tries) %>%
  add_markers(color = ~tries) %>% 
  layout(scene = list(size = 0.5, aspectratio = list(x = 1,y = 1, z = 1)))
```

Filter only the interpolation points and repeat the process again, this time we can see much more. The searching gets a bit stucked at index_val between 0.85 - 0.9 and then proceeds to the upper hill.

```{r}
glb_obj <- holes_2d_geo %>% filter(info == "interpolation")

num_col <- ncol(glb_obj$basis[[1]])
num_row <- nrow(glb_obj$basis[[1]])

basis <- purrr::flatten_dbl(glb_obj$basis)  %>% matrix(ncol = num_row * num_col, byrow = TRUE)

v1 <- stats::prcomp(basis[,1:num_row]) %>%
  stats::predict() %>%
  tibble::as_tibble() %>%
  dplyr::select(PC1)

v2 <- stats::prcomp(basis[,(num_row + 1):(2*num_row)]) %>%
  stats::predict() %>%
  tibble::as_tibble() %>%
  dplyr::select(PC1) %>% rename(PC2 = PC1)

combined <- tibble::as_tibble(basis) %>%
  dplyr::bind_cols(v1) %>%
  dplyr::bind_cols(v2) %>%
  dplyr::bind_cols(glb_obj %>% dplyr::select(-basis))
```

```{r}
library(plotly)
plot_ly(combined, x = ~PC1, y = ~PC2, z = ~index_val, hoverlabel = ~tries) %>%
  add_markers(color = ~tries) %>% 
  layout(scene = list(width = 2, aspectratio = list(x = 1,y = 1, z= 1)))
```

### 2020-05-26 paper/paper.Rmd
reorder sections and plots

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 
@@
 library(ggrepel)
+```
 
-load(here::here("data/temp.rda"))
-load(here::here("data/data_mult.rda"))
-load(here::here("data/holes_1d_geo.rda"))
-load(here::here("data/holes_1d_geo_polish.rda"))
-load(here::here("data/holes_2d_geo.rda"))
-load(here::here("data/holes_2d_better.rda"))
-load(here::here("data/holes_2d_better_max_tries.rda"))
-load(here::here("data/holes_2d_better_random.rda"))
-load(here::here("data/holes_2d_geo_polish.rda"))
-load(here::here("data/holes_2d_better_polish.rda"))
-load(here::here("data/holes_2d_better_max_tries_polish.rda"))
-load(here::here("data/holes_2d_better_random_polish.rda"))
-load(here::here("data/holes_2d_pos.rda"))
-load(here::here("data/interrupt_yes.rda"))
-load(here::here("data/interrupt_no.rda"))
-
+```{r load-data}
+files <- paste0("data/", list.files(here::here("data")))
+purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 ```
@@
 
-```{r}
+```{r tour-path}
 knitr::include_graphics(path = here::here("figures/tour_path_new.png"))
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r algo,eval = FALSE, echo = TRUE}
 animation <- function(){
@@
 
-```{r}
+```{r origin-data}
 origin_dt_bi <- data_mult %>% 
@@
 
-```{r eval = FALSE}
+```{r code, eval = FALSE}
 set.seed(123456)
@@
 
-```{r echo = TRUE}
+```{r glb-obj-1, echo = TRUE}
 holes_1d_geo %>% head(5)
@@
 
-```{r echo = TRUE}
+```{r glb-obj-2, echo = TRUE}
 holes_2d_better %>% head(5) 
@@
 
-# Visual methods
+# Post-run static diagnostics plots
 
@@
 A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`.
-These features explain the definition of the global object for projection pursuit guided tour in Figure \@ref(fig:glb-obj).
+These features explain the definition of the global object for projection pursuit guided tour in Figure \ref{fig:glb-obj}.
 
-```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="this is xxxx"}
+```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}this is xxxx"}
 knitr::include_graphics(path = here::here("figures/global_obj.png"))
 ```
-
-*The examples and results below still need to be modified as the development / modification of both ferrn and tourr package*.
 
@@
 
-A comparison of using scatter plot (all point geometry) and my new design is shown in Figure \@ref(fig:explore-all-search).
+A comparison of using scatter plot (all point geometry) and my new design is shown in Figure \ref{fig:points}.
 As we can see the points for the first few tries on hte scatter plot are all squeezed in a tiny region, being nearly unseeable.
@@
 
-```{r explore-all-search, fig.cap="This is xxx"}
+```{r points, fig.cap="\\label{points} This is xxx"}
 dt <- holes_2d_better_max_tries %>%
@@
 
-EXAMPLE: prevention on the decrese of interpolated value - interruption
+*Example 1: Interruption* The left panel of Figure \ref{fig:interruption} presents the trace plot of the interpolated basis using `search_better` to optimise the holes index (2D projection using the larger dataset).
+Tour interpolates from the current basis to the target basis, which has been found to have a higher index value by projection pursuit.
+The target basis will then be sent to the projection pursuit as the current basis to find the next target basis.
+We can observe from the plot that there are basis with even higher index value on the interpolation path.
+These bases could be used to search for new basis in the next round.
+Therefore, an interruption is constructed to only take interpolated basis up to the point where the index value stops increasing and the last basis is taken as the current basis for the next round of search.
+After implementing the interruption, the tracing plot with the same configuration is shown on the right panel.
+Rather than interpolate to the target basis, the interruption interrupt the process after id = 60 and proceeds the next round of search from the interpolated basis.
+Taking advantage of the interpolated basis results in a higher index value in the end with fewer steps.
 
-EXAMPLE: comparison of interpolation in noisy index function
+```{r interruption, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without the interruption.
+The interruption stops the interpolation when the index value starts to decrease, see the difference at id being around 60.
+The implementation of the interuption finds an ending basis with higher index value using fewer steps.
+"}
+p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>% 
+  mutate(anno = c("current basis", "interpolated basis", "target basis"))
+
+p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") + 
+  geom_point(data = p1_anno) + 
+  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) + 
+  theme(legend.position = "none")
+
+p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) + 
+  theme(legend.position = "none")
+
+p1 | p2
+```
+
+*Example 2: Noisy index function* Add an example on noisy function
 
@@
 
-```{r}
+```{r interpolation}
 # points is always included in the newest version
@@
 
-EXAMPLE: polishing - add method : compare with max.tries and without for search_better.
-
-EXAMPLE: add alpha: compare the cooling system of the neighbourhood parameter in serach_better & search_posse.
-
-```{r echo = TRUE}
-nrow(holes_2d_better_max_tries)
-nrow(holes_2d_pos)
-```
-
-```{r}
-dt <- holes_2d_better_max_tries %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-p1 <- dt %>%
-  ggplot(aes(x = as.factor(tries), y = alpha, col = as.factor(alpha))) +
-  geom_jitter() +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1)) + 
-  theme(legend.position = "none")
-
-dt <- holes_2d_pos %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-p2 <- dt %>%
-  ggplot(aes(x = as.factor(tries), y = alpha, col = as.factor(alpha))) +
-  geom_jitter() +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1))  + 
-  theme(legend.position = "none")
-
-p1 | p2
-```
-
-## Explore matrix parameter
-
-The variable or object to explore could sometimes be a more than a scale, for example, in a tour, we would also be interested to explore the relative position of the projection basis in the vector space.
-This imposes difficulties in visualisation since as human, we are bounded to preceive at most three dimensions.
-Dimension reduction is reuqired to summarise the projection matrix to a managable number of variable to map to a plot.
-Principal copmonent analysis is used to extract the direction that reflects most variation in the data.
-
-```{r echo = TRUE}
-holes_1d_geo %>% explore_proj_pca() 
-```
-
-# Animating plots
-
-```{r echo = TRUE, eval = FALSE}
-holes_1d_geo %>% explore_proj_pca(animate = TRUE)
-```
-
-# Finding errors and developing improvements
-
-The visualisation methods introduced above allow us to assess the performance of each optimisation routine and thus provide improvments.
-
-## Interrupt
-
-The left panel of Figure \@ref(fig:interruption) presents the trace plot of the interpolated basis using `search_better` to optimise the holes index (2D projection using the larger dataset).
-Tour interpolates from the current basis to the target basis, which has been found to have a higher index value by projection pursuit.
-The target basis will then be sent to the projection pursuit as the current basis to find the next target basis.
-We can observe from the plot that there are basis with even higher index value on the interpolation path.
-These bases could be used to search for new basis in the next round.
-Therefore, an interruption is constructed to only take interpolated basis up to the point where the index value stops increasing and the last basis is taken as the current basis for the next round of search.
-After implementing the interruption, the tracing plot with the same configuration is shown on the right panel.
-Rather than interpolate to the target basis, the interruption interrupt the process after id = 60 and proceeds the next round of search from the interpolated basis.
-Taking advantage of the interpolated basis results in a higher index value in the end with fewer steps.
-
-```{r interruption, fig.cap = "\\label{interruption} Trace plots of the interpolated basis with and without the interruption.
-The interruption stops the interpolation when the index value starts to decrease, see the difference at id being around 60.
-The implementation of the interuption finds an ending basis with higher index value using fewer steps.
-"}
-p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>% 
-  mutate(anno = c("current basis", "interpolated basis", "target basis"))
-
-p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") + 
-  geom_point(data = p1_anno) + 
-  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80)
-
-p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80)
-
-p1 | p2
-```
-
-Further the case when the index function is not smooth, the interpolation may not attain -> modification
-
-## Polish
-
-In principle, all the optimisation routines should present the same output on the same problem.
-In Figure \@ref(fig:trace-compare), red dots shows the trace of the interpolated basis using `search_geodesic` and `search_better` (with max.tries = 100), respectively, on the 2D projection problem.
+*Example 3: Polish* In principle, all the optimisation routines should present the same output on the same problem.
+In Figure \ref{fig:trace-compare}, red dots shows the trace of the interpolated basis using `search_geodesic` and `search_better` (with max.tries = 100), respectively, on the 2D projection problem.
 We can observe that they attain slightly different ending index value, which is not ideal.
@@
 
-```{r}
+```{r polish}
 p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("search_geodesic")
+  explore_trace_interp(col = method) + ggtitle("search_geodesic")
 
 p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("search_better")
+  explore_trace_interp(col = method) + ggtitle("search_better")
 
@@
 
-```{r eval = FALSE, echo = TRUE}
+```{r sample-code, eval = FALSE, echo = TRUE}
 set.seed(123456)
@@
 `max.tries` is a hyperparameter that controls the maximum number of try without improvement and its default value is 25.
-As shown in Figure \@ref(fig:trace-compare), after polishing, both trials attain the same index value.
+As shown in Figure \ref{fig:trace-compare}, after polishing, both trials attain the same index value.
 However, a small `max.tries` of 25 is not sufficient for the algorithm to find the true maximum.
@@
 
-```{r trace-compare, fig.cap="Breakdown of index value when using different max.tries in search better in conjunction with search polish.
+```{r trace-compare, fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
 Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}
@@
 p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("max.tries = 25")
+  explore_trace_interp(col = method) + ggtitle("max.tries = 25")
 
 p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
-  explore_trace_interp(point = TRUE) + ggtitle("max.tries = 500")
+  explore_trace_interp(col = method) + ggtitle("max.tries = 500")
 
@@
 
-## posse?
+*Example 4: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in serach_better & search_posse.
 
-search_better annd search_geodesic methods are not doing good on 2d projection -> see the simple case above: abs best is 0.96 while they only get to 0.88-ish.
-Would be worse for higher dimension.
-thus search_posse.
-get 0.95 and then polish to 0.96 :)
+```{r nrow, echo = TRUE}
+nrow(holes_2d_better_max_tries)
+nrow(holes_2d_pos)
+```
+
+```{r alpha}
+dt <- holes_2d_better_max_tries %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+p1 <- dt %>%
+  ggplot(aes(x = as.factor(tries), y = alpha)) +
+  geom_jitter(aes( col = as.factor(alpha))) +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1)) + 
+  theme(legend.position = "none")
+
+dt <- holes_2d_pos %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+p2 <- dt %>%
+  ggplot(aes(x = as.factor(tries), y = alpha,)) +
+  geom_jitter(aes(col = as.factor(alpha))) +
+  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
+            aes(group = 1))  + 
+  theme(legend.position = "none")
+
+p1 | p2
+```
+
+## Explore matrix parameter
+
+The variable or object to explore could sometimes be a more than a scale, for example, in a tour, we would also be interested to explore the relative position of the projection basis in the vector space.
+This imposes difficulties in visualisation since as human, we are bounded to preceive at most three dimensions.
+Dimension reduction is reuqired to summarise the projection matrix to a managable number of variable to map to a plot.
+Principal copmonent analysis is used to extract the direction that reflects most variation in the data.
+
+*Example 5 PCA* refine the example of matrix parameter exploration
+
+```{r pca, echo = TRUE}
+holes_1d_geo %>% explore_proj_pca(col = info) 
+```
+
+# Real-time animated diagnostic plots
+
+```{r pca-animated, echo = TRUE, eval = FALSE}
+holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info)
+```
 
 # Vis package
+
+Everything is coded up in a package.
+
+\clearpage

### 2020-05-26 paper/paper.tex
reorder sections and plots

@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\includegraphics{paper_files/figure-latex/origin-data-1.pdf}
 
@@
 \begin{verbatim}
-## # A tibble: 5 x 7
-##   basis             index_val tries info           loop method           id
-##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>         <int>
-## 1 <dbl[,2] [6 x 2]>     0.804     1 start            NA <NA>              1
-## 2 <dbl[,2] [6 x 2]>     0.793     1 random_search     1 search_better     2
-## 3 <dbl[,2] [6 x 2]>     0.784     1 random_search     2 search_better     3
-## 4 <dbl[,2] [6 x 2]>     0.773     1 random_search     3 search_better     4
-## 5 <dbl[,2] [6 x 2]>     0.795     1 random_search     4 search_better     5
+## # A tibble: 5 x 8
+##   basis             index_val tries info           loop method       alpha    id
+##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>        <dbl> <int>
+## 1 <dbl[,2] [6 x 2]>     0.804     1 start            NA <NA>           0.5     1
+## 2 <dbl[,2] [6 x 2]>     0.793     1 random_search     1 search_bett~   0.5     2
+## 3 <dbl[,2] [6 x 2]>     0.784     1 random_search     2 search_bett~   0.5     3
+## 4 <dbl[,2] [6 x 2]>     0.773     1 random_search     3 search_bett~   0.5     4
+## 5 <dbl[,2] [6 x 2]>     0.795     1 random_search     4 search_bett~   0.5     5
 \end{verbatim}
 
-\hypertarget{visual-methods}{%
-\section{Visual methods}\label{visual-methods}}
+\hypertarget{post-run-static-diagnostics-plots}{%
+\section{Post-run static diagnostics
+plots}\label{post-run-static-diagnostics-plots}}
 
@@
 These features explain the definition of the global object for
-projection pursuit guided tour in Figure @ref(fig:glb-obj).
+projection pursuit guided tour in Figure \ref{fig:glb-obj}.
 
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{this is xxxx}\label{fig:glb-obj}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}this is xxxx}\label{fig:glb-obj}
 \end{figure}
-
-\emph{The examples and results below still need to be modified as the
-development / modification of both ferrn and tourr package}.
 
@@
 A comparison of using scatter plot (all point geometry) and my new
-design is shown in Figure @ref(fig:explore-all-search). As we can see
-the points for the first few tries on hte scatter plot are all squeezed
-in a tiny region, being nearly unseeable. While the plot on the right
-balances out the points in each \texttt{tries}. By linking the updating
-points for each \texttt{tries}, we observe that the largest improvement
-occurs in \texttt{tries\ =\ 5} where the number of searching points
-dramastically increased from 11 points in \texttt{tries\ =\ 4} to 86
-points in \texttt{tries\ =\ 5}.
+design is shown in Figure \ref{fig:points}. As we can see the points for
+the first few tries on hte scatter plot are all squeezed in a tiny
+region, being nearly unseeable. While the plot on the right balances out
+the points in each \texttt{tries}. By linking the updating points for
+each \texttt{tries}, we observe that the largest improvement occurs in
+\texttt{tries\ =\ 5} where the number of searching points dramastically
+increased from 11 points in \texttt{tries\ =\ 4} to 86 points in
+\texttt{tries\ =\ 5}.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/explore-all-search-1.pdf}
-\caption{This is xxx}
+\includegraphics{paper_files/figure-latex/points-1.pdf}
+\caption{\label{points} This is xxx}
 \end{figure}
@@
 
-EXAMPLE: prevention on the decrese of interpolated value - interruption
+\emph{Example 1: Interruption} The left panel of Figure
+\ref{fig:interruption} presents the trace plot of the interpolated basis
+using \texttt{search\_better} to optimise the holes index (2D projection
+using the larger dataset). Tour interpolates from the current basis to
+the target basis, which has been found to have a higher index value by
+projection pursuit. The target basis will then be sent to the projection
+pursuit as the current basis to find the next target basis. We can
+observe from the plot that there are basis with even higher index value
+on the interpolation path. These bases could be used to search for new
+basis in the next round. Therefore, an interruption is constructed to
+only take interpolated basis up to the point where the index value stops
+increasing and the last basis is taken as the current basis for the next
+round of search. After implementing the interruption, the tracing plot
+with the same configuration is shown on the right panel. Rather than
+interpolate to the target basis, the interruption interrupt the process
+after id = 60 and proceeds the next round of search from the
+interpolated basis. Taking advantage of the interpolated basis results
+in a higher index value in the end with fewer steps.
 
-EXAMPLE: comparison of interpolation in noisy index function
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/interruption-1.pdf}
+\caption{\label{interruption}Trace plots of the interpolated basis with
+and without the interruption. The interruption stops the interpolation
+when the index value starts to decrease, see the difference at id being
+around 60. The implementation of the interuption finds an ending basis
+with higher index value using fewer steps.}
+\end{figure}
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\emph{Example 2: Noisy index function} Add an example on noisy function
+
+\includegraphics{paper_files/figure-latex/interpolation-1.pdf}
 
@@
 
-EXAMPLE: polishing - add method : compare with max.tries and without for
-search\_better.
+\emph{Example 3: Polish} In principle, all the optimisation routines
+should present the same output on the same problem. In Figure
+\ref{fig:trace-compare}, red dots shows the trace of the interpolated
+basis using \texttt{search\_geodesic} and \texttt{search\_better} (with
+max.tries = 100), respectively, on the 2D projection problem. We can
+observe that they attain slightly different ending index value, which is
+not ideal. This motivates the creation of a polishing search that
+polishes the ending basis and achieves unity on different methods.
 
-EXAMPLE: add alpha: compare the cooling system of the neighbourhood
-parameter in serach\_better \& search\_posse.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{nrow}\NormalTok{(holes_2d_better_max_tries)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 1531
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{nrow}\NormalTok{(holes_2d_pos)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 959
-\end{verbatim}
-
-\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-\hypertarget{explore-matrix-parameter}{%
-\subsection{Explore matrix parameter}\label{explore-matrix-parameter}}
-
-The variable or object to explore could sometimes be a more than a
-scale, for example, in a tour, we would also be interested to explore
-the relative position of the projection basis in the vector space. This
-imposes difficulties in visualisation since as human, we are bounded to
-preceive at most three dimensions. Dimension reduction is reuqired to
-summarise the projection matrix to a managable number of variable to map
-to a plot. Principal copmonent analysis is used to extract the direction
-that reflects most variation in the data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{() }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## Warning: The `x` argument of `as_tibble.matrix()` must have column names if `.name_repair` is omitted as of tibble 2.0.0.
-## Using compatibility `.name_repair`.
-## This warning is displayed once every 8 hours.
-## Call `lifecycle::last_warnings()` to see where this warning was generated.
-\end{verbatim}
-
-\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-\hypertarget{animating-plots}{%
-\section{Animating plots}\label{animating-plots}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{finding-errors-and-developing-improvements}{%
-\section{Finding errors and developing
-improvements}\label{finding-errors-and-developing-improvements}}
-
-The visualisation methods introduced above allow us to assess the
-performance of each optimisation routine and thus provide improvments.
-
-\hypertarget{interrupt}{%
-\subsection{Interrupt}\label{interrupt}}
-
-The left panel of Figure @ref(fig:interruption) presents the trace plot
-of the interpolated basis using \texttt{search\_better} to optimise the
-holes index (2D projection using the larger dataset). Tour interpolates
-from the current basis to the target basis, which has been found to have
-a higher index value by projection pursuit. The target basis will then
-be sent to the projection pursuit as the current basis to find the next
-target basis. We can observe from the plot that there are basis with
-even higher index value on the interpolation path. These bases could be
-used to search for new basis in the next round. Therefore, an
-interruption is constructed to only take interpolated basis up to the
-point where the index value stops increasing and the last basis is taken
-as the current basis for the next round of search. After implementing
-the interruption, the tracing plot with the same configuration is shown
-on the right panel. Rather than interpolate to the target basis, the
-interruption interrupt the process after id = 60 and proceeds the next
-round of search from the interpolated basis. Taking advantage of the
-interpolated basis results in a higher index value in the end with fewer
-steps.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/interruption-1.pdf}
-\caption{\label{interruption} Trace plots of the interpolated basis with
-and without the interruption. The interruption stops the interpolation
-when the index value starts to decrease, see the difference at id being
-around 60. The implementation of the interuption finds an ending basis
-with higher index value using fewer steps.}
-\end{figure}
-
-Further the case when the index function is not smooth, the
-interpolation may not attain -\textgreater{} modification
-
-\hypertarget{polish}{%
-\subsection{Polish}\label{polish}}
-
-In principle, all the optimisation routines should present the same
-output on the same problem. In Figure @ref(fig:trace-compare), red dots
-shows the trace of the interpolated basis using
-\texttt{search\_geodesic} and \texttt{search\_better} (with max.tries =
-100), respectively, on the 2D projection problem. We can observe that
-they attain slightly different ending index value, which is not ideal.
-This motivates the creation of a polishing search that polishes the
-ending basis and achieves unity on different methods.
-
-\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{paper_files/figure-latex/polish-1.pdf}
 
@@
 improvement and its default value is 25. As shown in Figure
-@ref(fig:trace-compare), after polishing, both trials attain the same
+\ref{fig:trace-compare}, after polishing, both trials attain the same
 index value. However, a small \texttt{max.tries} of 25 is not sufficient
@@
 \includegraphics{paper_files/figure-latex/trace-compare-1.pdf}
-\caption{Breakdown of index value when using different max.tries in
-search better in conjunction with search polish. Both attain the same
-final index value after the polishing while using a max.tries 25 is not
-sufficient to find the ture maximum.}
+\caption{\label{trace-compare}Breakdown of index value when using
+different max.tries in search better in conjunction with search polish.
+Both attain the same final index value after the polishing while using a
+max.tries 25 is not sufficient to find the ture maximum.}
 \end{figure}
 
-\hypertarget{posse}{%
-\subsection{posse?}\label{posse}}
+\emph{Example 4: The neighbourhood parameter alpha} Add an example on
+comparing the neighbourhood parameter in serach\_better \&
+search\_posse.
 
-search\_better annd search\_geodesic methods are not doing good on 2d
-projection -\textgreater{} see the simple case above: abs best is 0.96
-while they only get to 0.88-ish. Would be worse for higher dimension.
-thus search\_posse. get 0.95 and then polish to 0.96 :)
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{nrow}\NormalTok{(holes_2d_better_max_tries)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 1529
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{nrow}\NormalTok{(holes_2d_pos)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 1056
+\end{verbatim}
+
+\includegraphics{paper_files/figure-latex/alpha-1.pdf}
+
+\hypertarget{explore-matrix-parameter}{%
+\subsection{Explore matrix parameter}\label{explore-matrix-parameter}}
+
+The variable or object to explore could sometimes be a more than a
+scale, for example, in a tour, we would also be interested to explore
+the relative position of the projection basis in the vector space. This
+imposes difficulties in visualisation since as human, we are bounded to
+preceive at most three dimensions. Dimension reduction is reuqired to
+summarise the projection matrix to a managable number of variable to map
+to a plot. Principal copmonent analysis is used to extract the direction
+that reflects most variation in the data.
+
+\emph{Example 5 PCA} refine the example of matrix parameter exploration
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ info) }
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{paper_files/figure-latex/pca-1.pdf}
+
+\hypertarget{real-time-animated-diagnostic-plots}{%
+\section{Real-time animated diagnostic
+plots}\label{real-time-animated-diagnostic-plots}}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =}\NormalTok{ info)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{vis-package}{%
+\section{Vis package}\label{vis-package}}
+
+Everything is coded up in a package.
+
+\clearpage
 

### 2020-05-26 paper/paper.Rmd
reword explore scalar param section and part of matrix param

@@
 The most interesting parameter to explore is the value of objective function.
-Given the structure of projection pursuit guided tour algorithm, whatever searching method we use, we could roughly divide the observation into three category: searching, updating and interpolating points.
+The points recorded in the global object can be divided into two broad categories: searching points and interpolating points
 
-- *Searching points* include the observations that ones recorded during the trials and errors in the searching algorithm to find the target basis.
-In one `tries`, there could be only a few searching points while towards the end of the algorithm, they can easily go to hundreds.
+- *Searching points* include the observations that are recorded in the searching algorithm in order to find the target basis.
+The points for target bases is also included in the searching points and there is one such point per `tries`.
 
-- *Updating points* are the observations for the target points and there are only one points per `tries`.
-We use "all searching points" to refer to the collection of both searching and updating points.
+- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
 
-- *interpolation points* exist specifically in the guided tour in order to provide continuous animated view from one target basis to another, thus they are not included in the searching procedure.
-There are multiple points in one tries but it doesn't have `loop` value.
+### Explore searching points
 
-### Explore all searching points
-
-As mentioned previously, the largest feature of exploring searching points is its uncertainty in the number per `tries`.
-This makes it difficult to map `id` on the x-axis since the first few tries that can be informative to the evolving of the algorithm occupies a small number of points on the graph while the large less informative search points towards the end occupies the vast majority of the plot space.
+As mentioned previously, the largest difficulties of exploring searching points is its unknown number of observations per `tries`.
+Mapping `id` on the x-axis will leave the `tries` with few observations a small space in the plot, while those `tries` with large number of search points towards the end occupying the vast majority of the space in the plot.
 
 This motivates the use of summarised statistics.
-At each iteration, rather than knowing the value of *every* points, we are more interested to know a general summary of all the points and more importantly, the point with the largest `index_val` since it is the one that prescribe the interpolation and future search.
-Therefore, boxplot is a suitable candidate.
+At each iteration, rather than knowing the index value of *every* points, we are more interested to know a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the geodesic interpolation and future searches.
 
-However, boxplot doesn't tells us the number of point in each box.
-By showing boxplot for all `tries`, we lose information on how long it takes for the search to find the target basis.
-We may still want to show the exact points of every point if there are only a few points.
-Therefore a cutoff point is placed as an argument to decise when to switch from point geometry to boxplot geometry.
+Boxplot is a suitable candidate that provides five points summary of the data, while it has one drawback: it doesn't report the number of point in each box.
+We may risk losing information on how long it takes for the search to find the target basis by displaying the boxplot alone for all `tries`.
+Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry.
+This is achieved via the `cutoff` argument.
+A line geometry is also added to link the points with the largest index value for each `tries`.
+This helps to visualise the improvement made by each `tries`.
 
-This design is easy to implement with the concept of gramma of graphic.
-With the idea of layed gramma, we only need to subset the data and map it onto the either a point geometry or boxplot geometry.
-We can also add a layer with line geometry that connect the points with the largest `index_val` in each tries.
-This would show us which `tries` has the largest improvement.
+*Example: exploring searching points* The data is sourced from a 2D projection of the larger dataset and `search_better` is used with `max.tries = 500`.
+In Figure \ref{fig:points-id} and \ref{fig:points-tries}, a comparison has been made on visualising the searching points with `id` and `tries` on the x-axis, colored by `tries`.
+In Figure \ref{fig:points-id}, the searching points of the first few `tries` are squeezed in a small width leaving the large uninteresting searching points in the last three `tries` taking a vast majority of the plotting space.
+While in Figure \ref{fig:points-tries} the data is spaced by `tries` evenly in the plot.
+Label at the bottom indicates the number of observations in each tries and facilitates the choice of cutoff to switch from point geometry to boxplot geometry (`cutoff = 15`).
+The line geometry suggest the largest improvement happens at `tries = 5`.
 
-A comparison of using scatter plot (all point geometry) and my new design is shown in Figure \ref{fig:points}.
-As we can see the points for the first few tries on hte scatter plot are all squeezed in a tiny region, being nearly unseeable.
-While the plot on the right balances out the points in each `tries`.
-By linking the updating points for each `tries`, we observe that the largest improvement occurs in `tries = 5` where the number of searching points dramastically increased from 11 points in `tries = 4` to 86 points in `tries = 5`.
-
-```{r points, fig.cap="\\label{points} This is xxx"}
+```{r points-id, fig.asp = 0.5, fig.cap = "\\label{points-id}this is a xxx"}
 dt <- holes_2d_better_max_tries %>%
@@
 
-p1 <- dt %>% 
+dt %>% 
   ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
@@
             aes(group = 1)) 
-  #theme(legend.position = "none")
+```
 
-dt_count <- dt %>%
-  group_by(as.factor(tries)) %>%
-  count() %>%
-  rename(tries = `as.factor(tries)`)
+```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
+The left plot doesn't efficiently use the plot space to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
 
-p2 <- dt %>%
-  ggplot(aes(x = as.factor(tries), y = index_val, col = as.factor(tries))) +
-  geom_boxplot(data = dt %>% filter(tries %in% which(dt_count$n > 15))) +
-  geom_point(data = dt%>% filter(tries %in% which(dt_count$n < 15))) +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1)) 
-  #theme(legend.position = "none")
-
-(p1 | p2) & theme(legend.position = "bottom")
+dt %>% explore_trace_search()
 
@@
 
-### Explore only interpolation points
+### Explore interpolating points
 
-While mapping `tries` to the x axis allows for viewing the magnitude of the interpolation and also its density, we can't really see the curvature of the interpolated path.
-Think this would be important when the index is no longer smooth and the interpolation will jump up and down.
-Thus the geom_point + geom_line specification is still preferred.
+Plotting the interpolating points as time series data allows us to diagnose characterstics of different configurations and index functions.
+Here we present two examples of using  plots to diagnose the tour algorithm and different index functions.
 
-*Example 1: Interruption* The left panel of Figure \ref{fig:interruption} presents the trace plot of the interpolated basis using `search_better` to optimise the holes index (2D projection using the larger dataset).
-Tour interpolates from the current basis to the target basis, which has been found to have a higher index value by projection pursuit.
-The target basis will then be sent to the projection pursuit as the current basis to find the next target basis.
-We can observe from the plot that there are basis with even higher index value on the interpolation path.
-These bases could be used to search for new basis in the next round.
-Therefore, an interruption is constructed to only take interpolated basis up to the point where the index value stops increasing and the last basis is taken as the current basis for the next round of search.
-After implementing the interruption, the tracing plot with the same configuration is shown on the right panel.
-Rather than interpolate to the target basis, the interruption interrupt the process after id = 60 and proceeds the next round of search from the interpolated basis.
-Taking advantage of the interpolated basis results in a higher index value in the end with fewer steps.
+```{r interpolation}
+# points is always included in the newest version
+# p1 <- holes_2d_better %>% 
+#   filter(info == "interpolation") %>% 
+#   ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
+#   geom_point() + 
+#   geom_line(aes(group = 1))
+# 
+# p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
+# 
+# (p1 | p2)  &
+#   theme(legend.position = "bottom")
 
-```{r interruption, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without the interruption.
-The interruption stops the interpolation when the index value starts to decrease, see the difference at id being around 60.
+```
+
+*Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
+As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to ahve a higher index value.
+After the interpolation, the target basis will become the current basis and send back to the projection pursuit algorithm to find the next target basis.
+From figure \ref{fig:interruption}, it is possible that there are bases with index value higher than the target basis on the interpolation path and these bases could be used to search for new basis in the next iteration.
+
+Thus an interruption is constructed to accpet the interpolating bases up to the one with the larget index value on the interpolation path, and that basis is taken as the current basis for the next iteration.
+After implementing this interruption, the tracing plot with the same configuration is shown on the lower panel.
+We can observe that rather than interpolating to the target basis at `id = 62`, the interpolation stops before the index value starts to decrease at `id = 60`.
+This implementation results in a higher index value in the end with fewer steps.
+
+```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without the interruption.
+The interruption stops the interpolation when the index value starts to decrease at id = 60.
 The implementation of the interuption finds an ending basis with higher index value using fewer steps.
@@
 p1 | p2
+#xxx up and down plot allows for displaying time series features while left and right display makes it easier for comparison? The same problem for plot of comparing between different searching methods
 ```
 
-*Example 2: Noisy index function* Add an example on noisy function
+*Example: Noisy index function* Add an example on noisy function
 
@@
 
-```{r interpolation}
-# points is always included in the newest version
-p1 <- holes_2d_better %>% 
-  filter(info == "interpolation") %>% 
-  ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
-  geom_point() + 
-  geom_line(aes(group = 1))
+xxx
 
-p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
+Currently, I have three things to put for the noisy index function.
 
-(p1 | p2)  &
-  theme(legend.position = "bottom")
+1) interpolating plot finds that search_geodesic is not doing a good job on optimising the noisy index; while search_better is.
 
-```
+2) The interpolating plot for noisy index function is not smooth but the target basis is always increasing.
+
+3) search_better_random sees to find both local and global maximum.
+
+Need to decise where to put this example. Do we wan to intoduce noisy function at this stage or we put it here because they are insights draw from the interpolating plots.
+
+xxx
 
@@
 
-For now we only include the iterator and `index_val` in the mapping of the plot, while more paramters of interest could be added to explore the effect their effect.
+At previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e.
+color, could be added to compare other parameters in the global object.
+Two examples are shown below to explore and compare different searching methods and neighbourhood parameter alpha.
 
-*Example 3: Polish* In principle, all the optimisation routines should present the same output on the same problem.
-In Figure \ref{fig:trace-compare}, red dots shows the trace of the interpolated basis using `search_geodesic` and `search_better` (with max.tries = 100), respectively, on the 2D projection problem.
-We can observe that they attain slightly different ending index value, which is not ideal.
+*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem.
+Figure \ref{fig:trace-compare} shows the interpolating plots for two different searching methods: search_geodesic and serach_better on a 2D projection problem for the larger dataset using the holes index.
+The two methods reaches different ending value in the red dots, which is not ideal.
 This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
 
-```{r polish}
-p1 <- bind_rows(holes_2d_geo, holes_2d_geo_polish) %>% 
-  explore_trace_interp(col = method) + ggtitle("search_geodesic")
+```{r trace-compare}
+p1 <- holes_2d_geo %>% 
+  explore_trace_interp(col = tries) + ggtitle("search_geodesic") + ylim(0.8, 0.96)
 
-p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
-  explore_trace_interp(col = method) + ggtitle("search_better")
+p2 <- holes_2d_better_max_tries %>% 
+  explore_trace_interp(col = tries) + ggtitle("search_better") + ylim(0.8, 0.96)
 
@@
 
-`search_polish` takes the ending basis of a given search as the starting basis and uses a brutal-force approach to sample a larger number of basis (`n_sample`) in the neightbourhood, controlled by `polish_alpha`.
+`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a larg number of basis (`n_sample`) in the neightbourhood, whose radius is controlled by `polish_alpha`.
 Among the `n_sample` basis, the one with the largest index value becomes the candidate.
-If its index value is larger than that of the starting basis, it becomes the center of the searching neightbourhood in the next round.
-If no basis is found to have larger index value than the starting basis, the searching neighbourhood will shrink and the search continues.
+If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration.
+If no basis is found to have larger index value than the current basis, the searching neighbourhood will be shrunk and the search continues.
 The polishing search ends when one of the four stopping criteria is satisfied:
@@
 1) the two basis can't be too close 
-2) the improvement can't be too small 
+2) the percentage improvement of the index function can't be too small 
 3) the searching neighbourhood can't be too small
-4) the number of iteration can't exceed the max.tries
+4) the number of iteration can't exceed the `max.tries`
 
@@
 
-```{r trace-compare, fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
+```{r polish, fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
 Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}
@@
 
-*Example 4: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in serach_better & search_posse.
+<!-- *Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in serach_better & search_posse. -->
 
 ```{r nrow, echo = TRUE}
-nrow(holes_2d_better_max_tries)
-nrow(holes_2d_pos)
+# nrow(holes_2d_better_max_tries)
+# nrow(holes_2d_pos)
 ```
@@
 ```{r alpha}
-dt <- holes_2d_better_max_tries %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-p1 <- dt %>%
-  ggplot(aes(x = as.factor(tries), y = alpha)) +
-  geom_jitter(aes( col = as.factor(alpha))) +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1)) + 
-  theme(legend.position = "none")
-
-dt <- holes_2d_pos %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-p2 <- dt %>%
-  ggplot(aes(x = as.factor(tries), y = alpha,)) +
-  geom_jitter(aes(col = as.factor(alpha))) +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1))  + 
-  theme(legend.position = "none")
-
-p1 | p2
+# p1 <- holes_2d_better_max_tries %>%
+#   filter(info != "interpolation") %>% 
+#   mutate(id = row_number()) %>% 
+#   explore_trace_parameter(var = alpha) +
+#   ylim(0.46, 0.505) + ggtitle("search_better")
+# 
+# p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+# 
+# p2 <- holes_2d_pos %>%
+#   filter(info != "interpolation") %>% 
+#   mutate(id = row_number()) %>% 
+#   explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")
+# 
+# p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+# 
+# (p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
 ```
@@
 
-The variable or object to explore could sometimes be a more than a scale, for example, in a tour, we would also be interested to explore the relative position of the projection basis in the vector space.
-This imposes difficulties in visualisation since as human, we are bounded to preceive at most three dimensions.
-Dimension reduction is reuqired to summarise the projection matrix to a managable number of variable to map to a plot.
-Principal copmonent analysis is used to extract the direction that reflects most variation in the data.
+Matrix parameter could also be interested to explore, for example, in the tour, we are interested to explore the relative position of the projection basis in the vector space.
+A projection basis is a vector/matrix whose row number is the number of variable in the given dataset.
+This imposes difficulties in visualisation since we are bounded to preceive at most three dimensions.
+Thus, principal component analysis is used to reduce the dimension of the projection basi and the first two principal component are mapped to the x and y axis of the plot.
+Another variable of interest could be mapped to the color aesthetics to see how it changes as the projection basis changes in the projected 2D space.
+[have a second read here]
 
-*Example 5 PCA* refine the example of matrix parameter exploration
+*Example: understand search_geodesic via mapping `info` to color* `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm works.
+Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
+The best one is picked and a line search is then run on the geodesic direction to find the target basis.
+The starting and target basis are then interpolated.
+In the next iteration, the target basis becomes the current basis and then procedures continues.
+[should probably reword this part with info levels xxx]
 
-```{r pca, echo = TRUE}
+```{r pca, fig.cap = "\\label{pca}this is xxx"}
 holes_1d_geo %>% explore_proj_pca(col = info)
+
+#xxx the info level need to be reorder to show the start and directional search points better.
 ```
+
+*Example: understand the polishing alpha via mapping alpha to color* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhod at each loop and an appropriate inital value in neighbourhood parameter alpha would avoid search on large vector space and concentrate on the ending basis.
+In Figure, two different initial alpha value is used in the polishing step.
+[some more illustration of the comparison of the two polish alpha here]
 

### 2020-05-26 paper/paper.tex
reword explore scalar param section and part of matrix param

@@
 The most interesting parameter to explore is the value of objective
-function. Given the structure of projection pursuit guided tour
-algorithm, whatever searching method we use, we could roughly divide the
-observation into three category: searching, updating and interpolating
-points.
+function. The points recorded in the global object can be divided into
+two broad categories: searching points and interpolating points
 
@@
 \item
-  \emph{Searching points} include the observations that ones recorded
-  during the trials and errors in the searching algorithm to find the
-  target basis. In one \texttt{tries}, there could be only a few
-  searching points while towards the end of the algorithm, they can
-  easily go to hundreds.
+  \emph{Searching points} include the observations that are recorded in
+  the searching algorithm in order to find the target basis. The points
+  for target bases is also included in the searching points and there is
+  one such point per \texttt{tries}.
 \item
-  \emph{Updating points} are the observations for the target points and
-  there are only one points per \texttt{tries}. We use ``all searching
-  points'' to refer to the collection of both searching and updating
-  points.
-\item
-  \emph{interpolation points} exist specifically in the guided tour in
-  order to provide continuous animated view from one target basis to
-  another, thus they are not included in the searching procedure. There
-  are multiple points in one tries but it doesn't have \texttt{loop}
-  value.
+  \emph{interpolating points} exist in the guided tour to produce
+  continuous animated view from one target basis to another and it
+  doesn't have \texttt{loop} value.
 \end{itemize}
 
-\hypertarget{explore-all-searching-points}{%
-\subsubsection{Explore all searching
-points}\label{explore-all-searching-points}}
+\hypertarget{explore-searching-points}{%
+\subsubsection{Explore searching
+points}\label{explore-searching-points}}
 
-As mentioned previously, the largest feature of exploring searching
-points is its uncertainty in the number per \texttt{tries}. This makes
-it difficult to map \texttt{id} on the x-axis since the first few tries
-that can be informative to the evolving of the algorithm occupies a
-small number of points on the graph while the large less informative
-search points towards the end occupies the vast majority of the plot
-space.
+As mentioned previously, the largest difficulties of exploring searching
+points is its unknown number of observations per \texttt{tries}. Mapping
+\texttt{id} on the x-axis will leave the \texttt{tries} with few
+observations a small space in the plot, while those \texttt{tries} with
+large number of search points towards the end occupying the vast
+majority of the space in the plot.
 
 This motivates the use of summarised statistics. At each iteration,
-rather than knowing the value of \emph{every} points, we are more
+rather than knowing the index value of \emph{every} points, we are more
 interested to know a general summary of all the points and more
-importantly, the point with the largest \texttt{index\_val} since it is
-the one that prescribe the interpolation and future search. Therefore,
-boxplot is a suitable candidate.
+importantly, the point with the largest \texttt{index\_val} since it
+prescribes the geodesic interpolation and future searches.
 
-However, boxplot doesn't tells us the number of point in each box. By
-showing boxplot for all \texttt{tries}, we lose information on how long
-it takes for the search to find the target basis. We may still want to
-show the exact points of every point if there are only a few points.
-Therefore a cutoff point is placed as an argument to decise when to
-switch from point geometry to boxplot geometry.
+Boxplot is a suitable candidate that provides five points summary of the
+data, while it has one drawback: it doesn't report the number of point
+in each box. We may risk losing information on how long it takes for the
+search to find the target basis by displaying the boxplot alone for all
+\texttt{tries}. Thus, the number of point for each \texttt{tries} is
+displayed at the bottom of each box and we provide options to switch
+\texttt{tries} with small number of points to a point geometry. This is
+achieved via the \texttt{cutoff} argument. A line geometry is also added
+to link the points with the largest index value for each \texttt{tries}.
+This helps to visualise the improvement made by each \texttt{tries}.
 
-This design is easy to implement with the concept of gramma of graphic.
-With the idea of layed gramma, we only need to subset the data and map
-it onto the either a point geometry or boxplot geometry. We can also add
-a layer with line geometry that connect the points with the largest
-\texttt{index\_val} in each tries. This would show us which
-\texttt{tries} has the largest improvement.
-
-A comparison of using scatter plot (all point geometry) and my new
-design is shown in Figure \ref{fig:points}. As we can see the points for
-the first few tries on hte scatter plot are all squeezed in a tiny
-region, being nearly unseeable. While the plot on the right balances out
-the points in each \texttt{tries}. By linking the updating points for
-each \texttt{tries}, we observe that the largest improvement occurs in
-\texttt{tries\ =\ 5} where the number of searching points dramastically
-increased from 11 points in \texttt{tries\ =\ 4} to 86 points in
+\emph{Example: exploring searching points} The data is sourced from a 2D
+projection of the larger dataset and \texttt{search\_better} is used
+with \texttt{max.tries\ =\ 500}. In Figure \ref{fig:points-id} and
+\ref{fig:points-tries}, a comparison has been made on visualising the
+searching points with \texttt{id} and \texttt{tries} on the x-axis,
+colored by \texttt{tries}. In Figure \ref{fig:points-id}, the searching
+points of the first few \texttt{tries} are squeezed in a small width
+leaving the large uninteresting searching points in the last three
+\texttt{tries} taking a vast majority of the plotting space. While in
+Figure \ref{fig:points-tries} the data is spaced by \texttt{tries}
+evenly in the plot. Label at the bottom indicates the number of
+observations in each tries and facilitates the choice of cutoff to
+switch from point geometry to boxplot geometry (\texttt{cutoff\ =\ 15}).
+The line geometry suggest the largest improvement happens at
 \texttt{tries\ =\ 5}.
@@
 \centering
-\includegraphics{paper_files/figure-latex/points-1.pdf}
-\caption{\label{points} This is xxx}
+\includegraphics{paper_files/figure-latex/points-id-1.pdf}
+\caption{\label{points-id}this is a xxx}
 \end{figure}
 
-\hypertarget{explore-only-interpolation-points}{%
-\subsubsection{Explore only interpolation
-points}\label{explore-only-interpolation-points}}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/points-tries-1.pdf}
+\caption{\label{points}A comparison of plotting the same search points
+with different plot designs. The left plot doesn't efficiently use the
+plot space to convey information from the plot while the right plot
+provides good summarisation of data and number of points in each tries.}
+\end{figure}
 
-While mapping \texttt{tries} to the x axis allows for viewing the
-magnitude of the interpolation and also its density, we can't really see
-the curvature of the interpolated path. Think this would be important
-when the index is no longer smooth and the interpolation will jump up
-and down. Thus the geom\_point + geom\_line specification is still
-preferred.
+\hypertarget{explore-interpolating-points}{%
+\subsubsection{Explore interpolating
+points}\label{explore-interpolating-points}}
 
-\emph{Example 1: Interruption} The left panel of Figure
-\ref{fig:interruption} presents the trace plot of the interpolated basis
-using \texttt{search\_better} to optimise the holes index (2D projection
-using the larger dataset). Tour interpolates from the current basis to
-the target basis, which has been found to have a higher index value by
-projection pursuit. The target basis will then be sent to the projection
-pursuit as the current basis to find the next target basis. We can
-observe from the plot that there are basis with even higher index value
-on the interpolation path. These bases could be used to search for new
-basis in the next round. Therefore, an interruption is constructed to
-only take interpolated basis up to the point where the index value stops
-increasing and the last basis is taken as the current basis for the next
-round of search. After implementing the interruption, the tracing plot
-with the same configuration is shown on the right panel. Rather than
-interpolate to the target basis, the interruption interrupt the process
-after id = 60 and proceeds the next round of search from the
-interpolated basis. Taking advantage of the interpolated basis results
-in a higher index value in the end with fewer steps.
+Plotting the interpolating points as time series data allows us to
+diagnose characterstics of different configurations and index functions.
+Here we present two examples of using plots to diagnose the tour
+algorithm and different index functions.
+
+\emph{Example: Interruption} This examples uses \texttt{search\_better}
+for a 2D projection on the larger dataset using the \texttt{holes}
+index. As mentioned previously, the interpolation starts from the
+current basis to the target basis, which has been found by the
+projection pursuit algorithm to ahve a higher index value. After the
+interpolation, the target basis will become the current basis and send
+back to the projection pursuit algorithm to find the next target basis.
+From figure \ref{fig:interruption}, it is possible that there are bases
+with index value higher than the target basis on the interpolation path
+and these bases could be used to search for new basis in the next
+iteration.
+
+Thus an interruption is constructed to accpet the interpolating bases up
+to the one with the larget index value on the interpolation path, and
+that basis is taken as the current basis for the next iteration. After
+implementing this interruption, the tracing plot with the same
+configuration is shown on the lower panel. We can observe that rather
+than interpolating to the target basis at \texttt{id\ =\ 62}, the
+interpolation stops before the index value starts to decrease at
+\texttt{id\ =\ 60}. This implementation results in a higher index value
+in the end with fewer steps.
 
@@
 and without the interruption. The interruption stops the interpolation
-when the index value starts to decrease, see the difference at id being
-around 60. The implementation of the interuption finds an ending basis
-with higher index value using fewer steps.}
+when the index value starts to decrease at id = 60. The implementation
+of the interuption finds an ending basis with higher index value using
+fewer steps.}
 \end{figure}
 
-\emph{Example 2: Noisy index function} Add an example on noisy function
+\emph{Example: Noisy index function} Add an example on noisy function
 
-\includegraphics{paper_files/figure-latex/interpolation-1.pdf}
+xxx
+
+Currently, I have three things to put for the noisy index function.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  interpolating plot finds that search\_geodesic is not doing a good job
+  on optimising the noisy index; while search\_better is.
+\item
+  The interpolating plot for noisy index function is not smooth but the
+  target basis is always increasing.
+\item
+  search\_better\_random sees to find both local and global maximum.
+\end{enumerate}
+
+Need to decise where to put this example. Do we wan to intoduce noisy
+function at this stage or we put it here because they are insights draw
+from the interpolating plots.
+
+xxx
 
@@
 
-For now we only include the iterator and \texttt{index\_val} in the
-mapping of the plot, while more paramters of interest could be added to
-explore the effect their effect.
+At previous two sections, only the iterator and the index value are
+mapped onto the x and y aesthetics of the plot; while more aesthetics
+i.e.~color, could be added to compare other parameters in the global
+object. Two examples are shown below to explore and compare different
+searching methods and neighbourhood parameter alpha.
 
-\emph{Example 3: Polish} In principle, all the optimisation routines
-should present the same output on the same problem. In Figure
-\ref{fig:trace-compare}, red dots shows the trace of the interpolated
-basis using \texttt{search\_geodesic} and \texttt{search\_better} (with
-max.tries = 100), respectively, on the 2D projection problem. We can
-observe that they attain slightly different ending index value, which is
-not ideal. This motivates the creation of a polishing search that
-polishes the ending basis and achieves unity on different methods.
+\emph{Example: Polish} In principle, all the optimisation routines
+should result in the same output on the same problem. Figure
+\ref{fig:trace-compare} shows the interpolating plots for two different
+searching methods: search\_geodesic and serach\_better on a 2D
+projection problem for the larger dataset using the holes index. The two
+methods reaches different ending value in the red dots, which is not
+ideal. This motivates the creation of a polishing search that polishes
+the ending basis and achieves unity on different methods.
 
-\includegraphics{paper_files/figure-latex/polish-1.pdf}
+\includegraphics{paper_files/figure-latex/trace-compare-1.pdf}
 
 \texttt{search\_polish} takes the ending basis of a given search as the
-starting basis and uses a brutal-force approach to sample a larger
-number of basis (\texttt{n\_sample}) in the neightbourhood, controlled
-by \texttt{polish\_alpha}. Among the \texttt{n\_sample} basis, the one
-with the largest index value becomes the candidate. If its index value
-is larger than that of the starting basis, it becomes the center of the
-searching neightbourhood in the next round. If no basis is found to have
-larger index value than the starting basis, the searching neighbourhood
-will shrink and the search continues. The polishing search ends when one
-of the four stopping criteria is satisfied:
+current basis and uses a brutal-force approach to sample a larg number
+of basis (\texttt{n\_sample}) in the neightbourhood, whose radius is
+controlled by \texttt{polish\_alpha}. Among the \texttt{n\_sample}
+basis, the one with the largest index value becomes the candidate. If
+its index value of the candidate basis is larger than that of the
+current basis, it becomes the current basis in the next iteration. If no
+basis is found to have larger index value than the current basis, the
+searching neighbourhood will be shrunk and the search continues. The
+polishing search ends when one of the four stopping criteria is
+satisfied:
 
@@
 \item
-  the improvement can't be too small
+  the percentage improvement of the index function can't be too small
 \item
@@
 \item
-  the number of iteration can't exceed the max.tries
+  the number of iteration can't exceed the \texttt{max.tries}
 \end{enumerate}
@@
 \centering
-\includegraphics{paper_files/figure-latex/trace-compare-1.pdf}
+\includegraphics{paper_files/figure-latex/polish-1.pdf}
 \caption{\label{trace-compare}Breakdown of index value when using
@@
 
-\emph{Example 4: The neighbourhood parameter alpha} Add an example on
-comparing the neighbourhood parameter in serach\_better \&
-search\_posse.
-
 \begin{Shaded}
 \begin{Highlighting}[]
-\KeywordTok{nrow}\NormalTok{(holes_2d_better_max_tries)}
+\CommentTok{# nrow(holes_2d_better_max_tries)}
+\CommentTok{# nrow(holes_2d_pos)}
 \end{Highlighting}
 \end{Shaded}
-
-\begin{verbatim}
-## [1] 1529
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{nrow}\NormalTok{(holes_2d_pos)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] 1056
-\end{verbatim}
-
-\includegraphics{paper_files/figure-latex/alpha-1.pdf}
 
@@
 
-The variable or object to explore could sometimes be a more than a
-scale, for example, in a tour, we would also be interested to explore
-the relative position of the projection basis in the vector space. This
-imposes difficulties in visualisation since as human, we are bounded to
-preceive at most three dimensions. Dimension reduction is reuqired to
-summarise the projection matrix to a managable number of variable to map
-to a plot. Principal copmonent analysis is used to extract the direction
-that reflects most variation in the data.
+Matrix parameter could also be interested to explore, for example, in
+the tour, we are interested to explore the relative position of the
+projection basis in the vector space. A projection basis is a
+vector/matrix whose row number is the number of variable in the given
+dataset. This imposes difficulties in visualisation since we are bounded
+to preceive at most three dimensions. Thus, principal component analysis
+is used to reduce the dimension of the projection basi and the first two
+principal component are mapped to the x and y axis of the plot. Another
+variable of interest could be mapped to the color aesthetics to see how
+it changes as the projection basis changes in the projected 2D space.
+{[}have a second read here{]}
 
-\emph{Example 5 PCA} refine the example of matrix parameter exploration
+\emph{Example: understand search\_geodesic via mapping \texttt{info} to
+color} \texttt{search\_geodesic} is a two-stage ascending algorithm with
+four different stages in the search and a PCA plot useful to understand
+how the algorithm works. Starting from the start basis, a directional
+search is conducted in a narrow neighbourhood on five random directions.
+The best one is picked and a line search is then run on the geodesic
+direction to find the target basis. The starting and target basis are
+then interpolated. In the next iteration, the target basis becomes the
+current basis and then procedures continues. {[}should probably reword
+this part with info levels xxx{]}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ info) }
-\end{Highlighting}
-\end{Shaded}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/pca-1.pdf}
+\caption{\label{pca}this is xxx}
+\end{figure}
 
-\includegraphics{paper_files/figure-latex/pca-1.pdf}
+\emph{Example: understand the polishing alpha via mapping alpha to
+color} \texttt{search\_polish} is a brutal-force algorithm that evaluate
+1000 points in the neighbourhod at each loop and an appropriate inital
+value in neighbourhood parameter alpha would avoid search on large
+vector space and concentrate on the ending basis. In Figure, two
+different initial alpha value is used in the polishing step. {[}some
+more illustration of the comparison of the two polish alpha here{]}
 

### 2020-05-31 source/explore_basis.Rmd
update explore_basis

@@
 load(here::here("data/holes_1d_geo_polish.rda"))
+load(here::here("data/holes_1d_geo_polish_default_alpha.rda"))
 load(here::here("data/holes_2d_geo.rda"))
@@
 ```{r}
-holes_1d_geo %>% explore_proj_pca() 
+holes_1d_geo %>% explore_proj_pca(col = info) 
 ```
@@
 ```{r}
-glb_obj <- bind_rows(holes_1d_geo, holes_1d_geo_polish)
-
-num_col <- ncol(glb_obj$basis[[1]])
-num_row <- nrow(glb_obj$basis[[1]])
-
-basis <- purrr::flatten_dbl(glb_obj$basis)  %>% matrix(ncol = num_row * num_col, byrow = TRUE)
-
-loadings <- basis %>%
-  stats::prcomp() %>%
-  stats::predict() %>%
-  tibble::as_tibble()
-
-combined <- tibble::as_tibble(basis) %>%
-  dplyr::bind_cols(loadings) %>%
-  dplyr::bind_cols(glb_obj %>% dplyr::select(-basis))
-
-combined %>% 
-  ggplot(aes(x = PC1, y = PC2, col = method)) +
-  geom_point() +
-  geom_point(data = combined %>% filter(method == "search_geodesic")) +
-  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
-  theme(aspect.ratio = 1)
+bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
+  explore_proj_pca(col = method)
 ```
@@
 ```{r}
-combined %>% 
-  ggplot(aes(x = PC1, y = PC2, col = as.factor(alpha))) +
-  geom_point() +
-  geom_point(data = combined %>% filter(method == "search_geodesic")) +
-  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
-  theme(aspect.ratio = 1)
+bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
+  explore_proj_pca(col = as.factor(alpha))
 ```
@@
 ```{r}
-combined %>% filter(alpha < 0.03) %>% 
-  ggplot(aes(x = PC1, y = PC2, col = as.factor(alpha))) +
-  geom_point() +
-  geom_point(data = combined %>% filter(method == "search_geodesic")) +
-  #stat_density(aes(fill = after_stat(nlevel)), geom = "polygon", alpha = 0.5) +
-  theme(aspect.ratio = 1)
-```
-
-```{r}
-combined %>% filter(info == "polish_best") %>% dplyr::select(index_val, alpha, loop)
+bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
+  explore_proj_pca(col = as.factor(alpha))
 ```

### 2020-05-31 paper/paper.Rmd
add noisy index function section; add polish alpha for search_goedesic section

@@
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
+
+source(here::here("source/indice.r"))
+source(here::here("source/sim_data.r"))
 ```
@@
 
-Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-diemsion projection of multivariate data.
+Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-dimension projection of multivariate data.
 A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine.
@@
 Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path.
-An illustration modified from [@buja2005computational] entails the geodesic interpolation bewteen the plane generated by projection pursuit.
+An illustration modified from [@buja2005computational] entails the geodesic interpolation between the plane generated by projection pursuit.
 
@@
 
-As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the com- putational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe.
+As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the computational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe.
 An ideal optimisation procedure needs to have following characteristics:
 
-- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentible
+- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentiable
 
-- *Being able to optimise with constraints*: the projection matrix is restricted to an orthornormal matrix.
+- *Being able to optimise with constraints*: the projection matrix is restricted to an orthonormal matrix.
 
@@
 @posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis.
-The neighbourhood is defined via the radius of the p-dimenionsal sphere, $c$.
+The neighbourhood is defined via the radius of the p-dimensional sphere, $c$.
 The new basis is taken as the target basis if it has higher index value, or the sampling continues.
-If no basis is found to have higher index value after a certain humber of tries $n$, the radius $c$ is halved.
-The algorithm stops when the maximum numbere of iteration is attained or the radius $c$ is less than a pre-determined number.
-[Pursuit package uses this method and it works greate!
+If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved.
+The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.
+[Pursuit package uses this method and it works great!
 But I don't think we implement this - although don't think it is too hard to do it].
@@
 @cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
-Since some indices could be non-differentiable, the computation of derivative is replaced by a psudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
+Since some indices could be non-differentiable, the computation of derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
 Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
@@
 
-*Think about how does your pacakge help people to understand optimisation*
+*Think about how does your package help people to understand optimisation*
 
  - diagnostic on stochastic optim
- - vis the prograssion of multi-parameter decision variable 
- - understanding learning rate - neightbourhood parameter
+ - vis the progression of multi-parameter decision variable 
+ - understanding learning rate - neighbourhood parameter
  - understand where the local & global maximum is found - trace plot - see if noisy function
@@
 
-The psudo-code below illustrates the implementation of guided tour in the tourr package.
+The pseudo-code below illustrates the implementation of guided tour in the tourr package.
 Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
@@
 
-Visualisatioin has been widely used for exploring and understanding data.
-Visualisation presents information in a graphicial manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation.
+Visualisation has been widely used for exploring and understanding data.
+Visualisation presents information in a graphical manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation.
 Visual diagnostics can be real-time or post-run.
 Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
-This section focuses on the definition and production of post-run diagnoistics and the next section discusses real-time diagnostics.
+This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.
 
@@
 
-The graphical system in R is established based on the conception gramma of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.
+The graphical system in R is established based on the conception grammar of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.
 A layer includes 1) the dataset that powers the plot; 2) a geometric object that represents the visual shape of the plot and 3) relevant statistical transformation if needed.
-An important concept in the gramma of graphic is *aesthetic mapping*.
+An important concept in the grammar of graphic is *aesthetic mapping*.
 Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object.
@@
 
-The concept of gramma of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
+The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
 The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
@@
 
-In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randonly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[needa rephrase this part].**
+In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**
 
-Two set of simulated data are used in the demonstration of the visualiation and diagnostics of the tour optimisation.
+Two set of simulated data are used in the demonstration of the visualisation and diagnostics of the tour optimisation.
 A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`).
-`x2` is the informative variable simulated from two bi-modal normal distribution centered at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`.
+`x2` is the informative variable simulated from two bi-modal normal distribution centred at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`.
 The data has been scaled to ensure `x2` has variance of 1.
@@
 Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure.
-The global structure consists of six columnes: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process.
-The example below presents the global object of a 1D projection of the small dataset with geodesic seraching method.
+The global structure consists of six columns: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process.
+The example below presents the global object of a 1D projection of the small dataset with geodesic searching method.
 
@@
 The parameter could be could be the decision variable, the value of the optimisation function or other parameter that facilitate the algorithm.
-In projectio pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`).
+In projection pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`).
 These are the variables that will be mapped into the diagnostic plot.
 The global object also has its time series nature since the observations are recorded as the algorithm progresses and thus can't be switch in rows.
-This time series structure is always hierachical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm.
+This time series structure is always hierarchical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm.
 Take our example, in the global object, each row has an `id` label which is the smallest unit the observation is ordered.
-A larger ordering unit is `tries` since it is updated everytime a new target basis is found.
+A larger ordering unit is `tries` since it is updated every time a new target basis is found.
 A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`.
@@
 
-```{r points-id, fig.asp = 0.5, fig.cap = "\\label{points-id}this is a xxx"}
+```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
+The left plot doesn't efficiently use the plot space to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
+
 dt <- holes_2d_better_max_tries %>%
@@
   mutate(id = row_number())
-
-dt %>% 
-  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
-  geom_point()  +
-  geom_line(data = dt %>% group_by(tries) %>% filter(index_val == max(index_val)),
-            aes(group = 1)) 
-```
-
-```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
-The left plot doesn't efficiently use the plot space to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
 
@@
 
-Plotting the interpolating points as time series data allows us to diagnose characterstics of different configurations and index functions.
+Plotting the interpolating points as time series data allows us to diagnose characteristics of different configurations and index functions.
 Here we present two examples of using  plots to diagnose the tour algorithm and different index functions.
@@
 *Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
-As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to ahve a higher index value.
+As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to have a higher index value.
 After the interpolation, the target basis will become the current basis and send back to the projection pursuit algorithm to find the next target basis.
@@
 
-Thus an interruption is constructed to accpet the interpolating bases up to the one with the larget index value on the interpolation path, and that basis is taken as the current basis for the next iteration.
+Thus an interruption is constructed to accept the interpolating bases up to the one with the larger index value on the interpolation path, and that basis is taken as the current basis for the next iteration.
 After implementing this interruption, the tracing plot with the same configuration is shown on the lower panel.
@@
 
-*Example: Noisy index function* Add an example on noisy function
+*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{fig:interruption}, is smooth, while this may not be the case for more complicated index function.
+`kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF).
+Several visualisations below shows some characteristic of this index function.
 
-<!-- The tracing plot is useful to explore the index value. Given a global object created, `explore_trace_all()` plots the index value of all the bases.  -->
+Figure \ref{fig:geodesic-better} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better`.
+Rather than a smooth interpolation, the interpolation path for `kol_cdf` shows a zig-zag pattern.
+Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method.
 
-xxx
+```{r geodesic-better,  fig.asp=0.5, fig.cap = "Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisy index: kol_cdf.
+The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
+p1 <- bind_rows(kol_cdf_1d_geodesic, kol_cdf_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
 
-Currently, I have three things to put for the noisy index function.
+p2 <- bind_rows(kol_cdf_1d_better, kol_cdf_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
 
-1) interpolating plot finds that search_geodesic is not doing a good job on optimising the noisy index; while search_better is.
+(p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
+```
 
-2) The interpolating plot for noisy index function is not smooth but the target basis is always increasing.
+Moving from a 1D projection on a dataset with only one informative variable, we now introduce another informative variable in the data and remain 1D projection.
+In this problem, there are two local maximum being `r calc_kol_cdf(matrix(c(0, 1, 0 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 1, 0, 0, 0, 0]$ and `r calc_kol_cdf(matrix(c(0, 0, 1 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 0, 1 ,0, 0, 0]$.
+As in Figure \ref{fig:1d-2var-different-seeds}, using `search_better` with different seeds, both local maximum can be found.
+A simulated annealing algorithm, `search_better_random` in Figure \ref{fig:1d-2var-better-random}, can always find the global maximum at a cost of a large number of `tries`.
 
-3) search_better_random sees to find both local and global maximum.
+```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
+Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
+p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp() + ggtitle("search_better, seed 123456")
+p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp() +  ggtitle("search_better, seed 12345")
 
-Need to decise where to put this example. Do we wan to intoduce noisy function at this stage or we put it here because they are insights draw from the interpolating plots.
+(p5 | p6) & theme(legend.position = "none") & ylim(0, 0.25)
+```
 
-xxx
+```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
+kol_cdf_1d_2var_better_random %>% 
+  explore_trace_interp() + 
+  ylim(0, 0.25) + 
+  theme(legend.position = "none") + 
+  ggtitle("search_better_random")
+```
 
@@
 
-*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem.
-Figure \ref{fig:trace-compare} shows the interpolating plots for two different searching methods: search_geodesic and serach_better on a 2D projection problem for the larger dataset using the holes index.
-The two methods reaches different ending value in the red dots, which is not ideal.
+*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem while this may not be the case in real application.
 This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
 
-```{r trace-compare}
-p1 <- holes_2d_geo %>% 
-  explore_trace_interp(col = tries) + ggtitle("search_geodesic") + ylim(0.8, 0.96)
-
-p2 <- holes_2d_better_max_tries %>% 
-  explore_trace_interp(col = tries) + ggtitle("search_better") + ylim(0.8, 0.96)
-
-(p1 | p2) &
-  theme(legend.position = "none") 
-```
-
-`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a larg number of basis (`n_sample`) in the neightbourhood, whose radius is controlled by `polish_alpha`.
+`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`.
 Among the `n_sample` basis, the one with the largest index value becomes the candidate.
@@
 
-```{r polish, fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
+```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
 Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}
@@
 
-<!-- *Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in serach_better & search_posse. -->
+*Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in search_better & search_posse.
 
@@
 
-```{r alpha}
-# p1 <- holes_2d_better_max_tries %>%
-#   filter(info != "interpolation") %>% 
-#   mutate(id = row_number()) %>% 
-#   explore_trace_parameter(var = alpha) +
-#   ylim(0.46, 0.505) + ggtitle("search_better")
-# 
-# p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
-# 
-# p2 <- holes_2d_pos %>%
-#   filter(info != "interpolation") %>% 
-#   mutate(id = row_number()) %>% 
-#   explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")
-# 
-# p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
-# 
-# (p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
+```{r alpha, fig.asp=0.5,eval = FALSE}
+p1 <- holes_2d_better_max_tries %>%
+  filter(info != "interpolation") %>%
+  mutate(id = row_number()) %>%
+  explore_trace_parameter(var = alpha) +
+  ylim(0.46, 0.505) + ggtitle("search_better")
+
+p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+
+p2 <- holes_2d_pos %>%
+  filter(info != "interpolation") %>%
+  mutate(id = row_number()) %>%
+  explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")
+
+p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+
+(p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
 ```
@@
 A projection basis is a vector/matrix whose row number is the number of variable in the given dataset.
-This imposes difficulties in visualisation since we are bounded to preceive at most three dimensions.
-Thus, principal component analysis is used to reduce the dimension of the projection basi and the first two principal component are mapped to the x and y axis of the plot.
+This imposes difficulties in visualisation since we are bounded to perceive at most three dimensions.
+Thus, principal component analysis is used to reduce the dimension of the projection basis and the first two principal component are mapped to the x and y axis of the plot.
 Another variable of interest could be mapped to the color aesthetics to see how it changes as the projection basis changes in the projected 2D space.
@@
 
-```{r pca, fig.cap = "\\label{pca}this is xxx"}
-holes_1d_geo %>% explore_proj_pca(col = info)
+```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic Coloring by info allows for better understanding of each stage in the geodesic search"}
+
+holes_1d_geo %>% explore_proj_pca(col = info) + 
+  theme(legend.position = "bottom")
 
@@
 
-*Example: understand the polishing alpha via mapping alpha to color* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhod at each loop and an appropriate inital value in neighbourhood parameter alpha would avoid search on large vector space and concentrate on the ending basis.
-In Figure, two different initial alpha value is used in the polishing step.
-[some more illustration of the comparison of the two polish alpha here]
+*Example: initial value for polishing alpha in search geodesic* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
+Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value.
+A default initial value for polishing step is 0.5 and we are interested in whether this is an appropriate initial value.
+The problem is a 1D projection of the small dataset using `search_geodesic` and followed by `search_polish`.
+Figure \ref{fig:polish-alpha} display all projection bases on the first two principal components, colored by the `polish_alpha` parameter in the polishing step.
+Rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, it searches a much larger vector space that are unnecessary.
+Thus a user-supplied smaller initial value for `polish_alpha` would be ideal.
+
+On the right panel of Figure \ref{fig:polish-alpha-tracing}, the `polish_alpha` is initialised as the projection distance between the last two target bases and we observe a concentrated searching space near the ending basis.
+With the supplied initial value for `polish_alpha`, the algorithm reaches the same ending index value with fewer iterations.
+
+```{r polish-alpha, fig.cap = "PCA plot of two different polish alpha initialisations.
+A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis."}
+p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
+  explore_proj_pca(col = as.factor(alpha))
+
+p2 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
+  explore_proj_pca(col = as.factor(alpha))
+
+p3 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>%
+  explore_trace_interp(col = method) + ylim(0.75, 0.95)
+
+p4 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
+  explore_trace_interp(col = method) +  ylim(0.75, 0.95)
+
+(p1 | p2) & theme(legend.position = "none")
+
+```
+
+```{r polish-alpha-tracing, fig.asp = 0.5,fig.cap="Comparing the trace of two different polish alpha initialisations. Both intialisations reach the same ending index values "}
+(p3 | p4) & theme(legend.position = "none")
+
+```
 
@@
 ```{r pca-animated, echo = TRUE, eval = FALSE}
-holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info)
+holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) + 
+  theme(legend.position = "bottom")
 ```

### 2020-05-31 paper/paper.tex
add noisy index function section; add polish alpha for search_goedesic section

@@
 coined by \citet{friedman1974projection} to detect ``interesting''
-low-diemsion projection of multivariate data. A projection pursuit
+low-dimension projection of multivariate data. A projection pursuit
 requires the definition of a projection pursuit index function and an
@@
 from \citep{buja2005computational} entails the geodesic interpolation
-bewteen the plane generated by projection pursuit.
+between the plane generated by projection pursuit.
 
@@
 maximizing the (one- and two-dimensional) projection index strongly
-influences both the statistical and the com- putational aspects of the
+influences both the statistical and the computational aspects of the
 procedure.'' The quality of the optimisation procedure largely affect
@@
   \emph{Being able to handle non-differentiable index function}: the
-  index function could be noisy and non-differentible
+  index function could be noisy and non-differentiable
 \item
   \emph{Being able to optimise with constraints}: the projection matrix
-  is restricted to an orthornormal matrix.
+  is restricted to an orthonormal matrix.
 \item
@@
 samples new basis in the neighbourhood of the current basis. The
-neighbourhood is defined via the radius of the p-dimenionsal sphere,
+neighbourhood is defined via the radius of the p-dimensional sphere,
 \(c\). The new basis is taken as the target basis if it has higher index
 value, or the sampling continues. If no basis is found to have higher
-index value after a certain humber of tries \(n\), the radius \(c\) is
-halved. The algorithm stops when the maximum numbere of iteration is
+index value after a certain number of tries \(n\), the radius \(c\) is
+halved. The algorithm stops when the maximum number of iteration is
 attained or the radius \(c\) is less than a pre-determined number.
-{[}Pursuit package uses this method and it works greate! But I don't
+{[}Pursuit package uses this method and it works great! But I don't
 think we implement this - although don't think it is too hard to do
@@
 and differentiable. Since some indices could be non-differentiable, the
-computation of derivative is replaced by a psudo-derivative of
+computation of derivative is replaced by a pseudo-derivative of
 evaluating five randomly generated directions in a tiny nearby
@@
 
-\emph{Think about how does your pacakge help people to understand
+\emph{Think about how does your package help people to understand
 optimisation}
@@
 \item
-  vis the prograssion of multi-parameter decision variable
+  vis the progression of multi-parameter decision variable
 \item
-  understanding learning rate - neightbourhood parameter
+  understanding learning rate - neighbourhood parameter
 \item
@@
 
-The psudo-code below illustrates the implementation of guided tour in
+The pseudo-code below illustrates the implementation of guided tour in
 the tourr package. Given an projection pursuit index function and a
@@
 
-Visualisatioin has been widely used for exploring and understanding
-data. Visualisation presents information in a graphicial manner and
-often allows people to see information they would otherwise not seen in
-the reporting of numerical summarisation. Visual diagnostics can be
+Visualisation has been widely used for exploring and understanding data.
+Visualisation presents information in a graphical manner and often
+allows people to see information they would otherwise not seen in the
+reporting of numerical summarisation. Visual diagnostics can be
 real-time or post-run. Real time diagnostic directly uses the data
@@
 doesn't need to store the data. This section focuses on the definition
-and production of post-run diagnoistics and the next section discusses
+and production of post-run diagnostics and the next section discusses
 real-time diagnostics.
@@
 
-The graphical system in R is established based on the conception gramma
+The graphical system in R is established based on the conception grammar
 of graphic \citep{wickham2010layered}, where a graphic is defined using
@@
 shape of the plot and 3) relevant statistical transformation if needed.
-An important concept in the gramma of graphic is \emph{aesthetic
+An important concept in the grammar of graphic is \emph{aesthetic
 mapping}. Aesthetic mapping links the variable in a dataset to
@@
 
-The concept of gramma of graphic requires a dataset to be supplied in
+The concept of grammar of graphic requires a dataset to be supplied in
 making a plot, thus a global object needs to be created in the iterative
@@
 target basis generated by the projection pursuit can be accessed later
-via \texttt{save\_history()}, interpolating bases and those randonly
+via \texttt{save\_history()}, interpolating bases and those randomly
 nearby bases generated in the optimisation are not stored. This creates
 difficulties for fully understand the behaviour of the optimisation and
-interpolation of tour in complex scenario \textbf{{[}needa rephrase this
-part{]}.}
+interpolation of tour in complex scenario \textbf{{[}need a rephrase
+this part{]}.}
 
 Two set of simulated data are used in the demonstration of the
-visualiation and diagnostics of the tour optimisation. A small dataset
+visualisation and diagnostics of the tour optimisation. A small dataset
 consists of 1000 randomly simulated observations of five variables
@@
 \texttt{x2} is the informative variable simulated from two bi-modal
-normal distribution centered at -3 and 3 with variance being 1 and the
+normal distribution centred at -3 and 3 with variance being 1 and the
 other four are simulated from \texttt{N(0,\ 1)}. The data has been
@@
 generated will be stored in a global structure. The global structure
-consists of six columnes: \texttt{basis}, \texttt{index\_val},
+consists of six columns: \texttt{basis}, \texttt{index\_val},
 \texttt{tries}, \texttt{info}, \texttt{loop}, \texttt{id} and captured
@@
 presents the global object of a 1D projection of the small dataset with
-geodesic seraching method.
+geodesic searching method.
 
@@
 could be the decision variable, the value of the optimisation function
-or other parameter that facilitate the algorithm. In projectio pursuit
+or other parameter that facilitate the algorithm. In projection pursuit
 guided tour, apart from the decision variable projection basis, the
@@
 the observations are recorded as the algorithm progresses and thus can't
-be switch in rows. This time series structure is always hierachical and
+be switch in rows. This time series structure is always hierarchical and
 nested since each row can be labelled with the smallest unit \texttt{id}
@@
 smallest unit the observation is ordered. A larger ordering unit is
-\texttt{tries} since it is updated everytime a new target basis is
+\texttt{tries} since it is updated every time a new target basis is
 found. A nested ordering unit is \texttt{loop}, which increases by one
@@
 \centering
-\includegraphics{paper_files/figure-latex/points-id-1.pdf}
-\caption{\label{points-id}this is a xxx}
-\end{figure}
-
-\begin{figure}
-\centering
 \includegraphics{paper_files/figure-latex/points-tries-1.pdf}
@@
 Plotting the interpolating points as time series data allows us to
-diagnose characterstics of different configurations and index functions.
-Here we present two examples of using plots to diagnose the tour
-algorithm and different index functions.
+diagnose characteristics of different configurations and index
+functions. Here we present two examples of using plots to diagnose the
+tour algorithm and different index functions.
 
@@
 current basis to the target basis, which has been found by the
-projection pursuit algorithm to ahve a higher index value. After the
+projection pursuit algorithm to have a higher index value. After the
 interpolation, the target basis will become the current basis and send
@@
 
-Thus an interruption is constructed to accpet the interpolating bases up
-to the one with the larget index value on the interpolation path, and
+Thus an interruption is constructed to accept the interpolating bases up
+to the one with the larger index value on the interpolation path, and
 that basis is taken as the current basis for the next iteration. After
@@
 
-\emph{Example: Noisy index function} Add an example on noisy function
+\emph{Example: Noisy index function} The interpolation path of holes
+index, as seen in Figure \ref{fig:interruption}, is smooth, while this
+may not be the case for more complicated index function.
+\texttt{kol\_cdf} index, an index function based on Kolmogorov test,
+compares the difference between a projection matrix and a randomly
+generated normal distribution based on cumulated distribution function
+(CDF). Several visualisations below shows some characteristic of this
+index function.
 
-xxx
+Figure \ref{fig:geodesic-better} compares the tracing plot of the
+interpolating points for \texttt{search\_geodesic} and
+\texttt{search\_better}. Rather than a smooth interpolation, the
+interpolation path for \texttt{kol\_cdf} shows a zig-zag pattern.
+Polishing step has done much more work to reach the final index value
+for \texttt{search\_geodesic} than \texttt{search\_better} and this
+indicates that noisy index's favour of a random search method than
+ascent method.
 
-Currently, I have three things to put for the noisy index function.
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/geodesic-better-1.pdf}
+\caption{Comparison of two different searching methods: search\_geodesic
+and search\_better on 1D projection problem for a noisy index: kol\_cdf.
+The geodesic search rely heavily on the polishing step to find the final
+index value while search better works well.}
+\end{figure}
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  interpolating plot finds that search\_geodesic is not doing a good job
-  on optimising the noisy index; while search\_better is.
-\item
-  The interpolating plot for noisy index function is not smooth but the
-  target basis is always increasing.
-\item
-  search\_better\_random sees to find both local and global maximum.
-\end{enumerate}
+Moving from a 1D projection on a dataset with only one informative
+variable, we now introduce another informative variable in the data and
+remain 1D projection. In this problem, there are two local maximum being
+0.17 when the projection matrix is \([0, 1, 0, 0, 0, 0]\) and 0.228 when
+the projection matrix is \([0, 0, 1 ,0, 0, 0]\). As in Figure
+\ref{fig:1d-2var-different-seeds}, using \texttt{search\_better} with
+different seeds, both local maximum can be found. A simulated annealing
+algorithm, \texttt{search\_better\_random} in Figure
+\ref{fig:1d-2var-better-random}, can always find the global maximum at a
+cost of a large number of \texttt{tries}.
 
-Need to decise where to put this example. Do we wan to intoduce noisy
-function at this stage or we put it here because they are insights draw
-from the interpolating plots.
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
+\caption{The trace plot search better in a 1D projection problem with
+two informative variables using different seeds (without polishing).
+Since there are two informative variables, setting different value for
+seed will lead search better to find either of the local maximum.}
+\end{figure}
 
-xxx
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
+\caption{Using search better random for the problem above will result in
+finding the global maximum but much larger number of iteration is
+needed.}
+\end{figure}
 
@@
 \emph{Example: Polish} In principle, all the optimisation routines
-should result in the same output on the same problem. Figure
-\ref{fig:trace-compare} shows the interpolating plots for two different
-searching methods: search\_geodesic and serach\_better on a 2D
-projection problem for the larger dataset using the holes index. The two
-methods reaches different ending value in the red dots, which is not
-ideal. This motivates the creation of a polishing search that polishes
-the ending basis and achieves unity on different methods.
-
-\includegraphics{paper_files/figure-latex/trace-compare-1.pdf}
+should result in the same output on the same problem while this may not
+be the case in real application. This motivates the creation of a
+polishing search that polishes the ending basis and achieves unity on
+different methods.
 
 \texttt{search\_polish} takes the ending basis of a given search as the
-current basis and uses a brutal-force approach to sample a larg number
-of basis (\texttt{n\_sample}) in the neightbourhood, whose radius is
+current basis and uses a brutal-force approach to sample a large number
+of basis (\texttt{n\_sample}) in the neighbourhood, whose radius is
 controlled by \texttt{polish\_alpha}. Among the \texttt{n\_sample}
@@
 
+\emph{Example: The neighbourhood parameter alpha} Add an example on
+comparing the neighbourhood parameter in search\_better \&
+search\_posse.
+
 \begin{Shaded}
@@
 dataset. This imposes difficulties in visualisation since we are bounded
-to preceive at most three dimensions. Thus, principal component analysis
-is used to reduce the dimension of the projection basi and the first two
-principal component are mapped to the x and y axis of the plot. Another
-variable of interest could be mapped to the color aesthetics to see how
-it changes as the projection basis changes in the projected 2D space.
-{[}have a second read here{]}
+to perceive at most three dimensions. Thus, principal component analysis
+is used to reduce the dimension of the projection basis and the first
+two principal component are mapped to the x and y axis of the plot.
+Another variable of interest could be mapped to the color aesthetics to
+see how it changes as the projection basis changes in the projected 2D
+space. {[}have a second read here{]}
 
@@
 \includegraphics{paper_files/figure-latex/pca-1.pdf}
-\caption{\label{pca}this is xxx}
+\caption{\label{pca}PCA plot of search geodesic Coloring by info allows
+for better understanding of each stage in the geodesic search}
 \end{figure}
 
-\emph{Example: understand the polishing alpha via mapping alpha to
-color} \texttt{search\_polish} is a brutal-force algorithm that evaluate
-1000 points in the neighbourhod at each loop and an appropriate inital
-value in neighbourhood parameter alpha would avoid search on large
-vector space and concentrate on the ending basis. In Figure, two
-different initial alpha value is used in the polishing step. {[}some
-more illustration of the comparison of the two polish alpha here{]}
+\emph{Example: initial value for polishing alpha in search geodesic}
+\texttt{search\_polish} is a brutal-force algorithm that evaluate 1000
+points in the neighbourhood at each loop. Setting an appropriate initial
+value for polish\_alpha would avoid wasting search on large vector space
+that are not likely to produce higher index value. A default initial
+value for polishing step is 0.5 and we are interested in whether this is
+an appropriate initial value. The problem is a 1D projection of the
+small dataset using \texttt{search\_geodesic} and followed by
+\texttt{search\_polish}. Figure \ref{fig:polish-alpha} display all
+projection bases on the first two principal components, colored by the
+\texttt{polish\_alpha} parameter in the polishing step. Rather than
+concentrating on the ending basis from \texttt{search\_geodesic} as what
+polishing step is designed, it searches a much larger vector space that
+are unnecessary. Thus a user-supplied smaller initial value for
+\texttt{polish\_alpha} would be ideal.
+
+On the right panel of Figure \ref{fig:polish-alpha-tracing}, the
+\texttt{polish\_alpha} is initialised as the projection distance between
+the last two target bases and we observe a concentrated searching space
+near the ending basis. With the supplied initial value for
+\texttt{polish\_alpha}, the algorithm reaches the same ending index
+value with fewer iterations.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/polish-alpha-1.pdf}
+\caption{PCA plot of two different polish alpha initialisations. A
+default polish alpha = 0.5 searches a larger space that is unncessary
+while a small customised initial value of polish alpha will search near
+the ending basis.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/polish-alpha-tracing-1.pdf}
+\caption{Comparing the trace of two different polish alpha
+initialisations. Both intialisations reach the same ending index values}
+\end{figure}
 
@@
 \begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =}\NormalTok{ info)}
+\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =}\NormalTok{ info) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
 \end{Highlighting}

### 2020-06-08 paper/paper.Rmd
add DFO in introduction; rewrite introduction + vis diagnostics

@@
 
-## Tour
+## Derivative free optimisation
 
-Tour provides a way to explore multivariate data interactively via an established tour path.
-A tour path is formed by interpolating between randomly generated plane.
-Different types of tours are available depends on the purpose of the exploration i.e.
-a grand tour is suitable for randomly exploring the data from different angles; a guided tour detects a particular structure in the data.
-a manual tour allows user to manually control the projection [@cook1997manual].
+Given an objective function $f$, one way of optimising it is to equate its gradient to zero.
+In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
+The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which gains its popularity due to its simplicity in use and reliability in complicated practical problem.
+Coined by @hooke1961direct, direct search methods don't require any gradient or hessian information and evaluate $f$ directly.
+A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability.
+[feel like this sentence can be expanded to include more information.
+xxx].
+Further, direct search methods can be classified as *stochastic* and *deterministic* depends on whether a random sample is used in the search.
+The stochastic version of  direct-search method is the focus of this paper.
 
-## Guided tour
+## Projection pursuit guided tour
 
-Guided tour is usually used in conjunction projection pursuit, a method coined by @friedman1974projection to detect "interesting" low-dimension projection of multivariate data.
-A projection pursuit requires the definition of a projection pursuit index function and an optimisation routine.
-The projection pursuit index measures the "interestingness" of data defined as its departure from normality.
-Numerous indices have been proposed in the literature, including lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
-An optimisation routine is required to find the projection basis (thus projection) that maximises the projection pursuit index.
-The discussion of existing optimisation procedure will be discussed in the next section.
+The optimisation problem we're interested in is in the context of projection pursuit.
+Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e.
+clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
+A Projection Pursuit Index (PPI) is a function of the projection matrix (a.k.a projection angle, projection basis) measuring the "interestingness" of data and we refer the value of PPI function as *index value*.
+In the literature, the indices being proposed include lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+[Any literature on holes index?
+xxx]
 
-Guided tour creates visualisation for the projections found by projection pursuit by constructing a tour path.
-An illustration modified from [@buja2005computational] entails the geodesic interpolation between the plane generated by projection pursuit.
+ As @friedman1974projection noted "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." A suitable optimisation procedure is needed to find the projection angle that maximises the PPI and the quality of the optimisation largely affect the interesting projections one could possibly observe.
 
-```{r tour-path}
-knitr::include_graphics(path = here::here("figures/tour_path_new.png"))
+Projection pursuit is usually used in conjunction with a tour method called *guided tour*.
+Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path*, like movie frames in real time.
+Starting from a randomly generated projection basis, projection pursuit searches for a target basis with higher index value and guided tour interpolates geodesically between the initial basis and the target basis.
+This series of bases forms a tour path and the projection of the data on these bases is played in real time.
+This search-and-interpolateprocess is repeated until no basis can be found with higher index value.
+Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
+Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+
+```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
 ```
 
-## Optimisation methods in projection pursuit literature
+[doubt if I need the following. xxx]
+<!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis.
+An illustration modified from [@buja2005computational] vividly depicts the tour path in guided tour.
+-->
 
-As @friedman1974projection said "..., the technique use for maximizing the (one- and two-dimensional) projection index strongly influences both the statistical and the computational aspects of the procedure." The quality of the optimisation procedure largely affect the tour view and thus, the interesting projection one could possibly observe.
-An ideal optimisation procedure needs to have following characteristics:
+<!-- The pseudo-code below illustrates the implementation of guided tour in the tourr package.
+Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
+This information is then used to compute a series of interpolating bases inside the `tour()` function.
+All the basis will be sent to create animation for visualising the tour in the `animate()` function.
+-->
 
-- *Being able to handle non-differentiable index function*: the index function could be noisy and non-differentiable
+<!-- ```{r algo,eval = FALSE, echo = TRUE} -->
+<!-- animation <- function(){ -->
 
-- *Being able to optimise with constraints*: the projection matrix is restricted to an orthonormal matrix.
+<!--   # compute projection basis  -->
+<!--   tour <- function(){ -->
 
-- *Being able to reveal both local and global maximum*: Although the primary interest is to find the global maximum, distinct local structures are also of our interest.
+<!--     # construct bases on the tour path -->
+<!--     new_geodesic_path <- function(){ -->
+<!--       tour_path <- function(){ -->
 
-Below present three existing methods in tour.
+<!--         # GENERATOR: generate projection basis via projection pursuit -->
+<!--         guided_tour <- function(){ -->
+<!--           generator <- function(){ -->
 
-@posse1995projection presented a random search algorithm that samples new basis in the neighbourhood of the current basis.
-The neighbourhood is defined via the radius of the p-dimensional sphere, $c$.
+<!--             # define projection pursuit index -->
+<!--             # generate the target basis from the current basis via optimisation -->
+<!--           } -->
+<!--         } -->
+
+<!--         # prepare geodesic information needed for interpolating along the tour path -->
+<!--       } -->
+<!--     } -->
+
+<!--     # INTERPOLATOR: interpolate between the current and target basis  -->
+<!--     function(){ -->
+<!--       # generate interpolating bases on the geodesic path -->
+<!--     } -->
+<!--   } -->
+
+<!--   # animate according to different display methods -->
+<!-- } -->
+<!-- ``` -->
+
+## Optimisation in projection pursuit
+
+Below listed several features characterise the optimisation procedures needed in projection pursuit
+
+- *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.
+
+- *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.
+
+- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reeal structures in the data that are of our interest.
+
+[thinking if the following three should be presented as algorithms or plain description is fine. xxx]
+
+There are three existing methods for optimisating PPI function and we review them below.
+@posse1995projection used a stochastic direct search method,  a random search algorithm to sample new bases in the neighbourhood of the current basis defined by the radius of the p-dimensional sphere, $c$.
 The new basis is taken as the target basis if it has higher index value, or the sampling continues.
@@
 The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.
-[Pursuit package uses this method and it works great!
-But I don't think we implement this - although don't think it is too hard to do it].
 
@@
 If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.
-
-<!-- Because of the fact that projection pursuit index can be non-differentiable and that the projection basis is constrained to be orthonormal, the optimisation need to be a derivative-free, constrained procedure.
-Also, local maxima is equally interested as global maxima since it reveals distinct local structure in the data.
--->
 
@@
 
-# Iterative algorithm and its diagnostics
+# Visual diagnostic system
 
-## Tour components
-
-Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
-A tour includes two major components: a *generator* that generating the projection basis according to projection pursuit and an *interpolator* that performing geodesic interpolation between the projection basis.
-
-The pseudo-code below illustrates the implementation of guided tour in the tourr package.
-Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
-This information is then used to compute a series of interpolating bases inside the `tour()` function.
-All the basis will be sent to create animation for visualising the tour in the `animate()` function.
-
-```{r algo,eval = FALSE, echo = TRUE}
-animation <- function(){
-
-  # compute projection basis 
-  tour <- function(){
-
-    # construct bases on the tour path
-    new_geodesic_path <- function(){
-      tour_path <- function(){
-
-        # GENERATOR: generate projection basis via projection pursuit
-        guided_tour <- function(){
-          generator <- function(){
-
-            # define projection pursuit index
-            # generate the target basis from the current basis via optimisation
-          }
-        }
-
-        # prepare geodesic information needed for interpolating along the tour path
-      }
-    }
-
-    # INTERPOLATOR: interpolate between the current and target basis 
-    function(){
-      # generate interpolating bases on the geodesic path
-    }
-  }
-
-  # animate according to different display methods
-}
-```
-
-## Diagnostics
-
-Visualisation has been widely used for exploring and understanding data.
-Visualisation presents information in a graphical manner and often allows people to see information they would otherwise not seen in the reporting of numerical summarisation.
-Visual diagnostics can be real-time or post-run.
+## Visual diagnostics 
+<!-- Visual diagnostics can be real-time or post-run.
 Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
 This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.
+-->
 
-Post-run diagnostics requires the data produced during the algorithm to be stored/ saved in order to produce plot diagnostics.
+<!-- Visualisation has been widely used for exploring and understanding data. Presenting information in a graphical manner often allows people to see information they would otherwise not seen. -->
 
-The graphical system in R is established based on the conception grammar of graphic [@wickham2010layered], where a graphic is defined using stacked layers in a coordination system.
-A layer includes 1) the dataset that powers the plot; 2) a geometric object that represents the visual shape of the plot and 3) relevant statistical transformation if needed.
+Random search methods has a black-box machinism and focuses solely on finding the global maximum point, while the projection pursuit problem we have is aiming at *exploring* the data and thus is interested in how the algorithm finds its maximum.
+This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
+
+The necessity of developing such a system rather than simply producing different diagnostic plots is because there is always needs to develop a function specific for diagnosing one particular variables and these functions developed are not reusable at all.
+Thus we want to establish a set of rules that prescribe the diagnostic of iterative algorithms systematically.
+
+The idea of unifying all the diagnostic plots under one framework is inspired by the concept of *gramma of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+In gramma of graphic, plots are not defined by its appearance (i.e.
+boxplot, histogram, scatter plot etc) but by "stacked layers".
+By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot, instead, it defines layers and allow users to specify the plot through these defined layers.
+
+[Do I need this paragraph to details the gramma of graphic? xxx]
+<!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
 An important concept in the grammar of graphic is *aesthetic mapping*.
@@
 For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot.
-To create a boxplot, we first map one variable on the x-axis and then map the five point summary of another variable on the y-axis.
-This computation of five point summary from the origin variable is statistical transformation.
-This definition of graphic through layers provides advantages to produce complex plots since all the plots, however complex, can be decomposed into basic geometric objects shaped by the variable supplied.
+Another example of creating a boxplot involves 1) mapping one variable on the x-axis and 2) transforming another variable to its five-point-summary and mapping the five-point-summary to the y-axis.
+-->
 
-This idea of decompositing a graph to basic elements inspires me to characterise the diagnostics of iterative algorithm in a similar fashion.
-The optimisation in many machine learning algorithms these days is iterative in nature, while remains as a black box.
-Being able to diagnose it visually allows researchers to have the tool to unfold the myth and thus provides more understand to the algorithm.
+## Glocal object
 
-The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
+Ggplot prescribes how to make a graphic but it doesn't prescribe how to run a diagnostic.
+[need to have a paragraph or two here to describe how I come up with the design of four components in the global object - need to related to the component in optimisation.
+xxx].
+
+<!-- The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
 The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
 The next section shows how a global object is created in projection pursuit guided tour.
+-->
 
-## Data Structure
+<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
 
-In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**
+A global object used for diagnosing an optimisation is defined with four components: decision variable, value of objective function, iterators and other parameter of interest.
+In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is index value.
+There are three iterators: `id` is the smallest ordering unit that increases by one for each observatoin; `tries` is a larger order unit, which is updated once a search-and-interpolate step is finished.
+A third iterator `loop` is used to record the number of repitition in the search step and starts over from one at the beginning of a new `tries`.
+Since there are different methods used in the projection pursuit guided tour, we are interested to compare between different methods, thus `method` is created as another parameter of interest.
+The neighbourhood parameter alpha is always interested in the direct search methods since it controls the size of the sampling space, thus `alpha` is another parameter of interest.
+`info` is created to label different stage in the searching process and it is the third parameter of interest.
+A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
 
-Two set of simulated data are used in the demonstration of the visualisation and diagnostics of the tour optimisation.
+```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The global object in projection pursuit guided tour."}
+knitr::include_graphics(path = here::here("figures/global_obj.png"))
+```
+
+# Diagnostics plots
+
+## Simulated data
+
+Two set of simulated data are used in the demo of the visual diagnostics in projection pursuit guided tour.
 A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`).
-`x2` is the informative variable simulated from two bi-modal normal distribution centred at -3 and 3 with variance being 1 and the other four are simulated from `N(0, 1)`.
+`x2` is the informative variable simulated from two bi-modal normal distribution centred at -3 and 3 with variance of one and the other four are simulated from non-informative standard random normal distributions.
 The data has been scaled to ensure `x2` has variance of 1.
+The goal is to find the 1D projection of bi-modal shape, which correspondes to the projection matrix (vector) of `[0, 1, 0, 0, 0]`.
 
 A larger dataset contains more informative variables (`x3` to `x7`) of different types.
-`x3` takes 500 positive one and 500 negative one.
-The distribution of all the variables except `x3` is plotted below.
-*[should I introduce the dist for each var?]*
+The distribution of all the variables except `x3` is plotted below and `x3` takes 500 positive one and 500 negative one.
+[should I introduce the dist for each var?
+xxx] [also feel like we don't really use `x3` to `x6`, should we not mention about these?
+xxx]
 
@@
 
-# Post-run static diagnostics plots
-
-Given all the information of interest in a global object, the next thing is to design its visual representation that can diagnose our algorithm.
-In an iterative algorithm, one would be interested to know how one parameter changes as the algorithm iterates.
-The parameter could be could be the decision variable, the value of the optimisation function or other parameter that facilitate the algorithm.
-In projection pursuit guided tour, apart from the decision variable projection basis, the value of index function (`index_val`), we are also interested to explore the effect the different searching methods, the neighbourhood parameter alpha, and the stage a particular observation is in (`info`).
-These are the variables that will be mapped into the diagnostic plot.
-The global object also has its time series nature since the observations are recorded as the algorithm progresses and thus can't be switch in rows.
-This time series structure is always hierarchical and nested since each row can be labelled with the smallest unit `id` and up by the iterative structure of the algorithm.
-Take our example, in the global object, each row has an `id` label which is the smallest unit the observation is ordered.
-A larger ordering unit is `tries` since it is updated every time a new target basis is found.
-A nested ordering unit is `loop`, which increases by one as searching method iterates and starts over at a new `tries`.
-These features explain the definition of the global object for projection pursuit guided tour in Figure \ref{fig:glb-obj}.
-
-```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}this is xxxx"}
-knitr::include_graphics(path = here::here("figures/global_obj.png"))
-```
-
 ## Explore scalar parameters
@@
 *Example: exploring searching points* The data is sourced from a 2D projection of the larger dataset and `search_better` is used with `max.tries = 500`.
-In Figure \ref{fig:points-id} and \ref{fig:points-tries}, a comparison has been made on visualising the searching points with `id` and `tries` on the x-axis, colored by `tries`.
+In Figure \ref{fig:points-id} and \ref{fig:points}, a comparison has been made on visualising the searching points with `id` and `tries` on the x-axis, colored by `tries`.
 In Figure \ref{fig:points-id}, the searching points of the first few `tries` are squeezed in a small width leaving the large uninteresting searching points in the last three `tries` taking a vast majority of the plotting space.
-While in Figure \ref{fig:points-tries} the data is spaced by `tries` evenly in the plot.
+While in Figure \ref{fig:points} the data is spaced by `tries` evenly in the plot.
 Label at the bottom indicates the number of observations in each tries and facilitates the choice of cutoff to switch from point geometry to boxplot geometry (`cutoff = 15`).
@@
 
-Matrix parameter could also be interested to explore, for example, in the tour, we are interested to explore the relative position of the projection basis in the vector space.
-A projection basis is a vector/matrix whose row number is the number of variable in the given dataset.
+Sometimes, a parameter to explore could be a vector or matrix instead of scalar, for example, in tour, we are interested to explore the projection basis in the vector space.
 This imposes difficulties in visualisation since we are bounded to perceive at most three dimensions.
-Thus, principal component analysis is used to reduce the dimension of the projection basis and the first two principal component are mapped to the x and y axis of the plot.
-Another variable of interest could be mapped to the color aesthetics to see how it changes as the projection basis changes in the projected 2D space.
-[have a second read here]
+Thus, principal component analysis is used to reduce the dimension of projection bases and the first two principal components are mapped to the x and y axis of the plot.
+Additional variable of interest could be mapped to the color aesthetics to for exploration.
 

### 2020-06-08 paper/paper.tex
add DFO in introduction; rewrite introduction + vis diagnostics

@@
 
-\hypertarget{tour}{%
-\subsection{Tour}\label{tour}}
+\hypertarget{derivative-free-optimisation}{%
+\subsection{Derivative free
+optimisation}\label{derivative-free-optimisation}}
 
-Tour provides a way to explore multivariate data interactively via an
-established tour path. A tour path is formed by interpolating between
-randomly generated plane. Different types of tours are available depends
-on the purpose of the exploration i.e.~a grand tour is suitable for
-randomly exploring the data from different angles; a guided tour detects
-a particular structure in the data. a manual tour allows user to
-manually control the projection \citep{cook1997manual}.
+Given an objective function \(f\), one way of optimising it is to equate
+its gradient to zero. In modern optimisation problems, gradient
+information can be hard to evaluate or sometimes even impossible and
+Derivative-Free Optimisation (DFO) methods can be useful to approach
+these problems. The two most common methods in DFO are
+\emph{direct-search methods} and \emph{model-based methods} and this
+paper dedicates to the discussion of direct-search methods, which gains
+its popularity due to its simplicity in use and reliability in
+complicated practical problem. Coined by \citet{hooke1961direct}, direct
+search methods don't require any gradient or hessian information and
+evaluate \(f\) directly. A well-known example of it is the Nelder-Mead
+algorithm \citep{nelder1965simplex} and it enjoys the popularity due to
+its simplicity and reliability. {[}feel like this sentence can be
+expanded to include more information. xxx{]}. Further, direct search
+methods can be classified as \emph{stochastic} and \emph{deterministic}
+depends on whether a random sample is used in the search. The stochastic
+version of direct-search method is the focus of this paper.
 
-\hypertarget{guided-tour}{%
-\subsection{Guided tour}\label{guided-tour}}
+\hypertarget{projection-pursuit-guided-tour}{%
+\subsection{Projection pursuit guided
+tour}\label{projection-pursuit-guided-tour}}
 
-Guided tour is usually used in conjunction projection pursuit, a method
-coined by \citet{friedman1974projection} to detect ``interesting''
-low-dimension projection of multivariate data. A projection pursuit
-requires the definition of a projection pursuit index function and an
-optimisation routine. The projection pursuit index measures the
-``interestingness'' of data defined as its departure from normality.
-Numerous indices have been proposed in the literature, including
-lengendre index \citep{friedman1974projection}, hermite index
-\citep{hall1989polynomial}, natural hermite index
+The optimisation problem we're interested in is in the context of
+projection pursuit. Coined by \citet{friedman1974projection}, projection
+pursuit is a method that detects the interesting structure
+(i.e.~clustering, outliers and skewness) of multivariate data via
+projecting it in lower dimensions. A Projection Pursuit Index (PPI) is a
+function of the projection matrix (a.k.a projection angle, projection
+basis) measuring the ``interestingness'' of data and we refer the value
+of PPI function as \emph{index value}. In the literature, the indices
+being proposed include lengendre index \citep{friedman1974projection},
+hermite index \citep{hall1989polynomial}, natural hermite index
 \citep{cook1993projection}, chi-square index
 \citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
-index \citep{lee2010projection}. An optimisation routine is required to
-find the projection basis (thus projection) that maximises the
-projection pursuit index. The discussion of existing optimisation
-procedure will be discussed in the next section.
+index \citep{lee2010projection}. {[}Any literature on holes index?
+xxx{]}
 
-Guided tour creates visualisation for the projections found by
-projection pursuit by constructing a tour path. An illustration modified
-from \citep{buja2005computational} entails the geodesic interpolation
-between the plane generated by projection pursuit.
+As \citet{friedman1974projection} noted ``\ldots{}, the technique use
+for maximising the projection index strongly influences both the
+statistical and the computational aspects of the procedure.'' A suitable
+optimisation procedure is needed to find the projection angle that
+maximises the PPI and the quality of the optimisation largely affect the
+interesting projections one could possibly observe.
 
-\includegraphics[width=7.75in]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_new}
+Projection pursuit is usually used in conjunction with a tour method
+called \emph{guided tour}. Tour explores the multivariate data
+\emph{interactively} via playing a series of projections, that form a
+\emph{tour path}, like movie frames in real time. Starting from a
+randomly generated projection basis, projection pursuit searches for a
+target basis with higher index value and guided tour interpolates
+geodesically between the initial basis and the target basis. This series
+of bases forms a tour path and the projection of the data on these bases
+is played in real time. This search-and-interpolateprocess is repeated
+until no basis can be found with higher index value. Modified from
+\citet{buja2005computational}, Figure \ref{tour-path} vividly depicts
+the tour path in guided tour. Guided tour, along with other types of
+tour, has been implemented in the \emph{tourr} package in R, available
+on the Comprehensive R Archive Network at
+\url{https://cran.r-project.org/web/packages/tourr/}
+\citep{wickham2011tourrpackage}.
 
-\hypertarget{optimisation-methods-in-projection-pursuit-literature}{%
-\subsection{Optimisation methods in projection pursuit
-literature}\label{optimisation-methods-in-projection-pursuit-literature}}
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\end{figure}
 
-As \citet{friedman1974projection} said ``\ldots{}, the technique use for
-maximizing the (one- and two-dimensional) projection index strongly
-influences both the statistical and the computational aspects of the
-procedure.'' The quality of the optimisation procedure largely affect
-the tour view and thus, the interesting projection one could possibly
-observe. An ideal optimisation procedure needs to have following
-characteristics:
+{[}doubt if I need the following. xxx{]}
+
+\hypertarget{optimisation-in-projection-pursuit}{%
+\subsection{Optimisation in projection
+pursuit}\label{optimisation-in-projection-pursuit}}
+
+Below listed several features characterise the optimisation procedures
+needed in projection pursuit
 
@@
 \item
-  \emph{Being able to handle non-differentiable index function}: the
-  index function could be noisy and non-differentiable
+  \emph{Being able to handle non-differentiable PPI function}: The PPI
+  function could be non-differentiable, thus derivative free methods are
+  preferred.
 \item
-  \emph{Being able to optimise with constraints}: the projection matrix
-  is restricted to an orthonormal matrix.
+  \emph{Being able to optimise with constraints}: The constraint comes
+  from projection matrix being an orthonormal matrix.
 \item
-  \emph{Being able to reveal both local and global maximum}: Although
-  the primary interest is to find the global maximum, distinct local
-  structures are also of our interest.
+  \emph{Being able to find both local and global maximum}: Although the
+  primary interest is to find the global maximum, local maximum could
+  also reeal structures in the data that are of our interest.
 \end{itemize}
 
-Below present three existing methods in tour.
+{[}thinking if the following three should be presented as algorithms or
+plain description is fine. xxx{]}
 
-\citet{posse1995projection} presented a random search algorithm that
-samples new basis in the neighbourhood of the current basis. The
-neighbourhood is defined via the radius of the p-dimensional sphere,
-\(c\). The new basis is taken as the target basis if it has higher index
-value, or the sampling continues. If no basis is found to have higher
-index value after a certain number of tries \(n\), the radius \(c\) is
-halved. The algorithm stops when the maximum number of iteration is
-attained or the radius \(c\) is less than a pre-determined number.
-{[}Pursuit package uses this method and it works great! But I don't
-think we implement this - although don't think it is too hard to do
-it{]}.
+There are three existing methods for optimisating PPI function and we
+review them below. \citet{posse1995projection} used a stochastic direct
+search method, a random search algorithm to sample new bases in the
+neighbourhood of the current basis defined by the radius of the
+p-dimensional sphere, \(c\). The new basis is taken as the target basis
+if it has higher index value, or the sampling continues. If no basis is
+found to have higher index value after a certain number of tries \(n\),
+the radius \(c\) is halved. The algorithm stops when the maximum number
+of iteration is attained or the radius \(c\) is less than a
+pre-determined number.
 
@@
 
-\hypertarget{iterative-algorithm-and-its-diagnostics}{%
-\section{Iterative algorithm and its
-diagnostics}\label{iterative-algorithm-and-its-diagnostics}}
+\hypertarget{visual-diagnostic-system}{%
+\section{Visual diagnostic system}\label{visual-diagnostic-system}}
 
-\hypertarget{tour-components}{%
-\subsection{Tour components}\label{tour-components}}
+\hypertarget{visual-diagnostics}{%
+\subsection{Visual diagnostics}\label{visual-diagnostics}}
 
-Guided tour, along with other types of tour, has been implemented in the
-\emph{tourr} package in R, available on the Comprehensive R Archive
-Network at \url{https://cran.r-project.org/web/packages/tourr/}
-\citep{wickham2011tourrpackage}. A tour includes two major components: a
-\emph{generator} that generating the projection basis according to
-projection pursuit and an \emph{interpolator} that performing geodesic
-interpolation between the projection basis.
+Random search methods has a black-box machinism and focuses solely on
+finding the global maximum point, while the projection pursuit problem
+we have is aiming at \emph{exploring} the data and thus is interested in
+how the algorithm finds its maximum. This motivates us \textbf{to
+develop a visual diagnostic system for exploring the optimisation
+searching path}.
 
-The pseudo-code below illustrates the implementation of guided tour in
-the tourr package. Given an projection pursuit index function and a
-randomly generated projection basis (current basis), the optimisation
-procedure produces a target basis inside \texttt{generator()} Both the
-current basis and the target basis will be supplied to
-\texttt{tour\_path()} to prepare information needed for constructing a
-geodesic path. This information is then used to compute a series of
-interpolating bases inside the \texttt{tour()} function. All the basis
-will be sent to create animation for visualising the tour in the
-\texttt{animate()} function.
+The necessity of developing such a system rather than simply producing
+different diagnostic plots is because there is always needs to develop a
+function specific for diagnosing one particular variables and these
+functions developed are not reusable at all. Thus we want to establish a
+set of rules that prescribe the diagnostic of iterative algorithms
+systematically.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{animation <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
+The idea of unifying all the diagnostic plots under one framework is
+inspired by the concept of \emph{gramma of
+graphic}\citep{wickham2010layered}, which powers the primary graphical
+system in R, ggplot2 \citep{ggplot2}. In gramma of graphic, plots are
+not defined by its appearance (i.e.~boxplot, histogram, scatter plot
+etc) but by ``stacked layers''. By this design, ggplot doesn't need to
+develop a gazillion of functions that each produces a different type of
+plot, instead, it defines layers and allow users to specify the plot
+through these defined layers.
 
-  \CommentTok{# compute projection basis }
-\NormalTok{  tour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
+{[}Do I need this paragraph to details the gramma of graphic? xxx{]}
 
-    \CommentTok{# construct bases on the tour path}
-\NormalTok{    new_geodesic_path <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
-\NormalTok{      tour_path <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
+\hypertarget{glocal-object}{%
+\subsection{Glocal object}\label{glocal-object}}
 
-        \CommentTok{# GENERATOR: generate projection basis via projection pursuit}
-\NormalTok{        guided_tour <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
-\NormalTok{          generator <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{()\{}
+Ggplot prescribes how to make a graphic but it doesn't prescribe how to
+run a diagnostic. {[}need to have a paragraph or two here to describe
+how I come up with the design of four components in the global object -
+need to related to the component in optimisation. xxx{]}.
 
-            \CommentTok{# define projection pursuit index}
-            \CommentTok{# generate the target basis from the current basis via optimisation}
-\NormalTok{          \}}
-\NormalTok{        \}}
+A global object used for diagnosing an optimisation is defined with four
+components: decision variable, value of objective function, iterators
+and other parameter of interest. In the projection pursuit guided tour,
+the decision variable is the projection matrix, the value of objective
+function is index value. There are three iterators: \texttt{id} is the
+smallest ordering unit that increases by one for each observatoin;
+\texttt{tries} is a larger order unit, which is updated once a
+search-and-interpolate step is finished. A third iterator \texttt{loop}
+is used to record the number of repitition in the search step and starts
+over from one at the beginning of a new \texttt{tries}. Since there are
+different methods used in the projection pursuit guided tour, we are
+interested to compare between different methods, thus \texttt{method} is
+created as another parameter of interest. The neighbourhood parameter
+alpha is always interested in the direct search methods since it
+controls the size of the sampling space, thus \texttt{alpha} is another
+parameter of interest. \texttt{info} is created to label different stage
+in the searching process and it is the third parameter of interest. A
+sketch of the global object for projection pursuit guided tour is
+presented in Figure \ref{fig:glb-obj}.
 
-        \CommentTok{# prepare geodesic information needed for interpolating along the tour path}
-\NormalTok{      \}}
-\NormalTok{    \}}
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
+\end{figure}
 
-    \CommentTok{# INTERPOLATOR: interpolate between the current and target basis }
-    \ControlFlowTok{function}\NormalTok{()\{}
-      \CommentTok{# generate interpolating bases on the geodesic path}
-\NormalTok{    \}}
-\NormalTok{  \}}
+\hypertarget{diagnostics-plots}{%
+\section{Diagnostics plots}\label{diagnostics-plots}}
 
-  \CommentTok{# animate according to different display methods}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{simulated-data}{%
+\subsection{Simulated data}\label{simulated-data}}
 
-\hypertarget{diagnostics}{%
-\subsection{Diagnostics}\label{diagnostics}}
-
-Visualisation has been widely used for exploring and understanding data.
-Visualisation presents information in a graphical manner and often
-allows people to see information they would otherwise not seen in the
-reporting of numerical summarisation. Visual diagnostics can be
-real-time or post-run. Real time diagnostic directly uses the data
-produced in the algorithm to produce visual representation and thus
-doesn't need to store the data. This section focuses on the definition
-and production of post-run diagnostics and the next section discusses
-real-time diagnostics.
-
-Post-run diagnostics requires the data produced during the algorithm to
-be stored/ saved in order to produce plot diagnostics.
-
-The graphical system in R is established based on the conception grammar
-of graphic \citep{wickham2010layered}, where a graphic is defined using
-stacked layers in a coordination system. A layer includes 1) the dataset
-that powers the plot; 2) a geometric object that represents the visual
-shape of the plot and 3) relevant statistical transformation if needed.
-An important concept in the grammar of graphic is \emph{aesthetic
-mapping}. Aesthetic mapping links the variable in a dataset to
-information needed to produce a geometric object. For example, we map
-one variable on the x-axis and another on the y-axis to create a
-scatterplot. To create a boxplot, we first map one variable on the
-x-axis and then map the five point summary of another variable on the
-y-axis. This computation of five point summary from the origin variable
-is statistical transformation. This definition of graphic through layers
-provides advantages to produce complex plots since all the plots,
-however complex, can be decomposed into basic geometric objects shaped
-by the variable supplied.
-
-This idea of decompositing a graph to basic elements inspires me to
-characterise the diagnostics of iterative algorithm in a similar
-fashion. The optimisation in many machine learning algorithms these days
-is iterative in nature, while remains as a black box. Being able to
-diagnose it visually allows researchers to have the tool to unfold the
-myth and thus provides more understand to the algorithm.
-
-The concept of grammar of graphic requires a dataset to be supplied in
-making a plot, thus a global object needs to be created in the iterative
-algorithm. The variables will be mapped into element of a graphic to
-explore and thus should contain all the parameters of interest. The next
-section shows how a global object is created in projection pursuit
-guided tour.
-
-\hypertarget{data-structure}{%
-\subsection{Data Structure}\label{data-structure}}
-
-In the current implementation of the \texttt{tourr} package, while the
-target basis generated by the projection pursuit can be accessed later
-via \texttt{save\_history()}, interpolating bases and those randomly
-nearby bases generated in the optimisation are not stored. This creates
-difficulties for fully understand the behaviour of the optimisation and
-interpolation of tour in complex scenario \textbf{{[}need a rephrase
-this part{]}.}
-
-Two set of simulated data are used in the demonstration of the
-visualisation and diagnostics of the tour optimisation. A small dataset
-consists of 1000 randomly simulated observations of five variables
-(\texttt{x1}, \texttt{x2}, \texttt{x8}, \texttt{x9}, \texttt{x10}).
-\texttt{x2} is the informative variable simulated from two bi-modal
-normal distribution centred at -3 and 3 with variance being 1 and the
-other four are simulated from \texttt{N(0,\ 1)}. The data has been
-scaled to ensure \texttt{x2} has variance of 1.
+Two set of simulated data are used in the demo of the visual diagnostics
+in projection pursuit guided tour. A small dataset consists of 1000
+randomly simulated observations of five variables (\texttt{x1},
+\texttt{x2}, \texttt{x8}, \texttt{x9}, \texttt{x10}). \texttt{x2} is the
+informative variable simulated from two bi-modal normal distribution
+centred at -3 and 3 with variance of one and the other four are
+simulated from non-informative standard random normal distributions. The
+data has been scaled to ensure \texttt{x2} has variance of 1. The goal
+is to find the 1D projection of bi-modal shape, which correspondes to
+the projection matrix (vector) of \texttt{{[}0,\ 1,\ 0,\ 0,\ 0{]}}.
 
 A larger dataset contains more informative variables (\texttt{x3} to
-\texttt{x7}) of different types. \texttt{x3} takes 500 positive one and
-500 negative one. The distribution of all the variables except
-\texttt{x3} is plotted below. \emph{{[}should I introduce the dist for
-each var?{]}}
+\texttt{x7}) of different types. The distribution of all the variables
+except \texttt{x3} is plotted below and \texttt{x3} takes 500 positive
+one and 500 negative one. {[}should I introduce the dist for each var?
+xxx{]} {[}also feel like we don't really use \texttt{x3} to \texttt{x6},
+should we not mention about these? xxx{]}
 
@@
 
-\hypertarget{post-run-static-diagnostics-plots}{%
-\section{Post-run static diagnostics
-plots}\label{post-run-static-diagnostics-plots}}
-
-Given all the information of interest in a global object, the next thing
-is to design its visual representation that can diagnose our algorithm.
-In an iterative algorithm, one would be interested to know how one
-parameter changes as the algorithm iterates. The parameter could be
-could be the decision variable, the value of the optimisation function
-or other parameter that facilitate the algorithm. In projection pursuit
-guided tour, apart from the decision variable projection basis, the
-value of index function (\texttt{index\_val}), we are also interested to
-explore the effect the different searching methods, the neighbourhood
-parameter alpha, and the stage a particular observation is in
-(\texttt{info}). These are the variables that will be mapped into the
-diagnostic plot. The global object also has its time series nature since
-the observations are recorded as the algorithm progresses and thus can't
-be switch in rows. This time series structure is always hierarchical and
-nested since each row can be labelled with the smallest unit \texttt{id}
-and up by the iterative structure of the algorithm. Take our example, in
-the global object, each row has an \texttt{id} label which is the
-smallest unit the observation is ordered. A larger ordering unit is
-\texttt{tries} since it is updated every time a new target basis is
-found. A nested ordering unit is \texttt{loop}, which increases by one
-as searching method iterates and starts over at a new \texttt{tries}.
-These features explain the definition of the global object for
-projection pursuit guided tour in Figure \ref{fig:glb-obj}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}this is xxxx}\label{fig:glb-obj}
-\end{figure}
-
 \hypertarget{explore-scalar-parameters}{%
@@
 
-\emph{Example: exploring searching points} The data is sourced from a 2D
-projection of the larger dataset and \texttt{search\_better} is used
-with \texttt{max.tries\ =\ 500}. In Figure \ref{fig:points-id} and
-\ref{fig:points-tries}, a comparison has been made on visualising the
-searching points with \texttt{id} and \texttt{tries} on the x-axis,
-colored by \texttt{tries}. In Figure \ref{fig:points-id}, the searching
-points of the first few \texttt{tries} are squeezed in a small width
-leaving the large uninteresting searching points in the last three
-\texttt{tries} taking a vast majority of the plotting space. While in
-Figure \ref{fig:points-tries} the data is spaced by \texttt{tries}
-evenly in the plot. Label at the bottom indicates the number of
-observations in each tries and facilitates the choice of cutoff to
-switch from point geometry to boxplot geometry (\texttt{cutoff\ =\ 15}).
-The line geometry suggest the largest improvement happens at
-\texttt{tries\ =\ 5}.
-
 \begin{figure}
@@
 \end{figure}
+
+\emph{Example: exploring searching points} The data is sourced from a 2D
+projection of the larger dataset and \texttt{search\_better} is used
+with \texttt{max.tries\ =\ 500}. In Figure \ref{fig:points-id} and
+\ref{fig:points}, a comparison has been made on visualising the
+searching points with \texttt{id} and \texttt{tries} on the x-axis,
+colored by \texttt{tries}. In Figure \ref{fig:points-id}, the searching
+points of the first few \texttt{tries} are squeezed in a small width
+leaving the large uninteresting searching points in the last three
+\texttt{tries} taking a vast majority of the plotting space. While in
+Figure \ref{fig:points} the data is spaced by \texttt{tries} evenly in
+the plot. Label at the bottom indicates the number of observations in
+each tries and facilitates the choice of cutoff to switch from point
+geometry to boxplot geometry (\texttt{cutoff\ =\ 15}). The line geometry
+suggest the largest improvement happens at \texttt{tries\ =\ 5}.
 
@@
 
-Matrix parameter could also be interested to explore, for example, in
-the tour, we are interested to explore the relative position of the
-projection basis in the vector space. A projection basis is a
-vector/matrix whose row number is the number of variable in the given
-dataset. This imposes difficulties in visualisation since we are bounded
-to perceive at most three dimensions. Thus, principal component analysis
-is used to reduce the dimension of the projection basis and the first
-two principal component are mapped to the x and y axis of the plot.
-Another variable of interest could be mapped to the color aesthetics to
-see how it changes as the projection basis changes in the projected 2D
-space. {[}have a second read here{]}
+Sometimes, a parameter to explore could be a vector or matrix instead of
+scalar, for example, in tour, we are interested to explore the
+projection basis in the vector space. This imposes difficulties in
+visualisation since we are bounded to perceive at most three dimensions.
+Thus, principal component analysis is used to reduce the dimension of
+projection bases and the first two principal components are mapped to
+the x and y axis of the plot. Additional variable of interest could be
+mapped to the color aesthetics to for exploration.
 

### 2020-06-09 paper/paper.Rmd
more text editing - whole paper except 1.3 & 1.4

@@
 - that do not appear in the title
-abstract: |
-  1. check consistency of using PP for projection pursuit and PPI for projection pursuit index 
+abstract: 
 bibliography: biblio.bib
@@
 The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which gains its popularity due to its simplicity in use and reliability in complicated practical problem.
-Coined by @hooke1961direct, direct search methods don't require any gradient or hessian information and evaluate $f$ directly.
+Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and evaluate $f$ directly.
 A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability.
@@
 A Projection Pursuit Index (PPI) is a function of the projection matrix (a.k.a projection angle, projection basis) measuring the "interestingness" of data and we refer the value of PPI function as *index value*.
-In the literature, the indices being proposed include lengendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+In the literature, the indices being proposed include Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 [Any literature on holes index?
@@
 This series of bases forms a tour path and the projection of the data on these bases is played in real time.
-This search-and-interpolateprocess is repeated until no basis can be found with higher index value.
+This search-and-interpolate process is repeated until no basis can be found with higher index value.
 Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
@@
 
-- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reeal structures in the data that are of our interest.
+- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
 
@@
 
-Random search methods has a black-box machinism and focuses solely on finding the global maximum point, while the projection pursuit problem we have is aiming at *exploring* the data and thus is interested in how the algorithm finds its maximum.
+Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit problem we have aims at *exploring* the data and thus is interested in how the algorithm finds its maximum.
 This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
 
-The necessity of developing such a system rather than simply producing different diagnostic plots is because there is always needs to develop a function specific for diagnosing one particular variables and these functions developed are not reusable at all.
-Thus we want to establish a set of rules that prescribe the diagnostic of iterative algorithms systematically.
+The necessity of developing such a system rather than simply producing different diagnostic plots is because the diagnostics of each variable requires a different function and these functions can't be scaled to other problems.
+Thus we want to establish a set of rules that can generalise the diagnostic of iterative algorithms.
 
-The idea of unifying all the diagnostic plots under one framework is inspired by the concept of *gramma of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
-In gramma of graphic, plots are not defined by its appearance (i.e.
+The idea of generalising all the diagnostic plots under one framework is inspired by the concept of *grammar of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+In grammar of graphic, plots are not defined by its appearance (i.e.
 boxplot, histogram, scatter plot etc) but by "stacked layers".
-By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot, instead, it defines layers and allow users to specify the plot through these defined layers.
+By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot.
+Instead, it creates aesthetic mapping from the variables to the geometric objects and builds the plot through different layers.
 
-[Do I need this paragraph to details the gramma of graphic? xxx]
+[Do I need this paragraph to details the grammar of graphic? xxx]
 <!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
@@
 
-## Glocal object
+## Global object
 
-Ggplot prescribes how to make a graphic but it doesn't prescribe how to run a diagnostic.
-[need to have a paragraph or two here to describe how I come up with the design of four components in the global object - need to related to the component in optimisation.
-xxx].
+Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
+Given an optimisation algorithm, two primary variables of interest are the *decision variable* and the *value of the objective function*.
+*iterators* indexes the data collected and this order has time series feature that prescribes the progression of the optimisation.
+Given the complexity of the algorithm, there could be multiple iterators.
+There are other parameters that can't be classified as one of the three categories but are also of our interest.
+They are defined under the fourth category: *other parameter of interest*.
+This could include variables that are specific for one particular algorithm.
+Below we show an example of defining the global object of projection pursuit guided tour.
 
@@
 
-A global object used for diagnosing an optimisation is defined with four components: decision variable, value of objective function, iterators and other parameter of interest.
-In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is index value.
-There are three iterators: `id` is the smallest ordering unit that increases by one for each observatoin; `tries` is a larger order unit, which is updated once a search-and-interpolate step is finished.
-A third iterator `loop` is used to record the number of repitition in the search step and starts over from one at the beginning of a new `tries`.
-Since there are different methods used in the projection pursuit guided tour, we are interested to compare between different methods, thus `method` is created as another parameter of interest.
-The neighbourhood parameter alpha is always interested in the direct search methods since it controls the size of the sampling space, thus `alpha` is another parameter of interest.
-`info` is created to label different stage in the searching process and it is the third parameter of interest.
+In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is the index value.
+There are three iterators: `id` is the smallest ordering unit that increases by one for each observation; `tries` is updated once a search-and-interpolate step is finished.
+A third iterator `loop` is used to record the number of repetition in the search step and starts over from one at the beginning of a new `tries`.
+Three other parameters of our interest includes `method`, `alpha` and `info`.
+`method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search.
+The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses.
+`info` labels different stages in the searching process.
 A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
@@
 The data has been scaled to ensure `x2` has variance of 1.
-The goal is to find the 1D projection of bi-modal shape, which correspondes to the projection matrix (vector) of `[0, 1, 0, 0, 0]`.
+The goal is to find the 1D projection of bi-modal shape, which corresponds to the projection matrix (vector) of `[0, 1, 0, 0, 0]`.
 
 A larger dataset contains more informative variables (`x3` to `x7`) of different types.
-The distribution of all the variables except `x3` is plotted below and `x3` takes 500 positive one and 500 negative one.
+The distribution of all the variables except `x3` is plotted in Figure \ref{origin-data} and `x3` takes 500 positive one and 500 negative one.
 [should I introduce the dist for each var?
@@
 
-```{r origin-data}
+```{r origin-data, fig.cap="\\label{origin-data} The distribution of simulated data except x3"}
 origin_dt_bi <- data_mult %>% 
@@
 
-Once the dataset is sent to the tourr package, all the information generated will be stored in a global structure.
-The global structure consists of six columns: `basis`, `index_val`, `tries`, `info`, `loop`, `id` and captured all the basis generated during whole tour process.
-The example below presents the global object of a 1D projection of the small dataset with geodesic searching method.
+In tour, when the optimisation ends the global object will be stored and printed and can be turned off via `print = FALSE`.
+Additional messages during the optimisation can be displayed via `verbose = TRUE`.
+Below presented the global object of the 1D projection using the small dataset.
+[the printing is not ideal as it doesn't show all the columns.
+xxx]
 
@@
 
-```{r glb-obj-1, echo = TRUE}
+```{r glb-obj-1}
 holes_1d_geo %>% head(5)
@@
 
-`tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in.
+<!-- `tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in.
 This would include the `interpolation` stage and the detailed stage in the optimisation i.e.
@@
 `id` creates a sequential order of the basis.
-This information will be stored and printed when the optimisation ends and can be turned off via `print = FALSE`.
-Additional messages during the optimisation can be displayed via `verbose = TRUE`.
-Another examples is a 2D projection of the larger dataset with two informative variable (`x2` and `x7`) using search_better method.
+-->
+
+<!--  Another examples is a 2D projection of the larger dataset with two informative variable (`x2` and `x7`) using search_better method.
 Notice in this example, the dimension of the bases becomes 6 by 2.
+-->
 
-```{r glb-obj-2, echo = TRUE}
-holes_2d_better %>% head(5) 
-```
+<!-- ```{r glb-obj-2, echo = TRUE} -->
+<!-- holes_2d_better %>% head(5)  -->
+<!-- ``` -->
 
@@
 
-The most interesting parameter to explore is the value of objective function.
-The points recorded in the global object can be divided into two broad categories: searching points and interpolating points
+In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
 
-- *Searching points* include the observations that are recorded in the searching algorithm in order to find the target basis.
+- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
 The points for target bases is also included in the searching points and there is one such point per `tries`.
@@
 
-As mentioned previously, the largest difficulties of exploring searching points is its unknown number of observations per `tries`.
-Mapping `id` on the x-axis will leave the `tries` with few observations a small space in the plot, while those `tries` with large number of search points towards the end occupying the vast majority of the space in the plot.
-
-This motivates the use of summarised statistics.
+The largest difficulties of exploring searching points is its unknown number of observations per `tries`.
+Mapping `id` on the x-axis will leave the `tries` with few observations a small space in the plot, while those `tries` with large number of search points (usually towards the end) occupy the vast majority of the plot space.
+This motivates to summarise the points in each `tries`.
 At each iteration, rather than knowing the index value of *every* points, we are more interested to know a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the geodesic interpolation and future searches.
 
-Boxplot is a suitable candidate that provides five points summary of the data, while it has one drawback: it doesn't report the number of point in each box.
-We may risk losing information on how long it takes for the search to find the target basis by displaying the boxplot alone for all `tries`.
-Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry.
-This is achieved via the `cutoff` argument.
-A line geometry is also added to link the points with the largest index value for each `tries`.
+Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box.
+We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
+Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry, which is achieved via the `cutoff` argument.
+A line geometry is also added to link the points with the largest index value in each `tries`.
 This helps to visualise the improvement made by each `tries`.
 
-*Example: exploring searching points* The data is sourced from a 2D projection of the larger dataset and `search_better` is used with `max.tries = 500`.
-In Figure \ref{fig:points-id} and \ref{fig:points}, a comparison has been made on visualising the searching points with `id` and `tries` on the x-axis, colored by `tries`.
-In Figure \ref{fig:points-id}, the searching points of the first few `tries` are squeezed in a small width leaving the large uninteresting searching points in the last three `tries` taking a vast majority of the plotting space.
-While in Figure \ref{fig:points} the data is spaced by `tries` evenly in the plot.
+*Example: exploring searching points* The larger dataset is used for 2D projection with `search_better` and `max.tries = 500`.
+In Figure \ref{points}, the searching points with `id` and `tries` on the x-axis, colored by `tries`.
 Label at the bottom indicates the number of observations in each tries and facilitates the choice of cutoff to switch from point geometry to boxplot geometry (`cutoff = 15`).
-The line geometry suggest the largest improvement happens at `tries = 5`.
+The line geometry suggests the largest improvement happens at `tries = 5`.
 
@@
 Plotting the interpolating points as time series data allows us to diagnose characteristics of different configurations and index functions.
-Here we present two examples of using  plots to diagnose the tour algorithm and different index functions.
+Here we present two examples of using  plots to diagnose tour algorithms and different index functions.
 
@@
 *Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
-As mentioned previously, the interpolation starts from the current basis to the target basis, which has been found by the projection pursuit algorithm to have a higher index value.
-After the interpolation, the target basis will become the current basis and send back to the projection pursuit algorithm to find the next target basis.
-From figure \ref{fig:interruption}, it is possible that there are bases with index value higher than the target basis on the interpolation path and these bases could be used to search for new basis in the next iteration.
+In the interpolation stage, continuous frames will be constructed to bridge the project on the current basis to the target basis.
+The target basis will become the current basis in the next iteration and the searching stage continues to find the next target basis.
+In the left panel of Figure \ref{interruption}, we observe that  there are interpolating bases with higher index values than the target basis and these bases could be used to search for new basis in the next searching stage.
 
-Thus an interruption is constructed to accept the interpolating bases up to the one with the larger index value on the interpolation path, and that basis is taken as the current basis for the next iteration.
-After implementing this interruption, the tracing plot with the same configuration is shown on the lower panel.
-We can observe that rather than interpolating to the target basis at `id = 62`, the interpolation stops before the index value starts to decrease at `id = 60`.
+An interruption is then constructed to accept the interpolating bases up to the one with the largest index value and use that basis as the current basis for the next searching stage.
+After implementing this interruption, the tracing plot with the same configuration is shown on the right panel of Figure \ref{interruption}.
+In the third `tries`, rather than interpolating fully to the target basis at `id = 62`, it stops before the index value starts to decrease at `id = 60`.
 This implementation results in a higher index value in the end with fewer steps.
@@
 
-*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{fig:interruption}, is smooth, while this may not be the case for more complicated index function.
+*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
 `kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF).
-Several visualisations below shows some characteristic of this index function.
+Several diagnostic visualisations below show the characteristic of this index function.
 
-Figure \ref{fig:geodesic-better} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better`.
-Rather than a smooth interpolation, the interpolation path for `kol_cdf` shows a zig-zag pattern.
-Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method.
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better` on 1D projection for `kol_cdf` index and the interpolation path shows a zig-zag pattern.
+Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method [the order of the examples needs to be rearranged since I haven't introduced polishing here.
+xxx].
 
-```{r geodesic-better,  fig.asp=0.5, fig.cap = "Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisy index: kol_cdf.
+```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisier index: kol_cdf.
 The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
@@
 
-Moving from a 1D projection on a dataset with only one informative variable, we now introduce another informative variable in the data and remain 1D projection.
-In this problem, there are two local maximum being `r calc_kol_cdf(matrix(c(0, 1, 0 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 1, 0, 0, 0, 0]$ and `r calc_kol_cdf(matrix(c(0, 0, 1 ,0, 0, 0)), data = data_mult[,c(1,2,7:10)])` when the projection matrix is $[0, 0, 1 ,0, 0, 0]$.
-As in Figure \ref{fig:1d-2var-different-seeds}, using `search_better` with different seeds, both local maximum can be found.
-A simulated annealing algorithm, `search_better_random` in Figure \ref{fig:1d-2var-better-random}, can always find the global maximum at a cost of a large number of `tries`.
+The second example for `kol_cdf` index uses 1D projection on the larger dataset.
+Since there are two informative variables and the projection is 1D, there are two local maximum  when the projection matrix at $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
+As in Figure \ref{1d-2var-different-seeds}, different local maximum can be found using `search_better` with different seeds while `search_better_random` can always find the global maximum, as in Figure \ref{1d-2var-better-random}, although at a high cost of number of points evaluated.
 
-```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
+```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="\\label{1d-2var-different-seeds}The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
 Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
@@
 
-```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
+```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "\\label{1d-2var-better-random}Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
 kol_cdf_1d_2var_better_random %>% 
@@
 
-At previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e.
+<!-- In the previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e.
 color, could be added to compare other parameters in the global object.
-Two examples are shown below to explore and compare different searching methods and neighbourhood parameter alpha.
+-->
 
-*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem while this may not be the case in real application.
+<!-- Two examples are shown below to explore and compare different searching methods and neighbourhood parameter alpha.  -->
+
+*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem while this may not be true in real application.
 This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
@@
 `search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`.
-Among the `n_sample` basis, the one with the largest index value becomes the candidate.
+Among the `n_sample` basis, the one with the largest index value becomes the candidate basis.
 If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration.
@@
 4) the number of iteration can't exceed the `max.tries`
+
+[should the stopping criteria be more detailed? xxx]
 
@@
 `max.tries` is a hyperparameter that controls the maximum number of try without improvement and its default value is 25.
-As shown in Figure \ref{fig:trace-compare}, after polishing, both trials attain the same index value.
-However, a small `max.tries` of 25 is not sufficient for the algorithm to find the true maximum.
-This is because 25 tries is not sufficient for the 2D searching space.
+As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while a small `max.tries` of 25 is not sufficient for `search_better` to find the global maximum.
+The `max.tries` argument needs to be adjusted to ensure it is sufficient to explore the parameter space.
 
@@
 
-*Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in search_better & search_posse.
+<!-- *Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in search_better & search_posse. -->
 
-```{r nrow, echo = TRUE}
-# nrow(holes_2d_better_max_tries)
-# nrow(holes_2d_pos)
-```
+<!-- ```{r nrow, echo = TRUE} -->
+<!-- # nrow(holes_2d_better_max_tries) -->
+<!-- # nrow(holes_2d_pos) -->
+<!-- ``` -->
 
-```{r alpha, fig.asp=0.5,eval = FALSE}
-p1 <- holes_2d_better_max_tries %>%
-  filter(info != "interpolation") %>%
-  mutate(id = row_number()) %>%
-  explore_trace_parameter(var = alpha) +
-  ylim(0.46, 0.505) + ggtitle("search_better")
+<!-- ```{r alpha, fig.asp=0.5,eval = FALSE} -->
+<!-- p1 <- holes_2d_better_max_tries %>% -->
+<!--   filter(info != "interpolation") %>% -->
+<!--   mutate(id = row_number()) %>% -->
+<!--   explore_trace_parameter(var = alpha) + -->
+<!--   ylim(0.46, 0.505) + ggtitle("search_better") -->
 
-p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+<!-- p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96) -->
 
-p2 <- holes_2d_pos %>%
-  filter(info != "interpolation") %>%
-  mutate(id = row_number()) %>%
-  explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse")
+<!-- p2 <- holes_2d_pos %>% -->
+<!--   filter(info != "interpolation") %>% -->
+<!--   mutate(id = row_number()) %>% -->
+<!--   explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse") -->
 
-p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96)
+<!-- p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96) -->
 
-(p1 | p2) / (p3 | p4) + plot_layout(guides = "collect")
-```
+<!-- (p1 | p2) / (p3 | p4) + plot_layout(guides = "collect") -->
+<!-- ``` -->
 
@@
 
-Sometimes, a parameter to explore could be a vector or matrix instead of scalar, for example, in tour, we are interested to explore the projection basis in the vector space.
-This imposes difficulties in visualisation since we are bounded to perceive at most three dimensions.
+The parameter to explore can be a vector or matrix instead of scalar, for example, the projection basis in tour.
+This imposes difficulties in visualisation since humans are bounded to perceive at most three dimensions in a plot.
 Thus, principal component analysis is used to reduce the dimension of projection bases and the first two principal components are mapped to the x and y axis of the plot.
-Additional variable of interest could be mapped to the color aesthetics to for exploration.
+Additional variable of interest could be mapped to the color aesthetics for exploration.
 
-*Example: understand search_geodesic via mapping `info` to color* `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm works.
+*Example: understand search_geodesic* [feel like this example is merely explaining search geodesic algorithm, so maybe introduce the animated plot here?
+xxx] `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
 Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
 The best one is picked and a line search is then run on the geodesic direction to find the target basis.
-The starting and target basis are then interpolated.
+The starting and target bases are then interpolated.
 In the next iteration, the target basis becomes the current basis and then procedures continues.
-[should probably reword this part with info levels xxx]
 
@@
 
-*Example: initial value for polishing alpha in search geodesic* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
+*Example: initial value for polishing alpha* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
 Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value.
-A default initial value for polishing step is 0.5 and we are interested in whether this is an appropriate initial value.
+The default initial value for polishing step is 0.5 and we are interested in whether this is an appropriate initial value to use after `search_geodesic`.
 The problem is a 1D projection of the small dataset using `search_geodesic` and followed by `search_polish`.
-Figure \ref{fig:polish-alpha} display all projection bases on the first two principal components, colored by the `polish_alpha` parameter in the polishing step.
-Rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, it searches a much larger vector space that are unnecessary.
-Thus a user-supplied smaller initial value for `polish_alpha` would be ideal.
+The top-left panel of Figure \ref{polish-alpha} displays all the projection bases on the first two principal components, colored by the `polish_alpha`.
+We can observe that rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, `search_polish` searches a much larger vector space, which is unnecessary.
+Thus a customised smaller initial value for `polish_alpha` would be ideal.
+One way to do this is to initialised `polish_alpha` as the projection distance between the last two target bases.
+The top-right panel of Figure \ref{polish-alpha} shows a more desirable concentrated searching space near the ending basis.
+Both specifications of initial value allow the searches to reach the same ending index values.
 
-On the right panel of Figure \ref{fig:polish-alpha-tracing}, the `polish_alpha` is initialised as the projection distance between the last two target bases and we observe a concentrated searching space near the ending basis.
-With the supplied initial value for `polish_alpha`, the algorithm reaches the same ending index value with fewer iterations.
-
-```{r polish-alpha, fig.cap = "PCA plot of two different polish alpha initialisations.
-A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis."}
+```{r polish-alpha, fig.cap = "\\label{polish-alpha}PCA plot of two different polish alpha initialisations.
+A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis.
+Both intialisations reach the same ending index values.
+"}
 p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
@@
 
-(p1 | p2) & theme(legend.position = "none")
+((p1 | p2) /(p3 | p4)) +  plot_layout(heights = c(2, 1)) & theme(legend.position = "none")
 
@@
 
-```{r polish-alpha-tracing, fig.asp = 0.5,fig.cap="Comparing the trace of two different polish alpha initialisations. Both intialisations reach the same ending index values "}
-(p3 | p4) & theme(legend.position = "none")
-
-```
-
-# Real-time animated diagnostic plots
+# Animated diagnostic plots
 

### 2020-06-09 paper/paper.tex
more text editing - whole paper except 1.3 & 1.4

@@
 \begin{abstract}
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  check consistency of using PP for projection pursuit and PPI for
-  projection pursuit index
-\end{enumerate}
+
 \end{abstract}
@@
 complicated practical problem. Coined by \citet{hooke1961direct}, direct
-search methods don't require any gradient or hessian information and
+search methods don't require any gradient or Hessian information and
 evaluate \(f\) directly. A well-known example of it is the Nelder-Mead
@@
 of PPI function as \emph{index value}. In the literature, the indices
-being proposed include lengendre index \citep{friedman1974projection},
+being proposed include Legendre index \citep{friedman1974projection},
 hermite index \citep{hall1989polynomial}, natural hermite index
@@
 of bases forms a tour path and the projection of the data on these bases
-is played in real time. This search-and-interpolateprocess is repeated
+is played in real time. This search-and-interpolate process is repeated
 until no basis can be found with higher index value. Modified from
@@
   primary interest is to find the global maximum, local maximum could
-  also reeal structures in the data that are of our interest.
+  also reveal structures in the data that are of our interest.
 \end{itemize}
@@
 
-Random search methods has a black-box machinism and focuses solely on
+Random search methods has a black-box mechanism and focuses solely on
 finding the global maximum point, while the projection pursuit problem
-we have is aiming at \emph{exploring} the data and thus is interested in
-how the algorithm finds its maximum. This motivates us \textbf{to
-develop a visual diagnostic system for exploring the optimisation
-searching path}.
+we have aims at \emph{exploring} the data and thus is interested in how
+the algorithm finds its maximum. This motivates us \textbf{to develop a
+visual diagnostic system for exploring the optimisation searching path}.
 
 The necessity of developing such a system rather than simply producing
-different diagnostic plots is because there is always needs to develop a
-function specific for diagnosing one particular variables and these
-functions developed are not reusable at all. Thus we want to establish a
-set of rules that prescribe the diagnostic of iterative algorithms
-systematically.
+different diagnostic plots is because the diagnostics of each variable
+requires a different function and these functions can't be scaled to
+other problems. Thus we want to establish a set of rules that can
+generalise the diagnostic of iterative algorithms.
 
-The idea of unifying all the diagnostic plots under one framework is
-inspired by the concept of \emph{gramma of
+The idea of generalising all the diagnostic plots under one framework is
+inspired by the concept of \emph{grammar of
 graphic}\citep{wickham2010layered}, which powers the primary graphical
-system in R, ggplot2 \citep{ggplot2}. In gramma of graphic, plots are
+system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
 not defined by its appearance (i.e.~boxplot, histogram, scatter plot
@@
 develop a gazillion of functions that each produces a different type of
-plot, instead, it defines layers and allow users to specify the plot
-through these defined layers.
+plot. Instead, it creates aesthetic mapping from the variables to the
+geometric objects and builds the plot through different layers.
 
-{[}Do I need this paragraph to details the gramma of graphic? xxx{]}
+{[}Do I need this paragraph to details the grammar of graphic? xxx{]}
 
-\hypertarget{glocal-object}{%
-\subsection{Glocal object}\label{glocal-object}}
+\hypertarget{global-object}{%
+\subsection{Global object}\label{global-object}}
 
-Ggplot prescribes how to make a graphic but it doesn't prescribe how to
-run a diagnostic. {[}need to have a paragraph or two here to describe
-how I come up with the design of four components in the global object -
-need to related to the component in optimisation. xxx{]}.
+Ggplot requires a data frame that contains all the variables to plot and
+a \emph{global object} is constructed as the data frame supplied to the
+visual diagnostic plots to better suit the characters of iterative
+optimisation algorithms. Given an optimisation algorithm, two primary
+variables of interest are the \emph{decision variable} and the
+\emph{value of the objective function}. \emph{iterators} indexes the
+data collected and this order has time series feature that prescribes
+the progression of the optimisation. Given the complexity of the
+algorithm, there could be multiple iterators. There are other parameters
+that can't be classified as one of the three categories but are also of
+our interest. They are defined under the fourth category: \emph{other
+parameter of interest}. This could include variables that are specific
+for one particular algorithm. Below we show an example of defining the
+global object of projection pursuit guided tour.
 
-A global object used for diagnosing an optimisation is defined with four
-components: decision variable, value of objective function, iterators
-and other parameter of interest. In the projection pursuit guided tour,
-the decision variable is the projection matrix, the value of objective
-function is index value. There are three iterators: \texttt{id} is the
-smallest ordering unit that increases by one for each observatoin;
-\texttt{tries} is a larger order unit, which is updated once a
-search-and-interpolate step is finished. A third iterator \texttt{loop}
-is used to record the number of repitition in the search step and starts
-over from one at the beginning of a new \texttt{tries}. Since there are
-different methods used in the projection pursuit guided tour, we are
-interested to compare between different methods, thus \texttt{method} is
-created as another parameter of interest. The neighbourhood parameter
-alpha is always interested in the direct search methods since it
-controls the size of the sampling space, thus \texttt{alpha} is another
-parameter of interest. \texttt{info} is created to label different stage
-in the searching process and it is the third parameter of interest. A
-sketch of the global object for projection pursuit guided tour is
-presented in Figure \ref{fig:glb-obj}.
+In the projection pursuit guided tour, the decision variable is the
+projection matrix, the value of objective function is the index value.
+There are three iterators: \texttt{id} is the smallest ordering unit
+that increases by one for each observation; \texttt{tries} is updated
+once a search-and-interpolate step is finished. A third iterator
+\texttt{loop} is used to record the number of repetition in the search
+step and starts over from one at the beginning of a new \texttt{tries}.
+Three other parameters of our interest includes \texttt{method},
+\texttt{alpha} and \texttt{info}. \texttt{method} identifies the name of
+the searching method used and we are interested in comparing the
+performance between different algorithms under direct search. The
+neighbourhood parameter \texttt{alpha} controls the size of the sampling
+space and we are interested to understand how the searching space
+shrinks as the algorithm progresses. \texttt{info} labels different
+stages in the searching process. A sketch of the global object for
+projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
 
@@
 data has been scaled to ensure \texttt{x2} has variance of 1. The goal
-is to find the 1D projection of bi-modal shape, which correspondes to
-the projection matrix (vector) of \texttt{{[}0,\ 1,\ 0,\ 0,\ 0{]}}.
+is to find the 1D projection of bi-modal shape, which corresponds to the
+projection matrix (vector) of \texttt{{[}0,\ 1,\ 0,\ 0,\ 0{]}}.
 
@@
 \texttt{x7}) of different types. The distribution of all the variables
-except \texttt{x3} is plotted below and \texttt{x3} takes 500 positive
-one and 500 negative one. {[}should I introduce the dist for each var?
-xxx{]} {[}also feel like we don't really use \texttt{x3} to \texttt{x6},
-should we not mention about these? xxx{]}
+except \texttt{x3} is plotted in Figure \ref{origin-data} and
+\texttt{x3} takes 500 positive one and 500 negative one. {[}should I
+introduce the dist for each var? xxx{]} {[}also feel like we don't
+really use \texttt{x3} to \texttt{x6}, should we not mention about
+these? xxx{]}
 
+\begin{figure}
+\centering
 \includegraphics{paper_files/figure-latex/origin-data-1.pdf}
+\caption{\label{origin-data} The distribution of simulated data except
+x3}
+\end{figure}
 
-Once the dataset is sent to the tourr package, all the information
-generated will be stored in a global structure. The global structure
-consists of six columns: \texttt{basis}, \texttt{index\_val},
-\texttt{tries}, \texttt{info}, \texttt{loop}, \texttt{id} and captured
-all the basis generated during whole tour process. The example below
-presents the global object of a 1D projection of the small dataset with
-geodesic searching method.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+In tour, when the optimisation ends the global object will be stored and
+printed and can be turned off via \texttt{print\ =\ FALSE}. Additional
+messages during the optimisation can be displayed via
+\texttt{verbose\ =\ TRUE}. Below presented the global object of the 1D
+projection using the small dataset. {[}the printing is not ideal as it
+doesn't show all the columns. xxx{]}
 
@@
 
-\texttt{tries} has an increment of one once the generator is called
-(equivalently a new target basis is generated); \texttt{info} records
-the stage the basis is in. This would include the \texttt{interpolation}
-stage and the detailed stage in the optimisation i.e.
-\texttt{direction\_search}, \texttt{best\_direction\_search},
-\texttt{line\_search}and \texttt{best\_line\_search} for geodesic
-searching (\texttt{search\_geodesic}); \texttt{random\_search} and
-\texttt{new\_basis} for simulating annealing (\texttt{search\_better}).
-\texttt{loop} is the counter used for the optimisation procedure and
-thus will be \texttt{NA} for interpolation steps. \texttt{id} creates a
-sequential order of the basis. This information will be stored and
-printed when the optimisation ends and can be turned off via
-\texttt{print\ =\ FALSE}. Additional messages during the optimisation
-can be displayed via \texttt{verbose\ =\ TRUE}. Another examples is a 2D
-projection of the larger dataset with two informative variable
-(\texttt{x2} and \texttt{x7}) using search\_better method. Notice in
-this example, the dimension of the bases becomes 6 by 2.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_2d_better }\OperatorTok{%>%}\StringTok{ }\KeywordTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## # A tibble: 5 x 8
-##   basis             index_val tries info           loop method       alpha    id
-##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>        <dbl> <int>
-## 1 <dbl[,2] [6 x 2]>     0.804     1 start            NA <NA>           0.5     1
-## 2 <dbl[,2] [6 x 2]>     0.793     1 random_search     1 search_bett~   0.5     2
-## 3 <dbl[,2] [6 x 2]>     0.784     1 random_search     2 search_bett~   0.5     3
-## 4 <dbl[,2] [6 x 2]>     0.773     1 random_search     3 search_bett~   0.5     4
-## 5 <dbl[,2] [6 x 2]>     0.795     1 random_search     4 search_bett~   0.5     5
-\end{verbatim}
-
 \hypertarget{explore-scalar-parameters}{%
@@
 
-The most interesting parameter to explore is the value of objective
-function. The points recorded in the global object can be divided into
-two broad categories: searching points and interpolating points
+In tour, all the points recorded in the global object can be divided
+into two broad categories: searching points and interpolating points
 
@@
 \item
-  \emph{Searching points} include the observations that are recorded in
-  the searching algorithm in order to find the target basis. The points
-  for target bases is also included in the searching points and there is
-  one such point per \texttt{tries}.
+  \emph{Searching points} include the observations recorded in the
+  searching algorithm to find the target basis. The points for target
+  bases is also included in the searching points and there is one such
+  point per \texttt{tries}.
 \item
@@
 
-As mentioned previously, the largest difficulties of exploring searching
-points is its unknown number of observations per \texttt{tries}. Mapping
-\texttt{id} on the x-axis will leave the \texttt{tries} with few
-observations a small space in the plot, while those \texttt{tries} with
-large number of search points towards the end occupying the vast
-majority of the space in the plot.
-
-This motivates the use of summarised statistics. At each iteration,
-rather than knowing the index value of \emph{every} points, we are more
-interested to know a general summary of all the points and more
-importantly, the point with the largest \texttt{index\_val} since it
-prescribes the geodesic interpolation and future searches.
+The largest difficulties of exploring searching points is its unknown
+number of observations per \texttt{tries}. Mapping \texttt{id} on the
+x-axis will leave the \texttt{tries} with few observations a small space
+in the plot, while those \texttt{tries} with large number of search
+points (usually towards the end) occupy the vast majority of the plot
+space. This motivates to summarise the points in each \texttt{tries}. At
+each iteration, rather than knowing the index value of \emph{every}
+points, we are more interested to know a general summary of all the
+points and more importantly, the point with the largest
+\texttt{index\_val} since it prescribes the geodesic interpolation and
+future searches.
 
 Boxplot is a suitable candidate that provides five points summary of the
-data, while it has one drawback: it doesn't report the number of point
-in each box. We may risk losing information on how long it takes for the
-search to find the target basis by displaying the boxplot alone for all
+data, however it has one drawback: it doesn't report the number of point
+in each box. We may risk losing information on how many points it takes
+to find the target basis by displaying the boxplot alone for all
 \texttt{tries}. Thus, the number of point for each \texttt{tries} is
 displayed at the bottom of each box and we provide options to switch
-\texttt{tries} with small number of points to a point geometry. This is
+\texttt{tries} with small number of points to a point geometry, which is
 achieved via the \texttt{cutoff} argument. A line geometry is also added
-to link the points with the largest index value for each \texttt{tries}.
+to link the points with the largest index value in each \texttt{tries}.
 This helps to visualise the improvement made by each \texttt{tries}.
+
+\emph{Example: exploring searching points} The larger dataset is used
+for 2D projection with \texttt{search\_better} and
+\texttt{max.tries\ =\ 500}. In Figure \ref{points}, the searching points
+with \texttt{id} and \texttt{tries} on the x-axis, colored by
+\texttt{tries}. Label at the bottom indicates the number of observations
+in each tries and facilitates the choice of cutoff to switch from point
+geometry to boxplot geometry (\texttt{cutoff\ =\ 15}). The line geometry
+suggests the largest improvement happens at \texttt{tries\ =\ 5}.
 
@@
 
-\emph{Example: exploring searching points} The data is sourced from a 2D
-projection of the larger dataset and \texttt{search\_better} is used
-with \texttt{max.tries\ =\ 500}. In Figure \ref{fig:points-id} and
-\ref{fig:points}, a comparison has been made on visualising the
-searching points with \texttt{id} and \texttt{tries} on the x-axis,
-colored by \texttt{tries}. In Figure \ref{fig:points-id}, the searching
-points of the first few \texttt{tries} are squeezed in a small width
-leaving the large uninteresting searching points in the last three
-\texttt{tries} taking a vast majority of the plotting space. While in
-Figure \ref{fig:points} the data is spaced by \texttt{tries} evenly in
-the plot. Label at the bottom indicates the number of observations in
-each tries and facilitates the choice of cutoff to switch from point
-geometry to boxplot geometry (\texttt{cutoff\ =\ 15}). The line geometry
-suggest the largest improvement happens at \texttt{tries\ =\ 5}.
-
 \hypertarget{explore-interpolating-points}{%
@@
 diagnose characteristics of different configurations and index
-functions. Here we present two examples of using plots to diagnose the
-tour algorithm and different index functions.
+functions. Here we present two examples of using plots to diagnose tour
+algorithms and different index functions.
 
@@
 for a 2D projection on the larger dataset using the \texttt{holes}
-index. As mentioned previously, the interpolation starts from the
-current basis to the target basis, which has been found by the
-projection pursuit algorithm to have a higher index value. After the
-interpolation, the target basis will become the current basis and send
-back to the projection pursuit algorithm to find the next target basis.
-From figure \ref{fig:interruption}, it is possible that there are bases
-with index value higher than the target basis on the interpolation path
-and these bases could be used to search for new basis in the next
-iteration.
+index. In the interpolation stage, continuous frames will be constructed
+to bridge the project on the current basis to the target basis. The
+target basis will become the current basis in the next iteration and the
+searching stage continues to find the next target basis. In the left
+panel of Figure \ref{interruption}, we observe that there are
+interpolating bases with higher index values than the target basis and
+these bases could be used to search for new basis in the next searching
+stage.
 
-Thus an interruption is constructed to accept the interpolating bases up
-to the one with the larger index value on the interpolation path, and
-that basis is taken as the current basis for the next iteration. After
-implementing this interruption, the tracing plot with the same
-configuration is shown on the lower panel. We can observe that rather
-than interpolating to the target basis at \texttt{id\ =\ 62}, the
-interpolation stops before the index value starts to decrease at
-\texttt{id\ =\ 60}. This implementation results in a higher index value
-in the end with fewer steps.
+An interruption is then constructed to accept the interpolating bases up
+to the one with the largest index value and use that basis as the
+current basis for the next searching stage. After implementing this
+interruption, the tracing plot with the same configuration is shown on
+the right panel of Figure \ref{interruption}. In the third
+\texttt{tries}, rather than interpolating fully to the target basis at
+\texttt{id\ =\ 62}, it stops before the index value starts to decrease
+at \texttt{id\ =\ 60}. This implementation results in a higher index
+value in the end with fewer steps.
 
@@
 \emph{Example: Noisy index function} The interpolation path of holes
-index, as seen in Figure \ref{fig:interruption}, is smooth, while this
-may not be the case for more complicated index function.
-\texttt{kol\_cdf} index, an index function based on Kolmogorov test,
-compares the difference between a projection matrix and a randomly
-generated normal distribution based on cumulated distribution function
-(CDF). Several visualisations below shows some characteristic of this
-index function.
+index, as seen in Figure \ref{interruption}, is smooth, while this may
+not be the case for more complicated index functions. \texttt{kol\_cdf}
+index, an index function based on Kolmogorov test, compares the
+difference between a projection matrix and a randomly generated normal
+distribution based on cumulated distribution function (CDF). Several
+diagnostic visualisations below show the characteristic of this index
+function.
 
-Figure \ref{fig:geodesic-better} compares the tracing plot of the
-interpolating points for \texttt{search\_geodesic} and
-\texttt{search\_better}. Rather than a smooth interpolation, the
-interpolation path for \texttt{kol\_cdf} shows a zig-zag pattern.
-Polishing step has done much more work to reach the final index value
-for \texttt{search\_geodesic} than \texttt{search\_better} and this
-indicates that noisy index's favour of a random search method than
-ascent method.
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating
+points for \texttt{search\_geodesic} and \texttt{search\_better} on 1D
+projection for \texttt{kol\_cdf} index and the interpolation path shows
+a zig-zag pattern. Polishing step has done much more work to reach the
+final index value for \texttt{search\_geodesic} than
+\texttt{search\_better} and this indicates that noisy index's favour of
+a random search method than ascent method {[}the order of the examples
+needs to be rearranged since I haven't introduced polishing here.
+xxx{]}.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/geodesic-better-1.pdf}
-\caption{Comparison of two different searching methods: search\_geodesic
-and search\_better on 1D projection problem for a noisy index: kol\_cdf.
-The geodesic search rely heavily on the polishing step to find the final
-index value while search better works well.}
+\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
+\caption{\label{kol-cdf}Comparison of two different searching methods:
+search\_geodesic and search\_better on 1D projection problem for a
+noisier index: kol\_cdf. The geodesic search rely heavily on the
+polishing step to find the final index value while search better works
+well.}
 \end{figure}
 
-Moving from a 1D projection on a dataset with only one informative
-variable, we now introduce another informative variable in the data and
-remain 1D projection. In this problem, there are two local maximum being
-0.17 when the projection matrix is \([0, 1, 0, 0, 0, 0]\) and 0.228 when
-the projection matrix is \([0, 0, 1 ,0, 0, 0]\). As in Figure
-\ref{fig:1d-2var-different-seeds}, using \texttt{search\_better} with
-different seeds, both local maximum can be found. A simulated annealing
-algorithm, \texttt{search\_better\_random} in Figure
-\ref{fig:1d-2var-better-random}, can always find the global maximum at a
-cost of a large number of \texttt{tries}.
+The second example for \texttt{kol\_cdf} index uses 1D projection on the
+larger dataset. Since there are two informative variables and the
+projection is 1D, there are two local maximum when the projection matrix
+at \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\). As in Figure
+\ref{1d-2var-different-seeds}, different local maximum can be found
+using \texttt{search\_better} with different seeds while
+\texttt{search\_better\_random} can always find the global maximum, as
+in Figure \ref{1d-2var-better-random}, although at a high cost of number
+of points evaluated.
 
@@
 \includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
-\caption{The trace plot search better in a 1D projection problem with
-two informative variables using different seeds (without polishing).
-Since there are two informative variables, setting different value for
-seed will lead search better to find either of the local maximum.}
+\caption{\label{1d-2var-different-seeds}The trace plot search better in
+a 1D projection problem with two informative variables using different
+seeds (without polishing). Since there are two informative variables,
+setting different value for seed will lead search better to find either
+of the local maximum.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
-\caption{Using search better random for the problem above will result in
-finding the global maximum but much larger number of iteration is
-needed.}
+\caption{\label{1d-2var-better-random}Using search better random for the
+problem above will result in finding the global maximum but much larger
+number of iteration is needed.}
 \end{figure}
@@
 
-At previous two sections, only the iterator and the index value are
-mapped onto the x and y aesthetics of the plot; while more aesthetics
-i.e.~color, could be added to compare other parameters in the global
-object. Two examples are shown below to explore and compare different
-searching methods and neighbourhood parameter alpha.
-
 \emph{Example: Polish} In principle, all the optimisation routines
 should result in the same output on the same problem while this may not
-be the case in real application. This motivates the creation of a
-polishing search that polishes the ending basis and achieves unity on
-different methods.
+be true in real application. This motivates the creation of a polishing
+search that polishes the ending basis and achieves unity on different
+methods.
 
@@
 controlled by \texttt{polish\_alpha}. Among the \texttt{n\_sample}
-basis, the one with the largest index value becomes the candidate. If
-its index value of the candidate basis is larger than that of the
+basis, the one with the largest index value becomes the candidate basis.
+If its index value of the candidate basis is larger than that of the
 current basis, it becomes the current basis in the next iteration. If no
@@
 \end{enumerate}
+
+{[}should the stopping criteria be more detailed? xxx{]}
 
@@
 improvement and its default value is 25. As shown in Figure
-\ref{fig:trace-compare}, after polishing, both trials attain the same
-index value. However, a small \texttt{max.tries} of 25 is not sufficient
-for the algorithm to find the true maximum. This is because 25 tries is
-not sufficient for the 2D searching space.
+\ref{trace-compare}, both trials attain the same index value after
+polishing while a small \texttt{max.tries} of 25 is not sufficient for
+\texttt{search\_better} to find the global maximum. The
+\texttt{max.tries} argument needs to be adjusted to ensure it is
+sufficient to explore the parameter space.
 
@@
 
-\emph{Example: The neighbourhood parameter alpha} Add an example on
-comparing the neighbourhood parameter in search\_better \&
-search\_posse.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{# nrow(holes_2d_better_max_tries)}
-\CommentTok{# nrow(holes_2d_pos)}
-\end{Highlighting}
-\end{Shaded}
-
 \hypertarget{explore-matrix-parameter}{%
@@
 
-Sometimes, a parameter to explore could be a vector or matrix instead of
-scalar, for example, in tour, we are interested to explore the
-projection basis in the vector space. This imposes difficulties in
-visualisation since we are bounded to perceive at most three dimensions.
-Thus, principal component analysis is used to reduce the dimension of
-projection bases and the first two principal components are mapped to
-the x and y axis of the plot. Additional variable of interest could be
-mapped to the color aesthetics to for exploration.
+The parameter to explore can be a vector or matrix instead of scalar,
+for example, the projection basis in tour. This imposes difficulties in
+visualisation since humans are bounded to perceive at most three
+dimensions in a plot. Thus, principal component analysis is used to
+reduce the dimension of projection bases and the first two principal
+components are mapped to the x and y axis of the plot. Additional
+variable of interest could be mapped to the color aesthetics for
+exploration.
 
-\emph{Example: understand search\_geodesic via mapping \texttt{info} to
-color} \texttt{search\_geodesic} is a two-stage ascending algorithm with
-four different stages in the search and a PCA plot useful to understand
-how the algorithm works. Starting from the start basis, a directional
-search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and a line search is then run on the geodesic
-direction to find the target basis. The starting and target basis are
-then interpolated. In the next iteration, the target basis becomes the
-current basis and then procedures continues. {[}should probably reword
-this part with info levels xxx{]}
+\emph{Example: understand search\_geodesic} {[}feel like this example is
+merely explaining search geodesic algorithm, so maybe introduce the
+animated plot here? xxx{]} \texttt{search\_geodesic} is a two-stage
+ascending algorithm with four different stages in the search and a PCA
+plot useful to understand how the algorithm progresses and the relative
+position of each basis in the PCA projected 2D space. Starting from the
+start basis, a directional search is conducted in a narrow neighbourhood
+on five random directions. The best one is picked and a line search is
+then run on the geodesic direction to find the target basis. The
+starting and target bases are then interpolated. In the next iteration,
+the target basis becomes the current basis and then procedures
+continues.
 
@@
 
-\emph{Example: initial value for polishing alpha in search geodesic}
+\emph{Example: initial value for polishing alpha}
 \texttt{search\_polish} is a brutal-force algorithm that evaluate 1000
@@
 value for polish\_alpha would avoid wasting search on large vector space
-that are not likely to produce higher index value. A default initial
+that are not likely to produce higher index value. The default initial
 value for polishing step is 0.5 and we are interested in whether this is
-an appropriate initial value. The problem is a 1D projection of the
-small dataset using \texttt{search\_geodesic} and followed by
-\texttt{search\_polish}. Figure \ref{fig:polish-alpha} display all
-projection bases on the first two principal components, colored by the
-\texttt{polish\_alpha} parameter in the polishing step. Rather than
-concentrating on the ending basis from \texttt{search\_geodesic} as what
-polishing step is designed, it searches a much larger vector space that
-are unnecessary. Thus a user-supplied smaller initial value for
-\texttt{polish\_alpha} would be ideal.
-
-On the right panel of Figure \ref{fig:polish-alpha-tracing}, the
-\texttt{polish\_alpha} is initialised as the projection distance between
-the last two target bases and we observe a concentrated searching space
-near the ending basis. With the supplied initial value for
-\texttt{polish\_alpha}, the algorithm reaches the same ending index
-value with fewer iterations.
+an appropriate initial value to use after \texttt{search\_geodesic}. The
+problem is a 1D projection of the small dataset using
+\texttt{search\_geodesic} and followed by \texttt{search\_polish}. The
+top-left panel of Figure \ref{polish-alpha} displays all the projection
+bases on the first two principal components, colored by the
+\texttt{polish\_alpha}. We can observe that rather than concentrating on
+the ending basis from \texttt{search\_geodesic} as what polishing step
+is designed, \texttt{search\_polish} searches a much larger vector
+space, which is unnecessary. Thus a customised smaller initial value for
+\texttt{polish\_alpha} would be ideal. One way to do this is to
+initialised \texttt{polish\_alpha} as the projection distance between
+the last two target bases. The top-right panel of Figure
+\ref{polish-alpha} shows a more desirable concentrated searching space
+near the ending basis. Both specifications of initial value allow the
+searches to reach the same ending index values.
 
@@
 \includegraphics{paper_files/figure-latex/polish-alpha-1.pdf}
-\caption{PCA plot of two different polish alpha initialisations. A
-default polish alpha = 0.5 searches a larger space that is unncessary
-while a small customised initial value of polish alpha will search near
-the ending basis.}
+\caption{\label{polish-alpha}PCA plot of two different polish alpha
+initialisations. A default polish alpha = 0.5 searches a larger space
+that is unncessary while a small customised initial value of polish
+alpha will search near the ending basis. Both intialisations reach the
+same ending index values.}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/polish-alpha-tracing-1.pdf}
-\caption{Comparing the trace of two different polish alpha
-initialisations. Both intialisations reach the same ending index values}
-\end{figure}
-
-\hypertarget{real-time-animated-diagnostic-plots}{%
-\section{Real-time animated diagnostic
-plots}\label{real-time-animated-diagnostic-plots}}
+\hypertarget{animated-diagnostic-plots}{%
+\section{Animated diagnostic plots}\label{animated-diagnostic-plots}}
 

### 2020-06-09 source/geozoo.Rmd
a tiny example of geozoo

---
title: "geozoo"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(geozoo)

files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

```

```{r}
pca <- holes_1d_geo %>% ferrn::compute_pca()

interp <- pca$combined %>% filter(info == "interpolation") %>% dplyr::select(V1:V5) %>% as.matrix()

a <- sphere.hollow(p = 5, n = 1000)
a$points <- a$points %>% rbind(interp)
color <- c(rep("black", 1000), rep("red", nrow(interp)))
print(a, col = color)
```

### 2020-06-22 paper/paper.tex
Comments/questions added for discussion on Wed 24 June

@@
 \usepackage{hyperref}
+\usepackage{todonotes}
 \providecommand{\tightlist}{%
@@
 
+In an optimization problem the goal is to find the best solution
+within the space of all feasible solutions which typically is represented
+by a set of constraints. The problem consists on optimizing an objective
+function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given
+by the problem constraints to either minimize or maximize a function.... Gradient based
+optimization has been typically used to solve such problems.
+However, there are many situations where derivatives
+of an objective function are unavailable or unreliable and therefore
+traditional methods based on derivatives are not the best option to solve an optimization problem....
+\todo[inline]{Intro to optimization to set the scene for the paper?}
+
 \hypertarget{derivative-free-optimisation}{%
@@
 optimisation}\label{derivative-free-optimisation}}
+
+\todo[inline]{How about adding more details into derivative free methods?}
 
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
-\end{figure}
+%\begin{figure}
+% \includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\todo[inline]{I commented out the figure: Can you please update the path above to be a relative path instead?Thanks!}
+%\end{figure}
 
@@
 pursuit}\label{optimisation-in-projection-pursuit}}
+
+\todo[inline]{Are we using  projection pursuit/guided tour to better understand
+the convergency of optimization algorithms visually in combination with the
+algorithms discussed below? Or we are focusing on the optimisation problem
+only within the project pursuit context? Some of the problems listed below
+are also applicable to optimization problem in general too }
 
@@
 
+\todo[inline]{I would expland this section more as the core contribution}
 \hypertarget{visual-diagnostics}{%
@@
 
+\todo[inline]{I would add the maths here to describe the
+simulations}
+
 Two set of simulated data are used in the demo of the visual diagnostics
@@
 \subsection{Explore scalar parameters}\label{explore-scalar-parameters}}
+
+\todo[inline]{I was wondering if it would be worth to talk
+with a bit more detail about the tour for the audience that might
+not be familiar with it?}
 

### 2020-06-23 paper/paper.Rmd
update

@@
 
+*In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
+The problem consists on optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function....
+Gradient based optimization has been typically used to solve such problems.
+However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem....*
+
+[Intro to optimization to set the scene for the paper? ppp]
+
 ## Derivative free optimisation
@@
 The stochastic version of  direct-search method is the focus of this paper.
+
+[How about adding more details into derivative free methods? ppp]
 

### 2020-06-23 paper/paper.Rmd
add abstract and some re-arrangement of paragraphs

@@
 keywords:
-- 3 to 6 keywords
-- that do not appear in the title
-abstract: 
+- optimisation
+- projection pursuit 
+- guided tourr
+- visual 
+- diagnostics 
+- R
+abstract: Friedman & Tukey commented on their intial paper on projection pursuit in 1974 that "the  technique  use  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
+In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
+This diagnostic system can be applied to more general to help practitioner to unveil the black-box in randomised iteartive (optimisation) algorithms.
+An R package, ferrn, has been created to implement this diagnostic system.
 bibliography: biblio.bib
@@
 In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
-The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which gains its popularity due to its simplicity in use and reliability in complicated practical problem.
-Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and evaluate $f$ directly.
-A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability.
-[feel like this sentence can be expanded to include more information.
-xxx].
-Further, direct search methods can be classified as *stochastic* and *deterministic* depends on whether a random sample is used in the search.
-The stochastic version of  direct-search method is the focus of this paper.
+One common class of methods in DFO is *Direcct-search methods*.
+Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems.
+Depends on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.
+The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.
 
@@
 
-## Projection pursuit guided tour
+<!-- The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which .  -->
 
-The optimisation problem we're interested in is in the context of projection pursuit.
-Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e.
-clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-A Projection Pursuit Index (PPI) is a function of the projection matrix (a.k.a projection angle, projection basis) measuring the "interestingness" of data and we refer the value of PPI function as *index value*.
-In the literature, the indices being proposed include Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
-[Any literature on holes index?
-xxx]
-
- As @friedman1974projection noted "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." A suitable optimisation procedure is needed to find the projection angle that maximises the PPI and the quality of the optimisation largely affect the interesting projections one could possibly observe.
-
-Projection pursuit is usually used in conjunction with a tour method called *guided tour*.
-Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path*, like movie frames in real time.
-Starting from a randomly generated projection basis, projection pursuit searches for a target basis with higher index value and guided tour interpolates geodesically between the initial basis and the target basis.
-This series of bases forms a tour path and the projection of the data on these bases is played in real time.
-This search-and-interpolate process is repeated until no basis can be found with higher index value.
-Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
-Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
-
-```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
-```
-
-[doubt if I need the following. xxx]
-<!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis.
-An illustration modified from [@buja2005computational] vividly depicts the tour path in guided tour.
+<!-- A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability.
+[feel like this sentence can be expanded to include more information.
+xxx].
 -->
 
-<!-- The pseudo-code below illustrates the implementation of guided tour in the tourr package.
-Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
-This information is then used to compute a series of interpolating bases inside the `tour()` function.
-All the basis will be sent to create animation for visualising the tour in the `animate()` function.
--->
+section 1: DFO
 
-<!-- ```{r algo,eval = FALSE, echo = TRUE} -->
-<!-- animation <- function(){ -->
+section 2: idea of diagnostic system & global object
 
-<!--   # compute projection basis  -->
-<!--   tour <- function(){ -->
+section 3: general intro to tourr
 
-<!--     # construct bases on the tour path -->
-<!--     new_geodesic_path <- function(){ -->
-<!--       tour_path <- function(){ -->
+section 4: apply diagnostic system + simulated data
 
-<!--         # GENERATOR: generate projection basis via projection pursuit -->
-<!--         guided_tour <- function(){ -->
-<!--           generator <- function(){ -->
+section 5: plots
 
-<!--             # define projection pursuit index -->
-<!--             # generate the target basis from the current basis via optimisation -->
-<!--           } -->
-<!--         } -->
-
-<!--         # prepare geodesic information needed for interpolating along the tour path -->
-<!--       } -->
-<!--     } -->
-
-<!--     # INTERPOLATOR: interpolate between the current and target basis  -->
-<!--     function(){ -->
-<!--       # generate interpolating bases on the geodesic path -->
-<!--     } -->
-<!--   } -->
-
-<!--   # animate according to different display methods -->
-<!-- } -->
-<!-- ``` -->
-
-## Optimisation in projection pursuit
-
-Below listed several features characterise the optimisation procedures needed in projection pursuit
-
-- *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.
-
-- *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.
-
-- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
-
-[thinking if the following three should be presented as algorithms or plain description is fine. xxx]
-
-There are three existing methods for optimisating PPI function and we review them below.
-@posse1995projection used a stochastic direct search method,  a random search algorithm to sample new bases in the neighbourhood of the current basis defined by the radius of the p-dimensional sphere, $c$.
-The new basis is taken as the target basis if it has higher index value, or the sampling continues.
-If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved.
-The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.
-
-@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
-Since some indices could be non-differentiable, the computation of derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
-Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
-
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
-Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$.
-The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis.
-If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.
-
-## problems and difficulties in PP optimisation
+## Motivation to visaul diagnostics
 
@@
 By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot.
-Instead, it creates aesthetic mapping from the variables to the geometric objects and builds the plot through different layers.
+Instead, it aesthetically maps the variables to the geometric objects and builds the plot through different layers.
 
-[Do I need this paragraph to details the grammar of graphic? xxx]
+<!-- [Do I need this paragraph to details the grammar of graphic? xxx - don't think so; not particularly relevant - assume most people know it roughly] -->
 <!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
@@
 
+# Projection pursuit guided tour
+
+The optimisation problem we're interested in is in the context of projection pursuit.
+Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e.
+clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
+A Projection Pursuit Index (PPI) is a function of the projection matrix (a.k.a projection angle, projection basis) measuring the "interestingness" of data and we refer the value of PPI function as *index value*.
+In the literature, the indices being proposed include Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+[Any literature on holes index?
+xxx]
+
+ As @friedman1974projection noted "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." A suitable optimisation procedure is needed to find the projection angle that maximises the PPI and the quality of the optimisation largely affect the interesting projections one could possibly observe.
+
+Projection pursuit is usually used in conjunction with a tour method called *guided tour*.
+Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path*, like movie frames in real time.
+Starting from a randomly generated projection basis, projection pursuit searches for a target basis with higher index value and guided tour interpolates geodesically between the initial basis and the target basis.
+This series of bases forms a tour path and the projection of the data on these bases is played in real time.
+This search-and-interpolate process is repeated until no basis can be found with higher index value.
+Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
+Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+
+```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
+```
+
+[doubt if I need the following. xxx - maybe an algorithmatic version of guide tour would be helpful for non-expert to understand the method?]
+<!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis.
+An illustration modified from [@buja2005computational] vividly depicts the tour path in guided tour.
+-->
+
+<!-- The pseudo-code below illustrates the implementation of guided tour in the tourr package.
+Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
+This information is then used to compute a series of interpolating bases inside the `tour()` function.
+All the basis will be sent to create animation for visualising the tour in the `animate()` function.
+-->
+
+<!-- ```{r algo,eval = FALSE, echo = TRUE} -->
+<!-- animation <- function(){ -->
+
+<!--   # compute projection basis  -->
+<!--   tour <- function(){ -->
+
+<!--     # construct bases on the tour path -->
+<!--     new_geodesic_path <- function(){ -->
+<!--       tour_path <- function(){ -->
+
+<!--         # GENERATOR: generate projection basis via projection pursuit -->
+<!--         guided_tour <- function(){ -->
+<!--           generator <- function(){ -->
+
+<!--             # define projection pursuit index -->
+<!--             # generate the target basis from the current basis via optimisation -->
+<!--           } -->
+<!--         } -->
+
+<!--         # prepare geodesic information needed for interpolating along the tour path -->
+<!--       } -->
+<!--     } -->
+
+<!--     # INTERPOLATOR: interpolate between the current and target basis  -->
+<!--     function(){ -->
+<!--       # generate interpolating bases on the geodesic path -->
+<!--     } -->
+<!--   } -->
+
+<!--   # animate according to different display methods -->
+<!-- } -->
+<!-- ``` -->
+
+## Optimisation in projection pursuit
+
+[thinking if the following three should be presented as algorithms or plain description is fine. xxx]
+
+There are three existing methods for optimisating PPI function and we review them below.
+@posse1995projection used a stochastic direct search method,  a random search algorithm to sample new bases in the neighbourhood of the current basis defined by the radius of the p-dimensional sphere, $c$.
+The new basis is taken as the target basis if it has higher index value, or the sampling continues.
+If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved.
+The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.
+
+@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
+Since some indices could be non-differentiable, the computation of derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
+Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
+
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
+Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$.
+The simulated annealing algorithm works as follows.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis.
+If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
+If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.
+
+Below listed several features characterise the optimisation procedures needed in projection pursuit
+
+- *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.
+
+- *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.
+
+- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
+
+# Visual diagnostic in guided tour
+
+## Global object
+
 In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is the index value.
@@
 ```
-
-# Diagnostics plots
 
@@
 <!-- ``` -->
+
+# Diagnostics plots
 

### 2020-06-23 paper/paper.Rmd
update patricia's notes with [... ppp]

@@
 
+[Are we using  projection pursuit/guided tour to better understand the convergency of optimization algorithms visually in combination with the algorithms discussed below?
+Or we are focusing on the optimisation problem only within the project pursuit context?
+Some of the problems listed below are also applicable to optimization problem in general too.
+ppp]
+
 Below listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific for PP optimisation.
@@
 # Visual diagnostic system
+
+[I would expland this section more as the core contribution. ppp]
 
@@
 ## Simulated data
+
+[I would add the maths here to describe the simulations. ppp]
 

### 2020-06-23 paper/paper.Rmd
add a section for more general introduction

@@
 
-*In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
+In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
 The problem consists on optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function....
 Gradient based optimization has been typically used to solve such problems.
-However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem....*
+However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.
 
-[Intro to optimization to set the scene for the paper? ppp]
+Derivative free methods provides another option to optimise the objective function without evaluating any gradient or hessian information and a parciular class of methods: direct search, has gained its popularity through its conceptual simplicity.
+However, the whole searching process in the algorithm remains a black-box.
+Plots are usually used to evaluate and compare the performance of different algorithms but it can easily become tedious because the code will have to be modified significantly when comparing different parameters in the algorithms.
+For example, a categorical variable with 5 levels can be easily mapped onto color while mapping another categorical variable with 30 levels will not make the plot informative.
+Thus the plot needs to be re-designed to better suits the characteristics of the parameter (whether the parameter is a scalar or a matrix?
+whether the parameter is quantitative or categorical?
+If categorical, how many levels does the parameter have?).
+This motivates the  design of a visual diagnostic framework for optimisation algorithms based on the idea of a *global object*.
 
-## Derivative free optimisation
+The paper is organised as follows.
+Section \ref{DFO} gives a general literature review of optimisation, specifically derivative free optimisation.
+Section \ref{vis-diag} presents the new idea of constructing a
+systematic visual framework that diagnoses the components of an optimisation procedure (parameters, searching path, etc).
+The rest of the paper serves as a comprehensive example of using the visual diagnostics on one particular problem: *projection pursuit guided tour*.
+Some background knowledge of projection pursuit guided tour is provided in Section \ref{tour}.
+Section \ref{apply} applys the concepts proposed in section \ref{vis-diag} in the tour problem and sets up the data.
+The last section, Section \ref{plots}, presents the visual diagnostic plots and explains how they can help to understand different aspects of the optimisation in tour.
+
+# Derivative free optimisation {#DFO}
 
@@
 In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
-One common class of methods in DFO is *Direcct-search methods*.
+One common class of methods in DFO is *Direct-search methods*.
 Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems.
@@
 
-section 1: DFO
-
-section 2: idea of diagnostic system & global object
-
-section 3: general intro to tourr
-
-section 4: apply diagnostic system + simulated data
-
-section 5: plots
-
-## Motivation to visaul diagnostics
+## Difficulties
 
@@
 
-# Visual diagnostic system
+# Visual diagnostic system {#vis-diag}
 
@@
 
-# Projection pursuit guided tour
+# Projection pursuit guided tour {#tour}
 
@@
 
-# Visual diagnostic in guided tour
+# Visual diagnostic in guided tour {#apply}
 
@@
 
-# Diagnostics plots
+# Diagnostics plots {#plots}
 

### 2020-06-23 paper/paper.tex
add a section for more general introduction

--- 2020-06-09 paper/paper.tex
@@
 \begin{abstract}
-
+Friedman \& Tukey commented on their intial paper on projection pursuit
+in 1974 that ``the technique use for maximising the projection index
+strongly influences both the statistical and the computational aspects
+of the procedure.'' While many projection pursuit indices have been
+proposed in the literature, few concerns the optimisation procedure. In
+this paper, we developed a system of diagnostics aiming to visually
+learn how the optimisation procedures find its way towards the optimum.
+This diagnostic system can be applied to more general to help
+practitioner to unveil the black-box in randomised iteartive
+(optimisation) algorithms. An R package, ferrn, has been created to
+implement this diagnostic system.
 \end{abstract}
@@
 \noindent%
-{\it Keywords:} 3 to 6 keywords, that do not appear in the title
+{\it Keywords:} optimisation, projection pursuit, guided tourr, visual, diagnostics, R
 \vfill
@@
 
-\hypertarget{derivative-free-optimisation}{%
-\subsection{Derivative free
-optimisation}\label{derivative-free-optimisation}}
+In an optimization problem the goal is to find the best solution within
+the space of all feasible solutions which typically is represented by a
+set of constraints. The problem consists on optimizing an objective
+function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
+spaced given by the problem constraints to either minimize or maximize a
+function\ldots{}. Gradient based optimization has been typically used to
+solve such problems. However, there are many situations where
+derivatives of an objective function are unavailable or unreliable and
+therefore traditional methods based on derivatives are not the best
+option to solve an optimization problem.
+
+Derivative free methods provides another option to optimise the
+objective function without evaluating any gradient or hessian
+information and a parciular class of methods: direct search, has gained
+its popularity through its conceptual simplicity. However, the whole
+searching process in the algorithm remains a black-box. Plots are
+usually used to evaluate and compare the performance of different
+algorithms but it can easily become tedious because the code will have
+to be modified significantly when comparing different parameters in the
+algorithms. For example, a categorical variable with 5 levels can be
+easily mapped onto color while mapping another categorical variable with
+30 levels will not make the plot informative. Thus the plot needs to be
+re-designed to better suits the characteristics of the parameter
+(whether the parameter is a scalar or a matrix? whether the parameter is
+quantitative or categorical? If categorical, how many levels does the
+parameter have?). This motivates the design of a visual diagnostic
+framework for optimisation algorithms based on the idea of a
+\emph{global object}.
+
+The paper is organised as follows. Section \ref{DFO} gives a general
+literature review of optimisation, specifically derivative free
+optimisation. Section \ref{vis-diag} presents the new idea of
+constructing a systematic visual framework that diagnoses the components
+of an optimisation procedure (parameters, searching path, etc). The rest
+of the paper serves as a comprehensive example of using the visual
+diagnostics on one particular problem: \emph{projection pursuit guided
+tour}. Some background knowledge of projection pursuit guided tour is
+provided in Section \ref{tour}. Section \ref{apply} applys the concepts
+proposed in section \ref{vis-diag} in the tour problem and sets up the
+data. The last section, Section \ref{plots}, presents the visual
+diagnostic plots and explains how they can help to understand different
+aspects of the optimisation in tour.
+
+\hypertarget{DFO}{%
+\section{Derivative free optimisation}\label{DFO}}
 
@@
 Derivative-Free Optimisation (DFO) methods can be useful to approach
-these problems. The two most common methods in DFO are
-\emph{direct-search methods} and \emph{model-based methods} and this
-paper dedicates to the discussion of direct-search methods, which gains
-its popularity due to its simplicity in use and reliability in
-complicated practical problem. Coined by \citet{hooke1961direct}, direct
-search methods don't require any gradient or Hessian information and
-evaluate \(f\) directly. A well-known example of it is the Nelder-Mead
-algorithm \citep{nelder1965simplex} and it enjoys the popularity due to
-its simplicity and reliability. {[}feel like this sentence can be
-expanded to include more information. xxx{]}. Further, direct search
-methods can be classified as \emph{stochastic} and \emph{deterministic}
-depends on whether a random sample is used in the search. The stochastic
-version of direct-search method is the focus of this paper.
+these problems. One common class of methods in DFO is
+\emph{Direct-search methods}. Coined by \citet{hooke1961direct}, direct
+search methods don't require any gradient or Hessian information and has
+gained its popularity through its simplicity in use and reliability in
+solving complicated practical problems. Depends on whether a random
+sample is used in the search, this class of methods can be further
+classified as \emph{stochastic} or \emph{deterministic}. The stochastic
+version of the direct-search methods will be the main optimisation
+procedure analysed in this paper.
 
-\hypertarget{projection-pursuit-guided-tour}{%
-\subsection{Projection pursuit guided
-tour}\label{projection-pursuit-guided-tour}}
+{[}How about adding more details into derivative free methods? ppp{]}
 
-The optimisation problem we're interested in is in the context of
-projection pursuit. Coined by \citet{friedman1974projection}, projection
-pursuit is a method that detects the interesting structure
-(i.e.~clustering, outliers and skewness) of multivariate data via
-projecting it in lower dimensions. A Projection Pursuit Index (PPI) is a
-function of the projection matrix (a.k.a projection angle, projection
-basis) measuring the ``interestingness'' of data and we refer the value
-of PPI function as \emph{index value}. In the literature, the indices
-being proposed include Legendre index \citep{friedman1974projection},
-hermite index \citep{hall1989polynomial}, natural hermite index
-\citep{cook1993projection}, chi-square index
-\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
-index \citep{lee2010projection}. {[}Any literature on holes index?
-xxx{]}
+\hypertarget{difficulties}{%
+\subsection{Difficulties}\label{difficulties}}
 
-As \citet{friedman1974projection} noted ``\ldots{}, the technique use
-for maximising the projection index strongly influences both the
-statistical and the computational aspects of the procedure.'' A suitable
-optimisation procedure is needed to find the projection angle that
-maximises the PPI and the quality of the optimisation largely affect the
-interesting projections one could possibly observe.
-
-Projection pursuit is usually used in conjunction with a tour method
-called \emph{guided tour}. Tour explores the multivariate data
-\emph{interactively} via playing a series of projections, that form a
-\emph{tour path}, like movie frames in real time. Starting from a
-randomly generated projection basis, projection pursuit searches for a
-target basis with higher index value and guided tour interpolates
-geodesically between the initial basis and the target basis. This series
-of bases forms a tour path and the projection of the data on these bases
-is played in real time. This search-and-interpolate process is repeated
-until no basis can be found with higher index value. Modified from
-\citet{buja2005computational}, Figure \ref{tour-path} vividly depicts
-the tour path in guided tour. Guided tour, along with other types of
-tour, has been implemented in the \emph{tourr} package in R, available
-on the Comprehensive R Archive Network at
-\url{https://cran.r-project.org/web/packages/tourr/}
-\citep{wickham2011tourrpackage}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
-\end{figure}
-
-{[}doubt if I need the following. xxx{]}
-
-\hypertarget{optimisation-in-projection-pursuit}{%
-\subsection{Optimisation in projection
-pursuit}\label{optimisation-in-projection-pursuit}}
-
-Below listed several features characterise the optimisation procedures
-needed in projection pursuit
-
-\begin{itemize}
-\item
-  \emph{Being able to handle non-differentiable PPI function}: The PPI
-  function could be non-differentiable, thus derivative free methods are
-  preferred.
-\item
-  \emph{Being able to optimise with constraints}: The constraint comes
-  from projection matrix being an orthonormal matrix.
-\item
-  \emph{Being able to find both local and global maximum}: Although the
-  primary interest is to find the global maximum, local maximum could
-  also reveal structures in the data that are of our interest.
-\end{itemize}
-
-{[}thinking if the following three should be presented as algorithms or
-plain description is fine. xxx{]}
-
-There are three existing methods for optimisating PPI function and we
-review them below. \citet{posse1995projection} used a stochastic direct
-search method, a random search algorithm to sample new bases in the
-neighbourhood of the current basis defined by the radius of the
-p-dimensional sphere, \(c\). The new basis is taken as the target basis
-if it has higher index value, or the sampling continues. If no basis is
-found to have higher index value after a certain number of tries \(n\),
-the radius \(c\) is halved. The algorithm stops when the maximum number
-of iteration is attained or the radius \(c\) is less than a
-pre-determined number.
-
-\citet{cook1995grand} explained the use of a gradient ascent
-optimisation with the assumption that the index function is continuous
-and differentiable. Since some indices could be non-differentiable, the
-computation of derivative is replaced by a pseudo-derivative of
-evaluating five randomly generated directions in a tiny nearby
-neighbourhood. Taking a step on the straight derivative direction has
-been modified to maximise the projection pursuit index along the
-geodesic direction.
-
-Simulated annealing
-\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
-non-derivative procedure based on a non-increasing cooling scheme
-\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
-is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
-algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
-and a randomly generated orthonormal basis \(B\), a candidate basis is
-constructed as \(B_j = (1- \alpha)B_i + \alpha B\) where \(B_i\) is the
-current basis. If the index value of the candidate basis is larger than
-the one of the current basis, the candidate basis becomes the target
-basis. If it is smaller, the candidate is accepted with probability
-\(A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)\) where
-\(I(.)\) is the index function.
-
-\hypertarget{problems-and-difficulties-in-pp-optimisation}{%
-\subsection{problems and difficulties in PP
-optimisation}\label{problems-and-difficulties-in-pp-optimisation}}
+{[}Are we using projection pursuit/guided tour to better understand the
+convergency of optimization algorithms visually in combination with the
+algorithms discussed below? Or we are focusing on the optimisation
+problem only within the project pursuit context? Some of the problems
+listed below are also applicable to optimization problem in general too.
+ppp{]}
 
@@
 
-\hypertarget{visual-diagnostic-system}{%
-\section{Visual diagnostic system}\label{visual-diagnostic-system}}
+\hypertarget{vis-diag}{%
+\section{Visual diagnostic system}\label{vis-diag}}
+
+{[}I would expland this section more as the core contribution. ppp{]}
 
@@
 develop a gazillion of functions that each produces a different type of
-plot. Instead, it creates aesthetic mapping from the variables to the
-geometric objects and builds the plot through different layers.
-
-{[}Do I need this paragraph to details the grammar of graphic? xxx{]}
+plot. Instead, it aesthetically maps the variables to the geometric
+objects and builds the plot through different layers.
 
@@
 global object of projection pursuit guided tour.
+
+\hypertarget{tour}{%
+\section{Projection pursuit guided tour}\label{tour}}
+
+The optimisation problem we're interested in is in the context of
+projection pursuit. Coined by \citet{friedman1974projection}, projection
+pursuit is a method that detects the interesting structure
+(i.e.~clustering, outliers and skewness) of multivariate data via
+projecting it in lower dimensions. A Projection Pursuit Index (PPI) is a
+function of the projection matrix (a.k.a projection angle, projection
+basis) measuring the ``interestingness'' of data and we refer the value
+of PPI function as \emph{index value}. In the literature, the indices
+being proposed include Legendre index \citep{friedman1974projection},
+hermite index \citep{hall1989polynomial}, natural hermite index
+\citep{cook1993projection}, chi-square index
+\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
+index \citep{lee2010projection}. {[}Any literature on holes index?
+xxx{]}
+
+As \citet{friedman1974projection} noted ``\ldots{}, the technique use
+for maximising the projection index strongly influences both the
+statistical and the computational aspects of the procedure.'' A suitable
+optimisation procedure is needed to find the projection angle that
+maximises the PPI and the quality of the optimisation largely affect the
+interesting projections one could possibly observe.
+
+Projection pursuit is usually used in conjunction with a tour method
+called \emph{guided tour}. Tour explores the multivariate data
+\emph{interactively} via playing a series of projections, that form a
+\emph{tour path}, like movie frames in real time. Starting from a
+randomly generated projection basis, projection pursuit searches for a
+target basis with higher index value and guided tour interpolates
+geodesically between the initial basis and the target basis. This series
+of bases forms a tour path and the projection of the data on these bases
+is played in real time. This search-and-interpolate process is repeated
+until no basis can be found with higher index value. Modified from
+\citet{buja2005computational}, Figure \ref{tour-path} vividly depicts
+the tour path in guided tour. Guided tour, along with other types of
+tour, has been implemented in the \emph{tourr} package in R, available
+on the Comprehensive R Archive Network at
+\url{https://cran.r-project.org/web/packages/tourr/}
+\citep{wickham2011tourrpackage}.
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\end{figure}
+
+{[}doubt if I need the following. xxx - maybe an algorithmatic version
+of guide tour would be helpful for non-expert to understand the
+method?{]}
+
+\hypertarget{optimisation-in-projection-pursuit}{%
+\subsection{Optimisation in projection
+pursuit}\label{optimisation-in-projection-pursuit}}
+
+{[}thinking if the following three should be presented as algorithms or
+plain description is fine. xxx{]}
+
+There are three existing methods for optimisating PPI function and we
+review them below. \citet{posse1995projection} used a stochastic direct
+search method, a random search algorithm to sample new bases in the
+neighbourhood of the current basis defined by the radius of the
+p-dimensional sphere, \(c\). The new basis is taken as the target basis
+if it has higher index value, or the sampling continues. If no basis is
+found to have higher index value after a certain number of tries \(n\),
+the radius \(c\) is halved. The algorithm stops when the maximum number
+of iteration is attained or the radius \(c\) is less than a
+pre-determined number.
+
+\citet{cook1995grand} explained the use of a gradient ascent
+optimisation with the assumption that the index function is continuous
+and differentiable. Since some indices could be non-differentiable, the
+computation of derivative is replaced by a pseudo-derivative of
+evaluating five randomly generated directions in a tiny nearby
+neighbourhood. Taking a step on the straight derivative direction has
+been modified to maximise the projection pursuit index along the
+geodesic direction.
+
+Simulated annealing
+\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
+non-derivative procedure based on a non-increasing cooling scheme
+\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
+is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
+algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
+and a randomly generated orthonormal basis \(B\), a candidate basis is
+constructed as \(B_j = (1- \alpha)B_i + \alpha B\) where \(B_i\) is the
+current basis. If the index value of the candidate basis is larger than
+the one of the current basis, the candidate basis becomes the target
+basis. If it is smaller, the candidate is accepted with probability
+\(A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)\) where
+\(I(.)\) is the index function.
+
+Below listed several features characterise the optimisation procedures
+needed in projection pursuit
+
+\begin{itemize}
+\item
+  \emph{Being able to handle non-differentiable PPI function}: The PPI
+  function could be non-differentiable, thus derivative free methods are
+  preferred.
+\item
+  \emph{Being able to optimise with constraints}: The constraint comes
+  from projection matrix being an orthonormal matrix.
+\item
+  \emph{Being able to find both local and global maximum}: Although the
+  primary interest is to find the global maximum, local maximum could
+  also reveal structures in the data that are of our interest.
+\end{itemize}
+
+\hypertarget{apply}{%
+\section{Visual diagnostic in guided tour}\label{apply}}
+
+\hypertarget{global-object-1}{%
+\subsection{Global object}\label{global-object-1}}
 
@@
 
-\hypertarget{diagnostics-plots}{%
-\section{Diagnostics plots}\label{diagnostics-plots}}
-
 \hypertarget{simulated-data}{%
 \subsection{Simulated data}\label{simulated-data}}
+
+{[}I would add the maths here to describe the simulations. ppp{]}
 
@@
 \end{verbatim}
+
+\hypertarget{plots}{%
+\section{Diagnostics plots}\label{plots}}
 

### 2020-06-23 paper/paper.Rmd
fix typo

@@
 
-Derivative free methods provides another option to optimise the objective function without evaluating any gradient or hessian information and a parciular class of methods: direct search, has gained its popularity through its conceptual simplicity.
+Derivative free methods provides another option to optimise the objective function without evaluating any gradient or Hessian information and a particular class of methods: direct search, has gained its popularity through its conceptual simplicity.
 However, the whole searching process in the algorithm remains a black-box.
@@
 Some background knowledge of projection pursuit guided tour is provided in Section \ref{tour}.
-Section \ref{apply} applys the concepts proposed in section \ref{vis-diag} in the tour problem and sets up the data.
+Section \ref{apply} applies the concepts proposed in section \ref{vis-diag} in the tour problem and sets up the data.
 The last section, Section \ref{plots}, presents the visual diagnostic plots and explains how they can help to understand different aspects of the optimisation in tour.
@@
 
-[Are we using  projection pursuit/guided tour to better understand the convergency of optimization algorithms visually in combination with the algorithms discussed below?
+[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
 Or we are focusing on the optimisation problem only within the project pursuit context?
@@
 
-[I would expland this section more as the core contribution. ppp]
+[I would expand this section more as the core contribution. ppp]
 

### 2020-06-23 paper/paper.tex
fix typo

@@
 Derivative free methods provides another option to optimise the
-objective function without evaluating any gradient or hessian
-information and a parciular class of methods: direct search, has gained
+objective function without evaluating any gradient or Hessian
+information and a particular class of methods: direct search, has gained
 its popularity through its conceptual simplicity. However, the whole
@@
 tour}. Some background knowledge of projection pursuit guided tour is
-provided in Section \ref{tour}. Section \ref{apply} applys the concepts
+provided in Section \ref{tour}. Section \ref{apply} applies the concepts
 proposed in section \ref{vis-diag} in the tour problem and sets up the
@@
 {[}Are we using projection pursuit/guided tour to better understand the
-convergency of optimization algorithms visually in combination with the
+convergence of optimization algorithms visually in combination with the
 algorithms discussed below? Or we are focusing on the optimisation
@@
 
-{[}I would expland this section more as the core contribution. ppp{]}
+{[}I would expand this section more as the core contribution. ppp{]}
 

### 2020-06-23 paper/paper.Rmd
write up the skeleton of diagnostic plot in theory and rearrange examples

@@
 
-## Visual diagnostics 
+## Origin idea
 <!-- Visual diagnostics can be real-time or post-run.
@@
 
+An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
+[the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part.
+]
+
+In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
+
+- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
+The points for target bases is also included in the searching points and there is one such point per `tries`.
+
+- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
+
 <!-- The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
@@
 
+## Visual diagnostics
+
+Follow the research question we raised earlier, the purpose of visual diagnostics is to understand
+
+- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.
+
+- How does the searching space look like, that is, geometrically, where are the points located in the space.
+
+Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
+The second question can be addressed via visualising the rotating high dimensional space.
+This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time.
+The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
+Thus we can do a similar rotation on the screen to precieve even higher dimension.
+[there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
+Thus animated visualisation is needed to preceive the optimisation path in the searching space.
+
+The plot design of visual diagnostics depends on the characteristics of the variables to plot.
+If the searching points are of interest [point geom is not good; need summarise.
+While updating points are more manageable -> point geom].
+However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
+[stratefy may help?
+- may need more work here.]
+
+Exploring the value of an objective function is cna be a simple task.
+Because the iterator has a time series ordering feature, it goes onto the x-axis as convention.
+The value of hte objective function goes to the y-axis.
+Other parameter of interest can be represented using color to present more information.
+[This is the case for our purpose with method, but when the level goes higher or the parameter is quantitative, color may not be a good option.
+]
+
+An option to explore of the searching space is to explore a reduced projection 2D space via principle component analysis or other dimensional reduction methods.
+In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
+[need to think further about how different type of variables can be shown].
+While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
+Geozoo is an option.
+It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
+
 # Projection pursuit guided tour {#tour}
+
+From Section \ref{tour}, we presents a comprehensive case study on how to use the visual diagnostics to explore the optimisation in a specific problem: projection pursuit guided tour.
+Section \ref{tour} aims to provide non-experts with an overview of the problem content and the existing optimisation procedures used in projection pursuit guided tour.
+For those who are already familiar with the technique, this section can be skipped.
 
@@
 
+In this section, we present the simulated data along with how the global object can be applied in guided tour.
+
 ## Global object
@@
 
-## Explore scalar parameters
+Below we will present several examples to diagnose different aspect of the tour optimisation.
+We will first provide examples on producing 1) static plots to explore the value of objective function and 2) animated plots to explore the searching space and then a more sophisticated example that combines both to understand how to optimise a noisy and complex index function.
+In both section, we will first provide a toy example that is easy to grasp and one or two examples that help us understand the optimisation algorithm and parameter choice better.
 
-In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
+## Explore the value of objective function {#static}
 
-- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
-The points for target bases is also included in the searching points and there is one such point per `tries`.
-
-- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
-
-### Explore searching points
+### Toy example: exploring searching points
 
@@
 
-### Explore interpolating points
-
-Plotting the interpolating points as time series data allows us to diagnose characteristics of different configurations and index functions.
-Here we present two examples of using  plots to diagnose tour algorithms and different index functions.
+### A more complex example: adding an interruption in the interpolation stage
 
 ```{r interpolation}
-# points is always included in the newest version
+#points is always included in the newest version
 # p1 <- holes_2d_better %>%
@@
 #   theme(legend.position = "bottom")
-
+# 
 ```
@@
 
-*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
-`kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF).
-Several diagnostic visualisations below show the characteristic of this index function.
-
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better` on 1D projection for `kol_cdf` index and the interpolation path shows a zig-zag pattern.
-Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method [the order of the examples needs to be rearranged since I haven't introduced polishing here.
-xxx].
-
-```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisier index: kol_cdf.
-The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
-p1 <- bind_rows(kol_cdf_1d_geodesic, kol_cdf_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
-
-p2 <- bind_rows(kol_cdf_1d_better, kol_cdf_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
-
-(p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
-```
-
-The second example for `kol_cdf` index uses 1D projection on the larger dataset.
-Since there are two informative variables and the projection is 1D, there are two local maximum  when the projection matrix at $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
-As in Figure \ref{1d-2var-different-seeds}, different local maximum can be found using `search_better` with different seeds while `search_better_random` can always find the global maximum, as in Figure \ref{1d-2var-better-random}, although at a high cost of number of points evaluated.
-
-```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="\\label{1d-2var-different-seeds}The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
-Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
-p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp() + ggtitle("search_better, seed 123456")
-p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp() +  ggtitle("search_better, seed 12345")
-
-(p5 | p6) & theme(legend.position = "none") & ylim(0, 0.25)
-```
-
-```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "\\label{1d-2var-better-random}Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
-kol_cdf_1d_2var_better_random %>% 
-  explore_trace_interp() + 
-  ylim(0, 0.25) + 
-  theme(legend.position = "none") + 
-  ggtitle("search_better_random")
-```
-
-### Adding more variables to the mapping
+### Another example: polishing the final projection basis
 
@@
 
-## Explore matrix parameter
+## Explore searching space {#animated}
 
@@
 Additional variable of interest could be mapped to the color aesthetics for exploration.
+
+### A toy example: understand different stage of search_geodesic
 
@@
 ```
+
+### A more complex example: Choosing the initial value for polishing parameter
 
@@
 
-# Animated diagnostic plots
+<!-- # Animated diagnostic plots -->
 
-```{r pca-animated, echo = TRUE, eval = FALSE}
-holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) + 
-  theme(legend.position = "bottom")
+<!-- ```{r pca-animated, echo = TRUE, eval = FALSE} -->
+<!-- holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) +  -->
+<!--   theme(legend.position = "bottom") -->
+<!-- ``` -->
+
+## A comprehensive example of diagnosing a noisy index function
+
+*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
+`kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF).
+Several diagnostic visualisations below show the characteristic of this index function.
+
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better` on 1D projection for `kol_cdf` index and the interpolation path shows a zig-zag pattern.
+Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method [the order of the examples needs to be rearranged since I haven't introduced polishing here.
+xxx].
+
+```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisier index: kol_cdf.
+The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
+p1 <- bind_rows(kol_cdf_1d_geodesic, kol_cdf_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
+
+p2 <- bind_rows(kol_cdf_1d_better, kol_cdf_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
+
+(p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
+```
+
+The second example for `kol_cdf` index uses 1D projection on the larger dataset.
+Since there are two informative variables and the projection is 1D, there are two local maximum  when the projection matrix at $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
+As in Figure \ref{1d-2var-different-seeds}, different local maximum can be found using `search_better` with different seeds while `search_better_random` can always find the global maximum, as in Figure \ref{1d-2var-better-random}, although at a high cost of number of points evaluated.
+
+```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="\\label{1d-2var-different-seeds}The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
+Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
+p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp() + ggtitle("search_better, seed 123456")
+p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp() +  ggtitle("search_better, seed 12345")
+
+(p5 | p6) & theme(legend.position = "none") & ylim(0, 0.25)
+```
+
+```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "\\label{1d-2var-better-random}Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
+kol_cdf_1d_2var_better_random %>% 
+  explore_trace_interp() + 
+  ylim(0, 0.25) + 
+  theme(legend.position = "none") + 
+  ggtitle("search_better_random")
 ```

### 2020-06-23 paper/paper.tex
write up the skeleton of diagnostic plot in theory and rearrange examples

@@
 
-\hypertarget{visual-diagnostics}{%
-\subsection{Visual diagnostics}\label{visual-diagnostics}}
+\hypertarget{origin-idea}{%
+\subsection{Origin idea}\label{origin-idea}}
 
@@
 
+An optimisation procedure can easily generate hundreds or thousands
+sampling points and the question becomes what are the points we want to
+explore\ldots{} {[}the classification of searching or updating points
+goes here and people should be clear what part of the points goes to the
+plot - this is the data part not the plot part. {]}
+
+In tour, all the points recorded in the global object can be divided
+into two broad categories: searching points and interpolating points
+
+\begin{itemize}
+\item
+  \emph{Searching points} include the observations recorded in the
+  searching algorithm to find the target basis. The points for target
+  bases is also included in the searching points and there is one such
+  point per \texttt{tries}.
+\item
+  \emph{interpolating points} exist in the guided tour to produce
+  continuous animated view from one target basis to another and it
+  doesn't have \texttt{loop} value.
+\end{itemize}
+
+\hypertarget{visual-diagnostics}{%
+\subsection{Visual diagnostics}\label{visual-diagnostics}}
+
+Follow the research question we raised earlier, the purpose of visual
+diagnostics is to understand
+
+\begin{itemize}
+\item
+  Whether the procedure has provided an optimum and how the value of the
+  objective function changes through the optimisation.
+\item
+  How does the searching space look like, that is, geometrically, where
+  are the points located in the space.
+\end{itemize}
+
+Overall, the first question can be answered using a static plot with
+x-axis showing the progression of the optimisation and y-axis showing
+the value of the objective function. The second question can be
+addressed via visualising the rotating high dimensional space. This
+seems to be a hard task to see beyond 3D, but the fact is that we never
+really see all the six faces of a cube at one time. The way humans
+understand the dimension of a 3D object is either through rotating the
+physical 3D object or using shade and line type to annotate the 3D
+object in a paper or electronically. Thus we can do a similar rotation
+on the screen to precieve even higher dimension. {[}there is likely to
+be a learning curve - remember when learning geometry in school, it
+takes a while to get used to see 3D cube on a paper{]}. Thus animated
+visualisation is needed to preceive the optimisation path in the
+searching space.
+
+The plot design of visual diagnostics depends on the characteristics of
+the variables to plot. If the searching points are of interest {[}point
+geom is not good; need summarise. While updating points are more
+manageable -\textgreater{} point geom{]}. However, one must realise that
+with hundreds or thousands of searching points, exploring the sample
+space in animation could be slow and this is because of the time it
+takes to render hundreds point. {[}stratefy may help? - may need more
+work here.{]}
+
+Exploring the value of an objective function is cna be a simple task.
+Because the iterator has a time series ordering feature, it goes onto
+the x-axis as convention. The value of hte objective function goes to
+the y-axis. Other parameter of interest can be represented using color
+to present more information. {[}This is the case for our purpose with
+method, but when the level goes higher or the parameter is quantitative,
+color may not be a good option. {]}
+
+An option to explore of the searching space is to explore a reduced
+projection 2D space via principle component analysis or other
+dimensional reduction methods. In this way, the first two principle
+components will take up the x and y axis and all the other information
+will be mapped using other aesthetic attribution. {[}need to think
+further about how different type of variables can be shown{]}. While
+explore the reduced space is an initial attempt to understand the
+searching space, there are existing technology for rotating a higher
+dimensional space for visualisation. Geozoo is an option. It generates
+random points on the high dimensional space and we can overlay it with
+the points on the optimisation path to visualise the spread of it on the
+high-D sphere.
+
 \hypertarget{tour}{%
 \section{Projection pursuit guided tour}\label{tour}}
+
+From Section \ref{tour}, we presents a comprehensive case study on how
+to use the visual diagnostics to explore the optimisation in a specific
+problem: projection pursuit guided tour. Section \ref{tour} aims to
+provide non-experts with an overview of the problem content and the
+existing optimisation procedures used in projection pursuit guided tour.
+For those who are already familiar with the technique, this section can
+be skipped.
 
@@
 
+In this section, we present the simulated data along with how the global
+object can be applied in guided tour.
+
 \hypertarget{global-object-1}{%
@@
 
-\hypertarget{explore-scalar-parameters}{%
-\subsection{Explore scalar parameters}\label{explore-scalar-parameters}}
+Below we will present several examples to diagnose different aspect of
+the tour optimisation. We will first provide examples on producing 1)
+static plots to explore the value of objective function and 2) animated
+plots to explore the searching space and then a more sophisticated
+example that combines both to understand how to optimise a noisy and
+complex index function. In both section, we will first provide a toy
+example that is easy to grasp and one or two examples that help us
+understand the optimisation algorithm and parameter choice better.
 
-In tour, all the points recorded in the global object can be divided
-into two broad categories: searching points and interpolating points
+\hypertarget{static}{%
+\subsection{Explore the value of objective function}\label{static}}
 
-\begin{itemize}
-\item
-  \emph{Searching points} include the observations recorded in the
-  searching algorithm to find the target basis. The points for target
-  bases is also included in the searching points and there is one such
-  point per \texttt{tries}.
-\item
-  \emph{interpolating points} exist in the guided tour to produce
-  continuous animated view from one target basis to another and it
-  doesn't have \texttt{loop} value.
-\end{itemize}
-
-\hypertarget{explore-searching-points}{%
-\subsubsection{Explore searching
-points}\label{explore-searching-points}}
+\hypertarget{toy-example-exploring-searching-points}{%
+\subsubsection{Toy example: exploring searching
+points}\label{toy-example-exploring-searching-points}}
 
@@
 
-\hypertarget{explore-interpolating-points}{%
-\subsubsection{Explore interpolating
-points}\label{explore-interpolating-points}}
-
-Plotting the interpolating points as time series data allows us to
-diagnose characteristics of different configurations and index
-functions. Here we present two examples of using plots to diagnose tour
-algorithms and different index functions.
+\hypertarget{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}{%
+\subsubsection{A more complex example: adding an interruption in the
+interpolation
+stage}\label{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}}
 
@@
 
-\emph{Example: Noisy index function} The interpolation path of holes
-index, as seen in Figure \ref{interruption}, is smooth, while this may
-not be the case for more complicated index functions. \texttt{kol\_cdf}
-index, an index function based on Kolmogorov test, compares the
-difference between a projection matrix and a randomly generated normal
-distribution based on cumulated distribution function (CDF). Several
-diagnostic visualisations below show the characteristic of this index
-function.
-
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating
-points for \texttt{search\_geodesic} and \texttt{search\_better} on 1D
-projection for \texttt{kol\_cdf} index and the interpolation path shows
-a zig-zag pattern. Polishing step has done much more work to reach the
-final index value for \texttt{search\_geodesic} than
-\texttt{search\_better} and this indicates that noisy index's favour of
-a random search method than ascent method {[}the order of the examples
-needs to be rearranged since I haven't introduced polishing here.
-xxx{]}.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{kol-cdf}Comparison of two different searching methods:
-search\_geodesic and search\_better on 1D projection problem for a
-noisier index: kol\_cdf. The geodesic search rely heavily on the
-polishing step to find the final index value while search better works
-well.}
-\end{figure}
-
-The second example for \texttt{kol\_cdf} index uses 1D projection on the
-larger dataset. Since there are two informative variables and the
-projection is 1D, there are two local maximum when the projection matrix
-at \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\). As in Figure
-\ref{1d-2var-different-seeds}, different local maximum can be found
-using \texttt{search\_better} with different seeds while
-\texttt{search\_better\_random} can always find the global maximum, as
-in Figure \ref{1d-2var-better-random}, although at a high cost of number
-of points evaluated.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
-\caption{\label{1d-2var-different-seeds}The trace plot search better in
-a 1D projection problem with two informative variables using different
-seeds (without polishing). Since there are two informative variables,
-setting different value for seed will lead search better to find either
-of the local maximum.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
-\caption{\label{1d-2var-better-random}Using search better random for the
-problem above will result in finding the global maximum but much larger
-number of iteration is needed.}
-\end{figure}
-
-\hypertarget{adding-more-variables-to-the-mapping}{%
-\subsubsection{Adding more variables to the
-mapping}\label{adding-more-variables-to-the-mapping}}
+\hypertarget{another-example-polishing-the-final-projection-basis}{%
+\subsubsection{Another example: polishing the final projection
+basis}\label{another-example-polishing-the-final-projection-basis}}
 
@@
 
-\hypertarget{explore-matrix-parameter}{%
-\subsection{Explore matrix parameter}\label{explore-matrix-parameter}}
+\hypertarget{animated}{%
+\subsection{Explore searching space}\label{animated}}
 
@@
 exploration.
+
+\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
+\subsubsection{A toy example: understand different stage of
+search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
 
@@
 \end{figure}
+
+\hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
+\subsubsection{A more complex example: Choosing the initial value for
+polishing
+parameter}\label{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}}
 
@@
 
-\hypertarget{animated-diagnostic-plots}{%
-\section{Animated diagnostic plots}\label{animated-diagnostic-plots}}
+\hypertarget{a-comprehensive-example-of-diagnosing-noisy-index-function}{%
+\subsection{A comprehensive example of diagnosing noisy index
+function}\label{a-comprehensive-example-of-diagnosing-noisy-index-function}}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{holes_1d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{explore_proj_pca}\NormalTok{(}\DataTypeTok{animate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =}\NormalTok{ info) }\OperatorTok{+}\StringTok{ }
-\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\emph{Example: Noisy index function} The interpolation path of holes
+index, as seen in Figure \ref{interruption}, is smooth, while this may
+not be the case for more complicated index functions. \texttt{kol\_cdf}
+index, an index function based on Kolmogorov test, compares the
+difference between a projection matrix and a randomly generated normal
+distribution based on cumulated distribution function (CDF). Several
+diagnostic visualisations below show the characteristic of this index
+function.
+
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating
+points for \texttt{search\_geodesic} and \texttt{search\_better} on 1D
+projection for \texttt{kol\_cdf} index and the interpolation path shows
+a zig-zag pattern. Polishing step has done much more work to reach the
+final index value for \texttt{search\_geodesic} than
+\texttt{search\_better} and this indicates that noisy index's favour of
+a random search method than ascent method {[}the order of the examples
+needs to be rearranged since I haven't introduced polishing here.
+xxx{]}.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
+\caption{\label{kol-cdf}Comparison of two different searching methods:
+search\_geodesic and search\_better on 1D projection problem for a
+noisier index: kol\_cdf. The geodesic search rely heavily on the
+polishing step to find the final index value while search better works
+well.}
+\end{figure}
+
+The second example for \texttt{kol\_cdf} index uses 1D projection on the
+larger dataset. Since there are two informative variables and the
+projection is 1D, there are two local maximum when the projection matrix
+at \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\). As in Figure
+\ref{1d-2var-different-seeds}, different local maximum can be found
+using \texttt{search\_better} with different seeds while
+\texttt{search\_better\_random} can always find the global maximum, as
+in Figure \ref{1d-2var-better-random}, although at a high cost of number
+of points evaluated.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
+\caption{\label{1d-2var-different-seeds}The trace plot search better in
+a 1D projection problem with two informative variables using different
+seeds (without polishing). Since there are two informative variables,
+setting different value for seed will lead search better to find either
+of the local maximum.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
+\caption{\label{1d-2var-better-random}Using search better random for the
+problem above will result in finding the global maximum but much larger
+number of iteration is needed.}
+\end{figure}
 

### 2020-06-24 paper/paper.Rmd
rearrange the structure of the paper as agreed in the meeting: design and plots go tgt

@@
 
-# Visual diagnostic system {#vis-diag}
-
-[I would expand this section more as the core contribution. ppp]
-
-## Origin idea
-<!-- Visual diagnostics can be real-time or post-run.
-Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
-This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.
--->
-
-<!-- Visualisation has been widely used for exploring and understanding data. Presenting information in a graphical manner often allows people to see information they would otherwise not seen. -->
-
-Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit problem we have aims at *exploring* the data and thus is interested in how the algorithm finds its maximum.
-This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
-
-The necessity of developing such a system rather than simply producing different diagnostic plots is because the diagnostics of each variable requires a different function and these functions can't be scaled to other problems.
-Thus we want to establish a set of rules that can generalise the diagnostic of iterative algorithms.
-
-The idea of generalising all the diagnostic plots under one framework is inspired by the concept of *grammar of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
-In grammar of graphic, plots are not defined by its appearance (i.e.
-boxplot, histogram, scatter plot etc) but by "stacked layers".
-By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot.
-Instead, it aesthetically maps the variables to the geometric objects and builds the plot through different layers.
-
-<!-- [Do I need this paragraph to details the grammar of graphic? xxx - don't think so; not particularly relevant - assume most people know it roughly] -->
-<!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
-An important concept in the grammar of graphic is *aesthetic mapping*.
-Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object.
-For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot.
-Another example of creating a boxplot involves 1) mapping one variable on the x-axis and 2) transforming another variable to its five-point-summary and mapping the five-point-summary to the y-axis.
--->
-
-## Global object
-
-Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
-Given an optimisation algorithm, two primary variables of interest are the *decision variable* and the *value of the objective function*.
-*iterators* indexes the data collected and this order has time series feature that prescribes the progression of the optimisation.
-Given the complexity of the algorithm, there could be multiple iterators.
-There are other parameters that can't be classified as one of the three categories but are also of our interest.
-They are defined under the fourth category: *other parameter of interest*.
-This could include variables that are specific for one particular algorithm.
-Below we show an example of defining the global object of projection pursuit guided tour.
-
-An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
-[the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part.
-]
-
-In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
-
-- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
-The points for target bases is also included in the searching points and there is one such point per `tries`.
-
-- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
-
-<!-- The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
-The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
-The next section shows how a global object is created in projection pursuit guided tour.
--->
-
-<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
-
-## Visual diagnostics
-
-Follow the research question we raised earlier, the purpose of visual diagnostics is to understand
-
-- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.
-
-- How does the searching space look like, that is, geometrically, where are the points located in the space.
-
-Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
-The second question can be addressed via visualising the rotating high dimensional space.
-This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time.
-The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
-Thus we can do a similar rotation on the screen to precieve even higher dimension.
-[there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
-Thus animated visualisation is needed to preceive the optimisation path in the searching space.
-
-The plot design of visual diagnostics depends on the characteristics of the variables to plot.
-If the searching points are of interest [point geom is not good; need summarise.
-While updating points are more manageable -> point geom].
-However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
-[stratefy may help?
-- may need more work here.]
-
-Exploring the value of an objective function is cna be a simple task.
-Because the iterator has a time series ordering feature, it goes onto the x-axis as convention.
-The value of hte objective function goes to the y-axis.
-Other parameter of interest can be represented using color to present more information.
-[This is the case for our purpose with method, but when the level goes higher or the parameter is quantitative, color may not be a good option.
-]
-
-An option to explore of the searching space is to explore a reduced projection 2D space via principle component analysis or other dimensional reduction methods.
-In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
-[need to think further about how different type of variables can be shown].
-While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
-Geozoo is an option.
-It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
-
 # Projection pursuit guided tour {#tour}
@@
 
-# Visual diagnostic in guided tour {#apply}
+# Visual diagnostic system {#vis-diag}
 
-In this section, we present the simulated data along with how the global object can be applied in guided tour.
+[I would expand this section more as the core contribution. ppp]
+
+## Origin idea
+<!-- Visual diagnostics can be real-time or post-run.
+Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
+This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.
+-->
+
+<!-- Visualisation has been widely used for exploring and understanding data. Presenting information in a graphical manner often allows people to see information they would otherwise not seen. -->
+
+Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit problem we have aims at *exploring* the data and thus is interested in how the algorithm finds its maximum.
+This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
+
+The necessity of developing such a system rather than simply producing different diagnostic plots is because the diagnostics of each variable requires a different function and these functions can't be scaled to other problems.
+Thus we want to establish a set of rules that can generalise the diagnostic of iterative algorithms.
+
+The idea of generalising all the diagnostic plots under one framework is inspired by the concept of *grammar of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+In grammar of graphic, plots are not defined by its appearance (i.e.
+boxplot, histogram, scatter plot etc) but by "stacked layers".
+By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot.
+Instead, it aesthetically maps the variables to the geometric objects and builds the plot through different layers.
+
+<!-- [Do I need this paragraph to details the grammar of graphic? xxx - don't think so; not particularly relevant - assume most people know it roughly] -->
+<!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
+An important concept in the grammar of graphic is *aesthetic mapping*.
+Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object.
+For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot.
+Another example of creating a boxplot involves 1) mapping one variable on the x-axis and 2) transforming another variable to its five-point-summary and mapping the five-point-summary to the y-axis.
+-->
 
 ## Global object
+
+Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
+Given an optimisation algorithm, two primary variables of interest are the *decision variable* and the *value of the objective function*.
+*iterators* indexes the data collected and this order has time series feature that prescribes the progression of the optimisation.
+Given the complexity of the algorithm, there could be multiple iterators.
+There are other parameters that can't be classified as one of the three categories but are also of our interest.
+They are defined under the fourth category: *other parameter of interest*.
+This could include variables that are specific for one particular algorithm.
+Below we show an example of defining the global object of projection pursuit guided tour.
+
+An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
+[the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part.
+]
+
+In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
+
+- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
+The points for target bases is also included in the searching points and there is one such point per `tries`.
+
+- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
+
+<!-- The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
+The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
+The next section shows how a global object is created in projection pursuit guided tour.
+-->
+
+<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
 
@@
 
-## Simulated data
+### Simulated data
 
@@
 
-# Diagnostics plots {#plots}
+## Visual diagnostics
 
@@
 
-## Explore the value of objective function {#static}
+Follow the research question we raised earlier, the purpose of visual diagnostics is to understand
 
-### Toy example: exploring searching points
+- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.
+
+- How does the searching space look like, that is, geometrically, where are the points located in the space.
+
+Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
+The second question can be addressed via visualising the rotating high dimensional space.
+This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time.
+The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
+Thus we can do a similar rotation on the screen to precieve even higher dimension.
+[there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
+Thus animated visualisation is needed to preceive the optimisation path in the searching space.
+
+The plot design of visual diagnostics depends on the characteristics of the variables to plot.
+If the searching points are of interest [point geom is not good; need summarise.
+While updating points are more manageable -> point geom].
+However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
+[stratefy may help?
+- may need more work here.]
+
+### Explore the value of objective function {#static}
+
+#### Toy example: exploring searching points
 
@@
 
-### A more complex example: adding an interruption in the interpolation stage
+#### A more complex example: adding an interruption in the interpolation stage
 
@@
 
-### Another example: polishing the final projection basis
+#### Another example: polishing the final projection basis
 
@@
 
-## Explore searching space {#animated}
+Exploring the value of an objective function is cna be a simple task.
+Because the iterator has a time series ordering feature, it goes onto the x-axis as convention.
+The value of hte objective function goes to the y-axis.
+Other parameter of interest can be represented using color to present more information.
+[This is the case for our purpose with method, but when the level goes higher or the parameter is quantitative, color may not be a good option.
+]
+
+An option to explore of the searching space is to explore a reduced projection 2D space via principle component analysis or other dimensional reduction methods.
+In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
+[need to think further about how different type of variables can be shown].
+While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
+Geozoo is an option.
+It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
+
+### Explore searching space {#animated}
 
@@
 
-### A toy example: understand different stage of search_geodesic
+#### A toy example: understand different stage of search_geodesic
 
@@
 
-### A more complex example: Choosing the initial value for polishing parameter
+#### A more complex example: Choosing the initial value for polishing parameter
 
@@
 
-## A comprehensive example of diagnosing a noisy index function
+### A comprehensive example of diagnosing a noisy index function
 
@@
 
-# Vis package
+# Implementation: Ferrn pacakge
 
-Everything is coded up in a package.
+Everything is coded up in a package. Package structure
+
+# Conclusion
 

### 2020-06-24 paper/paper.tex
rearrange the structure of the paper as agreed in the meeting: design and plots go tgt

@@
 
-\hypertarget{vis-diag}{%
-\section{Visual diagnostic system}\label{vis-diag}}
-
-{[}I would expand this section more as the core contribution. ppp{]}
-
-\hypertarget{origin-idea}{%
-\subsection{Origin idea}\label{origin-idea}}
-
-Random search methods has a black-box mechanism and focuses solely on
-finding the global maximum point, while the projection pursuit problem
-we have aims at \emph{exploring} the data and thus is interested in how
-the algorithm finds its maximum. This motivates us \textbf{to develop a
-visual diagnostic system for exploring the optimisation searching path}.
-
-The necessity of developing such a system rather than simply producing
-different diagnostic plots is because the diagnostics of each variable
-requires a different function and these functions can't be scaled to
-other problems. Thus we want to establish a set of rules that can
-generalise the diagnostic of iterative algorithms.
-
-The idea of generalising all the diagnostic plots under one framework is
-inspired by the concept of \emph{grammar of
-graphic}\citep{wickham2010layered}, which powers the primary graphical
-system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
-not defined by its appearance (i.e.~boxplot, histogram, scatter plot
-etc) but by ``stacked layers''. By this design, ggplot doesn't need to
-develop a gazillion of functions that each produces a different type of
-plot. Instead, it aesthetically maps the variables to the geometric
-objects and builds the plot through different layers.
-
-\hypertarget{global-object}{%
-\subsection{Global object}\label{global-object}}
-
-Ggplot requires a data frame that contains all the variables to plot and
-a \emph{global object} is constructed as the data frame supplied to the
-visual diagnostic plots to better suit the characters of iterative
-optimisation algorithms. Given an optimisation algorithm, two primary
-variables of interest are the \emph{decision variable} and the
-\emph{value of the objective function}. \emph{iterators} indexes the
-data collected and this order has time series feature that prescribes
-the progression of the optimisation. Given the complexity of the
-algorithm, there could be multiple iterators. There are other parameters
-that can't be classified as one of the three categories but are also of
-our interest. They are defined under the fourth category: \emph{other
-parameter of interest}. This could include variables that are specific
-for one particular algorithm. Below we show an example of defining the
-global object of projection pursuit guided tour.
-
-An optimisation procedure can easily generate hundreds or thousands
-sampling points and the question becomes what are the points we want to
-explore\ldots{} {[}the classification of searching or updating points
-goes here and people should be clear what part of the points goes to the
-plot - this is the data part not the plot part. {]}
-
-In tour, all the points recorded in the global object can be divided
-into two broad categories: searching points and interpolating points
-
-\begin{itemize}
-\item
-  \emph{Searching points} include the observations recorded in the
-  searching algorithm to find the target basis. The points for target
-  bases is also included in the searching points and there is one such
-  point per \texttt{tries}.
-\item
-  \emph{interpolating points} exist in the guided tour to produce
-  continuous animated view from one target basis to another and it
-  doesn't have \texttt{loop} value.
-\end{itemize}
-
-\hypertarget{visual-diagnostics}{%
-\subsection{Visual diagnostics}\label{visual-diagnostics}}
-
-Follow the research question we raised earlier, the purpose of visual
-diagnostics is to understand
-
-\begin{itemize}
-\item
-  Whether the procedure has provided an optimum and how the value of the
-  objective function changes through the optimisation.
-\item
-  How does the searching space look like, that is, geometrically, where
-  are the points located in the space.
-\end{itemize}
-
-Overall, the first question can be answered using a static plot with
-x-axis showing the progression of the optimisation and y-axis showing
-the value of the objective function. The second question can be
-addressed via visualising the rotating high dimensional space. This
-seems to be a hard task to see beyond 3D, but the fact is that we never
-really see all the six faces of a cube at one time. The way humans
-understand the dimension of a 3D object is either through rotating the
-physical 3D object or using shade and line type to annotate the 3D
-object in a paper or electronically. Thus we can do a similar rotation
-on the screen to precieve even higher dimension. {[}there is likely to
-be a learning curve - remember when learning geometry in school, it
-takes a while to get used to see 3D cube on a paper{]}. Thus animated
-visualisation is needed to preceive the optimisation path in the
-searching space.
-
-The plot design of visual diagnostics depends on the characteristics of
-the variables to plot. If the searching points are of interest {[}point
-geom is not good; need summarise. While updating points are more
-manageable -\textgreater{} point geom{]}. However, one must realise that
-with hundreds or thousands of searching points, exploring the sample
-space in animation could be slow and this is because of the time it
-takes to render hundreds point. {[}stratefy may help? - may need more
-work here.{]}
-
-Exploring the value of an objective function is cna be a simple task.
-Because the iterator has a time series ordering feature, it goes onto
-the x-axis as convention. The value of hte objective function goes to
-the y-axis. Other parameter of interest can be represented using color
-to present more information. {[}This is the case for our purpose with
-method, but when the level goes higher or the parameter is quantitative,
-color may not be a good option. {]}
-
-An option to explore of the searching space is to explore a reduced
-projection 2D space via principle component analysis or other
-dimensional reduction methods. In this way, the first two principle
-components will take up the x and y axis and all the other information
-will be mapped using other aesthetic attribution. {[}need to think
-further about how different type of variables can be shown{]}. While
-explore the reduced space is an initial attempt to understand the
-searching space, there are existing technology for rotating a higher
-dimensional space for visualisation. Geozoo is an option. It generates
-random points on the high dimensional space and we can overlay it with
-the points on the optimisation path to visualise the spread of it on the
-high-D sphere.
-
 \hypertarget{tour}{%
@@
 
-\hypertarget{apply}{%
-\section{Visual diagnostic in guided tour}\label{apply}}
+\hypertarget{vis-diag}{%
+\section{Visual diagnostic system}\label{vis-diag}}
 
-In this section, we present the simulated data along with how the global
-object can be applied in guided tour.
+{[}I would expand this section more as the core contribution. ppp{]}
 
-\hypertarget{global-object-1}{%
-\subsection{Global object}\label{global-object-1}}
+\hypertarget{origin-idea}{%
+\subsection{Origin idea}\label{origin-idea}}
+
+Random search methods has a black-box mechanism and focuses solely on
+finding the global maximum point, while the projection pursuit problem
+we have aims at \emph{exploring} the data and thus is interested in how
+the algorithm finds its maximum. This motivates us \textbf{to develop a
+visual diagnostic system for exploring the optimisation searching path}.
+
+The necessity of developing such a system rather than simply producing
+different diagnostic plots is because the diagnostics of each variable
+requires a different function and these functions can't be scaled to
+other problems. Thus we want to establish a set of rules that can
+generalise the diagnostic of iterative algorithms.
+
+The idea of generalising all the diagnostic plots under one framework is
+inspired by the concept of \emph{grammar of
+graphic}\citep{wickham2010layered}, which powers the primary graphical
+system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
+not defined by its appearance (i.e.~boxplot, histogram, scatter plot
+etc) but by ``stacked layers''. By this design, ggplot doesn't need to
+develop a gazillion of functions that each produces a different type of
+plot. Instead, it aesthetically maps the variables to the geometric
+objects and builds the plot through different layers.
+
+\hypertarget{global-object}{%
+\subsection{Global object}\label{global-object}}
+
+Ggplot requires a data frame that contains all the variables to plot and
+a \emph{global object} is constructed as the data frame supplied to the
+visual diagnostic plots to better suit the characters of iterative
+optimisation algorithms. Given an optimisation algorithm, two primary
+variables of interest are the \emph{decision variable} and the
+\emph{value of the objective function}. \emph{iterators} indexes the
+data collected and this order has time series feature that prescribes
+the progression of the optimisation. Given the complexity of the
+algorithm, there could be multiple iterators. There are other parameters
+that can't be classified as one of the three categories but are also of
+our interest. They are defined under the fourth category: \emph{other
+parameter of interest}. This could include variables that are specific
+for one particular algorithm. Below we show an example of defining the
+global object of projection pursuit guided tour.
+
+An optimisation procedure can easily generate hundreds or thousands
+sampling points and the question becomes what are the points we want to
+explore\ldots{} {[}the classification of searching or updating points
+goes here and people should be clear what part of the points goes to the
+plot - this is the data part not the plot part. {]}
+
+In tour, all the points recorded in the global object can be divided
+into two broad categories: searching points and interpolating points
+
+\begin{itemize}
+\item
+  \emph{Searching points} include the observations recorded in the
+  searching algorithm to find the target basis. The points for target
+  bases is also included in the searching points and there is one such
+  point per \texttt{tries}.
+\item
+  \emph{interpolating points} exist in the guided tour to produce
+  continuous animated view from one target basis to another and it
+  doesn't have \texttt{loop} value.
+\end{itemize}
 
@@
 \hypertarget{simulated-data}{%
-\subsection{Simulated data}\label{simulated-data}}
+\subsubsection{Simulated data}\label{simulated-data}}
 
@@
 
-\hypertarget{plots}{%
-\section{Diagnostics plots}\label{plots}}
+\hypertarget{visual-diagnostics}{%
+\subsection{Visual diagnostics}\label{visual-diagnostics}}
 
@@
 
+Follow the research question we raised earlier, the purpose of visual
+diagnostics is to understand
+
+\begin{itemize}
+\item
+  Whether the procedure has provided an optimum and how the value of the
+  objective function changes through the optimisation.
+\item
+  How does the searching space look like, that is, geometrically, where
+  are the points located in the space.
+\end{itemize}
+
+Overall, the first question can be answered using a static plot with
+x-axis showing the progression of the optimisation and y-axis showing
+the value of the objective function. The second question can be
+addressed via visualising the rotating high dimensional space. This
+seems to be a hard task to see beyond 3D, but the fact is that we never
+really see all the six faces of a cube at one time. The way humans
+understand the dimension of a 3D object is either through rotating the
+physical 3D object or using shade and line type to annotate the 3D
+object in a paper or electronically. Thus we can do a similar rotation
+on the screen to precieve even higher dimension. {[}there is likely to
+be a learning curve - remember when learning geometry in school, it
+takes a while to get used to see 3D cube on a paper{]}. Thus animated
+visualisation is needed to preceive the optimisation path in the
+searching space.
+
+The plot design of visual diagnostics depends on the characteristics of
+the variables to plot. If the searching points are of interest {[}point
+geom is not good; need summarise. While updating points are more
+manageable -\textgreater{} point geom{]}. However, one must realise that
+with hundreds or thousands of searching points, exploring the sample
+space in animation could be slow and this is because of the time it
+takes to render hundreds point. {[}stratefy may help? - may need more
+work here.{]}
+
 \hypertarget{static}{%
-\subsection{Explore the value of objective function}\label{static}}
+\subsubsection{Explore the value of objective function}\label{static}}
 
 \hypertarget{toy-example-exploring-searching-points}{%
-\subsubsection{Toy example: exploring searching
+\paragraph{Toy example: exploring searching
 points}\label{toy-example-exploring-searching-points}}
@@
 \hypertarget{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}{%
-\subsubsection{A more complex example: adding an interruption in the
+\paragraph{A more complex example: adding an interruption in the
 interpolation
@@
 \hypertarget{another-example-polishing-the-final-projection-basis}{%
-\subsubsection{Another example: polishing the final projection
+\paragraph{Another example: polishing the final projection
 basis}\label{another-example-polishing-the-final-projection-basis}}
@@
 
+Exploring the value of an objective function is cna be a simple task.
+Because the iterator has a time series ordering feature, it goes onto
+the x-axis as convention. The value of hte objective function goes to
+the y-axis. Other parameter of interest can be represented using color
+to present more information. {[}This is the case for our purpose with
+method, but when the level goes higher or the parameter is quantitative,
+color may not be a good option. {]}
+
+An option to explore of the searching space is to explore a reduced
+projection 2D space via principle component analysis or other
+dimensional reduction methods. In this way, the first two principle
+components will take up the x and y axis and all the other information
+will be mapped using other aesthetic attribution. {[}need to think
+further about how different type of variables can be shown{]}. While
+explore the reduced space is an initial attempt to understand the
+searching space, there are existing technology for rotating a higher
+dimensional space for visualisation. Geozoo is an option. It generates
+random points on the high dimensional space and we can overlay it with
+the points on the optimisation path to visualise the spread of it on the
+high-D sphere.
+
 \hypertarget{animated}{%
-\subsection{Explore searching space}\label{animated}}
+\subsubsection{Explore searching space}\label{animated}}
 
@@
 \hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
-\subsubsection{A toy example: understand different stage of
+\paragraph{A toy example: understand different stage of
 search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
@@
 \hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
-\subsubsection{A more complex example: Choosing the initial value for
+\paragraph{A more complex example: Choosing the initial value for
 polishing
@@
 
-\hypertarget{a-comprehensive-example-of-diagnosing-noisy-index-function}{%
-\subsection{A comprehensive example of diagnosing noisy index
-function}\label{a-comprehensive-example-of-diagnosing-noisy-index-function}}
+\hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
+\subsubsection{A comprehensive example of diagnosing a noisy index
+function}\label{a-comprehensive-example-of-diagnosing-a-noisy-index-function}}
 
@@
 
-\hypertarget{vis-package}{%
-\section{Vis package}\label{vis-package}}
+\hypertarget{implementation-ferrn-pacakge}{%
+\section{Implementation: Ferrn
+pacakge}\label{implementation-ferrn-pacakge}}
 
-Everything is coded up in a package.
+Everything is coded up in a package. Package structure
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 

### 2020-06-24 source/geozoo.Rmd
animated searching space exploration for noisy function example

@@
 library(geozoo)
+library(ferrn)
 
@@
 ```{r}
+vis <- function(data, color = NULL, ...){
+
+  row <- data$basis[[1]] %>% nrow()
+  col <- data$basis[[1]] %>% ncol()
+  dim <- row * col
+
+  path <- data %>% pull(basis) %>% purrr::flatten_dbl() %>% matrix(ncol = dim, byrow = TRUE)
+
+  a <- sphere.hollow(p = dim, n = 1000)
+  a$points <- a$points %>% rbind(path)
+
+  if(is.null(color)){
+    color <- c(rep("grey", 1000), rep("red", nrow(data) - 1), "green")
+  }
+
+  print(a, col = color)
+
+}
+```
+
+1. 1D projection using search_geodesic
+
+- The final projection is colored green, with rotation, it is always posed on the opposite direction of V2 with same magnitude.
+
+```{r}
+# final basis 
+holes_1d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis)
+# final basis after polish
+holes_1d_geo_polish %>% tail(1) %>% pull(basis)
+
+dt <- holes_1d_geo %>% filter(info == "interpolation")
+color <-  c(rep("grey", 1000), rep("red", nrow(dt) - 1), "green")
+vis(dt, color = color)
+
+interp <- holes_1d_geo %>% filter(info == "interpolation")
+polish <- holes_1d_geo_polish %>% filter(info == "loop_best")
+dt_bind <- bind_rows(interp,polish)
+
+color <- c(rep("grey", 1000), rep("red", nrow(interp) - 1),"green", rep("blue", nrow(polish)))
+vis(dt_bind, color = color)
+
+```
+
+2. Comparing two search better paths
+
+```{r}
+# a small example 
+interp1 %>% tail(1) %>% pull(basis)
+interp2 %>% tail(1) %>% pull(basis)
+
+# different seeds lead to different variables found
+interp1 <- kol_cdf_1d_2var_better %>% filter(info == "interpolation")
+interp2 <- kol_cdf_1d_2var_better_2 %>% filter(info == "interpolation")
+dt_bind <- bind_rows(interp1, interp2)
+
+color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
+color <- c(rep("grey", 1000), 
+           rep(color_hex[[1]], nrow(interp1)-1), color_hex[[3]],
+           rep(color_hex[[2]], nrow(interp2)-1), color_hex[[3]])
+
+vis(dt_bind, color)
+```
+
+```{r}
+# better random finds global
+
+# explore the full parameter space 
+interp1 <- kol_cdf_1d_2var_better %>% filter(info == "interpolation")
+interp2 <- kol_cdf_1d_2var_better_2 %>% filter(info == "interpolation")
+interp3 <- kol_cdf_1d_2var_better_random %>% filter(info == "interpolation")
+
+polish1 <- kol_cdf_1d_2var_better_polish %>% filter(info == "polish_best")
+polish2 <- kol_cdf_1d_2var_better_2_polish %>% filter(info == "polish_best")
+polish3 <- kol_cdf_1d_2var_better_random_polish %>% filter(info == "polish_best")
+
+dt_bind <- bind_rows(interp1, polish1, 
+                     interp2, polish2,
+                     interp3, polish3)
+
+color_hex <- RColorBrewer::brewer.pal(8, "Dark2")
+color <- c(rep("grey", 1000), 
+           rep(color_hex[[1]], nrow(interp1) + nrow(polish1)-1), "green",
+           rep(color_hex[[2]], nrow(interp2) + nrow(polish2)-1), "green",
+           rep(color_hex[[3]], nrow(interp3) + nrow(polish3)-1), "green"
+           )
+
+vis(dt_bind, color)
+
+# explore the reduced 2D space via PCAs
+interp1 <- kol_cdf_1d_2var_better 
+interp2 <- kol_cdf_1d_2var_better_2
+interp3 <- kol_cdf_1d_2var_better_random
+
+pca <- bind_rows(interp1, interp2, interp3) %>% 
+  mutate(algorithm = c(rep("better1", nrow(interp1)), 
+                       rep("better2", nrow(interp2)), 
+                       rep("better_random", nrow(interp3)))) %>% 
+  ferrn::compute_pca()
+
+pca$combined %>% 
+  ggplot(aes(x = PC1, y = PC2, color = algorithm)) + 
+  geom_point() + 
+  theme(aspect.ratio = 1)
+
+interp3 %>% filter(tries > 40) %>% explore_trace_search()
+interp3 %>% explore_trace_search()
+```
+
+```{r}
+# finds V8
+holes_2d_better %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis)
+# finds V3 and V8
+holes_2d_better_max_tries %>% filter(info == "interpolation")%>% tail(1) %>% pull(basis)
+
+interp1 <- holes_2d_better %>% filter(info == "interpolation")
+polish1 <- holes_2d_better_polish %>% filter(info == "polish_best")
+interp2 <- holes_2d_better_max_tries %>% filter(info == "interpolation")
+polish2 <- holes_2d_better_max_tries_polish %>% filter(info == "polish_best")
+dt_bind <- bind_rows(interp1, 
+                     polish1, 
+                     interp2, 
+                     polish2
+                     )
+
+color_hex <- scales::brewer_pal(palette = "Dark2")(8)
+scales::show_col(color_hex)
+# color <- c(rep("grey", 1000), 
+#            rep(color_hex[1], nrow(interp1)-1), color_hex[2], rep(color_hex[3], nrow(polish1)-1), color_hex[4],
+#            rep(color_hex[5], nrow(interp2)-1), color_hex[6], rep(color_hex[7], nrow(polish2)-1), color_hex[8]
+#            )
+
+color <- c(rep("grey", 1000), 
+           rep(color_hex[1], nrow(interp1)-1), "green", rep("red", nrow(polish1)-1), "blue",
+           rep(color_hex[5], nrow(interp2)-1), "green", rep("lightblue", nrow(polish2)-1), "blue"
+           )
+
+dt_bind <- bind_rows(interp2, polish2)
+color <- c(rep("grey", 1000), 
+           rep(color_hex[1], nrow(interp2)), rep("red", nrow(polish2)-1), "blue")
+vis(dt_bind, color)
+
+```
+
+```{r}
 pca <- holes_1d_geo %>% ferrn::compute_pca()
 
-interp <- pca$combined %>% filter(info == "interpolation") %>% dplyr::select(V1:V5) %>% as.matrix()
+interp <- pca$combined %>% dplyr::select(V1:V5) %>% as.matrix()
 
@@
 a$points <- a$points %>% rbind(interp)
-color <- c(rep("black", 1000), rep("red", nrow(interp)))
+color <- c(rep("black", 1000), "green",rep("red", nrow(interp)-1))
 print(a, col = color)
 ```
+
+$\mathcal{S}_{p-1}$ minus one because

### 2020-07-06 paper/paper.Rmd
define projection and index function

@@
 bibliography: biblio.bib
+preamble: >
+  \usepackage{amssymb}
 output: rticles::asa_article
@@
 Section \ref{tour} aims to provide non-experts with an overview of the problem content and the existing optimisation procedures used in projection pursuit guided tour.
-For those who are already familiar with the technique, this section can be skipped.
+For those who are already familiar with the techniques, feel free to skip this section.
 
@@
 clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-A Projection Pursuit Index (PPI) is a function of the projection matrix (a.k.a projection angle, projection basis) measuring the "interestingness" of data and we refer the value of PPI function as *index value*.
-In the literature, the indices being proposed include Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+Let $D_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $P = D \cdot A$, where $A_{p\times d}$ is the projection basis and $P_{n \times p}$ is the projected data.
+Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $A$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or indice function,  that can be used to measure the "interestingness" of the projection.
+A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 [Any literature on holes index?

### 2020-07-06 paper/paper.tex
define projection and index function

@@
 %% load any required packages here
+\usepackage{amssymb}
 
@@
 existing optimisation procedures used in projection pursuit guided tour.
-For those who are already familiar with the technique, this section can
-be skipped.
+For those who are already familiar with the techniques, feel free to
+skip this section.
 
@@
 (i.e.~clustering, outliers and skewness) of multivariate data via
-projecting it in lower dimensions. A Projection Pursuit Index (PPI) is a
-function of the projection matrix (a.k.a projection angle, projection
-basis) measuring the ``interestingness'' of data and we refer the value
-of PPI function as \emph{index value}. In the literature, the indices
-being proposed include Legendre index \citep{friedman1974projection},
-hermite index \citep{hall1989polynomial}, natural hermite index
+projecting it in lower dimensions. Let \(D_{n \times p}\) be a data
+matrix, an n-d projection can be seen as a linear transformation:
+\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(P = D \cdot A\),
+where \(A_{p\times d}\) is the projection basis and \(P_{n \times p}\)
+is the projected data. Define
+\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
+that maps the projection basis \(A\) onto a scalar value \(I\), this
+function is commonly known as the projection pursuit index (PPI)
+function, or indice function, that can be used to measure the
+``interestingness'' of the projection. A number of indice functions have
+been proposed in the literature and this includes Legendre index
+\citep{friedman1974projection}, hermite index
+\citep{hall1989polynomial}, natural hermite index
 \citep{cook1993projection}, chi-square index

### 2020-07-06 paper/paper.Rmd
formation of problem using math notation

@@
 Projection pursuit is usually used in conjunction with a tour method called *guided tour*.
-Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path*, like movie frames in real time.
-Starting from a randomly generated projection basis, projection pursuit searches for a target basis with higher index value and guided tour interpolates geodesically between the initial basis and the target basis.
-This series of bases forms a tour path and the projection of the data on these bases is played in real time.
-This search-and-interpolate process is repeated until no basis can be found with higher index value.
+Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path* and guided tour uses the path that is geodesically the shortest.
+Details of the mathematical construction of a tour path can be found in @buja2005computational.
 Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
@@
 [thinking if the following three should be presented as algorithms or plain description is fine. xxx]
+
+Now we formulate the problem in mathmetical notation.
+Projection pursuit finds the projection basis $A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:
+
+\begin{align}
+I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
+\end{align}
+
+via an iterative search: $\{A^1, A^2, \cdots \cdots A^J\}$.
+Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}$$
 

### 2020-07-06 paper/paper.tex
formation of problem using math notation

@@
 \emph{interactively} via playing a series of projections, that form a
-\emph{tour path}, like movie frames in real time. Starting from a
-randomly generated projection basis, projection pursuit searches for a
-target basis with higher index value and guided tour interpolates
-geodesically between the initial basis and the target basis. This series
-of bases forms a tour path and the projection of the data on these bases
-is played in real time. This search-and-interpolate process is repeated
-until no basis can be found with higher index value. Modified from
+\emph{tour path} and guided tour uses the path that is geodesically the
+shortest. Details of the mathematical construction of a tour path can be
+found in \citet{buja2005computational}. Modified from
 \citet{buja2005computational}, Figure \ref{tour-path} vividly depicts
@@
 plain description is fine. xxx{]}
+
+Now we formulate the problem in mathmetical notation. Projection pursuit
+finds the projection basis
+\(A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where \(\mathbf{a}_i\)
+is a \(p \times 1\) vector, satisfying the following optimisation
+problem:
+
+\begin{align}
+I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
+\end{align}
+
+via an iterative search: \(\{A^1, A^2, \cdots \cdots A^J\}\). Guided
+tour interpolates geodesically between each basis found by projection
+pursuit \(A^j\) to form a tour path:
+\[\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}\]
 

### 2020-07-06 paper/paper.Rmd
an attempt to write the three algorithms in math

@@
   \usepackage{amssymb}
+  \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 output: rticles::asa_article
@@
 There are three existing methods for optimisating PPI function and we review them below.
-@posse1995projection used a stochastic direct search method,  a random search algorithm to sample new bases in the neighbourhood of the current basis defined by the radius of the p-dimensional sphere, $c$.
-The new basis is taken as the target basis if it has higher index value, or the sampling continues.
-If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved.
-The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.
+
+@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{(j +1)}$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$.
+If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops.
+If no basis is found to have higher index value after the maximum number of tries $i_{max}$, the algorithm stops with nothing outputted.
+$h$ is the halfing parameter meaning if $h > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search.
+The algorithm for a random search is summarised in Algorithm \ref{random-search}.
+
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $h = 0$, $loop = 1, i_{max}$\;
+  \While{$i < i_{max}$}{
+    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
+    $I^{(j + 1)} = f(A^{(j+1)})$\;
+    \eIf{$I^{(j+1)} > I^{j}$}{
+      $A^{j+1} = A^{(j+1)}$\;
+      }{
+      $h = h + 1$\;
+      }
+    $i = i + 1$\;
+  }
+  \caption{random search: my first algorithm}
+  \label{random-search}
+\end{algorithm}
 
@@
 
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $loop = 1, i_{max}$\;
+  \While{$i < i_{max}$}{
+    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
+    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
+    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
+    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
+    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
+    ...\;
+    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
+    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
+      \If{$p_{diff} > 0.001$}{
+        $A^{j+1} = A^{(J_2)}$\;
+        $I^{j+1} = I^{(J_2)}$\;
+        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+      }
+    $i = i + 1$\;
+  }
+  \caption{searchgeodesic}
+  \label{search_geodesic}
+\end{algorithm}
+
 Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
@@
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.
+If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$ where $I^j$ and $I^{(j+1)}$ are the index value of the current and candidate basis respectively.
+The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
+
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $T_0$, $\alpha$\;
+  \While{$i < i_{max}$}{
+    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
+    $I^{(j+1)} = f(A^{(j+1)})$\;
+    $T(i) = \frac{T_0}{\log(i + 1)}$\;
+      \eIf{$I^{(j+1)} > I^{j}$}{
+        $A^{j+1} = A^{(j+1)}$\;
+        $I^{j+1} = I^{(j+1)}$\;
+        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
+      }{
+        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
+        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
+        \If{$\text{prob} > \text{rand}$}{
+          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+        }
+      }
+    $i = i + 1$\;
+  }
+  \caption{simulated annealing}
+  \label{simulated_annealing}
+\end{algorithm}
+
+\newpage
 

### 2020-07-06 paper/paper.tex
an attempt to write the three algorithms in math

@@
 %% load any required packages here
-\usepackage{amssymb}
+\usepackage{amssymb} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 There are three existing methods for optimisating PPI function and we
-review them below. \citet{posse1995projection} used a stochastic direct
-search method, a random search algorithm to sample new bases in the
-neighbourhood of the current basis defined by the radius of the
-p-dimensional sphere, \(c\). The new basis is taken as the target basis
-if it has higher index value, or the sampling continues. If no basis is
-found to have higher index value after a certain number of tries \(n\),
-the radius \(c\) is halved. The algorithm stops when the maximum number
-of iteration is attained or the radius \(c\) is less than a
-pre-determined number.
+review them below.
+
+\citet{posse1995projection} used a stochastic direct search method, a
+random search algorithm to sample a candidate bases \(A^{(j +1)}\) in
+the neighbourhood of the current basis \(A^j\) defined by the radius of
+the p-dimensional sphere, \(\alpha\). If the candidate basis has a
+higher index value than the current basis, it is outputed as the target
+basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
+If no basis is found to have higher index value after the maximum number
+of tries \(i_{max}\), the algorithm stops with nothing outputted. \(h\)
+is the halfing parameter meaning if \(h > 30\), the nieghbourhood
+parameter \(\alpha\) will be reduced by half in the next iteration of
+the random search. The algorithm for a random search is summarised in
+Algorithm \ref{random-search}.
+
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $h = 0$, $loop = 1, i_{max}$\;
+  \While{$i < i_{max}$}{
+    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
+    $I^{(j + 1)} = f(A^{(j+1)})$\;
+    \eIf{$I^{(j+1)} > I^{j}$}{
+      $A^{j+1} = A^{(j+1)}$\;
+      }{
+      $h = h + 1$\;
+      }
+    $i = i + 1$\;
+  }
+  \caption{random search: my first algorithm}
+  \label{random-search}
+\end{algorithm}
 
@@
 
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $loop = 1, i_{max}$\;
+  \While{$i < i_{max}$}{
+    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
+    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
+    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
+    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
+    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
+    ...\;
+    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
+    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
+      \If{$p_{diff} > 0.001$}{
+        $A^{j+1} = A^{(J_2)}$\;
+        $I^{j+1} = I^{(J_2)}$\;
+        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+      }
+    $i = i + 1$\;
+  }
+  \caption{searchgeodesic}
+  \label{search_geodesic}
+\end{algorithm}
+
 Simulated annealing
@@
 algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
-and a randomly generated orthonormal basis \(B\), a candidate basis is
-constructed as \(B_j = (1- \alpha)B_i + \alpha B\) where \(B_i\) is the
-current basis. If the index value of the candidate basis is larger than
-the one of the current basis, the candidate basis becomes the target
-basis. If it is smaller, the candidate is accepted with probability
-\(A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)\) where
-\(I(.)\) is the index function.
+and a randomly generated orthonormal basis \(A_{rand}\), a candidate
+basis is constructed as
+\(A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
+of the candidate basis is larger than the one of the current basis, the
+candidate basis becomes the target basis. If it is smaller, the
+candidate is accepted with probability
+\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}\)
+where \(I^j\) and \(I^{(j+1)}\) are the index value of the current and
+candidate basis respectively. The algorithm can be summarised as in
+Algorithm \ref{simulated_annealing}.
+
+\newpage
+\begin{algorithm}[H]
+\SetAlgoLined
+  initialisation: $T_0$, $\alpha$\;
+  \While{$i < i_{max}$}{
+    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
+    $I^{(j+1)} = f(A^{(j+1)})$\;
+    $T(i) = \frac{T_0}{\log(i + 1)}$\;
+      \eIf{$I^{(j+1)} > I^{j}$}{
+        $A^{j+1} = A^{(j+1)}$\;
+        $I^{j+1} = I^{(j+1)}$\;
+        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
+      }{
+        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
+        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
+        \If{$\text{prob} > \text{rand}$}{
+          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+        }
+      }
+    $i = i + 1$\;
+  }
+  \caption{simulated annealing}
+  \label{simulated_annealing}
+\end{algorithm}
+
+\newpage
 

### 2020-07-07 paper/paper.Rmd
notation for global object

@@
 preamble: >
-  \usepackage{amssymb}
+  \usepackage{amssymb, amsmath, mathtools}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 Given an optimisation algorithm, two primary variables of interest are the *decision variable* and the *value of the objective function*.
-*iterators* indexes the data collected and this order has time series feature that prescribes the progression of the optimisation.
+To align with the notation used previously for projection pursuit, we denote the decision variable to be $\mathbf{A}$, which are matrices in projection pursuit algorithms and value of the objective function to be $I$, the index function.
+
+*Iterators* indexes the data collected and it has a time series feature that prescribes the progression of the optimisation.
 Given the complexity of the algorithm, there could be multiple iterators.
-There are other parameters that can't be classified as one of the three categories but are also of our interest.
-They are defined under the fourth category: *other parameter of interest*.
+The simpliest iterator: *index per observation*, is a unique identifier for each observation in the data.
+For each level of iteration, we design two indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
+*index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
+Depend on the number of iteration level ($L$), the total number of iterators will differ and can be calculated as $1 + 2L$.
+In projection pursuit optimisation algorithms, there is only one level of iteartion and hence exists three iterators: $t_1$, $t_2$ and $t_3$.
+A sketch of the design of iterators is provided in Figure \ref{iterators}.
+
+There could exist other parameters that are also of our interest but can't be classified as one of the three categories.
+They are defined under the fourth category: *other parameter of interest: $O$*.
 This could include variables that are specific for one particular algorithm.
 Below we show an example of defining the global object of projection pursuit guided tour.
+
+Now we formulate the idea of global object in terms of mathmetical notation.
+Superscript $j$ is used to denote the number of iteration.
+Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left.
+The output basis doesn't have subscript.
+For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$.
+In the next iteration (iteration 2), $\mathbf{A}^1$ becomes the current basis, projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$ and the tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$.
+Below we arrange all the bases in a format that each coloumn represents an iteration in Equation \ref{eq:matrixA}.
+
+When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector with its natual ordering as in Equation \ref{eq:vectorA}.
+Hence the full data structure can be written as in Equation \ref{eq:data-structure}
+
+```{r iterators, out.width = "100%", out.height = "25%", fig.cap = "\\label{iterators} A sketch of the design of iterators in iterative algorithms."}
+knitr::include_graphics(here::here("figures", "iterators.png"))
+```
+
+\begin{equation}
+\left[
+\begin{array}{ccccccc}
+/ & \mathbf{A}^1_1 & \mathbf{A}^2_1 & \cdots & \mathbf{A}^j_1 &\cdots & \mathbf{A}^J_1 \\
+/ & \mathbf{A}^1_2 & \mathbf{A}^2_2 & \cdots & \mathbf{A}^j_2 &\cdots & \mathbf{A}^J_2 \\
+\vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
+/ &\mathbf{A}^1_{l_1} & \mathbf{A}^2_{l_2} & \cdots & \mathbf{A}^j_{l_h} &\cdots & \mathbf{A}^J_{L_H}\\
+\hline
+\mathbf{A}^0 & \mathbf{A}^1 & \mathbf{A}^2 & \cdots & \mathbf{A}^j  &\cdots & \mathbf{A}^J\\
+\hline
+/ &{}_{1}\mathbf{A}^1 & {}_{1}\mathbf{A}^2 & \cdots &{}_{1}\mathbf{A}^j & \cdots & /\\
+/ &{}_{2}\mathbf{A}^1 & {}_{2}\mathbf{A}^2 & \cdots &{}_{2}\mathbf{A}^j & \cdots & / \\
+\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & / \\
+/ &{}_{k_1}\mathbf{A}^1 & {}_{k_2}\mathbf{A}^2 & \cdots &{}_{k_m}\mathbf{A}^j & \cdots & / \\
+\end{array}
+\right]
+= 
+\left[
+\begin{array}{c}
+\vphantom{A^J_1}\text{candidate basis} \\
+\vphantom{A^J_2}\text{candidate basis} \\
+\vphantom{\vdots} \vdots \\
+\vphantom{A^J_{L_H}}\text{candidate basis} \\
+\hline
+\vphantom{A^J}\text{output basis} \\
+\hline
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\vphantom{\text{NULL}}\vdots \\
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\end{array}
+\right]
+\label{eq:matrixA}
+\end{equation}
+
+\begin{equation}
+\begin{bsmallmatrix}
+\mathbf{A}^0 &\mathbf{A}^1_1 &  \mathbf{A}^1_{2} & \cdots& \mathbf{A}^1_{l_1} & \mathbf{A}^1 & {}_{1}\mathbf{A}^1& {}_{2}\mathbf{A}^1 &\cdots& {}_{k_1}\mathbf{A}^1 & \mathbf{A}^2_1 & \mathbf{A}^2_{2} & \cdots& \mathbf{A}^2_{l_2} & \mathbf{A}^2 & {}_{1}\mathbf{A}^2 &{}_{2}\mathbf{A}^2 &\cdots& {}_{k_2}\mathbf{A}^2 & \cdots & \mathbf{A}^J
+\end{bsmallmatrix} 
+\label{eq:vectorA}
+\end{equation}
+
+\begin{equation}
+\left[
+\begin{array}{cccccccc}
+\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
+\hline
+\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
+\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
+\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
+\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
+\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
+\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
+{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
+{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
+\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
+{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
+\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
+\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
+\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
+\end{array}
+\right]
+\label{eq:data-structure}
+\end{equation}
+
+where $n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$ and $n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$
 

### 2020-07-07 paper/paper.tex
notation for global object

@@
 %% load any required packages here
-\usepackage{amssymb} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 variables of interest are the \emph{decision variable} and the
-\emph{value of the objective function}. \emph{iterators} indexes the
-data collected and this order has time series feature that prescribes
-the progression of the optimisation. Given the complexity of the
-algorithm, there could be multiple iterators. There are other parameters
-that can't be classified as one of the three categories but are also of
-our interest. They are defined under the fourth category: \emph{other
-parameter of interest}. This could include variables that are specific
-for one particular algorithm. Below we show an example of defining the
-global object of projection pursuit guided tour.
+\emph{value of the objective function}. To align with the notation used
+previously for projection pursuit, we denote the decision variable to be
+\(\mathbf{A}\), which are matrices in projection pursuit algorithms and
+value of the objective function to be \(I\), the index function.
+
+\emph{Iterators} indexes the data collected and it has a time series
+feature that prescribes the progression of the optimisation. Given the
+complexity of the algorithm, there could be multiple iterators. The
+simpliest iterator: \emph{index per observation}, is a unique identifier
+for each observation in the data. For each level of iteration, we design
+two indices: \emph{index per iteration} is fixed for each observations
+in one iteration and has an increment of one once a new iteration
+starts. \emph{index within iteration} has an increment of one for each
+observation in an iteartion and starts over from one once a new
+iteration starts. Depend on the number of iteration level (\(L\)), the
+total number of iterators will differ and can be calculated as
+\(1 + 2L\). In projection pursuit optimisation algorithms, there is only
+one level of iteartion and hence exists three iterators: \(t_1\),
+\(t_2\) and \(t_3\). A sketch of the design of iterators is provided in
+Figure \ref{iterators}.
+
+There could exist other parameters that are also of our interest but
+can't be classified as one of the three categories. They are defined
+under the fourth category: \emph{other parameter of interest: \(O\)}.
+This could include variables that are specific for one particular
+algorithm. Below we show an example of defining the global object of
+projection pursuit guided tour.
+
+Now we formulate the idea of global object in terms of mathmetical
+notation. Superscript \(j\) is used to denote the number of iteration.
+Candidate basis are denoted using subscript on the right and
+interpolation basis are denoted using subscript on the left. The output
+basis doesn't have subscript. For example, \(\mathbf{A}^1_{2}\) means
+the second candidate basis in the first iteration and
+\({}_{2}\mathbf{A}^1\) means the second interpolation basis that forms
+the tour path from the current basis \(\mathbf{A}^0\) to the target
+basis \(\mathbf{A}^1\). In the next iteration (iteration 2),
+\(\mathbf{A}^1\) becomes the current basis, projection pursuit searches
+through
+\(\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}\) to
+find the target basis \(\mathbf{A}^2\) and the tour path from
+\(\mathbf{A}^1\) to \(\mathbf{A}^2\) is constructed using the
+interpolation bases
+\({}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2\).
+Below we arrange all the bases in a format that each coloumn represents
+an iteration in Equation \ref{eq:matrixA}.
+
+When aggregated with information of index value and other interesting
+parameter, it is easier to write the basis as a column vector with its
+natual ordering as in Equation \ref{eq:vectorA}. Hence the full data
+structure can be written as in Equation \ref{eq:data-structure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/iterators} \caption{\label{iterators} A sketch of the design of iterators in iterative algorithms.}\label{fig:iterators}
+\end{figure}
+
+\begin{equation}
+\left[
+\begin{array}{ccccccc}
+/ & \mathbf{A}^1_1 & \mathbf{A}^2_1 & \cdots & \mathbf{A}^j_1 &\cdots & \mathbf{A}^J_1 \\
+/ & \mathbf{A}^1_2 & \mathbf{A}^2_2 & \cdots & \mathbf{A}^j_2 &\cdots & \mathbf{A}^J_2 \\
+\vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
+/ &\mathbf{A}^1_{l_1} & \mathbf{A}^2_{l_2} & \cdots & \mathbf{A}^j_{l_h} &\cdots & \mathbf{A}^J_{L_H}\\
+\hline
+\mathbf{A}^0 & \mathbf{A}^1 & \mathbf{A}^2 & \cdots & \mathbf{A}^j  &\cdots & \mathbf{A}^J\\
+\hline
+/ &{}_{1}\mathbf{A}^1 & {}_{1}\mathbf{A}^2 & \cdots &{}_{1}\mathbf{A}^j & \cdots & /\\
+/ &{}_{2}\mathbf{A}^1 & {}_{2}\mathbf{A}^2 & \cdots &{}_{2}\mathbf{A}^j & \cdots & / \\
+\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & / \\
+/ &{}_{k_1}\mathbf{A}^1 & {}_{k_2}\mathbf{A}^2 & \cdots &{}_{k_m}\mathbf{A}^j & \cdots & / \\
+\end{array}
+\right]
+= 
+\left[
+\begin{array}{c}
+\vphantom{A^J_1}\text{candidate basis} \\
+\vphantom{A^J_2}\text{candidate basis} \\
+\vphantom{\vdots} \vdots \\
+\vphantom{A^J_{L_H}}\text{candidate basis} \\
+\hline
+\vphantom{A^J}\text{output basis} \\
+\hline
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\vphantom{\text{NULL}}\vdots \\
+\vphantom{\text{NULL}}\text{interpolation basis} \\
+\end{array}
+\right]
+\label{eq:matrixA}
+\end{equation}
+
+\begin{equation}
+\begin{bsmallmatrix}
+\mathbf{A}^0 &\mathbf{A}^1_1 &  \mathbf{A}^1_{2} & \cdots& \mathbf{A}^1_{l_1} & \mathbf{A}^1 & {}_{1}\mathbf{A}^1& {}_{2}\mathbf{A}^1 &\cdots& {}_{k_1}\mathbf{A}^1 & \mathbf{A}^2_1 & \mathbf{A}^2_{2} & \cdots& \mathbf{A}^2_{l_2} & \mathbf{A}^2 & {}_{1}\mathbf{A}^2 &{}_{2}\mathbf{A}^2 &\cdots& {}_{k_2}\mathbf{A}^2 & \cdots & \mathbf{A}^J
+\end{bsmallmatrix} 
+\label{eq:vectorA}
+\end{equation}
+
+\begin{equation}
+\left[
+\begin{array}{cccccccc}
+\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
+\hline
+\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
+\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
+\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
+\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
+\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
+\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
+{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
+{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
+\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
+{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
+\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
+\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
+\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
+\end{array}
+\right]
+\label{eq:data-structure}
+\end{equation}
+
+where \(n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\) and
+\(n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)
 

### 2020-07-07 paper/paper.Rmd
make sure notation in the algorithm before matches with the ones in the global object

@@
 Now we formulate the problem in mathmetical notation.
-Projection pursuit finds the projection basis $A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:
+Projection pursuit finds the projection basis $\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:
 
 \begin{align}
-I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
+I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
 \end{align}
 
-via an iterative search: $\{A^1, A^2, \cdots \cdots A^J\}$.
-Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}$$
+via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
+Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}$$
+
+Here we use left subscript for the index of the interpolation at each iteration i,e, ${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration.
+The right subcript is reserved for the index of searching points in the optimisation algorithm.
 
@@
 
-@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{(j +1)}$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$.
+@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{j +1}_l$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$.
 If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops.
-If no basis is found to have higher index value after the maximum number of tries $i_{max}$, the algorithm stops with nothing outputted.
-$h$ is the halfing parameter meaning if $h > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search.
+If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted.
+$c$ is the halfing parameter meaning if $c > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search.
 The algorithm for a random search is summarised in Algorithm \ref{random-search}.
@@
 \SetAlgoLined
-  initialisation: $h = 0$, $loop = 1, i_{max}$\;
-  \While{$i < i_{max}$}{
-    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
-    $I^{(j + 1)} = f(A^{(j+1)})$\;
-    \eIf{$I^{(j+1)} > I^{j}$}{
-      $A^{j+1} = A^{(j+1)}$\;
+  initialisation: $c = 0$, $l = 1$ $l_{\max}$, $\mathbf{A}^{j}$\;
+  \While{$l < l_{\max}$}{
+    $\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$\;
+    $I^{j+1}_l = f(\mathbf{A}^{j+1}_l)$\;
+    \eIf{$I^{j+1}_l > I^{j}$}{
+      $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
       }{
-      $h = h + 1$\;
+      $c = c + 1$\;
       }
-    $i = i + 1$\;
+    $l = l + 1$\;
   }
@@
 \SetAlgoLined
-  initialisation: $loop = 1, i_{max}$\;
-  \While{$i < i_{max}$}{
+  initialisation: $l_{\max}$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+  \While{$l < l_{\max}$}{
     Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
-    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
+    Find $A^{J_1}_l$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
     Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
@@
   }
-  \caption{searchgeodesic}
+  \caption{searchgeodesic: this part has problem!}
   \label{search_geodesic}
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$.
-Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$.
+Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{log(l + 1)}$.
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}$.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$ where $I^j$ and $I^{(j+1)}$ are the index value of the current and candidate basis respectively.
+If it is smaller, the candidate is accepted with probability $\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively.
 The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
@@
 \SetAlgoLined
-  initialisation: $T_0$, $\alpha$\;
-  \While{$i < i_{max}$}{
-    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
-    $I^{(j+1)} = f(A^{(j+1)})$\;
-    $T(i) = \frac{T_0}{\log(i + 1)}$\;
-      \eIf{$I^{(j+1)} > I^{j}$}{
-        $A^{j+1} = A^{(j+1)}$\;
-        $I^{j+1} = I^{(j+1)}$\;
-        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
+  initialisation: $T_0$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+  \While{$l < l_{\max}$}{
+    $\mathbf{A}^{j+1}_{l} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
+    $I^{j+1}_l = f(\mathbf{A}^{j+1}_{l})$\;
+    $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I^{j+1}_l > I^{j}$}{
+        $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
+        $I^{j+1} = I^{j+1}_l$\;
+        \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$,$l$ and relevant metadata}
       }{
-        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
+        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$\;
         Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
         \If{$\text{prob} > \text{rand}$}{
-          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+          \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$, $l$ and relevant metadata}
         }
       }
-    $i = i + 1$\;
+    $l = l + 1$\;
   }
@@
 
-*Iterators* indexes the data collected and it has a time series feature that prescribes the progression of the optimisation.
+*Iterators* indexes the data collected and has a time series feature that prescribes the its natural ordering.
 Given the complexity of the algorithm, there could be multiple iterators.
@@
 *index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
+A sketch of the difference between these three iterators is provided in Figure \ref{iterators}.
 Depend on the number of iteration level ($L$), the total number of iterators will differ and can be calculated as $1 + 2L$.
-In projection pursuit optimisation algorithms, there is only one level of iteartion and hence exists three iterators: $t_1$, $t_2$ and $t_3$.
-A sketch of the design of iterators is provided in Figure \ref{iterators}.
-
-There could exist other parameters that are also of our interest but can't be classified as one of the three categories.
-They are defined under the fourth category: *other parameter of interest: $O$*.
-This could include variables that are specific for one particular algorithm.
-Below we show an example of defining the global object of projection pursuit guided tour.
-
-Now we formulate the idea of global object in terms of mathmetical notation.
-Superscript $j$ is used to denote the number of iteration.
-Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left.
-The output basis doesn't have subscript.
-For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$.
-In the next iteration (iteration 2), $\mathbf{A}^1$ becomes the current basis, projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$ and the tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$.
-Below we arrange all the bases in a format that each coloumn represents an iteration in Equation \ref{eq:matrixA}.
-
-When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector with its natual ordering as in Equation \ref{eq:vectorA}.
-Hence the full data structure can be written as in Equation \ref{eq:data-structure}
 
@@
 ```
+
+In projection pursuit optimisation algorithms, there is one level of iteartion and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
+
+There could exist other parameters that are also of our interest but can't be classified as one of the three categories.
+They are defined under the fourth category: *other parameter of interest: $O$*.
+In projection pursuit, three other parameters of our interest includes `method`, `alpha` and `info`.
+`method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search.
+The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses.
+`info` labels different stages in the searching process.
+A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
+
+```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The global object in projection pursuit guided tour."}
+knitr::include_graphics(path = here::here("figures/global_obj.png"))
+```
+
+## Notation
+
+Now we formulate the idea of global object using mathmetical notation.
+Let $\mathbf{A}^0$ denotes the starting basis.
+Superscript $j$ is used to denote the number of iteration (index per iteration).
+Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left.
+For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$.
+In iteration two, $\mathbf{A}^1$ becomes the current basis and projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$.
+The tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$.
+All the bases together can be arranged in a format that each coloumn represents an iteration as in Equation \ref{eq:matrixA}.
 
@@
 
+When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector as in Equation \ref{eq:vectorA} and the full data structure is presneted in Equation \ref{eq:data-structure}.
+
 \begin{equation}
@@
 \left[
-\begin{array}{cccccccc}
+\begin{array}{c|c|ccc|ccc}
 \mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
@@
 
-An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
+<!-- An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
 [the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part.
-]
+] -->
 
-In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points
+<!-- In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points -->
 
-- *Searching points* include the observations recorded in the searching algorithm to find the target basis.
+<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
 The points for target bases is also included in the searching points and there is one such point per `tries`.
+-->
 
-- *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value.
+<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
 
@@
 
-In the projection pursuit guided tour, the decision variable is the projection matrix, the value of objective function is the index value.
-There are three iterators: `id` is the smallest ordering unit that increases by one for each observation; `tries` is updated once a search-and-interpolate step is finished.
-A third iterator `loop` is used to record the number of repetition in the search step and starts over from one at the beginning of a new `tries`.
-Three other parameters of our interest includes `method`, `alpha` and `info`.
-`method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search.
-The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses.
-`info` labels different stages in the searching process.
-A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
-
-```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The global object in projection pursuit guided tour."}
-knitr::include_graphics(path = here::here("figures/global_obj.png"))
-```
-
-### Simulated data
+## Simulated data
 

### 2020-07-07 paper/paper.tex
make sure notation in the algorithm before matches with the ones in the global object

@@
 finds the projection basis
-\(A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where \(\mathbf{a}_i\)
-is a \(p \times 1\) vector, satisfying the following optimisation
-problem:
+\(\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
+\(\mathbf{a}_i\) is a \(p \times 1\) vector, satisfying the following
+optimisation problem:
 
 \begin{align}
-I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
+I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
 \end{align}
 
-via an iterative search: \(\{A^1, A^2, \cdots \cdots A^J\}\). Guided
-tour interpolates geodesically between each basis found by projection
-pursuit \(A^j\) to form a tour path:
-\[\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}\]
+via an iterative search:
+\(\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
+Guided tour interpolates geodesically between each basis found by
+projection pursuit \(A^j\) to form a tour path:
+\[\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}\]
+
+Here we use left subscript for the index of the interpolation at each
+iteration i,e, \({}_{2}\mathbf{A}^1\) means the second interpolation
+basis in the first iteration. The right subcript is reserved for the
+index of searching points in the optimisation algorithm.
 
@@
 \citet{posse1995projection} used a stochastic direct search method, a
-random search algorithm to sample a candidate bases \(A^{(j +1)}\) in
+random search algorithm to sample a candidate bases \(A^{j +1}_l\) in
 the neighbourhood of the current basis \(A^j\) defined by the radius of
@@
 If no basis is found to have higher index value after the maximum number
-of tries \(i_{max}\), the algorithm stops with nothing outputted. \(h\)
-is the halfing parameter meaning if \(h > 30\), the nieghbourhood
+of tries \(l_{\max}\), the algorithm stops with nothing outputted. \(c\)
+is the halfing parameter meaning if \(c > 30\), the nieghbourhood
 parameter \(\alpha\) will be reduced by half in the next iteration of
@@
 non-derivative procedure based on a non-increasing cooling scheme
-\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
-is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
+\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
+is defined as \(T(l) = \frac{T_0}{log(l + 1)}\). The simulated annealing
 algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
@@
 basis is constructed as
-\(A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
+\(A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
 of the candidate basis is larger than the one of the current basis, the
@@
 candidate is accepted with probability
-\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}\)
-where \(I^j\) and \(I^{(j+1)}\) are the index value of the current and
+\(\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
+where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
 candidate basis respectively. The algorithm can be summarised as in
@@
 
-\emph{Iterators} indexes the data collected and it has a time series
-feature that prescribes the progression of the optimisation. Given the
-complexity of the algorithm, there could be multiple iterators. The
-simpliest iterator: \emph{index per observation}, is a unique identifier
-for each observation in the data. For each level of iteration, we design
-two indices: \emph{index per iteration} is fixed for each observations
-in one iteration and has an increment of one once a new iteration
-starts. \emph{index within iteration} has an increment of one for each
+\emph{Iterators} indexes the data collected and has a time series
+feature that prescribes the its natural ordering. Given the complexity
+of the algorithm, there could be multiple iterators. The simpliest
+iterator: \emph{index per observation}, is a unique identifier for each
+observation in the data. For each level of iteration, we design two
+indices: \emph{index per iteration} is fixed for each observations in
+one iteration and has an increment of one once a new iteration starts.
+\emph{index within iteration} has an increment of one for each
 observation in an iteartion and starts over from one once a new
-iteration starts. Depend on the number of iteration level (\(L\)), the
-total number of iterators will differ and can be calculated as
-\(1 + 2L\). In projection pursuit optimisation algorithms, there is only
-one level of iteartion and hence exists three iterators: \(t_1\),
-\(t_2\) and \(t_3\). A sketch of the design of iterators is provided in
-Figure \ref{iterators}.
+iteration starts. A sketch of the difference between these three
+iterators is provided in Figure \ref{iterators}. Depend on the number of
+iteration level (\(L\)), the total number of iterators will differ and
+can be calculated as \(1 + 2L\).
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/iterators} \caption{\label{iterators} A sketch of the design of iterators in iterative algorithms.}\label{fig:iterators}
+\end{figure}
+
+In projection pursuit optimisation algorithms, there is one level of
+iteartion and hence exists three iterators: \texttt{id} indices each
+observation by a unique number; \texttt{tries} is the index per
+iteration iterator that gets updated once a search-and-interpolate step
+is finished; and \texttt{loop} is the index within iteration iterator
+and starts over from one at the beginning of a new \texttt{tries}.
 
@@
 can't be classified as one of the three categories. They are defined
-under the fourth category: \emph{other parameter of interest: \(O\)}.
-This could include variables that are specific for one particular
-algorithm. Below we show an example of defining the global object of
-projection pursuit guided tour.
-
-Now we formulate the idea of global object in terms of mathmetical
-notation. Superscript \(j\) is used to denote the number of iteration.
-Candidate basis are denoted using subscript on the right and
-interpolation basis are denoted using subscript on the left. The output
-basis doesn't have subscript. For example, \(\mathbf{A}^1_{2}\) means
-the second candidate basis in the first iteration and
-\({}_{2}\mathbf{A}^1\) means the second interpolation basis that forms
-the tour path from the current basis \(\mathbf{A}^0\) to the target
-basis \(\mathbf{A}^1\). In the next iteration (iteration 2),
-\(\mathbf{A}^1\) becomes the current basis, projection pursuit searches
-through
+under the fourth category: \emph{other parameter of interest: \(O\)}. In
+projection pursuit, three other parameters of our interest includes
+\texttt{method}, \texttt{alpha} and \texttt{info}. \texttt{method}
+identifies the name of the searching method used and we are interested
+in comparing the performance between different algorithms under direct
+search. The neighbourhood parameter \texttt{alpha} controls the size of
+the sampling space and we are interested to understand how the searching
+space shrinks as the algorithm progresses. \texttt{info} labels
+different stages in the searching process. A sketch of the global object
+for projection pursuit guided tour is presented in Figure
+\ref{fig:glb-obj}.
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
+\end{figure}
+
+\hypertarget{notation}{%
+\subsection{Notation}\label{notation}}
+
+Now we formulate the idea of global object using mathmetical notation.
+Let \(\mathbf{A}^0\) denotes the starting basis. Superscript \(j\) is
+used to denote the number of iteration (index per iteration). Candidate
+basis are denoted using subscript on the right and interpolation basis
+are denoted using subscript on the left. For example,
+\(\mathbf{A}^1_{2}\) means the second candidate basis in the first
+iteration and \({}_{2}\mathbf{A}^1\) means the second interpolation
+basis that forms the tour path from the current basis \(\mathbf{A}^0\)
+to the target basis \(\mathbf{A}^1\). In iteration two, \(\mathbf{A}^1\)
+becomes the current basis and projection pursuit searches through
 \(\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}\) to
-find the target basis \(\mathbf{A}^2\) and the tour path from
+find the target basis \(\mathbf{A}^2\). The tour path from
 \(\mathbf{A}^1\) to \(\mathbf{A}^2\) is constructed using the
@@
 \({}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2\).
-Below we arrange all the bases in a format that each coloumn represents
-an iteration in Equation \ref{eq:matrixA}.
-
-When aggregated with information of index value and other interesting
-parameter, it is easier to write the basis as a column vector with its
-natual ordering as in Equation \ref{eq:vectorA}. Hence the full data
-structure can be written as in Equation \ref{eq:data-structure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/iterators} \caption{\label{iterators} A sketch of the design of iterators in iterative algorithms.}\label{fig:iterators}
-\end{figure}
+All the bases together can be arranged in a format that each coloumn
+represents an iteration as in Equation \ref{eq:matrixA}.
 
@@
 
+When aggregated with information of index value and other interesting
+parameter, it is easier to write the basis as a column vector as in
+Equation \ref{eq:vectorA} and the full data structure is presneted in
+Equation \ref{eq:data-structure}.
+
 \begin{equation}
@@
 
-An optimisation procedure can easily generate hundreds or thousands
-sampling points and the question becomes what are the points we want to
-explore\ldots{} {[}the classification of searching or updating points
-goes here and people should be clear what part of the points goes to the
-plot - this is the data part not the plot part. {]}
-
-In tour, all the points recorded in the global object can be divided
-into two broad categories: searching points and interpolating points
-
-\begin{itemize}
-\item
-  \emph{Searching points} include the observations recorded in the
-  searching algorithm to find the target basis. The points for target
-  bases is also included in the searching points and there is one such
-  point per \texttt{tries}.
-\item
-  \emph{interpolating points} exist in the guided tour to produce
-  continuous animated view from one target basis to another and it
-  doesn't have \texttt{loop} value.
-\end{itemize}
-
-In the projection pursuit guided tour, the decision variable is the
-projection matrix, the value of objective function is the index value.
-There are three iterators: \texttt{id} is the smallest ordering unit
-that increases by one for each observation; \texttt{tries} is updated
-once a search-and-interpolate step is finished. A third iterator
-\texttt{loop} is used to record the number of repetition in the search
-step and starts over from one at the beginning of a new \texttt{tries}.
-Three other parameters of our interest includes \texttt{method},
-\texttt{alpha} and \texttt{info}. \texttt{method} identifies the name of
-the searching method used and we are interested in comparing the
-performance between different algorithms under direct search. The
-neighbourhood parameter \texttt{alpha} controls the size of the sampling
-space and we are interested to understand how the searching space
-shrinks as the algorithm progresses. \texttt{info} labels different
-stages in the searching process. A sketch of the global object for
-projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
-\end{figure}
-
 \hypertarget{simulated-data}{%
-\subsubsection{Simulated data}\label{simulated-data}}
+\subsection{Simulated data}\label{simulated-data}}
 

### 2020-07-07 paper/paper.Rmd
small tweaks here and there

@@
 clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-Let $D_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $P = D \cdot A$, where $A_{p\times d}$ is the projection basis and $P_{n \times p}$ is the projected data.
-Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $A$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or indice function,  that can be used to measure the "interestingness" of the projection.
+Let $\mathbf{D}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{D} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times p}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
 A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
@@
 \begin{align}
-I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
+&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
 s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
@@
 via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
-Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}$$
+Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}$$
 
-Here we use left subscript for the index of the interpolation at each iteration i,e, ${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration.
+Where $\mathbf{A}^0$ is the starting basis and $\mathbf{A}^J$ is the final basis that has the largest index value in the searching space.
+Here we use left subscript to denote the index of the interpolation bases in each iteration i.e.
+${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration.
 The right subcript is reserved for the index of searching points in the optimisation algorithm.
@@
 There are three existing methods for optimisating PPI function and we review them below.
-
-@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{j +1}_l$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$.
-If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops.
+@posse1995projection proposed a stochastic direct search method, a random search algorithm.
+Given a current basis $\mathbf{A}^{j}$, a candidate basis $\mathbf{A}^{j +1}_l$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}^{j}$ by $$\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$$.
+If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $\mathbf{A}^{j + 1}$ along with other metadata and the current iteration stops.
 If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted.
-$c$ is the halfing parameter meaning if $c > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search.
+$c$ is the halfing parameter and when $c > 30$, the nieghbourhood parameter $\alpha$ in the next iteration will be reduced by half.
 The algorithm for a random search is summarised in Algorithm \ref{random-search}.
@@
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively.
+If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively.
 The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
@@
 \begin{array}{c|c|ccc|ccc}
-\mathbf{A} & I & j & l & n & o_1 & \cdots &o_p\\
+\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
 \hline
-\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
 \mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
@@
 \mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
+\hline 
 \mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
+\hline
 {}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
@@
 {}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
+\hline
 \vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
-\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n^{*} &o_{1, n-1}& \cdots & o_{p, n-1}\\
-\mathbf{A}^J & I^J & J & l_H + 1 & n^{**} & o_{1, n}& \cdots & o_{p, n}\\
+\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n-1 &o_{1, n-1}& \cdots & o_{p, n-1}\\
+\mathbf{A}^J & I^J & J & l_H + 1 & n & o_{1, n}& \cdots & o_{p, n}\\
 \end{array}
@@
 
-where $n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$ and $n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$
+where $n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$
 
@@
 
-Below we will present several examples to diagnose different aspect of the tour optimisation.
-We will first provide examples on producing 1) static plots to explore the value of objective function and 2) animated plots to explore the searching space and then a more sophisticated example that combines both to understand how to optimise a noisy and complex index function.
-In both section, we will first provide a toy example that is easy to grasp and one or two examples that help us understand the optimisation algorithm and parameter choice better.
+Below we will present several examples to diagnose different aspects of the projection pursuit optimisation. We will present:
 
-Follow the research question we raised earlier, the purpose of visual diagnostics is to understand
+1) static plots to explore the index value,
 
-- Whether the procedure has provided an optimum and how the value of the objective function changes through the optimisation.
+2) animated plots to explore the projection basis and,
 
-- How does the searching space look like, that is, geometrically, where are the points located in the space.
+3) a self-contained example to understand how to optimise a noisy and complex index function.
+
+In the first two sections, we will first provide a toy example that is easy to grasp and then more examples that can help us to understand the algorithm and parameter choice.
+Remeber the research question we raised earlier, the purpose of visual diagnostics is to understand:
+
+- Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm?
+
+- How does the searching space look like, that is, geometrically, where are the projection bases located in the space?
 

### 2020-07-07 paper/paper.tex
small tweaks here and there

@@
 (i.e.~clustering, outliers and skewness) of multivariate data via
-projecting it in lower dimensions. Let \(D_{n \times p}\) be a data
-matrix, an n-d projection can be seen as a linear transformation:
-\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(P = D \cdot A\),
-where \(A_{p\times d}\) is the projection basis and \(P_{n \times p}\)
-is the projected data. Define
+projecting it in lower dimensions. Let \(\mathbf{D}_{n \times p}\) be a
+data matrix, an n-d projection can be seen as a linear transformation:
+\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
+\(\mathbf{P} = \mathbf{D} \cdot \mathbf{A}\), where
+\(\mathbf{P}_{n \times p}\) is the projected data and
+\(\mathbf{A}_{p\times d}\) is the projection basis Define
 \(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
-that maps the projection basis \(A\) onto a scalar value \(I\), this
-function is commonly known as the projection pursuit index (PPI)
-function, or indice function, that can be used to measure the
+that maps the projection basis \(\mathbf{A}\) onto a scalar value \(I\),
+this function is commonly known as the projection pursuit index (PPI)
+function, or index function and can be used to measure the
 ``interestingness'' of the projection. A number of indice functions have
@@
 projection pursuit \(A^j\) to form a tour path:
-\[\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}\]
-
-Here we use left subscript for the index of the interpolation at each
-iteration i,e, \({}_{2}\mathbf{A}^1\) means the second interpolation
-basis in the first iteration. The right subcript is reserved for the
-index of searching points in the optimisation algorithm.
+\[\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}\]
+
+Where \(\mathbf{A}^0\) is the starting basis and \(\mathbf{A}^J\) is the
+final basis that has the largest index value in the searching space.
+Here we use left subscript to denote the index of the interpolation
+bases in each iteration i.e. \({}_{2}\mathbf{A}^1\) means the second
+interpolation basis in the first iteration. The right subcript is
+reserved for the index of searching points in the optimisation
+algorithm.
 
 There are three existing methods for optimisating PPI function and we
-review them below.
-
-\citet{posse1995projection} used a stochastic direct search method, a
-random search algorithm to sample a candidate bases \(A^{j +1}_l\) in
-the neighbourhood of the current basis \(A^j\) defined by the radius of
-the p-dimensional sphere, \(\alpha\). If the candidate basis has a
-higher index value than the current basis, it is outputed as the target
-basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
-If no basis is found to have higher index value after the maximum number
-of tries \(l_{\max}\), the algorithm stops with nothing outputted. \(c\)
-is the halfing parameter meaning if \(c > 30\), the nieghbourhood
-parameter \(\alpha\) will be reduced by half in the next iteration of
-the random search. The algorithm for a random search is summarised in
-Algorithm \ref{random-search}.
+review them below. \citet{posse1995projection} proposed a stochastic
+direct search method, a random search algorithm. Given a current basis
+\(\mathbf{A}^{j}\), a candidate basis \(\mathbf{A}^{j +1}_l\) is sampled
+in the neighbourhood defined by the radius of the p-dimensional sphere,
+\(\alpha\) of the current basis \(\mathbf{A}^{j}\) by
+\[\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}\]. If
+the candidate basis has a higher index value than the current basis, it
+is outputed as the target basis \(\mathbf{A}^{j + 1}\) along with other
+metadata and the current iteration stops. If no basis is found to have
+higher index value after the maximum number of tries \(l_{\max}\), the
+algorithm stops with nothing outputted. \(c\) is the halfing parameter
+and when \(c > 30\), the nieghbourhood parameter \(\alpha\) in the next
+iteration will be reduced by half. The algorithm for a random search is
+summarised in Algorithm \ref{random-search}.
 
@@
 candidate is accepted with probability
-\(\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
+\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
 where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
@@
 
-where \(n* = (J-2) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\) and
-\(n^{**} = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)
+where \(n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)
 
@@
 
-Below we will present several examples to diagnose different aspect of
-the tour optimisation. We will first provide examples on producing 1)
-static plots to explore the value of objective function and 2) animated
-plots to explore the searching space and then a more sophisticated
-example that combines both to understand how to optimise a noisy and
-complex index function. In both section, we will first provide a toy
-example that is easy to grasp and one or two examples that help us
-understand the optimisation algorithm and parameter choice better.
-
-Follow the research question we raised earlier, the purpose of visual
-diagnostics is to understand
+Below we will present several examples to diagnose different aspects of
+the projection pursuit optimisation. We will present:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  static plots to explore the index value,
+\item
+  animated plots to explore the projection basis and,
+\item
+  a self-contained example to understand how to optimise a noisy and
+  complex index function.
+\end{enumerate}
+
+In the first two sections, we will first provide a toy example that is
+easy to grasp and then more examples that can help us to understand the
+algorithm and parameter choice. Remeber the research question we raised
+earlier, the purpose of visual diagnostics is to understand:
 
@@
 \item
-  Whether the procedure has provided an optimum and how the value of the
-  objective function changes through the optimisation.
+  Whether the algorithm has successfully found the maximum and how the
+  index value changes throughout the algorithm?
 \item
   How does the searching space look like, that is, geometrically, where
-  are the points located in the space.
+  are the projection bases located in the space?
 \end{itemize}

### 2020-07-07 source/geozoo.Rmd
add lines between points in geozoo animation

@@
 
-```{r}
-vis <- function(data, color = NULL, ...){
+```{r echo = FALSE}
+vis <- function(data, color = NULL, edges = NULL, edges.col = NULL, ...){
 
@@
   dim <- row * col
+  n_bg <- 1000
 
@@
 
-  a <- sphere.hollow(p = dim, n = 1000)
+  a <- sphere.hollow(p = dim, n = n_bg)
   a$points <- a$points %>% rbind(path)
@@
   if(is.null(color)){
-    color <- c(rep("grey", 1000), rep("red", nrow(data) - 1), "green")
+    color <- c(rep("grey", 1000), rep("red", nrow(data)))
   }
 
-  print(a, col = color)
+  # if(is.null(edges)){
+  #   start <- n_bg + 1
+  #   end <- nrow(a$points)
+  #   edges <- matrix(c(start: (end -1), 
+  #                     (start + 1): end), ncol = 2)
+  # }
+  # 
+  # if(is.null(edges.col)){
+  #   edges.col <- "red"
+  # }
+
+  print(a, col = color, edges = edges, edges.col = edges.col)
 
@@
 # final basis 
-holes_1d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis)
+holes_1d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
 # final basis after polish
-holes_1d_geo_polish %>% tail(1) %>% pull(basis)
+holes_1d_geo_polish %>% tail(1) %>% pull(basis) %>% .[[1]]
 
+# searching path and final basis
 dt <- holes_1d_geo %>% filter(info == "interpolation")
-color <-  c(rep("grey", 1000), rep("red", nrow(dt) - 1), "green")
-vis(dt, color = color)
+color <- hue_pal()(1)
+color_vec <-  c(rep("grey", 1000), rep(color, nrow(dt)))
+vis(dt, color = color_vec, edges.col = color)
 
+# searching path, final basis and polish
 interp <- holes_1d_geo %>% filter(info == "interpolation")
@@
 dt_bind <- bind_rows(interp,polish) 
-
-color <- c(rep("grey", 1000), rep("red", nrow(interp) - 1),"green", rep("blue", nrow(polish)))
-vis(dt_bind, color = color)
+color1 <-  hue_pal()(2)[1]
+color2 <-  hue_pal()(2)[2]
+color_vec <- c(rep("grey", 1000), rep(color1, nrow(interp)), rep(color2, nrow(polish)))
+edges.col <- c(rep(color1, nrow(interp)), rep(color2, nrow(polish)))
+vis(dt_bind, color = color_vec, edges.col = edges.col)
 
@@
 ```{r}
-# a small example 
-interp1 %>% tail(1) %>% pull(basis)
-interp2 %>% tail(1) %>% pull(basis)
-
 # different seeds lead to different variables found
@@
 
+# a small example 
+interp1 %>% tail(1) %>% pull(basis) %>% .[[1]]
+interp2 %>% tail(1) %>% pull(basis) %>% .[[1]]
+
 color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
-color <- c(rep("grey", 1000), 
-           rep(color_hex[[1]], nrow(interp1)-1), color_hex[[3]],
-           rep(color_hex[[2]], nrow(interp2)-1), color_hex[[3]])
+color_vec<- c(rep("grey", 1000), 
+           rep(color_hex[[1]], nrow(interp1)),
+           rep(color_hex[[2]], nrow(interp2)))
+edges <- matrix(
+  c((1000 + 1): (1000 + nrow(interp1)-1), (1000 + nrow(interp1) + 1): (1000 + nrow(dt_bind)-1), 
+    (1000 + 2): (1000 + nrow(interp1)), (1000 + nrow(interp1) + 2): (1000 + nrow(dt_bind))), 
+  ncol = 2
+)
+edges.col <- c(rep(color_hex[[1]], nrow(interp1)), 
+               rep(color_hex[[2]], nrow(interp2)))
 
-vis(dt_bind, color)
+vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)
 ```
@@
 ```{r}
+# don't think adding the lines is particularly useful tho
+# explore the full parameter space 
+interp3 <- kol_cdf_1d_2var_better_random %>% filter(info == "interpolation")
+polish3 <- kol_cdf_1d_2var_better_random_polish %>% filter(info == "polish_best")
+n_interp3 <- nrow(interp3) + nrow(polish3)
+
+dt_bind <- bind_rows(interp3, polish3)
+
+color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
+color_vec <- c(rep("grey", 1000), 
+           rep(color_hex[[3]], n_interp3))
+
+edges <- matrix(
+  c((1000 + 1): (1000 + nrow(dt_bind)-1),
+
+    (1000 + 2): (1000 + nrow(dt_bind))),
+
+  ncol = 2
+)
+
+edges.col <- c(rep(color_hex[[3]], n_interp3))
+
+vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)
+```
+
+```{r}
+# combined all three together
 # better random finds global 
-
 # explore the full parameter space 
@@
 
+n_interp1 <- nrow(interp1) + nrow(polish1)
+n_interp2 <- nrow(interp2) + nrow(polish2)
+n_interp3 <- nrow(interp3) + nrow(polish3)
+
 dt_bind <- bind_rows(interp1, polish1, 
@@
 
-color_hex <- RColorBrewer::brewer.pal(8, "Dark2")
-color <- c(rep("grey", 1000), 
-           rep(color_hex[[1]], nrow(interp1) + nrow(polish1)-1), "green",
-           rep(color_hex[[2]], nrow(interp2) + nrow(polish2)-1), "green",
-           rep(color_hex[[3]], nrow(interp3) + nrow(polish3)-1), "green"
+color_hex <- RColorBrewer::brewer.pal(4, "Dark2")
+color_vec <- c(rep("grey", 1000), 
+           rep(color_hex[[1]], n_interp1), 
+           rep(color_hex[[2]], n_interp2), 
+           rep(color_hex[[3]], n_interp3))
+
+edges <- matrix(
+  c((1000 + 1): (1000 + n_interp1), 
+    (1000 + n_interp1 + 1): (1000 + n_interp1 + n_interp2-1), 
+    (1000 + n_interp1 + n_interp2 + 1): (1000 + nrow(dt_bind)-1),
+
+    (1000 + 2): (1000 + n_interp1 + 1), 
+    (1000 + n_interp1 + 2): (1000 + n_interp1 + n_interp2), 
+    (1000 + n_interp1 + n_interp2 + 2): (1000 + nrow(dt_bind))),
+
+  ncol = 2
 )
 
-vis(dt_bind, color)
+edges.col <- c(rep(color_hex[[1]], n_interp1), 
+               rep(color_hex[[2]], n_interp2),
+               rep(color_hex[[3]], n_interp3))
+
+vis(dt_bind, color_vec, edges = edges, edges.col = edges.col)
 
@@
                        rep("better2", nrow(interp2)), 
-                       rep("better_random", nrow(interp3)))) %>% 
+                       rep("better_random", nrow(interp3))),
+         algorithm = fct_relevel(algorithm, c("better1", "better2", "better_random"))) %>% 
   ferrn::compute_pca()
@@
   geom_point() + 
+  scale_color_brewer(palette = "Dark2") + 
   theme(aspect.ratio = 1)

### 2020-07-07 paper/paper.Rmd
match the color in the last example with geozoo color

@@
 Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
-p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp() + ggtitle("search_better, seed 123456")
-p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp() +  ggtitle("search_better, seed 12345")
+color <- RColorBrewer::brewer.pal(3, "Dark2")
+p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
+  scale_color_manual(values = color[2]) + ggtitle("search_better, seed 123456")
+p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
+  scale_color_manual(values = color[1]) + ggtitle("search_better, seed 12345")
 
@@
 kol_cdf_1d_2var_better_random %>% 
-  explore_trace_interp() + 
+  explore_trace_interp(col = 0) + 
   ylim(0, 0.25) + 
+  scale_color_manual(values = color[3]) + 
   theme(legend.position = "none") + 

### 2020-07-10 paper/paper.Rmd
modify notation as per wednesday meeting (16)

@@
 clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-Let $\mathbf{D}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{D} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times p}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
+Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times p}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
 A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
 \end{align}
 
-via an iterative search: $\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
+via an iterative search: $\{\mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
 Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}$$
@@
 Now we formulate the idea of global object using mathmetical notation.
+Let $$\mathbf{A}_{j,l,k}$$ denotes the projection matrix in iteration $j$.
+If it is searching basis produced in the optimisation, then it has a subscript $l$.
+If it is an interpolation basis, it has a $k$ subscript.
+For example $\mathbf{A}_{1,2,/}$ means the second searching basis in the first iteration, $\mathbf{A}_{1, /, 2}$ means the second interpolation basis that on the tour path from the current basis $\mathbf{A}^0$ to the target basis found in iteration 1: $\mathbf{A}^1$.
+The target basis found in each iteration, $\mathbf{A}_{j, l_j, /}$ can also be simplified using a superscript as $\mathbf{A}^j$.
+
 Let $\mathbf{A}^0$ denotes the starting basis.
@@
 \begin{array}{ccccccc}
-/ & \mathbf{A}^1_1 & \mathbf{A}^2_1 & \cdots & \mathbf{A}^j_1 &\cdots & \mathbf{A}^J_1 \\
-/ & \mathbf{A}^1_2 & \mathbf{A}^2_2 & \cdots & \mathbf{A}^j_2 &\cdots & \mathbf{A}^J_2 \\
+\ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{j,\sum_{i = 1}^{j-1}l_i +1,\ast} &\cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1 ,\ast} \\
+\ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{j,\sum_{i = 1}^{j-1}l_i +2,\ast} &\cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2, \ast} \\
 \vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
-/ &\mathbf{A}^1_{l_1} & \mathbf{A}^2_{l_2} & \cdots & \mathbf{A}^j_{l_h} &\cdots & \mathbf{A}^J_{L_H}\\
+\ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{j, \sum_{i = 1}^{j-1}l_i + l, \ast}  &\cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i + l, \ast}\\
+\vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
+\mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{j, \sum_{i = 1}^{j}l_i , \ast}  &\cdots & \mathbf{A}_{J, \sum_{i = 1}^{J} l_i, \ast}\\
 \hline
-\mathbf{A}^0 & \mathbf{A}^1 & \mathbf{A}^2 & \cdots & \mathbf{A}^j  &\cdots & \mathbf{A}^J\\
-\hline
-/ &{}_{1}\mathbf{A}^1 & {}_{1}\mathbf{A}^2 & \cdots &{}_{1}\mathbf{A}^j & \cdots & /\\
-/ &{}_{2}\mathbf{A}^1 & {}_{2}\mathbf{A}^2 & \cdots &{}_{2}\mathbf{A}^j & \cdots & / \\
-\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & / \\
-/ &{}_{k_1}\mathbf{A}^1 & {}_{k_2}\mathbf{A}^2 & \cdots &{}_{k_m}\mathbf{A}^j & \cdots & / \\
+\ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{j, \ast,\sum_{i = 1}^{j-1}k_i + 1} & \cdots & \ast\\
+\ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{j, \ast, \sum_{i = 1}^{j-1}k_i +2} & \cdots & \ast \\
+\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & \vdots \\
+\ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{j, \ast, \sum_{i = 1}^{j-1}k_i +k} & \cdots & \ast \\
+\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & \vdots \\
+\ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{j, \ast,\sum_{i = 1}^{j}k_i} & \cdots & \ast \\
 \end{array}
@@
 \vphantom{\vdots} \vdots \\
-\vphantom{A^J_{L_H}}\text{candidate basis} \\
-\hline
-\vphantom{A^J}\text{output basis} \\
+\vphantom{A^J_2}\text{candidate basis} \\
+\vphantom{\vdots} \vdots \\
+\vphantom{A^J}\text{current/target basis} \\
 \hline
@@
 \vphantom{\text{NULL}}\text{interpolation basis} \\
+\vphantom{\vdots} \vdots \\
+\vphantom{\text{NULL}}\text{interpolation basis} \\
 \end{array}
 \right]
-\label{eq:matrixA}
+\label{matrixA}
 \end{equation}
@@
 
+All the basis can also be denoted usign a single subscript by their natural ordering as $\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n$.
+
 \begin{equation}
-\begin{bsmallmatrix}
-\mathbf{A}^0 &\mathbf{A}^1_1 &  \mathbf{A}^1_{2} & \cdots& \mathbf{A}^1_{l_1} & \mathbf{A}^1 & {}_{1}\mathbf{A}^1& {}_{2}\mathbf{A}^1 &\cdots& {}_{k_1}\mathbf{A}^1 & \mathbf{A}^2_1 & \mathbf{A}^2_{2} & \cdots& \mathbf{A}^2_{l_2} & \mathbf{A}^2 & {}_{1}\mathbf{A}^2 &{}_{2}\mathbf{A}^2 &\cdots& {}_{k_2}\mathbf{A}^2 & \cdots & \mathbf{A}^J
-\end{bsmallmatrix} 
+\begin{array}{cccccccc}
+\bigl[\mathbf{A}_{0, \ast, \ast} & 
+\mathbf{A}_{1, 1, \ast} & 
+\mathbf{A}_{1,2,\ast} & 
+\cdots & 
+\mathbf{A}_{1,l,\ast} & 
+\cdots & 
+\mathbf{A}_{1,l_1, \ast}   \\ 
+\mathbf{A}_{1, \ast, 1} & 
+\mathbf{A}_{1, \ast, 2} & 
+\cdots &
+\mathbf{A}_{1, \ast, k} &
+\cdots &
+\mathbf{A}_{1, \ast, k_1} &
+\cdots &
+\mathbf{A}_{J, l_J, \ast} \bigr]
+\end{array} 
 \label{eq:vectorA}
@@
 \left[
-\begin{array}{c|c|ccc|ccc}
-\mathbf{A}^0 & I^0 & 0 & / & 1 & / & \cdots & /\\
+\begin{array}{c|c|cccc|ccc}
+\mathbf{A} & I & j &  l  & k & n & S_{1}& \cdots & S_{p}\\
 \hline
-\mathbf{A}^1_1 & I^1_1 & 1 & 1 & 2 & o_{1, 1} & \cdots & o_{p, 1}\\
-\mathbf{A}^1_2 & I^1_2 & 1 & 2 & 3 & o_{1, 2} & \cdots & o_{p, 2}\\
-\vdots &\vdots & \vdots &\vdots &\vdots &\vdots &\cdots &\vdots\\
-\mathbf{A}^1_{l_1} & I^1_{l_1} & 1 & l_1 & l_1+1 & \vdots & \cdots & \vdots\\
+\mathbf{A}_0 & I_0 & 0 & \ast & \ast & 1 & \ast & \cdots & \ast\\
 \hline
-\mathbf{A}^1 & I^1 & 1 & l_1 + 1 & l_1 + 2 & \vdots & \cdots & \vdots\\
+\mathbf{A}_1 & I_1 & 1 & 1 & \ast& 2 & S_{1, 1} & \cdots & S_{p, 1}\\
+\mathbf{A}_2 & I_2 & 1 & 2 &\ast & 3 & S_{1, 2} & \cdots & S_{p, 2}\\
+\vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots\\
+\vdots & \vdots & 1 & l_1 & \ast & \vdots & \vdots & \cdots & \vdots\\
 \hline
-{}_{1}\mathbf{A}^1 &{}_{1}I^1 & 1 & / & l_1 + 3 &\vdots & \cdots & \vdots\\
-{}_{2}\mathbf{A}^1 &{}_{2}I^1 & 1 & / & l_1 + 4&\vdots & \cdots & \vdots\\
-\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
-{}_{k_1}\mathbf{A}^1 &{}_{k_1}I^1 & 1 & / & l_1 +k_1 + 2 &\vdots & \cdots & \vdots\\
+\vdots & \vdots & 1 & \ast & 1& \vdots &\vdots & \cdots & \vdots\\
+\vdots &\vdots& 1 & \ast & 2& \vdots&\vdots & \cdots & \vdots\\
+\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
+\vdots &\vdots & 1 & \ast & k_1&\vdots &\vdots & \cdots & \vdots\\
 \hline
-\vdots &\vdots &\vdots  &\vdots &\vdots &\vdots &\cdots &\vdots \\
-\mathbf{A}^J_{l_H} & I^J_{l_H} & J & l_H & n-1 &o_{1, n-1}& \cdots & o_{p, n-1}\\
-\mathbf{A}^J & I^J & J & l_H + 1 & n & o_{1, n}& \cdots & o_{p, n}\\
+\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
+\mathbf{A}_n & I_n & J &  l_J & \ast & n & S_{1, n}& \cdots & S_{p, n}\\
 \end{array}
@@
 
-where $n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1$
+where $n = 1+ \sum_{i = 1}^J l_i + \sum_{i = 1}^{J-1} k_i$
 

### 2020-07-10 paper/paper.tex
modify notation as per wednesday meeting (16)

@@
 (i.e.~clustering, outliers and skewness) of multivariate data via
-projecting it in lower dimensions. Let \(\mathbf{D}_{n \times p}\) be a
+projecting it in lower dimensions. Let \(\mathbf{X}_{n \times p}\) be a
 data matrix, an n-d projection can be seen as a linear transformation:
 \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
-\(\mathbf{P} = \mathbf{D} \cdot \mathbf{A}\), where
+\(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{P}_{n \times p}\) is the projected data and
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
+s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
 \end{align}
@@
 via an iterative search:
-\(\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
-Guided tour interpolates geodesically between each basis found by
-projection pursuit \(A^j\) to form a tour path:
+\(\{\mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\). Guided
+tour interpolates geodesically between each basis found by projection
+pursuit \(A^j\) to form a tour path:
 \[\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}\]
@@
 Now we formulate the idea of global object using mathmetical notation.
+Let \[\mathbf{A}_{j,l,k}\] denotes the projection matrix in iteration
+\(j\). If it is searching basis produced in the optimisation, then it
+has a subscript \(l\). If it is an interpolation basis, it has a \(k\)
+subscript. For example \(\mathbf{A}_{1,2,/}\) means the second searching
+basis in the first iteration, \(\mathbf{A}_{1, /, 2}\) means the second
+interpolation basis that on the tour path from the current basis
+\(\mathbf{A}^0\) to the target basis found in iteration 1:
+\(\mathbf{A}^1\). The target basis found in each iteration,
+\(\mathbf{A}_{j, l_j, /}\) can also be simplified using a superscript as
+\(\mathbf{A}^j\).
+
 Let \(\mathbf{A}^0\) denotes the starting basis. Superscript \(j\) is
@@
 
+All the basis can also be denoted usign a single subscript by their
+natural ordering as
+\(\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n\).
+
 \begin{equation}
-\begin{bsmallmatrix}
-\mathbf{A}^0 &\mathbf{A}^1_1 &  \mathbf{A}^1_{2} & \cdots& \mathbf{A}^1_{l_1} & \mathbf{A}^1 & {}_{1}\mathbf{A}^1& {}_{2}\mathbf{A}^1 &\cdots& {}_{k_1}\mathbf{A}^1 & \mathbf{A}^2_1 & \mathbf{A}^2_{2} & \cdots& \mathbf{A}^2_{l_2} & \mathbf{A}^2 & {}_{1}\mathbf{A}^2 &{}_{2}\mathbf{A}^2 &\cdots& {}_{k_2}\mathbf{A}^2 & \cdots & \mathbf{A}^J
-\end{bsmallmatrix} 
+\begin{array}{cccccccc}
+\bigl[\mathbf{A}_{0, \ast, \ast} & 
+\mathbf{A}_{1, 1, \ast} & 
+\mathbf{A}_{1,2,\ast} & 
+\cdots & 
+\mathbf{A}_{1,l,\ast} & 
+\cdots & 
+\mathbf{A}_{1,l_1, \ast}   \\ 
+\mathbf{A}_{1, \ast, 1} & 
+\mathbf{A}_{1, \ast, 2} & 
+\cdots &
+\mathbf{A}_{1, \ast, k} &
+\cdots &
+\mathbf{A}_{1, \ast, k_1} &
+\cdots &
+\mathbf{A}_{J, l_J, \ast} \bigr]
+\end{array} 
 \label{eq:vectorA}
@@
 
-where \(n = (J-1) + \sum_{h = 1}^H l_h + \sum_{m = 1}^M k_m + 1\)
+where \(n = 1+ \sum_{i = 1}^J l_i + \sum_{i = 1}^{J-1} k_i\)
 

### 2020-07-13 paper/paper.Rmd
more notation on the searching space and kol_cdf index

@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools}
+  \usepackage{amssymb, amsmath, mathtools, bbm}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times p}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
+Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
 A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
@@
 
-Below we will present several examples to diagnose different aspects of the projection pursuit optimisation. We will present:
+Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation. We will present:
 
@@
 
-3) a self-contained example to understand how to optimise a noisy and complex index function.
+3) a self-contained example to optimise a complex index function.
 
@@
 
-#### Toy example: exploring searching points
+It is a convention to map the time index of a time series on the x-axis and value to the y-axis.
+In our diagnostic plots, it is to map the iterator on the x-axis and index value to the y-axis.
+The next question is to ask, given four indices being defined in the global object, which one should be used?
+*It depends on the *.
 
-The largest difficulties of exploring searching points is its unknown number of observations per `tries`.
-Mapping `id` on the x-axis will leave the `tries` with few observations a small space in the plot, while those `tries` with large number of search points (usually towards the end) occupy the vast majority of the plot space.
-This motivates to summarise the points in each `tries`.
-At each iteration, rather than knowing the index value of *every* points, we are more interested to know a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the geodesic interpolation and future searches.
+As the searching goes towards the end, it is easy to have hundreds of candidate bases in each iteration with very close index value.
+When plotting, these iteration will occupy the vast majority of the plot space.
+This motivates to summarise the points in each iteration (`tries`).
+At each iteration, rather than knowing the index value of *every* points, we are more interested to have a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the next geodesic interpolation and future searches.
+
+Using the notation we defined earlier, the plot can be defined as plotting $j$ on the x-axis and the five point summary of the following groups on the y-axis:
+
+$$\mathbbm{1}(I_{j = j^*})$$
 
@@
 
-*Example: exploring searching points* The larger dataset is used for 2D projection with `search_better` and `max.tries = 500`.
+#### Toy example: exploring searching points
+
+The larger dataset is used for 2D projection with `search_better` and `max.tries = 500`.
 In Figure \ref{points}, the searching points with `id` and `tries` on the x-axis, colored by `tries`.
@@
 
-#### A more complex example: adding an interruption in the interpolation stage
+\newpage
+
+The number of interpolation points is much managable than the searching points and always shows a pattern of gradual increase in the index value.
+Hence a line plot with interpolating order on the x-axis will be desirable.
+Using notation, the x-axis is a sequence $$x = \{1, 2, \cdots, k_1, \cdots, \sum_{i=1}^Jk_i\}$$ and y-axis are the index values defined by $y = I_{x}$.
+
+#### A more complex example: Interruption
 
@@
 
-*Example: Interruption* This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
+This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
 In the interpolation stage, continuous frames will be constructed to bridge the project on the current basis to the target basis.
@@
 
-Exploring the value of an objective function is cna be a simple task.
-Because the iterator has a time series ordering feature, it goes onto the x-axis as convention.
-The value of hte objective function goes to the y-axis.
-Other parameter of interest can be represented using color to present more information.
-[This is the case for our purpose with method, but when the level goes higher or the parameter is quantitative, color may not be a good option.
-]
+\newpage
 
-An option to explore of the searching space is to explore a reduced projection 2D space via principle component analysis or other dimensional reduction methods.
+### Explore searching space {#animated}
+
+In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot.
+An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
+We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
+This gives a matrix of dimension $n$ by $(p*d)$.
+Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
+By definition, the principal components maximise the variance of the projection bases in the reduced space.
+Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
+Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix
+
+$$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$
+
+has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
+The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$
+
+The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
+THis means all the iterators and interesting other variables defined in the global object can be binded with the these two projections and can be mapped to other aesthetics (i.e.
+color).
+
 In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
@@
 It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
-
-### Explore searching space {#animated}
-
-The parameter to explore can be a vector or matrix instead of scalar, for example, the projection basis in tour.
-This imposes difficulties in visualisation since humans are bounded to perceive at most three dimensions in a plot.
-Thus, principal component analysis is used to reduce the dimension of projection bases and the first two principal components are mapped to the x and y axis of the plot.
-Additional variable of interest could be mapped to the color aesthetics for exploration.
 
@@
 
-holes_1d_geo %>% explore_proj_pca(col = info) + 
-  theme(legend.position = "bottom")
+holes_1d_geo %>% explore_proj_pca(col = info)
 
@@
 
+\newpage
+
 ### A comprehensive example of diagnosing a noisy index function
 
-*Example: Noisy index function* The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
-`kol_cdf` index, an index function based on Kolmogorov test, compares the difference between a projection matrix and a randomly generated normal distribution based on cumulated distribution function (CDF).
-Several diagnostic visualisations below show the characteristic of this index function.
+The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
+`kol_cdf` index, an 1D projection index function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
+Denotes the ECDF function as $F(u)$ with subscript incidating the variable,  the Kolmogorove statistics defined by
 
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating points for `search_geodesic` and `search_better` on 1D projection for `kol_cdf` index and the interpolation path shows a zig-zag pattern.
-Polishing step has done much more work to reach the final index value for `search_geodesic` than `search_better` and this indicates that noisy index's favour of a random search method than ascent method [the order of the examples needs to be rearranged since I haven't introduced polishing here.
-xxx].
+$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
+
+can be seen as a function of the projection matrix $\mathbf{A}_{p \times 1}$ and hence a valid index function.
+
+#### Explore index value
+
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating points when using different optimisation algorithms: `search_geodesic` and `search_better`. One can observe that
+
+- The index value of `kol_cdf` index is much smaller than that of holes index
+- The link of index values from interpolation bases  are no longer smooth
+- Both algorithms reach a similar final index value after polishing
+
+Polishing step has done much more work to find the final index value ub `search_geodesic` than `search_better` and this indicates `kol_cdf` function favours of a random search method than ascent method.
 
@@
 
-The second example for `kol_cdf` index uses 1D projection on the larger dataset.
-Since there are two informative variables and the projection is 1D, there are two local maximum  when the projection matrix at $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
-As in Figure \ref{1d-2var-different-seeds}, different local maximum can be found using `search_better` with different seeds while `search_better_random` can always find the global maximum, as in Figure \ref{1d-2var-better-random}, although at a high cost of number of points evaluated.
+Now we enlarge the dataset to include two informative variables: `x2` and `x3` and remain 1D projection.
+In this case, two local maximum appear when projection matrix being $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
+
+Using different seeds in `search_better` allows us to find both local maximum as in Figure \ref{1d-2var-different-seeds}.
+Comparing the maximum of both, we can see that the global maximum happens when `x2` is found.
+It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds?
+`search_better_random` manages to do it via a metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
 
@@
 
+#### Explore searching space
+
+We can also plot the searching points of all the three algorithms in the searching space and explore their relative position against each other using principal components.
+As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhausive search.
+The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end.
+
+```{r 1d-2var-explore-proj-pca, fig.cap = "\\label{1d-2var-explore-proj-pca} The projected projection basis using principal components.
+The bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhausive search.
+The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end."}
+interp1 <- kol_cdf_1d_2var_better 
+interp2 <- kol_cdf_1d_2var_better_2
+interp3 <- kol_cdf_1d_2var_better_random
+
+bind_rows(interp1, interp2, interp3) %>% 
+  mutate(algorithm = c(rep("better1", nrow(interp1)), 
+                       rep("better2", nrow(interp2)), 
+                       rep("better_random", nrow(interp3))),
+         algorithm = fct_relevel(algorithm, c("better1", "better2", "better_random"))) %>% 
+  explore_proj_pca(col = algorithm) + 
+  scale_color_brewer(palette = "Dark2")
+
+```
+
 # Implementation: Ferrn pacakge

### 2020-07-13 paper/paper.tex
more notation on the searching space and kol_cdf index

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, bbm} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
-\(\mathbf{P}_{n \times p}\) is the projected data and
+\(\mathbf{P}_{n \times d}\) is the projected data and
 \(\mathbf{A}_{p\times d}\) is the projection basis Define
@@
 
-Below we will present several examples to diagnose different aspects of
-the projection pursuit optimisation. We will present:
+Below we will present several examples of diagnosing different aspects
+of the projection pursuit optimisation. We will present:
 
@@
 \item
-  a self-contained example to understand how to optimise a noisy and
-  complex index function.
+  a self-contained example to optimise a complex index function.
 \end{enumerate}
@@
 
-\hypertarget{toy-example-exploring-searching-points}{%
-\paragraph{Toy example: exploring searching
-points}\label{toy-example-exploring-searching-points}}
-
-The largest difficulties of exploring searching points is its unknown
-number of observations per \texttt{tries}. Mapping \texttt{id} on the
-x-axis will leave the \texttt{tries} with few observations a small space
-in the plot, while those \texttt{tries} with large number of search
-points (usually towards the end) occupy the vast majority of the plot
-space. This motivates to summarise the points in each \texttt{tries}. At
-each iteration, rather than knowing the index value of \emph{every}
-points, we are more interested to know a general summary of all the
-points and more importantly, the point with the largest
-\texttt{index\_val} since it prescribes the geodesic interpolation and
-future searches.
+It is a convention to map the time index of a time series on the x-axis
+and value to the y-axis. In our diagnostic plots, it is to map the
+iterator on the x-axis and index value to the y-axis. The next question
+is to ask, given four indices being defined in the global object, which
+one should be used? \emph{It depends on the }.
+
+As the searching goes towards the end, it is easy to have hundreds of
+candidate bases in each iteration with very close index value. When
+plotting, these iteration will occupy the vast majority of the plot
+space. This motivates to summarise the points in each iteration
+(\texttt{tries}). At each iteration, rather than knowing the index value
+of \emph{every} points, we are more interested to have a general summary
+of all the points and more importantly, the point with the largest
+\texttt{index\_val} since it prescribes the next geodesic interpolation
+and future searches.
+
+Using the notation we defined earlier, the plot can be defined as
+plotting \(j\) on the x-axis and the five point summary of the following
+groups on the y-axis:
+
+\[\mathbbm{1}(I_{j = j^*})\]
 
@@
 
-\emph{Example: exploring searching points} The larger dataset is used
-for 2D projection with \texttt{search\_better} and
-\texttt{max.tries\ =\ 500}. In Figure \ref{points}, the searching points
-with \texttt{id} and \texttt{tries} on the x-axis, colored by
-\texttt{tries}. Label at the bottom indicates the number of observations
-in each tries and facilitates the choice of cutoff to switch from point
-geometry to boxplot geometry (\texttt{cutoff\ =\ 15}). The line geometry
-suggests the largest improvement happens at \texttt{tries\ =\ 5}.
+\hypertarget{toy-example-exploring-searching-points}{%
+\paragraph{Toy example: exploring searching
+points}\label{toy-example-exploring-searching-points}}
+
+The larger dataset is used for 2D projection with
+\texttt{search\_better} and \texttt{max.tries\ =\ 500}. In Figure
+\ref{points}, the searching points with \texttt{id} and \texttt{tries}
+on the x-axis, colored by \texttt{tries}. Label at the bottom indicates
+the number of observations in each tries and facilitates the choice of
+cutoff to switch from point geometry to boxplot geometry
+(\texttt{cutoff\ =\ 15}). The line geometry suggests the largest
+improvement happens at \texttt{tries\ =\ 5}.
 
@@
 
-\hypertarget{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}{%
-\paragraph{A more complex example: adding an interruption in the
-interpolation
-stage}\label{a-more-complex-example-adding-an-interruption-in-the-interpolation-stage}}
-
-\emph{Example: Interruption} This examples uses \texttt{search\_better}
-for a 2D projection on the larger dataset using the \texttt{holes}
-index. In the interpolation stage, continuous frames will be constructed
-to bridge the project on the current basis to the target basis. The
-target basis will become the current basis in the next iteration and the
-searching stage continues to find the next target basis. In the left
-panel of Figure \ref{interruption}, we observe that there are
-interpolating bases with higher index values than the target basis and
-these bases could be used to search for new basis in the next searching
-stage.
+\newpage
+
+The number of interpolation points is much managable than the searching
+points and always shows a pattern of gradual increase in the index
+value. Hence a line plot with interpolating order on the x-axis will be
+desirable. Using notation, the x-axis is a sequence
+\[x = \{1, 2, \cdots, k_1, \cdots, \sum_{i=1}^Jk_i\}\] and y-axis are
+the index values defined by \(y = I_{x}\).
+
+\hypertarget{a-more-complex-example-interruption}{%
+\paragraph{A more complex example:
+Interruption}\label{a-more-complex-example-interruption}}
+
+This examples uses \texttt{search\_better} for a 2D projection on the
+larger dataset using the \texttt{holes} index. In the interpolation
+stage, continuous frames will be constructed to bridge the project on
+the current basis to the target basis. The target basis will become the
+current basis in the next iteration and the searching stage continues to
+find the next target basis. In the left panel of Figure
+\ref{interruption}, we observe that there are interpolating bases with
+higher index values than the target basis and these bases could be used
+to search for new basis in the next searching stage.
 
@@
 
-Exploring the value of an objective function is cna be a simple task.
-Because the iterator has a time series ordering feature, it goes onto
-the x-axis as convention. The value of hte objective function goes to
-the y-axis. Other parameter of interest can be represented using color
-to present more information. {[}This is the case for our purpose with
-method, but when the level goes higher or the parameter is quantitative,
-color may not be a good option. {]}
-
-An option to explore of the searching space is to explore a reduced
-projection 2D space via principle component analysis or other
-dimensional reduction methods. In this way, the first two principle
-components will take up the x and y axis and all the other information
-will be mapped using other aesthetic attribution. {[}need to think
-further about how different type of variables can be shown{]}. While
-explore the reduced space is an initial attempt to understand the
-searching space, there are existing technology for rotating a higher
-dimensional space for visualisation. Geozoo is an option. It generates
-random points on the high dimensional space and we can overlay it with
-the points on the optimisation path to visualise the spread of it on the
-high-D sphere.
+\newpage
 
@@
 
-The parameter to explore can be a vector or matrix instead of scalar,
-for example, the projection basis in tour. This imposes difficulties in
-visualisation since humans are bounded to perceive at most three
-dimensions in a plot. Thus, principal component analysis is used to
-reduce the dimension of projection bases and the first two principal
-components are mapped to the x and y axis of the plot. Additional
-variable of interest could be mapped to the color aesthetics for
-exploration.
+In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
+are usually of dimension \(p \times d\) and hence can't be visualised in
+a 2D plot. An option to explore the searching space of these bases is to
+explore a reduced space via principal component analysis (PCA). We first
+flat each projection bases \(\mathbf{A}\) to a row vector and stack them
+by rows. This gives a matrix of dimension \(n\) by \((p*d)\). Centering
+and scaling are usually done before performing PCA and we denotes the
+pre-processed matrix as \(\tilde{\mathbf{A}}_{n \times (p \times d)}\).
+By definition, the principal components maximise the variance of the
+projection bases in the reduced space. Using lagrange multipler, this is
+equivalent to find the eigenvector of the variance-covariance matrix:
+\(\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}\). Decomposing
+\(\tilde{\mathbf{A}}\) using singular value decomposition (SVD) gives
+\[\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}\] and
+hence the variance-covariance matrix
+
+\[\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}\]
+
+has eigenvector
+\(\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]\).
+The projection mapping to the x- and y-axis uses the first two principal
+components and hence
+\[\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2\]
+
+The resulting two vectors are of dimension \(n \times 1\), which is
+compatible with the dimension of the global object. THis means all the
+iterators and interesting other variables defined in the global object
+can be binded with the these two projections and can be mapped to other
+aesthetics (i.e.~color).
+
+In this way, the first two principle components will take up the x and y
+axis and all the other information will be mapped using other aesthetic
+attribution. {[}need to think further about how different type of
+variables can be shown{]}. While explore the reduced space is an initial
+attempt to understand the searching space, there are existing technology
+for rotating a higher dimensional space for visualisation. Geozoo is an
+option. It generates random points on the high dimensional space and we
+can overlay it with the points on the optimisation path to visualise the
+spread of it on the high-D sphere.
 
@@
 
+\newpage
+
 \hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
@@
 
-\emph{Example: Noisy index function} The interpolation path of holes
-index, as seen in Figure \ref{interruption}, is smooth, while this may
-not be the case for more complicated index functions. \texttt{kol\_cdf}
-index, an index function based on Kolmogorov test, compares the
-difference between a projection matrix and a randomly generated normal
-distribution based on cumulated distribution function (CDF). Several
-diagnostic visualisations below show the characteristic of this index
-function.
+The interpolation path of holes index, as seen in Figure
+\ref{interruption}, is smooth, while this may not be the case for more
+complicated index functions. \texttt{kol\_cdf} index, an 1D projection
+index function based on Kolmogorov test, compares the difference between
+the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly
+generated normal distribution, \(y_n\) based on the empirical cumulated
+distribution function (ECDF). Denotes the ECDF function as \(F(u)\) with
+subscript incidating the variable, the Kolmogorove statistics defined by
+
+\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
+
+can be seen as a function of the projection matrix
+\(\mathbf{A}_{p \times 1}\) and hence a valid index function.
+
+\hypertarget{explore-index-value}{%
+\paragraph{Explore index value}\label{explore-index-value}}
 
 Figure \ref{kol-cdf} compares the tracing plot of the interpolating
-points for \texttt{search\_geodesic} and \texttt{search\_better} on 1D
-projection for \texttt{kol\_cdf} index and the interpolation path shows
-a zig-zag pattern. Polishing step has done much more work to reach the
-final index value for \texttt{search\_geodesic} than
-\texttt{search\_better} and this indicates that noisy index's favour of
-a random search method than ascent method {[}the order of the examples
-needs to be rearranged since I haven't introduced polishing here.
-xxx{]}.
+points when using different optimisation algorithms:
+\texttt{search\_geodesic} and \texttt{search\_better}. One can observe
+that
+
+\begin{itemize}
+\tightlist
+\item
+  The index value of \texttt{kol\_cdf} index is much smaller than that
+  of holes index
+\item
+  The link of index values from interpolation bases are no longer smooth
+\item
+  Both algorithms reach a similar final index value after polishing
+\end{itemize}
+
+Polishing step has done much more work to find the final index value ub
+\texttt{search\_geodesic} than \texttt{search\_better} and this
+indicates \texttt{kol\_cdf} function favours of a random search method
+than ascent method.
 
@@
 
-The second example for \texttt{kol\_cdf} index uses 1D projection on the
-larger dataset. Since there are two informative variables and the
-projection is 1D, there are two local maximum when the projection matrix
-at \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\). As in Figure
-\ref{1d-2var-different-seeds}, different local maximum can be found
-using \texttt{search\_better} with different seeds while
-\texttt{search\_better\_random} can always find the global maximum, as
-in Figure \ref{1d-2var-better-random}, although at a high cost of number
-of points evaluated.
+Now we enlarge the dataset to include two informative variables:
+\texttt{x2} and \texttt{x3} and remain 1D projection. In this case, two
+local maximum appear when projection matrix being \([0, 1, 0, 0, 0, 0]\)
+and \([0, 0, 1 ,0, 0, 0]\).
+
+Using different seeds in \texttt{search\_better} allows us to find both
+local maximum as in Figure \ref{1d-2var-different-seeds}. Comparing the
+maximum of both, we can see that the global maximum happens when
+\texttt{x2} is found. It is natural to ask then if there is an algorithm
+that can find the global maximum without trying on different seeds?
+\texttt{search\_better\_random} manages to do it via a
+metropolis-hasting random search as shown in Figure
+\ref{1d-2var-better-random}, although at a higher cost of number of
+points to evaluate.
 
@@
 
+\hypertarget{explore-searching-space}{%
+\paragraph{Explore searching space}\label{explore-searching-space}}
+
+We can also plot the searching points of all the three algorithms in the
+searching space and explore their relative position against each other
+using principal components. As shown in Figure
+\ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only
+search a proportion of the searching space while better\_random produces
+a more exhausive search. The large overlapping of better1 and
+better\_random is explained by the fact that both algorithms finds x2 in
+the end.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/1d-2var-explore-proj-pca-1.pdf}
+\caption{\label{1d-2var-explore-proj-pca} The projected projection basis
+using principal components. The bases from better1 and better2 only
+search a proportion of the searching space while better\_random produces
+a more exhausive search. The large overlapping of better1 and
+better\_random is explained by the fact that both algorithms finds x2 in
+the end.}
+\end{figure}
+
 \hypertarget{implementation-ferrn-pacakge}{%

### 2020-07-14 paper/paper.Rmd
finish notation for plots and more plots for exploring searching space

@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools, bbm}
+  \usepackage{amssymb, amsmath, mathtools, dsfont}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 \begin{array}{ccccccc}
-\ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{j,\sum_{i = 1}^{j-1}l_i +1,\ast} &\cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1 ,\ast} \\
-\ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{j,\sum_{i = 1}^{j-1}l_i +2,\ast} &\cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2, \ast} \\
-\vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
-\ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{j, \sum_{i = 1}^{j-1}l_i + l, \ast}  &\cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i + l, \ast}\\
-\vdots & \vdots & \vdots & \ddots & \vdots & \ddots &\vdots   \\
-\mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{j, \sum_{i = 1}^{j}l_i , \ast}  &\cdots & \mathbf{A}_{J, \sum_{i = 1}^{J} l_i, \ast}\\
+\ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +1 ,\ast} \\
+\ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +2, \ast} \\
+\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
+\ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{J+1, \sum_{i = 1}^{J-1}l_i + l, \ast}  & \mathbf{A}_{J + 1, \sum_{i = 1}^{J}l_i + l, \ast}\\
+\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
+\ast & \mathbf{A}_{1, l_1-1, \ast} & \mathbf{A}_{2, l_1 + l_2-1, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i -1, \ast} & \mathbf{A}_{J+1, \sum_{i = 1}^{J+1} l_i-1, \ast}\\
+\mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J}l_i , \ast} & \ast\\
 \hline
-\ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{j, \ast,\sum_{i = 1}^{j-1}k_i + 1} & \cdots & \ast\\
-\ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{j, \ast, \sum_{i = 1}^{j-1}k_i +2} & \cdots & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{j, \ast, \sum_{i = 1}^{j-1}k_i +k} & \cdots & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots  & \ddots & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{j, \ast,\sum_{i = 1}^{j}k_i} & \cdots & \ast \\
+\ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J-1}k_i + 1}  & \ast\\
+\ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +2}  & \ast \\
+\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
+\ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +k}  & \ast \\
+\vdots & \vdots & \vdots & \ddots & \vdots   & \vdots \\
+\ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J}k_i} & \ast \\
 \end{array}
@@
 \vphantom{\vdots} \vdots \\
+\vphantom{A^J_2}\text{candidate basis} \\
 \vphantom{A^J}\text{current/target basis} \\
@@
 \right]
-\label{matrixA}
+\label{eq:matrixA}
 \end{equation}
 
-When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector as in Equation \ref{eq:vectorA} and the full data structure is presneted in Equation \ref{eq:data-structure}.
-
-All the basis can also be denoted usign a single subscript by their natural ordering as $\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n$.
+When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector denoting by a single subscript by their natural ordering i.e.
+all the bases in iteration 1 along with the starting basis as shown in Equation \ref{eq:vectorA}, can be denoted as $\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n$ where $n = 1 + l_1 + k_1$.
+The full data structure can thus be presneted as in Equation \ref{eq:data-structure}.
 
 \begin{equation}
-\begin{array}{cccccccc}
-\bigl[\mathbf{A}_{0, \ast, \ast} & 
+\begin{array}{cccccccccccc}
+\bigl[
+\mathbf{A}_{0, \ast, \ast} & 
 \mathbf{A}_{1, 1, \ast} & 
@@
 \cdots & 
-\mathbf{A}_{1,l,\ast} & 
-\cdots & 
-\mathbf{A}_{1,l_1, \ast}   \\ 
+\mathbf{A}_{1,l_1, \ast}   &
 \mathbf{A}_{1, \ast, 1} & 
@@
 \cdots &
-\mathbf{A}_{1, \ast, k} &
-\cdots &
 \mathbf{A}_{1, \ast, k_1} &
-\cdots &
-\mathbf{A}_{J, l_J, \ast} \bigr]
+\bigr]
 \end{array} 
@@
 \hline
-\mathbf{A}_1 & I_1 & 1 & 1 & \ast& 2 & S_{1, 1} & \cdots & S_{p, 1}\\
-\mathbf{A}_2 & I_2 & 1 & 2 &\ast & 3 & S_{1, 2} & \cdots & S_{p, 2}\\
+\mathbf{A}_1 & I_1 & 1 & 1 & \ast& 2 & S_{1, 1} & \cdots & S_{1, p}\\
+\mathbf{A}_2 & I_2 & 1 & 2 &\ast & 3 & S_{2, 1} & \cdots & S_{2, p}\\
 \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots\\
@@
 \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
-\mathbf{A}_n & I_n & J &  l_J & \ast & n & S_{1, n}& \cdots & S_{p, n}\\
+\mathbf{A}_n & I_n & J &  l_J & \ast & n & S_{n, 1}& \cdots & S_{n, p}\\
 \end{array}
@@
 
-where $n = 1+ \sum_{i = 1}^J l_i + \sum_{i = 1}^{J-1} k_i$
+where $n = 1 + \sum_{i = 1}^{J+1} l_i -1 + \sum_{i = 1}^{J} k_i=  \sum_{i = 1}^{J+1} l_i + \sum_{i = 1}^{J} k_i$
 
@@
 
-Using the notation we defined earlier, the plot can be defined as plotting $j$ on the x-axis and the five point summary of the following groups on the y-axis:
+Using the notation we defined earlier, the plot can be defined as plotting
 
-$$\mathbbm{1}(I_{j = j^*})$$
+$$x = \{1, 2, \cdots J + 1\}$$
+
+against the five point summary of each $j^*$ on the y-axis:
+
+$$y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$$
+
+where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$. $\mathds{1}(.)$ is the identity operator.
 
@@
 Hence a line plot with interpolating order on the x-axis will be desirable.
-Using notation, the x-axis is a sequence $$x = \{1, 2, \cdots, k_1, \cdots, \sum_{i=1}^Jk_i\}$$ and y-axis are the index values defined by $y = I_{x}$.
+Using notation, the x-axis is a sequence $$x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}$$ and y-axis are the index values defined by $$y = \{I_{jlk}: k = x\}$$
 
@@
 
-#### Another example: polishing the final projection basis
+#### Another example: Polish
 
@@
 
-*Example: Polish* In principle, all the optimisation routines should result in the same output on the same problem while this may not be true in real application.
-This motivates the creation of a polishing search that polishes the ending basis and achieves unity on different methods.
+In principle, all the optimisation routines should result in the same output for the same problem while this may not be true in real application.
+This motivates the creation of a polishing search that polishes the final basis and achieves unity on different methods.
 
-`search_polish` takes the ending basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`.
+`search_polish` takes the final basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`.
 Among the `n_sample` basis, the one with the largest index value becomes the candidate basis.
@@
 ```
+
+Polishing points form on itself a global structure with candidate bases and output bases.
+If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$
+
+Then the index plot incorporating both the interpolating points and polishing points can be defined as
+
+$$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$
+$$y = \{I_{jlk}: k = x\}$$
 

### 2020-07-14 paper/paper.tex
finish notation for plots and more plots for exploring searching space

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, bbm} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, dsfont} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 \right]
-\label{matrixA}
+\label{eq:matrixA}
 \end{equation}
@@
 When aggregated with information of index value and other interesting
-parameter, it is easier to write the basis as a column vector as in
-Equation \ref{eq:vectorA} and the full data structure is presneted in
-Equation \ref{eq:data-structure}.
-
-All the basis can also be denoted usign a single subscript by their
-natural ordering as
-\(\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n\).
+parameter, it is easier to write the basis as a column vector denoting
+by a single subscript by their natural ordering i.e.~all the bases in
+iteration 1 along with the starting basis as shown in Equation
+\ref{eq:vectorA}, can be denoted as
+\(\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n\) where
+\(n = 1 + l_1 + k_1\). The full data structure can thus be presneted as
+in Equation \ref{eq:data-structure}.
 
 \begin{equation}
-\begin{array}{cccccccc}
-\bigl[\mathbf{A}_{0, \ast, \ast} & 
+\begin{array}{cccccccccccc}
+\bigl[
+\mathbf{A}_{0, \ast, \ast} & 
 \mathbf{A}_{1, 1, \ast} & 
@@
 
-where \(n = 1+ \sum_{i = 1}^J l_i + \sum_{i = 1}^{J-1} k_i\)
+where
+\(n = 1 + \sum_{i = 1}^{J+1} l_i -1 + \sum_{i = 1}^{J} k_i= \sum_{i = 1}^{J+1} l_i + \sum_{i = 1}^{J} k_i\)
 
@@
 Using the notation we defined earlier, the plot can be defined as
-plotting \(j\) on the x-axis and the five point summary of the following
-groups on the y-axis:
-
-\[\mathbbm{1}(I_{j = j^*})\]
+plotting
+
+\[x = \{1, 2, \cdots J + 1\}\]
+
+against the five point summary of each \(j^*\) on the y-axis:
+
+\[y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\]
+
+where \(q\) takes one of \(0, 25, 50, 75, 100\) and \(Q_q(x)\) finds the
+qth-quantile for the vector \(x\). \(\mathds{1}(.)\) is the identity
+operator.
 
@@
 desirable. Using notation, the x-axis is a sequence
-\[x = \{1, 2, \cdots, k_1, \cdots, \sum_{i=1}^Jk_i\}\] and y-axis are
-the index values defined by \(y = I_{x}\).
+\[x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}\] and y-axis are the index values
+defined by \[y = \{I_{jlk}: k = x\}\]
 
@@
 
-\hypertarget{another-example-polishing-the-final-projection-basis}{%
-\paragraph{Another example: polishing the final projection
-basis}\label{another-example-polishing-the-final-projection-basis}}
-
-\emph{Example: Polish} In principle, all the optimisation routines
-should result in the same output on the same problem while this may not
-be true in real application. This motivates the creation of a polishing
-search that polishes the ending basis and achieves unity on different
-methods.
-
-\texttt{search\_polish} takes the ending basis of a given search as the
+\hypertarget{another-example-polish}{%
+\paragraph{Another example: Polish}\label{another-example-polish}}
+
+In principle, all the optimisation routines should result in the same
+output for the same problem while this may not be true in real
+application. This motivates the creation of a polishing search that
+polishes the final basis and achieves unity on different methods.
+
+\texttt{search\_polish} takes the final basis of a given search as the
 current basis and uses a brutal-force approach to sample a large number
@@
 \end{Shaded}
+
+Polishing points form on itself a global structure with candidate bases
+and output bases. If we're allowed to abuse the notation a bit, we
+denote \[A_{J, \ast, \sum_{i = 1}^J k_i + 1}\] as the first polishing
+basis and \[A_{J, \ast, \sum_{i = 1}^J k_i + M}\] as the last polishing
+basis, where
+\[M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})\]
+
+Then the index plot incorporating both the interpolating points and
+polishing points can be defined as
+
+\[x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1) \cdots,  \sum_{i = 1}^J k_i + M\}\]
+\[y = \{I_{jlk}: k = x\}\]
 

### 2020-07-20 paper/paper.Rmd
move notation to the first intro of PP; modify notation in the 3 algos

@@
 clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation: $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis  Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or index function and can be used to measure the "interestingness" of the projection.
-A number of indice functions have been proposed in the literature and this includes Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
-[Any literature on holes index?
-xxx]
+Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
+Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and can be used to measure the "interestingness" of the projection.
+A number of indice functions have been proposed in the literature including Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 
@@
 Details of the mathematical construction of a tour path can be found in @buja2005computational.
-Modified from @buja2005computational, Figure \ref{tour-path} vividly depicts the tour path in guided tour.
 Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 
-```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
-```
-
-[doubt if I need the following. xxx - maybe an algorithmatic version of guide tour would be helpful for non-expert to understand the method?]
 <!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis.
@@
 
-## Optimisation in projection pursuit
+## Notation
 
-[thinking if the following three should be presented as algorithms or plain description is fine. xxx]
+For a projection basis, let the three-subscript notation $$\mathbf{A}_{jlk}$$ represent the projection matrix in iteration $j$ with either a searching index $l$ or an interpolation index $k$.
+The placeholder $*$ will be the substitue for the non-applicable index.
+To give two examples, $\mathbf{A}_{12*}$ denotes the second searching basis in the first iteration; $\mathbf{A}_{1*2}$ denotes the second interpolation basis in the first iteration.
 
-Now we formulate the problem in mathmetical notation.
-Projection pursuit finds the projection basis $\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfying the following optimisation problem:
+If we are being pedantic, all the target bases, except the starting basis, should have a searching index since it is the last candidate basis that successfully allows the optimisation algorithm to end and output a new basis for interpolation.
+If we denotes the length of each search (number of basis searched in each iteration) as $l_1, l_2, \cdots, l_J$ and the length of each interpolation is $k_1, k_2, \cdots, k_J$ , then the searching index for target basis in each iteration will become $l_1, l_1 + l_2, \cdots, l_1 +l_2+\cdots l_J$.
+To avoid this complex notation involving opteration in the subscript, we simplify it using a superscript as $$\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}$$ where $\mathbf{A}^{0}$ is the starting basis.
+Notice that in the case when the complex notation can't be avoided, we shall separate each index using a comma and  write the index in brackets if necessary.
+Using the notation above, all the bases on the interpolation path (including target and interpolation bases) can be written as in Equation \ref{eq:interp-path}.
+
+\begin{equation}
+\{ 
+\mathbf{A}^0, 
+\mathbf{A}^1, 
+\mathbf{A}_{1 \ast 1}, 
+\mathbf{A}_{1 \ast 2},
+\cdots, 
+\mathbf{A}^2, 
+\mathbf{A}_{2 \ast 1}, 
+\mathbf{A}_{2 \ast 2}, 
+\cdots, 
+\mathbf{A}^J,
+\mathbf{A}_{J \ast 1}, 
+\mathbf{A}_{J \ast 2}, 
+\cdots, 
+\mathbf{A}_{J \ast K},
+\}
+\label{eq:interp-path}
+\end{equation}
+
+where $K$ denotes the last interpolating index.
+A visual representation of the interpolation path ( for iteration one) modified from @buja2005computational is shown in Figure \ref{tour-path}.
+If we use a dot to represent a plane in the searching space, the searching points along with the current and target basis in iteration one can sketched in Figure \ref{searching-points}.
+Combine the two above, we can write all the bases in the first iteration (including the starting basis) in its natural ordering as in Equation \ref{eq:vectorA}.
+
+\begin{equation}
+\{
+\mathbf{A}^{0}, 
+\mathbf{A}_{1 1\ast}, 
+\mathbf{A}_{1 2\ast}, 
+\cdots, 
+\mathbf{A}^{1},
+\mathbf{A}_{1\ast 1}, 
+\mathbf{A}_{1\ast 2},  
+\cdots, 
+\mathbf{A}_{1\ast k_1}
+\}
+\label{eq:vectorA}
+\end{equation}
+
+```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
+```
+
+```{r searching-points, out.width="100%", out.height="60%", fig.cap="\\label{searching-points}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
+```
+
+## Optimisation problem and existing algorithms
+
+Now we begin to formulate the optimisation problem.
+Given a randomly generated starting basis $\mathbf{A}^0$, projection pursuit finds the final projection basis $\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfies the following optimisation problem:
 
@@
 
-via an iterative search: $\{\mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
-Guided tour interpolates geodesically between each basis found by projection pursuit $A^j$ to form a tour path: $$\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}$$
-
-Where $\mathbf{A}^0$ is the starting basis and $\mathbf{A}^J$ is the final basis that has the largest index value in the searching space.
-Here we use left subscript to denote the index of the interpolation bases in each iteration i.e.
-${}_{2}\mathbf{A}^1$ means the second interpolation basis in the first iteration.
-The right subcript is reserved for the index of searching points in the optimisation algorithm.
+via an iterative search of target basis: $\{\mathbf{A}^0, \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
 
 There are three existing methods for optimisating PPI function and we review them below.
+
 @posse1995projection proposed a stochastic direct search method, a random search algorithm.
-Given a current basis $\mathbf{A}^{j}$, a candidate basis $\mathbf{A}^{j +1}_l$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}^{j}$ by $$\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$$.
-If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $\mathbf{A}^{j + 1}$ along with other metadata and the current iteration stops.
+Given a current basis $\mathbf{A}^{j-1}$, a candidate basis $\mathbf{A}_{jl*}$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}^{j}$ by $\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}$.
+If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $\mathbf{A}^{j}$ along with other metadata and the current iteration stops.
 If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted.
@@
 
-\newpage
-\begin{algorithm}[H]
+\begin{algorithm}
 \SetAlgoLined
-  initialisation: $c = 0$, $l = 1$ $l_{\max}$, $\mathbf{A}^{j}$\;
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+  initialisation\;
   \While{$l < l_{\max}$}{
-    $\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$\;
-    $I^{j+1}_l = f(\mathbf{A}^{j+1}_l)$\;
-    \eIf{$I^{j+1}_l > I^{j}$}{
-      $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
+    Generate $\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{j, l,*}$ is orthonormal\;
+    Compute $I_{jl*}  = f(\mathbf{A}_{jl*})$\;
+    \eIf{$I_{jl*} > I^{j-1}$}{
+      $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
       }{
@@
   }
-  \caption{random search: my first algorithm}
+  \caption{random search}
   \label{random-search}
@@
 Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
+See Algorithm \ref{search-geodesic} for a summarisation of the steps.
 
-\newpage
-\begin{algorithm}[H]
+\begin{algorithm}
 \SetAlgoLined
-  initialisation: $l_{\max}$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+\SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+  initialisation\;
   \While{$l < l_{\max}$}{
-    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
-    Find $A^{J_1}_l$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
-    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
-    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
-    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
-    ...\;
-    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
-    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
-      \If{$p_{diff} > 0.001$}{
-        $A^{j+1} = A^{(J_2)}$\;
-        $I^{j+1} = I^{(J_2)}$\;
-        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+    Generate ten bases in five random directions: $\mathbf{A}_{jl*}: \mathbf{A}_{j,(l + 9), *}$ within a small neighbourhood $\text{dist}$\;
+    Find the direction with the largest index value: $\mathbf{A}_{j, l_{d}, *}$\;
+    Construct the geodesic from $\mathbf{A}^{j-1}$ to $\mathbf{A}_{j, l_{d}, *}$\;
+    Find $\mathbf{A}_{j,l_g, *}$ on the geodesic that has the largest index value \;
+    Compute $I_{j,l_g, *} = f(\mathbf{A}_{j,l_g, *})$, $p_{\text{diff}} = (I_{j,l_g, *} - I^{j-1})/I_{j,l_g, *}$\;
+      \If{$p_{\text{diff}} > 0.001$}{
+        $A^{j} = A_{j,l_g, *}$, $I^{j} = I_{j,l_g, *}$\;
       }
-    $i = i + 1$\;
+    $l = l + 1$\;
   }
-  \caption{searchgeodesic: this part has problem!}
-  \label{search_geodesic}
+  \caption{search geodesic}
+  \label{search-geodesic}
 \end{algorithm}
@@
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}$.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$ where $I^j$ and $I^{j+1}_l$ are the index value of the current and candidate basis respectively.
+If it is smaller, the candidate is accepted with probability $P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}$ where $I^{j-1}$ and $I_{jl*}$ are the index value of the current and candidate basis respectively.
 The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
 
-\newpage
-\begin{algorithm}[H]
+\begin{algorithm}
 \SetAlgoLined
-  initialisation: $T_0$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+\SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+  initialisation\;
   \While{$l < l_{\max}$}{
-    $\mathbf{A}^{j+1}_{l} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
-    $I^{j+1}_l = f(\mathbf{A}^{j+1}_{l})$\;
-    $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I^{j+1}_l > I^{j}$}{
-        $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
-        $I^{j+1} = I^{j+1}_l$\;
-        \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$,$l$ and relevant metadata}
+    Generate $\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{jl*}$ is orthonormal \;
+    Commpute $I_{jl*} = f(\mathbf{A}_{jl*})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{jl*} > I^{j-1}$}{
+        $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
       }{
-        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$\;
-        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
-        \If{$\text{prob} > \text{rand}$}{
-          \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$, $l$ and relevant metadata}
+        Compute $P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}$\;
+        Draw $D$ from a uniform distribution: $D \sim \text{Unif(0, 1)}$\;
+        \If{$P > D$}{
+          $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
         }
@@
 \end{algorithm}
-
-\newpage
 
@@
 Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
-Given an optimisation algorithm, two primary variables of interest are the *decision variable* and the *value of the objective function*.
-To align with the notation used previously for projection pursuit, we denote the decision variable to be $\mathbf{A}$, which are matrices in projection pursuit algorithms and value of the objective function to be $I$, the index function.
+Given an optimisation algorithm, two primary variables of interest are the *decision variable: $\mathbf{A}$* and the *value of the objective function: $I$*.
 
-*Iterators* indexes the data collected and has a time series feature that prescribes the its natural ordering.
-Given the complexity of the algorithm, there could be multiple iterators.
-The simpliest iterator: *index per observation*, is a unique identifier for each observation in the data.
-For each level of iteration, we design two indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
+*Iterators* indexes the data collected and has a time series feature that prescribes the its natural ordering of the decision varible in the searching.
+The simpliest iterator, *index per observation: $t$*, is a unique identifier for each observation in the data.
+For each level of iteration, we design two types of indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
 *index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
 A sketch of the difference between these three iterators is provided in Figure \ref{iterators}.
-Depend on the number of iteration level ($L$), the total number of iterators will differ and can be calculated as $1 + 2L$.
 
@@
 There could exist other parameters that are also of our interest but can't be classified as one of the three categories.
-They are defined under the fourth category: *other parameter of interest: $O$*.
+They are defined under the fourth category: *other parameter of interest: $S$*.
 In projection pursuit, three other parameters of our interest includes `method`, `alpha` and `info`.
@@
 
-## Notation
+<!-- \begin{equation} -->
+<!-- \left[ -->
+<!-- \begin{array}{ccccccc} -->
+<!-- \ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +1 ,\ast} \\ -->
+<!-- \ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +2, \ast} \\ -->
+<!-- \vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\ -->
+<!-- \ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{J+1, \sum_{i = 1}^{J-1}l_i + l, \ast}  & \mathbf{A}_{J + 1, \sum_{i = 1}^{J}l_i + l, \ast}\\ -->
+<!-- \vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\ -->
+<!-- \ast & \mathbf{A}_{1, l_1-1, \ast} & \mathbf{A}_{2, l_1 + l_2-1, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i -1, \ast} & \mathbf{A}_{J+1, \sum_{i = 1}^{J+1} l_i-1, \ast}\\ -->
+<!-- \mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J}l_i , \ast} & \ast\\ -->
+<!-- \hline -->
+<!-- \ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J-1}k_i + 1}  & \ast\\ -->
+<!-- \ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +2}  & \ast \\ -->
+<!-- \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\ -->
+<!-- \ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +k}  & \ast \\ -->
+<!-- \vdots & \vdots & \vdots & \ddots & \vdots   & \vdots \\ -->
+<!-- \ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J}k_i} & \ast \\ -->
+<!-- \end{array} -->
+<!-- \right] -->
+<!-- =  -->
+<!-- \left[ -->
+<!-- \begin{array}{c} -->
+<!-- \vphantom{A^J_1}\text{candidate basis} \\ -->
+<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
+<!-- \vphantom{\vdots} \vdots \\ -->
+<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
+<!-- \vphantom{\vdots} \vdots \\ -->
+<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
+<!-- \vphantom{A^J}\text{current/target basis} \\ -->
+<!-- \hline -->
+<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
+<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
+<!-- \vphantom{\text{NULL}}\vdots \\ -->
+<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
+<!-- \vphantom{\vdots} \vdots \\ -->
+<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
+<!-- \end{array} -->
+<!-- \right] -->
+<!-- \label{eq:matrixA} -->
+<!-- \end{equation} -->
 
-Now we formulate the idea of global object using mathmetical notation.
-Let $$\mathbf{A}_{j,l,k}$$ denotes the projection matrix in iteration $j$.
-If it is searching basis produced in the optimisation, then it has a subscript $l$.
-If it is an interpolation basis, it has a $k$ subscript.
-For example $\mathbf{A}_{1,2,/}$ means the second searching basis in the first iteration, $\mathbf{A}_{1, /, 2}$ means the second interpolation basis that on the tour path from the current basis $\mathbf{A}^0$ to the target basis found in iteration 1: $\mathbf{A}^1$.
-The target basis found in each iteration, $\mathbf{A}_{j, l_j, /}$ can also be simplified using a superscript as $\mathbf{A}^j$.
+To write down the notation for the data structure, it is easier to write the bases as a column vector denoting by a single subscript by their natural ordering:
 
-Let $\mathbf{A}^0$ denotes the starting basis.
-Superscript $j$ is used to denote the number of iteration (index per iteration).
-Candidate basis are denoted using subscript on the right and interpolation basis are denoted using subscript on the left.
-For example, $\mathbf{A}^1_{2}$ means the second candidate basis in the first iteration and ${}_{2}\mathbf{A}^1$ means the second interpolation basis that forms the tour path from the current basis $\mathbf{A}^0$ to the target basis $\mathbf{A}^1$.
-In iteration two, $\mathbf{A}^1$ becomes the current basis and projection pursuit searches through $\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}$ to find the target basis $\mathbf{A}^2$.
-The tour path from $\mathbf{A}^1$ to $\mathbf{A}^2$ is constructed using the interpolation bases ${}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2$.
-All the bases together can be arranged in a format that each coloumn represents an iteration as in Equation \ref{eq:matrixA}.
+\begin{equation}
+\{
+\mathbf{A}_0, 
+\mathbf{A}_1, 
+\mathbf{A}_2, 
+\cdots,
+\mathbf{A}_T
+\label{vector-basis}
+\}
+\end{equation}
+
+where $T$ is the total number of plane serached by the optimisation algorithm including all the target planes, all the interpolating planes and all the candidate planes but NOT the starting plane.
+The full data structure can thus be shown as in Equation \ref{eq:data-structure}.
 
@@
 \left[
-\begin{array}{ccccccc}
-\ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +1 ,\ast} \\
-\ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +2, \ast} \\
-\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
-\ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{J+1, \sum_{i = 1}^{J-1}l_i + l, \ast}  & \mathbf{A}_{J + 1, \sum_{i = 1}^{J}l_i + l, \ast}\\
-\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
-\ast & \mathbf{A}_{1, l_1-1, \ast} & \mathbf{A}_{2, l_1 + l_2-1, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i -1, \ast} & \mathbf{A}_{J+1, \sum_{i = 1}^{J+1} l_i-1, \ast}\\
-\mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J}l_i , \ast} & \ast\\
+\begin{array}{c|cc|ccc|ccc}
+t & \mathbf{A} & I & j &  l  & k & S_{1}& \cdots & S_{p}\\
 \hline
-\ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J-1}k_i + 1}  & \ast\\
-\ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +2}  & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +k}  & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots   & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J}k_i} & \ast \\
-\end{array}
-\right]
-= 
-\left[
-\begin{array}{c}
-\vphantom{A^J_1}\text{candidate basis} \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{A^J}\text{current/target basis} \\
+1 & \mathbf{A}_0 & I_0 & 0 & \ast & \ast & \ast & \cdots & \ast\\
 \hline
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\text{NULL}}\vdots \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\end{array}
-\right]
-\label{eq:matrixA}
-\end{equation}
-
-When aggregated with information of index value and other interesting parameter, it is easier to write the basis as a column vector denoting by a single subscript by their natural ordering i.e.
-all the bases in iteration 1 along with the starting basis as shown in Equation \ref{eq:vectorA}, can be denoted as $\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n$ where $n = 1 + l_1 + k_1$.
-The full data structure can thus be presneted as in Equation \ref{eq:data-structure}.
-
-\begin{equation}
-\begin{array}{cccccccccccc}
-\bigl[
-\mathbf{A}_{0, \ast, \ast} & 
-\mathbf{A}_{1, 1, \ast} & 
-\mathbf{A}_{1,2,\ast} & 
-\cdots & 
-\mathbf{A}_{1,l_1, \ast}   &
-\mathbf{A}_{1, \ast, 1} & 
-\mathbf{A}_{1, \ast, 2} & 
-\cdots &
-\mathbf{A}_{1, \ast, k_1} &
-\bigr]
-\end{array} 
-\label{eq:vectorA}
-\end{equation}
-
-\begin{equation}
-\left[
-\begin{array}{c|c|cccc|ccc}
-\mathbf{A} & I & j &  l  & k & n & S_{1}& \cdots & S_{p}\\
+2 & \mathbf{A}_1 & I_1 & 1 & 1 & \ast & S_{1, 1} & \cdots & S_{1, p}\\
+3 & \mathbf{A}_2 & I_2 & 1 & 2 & \ast & S_{2, 1} & \cdots & S_{2, p}\\
+\vdots & \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
+\vdots & \vdots & \vdots & 1 & l_1 & \ast & \vdots &  \cdots & \vdots\\
 \hline
-\mathbf{A}_0 & I_0 & 0 & \ast & \ast & 1 & \ast & \cdots & \ast\\
+\vdots &\vdots & \vdots & 1 & \ast & 1& \vdots &  \cdots & \vdots\\
+\vdots &\vdots &\vdots& 1 & \ast & 2& \vdots& \cdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+\vdots &\vdots &\vdots & 1 & \ast & k_1&\vdots & \cdots & \vdots\\
 \hline
-\mathbf{A}_1 & I_1 & 1 & 1 & \ast& 2 & S_{1, 1} & \cdots & S_{1, p}\\
-\mathbf{A}_2 & I_2 & 1 & 2 &\ast & 3 & S_{2, 1} & \cdots & S_{2, p}\\
-\vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots\\
-\vdots & \vdots & 1 & l_1 & \ast & \vdots & \vdots & \cdots & \vdots\\
-\hline
-\vdots & \vdots & 1 & \ast & 1& \vdots &\vdots & \cdots & \vdots\\
-\vdots &\vdots& 1 & \ast & 2& \vdots&\vdots & \cdots & \vdots\\
-\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
-\vdots &\vdots & 1 & \ast & k_1&\vdots &\vdots & \cdots & \vdots\\
-\hline
-\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
-\mathbf{A}_n & I_n & J &  l_J & \ast & n & S_{n, 1}& \cdots & S_{n, p}\\
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+T+1 & \mathbf{A}_T & I_T & J &  l_J & \ast & S_{T, 1}& \cdots & S_{T, p}\\
 \end{array}
@@
 \end{equation}
-
-where $n = 1 + \sum_{i = 1}^{J+1} l_i -1 + \sum_{i = 1}^{J} k_i=  \sum_{i = 1}^{J+1} l_i + \sum_{i = 1}^{J} k_i$
 

### 2020-07-20 paper/paper.tex
move notation to the first intro of PP; modify notation in the 3 algos

@@
 projecting it in lower dimensions. Let \(\mathbf{X}_{n \times p}\) be a
-data matrix, an n-d projection can be seen as a linear transformation:
+data matrix, an n-d projection can be seen as a linear transformation
 \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
@@
 \(\mathbf{P}_{n \times d}\) is the projected data and
-\(\mathbf{A}_{p\times d}\) is the projection basis Define
+\(\mathbf{A}_{p\times d}\) is the projection basis. Define
 \(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
@@
 this function is commonly known as the projection pursuit index (PPI)
-function, or index function and can be used to measure the
+function, or the index function and can be used to measure the
 ``interestingness'' of the projection. A number of indice functions have
-been proposed in the literature and this includes Legendre index
+been proposed in the literature including Legendre index
 \citep{friedman1974projection}, hermite index
@@
 \citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
-index \citep{lee2010projection}. {[}Any literature on holes index?
-xxx{]}
+index \citep{lee2010projection}.
 
@@
 shortest. Details of the mathematical construction of a tour path can be
-found in \citet{buja2005computational}. Modified from
-\citet{buja2005computational}, Figure \ref{tour-path} vividly depicts
-the tour path in guided tour. Guided tour, along with other types of
-tour, has been implemented in the \emph{tourr} package in R, available
-on the Comprehensive R Archive Network at
+found in \citet{buja2005computational}. Guided tour, along with other
+types of tour, has been implemented in the \emph{tourr} package in R,
+available on the Comprehensive R Archive Network at
 \url{https://cran.r-project.org/web/packages/tourr/}
@@
 
+\hypertarget{notation}{%
+\subsection{Notation}\label{notation}}
+
+For a projection basis, let the three-subscript notation
+\[\mathbf{A}_{jlk}\] represent the projection matrix in iteration \(j\)
+with either a searching index \(l\) or an interpolation index \(k\). The
+placeholder \(*\) will be the substitue for the non-applicable index. To
+give two examples, \(\mathbf{A}_{12*}\) denotes the second searching
+basis in the first iteration; \(\mathbf{A}_{1*2}\) denotes the second
+interpolation basis in the first iteration.
+
+If we are being pedantic, all the target bases, except the starting
+basis, should have a searching index since it is the last candidate
+basis that successfully allows the optimisation algorithm to end and
+output a new basis for interpolation. If we denotes the length of each
+search (number of basis searched in each iteration) as
+\(l_1, l_2, \cdots, l_J\) and the length of each interpolation is
+\(k_1, k_2, \cdots, k_J\) , then the searching index for target basis in
+each iteration will become
+\(l_1, l_1 + l_2, \cdots, l_1 +l_2+\cdots l_J\). To avoid this complex
+notation involving opteration in the subscript, we simplify it using a
+superscript as
+\[\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}\]
+where \(\mathbf{A}^{0}\) is the starting basis. Notice that in the case
+when the complex notation can't be avoided, we shall separate each index
+using a comma and write the index in brackets if necessary. Using the
+notation above, all the bases on the interpolation path (including
+target and interpolation bases) can be written as in Equation
+\ref{eq:interp-path}.
+
+\begin{equation}
+\{ 
+\mathbf{A}^0, 
+\mathbf{A}^1, 
+\mathbf{A}_{1 \ast 1}, 
+\mathbf{A}_{1 \ast 2},
+\cdots, 
+\mathbf{A}^2, 
+\mathbf{A}_{2 \ast 1}, 
+\mathbf{A}_{2 \ast 2}, 
+\cdots, 
+\mathbf{A}^J,
+\mathbf{A}_{J \ast 1}, 
+\mathbf{A}_{J \ast 2}, 
+\cdots, 
+\mathbf{A}_{J \ast K},
+\}
+\label{eq:interp-path}
+\end{equation}
+
+where \(K\) denotes the last interpolating index. A visual
+representation of the interpolation path ( for iteration one) modified
+from \citet{buja2005computational} is shown in Figure \ref{tour-path}.
+If we use a dot to represent a plane in the searching space, the
+searching points along with the current and target basis in iteration
+one can sketched in Figure \ref{searching-points}. Combine the two
+above, we can write all the bases in the first iteration (including the
+starting basis) in its natural ordering as in Equation \ref{eq:vectorA}.
+
+\begin{equation}
+\{
+\mathbf{A}^{0}, 
+\mathbf{A}_{1 1\ast}, 
+\mathbf{A}_{1 2\ast}, 
+\cdots, 
+\mathbf{A}^{1},
+\mathbf{A}_{1\ast 1}, 
+\mathbf{A}_{1\ast 2},  
+\cdots, 
+\mathbf{A}_{1\ast k_1}
+\}
+\label{eq:vectorA}
+\end{equation}
+
 \begin{figure}
@@
 
-{[}doubt if I need the following. xxx - maybe an algorithmatic version
-of guide tour would be helpful for non-expert to understand the
-method?{]}
-
-\hypertarget{optimisation-in-projection-pursuit}{%
-\subsection{Optimisation in projection
-pursuit}\label{optimisation-in-projection-pursuit}}
-
-{[}thinking if the following three should be presented as algorithms or
-plain description is fine. xxx{]}
-
-Now we formulate the problem in mathmetical notation. Projection pursuit
-finds the projection basis
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{searching-points}An illustration of the tour path}\label{fig:searching-points}
+\end{figure}
+
+\hypertarget{optimisation-problem-and-existing-algorithms}{%
+\subsection{Optimisation problem and existing
+algorithms}\label{optimisation-problem-and-existing-algorithms}}
+
+Now we begin to formulate the optimisation problem. Given a randomly
+generated starting basis \(\mathbf{A}^0\), projection pursuit finds the
+final projection basis
 \(\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
-\(\mathbf{a}_i\) is a \(p \times 1\) vector, satisfying the following
+\(\mathbf{a}_i\) is a \(p \times 1\) vector, satisfies the following
 optimisation problem:
@@
 
-via an iterative search:
-\(\{\mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\). Guided
-tour interpolates geodesically between each basis found by projection
-pursuit \(A^j\) to form a tour path:
-\[\{ \mathbf{A}^0, \mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, \mathbf{A}^{J-1}, {}_{1}\mathbf{A}^{J-1}, \cdots {}_{k_m}\mathbf{A}^{J-1} ,\mathbf{A}^J \}\]
-
-Where \(\mathbf{A}^0\) is the starting basis and \(\mathbf{A}^J\) is the
-final basis that has the largest index value in the searching space.
-Here we use left subscript to denote the index of the interpolation
-bases in each iteration i.e. \({}_{2}\mathbf{A}^1\) means the second
-interpolation basis in the first iteration. The right subcript is
-reserved for the index of searching points in the optimisation
-algorithm.
+via an iterative search of target basis:
+\(\{\mathbf{A}^0, \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
 
 There are three existing methods for optimisating PPI function and we
-review them below. \citet{posse1995projection} proposed a stochastic
-direct search method, a random search algorithm. Given a current basis
-\(\mathbf{A}^{j}\), a candidate basis \(\mathbf{A}^{j +1}_l\) is sampled
-in the neighbourhood defined by the radius of the p-dimensional sphere,
-\(\alpha\) of the current basis \(\mathbf{A}^{j}\) by
-\[\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}\]. If
+review them below.
+
+\citet{posse1995projection} proposed a stochastic direct search method,
+a random search algorithm. Given a current basis \(\mathbf{A}^{j-1}\), a
+candidate basis \(\mathbf{A}_{jl*}\) is sampled in the neighbourhood
+defined by the radius of the p-dimensional sphere, \(\alpha\) of the
+current basis \(\mathbf{A}^{j}\) by
+\(\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}\). If
 the candidate basis has a higher index value than the current basis, it
-is outputed as the target basis \(\mathbf{A}^{j + 1}\) along with other
+is outputed as the target basis \(\mathbf{A}^{j}\) along with other
 metadata and the current iteration stops. If no basis is found to have
@@
 been modified to maximise the projection pursuit index along the
-geodesic direction.
-
-\newpage
-\begin{algorithm}[H]
+geodesic direction. See Algorithm \ref{search-geodesic} for a
+summarisation of the steps.
+
+\begin{algorithm}
 \SetAlgoLined
-  initialisation: $l_{\max}$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+\SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+  initialisation\;
   \While{$l < l_{\max}$}{
-    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
-    Find $A^{J_1}_l$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
-    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
-    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
-    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
-    ...\;
-    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
-    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
-      \If{$p_{diff} > 0.001$}{
-        $A^{j+1} = A^{(J_2)}$\;
-        $I^{j+1} = I^{(J_2)}$\;
-        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
+    Generate ten bases in five random directions: $\mathbf{A}_{jl*}: \mathbf{A}_{j,(l + 9), *}$ within a small neighbourhood $\text{dist}$\;
+    Find the direction with the largest index value: $\mathbf{A}_{j, l_{d}, *}$\;
+    Construct the geodesic from $\mathbf{A}^{j-1}$ to $\mathbf{A}_{j, l_{d}, *}$\;
+    Find $\mathbf{A}_{j,l_g, *}$ on the geodesic that has the largest index value \;
+    Compute $I_{j,l_g, *} = f(\mathbf{A}_{j,l_g, *})$, $p_{\text{diff}} = (I_{j,l_g, *} - I^{j-1})/I_{j,l_g, *}$\;
+      \If{$p_{\text{diff}} > 0.001$}{
+        $A^{j} = A_{j,l_g, *}$, $I^{j} = I_{j,l_g, *}$\;
       }
-    $i = i + 1$\;
+    $l = l + 1$\;
   }
-  \caption{searchgeodesic: this part has problem!}
-  \label{search_geodesic}
+  \caption{search geodesic}
+  \label{search-geodesic}
 \end{algorithm}
@@
 basis is constructed as
-\(A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
-of the candidate basis is larger than the one of the current basis, the
-candidate basis becomes the target basis. If it is smaller, the
-candidate is accepted with probability
-\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
-where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
+\(\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}\).
+If the index value of the candidate basis is larger than the one of the
+current basis, the candidate basis becomes the target basis. If it is
+smaller, the candidate is accepted with probability
+\(P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}\)
+where \(I^{j-1}\) and \(I_{jl*}\) are the index value of the current and
 candidate basis respectively. The algorithm can be summarised as in
@@
 
-\newpage
-\begin{algorithm}[H]
+\begin{algorithm}
 \SetAlgoLined
-  initialisation: $T_0$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
+\SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{output}
+    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+  initialisation\;
   \While{$l < l_{\max}$}{
-    $\mathbf{A}^{j+1}_{l} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
-    $I^{j+1}_l = f(\mathbf{A}^{j+1}_{l})$\;
-    $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I^{j+1}_l > I^{j}$}{
-        $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
-        $I^{j+1} = I^{j+1}_l$\;
-        \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$,$l$ and relevant metadata}
+    Generate $\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{jl*}$ is orthonormal \;
+    Commpute $I_{jl*} = f(\mathbf{A}_{jl*})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{jl*} > I^{j-1}$}{
+        $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
       }{
-        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$\;
-        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
-        \If{$\text{prob} > \text{rand}$}{
-          \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$, $l$ and relevant metadata}
+        Compute $P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}$\;
+        Draw $D$ from a uniform distribution: $D \sim \text{Unif(0, 1)}$\;
+        \If{$P > D$}{
+          $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
         }
@@
 optimisation algorithms. Given an optimisation algorithm, two primary
-variables of interest are the \emph{decision variable} and the
-\emph{value of the objective function}. To align with the notation used
-previously for projection pursuit, we denote the decision variable to be
-\(\mathbf{A}\), which are matrices in projection pursuit algorithms and
-value of the objective function to be \(I\), the index function.
+variables of interest are the \emph{decision variable: \(\mathbf{A}\)}
+and the \emph{value of the objective function: \(I\)}.
 
 \emph{Iterators} indexes the data collected and has a time series
-feature that prescribes the its natural ordering. Given the complexity
-of the algorithm, there could be multiple iterators. The simpliest
-iterator: \emph{index per observation}, is a unique identifier for each
-observation in the data. For each level of iteration, we design two
-indices: \emph{index per iteration} is fixed for each observations in
-one iteration and has an increment of one once a new iteration starts.
-\emph{index within iteration} has an increment of one for each
-observation in an iteartion and starts over from one once a new
-iteration starts. A sketch of the difference between these three
-iterators is provided in Figure \ref{iterators}. Depend on the number of
-iteration level (\(L\)), the total number of iterators will differ and
-can be calculated as \(1 + 2L\).
+feature that prescribes the its natural ordering of the decision varible
+in the searching. The simpliest iterator, \emph{index per observation:
+\(t\)}, is a unique identifier for each observation in the data. For
+each level of iteration, we design two types of indices: \emph{index per
+iteration} is fixed for each observations in one iteration and has an
+increment of one once a new iteration starts. \emph{index within
+iteration} has an increment of one for each observation in an iteartion
+and starts over from one once a new iteration starts. A sketch of the
+difference between these three iterators is provided in Figure
+\ref{iterators}.
 
@@
 can't be classified as one of the three categories. They are defined
-under the fourth category: \emph{other parameter of interest: \(O\)}. In
+under the fourth category: \emph{other parameter of interest: \(S\)}. In
 projection pursuit, three other parameters of our interest includes
@@
 
-\hypertarget{notation}{%
-\subsection{Notation}\label{notation}}
-
-Now we formulate the idea of global object using mathmetical notation.
-Let \[\mathbf{A}_{j,l,k}\] denotes the projection matrix in iteration
-\(j\). If it is searching basis produced in the optimisation, then it
-has a subscript \(l\). If it is an interpolation basis, it has a \(k\)
-subscript. For example \(\mathbf{A}_{1,2,/}\) means the second searching
-basis in the first iteration, \(\mathbf{A}_{1, /, 2}\) means the second
-interpolation basis that on the tour path from the current basis
-\(\mathbf{A}^0\) to the target basis found in iteration 1:
-\(\mathbf{A}^1\). The target basis found in each iteration,
-\(\mathbf{A}_{j, l_j, /}\) can also be simplified using a superscript as
-\(\mathbf{A}^j\).
-
-Let \(\mathbf{A}^0\) denotes the starting basis. Superscript \(j\) is
-used to denote the number of iteration (index per iteration). Candidate
-basis are denoted using subscript on the right and interpolation basis
-are denoted using subscript on the left. For example,
-\(\mathbf{A}^1_{2}\) means the second candidate basis in the first
-iteration and \({}_{2}\mathbf{A}^1\) means the second interpolation
-basis that forms the tour path from the current basis \(\mathbf{A}^0\)
-to the target basis \(\mathbf{A}^1\). In iteration two, \(\mathbf{A}^1\)
-becomes the current basis and projection pursuit searches through
-\(\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}\) to
-find the target basis \(\mathbf{A}^2\). The tour path from
-\(\mathbf{A}^1\) to \(\mathbf{A}^2\) is constructed using the
-interpolation bases
-\({}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2\).
-All the bases together can be arranged in a format that each coloumn
-represents an iteration as in Equation \ref{eq:matrixA}.
+To write down the notation for the data structure, it is easier to write
+the bases as a column vector denoting by a single subscript by their
+natural ordering:
+
+\begin{equation}
+\{
+\mathbf{A}_0, 
+\mathbf{A}_1, 
+\mathbf{A}_2, 
+\cdots,
+\mathbf{A}_T
+\label{vector-basis}
+\}
+\end{equation}
+
+where \(T\) is the total number of plane serached by the optimisation
+algorithm including all the target planes, all the interpolating planes
+and all the candidate planes but NOT the starting plane. The full data
+structure can thus be shown as in Equation \ref{eq:data-structure}.
 
@@
 \left[
-\begin{array}{ccccccc}
-\ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +1 ,\ast} \\
-\ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +2, \ast} \\
-\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
-\ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{J+1, \sum_{i = 1}^{J-1}l_i + l, \ast}  & \mathbf{A}_{J + 1, \sum_{i = 1}^{J}l_i + l, \ast}\\
-\vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\
-\ast & \mathbf{A}_{1, l_1-1, \ast} & \mathbf{A}_{2, l_1 + l_2-1, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i -1, \ast} & \mathbf{A}_{J+1, \sum_{i = 1}^{J+1} l_i-1, \ast}\\
-\mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J}l_i , \ast} & \ast\\
+\begin{array}{c|cc|ccc|ccc}
+t & \mathbf{A} & I & j &  l  & k & S_{1}& \cdots & S_{p}\\
 \hline
-\ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J-1}k_i + 1}  & \ast\\
-\ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +2}  & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +k}  & \ast \\
-\vdots & \vdots & \vdots & \ddots & \vdots   & \vdots \\
-\ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J}k_i} & \ast \\
-\end{array}
-\right]
-= 
-\left[
-\begin{array}{c}
-\vphantom{A^J_1}\text{candidate basis} \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{A^J_2}\text{candidate basis} \\
-\vphantom{A^J}\text{current/target basis} \\
+1 & \mathbf{A}_0 & I_0 & 0 & \ast & \ast & \ast & \cdots & \ast\\
 \hline
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\text{NULL}}\vdots \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\vphantom{\vdots} \vdots \\
-\vphantom{\text{NULL}}\text{interpolation basis} \\
-\end{array}
-\right]
-\label{eq:matrixA}
-\end{equation}
-
-When aggregated with information of index value and other interesting
-parameter, it is easier to write the basis as a column vector denoting
-by a single subscript by their natural ordering i.e.~all the bases in
-iteration 1 along with the starting basis as shown in Equation
-\ref{eq:vectorA}, can be denoted as
-\(\mathbf{A}_0, \mathbf{A}_1, \cdots, \mathbf{A}_n\) where
-\(n = 1 + l_1 + k_1\). The full data structure can thus be presneted as
-in Equation \ref{eq:data-structure}.
-
-\begin{equation}
-\begin{array}{cccccccccccc}
-\bigl[
-\mathbf{A}_{0, \ast, \ast} & 
-\mathbf{A}_{1, 1, \ast} & 
-\mathbf{A}_{1,2,\ast} & 
-\cdots & 
-\mathbf{A}_{1,l_1, \ast}   &
-\mathbf{A}_{1, \ast, 1} & 
-\mathbf{A}_{1, \ast, 2} & 
-\cdots &
-\mathbf{A}_{1, \ast, k_1} &
-\bigr]
-\end{array} 
-\label{eq:vectorA}
-\end{equation}
-
-\begin{equation}
-\left[
-\begin{array}{c|c|cccc|ccc}
-\mathbf{A} & I & j &  l  & k & n & S_{1}& \cdots & S_{p}\\
+2 & \mathbf{A}_1 & I_1 & 1 & 1 & \ast & S_{1, 1} & \cdots & S_{1, p}\\
+3 & \mathbf{A}_2 & I_2 & 1 & 2 & \ast & S_{2, 1} & \cdots & S_{2, p}\\
+\vdots & \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
+\vdots & \vdots & \vdots & 1 & l_1 & \ast & \vdots &  \cdots & \vdots\\
 \hline
-\mathbf{A}_0 & I_0 & 0 & \ast & \ast & 1 & \ast & \cdots & \ast\\
+\vdots &\vdots & \vdots & 1 & \ast & 1& \vdots &  \cdots & \vdots\\
+\vdots &\vdots &\vdots& 1 & \ast & 2& \vdots& \cdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+\vdots &\vdots &\vdots & 1 & \ast & k_1&\vdots & \cdots & \vdots\\
 \hline
-\mathbf{A}_1 & I_1 & 1 & 1 & \ast& 2 & S_{1, 1} & \cdots & S_{1, p}\\
-\mathbf{A}_2 & I_2 & 1 & 2 &\ast & 3 & S_{2, 1} & \cdots & S_{2, p}\\
-\vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots\\
-\vdots & \vdots & 1 & l_1 & \ast & \vdots & \vdots & \cdots & \vdots\\
-\hline
-\vdots & \vdots & 1 & \ast & 1& \vdots &\vdots & \cdots & \vdots\\
-\vdots &\vdots& 1 & \ast & 2& \vdots&\vdots & \cdots & \vdots\\
-\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
-\vdots &\vdots & 1 & \ast & k_1&\vdots &\vdots & \cdots & \vdots\\
-\hline
-\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\vdots &\cdots &\vdots \\
-\mathbf{A}_n & I_n & J &  l_J & \ast & n & S_{n, 1}& \cdots & S_{n, p}\\
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+T+1 & \mathbf{A}_T & I_T & J &  l_J & \ast & S_{T, 1}& \cdots & S_{T, p}\\
 \end{array}
@@
 \end{equation}
-
-where
-\(n = 1 + \sum_{i = 1}^{J+1} l_i -1 + \sum_{i = 1}^{J} k_i= \sum_{i = 1}^{J+1} l_i + \sum_{i = 1}^{J} k_i\)
 
@@
 
-\[x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1) \cdots,  \sum_{i = 1}^J k_i + M\}\]
+\[x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}\]
 \[y = \{I_{jlk}: k = x\}\]

### 2020-07-20 paper/paper.Rmd
tweaks on the definition of global object, mainly notation

@@
 If we are being pedantic, all the target bases, except the starting basis, should have a searching index since it is the last candidate basis that successfully allows the optimisation algorithm to end and output a new basis for interpolation.
-If we denotes the length of each search (number of basis searched in each iteration) as $l_1, l_2, \cdots, l_J$ and the length of each interpolation is $k_1, k_2, \cdots, k_J$ , then the searching index for target basis in each iteration will become $l_1, l_1 + l_2, \cdots, l_1 +l_2+\cdots l_J$.
-To avoid this complex notation involving opteration in the subscript, we simplify it using a superscript as $$\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}$$ where $\mathbf{A}^{0}$ is the starting basis.
+If we denotes the length of each search (number of basis searched in each iteration) as $l_1, l_2, \cdots, l_J$ and the length of each interpolation is $k_1, k_2, \cdots, k_J$ , then the searching index for target basis in each iteration will become $\mathbf{A}_{1l_1*}, \mathbf{A}_{2l_2*}, \cdots, \mathbf{A}_{Jl_J*}$.
+We simplify this notation for target bases by using a superscript as $$\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}$$ where $\mathbf{A}^{0}$ is the starting basis.
 Notice that in the case when the complex notation can't be avoided, we shall separate each index using a comma and  write the index in brackets if necessary.
@@
 
-## Origin idea
+## Motivation
 <!-- Visual diagnostics can be real-time or post-run.
@@
 Given an optimisation algorithm, two primary variables of interest are the *decision variable: $\mathbf{A}$* and the *value of the objective function: $I$*.
+To further simplify the notation for $\mathbf{A}$, we write the bases as a column vector denoting by a single subscript by their natural ordering as in Equation \ref{vector-basis}.
+
+\begin{equation}
+\{
+\mathbf{A}_0, 
+\mathbf{A}_1, 
+\mathbf{A}_2, 
+\cdots,
+\mathbf{A}_T
+\label{vector-basis}
+\}
+\end{equation}
+
+where $T$ is the total number of plane serached by the optimisation algorithm including all the target planes, all the interpolating planes and all the candidate planes but NOT the starting plane.
 
@@
 The simpliest iterator, *index per observation: $t$*, is a unique identifier for each observation in the data.
-For each level of iteration, we design two types of indices: *index per iteration* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
-*index within iteration* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
+For each level of iteration, we design two types of indices: *index per iteration: $j$* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
+*index within iteration: $l$* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
 A sketch of the difference between these three iterators is provided in Figure \ref{iterators}.
+In projection pursuit optimisation algorithms, there is one level of iteartion and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
+We give the interpolating basis a different index $k$ for projection pursuit guided tour.
+It is similar in nature to $l$ but is specific for interpolating bases, which is usually not part of the optimisation.
 
@@
 ```
-
-In projection pursuit optimisation algorithms, there is one level of iteartion and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
 
@@
 A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
+The full data structure can thus be shown as in Equation \ref{eq:data-structure}.
+
+\begin{equation}
+\left[
+\begin{array}{c|cc|ccc|ccc}
+t & \mathbf{A} & I & j &  l  & k & S_{1}& \cdots & S_{p}\\
+\hline
+1 & \mathbf{A}_0 & I_0 & 0 & \ast & \ast & \ast & \cdots & \ast\\
+\hline
+2 & \mathbf{A}_1 & I_1 & 1 & 1 & \ast & S_{1, 1} & \cdots & S_{1, p}\\
+3 & \mathbf{A}_2 & I_2 & 1 & 2 & \ast & S_{2, 1} & \cdots & S_{2, p}\\
+\vdots & \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
+\vdots & \vdots & \vdots & 1 & l_1 & \ast & \vdots &  \cdots & \vdots\\
+\hline
+\vdots &\vdots & \vdots & 1 & \ast & 1& \vdots &  \cdots & \vdots\\
+\vdots &\vdots &\vdots& 1 & \ast & 2& \vdots& \cdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+\vdots &\vdots &\vdots & 1 & \ast & k_1&\vdots & \cdots & \vdots\\
+\hline
+\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+T+1 & \mathbf{A}_T & I_T & J &  l_J & \ast & S_{T, 1}& \cdots & S_{T, p}\\
+\end{array}
+\right]
+\label{eq:data-structure}
+\end{equation}
 
@@
 <!-- \end{equation} -->
-
-To write down the notation for the data structure, it is easier to write the bases as a column vector denoting by a single subscript by their natural ordering:
-
-\begin{equation}
-\{
-\mathbf{A}_0, 
-\mathbf{A}_1, 
-\mathbf{A}_2, 
-\cdots,
-\mathbf{A}_T
-\label{vector-basis}
-\}
-\end{equation}
-
-where $T$ is the total number of plane serached by the optimisation algorithm including all the target planes, all the interpolating planes and all the candidate planes but NOT the starting plane.
-The full data structure can thus be shown as in Equation \ref{eq:data-structure}.
-
-\begin{equation}
-\left[
-\begin{array}{c|cc|ccc|ccc}
-t & \mathbf{A} & I & j &  l  & k & S_{1}& \cdots & S_{p}\\
-\hline
-1 & \mathbf{A}_0 & I_0 & 0 & \ast & \ast & \ast & \cdots & \ast\\
-\hline
-2 & \mathbf{A}_1 & I_1 & 1 & 1 & \ast & S_{1, 1} & \cdots & S_{1, p}\\
-3 & \mathbf{A}_2 & I_2 & 1 & 2 & \ast & S_{2, 1} & \cdots & S_{2, p}\\
-\vdots & \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
-\vdots & \vdots & \vdots & 1 & l_1 & \ast & \vdots &  \cdots & \vdots\\
-\hline
-\vdots &\vdots & \vdots & 1 & \ast & 1& \vdots &  \cdots & \vdots\\
-\vdots &\vdots &\vdots& 1 & \ast & 2& \vdots& \cdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-\vdots &\vdots &\vdots & 1 & \ast & k_1&\vdots & \cdots & \vdots\\
-\hline
-\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-T+1 & \mathbf{A}_T & I_T & J &  l_J & \ast & S_{T, 1}& \cdots & S_{T, p}\\
-\end{array}
-\right]
-\label{eq:data-structure}
-\end{equation}
 

### 2020-07-20 paper/paper.tex
tweaks on the definition of global object, mainly notation

@@
 each iteration will become
-\(l_1, l_1 + l_2, \cdots, l_1 +l_2+\cdots l_J\). To avoid this complex
-notation involving opteration in the subscript, we simplify it using a
-superscript as
+\(\mathbf{A}_{1l_1*}, \mathbf{A}_{2l_2*}, \cdots, \mathbf{A}_{Jl_J*}\).
+We simplify this notation for target bases by using a superscript as
 \[\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}\]
@@
 
-\hypertarget{origin-idea}{%
-\subsection{Origin idea}\label{origin-idea}}
+\hypertarget{motivation}{%
+\subsection{Motivation}\label{motivation}}
 
@@
 variables of interest are the \emph{decision variable: \(\mathbf{A}\)}
-and the \emph{value of the objective function: \(I\)}.
+and the \emph{value of the objective function: \(I\)}. To further
+simplify the notation for \(\mathbf{A}\), we write the bases as a column
+vector denoting by a single subscript by their natural ordering as in
+Equation \ref{vector-basis}.
+
+\begin{equation}
+\{
+\mathbf{A}_0, 
+\mathbf{A}_1, 
+\mathbf{A}_2, 
+\cdots,
+\mathbf{A}_T
+\label{vector-basis}
+\}
+\end{equation}
+
+where \(T\) is the total number of plane serached by the optimisation
+algorithm including all the target planes, all the interpolating planes
+and all the candidate planes but NOT the starting plane.
 
@@
 each level of iteration, we design two types of indices: \emph{index per
-iteration} is fixed for each observations in one iteration and has an
-increment of one once a new iteration starts. \emph{index within
-iteration} has an increment of one for each observation in an iteartion
-and starts over from one once a new iteration starts. A sketch of the
-difference between these three iterators is provided in Figure
-\ref{iterators}.
+iteration: \(j\)} is fixed for each observations in one iteration and
+has an increment of one once a new iteration starts. \emph{index within
+iteration: \(l\)} has an increment of one for each observation in an
+iteartion and starts over from one once a new iteration starts. A sketch
+of the difference between these three iterators is provided in Figure
+\ref{iterators}. In projection pursuit optimisation algorithms, there is
+one level of iteartion and hence exists three iterators: \texttt{id}
+indices each observation by a unique number; \texttt{tries} is the index
+per iteration iterator that gets updated once a search-and-interpolate
+step is finished; and \texttt{loop} is the index within iteration
+iterator and starts over from one at the beginning of a new
+\texttt{tries}. We give the interpolating basis a different index \(k\)
+for projection pursuit guided tour. It is similar in nature to \(l\) but
+is specific for interpolating bases, which is usually not part of the
+optimisation.
 
@@
 \end{figure}
-
-In projection pursuit optimisation algorithms, there is one level of
-iteartion and hence exists three iterators: \texttt{id} indices each
-observation by a unique number; \texttt{tries} is the index per
-iteration iterator that gets updated once a search-and-interpolate step
-is finished; and \texttt{loop} is the index within iteration iterator
-and starts over from one at the beginning of a new \texttt{tries}.
 
@@
 for projection pursuit guided tour is presented in Figure
-\ref{fig:glb-obj}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
-\end{figure}
-
-To write down the notation for the data structure, it is easier to write
-the bases as a column vector denoting by a single subscript by their
-natural ordering:
-
-\begin{equation}
-\{
-\mathbf{A}_0, 
-\mathbf{A}_1, 
-\mathbf{A}_2, 
-\cdots,
-\mathbf{A}_T
-\label{vector-basis}
-\}
-\end{equation}
-
-where \(T\) is the total number of plane serached by the optimisation
-algorithm including all the target planes, all the interpolating planes
-and all the candidate planes but NOT the starting plane. The full data
-structure can thus be shown as in Equation \ref{eq:data-structure}.
+\ref{fig:glb-obj}. The full data structure can thus be shown as in
+Equation \ref{eq:data-structure}.
 
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
+\end{figure}
+
 \hypertarget{simulated-data}{%

### 2020-07-20 paper/paper.Rmd
add math for data simulation and reference to tidy data

@@
 
-[I would add the maths here to describe the simulations. ppp]
+We first simulate some random variables of size 1000 with different structures.
+`x1`, `x8`, `x9` and `x10` are simulated from normal distribution iwth zero mean and variance of one as in equation \ref{eq:sim-norm}.
+When using projection pursuit to explore the data structure based on its departure from normality, the entry in the projection basis for these variables should be close to zero in theory.
+`x2` to `x7` are mixture of normal distribution with different weights and locations.
+Equation \ref{eq:sim-x2} to \ref{eq:sim-x7} show the distribution of each variables and Figure \ref{origin-data} shows the histogram of each variable except `x3`.
+All the variables are then scaled to ensure the mixture has variance of one.
 
-Two set of simulated data are used in the demo of the visual diagnostics in projection pursuit guided tour.
-A small dataset consists of 1000 randomly simulated observations of five variables (`x1`, `x2`, `x8`, `x9`, `x10`).
-`x2` is the informative variable simulated from two bi-modal normal distribution centred at -3 and 3 with variance of one and the other four are simulated from non-informative standard random normal distributions.
-The data has been scaled to ensure `x2` has variance of 1.
-The goal is to find the 1D projection of bi-modal shape, which corresponds to the projection matrix (vector) of `[0, 1, 0, 0, 0]`.
-
-A larger dataset contains more informative variables (`x3` to `x7`) of different types.
-The distribution of all the variables except `x3` is plotted in Figure \ref{origin-data} and `x3` takes 500 positive one and 500 negative one.
-[should I introduce the dist for each var?
-xxx] [also feel like we don't really use `x3` to `x6`, should we not mention about these?
-xxx]
+\begin{align}
+x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \label{eq:sim-norm} \\
+x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\label{eq:sim-x2}\\
+\Pr(x_3) &= 
+\begin{cases}
+0.5 & \text{if $x_3 = -1$ or $1$}\\
+0 & \text{otherwise}
+\end{cases}\label{eq:sim-x3}\\
+x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1) \label{eq:sim-x4}\\
+x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\label{eq:sim-x5}\\
+x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\label{eq:sim-x6}\\
+x_7 &\sim 0.5 \mathcal{N}(-5, 1) + 0.5 \mathcal{N}(5, 1) 
+\label{eq:sim-x7}
+\end{align}
 
@@
 
-In tour, when the optimisation ends the global object will be stored and printed and can be turned off via `print = FALSE`.
-Additional messages during the optimisation can be displayed via `verbose = TRUE`.
-Below presented the global object of the 1D projection using the small dataset.
-[the printing is not ideal as it doesn't show all the columns.
-xxx]
+We form our first dataset using variable `x1`, `x2`, `x8`, `x9` and `x10` and supply it to one of the above algorithm, say `search_geodesic` in the `tourr` package.
+When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
+Additional messages during the optimisation can be displayed by `verbose = TRUE`.
+Below shows the first five rows of the global object.
+The global object meets the definition of tidy data introduced in [@wickham2014tidy] where each observation forms a row, each variable forms a column and each type of observational unit forms a table.
+The adoption of tidy data makes it easier for data wrangling and visualisation, which powers the visual diagnostics system introduced in this paper.
 
@@
 set.seed(123456)
-holes_1d_geo <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                           search_f =  search_geodesic),
+holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+                                           search_f =  search_better),
                   rescale = FALSE, verbose = TRUE)
@@
 
+\newpage
+
 ```{r glb-obj-1}
-holes_1d_geo %>% head(5)
+print(holes_1d_better %>% head(5))
 ```

### 2020-07-20 paper/paper.tex
add math for data simulation and reference to tidy data

@@
 
-{[}I would add the maths here to describe the simulations. ppp{]}
-
-Two set of simulated data are used in the demo of the visual diagnostics
-in projection pursuit guided tour. A small dataset consists of 1000
-randomly simulated observations of five variables (\texttt{x1},
-\texttt{x2}, \texttt{x8}, \texttt{x9}, \texttt{x10}). \texttt{x2} is the
-informative variable simulated from two bi-modal normal distribution
-centred at -3 and 3 with variance of one and the other four are
-simulated from non-informative standard random normal distributions. The
-data has been scaled to ensure \texttt{x2} has variance of 1. The goal
-is to find the 1D projection of bi-modal shape, which corresponds to the
-projection matrix (vector) of \texttt{{[}0,\ 1,\ 0,\ 0,\ 0{]}}.
-
-A larger dataset contains more informative variables (\texttt{x3} to
-\texttt{x7}) of different types. The distribution of all the variables
-except \texttt{x3} is plotted in Figure \ref{origin-data} and
-\texttt{x3} takes 500 positive one and 500 negative one. {[}should I
-introduce the dist for each var? xxx{]} {[}also feel like we don't
-really use \texttt{x3} to \texttt{x6}, should we not mention about
-these? xxx{]}
+We first simulate some random variables of size 1000 with different
+structures. \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are
+simulated from normal distribution iwth zero mean and variance of one as
+in equation \ref{eq:sim-norm}. When using projection pursuit to explore
+the data structure based on its departure from normality, the entry in
+the projection basis for these variables should be close to zero in
+theory. \texttt{x2} to \texttt{x7} are mixture of normal distribution
+with different weights and locations. Equation \ref{eq:sim-x2} to
+\ref{eq:sim-x7} show the distribution of each variables and Figure
+\ref{origin-data} shows the histogram of each variable except
+\texttt{x3}. All the variables are then scaled to ensure the mixture has
+variance of one.
+
+\begin{align}
+x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \label{eq:sim-norm} \\
+x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\label{eq:sim-x2}\\
+\Pr(x_3) &= 
+\begin{cases}
+0.5 & \text{if $x_3 = -1$ or $1$}\\
+0 & \text{otherwise}
+\end{cases}\label{eq:sim-x3}\\
+x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1) \label{eq:sim-x4}\\
+x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\label{eq:sim-x5}\\
+x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\label{eq:sim-x6}\\
+x_7 &\sim 0.5 \mathcal{N}(-5, 1) + 0.5 \mathcal{N}(5, 1) 
+\label{eq:sim-x7}
+\end{align}
 
@@
 
-In tour, when the optimisation ends the global object will be stored and
-printed and can be turned off via \texttt{print\ =\ FALSE}. Additional
-messages during the optimisation can be displayed via
-\texttt{verbose\ =\ TRUE}. Below presented the global object of the 1D
-projection using the small dataset. {[}the printing is not ideal as it
-doesn't show all the columns. xxx{]}
+We form our first dataset using variable \texttt{x1}, \texttt{x2},
+\texttt{x8}, \texttt{x9} and \texttt{x10} and supply it to one of the
+above algorithm, say \texttt{search\_geodesic} in the \texttt{tourr}
+package. When the optimisation ends, the global object will be stored
+and printed (it can be turned off by supplying argument
+\texttt{print\ =\ FALSE}). Additional messages during the optimisation
+can be displayed by \texttt{verbose\ =\ TRUE}. Below shows the first
+five rows of the global object. The global object meets the definition
+of tidy data introduced in \citep{wickham2014tidy} where each
+observation forms a row, each variable forms a column and each type of
+observational unit forms a table. The adoption of tidy data makes it
+easier for data wrangling and visualisation, which powers the visual
+diagnostics system introduced in this paper.
+
+\newpage
 
@@
 ##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>        <dbl> <int>
-## 1 <dbl[,1] [5 x ~     0.749     1 start             NA <NA>            0.5     1
-## 2 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       2
-## 3 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       3
-## 4 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       4
-## 5 <dbl[,1] [5 x ~     0.749     1 direction_sea~     1 search_geode~  NA       5
+## 1 <dbl[,1] [5 x 1]>     0.749     1 start            NA <NA>           0.5     1
+## 2 <dbl[,1] [5 x 1]>     0.730     1 random_search     1 search_bett~   0.5     2
+## 3 <dbl[,1] [5 x 1]>     0.743     1 random_search     2 search_bett~   0.5     3
+## 4 <dbl[,1] [5 x 1]>     0.736     1 random_search     3 search_bett~   0.5     4
+## 5 <dbl[,1] [5 x 1]>     0.747     1 random_search     4 search_bett~   0.5     5
 \end{verbatim}

### 2020-07-20 paper/paper.Rmd
new version of plot definition

@@
 
-Overall, the first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
-The second question can be addressed via visualising the rotating high dimensional space.
-This seems to be a hard task to see beyond 3D, but the fact is that we never really see all the six faces of a cube at one time.
-The way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
+The first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
+The second question can be addressed via visualising the rotating high dimensional space or its projection on the reduced 2D space.
+Thus animated visualisation is needed to preceive the optimisation path in the searching space.
+
+<!-- [This seems to be a hard task to see the rotation of high dimensional space, but the way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
 Thus we can do a similar rotation on the screen to precieve even higher dimension.
-[there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
-Thus animated visualisation is needed to preceive the optimisation path in the searching space.
-
-The plot design of visual diagnostics depends on the characteristics of the variables to plot.
+there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
+-->
+
+Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
+To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.
+
+<!-- The plot design of visual diagnostics depends on the characteristics of the variables to plot.
 If the searching points are of interest [point geom is not good; need summarise.
@@
 [stratefy may help?
-- may need more work here.]
+- may need more work here.] -->
 
@@
 At each iteration, rather than knowing the index value of *every* points, we are more interested to have a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the next geodesic interpolation and future searches.
-
-Using the notation we defined earlier, the plot can be defined as plotting
-
-$$x = \{1, 2, \cdots J + 1\}$$
-
-against the five point summary of each $j^*$ on the y-axis:
-
-$$y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$$
-
-where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$. $\mathds{1}(.)$ is the identity operator.
 
@@
 This helps to visualise the improvement made by each `tries`.
+With gramma of graphics [@wickham2010layered], the plot for exploring index value can be defined in three layers as follows:
+
+- Layer 1: boxplot geom
+  - data: group by `tries` and filter the observations in the group that have count greater than `cutoff = 15`.
+  - x: `tries` is mapped to the x-axis
+  - y: the index value after statistical transformation $y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$ is mapped to the y-axis where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$.
+$\mathds{1}(.)$ is the identity operator.
+
+- Layer 2: point geom
+  - data: group by `tries` and filter the observations in the group that have count less than `cutoff = 15`.
+  - x: `tries` is mapped to the x-axis
+  - y: `index_val` is mapped to the y-axis
+
+- Layer 3: line geom
+  - data: filter the points with the highest index value in each `tries`
+  - x: `tries` is mapped to the x-axis
+  - y: `index_val` is mapped to the y-axis
 
@@
 
-\newpage
-
-The number of interpolation points is much managable than the searching points and always shows a pattern of gradual increase in the index value.
-Hence a line plot with interpolating order on the x-axis will be desirable.
-Using notation, the x-axis is a sequence $$x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}$$ and y-axis are the index values defined by $$y = \{I_{jlk}: k = x\}$$
+Sometimes, it is more interesting to explore the index value on the interpolation path and the plot definition is slightly varied.
+
+- Layer 1: point geom
+  - data: filter the observations with `info` being `interpolation` and mutate `id` to be the row number of the subsetted tibble
+  - x: `id` is mapped to the x-axis
+  - y: `index_val` is mapped to the y-axis
+
+- Layer 2: line geom
+  - using line goemetry for the same data and aesthetics
 
@@
 
-Polishing points form on itself a global structure with candidate bases and output bases.
-If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$
-
-Then the index plot incorporating both the interpolating points and polishing points can be defined as
-
-$$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$
-$$y = \{I_{jlk}: k = x\}$$
+Slight variation of plot definition due to the addition of polishing points is as follows:
+
+- Layer 1: point geom
+  - data: filter the observations with `info` being `interpolation`; bind the global object from optimisation and interpolation and form polishing; mutate `id` to be the row number of the binded tibble.
+  - x: `id` is mapped to the x-axis
+  - y: `index_val` is mapped to the y-axis 
+  - color: `method` is mapped to the color aesthetic
+
+- Layer 2: line geom
+  - using line goemetry for the same data and aesthetics
+
+<!-- Polishing points form on itself a global structure with candidate bases and output bases.
+If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$ -->
+
+<!-- Then the index plot incorporating both the interpolating points and polishing points can be defined as  -->
+
+<!-- $$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$ -->
+<!-- $$y = \{I_{jlk}: k = x\}$$ -->
 
@@
 An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
-We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
+The visualisation can thus be defined as
+
+<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
 This gives a matrix of dimension $n$ by $(p*d)$.
@@
 Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
-Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix
-
-$$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$
-
-has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
-The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$
-
-The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
+Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix    -->
+
+<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
+
+<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
+The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
+
+<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
 THis means all the iterators and interesting other variables defined in the global object can be binded with the these two projections and can be mapped to other aesthetics (i.e.
 color).
-
-In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
+-->
+
+- Layer 1: point geom
+  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
+  - x: the projected basis on the first principal component
+  - y: the projected basis on the second principal component
+  - color: an variable of interest is mapped onto color
+
+<!-- In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
 [need to think further about how different type of variables can be shown].
+-->
+
 While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.

### 2020-07-20 paper/paper.tex
new version of plot definition

@@
 
-Overall, the first question can be answered using a static plot with
-x-axis showing the progression of the optimisation and y-axis showing
-the value of the objective function. The second question can be
-addressed via visualising the rotating high dimensional space. This
-seems to be a hard task to see beyond 3D, but the fact is that we never
-really see all the six faces of a cube at one time. The way humans
-understand the dimension of a 3D object is either through rotating the
-physical 3D object or using shade and line type to annotate the 3D
-object in a paper or electronically. Thus we can do a similar rotation
-on the screen to precieve even higher dimension. {[}there is likely to
-be a learning curve - remember when learning geometry in school, it
-takes a while to get used to see 3D cube on a paper{]}. Thus animated
-visualisation is needed to preceive the optimisation path in the
-searching space.
-
-The plot design of visual diagnostics depends on the characteristics of
-the variables to plot. If the searching points are of interest {[}point
-geom is not good; need summarise. While updating points are more
-manageable -\textgreater{} point geom{]}. However, one must realise that
-with hundreds or thousands of searching points, exploring the sample
-space in animation could be slow and this is because of the time it
-takes to render hundreds point. {[}stratefy may help? - may need more
-work here.{]}
+The first question can be answered using a static plot with x-axis
+showing the progression of the optimisation and y-axis showing the value
+of the objective function. The second question can be addressed via
+visualising the rotating high dimensional space or its projection on the
+reduced 2D space. Thus animated visualisation is needed to preceive the
+optimisation path in the searching space.
+
+Since the global object is already tidy, not further tidying steps is
+needed, while certain data wrangling steps
+\citep{wickham2016rfordatascience} are still needed to transform the
+global object into desirable format for one particular visualisation. To
+emphasize on this good practice of data analysis, we will describe the
+transformation steps needed for each diagnostic plots before stepping
+into visualisation.
 
@@
 and future searches.
-
-Using the notation we defined earlier, the plot can be defined as
-plotting
-
-\[x = \{1, 2, \cdots J + 1\}\]
-
-against the five point summary of each \(j^*\) on the y-axis:
-
-\[y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\]
-
-where \(q\) takes one of \(0, 25, 50, 75, 100\) and \(Q_q(x)\) finds the
-qth-quantile for the vector \(x\). \(\mathds{1}(.)\) is the identity
-operator.
 
@@
 This helps to visualise the improvement made by each \texttt{tries}.
+With gramma of graphics \citep{wickham2010layered}, the plot for
+exploring index value can be defined in three layers as follows:
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: boxplot geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: group by \texttt{tries} and filter the observations in the
+    group that have count greater than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \texttt{tries} is mapped to the x-axis
+  \item
+    y: the index value after statistical transformation
+    \(y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\) is
+    mapped to the y-axis where \(q\) takes one of \(0, 25, 50, 75, 100\)
+    and \(Q_q(x)\) finds the qth-quantile for the vector \(x\).
+    \(\mathds{1}(.)\) is the identity operator.
+  \end{itemize}
+\item
+  Layer 2: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: group by \texttt{tries} and filter the observations in the
+    group that have count less than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \texttt{tries} is mapped to the x-axis
+  \item
+    y: \texttt{index\_val} is mapped to the y-axis
+  \end{itemize}
+\item
+  Layer 3: line geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: filter the points with the highest index value in each
+    \texttt{tries}
+  \item
+    x: \texttt{tries} is mapped to the x-axis
+  \item
+    y: \texttt{index\_val} is mapped to the y-axis
+  \end{itemize}
+\end{itemize}
 
@@
 
-\newpage
-
-The number of interpolation points is much managable than the searching
-points and always shows a pattern of gradual increase in the index
-value. Hence a line plot with interpolating order on the x-axis will be
-desirable. Using notation, the x-axis is a sequence
-\[x = \{1, 2, \cdots,\sum_{i=1}^Jk_i\}\] and y-axis are the index values
-defined by \[y = \{I_{jlk}: k = x\}\]
+Sometimes, it is more interesting to explore the index value on the
+interpolation path and the plot definition is slightly varied.
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: filter the observations with \texttt{info} being
+    \texttt{interpolation} and mutate \texttt{id} to be the row number
+    of the subsetted tibble
+  \item
+    x: \texttt{id} is mapped to the x-axis
+  \item
+    y: \texttt{index\_val} is mapped to the y-axis
+  \end{itemize}
+\item
+  Layer 2: line geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    using line goemetry for the same data and aesthetics
+  \end{itemize}
+\end{itemize}
 
@@
 
-Polishing points form on itself a global structure with candidate bases
-and output bases. If we're allowed to abuse the notation a bit, we
-denote \[A_{J, \ast, \sum_{i = 1}^J k_i + 1}\] as the first polishing
-basis and \[A_{J, \ast, \sum_{i = 1}^J k_i + M}\] as the last polishing
-basis, where
-\[M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})\]
-
-Then the index plot incorporating both the interpolating points and
-polishing points can be defined as
-
-\[x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}\]
-\[y = \{I_{jlk}: k = x\}\]
+Slight variation of plot definition due to the addition of polishing
+points is as follows:
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: filter the observations with \texttt{info} being
+    \texttt{interpolation}; bind the global object from optimisation and
+    interpolation and form polishing; mutate \texttt{id} to be the row
+    number of the binded tibble.
+  \item
+    x: \texttt{id} is mapped to the x-axis
+  \item
+    y: \texttt{index\_val} is mapped to the y-axis
+  \item
+    color: \texttt{method} is mapped to the color aesthetic
+  \end{itemize}
+\item
+  Layer 2: line geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    using line goemetry for the same data and aesthetics
+  \end{itemize}
+\end{itemize}
 
@@
 a 2D plot. An option to explore the searching space of these bases is to
-explore a reduced space via principal component analysis (PCA). We first
-flat each projection bases \(\mathbf{A}\) to a row vector and stack them
-by rows. This gives a matrix of dimension \(n\) by \((p*d)\). Centering
-and scaling are usually done before performing PCA and we denotes the
-pre-processed matrix as \(\tilde{\mathbf{A}}_{n \times (p \times d)}\).
-By definition, the principal components maximise the variance of the
-projection bases in the reduced space. Using lagrange multipler, this is
-equivalent to find the eigenvector of the variance-covariance matrix:
-\(\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}\). Decomposing
-\(\tilde{\mathbf{A}}\) using singular value decomposition (SVD) gives
-\[\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}\] and
-hence the variance-covariance matrix
-
-\[\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}\]
-
-has eigenvector
-\(\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]\).
-The projection mapping to the x- and y-axis uses the first two principal
-components and hence
-\[\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2\]
-
-The resulting two vectors are of dimension \(n \times 1\), which is
-compatible with the dimension of the global object. THis means all the
-iterators and interesting other variables defined in the global object
-can be binded with the these two projections and can be mapped to other
-aesthetics (i.e.~color).
-
-In this way, the first two principle components will take up the x and y
-axis and all the other information will be mapped using other aesthetic
-attribution. {[}need to think further about how different type of
-variables can be shown{]}. While explore the reduced space is an initial
-attempt to understand the searching space, there are existing technology
-for rotating a higher dimensional space for visualisation. Geozoo is an
-option. It generates random points on the high dimensional space and we
-can overlay it with the points on the optimisation path to visualise the
-spread of it on the high-D sphere.
+explore a reduced space via principal component analysis (PCA). The
+visualisation can thus be defined as
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: subset the basis of interest and arrange into a matrix format;
+    perform PCA on the basis matrix and compute the projected basis on
+    the first two principal components; bind the variables from the
+    original global object and form a tibble
+  \item
+    x: the projected basis on the first principal component
+  \item
+    y: the projected basis on the second principal component
+  \item
+    color: an variable of interest is mapped onto color
+  \end{itemize}
+\end{itemize}
+
+While explore the reduced space is an initial attempt to understand the
+searching space, there are existing technology for rotating a higher
+dimensional space for visualisation. Geozoo is an option. It generates
+random points on the high dimensional space and we can overlay it with
+the points on the optimisation path to visualise the spread of it on the
+high-D sphere.
 

### 2020-07-21 paper/paper.Rmd
some re-writing of exploring index value and update plots

@@
 
-## Visual diagnostics
+# Visual diagnostics
 
@@
 
-### Explore the value of objective function {#static}
-
-It is a convention to map the time index of a time series on the x-axis and value to the y-axis.
-In our diagnostic plots, it is to map the iterator on the x-axis and index value to the y-axis.
-The next question is to ask, given four indices being defined in the global object, which one should be used?
-*It depends on the *.
-
-As the searching goes towards the end, it is easy to have hundreds of candidate bases in each iteration with very close index value.
-When plotting, these iteration will occupy the vast majority of the plot space.
-This motivates to summarise the points in each iteration (`tries`).
-At each iteration, rather than knowing the index value of *every* points, we are more interested to have a general summary of all the points and more importantly, the point with the largest `index_val` since it prescribes the next geodesic interpolation and future searches.
+## Explore the value of objective function {#static}
+
+### Searching points
+
+A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimal progressively.
+We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
+In the plot, points from those iterations towards the end will occupy the vast majority of the plot space.
+This motivates to use summarisation.
+Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).
 
@@
 We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
-Thus, the number of point for each `tries` is displayed at the bottom of each box and we provide options to switch `tries` with small number of points to a point geometry, which is achieved via the `cutoff` argument.
-A line geometry is also added to link the points with the largest index value in each `tries`.
-This helps to visualise the improvement made by each `tries`.
-With gramma of graphics [@wickham2010layered], the plot for exploring index value can be defined in three layers as follows:
+Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via an `cutoff` argument.
+A line geometry is also added to link the points with the largest index value in each iteration.
+This helps to visualise the improvement made in each iteration.
+Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:
 
@@
 
-The larger dataset is used for 2D projection with `search_better` and `max.tries = 500`.
-In Figure \ref{points}, the searching points with `id` and `tries` on the x-axis, colored by `tries`.
-Label at the bottom indicates the number of observations in each tries and facilitates the choice of cutoff to switch from point geometry to boxplot geometry (`cutoff = 15`).
-The line geometry suggests the largest improvement happens at `tries = 5`.
+We choose variable `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
+Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
+The index value of the searching points are shown in Figure \ref{points}.
+Label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
+We learn that the `search_better` quickly finds better projection basis with higher index value at first and the takes longer to find a better one later.
 
@@
 
-Sometimes, it is more interesting to explore the index value on the interpolation path and the plot definition is slightly varied.
+### Interpolating points
+
+Sometimes, rather than explore the searching points, we may be interested in explore the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
+Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed.
+The slightly modified plot definition is shown below:
 
@@
 
-This examples uses `search_better` for a 2D projection on the larger dataset using the `holes` index.
-In the interpolation stage, continuous frames will be constructed to bridge the project on the current basis to the target basis.
-The target basis will become the current basis in the next iteration and the searching stage continues to find the next target basis.
-In the left panel of Figure \ref{interruption}, we observe that  there are interpolating bases with higher index values than the target basis and these bases could be used to search for new basis in the next searching stage.
-
-An interruption is then constructed to accept the interpolating bases up to the one with the largest index value and use that basis as the current basis for the next searching stage.
-After implementing this interruption, the tracing plot with the same configuration is shown on the right panel of Figure \ref{interruption}.
-In the third `tries`, rather than interpolating fully to the target basis at `id = 62`, it stops before the index value starts to decrease at `id = 60`.
-This implementation results in a higher index value in the end with fewer steps.
+We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for `holes` index.
+From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with higher index value than the target basis on the interpolation path.
+In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
+Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
+After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{interruption}.
 
@@
 
-#### Another example: Polish
+### Polishing points
 
@@
 
-In principle, all the optimisation routines should result in the same output for the same problem while this may not be true in real application.
-This motivates the creation of a polishing search that polishes the final basis and achieves unity on different methods.
-
-`search_polish` takes the final basis of a given search as the current basis and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood, whose radius is controlled by `polish_alpha`.
-Among the `n_sample` basis, the one with the largest index value becomes the candidate basis.
-If its index value of the candidate basis is larger than that of the current basis, it becomes the current basis in the next iteration.
-If no basis is found to have larger index value than the current basis, the searching neighbourhood will be shrunk and the search continues.
+In principle, all the optimisation routines should result in the same output for the same problem while this may not be the case in real application.
+This motivates the creation of a polishing search that polishes the final basis found and achieves unity across different methods.
+
+`search_polish` takes the final basis of a given search as a start and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood.
+Among those sampled basis, the one with the largest index value is chosen to be compared with the current basis.
+If its index value is larger than that of the current basis, it becomes the current basis in the next iteration.
+If no basis is found to have larger index value, the searching neighbourhood will shrink and the search continues.
 The polishing search ends when one of the four stopping criteria is satisfied:
 
-1) the two basis can't be too close 
-2) the percentage improvement of the index function can't be too small 
+1) the chosen basis can't be too close to the current basis 
+2) the percentage improvement of the index value can't be too small 
 3) the searching neighbourhood can't be too small
@@
 
-[should the stopping criteria be more detailed? xxx]
-
 The usage of search_polish is as follows.
-After the first tour, the final basis from the interpolation is extracted and supplied into a new tour with the `start` argument and `search_polish` as the searching function in the guided_tour.
+After the first search, the final basis from the interpolation is extracted and supplied to the second search as the `start` argument.
+`search_polish` is used as the search function.
 All the other arguments should remain the same.
@@
 
-The following example conducted a 2D projection on the larger dataset using search better with different configurations.
-`max.tries` is a hyperparameter that controls the maximum number of try without improvement and its default value is 25.
-As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while a small `max.tries` of 25 is not sufficient for `search_better` to find the global maximum.
-The `max.tries` argument needs to be adjusted to ensure it is sufficient to explore the parameter space.
+#### Another example: Polish
+
+Again using the same data, we are interested to compare the effect of different `max.tries` in the 2D projection setting.
+`max.tries` is a hyperparameter that controls the maximum number of try before the search ends.
+The default value of 25 is suitable for 1D projection while we suspect it may not be a good option for the 2D case and hence want to compare it with an alternative, 500.
+As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to succiciently explore the parameter space.
 
@@
 
-### Explore searching space {#animated}
+## Explore searching space {#animated}
 
@@
 
-While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
-Geozoo is an option.
-It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
-
-#### A toy example: understand different stage of search_geodesic
+### A toy example: understand different stage of search_geodesic
 
@@
 
-#### A more complex example: Choosing the initial value for polishing parameter
+### A more complex example: Choosing the initial value for polishing parameter
 
@@
 
+While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
+Geozoo is an option.
+It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
+
+[add example from geozoo]
+
 <!-- # Animated diagnostic plots -->
@@
 
-### A comprehensive example of diagnosing a noisy index function
+## A comprehensive example of diagnosing a noisy index function
 
@@
 
-#### Explore index value
+### Explore index value
 
@@
 
-#### Explore searching space
+### Explore searching space
 

### 2020-07-21 paper/paper.tex
some re-writing of exploring index value and update plots

@@
 \hypertarget{visual-diagnostics}{%
-\subsection{Visual diagnostics}\label{visual-diagnostics}}
+\section{Visual diagnostics}\label{visual-diagnostics}}
 
@@
 \hypertarget{static}{%
-\subsubsection{Explore the value of objective function}\label{static}}
-
-It is a convention to map the time index of a time series on the x-axis
-and value to the y-axis. In our diagnostic plots, it is to map the
-iterator on the x-axis and index value to the y-axis. The next question
-is to ask, given four indices being defined in the global object, which
-one should be used? \emph{It depends on the }.
-
-As the searching goes towards the end, it is easy to have hundreds of
-candidate bases in each iteration with very close index value. When
-plotting, these iteration will occupy the vast majority of the plot
-space. This motivates to summarise the points in each iteration
-(\texttt{tries}). At each iteration, rather than knowing the index value
-of \emph{every} points, we are more interested to have a general summary
-of all the points and more importantly, the point with the largest
-\texttt{index\_val} since it prescribes the next geodesic interpolation
-and future searches.
+\subsection{Explore the value of objective function}\label{static}}
+
+\hypertarget{searching-points}{%
+\subsubsection{Searching points}\label{searching-points}}
+
+A primary interest of diagnosing an optimisation algorithm is to study
+how it finds its optimal progressively. We could plot the index value
+across its natural ordering, however, different iterations may have
+different number of points and, towards the end of the search there
+could easily be hundreds of bases being tested before the target basis
+is found. In the plot, points from those iterations towards the end will
+occupy the vast majority of the plot space. This motivates to use
+summarisation. Rather than knowing the index value of \emph{every}
+basis, we are more interested to have a general summary of all the index
+value in that iteration and more importantly, the basis with the largest
+index value (since it prescribes the next geodesic interpolation and
+future searches).
 
@@
 to find the target basis by displaying the boxplot alone for all
-\texttt{tries}. Thus, the number of point for each \texttt{tries} is
-displayed at the bottom of each box and we provide options to switch
-\texttt{tries} with small number of points to a point geometry, which is
-achieved via the \texttt{cutoff} argument. A line geometry is also added
-to link the points with the largest index value in each \texttt{tries}.
-This helps to visualise the improvement made by each \texttt{tries}.
-With gramma of graphics \citep{wickham2010layered}, the plot for
-exploring index value can be defined in three layers as follows:
+\texttt{tries}. Thus, the number of point in each iteration is displayed
+at the bottom of each box and we provide options to switch iteration
+with small number of points to a point geometry, which is achieved via
+an \texttt{cutoff} argument. A line geometry is also added to link the
+points with the largest index value in each iteration. This helps to
+visualise the improvement made in each iteration. Using the concept of
+\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
+exploring index value can be defined in three layers as:
 
@@
 
-The larger dataset is used for 2D projection with
-\texttt{search\_better} and \texttt{max.tries\ =\ 500}. In Figure
-\ref{points}, the searching points with \texttt{id} and \texttt{tries}
-on the x-axis, colored by \texttt{tries}. Label at the bottom indicates
-the number of observations in each tries and facilitates the choice of
-cutoff to switch from point geometry to boxplot geometry
-(\texttt{cutoff\ =\ 15}). The line geometry suggests the largest
-improvement happens at \texttt{tries\ =\ 5}.
+We choose variable \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
+\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
+Parameter \texttt{search\_f\ =\ tour::search\_better} and
+\texttt{max.tries\ =\ 500} is used. The index value of the searching
+points are shown in Figure \ref{points}. Label at the bottom indicates
+the number of observations in each iteration and facilitates the choice
+of \texttt{cutoff} argument (by default \texttt{cutoff\ =\ 15}). We
+learn that the \texttt{search\_better} quickly finds better projection
+basis with higher index value at first and the takes longer to find a
+better one later.
 
@@
 
-Sometimes, it is more interesting to explore the index value on the
-interpolation path and the plot definition is slightly varied.
+\hypertarget{interpolating-points}{%
+\subsubsection{Interpolating points}\label{interpolating-points}}
+
+Sometimes, rather than explore the searching points, we may be
+interested in explore the points on the interpolation path (target and
+interpolating points) since these points will be played by the tour
+animation. Since interpolation paths are geodesically the shortest, a
+summarisation using boxplot geometry is no longer needed. The slightly
+modified plot definition is shown below:
 
@@
 
-This examples uses \texttt{search\_better} for a 2D projection on the
-larger dataset using the \texttt{holes} index. In the interpolation
-stage, continuous frames will be constructed to bridge the project on
-the current basis to the target basis. The target basis will become the
-current basis in the next iteration and the searching stage continues to
-find the next target basis. In the left panel of Figure
-\ref{interruption}, we observe that there are interpolating bases with
-higher index values than the target basis and these bases could be used
-to search for new basis in the next searching stage.
-
-An interruption is then constructed to accept the interpolating bases up
-to the one with the largest index value and use that basis as the
-current basis for the next searching stage. After implementing this
-interruption, the tracing plot with the same configuration is shown on
-the right panel of Figure \ref{interruption}. In the third
-\texttt{tries}, rather than interpolating fully to the target basis at
-\texttt{id\ =\ 62}, it stops before the index value starts to decrease
-at \texttt{id\ =\ 60}. This implementation results in a higher index
-value in the end with fewer steps.
+We use the same dataset as the toy example above to explore the search
+function \texttt{search\_better} and we want to learn how the index
+value changes on the interpolation path for \texttt{holes} index. From
+the left panel of Figure \ref{interruption}, we observe that when
+interpolating from the current basis to the target basis, the index
+value may not be monotone: we could reach a basis with higher index
+value than the target basis on the interpolation path. In this sense, we
+would be better off using the basis with the highest index value on the
+interpolation path as the current basis for the next iteration (rather
+than using the target basis).
+
+Hence, an interruption is constructed to accept the interpolating bases
+only up to the one with the largest index value. After implementing this
+interruption, the search finds higher final index value with fewer steps
+as shown in the right panel of Figure \ref{interruption}.
 
@@
 
-\hypertarget{another-example-polish}{%
-\paragraph{Another example: Polish}\label{another-example-polish}}
+\hypertarget{polishing-points}{%
+\subsubsection{Polishing points}\label{polishing-points}}
 
 In principle, all the optimisation routines should result in the same
-output for the same problem while this may not be true in real
+output for the same problem while this may not be the case in real
 application. This motivates the creation of a polishing search that
-polishes the final basis and achieves unity on different methods.
-
-\texttt{search\_polish} takes the final basis of a given search as the
-current basis and uses a brutal-force approach to sample a large number
-of basis (\texttt{n\_sample}) in the neighbourhood, whose radius is
-controlled by \texttt{polish\_alpha}. Among the \texttt{n\_sample}
-basis, the one with the largest index value becomes the candidate basis.
-If its index value of the candidate basis is larger than that of the
-current basis, it becomes the current basis in the next iteration. If no
-basis is found to have larger index value than the current basis, the
-searching neighbourhood will be shrunk and the search continues. The
-polishing search ends when one of the four stopping criteria is
-satisfied:
+polishes the final basis found and achieves unity across different
+methods.
+
+\texttt{search\_polish} takes the final basis of a given search as a
+start and uses a brutal-force approach to sample a large number of basis
+(\texttt{n\_sample}) in the neighbourhood. Among those sampled basis,
+the one with the largest index value is chosen to be compared with the
+current basis. If its index value is larger than that of the current
+basis, it becomes the current basis in the next iteration. If no basis
+is found to have larger index value, the searching neighbourhood will
+shrink and the search continues. The polishing search ends when one of
+the four stopping criteria is satisfied:
 
@@
 \item
-  the two basis can't be too close
-\item
-  the percentage improvement of the index function can't be too small
+  the chosen basis can't be too close to the current basis
+\item
+  the percentage improvement of the index value can't be too small
 \item
@@
 
-{[}should the stopping criteria be more detailed? xxx{]}
-
-The usage of search\_polish is as follows. After the first tour, the
-final basis from the interpolation is extracted and supplied into a new
-tour with the \texttt{start} argument and \texttt{search\_polish} as the
-searching function in the guided\_tour. All the other arguments should
-remain the same.
+The usage of search\_polish is as follows. After the first search, the
+final basis from the interpolation is extracted and supplied to the
+second search as the \texttt{start} argument. \texttt{search\_polish} is
+used as the search function. All the other arguments should remain the
+same.
 
@@
 
-The following example conducted a 2D projection on the larger dataset
-using search better with different configurations. \texttt{max.tries} is
-a hyperparameter that controls the maximum number of try without
-improvement and its default value is 25. As shown in Figure
-\ref{trace-compare}, both trials attain the same index value after
-polishing while a small \texttt{max.tries} of 25 is not sufficient for
-\texttt{search\_better} to find the global maximum. The
-\texttt{max.tries} argument needs to be adjusted to ensure it is
-sufficient to explore the parameter space.
+\hypertarget{another-example-polish}{%
+\paragraph{Another example: Polish}\label{another-example-polish}}
+
+Again using the same data, we are interested to compare the effect of
+different \texttt{max.tries} in the 2D projection setting.
+\texttt{max.tries} is a hyperparameter that controls the maximum number
+of try before the search ends. The default value of 25 is suitable for
+1D projection while we suspect it may not be a good option for the 2D
+case and hence want to compare it with an alternative, 500. As shown in
+Figure \ref{trace-compare}, both trials attain the same index value
+after polishing while the small \texttt{max.tries} of 25 is not
+sufficient for \texttt{search\_better} to find its global maximum and we
+will need to adjust the \texttt{max.tries} argument for the search to
+succiciently explore the parameter space.
 
@@
 \hypertarget{animated}{%
-\subsubsection{Explore searching space}\label{animated}}
+\subsection{Explore searching space}\label{animated}}
 
@@
 \hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
-\paragraph{A toy example: understand different stage of
+\subsubsection{A toy example: understand different stage of
 search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
@@
 \hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
-\paragraph{A more complex example: Choosing the initial value for
+\subsubsection{A more complex example: Choosing the initial value for
 polishing
@@
 \hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
-\subsubsection{A comprehensive example of diagnosing a noisy index
+\subsection{A comprehensive example of diagnosing a noisy index
 function}\label{a-comprehensive-example-of-diagnosing-a-noisy-index-function}}
@@
 \hypertarget{explore-index-value}{%
-\paragraph{Explore index value}\label{explore-index-value}}
+\subsubsection{Explore index value}\label{explore-index-value}}
 
@@
 \hypertarget{explore-searching-space}{%
-\paragraph{Explore searching space}\label{explore-searching-space}}
+\subsubsection{Explore searching space}\label{explore-searching-space}}
 

### 2020-07-27 paper/paper.Rmd
fix notation as per meeting 18

@@
 Plots are usually used to evaluate and compare the performance of different algorithms but it can easily become tedious because the code will have to be modified significantly when comparing different parameters in the algorithms.
-For example, a categorical variable with 5 levels can be easily mapped onto color while mapping another categorical variable with 30 levels will not make the plot informative.
+For example, a categorical variable with 5 levels can be easily mapped onto colour while mapping another categorical variable with 30 levels will not make the plot informative.
 Thus the plot needs to be re-designed to better suits the characteristics of the parameter (whether the parameter is a scalar or a matrix?
@@
 Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and can be used to measure the "interestingness" of the projection.
-A number of indice functions have been proposed in the literature including Legendre index [@friedman1974projection], hermite index [@hall1989polynomial], natural hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+A number of index functions have been proposed in the literature including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 
@@
 Details of the mathematical construction of a tour path can be found in @buja2005computational.
-Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+[Maybe add the sketch for illustrating tour path.] Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 
@@
 
-## Notation
-
-For a projection basis, let the three-subscript notation $$\mathbf{A}_{jlk}$$ represent the projection matrix in iteration $j$ with either a searching index $l$ or an interpolation index $k$.
-The placeholder $*$ will be the substitue for the non-applicable index.
-To give two examples, $\mathbf{A}_{12*}$ denotes the second searching basis in the first iteration; $\mathbf{A}_{1*2}$ denotes the second interpolation basis in the first iteration.
-
-If we are being pedantic, all the target bases, except the starting basis, should have a searching index since it is the last candidate basis that successfully allows the optimisation algorithm to end and output a new basis for interpolation.
-If we denotes the length of each search (number of basis searched in each iteration) as $l_1, l_2, \cdots, l_J$ and the length of each interpolation is $k_1, k_2, \cdots, k_J$ , then the searching index for target basis in each iteration will become $\mathbf{A}_{1l_1*}, \mathbf{A}_{2l_2*}, \cdots, \mathbf{A}_{Jl_J*}$.
-We simplify this notation for target bases by using a superscript as $$\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}$$ where $\mathbf{A}^{0}$ is the starting basis.
-Notice that in the case when the complex notation can't be avoided, we shall separate each index using a comma and  write the index in brackets if necessary.
-Using the notation above, all the bases on the interpolation path (including target and interpolation bases) can be written as in Equation \ref{eq:interp-path}.
-
-\begin{equation}
-\{ 
-\mathbf{A}^0, 
-\mathbf{A}^1, 
-\mathbf{A}_{1 \ast 1}, 
-\mathbf{A}_{1 \ast 2},
-\cdots, 
-\mathbf{A}^2, 
-\mathbf{A}_{2 \ast 1}, 
-\mathbf{A}_{2 \ast 2}, 
-\cdots, 
-\mathbf{A}^J,
-\mathbf{A}_{J \ast 1}, 
-\mathbf{A}_{J \ast 2}, 
-\cdots, 
-\mathbf{A}_{J \ast K},
-\}
-\label{eq:interp-path}
-\end{equation}
-
-where $K$ denotes the last interpolating index.
-A visual representation of the interpolation path ( for iteration one) modified from @buja2005computational is shown in Figure \ref{tour-path}.
-If we use a dot to represent a plane in the searching space, the searching points along with the current and target basis in iteration one can sketched in Figure \ref{searching-points}.
-Combine the two above, we can write all the bases in the first iteration (including the starting basis) in its natural ordering as in Equation \ref{eq:vectorA}.
-
-\begin{equation}
-\{
-\mathbf{A}^{0}, 
-\mathbf{A}_{1 1\ast}, 
-\mathbf{A}_{1 2\ast}, 
-\cdots, 
-\mathbf{A}^{1},
-\mathbf{A}_{1\ast 1}, 
-\mathbf{A}_{1\ast 2},  
-\cdots, 
-\mathbf{A}_{1\ast k_1}
-\}
-\label{eq:vectorA}
-\end{equation}
-
 ```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
@@
 
-```{r searching-points, out.width="100%", out.height="60%", fig.cap="\\label{searching-points}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
-```
-
 ## Optimisation problem and existing algorithms
@@
 \begin{align}
-&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{A}) \\
+&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
 s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
@@
 
-There are three existing methods for optimisating PPI function and we review them below.
+There are three existing methods for optimising PPI function and we review them below.
 
 @posse1995projection proposed a stochastic direct search method, a random search algorithm.
-Given a current basis $\mathbf{A}^{j-1}$, a candidate basis $\mathbf{A}_{jl*}$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}^{j}$ by $\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}$.
-If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $\mathbf{A}^{j}$ along with other metadata and the current iteration stops.
+Given a current basis $\mathbf{A}_{\text{cur}}$, a candidate basis $\mathbf{A}_{t}$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}_{t}$ by $\mathbf{A}_{t} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$.
+If the candidate basis has a higher index value than the current basis, it is outputted as the target basis $\mathbf{A}_{t}$ along with other metadata and the current iteration stops.
 If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted.
-$c$ is the halfing parameter and when $c > 30$, the nieghbourhood parameter $\alpha$ in the next iteration will be reduced by half.
+$c$ is the halfing parameter and when $c > 30$, the neighbourhood parameter $\alpha$ in the next iteration will be reduced by half.
 The algorithm for a random search is summarised in Algorithm \ref{random-search}.
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
-    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
+    \Output{The target basis: $\mathbf{A}_{t+1}$}
   initialisation\;
+  Set $l = 1$\;
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{j, l,*}$ is orthonormal\;
-    Compute $I_{jl*}  = f(\mathbf{A}_{jl*})$\;
-    \eIf{$I_{jl*} > I^{j-1}$}{
-      $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
+    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{l}$ is orthonormal\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    \eIf{$I_{l} > I_{\text{cur}}$}{
+      \KwRet{$\mathbf{A}_{l}$} \;
       }{
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
-    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
+    \Output{The target basis: $\mathbf{A}_{**}$}
   initialisation\;
+  Set $l = 1$\;
   \While{$l < l_{\max}$}{
-    Generate ten bases in five random directions: $\mathbf{A}_{jl*}: \mathbf{A}_{j,(l + 9), *}$ within a small neighbourhood $\text{dist}$\;
-    Find the direction with the largest index value: $\mathbf{A}_{j, l_{d}, *}$\;
-    Construct the geodesic from $\mathbf{A}^{j-1}$ to $\mathbf{A}_{j, l_{d}, *}$\;
-    Find $\mathbf{A}_{j,l_g, *}$ on the geodesic that has the largest index value \;
-    Compute $I_{j,l_g, *} = f(\mathbf{A}_{j,l_g, *})$, $p_{\text{diff}} = (I_{j,l_g, *} - I^{j-1})/I_{j,l_g, *}$\;
+    Generate ten bases in five random directions: $\mathbf{A}_{l}: \mathbf{A}_{l+9}$ within a small neighbourhood $\text{dist}$\;
+    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $l < * < l+ 9$\;
+    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
+    Find $\mathbf{A}_{**}$ on the geodesic $\mathcal{G}$ that has the largest index value \;
+    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
       \If{$p_{\text{diff}} > 0.001$}{
-        $A^{j} = A_{j,l_g, *}$, $I^{j} = I_{j,l_g, *}$\;
+         \KwRet{$\mathbf{A}_{**}$} \;
       }
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$.
-Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{log(l + 1)}$.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(t)$.
+Given an initial $T_0$, the temperature at iteration $t$ is defined as $T(t) = \frac{T_0}{log(t + 1)}$.
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}$.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}$ where $I^{j-1}$ and $I_{jl*}$ are the index value of the current and candidate basis respectively.
+If it is smaller, the candidate is accepted with probability $P= \min\left\{\exp\left[\frac{I_{t+1} - I_{t}}{T(t)}\right],1\right\}$ where $I_{t}$ and $I_{t+1}$ are the index value of the current and candidate basis respectively.
 The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}^{j-1}$; The index function: $f$}
+    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
     \Output{The global object; The target basis: $\mathbf{A}^{j}$}
@@
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{jl*}$ is orthonormal \;
-    Commpute $I_{jl*} = f(\mathbf{A}_{jl*})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I_{jl*} > I^{j-1}$}{
-        $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
+    Set $l = 0$\;
+    Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
+    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{l} > I_{\text{cur}}$}{
+        \KwRet{$\mathbf{A}_{l}$} \;
       }{
-        Compute $P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}$\;
+        Compute $P= \min\left\{\exp\left[\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
         Draw $D$ from a uniform distribution: $D \sim \text{Unif(0, 1)}$\;
         \If{$P > D$}{
-          $A^{j} = A_{jl*}$, $I^{j} = I_{jl*}$\;
+           \KwRet{$\mathbf{A}_{l}$} \;
         }
@@
 Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
-Given an optimisation algorithm, two primary variables of interest are the *decision variable: $\mathbf{A}$* and the *value of the objective function: $I$*.
-To further simplify the notation for $\mathbf{A}$, we write the bases as a column vector denoting by a single subscript by their natural ordering as in Equation \ref{vector-basis}.
-
-\begin{equation}
-\{
-\mathbf{A}_0, 
-\mathbf{A}_1, 
-\mathbf{A}_2, 
-\cdots,
-\mathbf{A}_T
-\label{vector-basis}
-\}
-\end{equation}
-
-where $T$ is the total number of plane serached by the optimisation algorithm including all the target planes, all the interpolating planes and all the candidate planes but NOT the starting plane.
-
-*Iterators* indexes the data collected and has a time series feature that prescribes the its natural ordering of the decision varible in the searching.
-The simpliest iterator, *index per observation: $t$*, is a unique identifier for each observation in the data.
-For each level of iteration, we design two types of indices: *index per iteration: $j$* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
-*index within iteration: $l$* has an increment of one for each observation in an iteartion and starts over from one once a new iteration starts.
-A sketch of the difference between these three iterators is provided in Figure \ref{iterators}.
-In projection pursuit optimisation algorithms, there is one level of iteartion and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
+Given an optimisation algorithm, two primary variables of interest are the *decision variable: $\mathbf{A}$* and the *value of the objective function: $I$*, which is the projection basis and index value repsectively in projection pursuit.
+*State: $S$* is another variable of great importance since it labels $\mathbf{A}$ and $I$ by the stage they are in the whole search-and-interpolate process.
+
+*Iterators* indexes the data collected and prescribes the the natural ordering of each observation in different levels.
+*$t$* is a unique identifier that has an increment of one for each observation and *$j$* is the counter for each search-and-interpolate round.
+[add a bit more for *$l$* after decide whether to combine $l$ and $k$.
+]
+
+<!-- For each level of iteration, we design two types of indices: *index per iteration: $j$* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
+*index within iteration: $l$* has an increment of one for each observation in an iteration and starts over from one once a new iteration starts.
+-->
+
+<!-- In projection pursuit optimisation algorithms, there is one level of iteration and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
 We give the interpolating basis a different index $k$ for projection pursuit guided tour.
 It is similar in nature to $l$ but is specific for interpolating bases, which is usually not part of the optimisation.
-
-```{r iterators, out.width = "100%", out.height = "25%", fig.cap = "\\label{iterators} A sketch of the design of iterators in iterative algorithms."}
-knitr::include_graphics(here::here("figures", "iterators.png"))
-```
-
-There could exist other parameters that are also of our interest but can't be classified as one of the three categories.
-They are defined under the fourth category: *other parameter of interest: $S$*.
-In projection pursuit, three other parameters of our interest includes `method`, `alpha` and `info`.
-`method` identifies the name of the searching method used and we are interested in comparing the performance between different algorithms under direct search.
-The neighbourhood parameter `alpha` controls the size of the sampling space and we are interested to understand how the searching space shrinks as the algorithm progresses.
-`info` labels different stages in the searching process.
+-->
+
+<!-- # ```{r iterators, out.width = "100%", out.height = "25%", fig.cap = "\\label{iterators} A sketch of the design of iterators in iterative algorithms."} -->
+<!-- # knitr::include_graphics(here::here("figures", "iterators.png")) -->
+<!-- # ``` -->
+
+There are other parameters that are of our interest and we denote them as *$V_{p}$*.
+In projection pursuit, this includes $V_1 = \text{method}$, denoting the name of the searching method used and $V_2 = \text{alpha}$, which is a neighbourhood parameter that controls the size in sampling candidate bases.
 A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
@@
 \left[
-\begin{array}{c|cc|ccc|ccc}
-t & \mathbf{A} & I & j &  l  & k & S_{1}& \cdots & S_{p}\\
+\begin{array}{c|ccc|ccc|ccc}
+t & \mathbf{A} & I & S & j &  l  & k & V_{1}& \cdots & V_{p}\\
 \hline
-1 & \mathbf{A}_0 & I_0 & 0 & \ast & \ast & \ast & \cdots & \ast\\
+1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & \ast & V_{1, 1} & \cdots & V_{p, 1}\\
 \hline
-2 & \mathbf{A}_1 & I_1 & 1 & 1 & \ast & S_{1, 1} & \cdots & S_{1, p}\\
-3 & \mathbf{A}_2 & I_2 & 1 & 2 & \ast & S_{2, 1} & \cdots & S_{2, p}\\
-\vdots & \vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
-\vdots & \vdots & \vdots & 1 & l_1 & \ast & \vdots &  \cdots & \vdots\\
+2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1 & \ast & V_{1, 2} & \cdots & V_{p, 2}\\
+3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2 & \ast & V_{1, 3} & \cdots & V_{p, 3}\\
+\vdots & \vdots &\vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
+\vdots & \vdots & \vdots &\vdots & 2 & l_1 & \ast & \vdots &  \cdots & \vdots\\
 \hline
-\vdots &\vdots & \vdots & 1 & \ast & 1& \vdots &  \cdots & \vdots\\
-\vdots &\vdots &\vdots& 1 & \ast & 2& \vdots& \cdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-\vdots &\vdots &\vdots & 1 & \ast & k_1&\vdots & \cdots & \vdots\\
+\vdots &\vdots & \vdots &\vdots & 2 & \ast & 1& \vdots &  \cdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots & 2 & \ast & 2& \vdots& \cdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+\vdots &\vdots &\vdots &\vdots & 2 & \ast & k_1&\vdots & \cdots & \vdots\\
 \hline
-\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-T+1 & \mathbf{A}_T & I_T & J &  l_J & \ast & S_{T, 1}& \cdots & S_{T, p}\\
+\vdots &\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
+T & \mathbf{A}_T & I_T &S_T  & J &  l_J & \ast & V_{1, T}& \cdots & V_{p, T}\\
 \end{array}
@@
 We first simulate some random variables of size 1000 with different structures.
-`x1`, `x8`, `x9` and `x10` are simulated from normal distribution iwth zero mean and variance of one as in equation \ref{eq:sim-norm}.
+`x1`, `x8`, `x9` and `x10` are simulated from normal distribution with zero mean and variance of one as in equation \ref{eq:sim-norm}.
 When using projection pursuit to explore the data structure based on its departure from normality, the entry in the projection basis for these variables should be close to zero in theory.
@@
 In the first two sections, we will first provide a toy example that is easy to grasp and then more examples that can help us to understand the algorithm and parameter choice.
-Remeber the research question we raised earlier, the purpose of visual diagnostics is to understand:
+Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:
 
@@
 The second question can be addressed via visualising the rotating high dimensional space or its projection on the reduced 2D space.
-Thus animated visualisation is needed to preceive the optimisation path in the searching space.
+Thus animated visualisation is needed to perceive the optimisation path in the searching space.
 
@@
 - Layer 1: boxplot geom
-  - data: group by `tries` and filter the observations in the group that have count greater than `cutoff = 15`.
-  - x: `tries` is mapped to the x-axis
-  - y: the index value after statistical transformation $y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}$ is mapped to the y-axis where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$.
+  - data: group by $j$ and filter the observations in the group that have count greater than `cutoff = 15`.
+  - x: $j$ is mapped to the x-axis
+  - y: the statistical transformed index value: $Q_q(I_{T} \cdot \mathds{1} (S = \text{search}))$ is mapped to the y-axis where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$.
 $\mathds{1}(.)$ is the identity operator.
@@
 - Layer 2: point geom
-  - data: group by `tries` and filter the observations in the group that have count less than `cutoff = 15`.
-  - x: `tries` is mapped to the x-axis
-  - y: `index_val` is mapped to the y-axis
+  - data: group by $j$ and filter the observations in the group that have count less than `cutoff = 15`.
+  - x: $j$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
 
 - Layer 3: line geom
-  - data: filter the points with the highest index value in each `tries`
-  - x: `tries` is mapped to the x-axis
-  - y: `index_val` is mapped to the y-axis
+  - data: filter the points with the highest index value in group $j$
+  - x: $j$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
 
@@
 - Layer 1: point geom
-  - data: filter the observations with `info` being `interpolation` and mutate `id` to be the row number of the subsetted tibble
-  - x: `id` is mapped to the x-axis
-  - y: `index_val` is mapped to the y-axis
+  - data: filter the observations with $S = \text{interpolation}$ and mutate $t$ to be the row number of the subsetted tibble
+  - x: $t$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
 
 - Layer 2: line geom
-  - using line goemetry for the same data and aesthetics
+  - using line geometry for the same data and aesthetics
 
@@
 <!-- In the previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e.
-color, could be added to compare other parameters in the global object.
+colour, could be added to compare other parameters in the global object.
 -->
@@
 - Layer 1: point geom
-  - data: filter the observations with `info` being `interpolation`; bind the global object from optimisation and interpolation and form polishing; mutate `id` to be the row number of the binded tibble.
-  - x: `id` is mapped to the x-axis
-  - y: `index_val` is mapped to the y-axis 
-  - color: `method` is mapped to the color aesthetic
+  - data: filter the observations with  $S = \text{interpolation}$; bind the global object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
+  - x: $t$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis 
+  - colour: $V$ is mapped to the colour aesthetic
 
 - Layer 2: line geom
-  - using line goemetry for the same data and aesthetics
 
@@
 The default value of 25 is suitable for 1D projection while we suspect it may not be a good option for the 2D case and hence want to compare it with an alternative, 500.
-As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to succiciently explore the parameter space.
+As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to sufficiently explore the parameter space.
 
@@
 <!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
-THis means all the iterators and interesting other variables defined in the global object can be binded with the these two projections and can be mapped to other aesthetics (i.e.
-color).
+THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
+colour).
 -->
@@
   - y: the projected basis on the second principal component
-  - color: an variable of interest is mapped onto color
+  - colour: $V$ is mapped to the colour aesthetic
 
@@
 
-```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic Coloring by info allows for better understanding of each stage in the geodesic search"}
+```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
 
@@
 The problem is a 1D projection of the small dataset using `search_geodesic` and followed by `search_polish`.
-The top-left panel of Figure \ref{polish-alpha} displays all the projection bases on the first two principal components, colored by the `polish_alpha`.
+The top-left panel of Figure \ref{polish-alpha} displays all the projection bases on the first two principal components, coloured by the `polish_alpha`.
 We can observe that rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, `search_polish` searches a much larger vector space, which is unnecessary.
@@
 `kol_cdf` index, an 1D projection index function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as $F(u)$ with subscript incidating the variable,  the Kolmogorove statistics defined by
+Denotes the ECDF function as $F(u)$ with subscript indicating the variable,  the Kolmogorov statistics defined by
 
@@
 
-Polishing step has done much more work to find the final index value ub `search_geodesic` than `search_better` and this indicates `kol_cdf` function favours of a random search method than ascent method.
+Polishing step has done much more work to find the final index value in `search_geodesic` than `search_better` and this indicates `kol_cdf` function favours of a random search method than ascent method.
 
@@
 Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
-color <- RColorBrewer::brewer.pal(3, "Dark2")
+colour <- RColorBrewer::brewer.pal(3, "Dark2")
 p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
-  scale_color_manual(values = color[2]) + ggtitle("search_better, seed 123456")
+  scale_colour_manual(values = colour[2]) + ggtitle("search_better, seed 123456")
 p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
-  scale_color_manual(values = color[1]) + ggtitle("search_better, seed 12345")
+  scale_colour_manual(values = colour[1]) + ggtitle("search_better, seed 12345")
 
@@
   ylim(0, 0.25) + 
-  scale_color_manual(values = color[3]) + 
+  scale_colour_manual(values = colour[3]) + 
   theme(legend.position = "none") + 
@@
 We can also plot the searching points of all the three algorithms in the searching space and explore their relative position against each other using principal components.
-As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhausive search.
+As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhaustive search.
 The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end.
@@
   explore_proj_pca(col = algorithm) + 
-  scale_color_brewer(palette = "Dark2")
-
-```
-
-# Implementation: Ferrn pacakge
+  scale_colour_brewer(palette = "Dark2")
+
+```
+
+# Implementation: Ferrn package
 

### 2020-07-27 paper/paper.tex
fix notation as per meeting 18

@@
 algorithms. For example, a categorical variable with 5 levels can be
-easily mapped onto color while mapping another categorical variable with
-30 levels will not make the plot informative. Thus the plot needs to be
-re-designed to better suits the characteristics of the parameter
+easily mapped onto colour while mapping another categorical variable
+with 30 levels will not make the plot informative. Thus the plot needs
+to be re-designed to better suits the characteristics of the parameter
 (whether the parameter is a scalar or a matrix? whether the parameter is
@@
 function, or the index function and can be used to measure the
-``interestingness'' of the projection. A number of indice functions have
+``interestingness'' of the projection. A number of index functions have
 been proposed in the literature including Legendre index
-\citep{friedman1974projection}, hermite index
-\citep{hall1989polynomial}, natural hermite index
+\citep{friedman1974projection}, Hermite index
+\citep{hall1989polynomial}, natural Hermite index
 \citep{cook1993projection}, chi-square index
@@
 shortest. Details of the mathematical construction of a tour path can be
-found in \citet{buja2005computational}. Guided tour, along with other
-types of tour, has been implemented in the \emph{tourr} package in R,
-available on the Comprehensive R Archive Network at
+found in \citet{buja2005computational}. {[}Maybe add the sketch for
+illustrating tour path.{]} Guided tour, along with other types of tour,
+has been implemented in the \emph{tourr} package in R, available on the
+Comprehensive R Archive Network at
 \url{https://cran.r-project.org/web/packages/tourr/}
@@
 
-\hypertarget{notation}{%
-\subsection{Notation}\label{notation}}
-
-For a projection basis, let the three-subscript notation
-\[\mathbf{A}_{jlk}\] represent the projection matrix in iteration \(j\)
-with either a searching index \(l\) or an interpolation index \(k\). The
-placeholder \(*\) will be the substitue for the non-applicable index. To
-give two examples, \(\mathbf{A}_{12*}\) denotes the second searching
-basis in the first iteration; \(\mathbf{A}_{1*2}\) denotes the second
-interpolation basis in the first iteration.
-
-If we are being pedantic, all the target bases, except the starting
-basis, should have a searching index since it is the last candidate
-basis that successfully allows the optimisation algorithm to end and
-output a new basis for interpolation. If we denotes the length of each
-search (number of basis searched in each iteration) as
-\(l_1, l_2, \cdots, l_J\) and the length of each interpolation is
-\(k_1, k_2, \cdots, k_J\) , then the searching index for target basis in
-each iteration will become
-\(\mathbf{A}_{1l_1*}, \mathbf{A}_{2l_2*}, \cdots, \mathbf{A}_{Jl_J*}\).
-We simplify this notation for target bases by using a superscript as
-\[\mathbf{A}^{0}, \mathbf{A}^{1}, \mathbf{A}^{2}, \cdots, \mathbf{A}^{J}\]
-where \(\mathbf{A}^{0}\) is the starting basis. Notice that in the case
-when the complex notation can't be avoided, we shall separate each index
-using a comma and write the index in brackets if necessary. Using the
-notation above, all the bases on the interpolation path (including
-target and interpolation bases) can be written as in Equation
-\ref{eq:interp-path}.
-
-\begin{equation}
-\{ 
-\mathbf{A}^0, 
-\mathbf{A}^1, 
-\mathbf{A}_{1 \ast 1}, 
-\mathbf{A}_{1 \ast 2},
-\cdots, 
-\mathbf{A}^2, 
-\mathbf{A}_{2 \ast 1}, 
-\mathbf{A}_{2 \ast 2}, 
-\cdots, 
-\mathbf{A}^J,
-\mathbf{A}_{J \ast 1}, 
-\mathbf{A}_{J \ast 2}, 
-\cdots, 
-\mathbf{A}_{J \ast K},
-\}
-\label{eq:interp-path}
-\end{equation}
-
-where \(K\) denotes the last interpolating index. A visual
-representation of the interpolation path ( for iteration one) modified
-from \citet{buja2005computational} is shown in Figure \ref{tour-path}.
-If we use a dot to represent a plane in the searching space, the
-searching points along with the current and target basis in iteration
-one can sketched in Figure \ref{searching-points}. Combine the two
-above, we can write all the bases in the first iteration (including the
-starting basis) in its natural ordering as in Equation \ref{eq:vectorA}.
-
-\begin{equation}
-\{
-\mathbf{A}^{0}, 
-\mathbf{A}_{1 1\ast}, 
-\mathbf{A}_{1 2\ast}, 
-\cdots, 
-\mathbf{A}^{1},
-\mathbf{A}_{1\ast 1}, 
-\mathbf{A}_{1\ast 2},  
-\cdots, 
-\mathbf{A}_{1\ast k_1}
-\}
-\label{eq:vectorA}
-\end{equation}
-
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/tour_path_increasing} \caption{\label{searching-points}An illustration of the tour path}\label{fig:searching-points}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
 \end{figure}
@@
 
-There are three existing methods for optimisating PPI function and we
+There are three existing methods for optimising PPI function and we
 review them below.
@@
 \citet{posse1995projection} proposed a stochastic direct search method,
-a random search algorithm. Given a current basis \(\mathbf{A}^{j-1}\), a
-candidate basis \(\mathbf{A}_{jl*}\) is sampled in the neighbourhood
-defined by the radius of the p-dimensional sphere, \(\alpha\) of the
-current basis \(\mathbf{A}^{j}\) by
-\(\mathbf{A}_{jl*} = \mathbf{A}^{j-1} + \alpha \mathbf{A}_{rand}\). If
-the candidate basis has a higher index value than the current basis, it
-is outputed as the target basis \(\mathbf{A}^{j}\) along with other
+a random search algorithm. Given a current basis
+\(\mathbf{A}_{\text{cur}}\), a candidate basis \(\mathbf{A}_{t}\) is
+sampled in the neighbourhood defined by the radius of the p-dimensional
+sphere, \(\alpha\) of the current basis \(\mathbf{A}_{t}\) by
+\(\mathbf{A}_{t} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}\).
+If the candidate basis has a higher index value than the current basis,
+it is outputted as the target basis \(\mathbf{A}_{t}\) along with other
 metadata and the current iteration stops. If no basis is found to have
@@
 algorithm stops with nothing outputted. \(c\) is the halfing parameter
-and when \(c > 30\), the nieghbourhood parameter \(\alpha\) in the next
+and when \(c > 30\), the neighbourhood parameter \(\alpha\) in the next
 iteration will be reduced by half. The algorithm for a random search is
@@
 non-derivative procedure based on a non-increasing cooling scheme
-\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
-is defined as \(T(l) = \frac{T_0}{log(l + 1)}\). The simulated annealing
+\(T(t)\). Given an initial \(T_0\), the temperature at iteration \(t\)
+is defined as \(T(t) = \frac{T_0}{log(t + 1)}\). The simulated annealing
 algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
@@
 basis is constructed as
-\(\mathbf{A}_{jl*} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}\).
+\(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}\).
 If the index value of the candidate basis is larger than the one of the
@@
 smaller, the candidate is accepted with probability
-\(P= \min\left\{\exp\left[\frac{I^{j-1} - I_{jl*}}{T(l)}\right],1\right\}\)
-where \(I^{j-1}\) and \(I_{jl*}\) are the index value of the current and
+\(P= \min\left\{\exp\left[\frac{I_{t+1} - I_{t}}{T(t)}\right],1\right\}\)
+where \(I_{t}\) and \(I_{t+1}\) are the index value of the current and
 candidate basis respectively. The algorithm can be summarised as in
@@
 variables of interest are the \emph{decision variable: \(\mathbf{A}\)}
-and the \emph{value of the objective function: \(I\)}. To further
-simplify the notation for \(\mathbf{A}\), we write the bases as a column
-vector denoting by a single subscript by their natural ordering as in
-Equation \ref{vector-basis}.
-
-\begin{equation}
-\{
-\mathbf{A}_0, 
-\mathbf{A}_1, 
-\mathbf{A}_2, 
-\cdots,
-\mathbf{A}_T
-\label{vector-basis}
-\}
-\end{equation}
-
-where \(T\) is the total number of plane serached by the optimisation
-algorithm including all the target planes, all the interpolating planes
-and all the candidate planes but NOT the starting plane.
-
-\emph{Iterators} indexes the data collected and has a time series
-feature that prescribes the its natural ordering of the decision varible
-in the searching. The simpliest iterator, \emph{index per observation:
-\(t\)}, is a unique identifier for each observation in the data. For
-each level of iteration, we design two types of indices: \emph{index per
-iteration: \(j\)} is fixed for each observations in one iteration and
-has an increment of one once a new iteration starts. \emph{index within
-iteration: \(l\)} has an increment of one for each observation in an
-iteartion and starts over from one once a new iteration starts. A sketch
-of the difference between these three iterators is provided in Figure
-\ref{iterators}. In projection pursuit optimisation algorithms, there is
-one level of iteartion and hence exists three iterators: \texttt{id}
-indices each observation by a unique number; \texttt{tries} is the index
-per iteration iterator that gets updated once a search-and-interpolate
-step is finished; and \texttt{loop} is the index within iteration
-iterator and starts over from one at the beginning of a new
-\texttt{tries}. We give the interpolating basis a different index \(k\)
-for projection pursuit guided tour. It is similar in nature to \(l\) but
-is specific for interpolating bases, which is usually not part of the
-optimisation.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/iterators} \caption{\label{iterators} A sketch of the design of iterators in iterative algorithms.}\label{fig:iterators}
-\end{figure}
-
-There could exist other parameters that are also of our interest but
-can't be classified as one of the three categories. They are defined
-under the fourth category: \emph{other parameter of interest: \(S\)}. In
-projection pursuit, three other parameters of our interest includes
-\texttt{method}, \texttt{alpha} and \texttt{info}. \texttt{method}
-identifies the name of the searching method used and we are interested
-in comparing the performance between different algorithms under direct
-search. The neighbourhood parameter \texttt{alpha} controls the size of
-the sampling space and we are interested to understand how the searching
-space shrinks as the algorithm progresses. \texttt{info} labels
-different stages in the searching process. A sketch of the global object
-for projection pursuit guided tour is presented in Figure
+and the \emph{value of the objective function: \(I\)}, which is the
+projection basis and index value repsectively in projection pursuit.
+\emph{State: \(S\)} is another variable of great importance since it
+labels \(\mathbf{A}\) and \(I\) by the stage they are in the whole
+search-and-interpolate process.
+
+\emph{Iterators} indexes the data collected and prescribes the the
+natural ordering of each observation in different levels. \emph{\(t\)}
+is a unique identifier that has an increment of one for each observation
+and \emph{\(j\)} is the counter for each search-and-interpolate round.
+{[}add a bit more for \emph{\(l\)} after decide whether to combine \(l\)
+and \(k\). {]}
+
+There are other parameters that are of our interest and we denote them
+as \emph{\(V_{p}\)}. In projection pursuit, this includes
+\(V_1 = \text{method}\), denoting the name of the searching method used
+and \(V_2 = \text{alpha}\), which is a neighbourhood parameter that
+controls the size in sampling candidate bases. A sketch of the global
+object for projection pursuit guided tour is presented in Figure
 \ref{fig:glb-obj}. The full data structure can thus be shown as in
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
 \end{figure}
@@
 structures. \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are
-simulated from normal distribution iwth zero mean and variance of one as
+simulated from normal distribution with zero mean and variance of one as
 in equation \ref{eq:sim-norm}. When using projection pursuit to explore
@@
 easy to grasp and then more examples that can help us to understand the
-algorithm and parameter choice. Remeber the research question we raised
+algorithm and parameter choice. Remember the research question we raised
 earlier, the purpose of visual diagnostics is to understand:
@@
 visualising the rotating high dimensional space or its projection on the
-reduced 2D space. Thus animated visualisation is needed to preceive the
+reduced 2D space. Thus animated visualisation is needed to perceive the
 optimisation path in the searching space.
@@
   \item
-    data: group by \texttt{tries} and filter the observations in the
-    group that have count greater than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \texttt{tries} is mapped to the x-axis
-  \item
-    y: the index value after statistical transformation
-    \(y = \{Q_q(I_{jlk} \cdot \mathds{1} (j = x, k = \ast))\}\) is
-    mapped to the y-axis where \(q\) takes one of \(0, 25, 50, 75, 100\)
-    and \(Q_q(x)\) finds the qth-quantile for the vector \(x\).
+    data: group by \(j\) and filter the observations in the group that
+    have count greater than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: the statistical transformed index value:
+    \(Q_q(I_{T} \cdot \mathds{1} (S = \text{search}))\) is mapped to the
+    y-axis where \(q\) takes one of \(0, 25, 50, 75, 100\) and
+    \(Q_q(x)\) finds the qth-quantile for the vector \(x\).
     \(\mathds{1}(.)\) is the identity operator.
@@
   \item
-    data: group by \texttt{tries} and filter the observations in the
-    group that have count less than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \texttt{tries} is mapped to the x-axis
-  \item
-    y: \texttt{index\_val} is mapped to the y-axis
+    data: group by \(j\) and filter the observations in the group that
+    have count less than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
   \end{itemize}
@@
   \item
-    data: filter the points with the highest index value in each
-    \texttt{tries}
-  \item
-    x: \texttt{tries} is mapped to the x-axis
-  \item
-    y: \texttt{index\_val} is mapped to the y-axis
+    data: filter the points with the highest index value in group \(j\)
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
   \end{itemize}
@@
   \item
-    data: filter the observations with \texttt{info} being
-    \texttt{interpolation} and mutate \texttt{id} to be the row number
-    of the subsetted tibble
-  \item
-    x: \texttt{id} is mapped to the x-axis
-  \item
-    y: \texttt{index\_val} is mapped to the y-axis
+    data: filter the observations with \(S = \text{interpolation}\) and
+    mutate \(t\) to be the row number of the subsetted tibble
+  \item
+    x: \(t\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
   \end{itemize}
@@
   \item
-    using line goemetry for the same data and aesthetics
+    using line geometry for the same data and aesthetics
   \end{itemize}
@@
 interpolation path as the current basis for the next iteration (rather
-than using the target basis).
-
-Hence, an interruption is constructed to accept the interpolating bases
-only up to the one with the largest index value. After implementing this
-interruption, the search finds higher final index value with fewer steps
-as shown in the right panel of Figure \ref{interruption}.
+than using the target basis). Hence, an interruption is constructed to
+accept the interpolating bases only up to the one with the largest index
+value. After implementing this interruption, the search finds higher
+final index value with fewer steps as shown in the right panel of Figure
+\ref{interruption}.
 
@@
   \item
-    data: filter the observations with \texttt{info} being
-    \texttt{interpolation}; bind the global object from optimisation and
-    interpolation and form polishing; mutate \texttt{id} to be the row
-    number of the binded tibble.
-  \item
-    x: \texttt{id} is mapped to the x-axis
-  \item
-    y: \texttt{index\_val} is mapped to the y-axis
-  \item
-    color: \texttt{method} is mapped to the color aesthetic
+    data: filter the observations with \(S = \text{interpolation}\);
+    bind the global object from optimisation and interpolation and form
+    polishing; mutate \(t\) to be the row number of the binded tibble.
+  \item
+    x: \(t\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
+  \item
+    colour: \(V\) is mapped to the colour aesthetic
   \end{itemize}
@@
   Layer 2: line geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    using line goemetry for the same data and aesthetics
-  \end{itemize}
 \end{itemize}
@@
 will need to adjust the \texttt{max.tries} argument for the search to
-succiciently explore the parameter space.
+sufficiently explore the parameter space.
 
@@
   \item
-    color: an variable of interest is mapped onto color
-  \end{itemize}
-\end{itemize}
-
-While explore the reduced space is an initial attempt to understand the
-searching space, there are existing technology for rotating a higher
-dimensional space for visualisation. Geozoo is an option. It generates
-random points on the high dimensional space and we can overlay it with
-the points on the optimisation path to visualise the spread of it on the
-high-D sphere.
+    colour: \(V\) is mapped to the colour aesthetic
+  \end{itemize}
+\end{itemize}
 
@@
 \includegraphics{paper_files/figure-latex/pca-1.pdf}
-\caption{\label{pca}PCA plot of search geodesic Coloring by info allows
+\caption{\label{pca}PCA plot of search geodesic colouring by info allows
 for better understanding of each stage in the geodesic search}
@@
 top-left panel of Figure \ref{polish-alpha} displays all the projection
-bases on the first two principal components, colored by the
+bases on the first two principal components, coloured by the
 \texttt{polish\_alpha}. We can observe that rather than concentrating on
@@
 
+While explore the reduced space is an initial attempt to understand the
+searching space, there are existing technology for rotating a higher
+dimensional space for visualisation. Geozoo is an option. It generates
+random points on the high dimensional space and we can overlay it with
+the points on the optimisation path to visualise the spread of it on the
+high-D sphere.
+
+{[}add example from geozoo{]}
+
 \newpage
@@
 distribution function (ECDF). Denotes the ECDF function as \(F(u)\) with
-subscript incidating the variable, the Kolmogorove statistics defined by
+subscript indicating the variable, the Kolmogorov statistics defined by
 
@@
 
-Polishing step has done much more work to find the final index value ub
+Polishing step has done much more work to find the final index value in
 \texttt{search\_geodesic} than \texttt{search\_better} and this
@@
 search a proportion of the searching space while better\_random produces
-a more exhausive search. The large overlapping of better1 and
+a more exhaustive search. The large overlapping of better1 and
 better\_random is explained by the fact that both algorithms finds x2 in
@@
 
-\hypertarget{implementation-ferrn-pacakge}{%
+\hypertarget{implementation-ferrn-package}{%
 \section{Implementation: Ferrn
-pacakge}\label{implementation-ferrn-pacakge}}
+package}\label{implementation-ferrn-package}}
 

### 2020-07-29 paper/paper.Rmd
modify based on meeting 19 and polish section 3 and 4 (except 4.1)

@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools, dsfont}
+  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 
-The paper is organised as follows.
-Section \ref{DFO} gives a general literature review of optimisation, specifically derivative free optimisation.
-Section \ref{vis-diag} presents the new idea of constructing a
-systematic visual framework that diagnoses the components of an optimisation procedure (parameters, searching path, etc).
-The rest of the paper serves as a comprehensive example of using the visual diagnostics on one particular problem: *projection pursuit guided tour*.
-Some background knowledge of projection pursuit guided tour is provided in Section \ref{tour}.
-Section \ref{apply} applies the concepts proposed in section \ref{vis-diag} in the tour problem and sets up the data.
-The last section, Section \ref{plots}, presents the visual diagnostic plots and explains how they can help to understand different aspects of the optimisation in tour.
+The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
+Section \ref{tour} introduces the optimisation problem in a specific problem: projection pursuit guided tour and existing derivative free algorithms for solving it. 
+Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveil the blackbox of optimisation.
+Section \ref{vis-diag} focues on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples.
+Finally, \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
 
@@
 
+\newpage
+
 # Projection pursuit guided tour {#tour}
 
-From Section \ref{tour}, we presents a comprehensive case study on how to use the visual diagnostics to explore the optimisation in a specific problem: projection pursuit guided tour.
-Section \ref{tour} aims to provide non-experts with an overview of the problem content and the existing optimisation procedures used in projection pursuit guided tour.
-For those who are already familiar with the techniques, feel free to skip this section.
-
-The optimisation problem we're interested in is in the context of projection pursuit.
-Coined by @friedman1974projection, projection pursuit is a method that detects the interesting structure (i.e.
-clustering, outliers and skewness) of multivariate data via projecting it in lower dimensions.
-Let $\mathbf{X}_{n \times p}$ be a data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
-Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and can be used to measure the "interestingness" of the projection.
-A number of index functions have been proposed in the literature including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
-
- As @friedman1974projection noted "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." A suitable optimisation procedure is needed to find the projection angle that maximises the PPI and the quality of the optimisation largely affect the interesting projections one could possibly observe.
-
-Projection pursuit is usually used in conjunction with a tour method called *guided tour*.
-Tour explores the multivariate data *interactively* via playing a series of projections, that form a *tour path* and guided tour uses the path that is geodesically the shortest.
-Details of the mathematical construction of a tour path can be found in @buja2005computational.
-[Maybe add the sketch for illustrating tour path.] Guided tour, along with other types of tour, has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+The optimisation problem we're interested in is in the context of projection pursuit guided tour.
+Coined by @friedman1974projection, projection pursuit detects interesting structures (i.e.
+clustering, outliers and skewness) in multivariate data via projecting it in lower dimensions.
+Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
+Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of the projection.
+A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+
+In their initial paper, @friedman1974projection noted that "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find bases that give interesting projections.
+While we leave the formal construction the optimisation problem and existing algorithms in section \ref{tour-optim}, we want to emphasizes here that given a random starting (current) basis, projection pursuit repetitively searches for candidate bases nearby and outputs a target basis once it finds one with higher index value than the current basis.
+The target basis becomes the current basis in the next iteration and the searching continues until no better basis can be found or the termination criteria are reached.
+
+A tour method, guided tour is usually used in conjunction with projection pursuit to interactively explore multivariate data.
+Tour plays an animation on the projected data which is constructed using the bases outputted by projection pursuit and their interpolations.
+Guided tour constructs the interpolation using the shortest geodesic path between two outputted bases and we refer readers to @buja2005computational for details of the mathematical construction of the interpolation basis  Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by projection pursuit and the white frames, which interpolates two blue frames.
+The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 
@@
 ```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_increasing.png"))
-```
-
-## Optimisation problem and existing algorithms
+knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.jpeg"))
+```
+
+## Optimisation problem and existing algorithms {#tour-optim}
 
 Now we begin to formulate the optimisation problem.
-Given a randomly generated starting basis $\mathbf{A}^0$, projection pursuit finds the final projection basis $\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i$ is a $p \times 1$ vector, satisfies the following optimisation problem:
+Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i \in \mathbb{R}^{p}$, satisfies the following optimisation problem:
 
@@
 
-via an iterative search of target basis: $\{\mathbf{A}^0, \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}$.
-
-There are three existing methods for optimising PPI function and we review them below.
-
-@posse1995projection proposed a stochastic direct search method, a random search algorithm.
-Given a current basis $\mathbf{A}_{\text{cur}}$, a candidate basis $\mathbf{A}_{t}$ is sampled in the neighbourhood defined by the radius of the p-dimensional sphere, $\alpha$ of the current basis $\mathbf{A}_{t}$ by $\mathbf{A}_{t} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$.
-If the candidate basis has a higher index value than the current basis, it is outputted as the target basis $\mathbf{A}_{t}$ along with other metadata and the current iteration stops.
-If no basis is found to have higher index value after the maximum number of tries $l_{\max}$, the algorithm stops with nothing outputted.
-$c$ is the halfing parameter and when $c > 30$, the neighbourhood parameter $\alpha$ in the next iteration will be reduced by half.
-The algorithm for a random search is summarised in Algorithm \ref{random-search}.
+There are three existing algorithms.
+
+@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{rand}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path.
+A new round of search continues to find a better basis after the interpolation finishes.
+The halfing parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
+When the search needs to sample more than $c$ number of basis to find an outputted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
+If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops.
+The algorithm is summarised in Algorithm \ref{random-search} for one iteration.
 
@@
     \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
-    \Output{The target basis: $\mathbf{A}_{t+1}$}
+    \Output{The target basis: $\mathbf{A}_{l}$}
   initialisation\;
-  Set $l = 1$\;
+  Set $l = 1$ and $c = 0$\;
   \While{$l < l_{\max}$}{
@@
 
-@cook1995grand explained the use of a gradient ascent optimisation with the assumption that the index function is continuous and differentiable.
-Since some indices could be non-differentiable, the computation of derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
-Taking a step on the straight derivative direction has been modified to maximise the projection pursuit index along the geodesic direction.
-See Algorithm \ref{search-geodesic} for a summarisation of the steps.
+@cook1995grand explained the use of a gradient ascent optimisation with an assumption that the index function is continuous and differentiable.
+Since some indices could be non-differentiable, the derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
+Once the best direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic between the current basis $\mathbf{A}_{\text{cur}}$ and the best directional basis $\mathbf{A}_{*}$.
+If the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value, $\mathbf{A}_{**}$ is outputted for the current iteration to construct the interpolation or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates.
+Algorithm \ref{search-geodesic} summarise the steps in geodesic search.
 
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(t)$.
-Given an initial $T_0$, the temperature at iteration $t$ is defined as $T(t) = \frac{T_0}{log(t + 1)}$.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$.
+Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}$.
-If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis.
-If it is smaller, the candidate is accepted with probability $P= \min\left\{\exp\left[\frac{I_{t+1} - I_{t}}{T(t)}\right],1\right\}$ where $I_{t}$ and $I_{t+1}$ are the index value of the current and candidate basis respectively.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is sampled by $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}$.
+If the index value of the candidate basis is larger than the one of the current basis, the candidate basis is outputted.
+While if it is smaller, the candidate has another chance of being accepted with probability $P= \min\left\{\exp\left[\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}$ where $I_{\cdot}$ denotes the index value of a given basis.
 The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
-    \Output{The global object; The target basis: $\mathbf{A}^{j}$}
+    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$; Initial termperature: $T_0$}
+    \Output{The global object; The target basis: $\mathbf{A}_{l}$}
   initialisation\;
+  Set $l = 1$\;
   \While{$l < l_{\max}$}{
-    Set $l = 0$\;
     Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
@@
 
-Below listed several features characterise the optimisation procedures needed in projection pursuit
-
-- *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.
-
-- *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.
-
-- *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
+Notice that the three algorithms above constitute one search. The outputted basis is then sent to construct the interpolation and one such search-and-interpolate step is one round.
+
+<!-- Below listed several features characterise the optimisation procedures needed in projection pursuit -->
+
+<!-- - *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.  -->
+
+<!-- - *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.  -->
+
+<!-- - *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
+-->
 
@@
 
-Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit problem we have aims at *exploring* the data and thus is interested in how the algorithm finds its maximum.
+Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit explores the data with an aim to discover the potential underlying structure and thus is interested in how the algorithm finds its maximum.
 This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
@@
 
-Ggplot requires a data frame that contains all the variables to plot and a *global object* is constructed as the data frame supplied to the visual diagnostic plots to better suit the characters of iterative optimisation algorithms.
-Given an optimisation algorithm, two primary variables of interest are the *decision variable: $\mathbf{A}$* and the *value of the objective function: $I$*, which is the projection basis and index value repsectively in projection pursuit.
-*State: $S$* is another variable of great importance since it labels $\mathbf{A}$ and $I$ by the stage they are in the whole search-and-interpolate process.
-
-*Iterators* indexes the data collected and prescribes the the natural ordering of each observation in different levels.
-*$t$* is a unique identifier that has an increment of one for each observation and *$j$* is the counter for each search-and-interpolate round.
-[add a bit more for *$l$* after decide whether to combine $l$ and $k$.
-]
+Tidy data, as introduced in [@wickham2014tidy], has three criteria:  1) each observation forms a row, 2) each variable forms a column and 3) each type of observational unit forms a table.
+The adoption of tidy data makes it easier for data wrangling and visualisation.
+For our visual diagnostics, global object is a such data construction with key components in optimisation algorithms and is designed for making diagnostic plots easier.
+In the optimisation algorithms for projection pursuit, two key elements are the projection basis: $\mathbf{A}$ and the index value: $I$.
+State: $S$ labels the observation with detailed stage of searching or interpolation.
+
+Multiple iterators are needed to index the data collected at different levels.
+$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same for within one round and has an increment of one once a new round starts.
+$l$ is the counter for each search/interpolation allowing us to know how many basis has the algorithm searched before finding one to output.
 
@@
 There are other parameters that are of our interest and we denote them as *$V_{p}$*.
-In projection pursuit, this includes $V_1 = \text{method}$, denoting the name of the searching method used and $V_2 = \text{alpha}$, which is a neighbourhood parameter that controls the size in sampling candidate bases.
-A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
-The full data structure can thus be shown as in Equation \ref{eq:data-structure}.
+In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, which is the neighbourhood parameter that controls the size in sampling candidate bases.
+The data structure can thus be shown as in Equation \ref{eq:data-structure}.
 
@@
 \left[
-\begin{array}{c|ccc|ccc|ccc}
-t & \mathbf{A} & I & S & j &  l  & k & V_{1}& \cdots & V_{p}\\
-\hline
-1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & \ast & V_{1, 1} & \cdots & V_{p, 1}\\
-\hline
-2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1 & \ast & V_{1, 2} & \cdots & V_{p, 2}\\
-3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2 & \ast & V_{1, 3} & \cdots & V_{p, 3}\\
-\vdots & \vdots &\vdots &\vdots & \vdots &\vdots & \vdots &\vdots &\cdots &\vdots\\
-\vdots & \vdots & \vdots &\vdots & 2 & l_1 & \ast & \vdots &  \cdots & \vdots\\
-\hline
-\vdots &\vdots & \vdots &\vdots & 2 & \ast & 1& \vdots &  \cdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots & 2 & \ast & 2& \vdots& \cdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-\vdots &\vdots &\vdots &\vdots & 2 & \ast & k_1&\vdots & \cdots & \vdots\\
-\hline
-\vdots &\vdots &\vdots &\vdots &\vdots  &\vdots & \vdots &\vdots &\cdots &\vdots \\
-T & \mathbf{A}_T & I_T &S_T  & J &  l_J & \ast & V_{1, T}& \cdots & V_{p, T}\\
+\begin{array}{c|ccc|cc|cc}
+t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
+\hline
+1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{21}\\
+\hline
+2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{12}  & V_{22}\\
+3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{13}  & V_{23}\\
+\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
+\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
+\hline
+\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
+\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
+\hline
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+\hline
+\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{1T}& V_{2T}\\
+\hline
+\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
+\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
+\hline
+\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{1T^\prime}& V_{2T^\prime}\\
+\end{array}
+\right]
+= 
+\left[
+\begin{array}{c}
+\text{column name} \\
+\hline
+\text{search (start basis)} \\
+\hline
+\text{search} \\
+\text{search} \\
+\vdots \\
+\text{search (outputed basis)} \\
+\hline
+\text{interpolate} \\
+\text{interpolate} \\
+\vdots \\
+\text{interpolate} \\
+\hline
+\vdots \\
+\hline
+\text{search} \\
+\vdots \\
+\text{search (final basis)} \\
+\hline
+\text{interpolate} \\
+\vdots \\
+\text{interpolate} \\
+\hline
+\text{search (no output)} \\
+\vdots \\
+\text{search (no output)} \\
 \end{array}
@@
 \end{equation}
+
+where $T^{\prime} = T + k_{J}+ l_{J+1}$.
+Note that we deliberately denote the last round of search as $j = J+1$ and in that round there is no output/interpolation basis and the algorithm terminates.
+This notation allows us to denote the last complete search-and-interpolate round as round $J$ and hence the final basis is $A_T$ and highest index value found is $I_T$.
+
+[outside the paper: I find the notation of current/target basis is confusing because the target basis in round $j$ becomes the current basis in round $j+1$.
+Also, when we start to have polish, the target basis may not be the current basis in the next round...
+The place where current/target is most appropriate is probably when describing the interpolation where the first one is always the current basis and the last is always the target basis.
+I think it is better to leave this language in the code]
+
+A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
+
+[I feel this sketch was initially useful but now since we have the data matrix and the printed output of the global object, it doesn't any additional information.
+I'm still keeping it here but we may need to remove it if it's not useful :(]
 
@@
 
-We first simulate some random variables of size 1000 with different structures.
+We simulate some random variables of size 1000 with different structures.
 `x1`, `x8`, `x9` and `x10` are simulated from normal distribution with zero mean and variance of one as in equation \ref{eq:sim-norm}.
 When using projection pursuit to explore the data structure based on its departure from normality, the entry in the projection basis for these variables should be close to zero in theory.
-`x2` to `x7` are mixture of normal distribution with different weights and locations.
-Equation \ref{eq:sim-x2} to \ref{eq:sim-x7} show the distribution of each variables and Figure \ref{origin-data} shows the histogram of each variable except `x3`.
-All the variables are then scaled to ensure the mixture has variance of one.
+`x2` to `x7` are mixture of normal distributions with different weights and locations.
+Equation \ref{eq:sim-x2} to \ref{eq:sim-x7} outlines the distribution where each variable is simulated from and Figure \ref{origin-data} shows the histogram of each variable except `x3`.
+All the variables are then scaled to have variance of one before running the projection pursuit.
 
@@
 
-We form our first dataset using variable `x1`, `x2`, `x8`, `x9` and `x10` and supply it to one of the above algorithm, say `search_geodesic` in the `tourr` package.
+We form our first dataset using variable `x1`, `x2`, `x8`, `x9` and `x10` and run the guided tour with optimiser `search_better`.
 When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
-Additional messages during the optimisation can be displayed by `verbose = TRUE`.
-Below shows the first five rows of the global object.
-The global object meets the definition of tidy data introduced in [@wickham2014tidy] where each observation forms a row, each variable forms a column and each type of observational unit forms a table.
-The adoption of tidy data makes it easier for data wrangling and visualisation, which powers the visual diagnostics system introduced in this paper.
+Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.
+Below shows the first ten rows of the global object.
+Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.
+
+[I notice that in the implementation the loop index for interpolation starts from 0 rather than 1.
+Need more investigation to see if it is easy to change it to 1 or we will have to change the definition of the data structure.]
 
@@
 ```{r glb-obj-1}
-print(holes_1d_better %>% head(5))
+holes_1d_better %>% 
+  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
+  head(10)
 ```
@@
 
-# Visual diagnostics
+# Visual diagnostic plots {#vis-diag-plots}
 
@@
   - x: $j$ is mapped to the x-axis
-  - y: the statistical transformed index value: $Q_q(I_{T} \cdot \mathds{1} (S = \text{search}))$ is mapped to the y-axis where $q$ takes one of $0, 25, 50, 75, 100$ and $Q_q(x)$ finds the qth-quantile for the vector $x$.
-$\mathds{1}(.)$ is the identity operator.
+  - y: the statistical transformed index value: $Q_q(I^{\prime}_t)$ is mapped to the y-axis where $Q_q(x)$ finds the $0, 25, 50, 75, 100$th-quantile for the vector $x$ and $I^{\prime}_t$ is the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
 
@@
 
-# Implementation: Ferrn package
+# Implementation: Ferrn package {#implementation}
 

### 2020-07-29 paper/paper.tex
modify based on meeting 19 and polish section 3 and 4 (except 4.1)

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, dsfont, bbm} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 
-The paper is organised as follows. Section \ref{DFO} gives a general
-literature review of optimisation, specifically derivative free
-optimisation. Section \ref{vis-diag} presents the new idea of
-constructing a systematic visual framework that diagnoses the components
-of an optimisation procedure (parameters, searching path, etc). The rest
-of the paper serves as a comprehensive example of using the visual
-diagnostics on one particular problem: \emph{projection pursuit guided
-tour}. Some background knowledge of projection pursuit guided tour is
-provided in Section \ref{tour}. Section \ref{apply} applies the concepts
-proposed in section \ref{vis-diag} in the tour problem and sets up the
-data. The last section, Section \ref{plots}, presents the visual
-diagnostic plots and explains how they can help to understand different
-aspects of the optimisation in tour.
+The remainder of the paper is organised as follows. Section \ref{DFO}
+gives a general overview of optimisation methods, specifically
+derivative free optimisation. Section \ref{tour} introduces the
+optimisation problem in a specific problem: projection pursuit guided
+tour and existing derivative free algorithms for solving it. Section
+\ref{vis-diag} and \ref{vis-diag-plots} together present the new
+construction of a visual diagnostic framework that unveil the blackbox
+of optimisation. Section \ref{vis-diag} focues on motivational ideas and
+data structure and Section \ref{vis-diag-plots} provides definitions of
+different diagnostic plots using grammar of graphic and examples.
+Finally, \ref{implementation} mentions the R package: ferrn that
+implements the visual diagnostics described above.
 
@@
 
+\newpage
+
 \hypertarget{tour}{%
@@
 
-From Section \ref{tour}, we presents a comprehensive case study on how
-to use the visual diagnostics to explore the optimisation in a specific
-problem: projection pursuit guided tour. Section \ref{tour} aims to
-provide non-experts with an overview of the problem content and the
-existing optimisation procedures used in projection pursuit guided tour.
-For those who are already familiar with the techniques, feel free to
-skip this section.
-
 The optimisation problem we're interested in is in the context of
-projection pursuit. Coined by \citet{friedman1974projection}, projection
-pursuit is a method that detects the interesting structure
-(i.e.~clustering, outliers and skewness) of multivariate data via
-projecting it in lower dimensions. Let \(\mathbf{X}_{n \times p}\) be a
-data matrix, an n-d projection can be seen as a linear transformation
-\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
+projection pursuit guided tour. Coined by
+\citet{friedman1974projection}, projection pursuit detects interesting
+structures (i.e.~clustering, outliers and skewness) in multivariate data
+via projecting it in lower dimensions. Let \(\mathbf{X}_{n \times p}\)
+be the data matrix, an n-d projection can be seen as a linear
+transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
 \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
@@
 \(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
-that maps the projection basis \(\mathbf{A}\) onto a scalar value \(I\),
-this function is commonly known as the projection pursuit index (PPI)
-function, or the index function and can be used to measure the
+that maps the projection basis \(\mathbf{A}\) onto a scalar index value
+\(I\), this function is commonly known as the projection pursuit index
+(PPI) function, or the index function and is used to measure the
 ``interestingness'' of the projection. A number of index functions have
-been proposed in the literature including Legendre index
-\citep{friedman1974projection}, Hermite index
+been proposed in the literature to detect different data structures,
+including Legendre index \citep{friedman1974projection}, Hermite index
 \citep{hall1989polynomial}, natural Hermite index
@@
 
-As \citet{friedman1974projection} noted ``\ldots{}, the technique use
-for maximising the projection index strongly influences both the
-statistical and the computational aspects of the procedure.'' A suitable
-optimisation procedure is needed to find the projection angle that
-maximises the PPI and the quality of the optimisation largely affect the
-interesting projections one could possibly observe.
-
-Projection pursuit is usually used in conjunction with a tour method
-called \emph{guided tour}. Tour explores the multivariate data
-\emph{interactively} via playing a series of projections, that form a
-\emph{tour path} and guided tour uses the path that is geodesically the
-shortest. Details of the mathematical construction of a tour path can be
-found in \citet{buja2005computational}. {[}Maybe add the sketch for
-illustrating tour path.{]} Guided tour, along with other types of tour,
-has been implemented in the \emph{tourr} package in R, available on the
-Comprehensive R Archive Network at
+In their initial paper, \citet{friedman1974projection} noted that
+``\ldots{}, the technique use for maximising the projection index
+strongly influences both the statistical and the computational aspects
+of the procedure.'' Hence, effective optimisation algorithms are
+necessary for projection pursuit to find bases that give interesting
+projections. While we leave the formal construction the optimisation
+problem and existing algorithms in section \ref{tour-optim}, we want to
+emphasizes here that given a random starting (current) basis, projection
+pursuit repetitively searches for candidate bases nearby and outputs a
+target basis once it finds one with higher index value than the current
+basis. The target basis becomes the current basis in the next iteration
+and the searching continues until no better basis can be found or the
+termination criteria are reached.
+
+A tour method, guided tour is usually used in conjunction with
+projection pursuit to interactively explore multivariate data. Tour
+plays an animation on the projected data which is constructed using the
+bases outputted by projection pursuit and their interpolations. Guided
+tour constructs the interpolation using the shortest geodesic path
+between two outputted bases and we refer readers to
+\citet{buja2005computational} for details of the mathematical
+construction of the interpolation basis Figure \ref{tour-path} shows a
+sketch of the tour path consisting of the blue frames produced by
+projection pursuit and the white frames, which interpolates two blue
+frames. The tour method has been implemented in the \emph{tourr} package
+in R, available on the Comprehensive R Archive Network at
 \url{https://cran.r-project.org/web/packages/tourr/}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_increasing} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
 \end{figure}
 
-\hypertarget{optimisation-problem-and-existing-algorithms}{%
+\hypertarget{tour-optim}{%
 \subsection{Optimisation problem and existing
-algorithms}\label{optimisation-problem-and-existing-algorithms}}
+algorithms}\label{tour-optim}}
 
 Now we begin to formulate the optimisation problem. Given a randomly
-generated starting basis \(\mathbf{A}^0\), projection pursuit finds the
+generated starting basis \(\mathbf{A}_1\), projection pursuit finds the
 final projection basis
-\(\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
-\(\mathbf{a}_i\) is a \(p \times 1\) vector, satisfies the following
+\(\mathbf{A}_T = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
+\(\mathbf{a}_i \in \mathbb{R}^{p}\), satisfies the following
 optimisation problem:
@@
 
-via an iterative search of target basis:
-\(\{\mathbf{A}^0, \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
-
-There are three existing methods for optimising PPI function and we
-review them below.
-
-\citet{posse1995projection} proposed a stochastic direct search method,
-a random search algorithm. Given a current basis
-\(\mathbf{A}_{\text{cur}}\), a candidate basis \(\mathbf{A}_{t}\) is
-sampled in the neighbourhood defined by the radius of the p-dimensional
-sphere, \(\alpha\) of the current basis \(\mathbf{A}_{t}\) by
-\(\mathbf{A}_{t} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}\).
-If the candidate basis has a higher index value than the current basis,
-it is outputted as the target basis \(\mathbf{A}_{t}\) along with other
-metadata and the current iteration stops. If no basis is found to have
-higher index value after the maximum number of tries \(l_{\max}\), the
-algorithm stops with nothing outputted. \(c\) is the halfing parameter
-and when \(c > 30\), the neighbourhood parameter \(\alpha\) in the next
-iteration will be reduced by half. The algorithm for a random search is
-summarised in Algorithm \ref{random-search}.
+There are three existing algorithms.
+
+\citet{posse1995projection} proposed a random search algorithm that
+samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the
+current basis \(\mathbf{A}_{\text{cur}}\) by
+\(\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}\),
+where \(\alpha\) controls the radius of the sampling neighbourhood and
+\(\mathbf{A}_{rand}\) is a randomly generated matrix with the same
+dimension as \(\mathbf{A}_{\text{cur}}\). The optimiser keeps sampling
+bases near the current basis until it finds one with higher index value
+than the current basis and then outputs it for guided tour to construct
+the interpolation path. A new round of search continues to find a better
+basis after the interpolation finishes. The halfing parameter \(c\) with
+default value of 30 is designed to adjust the searching neighbourhood
+\(\alpha\). When the search needs to sample more than \(c\) number of
+basis to find an outputted basis, the neighbourhood parameter \(\alpha\)
+will be reduced by half in the next iteration. If the optimiser can't
+find a better basis within the maximum number of tries \(l_{\max}\), the
+algorithm stops. The algorithm is summarised in Algorithm
+\ref{random-search} for one iteration.
 
@@
 \citet{cook1995grand} explained the use of a gradient ascent
-optimisation with the assumption that the index function is continuous
+optimisation with an assumption that the index function is continuous
 and differentiable. Since some indices could be non-differentiable, the
-computation of derivative is replaced by a pseudo-derivative of
-evaluating five randomly generated directions in a tiny nearby
-neighbourhood. Taking a step on the straight derivative direction has
-been modified to maximise the projection pursuit index along the
-geodesic direction. See Algorithm \ref{search-geodesic} for a
-summarisation of the steps.
+derivative is replaced by a pseudo-derivative of evaluating five
+randomly generated directions in a tiny nearby neighbourhood. Once the
+best direction is determined, the optimiser finds the best projection
+\(\mathbf{A}_{**}\) on the geodesic between the current basis
+\(\mathbf{A}_{\text{cur}}\) and the best directional basis
+\(\mathbf{A}_{*}\). If the percentage change in the index value between
+\(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a
+threshold value, \(\mathbf{A}_{**}\) is outputted for the current
+iteration to construct the interpolation or the algorithm repeats the
+above steps until \(l_{\max}\) is reached and the searching terminates.
+Algorithm \ref{search-geodesic} summarise the steps in geodesic search.
 
@@
 non-derivative procedure based on a non-increasing cooling scheme
-\(T(t)\). Given an initial \(T_0\), the temperature at iteration \(t\)
-is defined as \(T(t) = \frac{T_0}{log(t + 1)}\). The simulated annealing
-algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
-and a randomly generated orthonormal basis \(A_{rand}\), a candidate
-basis is constructed as
+\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
+is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). The simulated
+annealing algorithm works as follows. Given a neighbourhood parameter
+\(\alpha\) and a randomly generated orthonormal basis \(A_{rand}\), a
+candidate basis is sampled by
 \(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}\).
 If the index value of the candidate basis is larger than the one of the
-current basis, the candidate basis becomes the target basis. If it is
-smaller, the candidate is accepted with probability
-\(P= \min\left\{\exp\left[\frac{I_{t+1} - I_{t}}{T(t)}\right],1\right\}\)
-where \(I_{t}\) and \(I_{t+1}\) are the index value of the current and
-candidate basis respectively. The algorithm can be summarised as in
-Algorithm \ref{simulated_annealing}.
+current basis, the candidate basis is outputted. While if it is smaller,
+the candidate has another chance of being accepted with probability
+\(P= \min\left\{\exp\left[\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}\)
+where \(I_{\cdot}\) denotes the index value of a given basis. The
+algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
 
@@
 
-Below listed several features characterise the optimisation procedures
-needed in projection pursuit
-
-\begin{itemize}
-\item
-  \emph{Being able to handle non-differentiable PPI function}: The PPI
-  function could be non-differentiable, thus derivative free methods are
-  preferred.
-\item
-  \emph{Being able to optimise with constraints}: The constraint comes
-  from projection matrix being an orthonormal matrix.
-\item
-  \emph{Being able to find both local and global maximum}: Although the
-  primary interest is to find the global maximum, local maximum could
-  also reveal structures in the data that are of our interest.
-\end{itemize}
+Notice that the three algorithms above constitute one search. The
+outputted basis is then sent to construct the interpolation and one such
+search-and-interpolate step is one round.
 
@@
 Random search methods has a black-box mechanism and focuses solely on
-finding the global maximum point, while the projection pursuit problem
-we have aims at \emph{exploring} the data and thus is interested in how
-the algorithm finds its maximum. This motivates us \textbf{to develop a
-visual diagnostic system for exploring the optimisation searching path}.
+finding the global maximum point, while the projection pursuit explores
+the data with an aim to discover the potential underlying structure and
+thus is interested in how the algorithm finds its maximum. This
+motivates us \textbf{to develop a visual diagnostic system for exploring
+the optimisation searching path}.
 
@@
 
-Ggplot requires a data frame that contains all the variables to plot and
-a \emph{global object} is constructed as the data frame supplied to the
-visual diagnostic plots to better suit the characters of iterative
-optimisation algorithms. Given an optimisation algorithm, two primary
-variables of interest are the \emph{decision variable: \(\mathbf{A}\)}
-and the \emph{value of the objective function: \(I\)}, which is the
-projection basis and index value repsectively in projection pursuit.
-\emph{State: \(S\)} is another variable of great importance since it
-labels \(\mathbf{A}\) and \(I\) by the stage they are in the whole
-search-and-interpolate process.
-
-\emph{Iterators} indexes the data collected and prescribes the the
-natural ordering of each observation in different levels. \emph{\(t\)}
-is a unique identifier that has an increment of one for each observation
-and \emph{\(j\)} is the counter for each search-and-interpolate round.
-{[}add a bit more for \emph{\(l\)} after decide whether to combine \(l\)
-and \(k\). {]}
+Tidy data, as introduced in \citep{wickham2014tidy}, has three criteria:
+1) each observation forms a row, 2) each variable forms a column and 3)
+each type of observational unit forms a table. The adoption of tidy data
+makes it easier for data wrangling and visualisation. For our visual
+diagnostics, global object is a such data construction with key
+components in optimisation algorithms and is designed for making
+diagnostic plots easier. In the optimisation algorithms for projection
+pursuit, two key elements are the projection basis: \(\mathbf{A}\) and
+the index value: \(I\). State: \(S\) labels the observation with
+detailed stage of searching or interpolation.
+
+Multiple iterators are needed to index the data collected at different
+levels. \(t\) is a unique identifier that prescribes the natural
+ordering of each observation; \(j\) is the counter for each
+search-and-interpolate round, which remains the same for within one
+round and has an increment of one once a new round starts. \(l\) is the
+counter for each search/interpolation allowing us to know how many basis
+has the algorithm searched before finding one to output.
 
@@
 as \emph{\(V_{p}\)}. In projection pursuit, this includes
-\(V_1 = \text{method}\), denoting the name of the searching method used
-and \(V_2 = \text{alpha}\), which is a neighbourhood parameter that
-controls the size in sampling candidate bases. A sketch of the global
-object for projection pursuit guided tour is presented in Figure
-\ref{fig:glb-obj}. The full data structure can thus be shown as in
-Equation \ref{eq:data-structure}.
+\(V_1 = \text{method}\), which tags the name of the algorithm used and
+\(V_2 = \text{alpha}\), which is the neighbourhood parameter that
+controls the size in sampling candidate bases. The data structure can
+thus be shown as in Equation \ref{eq:data-structure}.
 
@@
 
+where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that we deliberately
+denote the last round of search as \(j = J+1\) and in that round there
+is no output/interpolation basis and the algorithm terminates. This
+notation allows us to denote the last complete search-and-interpolate
+round as round \(J\) and hence the final basis is \(A_T\) and highest
+index value found is \(I_T\).
+
+{[}outside the paper: I find the notation of current/target basis is
+confusing because the target basis in round \(j\) becomes the current
+basis in round \(j+1\). Also, when we start to have polish, the target
+basis may not be the current basis in the next round\ldots{} The place
+where current/target is most appropriate is probably when describing the
+interpolation where the first one is always the current basis and the
+last is always the target basis. I think it is better to leave this
+language in the code{]}
+
+A sketch of the global object for projection pursuit guided tour is
+presented in Figure \ref{fig:glb-obj}.
+
+{[}I feel this sketch was initially useful but now since we have the
+data matrix and the printed output of the global object, it doesn't any
+additional information. I'm still keeping it here but we may need to
+remove it if it's not useful :({]}
+
 \begin{figure}
@@
 
-We first simulate some random variables of size 1000 with different
+We simulate some random variables of size 1000 with different
 structures. \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are
@@
 the projection basis for these variables should be close to zero in
-theory. \texttt{x2} to \texttt{x7} are mixture of normal distribution
+theory. \texttt{x2} to \texttt{x7} are mixture of normal distributions
 with different weights and locations. Equation \ref{eq:sim-x2} to
-\ref{eq:sim-x7} show the distribution of each variables and Figure
-\ref{origin-data} shows the histogram of each variable except
-\texttt{x3}. All the variables are then scaled to ensure the mixture has
-variance of one.
+\ref{eq:sim-x7} outlines the distribution where each variable is
+simulated from and Figure \ref{origin-data} shows the histogram of each
+variable except \texttt{x3}. All the variables are then scaled to have
+variance of one before running the projection pursuit.
 
@@
 We form our first dataset using variable \texttt{x1}, \texttt{x2},
-\texttt{x8}, \texttt{x9} and \texttt{x10} and supply it to one of the
-above algorithm, say \texttt{search\_geodesic} in the \texttt{tourr}
-package. When the optimisation ends, the global object will be stored
-and printed (it can be turned off by supplying argument
-\texttt{print\ =\ FALSE}). Additional messages during the optimisation
-can be displayed by \texttt{verbose\ =\ TRUE}. Below shows the first
-five rows of the global object. The global object meets the definition
-of tidy data introduced in \citep{wickham2014tidy} where each
-observation forms a row, each variable forms a column and each type of
-observational unit forms a table. The adoption of tidy data makes it
-easier for data wrangling and visualisation, which powers the visual
-diagnostics system introduced in this paper.
+\texttt{x8}, \texttt{x9} and \texttt{x10} and run the guided tour with
+optimiser \texttt{search\_better}. When the optimisation ends, the
+global object will be stored and printed (it can be turned off by
+supplying argument \texttt{print\ =\ FALSE}). Additional messages during
+the optimisation can be displayed by argument \texttt{verbose\ =\ TRUE}.
+Below shows the first ten rows of the global object. Notice that the
+tibble object allows the list-column \texttt{basis} to be printed out
+nicely with the dimension of the projection basis readily available.
+
+{[}I notice that in the implementation the loop index for interpolation
+starts from 0 rather than 1. Need more investigation to see if it is
+easy to change it to 1 or we will have to change the definition of the
+data structure.{]}
 
@@
 \begin{verbatim}
-## # A tibble: 5 x 8
-##   basis             index_val tries info           loop method       alpha    id
-##   <list>                <dbl> <dbl> <chr>         <dbl> <chr>        <dbl> <int>
-## 1 <dbl[,1] [5 x 1]>     0.749     1 start            NA <NA>           0.5     1
-## 2 <dbl[,1] [5 x 1]>     0.730     1 random_search     1 search_bett~   0.5     2
-## 3 <dbl[,1] [5 x 1]>     0.743     1 random_search     2 search_bett~   0.5     3
-## 4 <dbl[,1] [5 x 1]>     0.736     1 random_search     3 search_bett~   0.5     4
-## 5 <dbl[,1] [5 x 1]>     0.747     1 random_search     4 search_bett~   0.5     5
+## # A tibble: 10 x 8
+##       id basis             index_val info         tries  loop method       alpha
+##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
+##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 <NA>           0.5
+##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
+##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
+##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
+##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
+##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
+##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
+##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     0 search_bett~  NA  
+##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     1 search_bett~  NA  
+## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA
 \end{verbatim}
 
-\hypertarget{visual-diagnostics}{%
-\section{Visual diagnostics}\label{visual-diagnostics}}
+\hypertarget{vis-diag-plots}{%
+\section{Visual diagnostic plots}\label{vis-diag-plots}}
 
@@
   \item
-    y: the statistical transformed index value:
-    \(Q_q(I_{T} \cdot \mathds{1} (S = \text{search}))\) is mapped to the
-    y-axis where \(q\) takes one of \(0, 25, 50, 75, 100\) and
-    \(Q_q(x)\) finds the qth-quantile for the vector \(x\).
-    \(\mathds{1}(.)\) is the identity operator.
+    y: the statistical transformed index value: \(Q_q(I^{\prime}_t)\) is
+    mapped to the y-axis where \(Q_q(x)\) finds the
+    \(0, 25, 50, 75, 100\)th-quantile for the vector \(x\) and
+    \(I^{\prime}_t\) is the index value of all the searching bases
+    defined in Matrix \ref{eq:data-structure}.
   \end{itemize}
@@
 
-\hypertarget{implementation-ferrn-package}{%
-\section{Implementation: Ferrn
-package}\label{implementation-ferrn-package}}
+\hypertarget{implementation}{%
+\section{Implementation: Ferrn package}\label{implementation}}
 

### 2020-07-29 paper/paper.Rmd
modifying 4.1 done

@@
 
+\newpage
+
 # Visual diagnostic system {#vis-diag}
-
-[I would expand this section more as the core contribution. ppp]
 
@@
 
-<!-- Visualisation has been widely used for exploring and understanding data. Presenting information in a graphical manner often allows people to see information they would otherwise not seen. -->
-
-Random search methods has a black-box mechanism and focuses solely on finding the global maximum point, while the projection pursuit explores the data with an aim to discover the potential underlying structure and thus is interested in how the algorithm finds its maximum.
-This motivates us **to develop a visual diagnostic system for exploring the optimisation searching path**.
-
-The necessity of developing such a system rather than simply producing different diagnostic plots is because the diagnostics of each variable requires a different function and these functions can't be scaled to other problems.
-Thus we want to establish a set of rules that can generalise the diagnostic of iterative algorithms.
-
-The idea of generalising all the diagnostic plots under one framework is inspired by the concept of *grammar of graphic*[@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+We can see from Figure \ref{tour-path} that only the data projected using the outputted bases and interpolated bases are animated by tour; none of the searching bases are presented in teh tour animation.
+The optimisation algorithms in projection pursuit, although has been demonstrated to find the global maximum in various literature we referred above, focuses solely on finding the global maximum and remains as black-box machanisms.
+There has been fewer tools available to explore how each algorithm iteratively finds its maximum and compare how different algorithms searches the parameter space.
+Visualisation has been widely used for exploring and understanding data.
+Presenting information in a graphical manner often allows people to see information they would otherwise not seen.
+This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with a aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed, only an appropriate optimisation algorithm allows us to find the projection basis that presents interesting underlying structure in the data**.
+
+The idea of generalised framwork for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
 In grammar of graphic, plots are not defined by its appearance (i.e.
 boxplot, histogram, scatter plot etc) but by "stacked layers".
-By this design, ggplot doesn't need to develop a gazillion of functions that each produces a different type of plot.
-Instead, it aesthetically maps the variables to the geometric objects and builds the plot through different layers.
+Using this design, ggplot doesn't have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
+Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
+There are different ways to represent the same data in rectangular form.
+Certain cleaning steps are needed, before using ggplot2, to bring the data into a tidy data format [@wickham2014tidy] where 1) each observation forms a row; 2) each variable forms a column and 3) each type of observational unit forms a table.
+In a tidy format, data wrangling and visualisation are greatly simplified.
+
+Hence the study **contributes to set up a tidy data structure for collecting observations from optimisation algorithms and prescribe a set of rules for generalising diagnostic plots for optimisation algorithms, specifically those used for projection pursuit.**
 
@@
 
-Tidy data, as introduced in [@wickham2014tidy], has three criteria:  1) each observation forms a row, 2) each variable forms a column and 3) each type of observational unit forms a table.
-The adoption of tidy data makes it easier for data wrangling and visualisation.
  For our visual diagnostics, global object is a such data construction with key components in optimisation algorithms and is designed for making diagnostic plots easier.

### 2020-07-29 paper/paper.tex
modifying 4.1 done

@@
 
+\newpage
+
 \hypertarget{vis-diag}{%
@@
 
-{[}I would expand this section more as the core contribution. ppp{]}
-
 \hypertarget{motivation}{%
@@
 
-Random search methods has a black-box mechanism and focuses solely on
-finding the global maximum point, while the projection pursuit explores
-the data with an aim to discover the potential underlying structure and
-thus is interested in how the algorithm finds its maximum. This
-motivates us \textbf{to develop a visual diagnostic system for exploring
-the optimisation searching path}.
-
-The necessity of developing such a system rather than simply producing
-different diagnostic plots is because the diagnostics of each variable
-requires a different function and these functions can't be scaled to
-other problems. Thus we want to establish a set of rules that can
-generalise the diagnostic of iterative algorithms.
-
-The idea of generalising all the diagnostic plots under one framework is
-inspired by the concept of \emph{grammar of
-graphic}\citep{wickham2010layered}, which powers the primary graphical
-system in R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are
-not defined by its appearance (i.e.~boxplot, histogram, scatter plot
-etc) but by ``stacked layers''. By this design, ggplot doesn't need to
-develop a gazillion of functions that each produces a different type of
-plot. Instead, it aesthetically maps the variables to the geometric
-objects and builds the plot through different layers.
+We can see from Figure \ref{tour-path} that only the data projected
+using the outputted bases and interpolated bases are animated by tour;
+none of the searching bases are presented in teh tour animation. The
+optimisation algorithms in projection pursuit, although has been
+demonstrated to find the global maximum in various literature we
+referred above, focuses solely on finding the global maximum and remains
+as black-box machanisms. There has been fewer tools available to explore
+how each algorithm iteratively finds its maximum and compare how
+different algorithms searches the parameter space. Visualisation has
+been widely used for exploring and understanding data. Presenting
+information in a graphical manner often allows people to see information
+they would otherwise not seen. This motivates us \textbf{to develop a
+visual framework to diagnose optimisation algorithms in projection
+pursuit with a aim to understand and compare different existing
+algorithms.} This study is meaningful because \textbf{as more complex
+index functions are proposed, only an appropriate optimisation algorithm
+allows us to find the projection basis that presents interesting
+underlying structure in the data}.
+
+The idea of generalised framwork for diagnostic plots is inspired by the
+concept of grammar of graphic \citep{wickham2010layered}, which powers
+the primary graphical system in R, ggplot2 \citep{ggplot2}. In grammar
+of graphic, plots are not defined by its appearance (i.e.~boxplot,
+histogram, scatter plot etc) but by ``stacked layers''. Using this
+design, ggplot doesn't have to develop a gazillion of functions that
+each produces a different type of plot from a different data structure.
+Instead, it aesthetically maps variables (and its statistical
+transformation) in a dataset to different geometric objects (points,
+lines, box-and-whisker etc) and builds the plot through overlaying
+different layers. There are different ways to represent the same data in
+rectangular form. Certain cleaning steps are needed, before using
+ggplot2, to bring the data into a tidy data format
+\citep{wickham2014tidy} where 1) each observation forms a row; 2) each
+variable forms a column and 3) each type of observational unit forms a
+table. In a tidy format, data wrangling and visualisation are greatly
+simplified.
+
+Hence the study \textbf{contributes to set up a tidy data structure for
+collecting observations from optimisation algorithms and prescribe a set
+of rules for generalising diagnostic plots for optimisation algorithms,
+specifically those used for projection pursuit.}
 
@@
 
-Tidy data, as introduced in \citep{wickham2014tidy}, has three criteria:
-1) each observation forms a row, 2) each variable forms a column and 3)
-each type of observational unit forms a table. The adoption of tidy data
-makes it easier for data wrangling and visualisation. For our visual
-diagnostics, global object is a such data construction with key
-components in optimisation algorithms and is designed for making
-diagnostic plots easier. In the optimisation algorithms for projection
-pursuit, two key elements are the projection basis: \(\mathbf{A}\) and
-the index value: \(I\). State: \(S\) labels the observation with
-detailed stage of searching or interpolation.
+For our visual diagnostics, global object is a such data construction
+with key components in optimisation algorithms and is designed for
+making diagnostic plots easier. In the optimisation algorithms for
+projection pursuit, two key elements are the projection basis:
+\(\mathbf{A}\) and the index value: \(I\). State: \(S\) labels the
+observation with detailed stage of searching or interpolation.
 

### 2020-07-29 paper/paper.Rmd
spelling check

@@
 Section \ref{tour} introduces the optimisation problem in a specific problem: projection pursuit guided tour and existing derivative free algorithms for solving it. 
-Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveil the blackbox of optimisation.
-Section \ref{vis-diag} focues on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples.
+Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveil the black-box of optimisation.
+Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples.
 Finally, \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$; Initial termperature: $T_0$}
+    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$; Initial temperature: $T_0$}
     \Output{The global object; The target basis: $\mathbf{A}_{l}$}
@@
 
-We can see from Figure \ref{tour-path} that only the data projected using the outputted bases and interpolated bases are animated by tour; none of the searching bases are presented in teh tour animation.
-The optimisation algorithms in projection pursuit, although has been demonstrated to find the global maximum in various literature we referred above, focuses solely on finding the global maximum and remains as black-box machanisms.
-There has been fewer tools available to explore how each algorithm iteratively finds its maximum and compare how different algorithms searches the parameter space.
+We can see from Figure \ref{tour-path} that only the data projected using the outputted bases and interpolated bases are animated by tour; none of the searching bases are presented in the tour animation.
+The optimisation algorithms in projection pursuit, although has been demonstrated to find the global maximum in various literature we referred above, focuses solely on finding the global maximum and remains as black-box mechanisms There has been fewer tools available to explore how each algorithm iteratively finds its maximum and compare how different algorithms searches the parameter space.
 Visualisation has been widely used for exploring and understanding data.
@@
 
-The idea of generalised framwork for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+The idea of generalised framework for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
 In grammar of graphic, plots are not defined by its appearance (i.e.
@@
 \vdots \\
-\text{search (outputed basis)} \\
+\text{search (outputted basis)} \\
 \hline

### 2020-07-30 paper/paper.Rmd
another read-through

@@
 
-The optimisation problem we're interested in is in the context of projection pursuit guided tour.
 Coined by @friedman1974projection, projection pursuit detects interesting structures (i.e.
-clustering, outliers and skewness) in multivariate data via projecting it in lower dimensions.
+clustering, outliers and skewness) in multivariate data via projecting onto lower dimensions.
 Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
-Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto a scalar index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of the projection.
+Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
 A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 
-In their initial paper, @friedman1974projection noted that "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find bases that give interesting projections.
-While we leave the formal construction the optimisation problem and existing algorithms in section \ref{tour-optim}, we want to emphasizes here that given a random starting (current) basis, projection pursuit repetitively searches for candidate bases nearby and outputs a target basis once it finds one with higher index value than the current basis.
-The target basis becomes the current basis in the next iteration and the searching continues until no better basis can be found or the termination criteria are reached.
+In their initial paper, @friedman1974projection noted that "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
+While we leave the formal construction of the optimisation problem and existing algorithms in section \ref{tour-optim}, we outlines the general idea here.
+Given a random starting (current) basis, projection pursuit repetitively searches for candidate bases nearby until it finds one with higher index value than the current basis.
+In the second round, that basis becomes the current basis and the repetitive sampling  continues.
+The process ends until no better basis can be found or one of the termination criteria is reached.
 
@@
 Tour plays an animation on the projected data which is constructed using the bases outputted by projection pursuit and their interpolations.
-Guided tour constructs the interpolation using the shortest geodesic path between two outputted bases and we refer readers to @buja2005computational for details of the mathematical construction of the interpolation basis  Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by projection pursuit and the white frames, which interpolates two blue frames.
+Guided tour constructs the interpolation using the shortest geodesic path between two outputted bases and we refer readers to @buja2005computational for the mathematical details.
+Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm and the white frames, which are the interpolations between two blue frames.
 The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
@@
 
-We can see from Figure \ref{tour-path} that only the data projected using the outputted bases and interpolated bases are animated by tour; none of the searching bases are presented in the tour animation.
-The optimisation algorithms in projection pursuit, although has been demonstrated to find the global maximum in various literature we referred above, focuses solely on finding the global maximum and remains as black-box mechanisms There has been fewer tools available to explore how each algorithm iteratively finds its maximum and compare how different algorithms searches the parameter space.
-Visualisation has been widely used for exploring and understanding data.
-Presenting information in a graphical manner often allows people to see information they would otherwise not seen.
-This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with a aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed, only an appropriate optimisation algorithm allows us to find the projection basis that presents interesting underlying structure in the data**.
+We can see from Figure \ref{tour-path} that only the data projected using the outputted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
+Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.
+
+Visualisation has been widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to see information they would otherwise not seen.
+This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with a aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
 
@@
 Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
+
 There are different ways to represent the same data in rectangular form.
@@
 
- For our visual diagnostics, global object is a such data construction with key components in optimisation algorithms and is designed for making diagnostic plots easier.
-In the optimisation algorithms for projection pursuit, two key elements are the projection basis: $\mathbf{A}$ and the index value: $I$.
-State: $S$ labels the observation with detailed stage of searching or interpolation.
+Global object is a data construction with key components from optimisation algorithms and is designed for making diagnostic plots easier.
+In the optimisation algorithms for projection pursuit, three key elements are 1) projection bases: $\mathbf{A}$, 2) index values: $I$ and 3) State: $S$, which labels the observation with detailed stage of searching or interpolation.
 
 Multiple iterators are needed to index the data collected at different levels.
-$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same for within one round and has an increment of one once a new round starts.
-$l$ is the counter for each search/interpolation allowing us to know how many basis has the algorithm searched before finding one to output.
+$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts.
+$l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
 
@@
 There are other parameters that are of our interest and we denote them as *$V_{p}$*.
-In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, which is the neighbourhood parameter that controls the size in sampling candidate bases.
+In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
 The data structure can thus be shown as in Equation \ref{eq:data-structure}.

### 2020-07-30 paper/paper.tex
another read-through

@@
 \ref{vis-diag} and \ref{vis-diag-plots} together present the new
-construction of a visual diagnostic framework that unveil the blackbox
-of optimisation. Section \ref{vis-diag} focues on motivational ideas and
-data structure and Section \ref{vis-diag-plots} provides definitions of
-different diagnostic plots using grammar of graphic and examples.
+construction of a visual diagnostic framework that unveil the black-box
+of optimisation. Section \ref{vis-diag} focuses on motivational ideas
+and data structure and Section \ref{vis-diag-plots} provides definitions
+of different diagnostic plots using grammar of graphic and examples.
 Finally, \ref{implementation} mentions the R package: ferrn that
@@
 
-The optimisation problem we're interested in is in the context of
-projection pursuit guided tour. Coined by
-\citet{friedman1974projection}, projection pursuit detects interesting
-structures (i.e.~clustering, outliers and skewness) in multivariate data
-via projecting it in lower dimensions. Let \(\mathbf{X}_{n \times p}\)
-be the data matrix, an n-d projection can be seen as a linear
-transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
-\(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
+Coined by \citet{friedman1974projection}, projection pursuit detects
+interesting structures (i.e.~clustering, outliers and skewness) in
+multivariate data via projecting onto lower dimensions. Let
+\(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be
+seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\)
+defined by \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{P}_{n \times d}\) is the projected data and
@@
 \(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
-that maps the projection basis \(\mathbf{A}\) onto a scalar index value
-\(I\), this function is commonly known as the projection pursuit index
-(PPI) function, or the index function and is used to measure the
-``interestingness'' of the projection. A number of index functions have
+that maps the projection basis \(\mathbf{A}\) onto an index value \(I\),
+this function is commonly known as the projection pursuit index (PPI)
+function, or the index function and is used to measure the
+``interestingness'' of a projection. A number of index functions have
 been proposed in the literature to detect different data structures,
@@
 of the procedure.'' Hence, effective optimisation algorithms are
-necessary for projection pursuit to find bases that give interesting
-projections. While we leave the formal construction the optimisation
-problem and existing algorithms in section \ref{tour-optim}, we want to
-emphasizes here that given a random starting (current) basis, projection
-pursuit repetitively searches for candidate bases nearby and outputs a
-target basis once it finds one with higher index value than the current
-basis. The target basis becomes the current basis in the next iteration
-and the searching continues until no better basis can be found or the
-termination criteria are reached.
+necessary for projection pursuit to find the bases that give interesting
+projections. While we leave the formal construction of the optimisation
+problem and existing algorithms in section \ref{tour-optim}, we outlines
+the general idea here. Given a random starting (current) basis,
+projection pursuit repetitively searches for candidate bases nearby
+until it finds one with higher index value than the current basis. In
+the second round, that basis becomes the current basis and the
+repetitive sampling continues. The process ends until no better basis
+can be found or one of the termination criteria is reached.
 
@@
 between two outputted bases and we refer readers to
-\citet{buja2005computational} for details of the mathematical
-construction of the interpolation basis Figure \ref{tour-path} shows a
-sketch of the tour path consisting of the blue frames produced by
-projection pursuit and the white frames, which interpolates two blue
-frames. The tour method has been implemented in the \emph{tourr} package
-in R, available on the Comprehensive R Archive Network at
+\citet{buja2005computational} for the mathematical details. Figure
+\ref{tour-path} shows a sketch of the tour path consisting of the blue
+frames produced by the projection pursuit optimisation algorithm and the
+white frames, which are the interpolations between two blue frames. The
+tour method has been implemented in the \emph{tourr} package in R,
+available on the Comprehensive R Archive Network at
 \url{https://cran.r-project.org/web/packages/tourr/}
@@
 We can see from Figure \ref{tour-path} that only the data projected
-using the outputted bases and interpolated bases are animated by tour;
-none of the searching bases are presented in teh tour animation. The
-optimisation algorithms in projection pursuit, although has been
-demonstrated to find the global maximum in various literature we
-referred above, focuses solely on finding the global maximum and remains
-as black-box machanisms. There has been fewer tools available to explore
-how each algorithm iteratively finds its maximum and compare how
-different algorithms searches the parameter space. Visualisation has
-been widely used for exploring and understanding data. Presenting
-information in a graphical manner often allows people to see information
-they would otherwise not seen. This motivates us \textbf{to develop a
-visual framework to diagnose optimisation algorithms in projection
-pursuit with a aim to understand and compare different existing
-algorithms.} This study is meaningful because \textbf{as more complex
-index functions are proposed, only an appropriate optimisation algorithm
-allows us to find the projection basis that presents interesting
-underlying structure in the data}.
-
-The idea of generalised framwork for diagnostic plots is inspired by the
-concept of grammar of graphic \citep{wickham2010layered}, which powers
-the primary graphical system in R, ggplot2 \citep{ggplot2}. In grammar
-of graphic, plots are not defined by its appearance (i.e.~boxplot,
-histogram, scatter plot etc) but by ``stacked layers''. Using this
-design, ggplot doesn't have to develop a gazillion of functions that
-each produces a different type of plot from a different data structure.
-Instead, it aesthetically maps variables (and its statistical
-transformation) in a dataset to different geometric objects (points,
-lines, box-and-whisker etc) and builds the plot through overlaying
-different layers. There are different ways to represent the same data in
-rectangular form. Certain cleaning steps are needed, before using
-ggplot2, to bring the data into a tidy data format
-\citep{wickham2014tidy} where 1) each observation forms a row; 2) each
-variable forms a column and 3) each type of observational unit forms a
-table. In a tidy format, data wrangling and visualisation are greatly
-simplified.
+using the outputted and interpolated bases are animated by tour; none of
+the searching bases sampled by the optimsiation algorithms are presented
+in the tour animation. Although these algorithms have been demonstrated
+to find the global maximum in various literature as we cited above, they
+focus solely on finding the global maximum and fewer tools are available
+to explore how each algorithm searches the parameter space and compare
+the results between different algorithms.
+
+Visualisation has been widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to see
+information they would otherwise not seen. This motivates us \textbf{to
+develop a visual framework to diagnose optimisation algorithms in
+projection pursuit with a aim to understand and compare different
+existing algorithms.} This study is meaningful because \textbf{as more
+complex index functions are proposed based on statistical theory, only
+computationally effective optimisation algorithms allow us to find the
+projection basis that presents interesting structure in the data}.
+
+The idea of generalised framework for diagnostic plots is inspired by
+the concept of grammar of graphic \citep{wickham2010layered}, which
+powers the primary graphical system in R, ggplot2 \citep{ggplot2}. In
+grammar of graphic, plots are not defined by its appearance
+(i.e.~boxplot, histogram, scatter plot etc) but by ``stacked layers''.
+Using this design, ggplot doesn't have to develop a gazillion of
+functions that each produces a different type of plot from a different
+data structure. Instead, it aesthetically maps variables (and its
+statistical transformation) in a dataset to different geometric objects
+(points, lines, box-and-whisker etc) and builds the plot through
+overlaying different layers.
+
+There are different ways to represent the same data in rectangular form.
+Certain cleaning steps are needed, before using ggplot2, to bring the
+data into a tidy data format \citep{wickham2014tidy} where 1) each
+observation forms a row; 2) each variable forms a column and 3) each
+type of observational unit forms a table. In a tidy format, data
+wrangling and visualisation are greatly simplified.
 
@@
 
-For our visual diagnostics, global object is a such data construction
-with key components in optimisation algorithms and is designed for
-making diagnostic plots easier. In the optimisation algorithms for
-projection pursuit, two key elements are the projection basis:
-\(\mathbf{A}\) and the index value: \(I\). State: \(S\) labels the
-observation with detailed stage of searching or interpolation.
+Global object is a data construction with key components from
+optimisation algorithms and is designed for making diagnostic plots
+easier. In the optimisation algorithms for projection pursuit, three key
+elements are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\)
+and 3) State: \(S\), which labels the observation with detailed stage of
+searching or interpolation.
 
@@
 ordering of each observation; \(j\) is the counter for each
-search-and-interpolate round, which remains the same for within one
-round and has an increment of one once a new round starts. \(l\) is the
+search-and-interpolate round, which remains the same within one round
+and has an increment of one once a new round starts. \(l\) is the
 counter for each search/interpolation allowing us to know how many basis
-has the algorithm searched before finding one to output.
+the algorithm has searched before finding one to output.
 
@@
 \(V_1 = \text{method}\), which tags the name of the algorithm used and
-\(V_2 = \text{alpha}\), which is the neighbourhood parameter that
-controls the size in sampling candidate bases. The data structure can
-thus be shown as in Equation \ref{eq:data-structure}.
+\(V_2 = \text{alpha}\), the neighbourhood parameter that controls the
+size in sampling candidate bases. The data structure can thus be shown
+as in Equation \ref{eq:data-structure}.
 

### 2020-08-02 paper/paper.Rmd
use standard quantile function notation

@@
   - x: $j$ is mapped to the x-axis
-  - y: the statistical transformed index value: $Q_q(I^{\prime}_t)$ is mapped to the y-axis where $Q_q(x)$ finds the $0, 25, 50, 75, 100$th-quantile for the vector $x$ and $I^{\prime}_t$ is the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
+  - y: the statistical transformed index value: $Q_{I^{\prime}_t}(q)$ is mapped to the y-axis where $Q_X(q)$, $q = 0, 25, 50, 75, 100$ finds the qth-quantile of $X$ and $I^{\prime}_t$ denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
 

### 2020-08-02 paper/paper.tex
use standard quantile function notation

@@
   \item
-    y: the statistical transformed index value: \(Q_q(I^{\prime}_t)\) is
-    mapped to the y-axis where \(Q_q(x)\) finds the
-    \(0, 25, 50, 75, 100\)th-quantile for the vector \(x\) and
-    \(I^{\prime}_t\) is the index value of all the searching bases
-    defined in Matrix \ref{eq:data-structure}.
+    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\)
+    is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\)
+    finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the
+    index value of all the searching bases defined in Matrix
+    \ref{eq:data-structure}.
   \end{itemize}

### 2020-08-04 paper/paper.tex
update paper.tex
minor corrections

@@
 \begin{abstract}
-Friedman \& Tukey commented on their intial paper on projection pursuit
-in 1974 that ``the technique use for maximising the projection index
+Friedman \& Tukey commented on their initial paper on projection pursuit
+in 1974 that ``the technique used for maximising the projection index
 strongly influences both the statistical and the computational aspects
@@
 learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied to more general to help
-practitioner to unveil the black-box in randomised iteartive
+This diagnostic system can be applied more generally to help
+practitioner to unveil the black-box in randomised iterative
 (optimisation) algorithms. An R package, ferrn, has been created to
@@
 the space of all feasible solutions which typically is represented by a
-set of constraints. The problem consists on optimizing an objective
+set of constraints. The problem consists in optimizing an objective
 function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
@@
 
-Derivative free methods provides another option to optimise the
+Derivative free methods provide another option to optimise the
 objective function without evaluating any gradient or Hessian
-information and a particular class of methods: direct search, has gained
-its popularity through its conceptual simplicity. However, the whole
+information and a particular class of methods, direct search, has gained
+popularity through its conceptual simplicity. However, the whole
 searching process in the algorithm remains a black-box. Plots are
@@
 with 30 levels will not make the plot informative. Thus the plot needs
-to be re-designed to better suits the characteristics of the parameter
+to be re-designed to better suit the characteristics of the parameter
 (whether the parameter is a scalar or a matrix? whether the parameter is
@@
 of different diagnostic plots using grammar of graphic and examples.
-Finally, \ref{implementation} mentions the R package: ferrn that
+Finally, Section \ref{implementation} mentions the R package: ferrn that
 implements the visual diagnostics described above.
@@
 gained its popularity through its simplicity in use and reliability in
-solving complicated practical problems. Depends on whether a random
+solving complicated practical problems. Depending on whether a random
 sample is used in the search, this class of methods can be further
@@
 algorithms discussed below? Or we are focusing on the optimisation
-problem only within the project pursuit context? Some of the problems
+problem only within the projection pursuit context? Some of the problems
 listed below are also applicable to optimization problem in general too.
@@
 
-Below listed several issues in projection pursuit optimisation. Some are
-general optimisation problems, while others are more specific for PP
+Below are listed several issues in projection pursuit optimisation. Some are
+general optimisation problems, while others are more specific to PP
 optimisation.
@@
   complex or the number of decision variables is large. Also, there are
-  discussions on how to avoid getting trapped in a local optimal in the
+  discussions on how to avoid getting trapped in a local optimum in the
   literature.
@@
   \emph{computation speed}: The optimisation procedure needs to be fast
-  to compute since tours produces real-time animation of the projected
+  to compute since tours produce real-time animation of the projected
   data.
@@
   algorithm, researchers usually set a seed to ensure the algorithm
-  producing the same result for every run. This practice supports
+  produces the same result for every run. This practice supports
   reproducibility, while less efforts has been made to guarantee
@@
   decision variable includes all the entries in the projection matrix,
-  which is high-dimensional. Researcher would be better off if they can
+  which is high-dimensional. Researchers would be better off if they could
   understand the relative position of different projection matrix in the
@@
   procedure usually involves iteratively finding projection bases that
-  maximises the index function, while tour requires geodesic
+  maximise the index function, while tour requires geodesic
   interpolation between these bases to produce a continuous view for the
@@
 \(\mathbf{A}_{p\times d}\) is the projection basis. Define
-\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) be an index function
+\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index function
 that maps the projection basis \(\mathbf{A}\) onto an index value \(I\),
@@
 In their initial paper, \citet{friedman1974projection} noted that
-``\ldots{}, the technique use for maximising the projection index
+``\ldots{}, the technique used for maximising the projection index
 strongly influences both the statistical and the computational aspects
@@
 projections. While we leave the formal construction of the optimisation
-problem and existing algorithms in section \ref{tour-optim}, we outlines
+problem and existing algorithms to section \ref{tour-optim}, we outline
 the general idea here. Given a random starting (current) basis,
-projection pursuit repetitively searches for candidate bases nearby
+projection pursuit repeatedly searches for candidate bases nearby
 until it finds one with higher index value than the current basis. In
@@
 the interpolation path. A new round of search continues to find a better
-basis after the interpolation finishes. The halfing parameter \(c\) with
+basis after the interpolation finishes. The halving parameter \(c\) with
 default value of 30 is designed to adjust the searching neighbourhood
@@
 annealing algorithm works as follows. Given a neighbourhood parameter
-\(\alpha\) and a randomly generated orthonormal basis \(A_{rand}\), a
+\(\alpha\) and a randomly generated orthonormal basis \(\mathbf{A}_{\text{rand}}\), a
 candidate basis is sampled by
-\(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}\).
+\(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{\text{rand}}\).
 If the index value of the candidate basis is larger than the one of the
@@
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
+    Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
     Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
@@
         Compute $P= \min\left\{\exp\left[\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-        Draw $D$ from a uniform distribution: $D \sim \text{Unif(0, 1)}$\;
-        \If{$P > D$}{
+        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+        \If{$P > U$}{
            \KwRet{$\mathbf{A}_{l}$} \;
@@
 Presenting information in a graphical format often allows people to see
-information they would otherwise not seen. This motivates us \textbf{to
+information they would otherwise not see. This motivates us \textbf{to
 develop a visual framework to diagnose optimisation algorithms in
-projection pursuit with a aim to understand and compare different
+projection pursuit with the aim to understand and compare different
 existing algorithms.} This study is meaningful because \textbf{as more
@@
 (i.e.~boxplot, histogram, scatter plot etc) but by ``stacked layers''.
-Using this design, ggplot doesn't have to develop a gazillion of
+Using this design, ggplot does not have to develop a gazillion of
 functions that each produces a different type of plot from a different
@@
 simulated from normal distribution with zero mean and variance of one as
-in equation \ref{eq:sim-norm}. When using projection pursuit to explore
+in equation \eqref{eq:sim-norm}. When using projection pursuit to explore
 the data structure based on its departure from normality, the entry in
@@
 theory. \texttt{x2} to \texttt{x7} are mixture of normal distributions
-with different weights and locations. Equation \ref{eq:sim-x2} to
-\ref{eq:sim-x7} outlines the distribution where each variable is
+with different weights and locations. Equation \eqref{eq:sim-x2} to
+\eqref{eq:sim-x7} outlines the distribution where each variable is
 simulated from and Figure \ref{origin-data} shows the histogram of each
 variable except \texttt{x3}. All the variables are then scaled to have
-variance of one before running the projection pursuit.
+unit variance before running the projection pursuit.
 
@@
 
-We form our first dataset using variable \texttt{x1}, \texttt{x2},
+We form our first dataset using variables \texttt{x1}, \texttt{x2},
 \texttt{x8}, \texttt{x9} and \texttt{x10} and run the guided tour with
@@
 A primary interest of diagnosing an optimisation algorithm is to study
-how it finds its optimal progressively. We could plot the index value
+how it finds its optimum progressively. We could plot the index value
 across its natural ordering, however, different iterations may have
@@
 Boxplot is a suitable candidate that provides five points summary of the
-data, however it has one drawback: it doesn't report the number of point
+data, however it has one drawback: it does not report the number of points
 in each box. We may risk losing information on how many points it takes
@@
 with small number of points to a point geometry, which is achieved via
-an \texttt{cutoff} argument. A line geometry is also added to link the
+a \texttt{cutoff} argument. A line geometry is also added to link the
 points with the largest index value in each iteration. This helps to
@@
 
-We choose variable \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
+We choose variables \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
 \texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
@@
 \texttt{max.tries\ =\ 500} is used. The index value of the searching
-points are shown in Figure \ref{points}. Label at the bottom indicates
+points are shown in Figure \ref{points}. The label at the bottom indicates
 the number of observations in each iteration and facilitates the choice
@@
 learn that the \texttt{search\_better} quickly finds better projection
-basis with higher index value at first and the takes longer to find a
+basis with higher index value at first and then takes longer to find a
 better one later.
@@
 \caption{\label{points}A comparison of plotting the same search points
-with different plot designs. The left plot doesn't efficiently use the
-plot space to convey information from the plot while the right plot
+with different plot designs. The left plot does not use the
+plot space efficiently to convey information from the plot while the right plot
 provides good summarisation of data and number of points in each tries.}
@@
 
-Sometimes, rather than explore the searching points, we may be
-interested in explore the points on the interpolation path (target and
+Sometimes, rather than exploring the searching points, we may be
+interested in exploring the points on the interpolation path (target and
 interpolating points) since these points will be played by the tour
@@
 function \texttt{search\_better} and we want to learn how the index
-value changes on the interpolation path for \texttt{holes} index. From
+value changes on the interpolation path for the \texttt{holes} index. From
 the left panel of Figure \ref{interruption}, we observe that when
@@
 accept the interpolating bases only up to the one with the largest index
-value. After implementing this interruption, the search finds higher
+value. After implementing this interruption, the search finds a higher
 final index value with fewer steps as shown in the right panel of Figure
@@
 \caption{\label{interruption}Trace plots of the interpolated basis with
-and without the interruption. The interruption stops the interpolation
+and without interruption. The interruption stops the interpolation
 when the index value starts to decrease at id = 60. The implementation
@@
 \item
-  the number of iteration can't exceed the \texttt{max.tries}
+  the number of iteration can't exceed \texttt{max.tries}
 \end{enumerate}
@@
 
-Slight variation of plot definition due to the addition of polishing
+A slight variation of the plot definition due to the addition of polishing
 points is as follows:
@@
 Both attain the same final index value after the polishing while using a
-max.tries 25 is not sufficient to find the ture maximum.}
+\texttt{max.tries} of 25 is not sufficient to find the true maximum.}
 \end{figure}
@@
 starting and target bases are then interpolated. In the next iteration,
-the target basis becomes the current basis and then procedures
+the target basis becomes the current basis and then the procedures
 continues.
@@
 \emph{Example: initial value for polishing alpha}
-\texttt{search\_polish} is a brutal-force algorithm that evaluate 1000
+\texttt{search\_polish} is a brute-force algorithm that evaluate 1000
 points in the neighbourhood at each loop. Setting an appropriate initial
@@
 \[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
-
 can be seen as a function of the projection matrix
@@
 \texttt{x2} and \texttt{x3} and remain 1D projection. In this case, two
-local maximum appear when projection matrix being \([0, 1, 0, 0, 0, 0]\)
+local maxima appear with projection matrix being \([0, 1, 0, 0, 0, 0]\)
 and \([0, 0, 1 ,0, 0, 0]\).
@@
 Using different seeds in \texttt{search\_better} allows us to find both
-local maximum as in Figure \ref{1d-2var-different-seeds}. Comparing the
+local maxima as in Figure \ref{1d-2var-different-seeds}. Comparing the
 maximum of both, we can see that the global maximum happens when
@@
 \texttt{search\_better\_random} manages to do it via a
-metropolis-hasting random search as shown in Figure
+Metropolis-Hasting random search as shown in Figure
 \ref{1d-2var-better-random}, although at a higher cost of number of
@@
 
-We can also plot the searching points of all the three algorithms in the
+We can also plot the searching points of all three algorithms in the
 searching space and explore their relative position against each other
@@
 a more exhaustive search. The large overlapping of better1 and
-better\_random is explained by the fact that both algorithms finds x2 in
+better\_random is explained by the fact that both algorithms find x2 in
 the end.

### 2020-08-06 paper/paper.Rmd
move changes made by Nicolas to Rmd

@@
 - R
-abstract: Friedman & Tukey commented on their intial paper on projection pursuit in 1974 that "the  technique  use  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
+abstract: Friedman & Tukey commented on their initial paper on projection pursuit in 1974 that "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
 In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied to more general to help practitioner to unveil the black-box in randomised iteartive (optimisation) algorithms.
+This diagnostic system can be applied to more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
 An R package, ferrn, has been created to implement this diagnostic system.
@@
 In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
-The problem consists on optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function....
+The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function....
 Gradient based optimization has been typically used to solve such problems.
@@
 
-Derivative free methods provides another option to optimise the objective function without evaluating any gradient or Hessian information and a particular class of methods: direct search, has gained its popularity through its conceptual simplicity.
+Derivative free methods provide another option to optimise the objective function without evaluating any gradient or Hessian information and a particular class of methods, direct search, has gained popularity through its conceptual simplicity.
 However, the whole searching process in the algorithm remains a black-box.
@@
 For example, a categorical variable with 5 levels can be easily mapped onto colour while mapping another categorical variable with 30 levels will not make the plot informative.
-Thus the plot needs to be re-designed to better suits the characteristics of the parameter (whether the parameter is a scalar or a matrix?
+Thus the plot needs to be re-designed to better suit the characteristics of the parameter (whether the parameter is a scalar or a matrix?
 whether the parameter is quantitative or categorical?
@@
 Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples.
-Finally, \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
+Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
 
@@
 Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems.
-Depends on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.
+Depending on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.
 The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.
@@
 [Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
-Or we are focusing on the optimisation problem only within the project pursuit context?
+Or we are focusing on the optimisation problem only within the projection pursuit context?
 Some of the problems listed below are also applicable to optimization problem in general too.
@@
 
-Below listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific for PP optimisation.
+Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.
 
 - *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
-Also, there are discussions on how to avoid getting trapped in a local optimal in the literature.
+Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.
 
@@
 
-- *computation speed*: The optimisation procedure needs to be fast to compute since tours produces real-time animation of the projected data.
-
-- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm producing the same result for every run.
+- *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data.
+
+- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
 This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
@@
 - *high-dimensional decision variable*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
-Researcher would be better off if they can understand the relative position of different projection matrix in the high-dimensional space.
-
-- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximises the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
+Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
+
+- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
 It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
@@
 Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
-Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
+Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
 A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 
-In their initial paper, @friedman1974projection noted that "..., the technique use for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
-While we leave the formal construction of the optimisation problem and existing algorithms in section \ref{tour-optim}, we outlines the general idea here.
-Given a random starting (current) basis, projection pursuit repetitively searches for candidate bases nearby until it finds one with higher index value than the current basis.
+In their initial paper, @friedman1974projection noted that "..., the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
+While we leave the formal construction of the optimisation problem and existing algorithms to section \ref{tour-optim}, we outline the general idea here.
+Given a random starting (current) basis, projection pursuit repeatedly searches for candidate bases nearby until it finds one with higher index value than the current basis.
 In the second round, that basis becomes the current basis and the repetitive sampling  continues.
@@
 A new round of search continues to find a better basis after the interpolation finishes.
-The halfing parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
+The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
 When the search needs to sample more than $c$ number of basis to find an outputted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
@@
 The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is sampled by $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{rand}$.
+Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{\text{rand}}$, a candidate basis is sampled by $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{\text{rand}}$.
 If the index value of the candidate basis is larger than the one of the current basis, the candidate basis is outputted.
@@
 Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see information they would otherwise not seen.
-This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with a aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
+Presenting information in a graphical format often allows people to see information they would otherwise not see.
+This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with the aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
 
@@
 boxplot, histogram, scatter plot etc) but by "stacked layers".
-Using this design, ggplot doesn't have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
+Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
 Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
@@
 Equation \ref{eq:sim-x2} to \ref{eq:sim-x7} outlines the distribution where each variable is simulated from and Figure \ref{origin-data} shows the histogram of each variable except `x3`.
-All the variables are then scaled to have variance of one before running the projection pursuit.
+All the variables are then scaled to have unit variance before running the projection pursuit.
 
@@
 
-We form our first dataset using variable `x1`, `x2`, `x8`, `x9` and `x10` and run the guided tour with optimiser `search_better`.
+We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10` and run the guided tour with optimiser `search_better`.
 When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
@@
 
-A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimal progressively.
+A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimum progressively.
 We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
@@
 
-Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it doesn't report the number of point in each box.
+Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it does not report the number of points in each box.
 We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
-Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via an `cutoff` argument.
+Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument.
 A line geometry is also added to link the points with the largest index value in each iteration.
@@
 
-We choose variable `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
+We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
 Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
 The index value of the searching points are shown in Figure \ref{points}.
-Label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
-We learn that the `search_better` quickly finds better projection basis with higher index value at first and the takes longer to find a better one later.
+The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
+We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.
 
 ```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
-The left plot doesn't efficiently use the plot space to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
+The left plot does not use the plot space efficiently to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
 
@@
 
-Sometimes, rather than explore the searching points, we may be interested in explore the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
+Sometimes, rather than exploring the searching points, we may be interested in exploring the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
 Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed.
@@
 
-We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for `holes` index.
-From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with higher index value than the target basis on the interpolation path.
+We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for the `holes` index.
+From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
 In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
@@
 
-```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without the interruption.
+```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without interruption.
 The interruption stops the interpolation when the index value starts to decrease at id = 60.
@@
 3) the searching neighbourhood can't be too small
-4) the number of iteration can't exceed the `max.tries`
+4) the number of iteration can't exceed `max.tries`
 
@@
 
-Slight variation of plot definition due to the addition of polishing points is as follows:
+A slight variation of the plot definition due to the addition of polishing points is as follows:
 
@@
 ```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
-Both attain the same final index value after the polishing while using a max.tries 25 is not sufficient to find the ture maximum."}
+Both attain the same final index value after the polishing while using a `max.tries` of 25 is not sufficient to find the true maximum."}
 
@@
 The starting and target bases are then interpolated.
-In the next iteration, the target basis becomes the current basis and then procedures continues.
+In the next iteration, the target basis becomes the current basis and then the procedures continues.
 
@@
 
-*Example: initial value for polishing alpha* `search_polish` is a brutal-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
+*Example: initial value for polishing alpha* `search_polish` is a brute-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
 Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value.
@@
 Now we enlarge the dataset to include two informative variables: `x2` and `x3` and remain 1D projection.
-In this case, two local maximum appear when projection matrix being $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
-
-Using different seeds in `search_better` allows us to find both local maximum as in Figure \ref{1d-2var-different-seeds}.
+In this case, two local maxima appear with projection matrix being $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
+
+Using different seeds in `search_better` allows us to find both local maxima d as in Figure \ref{1d-2var-different-seeds}.
 Comparing the maximum of both, we can see that the global maximum happens when `x2` is found.
 It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds?
-`search_better_random` manages to do it via a metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
+`search_better_random` manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
 
@@
 
-We can also plot the searching points of all the three algorithms in the searching space and explore their relative position against each other using principal components.
+We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.
 As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhaustive search.
-The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end.
+The large overlapping of better1 and better_random is explained by the fact that both algorithms find x2 in the end.
 

### 2020-08-06 paper/paper.tex
move changes made by Nicolas to Rmd

@@
 learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied more generally to help
+This diagnostic system can be applied to more generally to help
 practitioner to unveil the black-box in randomised iterative
@@
 
-Derivative free methods provide another option to optimise the
-objective function without evaluating any gradient or Hessian
-information and a particular class of methods, direct search, has gained
-popularity through its conceptual simplicity. However, the whole
-searching process in the algorithm remains a black-box. Plots are
-usually used to evaluate and compare the performance of different
-algorithms but it can easily become tedious because the code will have
-to be modified significantly when comparing different parameters in the
-algorithms. For example, a categorical variable with 5 levels can be
-easily mapped onto colour while mapping another categorical variable
-with 30 levels will not make the plot informative. Thus the plot needs
-to be re-designed to better suit the characteristics of the parameter
-(whether the parameter is a scalar or a matrix? whether the parameter is
-quantitative or categorical? If categorical, how many levels does the
-parameter have?). This motivates the design of a visual diagnostic
-framework for optimisation algorithms based on the idea of a
-\emph{global object}.
+Derivative free methods provide another option to optimise the objective
+function without evaluating any gradient or Hessian information and a
+particular class of methods, direct search, has gained popularity
+through its conceptual simplicity. However, the whole searching process
+in the algorithm remains a black-box. Plots are usually used to evaluate
+and compare the performance of different algorithms but it can easily
+become tedious because the code will have to be modified significantly
+when comparing different parameters in the algorithms. For example, a
+categorical variable with 5 levels can be easily mapped onto colour
+while mapping another categorical variable with 30 levels will not make
+the plot informative. Thus the plot needs to be re-designed to better
+suit the characteristics of the parameter (whether the parameter is a
+scalar or a matrix? whether the parameter is quantitative or
+categorical? If categorical, how many levels does the parameter have?).
+This motivates the design of a visual diagnostic framework for
+optimisation algorithms based on the idea of a \emph{global object}.
 
@@
 
-Below are listed several issues in projection pursuit optimisation. Some are
-general optimisation problems, while others are more specific to PP
+Below are listed several issues in projection pursuit optimisation. Some
+are general optimisation problems, while others are more specific too PP
 optimisation.
@@
   \emph{computation speed}: The optimisation procedure needs to be fast
-  to compute since tours produce real-time animation of the projected
-  data.
+  to compute since the tour produces real-time animation of the
+  projected data.
 \item
@@
   decision variable includes all the entries in the projection matrix,
-  which is high-dimensional. Researchers would be better off if they could
-  understand the relative position of different projection matrix in the
-  high-dimensional space.
+  which is high-dimensional. Researchers would be better off if they
+  could understand the relative position of different projection matrix
+  in the high-dimensional space.
 \item
@@
 \(\mathbf{A}_{p\times d}\) is the projection basis. Define
-\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index function
-that maps the projection basis \(\mathbf{A}\) onto an index value \(I\),
-this function is commonly known as the projection pursuit index (PPI)
-function, or the index function and is used to measure the
+\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index
+function that maps the projection basis \(\mathbf{A}\) onto an index
+value \(I\), this function is commonly known as the projection pursuit
+index (PPI) function, or the index function and is used to measure the
 ``interestingness'' of a projection. A number of index functions have
@@
 the general idea here. Given a random starting (current) basis,
-projection pursuit repeatedly searches for candidate bases nearby
-until it finds one with higher index value than the current basis. In
-the second round, that basis becomes the current basis and the
-repetitive sampling continues. The process ends until no better basis
-can be found or one of the termination criteria is reached.
+projection pursuit repeatedly searches for candidate bases nearby until
+it finds one with higher index value than the current basis. In the
+second round, that basis becomes the current basis and the repetitive
+sampling continues. The process ends until no better basis can be found
+or one of the termination criteria is reached.
 
@@
 annealing algorithm works as follows. Given a neighbourhood parameter
-\(\alpha\) and a randomly generated orthonormal basis \(\mathbf{A}_{\text{rand}}\), a
-candidate basis is sampled by
+\(\alpha\) and a randomly generated orthonormal basis
+\(A_{\text{rand}}\), a candidate basis is sampled by
 \(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{\text{rand}}\).
@@
 simulated from normal distribution with zero mean and variance of one as
-in equation \eqref{eq:sim-norm}. When using projection pursuit to explore
+in equation \ref{eq:sim-norm}. When using projection pursuit to explore
 the data structure based on its departure from normality, the entry in
@@
 theory. \texttt{x2} to \texttt{x7} are mixture of normal distributions
-with different weights and locations. Equation \eqref{eq:sim-x2} to
-\eqref{eq:sim-x7} outlines the distribution where each variable is
+with different weights and locations. Equation \ref{eq:sim-x2} to
+\ref{eq:sim-x7} outlines the distribution where each variable is
 simulated from and Figure \ref{origin-data} shows the histogram of each
@@
 Boxplot is a suitable candidate that provides five points summary of the
-data, however it has one drawback: it does not report the number of points
-in each box. We may risk losing information on how many points it takes
-to find the target basis by displaying the boxplot alone for all
+data, however it has one drawback: it does not report the number of
+points in each box. We may risk losing information on how many points it
+takes to find the target basis by displaying the boxplot alone for all
 \texttt{tries}. Thus, the number of point in each iteration is displayed
 at the bottom of each box and we provide options to switch iteration
-with small number of points to a point geometry, which is achieved via
-a \texttt{cutoff} argument. A line geometry is also added to link the
+with small number of points to a point geometry, which is achieved via a
+\texttt{cutoff} argument. A line geometry is also added to link the
 points with the largest index value in each iteration. This helps to
@@
 \texttt{max.tries\ =\ 500} is used. The index value of the searching
-points are shown in Figure \ref{points}. The label at the bottom indicates
-the number of observations in each iteration and facilitates the choice
-of \texttt{cutoff} argument (by default \texttt{cutoff\ =\ 15}). We
-learn that the \texttt{search\_better} quickly finds better projection
-basis with higher index value at first and then takes longer to find a
-better one later.
+points are shown in Figure \ref{points}. The label at the bottom
+indicates the number of observations in each iteration and facilitates
+the choice of \texttt{cutoff} argument (by default
+\texttt{cutoff\ =\ 15}). We learn that the \texttt{search\_better}
+quickly finds better projection basis with higher index value at first
+and then takes longer to find a better one later.
 
@@
 \caption{\label{points}A comparison of plotting the same search points
-with different plot designs. The left plot does not use the
-plot space efficiently to convey information from the plot while the right plot
+with different plot designs. The left plot does not use the plot space
+efficiently to convey information from the plot while the right plot
 provides good summarisation of data and number of points in each tries.}
@@
 function \texttt{search\_better} and we want to learn how the index
-value changes on the interpolation path for the \texttt{holes} index. From
-the left panel of Figure \ref{interruption}, we observe that when
+value changes on the interpolation path for the \texttt{holes} index.
+From the left panel of Figure \ref{interruption}, we observe that when
 interpolating from the current basis to the target basis, the index
-value may not be monotone: we could reach a basis with higher index
+value may not be monotone: we could reach a basis with a higher index
 value than the target basis on the interpolation path. In this sense, we
@@
 accept the interpolating bases only up to the one with the largest index
-value. After implementing this interruption, the search finds a higher
+value. After implementing this interruption, the search finds higher
 final index value with fewer steps as shown in the right panel of Figure
@@
 \caption{\label{interruption}Trace plots of the interpolated basis with
-and without interruption. The interruption stops the interpolation
-when the index value starts to decrease at id = 60. The implementation
-of the interuption finds an ending basis with higher index value using
-fewer steps.}
+and without interruption. The interruption stops the interpolation when
+the index value starts to decrease at id = 60. The implementation of the
+interuption finds an ending basis with higher index value using fewer
+steps.}
 \end{figure}
@@
 
-A slight variation of the plot definition due to the addition of polishing
-points is as follows:
+A slight variation of the plot definition due to the addition of
+polishing points is as follows:
 
@@
 \[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
+
 can be seen as a function of the projection matrix
@@
 Using different seeds in \texttt{search\_better} allows us to find both
-local maxima as in Figure \ref{1d-2var-different-seeds}. Comparing the
+local maxima d as in Figure \ref{1d-2var-different-seeds}. Comparing the
 maximum of both, we can see that the global maximum happens when
@@
 \texttt{search\_better\_random} manages to do it via a
-Metropolis-Hasting random search as shown in Figure
+Metropolis-hasting random search as shown in Figure
 \ref{1d-2var-better-random}, although at a higher cost of number of

### 2020-08-12 paper/paper.Rmd
write variable as X_{i2} rather than X_{2i}
elaborate Sec. 2 line search more; move difficulties to Sec. 3; elaborate tour & guided tour more

@@
 In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
-The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function....
+The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced space given by the problem constraints to either minimize or maximize a function....
 Gradient based optimization has been typically used to solve such problems.
@@
 The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
-Section \ref{tour} introduces the optimisation problem in a specific problem: projection pursuit guided tour and existing derivative free algorithms for solving it. 
-Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveil the black-box of optimisation.
-Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples.
+Section \ref{tour} introduces the optimisation problem in a specific context: projection pursuit guided tour and existing derivative free algorithms for solving it. 
+Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveils the black-box of optimisation.
+Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} gives definitions of different diagnostic plots using grammar of graphic and examples.
 Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
@@
 
-Given an objective function $f$, one way of optimising it is to equate its gradient to zero.
 In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and  Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
-One common class of methods in DFO is *Direct-search methods*.
-Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems.
-Depending on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.
-The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.
-
-[How about adding more details into derivative free methods? ppp]
-
-<!-- The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which .  -->
-
-<!-- A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability.
-[feel like this sentence can be expanded to include more information.
-xxx].
--->
-
-## Difficulties
-
-[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
-Or we are focusing on the optimisation problem only within the projection pursuit context?
-Some of the problems listed below are also applicable to optimization problem in general too.
-ppp]
-
-Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.
-
-- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
-Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.
-
-- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
-Stochastic optimisation method could be an alternative to solve these problems.
-
-- *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data.
-
-- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
-This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
-
-- *high-dimensional decision variable*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
-Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
-
-- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
-It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
-
-<!-- the interpolation can be seen as giving information on the direction of the derivative. -->
-
-<!-- - is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995).
-Any tools available to  better understand its searching process?
--->
-
-*Think about how does your package help people to understand optimisation*
-
- - diagnostic on stochastic optim
- - vis the progression of multi-parameter decision variable 
- - understanding learning rate - neighbourhood parameter
- - understand where the local & global maximum is found - trace plot - see if noisy function
+A traditional and still reliable in solving complicated practical problems, is *line search method* [@fletcher2013practical].
+In the context of a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
+
+\begin{equation}
+x^{(k + 1)} = x^{(k)} + \alpha_k* d^{(k)}
+\label{eq:line-search}
+\end{equation}
+
+where $d^{(k)}$ is the searching direction in iteration $k$, and $\alpha_k$ is the step-size.
+Strictly speaking, $\alpha_k$ is chosen by another minimisation of $f(x^{(k)} + \alpha* d^{(k)})$ with respect to $\alpha$ and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of $\alpha_k$ is attained [@curry1944method].
+In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation.
+A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(k+1)}> f^{(k)}$ and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
 
@@
 
-Coined by @friedman1974projection, projection pursuit detects interesting structures (i.e.
-clustering, outliers and skewness) in multivariate data via projecting onto lower dimensions.
+Modern development of the line search methods focuses on proposing different computation on the seraching direction: $d^{(k)}$ and various approximations on the step size: $\alpha_k$ catered for practical optimisation problems.
+The specific problem context we are interested in is called projection pursuit guided tour.
+Projection pursuit and guided tour are two separate methods in exporatory data analysis focusing on different aspects: coined by @friedman1974projection, projection pursuit detects interesting structures (i.e.
+clustering, outliers and skewness) in multivariate data via low dimensions projection; whilst guided tour is a particular variation in a broader class of data visualisation method called tour.
+
 Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
@@
 
-A tour method, guided tour is usually used in conjunction with projection pursuit to interactively explore multivariate data.
-Tour plays an animation on the projected data which is constructed using the bases outputted by projection pursuit and their interpolations.
-Guided tour constructs the interpolation using the shortest geodesic path between two outputted bases and we refer readers to @buja2005computational for the mathematical details.
-Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm and the white frames, which are the interpolations between two blue frames.
+Before introducing the guided tour, we shall be familiar with the general tour method [@cook2008grand].
+A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
+The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details.
+Iteratively choose different low dimension planes and interpolate between them forms a tour path.
+Different types of tour methods choose the low dimensional planes differently and we mention two other type of tour that are commonly used.
+A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data.
+Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.
+
+Guided tour chooses the planes produced by optimising the projection pursuit index function.
+Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, which are the interpolations between two blue frames.
 The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 
-<!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis.
-An illustration modified from [@buja2005computational] vividly depicts the tour path in guided tour.
--->
+```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.jpeg"))
+```
 
@@
 
-```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.jpeg"))
-```
-
 ## Optimisation problem and existing algorithms {#tour-optim}
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
+s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
 \end{align}
+
+Where $\delta_{ij}$ is the kronecker delta that takes 1 if $i = j$ and 0 otherwise.
+
+[you wanna give more information about the characteristics of the optimisation problem before introducing existing algorithms ]
+
+---
+
+If the gradient information was available, the decreasing condition is
+
+## Difficulties
+
+(feel this section should go after the general introduction of DFO and PPGT)
+
+[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
+Or we are focusing on the optimisation problem only within the projection pursuit context?
+Some of the problems listed below are also applicable to optimization problem in general too.
+ppp]
+
+Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.
+
+- *constraint optimisation*: our problem requires the project basis to be orthonormal
+
+- *Multivariate optimisation*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
+Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
+(multivariate optimisation)
+
+- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
+Stochastic optimisation method could be an alternative to solve these problems.
+(derivative information is not available)
+
+- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
+Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.
+(the exact $\alpha$ that minimised $f(x^{(k)}+ \alpha d^{(k)})$ can be hard to find or sometiems even impossible, imposing certain decrease, how every this doesn't guarantee global convergence)
+
+Computational considerations:
+
+- *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data.
+
+- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
+This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
+
+- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
+It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
+
+<!-- the interpolation can be seen as giving information on the direction of the derivative. -->
+
+<!-- - is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995).
+Any tools available to  better understand its searching process?
+-->
+
+*Think about how does your package help people to understand optimisation*
+
+ - diagnostic on stochastic optim
+ - vis the progression of multi-parameter decision variable 
+ - understanding learning rate - neighbourhood parameter
+ - understand where the local & global maximum is found - trace plot - see if noisy function
+
+---
 
@@
 \hline
-1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{21}\\
-\hline
-2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{12}  & V_{22}\\
-3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{13}  & V_{23}\\
+1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
+\hline
+2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
+3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{31}  & V_{32}\\
 \vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
@@
 \vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{1T}& V_{2T}\\
+T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
 \hline
@@
 \vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{1T^\prime}& V_{2T^\prime}\\
+T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
 \end{array}

### 2020-08-12 paper/paper.tex
write variable as X_{i2} rather than X_{2i}
elaborate Sec. 2 line search more; move difficulties to Sec. 3; elaborate tour & guided tour more

@@
 function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
-spaced given by the problem constraints to either minimize or maximize a
+space given by the problem constraints to either minimize or maximize a
 function\ldots{}. Gradient based optimization has been typically used to
@@
 derivative free optimisation. Section \ref{tour} introduces the
-optimisation problem in a specific problem: projection pursuit guided
+optimisation problem in a specific context: projection pursuit guided
 tour and existing derivative free algorithms for solving it. Section
 \ref{vis-diag} and \ref{vis-diag-plots} together present the new
-construction of a visual diagnostic framework that unveil the black-box
+construction of a visual diagnostic framework that unveils the black-box
 of optimisation. Section \ref{vis-diag} focuses on motivational ideas
-and data structure and Section \ref{vis-diag-plots} provides definitions
-of different diagnostic plots using grammar of graphic and examples.
+and data structure and Section \ref{vis-diag-plots} gives definitions of
+different diagnostic plots using grammar of graphic and examples.
 Finally, Section \ref{implementation} mentions the R package: ferrn that
@@
 
-Given an objective function \(f\), one way of optimising it is to equate
-its gradient to zero. In modern optimisation problems, gradient
-information can be hard to evaluate or sometimes even impossible and
-Derivative-Free Optimisation (DFO) methods can be useful to approach
-these problems. One common class of methods in DFO is
-\emph{Direct-search methods}. Coined by \citet{hooke1961direct}, direct
-search methods don't require any gradient or Hessian information and has
-gained its popularity through its simplicity in use and reliability in
-solving complicated practical problems. Depending on whether a random
-sample is used in the search, this class of methods can be further
-classified as \emph{stochastic} or \emph{deterministic}. The stochastic
-version of the direct-search methods will be the main optimisation
-procedure analysed in this paper.
-
-{[}How about adding more details into derivative free methods? ppp{]}
-
-\hypertarget{difficulties}{%
-\subsection{Difficulties}\label{difficulties}}
-
-{[}Are we using projection pursuit/guided tour to better understand the
-convergence of optimization algorithms visually in combination with the
-algorithms discussed below? Or we are focusing on the optimisation
-problem only within the projection pursuit context? Some of the problems
-listed below are also applicable to optimization problem in general too.
-ppp{]}
-
-Below are listed several issues in projection pursuit optimisation. Some
-are general optimisation problems, while others are more specific too PP
-optimisation.
-
-\begin{itemize}
-\item
-  \emph{Finding global maximum}: Although finding local maximum is
-  relatively easy with developed algorithms, it is generally hard to
-  guarantee global maximum in a problem where the objective function is
-  complex or the number of decision variables is large. Also, there are
-  discussions on how to avoid getting trapped in a local optimum in the
-  literature.
-\item
-  \emph{optimising non-smooth function}: When the objective function is
-  non-differentiable, derivative information can not be obtained, which
-  means traditional gradient- or Hessian- based methods are not
-  feasible. Stochastic optimisation method could be an alternative to
-  solve these problems.
-\item
-  \emph{computation speed}: The optimisation procedure needs to be fast
-  to compute since the tour produces real-time animation of the
-  projected data.
-\item
-  \emph{consistency result in stochastic optimisation}: In stochastic
-  algorithm, researchers usually set a seed to ensure the algorithm
-  produces the same result for every run. This practice supports
-  reproducibility, while less efforts has been made to guarantee
-  different seeds will provide the same result.
-\item
-  \emph{high-dimensional decision variable}: In projection pursuit, the
-  decision variable includes all the entries in the projection matrix,
-  which is high-dimensional. Researchers would be better off if they
-  could understand the relative position of different projection matrix
-  in the high-dimensional space.
-\item
-  \emph{role of interpolation in PP optimisation}: An optimisation
-  procedure usually involves iteratively finding projection bases that
-  maximise the index function, while tour requires geodesic
-  interpolation between these bases to produce a continuous view for the
-  users. It would be interesting to see if the interpolated bases could,
-  in reverse, help the optimisation reach faster convergence.
-\end{itemize}
-
-\emph{Think about how does your package help people to understand
-optimisation}
-
-\begin{itemize}
-\tightlist
-\item
-  diagnostic on stochastic optim
-\item
-  vis the progression of multi-parameter decision variable
-\item
-  understanding learning rate - neighbourhood parameter
-\item
-  understand where the local \& global maximum is found - trace plot -
-  see if noisy function
-\end{itemize}
+In modern optimisation problems, gradient information can be hard to
+evaluate or sometimes even impossible and Derivative-Free Optimisation
+(DFO) methods can be useful to approach these problems. A traditional
+and still reliable in solving complicated practical problems, is
+\emph{line search method} \citep{fletcher2013practical}. In the context
+of a simple one-dimensional problem of finding the \(x\) that minimises
+\(f(x)\), line search achieves the goal via an iterative algorithm in
+the form of Equation \ref{eq:line-search}.
+
+\begin{equation}
+x^{(k + 1)} = x^{(k)} + \alpha_k* d^{(k)}
+\label{eq:line-search}
+\end{equation}
+
+where \(d^{(k)}\) is the searching direction in iteration \(k\), and
+\(\alpha_k\) is the step-size. Strictly speaking, \(\alpha_k\) is chosen
+by another minimisation of \(f(x^{(k)} + \alpha* d^{(k)})\) with respect
+to \(\alpha\) and theoretical results have demonstrated the global
+convergence of the algorithm when the exact minimisation of \(\alpha_k\)
+is attained \citep{curry1944method}. In practice, this second
+minimisation is rarely implemented due to its computational demanding or
+even the existence of such a minimisation. A more realistic approach is
+to impose a mandatory decrease in the objective function for each
+iteration: \(f^{(k+1)}> f^{(k)}\) and despite we lose the guarantee on
+global convergence, this approach turns out to be efficient in practical
+problems.
 
@@
 
-Coined by \citet{friedman1974projection}, projection pursuit detects
-interesting structures (i.e.~clustering, outliers and skewness) in
-multivariate data via projecting onto lower dimensions. Let
-\(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be
-seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\)
-defined by \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
+Modern development of the line search methods focuses on proposing
+different computation on the seraching direction: \(d^{(k)}\) and
+various approximations on the step size: \(\alpha_k\) catered for
+practical optimisation problems. The specific problem context we are
+interested in is called projection pursuit guided tour. Projection
+pursuit and guided tour are two separate methods in exporatory data
+analysis focusing on different aspects: coined by
+\citet{friedman1974projection}, projection pursuit detects interesting
+structures (i.e.~clustering, outliers and skewness) in multivariate data
+via low dimensions projection; whilst guided tour is a particular
+variation in a broader class of data visualisation method called tour.
+
+Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection
+can be seen as a linear transformation
+\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
+\(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{P}_{n \times d}\) is the projected data and
@@
 
-A tour method, guided tour is usually used in conjunction with
-projection pursuit to interactively explore multivariate data. Tour
-plays an animation on the projected data which is constructed using the
-bases outputted by projection pursuit and their interpolations. Guided
-tour constructs the interpolation using the shortest geodesic path
-between two outputted bases and we refer readers to
-\citet{buja2005computational} for the mathematical details. Figure
-\ref{tour-path} shows a sketch of the tour path consisting of the blue
-frames produced by the projection pursuit optimisation algorithm and the
-white frames, which are the interpolations between two blue frames. The
-tour method has been implemented in the \emph{tourr} package in R,
-available on the Comprehensive R Archive Network at
+Before introducing the guided tour, we shall be familiar with the
+general tour method \citep{cook2008grand}. A tour produces animated
+visualisation of the high dimensional data via rotating low dimension
+planes. The smoothness of the animation is ensured by computing a series
+of intermediate planes between two low dimension planes via geodesic
+interpolation and we refer readers to \citet{buja2005computational} for
+the mathematical details. Iteratively choose different low dimension
+planes and interpolate between them forms a tour path. Different types
+of tour methods choose the low dimensional planes differently and we
+mention two other type of tour that are commonly used. A grand tour
+selects the planes randomly in the high dimensional space and hence
+serves as an initial exploration of the data. Manual control allows
+researches to fine-tuning an existing projection by gradually phase in
+and out one variable.
+
+Guided tour chooses the planes produced by optimising the projection
+pursuit index function. Figure \ref{tour-path} shows a sketch of the
+tour path consisting of the blue frames produced by the projection
+pursuit optimisation algorithm iteratively and the white frames, which
+are the interpolations between two blue frames. The tour method has been
+implemented in the \emph{tourr} package in R, available on the
+Comprehensive R Archive Network at
 \url{https://cran.r-project.org/web/packages/tourr/}
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
-s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
+s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
 \end{align}
+
+Where \(\delta_{ij}\) is the kronecker delta that takes 1 if \(i = j\)
+and 0 otherwise.
+
+{[}you wanna give more information about the characteristics of the
+optimisation problem before introducing existing algorithms {]}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
+
+If the gradient information was available, the decreasing condition is
+
+\hypertarget{difficulties}{%
+\subsection{Difficulties}\label{difficulties}}
+
+(feel this section should go after the general introduction of DFO and
+PPGT)
+
+{[}Are we using projection pursuit/guided tour to better understand the
+convergence of optimization algorithms visually in combination with the
+algorithms discussed below? Or we are focusing on the optimisation
+problem only within the projection pursuit context? Some of the problems
+listed below are also applicable to optimization problem in general too.
+ppp{]}
+
+Below are listed several issues in projection pursuit optimisation. Some
+are general optimisation problems, while others are more specific too PP
+optimisation.
+
+\begin{itemize}
+\item
+  \emph{constraint optimisation}: our problem requires the project basis
+  to be orthonormal
+\item
+  \emph{Multivariate optimisation}: In projection pursuit, the decision
+  variable includes all the entries in the projection matrix, which is
+  high-dimensional. Researchers would be better off if they could
+  understand the relative position of different projection matrix in the
+  high-dimensional space. (multivariate optimisation)
+\item
+  \emph{optimising non-smooth function}: When the objective function is
+  non-differentiable, derivative information can not be obtained, which
+  means traditional gradient- or Hessian- based methods are not
+  feasible. Stochastic optimisation method could be an alternative to
+  solve these problems. (derivative information is not available)
+\item
+  \emph{Finding global maximum}: Although finding local maximum is
+  relatively easy with developed algorithms, it is generally hard to
+  guarantee global maximum in a problem where the objective function is
+  complex or the number of decision variables is large. Also, there are
+  discussions on how to avoid getting trapped in a local optimum in the
+  literature. (the exact \(\alpha\) that minimised
+  \(f(x^{(k)}+ \alpha d^{(k)})\) can be hard to find or sometiems even
+  impossible, imposing certain decrease, how every this doesn't
+  guarantee global convergence)
+\end{itemize}
+
+Computational considerations:
+
+\begin{itemize}
+\item
+  \emph{computation speed}: The optimisation procedure needs to be fast
+  to compute since the tour produces real-time animation of the
+  projected data.
+\item
+  \emph{consistency result in stochastic optimisation}: In stochastic
+  algorithm, researchers usually set a seed to ensure the algorithm
+  produces the same result for every run. This practice supports
+  reproducibility, while less efforts has been made to guarantee
+  different seeds will provide the same result.
+\item
+  \emph{role of interpolation in PP optimisation}: An optimisation
+  procedure usually involves iteratively finding projection bases that
+  maximise the index function, while tour requires geodesic
+  interpolation between these bases to produce a continuous view for the
+  users. It would be interesting to see if the interpolated bases could,
+  in reverse, help the optimisation reach faster convergence.
+\end{itemize}
+
+\emph{Think about how does your package help people to understand
+optimisation}
+
+\begin{itemize}
+\tightlist
+\item
+  diagnostic on stochastic optim
+\item
+  vis the progression of multi-parameter decision variable
+\item
+  understanding learning rate - neighbourhood parameter
+\item
+  understand where the local \& global maximum is found - trace plot -
+  see if noisy function
+\end{itemize}
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 

### 2020-08-13 paper/paper.Rmd
describe some features of our optim prob; changes on the existing algorithms as per issue #5

@@
 
+[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
+Or we are focusing on the optimisation problem only within the projection pursuit context?
+Some of the problems listed below are also applicable to optimization problem in general too.
+ppp]
+
 \newpage
@@
 ```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.jpeg"))
+knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
 ```
@@
 
-## Optimisation problem and existing algorithms {#tour-optim}
+\newpage
+
+## Optimisation problem  {#tour-optim}
 
@@
 
-[you wanna give more information about the characteristics of the optimisation problem before introducing existing algorithms ]
-
----
-
-If the gradient information was available, the decreasing condition is
-
-## Difficulties
-
-(feel this section should go after the general introduction of DFO and PPGT)
-
-[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
-Or we are focusing on the optimisation problem only within the projection pursuit context?
-Some of the problems listed below are also applicable to optimization problem in general too.
-ppp]
-
-Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.
-
-- *constraint optimisation*: our problem requires the project basis to be orthonormal
-
-- *Multivariate optimisation*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
+There are several features of this optimisation that are worth noticing.
+First of all, it is a multivariate constraint optimisation problem.
+Since the decision variables are the entries of a projection basis, it is required to be orthonormal.
+It is also likely that the objective function is non-differentiable or the gradient information is simply not available.
+In this case, we will need to either use some approximation of the gradient or turn to derivative free methods.
+Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem need to be able to find the global maximum.
+Along the way, locla maximum may also be of our interest since they could present unexpected interesting projections.
+There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.
+
+<!-- Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.  -->
+
+<!-- - *constraint optimisation*: our problem requires the project basis to be orthonormal -->
+
+<!-- - *Multivariate optimisation*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
 Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
-(multivariate optimisation)
-
-- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
+(multivariate optimisation) -->
+
+<!-- - *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
 Stochastic optimisation method could be an alternative to solve these problems.
-(derivative information is not available)
-
-- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
+(derivative information is not available) -->
+
+<!-- - *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
 Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.
-(the exact $\alpha$ that minimised $f(x^{(k)}+ \alpha d^{(k)})$ can be hard to find or sometiems even impossible, imposing certain decrease, how every this doesn't guarantee global convergence)
-
-Computational considerations:
-
-- *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data.
-
-- *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
+(the exact $\alpha$ that minimised $f(x^{(k)}+ \alpha d^{(k)})$ can be hard to find or sometiems even impossible, imposing certain decrease, how every this doesn't guarantee global convergence) -->
+
+<!-- Computational considerations: -->
+
+<!-- - *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data. -->
+
+<!-- - *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
 This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
-
-- *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
+-->
+
+<!-- - *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
 It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
+-->
 
@@
 
-*Think about how does your package help people to understand optimisation*
-
- - diagnostic on stochastic optim
- - vis the progression of multi-parameter decision variable 
- - understanding learning rate - neighbourhood parameter
- - understand where the local & global maximum is found - trace plot - see if noisy function
-
----
-
-There are three existing algorithms.
-
-@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{rand}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+<!-- *Think about how does your package help people to understand optimisation* -->
+
+<!--  - diagnostic on stochastic optim -->
+<!--  - vis the progression of multi-parameter decision variable  -->
+<!--  - understanding learning rate - neighbourhood parameter -->
+<!--  - understand where the local & global maximum is found - trace plot - see if noisy function -->
+
+\newpage
+
+## Existing algorithms
+
+Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient acsent.
+
+@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
 The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path.
@@
 The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
-When the search needs to sample more than $c$ number of basis to find an outputted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
+When the search needs to sample more than $c$ number of basis to find an accpeted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
 If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops.
 The algorithm is summarised in Algorithm \ref{random-search} for one iteration.
+[mention orthonormalise to ensure the constaint is fulfilled; don't use derivative information but a random search]
 
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
-    \Output{The target basis: $\mathbf{A}_{l}$}
+    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
+    \Output{$\mathbf{A}_{l}$}
   initialisation\;
@@
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{l}$ is orthonormal\;
+    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthonormalise $\mathbf{A}_{l}$\;
     Compute $I_{l}  = f(\mathbf{A}_{l})$\;
@@
   }
+  We repeat this loop and if $c > 30$, half the $\alpha$. Reset $c$ to zero.
   \caption{random search}
@@
 
-@cook1995grand explained the use of a gradient ascent optimisation with an assumption that the index function is continuous and differentiable.
-Since some indices could be non-differentiable, the derivative is replaced by a pseudo-derivative of evaluating five randomly generated directions in a tiny nearby neighbourhood.
-Once the best direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic between the current basis $\mathbf{A}_{\text{cur}}$ and the best directional basis $\mathbf{A}_{*}$.
-If the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value, $\mathbf{A}_{**}$ is outputted for the current iteration to construct the interpolation or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] modifies `search_better` based on a non-increasing cooling scheme $T(l)$.
+Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probabiliby $$P= \min\left\{\exp\left[-\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{\cdot}$ denotes the index value of a given basis.
+This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
+This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
+a one dimenional projection of the dataset with two informative variables).
+The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+
+\begin{algorithm}
+\SetAlgoLined
+    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{l} > I_{\text{cur}}$}{
+        \KwRet{$\mathbf{A}_{l}$} \;
+      }{
+        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+        \If{$P > U$}{
+           \KwRet{$\mathbf{A}_{l}$} \;
+        }
+      }
+  \caption{simulated annealing}
+  \label{simulated_annealing}
+\end{algorithm}
+
+@cook1995grand used a gradient ascent algorithm with pseudo-derivative.
+Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the graident.
+The $\delta$ parameter is usually set to be tiny to ensure the geogesic direction is a good local approximation to the graident.
+Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate.
+$\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates.
 Algorithm \ref{search-geodesic} summarise the steps in geodesic search.
@@
   \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$}
-    \Output{The target basis: $\mathbf{A}_{**}$}
+    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n$, $\delta$}
+    \Output{$\mathbf{A}_{**}$}
   initialisation\;
@@
   \While{$l < l_{\max}$}{
-    Generate ten bases in five random directions: $\mathbf{A}_{l}: \mathbf{A}_{l+9}$ within a small neighbourhood $\text{dist}$\;
+    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{l}: \mathbf{A}_{l+9}$ within a small neighbourhood $\delta$\;
     Find the direction with the largest index value: $\mathbf{A}_{*}$ where $l < * < l+ 9$\;
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(l)$.
-Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-The simulated annealing algorithm works as follows.
-Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{\text{rand}}$, a candidate basis is sampled by $\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{\text{rand}}$.
-If the index value of the candidate basis is larger than the one of the current basis, the candidate basis is outputted.
-While if it is smaller, the candidate has another chance of being accepted with probability $P= \min\left\{\exp\left[\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}$ where $I_{\cdot}$ denotes the index value of a given basis.
-The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
-
-\begin{algorithm}
-\SetAlgoLined
-\SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$; Initial temperature: $T_0$}
-    \Output{The global object; The target basis: $\mathbf{A}_{l}$}
-  initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
-    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I_{l} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{l}$} \;
-      }{
-        Compute $P= \min\left\{\exp\left[\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-        \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{l}$} \;
-        }
-      }
-    $l = l + 1$\;
-  }
-  \caption{simulated annealing}
-  \label{simulated_annealing}
-\end{algorithm}
-
-Notice that the three algorithms above constitute one search. The outputted basis is then sent to construct the interpolation and one such search-and-interpolate step is one round.
-
 <!-- Below listed several features characterise the optimisation procedures needed in projection pursuit -->
@@
 
-We can see from Figure \ref{tour-path} that only the data projected using the outputted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
+We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
 Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.
@@
 \vdots \\
-\text{search (outputted basis)} \\
+\text{search (accepted basis)} \\
 \hline

### 2020-08-13 paper/paper.tex
describe some features of our optim prob; changes on the existing algorithms as per issue #5

@@
 
+{[}Are we using projection pursuit/guided tour to better understand the
+convergence of optimization algorithms visually in combination with the
+algorithms discussed below? Or we are focusing on the optimisation
+problem only within the projection pursuit context? Some of the problems
+listed below are also applicable to optimization problem in general too.
+ppp{]}
+
 \newpage
@@
 
+\newpage
+
 \hypertarget{tour-optim}{%
-\subsection{Optimisation problem and existing
-algorithms}\label{tour-optim}}
+\subsection{Optimisation problem}\label{tour-optim}}
 
@@
 
-{[}you wanna give more information about the characteristics of the
-optimisation problem before introducing existing algorithms {]}
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-If the gradient information was available, the decreasing condition is
-
-\hypertarget{difficulties}{%
-\subsection{Difficulties}\label{difficulties}}
-
-(feel this section should go after the general introduction of DFO and
-PPGT)
-
-{[}Are we using projection pursuit/guided tour to better understand the
-convergence of optimization algorithms visually in combination with the
-algorithms discussed below? Or we are focusing on the optimisation
-problem only within the projection pursuit context? Some of the problems
-listed below are also applicable to optimization problem in general too.
-ppp{]}
-
-Below are listed several issues in projection pursuit optimisation. Some
-are general optimisation problems, while others are more specific too PP
-optimisation.
-
-\begin{itemize}
-\item
-  \emph{constraint optimisation}: our problem requires the project basis
-  to be orthonormal
-\item
-  \emph{Multivariate optimisation}: In projection pursuit, the decision
-  variable includes all the entries in the projection matrix, which is
-  high-dimensional. Researchers would be better off if they could
-  understand the relative position of different projection matrix in the
-  high-dimensional space. (multivariate optimisation)
-\item
-  \emph{optimising non-smooth function}: When the objective function is
-  non-differentiable, derivative information can not be obtained, which
-  means traditional gradient- or Hessian- based methods are not
-  feasible. Stochastic optimisation method could be an alternative to
-  solve these problems. (derivative information is not available)
-\item
-  \emph{Finding global maximum}: Although finding local maximum is
-  relatively easy with developed algorithms, it is generally hard to
-  guarantee global maximum in a problem where the objective function is
-  complex or the number of decision variables is large. Also, there are
-  discussions on how to avoid getting trapped in a local optimum in the
-  literature. (the exact \(\alpha\) that minimised
-  \(f(x^{(k)}+ \alpha d^{(k)})\) can be hard to find or sometiems even
-  impossible, imposing certain decrease, how every this doesn't
-  guarantee global convergence)
-\end{itemize}
-
-Computational considerations:
-
-\begin{itemize}
-\item
-  \emph{computation speed}: The optimisation procedure needs to be fast
-  to compute since the tour produces real-time animation of the
-  projected data.
-\item
-  \emph{consistency result in stochastic optimisation}: In stochastic
-  algorithm, researchers usually set a seed to ensure the algorithm
-  produces the same result for every run. This practice supports
-  reproducibility, while less efforts has been made to guarantee
-  different seeds will provide the same result.
-\item
-  \emph{role of interpolation in PP optimisation}: An optimisation
-  procedure usually involves iteratively finding projection bases that
-  maximise the index function, while tour requires geodesic
-  interpolation between these bases to produce a continuous view for the
-  users. It would be interesting to see if the interpolated bases could,
-  in reverse, help the optimisation reach faster convergence.
-\end{itemize}
-
-\emph{Think about how does your package help people to understand
-optimisation}
-
-\begin{itemize}
-\tightlist
-\item
-  diagnostic on stochastic optim
-\item
-  vis the progression of multi-parameter decision variable
-\item
-  understanding learning rate - neighbourhood parameter
-\item
-  understand where the local \& global maximum is found - trace plot -
-  see if noisy function
-\end{itemize}
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-There are three existing algorithms.
+There are several features of this optimisation that are worth noticing.
+First of all, it is a multivariate constraint optimisation problem.
+Since the decision variables are the entries of a projection basis, it
+is required to be orthonormal. It is also likely that the objective
+function is non-differentiable or the gradient information is simply not
+available. In this case, we will need to either use some approximation
+of the gradient or turn to derivative free methods. Given the goal of
+projection pursuit as finding the basis with the largest index value,
+the optimisation problem need to be able to find the global maximum.
+Along the way, locla maximum may also be of our interest since they
+could present unexpected interesting projections. There is also one
+computational consideration: the optimisation procedure needs to be easy
+to compute since the tour animation is played in real-time.
+
+\newpage
+
+\hypertarget{existing-algorithms}{%
+\subsection{Existing algorithms}\label{existing-algorithms}}
+
+Below we introduce three possible algorithms: \texttt{search\_better}
+and \texttt{search\_better\_random} are derivative free methods that
+sample candidate basis in the neighbourhood whilst
+\texttt{search\_geodesic} approximates the gradient information and acts
+as a non-derivative version of the gradient acsent.
 
@@
 where \(\alpha\) controls the radius of the sampling neighbourhood and
-\(\mathbf{A}_{rand}\) is a randomly generated matrix with the same
-dimension as \(\mathbf{A}_{\text{cur}}\). The optimiser keeps sampling
-bases near the current basis until it finds one with higher index value
-than the current basis and then outputs it for guided tour to construct
-the interpolation path. A new round of search continues to find a better
-basis after the interpolation finishes. The halving parameter \(c\) with
-default value of 30 is designed to adjust the searching neighbourhood
-\(\alpha\). When the search needs to sample more than \(c\) number of
-basis to find an outputted basis, the neighbourhood parameter \(\alpha\)
-will be reduced by half in the next iteration. If the optimiser can't
-find a better basis within the maximum number of tries \(l_{\max}\), the
-algorithm stops. The algorithm is summarised in Algorithm
-\ref{random-search} for one iteration.
+\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
+same dimension as \(\mathbf{A}_{\text{cur}}\). The optimiser keeps
+sampling bases near the current basis until it finds one with higher
+index value than the current basis and then outputs it for guided tour
+to construct the interpolation path. A new round of search continues to
+find a better basis after the interpolation finishes. The halving
+parameter \(c\) with default value of 30 is designed to adjust the
+searching neighbourhood \(\alpha\). When the search needs to sample more
+than \(c\) number of basis to find an accpeted basis, the neighbourhood
+parameter \(\alpha\) will be reduced by half in the next iteration. If
+the optimiser can't find a better basis within the maximum number of
+tries \(l_{\max}\), the algorithm stops. The algorithm is summarised in
+Algorithm \ref{random-search} for one iteration. {[}mention
+orthonormalise to ensure the constaint is fulfilled; don't use
+derivative information but a random search{]}
 
@@
 
-\citet{cook1995grand} explained the use of a gradient ascent
-optimisation with an assumption that the index function is continuous
-and differentiable. Since some indices could be non-differentiable, the
-derivative is replaced by a pseudo-derivative of evaluating five
-randomly generated directions in a tiny nearby neighbourhood. Once the
-best direction is determined, the optimiser finds the best projection
-\(\mathbf{A}_{**}\) on the geodesic between the current basis
-\(\mathbf{A}_{\text{cur}}\) and the best directional basis
-\(\mathbf{A}_{*}\). If the percentage change in the index value between
-\(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a
-threshold value, \(\mathbf{A}_{**}\) is outputted for the current
-iteration to construct the interpolation or the algorithm repeats the
-above steps until \(l_{\max}\) is reached and the searching terminates.
-Algorithm \ref{search-geodesic} summarise the steps in geodesic search.
+Simulated annealing
+\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated}
+modifies \texttt{search\_better} based on a non-increasing cooling
+scheme \(T(l)\). Given an initial \(T_0\), the temperature at iteration
+\(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate
+basis fails to have an index value larger than the current basis,
+simulated annealing gives it a second chance to be accepted with
+probabiliby
+\[P= \min\left\{\exp\left[-\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}\]
+where \(I_{\cdot}\) denotes the index value of a given basis. This
+implementation allows the algorithm to jump out of a local maximum and
+enables a more holistic search of the whole parameter space. This
+feature is particularly useful when the dimension of the projected space
+is smaller than the number of informative variables in the dataset
+(i.e.~a one dimenional projection of the dataset with two informative
+variables). The algorithm can be written as replacing line 5-11 of
+Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+
+\begin{algorithm}
+\SetAlgoLined
+    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{l} > I_{\text{cur}}$}{
+        \KwRet{$\mathbf{A}_{l}$} \;
+      }{
+        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+        \If{$P > U$}{
+           \KwRet{$\mathbf{A}_{l}$} \;
+        }
+      }
+  \caption{simulated annealing}
+  \label{simulated_annealing}
+\end{algorithm}
+
+\citet{cook1995grand} used a gradient ascent algorithm with
+pseudo-derivative. Instead of computing the actual gradient of the index
+function with respect to the projection basis matrix,
+\texttt{search\_geodesic} samples \(2n\) bases that are randomly
+generated on a uniform open ball with the radius controlled by the
+\(\delta\) parameter and finds the most promising one to construct the
+geodesic direction as an approximation to the graident. The \(\delta\)
+parameter is usually set to be tiny to ensure the geogesic direction is
+a good local approximation to the graident. Once the searching direction
+is determined, the optimiser finds the best projection
+\(\mathbf{A}_{**}\) on the geodesic as the candidate.
+\(\mathbf{A}_{**}\) is outputted for the current iteration if the
+percentage change in the index value between \(\mathbf{A}_{**}\) and
+\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value or the
+algorithm repeats the above steps until \(l_{\max}\) is reached and the
+searching terminates. Algorithm \ref{search-geodesic} summarise the
+steps in geodesic search.
 
@@
 
-Simulated annealing
-\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
-non-derivative procedure based on a non-increasing cooling scheme
-\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
-is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). The simulated
-annealing algorithm works as follows. Given a neighbourhood parameter
-\(\alpha\) and a randomly generated orthonormal basis
-\(A_{\text{rand}}\), a candidate basis is sampled by
-\(\mathbf{A}_{t+1} = (1 - \alpha)\mathbf{A}_{t} + \alpha \mathbf{A}_{\text{rand}}\).
-If the index value of the candidate basis is larger than the one of the
-current basis, the candidate basis is outputted. While if it is smaller,
-the candidate has another chance of being accepted with probability
-\(P= \min\left\{\exp\left[\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}\)
-where \(I_{\cdot}\) denotes the index value of a given basis. The
-algorithm can be summarised as in Algorithm \ref{simulated_annealing}.
-
-\begin{algorithm}
-\SetAlgoLined
-\SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-    \Input{The current projection basis: $\mathbf{A}_{\text{cur}}$; The index function: $f$; Initial temperature: $T_0$}
-    \Output{The global object; The target basis: $\mathbf{A}_{l}$}
-  initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1 - \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ ensuring $\mathbf{A}_{l}$ is orthonormal \;
-    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I_{l} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{l}$} \;
-      }{
-        Compute $P= \min\left\{\exp\left[\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-        \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{l}$} \;
-        }
-      }
-    $l = l + 1$\;
-  }
-  \caption{simulated annealing}
-  \label{simulated_annealing}
-\end{algorithm}
-
-Notice that the three algorithms above constitute one search. The
-outputted basis is then sent to construct the interpolation and one such
-search-and-interpolate step is one round.
-
 \newpage
@@
 We can see from Figure \ref{tour-path} that only the data projected
-using the outputted and interpolated bases are animated by tour; none of
+using the accepted and interpolated bases are animated by tour; none of
 the searching bases sampled by the optimsiation algorithms are presented

### 2020-08-13 paper/paper.Rmd
remove the commented text that are no longer necessary

@@
 
-<!-- Below are listed several issues in projection pursuit optimisation. Some are general optimisation problems, while others are more specific too PP optimisation.  -->
-
 <!-- - *constraint optimisation*: our problem requires the project basis to be orthonormal -->
@@
 Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
-(multivariate optimisation) -->
+-->
 
@@
 
-<!-- Computational considerations: -->
-
 <!-- - *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data. -->
@@
 <!-- the interpolation can be seen as giving information on the direction of the derivative. -->
-
-<!-- - is `search_geodesic` widely known in the research community - doesn't seem to find paper introducing this except (Cook, 1995).
-Any tools available to  better understand its searching process?
--->
 
@@
 
-<!-- Below listed several features characterise the optimisation procedures needed in projection pursuit -->
-
-<!-- - *Being able to handle non-differentiable PPI function*: The PPI function could be non-differentiable, thus derivative free methods are preferred.  -->
-
-<!-- - *Being able to optimise with constraints*: The constraint comes from projection matrix being an orthonormal matrix.  -->
-
-<!-- - *Being able to find both local and global maximum*: Although the primary interest is to find the global maximum, local maximum could also reveal structures in the data that are of our interest.
--->
-
 \newpage
@@
 ## Motivation
-<!-- Visual diagnostics can be real-time or post-run.
-Real time diagnostic directly uses the data produced in the algorithm to produce visual representation and thus doesn't need to store the data.
-This section focuses on the definition and production of post-run diagnostics and the next section discusses real-time diagnostics.
--->
 
@@
 $l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
-
-<!-- For each level of iteration, we design two types of indices: *index per iteration: $j$* is fixed for each observations in one iteration and has an increment of one once a new iteration starts.
-*index within iteration: $l$* has an increment of one for each observation in an iteration and starts over from one once a new iteration starts.
--->
-
-<!-- In projection pursuit optimisation algorithms, there is one level of iteration and hence exists three iterators: `id` indices each observation by a unique number; `tries` is the index per iteration iterator that gets updated once a search-and-interpolate step is finished; and `loop` is the index within iteration iterator and starts over from one at the beginning of a new `tries`.
-We give the interpolating basis a different index $k$ for projection pursuit guided tour.
-It is similar in nature to $l$ but is specific for interpolating bases, which is usually not part of the optimisation.
--->
 
@@
 ```
-
-<!-- \begin{equation} -->
-<!-- \left[ -->
-<!-- \begin{array}{ccccccc} -->
-<!-- \ast & \mathbf{A}_{1,1,\ast} & \mathbf{A}_{2, l_1 +1,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +1,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +1 ,\ast} \\ -->
-<!-- \ast & \mathbf{A}_{1,2,\ast} & \mathbf{A}_{2,l_1 +2,\ast} & \cdots & \mathbf{A}_{J,\sum_{i = 1}^{J-1}l_i +2,\ast} & \mathbf{A}_{J+1,\sum_{i = 1}^{J}l_i +2, \ast} \\ -->
-<!-- \vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\ -->
-<!-- \ast & \mathbf{A}_{1, l, \ast} & \mathbf{A}_{2, l_1 + l, \ast} & \cdots & \mathbf{A}_{J+1, \sum_{i = 1}^{J-1}l_i + l, \ast}  & \mathbf{A}_{J + 1, \sum_{i = 1}^{J}l_i + l, \ast}\\ -->
-<!-- \vdots & \vdots & \vdots & \ddots & \vdots &\vdots   \\ -->
-<!-- \ast & \mathbf{A}_{1, l_1-1, \ast} & \mathbf{A}_{2, l_1 + l_2-1, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J-1}l_i -1, \ast} & \mathbf{A}_{J+1, \sum_{i = 1}^{J+1} l_i-1, \ast}\\ -->
-<!-- \mathbf{A}_{0,\ast,\ast} & \mathbf{A}_{1, l_1, \ast} & \mathbf{A}_{2, l_1 + l_2, \ast} & \cdots & \mathbf{A}_{J, \sum_{i = 1}^{J}l_i , \ast} & \ast\\ -->
-<!-- \hline -->
-<!-- \ast &\mathbf{A}_{1, \ast, 1} & \mathbf{A}_{2, \ast, k_1 + 1} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J-1}k_i + 1}  & \ast\\ -->
-<!-- \ast &\mathbf{A}_{1, \ast, 2} & \mathbf{A}_{2, \ast, k_1 + 2} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +2}  & \ast \\ -->
-<!-- \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\ -->
-<!-- \ast &\mathbf{A}_{1, \ast, k} & \mathbf{A}_{2, \ast, k_1 + k} & \cdots &\mathbf{A}_{J, \ast, \sum_{i = 1}^{J-1}k_i +k}  & \ast \\ -->
-<!-- \vdots & \vdots & \vdots & \ddots & \vdots   & \vdots \\ -->
-<!-- \ast &\mathbf{A}_{1, \ast, k_1} & \mathbf{A}_{2, \ast, k_1 + k_2} & \cdots &\mathbf{A}_{J, \ast,\sum_{i = 1}^{J}k_i} & \ast \\ -->
-<!-- \end{array} -->
-<!-- \right] -->
-<!-- = -->
-<!-- \left[ -->
-<!-- \begin{array}{c} -->
-<!-- \vphantom{A^J_1}\text{candidate basis} \\ -->
-<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
-<!-- \vphantom{\vdots} \vdots \\ -->
-<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
-<!-- \vphantom{\vdots} \vdots \\ -->
-<!-- \vphantom{A^J_2}\text{candidate basis} \\ -->
-<!-- \vphantom{A^J}\text{current/target basis} \\ -->
-<!-- \hline -->
-<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
-<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
-<!-- \vphantom{\text{NULL}}\vdots \\ -->
-<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
-<!-- \vphantom{\vdots} \vdots \\ -->
-<!-- \vphantom{\text{NULL}}\text{interpolation basis} \\ -->
-<!-- \end{array} -->
-<!-- \right] -->
-<!-- \label{eq:matrixA} -->
-<!-- \end{equation} -->
-
-<!-- An optimisation procedure can easily generate hundreds or thousands sampling points and the question becomes what are the points we want to explore...
-[the classification of searching or updating points goes here and people should be clear what part of the points goes to the plot - this is the data part not the plot part.
-] -->
-
-<!-- In tour, all the points recorded in the global object can be divided into two broad categories: searching points and interpolating points -->
-
-<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
-The points for target bases is also included in the searching points and there is one such point per `tries`.
--->
-
-<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
-
-<!-- The concept of grammar of graphic requires a dataset to be supplied in making a plot, thus a global object needs to be created in the iterative algorithm.
-The variables will be mapped into element of a graphic to explore and thus should contain all the parameters of interest.
-The next section shows how a global object is created in projection pursuit guided tour.
--->
-
-<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
 
@@
 
-<!-- `tries` has an increment of one once the generator is called (equivalently a new target basis is generated); `info` records the stage the basis is in.
-This would include the `interpolation` stage and the detailed stage in the optimisation i.e.
-`direction_search`, `best_direction_search`, `line_search`and `best_line_search` for geodesic searching (`search_geodesic`); `random_search` and `new_basis` for simulating annealing (`search_better`).
-`loop` is the counter used for the optimisation procedure and thus will be `NA` for interpolation steps.
-`id` creates a sequential order of the basis.
--->
-
-<!--  Another examples is a 2D projection of the larger dataset with two informative variable (`x2` and `x7`) using search_better method.
-Notice in this example, the dimension of the bases becomes 6 by 2.
--->
-
-<!-- ```{r glb-obj-2, echo = TRUE} -->
-<!-- holes_2d_better %>% head(5)  -->
-<!-- ``` -->
-
 # Visual diagnostic plots {#vis-diag-plots}
@@
 ## Explore the value of objective function {#static}
+
+<!-- In tour, all the points recorded in the global object can be roughly divided into two broad categories: searching points and interpolating points -->
+
+<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
+The points for target bases is also included in the searching points and there is one such point per `tries`.
+-->
+
+<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
+
+<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
 

### 2020-08-13 paper/paper.Rmd
modify according to issue #6

@@
 origin_dt_bi <- data_mult %>% 
-  dplyr::select(-x3) %>% 
+  dplyr::select(-c(x3, x8, x9, x10)) %>% 
   gather(names, values) %>%
@@
 
-We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10` and run the guided tour with optimiser `search_better`.
-When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
-Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.
-Below shows the first ten rows of the global object.
-Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.
+We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the global object.
 
@@
 ```
-
-\newpage
 
@@
 
+When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
+Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.
+Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.
+
 # Conclusion

### 2020-08-13 paper/paper.tex
modify according to issue #6

@@
 We form our first dataset using variables \texttt{x1}, \texttt{x2},
-\texttt{x8}, \texttt{x9} and \texttt{x10} and run the guided tour with
-optimiser \texttt{search\_better}. When the optimisation ends, the
-global object will be stored and printed (it can be turned off by
-supplying argument \texttt{print\ =\ FALSE}). Additional messages during
-the optimisation can be displayed by argument \texttt{verbose\ =\ TRUE}.
-Below shows the first ten rows of the global object. Notice that the
-tibble object allows the list-column \texttt{basis} to be printed out
-nicely with the dimension of the projection basis readily available.
+\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
+optimiser \texttt{search\_better} and below shows the first ten rows of
+the global object.
 
@@
 data structure.{]}
-
-\newpage
 
@@
 
+When the optimisation ends, the global object will be stored and printed
+(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
+Additional messages during the optimisation can be displayed by argument
+\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
+list-column \texttt{basis} to be printed out nicely with the dimension
+of the projection basis readily available.
+
 \hypertarget{conclusion}{%

### 2020-08-19 paper/paper.Rmd
changes in introduction as per meeting 21

@@
 
+<!-- We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
+Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.
+-->
+
+<!-- This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
+-->
+
+<!-- Hence the study **presents a gallery of diagnostic plots applicable to optimisation algorithms.** -->
+
+Visualisation has been widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to see information they would otherwise not see.
+This motivates our work of creating plots to diagnose optimisation algorithms in the context of projection pursuit guided tour, with the aim to understand and compare features of different existing algorithms.
+
 In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
-The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced space given by the problem constraints to either minimize or maximize a function....
-Gradient based optimization has been typically used to solve such problems.
-However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.
+The problem consists in optimizing an objective function $f: S \rightarrow \mathbb{R}$ with $S \in \mathbb{R}^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.
 
-Derivative free methods provide another option to optimise the objective function without evaluating any gradient or Hessian information and a particular class of methods, direct search, has gained popularity through its conceptual simplicity.
-However, the whole searching process in the algorithm remains a black-box.
-Plots are usually used to evaluate and compare the performance of different algorithms but it can easily become tedious because the code will have to be modified significantly when comparing different parameters in the algorithms.
-For example, a categorical variable with 5 levels can be easily mapped onto colour while mapping another categorical variable with 30 levels will not make the plot informative.
-Thus the plot needs to be re-designed to better suit the characteristics of the parameter (whether the parameter is a scalar or a matrix?
-whether the parameter is quantitative or categorical?
-If categorical, how many levels does the parameter have?).
-This motivates the  design of a visual diagnostic framework for optimisation algorithms based on the idea of a *global object*.
+Projection pursuit and guided tour are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
+Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
 
-The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
-Section \ref{tour} introduces the optimisation problem in a specific context: projection pursuit guided tour and existing derivative free algorithms for solving it. 
-Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveils the black-box of optimisation.
-Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} gives definitions of different diagnostic plots using grammar of graphic and examples.
-Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.
+The remainder of the paper is organised as follows. 
+Section \ref{optim} provides the literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour. 
+Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples. 
+Section \ref{application} shows the applicaiton of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
+Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
 
-# Derivative free optimisation {#DFO}
+# Derivative free optimisation {#optim}
 
@@
 
-## Motivation
-
-We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
-Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.
-
-Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see information they would otherwise not see.
-This motivates us **to develop a visual framework to diagnose optimisation algorithms in projection pursuit with the aim to understand and compare different existing algorithms.** This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
+## Global object
 
@@
 
-Hence the study **contributes to set up a tidy data structure for collecting observations from optimisation algorithms and prescribe a set of rules for generalising diagnostic plots for optimisation algorithms, specifically those used for projection pursuit.**
-
 <!-- [Do I need this paragraph to details the grammar of graphic? xxx - don't think so; not particularly relevant - assume most people know it roughly] -->
@@
 -->
-
-## Global object
 
@@
 
-# Visual diagnostic plots {#vis-diag-plots}
+## Visual diagnostic plots {#vis-diag-plots}
 
@@
 
+# Application {#application}
 # Implementation: Ferrn package {#implementation}

### 2020-08-19 paper/paper.tex
changes in introduction as per meeting 21

@@
 
+Visualisation has been widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to see
+information they would otherwise not see. This motivates our work of
+creating plots to diagnose optimisation algorithms in the context of
+projection pursuit guided tour, with the aim to understand and compare
+features of different existing algorithms.
+
 In an optimization problem the goal is to find the best solution within
@@
 set of constraints. The problem consists in optimizing an objective
-function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
-space given by the problem constraints to either minimize or maximize a
-function\ldots{}. Gradient based optimization has been typically used to
-solve such problems. However, there are many situations where
-derivatives of an objective function are unavailable or unreliable and
-therefore traditional methods based on derivatives are not the best
-option to solve an optimization problem.
-
-Derivative free methods provide another option to optimise the objective
-function without evaluating any gradient or Hessian information and a
-particular class of methods, direct search, has gained popularity
-through its conceptual simplicity. However, the whole searching process
-in the algorithm remains a black-box. Plots are usually used to evaluate
-and compare the performance of different algorithms but it can easily
-become tedious because the code will have to be modified significantly
-when comparing different parameters in the algorithms. For example, a
-categorical variable with 5 levels can be easily mapped onto colour
-while mapping another categorical variable with 30 levels will not make
-the plot informative. Thus the plot needs to be re-designed to better
-suit the characteristics of the parameter (whether the parameter is a
-scalar or a matrix? whether the parameter is quantitative or
-categorical? If categorical, how many levels does the parameter have?).
-This motivates the design of a visual diagnostic framework for
-optimisation algorithms based on the idea of a \emph{global object}.
-
-The remainder of the paper is organised as follows. Section \ref{DFO}
-gives a general overview of optimisation methods, specifically
-derivative free optimisation. Section \ref{tour} introduces the
-optimisation problem in a specific context: projection pursuit guided
-tour and existing derivative free algorithms for solving it. Section
-\ref{vis-diag} and \ref{vis-diag-plots} together present the new
-construction of a visual diagnostic framework that unveils the black-box
-of optimisation. Section \ref{vis-diag} focuses on motivational ideas
-and data structure and Section \ref{vis-diag-plots} gives definitions of
-different diagnostic plots using grammar of graphic and examples.
-Finally, Section \ref{implementation} mentions the R package: ferrn that
-implements the visual diagnostics described above.
-
-\hypertarget{DFO}{%
-\section{Derivative free optimisation}\label{DFO}}
+function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in
+a reduced space given by the problem constraints to either minimize or
+maximize a function.
+
+Projection pursuit and guided tour are exploratory data analysis tools
+that detect interesting structure of high dimensional data through
+projection on low dimensional space. Optimisation is applied here to
+search for the low dimensional space that finds the most interesting
+projection.
+
+The remainder of the paper is organised as follows. Section \ref{optim}
+provides the literature review of optimisation methods, specifically the
+line search methods used in projection pursuit guided tour. Section
+\ref{tour} reviews projection pursuit guided tour, form the introduces
+the optimisation problem, and introduces three main existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics design, from
+forming the data object to the definition of different diagnostic plots.
+Section \ref{application} shows some examples of how the diagnostic
+plots designed in section \ref{vis-diag} can be applied to understand
+and compare different algorithms and how they contribute to
+modifications that improve the algorithms. Finally, Section
+\ref{implementation} describes the R package: ferrn, that implements all
+the visual diagnostics above.
+
+\hypertarget{optim}{%
+\section{Derivative free optimisation}\label{optim}}
 
@@
 
-\hypertarget{motivation}{%
-\subsection{Motivation}\label{motivation}}
-
-We can see from Figure \ref{tour-path} that only the data projected
-using the accepted and interpolated bases are animated by tour; none of
-the searching bases sampled by the optimsiation algorithms are presented
-in the tour animation. Although these algorithms have been demonstrated
-to find the global maximum in various literature as we cited above, they
-focus solely on finding the global maximum and fewer tools are available
-to explore how each algorithm searches the parameter space and compare
-the results between different algorithms.
-
-Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see
-information they would otherwise not see. This motivates us \textbf{to
-develop a visual framework to diagnose optimisation algorithms in
-projection pursuit with the aim to understand and compare different
-existing algorithms.} This study is meaningful because \textbf{as more
-complex index functions are proposed based on statistical theory, only
-computationally effective optimisation algorithms allow us to find the
-projection basis that presents interesting structure in the data}.
+\hypertarget{global-object}{%
+\subsection{Global object}\label{global-object}}
 
@@
 wrangling and visualisation are greatly simplified.
-
-Hence the study \textbf{contributes to set up a tidy data structure for
-collecting observations from optimisation algorithms and prescribe a set
-of rules for generalising diagnostic plots for optimisation algorithms,
-specifically those used for projection pursuit.}
-
-\hypertarget{global-object}{%
-\subsection{Global object}\label{global-object}}
 
@@
 \hypertarget{vis-diag-plots}{%
-\section{Visual diagnostic plots}\label{vis-diag-plots}}
+\subsection{Visual diagnostic plots}\label{vis-diag-plots}}
 
@@
 
+\hypertarget{application}{%
+\section{Application}\label{application}}
+
 \hypertarget{implementation}{%

### 2020-08-19 paper/paper.Rmd
change the beginning of optimisation lit review

@@
 
-# Derivative free optimisation {#optim}
+\newpage
 
-In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and  Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
-A traditional and still reliable in solving complicated practical problems, is *line search method* [@fletcher2013practical].
-In the context of a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
+# Optimisation Methods {#optim}
+
+Given an optimisation problem, two basic approaches find the optimum based on different thinking.
+An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the close-form solution may not be available when the problem starts to become complex.
+An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution.
+An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.
+
+A traditional while often used in practice is an iteartive method called *line search method* [@fletcher2013practical].
+In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
 

### 2020-08-19 paper/paper.tex
change the beginning of optimisation lit review

@@
 line search methods used in projection pursuit guided tour. Section
-\ref{tour} reviews projection pursuit guided tour, form the introduces
-the optimisation problem, and introduces three main existing algorithms.
+\ref{tour} reviews projection pursuit guided tour, forms the
+optimisation problem, and introduces three main existing algorithms.
 Section \ref{vis-diag} presents the new visual diagnostics design, from
-forming the data object to the definition of different diagnostic plots.
-Section \ref{application} shows some examples of how the diagnostic
-plots designed in section \ref{vis-diag} can be applied to understand
-and compare different algorithms and how they contribute to
-modifications that improve the algorithms. Finally, Section
-\ref{implementation} describes the R package: ferrn, that implements all
-the visual diagnostics above.
+forming the data object to the definition of different diagnostic plots
+with some small examples. Section \ref{application} shows the
+applicaiton of how the diagnostic plots designed in section
+\ref{vis-diag} can be used to understand and compare different
+algorithms and how they contribute to modifications that improve the
+algorithms. Finally, Section \ref{implementation} describes the R
+package: ferrn, that implements all the visual diagnostics above.
+
+\newpage
 
 \hypertarget{optim}{%
-\section{Derivative free optimisation}\label{optim}}
-
-In modern optimisation problems, gradient information can be hard to
-evaluate or sometimes even impossible and Derivative-Free Optimisation
-(DFO) methods can be useful to approach these problems. A traditional
-and still reliable in solving complicated practical problems, is
-\emph{line search method} \citep{fletcher2013practical}. In the context
-of a simple one-dimensional problem of finding the \(x\) that minimises
-\(f(x)\), line search achieves the goal via an iterative algorithm in
-the form of Equation \ref{eq:line-search}.
+\section{Optimisation Methods}\label{optim}}
+
+Given an optimisation problem, two basic approaches find the optimum
+based on different thinking. An analytical approach aims to find the
+optimal solution in a finite number of steps, but a potential issue with
+it is that the close-form solution may not be available when the problem
+starts to become complex. An iterative approach, on the other hand,
+finds the optimum based on the idea of making progressive improvement to
+the current solution. An iterative method may end up finding a local
+optimum but the progressive nature of the algorithm allows the
+practitioner to decide when to stop if a desirable accuracy has been
+achieved.
+
+A traditional while often used in practice is an iteartive method called
+\emph{line search method} \citep{fletcher2013practical}. In a simple
+one-dimensional problem of finding the \(x\) that minimises \(f(x)\),
+line search achieves the goal via an iterative algorithm in the form of
+Equation \ref{eq:line-search}.
 

### 2020-08-19 paper/paper.Rmd
rearranging Sec.4&5 to 4. diagnostics and 5. application

@@
 
-# Visual diagnostic system {#vis-diag}
+# Visual diagnostics{#vis-diag}
 
@@
 
-## Simulated data
+## Plot definition
+
+Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
+To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.
+
+<!-- The plot design of visual diagnostics depends on the characteristics of the variables to plot.
+If the searching points are of interest [point geom is not good; need summarise.
+While updating points are more manageable -> point geom].
+However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
+[stratefy may help?
+- may need more work here.] -->
+
+### Explore the value of objective function {#static}
+
+<!-- In tour, all the points recorded in the global object can be roughly divided into two broad categories: searching points and interpolating points -->
+
+<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
+The points for target bases is also included in the searching points and there is one such point per `tries`.
+-->
+
+<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
+
+<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
+This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
+
+#### Searching points
+
+A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimum progressively.
+We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
+In the plot, points from those iterations towards the end will occupy the vast majority of the plot space.
+This motivates to use summarisation.
+Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).
+
+Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it does not report the number of points in each box.
+We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
+Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument.
+A line geometry is also added to link the points with the largest index value in each iteration.
+This helps to visualise the improvement made in each iteration.
+Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:
+
+- Layer 1: boxplot geom
+  - data: group by $j$ and filter the observations in the group that have count greater than `cutoff = 15`.
+  - x: $j$ is mapped to the x-axis
+  - y: the statistical transformed index value: $Q_{I^{\prime}_t}(q)$ is mapped to the y-axis where $Q_X(q)$, $q = 0, 25, 50, 75, 100$ finds the qth-quantile of $X$ and $I^{\prime}_t$ denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
+
+- Layer 2: point geom
+  - data: group by $j$ and filter the observations in the group that have count less than `cutoff = 15`.
+  - x: $j$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
+
+- Layer 3: line geom
+  - data: filter the points with the highest index value in group $j$
+  - x: $j$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
+
+##### Toy example: exploring searching points
+
+We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
+Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
+The index value of the searching points are shown in Figure \ref{points}.
+The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
+We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.
+
+```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
+The left plot does not use the plot space efficiently to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
+
+dt <- holes_2d_better_max_tries %>%
+  filter(info != "interpolation") %>% 
+  mutate(id = row_number())
+
+dt %>% explore_trace_search()
+
+```
+
+#### Interpolating points
+
+Sometimes, rather than exploring the searching points, we may be interested in exploring the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
+Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed.
+The slightly modified plot definition is shown below:
+
+- Layer 1: point geom
+  - data: filter the observations with $S = \text{interpolation}$ and mutate $t$ to be the row number of the subsetted tibble
+  - x: $t$ is mapped to the x-axis
+  - y: $I$ is mapped to the y-axis
+
+- Layer 2: line geom
+  - using line geometry for the same data and aesthetics
+
+##### Toy example
+
+### Explore searching space
+
+In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot.
+An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
+The visualisation can thus be defined as
+
+<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
+This gives a matrix of dimension $n$ by $(p*d)$.
+Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
+By definition, the principal components maximise the variance of the projection bases in the reduced space.
+Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
+Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix -->
+
+<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
+
+<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
+The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
+
+<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
+THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
+colour).
+-->
+
+- Layer 1: point geom
+  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
+  - x: the projected basis on the first principal component
+  - y: the projected basis on the second principal component
+  - colour: $V$ is mapped to the colour aesthetic
+
+<!-- In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
+[need to think further about how different type of variables can be shown].
+-->
+
+#### A toy example: understand different stage of search_geodesic
+
+`search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
+Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
+The best one is picked and a line search is then run on the geodesic direction to find the target basis.
+The starting and target bases are then interpolated.
+In the next iteration, the target basis becomes the current basis and then the procedures continues.
+
+```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+
+holes_1d_geo %>% explore_proj_pca(col = info)
+
+#xxx the info level need to be reorder to show the start and directional search points better.
+```
+
+\newpage
+
+# Application {#application}
+
+Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation.
+We will present more examples that can help us to understand the algorithm and parameter choice:
+
+1) ...
+2) ...
+3) ...
+4) ...
+
+Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:
+
+- Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm?
+
+- How does the searching space look like, that is, geometrically, where are the projection bases located in the space?
+
+(mention optim difficulties and split the simulated data section into the examples!!!)
 
@@
 
-## Visual diagnostic plots {#vis-diag-plots}
-
-Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation. We will present:
-
-1) static plots to explore the index value,
-
-2) animated plots to explore the projection basis and,
-
-3) a self-contained example to optimise a complex index function.
-
-In the first two sections, we will first provide a toy example that is easy to grasp and then more examples that can help us to understand the algorithm and parameter choice.
-Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:
-
-- Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm?
-
-- How does the searching space look like, that is, geometrically, where are the projection bases located in the space?
-
-The first question can be answered using a static plot with x-axis showing the progression of the optimisation and y-axis showing the value of the objective function.
-The second question can be addressed via visualising the rotating high dimensional space or its projection on the reduced 2D space.
-Thus animated visualisation is needed to perceive the optimisation path in the searching space.
-
-<!-- [This seems to be a hard task to see the rotation of high dimensional space, but the way humans understand the dimension of a 3D object is either through rotating the physical 3D object or using shade and line type to annotate the 3D object in a paper or electronically.
-Thus we can do a similar rotation on the screen to precieve even higher dimension.
-there is likely to be a learning curve - remember when learning geometry in school, it takes a while to get used to see 3D cube on a paper].
--->
-
-Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
-To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.
-
-<!-- The plot design of visual diagnostics depends on the characteristics of the variables to plot.
-If the searching points are of interest [point geom is not good; need summarise.
-While updating points are more manageable -> point geom].
-However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
-[stratefy may help?
-- may need more work here.] -->
-
-## Explore the value of objective function {#static}
-
-<!-- In tour, all the points recorded in the global object can be roughly divided into two broad categories: searching points and interpolating points -->
-
-<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
-The points for target bases is also included in the searching points and there is one such point per `tries`.
--->
-
-<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
-
-<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
-
-### Searching points
-
-A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimum progressively.
-We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
-In the plot, points from those iterations towards the end will occupy the vast majority of the plot space.
-This motivates to use summarisation.
-Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).
-
-Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it does not report the number of points in each box.
-We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
-Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument.
-A line geometry is also added to link the points with the largest index value in each iteration.
-This helps to visualise the improvement made in each iteration.
-Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:
-
-- Layer 1: boxplot geom
-  - data: group by $j$ and filter the observations in the group that have count greater than `cutoff = 15`.
-  - x: $j$ is mapped to the x-axis
-  - y: the statistical transformed index value: $Q_{I^{\prime}_t}(q)$ is mapped to the y-axis where $Q_X(q)$, $q = 0, 25, 50, 75, 100$ finds the qth-quantile of $X$ and $I^{\prime}_t$ denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
-
-- Layer 2: point geom
-  - data: group by $j$ and filter the observations in the group that have count less than `cutoff = 15`.
-  - x: $j$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-- Layer 3: line geom
-  - data: filter the points with the highest index value in group $j$
-  - x: $j$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-#### Toy example: exploring searching points
-
-We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
-Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
-The index value of the searching points are shown in Figure \ref{points}.
-The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
-We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.
-
-```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
-The left plot does not use the plot space efficiently to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
-
-dt <- holes_2d_better_max_tries %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-dt %>% explore_trace_search()
-
-```
-
-### Interpolating points
-
-Sometimes, rather than exploring the searching points, we may be interested in exploring the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
-Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed.
-The slightly modified plot definition is shown below:
-
-- Layer 1: point geom
-  - data: filter the observations with $S = \text{interpolation}$ and mutate $t$ to be the row number of the subsetted tibble
-  - x: $t$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-- Layer 2: line geom
-  - using line geometry for the same data and aesthetics
-
-#### A more complex example: Interruption
+## A more complex example: Interruption
 
@@
 
-### Polishing points
+## Another example: Polish
 
@@
 
-#### Another example: Polish
+---
 
@@
 
-\newpage
-
-## Explore searching space {#animated}
-
-In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot.
-An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
-The visualisation can thus be defined as
-
-<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
-This gives a matrix of dimension $n$ by $(p*d)$.
-Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
-By definition, the principal components maximise the variance of the projection bases in the reduced space.
-Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
-Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix -->
-
-<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
-
-<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
-The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
-
-<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
-THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
-colour).
--->
-
-- Layer 1: point geom
-  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
-  - x: the projected basis on the first principal component
-  - y: the projected basis on the second principal component
-  - colour: $V$ is mapped to the colour aesthetic
-
-<!-- In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
-[need to think further about how different type of variables can be shown].
--->
-
-### A toy example: understand different stage of search_geodesic
-
-*Example: understand search_geodesic* [feel like this example is merely explaining search geodesic algorithm, so maybe introduce the animated plot here?
-xxx] `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
-Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and a line search is then run on the geodesic direction to find the target basis.
-The starting and target bases are then interpolated.
-In the next iteration, the target basis becomes the current basis and then the procedures continues.
-
-```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-
-holes_1d_geo %>% explore_proj_pca(col = info)
-
-#xxx the info level need to be reorder to show the start and directional search points better.
-```
-
-### A more complex example: Choosing the initial value for polishing parameter
+## A more complex example: Choosing the initial value for polishing parameter
 
@@
 <!-- ``` -->
-
-\newpage
 
@@
 
-# Application {#application}
 # Implementation: Ferrn package {#implementation}

### 2020-08-19 paper/paper.tex
rearranging Sec.4&5 to 4. diagnostics and 5. application

@@
 \hypertarget{vis-diag}{%
-\section{Visual diagnostic system}\label{vis-diag}}
+\section{Visual diagnostics}\label{vis-diag}}
 
@@
 
-\hypertarget{simulated-data}{%
-\subsection{Simulated data}\label{simulated-data}}
+\hypertarget{plot-definition}{%
+\subsection{Plot definition}\label{plot-definition}}
+
+Since the global object is already tidy, not further tidying steps is
+needed, while certain data wrangling steps
+\citep{wickham2016rfordatascience} are still needed to transform the
+global object into desirable format for one particular visualisation. To
+emphasize on this good practice of data analysis, we will describe the
+transformation steps needed for each diagnostic plots before stepping
+into visualisation.
+
+\hypertarget{static}{%
+\subsubsection{Explore the value of objective function}\label{static}}
+
+\hypertarget{searching-points}{%
+\paragraph{Searching points}\label{searching-points}}
+
+A primary interest of diagnosing an optimisation algorithm is to study
+how it finds its optimum progressively. We could plot the index value
+across its natural ordering, however, different iterations may have
+different number of points and, towards the end of the search there
+could easily be hundreds of bases being tested before the target basis
+is found. In the plot, points from those iterations towards the end will
+occupy the vast majority of the plot space. This motivates to use
+summarisation. Rather than knowing the index value of \emph{every}
+basis, we are more interested to have a general summary of all the index
+value in that iteration and more importantly, the basis with the largest
+index value (since it prescribes the next geodesic interpolation and
+future searches).
+
+Boxplot is a suitable candidate that provides five points summary of the
+data, however it has one drawback: it does not report the number of
+points in each box. We may risk losing information on how many points it
+takes to find the target basis by displaying the boxplot alone for all
+\texttt{tries}. Thus, the number of point in each iteration is displayed
+at the bottom of each box and we provide options to switch iteration
+with small number of points to a point geometry, which is achieved via a
+\texttt{cutoff} argument. A line geometry is also added to link the
+points with the largest index value in each iteration. This helps to
+visualise the improvement made in each iteration. Using the concept of
+\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
+exploring index value can be defined in three layers as:
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: boxplot geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: group by \(j\) and filter the observations in the group that
+    have count greater than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\)
+    is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\)
+    finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the
+    index value of all the searching bases defined in Matrix
+    \ref{eq:data-structure}.
+  \end{itemize}
+\item
+  Layer 2: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: group by \(j\) and filter the observations in the group that
+    have count less than \texttt{cutoff\ =\ 15}.
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
+  \end{itemize}
+\item
+  Layer 3: line geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: filter the points with the highest index value in group \(j\)
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
+  \end{itemize}
+\end{itemize}
+
+\hypertarget{toy-example-exploring-searching-points}{%
+\subparagraph{Toy example: exploring searching
+points}\label{toy-example-exploring-searching-points}}
+
+We choose variables \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
+\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
+Parameter \texttt{search\_f\ =\ tour::search\_better} and
+\texttt{max.tries\ =\ 500} is used. The index value of the searching
+points are shown in Figure \ref{points}. The label at the bottom
+indicates the number of observations in each iteration and facilitates
+the choice of \texttt{cutoff} argument (by default
+\texttt{cutoff\ =\ 15}). We learn that the \texttt{search\_better}
+quickly finds better projection basis with higher index value at first
+and then takes longer to find a better one later.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/points-tries-1.pdf}
+\caption{\label{points}A comparison of plotting the same search points
+with different plot designs. The left plot does not use the plot space
+efficiently to convey information from the plot while the right plot
+provides good summarisation of data and number of points in each tries.}
+\end{figure}
+
+\hypertarget{interpolating-points}{%
+\paragraph{Interpolating points}\label{interpolating-points}}
+
+Sometimes, rather than exploring the searching points, we may be
+interested in exploring the points on the interpolation path (target and
+interpolating points) since these points will be played by the tour
+animation. Since interpolation paths are geodesically the shortest, a
+summarisation using boxplot geometry is no longer needed. The slightly
+modified plot definition is shown below:
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: filter the observations with \(S = \text{interpolation}\) and
+    mutate \(t\) to be the row number of the subsetted tibble
+  \item
+    x: \(t\) is mapped to the x-axis
+  \item
+    y: \(I\) is mapped to the y-axis
+  \end{itemize}
+\item
+  Layer 2: line geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    using line geometry for the same data and aesthetics
+  \end{itemize}
+\end{itemize}
+
+\hypertarget{toy-example}{%
+\subparagraph{Toy example}\label{toy-example}}
+
+\hypertarget{explore-searching-space}{%
+\subsubsection{Explore searching space}\label{explore-searching-space}}
+
+In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
+are usually of dimension \(p \times d\) and hence can't be visualised in
+a 2D plot. An option to explore the searching space of these bases is to
+explore a reduced space via principal component analysis (PCA). The
+visualisation can thus be defined as
+
+\begin{itemize}
+\tightlist
+\item
+  Layer 1: point geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: subset the basis of interest and arrange into a matrix format;
+    perform PCA on the basis matrix and compute the projected basis on
+    the first two principal components; bind the variables from the
+    original global object and form a tibble
+  \item
+    x: the projected basis on the first principal component
+  \item
+    y: the projected basis on the second principal component
+  \item
+    colour: \(V\) is mapped to the colour aesthetic
+  \end{itemize}
+\end{itemize}
+
+\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
+\paragraph{A toy example: understand different stage of
+search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
+
+\texttt{search\_geodesic} is a two-stage ascending algorithm with four
+different stages in the search and a PCA plot useful to understand how
+the algorithm progresses and the relative position of each basis in the
+PCA projected 2D space. Starting from the start basis, a directional
+search is conducted in a narrow neighbourhood on five random directions.
+The best one is picked and a line search is then run on the geodesic
+direction to find the target basis. The starting and target bases are
+then interpolated. In the next iteration, the target basis becomes the
+current basis and then the procedures continues.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/pca-1.pdf}
+\caption{\label{pca}PCA plot of search geodesic colouring by info allows
+for better understanding of each stage in the geodesic search}
+\end{figure}
+
+\newpage
+
+\hypertarget{application}{%
+\section{Application}\label{application}}
+
+Below we will present several examples of diagnosing different aspects
+of the projection pursuit optimisation. We will present more examples
+that can help us to understand the algorithm and parameter choice:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  \ldots{}
+\item
+  \ldots{}
+\item
+  \ldots{}
+\item
+  \ldots{}
+\end{enumerate}
+
+Remember the research question we raised earlier, the purpose of visual
+diagnostics is to understand:
+
+\begin{itemize}
+\item
+  Whether the algorithm has successfully found the maximum and how the
+  index value changes throughout the algorithm?
+\item
+  How does the searching space look like, that is, geometrically, where
+  are the projection bases located in the space?
+\end{itemize}
+
+(mention optim difficulties and split the simulated data section into
+the examples!!!)
 
@@
 
-\hypertarget{vis-diag-plots}{%
-\subsection{Visual diagnostic plots}\label{vis-diag-plots}}
-
-Below we will present several examples of diagnosing different aspects
-of the projection pursuit optimisation. We will present:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\item
-  static plots to explore the index value,
-\item
-  animated plots to explore the projection basis and,
-\item
-  a self-contained example to optimise a complex index function.
-\end{enumerate}
-
-In the first two sections, we will first provide a toy example that is
-easy to grasp and then more examples that can help us to understand the
-algorithm and parameter choice. Remember the research question we raised
-earlier, the purpose of visual diagnostics is to understand:
-
-\begin{itemize}
-\item
-  Whether the algorithm has successfully found the maximum and how the
-  index value changes throughout the algorithm?
-\item
-  How does the searching space look like, that is, geometrically, where
-  are the projection bases located in the space?
-\end{itemize}
-
-The first question can be answered using a static plot with x-axis
-showing the progression of the optimisation and y-axis showing the value
-of the objective function. The second question can be addressed via
-visualising the rotating high dimensional space or its projection on the
-reduced 2D space. Thus animated visualisation is needed to perceive the
-optimisation path in the searching space.
-
-Since the global object is already tidy, not further tidying steps is
-needed, while certain data wrangling steps
-\citep{wickham2016rfordatascience} are still needed to transform the
-global object into desirable format for one particular visualisation. To
-emphasize on this good practice of data analysis, we will describe the
-transformation steps needed for each diagnostic plots before stepping
-into visualisation.
-
-\hypertarget{static}{%
-\subsection{Explore the value of objective function}\label{static}}
-
-\hypertarget{searching-points}{%
-\subsubsection{Searching points}\label{searching-points}}
-
-A primary interest of diagnosing an optimisation algorithm is to study
-how it finds its optimum progressively. We could plot the index value
-across its natural ordering, however, different iterations may have
-different number of points and, towards the end of the search there
-could easily be hundreds of bases being tested before the target basis
-is found. In the plot, points from those iterations towards the end will
-occupy the vast majority of the plot space. This motivates to use
-summarisation. Rather than knowing the index value of \emph{every}
-basis, we are more interested to have a general summary of all the index
-value in that iteration and more importantly, the basis with the largest
-index value (since it prescribes the next geodesic interpolation and
-future searches).
-
-Boxplot is a suitable candidate that provides five points summary of the
-data, however it has one drawback: it does not report the number of
-points in each box. We may risk losing information on how many points it
-takes to find the target basis by displaying the boxplot alone for all
-\texttt{tries}. Thus, the number of point in each iteration is displayed
-at the bottom of each box and we provide options to switch iteration
-with small number of points to a point geometry, which is achieved via a
-\texttt{cutoff} argument. A line geometry is also added to link the
-points with the largest index value in each iteration. This helps to
-visualise the improvement made in each iteration. Using the concept of
-\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
-exploring index value can be defined in three layers as:
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: boxplot geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: group by \(j\) and filter the observations in the group that
-    have count greater than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\)
-    is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\)
-    finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the
-    index value of all the searching bases defined in Matrix
-    \ref{eq:data-structure}.
-  \end{itemize}
-\item
-  Layer 2: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: group by \(j\) and filter the observations in the group that
-    have count less than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\item
-  Layer 3: line geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: filter the points with the highest index value in group \(j\)
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{toy-example-exploring-searching-points}{%
-\paragraph{Toy example: exploring searching
-points}\label{toy-example-exploring-searching-points}}
-
-We choose variables \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
-\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
-Parameter \texttt{search\_f\ =\ tour::search\_better} and
-\texttt{max.tries\ =\ 500} is used. The index value of the searching
-points are shown in Figure \ref{points}. The label at the bottom
-indicates the number of observations in each iteration and facilitates
-the choice of \texttt{cutoff} argument (by default
-\texttt{cutoff\ =\ 15}). We learn that the \texttt{search\_better}
-quickly finds better projection basis with higher index value at first
-and then takes longer to find a better one later.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/points-tries-1.pdf}
-\caption{\label{points}A comparison of plotting the same search points
-with different plot designs. The left plot does not use the plot space
-efficiently to convey information from the plot while the right plot
-provides good summarisation of data and number of points in each tries.}
-\end{figure}
-
-\hypertarget{interpolating-points}{%
-\subsubsection{Interpolating points}\label{interpolating-points}}
-
-Sometimes, rather than exploring the searching points, we may be
-interested in exploring the points on the interpolation path (target and
-interpolating points) since these points will be played by the tour
-animation. Since interpolation paths are geodesically the shortest, a
-summarisation using boxplot geometry is no longer needed. The slightly
-modified plot definition is shown below:
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: filter the observations with \(S = \text{interpolation}\) and
-    mutate \(t\) to be the row number of the subsetted tibble
-  \item
-    x: \(t\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\item
-  Layer 2: line geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    using line geometry for the same data and aesthetics
-  \end{itemize}
-\end{itemize}
-
 \hypertarget{a-more-complex-example-interruption}{%
-\paragraph{A more complex example:
+\subsection{A more complex example:
 Interruption}\label{a-more-complex-example-interruption}}
@@
 
-\hypertarget{polishing-points}{%
-\subsubsection{Polishing points}\label{polishing-points}}
+\hypertarget{another-example-polish}{%
+\subsection{Another example: Polish}\label{another-example-polish}}
 
@@
 
-\hypertarget{another-example-polish}{%
-\paragraph{Another example: Polish}\label{another-example-polish}}
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 
@@
 
-\newpage
-
-\hypertarget{animated}{%
-\subsection{Explore searching space}\label{animated}}
-
-In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
-are usually of dimension \(p \times d\) and hence can't be visualised in
-a 2D plot. An option to explore the searching space of these bases is to
-explore a reduced space via principal component analysis (PCA). The
-visualisation can thus be defined as
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: subset the basis of interest and arrange into a matrix format;
-    perform PCA on the basis matrix and compute the projected basis on
-    the first two principal components; bind the variables from the
-    original global object and form a tibble
-  \item
-    x: the projected basis on the first principal component
-  \item
-    y: the projected basis on the second principal component
-  \item
-    colour: \(V\) is mapped to the colour aesthetic
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
-\subsubsection{A toy example: understand different stage of
-search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
-
-\emph{Example: understand search\_geodesic} {[}feel like this example is
-merely explaining search geodesic algorithm, so maybe introduce the
-animated plot here? xxx{]} \texttt{search\_geodesic} is a two-stage
-ascending algorithm with four different stages in the search and a PCA
-plot useful to understand how the algorithm progresses and the relative
-position of each basis in the PCA projected 2D space. Starting from the
-start basis, a directional search is conducted in a narrow neighbourhood
-on five random directions. The best one is picked and a line search is
-then run on the geodesic direction to find the target basis. The
-starting and target bases are then interpolated. In the next iteration,
-the target basis becomes the current basis and then the procedures
-continues.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/pca-1.pdf}
-\caption{\label{pca}PCA plot of search geodesic colouring by info allows
-for better understanding of each stage in the geodesic search}
-\end{figure}
-
 \hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
-\subsubsection{A more complex example: Choosing the initial value for
+\subsection{A more complex example: Choosing the initial value for
 polishing
@@
 
-\newpage
-
 \hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
@@
 
-\hypertarget{explore-searching-space}{%
-\subsubsection{Explore searching space}\label{explore-searching-space}}
+\hypertarget{explore-searching-space-1}{%
+\subsubsection{Explore searching
+space}\label{explore-searching-space-1}}
 
@@
 
-\hypertarget{application}{%
-\section{Application}\label{application}}
-
 \hypertarget{implementation}{%

### 2020-08-19 paper/paper.Rmd
change the notation for # of iteration to be j

@@
 \begin{equation}
-x^{(k + 1)} = x^{(k)} + \alpha_k* d^{(k)}
+x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)}
 \label{eq:line-search}
@@
 
-where $d^{(k)}$ is the searching direction in iteration $k$, and $\alpha_k$ is the step-size.
-Strictly speaking, $\alpha_k$ is chosen by another minimisation of $f(x^{(k)} + \alpha* d^{(k)})$ with respect to $\alpha$ and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of $\alpha_k$ is attained [@curry1944method].
+where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
+Strictly speaking, $\alpha_k$ is chosen by another minimisation of $f(x^{(j)} + \alpha* d^{(j)})$ with respect to $\alpha$ and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of $\alpha_j$ is attained [@curry1944method].
 In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation.
-A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(k+1)}> f^{(k)}$ and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
+A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$ and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
 
@@
 
-Modern development of the line search methods focuses on proposing different computation on the seraching direction: $d^{(k)}$ and various approximations on the step size: $\alpha_k$ catered for practical optimisation problems.
+Modern development of the line search methods focuses on proposing different computation on the seraching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems.
 The specific problem context we are interested in is called projection pursuit guided tour.
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] modifies `search_better` based on a non-increasing cooling scheme $T(l)$.
-Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probabiliby $$P= \min\left\{\exp\left[-\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{\cdot}$ denotes the index value of a given basis.
+Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] modifies `search_better` based on a non-increasing cooling scheme $T(j)$.
+Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probabiliby $$P= \min\left\{\exp\left[-\frac{I_{cur} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
 This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
@@
 \SetAlgoLined
-    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I_{l} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{l}$} \;
+    Compute $I_{j} = f(\mathbf{A}_{j})$ and $T(k) = \frac{T_0}{\log(j + 1)}$\;
+      \eIf{$I_{j} > I_{\text{cur}}$}{
+        \KwRet{$\mathbf{A}_{j}$} \;
       }{
-        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{j}}{T(j)}\right],1\right\}$\;
         Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
         \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{l}$} \;
+           \KwRet{$\mathbf{A}_{j}$} \;
         }
@@
   \SetKwInOut{Output}{output}
-    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n$, $\delta$}
+    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $j_{\max}$, $n$, $\delta$}
     \Output{$\mathbf{A}_{**}$}
   initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{l}: \mathbf{A}_{l+9}$ within a small neighbourhood $\delta$\;
-    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $l < * < l+ 9$\;
+  Set $j = 1$\;
+  \While{$j < j_{\max}$}{
+    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{j}: \mathbf{A}_{j+9}$ within a small neighbourhood $\delta$\;
+    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $j < * < j+ 9$\;
     Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
@@
       }
-    $l = l + 1$\;
+    $j = j + 1$\;
   }

### 2020-08-19 paper/paper.tex
change the notation for # of iteration to be j

@@
 
-where \(d^{(k)}\) is the searching direction in iteration \(k\), and
-\(\alpha_k\) is the step-size. Strictly speaking, \(\alpha_k\) is chosen
-by another minimisation of \(f(x^{(k)} + \alpha* d^{(k)})\) with respect
+where \(d^{(j)}\) is the searching direction in iteration \(j\), and
+\(\alpha_j\) is the step-size. Strictly speaking, \(\alpha_k\) is chosen
+by another minimisation of \(f(x^{(j)} + \alpha* d^{(j)})\) with respect
 to \(\alpha\) and theoretical results have demonstrated the global
-convergence of the algorithm when the exact minimisation of \(\alpha_k\)
+convergence of the algorithm when the exact minimisation of \(\alpha_j\)
 is attained \citep{curry1944method}. In practice, this second
@@
 to impose a mandatory decrease in the objective function for each
-iteration: \(f^{(k+1)}> f^{(k)}\) and despite we lose the guarantee on
+iteration: \(f^{(j+1)}> f^{(j)}\) and despite we lose the guarantee on
 global convergence, this approach turns out to be efficient in practical
@@
 Modern development of the line search methods focuses on proposing
-different computation on the seraching direction: \(d^{(k)}\) and
-various approximations on the step size: \(\alpha_k\) catered for
+different computation on the seraching direction: \(d^{(j)}\) and
+various approximations on the step size: \(\alpha_j\) catered for
 practical optimisation problems. The specific problem context we are
@@
 modifies \texttt{search\_better} based on a non-increasing cooling
-scheme \(T(l)\). Given an initial \(T_0\), the temperature at iteration
-\(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate
+scheme \(T(j)\). Given an initial \(T_0\), the temperature at iteration
+\(k\) is defined as \(T(j) = \frac{T_0}{\log(j + 1)}\). When a candidate
 basis fails to have an index value larger than the current basis,
@@
 probabiliby
-\[P= \min\left\{\exp\left[-\frac{I_{cur} - I_{l}}{T(l)}\right],1\right\}\]
-where \(I_{\cdot}\) denotes the index value of a given basis. This
+\[P= \min\left\{\exp\left[-\frac{I_{cur} - I_{j}}{T(j)}\right],1\right\}\]
+where \(I_{(\cdot)}\) denotes the index value of a given basis. This
 implementation allows the algorithm to jump out of a local maximum and

### 2020-08-19 paper/paper.Rmd
change the notation for optim constraint

@@
 Now we begin to formulate the optimisation problem.
-Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T = [\mathbf{a}_1, \cdots, \mathbf{a}_d]$, where $\mathbf{a}_i \in \mathbb{R}^{p}$, satisfies the following optimisation problem:
+Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
 
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
-s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
+&s.t.  \mathbf{A}^{\prime} \mathbf{A} = I_d
 \end{align}
 
-Where $\delta_{ij}$ is the kronecker delta that takes 1 if $i = j$ and 0 otherwise.
+where $I_d$ is the d-dimentional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be a orthonogal matrix with each column vector being orthornormal.
 

### 2020-08-19 paper/paper.tex
change the notation for optim constraint

@@
 generated starting basis \(\mathbf{A}_1\), projection pursuit finds the
-final projection basis
-\(\mathbf{A}_T = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
-\(\mathbf{a}_i \in \mathbb{R}^{p}\), satisfies the following
+final projection basis \(\mathbf{A}_T\) that satisfies the following
 optimisation problem:
@@
 &\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
-s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
+&s.t.  \mathbf{A}^{\prime} \mathbf{A} = I_d
 \end{align}
 
-Where \(\delta_{ij}\) is the kronecker delta that takes 1 if \(i = j\)
-and 0 otherwise.
+where \(I_d\) is the d-dimentional identity matrix. The constraint
+requires the projection basis \(\mathbf{A}\) to be a orthonogal matrix
+with each column vector being orthornormal.
 

### 2020-08-20 paper/paper.Rmd
Section 4.2.1 searching plots

@@
 
+\newpage
+
 ## Plot definition
 
-Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
+<!-- Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
 To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.
+-->
 
@@
 
-### Explore the value of objective function {#static}
+<!-- ### Explore the value of objective function {#static} -->
 
@@
 
-#### Searching points
+### Explore searching points
 
-A primary interest of diagnosing an optimisation algorithm is to study how it finds its optimum progressively.
-We could plot the index value across its natural ordering, however, different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
+A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum.
+One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
+When plotting the searching observations as points in a plot, the space each iteartion take will be proportional to the number of points in that iteration.
+
+<!-- different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
 In the plot, points from those iterations towards the end will occupy the vast majority of the plot space.
@@
 Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).
+-->
 
-Boxplot is a suitable candidate that provides five points summary of the data, however it has one drawback: it does not report the number of points in each box.
+<!-- .
 We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
@@
 This helps to visualise the improvement made in each iteration.
-Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as:
+Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as: -->
+
+Another option is to use summarisation for each iteration.
+Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis.
+This could be remedised by adding more layers using the concept of gramma of graphics.
+A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration.
+Further, a option to switch betwen displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
+This is avhieved via a `cutoff` parameter.
+Below we define the searching point plot based on four different component layers
 
@@
 
-##### Toy example: exploring searching points
+- Layer 4: label geom
+  - data: A count table of the number of observation in each iteration
+  - x: $j$ is mapped to the x-axis
+  - y: the y-axis for each label is $0.99 * \text{MIN}$, where $\text{MIN}$ is the smallest index value in the whole data object
+  - label: the number of observation is mapped to the label
 
-We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
+<!-- *Example: exploring searching points* We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
 Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
-The index value of the searching points are shown in Figure \ref{points}.
+The index value of the searching points are shown in Figure \ref{toy-search}.
 The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
 We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.
+-->
 
-```{r points-tries,fig.asp=0.5, fig.cap="\\label{points}A comparison of plotting the same search points with different plot designs.
-The left plot does not use the plot space efficiently to convey information from the plot while the right plot provides good summarisation of data and number of points in each tries."}
+Figure \ref{toy-search} presents a comparison between the two plotting options discussed above.
+Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasize the searches in the last few iterations.
+The second plot design evenly presents the summarised information for each iteration while allow for a switch to the full information for the iteration with small number of observations.
+
+```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
+In upper plot, points from the last three iteration spans the vast majority of the plotting space leaving the first few iterationsbeing squeezed together whilst the lower plot spaces each iteration evenly and present summary information easy to read."}
 
@@
 
-dt %>% explore_trace_search()
+p1 <- dt %>% explore_trace_search()
+
+p2 <- dt %>% ggplot() + 
+  geom_point(aes(x = id, y = index_val, col = as.factor(tries))) + 
+  theme(legend.position = "none")
+
+p2 / p1
 
@@
 
-#### Interpolating points
+\newpage
+
+### Explore interpolating points
 
@@
 
-##### Toy example
+*Toy example* A toy example is shown in Figure \ref{toy-interp}.
 
-### Explore searching space
+```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
+interp <- holes_2d_better_max_tries %>% 
+  filter(info == "interpolation") %>% 
+  mutate(id = row_number())
+
+interp %>% 
+  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
+  geom_point() + 
+  geom_line() + 
+  theme(legend.position = "bottom")
+```
+
+\newpage
+
+### Explore the searching space
 
@@
 
-#### A toy example: understand different stage of search_geodesic
-
-`search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
+*A toy example: understand different stage of search_geodesic* `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
 Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
@@
 In the next iteration, the target basis becomes the current basis and then the procedures continues.
+A toy example is shown in Figure \ref{toy-pca}.
 
-```{r pca, fig.cap = "\\label{pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
 

### 2020-08-20 paper/paper.tex
Section 4.2.1 searching plots

@@
 
+\newpage
+
 \hypertarget{plot-definition}{%
@@
 
-Since the global object is already tidy, not further tidying steps is
-needed, while certain data wrangling steps
-\citep{wickham2016rfordatascience} are still needed to transform the
-global object into desirable format for one particular visualisation. To
-emphasize on this good practice of data analysis, we will describe the
-transformation steps needed for each diagnostic plots before stepping
-into visualisation.
-
-\hypertarget{static}{%
-\subsubsection{Explore the value of objective function}\label{static}}
-
-\hypertarget{searching-points}{%
-\paragraph{Searching points}\label{searching-points}}
+\hypertarget{explore-searching-points}{%
+\subsubsection{Explore searching
+points}\label{explore-searching-points}}
 
 A primary interest of diagnosing an optimisation algorithm is to study
-how it finds its optimum progressively. We could plot the index value
-across its natural ordering, however, different iterations may have
-different number of points and, towards the end of the search there
-could easily be hundreds of bases being tested before the target basis
-is found. In the plot, points from those iterations towards the end will
-occupy the vast majority of the plot space. This motivates to use
-summarisation. Rather than knowing the index value of \emph{every}
-basis, we are more interested to have a general summary of all the index
-value in that iteration and more importantly, the basis with the largest
-index value (since it prescribes the next geodesic interpolation and
-future searches).
-
-Boxplot is a suitable candidate that provides five points summary of the
-data, however it has one drawback: it does not report the number of
-points in each box. We may risk losing information on how many points it
-takes to find the target basis by displaying the boxplot alone for all
-\texttt{tries}. Thus, the number of point in each iteration is displayed
-at the bottom of each box and we provide options to switch iteration
-with small number of points to a point geometry, which is achieved via a
-\texttt{cutoff} argument. A line geometry is also added to link the
-points with the largest index value in each iteration. This helps to
-visualise the improvement made in each iteration. Using the concept of
-\emph{gramma of graphics} \citep{wickham2010layered}, the plot for
-exploring index value can be defined in three layers as:
+how it progressively finds its optimum. One way of doing it is to plot
+the index value across its natural ordering \(t\), however, it usually
+takes the algorithm much longer to find a better basis than the current
+one towards the end of the search. When plotting the searching
+observations as points in a plot, the space each iteartion take will be
+proportional to the number of points in that iteration.
+
+Another option is to use summarisation for each iteration. Boxplot is a
+suitable candidate that provides five points summary of the data,
+however, there are two pieces of information missing from the boxplot:
+1) It does not report the number of points, and 2) the position of the
+last basis. This could be remedised by adding more layers using the
+concept of gramma of graphics. A label geometry is added at the bottom
+of the plot to show the number of points in each iteration and a line
+geometry links the last basis in each iteration. Further, a option to
+switch betwen displaying points and boxplot geometry is helpful since
+point geometry can be more intuitive for the iteration with few
+observations. This is avhieved via a \texttt{cutoff} parameter. Below we
+define the searching point plot based on four different component layers
 
@@
   \end{itemize}
-\end{itemize}
-
-\hypertarget{toy-example-exploring-searching-points}{%
-\subparagraph{Toy example: exploring searching
-points}\label{toy-example-exploring-searching-points}}
-
-We choose variables \texttt{x1}, \texttt{x2}, \texttt{x3}, \texttt{x8},
-\texttt{x9} and \texttt{x10} to perform a 2D projection with tour.
-Parameter \texttt{search\_f\ =\ tour::search\_better} and
-\texttt{max.tries\ =\ 500} is used. The index value of the searching
-points are shown in Figure \ref{points}. The label at the bottom
-indicates the number of observations in each iteration and facilitates
-the choice of \texttt{cutoff} argument (by default
-\texttt{cutoff\ =\ 15}). We learn that the \texttt{search\_better}
-quickly finds better projection basis with higher index value at first
-and then takes longer to find a better one later.
+\item
+  Layer 4: label geom
+
+  \begin{itemize}
+  \tightlist
+  \item
+    data: A count table of the number of observation in each iteration
+  \item
+    x: \(j\) is mapped to the x-axis
+  \item
+    y: the y-axis for each label is \(0.99 * \text{MIN}\), where
+    \(\text{MIN}\) is the smallest index value in the whole data object
+  \item
+    label: the number of observation is mapped to the label
+  \end{itemize}
+\end{itemize}
+
+Figure \ref{toy-search} presents a comparison between the two plotting
+options discussed above. Using the natural order to plot searching
+points distorts the point from the first few iterations and
+over-emphasize the searches in the last few iterations. The second plot
+design evenly presents the summarised information for each iteration
+while allow for a switch to the full information for the iteration with
+small number of observations.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/points-tries-1.pdf}
-\caption{\label{points}A comparison of plotting the same search points
-with different plot designs. The left plot does not use the plot space
-efficiently to convey information from the plot while the right plot
-provides good summarisation of data and number of points in each tries.}
+\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
+\caption{\label{toy-search}A comparison of plotting the same search
+points with different plot designs. In upper plot, points from the last
+three iteration spans the vast majority of the plotting space leaving
+the first few iterationsbeing squeezed together whilst the lower plot
+spaces each iteration evenly and present summary information easy to
+read.}
 \end{figure}
 
-\hypertarget{interpolating-points}{%
-\paragraph{Interpolating points}\label{interpolating-points}}
+\newpage
+
+\hypertarget{explore-interpolating-points}{%
+\subsubsection{Explore interpolating
+points}\label{explore-interpolating-points}}
 
@@
 
-\hypertarget{toy-example}{%
-\subparagraph{Toy example}\label{toy-example}}
-
-\hypertarget{explore-searching-space}{%
-\subsubsection{Explore searching space}\label{explore-searching-space}}
+\emph{Toy example} A toy example is shown in Figure \ref{toy-interp}.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
+\caption{\label{toy-interp} This is a toy example of plotting
+interpolated points}
+\end{figure}
+
+\newpage
+
+\hypertarget{explore-the-searching-space}{%
+\subsubsection{Explore the searching
+space}\label{explore-the-searching-space}}
 
@@
 
-\hypertarget{a-toy-example-understand-different-stage-of-search_geodesic}{%
-\paragraph{A toy example: understand different stage of
-search\_geodesic}\label{a-toy-example-understand-different-stage-of-search_geodesic}}
-
+\emph{A toy example: understand different stage of search\_geodesic}
 \texttt{search\_geodesic} is a two-stage ascending algorithm with four
@@
 then interpolated. In the next iteration, the target basis becomes the
-current basis and then the procedures continues.
+current basis and then the procedures continues. A toy example is shown
+in Figure \ref{toy-pca}.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/pca-1.pdf}
-\caption{\label{pca}PCA plot of search geodesic colouring by info allows
-for better understanding of each stage in the geodesic search}
+\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
+\caption{\label{toy-pca}PCA plot of search geodesic colouring by info
+allows for better understanding of each stage in the geodesic search}
 \end{figure}
@@
 
-\hypertarget{explore-searching-space-1}{%
-\subsubsection{Explore searching
-space}\label{explore-searching-space-1}}
+\hypertarget{explore-searching-space}{%
+\subsubsection{Explore searching space}\label{explore-searching-space}}
 

### 2020-08-20 paper/paper.Rmd
more changes in plot definition section

@@
 
-Sometimes, rather than exploring the searching points, we may be interested in exploring the points on the interpolation path (target and interpolating points) since these points will be played by the tour animation.
-Since interpolation paths are geodesically the shortest, a summarisation using boxplot geometry is no longer needed.
-The slightly modified plot definition is shown below:
+Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. The plot definition is as follows:
 
@@
 
-*Toy example* A toy example is shown in Figure \ref{toy-interp}.
+Figure \ref{toy-interp} shows the plot of interpolating points and it shows the algorithm starts from around 0.8 and with interpolation, gradually reaches over 0.95.
 
@@
 
-In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are usually of dimension $p \times d$ and hence can't be visualised in a 2D plot.
+In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are of dimension $p \times d$ with $p$ usually greater than two and hence can't be visualised in a 2D plot.
 An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
@@
 
-<!-- In this way, the first two principle components will take up the x and y axis and all the other information will be mapped using other aesthetic attribution.
-[need to think further about how different type of variables can be shown].
--->
-
-*A toy example: understand different stage of search_geodesic* `search_geodesic` is a two-stage ascending algorithm with four different stages in the search and a PCA plot useful to understand how the algorithm progresses and the relative position of each basis in the PCA projected 2D space.
+Figure 5 shows the searching space plot of a geodesic search, colored by different searching stages.
 Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and a line search is then run on the geodesic direction to find the target basis.
-The starting and target bases are then interpolated.
-In the next iteration, the target basis becomes the current basis and then the procedures continues.
-A toy example is shown in Figure \ref{toy-pca}.
+The best one is picked and an exact search finds the stepsize that maximises the index value and output the corresponding projection basis.
 

### 2020-08-20 paper/paper.tex
more changes in plot definition section

@@
 
-Sometimes, rather than exploring the searching points, we may be
-interested in exploring the points on the interpolation path (target and
-interpolating points) since these points will be played by the tour
-animation. Since interpolation paths are geodesically the shortest, a
-summarisation using boxplot geometry is no longer needed. The slightly
-modified plot definition is shown below:
+Sometimes, we may be interested in exploring the points on the
+interpolation path since these are the points that will be played by the
+tour animation. The plot definition is as follows:
 
@@
 
-\emph{Toy example} A toy example is shown in Figure \ref{toy-interp}.
+Figure \ref{toy-interp} shows the plot of interpolating points and it
+shows the algorithm starts from around 0.8 and with interpolation,
+gradually reaches over 0.95.
 
@@
 In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
-are usually of dimension \(p \times d\) and hence can't be visualised in
-a 2D plot. An option to explore the searching space of these bases is to
-explore a reduced space via principal component analysis (PCA). The
-visualisation can thus be defined as
+are of dimension \(p \times d\) with \(p\) usually greater than two and
+hence can't be visualised in a 2D plot. An option to explore the
+searching space of these bases is to explore a reduced space via
+principal component analysis (PCA). The visualisation can thus be
+defined as
 
@@
 
-\emph{A toy example: understand different stage of search\_geodesic}
-\texttt{search\_geodesic} is a two-stage ascending algorithm with four
-different stages in the search and a PCA plot useful to understand how
-the algorithm progresses and the relative position of each basis in the
-PCA projected 2D space. Starting from the start basis, a directional
+Figure 5 shows the searching space plot of a geodesic search, colored by
+different searching stages. Starting from the start basis, a directional
 search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and a line search is then run on the geodesic
-direction to find the target basis. The starting and target bases are
-then interpolated. In the next iteration, the target basis becomes the
-current basis and then the procedures continues. A toy example is shown
-in Figure \ref{toy-pca}.
+The best one is picked and an exact search finds the stepsize that
+maximises the index value and output the corresponding projection basis.
 

### 2020-08-20 paper/paper.Rmd
change of subsection titles

@@
 
-## Global object
+## Data structure
 
@@
 
-## Plot definition
+## Plot constructions
 
@@
 
-### Explore searching points
+### How hard is the optimiser working
 
@@
 ```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
-In upper plot, points from the last three iteration spans the vast majority of the plotting space leaving the first few iterationsbeing squeezed together whilst the lower plot spaces each iteration evenly and present summary information easy to read."}
+In upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
 
@@
 
-### Explore interpolating points
+### Examining the interpolation
 
@@
 
-### Explore the searching space
+### Understanding the optimiser's coverage of the search space
 
@@
 
-# Application {#application}
+# Diagnosing an optimiser {#application}
+
+[a paragraph that motivate the section: these optimisers are suspect to have problem, here is how we use the diag plot to diagnose them. ]
 
@@
 (mention optim difficulties and split the simulated data section into the examples!!!)
+
+## Simulation setup
 

### 2020-08-20 paper/paper.tex
change of subsection titles

@@
 
-\hypertarget{global-object}{%
-\subsection{Global object}\label{global-object}}
+\hypertarget{data-structure}{%
+\subsection{Data structure}\label{data-structure}}
 
@@
 
-\hypertarget{plot-definition}{%
-\subsection{Plot definition}\label{plot-definition}}
-
-\hypertarget{explore-searching-points}{%
-\subsubsection{Explore searching
-points}\label{explore-searching-points}}
+\hypertarget{plot-constructions}{%
+\subsection{Plot constructions}\label{plot-constructions}}
+
+\hypertarget{how-hard-is-the-optimiser-working}{%
+\subsubsection{How hard is the optimiser
+working}\label{how-hard-is-the-optimiser-working}}
 
@@
 points with different plot designs. In upper plot, points from the last
-three iteration spans the vast majority of the plotting space leaving
-the first few iterationsbeing squeezed together whilst the lower plot
-spaces each iteration evenly and present summary information easy to
+three iterations span the vast majority of the plotting space leaving
+the first few iterations being squeezed together whilst the lower plot
+spaces each iteration evenly and presents summary information easy to
 read.}
@@
 
-\hypertarget{explore-interpolating-points}{%
-\subsubsection{Explore interpolating
-points}\label{explore-interpolating-points}}
+\hypertarget{examining-the-interpolation}{%
+\subsubsection{Examining the
+interpolation}\label{examining-the-interpolation}}
 
@@
 
-\hypertarget{explore-the-searching-space}{%
-\subsubsection{Explore the searching
-space}\label{explore-the-searching-space}}
+\hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
+\subsubsection{Understanding the optimiser's coverage of the search
+space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
 
@@
 \hypertarget{application}{%
-\section{Application}\label{application}}
+\section{Diagnosing an optimiser}\label{application}}
+
+{[}a paragraph that motivate the section: these optimisers are suspect
+to have problem, here is how we use the diag plot to diagnose them. {]}
 
@@
 the examples!!!)
+
+\hypertarget{simulation-setup}{%
+\subsection{Simulation setup}\label{simulation-setup}}
 

### 2020-08-20 paper/paper.Rmd
motiational paragraph for data structure; mention of tidy data at the end of Sec 4.1

@@
 
-# Visual diagnostics{#vis-diag}
+# Visual diagnostics {#vis-diag}
+
+To be able to make diagnostic plot, the optimisation algorithm should populate a data structure that contains the key elements of the algorithm.
+When the algorithm runs, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object for future analysis.
 
 ## Data structure
-
-The idea of generalised framework for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
-In grammar of graphic, plots are not defined by its appearance (i.e.
-boxplot, histogram, scatter plot etc) but by "stacked layers".
-Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
-Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
-
-There are different ways to represent the same data in rectangular form.
-Certain cleaning steps are needed, before using ggplot2, to bring the data into a tidy data format [@wickham2014tidy] where 1) each observation forms a row; 2) each variable forms a column and 3) each type of observational unit forms a table.
-In a tidy format, data wrangling and visualisation are greatly simplified.
 
@@
 
-Global object is a data construction with key components from optimisation algorithms and is designed for making diagnostic plots easier.
-In the optimisation algorithms for projection pursuit, three key elements are 1) projection bases: $\mathbf{A}$, 2) index values: $I$ and 3) State: $S$, which labels the observation with detailed stage of searching or interpolation.
-
-Multiple iterators are needed to index the data collected at different levels.
+In the optimisation algorithms for projection pursuit, three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
+Multiple iterators are also  needed to index the data collected at different levels.
 $t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts.
 $l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
+There are other parameters that are of our interest and we denote them as *$V_{p}$*.
+In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
+A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
 
@@
 <!-- # ``` -->
-
-There are other parameters that are of our interest and we denote them as *$V_{p}$*.
-In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
-The data structure can thus be shown as in Equation \ref{eq:data-structure}.
 
@@
 
-A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.
-
-[I feel this sketch was initially useful but now since we have the data matrix and the printed output of the global object, it doesn't any additional information.
-I'm still keeping it here but we may need to remove it if it's not useful :(]
-
-```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The global object in projection pursuit guided tour."}
+```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."}
 knitr::include_graphics(path = here::here("figures/global_obj.png"))
 ```
+
+It is worth-noticing that the data structure constructed above meets the tidy data principal [@wickham2014tidy] that states
+
+1) each observation forms a row, 
+2) each variable forms a column, and 
+3) each type of observational unit forms a table
+
+The wrangling and visualisation of tidy data have been greatly simplified by the well-known dplyr[@dplyr] and ggplot2[@ggplot2] package.
 
@@
 
-<!-- Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation.
-To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation.
--->
+The idea of generalised framework for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+In grammar of graphic, plots are not defined by its appearance (i.e.
+boxplot, histogram, scatter plot etc) but by "stacked layers".
+Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
+Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
 
@@
 
-We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the global object.
+We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.
 
@@
 - Layer 1: point geom
-  - data: filter the observations with  $S = \text{interpolation}$; bind the global object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
+  - data: filter the observations with  $S = \text{interpolation}$; bind the data object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
   - x: $t$ is mapped to the x-axis
@@
 
-When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
+When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
 Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.

### 2020-08-20 paper/paper.tex
motiational paragraph for data structure; mention of tidy data at the end of Sec 4.1

@@
 
+To be able to make diagnostic plot, the optimisation algorithm should
+populate a data structure that contains the key elements of the
+algorithm. When the algorithm runs, key information regarding the
+decision variable, objective function and hyper-parameters needs to be
+recorded and stored as a data object for future analysis.
+
 \hypertarget{data-structure}{%
 \subsection{Data structure}\label{data-structure}}
+
+In the optimisation algorithms for projection pursuit, three main
+elements to record are 1) projection bases: \(\mathbf{A}\), 2) index
+values: \(I\), and 3) State: \(S\), which labels the observation with
+detailed stage in the optimisation. Multiple iterators are also needed
+to index the data collected at different levels. \(t\) is a unique
+identifier that prescribes the natural ordering of each observation;
+\(j\) is the counter for each search-and-interpolate round, which
+remains the same within one round and has an increment of one once a new
+round starts. \(l\) is the counter for each search/interpolation
+allowing us to know how many basis the algorithm has searched before
+finding one to output. There are other parameters that are of our
+interest and we denote them as \emph{\(V_{p}\)}. In projection pursuit,
+this includes \(V_1 = \text{method}\), which tags the name of the
+algorithm used and \(V_2 = \text{alpha}\), the neighbourhood parameter
+that controls the size in sampling candidate bases. A matrix notation of
+the data structure is presented in Equation \ref{eq:data-structure}.
+
+\begin{equation}
+\left[
+\begin{array}{c|ccc|cc|cc}
+t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
+\hline
+1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
+\hline
+2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
+3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{31}  & V_{32}\\
+\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
+\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
+\hline
+\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
+\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
+\hline
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+\hline
+\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
+\hline
+\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
+\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
+\hline
+\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
+\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
+T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
+\end{array}
+\right]
+= 
+\left[
+\begin{array}{c}
+\text{column name} \\
+\hline
+\text{search (start basis)} \\
+\hline
+\text{search} \\
+\text{search} \\
+\vdots \\
+\text{search (accepted basis)} \\
+\hline
+\text{interpolate} \\
+\text{interpolate} \\
+\vdots \\
+\text{interpolate} \\
+\hline
+\vdots \\
+\hline
+\text{search} \\
+\vdots \\
+\text{search (final basis)} \\
+\hline
+\text{interpolate} \\
+\vdots \\
+\text{interpolate} \\
+\hline
+\text{search (no output)} \\
+\vdots \\
+\text{search (no output)} \\
+\end{array}
+\right]
+\label{eq:data-structure}
+\end{equation}
+
+where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that we deliberately
+denote the last round of search as \(j = J+1\) and in that round there
+is no output/interpolation basis and the algorithm terminates. This
+notation allows us to denote the last complete search-and-interpolate
+round as round \(J\) and hence the final basis is \(A_T\) and highest
+index value found is \(I_T\).
+
+{[}outside the paper: I find the notation of current/target basis is
+confusing because the target basis in round \(j\) becomes the current
+basis in round \(j+1\). Also, when we start to have polish, the target
+basis may not be the current basis in the next round\ldots{} The place
+where current/target is most appropriate is probably when describing the
+interpolation where the first one is always the current basis and the
+last is always the target basis. I think it is better to leave this
+language in the code{]}
+
+\begin{figure}
+\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The data structure in projection pursuit guided tour.}\label{fig:glb-obj}
+\end{figure}
+
+It is worth-noticing that the data structure constructed above meets the
+tidy data principal \citep{wickham2014tidy} that states
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  each observation forms a row,
+\item
+  each variable forms a column, and
+\item
+  each type of observational unit forms a table
+\end{enumerate}
+
+The wrangling and visualisation of tidy data have been greatly
+simplified by the well-known dplyr\citep{dplyr} and
+ggplot2\citep{ggplot2} package.
+
+\newpage
+
+\hypertarget{plot-constructions}{%
+\subsection{Plot constructions}\label{plot-constructions}}
 
@@
 overlaying different layers.
-
-There are different ways to represent the same data in rectangular form.
-Certain cleaning steps are needed, before using ggplot2, to bring the
-data into a tidy data format \citep{wickham2014tidy} where 1) each
-observation forms a row; 2) each variable forms a column and 3) each
-type of observational unit forms a table. In a tidy format, data
-wrangling and visualisation are greatly simplified.
-
-Global object is a data construction with key components from
-optimisation algorithms and is designed for making diagnostic plots
-easier. In the optimisation algorithms for projection pursuit, three key
-elements are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\)
-and 3) State: \(S\), which labels the observation with detailed stage of
-searching or interpolation.
-
-Multiple iterators are needed to index the data collected at different
-levels. \(t\) is a unique identifier that prescribes the natural
-ordering of each observation; \(j\) is the counter for each
-search-and-interpolate round, which remains the same within one round
-and has an increment of one once a new round starts. \(l\) is the
-counter for each search/interpolation allowing us to know how many basis
-the algorithm has searched before finding one to output.
-
-There are other parameters that are of our interest and we denote them
-as \emph{\(V_{p}\)}. In projection pursuit, this includes
-\(V_1 = \text{method}\), which tags the name of the algorithm used and
-\(V_2 = \text{alpha}\), the neighbourhood parameter that controls the
-size in sampling candidate bases. The data structure can thus be shown
-as in Equation \ref{eq:data-structure}.
-
-\begin{equation}
-\left[
-\begin{array}{c|ccc|cc|cc}
-t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
-\hline
-1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
-\hline
-2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
-3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{31}  & V_{32}\\
-\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
-\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
-\hline
-\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
-\hline
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-\hline
-\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
-\hline
-\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
-\hline
-\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
-\end{array}
-\right]
-= 
-\left[
-\begin{array}{c}
-\text{column name} \\
-\hline
-\text{search (start basis)} \\
-\hline
-\text{search} \\
-\text{search} \\
-\vdots \\
-\text{search (accepted basis)} \\
-\hline
-\text{interpolate} \\
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\vdots \\
-\hline
-\text{search} \\
-\vdots \\
-\text{search (final basis)} \\
-\hline
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\text{search (no output)} \\
-\vdots \\
-\text{search (no output)} \\
-\end{array}
-\right]
-\label{eq:data-structure}
-\end{equation}
-
-where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that we deliberately
-denote the last round of search as \(j = J+1\) and in that round there
-is no output/interpolation basis and the algorithm terminates. This
-notation allows us to denote the last complete search-and-interpolate
-round as round \(J\) and hence the final basis is \(A_T\) and highest
-index value found is \(I_T\).
-
-{[}outside the paper: I find the notation of current/target basis is
-confusing because the target basis in round \(j\) becomes the current
-basis in round \(j+1\). Also, when we start to have polish, the target
-basis may not be the current basis in the next round\ldots{} The place
-where current/target is most appropriate is probably when describing the
-interpolation where the first one is always the current basis and the
-last is always the target basis. I think it is better to leave this
-language in the code{]}
-
-A sketch of the global object for projection pursuit guided tour is
-presented in Figure \ref{fig:glb-obj}.
-
-{[}I feel this sketch was initially useful but now since we have the
-data matrix and the printed output of the global object, it doesn't any
-additional information. I'm still keeping it here but we may need to
-remove it if it's not useful :({]}
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
-\end{figure}
-
-\newpage
-
-\hypertarget{plot-constructions}{%
-\subsection{Plot constructions}\label{plot-constructions}}
 
@@
 optimiser \texttt{search\_better} and below shows the first ten rows of
-the global object.
+the data object.
 
@@
     data: filter the observations with \(S = \text{interpolation}\);
-    bind the global object from optimisation and interpolation and form
+    bind the data object from optimisation and interpolation and form
     polishing; mutate \(t\) to be the row number of the binded tibble.
@@
 
-When the optimisation ends, the global object will be stored and printed
+When the optimisation ends, the data object will be stored and printed
 (it can be turned off by supplying argument \texttt{print\ =\ FALSE}).

### 2020-08-20 paper/paper.Rmd
example plots of interpolation

@@
 
-The idea of generalised framework for diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+With a constructed data object, the construction of diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
 In grammar of graphic, plots are not defined by its appearance (i.e.
@@
 
-```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
+```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
 In upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
@@
 
-p2 <- dt %>% ggplot() + 
-  geom_point(aes(x = id, y = index_val, col = as.factor(tries))) + 
-  theme(legend.position = "none")
-
-p2 / p1
+p1
 
@@
 
-Figure \ref{toy-interp} shows the plot of interpolating points and it shows the algorithm starts from around 0.8 and with interpolation, gradually reaches over 0.95.
+Figure \ref{toy-interp} presents the interpolation of three different tour paths.
+The upper plot shows a desirable interpolating in each iteration with the index value being progressively and monotonically increasing.
+While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
+The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern.
+In iteration three, the probablistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probablistic acceptance, the interpolation is now non-monotical.
+When the target basis has a higher index value, iteration eight reaches this basis by first decreases the index value.
+While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnose of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
 
-```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
+```{r toy-interp, fig.height = 8, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
 interp <- holes_2d_better_max_tries %>% 
@@
 
-interp %>% 
+p1 <- interp %>% 
   ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
@@
   theme(legend.position = "bottom")
+
+p2 <- interrupt_no %>% explore_trace_interp() + 
+  theme(legend.position = "bottom")
+
+p3 <- holes_2d_better_random %>% explore_trace_interp() + 
+  theme(legend.position = "bottom")
+
+p1 / p2 / p3
 ```

### 2020-08-20 paper/paper.tex
example plots of interpolation

@@
 
-The idea of generalised framework for diagnostic plots is inspired by
-the concept of grammar of graphic \citep{wickham2010layered}, which
-powers the primary graphical system in R, ggplot2 \citep{ggplot2}. In
-grammar of graphic, plots are not defined by its appearance
-(i.e.~boxplot, histogram, scatter plot etc) but by ``stacked layers''.
-Using this design, ggplot does not have to develop a gazillion of
-functions that each produces a different type of plot from a different
-data structure. Instead, it aesthetically maps variables (and its
-statistical transformation) in a dataset to different geometric objects
-(points, lines, box-and-whisker etc) and builds the plot through
+With a constructed data object, the construction of diagnostic plots is
+inspired by the concept of grammar of graphic
+\citep{wickham2010layered}, which powers the primary graphical system in
+R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are not defined
+by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by
+``stacked layers''. Using this design, ggplot does not have to develop a
+gazillion of functions that each produces a different type of plot from
+a different data structure. Instead, it aesthetically maps variables
+(and its statistical transformation) in a dataset to different geometric
+objects (points, lines, box-and-whisker etc) and builds the plot through
 overlaying different layers.
@@
 
-Figure \ref{toy-interp} shows the plot of interpolating points and it
-shows the algorithm starts from around 0.8 and with interpolation,
-gradually reaches over 0.95.
+Figure \ref{toy-interp} presents the interpolation of three different
+tour paths. The upper plot shows a desirable interpolating in each
+iteration with the index value being progressively and monotonically
+increasing. While in the middle plot, the increases towards the target
+basis is not monotonical in the last two iterations and interpolated
+basis with higher index value can be found on the tour path. The lower
+plot is constructed using \texttt{search\_better\_random}, where a basis
+with smaller index value has a probabilistic chance of being accepted
+and we can observe a much more involved pattern. In iteration three, the
+probablistic acceptance produces a monotonically decreasing
+interpolation whilst in iteration five, six and seven, with also a
+probablistic acceptance, the interpolation is now non-monotical. When
+the target basis has a higher index value, iteration eight reaches this
+basis by first decreases the index value. While the interpolation plot
+shows different possibilities of the change in index value on the tour
+path, the diagnose of the validity of each pattern and the related
+optimisation algorithms will be postponed to Section \ref{application}.
 

### 2020-08-24 paper/paper.Rmd
minor fixes section 1

@@
 In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied to more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
+This diagnostic system can be applied more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
 An R package, ferrn, has been created to implement this diagnostic system.
@@
 The remainder of the paper is organised as follows. 
-Section \ref{optim} provides the literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour. 
+Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour. 
 Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
 Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples. 
-Section \ref{application} shows the applicaiton of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
+Section \ref{application} shows the application of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
 Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
@@
 
-Modern development of the line search methods focuses on proposing different computation on the seraching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems.
+Modern development of the line search methods focuses on proposing different computation on the searching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems.
 The specific problem context we are interested in is called projection pursuit guided tour.

### 2020-08-24 numbat_talk/3min_talk_slide.Rmd
numbat 3 mins talk materials

---
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      navigation:
          ratio: '16:9'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# css: ["presdefault.css"] 
```

```{r}
files <- paste0("data/", list.files(here::here("data")))
purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))

source(here::here("source/indice.r"))
source(here::here("source/sim_data.r"))

library(tidyverse)
library(ferrn)
library(patchwork)
library(ggrepel)
```

#### Diagnostics of optimisation algorithms in projection pursuit guided tour

.pull-left[

**Tour:** exploratory data analysis tool that can detect the structure of data

Data: $X_{n \times p}$

Projection: $P_{n \times d} = X \cdot A_{p \times d}$

Define a projection pursuit index: $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$

Optimisation problem:

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
&s.t.  \mathbf{A}^{\prime} \mathbf{A} = I_d
\end{align}

Three existing algorithms:

- search_better
- search_better_random
- search_geodesic

]

.pull-right[
```{r fig.height=5}
colour <- RColorBrewer::brewer.pal(3, "Dark2")
max_1 <- kol_cdf_1d_2var_better %>% filter(index_val == max(index_val)) %>% pull(index_val)
max_2 <- kol_cdf_1d_2var_better_2 %>% filter(index_val == max(index_val)) %>% pull(index_val)

p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
  scale_colour_manual(values = colour[1]) + ggtitle("search better, seed 123456")
p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
  scale_colour_manual(values = colour[2]) + ggtitle("search better, seed 12345")

p7 <- kol_cdf_1d_2var_better_random %>% 
  explore_trace_interp(col = 0) + 
  ylim(0, 0.25) + 
  scale_colour_manual(values = colour[3]) + 
  theme(legend.position = "none") + 
  ggtitle("search better random")

((p5 | p6) / p7) &
  theme(legend.position = "none") & ylim(0, 0.25) & ylab("index value")

```

```{r fig.height=5}
interp1 <- kol_cdf_1d_2var_better 
interp2 <- kol_cdf_1d_2var_better_2
interp3 <- kol_cdf_1d_2var_better_random

dt <- bind_rows(interp1, interp2, interp3) %>% 
  mutate(algorithm = c(rep("search better 1", nrow(interp1)), 
                       rep("search better 2", nrow(interp2)), 
                       rep("search better random", nrow(interp3))),
         algorithm = fct_relevel(algorithm, c("search better 1", "search better 2", "search better random"))) %>% 
  compute_pca() %>% 
  `$`(combined)

dt %>% filter(info == "interpolation") %>% group_by(algorithm) %>% filter(id == max(id)) -> b

dt %>% 
  ggplot(aes(x = PC1, y = PC2, color = algorithm)) + 
  geom_point() +
  geom_path(alpha = 0.1) + 
  geom_point(data = b, size =10) + 
  theme(aspect.ratio = 1) + 
  scale_colour_brewer(palette = "Dark2") 
  #geom_point(data = dt %>% filter(info == "start"), color = "blue") +

```

```{r eval = FALSE}
a <- interp1 %>% compute_pca() %>% `$`(combined)

a %>% ggplot(aes(x = PC1, y = PC2, color = info)) + 
  geom_point() +  
  geom_path() + 
  geom_point(data = a %>% filter(info == "start")) + 
  theme(aspect.ratio = 1)

```

]

### 2020-08-26 paper/paper.Rmd
minor fixes sections 2 and 3

@@
 Given an optimisation problem, two basic approaches find the optimum based on different thinking.
-An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the close-form solution may not be available when the problem starts to become complex.
+An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the closed-form solution may not be available when the problem starts to become complex.
 An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution.
@@
 
-A traditional while often used in practice is an iteartive method called *line search method* [@fletcher2013practical].
+A traditional while often used in practice is an iterative method called *line search method* [@fletcher2013practical].
 In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
@@
 The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details.
-Iteratively choose different low dimension planes and interpolate between them forms a tour path.
-Different types of tour methods choose the low dimensional planes differently and we mention two other type of tour that are commonly used.
+Iteratively choosing different low dimension planes and interpolating between them forms a tour path.
+Different types of tour methods choose the low dimensional planes differently and we mention two other types of tour that are commonly used.
 A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data.
@@
 
-where $I_d$ is the d-dimentional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be a orthonogal matrix with each column vector being orthornormal.
+where $I_d$ is the d-dimensional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be an orthogonal matrix with each column vector being orthonormal.
 
@@
 In this case, we will need to either use some approximation of the gradient or turn to derivative free methods.
-Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem need to be able to find the global maximum.
-Along the way, locla maximum may also be of our interest since they could present unexpected interesting projections.
+Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem needs to be able to find the global maximum.
+Along the way, local maximum may also be of our interest since they could present unexpected interesting projections.
 There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.
@@
 
-Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient acsent.
+Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient ascent.
 
-@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
 The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path.
@@
 The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
-When the search needs to sample more than $c$ number of basis to find an accpeted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
+When the search needs to sample more than $c$ number of basis to find an accepted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
 If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops.
 The algorithm is summarised in Algorithm \ref{random-search} for one iteration.
-[mention orthonormalise to ensure the constaint is fulfilled; don't use derivative information but a random search]
+[mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]
 
@@
 
-Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] modifies `search_better` based on a non-increasing cooling scheme $T(j)$.
+Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] modifies `search_better` based on a non-increasing cooling scheme $T(j)$.
 Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probabiliby $$P= \min\left\{\exp\left[-\frac{I_{cur} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
 This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
 This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
-a one dimenional projection of the dataset with two informative variables).
+a one dimensional projection of the dataset with two informative variables).
 The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
@@
 @cook1995grand used a gradient ascent algorithm with pseudo-derivative.
-Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the graident.
-The $\delta$ parameter is usually set to be tiny to ensure the geogesic direction is a good local approximation to the graident.
+Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the gradient.
+The $\delta$ parameter is usually set to be tiny to ensure the geodesic direction is a good local approximation to the gradient.
 Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate.
 $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates.
-Algorithm \ref{search-geodesic} summarise the steps in geodesic search.
+Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
 
@@
 One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
-When plotting the searching observations as points in a plot, the space each iteartion take will be proportional to the number of points in that iteration.
+When plotting the searching observations as points in a plot, the space each iteration take will be proportional to the number of points in that iteration.
 

### 2020-08-26 paper/paper.Rmd
minor fixes section 4

@@
 
-It is worth-noticing that the data structure constructed above meets the tidy data principal [@wickham2014tidy] that states
+It is worth noticing that the data structure constructed above meets the tidy data principle [@wickham2014tidy] that states
 
@@
 One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
-When plotting the searching observations as points in a plot, the space each iteration take will be proportional to the number of points in that iteration.
+When plotting the searching observations as points in a plot, the space each iteration takes will be proportional to the number of points in that iteration.
 
@@
 Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis.
-This could be remedised by adding more layers using the concept of gramma of graphics.
+This could be remedied by adding more layers using the concept of grammar of graphics.
 A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration.
-Further, a option to switch betwen displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
-This is avhieved via a `cutoff` parameter.
+Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
+This is achieved via a `cutoff` parameter.
 Below we define the searching point plot based on four different component layers
@@
 - Layer 1: boxplot geom
-  - data: group by $j$ and filter the observations in the group that have count greater than `cutoff = 15`.
+  - data: group by $j$ and filter the observations in the groups that have count greater than `cutoff = 15`.
   - x: $j$ is mapped to the x-axis
@@
 Figure \ref{toy-search} presents a comparison between the two plotting options discussed above.
-Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasize the searches in the last few iterations.
-The second plot design evenly presents the summarised information for each iteration while allow for a switch to the full information for the iteration with small number of observations.
+Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
+The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.
 
 ```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
-In upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
+In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
 
@@
 Figure \ref{toy-interp} presents the interpolation of three different tour paths.
-The upper plot shows a desirable interpolating in each iteration with the index value being progressively and monotonically increasing.
+The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
 While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
 The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern.
-In iteration three, the probablistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probablistic acceptance, the interpolation is now non-monotical.
-When the target basis has a higher index value, iteration eight reaches this basis by first decreases the index value.
-While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnose of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
+In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
+When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
+While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
 
@@
 Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and an exact search finds the stepsize that maximises the index value and output the corresponding projection basis.
+The best one is picked and an exact search finds the stepsize that maximises the index value and outputs the corresponding projection basis.
 

### 2020-08-26 paper/paper.tex
update .pdf &.tex

@@
 learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied to more generally to help
+This diagnostic system can be applied more generally to help
 practitioner to unveil the black-box in randomised iterative
@@
 The remainder of the paper is organised as follows. Section \ref{optim}
-provides the literature review of optimisation methods, specifically the
+provides a literature review of optimisation methods, specifically the
 line search methods used in projection pursuit guided tour. Section
@@
 with some small examples. Section \ref{application} shows the
-applicaiton of how the diagnostic plots designed in section
+application of how the diagnostic plots designed in section
 \ref{vis-diag} can be used to understand and compare different
@@
 optimal solution in a finite number of steps, but a potential issue with
-it is that the close-form solution may not be available when the problem
-starts to become complex. An iterative approach, on the other hand,
-finds the optimum based on the idea of making progressive improvement to
-the current solution. An iterative method may end up finding a local
-optimum but the progressive nature of the algorithm allows the
-practitioner to decide when to stop if a desirable accuracy has been
-achieved.
-
-A traditional while often used in practice is an iteartive method called
+it is that the closed-form solution may not be available when the
+problem starts to become complex. An iterative approach, on the other
+hand, finds the optimum based on the idea of making progressive
+improvement to the current solution. An iterative method may end up
+finding a local optimum but the progressive nature of the algorithm
+allows the practitioner to decide when to stop if a desirable accuracy
+has been achieved.
+
+A traditional while often used in practice is an iterative method called
 \emph{line search method} \citep{fletcher2013practical}. In a simple
@@
 Modern development of the line search methods focuses on proposing
-different computation on the seraching direction: \(d^{(j)}\) and
+different computation on the searching direction: \(d^{(j)}\) and
 various approximations on the step size: \(\alpha_j\) catered for
@@
 interpolation and we refer readers to \citet{buja2005computational} for
-the mathematical details. Iteratively choose different low dimension
-planes and interpolate between them forms a tour path. Different types
+the mathematical details. Iteratively choosing different low dimension
+planes and interpolating between them forms a tour path. Different types
 of tour methods choose the low dimensional planes differently and we
-mention two other type of tour that are commonly used. A grand tour
+mention two other types of tour that are commonly used. A grand tour
 selects the planes randomly in the high dimensional space and hence
@@
 
-where \(I_d\) is the d-dimentional identity matrix. The constraint
-requires the projection basis \(\mathbf{A}\) to be a orthonogal matrix
-with each column vector being orthornormal.
+where \(I_d\) is the d-dimensional identity matrix. The constraint
+requires the projection basis \(\mathbf{A}\) to be an orthogonal matrix
+with each column vector being orthonormal.
 
@@
 projection pursuit as finding the basis with the largest index value,
-the optimisation problem need to be able to find the global maximum.
-Along the way, locla maximum may also be of our interest since they
+the optimisation problem needs to be able to find the global maximum.
+Along the way, local maximum may also be of our interest since they
 could present unexpected interesting projections. There is also one
@@
 \texttt{search\_geodesic} approximates the gradient information and acts
-as a non-derivative version of the gradient acsent.
+as a non-derivative version of the gradient ascent.
 
@@
 current basis \(\mathbf{A}_{\text{cur}}\) by
-\(\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{rand}\),
+\(\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\),
 where \(\alpha\) controls the radius of the sampling neighbourhood and
@@
 searching neighbourhood \(\alpha\). When the search needs to sample more
-than \(c\) number of basis to find an accpeted basis, the neighbourhood
+than \(c\) number of basis to find an accepted basis, the neighbourhood
 parameter \(\alpha\) will be reduced by half in the next iteration. If
@@
 Algorithm \ref{random-search} for one iteration. {[}mention
-orthonormalise to ensure the constaint is fulfilled; don't use
+orthonormalise to ensure the constraint is fulfilled; don't use
 derivative information but a random search{]}
@@
 Simulated annealing
-\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated}
+\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization}
 modifies \texttt{search\_better} based on a non-increasing cooling
@@
 simulated annealing gives it a second chance to be accepted with
-probabiliby
-\[P= \min\left\{\exp\left[-\frac{I_{cur} - I_{j}}{T(j)}\right],1\right\}\]
+probability
+\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}\]
 where \(I_{(\cdot)}\) denotes the index value of a given basis. This
@@
 is smaller than the number of informative variables in the dataset
-(i.e.~a one dimenional projection of the dataset with two informative
+(i.e.~a one dimensional projection of the dataset with two informative
 variables). The algorithm can be written as replacing line 5-11 of
@@
 \(\delta\) parameter and finds the most promising one to construct the
-geodesic direction as an approximation to the graident. The \(\delta\)
-parameter is usually set to be tiny to ensure the geogesic direction is
-a good local approximation to the graident. Once the searching direction
+geodesic direction as an approximation to the gradient. The \(\delta\)
+parameter is usually set to be tiny to ensure the geodesic direction is
+a good local approximation to the gradient. Once the searching direction
 is determined, the optimiser finds the best projection
@@
 algorithm repeats the above steps until \(l_{\max}\) is reached and the
-searching terminates. Algorithm \ref{search-geodesic} summarise the
+searching terminates. Algorithm \ref{search-geodesic} summarises the
 steps in geodesic search.
@@
 
-It is worth-noticing that the data structure constructed above meets the
-tidy data principal \citep{wickham2014tidy} that states
+It is worth noticing that the data structure constructed above meets the
+tidy data principle \citep{wickham2014tidy} that states
 
@@
 one towards the end of the search. When plotting the searching
-observations as points in a plot, the space each iteartion take will be
+observations as points in a plot, the space each iteration takes will be
 proportional to the number of points in that iteration.
@@
 1) It does not report the number of points, and 2) the position of the
-last basis. This could be remedised by adding more layers using the
-concept of gramma of graphics. A label geometry is added at the bottom
+last basis. This could be remedied by adding more layers using the
+concept of grammar of graphics. A label geometry is added at the bottom
 of the plot to show the number of points in each iteration and a line
-geometry links the last basis in each iteration. Further, a option to
-switch betwen displaying points and boxplot geometry is helpful since
+geometry links the last basis in each iteration. Further, an option to
+switch between displaying points and boxplot geometry is helpful since
 point geometry can be more intuitive for the iteration with few
-observations. This is avhieved via a \texttt{cutoff} parameter. Below we
+observations. This is achieved via a \texttt{cutoff} parameter. Below we
 define the searching point plot based on four different component layers
@@
   \item
-    data: group by \(j\) and filter the observations in the group that
+    data: group by \(j\) and filter the observations in the groups that
     have count greater than \texttt{cutoff\ =\ 15}.
@@
 points distorts the point from the first few iterations and
-over-emphasize the searches in the last few iterations. The second plot
+over-emphasizes the searches in the last few iterations. The second plot
 design evenly presents the summarised information for each iteration
-while allow for a switch to the full information for the iteration with
-small number of observations.
+while allowing for a switch to the full information for the iteration
+with small number of observations.
 
@@
 \caption{\label{toy-search}A comparison of plotting the same search
-points with different plot designs. In upper plot, points from the last
-three iterations span the vast majority of the plotting space leaving
-the first few iterations being squeezed together whilst the lower plot
-spaces each iteration evenly and presents summary information easy to
-read.}
+points with different plot designs. In the upper plot, points from the
+last three iterations span the vast majority of the plotting space
+leaving the first few iterations being squeezed together whilst the
+lower plot spaces each iteration evenly and presents summary information
+easy to read.}
 \end{figure}
@@
 Figure \ref{toy-interp} presents the interpolation of three different
-tour paths. The upper plot shows a desirable interpolating in each
+tour paths. The upper plot shows a desirable interpolation in each
 iteration with the index value being progressively and monotonically
@@
 and we can observe a much more involved pattern. In iteration three, the
-probablistic acceptance produces a monotonically decreasing
+probabilistic acceptance produces a monotonically decreasing
 interpolation whilst in iteration five, six and seven, with also a
-probablistic acceptance, the interpolation is now non-monotical. When
+probabilistic acceptance, the interpolation is now non-monotonical. When
 the target basis has a higher index value, iteration eight reaches this
-basis by first decreases the index value. While the interpolation plot
+basis by first decreasing the index value. While the interpolation plot
 shows different possibilities of the change in index value on the tour
-path, the diagnose of the validity of each pattern and the related
+path, the diagnosis of the validity of each pattern and the related
 optimisation algorithms will be postponed to Section \ref{application}.
@@
 The best one is picked and an exact search finds the stepsize that
-maximises the index value and output the corresponding projection basis.
+maximises the index value and outputs the corresponding projection
+basis.
 

### 2020-08-26 paper/paper.Rmd
the index function should be a mapping from the projected data space (R^{nxd}) rather than R^{pxd}

@@
 Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
-Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
+Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
 A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].

### 2020-08-26 paper/paper.tex
the index function should be a mapping from the projected data space (R^{nxd}) rather than R^{pxd}

@@
 \(\mathbf{A}_{p\times d}\) is the projection basis. Define
-\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index
+\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
 function that maps the projection basis \(\mathbf{A}\) onto an index

### 2020-08-31 paper/paper.Rmd
change on sec 5 motivation and simulation setup

@@
 
-[a paragraph that motivate the section: these optimisers are suspect to have problem, here is how we use the diag plot to diagnose them. ]
+<!-- [a paragraph that motivate the section: these optimisers are suspect to have problem, here is how we use the diag plot to diagnose them. ] -->
 
-Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation.
-We will present more examples that can help us to understand the algorithm and parameter choice:
+The three algorithms introduced in Section \ref{tour} have experienced several practical issues in application.
+[initial problems with tourr optim, 2-3 sentences].
+These issues can be studied and investigated using the diagnostic plots constructed above.
 
-1) ...
-2) ...
-3) ...
-4) ...
+<!-- Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation.
+We will present more examples that can help us to understand the algorithm and parameter choice:  -->
 
-Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:
+<!-- 1) ... -->
+<!-- 2) ... -->
+<!-- 3) ... -->
+<!-- 4) ... -->
 
-- Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm?
+<!-- Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:  -->
 
-- How does the searching space look like, that is, geometrically, where are the projection bases located in the space?
+<!-- - Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm? -->
 
-(mention optim difficulties and split the simulated data section into the examples!!!)
+<!-- - How does the searching space look like, that is, geometrically, where are the projection bases located in the space? -->
+
+<!-- (mention optim difficulties and split the simulated data section into the examples!!!) -->
 
@@
 
-We simulate some random variables of size 1000 with different structures.
-`x1`, `x8`, `x9` and `x10` are simulated from normal distribution with zero mean and variance of one as in equation \ref{eq:sim-norm}.
-When using projection pursuit to explore the data structure based on its departure from normality, the entry in the projection basis for these variables should be close to zero in theory.
-`x2` to `x7` are mixture of normal distributions with different weights and locations.
-Equation \ref{eq:sim-x2} to \ref{eq:sim-x7} outlines the distribution where each variable is simulated from and Figure \ref{origin-data} shows the histogram of each variable except `x3`.
-All the variables are then scaled to have unit variance before running the projection pursuit.
+Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Variable `x1`, `x8`, `x9` and `x10` are normal distributed with zero mean and unit variance and `x2` to `x7` are mixture of normal distributions with different weights and locations.
+Figure \ref{origin-data} plots the histogram of each variable except `x3`.
+The mixture variables are scaled to have unit variance before running the projection pursuit.
 
@@
 
-```{r origin-data, fig.cap="\\label{origin-data} The distribution of simulated data except x3"}
+```{r origin-data, fig.cap="\\label{origin-data} The histogram of simulated selected variables."}
 origin_dt_bi <- data_mult %>% 
@@
   facet_wrap(vars(names), ncol = 3)
-```
-
-We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.
-
-[I notice that in the implementation the loop index for interpolation starts from 0 rather than 1.
-Need more investigation to see if it is easy to change it to 1 or we will have to change the definition of the data structure.]
-
-```{r code, eval = FALSE}
-set.seed(123456)
-holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                           search_f =  search_better),
-                  rescale = FALSE, verbose = TRUE)
-
-#save(temp, file = here::here("vignettes", "data", "temp.rda"))
-```
-
-```{r glb-obj-1}
-holes_1d_better %>% 
-  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
-  head(10)
 ```
@@
 
+We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.
+
+```{r code, eval = FALSE}
+set.seed(123456)
+holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+                                           search_f =  search_better),
+                  rescale = FALSE, verbose = TRUE)
+
+#save(temp, file = here::here("vignettes", "data", "temp.rda"))
+```
+
+\newpage
+
+```{r glb-obj-1}
+holes_1d_better %>% 
+  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
+  head(10)
+```
+
 # Conclusion

### 2020-08-31 paper/paper.Rmd
changes on the plot definition, good vs. bad plots

@@
 
-```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
+```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
 In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
@@
 
-p1 <- dt %>% explore_trace_search()
-
-p1
+p1 <- dt %>% explore_trace_search() + 
+  ggtitle("satisfactory")
+
+p2 <- holes_2d_better %>% explore_trace_search() + 
+  ggtitle("not satisfactory")
+
+p1 / p2
 
@@
   geom_line() + 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") + 
+  ggtitle("satisfactory")
 
 p2 <- interrupt_no %>% explore_trace_interp() + 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") + 
+  ggtitle("not satisfactory")
 
 p3 <- holes_2d_better_random %>% explore_trace_interp() + 
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") + 
+  ggtitle("maybe satisfactory?")
 
@@
 
-```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-
-holes_1d_geo %>% explore_proj_pca(col = info)
+```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+# holes_1d_geo %>% explore_proj_pca(col = info) 
+# holes_1d_better%>% explore_proj_pca(col = info) 
+# 
+# bind_rows(holes_1d_geo, holes_1d_better) %>% explore_proj_pca(col = method)
+# 
+# 
+# dt <- bind_rows(holes_1d_better, holes_1d_geo) %>% 
+#   compute_pca() %>% 
+#   `$`(combined)
+# 
+# last <- dt %>% 
+#   filter(info == "interpolation") %>% 
+#     group_by(method) %>% 
+#   filter(id == max(id))
+# 
+# start <- dt %>% filter(id == 1)
+# 
+# dt %>% 
+#   ggplot(aes(x = PC1, y = PC2, col = info)) + 
+#   geom_point() + 
+#   #facet_wrap(vars(method)) + 
+#   theme(aspect.ratio = 1, legend.position = "bottom") + 
+#   geom_point(data = last, size = 7) + 
+#   geom_point(data = start, size = 7)
+# 
+# 
+# bind_rows(holes_1d_better, holes_1d_geo) %>%
+#   filter(info == "interpolation") %>%
+#   group_by(method) %>%
+#   mutate(id = row_number()) %>%
+#   ggplot(aes(x = id, y = index_val, group = method, col = as.factor(tries))) +
+#   geom_point() +
+#   geom_line() +
+#   facet_wrap(vars(method), ncol = 2)
+# 
+# ######
+# dt <- bind_rows(holes_2d_better, holes_2d_geo) %>% 
+#   compute_pca() %>% 
+#   `$`(combined)
+# 
+# last <- dt %>% 
+#   filter(info == "interpolation") %>% 
+#     group_by(method) %>% 
+#   filter(id == max(id))
+# 
+# start <- dt %>% filter(id == 1)
+# 
+# dt %>% filter(!is.na(info)) %>% 
+#   ggplot(aes(x = PC1, y = PC2, col = info)) + 
+#   geom_point() + 
+#   geom_point(data = last, size = 7) + 
+#   geom_point(data = start, size = 7) + 
+#   facet_wrap(vars(method)) + 
+#   theme(aspect.ratio = 1, legend.position = "bottom")
+
+#####
+dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"), holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>% mutate(info = ifelse(info == "start", "new_basis", info)) %>% 
+  compute_pca() %>% 
+  `$`(combined)
+
+last <- dt %>% 
+  filter(info == "interpolation") %>% 
+  group_by(algo) %>% 
+  filter(id == max(id))
+
+start <- dt %>% filter(id == 1)
+
+dt %>% filter(!is.na(info)) %>% 
+  ggplot() + 
+  geom_point(aes(x = PC1, y = PC2, col = info)) + 
+  geom_point(data = last, aes(x = PC1, y = PC2, col = info), size = 7) + 
+  geom_point(data = start,aes(x = PC1, y = PC2, col = info), size = 7) + 
+  facet_wrap(vars(algo)) + 
+  theme(aspect.ratio = 1, legend.position = "bottom")
+
+# bind_rows(holes_2d_better, holes_2d_geo) %>% 
+#   filter(info == "interpolation") %>% 
+#   group_by(method) %>% 
+#   mutate(id = row_number()) %>% 
+#   ggplot(aes(x = id, y = index_val, group = method, col = as.factor(tries))) + 
+#   geom_point() + 
+#   geom_line() + 
+#   facet_wrap(vars(method), ncol = 1)
+
+#holes_2d_geo %>% explore_trace_interp()
 

### 2020-08-31 paper/paper.Rmd
more changes of the section title, add new section on animated plots

@@
 
-## Data structure
+## Data structure for diagnostics
 
@@
 
-\newpage
-
-## Plot constructions
-
 With a constructed data object, the construction of diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
@@
 
-### How hard is the optimiser working
+## Check how hard the optimiser is working
 
@@
 
-### Examining the interpolation
+## Examining the interpolation progress
 
@@
 ```
+
+## Animating the diagnostic plots
+
+all the plots above can be animated if wanted - importance is how the data is moving around
+
+show an example of some snap shots  of the animation and provide a link... for animated pca plot
+
+## The tour looking at itself
+
+show the path of the tour on a sphere. (full data dimension animation...)
+
+include the tour animation (geozoo)
 
@@
 
-## A more complex example: Interruption
+## A problem of not monotonic
 
@@
 
-## Another example: Polish
+## Close but not close enough
 
@@
 
-## A comprehensive example of diagnosing a noisy index function
+## Seeing the signal in the noise
 
@@
 can be seen as a function of the projection matrix $\mathbf{A}_{p \times 1}$ and hence a valid index function.
-
-### Explore index value
 
@@
 
-### Explore searching space
-
 We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.

### 2020-08-31 paper/paper.tex
more changes of the section title, add new section on animated plots

@@
 
-\hypertarget{data-structure}{%
-\subsection{Data structure}\label{data-structure}}
+\hypertarget{data-structure-for-diagnostics}{%
+\subsection{Data structure for
+diagnostics}\label{data-structure-for-diagnostics}}
 
@@
 
-\newpage
-
-\hypertarget{plot-constructions}{%
-\subsection{Plot constructions}\label{plot-constructions}}
-
 With a constructed data object, the construction of diagnostic plots is
@@
 
-\hypertarget{how-hard-is-the-optimiser-working}{%
-\subsubsection{How hard is the optimiser
-working}\label{how-hard-is-the-optimiser-working}}
+\hypertarget{check-how-hard-the-optimiser-is-working}{%
+\subsection{Check how hard the optimiser is
+working}\label{check-how-hard-the-optimiser-is-working}}
 
@@
 
-\hypertarget{examining-the-interpolation}{%
-\subsubsection{Examining the
-interpolation}\label{examining-the-interpolation}}
+\hypertarget{examining-the-interpolation-progress}{%
+\subsection{Examining the interpolation
+progress}\label{examining-the-interpolation-progress}}
 
@@
 
+\hypertarget{animating-the-diagnostic-plots}{%
+\subsection{Animating the diagnostic
+plots}\label{animating-the-diagnostic-plots}}
+
+all the plots above can be animated if wanted - importance is how the
+data is moving around
+
+show an example of some snap shots of the animation and provide a
+link\ldots{} for animated pca plot
+
+\hypertarget{the-tour-looking-at-itself}{%
+\subsection{The tour looking at
+itself}\label{the-tour-looking-at-itself}}
+
+show the path of the tour on a sphere. (full data dimension
+animation\ldots{})
+
+include the tour animation (geozoo)
+
 \newpage
@@
 
-{[}a paragraph that motivate the section: these optimisers are suspect
-to have problem, here is how we use the diag plot to diagnose them. {]}
-
-Below we will present several examples of diagnosing different aspects
-of the projection pursuit optimisation. We will present more examples
-that can help us to understand the algorithm and parameter choice:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\tightlist
-\item
-  \ldots{}
-\item
-  \ldots{}
-\item
-  \ldots{}
-\item
-  \ldots{}
-\end{enumerate}
-
-Remember the research question we raised earlier, the purpose of visual
-diagnostics is to understand:
-
-\begin{itemize}
-\item
-  Whether the algorithm has successfully found the maximum and how the
-  index value changes throughout the algorithm?
-\item
-  How does the searching space look like, that is, geometrically, where
-  are the projection bases located in the space?
-\end{itemize}
-
-(mention optim difficulties and split the simulated data section into
-the examples!!!)
+The three algorithms introduced in Section \ref{tour} have experienced
+several practical issues in application. {[}initial problems with tourr
+optim, 2-3 sentences{]}. These issues can be studied and investigated
+using the diagnostic plots constructed above.
 
@@
 
-We simulate some random variables of size 1000 with different
-structures. \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are
-simulated from normal distribution with zero mean and variance of one as
-in equation \ref{eq:sim-norm}. When using projection pursuit to explore
-the data structure based on its departure from normality, the entry in
-the projection basis for these variables should be close to zero in
-theory. \texttt{x2} to \texttt{x7} are mixture of normal distributions
-with different weights and locations. Equation \ref{eq:sim-x2} to
-\ref{eq:sim-x7} outlines the distribution where each variable is
-simulated from and Figure \ref{origin-data} shows the histogram of each
-variable except \texttt{x3}. All the variables are then scaled to have
-unit variance before running the projection pursuit.
+Random variables with different structures has been simulated and the
+distribution of each is presented in Equation \ref{eq:sim-norm} to
+\ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and
+\texttt{x10} are normal distributed with zero mean and unit variance and
+\texttt{x2} to \texttt{x7} are mixture of normal distributions with
+different weights and locations. Figure \ref{origin-data} plots the
+histogram of each variable except \texttt{x3}. The mixture variables are
+scaled to have unit variance before running the projection pursuit.
 
@@
 \includegraphics{paper_files/figure-latex/origin-data-1.pdf}
-\caption{\label{origin-data} The distribution of simulated data except
-x3}
-\end{figure}
-
-We form our first dataset using variables \texttt{x1}, \texttt{x2},
-\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
-optimiser \texttt{search\_better} and below shows the first ten rows of
-the data object.
-
-{[}I notice that in the implementation the loop index for interpolation
-starts from 0 rather than 1. Need more investigation to see if it is
-easy to change it to 1 or we will have to change the definition of the
-data structure.{]}
-
-\begin{verbatim}
-## # A tibble: 10 x 8
-##       id basis             index_val info         tries  loop method       alpha
-##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
-##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 <NA>           0.5
-##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
-##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
-##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
-##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
-##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
-##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
-##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     0 search_bett~  NA  
-##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     1 search_bett~  NA  
-## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA
-\end{verbatim}
-
-\hypertarget{a-more-complex-example-interruption}{%
-\subsection{A more complex example:
-Interruption}\label{a-more-complex-example-interruption}}
+\caption{\label{origin-data} The histogram of simulated selected
+variables.}
+\end{figure}
+
+\hypertarget{a-problem-of-not-monotonic}{%
+\subsection{A problem of not
+monotonic}\label{a-problem-of-not-monotonic}}
 
@@
 
-\hypertarget{another-example-polish}{%
-\subsection{Another example: Polish}\label{another-example-polish}}
+\hypertarget{close-but-not-close-enough}{%
+\subsection{Close but not close
+enough}\label{close-but-not-close-enough}}
 
@@
 
-\hypertarget{a-comprehensive-example-of-diagnosing-a-noisy-index-function}{%
-\subsection{A comprehensive example of diagnosing a noisy index
-function}\label{a-comprehensive-example-of-diagnosing-a-noisy-index-function}}
+\hypertarget{seeing-the-signal-in-the-noise}{%
+\subsection{Seeing the signal in the
+noise}\label{seeing-the-signal-in-the-noise}}
 
@@
 \(\mathbf{A}_{p \times 1}\) and hence a valid index function.
-
-\hypertarget{explore-index-value}{%
-\subsubsection{Explore index value}\label{explore-index-value}}
 
@@
 
-\hypertarget{explore-searching-space}{%
-\subsubsection{Explore searching space}\label{explore-searching-space}}
-
 We can also plot the searching points of all three algorithms in the
@@
 
+We form our first dataset using variables \texttt{x1}, \texttt{x2},
+\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
+optimiser \texttt{search\_better} and below shows the first ten rows of
+the data object.
+
+\newpage
+
+\begin{verbatim}
+## # A tibble: 10 x 8
+##       id basis             index_val info         tries  loop method       alpha
+##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
+##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 <NA>           0.5
+##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
+##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
+##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
+##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
+##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
+##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
+##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     0 search_bett~  NA  
+##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     1 search_bett~  NA  
+## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA
+\end{verbatim}
+
 \hypertarget{conclusion}{%

### 2020-08-31 paper/paper.tex
more changes of the section title, add new section on animated plots

@@
 
-\hypertarget{examining-the-interpolation-progress}{%
-\subsection{Examining the interpolation
-progress}\label{examining-the-interpolation-progress}}
+\hypertarget{examining-the-optimisation-progress}{%
+\subsection{Examining the optimisation
+progress}\label{examining-the-optimisation-progress}}
 
@@
 \hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
-\subsubsection{Understanding the optimiser's coverage of the search
+\subsection{Understanding the optimiser's coverage of the search
 space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
@@
 
-\hypertarget{a-problem-of-not-monotonic}{%
+\hypertarget{a-problem-of-not-monotinic}{%
 \subsection{A problem of not
-monotonic}\label{a-problem-of-not-monotonic}}
+monotinic}\label{a-problem-of-not-monotinic}}
 
@@
 \end{figure}
-
-\hypertarget{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}{%
-\subsection{A more complex example: Choosing the initial value for
-polishing
-parameter}\label{a-more-complex-example-choosing-the-initial-value-for-polishing-parameter}}
 

### 2020-09-02 paper/paper.Rmd
more changes of the section title, add new section on animated plots
spelling
apply the botanical palette to plot construction

@@
 p1 <- dt %>% explore_trace_search() + 
-  ggtitle("satisfactory")
+  ggtitle("satisfactory") + 
+  scale_color_botanical(palette = "daisy")
 
 p2 <- holes_2d_better %>% explore_trace_search() + 
-  ggtitle("not satisfactory")
+  ggtitle("not satisfactory") + 
+  scale_color_botanical(palette = "daisy")
 
@@
 
-## Examining the interpolation progress
+## Examining the optimisation progress
 
@@
 ```{r toy-interp, fig.height = 8, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
-interp <- holes_2d_better_max_tries %>% 
-  filter(info == "interpolation") %>% 
-  mutate(id = row_number())
-
-p1 <- interp %>% 
-  ggplot(aes(x = id, y = index_val, col = as.factor(tries))) + 
-  geom_point() + 
-  geom_line() + 
-  theme(legend.position = "bottom") + 
-  ggtitle("satisfactory")
-
-p2 <- interrupt_no %>% explore_trace_interp() + 
-  theme(legend.position = "bottom") + 
-  ggtitle("not satisfactory")
-
-p3 <- holes_2d_better_random %>% explore_trace_interp() + 
-  theme(legend.position = "bottom") + 
-  ggtitle("maybe satisfactory?")
+p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
+  scale_color_botanical(palette = "fern_brown")
+
+p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
+  scale_color_botanical(palette = "fern_green")
+
+p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
+  scale_color_botanical(palette = "acacia")
 
@@
 
-### Understanding the optimiser's coverage of the search space
+## Understanding the optimiser's coverage of the search space
 
@@
 #####
+# theory <- matrix(c(0,0, 1,0,0,0, 0, 1, 0, 0,0,0), ncol = 2)
+# index <- holes()
+# index(as.matrix(data_mult) %*% theory)
+# 
+# 
+# theory <-orthonormalise(matrix(c(0,0.2, 0.7,0,0,0, 0, 0.7, 0.2, 0,0,0), ncol = 2))
+# theory 
+# index <- holes()
+# index(as.matrix(data_mult) %*% get_last_basis(holes_2d_better))
+
 dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"), holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>% mutate(info = ifelse(info == "start", "new_basis", info)) %>% 
@@
   `$`(combined)
-
 last <- dt %>% 
@@
   filter(id == max(id))
-
 start <- dt %>% filter(id == 1)
-
 dt %>% filter(!is.na(info)) %>% 
@@
 
-## A more complex example: Choosing the initial value for polishing parameter
-
 *Example: initial value for polishing alpha* `search_polish` is a brute-force algorithm that evaluate 1000 points in the neighbourhood at each loop.

### 2020-09-06 paper/paper.Rmd
a quick change  by apply the most recent version of ferrn pacakge (color)

@@
 library(ggrepel)
+library(tourr)
 ```
@@
 ```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-# holes_1d_geo %>% explore_proj_pca(col = info) 
-# holes_1d_better%>% explore_proj_pca(col = info) 
-# 
-# bind_rows(holes_1d_geo, holes_1d_better) %>% explore_proj_pca(col = method)
-# 
-# 
-# dt <- bind_rows(holes_1d_better, holes_1d_geo) %>% 
-#   compute_pca() %>% 
-#   `$`(combined)
-# 
-# last <- dt %>% 
-#   filter(info == "interpolation") %>% 
-#     group_by(method) %>% 
-#   filter(id == max(id))
-# 
-# start <- dt %>% filter(id == 1)
-# 
-# dt %>% 
-#   ggplot(aes(x = PC1, y = PC2, col = info)) + 
-#   geom_point() + 
-#   #facet_wrap(vars(method)) + 
-#   theme(aspect.ratio = 1, legend.position = "bottom") + 
-#   geom_point(data = last, size = 7) + 
-#   geom_point(data = start, size = 7)
-# 
-# 
-# bind_rows(holes_1d_better, holes_1d_geo) %>%
-#   filter(info == "interpolation") %>%
-#   group_by(method) %>%
-#   mutate(id = row_number()) %>%
-#   ggplot(aes(x = id, y = index_val, group = method, col = as.factor(tries))) +
-#   geom_point() +
-#   geom_line() +
-#   facet_wrap(vars(method), ncol = 2)
-# 
-# ######
-# dt <- bind_rows(holes_2d_better, holes_2d_geo) %>% 
-#   compute_pca() %>% 
-#   `$`(combined)
-# 
-# last <- dt %>% 
-#   filter(info == "interpolation") %>% 
-#     group_by(method) %>% 
-#   filter(id == max(id))
-# 
-# start <- dt %>% filter(id == 1)
-# 
-# dt %>% filter(!is.na(info)) %>% 
-#   ggplot(aes(x = PC1, y = PC2, col = info)) + 
-#   geom_point() + 
-#   geom_point(data = last, size = 7) + 
-#   geom_point(data = start, size = 7) + 
-#   facet_wrap(vars(method)) + 
-#   theme(aspect.ratio = 1, legend.position = "bottom")
-
-#####
-# theory <- matrix(c(0,0, 1,0,0,0, 0, 1, 0, 0,0,0), ncol = 2)
-# index <- holes()
-# index(as.matrix(data_mult) %*% theory)
-# 
-# 
-# theory <-orthonormalise(matrix(c(0,0.2, 0.7,0,0,0, 0, 0.7, 0.2, 0,0,0), ncol = 2))
-# theory 
-# index <- holes()
-# index(as.matrix(data_mult) %*% get_last_basis(holes_2d_better))
-
-dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"), holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>% mutate(info = ifelse(info == "start", "new_basis", info)) %>% 
+dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"), 
+                holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>% 
+ bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>% 
+  bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>% 
+  mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>% 
   compute_pca() %>% 
-  `$`(combined)
-last <- dt %>% 
-  filter(info == "interpolation") %>% 
-  group_by(algo) %>% 
-  filter(id == max(id))
-start <- dt %>% filter(id == 1)
-dt %>% filter(!is.na(info)) %>% 
-  ggplot() + 
-  geom_point(aes(x = PC1, y = PC2, col = info)) + 
-  geom_point(data = last, aes(x = PC1, y = PC2, col = info), size = 7) + 
-  geom_point(data = start,aes(x = PC1, y = PC2, col = info), size = 7) + 
-  facet_wrap(vars(algo)) + 
-  theme(aspect.ratio = 1, legend.position = "bottom")
-
-# bind_rows(holes_2d_better, holes_2d_geo) %>% 
-#   filter(info == "interpolation") %>% 
-#   group_by(method) %>% 
-#   mutate(id = row_number()) %>% 
-#   ggplot(aes(x = id, y = index_val, group = method, col = as.factor(tries))) + 
-#   geom_point() + 
-#   geom_line() + 
-#   facet_wrap(vars(method), ncol = 1)
-
-#holes_2d_geo %>% explore_trace_interp()
-
-#xxx the info level need to be reorder to show the start and directional search points better.
+  `$`(aug)
+
+dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
+#dt %>% explore_proj_pca(facet = algo, axis = 2)
+
 ```
@@
 
-p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>% 
-  explore_trace_interp(col = method) + ggtitle("max.tries = 25")
-
-p2 <- bind_rows(holes_2d_better_max_tries, holes_2d_better_max_tries_polish) %>% 
-  explore_trace_interp(col = method) + ggtitle("max.tries = 500")
-
-(p1 | p2) &
+holes_2d_better %>% mutate(max_tries = 25) %>% 
+  dplyr::bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500)) %>% 
+  explore_trace_interp(col = max_tries, facet = max_tries) + 
   theme(legend.position = "none")
@@
 
-*Example: initial value for polishing alpha* `search_polish` is a brute-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
+<!-- *Example: initial value for polishing alpha* `search_polish` is a brute-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
 Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value.
@@
 Both specifications of initial value allow the searches to reach the same ending index values.
-
-```{r polish-alpha, fig.cap = "\\label{polish-alpha}PCA plot of two different polish alpha initialisations.
+-->
+
+<!-- ```{r polish-alpha, fig.cap = "\\label{polish-alpha}PCA plot of two different polish alpha initialisations.
 A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis.
 Both intialisations reach the same ending index values.
-"}
-p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% 
-  explore_proj_pca(col = as.factor(alpha))
-
-p2 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
-  explore_proj_pca(col = as.factor(alpha))
-
-p3 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>%
-  explore_trace_interp(col = method) + ylim(0.75, 0.95)
-
-p4 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>% 
-  explore_trace_interp(col = method) +  ylim(0.75, 0.95)
-
-((p1 | p2) /(p3 | p4)) +  plot_layout(heights = c(2, 1)) & theme(legend.position = "none")
-
-```
+"} -->
+<!-- p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>%  -->
+<!--   explore_proj_pca(col = as.factor(alpha)) -->
+
+<!-- p2 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>%  -->
+<!--   explore_proj_pca(col = as.factor(alpha)) -->
+
+<!-- p3 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% -->
+<!--   explore_trace_interp(col = method) + ylim(0.75, 0.95) -->
+
+<!-- p4 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>%  -->
+<!--   explore_trace_interp(col = method) +  ylim(0.75, 0.95) -->
+
+<!-- ((p1 | p2) /(p3 | p4)) +  plot_layout(heights = c(2, 1)) & theme(legend.position = "none")  -->
+
+<!-- ``` -->
 
@@
 Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
-colour <- RColorBrewer::brewer.pal(3, "Dark2")
+colour <- ferrn::botanical_pal()(3)
 p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
@@
 
-```{r 1d-2var-explore-proj-pca, fig.cap = "\\label{1d-2var-explore-proj-pca} The projected projection basis using principal components.
+<!-- ```{r 1d-2var-explore-proj-pca, fig.cap = "\\label{1d-2var-explore-proj-pca} The projected projection basis using principal components.
 The bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhausive search.
-The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end."}
-interp1 <- kol_cdf_1d_2var_better 
-interp2 <- kol_cdf_1d_2var_better_2
-interp3 <- kol_cdf_1d_2var_better_random
-
-bind_rows(interp1, interp2, interp3) %>% 
-  mutate(algorithm = c(rep("better1", nrow(interp1)), 
-                       rep("better2", nrow(interp2)), 
-                       rep("better_random", nrow(interp3))),
-         algorithm = fct_relevel(algorithm, c("better1", "better2", "better_random"))) %>% 
-  explore_proj_pca(col = algorithm) + 
-  scale_colour_brewer(palette = "Dark2")
-
-```
+The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end."} -->
+<!-- interp1 <- kol_cdf_1d_2var_better  -->
+<!-- interp2 <- kol_cdf_1d_2var_better_2 -->
+<!-- interp3 <- kol_cdf_1d_2var_better_random  -->
+
+<!-- dt <- bind_rows(interp1, interp2, interp3) %>%  -->
+<!--   mutate(algorithm = c(rep("better 1", nrow(interp1)),  -->
+<!--                        rep("better 2", nrow(interp2)),  -->
+<!--                        rep("better random", nrow(interp3))), -->
+<!--          algorithm = fct_relevel(algorithm, c("better 1", "better 2", "better random")))  -->
+
+<!-- dt %>%  -->
+<!--   explore_proj_pca(col = algorithm, alpha = 0.5) +  -->
+<!--   scale_colour_manual(values = color(3))  -->
+<!-- ``` -->
 

### 2020-09-06 paper/paper.tex
apply the botanical palette to plot construction
a quick change  by apply the most recent version of ferrn pacakge (color)

@@
 
-\hypertarget{a-problem-of-not-monotinic}{%
+\hypertarget{a-problem-of-not-monotonic}{%
 \subsection{A problem of not
-monotinic}\label{a-problem-of-not-monotinic}}
+monotonic}\label{a-problem-of-not-monotonic}}
 
@@
 
-\emph{Example: initial value for polishing alpha}
-\texttt{search\_polish} is a brute-force algorithm that evaluate 1000
-points in the neighbourhood at each loop. Setting an appropriate initial
-value for polish\_alpha would avoid wasting search on large vector space
-that are not likely to produce higher index value. The default initial
-value for polishing step is 0.5 and we are interested in whether this is
-an appropriate initial value to use after \texttt{search\_geodesic}. The
-problem is a 1D projection of the small dataset using
-\texttt{search\_geodesic} and followed by \texttt{search\_polish}. The
-top-left panel of Figure \ref{polish-alpha} displays all the projection
-bases on the first two principal components, coloured by the
-\texttt{polish\_alpha}. We can observe that rather than concentrating on
-the ending basis from \texttt{search\_geodesic} as what polishing step
-is designed, \texttt{search\_polish} searches a much larger vector
-space, which is unnecessary. Thus a customised smaller initial value for
-\texttt{polish\_alpha} would be ideal. One way to do this is to
-initialised \texttt{polish\_alpha} as the projection distance between
-the last two target bases. The top-right panel of Figure
-\ref{polish-alpha} shows a more desirable concentrated searching space
-near the ending basis. Both specifications of initial value allow the
-searches to reach the same ending index values.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/polish-alpha-1.pdf}
-\caption{\label{polish-alpha}PCA plot of two different polish alpha
-initialisations. A default polish alpha = 0.5 searches a larger space
-that is unncessary while a small customised initial value of polish
-alpha will search near the ending basis. Both intialisations reach the
-same ending index values.}
-\end{figure}
-
 While explore the reduced space is an initial attempt to understand the
@@
 the end.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/1d-2var-explore-proj-pca-1.pdf}
-\caption{\label{1d-2var-explore-proj-pca} The projected projection basis
-using principal components. The bases from better1 and better2 only
-search a proportion of the searching space while better\_random produces
-a more exhausive search. The large overlapping of better1 and
-better\_random is explained by the fact that both algorithms finds x2 in
-the end.}
-\end{figure}
 

### 2020-09-09 paper/paper.Rmd
figured out a way to display the animation using 16 frames - bad plot for now

@@
 #dt %>% explore_proj_pca(facet = algo, axis = 2)
-
 ```
@@
 show an example of some snap shots  of the animation and provide a link... for animated pca plot
+
+```{r animated-pca, fig.height = 8, fig.cap = "\\label{animated-pca} 16 frames from animated PCA plot ." }
+rl <- lapply(list.files(here::here("figures/animated_pca/"),full.names = TRUE)[1:16], png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+gridExtra::grid.arrange(grobs=gl)
+```
 
@@
 
+1. how does the printed data object looks like?
+
+Do I talk about the work done in the tourr package?
+
 Everything is coded up in a package. Package structure
@@
 
+2.
+
 # Conclusion

### 2020-09-10 paper/paper.Rmd
a recent update on simulation based on recent development of tourr pacakge, polish related data is pending.

@@
 The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
-p1 <- bind_rows(kol_cdf_1d_geodesic, kol_cdf_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
-
-p2 <- bind_rows(kol_cdf_1d_better, kol_cdf_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
+# p1 <- bind_rows(kol_1d_geo, kol_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
+# 
+# p2 <- bind_rows(kol_1d_better, kol_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
+# 
+# (p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
+
+p1 <- kol_1d_geo %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
+
+p2 <- kol_1d_better %>% explore_trace_interp(col = method) + ggtitle("search_better")
 
@@
 colour <- ferrn::botanical_pal()(3)
-p5 <- kol_cdf_1d_2var_better %>% explore_trace_interp(col = 0) + 
+p5 <- kol_1d_2var_better %>% explore_trace_interp(col = 0) + 
   scale_colour_manual(values = colour[2]) + ggtitle("search_better, seed 123456")
-p6 <- kol_cdf_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
+p6 <- kol_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
   scale_colour_manual(values = colour[1]) + ggtitle("search_better, seed 12345")
@@
 ```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "\\label{1d-2var-better-random}Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
-kol_cdf_1d_2var_better_random %>% 
+kol_1d_2var_better_random %>% 
   explore_trace_interp(col = 0) + 

### 2020-09-10 paper/paper.tex
figured out a way to display the animation using 16 frames - bad plot for now
a recent update on simulation based on recent development of tourr pacakge, polish related data is pending.

@@
 
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/animated-pca-1.pdf}
+\caption{\label{animated-pca} 16 frames from animated PCA plot .}
+\end{figure}
+
 \hypertarget{the-tour-looking-at-itself}{%
@@
 
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  how does the printed data object looks like?
+\end{enumerate}
+
+Do I talk about the work done in the tourr package?
+
 Everything is coded up in a package. Package structure
@@
 ##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
-##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     0 search_bett~  NA  
-##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     1 search_bett~  NA  
-## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA
+##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     1 search_bett~  NA  
+##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA  
+## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     3 search_bett~  NA
 \end{verbatim}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{1}
+\item
+\end{enumerate}
 

### 2020-09-10 paper/paper.Rmd
update eixsting alogrithm: start the first algo by search_better; tune the language

@@
 
-Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate basis in the neighbourhood whilst `search_geodesic` approximates the gradient information and acts as a non-derivative version of the gradient ascent.
-
-@posse1995projection proposed a random search algorithm that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$, where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
-The optimiser keeps sampling bases near the current basis until it finds one with higher index value than the current basis and then outputs it for guided tour to construct the interpolation path.
-A new round of search continues to find a better basis after the interpolation finishes.
-The halving parameter $c$ with default value of 30 is designed to adjust the searching neighbourhood $\alpha$.
-When the search needs to sample more than $c$ number of basis to find an accepted basis, the neighbourhood parameter $\alpha$ will be reduced by half in the next iteration.
-If the optimiser can't find a better basis within the maximum number of tries $l_{\max}$, the algorithm stops.
-The algorithm is summarised in Algorithm \ref{random-search} for one iteration.
+Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate bases in the neighbourhood, whilst `search_geodesic` is an analogue of gradient ascent on the manifold of the projection basis.
+
+`search_better` is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+$\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the search terminates and outputs the basis for guided tour to construct an interpolation path.
+The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$.
+Another termination condition is when the maximum number of iteration $l_{\max}$ is reached.
+A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$.
+Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast.
+The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
 [mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]
@@
   initialisation\;
-  Set $l = 1$ and $c = 0$\;
+  Set $l = 1$\;
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthonormalise $\mathbf{A}_{l}$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonise $\mathbf{A}_{l}$\;
     Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    \If{$I_{l} > I_{\text{cur}}$}{
+      \KwRet{$\mathbf{A}_{l}$} \;
+      }
+    $l = l + 1$\;
+  }
+  \caption{random search}
+  \label{random-search}
+\end{algorithm}
+
+Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme $T(l)$.
+Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
+This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
+a one dimensional projection of the dataset with two informative variables).
+The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+
+\begin{algorithm}
+\SetAlgoLined
+    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
       \eIf{$I_{l} > I_{\text{cur}}$}{
@@
       }{
-      $c = c + 1$\;
-      }
-    $l = l + 1$\;
-  }
-  We repeat this loop and if $c > 30$, half the $\alpha$. Reset $c$ to zero.
-  \caption{random search}
-  \label{random-search}
-\end{algorithm}
-
-Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] modifies `search_better` based on a non-increasing cooling scheme $T(j)$.
-Given an initial $T_0$, the temperature at iteration $k$ is defined as $T(j) = \frac{T_0}{\log(j + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
-This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
-a one dimensional projection of the dataset with two informative variables).
-The algorithm can be written as replacing line 5-11 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
-
-\begin{algorithm}
-\SetAlgoLined
-    Compute $I_{j} = f(\mathbf{A}_{j})$ and $T(k) = \frac{T_0}{\log(j + 1)}$\;
-      \eIf{$I_{j} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{j}$} \;
-      }{
-        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{j}}{T(j)}\right],1\right\}$\;
+        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
         Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
         \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{j}$} \;
+           \KwRet{$\mathbf{A}_{l}$} \;
         }
@@
 
-@cook1995grand used a gradient ascent algorithm with pseudo-derivative.
-Instead of computing the actual gradient of the index function with respect to the projection basis matrix, `search_geodesic` samples $2n$ bases that are randomly generated on a uniform open ball with the radius controlled by the $\delta$ parameter and finds the most promising one to construct the geodesic direction as an approximation to the gradient.
-The $\delta$ parameter is usually set to be tiny to ensure the geodesic direction is a good local approximation to the gradient.
-Once the searching direction is determined, the optimiser finds the best projection $\mathbf{A}_{**}$ on the geodesic as the candidate.
-$\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value or the algorithm repeats the above steps until $l_{\max}$ is reached and the searching terminates.
+@cook1995grand used a gradient ascent algorithm on the manifold of the projection bases.
+In gradient ascent, one first find the direction for improvment via computing the gradient information.
+In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
+A geodesic is then constructed using the current basis and one of the $2n$ bases that have the highest index value.
+If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
+In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
+In `search_geodesic`, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+The optima $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value.
+As above, another termination condition is when $l_{\max}$ is reached.
 Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
@@
   \SetKwInOut{Output}{output}
-    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $j_{\max}$, $n$, $\delta$}
+    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n = 5$, $\delta$}
     \Output{$\mathbf{A}_{**}$}
   initialisation\;
-  Set $j = 1$\;
-  \While{$j < j_{\max}$}{
-    Generate $2n$ bases in $n$ random directions: $\mathbf{A}_{j}: \mathbf{A}_{j+9}$ within a small neighbourhood $\delta$\;
-    Find the direction with the largest index value: $\mathbf{A}_{*}$ where $j < * < j+ 9$\;
+  Set $l = 1$\;
+  \While{$l < l_{\max}$}{
+    Generate $2n$ bases in a small neighbourhood, $\delta$, of $\mathbf{A}_{\text{cur}}$ and ensure orthogonality \;
+    Find the one with the largest index value: $\mathbf{A}_{*}$\;
     Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
-    Find $\mathbf{A}_{**}$ on the geodesic $\mathcal{G}$ that has the largest index value \;
+    Optimise the index value on the geodesic $\mathcal{G}$ over a 90 degree window to produce the optima $\mathbf{A}_{**}$  \;
     Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
@@
       }
-    $j = j + 1$\;
+    $l = l + 1$\;
   }

### 2020-09-10 paper/paper.tex
update eixsting alogrithm: start the first algo by search_better; tune the language

@@
 and \texttt{search\_better\_random} are derivative free methods that
-sample candidate basis in the neighbourhood whilst
-\texttt{search\_geodesic} approximates the gradient information and acts
-as a non-derivative version of the gradient ascent.
-
-\citet{posse1995projection} proposed a random search algorithm that
-samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the
-current basis \(\mathbf{A}_{\text{cur}}\) by
-\(\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\),
+sample candidate bases in the neighbourhood, whilst
+\texttt{search\_geodesic} is an analogue of gradient ascent on the
+manifold of the projection basis.
+
+\texttt{search\_better} is a random search device that samples a
+candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
+basis \(\mathbf{A}_{\text{cur}}\) by
+\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
 where \(\alpha\) controls the radius of the sampling neighbourhood and
 \(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
-same dimension as \(\mathbf{A}_{\text{cur}}\). The optimiser keeps
-sampling bases near the current basis until it finds one with higher
-index value than the current basis and then outputs it for guided tour
-to construct the interpolation path. A new round of search continues to
-find a better basis after the interpolation finishes. The halving
-parameter \(c\) with default value of 30 is designed to adjust the
-searching neighbourhood \(\alpha\). When the search needs to sample more
-than \(c\) number of basis to find an accepted basis, the neighbourhood
-parameter \(\alpha\) will be reduced by half in the next iteration. If
-the optimiser can't find a better basis within the maximum number of
-tries \(l_{\max}\), the algorithm stops. The algorithm is summarised in
-Algorithm \ref{random-search} for one iteration. {[}mention
-orthonormalise to ensure the constraint is fulfilled; don't use
-derivative information but a random search{]}
+same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
+then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis
+\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
+for guided tour to construct an interpolation path. The next iteration
+of search begins after adjusting \(\alpha\) by a cooling parameter:
+\(\alpha_{j+1} = \alpha_j * \text{cooling}\). Another termination
+condition is when the maximum number of iteration \(l_{\max}\) is
+reached. A slightly different cooling scheme has been proposed by
+\citet{posse1995projection} to include a halving parameter \(c\). Rather
+than reducing the radius of the searching neighbourhood, \(\alpha\), at
+each iteration, Posse's design only adjust \(\alpha\) if the last search
+takes more than \(c\) times to find an accepted basis to avoid the
+searching space being reduced too fast. The algorithm of
+\texttt{search\_better} is summarised in Algorithm \ref{random-search}.
+{[}mention orthonormalise to ensure the constraint is fulfilled; don't
+use derivative information but a random search{]}
 
@@
   initialisation\;
-  Set $l = 1$ and $c = 0$\;
+  Set $l = 1$\;
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = \mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthonormalise $\mathbf{A}_{l}$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonise $\mathbf{A}_{l}$\;
     Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    \eIf{$I_{l} > I_{\text{cur}}$}{
+    \If{$I_{l} > I_{\text{cur}}$}{
       \KwRet{$\mathbf{A}_{l}$} \;
-      }{
-      $c = c + 1$\;
       }
@@
   }
-  We repeat this loop and if $c > 30$, half the $\alpha$. Reset $c$ to zero.
   \caption{random search}
@@
 Simulated annealing
-\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization}
-modifies \texttt{search\_better} based on a non-increasing cooling
-scheme \(T(j)\). Given an initial \(T_0\), the temperature at iteration
-\(k\) is defined as \(T(j) = \frac{T_0}{\log(j + 1)}\). When a candidate
-basis fails to have an index value larger than the current basis,
-simulated annealing gives it a second chance to be accepted with
+\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses
+the same sampling process as \texttt{search\_better} but allow a
+probabilistic acceptance of a basis with lower index value based on a
+cooling scheme \(T(l)\). Given an initial \(T_0\), the temperature at
+iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
+candidate basis fails to have an index value larger than the current
+basis, simulated annealing gives it a second chance to be accepted with
 probability
-\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{j}}{T(j)}\right],1\right\}\]
+\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\]
 where \(I_{(\cdot)}\) denotes the index value of a given basis. This
@@
 (i.e.~a one dimensional projection of the dataset with two informative
-variables). The algorithm can be written as replacing line 5-11 of
+variables). The algorithm can be written as replacing line 5-8 of
 Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
@@
 \SetAlgoLined
-    Compute $I_{j} = f(\mathbf{A}_{j})$ and $T(k) = \frac{T_0}{\log(j + 1)}$\;
-      \eIf{$I_{j} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{j}$} \;
+    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
+      \eIf{$I_{l} > I_{\text{cur}}$}{
+        \KwRet{$\mathbf{A}_{l}$} \;
       }{
-        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{j}}{T(j)}\right],1\right\}$\;
+        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
         Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
         \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{j}$} \;
+           \KwRet{$\mathbf{A}_{l}$} \;
         }
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm with
-pseudo-derivative. Instead of computing the actual gradient of the index
-function with respect to the projection basis matrix,
-\texttt{search\_geodesic} samples \(2n\) bases that are randomly
-generated on a uniform open ball with the radius controlled by the
-\(\delta\) parameter and finds the most promising one to construct the
-geodesic direction as an approximation to the gradient. The \(\delta\)
-parameter is usually set to be tiny to ensure the geodesic direction is
-a good local approximation to the gradient. Once the searching direction
-is determined, the optimiser finds the best projection
-\(\mathbf{A}_{**}\) on the geodesic as the candidate.
+\citet{cook1995grand} used a gradient ascent algorithm on the manifold
+of the projection bases. In gradient ascent, one first find the
+direction for improvment via computing the gradient information. In
+\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
+neighbourhood of the current basis, controlled by the neighbourhood
+parameter \(\delta\). A geodesic is then constructed using the current
+basis and one of the \(2n\) bases that have the highest index value. If
+the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed
+is an analogue of the gradient information in the curved space and can
+work as the searching direction. In gradient ascent, the next step is to
+conduct a line search to find the best improvement along the gradient
+direction. In \texttt{search\_geodesic}, this step is replaced by
+optimising the index value along the geodesic direction constructed
+before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\). The optima
 \(\mathbf{A}_{**}\) is outputted for the current iteration if the
 percentage change in the index value between \(\mathbf{A}_{**}\) and
-\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value or the
-algorithm repeats the above steps until \(l_{\max}\) is reached and the
-searching terminates. Algorithm \ref{search-geodesic} summarises the
-steps in geodesic search.
+\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value. As above,
+another termination condition is when \(l_{\max}\) is reached. Algorithm
+\ref{search-geodesic} summarises the steps in geodesic search.
 

### 2020-09-10 paper/paper.Rmd
Write opening for section 5

@@
 
-<!-- [a paragraph that motivate the section: these optimisers are suspect to have problem, here is how we use the diag plot to diagnose them. ] -->
-
-The three algorithms introduced in Section \ref{tour} have experienced several practical issues in application.
-[initial problems with tourr optim, 2-3 sentences].
-These issues can be studied and investigated using the diagnostic plots constructed above.
-
-<!-- Below we will present several examples of diagnosing different aspects of the projection pursuit optimisation.
-We will present more examples that can help us to understand the algorithm and parameter choice:  -->
-
-<!-- 1) ... -->
-<!-- 2) ... -->
-<!-- 3) ... -->
-<!-- 4) ... -->
-
-<!-- Remember the research question we raised earlier, the purpose of visual diagnostics is to understand:  -->
-
-<!-- - Whether the algorithm has successfully found the maximum and how the index value changes throughout the algorithm? -->
-
-<!-- - How does the searching space look like, that is, geometrically, where are the projection bases located in the space? -->
-
-<!-- (mention optim difficulties and split the simulated data section into the examples!!!) -->
+For a particular index function, the best algorithm to optimise is related to the character of the index and the data.
+If the index function is smooth and has single maximum, all of the three algorithms introduced above can find maximum.
+When multiple optima are presented, `search_better` may stuck in the local maximum.
+When the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+In this section, examples will be presented  to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.
 

### 2020-09-10 paper/paper.tex
Write opening for section 5

@@
 
-The three algorithms introduced in Section \ref{tour} have experienced
-several practical issues in application. {[}initial problems with tourr
-optim, 2-3 sentences{]}. These issues can be studied and investigated
-using the diagnostic plots constructed above.
+For a particular index function, the best algorithm to optimise is
+related to the character of the index and the data. If the index
+function is smooth and has single maximum, all of the three algorithms
+introduced above can find maximum. When multiple optima are presented,
+\texttt{search\_better} may stuck in the local maximum. When the index
+function is non-smooth, \texttt{search\_geodesic} may even fail to find
+the maximum. In this section, examples will be presented to outline how
+the diagnostic plots can be used to compare the performance of the
+algorithms in different scenarios.
 

### 2020-09-17 paper/paper.Rmd
some coarse writing on pca & animated & tour

@@
 library(tourr)
+library(gganimate)
 ```
@@
 
-In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are of dimension $p \times d$ with $p$ usually greater than two and hence can't be visualised in a 2D plot.
-An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
+The projection bases $\mathbf{A}_{p \times d}$ live on the surface of a $p \times d$ dimension sphere, where this dimension can easily go from 5 to 10.
+This makes viewing the original space hard.
+An easier alternative is to view the reduced space via some dimensional reduction methods i.e.
+Principal component analysis.
 The visualisation can thus be defined as
@@
 
-Figure 5 shows the searching space plot of a geodesic search, colored by different searching stages.
-Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and an exact search finds the stepsize that maximises the index value and outputs the corresponding projection basis.
+Figure \ref{toy-pca} shows the first two principal components of two searching paths.
+A Hollow sphere with the same dimension is overlaid to emphasize the imposition of orthonormal constraint.
+Starting from the same position, two searching methods takes different paths to get to the two final bases with the same data projection $Y$ (the two correspond to a rotation of the data in the projection).
+The distinction between the two is that `search_better()` also involve some random evaluation in the sphere due to its sampling process while `search_geodesic()` doesn't.
+This stochastic rather than deterministic approach can be useful in more complex scenario.
 
 ```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
-                holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
- bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
-  bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
-  mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
-  compute_pca() %>%
-  `$`(aug)
-
-dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
-#dt %>% explore_proj_pca(facet = algo, axis = 2)
+# dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
+#                 holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
+#  bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
+#   bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
+#   mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
+#   compute_pca() %>%
+#   `$`(aug)
+# 
+# 
+# dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
+# #dt %>% explore_proj_pca(facet = algo, axis = 2)
+
+path1 <- holes_1d_geo$basis %>% 
+  flatten_dbl() %>% 
+  matrix(ncol = 5, byrow = TRUE)
+path2 <- holes_1d_better$basis %>% 
+  flatten_dbl() %>% 
+  matrix(ncol = 5, byrow = TRUE)
+set.seed(1)
+sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
+theoretical <- matrix(c(0, 1, 0, 0, 0,
+                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
+
+dt <- rbind(sphere, path1, path2, theoretical)
+pca <- dt %>% prcomp()
+#summary(pca)
+pca_pred <- predict(pca) %>%
+  as_tibble() %>%
+  mutate(method = c(rep("sphere", nrow(sphere)),
+                    holes_1d_geo$method,
+                    holes_1d_better$method,
+                    rep("theoretical", 2))) %>%
+  group_by(method) %>%
+  mutate(id = row_number(),
+         id = ifelse(method == "sphere", 1, id)) %>%
+  ungroup()
+
+p <- pca_pred %>%
+  ggplot(aes(x = PC1, y = PC2, color = method))+
+  geom_point() +
+  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
+             size = 10) + 
+  theme(aspect.ratio = 1) +
+  scale_color_botanical(palette = "cherry")
+
+p
 ```
@@
 
-all the plots above can be animated if wanted - importance is how the data is moving around
-
-show an example of some snap shots  of the animation and provide a link... for animated pca plot
-
-```{r animated-pca, fig.height = 8, fig.cap = "\\label{animated-pca} 16 frames from animated PCA plot ." }
-rl <- lapply(list.files(here::here("figures/animated_pca/"),full.names = TRUE)[1:16], png::readPNG)
-gl <-  lapply(rl, grid::rasterGrob)
-gridExtra::grid.arrange(grobs=gl)
+An animated plot can be useful because it allows us to track the progression of the optimsiation, especially in the case of inspecting the search space.
+
+<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->
+
+<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->
+
+<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->
+
+<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->
+
+<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->
+
+```{r cache = TRUE, eval=FALSE}
+# rl <- lapply(list.files(here::here("figures/animated_pca_1d/"),full.names = TRUE)[62:78], png::readPNG)
+# gl <-  lapply(rl, grid::rasterGrob)
+# gridExtra::grid.arrange(grobs=gl)
+p +
+  transition_states(id) +
+  shadow_mark()
 ```
@@
 
-show the path of the tour on a sphere. (full data dimension animation...)
-
-include the tour animation (geozoo)
+As mentioned previously, the orthonormal constraint regulates the projection basis to be living on the surface of a unit sphere.
+The visualisation of the projection basis on its original $p \times d$ dimension can be done through an animated rotation of the $p \times d$ diemnsion sphere with the searching path overlaid.
+The high dimensional geometry can be simulated using the package `geozoo` and the animated rotation can be done via a regular grand tour or a slice tour [@laa2020slice].
+
+Compared with the PCA plot, this animated rotation display requires skills of the viewer to preceive the high dimensional space in rotation and what the viewers gain is a much holistic view of the path in the space in the sense that PCA does a further projection on the projection basis along with the hollow sphere, which gives the view that maixmises the variance.
+While the animated rotation display gives the view from different anlges and the PCA plot consitutes one frame from the animated rotaiton display.
 

### 2020-09-17 paper/paper.tex
some coarse writing on pca & animated & tour

@@
 
-In projection pursuit, the projection bases \(\mathbf{A}_{p \times d}\)
-are of dimension \(p \times d\) with \(p\) usually greater than two and
-hence can't be visualised in a 2D plot. An option to explore the
-searching space of these bases is to explore a reduced space via
-principal component analysis (PCA). The visualisation can thus be
+The projection bases \(\mathbf{A}_{p \times d}\) live on the surface of
+a \(p \times d\) dimension sphere, where this dimension can easily go
+from 5 to 10. This makes viewing the original space hard. An easier
+alternative is to view the reduced space via some dimensional reduction
+methods i.e.~Principal component analysis. The visualisation can thus be
 defined as
@@
 
-Figure 5 shows the searching space plot of a geodesic search, colored by
-different searching stages. Starting from the start basis, a directional
-search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and an exact search finds the stepsize that
-maximises the index value and outputs the corresponding projection
-basis.
+Figure \ref{toy-pca} shows the first two principal components of two
+searching paths. A Hollow sphere with the same dimension is overlaid to
+emphasize the imposition of orthonormal constraint. Starting from the
+same position, two searching methods takes different paths to get to the
+two final bases with the same data projection \(Y\) (the two correspond
+to a rotation of the data in the projection). The distinction between
+the two is that \texttt{search\_better()} also involve some random
+evaluation in the sphere due to its sampling process while
+\texttt{search\_geodesic()} doesn't. This stochastic rather than
+deterministic approach can be useful in more complex scenario.
 
@@
 
-all the plots above can be animated if wanted - importance is how the
-data is moving around
-
-show an example of some snap shots of the animation and provide a
-link\ldots{} for animated pca plot
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/animated-pca-1.pdf}
-\caption{\label{animated-pca} 16 frames from animated PCA plot .}
-\end{figure}
+An animated plot can be useful because it allows us to track the
+progression of the optimsiation, especially in the case of inspecting
+the search space.
 
@@
 
-show the path of the tour on a sphere. (full data dimension
-animation\ldots{})
-
-include the tour animation (geozoo)
+As mentioned previously, the orthonormal constraint regulates the
+projection basis to be living on the surface of a unit sphere. The
+visualisation of the projection basis on its original \(p \times d\)
+dimension can be done through an animated rotation of the \(p \times d\)
+diemnsion sphere with the searching path overlaid. The high dimensional
+geometry can be simulated using the package \texttt{geozoo} and the
+animated rotation can be done via a regular grand tour or a slice tour
+\citep{laa2020slice}.
+
+Compared with the PCA plot, this animated rotation display requires
+skills of the viewer to preceive the high dimensional space in rotation
+and what the viewers gain is a much holistic view of the path in the
+space in the sense that PCA does a further projection on the projection
+basis along with the hollow sphere, which gives the view that maixmises
+the variance. While the animated rotation display gives the view from
+different anlges and the PCA plot consitutes one frame from the animated
+rotaiton display.
 
@@
 ##    <int> <list>                <dbl> <chr>        <dbl> <dbl> <chr>        <dbl>
-##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 <NA>           0.5
+##  1     1 <dbl[,1] [5 x 1]>     0.749 new_basis        1     1 search_bett~   0.5
 ##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5

### 2020-09-17 paper/paper.Rmd
polish writing in Sec. 4.4 - 4.6

--- 2020-09-10 paper/paper.Rmd
@@
 
-In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are of dimension $p \times d$ with $p$ usually greater than two and hence can't be visualised in a 2D plot.
-An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA).
+Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
+Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go above five (5).
+Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d > 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.
+Principal component analysis.
+To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
+
 The visualisation can thus be defined as
@@
 
-Figure 5 shows the searching space plot of a geodesic search, colored by different searching stages.
-Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions.
-The best one is picked and an exact search finds the stepsize that maximises the index value and outputs the corresponding projection basis.
+Figure \ref{toy-pca} shows the first two principal components of two search paths.
+Starting from the same position, two searching methods take different paths to get its final bases.
+There are two theoretical bases because they correspond to the same data projection $Y$ with a 180 degree rotation.
+What differentiate the two optimisers is that `search_better()` also involve random evaluation of points in the sphere during its sampling process while `search_geodesic()` doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
 
 ```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
-                holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
- bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
-  bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
-  mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
-  compute_pca() %>%
-  `$`(aug)
-
-dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
-#dt %>% explore_proj_pca(facet = algo, axis = 2)
+path1 <- holes_1d_geo$basis %>% 
+  flatten_dbl() %>% 
+  matrix(ncol = 5, byrow = TRUE)
+path2 <- holes_1d_better$basis %>% 
+  flatten_dbl() %>% 
+  matrix(ncol = 5, byrow = TRUE)
+set.seed(1)
+sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
+theoretical <- matrix(c(0, 1, 0, 0, 0,
+                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
+
+dt <- rbind(sphere, path1, path2, theoretical)
+pca <- dt %>% prcomp()
+#summary(pca)
+pca_pred <- predict(pca) %>%
+  as_tibble() %>%
+  mutate(method = c(rep("sphere", nrow(sphere)),
+                    holes_1d_geo$method,
+                    holes_1d_better$method,
+                    rep("theoretical", 2))) %>%
+  group_by(method) %>%
+  mutate(id = row_number(),
+         id = ifelse(method == "sphere", 1, id)) %>%
+  ungroup()
+
+p <- pca_pred %>%
+  ggplot(aes(x = PC1, y = PC2, color = method))+
+  geom_point() +
+  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
+             size = 10) + 
+  geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
+  theme(aspect.ratio = 1) +
+  scale_color_botanical(palette = "cherry")
+
+p
 ```
@@
 
-all the plots above can be animated if wanted - importance is how the data is moving around
-
-show an example of some snap shots  of the animation and provide a link... for animated pca plot
-
-```{r animated-pca, fig.height = 8, fig.cap = "\\label{animated-pca} 16 frames from animated PCA plot ." }
-rl <- lapply(list.files(here::here("figures/animated_pca/"),full.names = TRUE)[1:16], png::readPNG)
-gl <-  lapply(rl, grid::rasterGrob)
-gridExtra::grid.arrange(grobs=gl)
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
+
+<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->
+
+<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->
+
+<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->
+
+<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->
+
+<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->
+
+```{r cache = TRUE, eval=FALSE}
+# rl <- lapply(list.files(here::here("figures/animated_pca_1d/"),full.names = TRUE)[62:78], png::readPNG)
+# gl <-  lapply(rl, grid::rasterGrob)
+# gridExtra::grid.arrange(grobs=gl)
+ani <- p +
+  transition_states(id) +
+  shadow_mark()
+
+animate(ani, nframes = 350)
+anim_save(ani, nframes = 350, filename = "figures/animated_pca_1d")
 ```
@@
 
-show the path of the tour on a sphere. (full data dimension animation...)
-
-include the tour animation (geozoo)
+Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a stereoscopic view of the search.
+To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation.
+This can be done via a regular grand tour or a slice tour [@laa2020slice], which emphasizes the points that are close to a section of the sphere.
+
+Compare to the PCA plot, the animated rotation (tour) display gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
+Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
+
+```{r eval = FALSE}
+set.seed(1)
+sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
+path1 <- holes_1d_geo$basis %>% 
+  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
+path2 <- holes_1d_better$basis %>% 
+  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
+dt <- rbind(sphere, path1, path2)
+colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
+pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
+color <- c(rep(pal[1], nrow(sphere)),
+           rep(pal[2], nrow(path1)),
+           rep(pal[3], nrow(path2)))
+animate_xy(dt, col = color, cex_slice = 0.8)
+animate_slice(dt, col = color)
+
+```
 

### 2020-09-17 paper/paper.tex
polish writing in Sec. 4.4 - 4.6

@@
 
-The projection bases \(\mathbf{A}_{p \times d}\) live on the surface of
-a \(p \times d\) dimension sphere, where this dimension can easily go
-from 5 to 10. This makes viewing the original space hard. An easier
-alternative is to view the reduced space via some dimensional reduction
-methods i.e.~Principal component analysis. The visualisation can thus be
-defined as
+Apart from checking the progression of an optimiser, another interesting
+aspect is to visualise how the search looks like in its parameter space.
+Given the orthonormality constraint, the projection bases
+\(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\)
+dimension sphere, where the dimension can easily go above five (5).
+Visualising the search paths on the original high diemnsional sphere
+would require skills from the viewer to preceive rotation of geometry in
+higher dimensional space (d \textgreater{} 3) while an easier
+alternative is to view the reduced space via some dimension reduction
+methods i.e.~Principal component analysis. To better visual the serach
+path as an embedding of a hollow sphere, random points on the high
+dimensional sphere is generated using package \texttt{geozoo} and PCA is
+conducted on both the bases and the points on the surface of the sphere.
+
+The visualisation can thus be defined as
 
@@
 Figure \ref{toy-pca} shows the first two principal components of two
-searching paths. A Hollow sphere with the same dimension is overlaid to
-emphasize the imposition of orthonormal constraint. Starting from the
-same position, two searching methods takes different paths to get to the
-two final bases with the same data projection \(Y\) (the two correspond
-to a rotation of the data in the projection). The distinction between
-the two is that \texttt{search\_better()} also involve some random
-evaluation in the sphere due to its sampling process while
-\texttt{search\_geodesic()} doesn't. This stochastic rather than
-deterministic approach can be useful in more complex scenario.
+search paths. Starting from the same position, two searching methods
+take different paths to get its final bases. There are two theoretical
+bases because they correspond to the same data projection \(Y\) with a
+180 degree rotation. What differentiate the two optimisers is that
+\texttt{search\_better()} also involve random evaluation of points in
+the sphere during its sampling process while \texttt{search\_geodesic()}
+doesn't and this stochastic rather than deterministic approach can be
+useful in complex scenario.
 
@@
 
-An animated plot can be useful because it allows us to track the
-progression of the optimsiation, especially in the case of inspecting
-the search space.
+Animating the plots introduced above is useful, especially in the case
+of PCA polot since it shows the bases found by the optimiser in its
+natural order.
 
@@
 
-As mentioned previously, the orthonormal constraint regulates the
-projection basis to be living on the surface of a unit sphere. The
-visualisation of the projection basis on its original \(p \times d\)
-dimension can be done through an animated rotation of the \(p \times d\)
-diemnsion sphere with the searching path overlaid. The high dimensional
-geometry can be simulated using the package \texttt{geozoo} and the
-animated rotation can be done via a regular grand tour or a slice tour
-\citep{laa2020slice}.
-
-Compared with the PCA plot, this animated rotation display requires
-skills of the viewer to preceive the high dimensional space in rotation
-and what the viewers gain is a much holistic view of the path in the
-space in the sense that PCA does a further projection on the projection
-basis along with the hollow sphere, which gives the view that maixmises
-the variance. While the animated rotation display gives the view from
-different anlges and the PCA plot consitutes one frame from the animated
-rotaiton display.
+Viewing the bases on the reduced space via PCA shed some lights on the
+space the optimisers have explored, the visualisation on the original
+\(p \times d\) dimension enables a stereoscopic view of the search. To
+view a high dimensional (\(d \ge 3\)) object on a screen, an approach is
+to play the rotation of the object in animation. This can be done via a
+regular grand tour or a slice tour \citep{laa2020slice}, which
+emphasizes the points that are close to a section of the sphere.
+
+Compare to the PCA plot, the animated rotation (tour) display gives a
+more holistic view of the search of the optimiser and these additional
+information gained from the animation is cruicial. This is because in
+essense, PCA presents one reduced space that maximises the variance and
+it inherently amplifies the spread of the randomly evaluated bases in
+the search since they have larger variance than the bases on the
+interpolation path. Hence when the interest is to view the interpolation
+path in the space, an animated rotation display provides the view of
+high dimension from different angles.
 

### 2020-09-24 paper/paper.Rmd
change generated points to be grey in tour animation

@@
 
-## Optimisation problem  {#tour-optim}
+## Optimisation in the tour {#tour-optim}
 
@@
 colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
+pal <- c("#D3D3D3","#D95F02","#7570B3", "#E7298A")
 color <- c(rep(pal[1], nrow(sphere)),
@@
            rep(pal[3], nrow(path2)))
-animate_xy(dt, col = color, cex_slice = 0.8)
-animate_slice(dt, col = color)
+cex <- c(rep(1, nrow(sphere)),
+         rep(2, nrow(path1)), 
+         rep(2, nrow(path2)))
+#save_history(animate_xy(dt, col = color, cex = cex))
+
+# save history and play to find the view similar to the PCA view
+
+animate_slice(dt, col = color, pch_other = 20)
 

### 2020-09-29 paper/paper.Rmd
shows a selected number of frames of the animated pca plot/ tour

--- 2020-09-17 paper/paper.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,root.dir = here::here())
 
@@
 
-```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
 path1 <- holes_1d_geo$basis %>% 
@@
 
-```{r cache = TRUE, eval=FALSE}
-# rl <- lapply(list.files(here::here("figures/animated_pca_1d/"),full.names = TRUE)[62:78], png::readPNG)
-# gl <-  lapply(rl, grid::rasterGrob)
-# gridExtra::grid.arrange(grobs=gl)
-ani <- p +
-  transition_states(id) +
-  shadow_mark()
-
-animate(ani, nframes = 350)
-anim_save(ani, nframes = 350, filename = "figures/animated_pca_1d")
+```{r fig.cap = "THis is the caption", fig.height=4, fig.align="center"}
+# ani <- p + theme(legend.position = "none") +
+#   transition_states(id) +
+#   shadow_mark()
+# 
+# animate(ani, nframes = 350, device = "png",
+#         renderer = file_renderer("figures/pca/",
+#                                  prefix = "pca", overwrite = TRUE))
+
+frames <- c("0002", "0058", "0078", "0117", "0172", "0350")
+ani <- paste0(here::here("figures/"), "pca/", "pca", frames, ".png")
+rl <- lapply(ani, png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+wrap_plots(gl) 
 ```
@@
 
-```{r eval = FALSE}
+```{r fig.cap="THis is te caption"}
 set.seed(1)
 sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-path1 <- holes_1d_geo$basis %>% 
+path1 <- holes_1d_better$basis %>% 
   flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-path2 <- holes_1d_better$basis %>% 
+path2 <- holes_1d_geo$basis %>% 
   flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-dt <- rbind(sphere, path1, path2)
+theoretical <- matrix(c(0, 1, 0, 0, 0,
+                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
+start <- get_start(holes_1d_better) %>% pull(basis) %>% .[[1]] %>% matrix(nrow = 1)
+dt <- rbind(sphere, path1, path2, theoretical, start)
 colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
+pal <- c("#D3D3D3",c("#524340",  #orchre
+                     "#B4B754",  # green
+                     "#F3B422" # yellow
+                     ))
 color <- c(rep(pal[1], nrow(sphere)),
            rep(pal[2], nrow(path1)),
-           rep(pal[3], nrow(path2)))
-animate_xy(dt, col = color, cex_slice = 0.8)
-animate_slice(dt, col = color)
-
+           rep(pal[3], nrow(path2)),
+           rep(pal[4], 2), 
+           pal[2] # for start
+           )
+cex <- c(rep(1, nrow(sphere)),
+         rep(1.5, nrow(path1)), 
+         rep(1.5, nrow(path2)), 
+         rep(5, 2), 
+         5 # for start
+         )
+
+#animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial
+
+# set.seed(123)
+# render(
+#   dt,
+#   tour_path = grand_tour(),
+#   dev = "png",
+#   display = display_xy(col=color,cex = cex, axes = "bottomleft"),
+#   rescale = FALSE,
+#   frames = 500,
+#   here::here("figures","tour", "tour%03d.png")
+# )
+
+frames <- c("001", "143", "078", "117", "172", "350")
+ani <- paste0(here::here("figures/"), "tour/", "tour", frames, ".png")
+rl <- lapply(ani, png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+wrap_plots(gl) 
 ```
@@
   dplyr::bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500)) %>% 
-  explore_trace_interp(col = max_tries, facet = max_tries) + 
+  explore_trace_interp(col = max_tries) + 
   theme(legend.position = "none")

### 2020-09-30 paper/paper.Rmd
A skeletal writing for the implementation section

@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm}
+  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 dt <- rbind(sphere, path1, path2, theoretical)
-pca <- dt %>% prcomp()
+pca <- dt %>% prcomp(scale. = TRUE)
 #summary(pca)
@@
 
-# Implementation: Ferrn package {#implementation}
+# Implementation {#implementation}
 
@@
 Everything is coded up in a package. Package structure
+
+The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package.
+The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.
 
@@
 Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.
-
-We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.
+Below presents a sample of the data object.
 
@@
 
-2.
+Once the data object has been obtained, `ferrn` provides four automatic diagnostic plots introduced in Section \ref{vis-diag}. The overall structure is ....
+
+## Extract pieces from the data object
+
+Several auxiliary functions provides a handy solution to extract certain component in the data object.
+For example, `get_best()` extracts the final basis found by the algorithm and `get_interp()` extracts the interpolation observations and renumber the id for a continuous numbering that being supplied to `explore_trace_interp()`.
+
+## Bind more information
+
+`bind_*` functions allow users to incorporate further useful information into the data object.
+Given a simulated dataset, the basis that can produce the most interesting projection is known and users can use `bind_theoretical()` to incorporate this best possible basis:
+
+```{r eval = FALSE, echo = TRUE}
+theo_best <- matrix(c(0, 1, 0, 0, 0), nrow = 5)
+holes_1d_better %>% 
+  bind_theoretical(matrix = theo_best, index = tourr::holes(), data_raw = data) 
+```
+
+`bind_random(_matrix)` generates a large number of points on the surface of a high dimensional sphere and attaches it to the original data object.
+This is particularly useful for exploring the search space in `explore_space_pca()` and `explore_space_tour()` since the space of the projection basis is a high dimensional sphere and overlaying the bases with random points on the sphere provides better perception to view these plots.
+
+```{r eval = FALSE, echo = TRUE}
+holes_1d_better %>% bind_random() %>% tail()
+```
+
+## Botanical color palette and reorder factor levels
+
+A handful of color from Australian domestic plants has been incorporated as the main color theme of the package, which inlcudes fern, cherry, banksia, acacia and daisy.
+
+\begin{table}
+\centering
+\begin{tabular}{|wl{5cm} wl{5cm}|}
+\hline
+\multicolumn{2}{|l|}{\textbf{Main functions}}\\
+\multicolumn{2}{|l|}{explore\_trace\_search()}\\
+\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
+\multicolumn{2}{|l|}{explore\_space\_pca() }\\
+\multicolumn{2}{|l|}{explore\_space\_tour()}\\
+& \\
+\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
+get\_best() & bind\_theoretical() \\
+get\_start() & bind\_random() \\
+get\_interp() & bind\_random\_matrix() \\
+get\_search\_count() & \\
+get\_basis\_matrix() & \\
+& \\
+relevel\_geo() & botanical\_palettes() \\
+relevel\_better() & botanical\_pal() \\
+& scale\_color\_botanical()\\
+\hline
+\end{tabular}
+\caption{A glossary of functions available in ferrn package.
+The main functions are four automatic plotting functions and auxiliary functions provides easier extraction (get\_*) and binding (bind\_*) of the data object.}
+\label{table:ferrn-glossary}
+\end{table}
 

### 2020-09-30 paper/paper.tex
change generated points to be grey in tour animation
shows a selected number of frames of the animated pca plot/ tour
A skeletal writing for the implementation section

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont, bbm} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 \hypertarget{tour-optim}{%
-\subsection{Optimisation problem}\label{tour-optim}}
+\subsection{Optimisation in the tour}\label{tour-optim}}
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1}
+
+}
+
+\caption{THis is the caption}\label{fig:unnamed-chunk-1}
+\end{figure}
+
 \hypertarget{the-tour-looking-at-itself}{%
@@
 high dimension from different angles.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{THis is te caption}
+\end{figure}
 
@@
 \hypertarget{implementation}{%
-\section{Implementation: Ferrn package}\label{implementation}}
+\section{Implementation}\label{implementation}}
 
@@
 Everything is coded up in a package. Package structure
+
+The implementation of this projection has been divided into two
+pacakges, where the data collection happens when a tour is running and
+hence implemented in the existing \texttt{tourr} package. The
+diagnostics of the algorithm via plots and other auxiliary functions
+have been implemented in a new pacakge, \texttt{ferrn}.
 
@@
 list-column \texttt{basis} to be printed out nicely with the dimension
-of the projection basis readily available.
-
-We form our first dataset using variables \texttt{x1}, \texttt{x2},
-\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
-optimiser \texttt{search\_better} and below shows the first ten rows of
+of the projection basis readily available. Below presents a sample of
 the data object.
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\item
-\end{enumerate}
+Once the data object has been obtained, \texttt{ferrn} provides four
+automatic diagnostic plots introduced in Section \ref{vis-diag}. The
+overall structure is \ldots{}.
+
+\hypertarget{extract-pieces-from-the-data-object}{%
+\subsection{Extract pieces from the data
+object}\label{extract-pieces-from-the-data-object}}
+
+Several auxiliary functions provides a handy solution to extract certain
+component in the data object. For example, \texttt{get\_best()} extracts
+the final basis found by the algorithm and \texttt{get\_interp()}
+extracts the interpolation observations and renumber the id for a
+continuous numbering that being supplied to
+\texttt{explore\_trace\_interp()}.
+
+\hypertarget{bind-more-information}{%
+\subsection{Bind more information}\label{bind-more-information}}
+
+\texttt{bind\_*} functions allow users to incorporate further useful
+information into the data object. Given a simulated dataset, the basis
+that can produce the most interesting projection is known and users can
+use \texttt{bind\_theoretical()} to incorporate this best possible
+basis:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{theo_best <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{)}
+\NormalTok{holes_1d_better }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\DataTypeTok{matrix =}\NormalTok{ theo_best, }\DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{data_raw =}\NormalTok{ data) }
+\end{Highlighting}
+\end{Shaded}
+
+\texttt{bind\_random(\_matrix)} generates a large number of points on
+the surface of a high dimensional sphere and attaches it to the original
+data object. This is particularly useful for exploring the search space
+in \texttt{explore\_space\_pca()} and \texttt{explore\_space\_tour()}
+since the space of the projection basis is a high dimensional sphere and
+overlaying the bases with random points on the sphere provides better
+perception to view these plots.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{holes_1d_better }\OperatorTok{%>%}\StringTok{ }\KeywordTok{bind_random}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{tail}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{botanical-color-palette-and-reorder-factor-levels}{%
+\subsection{Botanical color palette and reorder factor
+levels}\label{botanical-color-palette-and-reorder-factor-levels}}
+
+A handful of color from Australian domestic plants has been incorporated
+as the main color theme of the package, which inlcudes fern, cherry,
+banksia, acacia and daisy.
+
+\begin{table}
+\centering
+\begin{tabular}{|wl{5cm} wl{5cm}|}
+\hline
+\multicolumn{2}{|l|}{\textbf{Main functions}}\\
+\multicolumn{2}{|l|}{explore\_trace\_search()}\\
+\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
+\multicolumn{2}{|l|}{explore\_space\_pca() }\\
+\multicolumn{2}{|l|}{explore\_space\_tour()}\\
+& \\
+\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
+get\_best() & bind\_theoretical() \\
+get\_start() & bind\_random() \\
+get\_interp() & bind\_random\_matrix() \\
+get\_search\_count() & \\
+get\_basis\_matrix() & \\
+& \\
+relevel\_geo() & botanical\_palettes() \\
+relevel\_better() & botanical\_pal() \\
+& scale\_color\_botanical()\\
+\hline
+\end{tabular}
+\caption{A glossary of functions available in ferrn package.
+The main functions are four automatic plotting functions and auxiliary functions provides easier extraction (get\_*) and binding (bind\_*) of the data object.}
+\label{table:ferrn-glossary}
+\end{table}
 

### 2020-10-01 paper/paper.Rmd
write up for implementation section

--- 2020-09-29 paper/paper.Rmd
@@
 p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern_brown")
+  scale_color_botanical(palette = "fern")
 
 p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern_green")
+  scale_color_botanical(palette = "fern")
 
 p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "acacia")
+  scale_color_botanical(palette = "fern")
 
@@
 
-```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-path1 <- holes_1d_geo$basis %>% 
-  flatten_dbl() %>% 
-  matrix(ncol = 5, byrow = TRUE)
-path2 <- holes_1d_better$basis %>% 
-  flatten_dbl() %>% 
-  matrix(ncol = 5, byrow = TRUE)
-set.seed(1)
-sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-theoretical <- matrix(c(0, 1, 0, 0, 0,
-                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
-
-dt <- rbind(sphere, path1, path2, theoretical)
-pca <- dt %>% prcomp()
-#summary(pca)
-pca_pred <- predict(pca) %>%
-  as_tibble() %>%
-  mutate(method = c(rep("sphere", nrow(sphere)),
-                    holes_1d_geo$method,
-                    holes_1d_better$method,
-                    rep("theoretical", 2))) %>%
-  group_by(method) %>%
-  mutate(id = row_number(),
-         id = ifelse(method == "sphere", 1, id)) %>%
-  ungroup()
-
-p <- pca_pred %>%
-  ggplot(aes(x = PC1, y = PC2, color = method))+
-  geom_point() +
-  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
-             size = 10) + 
-  geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
-  theme(aspect.ratio = 1) + 
-  scale_color_botanical(palette = "cherry")
+```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+# path1 <- holes_1d_geo$basis %>% 
+#   flatten_dbl() %>% 
+#   matrix(ncol = 5, byrow = TRUE)
+# path2 <- holes_1d_better$basis %>% 
+#   flatten_dbl() %>% 
+#   matrix(ncol = 5, byrow = TRUE)
+# set.seed(1)
+# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
+# theoretical <- matrix(c(0, 1, 0, 0, 0,
+#                         0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
+# 
+# dt <- rbind(sphere, path1, path2, theoretical)
+# pca <- dt %>% prcomp(scale. = TRUE)
+# #summary(pca)
+# pca_pred <- predict(pca) %>%
+#   as_tibble() %>%
+#   mutate(method = c(rep("sphere", nrow(sphere)),
+#                     holes_1d_geo$method,
+#                     holes_1d_better$method,
+#                     rep("theoretical", 2))) %>%
+#   group_by(method) %>%
+#   mutate(id = row_number(),
+#          id = ifelse(method == "sphere", 1, id)) %>%
+#   ungroup()
+# 
+# p <- pca_pred %>%
+#   ggplot(aes(x = PC1, y = PC2, color = method))+
+#   geom_point() +
+#   geom_point(data = pca_pred %>% filter(method == "theoretical"), 
+#              size = 10) + 
+#   geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
+#   theme(aspect.ratio = 1) + 
+#   scale_color_botanical(palette = "cherry")
+# 
+# p
+
+p <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
+  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
+                   index = tourr::holes(), raw_data = data) %>% 
+  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
+                   index = tourr::holes(), raw_data  = data)%>%
+  explore_space_pca(col = method)  + 
+  scale_color_botanical(palette = "cherry") + 
+  theme(legend.position = "bottom")
 
@@
 #   shadow_mark()
+# 
+# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
+#                    index = tourr::holes(), raw_data = data) %>% 
+#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
+#                    index = tourr::holes(), raw_data  = data)%>%
+#   explore_space_pca(col = method, animate = TRUE)  + 
+#   scale_color_botanical(palette = "cherry") 
 # 
@@
 
-# Implementation: Ferrn package {#implementation}
-
-1. how does the printed data object looks like?
-
-Do I talk about the work done in the tourr package?
-
-Everything is coded up in a package. Package structure
+# Implementation {#implementation}
+
+(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
+The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package.
+The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.
 
@@
 
-2.
+\newpage
+
+Once the data object has been obtained, the package, `ferrn`, provides automatic diagnostic plots as described in Section \ref{vis-diag}.
+The structure of package functionality has been outlined in Table \ref{table:ferrn-glossary}.
+The four main functions plots, `explore_trace_search()`, `explore_trace_interp()`, `explore_space_pca()`, and `explore_space_tour()`, produce the diagnostic plots upon a simple supply of the data objects.
+Auxiliary functions, `get_*()` and `bind_*()`, provide handy solutions to extract certain components in the data object.
+They modulate the extraction and binding procedures in the main functions while being convenient to use on their own.
+
+```{r eval = FALSE, echo = TRUE}
+# Below are codes that generate the plots in Section 4 
+# using functions from ferrn package
+bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
+          holes_2d_better %>% mutate(max_tries = 25)) %>% 
+  explore_trace_search(group = max_tries)
+```
+
+```{r eval = FALSE, echo = TRUE}
+bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
+          interrupt_no %>% mutate(group = "no interruption"), 
+          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
+  mutate(group = fct_relevel(group, c("max_tries: 500", 
+                                      "no interruption", 
+                                      "better_random"))) %>% 
+  explore_trace_interp(group = group) + 
+  theme(legend.position = "none")
+```
+
+```{r eval = FALSE, echo = TRUE}
+bind_rows(holes_1d_geo, holes_1d_better) %>% 
+  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
+                   index = tourr::holes(), raw_data = data) %>% 
+  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
+                   index = tourr::holes(), raw_data  = data)%>%
+  explore_space_pca(col = method)  + 
+  scale_color_botanical(palette = "cherry")
+```
+
+```{r eval = FALSE, echo = TRUE}
+bind_rows(holes_1d_geo, holes_1d_better) %>% 
+  explore_space_tour(color = method)
+```
+
+## Extract pieces from the data object
+
+ `get_*()` extract and, in some cases, manipulate certain components from the existing data object, for example,  `get_best()` extracts the best basis found by the algorithm and `get_interp()` extracts the observations in the interpolation and prepare them ready to be supplied to `explore_trace_interp()`.
+
+## Bind external information
+
+`bind_*()` incorporate additional information that is outside the tour optimisation.
+Given a simulated dataset, the most interesting projection is known ahead of time, so does its basis.
+`bind_theoretical()` can be utilised to incorporate this best possible basis to the existing data object with the supply of the index function and original data for producing the index value.
+This binding is used in the space plots to compare the position of the theoretical best basis to the ones being searched by the algorithm.
+The following chunk presents its usage.
+
+```{r bind-theoretical-demo, eval = FALSE, echo = TRUE}
+theo_best <- matrix(c(0, 1, 0, 0, 0), nrow = 5)
+holes_1d_better %>% 
+  bind_theoretical(matrix = theo_best, index = tourr::holes(), raw_data = data) 
+```
+
+`bind_random()` and `bind_random_matrix()` generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object.
+This is also useful in the space plots since it preserves the spherical structure of the spaces of the projection bases, in the computation of PCA for `explore_space_pca()` and animation of `explore_space_tour()`.
+
+## Botanical color palette and reorder factor levels
+
+A collection of color palettes from Australian native plants has been created to match the theme of the package Quantitative palettes, daisy, banksia and cherry, are recommended for comparing between algorithms, while sequantial palettes, fern and acacia are better for trace plots.
+These color palettes can be accessed via
+
+- `scale_color_botanical(palette = "...")` for ggplot constructions, and
+
+- `explore_space_tour(..., color = botantical_palettes$...)` for a full space exploration
+
+When colored by `info`, users may want to map a particular color to a certain level to emphasize certain elements in the algorithm, `relevel_*()` allows the user to manipulate the level in the `info` column in the data object with the default being the chronological order that each level occurs in the optimisation.
+
+\begin{table}
+\centering
+\begin{tabular}{|wl{5cm} wl{6cm}|}
+\hline
+\multicolumn{2}{|l|}{\textbf{Main functions}}\\
+\multicolumn{2}{|l|}{explore\_trace\_search()}\\
+\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
+\multicolumn{2}{|l|}{explore\_space\_pca() }\\
+\multicolumn{2}{|l|}{explore\_space\_tour()}\\
+& \\
+\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
+get\_best() & bind\_theoretical() \\
+get\_start() & bind\_random() \\
+get\_interp() & bind\_random\_matrix() \\
+get\_search\_count() & \\
+get\_basis\_matrix() & \\
+& \\
+relevel\_geo() & botanical\_palettes:  fern, acacia\\
+relevel\_better() & \hspace{3mm} daisy, banksia, and cherry\\
+& botanical\_pal() \\
+& scale\_color\_botanical()\\
+\hline
+\end{tabular}
+\caption{A glossary of functions available in ferrn package. The main functions are four automatic plotting functions and auxiliary functions provides easier manipulation on the data object.}
+\label{table:ferrn-glossary}
+\end{table}
 

### 2020-10-01 paper/paper.tex
write up for implementation section

--- 2020-09-17 paper/paper.tex
@@
 \hypertarget{implementation}{%
-\section{Implementation: Ferrn package}\label{implementation}}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  how does the printed data object looks like?
-\end{enumerate}
-
-Do I talk about the work done in the tourr package?
-
-Everything is coded up in a package. Package structure
+\section{Implementation}\label{implementation}}
+
+(further notice on the implementation in the tour package will be
+discussed in the tour developer meeting.) The implementation of this
+projection has been divided into two pacakges, where the data collection
+happens when a tour is running and hence implemented in the existing
+\texttt{tourr} package. The diagnostics of the algorithm via plots and
+other auxiliary functions have been implemented in a new pacakge,
+\texttt{ferrn}.
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\item
-\end{enumerate}
+\newpage
+
+Once the data object has been obtained, the package, \texttt{ferrn},
+provides automatic diagnostic plots as described in Section
+\ref{vis-diag}. The structure of package functionality has been outlined
+in Table \ref{table:ferrn-glossary}. The four main functions plots,
+\texttt{explore\_trace\_search()}, \texttt{explore\_trace\_interp()},
+\texttt{explore\_space\_pca()}, and \texttt{explore\_space\_tour()},
+produce the diagnostic plots upon a simple supply of the data objects.
+Auxiliary functions, \texttt{get\_*()} and \texttt{bind\_*()}, provide
+handy solutions to extract certain components in the data object. They
+modulate the extraction and binding procedures in the main functions
+while being convenient to use on their own.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# Below are codes that generate the plots in Section 4 }
+\CommentTok{# using functions from ferrn package}
+\KeywordTok{bind_rows}\NormalTok{(holes_2d_better_max_tries }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{max_tries =} \DecValTok{500}\NormalTok{), }
+\NormalTok{          holes_2d_better }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{max_tries =} \DecValTok{25}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{explore_trace_search}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ max_tries)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{bind_rows}\NormalTok{(holes_2d_better_max_tries }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"max_tries: 500"}\NormalTok{), }
+\NormalTok{          interrupt_no }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"no interruption"}\NormalTok{), }
+\NormalTok{          holes_2d_better_random }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \StringTok{"better_random"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{group =} \KeywordTok{fct_relevel}\NormalTok{(group, }\KeywordTok{c}\NormalTok{(}\StringTok{"max_tries: 500"}\NormalTok{, }
+                                      \StringTok{"no interruption"}\NormalTok{, }
+                                      \StringTok{"better_random"}\NormalTok{))) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{explore_trace_interp}\NormalTok{(}\DataTypeTok{group =}\NormalTok{ group) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"none"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{bind_rows}\NormalTok{(holes_1d_geo, holes_1d_better) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{), }
+                   \DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data =}\NormalTok{ data) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{-1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{,}\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{), }
+                   \DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data  =}\NormalTok{ data)}\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{explore_space_pca}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ method)  }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{scale_color_botanical}\NormalTok{(}\DataTypeTok{palette =} \StringTok{"cherry"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{bind_rows}\NormalTok{(holes_1d_geo, holes_1d_better) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{explore_space_tour}\NormalTok{(}\DataTypeTok{color =}\NormalTok{ method)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{extract-pieces-from-the-data-object}{%
+\subsection{Extract pieces from the data
+object}\label{extract-pieces-from-the-data-object}}
+
+\texttt{get\_*()} extract and, in some cases, manipulate certain
+components from the existing data object, for example,
+\texttt{get\_best()} extracts the best basis found by the algorithm and
+\texttt{get\_interp()} extracts the observations in the interpolation
+and prepare them ready to be supplied to
+\texttt{explore\_trace\_interp()}.
+
+\hypertarget{bind-external-information}{%
+\subsection{Bind external information}\label{bind-external-information}}
+
+\texttt{bind\_*()} incorporate additional information that is outside
+the tour optimisation. Given a simulated dataset, the most interesting
+projection is known ahead of time, so does its basis.
+\texttt{bind\_theoretical()} can be utilised to incorporate this best
+possible basis to the existing data object with the supply of the index
+function and original data for producing the index value. This binding
+is used in the space plots to compare the position of the theoretical
+best basis to the ones being searched by the algorithm. The following
+chunk presents its usage.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{theo_best <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{), }\DataTypeTok{nrow =} \DecValTok{5}\NormalTok{)}
+\NormalTok{holes_1d_better }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{bind_theoretical}\NormalTok{(}\DataTypeTok{matrix =}\NormalTok{ theo_best, }\DataTypeTok{index =}\NormalTok{ tourr}\OperatorTok{::}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{raw_data =}\NormalTok{ data) }
+\end{Highlighting}
+\end{Shaded}
+
+\texttt{bind\_random()} and \texttt{bind\_random\_matrix()} generates a
+large number of points on the surface of a high dimensional sphere and
+attaches it to the existing data object. This is also useful in the
+space plots since it preserves the spherical structure of the spaces of
+the projection bases, in the computation of PCA for
+\texttt{explore\_space\_pca()} and animation of
+\texttt{explore\_space\_tour()}.
+
+\hypertarget{botanical-color-palette-and-reorder-factor-levels}{%
+\subsection{Botanical color palette and reorder factor
+levels}\label{botanical-color-palette-and-reorder-factor-levels}}
+
+A collection of color palettes from Australian native plants has been
+created to match the theme of the package Quantitative palettes, daisy,
+banksia and cherry, are recommended for comparing between algorithms,
+while sequantial palettes, fern and acacia are better for trace plots.
+These color palettes can be accessed via
+
+\begin{itemize}
+\item
+  \texttt{scale\_color\_botanical(palette\ =\ "...")} for ggplot
+  constructions, and
+\item
+  \texttt{explore\_space\_tour(...,\ color\ =\ botantical\_palettes\$...)}
+  for a full space exploration
+\end{itemize}
+
+When colored by \texttt{info}, users may want to map a particular color
+to a certain level to emphasize certain elements in the algorithm,
+\texttt{relevel\_*()} allows the user to manipulate the level in the
+\texttt{info} column in the data object with the default being the
+chronological order that each level occurs in the optimisation.
+
+\begin{table}
+\centering
+\begin{tabular}{|wl{5cm} wl{6cm}|}
+\hline
+\multicolumn{2}{|l|}{\textbf{Main functions}}\\
+\multicolumn{2}{|l|}{explore\_trace\_search()}\\
+\multicolumn{2}{|l|}{explore\_trace\_interp()}\\
+\multicolumn{2}{|l|}{explore\_space\_pca() }\\
+\multicolumn{2}{|l|}{explore\_space\_tour()}\\
+& \\
+\multicolumn{2}{|l|}{\textbf{Auxiliary functions}}\\
+get\_best() & bind\_theoretical() \\
+get\_start() & bind\_random() \\
+get\_interp() & bind\_random\_matrix() \\
+get\_search\_count() & \\
+get\_basis\_matrix() & \\
+& \\
+relevel\_geo() & botanical\_palettes:  fern, acacia\\
+relevel\_better() & \hspace{3mm} daisy, banksia, and cherry\\
+& botanical\_pal() \\
+& scale\_color\_botanical()\\
+\hline
+\end{tabular}
+\caption{A glossary of functions available in ferrn package. The main functions are four automatic plotting functions and auxiliary functions provides easier manipulation on the data object.}
+\label{table:ferrn-glossary}
+\end{table}
 

### 2020-10-03 paper/paper.Rmd
change functions in my pacakge to a list

--- 2020-09-29 paper/paper.Rmd
@@
 ---
-title: Title here
+title: "Visual diagnostics for constrained optimisation with application to guided tours"
 blinded: 0
@@
 
-```{r fig.cap = "THis is the caption", fig.height=4, fig.align="center"}
+```{r fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
+With animation, it is easier to track the progrssion from the start to finish in each algorithm.", fig.height=4, fig.align="center"}
 # ani <- p + theme(legend.position = "none") +
@@
 
-Compare to the PCA plot, the animated rotation (tour) display gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
 This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
@@
 
-```{r fig.cap="THis is te caption"}
+```{r toy-tour, fig.cap="\\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+The second frame on the top row view the space from a direction that is close to the one in PCA plot.
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles."}
 set.seed(1)
@@
 
-\newpage
-
 ```{r glb-obj-1}
@@
 
-2.
+Once the data object has been obtained, the package, `ferrn`, provides automatic diagnostic plots as described in Section \ref{vis-diag}. The structure of package functionality has been listed below.
+
+**Main functions**
+
+- `explore_trace_search()`: produce summary plots, as shown in Figure \ref{toy-search}
+- `explore_trace_interp()`: produce trace plots for the interpolation points, as shown in Figure \ref{toy-interp}
+- `explore_space_pca() `: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{toy-pca}.
+Animated version in Figure \ref{toy-pca-aniamted} can be acquired via the argument `animate = TRUE`
+- `explore_space_tour()`: produce animated tour view on the full space of the projection bases, as shown in Figure \ref{toy-tour}
+
+**Auxiliary functions**
+
+-  `get_*()` extract and, in some cases, manipulate certain components from the existing data object.
+    - `get_best()`: extract the best basis found by the algorithm
+    - `get_start()`: extract the starting basis
+    - `get_interp()`: extract the observations in the interpolation and prepare to be supplied to `explore_trace_interp()`. 
+    - `get_search_count()`: produce the summary table of the number of observation in each iteration and prepare to be supplied to `explore_trace_search()`
+    - `get_basis_matrix()`: flattern all the bases into a row and stack them to form a large matrix for `explore_space_tour()`
+
+- `bind_*()` incorporate additional information that is outside the tour optimisation.
+
+  - `bind_theoretical()`: incorporate the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
+  - `bind_random()`: generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object and output as a tibble object.
+`bind_random_matrix()` returns a matrix.
+
+  <!-- This binding is used in the space plots to compare the position of the theoretical best basis to the ones being searched by the algorithm. The following chunk presents its usage -->
+<!-- These are also useful in the space plots since it preserves the spherical structure of the spaces of the projection bases, in the computation of PCA for `explore_space_pca()` and animation of `explore_space_tour()`.
+-->
+
+- Color and relevel
+  - `botanical_palettes`:  a collection of color palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequantial palettes cotain fern and acacia.
+  - `botanical_pal()`: a color interpolator useful for `scale_color_botanical()`
+  - `scale_color_botanical()`: a ggplot construction for using botanical palettes. 
+  - `relevel_geo()`: manipulate the level in the `info` column in the data object when geodesic search is used
+  - `relevel_better()`: manipulate the level in the `info` column in the data object when search_better related search is used
+
+```{r eval = FALSE}
+# Below are codes that generate the plots in Section 4 
+# using functions from ferrn package
+bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
+          holes_2d_better %>% mutate(max_tries = 25)) %>% 
+  explore_trace_search(group = max_tries)
+```
+
+```{r eval = FALSE}
+bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
+          interrupt_no %>% mutate(group = "no interruption"), 
+          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
+  mutate(group = fct_relevel(group, c("max_tries: 500", 
+                                      "no interruption", 
+                                      "better_random"))) %>% 
+  explore_trace_interp(group = group) + 
+  theme(legend.position = "none")
+```
+
+```{r eval = FALSE}
+bind_rows(holes_1d_geo, holes_1d_better) %>% 
+  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
+                   index = tourr::holes(), raw_data = data) %>% 
+  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
+                   index = tourr::holes(), raw_data  = data)%>%
+  explore_space_pca(col = method)  + 
+  scale_color_botanical(palette = "cherry")
+```
+
+```{r eval = FALSE}
+bind_rows(holes_1d_geo, holes_1d_better) %>% 
+  explore_space_tour(color = method)
+```
 
@@
 
+This paper has illustrated setting up a data object that can be used for diagnosing a complex optimisation procedure.
+The ideas were illustrated using the optimisers available for projection pursuit guided tour.
+Here the constraint is the orthornormality condition of the projection bases.
+The approach used here could be broadly applied to understand other constrained optimisers.
+
+something about the diagnostic plots
+
+summarise what we have done and what might be done in the future.
+
 \clearpage

### 2020-10-03 paper/paper.tex
change functions in my pacakge to a list

--- 2020-09-17 paper/paper.tex
@@
 {
-  \title{\bf Title here}
+  \title{\bf Visual diagnostics for constrained optimisation with application to
+guided tours}
 
@@
   \begin{center}
-    {\LARGE\bf Title here}
+    {\LARGE\bf Visual diagnostics for constrained optimisation with application to
+guided tours}
   \end{center}
@@
 
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1}
+
+}
+
+\caption{\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
+With animation, it is easier to track the progrssion from the start to finish in each algorithm.}\label{fig:unnamed-chunk-1}
+\end{figure}
+
 \hypertarget{the-tour-looking-at-itself}{%
@@
 
-Compare to the PCA plot, the animated rotation (tour) display gives a
-more holistic view of the search of the optimiser and these additional
-information gained from the animation is cruicial. This is because in
-essense, PCA presents one reduced space that maximises the variance and
-it inherently amplifies the spread of the randomly evaluated bases in
-the search since they have larger variance than the bases on the
-interpolation path. Hence when the interest is to view the interpolation
-path in the space, an animated rotation display provides the view of
-high dimension from different angles.
+Compare to the PCA plot, the animated rotation (tour) display, in Figure
+\ref{toy-tour}, gives a more holistic view of the search of the
+optimiser and these additional information gained from the animation is
+cruicial. This is because in essense, PCA presents one reduced space
+that maximises the variance and it inherently amplifies the spread of
+the randomly evaluated bases in the search since they have larger
+variance than the bases on the interpolation path. Hence when the
+interest is to view the interpolation path in the space, an animated
+rotation display provides the view of high dimension from different
+angles.
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
+\caption{\label{toy-tour}A selected number of frames from the tour
+animation for viewing the 5D space of all the projection bases. The
+second frame on the top row view the space from a direction that is
+close to the one in PCA plot. The tour animation allows for a more
+holistic view of the full space in high dimensions from different
+angles.}
+\end{figure}
 
@@
 list-column \texttt{basis} to be printed out nicely with the dimension
-of the projection basis readily available.
-
-We form our first dataset using variables \texttt{x1}, \texttt{x2},
-\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
-optimiser \texttt{search\_better} and below shows the first ten rows of
+of the projection basis readily available. Below presents a sample of
 the data object.
-
-\newpage
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\setcounter{enumi}{1}
-\item
-\end{enumerate}
+Once the data object has been obtained, the package, \texttt{ferrn},
+provides automatic diagnostic plots as described in Section
+\ref{vis-diag}. The structure of package functionality has been listed
+below.
+
+\textbf{Main functions}
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{explore\_trace\_search()}: produce summary plots, as shown in
+  Figure \ref{toy-search}
+\item
+  \texttt{explore\_trace\_interp()}: produce trace plots for the
+  interpolation points, as shown in Figure \ref{toy-interp}
+\item
+  \texttt{explore\_space\_pca()}: produce plots of projection basis on
+  the reduced space by PCA, as shown in Figure \ref{toy-pca}. Animated
+  version in Figure \ref{toy-pca-aniamted} can be acquired via the
+  argument \texttt{animate\ =\ TRUE}
+\item
+  \texttt{explore\_space\_tour()}: produce animated tour view on the
+  full space of the projection bases, as shown in Figure \ref{toy-tour}
+\end{itemize}
+
+\textbf{Auxiliary functions}
+
+\begin{itemize}
+\item
+  \texttt{get\_*()} extract and, in some cases, manipulate certain
+  components from the existing data object.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{get\_best()}: extract the best basis found by the algorithm
+  \item
+    \texttt{get\_start()}: extract the starting basis
+  \item
+    \texttt{get\_interp()}: extract the observations in the
+    interpolation and prepare to be supplied to
+    \texttt{explore\_trace\_interp()}.
+  \item
+    \texttt{get\_search\_count()}: produce the summary table of the
+    number of observation in each iteration and prepare to be supplied
+    to \texttt{explore\_trace\_search()}
+  \item
+    \texttt{get\_basis\_matrix()}: flattern all the bases into a row and
+    stack them to form a large matrix for
+    \texttt{explore\_space\_tour()}
+  \end{itemize}
+\item
+  \texttt{bind\_*()} incorporate additional information that is outside
+  the tour optimisation.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{bind\_theoretical()}: incorporate the best possible basis to
+    the existing data object with the supply of the index function and
+    original data for producing the index value.
+  \item
+    \texttt{bind\_random()}: generates a large number of points on the
+    surface of a high dimensional sphere and attaches it to the existing
+    data object and output as a tibble object.
+    \texttt{bind\_random\_matrix()} returns a matrix.
+  \end{itemize}
+\item
+  Color and relevel
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{botanical\_palettes}: a collection of color palettes from
+    Australian native plants Quantitative palettes include daisy,
+    banksia and cherry and sequantial palettes cotain fern and acacia.
+  \item
+    \texttt{botanical\_pal()}: a color interpolator useful for
+    \texttt{scale\_color\_botanical()}
+  \item
+    \texttt{scale\_color\_botanical()}: a ggplot construction for using
+    botanical palettes.
+  \item
+    \texttt{relevel\_geo()}: manipulate the level in the \texttt{info}
+    column in the data object when geodesic search is used
+  \item
+    \texttt{relevel\_better()}: manipulate the level in the
+    \texttt{info} column in the data object when search\_better related
+    search is used
+  \end{itemize}
+\end{itemize}
 
@@
 
+This paper has illustrated setting up a data object that can be used for
+diagnosing a complex optimisation procedure. The ideas were illustrated
+using the optimisers available for projection pursuit guided tour. Here
+the constraint is the orthornormality condition of the projection bases.
+The approach used here could be broadly applied to understand other
+constrained optimisers.
+
+something about the diagnostic plots
+
+summarise what we have done and what might be done in the future.
+
 \clearpage

### 2020-10-07 paper/paper.Rmd
fixing on section 5 examples

@@
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
-output: rticles::asa_article
+output:
+  bookdown::pdf_book:
+    base_format: rticles::asa_article
+    fig_height: 4
 editor_options: 
@@
 
-\newpage
-
 # Optimisation Methods {#optim}
@@
 ppp]
-
-\newpage
 
@@
 
-```{r tour-path, out.width="100%", out.height="60%", fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
+```{r tour-path, out.width="60%", out.height="40%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"),)
 ```
@@
 <!--  - understand where the local & global maximum is found - trace plot - see if noisy function -->
-
-\newpage
 
@@
 
-\newpage
-
 # Visual diagnostics {#vis-diag}
@@
 
-\newpage
-
 ## Examining the optimisation progress
@@
 
-```{r toy-interp, fig.height = 8, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
+```{r toy-interp, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
 p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
@@
 
-\newpage
+\
 
@@
 
-```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
 # path1 <- holes_1d_geo$basis %>% 
@@
 
-\newpage
-
 # Diagnosing an optimiser {#application}
 
-For a particular index function, the best algorithm to optimise is related to the character of the index and the data.
-If the index function is smooth and has single maximum, all of the three algorithms introduced above can find maximum.
-When multiple optima are presented, `search_better` may stuck in the local maximum.
-When the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
+If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
+When multiple optima are presented, `search_better` may stuck in the local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
 In this section, examples will be presented  to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.
@@
 Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable `x1`, `x8`, `x9` and `x10` are normal distributed with zero mean and unit variance and `x2` to `x7` are mixture of normal distributions with different weights and locations.
-Figure \ref{origin-data} plots the histogram of each variable except `x3`.
-The mixture variables are scaled to have unit variance before running the projection pursuit.
+Variable `x1`, `x8`, `x9` and `x10` are normal distributed with zero mean and unit variance and `x2` to `x7` are mixtures of normal distributions with varied weights and locations.
+The mixture variables have been scaled to have an overall unit variance before running the projection pursuit.
 
@@
 
-```{r origin-data, fig.cap="\\label{origin-data} The histogram of simulated selected variables."}
-origin_dt_bi <- data_mult %>% 
-  dplyr::select(-c(x3, x8, x9, x10)) %>% 
-  gather(names, values) %>%
-  mutate(names = as_factor(names))
-
-origin_dt_bi %>%
-  ggplot(aes(x = values)) +
-  geom_histogram(binwidth = 0.3) +
-  geom_density(aes(y = 0.3 * ..count..)) +
-  facet_wrap(vars(names), ncol = 3)
-```
-
 ## A problem of not monotonic
-
-```{r interpolation}
-#points is always included in the newest version
-# p1 <- holes_2d_better %>%
-#   filter(info == "interpolation") %>%
-#   ggplot(aes(x = tries, y = index_val, col = as.factor(tries))) +
-#   geom_point() +
-#   geom_line(aes(group = 1))
-# 
-# p2 <- holes_2d_better %>% explore_trace_interp() + geom_point(aes(col = as.factor(tries)))
-# 
-# (p1 | p2)  &
-#   theme(legend.position = "bottom")
-# 
-```
 
@@
 p1 | p2
-#xxx up and down plot allows for displaying time series features while left and right display makes it easier for comparison? The same problem for plot of comparing between different searching methods
+
 ```
@@
 ```{r sample-code, eval = FALSE, echo = TRUE}
+library(tourr)
+library(ferrn)
+
 set.seed(123456)
-holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],tour_path = 
-                             guided_tour(holes(), d = 2, 
-                                         search_f = tourr:::search_geodesic),
+holes_2d_geo <- 
+  animate_xy(boa6,
+             tour_path = guided_tour(holes(), d = 2, 
+                                     search_f = search_geodesic),
              rescale = FALSE, verbose = TRUE)
 
-last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% 
-  tail(1) %>% pull(basis) %>% .[[1]]
+last_basis <- get_best(holes_2d_geo) %>% pull(basis) %>% .[[1]]
 
 set.seed(123456)
-holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)], tour_path = 
-                                    guided_tour(holes(), d = 2, 
-                                                search_f = tourr:::search_polish),
-                                  rescale = FALSE, verbose = TRUE, 
-                                  start = last_basis)
-```
-
-A slight variation of the plot definition due to the addition of polishing points is as follows:
-
-- Layer 1: point geom
-  - data: filter the observations with  $S = \text{interpolation}$; bind the data object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
-  - x: $t$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis 
-  - colour: $V$ is mapped to the colour aesthetic
-
-- Layer 2: line geom
-
-<!-- Polishing points form on itself a global structure with candidate bases and output bases.
-If we're allowed to abuse the notation a bit, we denote $$A_{J, \ast, \sum_{i = 1}^J k_i + 1}$$ as the first polishing basis and $$A_{J, \ast, \sum_{i = 1}^J k_i + M}$$ as the last polishing basis, where $$M = \sum_i S_{i, \text{method}} \mathds{1}(\text{method} = \text{polish})$$ -->
-
-<!-- Then the index plot incorporating both the interpolating points and polishing points can be defined as  -->
-
-<!-- $$x = \{1, 2, \cdots, \sum_{i = 1}^J k_i,(\sum_{i = 1}^J k_i + 1), \cdots,  \sum_{i = 1}^J k_i + M\}$$ -->
-<!-- $$y = \{I_{jlk}: k = x\}$$ -->
-
----
-
-Again using the same data, we are interested to compare the effect of different `max.tries` in the 2D projection setting.
+holes_2d_geo_polish <- 
+  animate_xy(boa6, 
+             tour_path = guided_tour(holes(), d = 2, 
+                                     search_f = search_polish),
+             rescale = FALSE, verbose = TRUE, start = last_basis)
+```
+
+Checking the final projection before and after applying the polish as in Figure \ref{geo-polish}, we can see that ...
+
+```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}On the left is the final projection from search geodesic while on the right is the polished result.
+Polishing helps to sharpen the edges of the clusters"}
+# set.seed(123456)
+# render(
+#   boa6,
+#   tour_path = guided_tour(holes(), d = 2, search_f = search_geodesic),
+#   dev = "png",
+#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   rescale = FALSE,
+#   frames = 100,
+#   file = here::here("figures","geo_polish", "before%03d.png")
+# )
+# 
+# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
+# 
+# set.seed(123456)
+# render(
+#   boa6,
+#   tour_path =  guided_tour(holes(), d = 2, search_f = search_polish, alpha = 0.1),
+#   dev = "png",
+#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   start = last_basis,
+#   rescale = FALSE,
+#   frames = 100,
+#   file = here::here("figures","geo_polish", "after%03d.png")
+# )
+
+before <- png::readPNG(here::here("figures","geo_polish", "before100.png"))
+after <- png::readPNG(here::here("figures","geo_polish", "after004.png"))
+gl <-  lapply(list(before, after), grid::rasterGrob)
+wrap_plots(gl) 
+```
+
+<!-- --- -->
+
+<!-- Again using the same data, we are interested to compare the effect of different `max.tries` in the 2D projection setting.
 `max.tries` is a hyperparameter that controls the maximum number of try before the search ends.
@@
 As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to sufficiently explore the parameter space.
-
-```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
-Both attain the same final index value after the polishing while using a `max.tries` of 25 is not sufficient to find the true maximum."}
-
-holes_2d_better %>% mutate(max_tries = 25) %>% 
-  dplyr::bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500)) %>% 
-  explore_trace_interp(col = max_tries) + 
-  theme(legend.position = "none")
-```
+-->
+
+<!-- ```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
+Both attain the same final index value after the polishing while using a `max.tries` of 25 is not sufficient to find the true maximum."} -->
+
+<!-- holes_2d_better %>% mutate(max_tries = 25) %>%  -->
+<!--   dplyr::bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500)) %>%  -->
+<!--   explore_trace_interp(col = max_tries) +  -->
+<!--   theme(legend.position = "none") -->
+<!-- ``` -->
 
@@
 
-[add example from geozoo]
-
 <!-- # Animated diagnostic plots -->
@@
 The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
-# p1 <- bind_rows(kol_1d_geo, kol_1d_geodesic_polish) %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
-# 
-# p2 <- bind_rows(kol_1d_better, kol_1d_better_polish) %>% explore_trace_interp(col = method) + ggtitle("search_better")
-# 
-# (p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
-
-p1 <- kol_1d_geo %>% explore_trace_interp(col = method)  + ggtitle("search_geodesic")
-
-p2 <- kol_1d_better %>% explore_trace_interp(col = method) + ggtitle("search_better")
-
-(p1 | p2 ) & theme(legend.position = "bottom") & ylim(0, 0.23)
+# set.seed(123456)
+# kol_1d_geo <- 
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
+#                                              search_f =  search_geodesic, max.tries = 100),
+#                rescale = FALSE, verbose = TRUE)
+# 
+# set.seed(123456)
+# kol_1d_better <- 
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
+#                                              search_f =  search_better, max.tries = 100),
+#                rescale = FALSE, verbose = TRUE)
+# save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
+# save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
+
+bind_rows(kol_1d_geo, kol_1d_better) %>% 
+  explore_trace_interp(col = method, group = method) +
+  facet_wrap(vars(method), ncol = 2) + 
+  scale_color_botanical() +
+  theme(legend.position = "none", aspect.ratio = 1)
+
 ```
@@
 `search_better_random` manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
-
-```{r 1d-2var-different-seeds,  fig.asp=0.5, fig.cap="\\label{1d-2var-different-seeds}The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
-Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum."}
-colour <- ferrn::botanical_pal()(3)
-p5 <- kol_1d_2var_better %>% explore_trace_interp(col = 0) + 
-  scale_colour_manual(values = colour[2]) + ggtitle("search_better, seed 123456")
-p6 <- kol_1d_2var_better_2 %>% explore_trace_interp(col = 0) + 
-  scale_colour_manual(values = colour[1]) + ggtitle("search_better, seed 12345")
-
-(p5 | p6) & theme(legend.position = "none") & ylim(0, 0.25)
-```
-
-```{r 1d-2var-better-random, fig.asp=0.5, fig.cap = "\\label{1d-2var-better-random}Using search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed."}
-kol_1d_2var_better_random %>% 
-  explore_trace_interp(col = 0) + 
-  ylim(0, 0.25) + 
-  scale_colour_manual(values = colour[3]) + 
-  theme(legend.position = "none") + 
-  ggtitle("search_better_random")
-```
-
+In Figure \ref{kol-pca},
+sing search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed.
 We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.
@@
 The large overlapping of better1 and better_random is explained by the fact that both algorithms find x2 in the end.
-
-<!-- ```{r 1d-2var-explore-proj-pca, fig.cap = "\\label{1d-2var-explore-proj-pca} The projected projection basis using principal components.
-The bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhausive search.
-The large overlapping of better1 and better_random is explained by the fact that both algorithms finds x2 in the end."} -->
-<!-- interp1 <- kol_cdf_1d_2var_better  -->
-<!-- interp2 <- kol_cdf_1d_2var_better_2 -->
-<!-- interp3 <- kol_cdf_1d_2var_better_random  -->
-
-<!-- dt <- bind_rows(interp1, interp2, interp3) %>%  -->
-<!--   mutate(algorithm = c(rep("better 1", nrow(interp1)),  -->
-<!--                        rep("better 2", nrow(interp2)),  -->
-<!--                        rep("better random", nrow(interp3))), -->
-<!--          algorithm = fct_relevel(algorithm, c("better 1", "better 2", "better random")))  -->
-
-<!-- dt %>%  -->
-<!--   explore_proj_pca(col = algorithm, alpha = 0.5) +  -->
-<!--   scale_colour_manual(values = color(3))  -->
-<!-- ``` -->
+The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
+Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum.
+
+```{r kol-pca, fig.cap = "\\label{kol-pca}THis is the caption"}
+
+# set.seed(123456)
+# kol_1d_2var_better <- 
+#   animate_dist(boa6, 
+#                tour_path = guided_tour(norm_kol(), d = 1,
+#                                        search_f = search_better, max.tries = 100),
+#                rescale = FALSE, verbose = TRUE)
+# 
+# 
+# set.seed(1234567)
+# kol_1d_2var_better_2 <- 
+#   animate_dist(boa6, 
+#                tour_path = guided_tour(norm_kol(), d = 1,
+#                                        search_f = search_better, max.tries = 100),
+#                rescale = FALSE, verbose = TRUE)
+# 
+# set.seed(12345678)
+# kol_1d_2var_better_2 <- 
+#   animate_dist(boa6, 
+#                tour_path = guided_tour(norm_kol(), d = 1,
+#                                        search_f = search_better, max.tries = 100),
+#                rescale = FALSE, verbose = TRUE)
+# 
+# 
+# 
+# 
+# set.seed(123456)
+# kol_1d_2var_better_random <- 
+#   animate_dist(boa6, 
+#                tour_path = guided_tour(norm_kol(), d = 1,
+#                                        search_f = search_better_random,
+#                                        max.tries = 2000, t0 = 1e-5),
+#                rescale = FALSE, verbose = TRUE)
+# 
+# save(kol_1d_2var_better, file = here::here("data", "kol_1d_2var_better.rda"))
+# save(kol_1d_2var_better_2, file = here::here("data", "kol_1d_2var_better_2.rda"))
+# save(kol_1d_2var_better_random, file = here::here("data", "kol_1d_2var_better_random.rda"))
+
+p1 <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
+          kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
+          kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+  explore_trace_interp(group = algo, col = algo) + 
+  facet_wrap(vars(algo), ncol = 3)  + 
+  scale_color_botanical(palette = "banksia")  + 
+  theme(legend.position = "none", aspect.ratio = 1)
+
+pca <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
+          kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
+          kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>% 
+   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>% 
+  compute_pca()
+
+random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-algo)
+theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-algo)
+
+p2 <- pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
+  ggplot(aes(x = PC1, y = PC2, col = algo)) + 
+  geom_point(data = random, col = "grey") +
+  geom_point(aes(col = algo), alpha = 0.1) + 
+  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
+  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
+  geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
+  geom_point(data = pca$aug %>% get_start(),  size = 5) +
+  facet_wrap(vars(algo)) +
+  theme_void() + 
+  theme(legend.position = "none", aspect.ratio = 1) + 
+  scale_color_botanical(palette = "banksia")
+
+wrap_plots(p1, p2, ncol = 1)
+
+```
 

### 2020-10-07 paper/paper.tex
fixing on section 5 examples

@@
 {
-  \title{\bf Visual diagnostics for constrained optimisation with application to
-guided tours}
+  \title{\bf Visual diagnostics for constrained optimisation with application to guided tours}
 
@@
   \begin{center}
-    {\LARGE\bf Visual diagnostics for constrained optimisation with application to
-guided tours}
+    {\LARGE\bf Visual diagnostics for constrained optimisation with application to guided tours}
   \end{center}
@@
 \begin{abstract}
-Friedman \& Tukey commented on their initial paper on projection pursuit
-in 1974 that ``the technique used for maximising the projection index
-strongly influences both the statistical and the computational aspects
-of the procedure.'' While many projection pursuit indices have been
-proposed in the literature, few concerns the optimisation procedure. In
-this paper, we developed a system of diagnostics aiming to visually
-learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied more generally to help
-practitioner to unveil the black-box in randomised iterative
-(optimisation) algorithms. An R package, ferrn, has been created to
-implement this diagnostic system.
+Friedman \& Tukey commented on their initial paper on projection pursuit in 1974 that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
+In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
+This diagnostic system can be applied more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
+An R package, ferrn, has been created to implement this diagnostic system.
 \end{abstract}
@@
 Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see
-information they would otherwise not see. This motivates our work of
-creating plots to diagnose optimisation algorithms in the context of
-projection pursuit guided tour, with the aim to understand and compare
-features of different existing algorithms.
-
-In an optimization problem the goal is to find the best solution within
-the space of all feasible solutions which typically is represented by a
-set of constraints. The problem consists in optimizing an objective
-function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in
-a reduced space given by the problem constraints to either minimize or
-maximize a function.
-
-Projection pursuit and guided tour are exploratory data analysis tools
-that detect interesting structure of high dimensional data through
-projection on low dimensional space. Optimisation is applied here to
-search for the low dimensional space that finds the most interesting
-projection.
-
-The remainder of the paper is organised as follows. Section \ref{optim}
-provides a literature review of optimisation methods, specifically the
-line search methods used in projection pursuit guided tour. Section
-\ref{tour} reviews projection pursuit guided tour, forms the
-optimisation problem, and introduces three main existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics design, from
-forming the data object to the definition of different diagnostic plots
-with some small examples. Section \ref{application} shows the
-application of how the diagnostic plots designed in section
-\ref{vis-diag} can be used to understand and compare different
-algorithms and how they contribute to modifications that improve the
-algorithms. Finally, Section \ref{implementation} describes the R
-package: ferrn, that implements all the visual diagnostics above.
-
-\newpage
+Presenting information in a graphical format often allows people to see information they would otherwise not see.
+This motivates our work of creating plots to diagnose optimisation algorithms in the context of projection pursuit guided tour, with the aim to understand and compare features of different existing algorithms.
+
+In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
+The problem consists in optimizing an objective function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in a reduced space given by the problem constraints to either minimize or maximize a function.
+
+Projection pursuit and guided tour are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
+Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
+
+The remainder of the paper is organised as follows.
+Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour.
+Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples.
+Section \ref{application} shows the application of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
+Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
 
@@
 
-Given an optimisation problem, two basic approaches find the optimum
-based on different thinking. An analytical approach aims to find the
-optimal solution in a finite number of steps, but a potential issue with
-it is that the closed-form solution may not be available when the
-problem starts to become complex. An iterative approach, on the other
-hand, finds the optimum based on the idea of making progressive
-improvement to the current solution. An iterative method may end up
-finding a local optimum but the progressive nature of the algorithm
-allows the practitioner to decide when to stop if a desirable accuracy
-has been achieved.
-
-A traditional while often used in practice is an iterative method called
-\emph{line search method} \citep{fletcher2013practical}. In a simple
-one-dimensional problem of finding the \(x\) that minimises \(f(x)\),
-line search achieves the goal via an iterative algorithm in the form of
-Equation \ref{eq:line-search}.
+Given an optimisation problem, two basic approaches find the optimum based on different thinking.
+An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the closed-form solution may not be available when the problem starts to become complex.
+An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution.
+An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.
+
+A traditional while often used in practice is an iterative method called \emph{line search method} \citep{fletcher2013practical}.
+In a simple one-dimensional problem of finding the \(x\) that minimises \(f(x)\), line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
 
@@
 
-where \(d^{(j)}\) is the searching direction in iteration \(j\), and
-\(\alpha_j\) is the step-size. Strictly speaking, \(\alpha_k\) is chosen
-by another minimisation of \(f(x^{(j)} + \alpha* d^{(j)})\) with respect
-to \(\alpha\) and theoretical results have demonstrated the global
-convergence of the algorithm when the exact minimisation of \(\alpha_j\)
-is attained \citep{curry1944method}. In practice, this second
-minimisation is rarely implemented due to its computational demanding or
-even the existence of such a minimisation. A more realistic approach is
-to impose a mandatory decrease in the objective function for each
-iteration: \(f^{(j+1)}> f^{(j)}\) and despite we lose the guarantee on
-global convergence, this approach turns out to be efficient in practical
-problems.
-
-{[}Are we using projection pursuit/guided tour to better understand the
-convergence of optimization algorithms visually in combination with the
-algorithms discussed below? Or we are focusing on the optimisation
-problem only within the projection pursuit context? Some of the problems
-listed below are also applicable to optimization problem in general too.
+where \(d^{(j)}\) is the searching direction in iteration \(j\), and \(\alpha_j\) is the step-size.
+Strictly speaking, \(\alpha_k\) is chosen by another minimisation of \(f(x^{(j)} + \alpha* d^{(j)})\) with respect to \(\alpha\) and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of \(\alpha_j\) is attained \citep{curry1944method}.
+In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation.
+A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\) and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
+
+{[}Are we using projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
+Or we are focusing on the optimisation problem only within the projection pursuit context?
+Some of the problems listed below are also applicable to optimization problem in general too.
 ppp{]}
-
-\newpage
 
@@
 
-Modern development of the line search methods focuses on proposing
-different computation on the searching direction: \(d^{(j)}\) and
-various approximations on the step size: \(\alpha_j\) catered for
-practical optimisation problems. The specific problem context we are
-interested in is called projection pursuit guided tour. Projection
-pursuit and guided tour are two separate methods in exporatory data
-analysis focusing on different aspects: coined by
-\citet{friedman1974projection}, projection pursuit detects interesting
-structures (i.e.~clustering, outliers and skewness) in multivariate data
-via low dimensions projection; whilst guided tour is a particular
-variation in a broader class of data visualisation method called tour.
-
-Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection
-can be seen as a linear transformation
-\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
-\(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
-\(\mathbf{P}_{n \times d}\) is the projected data and
-\(\mathbf{A}_{p\times d}\) is the projection basis. Define
-\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
-function that maps the projection basis \(\mathbf{A}\) onto an index
-value \(I\), this function is commonly known as the projection pursuit
-index (PPI) function, or the index function and is used to measure the
-``interestingness'' of a projection. A number of index functions have
-been proposed in the literature to detect different data structures,
-including Legendre index \citep{friedman1974projection}, Hermite index
-\citep{hall1989polynomial}, natural Hermite index
-\citep{cook1993projection}, chi-square index
-\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
-index \citep{lee2010projection}.
-
-In their initial paper, \citet{friedman1974projection} noted that
-``\ldots{}, the technique used for maximising the projection index
-strongly influences both the statistical and the computational aspects
-of the procedure.'' Hence, effective optimisation algorithms are
-necessary for projection pursuit to find the bases that give interesting
-projections. While we leave the formal construction of the optimisation
-problem and existing algorithms to section \ref{tour-optim}, we outline
-the general idea here. Given a random starting (current) basis,
-projection pursuit repeatedly searches for candidate bases nearby until
-it finds one with higher index value than the current basis. In the
-second round, that basis becomes the current basis and the repetitive
-sampling continues. The process ends until no better basis can be found
-or one of the termination criteria is reached.
-
-Before introducing the guided tour, we shall be familiar with the
-general tour method \citep{cook2008grand}. A tour produces animated
-visualisation of the high dimensional data via rotating low dimension
-planes. The smoothness of the animation is ensured by computing a series
-of intermediate planes between two low dimension planes via geodesic
-interpolation and we refer readers to \citet{buja2005computational} for
-the mathematical details. Iteratively choosing different low dimension
-planes and interpolating between them forms a tour path. Different types
-of tour methods choose the low dimensional planes differently and we
-mention two other types of tour that are commonly used. A grand tour
-selects the planes randomly in the high dimensional space and hence
-serves as an initial exploration of the data. Manual control allows
-researches to fine-tuning an existing projection by gradually phase in
-and out one variable.
-
-Guided tour chooses the planes produced by optimising the projection
-pursuit index function. Figure \ref{tour-path} shows a sketch of the
-tour path consisting of the blue frames produced by the projection
-pursuit optimisation algorithm iteratively and the white frames, which
-are the interpolations between two blue frames. The tour method has been
-implemented in the \emph{tourr} package in R, available on the
-Comprehensive R Archive Network at
-\url{https://cran.r-project.org/web/packages/tourr/}
-\citep{wickham2011tourrpackage}.
-
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.6\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001} \caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
-\end{figure}
-
-\newpage
+Modern development of the line search methods focuses on proposing different computation on the searching direction: \(d^{(j)}\) and various approximations on the step size: \(\alpha_j\) catered for practical optimisation problems.
+The specific problem context we are interested in is called projection pursuit guided tour.
+Projection pursuit and guided tour are two separate methods in exporatory data analysis focusing on different aspects: coined by \citet{friedman1974projection}, projection pursuit detects interesting structures (i.e.~clustering, outliers and skewness) in multivariate data via low dimensions projection; whilst guided tour is a particular variation in a broader class of data visualisation method called tour.
+
+Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{P}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection basis.
+Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projection basis \(\mathbf{A}\) onto an index value \(I\), this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the ``interestingness'' of a projection.
+A number of index functions have been proposed in the literature to detect different data structures, including Legendre index \citep{friedman1974projection}, Hermite index \citep{hall1989polynomial}, natural Hermite index \citep{cook1993projection}, chi-square index \citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA index \citep{lee2010projection}.
+
+In their initial paper, \citet{friedman1974projection} noted that ``\ldots{}, the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
+While we leave the formal construction of the optimisation problem and existing algorithms to section \ref{tour-optim}, we outline the general idea here.
+Given a random starting (current) basis, projection pursuit repeatedly searches for candidate bases nearby until it finds one with higher index value than the current basis.
+In the second round, that basis becomes the current basis and the repetitive sampling continues.
+The process ends until no better basis can be found or one of the termination criteria is reached.
+
+Before introducing the guided tour, we shall be familiar with the general tour method \citep{cook2008grand}.
+A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
+The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to \citet{buja2005computational} for the mathematical details.
+Iteratively choosing different low dimension planes and interpolating between them forms a tour path.
+Different types of tour methods choose the low dimensional planes differently and we mention two other types of tour that are commonly used.
+A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data.
+Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.
+
+Guided tour chooses the planes produced by optimising the projection pursuit index function.
+Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, which are the interpolations between two blue frames.
+The tour method has been implemented in the \emph{tourr} package in R, available on the Comprehensive R Archive Network at \url{https://cran.r-project.org/web/packages/tourr/} \citep{wickham2011tourrpackage}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.6\linewidth,height=0.4\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001}
+
+}
+
+\caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\end{figure}
 
@@
 
-Now we begin to formulate the optimisation problem. Given a randomly
-generated starting basis \(\mathbf{A}_1\), projection pursuit finds the
-final projection basis \(\mathbf{A}_T\) that satisfies the following
-optimisation problem:
+Now we begin to formulate the optimisation problem.
+Given a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit finds the final projection basis \(\mathbf{A}_T\) that satisfies the following optimisation problem:
 
@@
 
-where \(I_d\) is the d-dimensional identity matrix. The constraint
-requires the projection basis \(\mathbf{A}\) to be an orthogonal matrix
-with each column vector being orthonormal.
+where \(I_d\) is the d-dimensional identity matrix. The constraint requires the projection basis \(\mathbf{A}\) to be an orthogonal matrix with each column vector being orthonormal.
 
@@
 First of all, it is a multivariate constraint optimisation problem.
-Since the decision variables are the entries of a projection basis, it
-is required to be orthonormal. It is also likely that the objective
-function is non-differentiable or the gradient information is simply not
-available. In this case, we will need to either use some approximation
-of the gradient or turn to derivative free methods. Given the goal of
-projection pursuit as finding the basis with the largest index value,
-the optimisation problem needs to be able to find the global maximum.
-Along the way, local maximum may also be of our interest since they
-could present unexpected interesting projections. There is also one
-computational consideration: the optimisation procedure needs to be easy
-to compute since the tour animation is played in real-time.
-
-\newpage
+Since the decision variables are the entries of a projection basis, it is required to be orthonormal.
+It is also likely that the objective function is non-differentiable or the gradient information is simply not available.
+In this case, we will need to either use some approximation of the gradient or turn to derivative free methods.
+Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem needs to be able to find the global maximum.
+Along the way, local maximum may also be of our interest since they could present unexpected interesting projections.
+There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.
 
@@
 
-Below we introduce three possible algorithms: \texttt{search\_better}
-and \texttt{search\_better\_random} are derivative free methods that
-sample candidate bases in the neighbourhood, whilst
-\texttt{search\_geodesic} is an analogue of gradient ascent on the
-manifold of the projection basis.
-
-\texttt{search\_better} is a random search device that samples a
-candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
-basis \(\mathbf{A}_{\text{cur}}\) by
-\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
-where \(\alpha\) controls the radius of the sampling neighbourhood and
-\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
-same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
-then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis
-\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
-for guided tour to construct an interpolation path. The next iteration
-of search begins after adjusting \(\alpha\) by a cooling parameter:
-\(\alpha_{j+1} = \alpha_j * \text{cooling}\). Another termination
-condition is when the maximum number of iteration \(l_{\max}\) is
-reached. A slightly different cooling scheme has been proposed by
-\citet{posse1995projection} to include a halving parameter \(c\). Rather
-than reducing the radius of the searching neighbourhood, \(\alpha\), at
-each iteration, Posse's design only adjust \(\alpha\) if the last search
-takes more than \(c\) times to find an accepted basis to avoid the
-searching space being reduced too fast. The algorithm of
-\texttt{search\_better} is summarised in Algorithm \ref{random-search}.
-{[}mention orthonormalise to ensure the constraint is fulfilled; don't
-use derivative information but a random search{]}
+Below we introduce three possible algorithms: \texttt{search\_better} and \texttt{search\_better\_random} are derivative free methods that sample candidate bases in the neighbourhood, whilst \texttt{search\_geodesic} is an analogue of gradient ascent on the manifold of the projection basis.
+
+\texttt{search\_better} is a random search device that samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current basis \(\mathbf{A}_{\text{cur}}\) by \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\) where \(\alpha\) controls the radius of the sampling neighbourhood and \(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the same dimension as \(\mathbf{A}_{\text{cur}}\).
+\(\mathbf{A}_{l}\) is then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis \(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis for guided tour to construct an interpolation path.
+The next iteration of search begins after adjusting \(\alpha\) by a cooling parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\).
+Another termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
+A slightly different cooling scheme has been proposed by \citet{posse1995projection} to include a halving parameter \(c\).
+Rather than reducing the radius of the searching neighbourhood, \(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if the last search takes more than \(c\) times to find an accepted basis to avoid the searching space being reduced too fast.
+The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
+{[}mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search{]}
 
@@
 
-Simulated annealing
-\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses
-the same sampling process as \texttt{search\_better} but allow a
-probabilistic acceptance of a basis with lower index value based on a
-cooling scheme \(T(l)\). Given an initial \(T_0\), the temperature at
-iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
-candidate basis fails to have an index value larger than the current
-basis, simulated annealing gives it a second chance to be accepted with
-probability
-\[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\]
-where \(I_{(\cdot)}\) denotes the index value of a given basis. This
-implementation allows the algorithm to jump out of a local maximum and
-enables a more holistic search of the whole parameter space. This
-feature is particularly useful when the dimension of the projected space
-is smaller than the number of informative variables in the dataset
-(i.e.~a one dimensional projection of the dataset with two informative
-variables). The algorithm can be written as replacing line 5-8 of
-Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+Simulated annealing \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme \(T(l)\).
+Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
+This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
+This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.~a one dimensional projection of the dataset with two informative variables).
+The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
 
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm on the manifold
-of the projection bases. In gradient ascent, one first find the
-direction for improvment via computing the gradient information. In
-\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
-neighbourhood of the current basis, controlled by the neighbourhood
-parameter \(\delta\). A geodesic is then constructed using the current
-basis and one of the \(2n\) bases that have the highest index value. If
-the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed
-is an analogue of the gradient information in the curved space and can
-work as the searching direction. In gradient ascent, the next step is to
-conduct a line search to find the best improvement along the gradient
-direction. In \texttt{search\_geodesic}, this step is replaced by
-optimising the index value along the geodesic direction constructed
-before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\). The optima
-\(\mathbf{A}_{**}\) is outputted for the current iteration if the
-percentage change in the index value between \(\mathbf{A}_{**}\) and
-\(\mathbf{A}_{\text{cur}}\) is greater than a threshold value. As above,
-another termination condition is when \(l_{\max}\) is reached. Algorithm
-\ref{search-geodesic} summarises the steps in geodesic search.
+\citet{cook1995grand} used a gradient ascent algorithm on the manifold of the projection bases.
+In gradient ascent, one first find the direction for improvment via computing the gradient information.
+In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
+A geodesic is then constructed using the current basis and one of the \(2n\) bases that have the highest index value.
+If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
+In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
+In \texttt{search\_geodesic}, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
+The optima \(\mathbf{A}_{**}\) is outputted for the current iteration if the percentage change in the index value between \(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a threshold value.
+As above, another termination condition is when \(l_{\max}\) is reached.
+Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
 
@@
 
-\newpage
-
 \hypertarget{vis-diag}{%
@@
 
-To be able to make diagnostic plot, the optimisation algorithm should
-populate a data structure that contains the key elements of the
-algorithm. When the algorithm runs, key information regarding the
-decision variable, objective function and hyper-parameters needs to be
-recorded and stored as a data object for future analysis.
+To be able to make diagnostic plot, the optimisation algorithm should populate a data structure that contains the key elements of the algorithm.
+When the algorithm runs, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object for future analysis.
 
 \hypertarget{data-structure-for-diagnostics}{%
-\subsection{Data structure for
-diagnostics}\label{data-structure-for-diagnostics}}
-
-In the optimisation algorithms for projection pursuit, three main
-elements to record are 1) projection bases: \(\mathbf{A}\), 2) index
-values: \(I\), and 3) State: \(S\), which labels the observation with
-detailed stage in the optimisation. Multiple iterators are also needed
-to index the data collected at different levels. \(t\) is a unique
-identifier that prescribes the natural ordering of each observation;
-\(j\) is the counter for each search-and-interpolate round, which
-remains the same within one round and has an increment of one once a new
-round starts. \(l\) is the counter for each search/interpolation
-allowing us to know how many basis the algorithm has searched before
-finding one to output. There are other parameters that are of our
-interest and we denote them as \emph{\(V_{p}\)}. In projection pursuit,
-this includes \(V_1 = \text{method}\), which tags the name of the
-algorithm used and \(V_2 = \text{alpha}\), the neighbourhood parameter
-that controls the size in sampling candidate bases. A matrix notation of
-the data structure is presented in Equation \ref{eq:data-structure}.
+\subsection{Data structure for diagnostics}\label{data-structure-for-diagnostics}}
+
+In the optimisation algorithms for projection pursuit, three main elements to record are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3) State: \(S\), which labels the observation with detailed stage in the optimisation.
+Multiple iterators are also needed to index the data collected at different levels.
+\(t\) is a unique identifier that prescribes the natural ordering of each observation; \(j\) is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts.
+\(l\) is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
+There are other parameters that are of our interest and we denote them as \emph{\(V_{p}\)}.
+In projection pursuit, this includes \(V_1 = \text{method}\), which tags the name of the algorithm used and \(V_2 = \text{alpha}\), the neighbourhood parameter that controls the size in sampling candidate bases.
+A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
 
@@
 
-where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that we deliberately
-denote the last round of search as \(j = J+1\) and in that round there
-is no output/interpolation basis and the algorithm terminates. This
-notation allows us to denote the last complete search-and-interpolate
-round as round \(J\) and hence the final basis is \(A_T\) and highest
-index value found is \(I_T\).
-
-{[}outside the paper: I find the notation of current/target basis is
-confusing because the target basis in round \(j\) becomes the current
-basis in round \(j+1\). Also, when we start to have polish, the target
-basis may not be the current basis in the next round\ldots{} The place
-where current/target is most appropriate is probably when describing the
-interpolation where the first one is always the current basis and the
-last is always the target basis. I think it is better to leave this
-language in the code{]}
+where \(T^{\prime} = T + k_{J}+ l_{J+1}\).
+Note that we deliberately denote the last round of search as \(j = J+1\) and in that round there is no output/interpolation basis and the algorithm terminates.
+This notation allows us to denote the last complete search-and-interpolate round as round \(J\) and hence the final basis is \(A_T\) and highest index value found is \(I_T\).
+
+{[}outside the paper: I find the notation of current/target basis is confusing because the target basis in round \(j\) becomes the current basis in round \(j+1\).
+Also, when we start to have polish, the target basis may not be the current basis in the next round\ldots{} The place where current/target is most appropriate is probably when describing the interpolation where the first one is always the current basis and the last is always the target basis.
+I think it is better to leave this language in the code{]}
 
@@
 
-It is worth noticing that the data structure constructed above meets the
-tidy data principle \citep{wickham2014tidy} that states
+It is worth noticing that the data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that states
 
@@
 
-The wrangling and visualisation of tidy data have been greatly
-simplified by the well-known dplyr\citep{dplyr} and
-ggplot2\citep{ggplot2} package.
-
-With a constructed data object, the construction of diagnostic plots is
-inspired by the concept of grammar of graphic
-\citep{wickham2010layered}, which powers the primary graphical system in
-R, ggplot2 \citep{ggplot2}. In grammar of graphic, plots are not defined
-by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by
-``stacked layers''. Using this design, ggplot does not have to develop a
-gazillion of functions that each produces a different type of plot from
-a different data structure. Instead, it aesthetically maps variables
-(and its statistical transformation) in a dataset to different geometric
-objects (points, lines, box-and-whisker etc) and builds the plot through
-overlaying different layers.
+The wrangling and visualisation of tidy data have been greatly simplified by the well-known dplyr\citep{dplyr} and ggplot2\citep{ggplot2} package.
+
+With a constructed data object, the construction of diagnostic plots is inspired by the concept of grammar of graphic \citep{wickham2010layered}, which powers the primary graphical system in R, ggplot2 \citep{ggplot2}.
+In grammar of graphic, plots are not defined by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by ``stacked layers''.
+Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
+Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
 
 \hypertarget{check-how-hard-the-optimiser-is-working}{%
-\subsection{Check how hard the optimiser is
-working}\label{check-how-hard-the-optimiser-is-working}}
-
-A primary interest of diagnosing an optimisation algorithm is to study
-how it progressively finds its optimum. One way of doing it is to plot
-the index value across its natural ordering \(t\), however, it usually
-takes the algorithm much longer to find a better basis than the current
-one towards the end of the search. When plotting the searching
-observations as points in a plot, the space each iteration takes will be
-proportional to the number of points in that iteration.
-
-Another option is to use summarisation for each iteration. Boxplot is a
-suitable candidate that provides five points summary of the data,
-however, there are two pieces of information missing from the boxplot:
-1) It does not report the number of points, and 2) the position of the
-last basis. This could be remedied by adding more layers using the
-concept of grammar of graphics. A label geometry is added at the bottom
-of the plot to show the number of points in each iteration and a line
-geometry links the last basis in each iteration. Further, an option to
-switch between displaying points and boxplot geometry is helpful since
-point geometry can be more intuitive for the iteration with few
-observations. This is achieved via a \texttt{cutoff} parameter. Below we
-define the searching point plot based on four different component layers
+\subsection{Check how hard the optimiser is working}\label{check-how-hard-the-optimiser-is-working}}
+
+A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum.
+One way of doing it is to plot the index value across its natural ordering \(t\), however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
+When plotting the searching observations as points in a plot, the space each iteration takes will be proportional to the number of points in that iteration.
+
+Another option is to use summarisation for each iteration.
+Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis.
+This could be remedied by adding more layers using the concept of grammar of graphics.
+A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration.
+Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
+This is achieved via a \texttt{cutoff} parameter.
+Below we define the searching point plot based on four different component layers
 
@@
   \item
-    data: group by \(j\) and filter the observations in the groups that
-    have count greater than \texttt{cutoff\ =\ 15}.
+    data: group by \(j\) and filter the observations in the groups that have count greater than \texttt{cutoff\ =\ 15}.
   \item
@@
   \item
-    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\)
-    is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\)
-    finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the
-    index value of all the searching bases defined in Matrix
-    \ref{eq:data-structure}.
+    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\) is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\) finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
   \end{itemize}
@@
   \item
-    data: group by \(j\) and filter the observations in the group that
-    have count less than \texttt{cutoff\ =\ 15}.
+    data: group by \(j\) and filter the observations in the group that have count less than \texttt{cutoff\ =\ 15}.
   \item
@@
   \item
-    y: the y-axis for each label is \(0.99 * \text{MIN}\), where
-    \(\text{MIN}\) is the smallest index value in the whole data object
+    y: the y-axis for each label is \(0.99 * \text{MIN}\), where \(\text{MIN}\) is the smallest index value in the whole data object
   \item
@@
 
-Figure \ref{toy-search} presents a comparison between the two plotting
-options discussed above. Using the natural order to plot searching
-points distorts the point from the first few iterations and
-over-emphasizes the searches in the last few iterations. The second plot
-design evenly presents the summarised information for each iteration
-while allowing for a switch to the full information for the iteration
-with small number of observations.
+Figure \ref{toy-search} presents a comparison between the two plotting options discussed above.
+Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
+The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.
 
@@
 \includegraphics{paper_files/figure-latex/toy-search-1.pdf}
-\caption{\label{toy-search}A comparison of plotting the same search
-points with different plot designs. In the upper plot, points from the
-last three iterations span the vast majority of the plotting space
-leaving the first few iterations being squeezed together whilst the
-lower plot spaces each iteration evenly and presents summary information
-easy to read.}
-\end{figure}
-
-\newpage
+\caption{\label{fig:toy-search}\label{toy-search}A comparison of plotting the same search points with different plot designs.
+In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read.}
+\end{figure}
 
 \hypertarget{examining-the-optimisation-progress}{%
-\subsection{Examining the optimisation
-progress}\label{examining-the-optimisation-progress}}
-
-Sometimes, we may be interested in exploring the points on the
-interpolation path since these are the points that will be played by the
-tour animation. The plot definition is as follows:
+\subsection{Examining the optimisation progress}\label{examining-the-optimisation-progress}}
+
+Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. The plot definition is as follows:
 
@@
   \item
-    data: filter the observations with \(S = \text{interpolation}\) and
-    mutate \(t\) to be the row number of the subsetted tibble
+    data: filter the observations with \(S = \text{interpolation}\) and mutate \(t\) to be the row number of the subsetted tibble
   \item
@@
 
-Figure \ref{toy-interp} presents the interpolation of three different
-tour paths. The upper plot shows a desirable interpolation in each
-iteration with the index value being progressively and monotonically
-increasing. While in the middle plot, the increases towards the target
-basis is not monotonical in the last two iterations and interpolated
-basis with higher index value can be found on the tour path. The lower
-plot is constructed using \texttt{search\_better\_random}, where a basis
-with smaller index value has a probabilistic chance of being accepted
-and we can observe a much more involved pattern. In iteration three, the
-probabilistic acceptance produces a monotonically decreasing
-interpolation whilst in iteration five, six and seven, with also a
-probabilistic acceptance, the interpolation is now non-monotonical. When
-the target basis has a higher index value, iteration eight reaches this
-basis by first decreasing the index value. While the interpolation plot
-shows different possibilities of the change in index value on the tour
-path, the diagnosis of the validity of each pattern and the related
-optimisation algorithms will be postponed to Section \ref{application}.
+Figure \ref{toy-interp} presents the interpolation of three different tour paths.
+The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
+While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
+The lower plot is constructed using \texttt{search\_better\_random}, where a basis with smaller index value has a probabilistic chance of being accepted and we can observe a much more involved pattern.
+In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
+When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
+While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
 
@@
 \includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
-\caption{\label{toy-interp} This is a toy example of plotting
-interpolated points}
-\end{figure}
-
-\newpage
+\caption{\label{fig:toy-interp}\label{toy-interp} This is a toy example of plotting interpolated points}
+\end{figure}
+
+~
 
 \hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
-\subsection{Understanding the optimiser's coverage of the search
-space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
-
-Apart from checking the progression of an optimiser, another interesting
-aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases
-\(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\)
-dimension sphere, where the dimension can easily go above five (5).
-Visualising the search paths on the original high diemnsional sphere
-would require skills from the viewer to preceive rotation of geometry in
-higher dimensional space (d \textgreater{} 3) while an easier
-alternative is to view the reduced space via some dimension reduction
-methods i.e.~Principal component analysis. To better visual the serach
-path as an embedding of a hollow sphere, random points on the high
-dimensional sphere is generated using package \texttt{geozoo} and PCA is
-conducted on both the bases and the points on the surface of the sphere.
+\subsection{Understanding the optimiser's coverage of the search space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
+
+Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
+Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go above five (5).
+Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d \textgreater{} 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.~Principal component analysis.
+To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
 
@@
   \item
-    data: subset the basis of interest and arrange into a matrix format;
-    perform PCA on the basis matrix and compute the projected basis on
-    the first two principal components; bind the variables from the
-    original global object and form a tibble
+    data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
   \item
@@
 
-Figure \ref{toy-pca} shows the first two principal components of two
-search paths. Starting from the same position, two searching methods
-take different paths to get its final bases. There are two theoretical
-bases because they correspond to the same data projection \(Y\) with a
-180 degree rotation. What differentiate the two optimisers is that
-\texttt{search\_better()} also involve random evaluation of points in
-the sphere during its sampling process while \texttt{search\_geodesic()}
-doesn't and this stochastic rather than deterministic approach can be
-useful in complex scenario.
+Figure \ref{toy-pca} shows the first two principal components of two search paths.
+Starting from the same position, two searching methods take different paths to get its final bases.
+There are two theoretical bases because they correspond to the same data projection \(Y\) with a 180 degree rotation.
+What differentiate the two optimisers is that \texttt{search\_better()} also involve random evaluation of points in the sphere during its sampling process while \texttt{search\_geodesic()} doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
 
@@
 \includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
-\caption{\label{toy-pca}PCA plot of search geodesic colouring by info
-allows for better understanding of each stage in the geodesic search}
+\caption{\label{fig:toy-pca}\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}
 \end{figure}
@@
 \hypertarget{animating-the-diagnostic-plots}{%
-\subsection{Animating the diagnostic
-plots}\label{animating-the-diagnostic-plots}}
-
-Animating the plots introduced above is useful, especially in the case
-of PCA polot since it shows the bases found by the optimiser in its
-natural order.
+\subsection{Animating the diagnostic plots}\label{animating-the-diagnostic-plots}}
+
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
 
@@
 \hypertarget{the-tour-looking-at-itself}{%
-\subsection{The tour looking at
-itself}\label{the-tour-looking-at-itself}}
-
-Viewing the bases on the reduced space via PCA shed some lights on the
-space the optimisers have explored, the visualisation on the original
-\(p \times d\) dimension enables a stereoscopic view of the search. To
-view a high dimensional (\(d \ge 3\)) object on a screen, an approach is
-to play the rotation of the object in animation. This can be done via a
-regular grand tour or a slice tour \citep{laa2020slice}, which
-emphasizes the points that are close to a section of the sphere.
-
-Compare to the PCA plot, the animated rotation (tour) display, in Figure
-\ref{toy-tour}, gives a more holistic view of the search of the
-optimiser and these additional information gained from the animation is
-cruicial. This is because in essense, PCA presents one reduced space
-that maximises the variance and it inherently amplifies the spread of
-the randomly evaluated bases in the search since they have larger
-variance than the bases on the interpolation path. Hence when the
-interest is to view the interpolation path in the space, an animated
-rotation display provides the view of high dimension from different
-angles.
+\subsection{The tour looking at itself}\label{the-tour-looking-at-itself}}
+
+Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a stereoscopic view of the search.
+To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation.
+This can be done via a regular grand tour or a slice tour \citep{laa2020slice}, which emphasizes the points that are close to a section of the sphere.
+
+Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
+Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
 
@@
 \includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
-\caption{\label{toy-tour}A selected number of frames from the tour
-animation for viewing the 5D space of all the projection bases. The
-second frame on the top row view the space from a direction that is
-close to the one in PCA plot. The tour animation allows for a more
-holistic view of the full space in high dimensions from different
-angles.}
-\end{figure}
-
-\newpage
+\caption{\label{fig:toy-tour}\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+The second frame on the top row view the space from a direction that is close to the one in PCA plot.
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}
+\end{figure}
 
@@
 
-For a particular index function, the best algorithm to optimise is
-related to the character of the index and the data. If the index
-function is smooth and has single maximum, all of the three algorithms
-introduced above can find maximum. When multiple optima are presented,
-\texttt{search\_better} may stuck in the local maximum. When the index
-function is non-smooth, \texttt{search\_geodesic} may even fail to find
-the maximum. In this section, examples will be presented to outline how
-the diagnostic plots can be used to compare the performance of the
-algorithms in different scenarios.
+For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
+If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
+When multiple optima are presented, \texttt{search\_better} may stuck in the local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
+In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.
 
@@
 
-Random variables with different structures has been simulated and the
-distribution of each is presented in Equation \ref{eq:sim-norm} to
-\ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and
-\texttt{x10} are normal distributed with zero mean and unit variance and
-\texttt{x2} to \texttt{x7} are mixture of normal distributions with
-different weights and locations. Figure \ref{origin-data} plots the
-histogram of each variable except \texttt{x3}. The mixture variables are
-scaled to have unit variance before running the projection pursuit.
+Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are normal distributed with zero mean and unit variance and \texttt{x2} to \texttt{x7} are mixtures of normal distributions with varied weights and locations.
+The mixture variables have been scaled to have an overall unit variance before running the projection pursuit.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/origin-data-1.pdf}
-\caption{\label{origin-data} The histogram of simulated selected
-variables.}
-\end{figure}
-
 \hypertarget{a-problem-of-not-monotonic}{%
-\subsection{A problem of not
-monotonic}\label{a-problem-of-not-monotonic}}
-
-We use the same dataset as the toy example above to explore the search
-function \texttt{search\_better} and we want to learn how the index
-value changes on the interpolation path for the \texttt{holes} index.
-From the left panel of Figure \ref{interruption}, we observe that when
-interpolating from the current basis to the target basis, the index
-value may not be monotone: we could reach a basis with a higher index
-value than the target basis on the interpolation path. In this sense, we
-would be better off using the basis with the highest index value on the
-interpolation path as the current basis for the next iteration (rather
-than using the target basis). Hence, an interruption is constructed to
-accept the interpolating bases only up to the one with the largest index
-value. After implementing this interruption, the search finds higher
-final index value with fewer steps as shown in the right panel of Figure
-\ref{interruption}.
+\subsection{A problem of not monotonic}\label{a-problem-of-not-monotonic}}
+
+We use the same dataset as the toy example above to explore the search function \texttt{search\_better} and we want to learn how the index value changes on the interpolation path for the \texttt{holes} index.
+From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
+In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
+Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
+After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{interruption}.
 
@@
 \includegraphics{paper_files/figure-latex/interruption-1.pdf}
-\caption{\label{interruption}Trace plots of the interpolated basis with
-and without interruption. The interruption stops the interpolation when
-the index value starts to decrease at id = 60. The implementation of the
-interuption finds an ending basis with higher index value using fewer
-steps.}
+\caption{\label{fig:interruption}\label{interruption}Trace plots of the interpolated basis with and without interruption.
+The interruption stops the interpolation when the index value starts to decrease at id = 60.
+The implementation of the interuption finds an ending basis with higher index value using fewer steps.}
 \end{figure}
@@
 \hypertarget{close-but-not-close-enough}{%
-\subsection{Close but not close
-enough}\label{close-but-not-close-enough}}
-
-In principle, all the optimisation routines should result in the same
-output for the same problem while this may not be the case in real
-application. This motivates the creation of a polishing search that
-polishes the final basis found and achieves unity across different
-methods.
-
-\texttt{search\_polish} takes the final basis of a given search as a
-start and uses a brutal-force approach to sample a large number of basis
-(\texttt{n\_sample}) in the neighbourhood. Among those sampled basis,
-the one with the largest index value is chosen to be compared with the
-current basis. If its index value is larger than that of the current
-basis, it becomes the current basis in the next iteration. If no basis
-is found to have larger index value, the searching neighbourhood will
-shrink and the search continues. The polishing search ends when one of
-the four stopping criteria is satisfied:
+\subsection{Close but not close enough}\label{close-but-not-close-enough}}
+
+In principle, all the optimisation routines should result in the same output for the same problem while this may not be the case in real application.
+This motivates the creation of a polishing search that polishes the final basis found and achieves unity across different methods.
+
+\texttt{search\_polish} takes the final basis of a given search as a start and uses a brutal-force approach to sample a large number of basis (\texttt{n\_sample}) in the neighbourhood.
+Among those sampled basis, the one with the largest index value is chosen to be compared with the current basis.
+If its index value is larger than that of the current basis, it becomes the current basis in the next iteration.
+If no basis is found to have larger index value, the searching neighbourhood will shrink and the search continues.
+The polishing search ends when one of the four stopping criteria is satisfied:
 
@@
 
-The usage of search\_polish is as follows. After the first search, the
-final basis from the interpolation is extracted and supplied to the
-second search as the \texttt{start} argument. \texttt{search\_polish} is
-used as the search function. All the other arguments should remain the
-same.
+The usage of search\_polish is as follows.
+After the first search, the final basis from the interpolation is extracted and supplied to the second search as the \texttt{start} argument.
+\texttt{search\_polish} is used as the search function.
+All the other arguments should remain the same.
 
@@
 \begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(tourr)}
+\KeywordTok{library}\NormalTok{(ferrn)}
+
 \KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
-\NormalTok{holes_2d_geo <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)],}\DataTypeTok{tour_path =}
-                             \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
-                                         \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_geodesic),}
+\NormalTok{holes_2d_geo <-}\StringTok{ }
+\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6,}
+             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
+                                     \DataTypeTok{search_f =}\NormalTok{ search_geodesic),}
              \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}
 
-\NormalTok{last_basis <-}\StringTok{ }\NormalTok{holes_2d_geo }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(info }\OperatorTok{==}\StringTok{ "interpolation"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{tail}\NormalTok{(}\DecValTok{1}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}
+\NormalTok{last_basis <-}\StringTok{ }\KeywordTok{get_best}\NormalTok{(holes_2d_geo) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}
 
 \KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
-\NormalTok{holes_2d_geo_polish <-}\StringTok{ }\KeywordTok{animate_xy}\NormalTok{(data_mult[,}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{, }\DecValTok{7}\OperatorTok{:}\DecValTok{10}\NormalTok{)], }\DataTypeTok{tour_path =}
-                                    \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
-                                                \DataTypeTok{search_f =}\NormalTok{ tourr}\OperatorTok{:::}\NormalTok{search_polish),}
-                                  \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }
-                                  \DataTypeTok{start =}\NormalTok{ last_basis)}
+\NormalTok{holes_2d_geo_polish <-}\StringTok{ }
+\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6, }
+             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
+                                     \DataTypeTok{search_f =}\NormalTok{ search_polish),}
+             \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{start =}\NormalTok{ last_basis)}
 \end{Highlighting}
@@
 
-A slight variation of the plot definition due to the addition of
-polishing points is as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: filter the observations with \(S = \text{interpolation}\);
-    bind the data object from optimisation and interpolation and form
-    polishing; mutate \(t\) to be the row number of the binded tibble.
-  \item
-    x: \(t\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \item
-    colour: \(V\) is mapped to the colour aesthetic
-  \end{itemize}
-\item
-  Layer 2: line geom
-\end{itemize}
-
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-Again using the same data, we are interested to compare the effect of
-different \texttt{max.tries} in the 2D projection setting.
-\texttt{max.tries} is a hyperparameter that controls the maximum number
-of try before the search ends. The default value of 25 is suitable for
-1D projection while we suspect it may not be a good option for the 2D
-case and hence want to compare it with an alternative, 500. As shown in
-Figure \ref{trace-compare}, both trials attain the same index value
-after polishing while the small \texttt{max.tries} of 25 is not
-sufficient for \texttt{search\_better} to find its global maximum and we
-will need to adjust the \texttt{max.tries} argument for the search to
-sufficiently explore the parameter space.
+Checking the final projection before and after applying the polish as in Figure \ref{geo-polish}, we can see that \ldots{}
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/polish-1.pdf}
-\caption{\label{trace-compare}Breakdown of index value when using
-different max.tries in search better in conjunction with search polish.
-Both attain the same final index value after the polishing while using a
-\texttt{max.tries} of 25 is not sufficient to find the true maximum.}
-\end{figure}
-
-While explore the reduced space is an initial attempt to understand the
-searching space, there are existing technology for rotating a higher
-dimensional space for visualisation. Geozoo is an option. It generates
-random points on the high dimensional space and we can overlay it with
-the points on the optimisation path to visualise the spread of it on the
-high-D sphere.
-
-{[}add example from geozoo{]}
+\includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
+\caption{\label{fig:geo-polish}\label{geo-polish}On the left is the final projection from search geodesic while on the right is the polished result.
+Polishing helps to sharpen the edges of the clusters}
+\end{figure}
+
+While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
+Geozoo is an option.
+It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
 
 \hypertarget{seeing-the-signal-in-the-noise}{%
-\subsection{Seeing the signal in the
-noise}\label{seeing-the-signal-in-the-noise}}
-
-The interpolation path of holes index, as seen in Figure
-\ref{interruption}, is smooth, while this may not be the case for more
-complicated index functions. \texttt{kol\_cdf} index, an 1D projection
-index function based on Kolmogorov test, compares the difference between
-the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly
-generated normal distribution, \(y_n\) based on the empirical cumulated
-distribution function (ECDF). Denotes the ECDF function as \(F(u)\) with
-subscript indicating the variable, the Kolmogorov statistics defined by
+\subsection{Seeing the signal in the noise}\label{seeing-the-signal-in-the-noise}}
+
+The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
+\texttt{kol\_cdf} index, an 1D projection index function based on Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
+Denotes the ECDF function as \(F(u)\) with subscript indicating the variable, the Kolmogorov statistics defined by
 
@@
 
-can be seen as a function of the projection matrix
-\(\mathbf{A}_{p \times 1}\) and hence a valid index function.
-
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating
-points when using different optimisation algorithms:
-\texttt{search\_geodesic} and \texttt{search\_better}. One can observe
-that
-
-\begin{itemize}
-\tightlist
-\item
-  The index value of \texttt{kol\_cdf} index is much smaller than that
-  of holes index
+can be seen as a function of the projection matrix \(\mathbf{A}_{p \times 1}\) and hence a valid index function.
+
+Figure \ref{kol-cdf} compares the tracing plot of the interpolating points when using different optimisation algorithms: \texttt{search\_geodesic} and \texttt{search\_better}. One can observe that
+
+\begin{itemize}
+\tightlist
+\item
+  The index value of \texttt{kol\_cdf} index is much smaller than that of holes index
 \item
@@
 
-Polishing step has done much more work to find the final index value in
-\texttt{search\_geodesic} than \texttt{search\_better} and this
-indicates \texttt{kol\_cdf} function favours of a random search method
-than ascent method.
+Polishing step has done much more work to find the final index value in \texttt{search\_geodesic} than \texttt{search\_better} and this indicates \texttt{kol\_cdf} function favours of a random search method than ascent method.
 
@@
 \includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{kol-cdf}Comparison of two different searching methods:
-search\_geodesic and search\_better on 1D projection problem for a
-noisier index: kol\_cdf. The geodesic search rely heavily on the
-polishing step to find the final index value while search better works
-well.}
-\end{figure}
-
-Now we enlarge the dataset to include two informative variables:
-\texttt{x2} and \texttt{x3} and remain 1D projection. In this case, two
-local maxima appear with projection matrix being \([0, 1, 0, 0, 0, 0]\)
-and \([0, 0, 1 ,0, 0, 0]\).
-
-Using different seeds in \texttt{search\_better} allows us to find both
-local maxima d as in Figure \ref{1d-2var-different-seeds}. Comparing the
-maximum of both, we can see that the global maximum happens when
-\texttt{x2} is found. It is natural to ask then if there is an algorithm
-that can find the global maximum without trying on different seeds?
-\texttt{search\_better\_random} manages to do it via a
-Metropolis-hasting random search as shown in Figure
-\ref{1d-2var-better-random}, although at a higher cost of number of
-points to evaluate.
+\caption{\label{fig:kol-cdf}\label{kol-cdf}Comparison of two different searching methods: search\_geodesic and search\_better on 1D projection problem for a noisier index: kol\_cdf.
+The geodesic search rely heavily on the polishing step to find the final index value while search better works well.}
+\end{figure}
+
+Now we enlarge the dataset to include two informative variables: \texttt{x2} and \texttt{x3} and remain 1D projection.
+In this case, two local maxima appear with projection matrix being \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\).
+
+Using different seeds in \texttt{search\_better} allows us to find both local maxima d as in Figure \ref{1d-2var-different-seeds}.
+Comparing the maximum of both, we can see that the global maximum happens when \texttt{x2} is found.
+It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds?
+\texttt{search\_better\_random} manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
+In Figure \ref{kol-pca},
+sing search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed.
+We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.
+As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better\_random produces a more exhaustive search.
+The large overlapping of better1 and better\_random is explained by the fact that both algorithms find x2 in the end.
+The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
+Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum.
 
@@
 \centering
-\includegraphics{paper_files/figure-latex/1d-2var-different-seeds-1.pdf}
-\caption{\label{1d-2var-different-seeds}The trace plot search better in
-a 1D projection problem with two informative variables using different
-seeds (without polishing). Since there are two informative variables,
-setting different value for seed will lead search better to find either
-of the local maximum.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/1d-2var-better-random-1.pdf}
-\caption{\label{1d-2var-better-random}Using search better random for the
-problem above will result in finding the global maximum but much larger
-number of iteration is needed.}
-\end{figure}
-
-We can also plot the searching points of all three algorithms in the
-searching space and explore their relative position against each other
-using principal components. As shown in Figure
-\ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only
-search a proportion of the searching space while better\_random produces
-a more exhaustive search. The large overlapping of better1 and
-better\_random is explained by the fact that both algorithms find x2 in
-the end.
+\includegraphics{paper_files/figure-latex/kol-pca-1.pdf}
+\caption{\label{fig:kol-pca}\label{kol-pca}THis is the caption}
+\end{figure}
 
@@
 
-(further notice on the implementation in the tour package will be
-discussed in the tour developer meeting.) The implementation of this
-projection has been divided into two pacakges, where the data collection
-happens when a tour is running and hence implemented in the existing
-\texttt{tourr} package. The diagnostics of the algorithm via plots and
-other auxiliary functions have been implemented in a new pacakge,
-\texttt{ferrn}.
-
-When the optimisation ends, the data object will be stored and printed
-(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
-Additional messages during the optimisation can be displayed by argument
-\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
-list-column \texttt{basis} to be printed out nicely with the dimension
-of the projection basis readily available. Below presents a sample of
-the data object.
+(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
+The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing \texttt{tourr} package.
+The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, \texttt{ferrn}.
+
+When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
+Additional messages during the optimisation can be displayed by argument \texttt{verbose\ =\ TRUE}.
+Notice that the tibble object allows the list-column \texttt{basis} to be printed out nicely with the dimension of the projection basis readily available.
+Below presents a sample of the data object.
 
@@
 
-Once the data object has been obtained, the package, \texttt{ferrn},
-provides automatic diagnostic plots as described in Section
-\ref{vis-diag}. The structure of package functionality has been listed
-below.
+Once the data object has been obtained, the package, \texttt{ferrn}, provides automatic diagnostic plots as described in Section \ref{vis-diag}.
+The structure of package functionality has been listed below.
 
@@
 \item
-  \texttt{explore\_trace\_search()}: produce summary plots, as shown in
-  Figure \ref{toy-search}
-\item
-  \texttt{explore\_trace\_interp()}: produce trace plots for the
-  interpolation points, as shown in Figure \ref{toy-interp}
-\item
-  \texttt{explore\_space\_pca()}: produce plots of projection basis on
-  the reduced space by PCA, as shown in Figure \ref{toy-pca}. Animated
-  version in Figure \ref{toy-pca-aniamted} can be acquired via the
-  argument \texttt{animate\ =\ TRUE}
-\item
-  \texttt{explore\_space\_tour()}: produce animated tour view on the
-  full space of the projection bases, as shown in Figure \ref{toy-tour}
+  \texttt{explore\_trace\_search()}: produce summary plots, as shown in Figure \ref{toy-search}
+\item
+  \texttt{explore\_trace\_interp()}: produce trace plots for the interpolation points, as shown in Figure \ref{toy-interp}
+\item
+  \texttt{explore\_space\_pca()}: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{toy-pca}.
+Animated version in Figure \ref{toy-pca-aniamted} can be acquired via the argument \texttt{animate\ =\ TRUE}
+\item
+  \texttt{explore\_space\_tour()}: produce animated tour view on the full space of the projection bases, as shown in Figure \ref{toy-tour}
 \end{itemize}
@@
 \item
-  \texttt{get\_*()} extract and, in some cases, manipulate certain
-  components from the existing data object.
+  \texttt{get\_*()} extract and, in some cases, manipulate certain components from the existing data object.
 
@@
   \item
-    \texttt{get\_interp()}: extract the observations in the
-    interpolation and prepare to be supplied to
-    \texttt{explore\_trace\_interp()}.
-  \item
-    \texttt{get\_search\_count()}: produce the summary table of the
-    number of observation in each iteration and prepare to be supplied
-    to \texttt{explore\_trace\_search()}
-  \item
-    \texttt{get\_basis\_matrix()}: flattern all the bases into a row and
-    stack them to form a large matrix for
-    \texttt{explore\_space\_tour()}
-  \end{itemize}
-\item
-  \texttt{bind\_*()} incorporate additional information that is outside
-  the tour optimisation.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \texttt{bind\_theoretical()}: incorporate the best possible basis to
-    the existing data object with the supply of the index function and
-    original data for producing the index value.
-  \item
-    \texttt{bind\_random()}: generates a large number of points on the
-    surface of a high dimensional sphere and attaches it to the existing
-    data object and output as a tibble object.
+    \texttt{get\_interp()}: extract the observations in the interpolation and prepare to be supplied to \texttt{explore\_trace\_interp()}.
+  \item
+    \texttt{get\_search\_count()}: produce the summary table of the number of observation in each iteration and prepare to be supplied to \texttt{explore\_trace\_search()}
+  \item
+    \texttt{get\_basis\_matrix()}: flattern all the bases into a row and stack them to form a large matrix for \texttt{explore\_space\_tour()}
+  \end{itemize}
+\item
+  \texttt{bind\_*()} incorporate additional information that is outside the tour optimisation.
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \texttt{bind\_theoretical()}: incorporate the best possible basis to the existing data object with the supply of the index function and original data for producing the index value.
+  \item
+    \texttt{bind\_random()}: generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object and output as a tibble object.
 \texttt{bind\_random\_matrix()} returns a matrix.
@@
   \item
-    \texttt{botanical\_palettes}: a collection of color palettes from
-    Australian native plants Quantitative palettes include daisy,
-    banksia and cherry and sequantial palettes cotain fern and acacia.
-  \item
-    \texttt{botanical\_pal()}: a color interpolator useful for
-    \texttt{scale\_color\_botanical()}
-  \item
-    \texttt{scale\_color\_botanical()}: a ggplot construction for using
-    botanical palettes.
-  \item
-    \texttt{relevel\_geo()}: manipulate the level in the \texttt{info}
-    column in the data object when geodesic search is used
-  \item
-    \texttt{relevel\_better()}: manipulate the level in the
-    \texttt{info} column in the data object when search\_better related
-    search is used
+    \texttt{botanical\_palettes}: a collection of color palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequantial palettes cotain fern and acacia.
+  \item
+    \texttt{botanical\_pal()}: a color interpolator useful for \texttt{scale\_color\_botanical()}
+  \item
+    \texttt{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.
+  \item
+    \texttt{relevel\_geo()}: manipulate the level in the \texttt{info} column in the data object when geodesic search is used
+  \item
+    \texttt{relevel\_better()}: manipulate the level in the \texttt{info} column in the data object when search\_better related search is used
   \end{itemize}
@@
 
-This paper has illustrated setting up a data object that can be used for
-diagnosing a complex optimisation procedure. The ideas were illustrated
-using the optimisers available for projection pursuit guided tour. Here
-the constraint is the orthornormality condition of the projection bases.
-The approach used here could be broadly applied to understand other
-constrained optimisers.
+This paper has illustrated setting up a data object that can be used for diagnosing a complex optimisation procedure.
+The ideas were illustrated using the optimisers available for projection pursuit guided tour.
+Here the constraint is the orthornormality condition of the projection bases.
+The approach used here could be broadly applied to understand other constrained optimisers.
 

### 2020-10-07 paper/paper.Rmd
refine 5.3 Close but not close enough

@@
 
-In principle, all the optimisation routines should result in the same output for the same problem while this may not be the case in real application.
-This motivates the creation of a polishing search that polishes the final basis found and achieves unity across different methods.
-
-`search_polish` takes the final basis of a given search as a start and uses a brutal-force approach to sample a large number of basis (`n_sample`) in the neighbourhood.
-Among those sampled basis, the one with the largest index value is chosen to be compared with the current basis.
-If its index value is larger than that of the current basis, it becomes the current basis in the next iteration.
-If no basis is found to have larger index value, the searching neighbourhood will shrink and the search continues.
-The polishing search ends when one of the four stopping criteria is satisfied:
-
-1) the chosen basis can't be too close to the current basis 
-2) the percentage improvement of the index value can't be too small 
-3) the searching neighbourhood can't be too small
-4) the number of iteration can't exceed `max.tries`
-
-The usage of search_polish is as follows.
-After the first search, the final basis from the interpolation is extracted and supplied to the second search as the `start` argument.
-`search_polish` is used as the search function.
-All the other arguments should remain the same.
+Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
+This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
+The following code demonstrates the usage of `search_polish`:
 
@@
 
-Checking the final projection before and after applying the polish as in Figure \ref{geo-polish}, we can see that ...
-
-```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}On the left is the final projection from search geodesic while on the right is the polished result.
-Polishing helps to sharpen the edges of the clusters"}
+Similar to `search_better` as a stochastic random search, `search_polish` has a different scheme of reducing the search neighbourhood.
+In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
+On the contrary, `search_polish` allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
+Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries`.
+These include:
+
+1) the distance between the candidate basis and the current basis needs to be larger than 1e-3;
+2) the percentage change of the index value need to be larger than 1e-5; and
+3) the alpha parameter on itself need to be larger than 0.01
+
+Figure \ref{geo-polish} presents the final projections found before and after applying `search_polish` using the demo code.
+Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables, namely `V1`, `V8`, `V9` and `V10`.
+When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
+
+```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}Projected data before and after polishing.
+Weights on axis other than V2 and V7 have been further reduced and the edges of each cluster have become more obvious."}
 # set.seed(123456)
@@
 
+# set.seed(123456)
+# holes_2d_geo <- animate_xy(boa6,tour_path =
+#                              guided_tour(holes(), d = 2,
+#                                          search_f = search_geodesic),
+#                            rescale = FALSE, verbose = TRUE)
+# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
+# 
+# set.seed(123456)
+# holes_2d_geo_polish <- animate_xy(boa6, tour_path =
+#                                     guided_tour(holes(), d = 2,
+#                                                 search_f = search_polish, alpha = 0.1),
+#                                   rescale = FALSE, verbose = TRUE,
+#                                   start = last_basis)
+# 
+# save(holes_2d_geo, file = here::here("data", "holes_2d_geo.rda"))
+# save(holes_2d_geo_polish, file = here::here("data", "holes_2d_geo_polish.rda"))
+
 before <- png::readPNG(here::here("figures","geo_polish", "before100.png"))
@@
 ```
-
-<!-- --- -->
-
-<!-- Again using the same data, we are interested to compare the effect of different `max.tries` in the 2D projection setting.
-`max.tries` is a hyperparameter that controls the maximum number of try before the search ends.
-The default value of 25 is suitable for 1D projection while we suspect it may not be a good option for the 2D case and hence want to compare it with an alternative, 500.
-As shown in Figure \ref{trace-compare}, both trials attain the same index value after polishing while the small `max.tries` of 25 is not sufficient for `search_better` to find its global maximum and we will need to adjust the `max.tries` argument for the search to sufficiently explore the parameter space.
--->
-
-<!-- ```{r polish, fig.asp=0.5,fig.cap="\\label{trace-compare}Breakdown of index value when using different max.tries in search better in conjunction with search polish.
-Both attain the same final index value after the polishing while using a `max.tries` of 25 is not sufficient to find the true maximum."} -->
-
-<!-- holes_2d_better %>% mutate(max_tries = 25) %>%  -->
-<!--   dplyr::bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500)) %>%  -->
-<!--   explore_trace_interp(col = max_tries) +  -->
-<!--   theme(legend.position = "none") -->
-<!-- ``` -->
-
-<!-- *Example: The neighbourhood parameter alpha* Add an example on comparing the neighbourhood parameter in search_better & search_posse. -->
-
-<!-- ```{r nrow, echo = TRUE} -->
-<!-- # nrow(holes_2d_better_max_tries) -->
-<!-- # nrow(holes_2d_pos) -->
-<!-- ``` -->
-
-<!-- ```{r alpha, fig.asp=0.5,eval = FALSE} -->
-<!-- p1 <- holes_2d_better_max_tries %>% -->
-<!--   filter(info != "interpolation") %>% -->
-<!--   mutate(id = row_number()) %>% -->
-<!--   explore_trace_parameter(var = alpha) + -->
-<!--   ylim(0.46, 0.505) + ggtitle("search_better") -->
-
-<!-- p3 <- holes_2d_better_max_tries %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96) -->
-
-<!-- p2 <- holes_2d_pos %>% -->
-<!--   filter(info != "interpolation") %>% -->
-<!--   mutate(id = row_number()) %>% -->
-<!--   explore_trace_parameter(var = alpha) + ylim(0.46, 0.505) + ggtitle("search_posse") -->
-
-<!-- p4 <- holes_2d_pos %>% explore_trace_interp(col = alpha) + ylim(0.8, 0.96) -->
-
-<!-- (p1 | p2) / (p3 | p4) + plot_layout(guides = "collect") -->
-<!-- ``` -->
-
-<!-- *Example: initial value for polishing alpha* `search_polish` is a brute-force algorithm that evaluate 1000 points in the neighbourhood at each loop.
-Setting an appropriate initial value for polish_alpha would avoid wasting search on large vector space that are not likely to produce higher index value.
-The default initial value for polishing step is 0.5 and we are interested in whether this is an appropriate initial value to use after `search_geodesic`.
-The problem is a 1D projection of the small dataset using `search_geodesic` and followed by `search_polish`.
-The top-left panel of Figure \ref{polish-alpha} displays all the projection bases on the first two principal components, coloured by the `polish_alpha`.
-We can observe that rather than concentrating on the ending basis from `search_geodesic` as what polishing step is designed, `search_polish` searches a much larger vector space, which is unnecessary.
-Thus a customised smaller initial value for `polish_alpha` would be ideal.
-One way to do this is to initialised `polish_alpha` as the projection distance between the last two target bases.
-The top-right panel of Figure \ref{polish-alpha} shows a more desirable concentrated searching space near the ending basis.
-Both specifications of initial value allow the searches to reach the same ending index values.
--->
-
-<!-- ```{r polish-alpha, fig.cap = "\\label{polish-alpha}PCA plot of two different polish alpha initialisations.
-A default polish alpha = 0.5 searches a larger space that is unncessary while a small customised initial value of polish alpha will search near the ending basis.
-Both intialisations reach the same ending index values.
-"} -->
-<!-- p1 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>%  -->
-<!--   explore_proj_pca(col = as.factor(alpha)) -->
-
-<!-- p2 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>%  -->
-<!--   explore_proj_pca(col = as.factor(alpha)) -->
-
-<!-- p3 <- bind_rows(holes_1d_geo, holes_1d_geo_polish_default_alpha) %>% -->
-<!--   explore_trace_interp(col = method) + ylim(0.75, 0.95) -->
-
-<!-- p4 <- bind_rows(holes_1d_geo, holes_1d_geo_polish) %>%  -->
-<!--   explore_trace_interp(col = method) +  ylim(0.75, 0.95) -->
-
-<!-- ((p1 | p2) /(p3 | p4)) +  plot_layout(heights = c(2, 1)) & theme(legend.position = "none")  -->
-
-<!-- ``` -->
-
-While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
-Geozoo is an option.
-It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
-
-<!-- # Animated diagnostic plots -->
-
-<!-- ```{r pca-animated, echo = TRUE, eval = FALSE} -->
-<!-- holes_1d_geo %>% explore_proj_pca(animate = TRUE, col = info) +  -->
-<!--   theme(legend.position = "bottom") -->
-<!-- ``` -->
 

### 2020-10-07 paper/paper.tex
refine 5.3 Close but not close enough

@@
 
-In principle, all the optimisation routines should result in the same output for the same problem while this may not be the case in real application.
-This motivates the creation of a polishing search that polishes the final basis found and achieves unity across different methods.
-
-\texttt{search\_polish} takes the final basis of a given search as a start and uses a brutal-force approach to sample a large number of basis (\texttt{n\_sample}) in the neighbourhood.
-Among those sampled basis, the one with the largest index value is chosen to be compared with the current basis.
-If its index value is larger than that of the current basis, it becomes the current basis in the next iteration.
-If no basis is found to have larger index value, the searching neighbourhood will shrink and the search continues.
-The polishing search ends when one of the four stopping criteria is satisfied:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\tightlist
-\item
-  the chosen basis can't be too close to the current basis
-\item
-  the percentage improvement of the index value can't be too small
-\item
-  the searching neighbourhood can't be too small
-\item
-  the number of iteration can't exceed \texttt{max.tries}
-\end{enumerate}
-
-The usage of search\_polish is as follows.
-After the first search, the final basis from the interpolation is extracted and supplied to the second search as the \texttt{start} argument.
-\texttt{search\_polish} is used as the search function.
-All the other arguments should remain the same.
+Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
+This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
+The following code demonstrates the usage of \texttt{search\_polish}:
 
@@
 
-Checking the final projection before and after applying the polish as in Figure \ref{geo-polish}, we can see that \ldots{}
+Similar to \texttt{search\_better} as a stochastic random search, \texttt{search\_polish} has a different scheme of reducing the search neighbourhood.
+In each search-interpolation iteration, \texttt{search\_better} has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
+On the contrary, \texttt{search\_polish} allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
+Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original \texttt{max.tries}.
+These include:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  the distance between the candidate basis and the current basis needs to be larger than 1e-3;
+\item
+  the percentage change of the index value need to be larger than 1e-5; and
+\item
+  the alpha parameter on itself need to be larger than 0.01
+\end{enumerate}
+
+Figure \ref{geo-polish} presents the final projections found before and after applying \texttt{search\_polish} using the demo code.
+Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables, namely \texttt{V1}, \texttt{V8}, \texttt{V9} and \texttt{V10}.
+When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
 
@@
 \includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
-\caption{\label{fig:geo-polish}\label{geo-polish}On the left is the final projection from search geodesic while on the right is the polished result.
-Polishing helps to sharpen the edges of the clusters}
+\caption{\label{fig:geo-polish}\label{geo-polish}Projected data before and after polishing.
+Weights on axis other than V2 and V7 have been further reduced and the edges of each cluster have become more obvious.}
 \end{figure}
-
-While explore the reduced space is an initial attempt to understand the searching space, there are existing technology for rotating a higher dimensional space for visualisation.
-Geozoo is an option.
-It generates random points on the high dimensional space and we can overlay it with the points on the optimisation path to visualise the spread of it on the high-D sphere.
 

### 2020-10-07 paper/paper.Rmd
changes here and there and everywhere, mainly in section 5l and conclusion

@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array}
+  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs}
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
@@
 
-The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
-`kol_cdf` index, an 1D projection index function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as $F(u)$ with subscript indicating the variable,  the Kolmogorov statistics defined by
+Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
+`norm_kol`, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
+Denotes the ECDF function as $F_{.}(u)$ with the subscript indicating the projection or the random normal variable,  the `norm_kol` index is defined by
 
@@
 
-can be seen as a function of the projection matrix $\mathbf{A}_{p \times 1}$ and hence a valid index function.
-
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating points when using different optimisation algorithms: `search_geodesic` and `search_better`. One can observe that
-
-- The index value of `kol_cdf` index is much smaller than that of holes index
-- The link of index values from interpolation bases  are no longer smooth
-- Both algorithms reach a similar final index value after polishing
-
-Polishing step has done much more work to find the final index value in `search_geodesic` than `search_better` and this indicates `kol_cdf` function favours of a random search method than ascent method.
-
-```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Comparison of two different searching methods: search_geodesic and search_better on 1D projection problem for a noisier index: kol_cdf.
-The geodesic search rely heavily on the polishing step to find the final index value while search better works well."}
+Figure \ref{kol-cdf} compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
+This time, the interpolated path is no longer smooth when using either algorithm.
+It is also obvious that `search_geodesic` fails to optimise this rough index since there is barely improvment of the index value.
+On the other hand, `search_better` is doing a relatively good job on the optimisation for the 5-variable dataset `boa5`.
+The theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)` and `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 4)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
+A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 4)`].
+At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+
+```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol.
+Search_geodesic has varely made progression while search_better has successfully found the informative "}
 # set.seed(123456)
@@
 
+index <- tourr::norm_kol()
+theo_best_index_val <- index(as.matrix(boa5) %*% matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1))
+
 bind_rows(kol_1d_geo, kol_1d_better) %>% 
@@
   facet_wrap(vars(method), ncol = 2) + 
+  geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) + 
   scale_color_botanical() +
@@
 
-Now we enlarge the dataset to include two informative variables: `x2` and `x3` and remain 1D projection.
-In this case, two local maxima appear with projection matrix being $[0, 1, 0, 0, 0, 0]$ and $[0, 0, 1 ,0, 0, 0]$.
-
-Using different seeds in `search_better` allows us to find both local maxima d as in Figure \ref{1d-2var-different-seeds}.
-Comparing the maximum of both, we can see that the global maximum happens when `x2` is found.
-It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds?
-`search_better_random` manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
-In Figure \ref{kol-pca},
-sing search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed.
-We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.
-As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better_random produces a more exhaustive search.
-The large overlapping of better1 and better_random is explained by the fact that both algorithms find x2 in the end.
-The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
-Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum.
-
-```{r kol-pca, fig.cap = "\\label{kol-pca}THis is the caption"}
+The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection.
+The dataset used is `boa6` with `x2` and `x7` being informative.
+The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found.
+
+A simulation with 20 different seeds shows that `search_better` can find either one of the informative variable but there's no guarantee on which seed finds which variable.
+`search_better_random` manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed.
+To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature $t_0$ so that the probability of accepting an inferior basis is not too high.
+Figure \ref{kol-pca-sim}, ....
+Figure \ref{kol-pca} shows the comparison, in terms of the trace plot and PCA plot, between `search_better_random` and two `search_better` algorithms where one finds `x2` and another finds `x7`.
+The trace plot shows that the global maximum has been found by `search_better_random` and one `search_better`, but not the other.
+The PCA plot further illustrate how the two algorithms finds the global maximum: `search_better_random` travels along the top right corner on the projected 2D circle while the path of `search_better` is much shorter and only goes pass in between the two theoretical bases.
+
+(there's more things going on about both algorithms experimenting with different seeds.)
+
+```{r}
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_sim <- list()
+# kol_1d_2var_better_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   set.seed(seed[i])
+#   kol_1d_2var_better_sim[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE)
+# 
+# 
+#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
+# 
+#   set.seed(1842)
+#   kol_1d_2var_better_sim_polish[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+# }
+
+# save(kol_1d_2var_better_sim, file = here::here("data", "kol_1d_2var_better_sim.rda"))
+# save(kol_1d_2var_better_sim_polish, file = here::here("data", "kol_1d_2var_better_sim_polish.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+better_list <- purrr::map2(kol_1d_2var_better_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+```
+
+```{r seed-table, eval = FALSE}
+best <- get_best(polish_tidy, group = seed)
+
+all_summary <- best$basis %>% 
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
+  as_tibble() %>% 
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
+  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
+  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
+  group_by(seed, IndexVal) %>% 
+  filter(abs(value) == max(abs(value))) %>% 
+  arrange(IndexVal) %>% ungroup()
+
+all_summary %>% knitr::kable(digits = 3, caption = "this is the caption", format='latex', booktabs=TRUE, escape=FALSE)
+
+```
+
+```{r eval = FALSE}
+pca <- all_tidy %>% 
+  mutate(seed = as.factor(seed)) %>% 
+  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
+  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>% 
+   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>%
+
+  compute_pca()
+
+random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
+theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
+
+pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
+  group_by(seed) %>% arrange(IndexVal) %>% 
+  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
+  geom_point(data = random, col = "grey") +
+  geom_point(aes(col = VarFound), alpha = 0.5) + 
+  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
+  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
+  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
+  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
+  #theme_void() + 
+  theme(aspect.ratio = 1) + 
+  scale_color_botanical(palette = "fern")  
+```
+
+```{r eval = FALSE}
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_sim <- list()
+# kol_1d_2var_better_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   set.seed(seed[i])
+#   kol_1d_2var_better_sim[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE)
+# 
+# 
+#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
+# 
+#   set.seed(1842)
+#   kol_1d_2var_better_sim_polish[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+# }
+
+# save(kol_1d_2var_better_sim, file = here::here("data", "kol_1d_2var_better_sim.rda"))
+# save(kol_1d_2var_better_sim_polish, file = here::here("data", "kol_1d_2var_better_sim_polish.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>% 
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
+  as_tibble() %>% 
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
+  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
+  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
+  group_by(seed, IndexVal) %>% 
+  filter(abs(value) == max(abs(value))) %>% 
+  arrange(IndexVal) %>% ungroup()
+
+all_tidy %>% mutate(seed = as.factor(seed)) %>% 
+  left_join(all_summary, by = "seed") %>% 
+  explore_trace_interp(group = seed, col = VarFound) + 
+  facet_wrap(vars(fct_reorder(seed, IndexVal)))  + 
+  #geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
+  #geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
+  scale_color_botanical(palette = "banksia")  + 
+  theme( aspect.ratio = 1)
+
+pca <- all_tidy %>% 
+  mutate(seed = as.factor(seed)) %>% 
+  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
+  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>% 
+   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>%
+
+  compute_pca()
+
+random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
+theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
+
+pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
+  group_by(seed) %>% arrange(IndexVal) %>% 
+  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
+  geom_point(data = random, col = "grey") +
+  geom_point(aes(col = VarFound), alpha = 0.1) + 
+  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
+  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
+  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
+  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
+  #theme_void() + 
+  theme(aspect.ratio = 1) + 
+  scale_color_botanical(palette = "fern") 
+```
+
+```{r kol-pca, fig.height = 5, fig.cap = "\\label{kol-pca}Comparison between search_better_random and two different search_better.
+In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
+serach_better_random and one of the serach_better have found the global maximum but the other fails.
+The lower panel shows the PCA plot for each of the three algorithms.
+The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
+The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
+This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance."}
 
@@
 # save(kol_1d_2var_better_random, file = here::here("data", "kol_1d_2var_better_random.rda"))
-
-p1 <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
-          kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
-          kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+# 
+# p1 <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
+#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
+#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+#   explore_trace_interp(group = algo, col = algo) + 
+#   facet_wrap(vars(algo), ncol = 3)  + 
+#   scale_color_botanical(palette = "banksia")  + 
+#   theme(legend.position = "none", aspect.ratio = 1)
+# 
+# pca <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
+#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
+#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+#   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+#                                         index = tourr::norm_kol(),
+#                                         raw_data = boa6) %>% 
+#    bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+#                                         index = tourr::norm_kol(),
+#                                         raw_data = boa6) %>% 
+#   compute_pca()
+
+dt <- bind_rows(all_tidy %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
+          all_tidy %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
+          kol_1d_2var_better_random %>% mutate(algo = "better random"))
+
+index <- tourr::norm_kol() 
+theo1 <- index(as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6))
+theo2 <- index(as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6))
+p1 <- dt %>% 
   explore_trace_interp(group = algo, col = algo) + 
   facet_wrap(vars(algo), ncol = 3)  + 
+  geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
+  geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
   scale_color_botanical(palette = "banksia")  + 
@@
 
-pca <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
-          kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
-          kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
+pca <- dt %>% 
   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
@@
   compute_pca()
-
 random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-algo)
@@
   geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
-  geom_point(data = pca$aug %>% get_start(),  size = 5) +
+  geom_point(data = pca$aug %>% get_start(),  size = 3) +
   facet_wrap(vars(algo)) +
@@
 
-something about the diagnostic plots
-
-summarise what we have done and what might be done in the future.
+<!-- something about the diagnostic plots  -->
+Four diagnostic plots have been introduced in the paper to investigate the progression and the projection space of an optimser, ranging from simple summary plot to advanced high dimensional animation.
+The implementation of these visualisations are designed to be easy-to-use with each plot can be produced with a simple supply of the data object and few additional arguments.
+More advanced users may decide to modify on top of the basic diagnostic pltos or even build their own.
+
+<!-- summarise what we have done and what might be done in the future. -->
+Most of the work in this project has been translated into code in two pacakges: the collection of the data object is implemented in the existing `tourr` package; manipulation and visualisation of the data object are implemented in the new `ferrn` package.
+Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions i.e.
+scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.
 

### 2020-10-07 paper/paper.tex
changes here and there and everywhere, mainly in section 5l and conclusion

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
 
@@
 
-The interpolation path of holes index, as seen in Figure \ref{interruption}, is smooth, while this may not be the case for more complicated index functions.
-\texttt{kol\_cdf} index, an 1D projection index function based on Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as \(F(u)\) with subscript indicating the variable, the Kolmogorov statistics defined by
+Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
+\texttt{norm\_kol}, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
+Denotes the ECDF function as \(F_{.}(u)\) with the subscript indicating the projection or the random normal variable, the \texttt{norm\_kol} index is defined by
 
@@
 
-can be seen as a function of the projection matrix \(\mathbf{A}_{p \times 1}\) and hence a valid index function.
-
-Figure \ref{kol-cdf} compares the tracing plot of the interpolating points when using different optimisation algorithms: \texttt{search\_geodesic} and \texttt{search\_better}. One can observe that
-
-\begin{itemize}
-\tightlist
-\item
-  The index value of \texttt{kol\_cdf} index is much smaller than that of holes index
-\item
-  The link of index values from interpolation bases are no longer smooth
-\item
-  Both algorithms reach a similar final index value after polishing
-\end{itemize}
-
-Polishing step has done much more work to find the final index value in \texttt{search\_geodesic} than \texttt{search\_better} and this indicates \texttt{kol\_cdf} function favours of a random search method than ascent method.
+Figure \ref{kol-cdf} compares the tracing plot of two algorithms: \texttt{search\_geodesic} and \texttt{search\_better}.
+This time, the interpolated path is no longer smooth when using either algorithm.
+It is also obvious that \texttt{search\_geodesic} fails to optimise this rough index since there is barely improvment of the index value.
+On the other hand, \texttt{search\_better} is doing a relatively good job on the optimisation for the 5-variable dataset \texttt{boa5}.
+The theoretical best basis {[}0, 1, 0, 0, 0{]} produces an index value of 0.182 and \texttt{search\_better} finds the final basis {[}-0.02939, -0.99715, 0.04509, -0.03248, 0.04176{]} with an index value of 0.174.
+A further polish step will give a marginal improvement of index value to 0.173 with a basis of {[}-0.077342, -0.996695, -0.014675, -0.004626, -0.019520{]}.
+At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
@@
 \includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{fig:kol-cdf}\label{kol-cdf}Comparison of two different searching methods: search\_geodesic and search\_better on 1D projection problem for a noisier index: kol\_cdf.
-The geodesic search rely heavily on the polishing step to find the final index value while search better works well.}
+\caption{\label{fig:kol-cdf}\label{kol-cdf}Trace plot of search\_geodesic and search\_better on 1D projection problem for a noisy index: norm\_kol.
+Search\_geodesic has varely made progression while search\_better has successfully found the informative}
 \end{figure}
 
-Now we enlarge the dataset to include two informative variables: \texttt{x2} and \texttt{x3} and remain 1D projection.
-In this case, two local maxima appear with projection matrix being \([0, 1, 0, 0, 0, 0]\) and \([0, 0, 1 ,0, 0, 0]\).
+The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection.
+The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative.
+The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found.
 
-Using different seeds in \texttt{search\_better} allows us to find both local maxima d as in Figure \ref{1d-2var-different-seeds}.
-Comparing the maximum of both, we can see that the global maximum happens when \texttt{x2} is found.
-It is natural to ask then if there is an algorithm that can find the global maximum without trying on different seeds?
-\texttt{search\_better\_random} manages to do it via a Metropolis-hasting random search as shown in Figure \ref{1d-2var-better-random}, although at a higher cost of number of points to evaluate.
-In Figure \ref{kol-pca},
-sing search better random for the problem above will result in finding the global maximum but much larger number of iteration is needed.
-We can also plot the searching points of all three algorithms in the searching space and explore their relative position against each other using principal components.
-As shown in Figure \ref{1d-2var-explore-proj-pca}, the bases from better1 and better2 only search a proportion of the searching space while better\_random produces a more exhaustive search.
-The large overlapping of better1 and better\_random is explained by the fact that both algorithms find x2 in the end.
-The trace plot search better in a 1D projection problem with two informative variables using different seeds (without polishing).
-Since there are two informative variables, setting different value for seed will lead search better to find either of the local maximum.
+A simulation with 20 different seeds shows that \texttt{search\_better} can find either one of the informative variable but there's no guarantee on which seed finds which variable.
+\texttt{search\_better\_random} manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed.
+To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature \(t_0\) so that the probability of accepting an inferior basis is not too high.
+Figure \ref{kol-pca-sim}, \ldots{}.
+Figure \ref{kol-pca} shows the comparison, in terms of the trace plot and PCA plot, between \texttt{search\_better\_random} and two \texttt{search\_better} algorithms where one finds \texttt{x2} and another finds \texttt{x7}.
+The trace plot shows that the global maximum has been found by \texttt{search\_better\_random} and one \texttt{search\_better}, but not the other.
+The PCA plot further illustrate how the two algorithms finds the global maximum: \texttt{search\_better\_random} travels along the top right corner on the projected 2D circle while the path of \texttt{search\_better} is much shorter and only goes pass in between the two theoretical bases.
+
+(there's more things going on about both algorithms experimenting with different seeds.)
 
@@
 \includegraphics{paper_files/figure-latex/kol-pca-1.pdf}
-\caption{\label{fig:kol-pca}\label{kol-pca}THis is the caption}
+\caption{\label{fig:kol-pca}\label{kol-pca}Comparison between search\_better\_random and two different search\_better.
+In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
+serach\_better\_random and one of the serach\_better have found the global maximum but the other fails.
+The lower panel shows the PCA plot for each of the three algorithms.
+The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
+The paths of two search\_better are relatively short and simple while the search\_better\_random has a more sophisticated path.
+This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance.}
 \end{figure}
@@
 
-something about the diagnostic plots
+Four diagnostic plots have been introduced in the paper to investigate the progression and the projection space of an optimser, ranging from simple summary plot to advanced high dimensional animation.
+The implementation of these visualisations are designed to be easy-to-use with each plot can be produced with a simple supply of the data object and few additional arguments.
+More advanced users may decide to modify on top of the basic diagnostic pltos or even build their own.
 
-summarise what we have done and what might be done in the future.
+Most of the work in this project has been translated into code in two pacakges: the collection of the data object is implemented in the existing \texttt{tourr} package; manipulation and visualisation of the data object are implemented in the new \texttt{ferrn} package.
+Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions i.e.~scagnostics, association, and information index \citep{laa2020using} and understand how the optimisers behave for index functions with different structures.
 

### 2020-10-08 paper/paper.Rmd
include the plots for comparing diff seeds

@@
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
-                   index = tourr::holes(), raw_data = data) %>% 
+                   index = tourr::holes(), raw_data = boa5) %>% 
   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
-                   index = tourr::holes(), raw_data  = data)%>%
+                   index = tourr::holes(), raw_data  = boa5)%>%
   explore_space_pca(col = method)  + 
@@
 
-```{r}
+```{r fig.height = 8}
 set.seed(123)
@@
 better_tidy <- do.call(dplyr::bind_rows, better_list)
-all_tidy <- bind_rows(better_tidy, polish_tidy)
-```
-
-```{r seed-table, eval = FALSE}
-best <- get_best(polish_tidy, group = seed)
+all_tidy_better <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(all_tidy_better, group = seed) %>% arrange(index_val)
 
@@
   rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
-  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
+  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
   pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
@@
 
-all_summary %>% knitr::kable(digits = 3, caption = "this is the caption", format='latex', booktabs=TRUE, escape=FALSE)
-
-```
-
-```{r eval = FALSE}
+pca <- all_tidy_better%>% 
+  mutate(seed = as.factor(seed)) %>% 
+  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
+  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>% 
+   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+                                        index = tourr::norm_kol(),
+                                        raw_data = boa6) %>%
+
+  compute_pca()
+
+random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
+theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
+
+pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
+  group_by(seed) %>% arrange(IndexVal) %>% 
+  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
+  geom_point(data = random, col = "grey", size = 0.3) +
+  geom_point(aes(col = VarFound), alpha = 0.1) + 
+  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
+  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
+  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
+  geom_point(data = pca$aug %>% get_start() %>% filter(method == "search_better_random"), size = 3) +
+  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
+  theme_void() + 
+  theme(aspect.ratio = 1, legend.position = "bottom") + 
+  scale_color_botanical(palette = "fern")  
+```
+
+```{r fig.height= 8}
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_random_sim <- list()
+# kol_1d_2var_better_random_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   cat("i = ", i, "seed = ", seed[i], "\n")
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better, max.tries = 500, t0 = 1e-6, alpha = 0.7),
+#                  rescale = FALSE, verbose = TRUE)
+# 
+# 
+#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
+# 
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim_polish[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 200),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+#   
+# }
+
+#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
+#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>% 
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
+  as_tibble() %>% 
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
+  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
+  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
+  group_by(seed, IndexVal) %>% 
+  filter(abs(value) == max(abs(value))) %>% 
+  arrange(IndexVal) %>% ungroup()
+
 pca <- all_tidy %>% 
@@
   ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
-  geom_point(data = random, col = "grey") +
-  geom_point(aes(col = VarFound), alpha = 0.5) + 
+  geom_point(data = random, col = "grey", size = 0.3) +
+  geom_point(aes(col = VarFound), alpha = 0.05) + 
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
   geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
+  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
   facet_wrap(vars(fct_reorder(seed, IndexVal))) +
-  #theme_void() + 
-  theme(aspect.ratio = 1) + 
-  scale_color_botanical(palette = "fern")  
-```
-
-```{r eval = FALSE}
-set.seed(123)
-seed <- sample.int(10000, 20)
-# kol_1d_2var_better_sim <- list()
-# kol_1d_2var_better_sim_polish <- list()
-# for (i in 1: length(seed)){
-#   set.seed(seed[i])
-#   kol_1d_2var_better_sim[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_better, max.tries = 100),
-#                  rescale = FALSE, verbose = TRUE)
-# 
-# 
-#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
-# 
-#   set.seed(1842)
-#   kol_1d_2var_better_sim_polish[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_polish, max.tries = 100),
-#                  rescale = FALSE, verbose = TRUE,
-#                  start = last_basis)
-# }
-
-# save(kol_1d_2var_better_sim, file = here::here("data", "kol_1d_2var_better_sim.rda"))
-# save(kol_1d_2var_better_sim_polish, file = here::here("data", "kol_1d_2var_better_sim_polish.rda"))
-
-polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-polish_tidy <- do.call(dplyr::bind_rows, polish_list)
-better_tidy <- do.call(dplyr::bind_rows, better_list)
-all_tidy <- bind_rows(better_tidy, polish_tidy)
-
-best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
-best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
-best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
-
-all_summary <- best$basis %>% 
-  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
-  as_tibble() %>% 
-  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
-  mutate(IndexVal = best$index_val, seed = as.factor(seed)) %>% 
-  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
-  group_by(seed, IndexVal) %>% 
-  filter(abs(value) == max(abs(value))) %>% 
-  arrange(IndexVal) %>% ungroup()
-
-all_tidy %>% mutate(seed = as.factor(seed)) %>% 
-  left_join(all_summary, by = "seed") %>% 
-  explore_trace_interp(group = seed, col = VarFound) + 
-  facet_wrap(vars(fct_reorder(seed, IndexVal)))  + 
-  #geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
-  #geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
-  scale_color_botanical(palette = "banksia")  + 
-  theme( aspect.ratio = 1)
-
-pca <- all_tidy %>% 
-  mutate(seed = as.factor(seed)) %>% 
-  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
-  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>% 
-   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>%
-
-  compute_pca()
-
-random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
-theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
-
-pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
-  group_by(seed) %>% arrange(IndexVal) %>% 
-  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
-  geom_point(data = random, col = "grey") +
-  geom_point(aes(col = VarFound), alpha = 0.1) + 
-  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
-  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 7) +
-  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
-  #theme_void() + 
-  theme(aspect.ratio = 1) + 
+  theme_void() + 
+  theme(aspect.ratio = 1, legend.position = "bottom") + 
   scale_color_botanical(palette = "fern") 
@@
 
-dt <- bind_rows(all_tidy %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
-          all_tidy %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
-          kol_1d_2var_better_random %>% mutate(algo = "better random"))
+dt <- bind_rows(all_tidy_better %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
+          all_tidy_better %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
+          all_tidy %>% filter(seed == 4761) %>% mutate(algo = "better random"))
 
@@
   geom_point(data = random, col = "grey") +
-  geom_point(aes(col = algo), alpha = 0.1) + 
+  geom_point(aes(col = algo), alpha = 0.05) + 
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
@@
   geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
-  geom_point(data = pca$aug %>% get_start(),  size = 3) +
+  geom_point(data = pca$aug %>% get_start() %>% filter(method %in% c("search_better_random", "search_better")),  size = 3) +
   facet_wrap(vars(algo)) +

### 2020-10-09 paper/paper.Rmd
remove codes, comments, list of plot elements that are unncessary, kol_1d_2var_beter_sim2 data

@@
 
-[Are we using  projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
-Or we are focusing on the optimisation problem only within the projection pursuit context?
-Some of the problems listed below are also applicable to optimization problem in general too.
-ppp]
-
 # Projection pursuit guided tour {#tour}
@@
 
-```{r tour-path, out.width="60%", out.height="40%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
-knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"),)
+```{r tour-path, out.height="30%", out.width="60%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
+knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
 ```
@@
 
-[outside the paper: I find the notation of current/target basis is confusing because the target basis in round $j$ becomes the current basis in round $j+1$.
-Also, when we start to have polish, the target basis may not be the current basis in the next round...
-The place where current/target is most appropriate is probably when describing the interpolation where the first one is always the current basis and the last is always the target basis.
-I think it is better to leave this language in the code]
-
 ```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."}
@@
 This is achieved via a `cutoff` parameter.
-Below we define the searching point plot based on four different component layers
-
-- Layer 1: boxplot geom
-  - data: group by $j$ and filter the observations in the groups that have count greater than `cutoff = 15`.
-  - x: $j$ is mapped to the x-axis
-  - y: the statistical transformed index value: $Q_{I^{\prime}_t}(q)$ is mapped to the y-axis where $Q_X(q)$, $q = 0, 25, 50, 75, 100$ finds the qth-quantile of $X$ and $I^{\prime}_t$ denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
-
-- Layer 2: point geom
-  - data: group by $j$ and filter the observations in the group that have count less than `cutoff = 15`.
-  - x: $j$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-- Layer 3: line geom
-  - data: filter the points with the highest index value in group $j$
-  - x: $j$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-- Layer 4: label geom
-  - data: A count table of the number of observation in each iteration
-  - x: $j$ is mapped to the x-axis
-  - y: the y-axis for each label is $0.99 * \text{MIN}$, where $\text{MIN}$ is the smallest index value in the whole data object
-  - label: the number of observation is mapped to the label
-
-<!-- *Example: exploring searching points* We choose variables `x1`, `x2`, `x3`, `x8`, `x9` and `x10` to perform a 2D projection with tour.
-Parameter `search_f = tour::search_better` and `max.tries = 500` is used.
-The index value of the searching points are shown in Figure \ref{toy-search}.
-The label at the bottom indicates the number of observations in each iteration and facilitates the choice of `cutoff` argument (by default `cutoff = 15`).
-We learn that the `search_better` quickly finds better projection basis with higher index value at first and then takes longer to find a better one later.
--->
 
@@
 
-```{r toy-search, fig.height=5.5, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
+```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
 In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
-
-dt <- holes_2d_better_max_tries %>%
-  filter(info != "interpolation") %>% 
-  mutate(id = row_number())
-
-p1 <- dt %>% explore_trace_search() + 
-  ggtitle("satisfactory") + 
-  scale_color_botanical(palette = "daisy")
-
-p2 <- holes_2d_better %>% explore_trace_search() + 
-  ggtitle("not satisfactory") + 
-  scale_color_botanical(palette = "daisy")
-
-p1 / p2
-
+bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
+          holes_2d_better %>% mutate(max_tries = 25)) %>% 
+  explore_trace_search(group = max_tries) + 
+  scale_color_botanical(palette = "daisy") +
+  facet_wrap(vars(max_tries), ncol = 2)
 ```
@@
 
-Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. The plot definition is as follows:
-
-- Layer 1: point geom
-  - data: filter the observations with $S = \text{interpolation}$ and mutate $t$ to be the row number of the subsetted tibble
-  - x: $t$ is mapped to the x-axis
-  - y: $I$ is mapped to the y-axis
-
-- Layer 2: line geom
-  - using line geometry for the same data and aesthetics
-
+Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
 Figure \ref{toy-interp} presents the interpolation of three different tour paths.
@@
 
-```{r toy-interp, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
-p1 <- holes_2d_better_max_tries %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern")
-
-p2 <- interrupt_no %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern")
-
-p3 <- holes_2d_better_random %>% explore_trace_interp() + theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern")
-
-p1 / p2 / p3
-```
-
-\
+```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
+bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
+          interrupt_no %>% mutate(group = "no interruption"), 
+          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
+  mutate(group = fct_relevel(group, c("max_tries: 500", 
+                                      "no interruption", 
+                                      "better_random"))) %>% 
+  explore_trace_interp(group = group) + 
+  theme(legend.position = "none") + 
+  scale_color_botanical(palette = "fern") + 
+  facet_wrap(vars(group), ncol = 3)
+```
 
@@
 To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
-
-The visualisation can thus be defined as
 
@@
 
-- Layer 1: point geom
-  - data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
-  - x: the projected basis on the first principal component
-  - y: the projected basis on the second principal component
-  - colour: $V$ is mapped to the colour aesthetic
-
 Figure \ref{toy-pca} shows the first two principal components of two search paths.
@@
 
-```{r toy-pca, fig.height = 4, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
+```{r toy-pca, fig.height = 3, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
 # path1 <- holes_1d_geo$basis %>% 
@@
 
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
-
-<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->
-
-<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->
-
-<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->
-
-<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->
-
-<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->
-
-```{r fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progrssion from the start to finish in each algorithm.", fig.height=4, fig.align="center"}
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, ...
+
+```{r fig.height = 3, fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
+With animation, it is easier to track the progrssion from the start to finish in each algorithm."}
 # ani <- p + theme(legend.position = "none") +
@@
 
+# 
+# bind_rows(holes_1d_geo, holes_1d_better) %>% 
+#   explore_space_tour(color = method)
 #animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial
@@
 
-<!-- In the previous two sections, only the iterator and the index value are mapped onto the x and y aesthetics of the plot; while more aesthetics i.e.
-colour, could be added to compare other parameters in the global object.
--->
-
-<!-- Two examples are shown below to explore and compare different searching methods and neighbourhood parameter alpha.  -->
-
 Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
 This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
-The following code demonstrates the usage of `search_polish`:
-
-```{r sample-code, eval = FALSE, echo = TRUE}
-library(tourr)
-library(ferrn)
-
-set.seed(123456)
-holes_2d_geo <- 
-  animate_xy(boa6,
-             tour_path = guided_tour(holes(), d = 2, 
-                                     search_f = search_geodesic),
-             rescale = FALSE, verbose = TRUE)
-
-last_basis <- get_best(holes_2d_geo) %>% pull(basis) %>% .[[1]]
-
-set.seed(123456)
-holes_2d_geo_polish <- 
-  animate_xy(boa6, 
-             tour_path = guided_tour(holes(), d = 2, 
-                                     search_f = search_polish),
-             rescale = FALSE, verbose = TRUE, start = last_basis)
-```
 
@@
 
-```{r kol-cdf,fig.asp=0.5, fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol.
+```{r kol-cdf,fig.asp=0.5, fig.height = 3,  fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol.
 Search_geodesic has varely made progression while search_better has successfully found the informative "}
@@
 
-```{r fig.height= 8}
+```{r fig.height = 8}
 set.seed(123)
 seed <- sample.int(10000, 20)
-# kol_1d_2var_better_random_sim <- list()
-# kol_1d_2var_better_random_sim_polish <- list()
+# kol_1d_2var_better_random_sim2 <- list()
+# kol_1d_2var_better_random_sim_polish2 <- list()
 # for (i in 1: length(seed)){
@@
 #   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim[[i]] <-
+#   kol_1d_2var_better_random_sim2[[i]] <-
 #     animate_dist(boa6,
 #                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_better, max.tries = 500, t0 = 1e-6, alpha = 0.7),
+#                                          search_f = search_better_random, max.tries = 200),
 #                  rescale = FALSE, verbose = TRUE)
@@
 # 
-#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
+#   last_basis <- get_best(kol_1d_2var_better_random_sim2[[i]])$basis %>% .[[1]]
 # 
 #   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim_polish[[i]] <-
+#   kol_1d_2var_better_random_sim_polish2[[i]] <-
 #     animate_dist(boa6,
@@
 # }
-
-#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
-#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))
-
-polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+# save(kol_1d_2var_better_random_sim2, file = here::here("data", "kol_1d_2var_better_sim2.rda"))
+# save(kol_1d_2var_better_random_sim_polish2, file = here::here("data", "kol_1d_2var_better_sim_polish2.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish2,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+better_list <- purrr::map2(kol_1d_2var_better_random_sim2, seed,  ~.x %>% dplyr ::mutate(seed = .y)) 
 polish_tidy <- do.call(dplyr::bind_rows, polish_list)
@@
 
-best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
-best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
 best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
@@
 
-```{r kol-pca, fig.height = 5, fig.cap = "\\label{kol-pca}Comparison between search_better_random and two different search_better.
-In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
-serach_better_random and one of the serach_better have found the global maximum but the other fails.
-The lower panel shows the PCA plot for each of the three algorithms.
-The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
-The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
-This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance."}
-
-# set.seed(123456)
-# kol_1d_2var_better <- 
+```{r fig.height= 8}
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_random_sim <- list()
+# kol_1d_2var_better_random_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   cat("i = ", i, "seed = ", seed[i], "\n")
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim[[i]] <-
 #     animate_dist(boa6,
 #                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                        search_f = search_better, max.tries = 100),
+#                                          search_f = search_better_random, max.tries = 500, t0 = 1e-6, alpha = 0.7),
 #                  rescale = FALSE, verbose = TRUE)
@@
 # 
-# set.seed(1234567)
-# kol_1d_2var_better_2 <- 
+#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
+# 
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim_polish[[i]] <-
 #     animate_dist(boa6,
 #                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                        search_f = search_better, max.tries = 100),
-#                rescale = FALSE, verbose = TRUE)
-# 
-# set.seed(12345678)
-# kol_1d_2var_better_2 <- 
-#   animate_dist(boa6, 
-#                tour_path = guided_tour(norm_kol(), d = 1,
-#                                        search_f = search_better, max.tries = 100),
-#                rescale = FALSE, verbose = TRUE)
-# 
-# 
-# 
-# 
-# set.seed(123456)
-# kol_1d_2var_better_random <- 
-#   animate_dist(boa6, 
-#                tour_path = guided_tour(norm_kol(), d = 1,
-#                                        search_f = search_better_random,
-#                                        max.tries = 2000, t0 = 1e-5),
-#                rescale = FALSE, verbose = TRUE)
-# 
-# save(kol_1d_2var_better, file = here::here("data", "kol_1d_2var_better.rda"))
-# save(kol_1d_2var_better_2, file = here::here("data", "kol_1d_2var_better_2.rda"))
-# save(kol_1d_2var_better_random, file = here::here("data", "kol_1d_2var_better_random.rda"))
-# 
-# p1 <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
-#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
-#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
-#   explore_trace_interp(group = algo, col = algo) + 
-#   facet_wrap(vars(algo), ncol = 3)  + 
-#   scale_color_botanical(palette = "banksia")  + 
-#   theme(legend.position = "none", aspect.ratio = 1)
-# 
-# pca <- bind_rows(kol_1d_2var_better %>% mutate(algo = "better 1"),
-#           kol_1d_2var_better_2 %>% mutate(algo = "better 2"),
-#           kol_1d_2var_better_random %>% mutate(algo = "better random")) %>% 
-#   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-#                                         index = tourr::norm_kol(),
-#                                         raw_data = boa6) %>% 
-#    bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-#                                         index = tourr::norm_kol(),
-#                                         raw_data = boa6) %>% 
-#   compute_pca()
-
-dt <- bind_rows(all_tidy_better %>% filter(seed == 1842) %>% mutate(algo = "better V2"),
-          all_tidy_better %>% filter(seed == 9982) %>% mutate(algo = "better V7"),
-          all_tidy %>% filter(seed == 4761) %>% mutate(algo = "better random"))
-
-index <- tourr::norm_kol() 
-theo1 <- index(as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6))
-theo2 <- index(as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6))
-p1 <- dt %>% 
-  explore_trace_interp(group = algo, col = algo) + 
-  facet_wrap(vars(algo), ncol = 3)  + 
-  geom_hline(yintercept = theo1, alpha = 0.5, linetype = 2) + 
-  geom_hline(yintercept = theo2, alpha = 0.5, linetype = 2) + 
-  scale_color_botanical(palette = "banksia")  + 
-  theme(legend.position = "none", aspect.ratio = 1)
-
-pca <- dt %>% 
+#                                          search_f = search_polish, max.tries = 200),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+#   
+# }
+
+#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
+#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>% 
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
+  as_tibble() %>% 
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
+  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
+  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
+  group_by(seed, IndexVal) %>% 
+  filter(abs(value) == max(abs(value))) %>% 
+  arrange(IndexVal) %>% ungroup()
+
+pca <- all_tidy %>% 
+  mutate(seed = as.factor(seed)) %>% 
+  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
@@
                                         raw_data = boa6) %>%
+
   compute_pca()
-random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-algo)
-theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-algo)
-
-p2 <- pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
-  ggplot(aes(x = PC1, y = PC2, col = algo)) + 
-  geom_point(data = random, col = "grey") +
-  geom_point(aes(col = algo), alpha = 0.05) + 
+
+random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
+theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
+
+pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
+  group_by(seed) %>% arrange(IndexVal) %>% 
+  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
+  geom_point(data = random, col = "grey", size = 0.3) +
+  geom_point(aes(col = VarFound), alpha = 0.05) + 
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
   geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = algo), size = 5) +
-  geom_point(data = pca$aug %>% get_start() %>% filter(method %in% c("search_better_random", "search_better")),  size = 3) +
-  facet_wrap(vars(algo)) +
+  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
+  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
   theme_void() + 
-  theme(legend.position = "none", aspect.ratio = 1) + 
-  scale_color_botanical(palette = "banksia")
-
-wrap_plots(p1, p2, ncol = 1)
-
-```
+  theme(aspect.ratio = 1, legend.position = "bottom") + 
+  scale_color_botanical(palette = "fern") 
+```
+
+Comparison between search_better_random and two different search_better.
+In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
+serach_better_random and one of the serach_better have found the global maximum but the other fails.
+The lower panel shows the PCA plot for each of the three algorithms.
+The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
+The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
+This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
 
@@
 
-(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
 The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package.
@@
   dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
-  head(10) 
+  head(5) 
 ```
@@
 
-  <!-- This binding is used in the space plots to compare the position of the theoretical best basis to the ones being searched by the algorithm. The following chunk presents its usage -->
-<!-- These are also useful in the space plots since it preserves the spherical structure of the spaces of the projection bases, in the computation of PCA for `explore_space_pca()` and animation of `explore_space_tour()`.
--->
-
 - Color and relevel
@@
 
-```{r eval = FALSE}
-# Below are codes that generate the plots in Section 4 
-# using functions from ferrn package
-bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
-          holes_2d_better %>% mutate(max_tries = 25)) %>% 
-  explore_trace_search(group = max_tries)
-```
-
-```{r eval = FALSE}
-bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
-          interrupt_no %>% mutate(group = "no interruption"), 
-          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
-  mutate(group = fct_relevel(group, c("max_tries: 500", 
-                                      "no interruption", 
-                                      "better_random"))) %>% 
-  explore_trace_interp(group = group) + 
-  theme(legend.position = "none")
-```
-
-```{r eval = FALSE}
-bind_rows(holes_1d_geo, holes_1d_better) %>% 
-  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
-                   index = tourr::holes(), raw_data = data) %>% 
-  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
-                   index = tourr::holes(), raw_data  = data)%>%
-  explore_space_pca(col = method)  + 
-  scale_color_botanical(palette = "cherry")
-```
-
-```{r eval = FALSE}
-bind_rows(holes_1d_geo, holes_1d_better) %>% 
-  explore_space_tour(color = method)
-```
-
 # Conclusion

### 2020-10-09 paper/paper.tex
include the plots for comparing diff seeds
remove codes, comments, list of plot elements that are unncessary, kol_1d_2var_beter_sim2 data

@@
 \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
-
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 
@@
 
-{[}Are we using projection pursuit/guided tour to better understand the convergence of optimization algorithms visually in combination with the algorithms discussed below?
-Or we are focusing on the optimisation problem only within the projection pursuit context?
-Some of the problems listed below are also applicable to optimization problem in general too.
-ppp{]}
-
 \hypertarget{tour}{%
@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.4\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001}
 
@@
 
-{[}outside the paper: I find the notation of current/target basis is confusing because the target basis in round \(j\) becomes the current basis in round \(j+1\).
-Also, when we start to have polish, the target basis may not be the current basis in the next round\ldots{} The place where current/target is most appropriate is probably when describing the interpolation where the first one is always the current basis and the last is always the target basis.
-I think it is better to leave this language in the code{]}
-
 \begin{figure}
@@
 This is achieved via a \texttt{cutoff} parameter.
-Below we define the searching point plot based on four different component layers
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: boxplot geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: group by \(j\) and filter the observations in the groups that have count greater than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: the statistical transformed index value: \(Q_{I^{\prime}_t}(q)\) is mapped to the y-axis where \(Q_X(q)\), \(q = 0, 25, 50, 75, 100\) finds the qth-quantile of \(X\) and \(I^{\prime}_t\) denotes the index value of all the searching bases defined in Matrix \ref{eq:data-structure}.
-  \end{itemize}
-\item
-  Layer 2: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: group by \(j\) and filter the observations in the group that have count less than \texttt{cutoff\ =\ 15}.
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\item
-  Layer 3: line geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: filter the points with the highest index value in group \(j\)
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\item
-  Layer 4: label geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: A count table of the number of observation in each iteration
-  \item
-    x: \(j\) is mapped to the x-axis
-  \item
-    y: the y-axis for each label is \(0.99 * \text{MIN}\), where \(\text{MIN}\) is the smallest index value in the whole data object
-  \item
-    label: the number of observation is mapped to the label
-  \end{itemize}
-\end{itemize}
 
@@
 
-Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation. The plot definition is as follows:
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: filter the observations with \(S = \text{interpolation}\) and mutate \(t\) to be the row number of the subsetted tibble
-  \item
-    x: \(t\) is mapped to the x-axis
-  \item
-    y: \(I\) is mapped to the y-axis
-  \end{itemize}
-\item
-  Layer 2: line geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    using line geometry for the same data and aesthetics
-  \end{itemize}
-\end{itemize}
-
+Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
 Figure \ref{toy-interp} presents the interpolation of three different tour paths.
@@
 
-~
-
 \hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
@@
 To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
-
-The visualisation can thus be defined as
-
-\begin{itemize}
-\tightlist
-\item
-  Layer 1: point geom
-
-  \begin{itemize}
-  \tightlist
-  \item
-    data: subset the basis of interest and arrange into a matrix format; perform PCA on the basis matrix and compute the projected basis on the first two principal components; bind the variables from the original global object and form a tibble
-  \item
-    x: the projected basis on the first principal component
-  \item
-    y: the projected basis on the second principal component
-  \item
-    colour: \(V\) is mapped to the colour aesthetic
-  \end{itemize}
-\end{itemize}
 
@@
 
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, \ldots{}
 
 \begin{figure}
-
-{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-1-1}
-
-}
-
-\caption{\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progrssion from the start to finish in each algorithm.}\label{fig:unnamed-chunk-1}
+\centering
+\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:unnamed-chunk-1}\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
+With animation, it is easier to track the progrssion from the start to finish in each algorithm.}
 \end{figure}
@@
 This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
-The following code demonstrates the usage of \texttt{search\_polish}:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(tourr)}
-\KeywordTok{library}\NormalTok{(ferrn)}
-
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
-\NormalTok{holes_2d_geo <-}\StringTok{ }
-\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6,}
-             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
-                                     \DataTypeTok{search_f =}\NormalTok{ search_geodesic),}
-             \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{)}
-
-\NormalTok{last_basis <-}\StringTok{ }\KeywordTok{get_best}\NormalTok{(holes_2d_geo) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{pull}\NormalTok{(basis) }\OperatorTok{%>%}\StringTok{ }\NormalTok{.[[}\DecValTok{1}\NormalTok{]]}
-
-\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123456}\NormalTok{)}
-\NormalTok{holes_2d_geo_polish <-}\StringTok{ }
-\StringTok{  }\KeywordTok{animate_xy}\NormalTok{(boa6, }
-             \DataTypeTok{tour_path =} \KeywordTok{guided_tour}\NormalTok{(}\KeywordTok{holes}\NormalTok{(), }\DataTypeTok{d =} \DecValTok{2}\NormalTok{, }
-                                     \DataTypeTok{search_f =}\NormalTok{ search_polish),}
-             \DataTypeTok{rescale =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{start =}\NormalTok{ last_basis)}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/kol-pca-1.pdf}
-\caption{\label{fig:kol-pca}\label{kol-pca}Comparison between search\_better\_random and two different search\_better.
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+Comparison between search\_better\_random and two different search\_better.
 In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
@@
 The paths of two search\_better are relatively short and simple while the search\_better\_random has a more sophisticated path.
-This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance.}
-\end{figure}
+This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
 
@@
 
-(further notice on the implementation in the tour package will be discussed in the tour developer meeting.)
 The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing \texttt{tourr} package.
@@
 \begin{verbatim}
-## # A tibble: 10 x 8
+## # A tibble: 5 x 8
 ##      id basis             index_val info          tries  loop method       alpha
@@
 ## 1     1 <dbl[,1] [5 x 1]>     0.749 new_basis         1     1 search_bett~   0.5
-##  2     2 <dbl[,1] [5 x 1]>     0.730 random_sear~     2     1 search_bett~   0.5
-##  3     3 <dbl[,1] [5 x 1]>     0.743 random_sear~     2     2 search_bett~   0.5
-##  4     4 <dbl[,1] [5 x 1]>     0.736 random_sear~     2     3 search_bett~   0.5
-##  5     5 <dbl[,1] [5 x 1]>     0.747 random_sear~     2     4 search_bett~   0.5
-##  6     6 <dbl[,1] [5 x 1]>     0.725 random_sear~     2     5 search_bett~   0.5
-##  7     7 <dbl[,1] [5 x 1]>     0.752 new_basis        2     6 search_bett~   0.5
-##  8     8 <dbl[,1] [5 x 1]>     0.749 interpolati~     2     1 search_bett~  NA  
-##  9     9 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     2 search_bett~  NA  
-## 10    10 <dbl[,1] [5 x 1]>     0.750 interpolati~     2     3 search_bett~  NA
+## 2     2 <dbl[,1] [5 x 1]>     0.730 random_search     2     1 search_bett~   0.5
+## 3     3 <dbl[,1] [5 x 1]>     0.743 random_search     2     2 search_bett~   0.5
+## 4     4 <dbl[,1] [5 x 1]>     0.736 random_search     2     3 search_bett~   0.5
+## 5     5 <dbl[,1] [5 x 1]>     0.747 random_search     2     4 search_bett~   0.5
 \end{verbatim}

### 2020-10-09 paper/paper.Rmd
edit fig caption

@@
 
-```{r tour-path, out.height="30%", out.width="60%", fig.align='center', fig.cap="\\label{tour-path}An illustration of the tour path"}
+```{r tour-path, out.height="30%", out.width="60%", fig.align='center', fig.cap="\\label{tour-path}Each square (frame) represents the projected data with a corresponding basis.
+Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames."}
 knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
@@
 
-```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."}
-knitr::include_graphics(path = here::here("figures/global_obj.png"))
-```
+<!-- ```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."} -->
+<!-- knitr::include_graphics(path = here::here("figures/global_obj.png")) -->
+<!-- ``` -->
 
@@
 
-```{r toy-search, fig.height=3, fig.cap="\\label{toy-search}A comparison of plotting the same search points with different plot designs.
-In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read."}
+```{r toy-search, fig.asp=0.5,  fig.cap="\\label{toy-search}A comparison of \`search_better\` with different parameter value on \`max_tries\`.
+A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
+With \`max_tries = 500\` specification, the algorithm is working much harder towards to the end."}
 bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
@@
   scale_color_botanical(palette = "daisy") +
-  facet_wrap(vars(max_tries), ncol = 2)
+  facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
+  xlab("iteration")
 ```
@@
 
-```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} This is a toy example of plotting interpolated points"}
-bind_rows(holes_2d_better_max_tries %>% mutate(group = "max_tries: 500"), 
-          interrupt_no %>% mutate(group = "no interruption"), 
-          holes_2d_better_random %>% mutate(group = "better_random")) %>% 
-  mutate(group = fct_relevel(group, c("max_tries: 500", 
-                                      "no interruption", 
-                                      "better_random"))) %>% 
+```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+The color represents the number of iteration.
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies."}
+bind_rows(holes_2d_better_max_tries %>% mutate(group = "Algorithm 1"), 
+          interrupt_no %>% mutate(group = "Algorithm 2"), 
+          holes_2d_better_random %>% mutate(group = "Algorithm 3")) %>% 
+  mutate(group = fct_relevel(group, c("Algorithm 1", 
+                                      "Algorithm 2", 
+                                      "Algorithm 3"))) %>% 
   explore_trace_interp(group = group) + 
   theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern") + 
-  facet_wrap(vars(group), ncol = 3)
+  scale_color_botanical(palette = "fern")
 ```
@@
 
-```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}Trace plots of the interpolated basis with and without interruption.
-The interruption stops the interpolation when the index value starts to decrease at id = 60.
-The implementation of the interuption finds an ending basis with higher index value using fewer steps.
-"}
+```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}A comparison of the trace plot with and without the interruption.
+When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
+The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration."}
 p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>% 
@@
 
-p1 | p2
+(p1 | p2) & scale_color_botanical(palette = "fern")
 
@@
 
-```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}Projected data before and after polishing.
-Weights on axis other than V2 and V7 have been further reduced and the edges of each cluster have become more obvious."}
+```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}Two-D projection on \`boa6\` data with holes index optimised by \`search_geodesic\`.
+The left panel shows the final projected data before polish and the right panel shows the one after.
+The separation of the clusters on the y axis becomes sharper after the polish."}
 # set.seed(123456)
@@
 
-```{r kol-cdf,fig.asp=0.5, fig.height = 3,  fig.cap = "\\label{kol-cdf}Trace plot of search_geodesic and search_better on 1D projection problem for a noisy index: norm_kol.
-Search_geodesic has varely made progression while search_better has successfully found the informative "}
+```{r kol-cdf,fig.asp=0.5,fig.cap = "\\label{kol-cdf}One-D projection on \`boa5\` data with noisy index \`norm_kol\` optimised by \`search_geodesic\` and \`search_better\`.
+The grey dashed line represents the index value of the theoreical best basis.
+\`search_geodesic\` fails to optimse the noisy index while \`search_better\` has made reasonable improvments to reach the index value close to the one of theoretical best basis."}
 # set.seed(123456)
@@
 To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature $t_0$ so that the probability of accepting an inferior basis is not too high.
-Figure \ref{kol-pca-sim}, ....
-Figure \ref{kol-pca} shows the comparison, in terms of the trace plot and PCA plot, between `search_better_random` and two `search_better` algorithms where one finds `x2` and another finds `x7`.
+Figure \ref{pca-better-default}, ....
+Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between `search_better_random` and two `search_better` algorithms where one finds `x2` and another finds `x7`.
 The trace plot shows that the global maximum has been found by `search_better_random` and one `search_better`, but not the other.
@@
 
-```{r fig.height = 8}
+```{r pca-better-default, fig.height = 6, fig.cap="\\label{pca-better-default}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better\` with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced into 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when \`V2\` is found and the global maximum where \`V7\` is found.
+The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point."}
 set.seed(123)
@@
 
-pca <- all_tidy_better%>% 
+pca <- all_tidy_better %>% 
   mutate(seed = as.factor(seed)) %>% 
   left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
+  mutate(VarFound = fct_relevel(VarFound, c("V7", "V2"))) %>% 
   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
@@
 
-```{r fig.height = 8}
+```{r pca-better-random-default, fig.height = 6, fig.cap = "\\label{pca-better-random-default}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better_random\` with the same seeds.
+\`search_better_random\` has a probablistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
+"}
 set.seed(123)
@@
   left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
+  mutate(VarFound = fct_relevel(VarFound, c("V7", "V2"))) %>% 
   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
@@
 
-```{r fig.height= 8}
+there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search_better_random and two different search_better.
+In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
+serach_better_random and one of the serach_better have found the global maximum but the other fails.
+The lower panel shows the PCA plot for each of the three algorithms.
+The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
+The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
+This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
+
+```{r pca-better-random-tuned, fig.height= 6, fig.cap = "\\label{pca-better-random-tuned}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better_random\` with the same seeds.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \`alpha\` from 0.5 to 0.7.
+This time, all the 20 incidents have found the global maximum."}
 set.seed(123)
@@
 
-pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
+pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
   group_by(seed) %>% arrange(IndexVal) %>% 
@@
 
-Comparison between search_better_random and two different search_better.
-In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
-serach_better_random and one of the serach_better have found the global maximum but the other fails.
-The lower panel shows the PCA plot for each of the three algorithms.
-The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
-The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
-This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
-
 # Implementation {#implementation}

### 2020-10-09 paper/paper.tex
edit fig caption

@@
 
-\caption{\label{tour-path}An illustration of the tour path}\label{fig:tour-path}
+\caption{\label{tour-path}Each square (frame) represents the projected data with a corresponding basis.
+Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames.}\label{fig:tour-path}
 \end{figure}
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The data structure in projection pursuit guided tour.}\label{fig:glb-obj}
-\end{figure}
-
 It is worth noticing that the data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that states
@@
 \includegraphics{paper_files/figure-latex/toy-search-1.pdf}
-\caption{\label{fig:toy-search}\label{toy-search}A comparison of plotting the same search points with different plot designs.
-In the upper plot, points from the last three iterations span the vast majority of the plotting space leaving the first few iterations being squeezed together whilst the lower plot spaces each iteration evenly and presents summary information easy to read.}
+\caption{\label{fig:toy-search}\label{toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
+A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
+With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
-\caption{\label{fig:toy-interp}\label{toy-interp} This is a toy example of plotting interpolated points}
+\caption{\label{fig:toy-interp}\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+The color represents the number of iteration.
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/interruption-1.pdf}
-\caption{\label{fig:interruption}\label{interruption}Trace plots of the interpolated basis with and without interruption.
-The interruption stops the interpolation when the index value starts to decrease at id = 60.
-The implementation of the interuption finds an ending basis with higher index value using fewer steps.}
+\caption{\label{fig:interruption}\label{interruption}A comparison of the trace plot with and without the interruption.
+When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
+The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
-\caption{\label{fig:geo-polish}\label{geo-polish}Projected data before and after polishing.
-Weights on axis other than V2 and V7 have been further reduced and the edges of each cluster have become more obvious.}
+\caption{\label{fig:geo-polish}\label{geo-polish}Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
+The left panel shows the final projected data before polish and the right panel shows the one after.
+The separation of the clusters on the y axis becomes sharper after the polish.}
 \end{figure}
@@
 \includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{fig:kol-cdf}\label{kol-cdf}Trace plot of search\_geodesic and search\_better on 1D projection problem for a noisy index: norm\_kol.
-Search\_geodesic has varely made progression while search\_better has successfully found the informative}
+\caption{\label{fig:kol-cdf}\label{kol-cdf}One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
+The grey dashed line represents the index value of the theoreical best basis.
+\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}
 \end{figure}
@@
 To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature \(t_0\) so that the probability of accepting an inferior basis is not too high.
-Figure \ref{kol-pca-sim}, \ldots{}.
-Figure \ref{kol-pca} shows the comparison, in terms of the trace plot and PCA plot, between \texttt{search\_better\_random} and two \texttt{search\_better} algorithms where one finds \texttt{x2} and another finds \texttt{x7}.
+Figure \ref{pca-better-default}, \ldots{}.
+Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between \texttt{search\_better\_random} and two \texttt{search\_better} algorithms where one finds \texttt{x2} and another finds \texttt{x7}.
 The trace plot shows that the global maximum has been found by \texttt{search\_better\_random} and one \texttt{search\_better}, but not the other.
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/pca-better-default-1.pdf}
+\caption{\label{fig:pca-better-default}\label{pca-better-default}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced into 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
+The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}
+\end{figure}
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/pca-better-random-default-1.pdf}
+\caption{\label{fig:pca-better-random-default}\label{pca-better-random-default}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
+\texttt{search\_better\_random} has a probablistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.}
+\end{figure}
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
-
-Comparison between search\_better\_random and two different search\_better.
+there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search\_better\_random and two different search\_better.
 In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
@@
 This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/pca-better-random-tuned-1.pdf}
+\caption{\label{fig:pca-better-random-tuned}\label{pca-better-random-tuned}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.
+This time, all the 20 incidents have found the global maximum.}
+\end{figure}
 

### 2020-10-09 paper/paper.Rmd
write up noisy index diff seeds example

@@
 
-```{r load-data}
+```{r external, include = FALSE, cache = FALSE}
+library(knitr)
 files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
-
-source(here::here("source/indice.r"))
-source(here::here("source/sim_data.r"))
+read_chunk(here::here('scripts/vis-diag.R'))
 ```
@@
 Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate bases in the neighbourhood, whilst `search_geodesic` is an analogue of gradient ascent on the manifold of the projection basis.
-
-`search_better` is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
-$\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the search terminates and outputs the basis for guided tour to construct an interpolation path.
-The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$.
-Another termination condition is when the maximum number of iteration $l_{\max}$ is reached.
-A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$.
-Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast.
-The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
-[mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]
 
@@
 
-Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme $T(l)$.
-Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
-This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
-a one dimensional projection of the dataset with two informative variables).
-The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+`search_better` is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+$\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the search terminates and outputs the basis for guided tour to construct an interpolation path.
+The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$.
+Another termination condition is when the maximum number of iteration $l_{\max}$ is reached.
+A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$.
+Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast.
+The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
+[mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]
 
@@
 
-@cook1995grand used a gradient ascent algorithm on the manifold of the projection bases.
-In gradient ascent, one first find the direction for improvment via computing the gradient information.
-In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
-A geodesic is then constructed using the current basis and one of the $2n$ bases that have the highest index value.
-If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
-In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
-In `search_geodesic`, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from $-\pi/4$ to $\pi/4$.
-The optima $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value.
-As above, another termination condition is when $l_{\max}$ is reached.
-Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
+Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme $T(l)$.
+Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
+This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
+a one dimensional projection of the dataset with two informative variables).
+The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
 
@@
 
+@cook1995grand used a gradient ascent algorithm on the manifold of the projection bases.
+In gradient ascent, one first find the direction for improvment via computing the gradient information.
+In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
+A geodesic is then constructed using the current basis and one of the $2n$ bases that have the highest index value.
+If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
+In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
+In `search_geodesic`, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+The optima $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value.
+As above, another termination condition is when $l_{\max}$ is reached.
+Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
+
 # Visual diagnostics {#vis-diag}
@@
 
-Figure \ref{toy-search} presents a comparison between the two plotting options discussed above.
+Figure \@ref(fig:toy-search) presents a comparison between the two plotting options discussed above.
 Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
@@
 
-```{r toy-search, fig.asp=0.5,  fig.cap="\\label{toy-search}A comparison of \`search_better\` with different parameter value on \`max_tries\`.
+```{r toy-search, fig.cap = "(ref:toy-search)"}
+```
+
+(ref:toy-search) A comparison of `search_better` with different parameter value on `max_tries`.
 A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
-With \`max_tries = 500\` specification, the algorithm is working much harder towards to the end."}
-bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500), 
-          holes_2d_better %>% mutate(max_tries = 25)) %>% 
-  explore_trace_search(group = max_tries) + 
-  scale_color_botanical(palette = "daisy") +
-  facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
-  xlab("iteration")
-```
+With `max_tries = 500` specification, the algorithm is working much harder towards to the end.
 
 ## Examining the optimisation progress
-
-Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
-Figure \ref{toy-interp} presents the interpolation of three different tour paths.
-The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
-While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
-The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern.
-In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
-When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
-While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
 
@@
 
+Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
+Figure \ref{toy-interp} presents the interpolation of three different tour paths.
+The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
+While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
+The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern.
+In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
+When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
+While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
+
 ## Understanding the optimiser's coverage of the search space
-
-Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go above five (5).
-Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d > 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.
-Principal component analysis.
-To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
-
-<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
-This gives a matrix of dimension $n$ by $(p*d)$.
-Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
-By definition, the principal components maximise the variance of the projection bases in the reduced space.
-Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
-Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix -->
-
-<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
-
-<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
-The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
-
-<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
-THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
-colour).
--->
-
-Figure \ref{toy-pca} shows the first two principal components of two search paths.
-Starting from the same position, two searching methods take different paths to get its final bases.
-There are two theoretical bases because they correspond to the same data projection $Y$ with a 180 degree rotation.
-What differentiate the two optimisers is that `search_better()` also involve random evaluation of points in the sphere during its sampling process while `search_geodesic()` doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
 
@@
 
+Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
+Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go above five (5).
+Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d > 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.
+Principal component analysis.
+To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
+
+<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
+This gives a matrix of dimension $n$ by $(p*d)$.
+Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
+By definition, the principal components maximise the variance of the projection bases in the reduced space.
+Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
+Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix -->
+
+<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
+
+<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
+The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
+
+<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
+THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
+colour).
+-->
+
+Figure \ref{toy-pca} shows the first two principal components of two search paths.
+Starting from the same position, two searching methods take different paths to get its final bases.
+There are two theoretical bases because they correspond to the same data projection $Y$ with a 180 degree rotation.
+What differentiate the two optimisers is that `search_better()` also involve random evaluation of points in the sphere during its sampling process while `search_geodesic()` doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
+
 ## Animating the diagnostic plots
-
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, ...
 
@@
 
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
+In Figure \ref{toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
+
 ## The tour looking at itself
-
-Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a stereoscopic view of the search.
-To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation.
-This can be done via a regular grand tour or a slice tour [@laa2020slice], which emphasizes the points that are close to a section of the sphere.
-
-Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
-This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
-Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
 
@@
 
+Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a stereoscopic view of the search.
+To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation.
+This can be done via a regular grand tour or a slice tour [@laa2020slice], which emphasizes the points that are close to a section of the sphere.
+
+Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
+Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
+
 # Diagnosing an optimiser {#application}
@@
 
-Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
-This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
-
-Similar to `search_better` as a stochastic random search, `search_polish` has a different scheme of reducing the search neighbourhood.
-In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
-On the contrary, `search_polish` allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
-Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries`.
-These include:
-
-1) the distance between the candidate basis and the current basis needs to be larger than 1e-3;
-2) the percentage change of the index value need to be larger than 1e-5; and
-3) the alpha parameter on itself need to be larger than 0.01
-
-Figure \ref{geo-polish} presents the final projections found before and after applying `search_polish` using the demo code.
-Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables, namely `V1`, `V8`, `V9` and `V10`.
-When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
-
 ```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}Two-D projection on \`boa6\` data with holes index optimised by \`search_geodesic\`.
@@
 
+Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
+This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
+
+Similar to `search_better` as a stochastic random search, `search_polish` has a different scheme of reducing the search neighbourhood.
+In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
+On the contrary, `search_polish` allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
+Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries`.
+These include:
+
+1) the distance between the candidate basis and the current basis needs to be larger than 1e-3;
+2) the percentage change of the index value need to be larger than 1e-5; and
+3) the alpha parameter on itself need to be larger than 0.01
+
+Figure \ref{geo-polish} presents the final projections found before and after applying `search_polish` using the demo code.
+Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables, namely `V1`, `V8`, `V9` and `V10`.
+When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
+
 ## Seeing the signal in the noise
-
-Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
-`norm_kol`, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as $F_{.}(u)$ with the subscript indicating the projection or the random normal variable,  the `norm_kol` index is defined by
-
-$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
-
-Figure \ref{kol-cdf} compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
-This time, the interpolated path is no longer smooth when using either algorithm.
-It is also obvious that `search_geodesic` fails to optimise this rough index since there is barely improvment of the index value.
-On the other hand, `search_better` is doing a relatively good job on the optimisation for the 5-variable dataset `boa5`.
-The theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)` and `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 4)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
-A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 4)`].
-At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
@@
 
-The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection.
-The dataset used is `boa6` with `x2` and `x7` being informative.
-The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found.
-
-A simulation with 20 different seeds shows that `search_better` can find either one of the informative variable but there's no guarantee on which seed finds which variable.
-`search_better_random` manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed.
-To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature $t_0$ so that the probability of accepting an inferior basis is not too high.
-Figure \ref{pca-better-default}, ....
-Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between `search_better_random` and two `search_better` algorithms where one finds `x2` and another finds `x7`.
-The trace plot shows that the global maximum has been found by `search_better_random` and one `search_better`, but not the other.
-The PCA plot further illustrate how the two algorithms finds the global maximum: `search_better_random` travels along the top right corner on the projected 2D circle while the path of `search_better` is much shorter and only goes pass in between the two theoretical bases.
-
-(there's more things going on about both algorithms experimenting with different seeds.)
+Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
+`norm_kol`, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
+Denotes the ECDF function as $F_{.}(u)$ with the subscript indicating the projection or the random normal variable,  the `norm_kol` index is defined by
+
+$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
+
+Figure \ref{kol-cdf} compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
+This time, the interpolated path is no longer smooth when using either algorithm.
+It is also obvious that `search_geodesic` fails to optimise this rough index since there is barely improvment of the index value.
+On the other hand, `search_better` is doing a relatively good job on the optimisation for the 5-variable dataset `boa5`.
+The theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)` and `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 4)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
+A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 4)`].
+At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
@@
 
-there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search_better_random and two different search_better.
-In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
-serach_better_random and one of the serach_better have found the global maximum but the other fails.
-The lower panel shows the PCA plot for each of the three algorithms.
-The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
-The paths of two search_better are relatively short and simple while the search_better_random has a more sophisticated path.
-This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
-
 ```{r pca-better-random-tuned, fig.height= 6, fig.cap = "\\label{pca-better-random-tuned}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better_random\` with the same seeds.
@@
 
+The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is  presented.
+The dataset used is `boa6` with `x2` and `x7` being informative.
+The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found.
+Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: `search_better` and `search_better_random`.
+The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.
+
+Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser `search_better`.
+This happens because `search_better` will always accept a basis with higher index value.
+Figure \ref{pca-better-random-default} shows the results when the optimiser `search_better_random` is used.
+With a probability acceptance on inferior basis, more simulation ends up finds the global maximum.
+An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}).
+When using `search_better` to optimise, it goes straight down the track to find `V2` whilst with `search_better_random`, it first has a tendency to hand towards `V2` but then makes a sharp turn back to the track that finds `V7`.
+
+In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated basis and hence represents a wider search.
+By this specification, all the 20 seends find the global maximum and a few simulations i.e.
+seed 4761 (row 4,  column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.
+
 # Implementation {#implementation}

### 2020-10-09 paper/paper.tex
write up noisy index diff seeds example

@@
 
-\texttt{search\_better} is a random search device that samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current basis \(\mathbf{A}_{\text{cur}}\) by \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\) where \(\alpha\) controls the radius of the sampling neighbourhood and \(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the same dimension as \(\mathbf{A}_{\text{cur}}\).
-\(\mathbf{A}_{l}\) is then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis \(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis for guided tour to construct an interpolation path.
-The next iteration of search begins after adjusting \(\alpha\) by a cooling parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\).
-Another termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
-A slightly different cooling scheme has been proposed by \citet{posse1995projection} to include a halving parameter \(c\).
-Rather than reducing the radius of the searching neighbourhood, \(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if the last search takes more than \(c\) times to find an accepted basis to avoid the searching space being reduced too fast.
-The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
-{[}mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search{]}
-
 \begin{algorithm}
@@
 
-Simulated annealing \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme \(T(l)\).
-Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
-This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.~a one dimensional projection of the dataset with two informative variables).
-The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+\texttt{search\_better} is a random search device that samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current basis \(\mathbf{A}_{\text{cur}}\) by \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\) where \(\alpha\) controls the radius of the sampling neighbourhood and \(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the same dimension as \(\mathbf{A}_{\text{cur}}\).
+\(\mathbf{A}_{l}\) is then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis \(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis for guided tour to construct an interpolation path.
+The next iteration of search begins after adjusting \(\alpha\) by a cooling parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\).
+Another termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
+A slightly different cooling scheme has been proposed by \citet{posse1995projection} to include a halving parameter \(c\).
+Rather than reducing the radius of the searching neighbourhood, \(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if the last search takes more than \(c\) times to find an accepted basis to avoid the searching space being reduced too fast.
+The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
+{[}mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search{]}
 
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm on the manifold of the projection bases.
-In gradient ascent, one first find the direction for improvment via computing the gradient information.
-In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
-A geodesic is then constructed using the current basis and one of the \(2n\) bases that have the highest index value.
-If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
-In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
-In \texttt{search\_geodesic}, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
-The optima \(\mathbf{A}_{**}\) is outputted for the current iteration if the percentage change in the index value between \(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a threshold value.
-As above, another termination condition is when \(l_{\max}\) is reached.
-Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
+Simulated annealing \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme \(T(l)\).
+Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
+This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
+This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.~a one dimensional projection of the dataset with two informative variables).
+The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
 
@@
 \end{algorithm}
+
+\citet{cook1995grand} used a gradient ascent algorithm on the manifold of the projection bases.
+In gradient ascent, one first find the direction for improvment via computing the gradient information.
+In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
+A geodesic is then constructed using the current basis and one of the \(2n\) bases that have the highest index value.
+If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
+In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
+In \texttt{search\_geodesic}, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
+The optima \(\mathbf{A}_{**}\) is outputted for the current iteration if the percentage change in the index value between \(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a threshold value.
+As above, another termination condition is when \(l_{\max}\) is reached.
+Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
 
@@
 
-Figure \ref{toy-search} presents a comparison between the two plotting options discussed above.
+Figure \ref{fig:toy-search} presents a comparison between the two plotting options discussed above.
 Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
@@
 \includegraphics{paper_files/figure-latex/toy-search-1.pdf}
-\caption{\label{fig:toy-search}\label{toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
+\caption{\label{fig:toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
 A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
@@
 \subsection{Examining the optimisation progress}\label{examining-the-optimisation-progress}}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
+\caption{\label{fig:toy-interp}\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+The color represents the number of iteration.
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}
+\end{figure}
 
@@
 
+\hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
+\subsection{Understanding the optimiser's coverage of the search space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
+
 \begin{figure}
 \centering
-\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
-\caption{\label{fig:toy-interp}\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
-The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}
+\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
+\caption{\label{fig:toy-pca}\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}
 \end{figure}
-
-\hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
-\subsection{Understanding the optimiser's coverage of the search space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
-\caption{\label{fig:toy-pca}\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}
-\end{figure}
-
 \hypertarget{animating-the-diagnostic-plots}{%
 \subsection{Animating the diagnostic plots}\label{animating-the-diagnostic-plots}}
-
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order. In Figure \ref{toy-pca-animated}, \ldots{}
 
@@
 
+Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
+In Figure \ref{toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
+
 \hypertarget{the-tour-looking-at-itself}{%
 \subsection{The tour looking at itself}\label{the-tour-looking-at-itself}}
+
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
+\caption{\label{fig:toy-tour}\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+The second frame on the top row view the space from a direction that is close to the one in PCA plot.
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}
+\end{figure}
 
@@
 Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
-\caption{\label{fig:toy-tour}\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row view the space from a direction that is close to the one in PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}
-\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
+\caption{\label{fig:geo-polish}\label{geo-polish}Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
+The left panel shows the final projected data before polish and the right panel shows the one after.
+The separation of the clusters on the y axis becomes sharper after the polish.}
+\end{figure}
+
 Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
@@
 
+\hypertarget{seeing-the-signal-in-the-noise}{%
+\subsection{Seeing the signal in the noise}\label{seeing-the-signal-in-the-noise}}
+
 \begin{figure}
 \centering
-\includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
-\caption{\label{fig:geo-polish}\label{geo-polish}Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
-The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.}
+\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
+\caption{\label{fig:kol-cdf}\label{kol-cdf}One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
+The grey dashed line represents the index value of the theoreical best basis.
+\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}
 \end{figure}
-
-\hypertarget{seeing-the-signal-in-the-noise}{%
-\subsection{Seeing the signal in the noise}\label{seeing-the-signal-in-the-noise}}
 
@@
 At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{fig:kol-cdf}\label{kol-cdf}One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
-The grey dashed line represents the index value of the theoreical best basis.
-\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}
-\end{figure}
-
-The second experiment with the noisy index is to understand how the optimisers perform when there are multiple informative variables in the case of 1D projection.
-The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative.
-The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found.
-
-A simulation with 20 different seeds shows that \texttt{search\_better} can find either one of the informative variable but there's no guarantee on which seed finds which variable.
-\texttt{search\_better\_random} manages to find the global maximum via a metropolis-hasting algorithm without being affected by the seed.
-To ensure the algorithm finish the search in a timely manner, it is necessary to set a reasonable initial temperature \(t_0\) so that the probability of accepting an inferior basis is not too high.
-Figure \ref{pca-better-default}, \ldots{}.
-Figure \ref{pca-better-random-default} shows the comparison, in terms of the trace plot and PCA plot, between \texttt{search\_better\_random} and two \texttt{search\_better} algorithms where one finds \texttt{x2} and another finds \texttt{x7}.
-The trace plot shows that the global maximum has been found by \texttt{search\_better\_random} and one \texttt{search\_better}, but not the other.
-The PCA plot further illustrate how the two algorithms finds the global maximum: \texttt{search\_better\_random} travels along the top right corner on the projected 2D circle while the path of \texttt{search\_better} is much shorter and only goes pass in between the two theoretical bases.
-
-(there's more things going on about both algorithms experimenting with different seeds.)
 
@@
 
-there's more text here Figure \ref{pca-better-random-tuned}, Comparison between search\_better\_random and two different search\_better.
-In the trace plot, the two dashed horizontal lines are the index values of the two theoretical bases.
-serach\_better\_random and one of the serach\_better have found the global maximum but the other fails.
-The lower panel shows the PCA plot for each of the three algorithms.
-The two yellow dots, again, correspond to the theoretical bases with the one on the right being the global maximum.
-The paths of two search\_better are relatively short and simple while the search\_better\_random has a more sophisticated path.
-This is due to the fact that the interpolation at each iteration are longer, which shows its ability to find new bases at a distance
-
 \begin{figure}
@@
 \end{figure}
+
+The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is presented.
+The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative.
+The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found.
+Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: \texttt{search\_better} and \texttt{search\_better\_random}.
+The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.
+
+Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser \texttt{search\_better}.
+This happens because \texttt{search\_better} will always accept a basis with higher index value.
+Figure \ref{pca-better-random-default} shows the results when the optimiser \texttt{search\_better\_random} is used.
+With a probability acceptance on inferior basis, more simulation ends up finds the global maximum.
+An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}).
+When using \texttt{search\_better} to optimise, it goes straight down the track to find \texttt{V2} whilst with \texttt{search\_better\_random}, it first has a tendency to hand towards \texttt{V2} but then makes a sharp turn back to the track that finds \texttt{V7}.
+
+In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated basis and hence represents a wider search.
+By this specification, all the 20 seends find the global maximum and a few simulations i.e.~seed 4761 (row 4, column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.
 

### 2020-10-09 scripts/vis-diag.R

@@ -1,0 +1,13 @@
+## ---- load-pkg
+library(ferrn)
+library(tidyverse)
+files <- paste0("data/", list.files(here::here("data")))
+purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
+
+## ---- toy-search
+bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500),
+          holes_2d_better %>% mutate(max_tries = 25)) %>%
+  explore_trace_search(group = max_tries) +
+  scale_color_botanical(palette = "daisy") +
+  facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
+  xlab("iteration")

### 2020-10-09 paper/paper.Rmd
separate scripts from rmd, clean up repo

@@
     base_format: rticles::asa_article
-    fig_height: 4
+    fig_height: 3
 editor_options: 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,root.dir = here::here())
-
-library(tidyverse)
-library(ferrn)
-library(patchwork)
-library(ggrepel)
-library(tourr)
-library(gganimate)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
+                      root.dir = here::here(), fig.path = here::here("figs/"), 
+                      fig.align = 'center')
 ```
@@
 library(knitr)
-files <- paste0("data/", list.files(here::here("data")))
-purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 read_chunk(here::here('scripts/vis-diag.R'))
+read_chunk(here::here('scripts/examples.R'))
+read_chunk(here::here('scripts/kol-better-default.R'))
+read_chunk(here::here('scripts/kol-better-random-default.R'))
+read_chunk(here::here('scripts/kol-better-random-tuned.R'))
+read_chunk(here::here('scripts/implementation.R'))
+```
+
+```{r load-pkg}
 ```
@@
 
-```{r tour-path, out.height="30%", out.width="60%", fig.align='center', fig.cap="\\label{tour-path}Each square (frame) represents the projected data with a corresponding basis.
-Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames."}
-knitr::include_graphics(path = here::here("figures/tour_path_keynote/tour_path_keynote.001.png"))
+```{r tour-path, out.height="30%", out.width="60%", fig.cap="(ref:tour-path)"}
+include_graphics(here::here("img/tour_path.png"))
 ```
+
+(ref:tour-path) Each square (frame) represents the projected data with a corresponding basis.
+Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames.
 
@@
 
-```{r toy-interp, fig.height = 3, fig.cap = "\\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
-The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies."}
-bind_rows(holes_2d_better_max_tries %>% mutate(group = "Algorithm 1"), 
-          interrupt_no %>% mutate(group = "Algorithm 2"), 
-          holes_2d_better_random %>% mutate(group = "Algorithm 3")) %>% 
-  mutate(group = fct_relevel(group, c("Algorithm 1", 
-                                      "Algorithm 2", 
-                                      "Algorithm 3"))) %>% 
-  explore_trace_interp(group = group) + 
-  theme(legend.position = "none") + 
-  scale_color_botanical(palette = "fern")
+```{r toy-interp,fig.cap = "(ref:toy-interp)"}
 ```
 
+(ref:toy-interp) The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+The color represents the number of iteration.
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.
+
 Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
-Figure \ref{toy-interp} presents the interpolation of three different tour paths.
+Figure \@ref(fig:toy-interp) presents the interpolation of three different tour paths.
 The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
@@
 
-```{r toy-pca, fig.height = 3, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
-# path1 <- holes_1d_geo$basis %>% 
-#   flatten_dbl() %>% 
-#   matrix(ncol = 5, byrow = TRUE)
-# path2 <- holes_1d_better$basis %>% 
-#   flatten_dbl() %>% 
-#   matrix(ncol = 5, byrow = TRUE)
-# set.seed(1)
-# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-# theoretical <- matrix(c(0, 1, 0, 0, 0,
-#                         0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
-# 
-# dt <- rbind(sphere, path1, path2, theoretical)
-# pca <- dt %>% prcomp(scale. = TRUE)
-# #summary(pca)
-# pca_pred <- predict(pca) %>%
-#   as_tibble() %>%
-#   mutate(method = c(rep("sphere", nrow(sphere)),
-#                     holes_1d_geo$method,
-#                     holes_1d_better$method,
-#                     rep("theoretical", 2))) %>%
-#   group_by(method) %>%
-#   mutate(id = row_number(),
-#          id = ifelse(method == "sphere", 1, id)) %>%
-#   ungroup()
-# 
-# p <- pca_pred %>%
-#   ggplot(aes(x = PC1, y = PC2, color = method))+
-#   geom_point() +
-#   geom_point(data = pca_pred %>% filter(method == "theoretical"), 
-#              size = 10) + 
-#   geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
-#   theme(aspect.ratio = 1) + 
-#   scale_color_botanical(palette = "cherry")
-# 
-# p
+```{r toy-pca, fig.cap = "(ref:toy-pca)"}
+```
 
-p <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
-  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
-                   index = tourr::holes(), raw_data = boa5) %>% 
-  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
-                   index = tourr::holes(), raw_data  = boa5)%>%
-  explore_space_pca(col = method)  + 
-  scale_color_botanical(palette = "cherry") + 
-  theme(legend.position = "bottom")
-
-p
-```
+(ref:toy-pca) PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search
 
@@
 
-Figure \ref{toy-pca} shows the first two principal components of two search paths.
+Figure \@ref(fig:toy-pca) shows the first two principal components of two search paths.
 Starting from the same position, two searching methods take different paths to get its final bases.
@@
 
-```{r fig.height = 3, fig.cap = "\\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progrssion from the start to finish in each algorithm."}
-# ani <- p + theme(legend.position = "none") +
-#   transition_states(id) +
-#   shadow_mark()
-# 
-# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>% 
-#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5), 
-#                    index = tourr::holes(), raw_data = data) %>% 
-#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5), 
-#                    index = tourr::holes(), raw_data  = data)%>%
-#   explore_space_pca(col = method, animate = TRUE)  + 
-#   scale_color_botanical(palette = "cherry") 
-# 
-# animate(ani, nframes = 350, device = "png",
-#         renderer = file_renderer("figures/pca/",
-#                                  prefix = "pca", overwrite = TRUE))
-
-frames <- c("0002", "0058", "0078", "0117", "0172", "0350")
-ani <- paste0(here::here("figures/"), "pca/", "pca", frames, ".png")
-rl <- lapply(ani, png::readPNG)
-gl <-  lapply(rl, grid::rasterGrob)
-wrap_plots(gl) 
+```{r toy-pca-aniamted, fig.cap = "(ref:toy-pca-aniamted)"}
 ```
 
+(ref:toy-pca-aniamted) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.
+
 Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
-In Figure \ref{toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
+In Figure \@ref(fig:toy-pca-animated), There should actually be more writing here to and also the this examples may need to be changed.
 
@@
 
-```{r toy-tour, fig.cap="\\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+```{r toy-tour, fig.cap="(ref:toy-tour)"}
+```
+
+(ref:toy-tour) A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
 The second frame on the top row view the space from a direction that is close to the one in PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles."}
-set.seed(1)
-sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-path1 <- holes_1d_better$basis %>% 
-  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-path2 <- holes_1d_geo$basis %>% 
-  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-theoretical <- matrix(c(0, 1, 0, 0, 0,
-                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)
-start <- get_start(holes_1d_better) %>% pull(basis) %>% .[[1]] %>% matrix(nrow = 1)
-dt <- rbind(sphere, path1, path2, theoretical, start)
-colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#D3D3D3",c("#524340",  #orchre
-                     "#B4B754",  # green
-                     "#F3B422" # yellow
-                     ))
-color <- c(rep(pal[1], nrow(sphere)),
-           rep(pal[2], nrow(path1)),
-           rep(pal[3], nrow(path2)),
-           rep(pal[4], 2), 
-           pal[2] # for start
-           )
-cex <- c(rep(1, nrow(sphere)),
-         rep(1.5, nrow(path1)), 
-         rep(1.5, nrow(path2)), 
-         rep(5, 2), 
-         5 # for start
-         )
-
-# 
-# bind_rows(holes_1d_geo, holes_1d_better) %>% 
-#   explore_space_tour(color = method)
-#animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial
-
-# set.seed(123)
-# render(
-#   dt,
-#   tour_path = grand_tour(),
-#   dev = "png",
-#   display = display_xy(col=color,cex = cex, axes = "bottomleft"),
-#   rescale = FALSE,
-#   frames = 500,
-#   here::here("figures","tour", "tour%03d.png")
-# )
-
-frames <- c("001", "143", "078", "117", "172", "350")
-ani <- paste0(here::here("figures/"), "tour/", "tour", frames, ".png")
-rl <- lapply(ani, png::readPNG)
-gl <-  lapply(rl, grid::rasterGrob)
-wrap_plots(gl) 
-```
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
 
@@
 We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for the `holes` index.
-From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
+From the left panel of Figure \@ref(fig:interruption), we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
 In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
 Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
-After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{interruption}.
+After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \@ref(fig:interruption).
 
-```{r interruption,fig.asp=0.5, fig.cap = "\\label{interruption}A comparison of the trace plot with and without the interruption.
+```{r interruption, fig.asp=0.5, fig.cap = "(ref:interruption)"}
+```
+
+(ref:interruption) A comparison of the trace plot with and without the interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration."}
-p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>% 
-  mutate(anno = c("current basis", "interpolated basis", "target basis"))
-
-p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") + 
-  geom_point(data = p1_anno) + 
-  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) + 
-  theme(legend.position = "none")
-
-p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) + 
-  theme(legend.position = "none")
-
-(p1 | p2) & scale_color_botanical(palette = "fern")
-
-```
+The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.
 
@@
 
-```{r geo-polish, fig.height= 2.5, fig.cap = "\\label{geo-polish}Two-D projection on \`boa6\` data with holes index optimised by \`search_geodesic\`.
+```{r polish, fig.height= 2.5, fig.cap = "(ref:polish)"}
+```
+
+(ref:polish) Two-D projection on `boa6` data with holes index optimised by `search_geodesic`.
 The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish."}
-# set.seed(123456)
-# render(
-#   boa6,
-#   tour_path = guided_tour(holes(), d = 2, search_f = search_geodesic),
-#   dev = "png",
-#   display = display_xy(axes = "bottomleft", verbose = TRUE),
-#   rescale = FALSE,
-#   frames = 100,
-#   file = here::here("figures","geo_polish", "before%03d.png")
-# )
-# 
-# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
-# 
-# set.seed(123456)
-# render(
-#   boa6,
-#   tour_path =  guided_tour(holes(), d = 2, search_f = search_polish, alpha = 0.1),
-#   dev = "png",
-#   display = display_xy(axes = "bottomleft", verbose = TRUE),
-#   start = last_basis,
-#   rescale = FALSE,
-#   frames = 100,
-#   file = here::here("figures","geo_polish", "after%03d.png")
-# )
-
-# set.seed(123456)
-# holes_2d_geo <- animate_xy(boa6,tour_path =
-#                              guided_tour(holes(), d = 2,
-#                                          search_f = search_geodesic),
-#                            rescale = FALSE, verbose = TRUE)
-# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
-# 
-# set.seed(123456)
-# holes_2d_geo_polish <- animate_xy(boa6, tour_path =
-#                                     guided_tour(holes(), d = 2,
-#                                                 search_f = search_polish, alpha = 0.1),
-#                                   rescale = FALSE, verbose = TRUE,
-#                                   start = last_basis)
-# 
-# save(holes_2d_geo, file = here::here("data", "holes_2d_geo.rda"))
-# save(holes_2d_geo_polish, file = here::here("data", "holes_2d_geo_polish.rda"))
-
-before <- png::readPNG(here::here("figures","geo_polish", "before100.png"))
-after <- png::readPNG(here::here("figures","geo_polish", "after004.png"))
-gl <-  lapply(list(before, after), grid::rasterGrob)
-wrap_plots(gl) 
-```
+The separation of the clusters on the y axis becomes sharper after the polish.
 
@@
 
-Figure \ref{geo-polish} presents the final projections found before and after applying `search_polish` using the demo code.
+Figure \@ref(fig:polish) presents the final projections found before and after applying `search_polish` using the demo code.
 Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables, namely `V1`, `V8`, `V9` and `V10`.
-When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
+When appears in the projected data in Figure \@ref(fig:polish), polish works to sharpen the edges of each cluster.
 
@@
 
-```{r kol-cdf,fig.asp=0.5,fig.cap = "\\label{kol-cdf}One-D projection on \`boa5\` data with noisy index \`norm_kol\` optimised by \`search_geodesic\` and \`search_better\`.
+```{r noisy-better-geo,fig.asp=0.5,fig.cap = "(ref:noisy-better-geo)"}
+```
+
+(ref:noisy-better-geo) One-D projection on `boa5` data with noisy index `norm_kol` optimised by `search_geodesic` and `search_better`.
 The grey dashed line represents the index value of the theoreical best basis.
-\`search_geodesic\` fails to optimse the noisy index while \`search_better\` has made reasonable improvments to reach the index value close to the one of theoretical best basis."}
-# set.seed(123456)
-# kol_1d_geo <- 
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
-#                                              search_f =  search_geodesic, max.tries = 100),
-#                rescale = FALSE, verbose = TRUE)
-# 
-# set.seed(123456)
-# kol_1d_better <- 
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
-#                                              search_f =  search_better, max.tries = 100),
-#                rescale = FALSE, verbose = TRUE)
-# save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
-# save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
-
-index <- tourr::norm_kol()
-theo_best_index_val <- index(as.matrix(boa5) %*% matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1))
-
-bind_rows(kol_1d_geo, kol_1d_better) %>% 
-  explore_trace_interp(col = method, group = method) +
-  facet_wrap(vars(method), ncol = 2) + 
-  geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) + 
-  scale_color_botanical() +
-  theme(legend.position = "none", aspect.ratio = 1)
-
-```
+`search_geodesic` fails to optimse the noisy index while `search_better` has made reasonable improvments to reach the index value close to the one of theoretical best basis.
 
@@
 
-Figure \ref{kol-cdf} compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
+Figure \@ref(fig:noisy-better-geo) compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
 This time, the interpolated path is no longer smooth when using either algorithm.
@@
 
-```{r pca-better-default, fig.height = 6, fig.cap="\\label{pca-better-default}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better\` with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced into 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when \`V2\` is found and the global maximum where \`V7\` is found.
-The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point."}
-set.seed(123)
-seed <- sample.int(10000, 20)
-# kol_1d_2var_better_sim <- list()
-# kol_1d_2var_better_sim_polish <- list()
-# for (i in 1: length(seed)){
-#   set.seed(seed[i])
-#   kol_1d_2var_better_sim[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_better, max.tries = 100),
-#                  rescale = FALSE, verbose = TRUE)
-# 
-# 
-#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
-# 
-#   set.seed(1842)
-#   kol_1d_2var_better_sim_polish[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_polish, max.tries = 100),
-#                  rescale = FALSE, verbose = TRUE,
-#                  start = last_basis)
-# }
-
-# save(kol_1d_2var_better_sim, file = here::here("data", "kol_1d_2var_better_sim.rda"))
-# save(kol_1d_2var_better_sim_polish, file = here::here("data", "kol_1d_2var_better_sim_polish.rda"))
-
-polish_list <- purrr::map2(kol_1d_2var_better_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-better_list <- purrr::map2(kol_1d_2var_better_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-polish_tidy <- do.call(dplyr::bind_rows, polish_list)
-better_tidy <- do.call(dplyr::bind_rows, better_list)
-all_tidy_better <- bind_rows(better_tidy, polish_tidy)
-
-best <- get_best(all_tidy_better, group = seed) %>% arrange(index_val)
-
-all_summary <- best$basis %>% 
-  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
-  as_tibble() %>% 
-  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
-  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
-  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
-  group_by(seed, IndexVal) %>% 
-  filter(abs(value) == max(abs(value))) %>% 
-  arrange(IndexVal) %>% ungroup()
-
-pca <- all_tidy_better %>% 
-  mutate(seed = as.factor(seed)) %>% 
-  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
-  mutate(VarFound = fct_relevel(VarFound, c("V7", "V2"))) %>% 
-  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>% 
-   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>%
-
-  compute_pca()
-
-random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
-theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
-
-pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
-  group_by(seed) %>% arrange(IndexVal) %>% 
-  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
-  geom_point(data = random, col = "grey", size = 0.3) +
-  geom_point(aes(col = VarFound), alpha = 0.1) + 
-  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
-  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
-  geom_point(data = pca$aug %>% get_start() %>% filter(method == "search_better_random"), size = 3) +
-  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
-  theme_void() + 
-  theme(aspect.ratio = 1, legend.position = "bottom") + 
-  scale_color_botanical(palette = "fern")  
+```{r kol-better-default, fig.height = 6, fig.cap="(ref:kol-better-default)"}
 ```
 
-```{r pca-better-random-default, fig.height = 6, fig.cap = "\\label{pca-better-random-default}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better_random\` with the same seeds.
-\`search_better_random\` has a probablistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
-"}
-set.seed(123)
-seed <- sample.int(10000, 20)
-# kol_1d_2var_better_random_sim2 <- list()
-# kol_1d_2var_better_random_sim_polish2 <- list()
-# for (i in 1: length(seed)){
-#   cat("i = ", i, "seed = ", seed[i], "\n")
-#   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim2[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_better_random, max.tries = 200),
-#                  rescale = FALSE, verbose = TRUE)
-# 
-# 
-#   last_basis <- get_best(kol_1d_2var_better_random_sim2[[i]])$basis %>% .[[1]]
-# 
-#   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim_polish2[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_polish, max.tries = 200),
-#                  rescale = FALSE, verbose = TRUE,
-#                  start = last_basis)
-# 
-# }
-# save(kol_1d_2var_better_random_sim2, file = here::here("data", "kol_1d_2var_better_sim2.rda"))
-# save(kol_1d_2var_better_random_sim_polish2, file = here::here("data", "kol_1d_2var_better_sim_polish2.rda"))
+(ref:kol-better-default) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better` with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced into 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
+The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.
 
-polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish2,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-better_list <- purrr::map2(kol_1d_2var_better_random_sim2, seed,  ~.x %>% dplyr ::mutate(seed = .y)) 
-polish_tidy <- do.call(dplyr::bind_rows, polish_list)
-better_tidy <- do.call(dplyr::bind_rows, better_list)
-all_tidy <- bind_rows(better_tidy, polish_tidy)
-
-best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
-
-all_summary <- best$basis %>% 
-  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
-  as_tibble() %>% 
-  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
-  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
-  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
-  group_by(seed, IndexVal) %>% 
-  filter(abs(value) == max(abs(value))) %>% 
-  arrange(IndexVal) %>% ungroup()
-
-pca <- all_tidy %>% 
-  mutate(seed = as.factor(seed)) %>% 
-  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
-  mutate(VarFound = fct_relevel(VarFound, c("V7", "V2"))) %>% 
-  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                   index = tourr::norm_kol(),
-                   raw_data = boa6) %>% 
-  bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                   index = tourr::norm_kol(),
-                   raw_data = boa6) %>%
-
-  compute_pca()
-
-random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
-theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
-
-pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>% 
-  group_by(seed) %>% arrange(IndexVal) %>% 
-  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
-  geom_point(data = random, col = "grey", size = 0.3) +
-  geom_point(aes(col = VarFound), alpha = 0.05) + 
-  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
-  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
-  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
-  theme_void() + 
-  theme(aspect.ratio = 1, legend.position = "bottom") + 
-  scale_color_botanical(palette = "fern") 
+```{r kol-better-random-default, fig.height = 6, fig.cap = "(ref: kol-better-random-default)"}
 ```
 
-```{r pca-better-random-tuned, fig.height= 6, fig.cap = "\\label{pca-better-random-tuned}One-D projection of \`norm_kol\` index on \`boa6\` data optimised by \`search_better_random\` with the same seeds.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \`alpha\` from 0.5 to 0.7.
-This time, all the 20 incidents have found the global maximum."}
-set.seed(123)
-seed <- sample.int(10000, 20)
-# kol_1d_2var_better_random_sim <- list()
-# kol_1d_2var_better_random_sim_polish <- list()
-# for (i in 1: length(seed)){
-#   cat("i = ", i, "seed = ", seed[i], "\n")
-#   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_better_random, max.tries = 500, t0 = 1e-6, alpha = 0.7),
-#                  rescale = FALSE, verbose = TRUE)
-# 
-# 
-#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
-# 
-#   set.seed(seed[i])
-#   kol_1d_2var_better_random_sim_polish[[i]] <-
-#     animate_dist(boa6,
-#                  tour_path = guided_tour(norm_kol(), d = 1,
-#                                          search_f = search_polish, max.tries = 200),
-#                  rescale = FALSE, verbose = TRUE,
-#                  start = last_basis)
-#   
-# }
+(ref: kol-better-random-default) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
+`search_better_random` has a probablistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
 
-#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
-#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))
+```{r kol-better-random-tuned, fig.height= 6, fig.cap = "(ref:kol-better-random-tuned)"}
+```
 
-polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y)) 
-polish_tidy <- do.call(dplyr::bind_rows, polish_list)
-better_tidy <- do.call(dplyr::bind_rows, better_list)
-all_tidy <- bind_rows(better_tidy, polish_tidy)
-
-best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
-best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
-best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
-
-all_summary <- best$basis %>% 
-  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>% 
-  as_tibble() %>% 
-  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>% 
-  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>% 
-  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>% 
-  group_by(seed, IndexVal) %>% 
-  filter(abs(value) == max(abs(value))) %>% 
-  arrange(IndexVal) %>% ungroup()
-
-pca <- all_tidy %>% 
-  mutate(seed = as.factor(seed)) %>% 
-  left_join(all_summary %>% select(seed, VarFound, IndexVal), by = "seed") %>%
-  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>% 
-   bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                                        index = tourr::norm_kol(),
-                                        raw_data = boa6) %>%
-
-  compute_pca()
-
-random <- pca$aug %>% filter(method == "randomly_generated") %>% select(-seed)
-theoretical <- pca$aug %>% filter(info == "theoretical") %>% select(-seed)
-
-pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>% 
-  group_by(seed) %>% arrange(IndexVal) %>% 
-  ggplot(aes(x = PC1, y = PC2, col = VarFound)) + 
-  geom_point(data = random, col = "grey", size = 0.3) +
-  geom_point(aes(col = VarFound), alpha = 0.05) + 
-  geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) + 
-  geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) + 
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
-  facet_wrap(vars(fct_reorder(seed, IndexVal))) +
-  theme_void() + 
-  theme(aspect.ratio = 1, legend.position = "bottom") + 
-  scale_color_botanical(palette = "fern") 
-```
+(ref:kol-better-random-tuned) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
+This time, all the 20 incidents have found the global maximum.
 
@@
 
-```{r code, eval = FALSE}
-set.seed(123456)
-holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                           search_f =  search_better),
-                  rescale = FALSE, verbose = TRUE)
-
-#save(temp, file = here::here("vignettes", "data", "temp.rda"))
-```
-
-```{r glb-obj-1}
-holes_1d_better %>% 
-  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>% 
-  head(5) 
+```{r data-object, fig.cap = "(ref:data-object)"}
 ```
@@
 
-- `explore_trace_search()`: produce summary plots, as shown in Figure \ref{toy-search}
-- `explore_trace_interp()`: produce trace plots for the interpolation points, as shown in Figure \ref{toy-interp}
-- `explore_space_pca() `: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{toy-pca}.
-Animated version in Figure \ref{toy-pca-aniamted} can be acquired via the argument `animate = TRUE`
-- `explore_space_tour()`: produce animated tour view on the full space of the projection bases, as shown in Figure \ref{toy-tour}
+- `explore_trace_search()`: produce summary plots, as shown in Figure \@ref(fig:toy-search)
+- `explore_trace_interp()`: produce trace plots for the interpolation points, as shown in Figure \@ref(fig:toy-interp)
+- `explore_space_pca() `: produce plots of projection basis on the reduced space by PCA, as shown in Figure \@ref(fig:toy-pca).
+Animated version in Figure \@ref(fig:toy-pca-aniamted) can be acquired via the argument `animate = TRUE`
+- `explore_space_tour()`: produce animated tour view on the full space of the projection bases, as shown in Figure \@ref(toy-tour)
 

### 2020-10-09 paper/paper.tex
separate scripts from rmd, clean up repo

@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/tour_path_keynote/tour_path_keynote.001}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/img/tour_path}
 
@@
 
-\caption{\label{tour-path}Each square (frame) represents the projected data with a corresponding basis.
+\caption{Each square (frame) represents the projected data with a corresponding basis.
 Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames.}\label{fig:tour-path}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-search-1.pdf}
-\caption{\label{fig:toy-search}A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-search-1}
+
+}
+
+\caption{A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
 A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
-With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}
+With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}\label{fig:toy-search}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-interp-1.pdf}
-\caption{\label{fig:toy-interp}\label{toy-interp} The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-interp-1}
+
+}
+
+\caption{The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
 The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}\label{fig:toy-interp}
 \end{figure}
@@
 Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
-Figure \ref{toy-interp} presents the interpolation of three different tour paths.
+Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths.
 The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-pca-1.pdf}
-\caption{\label{fig:toy-pca}\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-pca-1}
+
+}
+
+\caption{PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}\label{fig:toy-pca}
 \end{figure}
@@
 
-Figure \ref{toy-pca} shows the first two principal components of two search paths.
+Figure \ref{fig:toy-pca} shows the first two principal components of two search paths.
 Starting from the same position, two searching methods take different paths to get its final bases.
@@
 
-\begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:unnamed-chunk-1}\label{toy-pca-aniamted}A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progrssion from the start to finish in each algorithm.}
-\end{figure}
-
 Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
-In Figure \ref{toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
+In Figure \ref{fig:toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
 
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/toy-tour-1.pdf}
-\caption{\label{fig:toy-tour}\label{toy-tour}A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-tour-1}
+
+}
+
+\caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
 The second frame on the top row view the space from a direction that is close to the one in PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}\label{fig:toy-tour}
 \end{figure}
@@
 We use the same dataset as the toy example above to explore the search function \texttt{search\_better} and we want to learn how the index value changes on the interpolation path for the \texttt{holes} index.
-From the left panel of Figure \ref{interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
+From the left panel of Figure \ref{fig:interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
 In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
 Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
-After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{interruption}.
+After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
 
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/interruption-1.pdf}
-\caption{\label{fig:interruption}\label{interruption}A comparison of the trace plot with and without the interruption.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/interruption-1}
+
+}
+
+\caption{A comparison of the trace plot with and without the interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.}
+The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.}\label{fig:interruption}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/geo-polish-1.pdf}
-\caption{\label{fig:geo-polish}\label{geo-polish}Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/polish-1}
+
+}
+
+\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
 The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.}
+The separation of the clusters on the y axis becomes sharper after the polish.}\label{fig:polish}
 \end{figure}
@@
 
-Figure \ref{geo-polish} presents the final projections found before and after applying \texttt{search\_polish} using the demo code.
+Figure \ref{fig:polish} presents the final projections found before and after applying \texttt{search\_polish} using the demo code.
 Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables, namely \texttt{V1}, \texttt{V8}, \texttt{V9} and \texttt{V10}.
-When appears in the projected data in Figure \ref{geo-polish}, polish works to sharpen the edges of each cluster.
+When appears in the projected data in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
 
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/kol-cdf-1.pdf}
-\caption{\label{fig:kol-cdf}\label{kol-cdf}One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+
+}
+
+\caption{One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
 The grey dashed line represents the index value of the theoreical best basis.
-\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}
+\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 
-Figure \ref{kol-cdf} compares the tracing plot of two algorithms: \texttt{search\_geodesic} and \texttt{search\_better}.
+Figure \ref{fig:noisy-better-geo} compares the tracing plot of two algorithms: \texttt{search\_geodesic} and \texttt{search\_better}.
 This time, the interpolated path is no longer smooth when using either algorithm.
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/pca-better-default-1.pdf}
-\caption{\label{fig:pca-better-default}\label{pca-better-default}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-default-1}
+
+}
+
+\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
 Each point represents a basis in the original 6D space, reduced into 2D by PCA.
 The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
-The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}
+The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better-default}
 \end{figure}
@@
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/pca-better-random-default-1.pdf}
-\caption{\label{fig:pca-better-random-default}\label{pca-better-random-default}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
-\texttt{search\_better\_random} has a probablistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.}
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-random-default-1}
+
+}
+
+\caption{(ref: kol-better-random-default)}\label{fig:kol-better-random-default}
 \end{figure}
 
+(ref: kol-better-random-default) One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
+\texttt{search\_better\_random} has a probablistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
+
 \begin{figure}
-\centering
-\includegraphics{paper_files/figure-latex/pca-better-random-tuned-1.pdf}
-\caption{\label{fig:pca-better-random-tuned}\label{pca-better-random-tuned}One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-random-tuned-1}
+
+}
+
+\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
 Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.
-This time, all the 20 incidents have found the global maximum.}
+This time, all the 20 incidents have found the global maximum.}\label{fig:kol-better-random-tuned}
 \end{figure}
@@
 \item
-  \texttt{explore\_trace\_search()}: produce summary plots, as shown in Figure \ref{toy-search}
+  \texttt{explore\_trace\_search()}: produce summary plots, as shown in Figure \ref{fig:toy-search}
 \item
-  \texttt{explore\_trace\_interp()}: produce trace plots for the interpolation points, as shown in Figure \ref{toy-interp}
+  \texttt{explore\_trace\_interp()}: produce trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
 \item
-  \texttt{explore\_space\_pca()}: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{toy-pca}.
-Animated version in Figure \ref{toy-pca-aniamted} can be acquired via the argument \texttt{animate\ =\ TRUE}
+  \texttt{explore\_space\_pca()}: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+Animated version in Figure \ref{fig:toy-pca-aniamted} can be acquired via the argument \texttt{animate\ =\ TRUE}
 \item

### 2020-10-09 scripts/examples.R

@@ -1,0 +1,85 @@
+## ----interruption
+p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>%
+  mutate(anno = c("current basis", "interpolated basis", "target basis"))
+
+p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") +
+  geom_point(data = p1_anno) +
+  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) +
+  theme(legend.position = "none")
+
+p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
+  theme(legend.position = "none")
+
+(p1 | p2) & scale_color_botanical(palette = "fern")
+
+## ----polish
+# set.seed(123456)
+# render(
+#   boa6,
+#   tour_path = guided_tour(holes(), d = 2, search_f = search_geodesic),
+#   dev = "png",
+#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   rescale = FALSE,
+#   frames = 100,
+#   file = here::here("anim","polish", "before%03d.png")
+# )
+#
+# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
+#
+# set.seed(123456)
+# render(
+#   boa6,
+#   tour_path =  guided_tour(holes(), d = 2, search_f = search_polish, alpha = 0.1),
+#   dev = "png",
+#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   start = last_basis,
+#   rescale = FALSE,
+#   frames = 100,
+#   file = here::here("anim","polish", "after%03d.png")
+# )
+
+# set.seed(123456)
+# holes_2d_geo <-
+#   animate_xy(boa6, tour_path = guided_tour(holes(), d = 2,
+#                                            search_f =  search_geodesic),
+#                            rescale = FALSE, verbose = TRUE)
+#
+# last_basis <- get_best(holes_2d_geo)$basis %>% .[[1]]
+#
+# set.seed(123456)
+# holes_2d_geo_polish <-
+#   animate_xy(boa6, tour_path = guided_tour(holes(), d = 2,
+#                                            search_f =  search_polish),
[…]

### 2020-10-09 scripts/implementation.R

@@ -1,0 +1,10 @@
+## ----data-object
+# set.seed(123456)
+# holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
+#                                                               search_f =  search_better),
+#                                 rescale = FALSE, verbose = TRUE)
+# save(holes_1d_better, file = here::here("data/holes_1d_better.rda"))
+
+holes_1d_better %>%
+  dplyr::select(id, basis, index_val, info, tries, loop, method, alpha) %>%
+  head(5)

### 2020-10-09 scripts/kol-better-default.R

@@ -1,0 +1,76 @@
+## ----kol-better-default
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_sim <- list()
+# kol_1d_2var_better_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   set.seed(seed[i])
+#   kol_1d_2var_better_sim[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE)
+#
+#
+#   last_basis <- get_best(kol_1d_2var_better_sim[[i]])$basis %>% .[[1]]
+#
+#   set.seed(1842)
+#   kol_1d_2var_better_sim_polish[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 100),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+# }
+
+
+# save(kol_1d_2var_better_sim, file = here::here("data", "kol_1d_2var_better_sim.rda"))
+# save(kol_1d_2var_better_sim_polish, file = here::here("data", "kol_1d_2var_better_sim_polish.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y))
+better_list <- purrr::map2(kol_1d_2var_better_sim,seed,  ~.x %>% dplyr::mutate(seed = .y))
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy_better <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(all_tidy_better, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>%
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>%
+  as_tibble() %>%
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>%
+  mutate(IndexVal = best$index_val, seed = as.factor(best$seed)) %>%
+  pivot_longer(-c(seed, IndexVal), names_to = "VarFound", values_to = "value") %>%
[…]

### 2020-10-09 scripts/kol-better-random-default.R

@@ -1,0 +1,75 @@
+## ----kol-better-random-default
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_random_sim2 <- list()
+# kol_1d_2var_better_random_sim_polish2 <- list()
+# for (i in 1: length(seed)){
+#   cat("i = ", i, "seed = ", seed[i], "\n")
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim2[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better_random, max.tries = 200),
+#                  rescale = FALSE, verbose = TRUE)
+#
+#
+#   last_basis <- get_best(kol_1d_2var_better_random_sim2[[i]])$basis %>% .[[1]]
+#
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim_polish2[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 200),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+#
+# }
+# save(kol_1d_2var_better_random_sim2, file = here::here("data", "kol_1d_2var_better_sim2.rda"))
+# save(kol_1d_2var_better_random_sim_polish2, file = here::here("data", "kol_1d_2var_better_sim_polish2.rda"))
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish2,seed,  ~.x %>% dplyr::mutate(seed = .y))
+better_list <- purrr::map2(kol_1d_2var_better_random_sim2, seed,  ~.x %>% dplyr ::mutate(seed = .y))
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>%
+  purrr::flatten_dbl() %>% matrix(ncol = 6, byrow = TRUE) %>%
+  as_tibble() %>%
+  rename("V7" = "V3", "V8" = "V4", "V9" = "V5", "V10" = "V6") %>%
[…]

### 2020-10-09 scripts/kol-better-random-tuned.R

@@ -1,0 +1,79 @@
+## ---- kol-better-random-tuned
+set.seed(123)
+seed <- sample.int(10000, 20)
+# kol_1d_2var_better_random_sim <- list()
+# kol_1d_2var_better_random_sim_polish <- list()
+# for (i in 1: length(seed)){
+#   cat("i = ", i, "seed = ", seed[i], "\n")
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_better_random, max.tries = 500, t0 = 1e-6, alpha = 0.7),
+#                  rescale = FALSE, verbose = TRUE)
+#
+#
+#   last_basis <- get_best(kol_1d_2var_better_random_sim[[i]])$basis %>% .[[1]]
+#
+#   set.seed(seed[i])
+#   kol_1d_2var_better_random_sim_polish[[i]] <-
+#     animate_dist(boa6,
+#                  tour_path = guided_tour(norm_kol(), d = 1,
+#                                          search_f = search_polish, max.tries = 200),
+#                  rescale = FALSE, verbose = TRUE,
+#                  start = last_basis)
+#
+# }
+
+
+#save(kol_1d_2var_better_random_sim, file = here::here("data", "kol_1d_2var_better_random_sim.rda"))
+#save(kol_1d_2var_better_random_sim_polish, file = here::here("data", "kol_1d_2var_better_random_sim_polish.rda"))
+
+
+polish_list <- purrr::map2(kol_1d_2var_better_random_sim_polish,seed,  ~.x %>% dplyr::mutate(seed = .y))
+better_list <- purrr::map2(kol_1d_2var_better_random_sim,seed,  ~.x %>% dplyr::mutate(seed = .y))
+polish_tidy <- do.call(dplyr::bind_rows, polish_list)
+better_tidy <- do.call(dplyr::bind_rows, better_list)
+all_tidy <- bind_rows(better_tidy, polish_tidy)
+
+best <- get_best(polish_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(better_tidy, group = seed) %>% arrange(index_val)
+best <- get_best(all_tidy, group = seed) %>% arrange(index_val)
+
+all_summary <- best$basis %>%
[…]

### 2020-10-09 scripts/vis-diag.R

@@ -1,6 +1,9 @@
 ## ---- load-pkg
 library(ferrn)
 library(tidyverse)
+library(tourr)
+library(patchwork)
+library(ggrepel)
 files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
@@ -11,3 +14,96 @@ bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500),
   scale_color_botanical(palette = "daisy") +
   facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
   xlab("iteration")
+
+## ---- toy-interp
+bind_rows(holes_2d_better_max_tries %>% mutate(group = "Algorithm 1"),
+          interrupt_no %>% mutate(group = "Algorithm 2"),
+          holes_2d_better_random %>% mutate(group = "Algorithm 3")) %>%
+  mutate(group = fct_relevel(group, c("Algorithm 1",
+                                      "Algorithm 2",
+                                      "Algorithm 3"))) %>%
+  explore_trace_interp(group = group) +
+  theme(legend.position = "none") +
+  scale_color_botanical(palette = "fern")
+
+## ---- toy-pca
+bind_rows(holes_1d_geo, holes_1d_better) %>%
+  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+                   index = tourr::holes(), raw_data = boa5) %>%
+  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5),
+                   index = tourr::holes(), raw_data  = boa5)%>%
+  explore_space_pca(col = method)  +
+  scale_color_botanical(palette = "cherry") +
+  theme(legend.position = "bottom")
+
+## ----toy-pca-animated
+# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>%
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                    index = tourr::holes(), raw_data = data) %>%
+#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5),
+#                    index = tourr::holes(), raw_data  = data)%>%
+#   explore_space_pca(col = method, animate = TRUE)  +
+#   scale_color_botanical(palette = "cherry")
+#
[…]

### 2020-10-09 source/animated_n_tour_plot.R

@@ -1,250 +1,0 @@
-library(tidyverse)
-library(geozoo)
-library(tourr)
-library(ferrn)
-library(gganimate)
-files <- paste0("data/", list.files(here::here("data")))
-purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
-pca <- rbind(holes_1d_better,
-             holes_1d_geo) %>%
-  compute_pca()
-
-pca$aug %>%
-  ggplot(aes(x = PC1, y = PC2, col = as.factor(info))) +
-  geom_point() +
-  theme(aspect.ratio = 1) +
-  facet_wrap(vars(method))
-
-# 1) only sphere
-set.seed(1)
-sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-dt <- sphere
-colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
-color <- c(rep(pal[1], nrow(sphere)))
-animate_xy(dt, col = color, cex_slice = 0.8)
-animate_slice(dt, col = color, cex_slice = 0.8)
-
-# 2) sphere and  holes_1d_geo
-path1 <- holes_1d_geo$basis %>% flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-dt <- rbind(sphere, path1)
-colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
-color <- c(rep(pal[1], nrow(sphere)),
-           rep(pal[2], nrow(path1)))
-animate_xy(dt, col = color, cex_slice = 0.8)
-animate_slice(dt, col = color, cex_slice = 0.8)
-
-# 3) sphere and holes_1d_better
-path2 <- holes_1d_better$basis %>% flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-dt <- rbind(sphere, path2)
-colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
-color <- c(rep(pal[1], nrow(sphere)),
-           rep(pal[3], nrow(path2)))
-animate_xy(dt, col = color)
-animate_slice(dt, col = color, cex_slice = 1, cex_other = 1)
-
[…]

### 2020-10-09 source/indice.r

@@ -1,26 +1,0 @@
-calc_kol <- function(data, proj, sphere = FALSE) {
-
-  if(sphere){
-    data <- sphere_data(rescale(data))
-  }
-
-  mat <- as.matrix(data) %*% proj
-
-  mat_bin_count <- ash::bin1(mat %>% scale(), c(min(mat), max(mat)), 100)$nc
-  norm_bin_count <- ash::bin1(rnorm(nrow(mat)) %>% scale(), c(min(mat), max(mat)), 100)$nc
-  diff <- sum((mat_bin_count - norm_bin_count)^2)/nrow(mat)
-
-  diff
-}
-
-calc_kol_cdf <- function(data, proj, sphere = FALSE) {
-
-  if(sphere){
-    data <- sphere_data(rescale(data))
-  }
-
-  mat <- as.matrix(data) %*% proj
-  norm <- rnorm(nrow(mat))
-
-  as.numeric(ks.test(mat, norm)$statistic)
-}

### 2020-10-09 source/my_test_file.R

@@ -1,131 +1,0 @@
-library(tidyverse)
-library(tourr)
-set.seed(1234)
-x1 <- rnorm(1000, 0, 1)
-x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
-x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
-x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
-x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
-x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
-x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
-x8 <- rnorm(1000, 0, 1)
-x9 <- rnorm(1000, 0, 1)
-x10 <- rnorm(1000, 0, 1)
-
-data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
-                       x9 = x9, x10 = x10) %>% purrr::map_df(scale)
-
-data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
-                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
-                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
-
-###############################
-devtools::load_all()
-set.seed(123456)
-holes_1d_geo <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                                   search_f =  tourr:::search_geodesic),
-                     rescale = FALSE, verbose = TRUE)
-
-devtools::load_all()
-set.seed(123456)
-holes_1d_geo_polish <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                                                  search_f =  tourr:::search_polish),
-                                    rescale = FALSE, verbose = TRUE, start = last_basis)
-
-
-
-devtools::load_all()
-set.seed(123456)
-holes_1d_better <- animate_dist(data, tour_path = guided_tour(holes(), d = 1,
-                                                           search_f =  tourr:::search_better),
-                             rescale = FALSE, verbose = TRUE)
-
-############
-devtools::load_all()
-set.seed(123456)
[…]

### 2020-10-09 source/paper.R

@@ -1,203 +1,0 @@
-library(tidyverse)
-library(tourr)
-library(ferrn)
-set.seed(1234)
-x1 <- rnorm(1000, 0, 1)
-x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
-x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
-x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
-x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
-x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
-x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
-x8 <- rnorm(1000, 0, 1)
-x9 <- rnorm(1000, 0, 1)
-x10 <- rnorm(1000, 0, 1)
-
-data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
-                       x9 = x9, x10 = x10) %>% scale() %>% as_tibble()
-
-data_mult <- tibble::tibble(x1 = x1, x2 = x2, x7 = x7,
-                            x8 = x8, x9 = x9, x10 = x10) %>% scale() %>% as_tibble()
-
-get_alpha <- function(dt){
-  last <- dt %>% filter(info == "interpolation") %>% tail(1) %>% pull(index_val) %>% .[[1]]
-  second_last <- dt %>% filter(info == "interpolation") %>% tail(2) %>% pull(index_val) %>% .[[1]]
-
-  last - second_last
-}
-
-save(data_mult, file = here::here("data", "data_mult.rda"))
-save(data, file = here::here("data", "data.rda"))
-
-
-################################################
-# polish example
-set.seed(123456)
-holes_2d_better <- animate_xy(data_mult, tour_path = guided_tour(holes(), d = 2,
-                                                                 search_f =  search_better),
-                              rescale = FALSE, verbose = TRUE)
-
-set.seed(123456)
-holes_2d_better_polish <- animate_xy(data_mult, tour_path = guided_tour(holes(), d = 2,
-                                                                        search_f =  search_polish,
-                                                                        alpha = get_alpha(holes_2d_better)),
[…]

### 2020-10-09 source/sim_data.R

@@ -1,20 +1,0 @@
-library(tidyverse)
-set.seed(1234)
-x1 <- rnorm(1000, 0, 1)
-x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
-x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
-x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
-x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
-x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
-x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
-x8 <- rnorm(1000, 0, 1)
-x9 <- rnorm(1000, 0, 1)
-x10 <- rnorm(1000, 0, 1)
-
-data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
-                       x9 = x9, x10 = x10) %>% purrr::map_df(scale)
-
-data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
-                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
-                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
-

### 2020-10-09 vis_method.R

@@ -1,81 +1,0 @@
-library(tidyverse)
-library(tourr)
-
-set.seed(1234)
-x1 <- rnorm(1000, 0, 1)
-x2 <- c(rnorm(500, -3, 1), rnorm(500, 3, 1))
-x8 <- rnorm(1000, 0, 1)
-x9 <- rnorm(1000, 0, 1)
-x10 <- rnorm(1000, 0, 1)
-
-
-data <- tibble::tibble(x1 = x1, x2 = x2, x8 = x8,
-                       x9 = x9, x10 = x10) %>% purrr::map_df(scale) %>% as.matrix()
-########################
-
-set.seed(1234)
-x1 <- rnorm(1000, 0, 1)
-x2 <- sample(c(rnorm(500, -3, 1), rnorm(500, 3, 1)), size = 1000)
-x3 <- sample(c(rep(-1, 500), rep(1, 500)), size = 1000)
-x4 <- sample(c(rnorm(250, -3, 1), rnorm(750, 3, 1)), size = 1000)
-x5 <- sample(c(rnorm(330, -5, 1), rnorm(340, 0, 1), rnorm(330, 5, 1)), size = 1000)
-x6 <- sample(c(rnorm(450, -5, 1), rnorm(100, 0, 1), rnorm(450, 5, 1)), size = 1000)
-x7 <- sample(c(rnorm(500, -5, 1), rnorm(500, 5, 1)), size = 1000)
-x8 <- rnorm(1000, 0, 1)
-x9 <- rnorm(1000, 0, 1)
-x10 <- rnorm(1000, 0, 1)
-
-data_mult <- tibble::tibble(x1 = x1, x2 = x2, x3 = x3,
-                            x4 = x4, x5 = x5, x6 = x6, x7 = x7,
-                            x8 = x8, x9 = x9, x10 = x10) %>% map_df(scale)
-
-########################
-set.seed(123456)
-holes_2d_geo <- animate_xy(data_mult[,c(1,2, 7:10)],
-                           tour_path = guided_tour(holes(), d = 2, search_f = tourr:::search_geodesic),
-                           rescale = FALSE, verbose = TRUE)
-
-last_basis <- holes_2d_geo %>% filter(info == "interpolation") %>% tail(1) %>% pull(basis) %>% .[[1]]
-
-set.seed(123456)
-holes_2d_geo_polish <- animate_xy(data_mult[,c(1,2, 7:10)],
[…]

### 2020-10-10 paper/paper.Rmd
fix pca flip sign

@@
                       root.dir = here::here(), fig.path = here::here("figs/"), 
-                      fig.align = 'center')
+                      fig.align = 'center', cache = TRUE, cache.path = 'cache/')
 ```
@@
 
-```{r toy-pca-aniamted, fig.cap = "(ref:toy-pca-aniamted)"}
+```{r toy-pca-animated, fig.cap = "(ref:toy-pca-animated)"}
 ```
 
-(ref:toy-pca-aniamted) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.
+(ref:toy-pca-animated) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.
 
@@
 
-Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+Compare to the PCA plot, the animated rotation (tour) display, in Figure \@ref(fig:toy-tour), gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
 This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.

### 2020-10-10 paper/paper.tex
fix pca flip sign

@@
 
+\begin{figure}
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+
+}
+
+\caption{A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\end{figure}
+
 Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
@@
 
-Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
+Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{fig:toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
 This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.

### 2020-10-10 scripts/kol-better-default.R

@@ -68,7 +68,7 @@ pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>%
   geom_point(aes(col = VarFound), alpha = 0.1) +
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) +
   geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) +
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
+  geom_point(data = pca$aug %>% get_best(group = seed) %>% filter(method != "theoretical"), size = 5) +
   geom_point(data = pca$aug %>% get_start() %>% filter(method == "search_better_random"), size = 3) +
   facet_wrap(vars(fct_reorder(seed, IndexVal))) +
   theme_void() +

### 2020-10-10 scripts/kol-better-random-default.R

@@ -68,7 +68,7 @@ pca$aug %>% filter(method != "randomly_generated"& info != "theoretical") %>%
   geom_point(aes(col = VarFound), alpha = 0.05) +
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) +
   geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) +
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
+  geom_point(data = pca$aug %>% get_best(group = seed) %>% filter(method != "theoretical"), size = 5) +
   facet_wrap(vars(fct_reorder(seed, IndexVal))) +
   theme_void() +
   theme(aspect.ratio = 1, legend.position = "bottom") +

### 2020-10-10 scripts/kol-better-random-tuned.R

@@ -72,7 +72,7 @@ pca$aug %>% filter(method != "randomly_generated" & info != "theoretical") %>%
   geom_point(aes(col = VarFound), alpha = 0.05) +
   geom_point(data = theoretical, size = 5, col = botanical_palettes$cherry[3]) +
   geom_path(data = pca$aug %>% filter(info == "interpolation"), size = 2) +
-  geom_point(data = pca$aug %>% get_best(group = seed), size = 5) +
+  geom_point(data = pca$aug %>% get_best(group = seed) %>% filter(method != "theoretical"), size = 5) +
   facet_wrap(vars(fct_reorder(seed, IndexVal))) +
   theme_void() +
   theme(aspect.ratio = 1, legend.position = "bottom") +

### 2020-10-10 scripts/vis-diag.R

@@ -2,6 +2,7 @@
 library(ferrn)
 library(tidyverse)
 library(tourr)
+library(gganimate)
 library(patchwork)
 library(ggrepel)
 files <- paste0("data/", list.files(here::here("data")))
@@ -27,69 +28,77 @@ bind_rows(holes_2d_better_max_tries %>% mutate(group = "Algorithm 1"),
   scale_color_botanical(palette = "fern")
 
 ## ---- toy-pca
-bind_rows(holes_1d_geo, holes_1d_better) %>%
+pca <- bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5) %>%
-  bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5),
-                   index = tourr::holes(), raw_data  = boa5)%>%
-  explore_space_pca(col = method)  +
+  compute_pca(group = method)
+pca$aug %>%
+  ggplot(aes(x = PC1, y = PC2)) +
+  geom_point(data = pca$aug %>% filter(info == "randomly_generated"), col = "grey", size = 0.1) +
+  #geom_path(data =  pca$aug %>% filter(info == "interpolation"), aes(col = method), size = 2) +
+  geom_point(data =  pca$aug %>% filter(info != "randomly_generated"), aes(col = method)) +
+  geom_point(data = pca$aug %>% filter(info == "theoretical"), aes(col = method), size = 10) +
+  geom_point(data = get_start(pca$aug), aes(col = method), size = 5) +
   scale_color_botanical(palette = "cherry") +
-  theme(legend.position = "bottom")
+  theme_void() +
+  theme(aspect.ratio = 1, legend.position =  "bottom")
 
 ## ----toy-pca-animated
-# ani <- bind_rows(holes_1d_geo, holes_1d_better) %>%
-#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-#                    index = tourr::holes(), raw_data = data) %>%
-#   bind_theoretical(matrix(c(0, -1, 0, 0,0), nrow = 5),
-#                    index = tourr::holes(), raw_data  = data)%>%
-#   explore_space_pca(col = method, animate = TRUE)  +
-#   scale_color_botanical(palette = "cherry")
[…]

### 2020-10-13 paper/paper.Rmd
tuning languages here and there and everywhere

--- 2020-10-09 paper/paper.Rmd
@@
 - R
-abstract: Friedman & Tukey commented on their initial paper on projection pursuit in 1974 that "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
-In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
-An R package, ferrn, has been created to implement this diagnostic system.
+abstract: Friedman & Tukey commented on their initial paper on projection pursuit in 1974 that "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedures.
+In this paper, we illustrates the diagnostics of optimisers for projection pursuit guided tour via the construction of a data object and producing diagnostic plots.
+These diagnostics workflow can be applied generally on optimisers to assess its performance.
+An R package, ferrn, has been created to implement the diagnostics.
 bibliography: biblio.bib
@@
 library(knitr)
+library(tidyverse)
+library(ferrn)
 read_chunk(here::here('scripts/vis-diag.R'))
 read_chunk(here::here('scripts/examples.R'))
-read_chunk(here::here('scripts/kol-better-default.R'))
-read_chunk(here::here('scripts/kol-better-random-default.R'))
-read_chunk(here::here('scripts/kol-better-random-tuned.R'))
+read_chunk(here::here('scripts/kol_result.R'))
 read_chunk(here::here('scripts/implementation.R'))
@@
 
-<!-- We can see from Figure \ref{tour-path} that only the data projected using the accepted and interpolated bases are animated by tour; none of the searching bases sampled by the optimsiation algorithms are presented in the tour animation.
-Although these algorithms have been demonstrated to find the global maximum in various literature as we cited above, they focus solely on finding the global maximum and fewer tools are available to explore how each algorithm searches the parameter space and compare the results between different algorithms.
--->
-
-<!-- This study is meaningful because **as more complex index functions are proposed based on statistical theory, only computationally effective optimisation algorithms allow us to find the projection basis that presents interesting structure in the data**.
--->
-
-<!-- Hence the study **presents a gallery of diagnostic plots applicable to optimisation algorithms.** -->
-
 Visualisation has been widely used in exploratory data analysis.
 Presenting information in a graphical format often allows people to see information they would otherwise not see.
-This motivates our work of creating plots to diagnose optimisation algorithms in the context of projection pursuit guided tour, with the aim to understand and compare features of different existing algorithms.
+This motivates our work of creating plots to diagnose optimisation algorithms with the aim to understand and compare features of different algorithms.
 
-In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
+In an optimization problem the goal is to find the best solution within the space of all feasible solutions, which typically is represented by a set of constraints.
 The problem consists in optimizing an objective function $f: S \rightarrow \mathbb{R}$ with $S \in \mathbb{R}^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.
 
-Projection pursuit and guided tour are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
+The problem context of the optimisation that has been worked in this paper is projection pursuit guided tour, which are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
 Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
@@
 The remainder of the paper is organised as follows. 
-Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour. 
-Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples. 
-Section \ref{application} shows the application of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
+Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods. 
+Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem and its features, and introduces three main existing algorithms.
+Section \ref{vis-diag} presents new visual diagnostics designs. Data object is first defined to pinned down the data structure, several diagnostic plots are then defined with sample examples given.
+Section \ref{application} shows the application of how the diagnostic plots designed previously can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
 Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
@@
 
-Given an optimisation problem, two basic approaches find the optimum based on different thinking.
-An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the closed-form solution may not be available when the problem starts to become complex.
-An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution.
-An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.
-
-A traditional while often used in practice is an iterative method called *line search method* [@fletcher2013practical].
-In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
+Given an optimisation problem, two broad approaches find the optimum based on different thinking and hence have its own strength and deficits.
+An analytical approach aims to find the optimal solution in a finite number of steps, but the closed-form solution may not be available when the problem starts to become complex. 
+An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution and researchers can decide when to stop if a desirable accuracy has been achieved.
+However, a potential issue is that it may end up finding a local optimum.
+In this paper, the main focus is on a particular iterative method:  *line search method* [@fletcher2013practical].
+In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search conducts its search via an iterative up as in Equation \ref{eq:line-search}.
 
@@
 where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
-Strictly speaking, $\alpha_k$ is chosen by another minimisation of $f(x^{(j)} + \alpha* d^{(j)})$ with respect to $\alpha$ and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of $\alpha_j$ is attained [@curry1944method].
-In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation.
-A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$ and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
+Theoretical results have demonstrated the global convergence of the algorithm when $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
+However, in practice, this is rarely implemented due to its computational demanding or sometimes the existence of such a minimisation[@fletcher2013practical].
+A modified approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$ and this approach turns out to be efficient in practical problems despite the lost of guarantee on global convergence.
 
@@
 Modern development of the line search methods focuses on proposing different computation on the searching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems.
-The specific problem context we are interested in is called projection pursuit guided tour.
-Projection pursuit and guided tour are two separate methods in exporatory data analysis focusing on different aspects: coined by @friedman1974projection, projection pursuit detects interesting structures (i.e.
-clustering, outliers and skewness) in multivariate data via low dimensions projection; whilst guided tour is a particular variation in a broader class of data visualisation method called tour.
+The specific problem context discussed in this paper is projection pursuit guided tour.
+Projection pursuit and guided tour are two separate methods in exploratory data analysis focusing on different aspects: Projection pursuit, coined by @friedman1974projection, detects interesting structures (i.e.
+clustering, outliers and skewness) in multivariate data via low dimensions projection whilst guided tour is one particular variation in a broader class of data visualisation method, tour.
 
-Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
+Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
 Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
@@
 
-In their initial paper, @friedman1974projection noted that "..., the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
-While we leave the formal construction of the optimisation problem and existing algorithms to section \ref{tour-optim}, we outline the general idea here.
-Given a random starting (current) basis, projection pursuit repeatedly searches for candidate bases nearby until it finds one with higher index value than the current basis.
-In the second round, that basis becomes the current basis and the repetitive sampling  continues.
-The process ends until no better basis can be found or one of the termination criteria is reached.
-
-Before introducing the guided tour, we shall be familiar with the general tour method [@cook2008grand].
-A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
-The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details.
-Iteratively choosing different low dimension planes and interpolating between them forms a tour path.
-Different types of tour methods choose the low dimensional planes differently and we mention two other types of tour that are commonly used.
-A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data.
-Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.
-
-Guided tour chooses the planes produced by optimising the projection pursuit index function.
-Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, which are the interpolations between two blue frames.
-The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+As a general visualisation method, A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
+Different tours choose the low dimensional planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly in the high dimensional space and manual tour[@cook1997manual] chooses a sequence of bases that gradually phase in and out one variable to view the contribution of that variable in the visualisation.
+Guided tour, the main interest of this paper, chooses its projection with the aid of projection pursuit.
+Given a random start, projection pursuit iteratively finds bases with higher index value and guided tour constructs geodesic interpolation between two consecutive to form a tour path.
+Mathematical details of the geodesic interpolation can be found in @buja2005computational.
+Figure \@ref(fig:tour-path) shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, the interpolations between two blue frames.
+The tour method has been implemented in the `tourr` package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
 
@@
 (ref:tour-path) Each square (frame) represents the projected data with a corresponding basis.
-Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames.
-
-<!-- The pseudo-code below illustrates the implementation of guided tour in the tourr package.
-Given an projection pursuit index function and a randomly generated projection basis (current basis), the optimisation procedure produces a target basis inside `generator()`  Both the current basis and the target basis will be supplied to `tour_path()` to prepare information needed for constructing a geodesic path.
-This information is then used to compute a series of interpolating bases inside the `tour()` function.
-All the basis will be sent to create animation for visualising the tour in the `animate()` function.
--->
-
-<!-- ```{r algo,eval = FALSE, echo = TRUE} -->
-<!-- animation <- function(){ -->
-
-<!--   # compute projection basis  -->
-<!--   tour <- function(){ -->
-
-<!--     # construct bases on the tour path -->
-<!--     new_geodesic_path <- function(){ -->
-<!--       tour_path <- function(){ -->
-
-<!--         # GENERATOR: generate projection basis via projection pursuit -->
-<!--         guided_tour <- function(){ -->
-<!--           generator <- function(){ -->
-
-<!--             # define projection pursuit index -->
-<!--             # generate the target basis from the current basis via optimisation -->
-<!--           } -->
-<!--         } -->
-
-<!--         # prepare geodesic information needed for interpolating along the tour path -->
-<!--       } -->
-<!--     } -->
-
-<!--     # INTERPOLATOR: interpolate between the current and target basis  -->
-<!--     function(){ -->
-<!--       # generate interpolating bases on the geodesic path -->
-<!--     } -->
-<!--   } -->
-
-<!--   # animate according to different display methods -->
-<!-- } -->
-<!-- ``` -->
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.
 
@@
 
-Now we begin to formulate the optimisation problem.
+The optimisation problem is stated as follows.
 Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
@@
 
-where $I_d$ is the d-dimensional identity matrix. The constraint requires the projection basis $\mathbf{A}$ to be an orthogonal matrix with each column vector being orthonormal.
+where $I_d$ is the d-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be an orthogonal matrices.
 
-There are several features of this optimisation that are worth noticing.
-First of all, it is a multivariate constraint optimisation problem.
-Since the decision variables are the entries of a projection basis, it is required to be orthonormal.
-It is also likely that the objective function is non-differentiable or the gradient information is simply not available.
-In this case, we will need to either use some approximation of the gradient or turn to derivative free methods.
-Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem needs to be able to find the global maximum.
-Along the way, local maximum may also be of our interest since they could present unexpected interesting projections.
-There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.
-
-<!-- - *constraint optimisation*: our problem requires the project basis to be orthonormal -->
-
-<!-- - *Multivariate optimisation*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional.
-Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.
--->
-
-<!-- - *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible.
-Stochastic optimisation method could be an alternative to solve these problems.
-(derivative information is not available) -->
-
-<!-- - *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large.
-Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.
-(the exact $\alpha$ that minimised $f(x^{(k)}+ \alpha d^{(k)})$ can be hard to find or sometiems even impossible, imposing certain decrease, how every this doesn't guarantee global convergence) -->
-
-<!-- - *computation speed*: The optimisation procedure needs to be fast to compute since the tour produces real-time animation of the projected data. -->
-
-<!-- - *consistency result in stochastic optimisation*: In stochastic algorithm, researchers usually set a seed to ensure the algorithm produces the same result for every run.
-This practice supports reproducibility, while less efforts has been made to guarantee different seeds will provide the same result.
--->
-
-<!-- - *role of interpolation in PP optimisation*: An optimisation procedure usually involves iteratively finding projection bases that maximise the index function, while tour requires geodesic interpolation between these bases to produce a continuous view for the users.
-It would be interesting to see if the interpolated bases could, in reverse, help the optimisation reach faster convergence.
--->
-
-<!-- the interpolation can be seen as giving information on the direction of the derivative. -->
-
-<!-- *Think about how does your package help people to understand optimisation* -->
-
-<!--  - diagnostic on stochastic optim -->
-<!--  - vis the progression of multi-parameter decision variable  -->
-<!--  - understanding learning rate - neighbourhood parameter -->
-<!--  - understand where the local & global maximum is found - trace plot - see if noisy function -->
+Several features of this optimisation are worth noticing.
+First of all, this is a constraint optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
+It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
+Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
+Lastly, there is also one computational consideration: the optimisation procedure needs to be fast to compute since the tour animation is played in real-time.
 
@@
 
-Below we introduce three possible algorithms: `search_better` and `search_better_random` are derivative free methods that sample candidate bases in the neighbourhood, whilst `search_geodesic` is an analogue of gradient ascent on the manifold of the projection basis.
+Below we introduce three possible algorithms: `search_better`, `search_better_random`, and `search_geodesic`.
+The first two are derivative free methods that sample candidate bases in the neighbourhood whilst `search_geodesic` is an analogue of gradient ascent on the projection basis space.
 
@@
   \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonise $\mathbf{A}_{l}$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
     Compute $I_{l}  = f(\mathbf{A}_{l})$\;
@@
 The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$.
-Another termination condition is when the maximum number of iteration $l_{\max}$ is reached.
+The termination condition is when the maximum number of iteration $l_{\max}$ is reached.
+The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
 A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$.
 Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast.
-The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
-[mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search]
 
@@
 
-Simulated annealing [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme $T(l)$.
+Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
 Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
 This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.
-a one dimensional projection of the dataset with two informative variables).
+This feature is particularly useful when local maximum is presented.
 The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
@@
 
-@cook1995grand used a gradient ascent algorithm on the manifold of the projection bases.
-In gradient ascent, one first find the direction for improvment via computing the gradient information.
+@cook1995grand used a gradient ascent algorithm on the space of the projection bases.
+In gradient ascent, one first find the direction for improvement via computing the gradient information.
 In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
-A geodesic is then constructed using the current basis and one of the $2n$ bases that have the highest index value.
-If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
-In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
-In `search_geodesic`, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from $-\pi/4$ to $\pi/4$.
-The optima $\mathbf{A}_{**}$ is outputted for the current iteration if the percentage change in the index value between $\mathbf{A}_{**}$ and $\mathbf{A}_{\text{cur}}$ is greater than a threshold value.
-As above, another termination condition is when $l_{\max}$ is reached.
+A geodesic is then constructed using the current basis and the one in $2n$ bases with the highest index value.
+If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and works as the searching direction.
+The next step in gradient ascent is to conduct a line search to find the best improvement along the gradient direction and in `search_geodesic`, this is replaced by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the termination condition on percentage improvement.
+The procedure will also terminate if $l_{\max}$ is reached.
 Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
@@
 
-To be able to make diagnostic plot, the optimisation algorithm should populate a data structure that contains the key elements of the algorithm.
-When the algorithm runs, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object for future analysis.
+To be able to make diagnostics on the optimisers, the algorithms need to populate a data structure with key elements of the algorithm.
+When the algorithms run, key information regarding the decision variable, objective function and hyper-parameters, needs to be recorded and stored as a data object so that it is ready to be supplied to the plotting functions for diagnostics.
 
@@
 
-<!-- [Do I need this paragraph to details the grammar of graphic? xxx - don't think so; not particularly relevant - assume most people know it roughly] -->
-<!-- A layer includes 1) the data that powers the plot; 2) a geometric object that represents the visual shape of the data and 3) relevant statistical transformation that transform the information in the data to the information used to draw the geometric object.
-An important concept in the grammar of graphic is *aesthetic mapping*.
-Aesthetic mapping links the variable in a dataset to information needed to produce a geometric object.
-For example, we map one variable on the x-axis and another on the y-axis to create a scatterplot.
-Another example of creating a boxplot involves 1) mapping one variable on the x-axis and 2) transforming another variable to its five-point-summary and mapping the five-point-summary to the y-axis.
--->
+In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
+Possible values for `search_better` and `search_better_random` includes `random_search`,  `new_basis`, and `interpolation`.
+`search_geodesic` has a wider variety that includes `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
 
-In the optimisation algorithms for projection pursuit, three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
 Multiple iterators are also  needed to index the data collected at different levels.
-$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts.
-$l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
-There are other parameters that are of our interest and we denote them as *$V_{p}$*.
-In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
+$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
+$l$ is the counter for each search/interpolation, which provides the information of how many basis the algorithm has searched before finding one to return.
+There are other parameters of interest, depends on the particular problem content and they are denoted as *$V_{p}$*.
+Two most common examples include $V_1 = \text{method}$, which tags the name of the algorithm used, and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
 A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
@@
 where $T^{\prime} = T + k_{J}+ l_{J+1}$.
-Note that we deliberately denote the last round of search as $j = J+1$ and in that round there is no output/interpolation basis and the algorithm terminates.
-This notation allows us to denote the last complete search-and-interpolate round as round $J$ and hence the final basis is $A_T$ and highest index value found is $I_T$.
+Note that there is no output in iteration $J + 1$ since the optimiser can't find a better basis and the algorithm terminates.
+In this notation, final basis found is $A_T$ with the highest index value $I_T$.
 
@@
 
-It is worth noticing that the data structure constructed above meets the tidy data principle [@wickham2014tidy] that states
+The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation forms a row and each variable forms a column.
+With tidy data structure, data wrangling and visualisation have been significantly simplified by the well-developed packages like dplyr[@dplyr] and ggplot2[@ggplot2].
 
-1) each observation forms a row, 
-2) each variable forms a column, and 
-3) each type of observational unit forms a table
-
-The wrangling and visualisation of tidy data have been greatly simplified by the well-known dplyr[@dplyr] and ggplot2[@ggplot2] package.
-
-With a constructed data object, the construction of diagnostic plots is inspired by the concept of grammar of graphic [@wickham2010layered], which powers the primary graphical system in R, ggplot2 [@ggplot2].
+The construction of diagnostic plots uses the concept of grammar of graphic [@wickham2010layered] in ggplot2.
 In grammar of graphic, plots are not defined by its appearance (i.e.
-boxplot, histogram, scatter plot etc) but by "stacked layers".
-Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
-Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
-
-<!-- The plot design of visual diagnostics depends on the characteristics of the variables to plot.
-If the searching points are of interest [point geom is not good; need summarise.
-While updating points are more manageable -> point geom].
-However, one must realise that with hundreds or thousands of searching points, exploring the sample space in animation could be slow and this is because of the time it takes to render hundreds point.
-[stratefy may help?
-- may need more work here.] -->
-
-<!-- ### Explore the value of objective function {#static} -->
-
-<!-- In tour, all the points recorded in the global object can be roughly divided into two broad categories: searching points and interpolating points -->
-
-<!-- - *Searching points* include the observations recorded in the searching algorithm to find the target basis.
-The points for target bases is also included in the searching points and there is one such point per `tries`.
--->
-
-<!-- - *interpolating points* exist in the guided tour to produce continuous animated view from one target basis to another and it doesn't have `loop` value. -->
-
-<!-- In the current implementation of the `tourr` package, while the target basis generated by the projection pursuit can be accessed later via `save_history()`, interpolating bases and those randomly nearby bases generated in the optimisation are not stored.
-This creates difficulties for fully understand the behaviour of the optimisation and interpolation of tour in complex scenario **[need a rephrase this part].**  -->
+boxplot, histogram, scatter plot etc) but by stacked layers.
+In the construction of diagnostic plots, there are multiple elements one may wants to emphasize and there's no single plot name that would meet the need, the stacked layers concept, on the other hand, allow information to be overlaid on the each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
 
@@
 
-A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum.
-One way of doing it is to plot the index value across its natural ordering $t$, however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
-When plotting the searching observations as points in a plot, the space each iteration takes will be proportional to the number of points in that iteration.
+A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum.
+Directly plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes longer for an optimiser to find a better basis towards the end.
+Another option is to use summarisation for each iteration.
+Boxplot is a suitable candidate that can provide five points summary of each iteration.
+Other additional information not presented in the  boxplot and then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
+Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a `cutoff` parameter.
 
-<!-- different iterations may have different number of points and, towards the end of the search there could easily be hundreds of bases being tested before the target basis is found.
-In the plot, points from those iterations towards the end will occupy the vast majority of the plot space.
-This motivates to use summarisation.
-Rather than knowing the index value of *every* basis, we are more interested to have a general summary of all the index value in that iteration and more importantly, the basis with the largest index value (since it prescribes the next geodesic interpolation and future searches).
--->
-
-<!-- .
-We may risk losing information on how many points it takes to find the target basis by displaying the boxplot alone for all `tries`.
-Thus, the number of point in each iteration is displayed at the bottom of each box and we provide options to switch iteration with small number of points to a point geometry, which is achieved via a `cutoff` argument.
-A line geometry is also added to link the points with the largest index value in each iteration.
-This helps to visualise the improvement made in each iteration.
-Using the concept of *gramma of graphics* [@wickham2010layered], the plot for exploring index value can be defined in three layers as: -->
-
-Another option is to use summarisation for each iteration.
-Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis.
-This could be remedied by adding more layers using the concept of grammar of graphics.
-A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration.
-Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
-This is achieved via a `cutoff` parameter.
-
-Figure \@ref(fig:toy-search) presents a comparison between the two plotting options discussed above.
-Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
-The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.
+Figure \@ref(fig:toy-search) shows a sample of the plot constructed using `search_better` with different value on the parameter `max_tries`.
+Comparing the returned index value of each iteration shows that the termination at `max_tries = 25` is not sufficient for `search_better` to explore the parameter space and a value of 500 is preferred over 25 in this context.
 
@@
 (ref:toy-search) A comparison of `search_better` with different parameter value on `max_tries`.
-A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
-With `max_tries = 500` specification, the algorithm is working much harder towards to the end.
+A six variable dataset `boa6` is used with `holes` index on a 2D problem.
+A `max_tries = 500` since it allows the optimiser to find better basis with higher index value at iteration six.
 
-## Examining the optimisation progress
+## Examining the optimisation progress {#toy-interp}
 
@@
 
-(ref:toy-interp) The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+(ref:toy-interp) The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
 The color represents the number of iteration.
@@
 
-Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
-Figure \@ref(fig:toy-interp) presents the interpolation of three different tour paths.
-The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
-While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
-The lower plot is constructed using `search_better_random`, where a basis with smaller index value  has a probabilistic chance of being accepted and we can observe a much more involved pattern.
-In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
-When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
-While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
+Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
+Figure \@ref(fig:toy-interp) presents the interpolation of three different tour paths each with different curvature.
+The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
+The middle path has an increase-then-decreases pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
+The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost cases is a deliberate construction of `search_better_random` where an inferior basis can be accepted in a probability manner so as to avoid trapping in the local maximum.
 
@@
 
-(ref:toy-pca) PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search
+(ref:toy-pca) 1D projection on the 5 variable dataset `boa5` with two optimisers: `search_better` and `search_geodesic`.
+The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
+The underlying grey points are randomly generated on the 5D space and reduced to 2D via PCA along with all the search points presented.
+The enlarged color points that colored as the interpolation are the starting points of each algorithm.
+They are initially simulated with the same starting points but all the bases in `search_geodesic` has been flipped positive to ensure that bases with same projected images but a sign difference are represented by the same point in the plot.
 
 Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go above five (5).
-Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d > 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.
-Principal component analysis.
-To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
+Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go over two or three.
+While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (d > 3), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.
+principal component analysis.
+To better perceive the the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
 
-<!-- We first flat each projection bases $\mathbf{A}$ to a row vector and stack them by rows.
-This gives a matrix of dimension $n$ by $(p*d)$.
-Centering and scaling are usually done before performing PCA and we denotes the pre-processed matrix as $\tilde{\mathbf{A}}_{n \times (p \times d)}$.
-By definition, the principal components maximise the variance of the projection bases in the reduced space.
-Using lagrange multipler, this is equivalent to find the eigenvector of the variance-covariance matrix: $\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}}$.
-Decomposing $\tilde{\mathbf{A}}$ using singular value decomposition (SVD) gives $$\tilde{\mathbf{A}} = \mathbf{U} \mathbf{\lambda} \mathbf{V}$$ and hence the variance-covariance matrix -->
-
-<!-- $$\Sigma = \tilde{\mathbf{A}}^T\tilde{\mathbf{A}} = (\mathbf{V}^T \mathbf{\lambda}^T \mathbf{U}^T) (\mathbf{U} \mathbf{\lambda} \mathbf{V}) = \mathbf{V}^T \mathbf{\lambda}^2 \mathbf{V}$$ -->
-
-<!-- has eigenvector $\mathbf{V} = [\mathbf{v}_1, \mathbf{v}_2, \cdots, \mathbf{v}_{p\times d}]$.
-The projection mapping to the x- and y-axis uses the first two principal components and hence $$\tilde{\mathbf{A}} \mathbf{v}_1 \text{ and } \tilde{\mathbf{A}} \mathbf{v}_2$$ -->
-
-<!-- The resulting two vectors are of dimension $n \times 1$, which is compatible with the dimension of the global object.
-THis means all the iterators and interesting other variables defined in the global object can be bound with the these two projections and can be mapped to other aesthetics (i.e.
-colour).
--->
-
-Figure \@ref(fig:toy-pca) shows the first two principal components of two search paths.
-Starting from the same position, two searching methods take different paths to get its final bases.
-There are two theoretical bases because they correspond to the same data projection $Y$ with a 180 degree rotation.
-What differentiate the two optimisers is that `search_better()` also involve random evaluation of points in the sphere during its sampling process while `search_geodesic()` doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
+Figure \@ref(fig:toy-pca) plots the first two principal components of two search paths, one using `search_better` and another uses `search_geodesic`.
+The search in the space reduced by PCA matches with the optimiser description before where the random sampling in `search_better` is broader, controlled by `alpha` parameter, which is default to 0.5 while the directional search in `search_geodesic`, controlled by `delta` with a default of 0.01, is so tiny that it can barely seen.
 
@@
 
-```{r toy-pca-aniamted, fig.cap = "(ref:toy-pca-aniamted)"}
+```{r toy-pca-animated, fig.cap = "(ref:toy-pca-animated)"}
 ```
 
-(ref:toy-pca-aniamted) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.
+(ref:toy-pca-animated) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.
 
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
-In Figure \@ref(fig:toy-pca-animated), There should actually be more writing here to and also the this examples may need to be changed.
+Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
+Figure \@ref(fig:toy-pca-animated) shows six frames of an animated version of Figure \@ref(fig:toy-pca) and this time, it shows that `search_better` finds the optimum quicker than `search_geodesic`.
 
@@
 
-Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a stereoscopic view of the search.
-To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation.
-This can be done via a regular grand tour or a slice tour [@laa2020slice], which emphasizes the points that are close to a section of the sphere.
-
-Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
-This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
-Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
+Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a more holistic stereoscopical view of the search.
+To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
+Compare to the PCA plot, the animated rotation (tour) display in Figure \@ref(fig:toy-tour) gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and grand tour produces a sequence of angle that view the search from different direction.
+As an evidence, the last frame in Figure \@ref(fig:toy-tour) is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to one in Figure \@ref(fig:toy-pca).
 
@@
 If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are presented, `search_better` may stuck in the local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
-In this section, examples will be presented  to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.
+When multiple optima are presented, `search_better` may stuck at the local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+In this section, examples will be presented  to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
 
@@
 
-## A problem of not monotonic
+## A problem of not monotonic {#monotonic}
 
-We use the same dataset as the toy example above to explore the search function `search_better` and we want to learn how the index value changes on the interpolation path for the `holes` index.
-From the left panel of Figure \@ref(fig:interruption), we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
-In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
-Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
+In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
+This pattern is undesirable since the optimiser could have start the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
+This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
 After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \@ref(fig:interruption).
@@
 
-(ref:interruption) A comparison of the trace plot with and without the interruption.
+(ref:interruption) Two-D projection on `boa6` data with holes index optimised by `search_better`.
+A comparison of the trace plot with and without the interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.
+The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.
 
@@
 Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
-This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
+This motivates the polish search where the final basis is supplied as the start of a new guided tour to search for any local breakthrough.
 
 Similar to `search_better` as a stochastic random search, `search_polish` has a different scheme of reducing the search neighbourhood.
-In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
+In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by the cooling parameter only after an iteration finishes.
 On the contrary, `search_polish` allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
-Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries`.
+Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries` limit.
 These include:
@@
 
-Figure \@ref(fig:polish) presents the final projections found before and after applying `search_polish` using the demo code.
-Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables, namely `V1`, `V8`, `V9` and `V10`.
-When appears in the projected data in Figure \@ref(fig:polish), polish works to sharpen the edges of each cluster.
+Figure \@ref(fig:polish) presents the final projections found before and after applying `search_polish` on `search_geodesic`.
+Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables.
+In terms of the projected data as in Figure \@ref(fig:polish), polish works to sharpen the edges of each cluster.
 
@@
 (ref:noisy-better-geo) One-D projection on `boa5` data with noisy index `norm_kol` optimised by `search_geodesic` and `search_better`.
-The grey dashed line represents the index value of the theoreical best basis.
-`search_geodesic` fails to optimse the noisy index while `search_better` has made reasonable improvments to reach the index value close to the one of theoretical best basis.
+The grey dashed line represents the index value of the theoretical best basis.
+`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of theoretical best basis.
 
-Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
-`norm_kol`, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{P}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
+The index function, up until this point, are all smooth, while this is not the case for all the index functions.
+`norm_kol`, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{Y}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
 Denotes the ECDF function as $F_{.}(u)$ with the subscript indicating the projection or the random normal variable,  the `norm_kol` index is defined by
@@
 
-Figure \@ref(fig:noisy-better-geo) compares the tracing plot of two algorithms: `search_geodesic` and `search_better`.
-This time, the interpolated path is no longer smooth when using either algorithm.
-It is also obvious that `search_geodesic` fails to optimise this rough index since there is barely improvment of the index value.
-On the other hand, `search_better` is doing a relatively good job on the optimisation for the 5-variable dataset `boa5`.
-The theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)` and `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 4)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
-A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 4)`].
+Figure \@ref(fig:noisy-better-geo) compares the tracing plot of two optimisers: `search_geodesic` and `search_better`.
+This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely improvement of the index value.
+On the other hand, `search_better` is doing a relatively well.
+With theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
+A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 3)`].
 At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
-```{r kol-better-default, fig.height = 6, fig.cap="(ref:kol-better-default)"}
+```{r kol-better, fig.height = 6, fig.cap="(ref:kol-better)"}
 ```
 
-(ref:kol-better-default) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better` with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced into 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
-The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.
+(ref:kol-better) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better` with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced to 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
+The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.
 
-```{r kol-better-random-default, fig.height = 6, fig.cap = "(ref: kol-better-random-default)"}
+```{r kol-random, fig.height = 6, fig.cap = "(ref:kol-random)"}
 ```
 
-(ref: kol-better-random-default) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
-`search_better_random` has a probablistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
+(ref:kol-random) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
+`search_better_random` has a probabilistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the global maximum.
 
-```{r kol-better-random-tuned, fig.height= 6, fig.cap = "(ref:kol-better-random-tuned)"}
+```{r kol-random-tuned, fig.height= 6, fig.cap = "(ref:kol-random-tuned)"}
 ```
 
-(ref:kol-better-random-tuned) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
+(ref:kol-random-tuned) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with an larger searching neighbourhood of 0.7.
 Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
-This time, all the 20 incidents have found the global maximum.
 
 The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is  presented.
-The dataset used is `boa6` with `x2` and `x7` being informative.
-The two possible theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)` repectively and hence, the global maximum happens when variable `x7` is found.
-Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: `search_better` and `search_better_random`.
-The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.
+The dataset used is `boa6` where `x2` and `x7` are informative.
+The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
+Hence, the global maximum happens when variable `x7` is found.
+Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \@ref(fig:kol-better)) and `search_better_random` (Figure \@ref(fig:kol-random) and Figure \@ref(fig:kol-random-tuned)).
+Compared to Figure \@ref(fig:kol-better) and \@ref(fig:kol-random), Figure \@ref(fig:kol-random-tuned) further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.
+The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
 
-Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser `search_better`.
-This happens because `search_better` will always accept a basis with higher index value.
-Figure \ref{pca-better-random-default} shows the results when the optimiser `search_better_random` is used.
-With a probability acceptance on inferior basis, more simulation ends up finds the global maximum.
-An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}).
-When using `search_better` to optimise, it goes straight down the track to find `V2` whilst with `search_better_random`, it first has a tendency to hand towards `V2` but then makes a sharp turn back to the track that finds `V7`.
+In Figure \@ref(fig:kol-better), the starting point largely affects whether `search_better` will find the global maximum.
+The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find `V7` because they are born in Rome.
+On the other end, the ones on the first two rows can only find `V2`, largely due to the fact that it doesn't even have chances to search near `V7`.
 
-In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated basis and hence represents a wider search.
-By this specification, all the 20 seends find the global maximum and a few simulations i.e.
-seed 4761 (row 4,  column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.
+A comparison between `search_better` and `search_better_random` shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
+Two iconic cases that shows how `search_better_random` improves `search_better` are seeds 4761 (position (1,4) in Figure \@ref(fig:kol-better) and (2,2) in Figure \@ref(fig:kol-random)) and seed 1842 ((2,2) in Figure \@ref(fig:kol-better) and (4,5) in Figure \@ref(fig:kol-random)).
+In both cases, `search_better_random` accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
+However, this change of search direction may sometimes have an adverse effect on `search_better_random` and change the search to the neighbourhood of `V2`.
+This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \@ref(fig:kol-random)).
+The two cases that neither the algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \@ref(fig:kol-better) and position (1,5) and (1,3) in Figure \@ref(fig:kol-random)).
+This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near `V7` to be adequately explored.
+
+In Figure \ref{kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated basis and hence allows a wider search.
+Seed 9982 ((2,4) in Figure \@ref(fig:kol-random-tuned)) and seed 2463 ((3,2) in Figure \@ref(fig:kol-random-tuned)) benefit from this and find the global maximum.
+A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
+In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, the allows a near `V2` bases to switch to a near `V7` bases and results these cases to find the global maximum.
+While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near `V7` basis to near `V2` basis and it doesn't come back and leave those simulations in the local maximum.
+
+The conclusion from this experiment is that the usage of `search_better_random` and increase the search space are methods that can avoid trapping in local maximum but the best solution will depends on the starting points of the simulation and the seed used.
 
@@
 
-The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing `tourr` package.
-The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, `ferrn`.
+The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing `tourr` package while the optimiser diagnostics have been implemented in a new package, `ferrn`.
+When a guided tour is run, the users can choose if the data from optimisation should be collected via the `verbose` argument.
+Once the data object has been obtained, the package, `ferrn`, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
+The structure of package functionality has been listed below.
 
-When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument `print = FALSE`).
-Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.
-Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.
-Below presents a sample of the data object.
+- `explore_trace_search()`: produces summary plots, as shown in Figure \@ref(fig:toy-search)
+- `explore_trace_interp()`: produces trace plots for the interpolation points, as shown in Figure \@ref(fig:toy-interp)
+- `explore_space_pca() `: produces plots of projection basis on the reduced space by PCA, as shown in Figure \@ref(fig:toy-pca).
+Animated version in Figure \@ref(fig:toy-pca-animated) can be turned on via the argument `animate = TRUE`
+- `explore_space_tour()`: produces animated tour view on the full space of the projection bases, as shown in Figure \@ref(fig:toy-tour).
 
-```{r data-object, fig.cap = "(ref:data-object)"}
-```
+- `get_*()` extracts and manipulates certain components from the existing data object.
+  - `get_best()`: extracts the best basis found in the data object
+  - `get_start()`: extracts the starting basis
+  - `get_interp()`: extracts the observations in the interpolation
+  - `get_search_count()`: produces the summary table of the number of observation in each iteration
+  - `get_basis_matrix()`: flattens all the bases into a matrix
 
-Once the data object has been obtained, the package, `ferrn`, provides automatic diagnostic plots as described in Section \ref{vis-diag}. The structure of package functionality has been listed below.
+- `bind_*()` incorporates additional information outside the tour optimisation into the data object.
+  - `bind_theoretical()`: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
+  - `bind_random()`: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
+`bind_random_matrix()` binds the points to the basis matrix.
 
-**Main functions**
-
-- `explore_trace_search()`: produce summary plots, as shown in Figure \@ref(fig:toy-search)
-- `explore_trace_interp()`: produce trace plots for the interpolation points, as shown in Figure \@ref(fig:toy-interp)
-- `explore_space_pca() `: produce plots of projection basis on the reduced space by PCA, as shown in Figure \@ref(fig:toy-pca).
-Animated version in Figure \@ref(fig:toy-pca-aniamted) can be acquired via the argument `animate = TRUE`
-- `explore_space_tour()`: produce animated tour view on the full space of the projection bases, as shown in Figure \@ref(toy-tour)
-
-**Auxiliary functions**
-
--  `get_*()` extract and, in some cases, manipulate certain components from the existing data object.
-    - `get_best()`: extract the best basis found by the algorithm
-    - `get_start()`: extract the starting basis
-    - `get_interp()`: extract the observations in the interpolation and prepare to be supplied to `explore_trace_interp()`. 
-    - `get_search_count()`: produce the summary table of the number of observation in each iteration and prepare to be supplied to `explore_trace_search()`
-    - `get_basis_matrix()`: flattern all the bases into a row and stack them to form a large matrix for `explore_space_tour()`
-
-- `bind_*()` incorporate additional information that is outside the tour optimisation.
-
-  - `bind_theoretical()`: incorporate the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
-  - `bind_random()`: generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object and output as a tibble object.
-`bind_random_matrix()` returns a matrix.
-
-- Color and relevel
-  - `botanical_palettes`:  a collection of color palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequantial palettes cotain fern and acacia.
-  - `botanical_pal()`: a color interpolator useful for `scale_color_botanical()`
+- Color
+  - `botanical_palettes`:  a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+  - `botanical_pal()`: a color interpolator 
   - `scale_color_botanical()`: a ggplot construction for using botanical palettes.
-  - `relevel_geo()`: manipulate the level in the `info` column in the data object when geodesic search is used
-  - `relevel_better()`: manipulate the level in the `info` column in the data object when search_better related search is used
 
@@
 <!-- something about the diagnostic plots  -->
-Four diagnostic plots have been introduced in the paper to investigate the progression and the projection space of an optimser, ranging from simple summary plot to advanced high dimensional animation.
-The implementation of these visualisations are designed to be easy-to-use with each plot can be produced with a simple supply of the data object and few additional arguments.
-More advanced users may decide to modify on top of the basic diagnostic pltos or even build their own.
+Four diagnostic plots have been introduced to investigate the progression and the projection space of an optimiser.
+The implementation of these visualisations is designed to be easy-to-use with each plot can be produced with a simple supply of the data object.
+More advanced users may decide to modify on top of the basic plots or even build their own.
 
 <!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two pacakges: the collection of the data object is implemented in the existing `tourr` package; manipulation and visualisation of the data object are implemented in the new `ferrn` package.
-Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions i.e.
+Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing `tourr` package; manipulation and visualisation of the data object are implemented in the new `ferrn` package.
+Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.
 scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.

### 2020-10-13 paper/paper.tex
tuning languages here and there and everywhere

@@
 \begin{abstract}
-Friedman \& Tukey commented on their initial paper on projection pursuit in 1974 that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure.
-In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum.
-This diagnostic system can be applied more generally to help practitioner to unveil the black-box in randomised iterative (optimisation) algorithms.
-An R package, ferrn, has been created to implement this diagnostic system.
+Friedman \& Tukey commented on their initial paper on projection pursuit in 1974 that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedures.
+In this paper, we illustrates the diagnostics of optimisers for projection pursuit guided tour via the construction of a data object and producing diagnostic plots.
+These diagnostics workflow can be applied generally on optimisers to assess its performance.
+An R package, ferrn, has been created to implement the diagnostics.
 \end{abstract}
@@
 Presenting information in a graphical format often allows people to see information they would otherwise not see.
-This motivates our work of creating plots to diagnose optimisation algorithms in the context of projection pursuit guided tour, with the aim to understand and compare features of different existing algorithms.
+This motivates our work of creating plots to diagnose optimisation algorithms with the aim to understand and compare features of different algorithms.
 
-In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints.
+In an optimization problem the goal is to find the best solution within the space of all feasible solutions, which typically is represented by a set of constraints.
 The problem consists in optimizing an objective function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in a reduced space given by the problem constraints to either minimize or maximize a function.
 
-Projection pursuit and guided tour are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
+The problem context of the optimisation that has been worked in this paper is projection pursuit guided tour, which are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
 Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
@@
 The remainder of the paper is organised as follows.
-Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods used in projection pursuit guided tour.
-Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem, and introduces three main existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics design, from forming the data object to the definition of different diagnostic plots with some small examples.
-Section \ref{application} shows the application of how the diagnostic plots designed in section \ref{vis-diag} can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
+Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods.
+Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem and its features, and introduces three main existing algorithms.
+Section \ref{vis-diag} presents new visual diagnostics designs. Data object is first defined to pinned down the data structure, several diagnostic plots are then defined with sample examples given.
+Section \ref{application} shows the application of how the diagnostic plots designed previously can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
 Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
@@
 
-Given an optimisation problem, two basic approaches find the optimum based on different thinking.
-An analytical approach aims to find the optimal solution in a finite number of steps, but a potential issue with it is that the closed-form solution may not be available when the problem starts to become complex.
-An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution.
-An iterative method may end up finding a local optimum but the progressive nature of the algorithm allows the practitioner to decide when to stop if a desirable accuracy has been achieved.
-
-A traditional while often used in practice is an iterative method called \emph{line search method} \citep{fletcher2013practical}.
-In a simple one-dimensional problem of finding the \(x\) that minimises \(f(x)\), line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.
+Given an optimisation problem, two broad approaches find the optimum based on different thinking and hence have its own strength and deficits.
+An analytical approach aims to find the optimal solution in a finite number of steps, but the closed-form solution may not be available when the problem starts to become complex.
+An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution and researchers can decide when to stop if a desirable accuracy has been achieved.
+However, a potential issue is that it may end up finding a local optimum.
+In this paper, the main focus is on a particular iterative method: \emph{line search method} \citep{fletcher2013practical}.
+In a simple one-dimensional problem of finding the \(x\) that minimises \(f(x)\), line search conducts its search via an iterative up as in Equation \ref{eq:line-search}.
 
@@
 where \(d^{(j)}\) is the searching direction in iteration \(j\), and \(\alpha_j\) is the step-size.
-Strictly speaking, \(\alpha_k\) is chosen by another minimisation of \(f(x^{(j)} + \alpha* d^{(j)})\) with respect to \(\alpha\) and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of \(\alpha_j\) is attained \citep{curry1944method}.
-In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation.
-A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\) and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.
+Theoretical results have demonstrated the global convergence of the algorithm when \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
+However, in practice, this is rarely implemented due to its computational demanding or sometimes the existence of such a minimisation\citep{fletcher2013practical}.
+A modified approach is to impose a mandatory decrease in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\) and this approach turns out to be efficient in practical problems despite the lost of guarantee on global convergence.
 
@@
 Modern development of the line search methods focuses on proposing different computation on the searching direction: \(d^{(j)}\) and various approximations on the step size: \(\alpha_j\) catered for practical optimisation problems.
-The specific problem context we are interested in is called projection pursuit guided tour.
-Projection pursuit and guided tour are two separate methods in exporatory data analysis focusing on different aspects: coined by \citet{friedman1974projection}, projection pursuit detects interesting structures (i.e.~clustering, outliers and skewness) in multivariate data via low dimensions projection; whilst guided tour is a particular variation in a broader class of data visualisation method called tour.
+The specific problem context discussed in this paper is projection pursuit guided tour.
+Projection pursuit and guided tour are two separate methods in exploratory data analysis focusing on different aspects: Projection pursuit, coined by \citet{friedman1974projection}, detects interesting structures (i.e.~clustering, outliers and skewness) in multivariate data via low dimensions projection whilst guided tour is one particular variation in a broader class of data visualisation method, tour.
 
-Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{P}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection basis.
+Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection basis.
 Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projection basis \(\mathbf{A}\) onto an index value \(I\), this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the ``interestingness'' of a projection.
@@
 
-In their initial paper, \citet{friedman1974projection} noted that ``\ldots{}, the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' Hence, effective optimisation algorithms are necessary for projection pursuit to find the bases that give interesting projections.
-While we leave the formal construction of the optimisation problem and existing algorithms to section \ref{tour-optim}, we outline the general idea here.
-Given a random starting (current) basis, projection pursuit repeatedly searches for candidate bases nearby until it finds one with higher index value than the current basis.
-In the second round, that basis becomes the current basis and the repetitive sampling continues.
-The process ends until no better basis can be found or one of the termination criteria is reached.
-
-Before introducing the guided tour, we shall be familiar with the general tour method \citep{cook2008grand}.
-A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
-The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to \citet{buja2005computational} for the mathematical details.
-Iteratively choosing different low dimension planes and interpolating between them forms a tour path.
-Different types of tour methods choose the low dimensional planes differently and we mention two other types of tour that are commonly used.
-A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data.
-Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.
-
-Guided tour chooses the planes produced by optimising the projection pursuit index function.
-Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, which are the interpolations between two blue frames.
-The tour method has been implemented in the \emph{tourr} package in R, available on the Comprehensive R Archive Network at \url{https://cran.r-project.org/web/packages/tourr/} \citep{wickham2011tourrpackage}.
+As a general visualisation method, A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
+Different tours choose the low dimensional planes differently, for example, a grand tour\citep{cook2008grand} selects the planes randomly in the high dimensional space and manual tour\citep{cook1997manual} chooses a sequence of bases that gradually phase in and out one variable to view the contribution of that variable in the visualisation.
+Guided tour, the main interest of this paper, chooses its projection with the aid of projection pursuit.
+Given a random start, projection pursuit iteratively finds bases with higher index value and guided tour constructs geodesic interpolation between two consecutive to form a tour path.
+Mathematical details of the geodesic interpolation can be found in \citet{buja2005computational}.
+Figure \ref{fig:tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, the interpolations between two blue frames.
+The tour method has been implemented in the \texttt{tourr} package in R, available on the Comprehensive R Archive Network at \url{https://cran.r-project.org/web/packages/tourr/} \citep{wickham2011tourrpackage}.
 
@@
 \caption{Each square (frame) represents the projected data with a corresponding basis.
-Blue frames have bases found by an optimisation algorithm iteratively whilst the white frames are constructed by geodesic interpolation between two blue frames.}\label{fig:tour-path}
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
@@
 
-Now we begin to formulate the optimisation problem.
+The optimisation problem is stated as follows.
 Given a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit finds the final projection basis \(\mathbf{A}_T\) that satisfies the following optimisation problem:
@@
 
-where \(I_d\) is the d-dimensional identity matrix. The constraint requires the projection basis \(\mathbf{A}\) to be an orthogonal matrix with each column vector being orthonormal.
+where \(I_d\) is the d-dimensional identity matrix and the constraint requires the projection bases \(\mathbf{A}\) to be an orthogonal matrices.
 
-There are several features of this optimisation that are worth noticing.
-First of all, it is a multivariate constraint optimisation problem.
-Since the decision variables are the entries of a projection basis, it is required to be orthonormal.
-It is also likely that the objective function is non-differentiable or the gradient information is simply not available.
-In this case, we will need to either use some approximation of the gradient or turn to derivative free methods.
-Given the goal of projection pursuit as finding the basis with the largest index value, the optimisation problem needs to be able to find the global maximum.
-Along the way, local maximum may also be of our interest since they could present unexpected interesting projections.
-There is also one computational consideration: the optimisation procedure needs to be easy to compute since the tour animation is played in real-time.
+Several features of this optimisation are worth noticing.
+First of all, this is a constraint optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
+It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
+Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
+Lastly, there is also one computational consideration: the optimisation procedure needs to be fast to compute since the tour animation is played in real-time.
 
@@
 
-Below we introduce three possible algorithms: \texttt{search\_better} and \texttt{search\_better\_random} are derivative free methods that sample candidate bases in the neighbourhood, whilst \texttt{search\_geodesic} is an analogue of gradient ascent on the manifold of the projection basis.
+Below we introduce three possible algorithms: \texttt{search\_better}, \texttt{search\_better\_random}, and \texttt{search\_geodesic}.
+The first two are derivative free methods that sample candidate bases in the neighbourhood whilst \texttt{search\_geodesic} is an analogue of gradient ascent on the projection basis space.
 
@@
 The next iteration of search begins after adjusting \(\alpha\) by a cooling parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\).
-Another termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
+The termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
+The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
 A slightly different cooling scheme has been proposed by \citet{posse1995projection} to include a halving parameter \(c\).
 Rather than reducing the radius of the searching neighbourhood, \(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if the last search takes more than \(c\) times to find an accepted basis to avoid the searching space being reduced too fast.
-The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
-{[}mention orthonormalise to ensure the constraint is fulfilled; don't use derivative information but a random search{]}
 
@@
 
-Simulated annealing \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on a cooling scheme \(T(l)\).
+Simulated annealing (\texttt{search\_better\_random}) \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on the annealing \(T(l)\).
 Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{I_{\text{cur}} - I_{l}}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
+When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
 This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when the dimension of the projected space is smaller than the number of informative variables in the dataset (i.e.~a one dimensional projection of the dataset with two informative variables).
+This feature is particularly useful when local maximum is presented.
 The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm on the manifold of the projection bases.
-In gradient ascent, one first find the direction for improvment via computing the gradient information.
+\citet{cook1995grand} used a gradient ascent algorithm on the space of the projection bases.
+In gradient ascent, one first find the direction for improvement via computing the gradient information.
 In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
-A geodesic is then constructed using the current basis and one of the \(2n\) bases that have the highest index value.
-If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and can work as the searching direction.
-In gradient ascent, the next step is to conduct a line search to find the best improvement along the gradient direction.
-In \texttt{search\_geodesic}, this step is replaced by optimising the index value along the geodesic direction constructed before over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
-The optima \(\mathbf{A}_{**}\) is outputted for the current iteration if the percentage change in the index value between \(\mathbf{A}_{**}\) and \(\mathbf{A}_{\text{cur}}\) is greater than a threshold value.
-As above, another termination condition is when \(l_{\max}\) is reached.
+A geodesic is then constructed using the current basis and the one in \(2n\) bases with the highest index value.
+If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and works as the searching direction.
+The next step in gradient ascent is to conduct a line search to find the best improvement along the gradient direction and in \texttt{search\_geodesic}, this is replaced by optimising the index value along the geodesic direction over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
+The optima \(\mathbf{A}_{**}\) is returned for the current iteration if it meets the termination condition on percentage improvement.
+The procedure will also terminate if \(l_{\max}\) is reached.
 Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
@@
 
-In the optimisation algorithms for projection pursuit, three main elements to record are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3) State: \(S\), which labels the observation with detailed stage in the optimisation.
+In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3) State: \(S\), which labels the observation with detailed stage in the optimisation.
+Possible values for \texttt{search\_better} and \texttt{search\_better\_random} includes \texttt{random\_search}, \texttt{new\_basis}, and \texttt{interpolation}.
+\texttt{search\_geodesic} has a wider variety that includes \texttt{new\_basis}, \texttt{direction\_search}, \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and \texttt{interpolation}.
+
 Multiple iterators are also needed to index the data collected at different levels.
-\(t\) is a unique identifier that prescribes the natural ordering of each observation; \(j\) is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts.
-\(l\) is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.
-There are other parameters that are of our interest and we denote them as \emph{\(V_{p}\)}.
-In projection pursuit, this includes \(V_1 = \text{method}\), which tags the name of the algorithm used and \(V_2 = \text{alpha}\), the neighbourhood parameter that controls the size in sampling candidate bases.
+\(t\) is a unique identifier that prescribes the natural ordering of each observation; \(j\) is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
+\(l\) is the counter for each search/interpolation, which provides the information of how many basis the algorithm has searched before finding one to return.
+There are other parameters of interest, depends on the particular problem content and they are denoted as \emph{\(V_{p}\)}.
+Two most common examples include \(V_1 = \text{method}\), which tags the name of the algorithm used, and \(V_2 = \text{alpha}\), the neighbourhood parameter that controls the size in sampling candidate bases.
 A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
@@
 where \(T^{\prime} = T + k_{J}+ l_{J+1}\).
-Note that we deliberately denote the last round of search as \(j = J+1\) and in that round there is no output/interpolation basis and the algorithm terminates.
-This notation allows us to denote the last complete search-and-interpolate round as round \(J\) and hence the final basis is \(A_T\) and highest index value found is \(I_T\).
+Note that there is no output in iteration \(J + 1\) since the optimiser can't find a better basis and the algorithm terminates.
+In this notation, final basis found is \(A_T\) with the highest index value \(I_T\).
 
-It is worth noticing that the data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that states
+The data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that requires each observation forms a row and each variable forms a column.
+With tidy data structure, data wrangling and visualisation have been significantly simplified by the well-developed packages like dplyr\citep{dplyr} and ggplot2\citep{ggplot2}.
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi})}
-\tightlist
-\item
-  each observation forms a row,
-\item
-  each variable forms a column, and
-\item
-  each type of observational unit forms a table
-\end{enumerate}
-
-The wrangling and visualisation of tidy data have been greatly simplified by the well-known dplyr\citep{dplyr} and ggplot2\citep{ggplot2} package.
-
-With a constructed data object, the construction of diagnostic plots is inspired by the concept of grammar of graphic \citep{wickham2010layered}, which powers the primary graphical system in R, ggplot2 \citep{ggplot2}.
-In grammar of graphic, plots are not defined by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by ``stacked layers''.
-Using this design, ggplot does not have to develop a gazillion of functions that each produces a different type of plot from a different data structure.
-Instead, it aesthetically maps variables (and its statistical transformation) in a dataset to different geometric objects (points, lines, box-and-whisker etc) and builds the plot through overlaying different layers.
+The construction of diagnostic plots uses the concept of grammar of graphic \citep{wickham2010layered} in ggplot2.
+In grammar of graphic, plots are not defined by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by stacked layers.
+In the construction of diagnostic plots, there are multiple elements one may wants to emphasize and there's no single plot name that would meet the need, the stacked layers concept, on the other hand, allow information to be overlaid on the each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
 
@@
 
-A primary interest of diagnosing an optimisation algorithm is to study how it progressively finds its optimum.
-One way of doing it is to plot the index value across its natural ordering \(t\), however, it usually takes the algorithm much longer to find a better basis than the current one towards the end of the search.
-When plotting the searching observations as points in a plot, the space each iteration takes will be proportional to the number of points in that iteration.
+A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum.
+Directly plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes longer for an optimiser to find a better basis towards the end.
+Another option is to use summarisation for each iteration.
+Boxplot is a suitable candidate that can provide five points summary of each iteration.
+Other additional information not presented in the boxplot and then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
+Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a \texttt{cutoff} parameter.
 
-Another option is to use summarisation for each iteration.
-Boxplot is a suitable candidate that provides five points summary of the data, however, there are two pieces of information missing from the boxplot: 1) It does not report the number of points, and 2) the position of the last basis.
-This could be remedied by adding more layers using the concept of grammar of graphics.
-A label geometry is added at the bottom of the plot to show the number of points in each iteration and a line geometry links the last basis in each iteration.
-Further, an option to switch between displaying points and boxplot geometry is helpful since point geometry can be more intuitive for the iteration with few observations.
-This is achieved via a \texttt{cutoff} parameter.
-
-Figure \ref{fig:toy-search} presents a comparison between the two plotting options discussed above.
-Using the natural order to plot searching points distorts the point from the first few iterations and over-emphasizes the searches in the last few iterations.
-The second plot design evenly presents the summarised information for each iteration while allowing for a switch to the full information for the iteration with small number of observations.
+Figure \ref{fig:toy-search} shows a sample of the plot constructed using \texttt{search\_better} with different value on the parameter \texttt{max\_tries}.
+Comparing the returned index value of each iteration shows that the termination at \texttt{max\_tries\ =\ 25} is not sufficient for \texttt{search\_better} to explore the parameter space and a value of 500 is preferred over 25 in this context.
 
@@
 \caption{A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
-A switch of geometry from points to boxplot happens when the count in the iteration excess 15 to avoid overplotting.
-With \texttt{max\_tries\ =\ 500} specification, the algorithm is working much harder towards to the end.}\label{fig:toy-search}
+A six variable dataset \texttt{boa6} is used with \texttt{holes} index on a 2D problem.
+A \texttt{max\_tries\ =\ 500} since it allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
 \end{figure}
 
-\hypertarget{examining-the-optimisation-progress}{%
-\subsection{Examining the optimisation progress}\label{examining-the-optimisation-progress}}
+\hypertarget{toy-interp}{%
+\subsection{Examining the optimisation progress}\label{toy-interp}}
 
@@
 
-\caption{The resulting trace plot on the interpolated points has been ploted when using three different algorithms to optimise the index.
+\caption{The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
 The color represents the number of iteration.
@@
 
-Sometimes, we may be interested in exploring the points on the interpolation path since these are the points that will be played by the tour animation.
-Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths.
-The upper plot shows a desirable interpolation in each iteration with the index value being progressively and monotonically increasing.
-While in the middle plot, the increases towards the target basis is not monotonical in the last two iterations and interpolated basis with higher index value can be found on the tour path.
-The lower plot is constructed using \texttt{search\_better\_random}, where a basis with smaller index value has a probabilistic chance of being accepted and we can observe a much more involved pattern.
-In iteration three, the probabilistic acceptance produces a monotonically decreasing interpolation whilst in iteration five, six and seven, with also a probabilistic acceptance, the interpolation is now non-monotonical.
-When the target basis has a higher index value, iteration eight reaches this basis by first decreasing the index value.
-While the interpolation plot shows different possibilities of the change in index value on the tour path, the diagnosis of the validity of each pattern and the related optimisation algorithms will be postponed to Section \ref{application}.
+Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
+Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths each with different curvature.
+The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
+The middle path has an increase-then-decreases pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
+The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost cases is a deliberate construction of \texttt{search\_better\_random} where an inferior basis can be accepted in a probability manner so as to avoid trapping in the local maximum.
 
@@
 
-\caption{PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search}\label{fig:toy-pca}
+\caption{1D projection on the 5 variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
+The yellow point corresponds to the theoretical best basis {[}0, 1, 0, 0, 0{]} with \texttt{V2} being the only non-normal variable in the dataset.
+The underlying grey points are randomly generated on the 5D space and reduced to 2D via PCA along with all the search points presented.
+The enlarged color points that colored as the interpolation are the starting points of each algorithm.
+They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
 \end{figure}
@@
 Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go above five (5).
-Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d \textgreater{} 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e.~Principal component analysis.
-To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
+Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go over two or three.
+While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (d \textgreater{} 3), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.~principal component analysis.
+To better perceive the the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
 
-Figure \ref{fig:toy-pca} shows the first two principal components of two search paths.
-Starting from the same position, two searching methods take different paths to get its final bases.
-There are two theoretical bases because they correspond to the same data projection \(Y\) with a 180 degree rotation.
-What differentiate the two optimisers is that \texttt{search\_better()} also involve random evaluation of points in the sphere during its sampling process while \texttt{search\_geodesic()} doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.
+Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using \texttt{search\_better} and another uses \texttt{search\_geodesic}.
+The search in the space reduced by PCA matches with the optimiser description before where the random sampling in \texttt{search\_better} is broader, controlled by \texttt{alpha} parameter, which is default to 0.5 while the directional search in \texttt{search\_geodesic}, controlled by \texttt{delta} with a default of 0.01, is so tiny that it can barely seen.
 
@@
 
-\caption{A selected number of frames from the animated PCA plot. With animation, it is easier to track the progrssion from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\caption{A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
 \end{figure}
 
-Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.
-In Figure \ref{fig:toy-pca-animated}, There should actually be more writing here to and also the this examples may need to be changed.
+Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
+Figure \ref{fig:toy-pca-animated} shows six frames of an animated version of Figure \ref{fig:toy-pca} and this time, it shows that \texttt{search\_better} finds the optimum quicker than \texttt{search\_geodesic}.
 
@@
 
-Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a stereoscopic view of the search.
-To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation.
-This can be done via a regular grand tour or a slice tour \citep{laa2020slice}, which emphasizes the points that are close to a section of the sphere.
-
-Compare to the PCA plot, the animated rotation (tour) display, in Figure \ref{fig:toy-tour}, gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial.
-This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path.
-Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.
+Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a more holistic stereoscopical view of the search.
+To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
+Compare to the PCA plot, the animated rotation (tour) display in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and grand tour produces a sequence of angle that view the search from different direction.
+As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to one in Figure \ref{fig:toy-pca}.
 
@@
 If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are presented, \texttt{search\_better} may stuck in the local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
-In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of the algorithms in different scenarios.
+When multiple optima are presented, \texttt{search\_better} may stuck at the local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
+In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
 
@@
 
-\hypertarget{a-problem-of-not-monotonic}{%
-\subsection{A problem of not monotonic}\label{a-problem-of-not-monotonic}}
+\hypertarget{monotonic}{%
+\subsection{A problem of not monotonic}\label{monotonic}}
 
-We use the same dataset as the toy example above to explore the search function \texttt{search\_better} and we want to learn how the index value changes on the interpolation path for the \texttt{holes} index.
-From the left panel of Figure \ref{fig:interruption}, we observe that when interpolating from the current basis to the target basis, the index value may not be monotone: we could reach a basis with a higher index value than the target basis on the interpolation path.
-In this sense, we would be better off using the basis with the highest index value on the interpolation path as the current basis for the next iteration (rather than using the target basis).
-Hence, an interruption is constructed to accept the interpolating bases only up to the one with the largest index value.
+In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
+This pattern is undesirable since the optimiser could have start the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
+This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
 After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
@@
 
-\caption{A comparison of the trace plot with and without the interruption.
+\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_better}.
+A comparison of the trace plot with and without the interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the interpolated basis with the highest index value in an iteration.}\label{fig:interruption}
+The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.}\label{fig:interruption}
 \end{figure}
@@
 Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
-This motivates the polish search where the final basis is supplied to a new guided tour to search for any local breakthrough.
+This motivates the polish search where the final basis is supplied as the start of a new guided tour to search for any local breakthrough.
 
 Similar to \texttt{search\_better} as a stochastic random search, \texttt{search\_polish} has a different scheme of reducing the search neighbourhood.
-In each search-interpolation iteration, \texttt{search\_better} has a fixed neighbourhood parameter alpha and this alpha is reduced by another cooling parameter only after an iteration finishes.
+In each search-interpolation iteration, \texttt{search\_better} has a fixed neighbourhood parameter alpha and this alpha is reduced by the cooling parameter only after an iteration finishes.
 On the contrary, \texttt{search\_polish} allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
-Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original \texttt{max.tries}.
+Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original \texttt{max.tries} limit.
 These include:
@@
 
-Figure \ref{fig:polish} presents the final projections found before and after applying \texttt{search\_polish} using the demo code.
-Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables, namely \texttt{V1}, \texttt{V8}, \texttt{V9} and \texttt{V10}.
-When appears in the projected data in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
+Figure \ref{fig:polish} presents the final projections found before and after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
+Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables.
+In terms of the projected data as in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
 
@@
 \caption{One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
-The grey dashed line represents the index value of the theoreical best basis.
-\texttt{search\_geodesic} fails to optimse the noisy index while \texttt{search\_better} has made reasonable improvments to reach the index value close to the one of theoretical best basis.}\label{fig:noisy-better-geo}
+The grey dashed line represents the index value of the theoretical best basis.
+\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of theoretical best basis.}\label{fig:noisy-better-geo}
 \end{figure}
 
-Up until this point, the index functions are smooth, meaning the trace of interpolation points is smooth, while this is not the case for all the index functions.
-\texttt{norm\_kol}, a 1D projection function based on Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{P}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
+The index function, up until this point, are all smooth, while this is not the case for all the index functions.
+\texttt{norm\_kol}, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
 Denotes the ECDF function as \(F_{.}(u)\) with the subscript indicating the projection or the random normal variable, the \texttt{norm\_kol} index is defined by
@@
 
-Figure \ref{fig:noisy-better-geo} compares the tracing plot of two algorithms: \texttt{search\_geodesic} and \texttt{search\_better}.
-This time, the interpolated path is no longer smooth when using either algorithm.
-It is also obvious that \texttt{search\_geodesic} fails to optimise this rough index since there is barely improvment of the index value.
-On the other hand, \texttt{search\_better} is doing a relatively good job on the optimisation for the 5-variable dataset \texttt{boa5}.
-The theoretical best basis {[}0, 1, 0, 0, 0{]} produces an index value of 0.182 and \texttt{search\_better} finds the final basis {[}-0.02939, -0.99715, 0.04509, -0.03248, 0.04176{]} with an index value of 0.174.
-A further polish step will give a marginal improvement of index value to 0.173 with a basis of {[}-0.077342, -0.996695, -0.014675, -0.004626, -0.019520{]}.
+Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: \texttt{search\_geodesic} and \texttt{search\_better}.
+This time, the interpolated path is no longer smooth when using either algorithm and \texttt{search\_geodesic} fails to optimise this index with barely improvement of the index value.
+On the other hand, \texttt{search\_better} is doing a relatively well.
+With theoretical best basis {[}0, 1, 0, 0, 0{]} produces an index value of 0.182, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
+A further polish step will give a marginal improvement of index value to 0.175 with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}.
 At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-default-1}
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-1}
 
@@
 \caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced into 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points repsents the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
-The points produced by the seraching algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better-default}
+Each point represents a basis in the original 6D space, reduced to 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
+The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
 \end{figure}
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-random-default-1}
-
-}
-
-\caption{(ref: kol-better-random-default)}\label{fig:kol-better-random-default}
-\end{figure}
-
-(ref: kol-better-random-default) One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
-\texttt{search\_better\_random} has a probablistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the true global maximum.
-
-\begin{figure}
-
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-random-tuned-1}
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-random-1}
 
@@
 \caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.
-This time, all the 20 incidents have found the global maximum.}\label{fig:kol-better-random-tuned}
+\texttt{search\_better\_random} has a probabilistic acceptance implementation that would also a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more incidents find the global maximum.}\label{fig:kol-random}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
+
+}
+
+\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with an larger searching neighbourhood of 0.7.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.}\label{fig:kol-random-tuned}
 \end{figure}
@@
 The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is presented.
-The dataset used is \texttt{boa6} with \texttt{x2} and \texttt{x7} being informative.
-The two possible theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235 repectively and hence, the global maximum happens when variable \texttt{x7} is found.
-Simulation is done with 20 randomly generated seeds on different parameter specifications on two optimsers: \texttt{search\_better} and \texttt{search\_better\_random}.
-The data obbject is collected for each simulation and principal component analysis is applied to project the bases as points into 2D space with the orthonormality constraint imposed on the 6D space preserved by randomly generated basis on the surface of the 6D sphere.
+The dataset used is \texttt{boa6} where \texttt{x2} and \texttt{x7} are informative.
+The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235, respectively.
+Hence, the global maximum happens when variable \texttt{x7} is found.
+Simulation is done with 20 randomly generated seeds on two optimisers: \texttt{search\_better} (Figure \ref{fig:kol-better}) and \texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
+Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned} further increases the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge the search space.
+The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
 
-Figure \ref{pca-better-default} shows that half of the seeds end up getting trapped in the local maximum on optimiser \texttt{search\_better}.
-This happens because \texttt{search\_better} will always accept a basis with higher index value.
-Figure \ref{pca-better-random-default} shows the results when the optimiser \texttt{search\_better\_random} is used.
-With a probability acceptance on inferior basis, more simulation ends up finds the global maximum.
-An evidence of the superior of this probabilistic implementation is simulation 5107 (row 1 column 2 in Figure \ref{pca-better-default} and row 3 column 5 in Figure \ref{pca-better-random-default}).
-When using \texttt{search\_better} to optimise, it goes straight down the track to find \texttt{V2} whilst with \texttt{search\_better\_random}, it first has a tendency to hand towards \texttt{V2} but then makes a sharp turn back to the track that finds \texttt{V7}.
+In Figure \ref{fig:kol-better}, the starting point largely affects whether \texttt{search\_better} will find the global maximum.
+The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7} because they are born in Rome.
+On the other end, the ones on the first two rows can only find \texttt{V2}, largely due to the fact that it doesn't even have chances to search near \texttt{V7}.
 
-In the next simulation, as in Figure \ref{pca-better-random-tuned}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated basis and hence represents a wider search.
-By this specification, all the 20 seends find the global maximum and a few simulations i.e.~seed 4761 (row 4, column 5), 9209 (row 3, column 2), 2986 (row 2, column 3) from this double guarantee of wider search and probabilistic acceptance.
+A comparison between \texttt{search\_better} and \texttt{search\_better\_random} shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
+Two iconic cases that shows how \texttt{search\_better\_random} improves \texttt{search\_better} are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}).
+In both cases, \texttt{search\_better\_random} accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
+However, this change of search direction may sometimes have an adverse effect on \texttt{search\_better\_random} and change the search to the neighbourhood of \texttt{V2}.
+This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
+The two cases that neither the algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
+This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near \texttt{V7} to be adequately explored.
+
+In Figure \ref{kol-random-tuned}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated basis and hence allows a wider search.
+Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
+A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
+In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, the allows a near \texttt{V2} bases to switch to a near \texttt{V7} bases and results these cases to find the global maximum.
+While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near \texttt{V7} basis to near \texttt{V2} basis and it doesn't come back and leave those simulations in the local maximum.
+
+The conclusion from this experiment is that the usage of \texttt{search\_better\_random} and increase the search space are methods that can avoid trapping in local maximum but the best solution will depends on the starting points of the simulation and the seed used.
 
@@
 
-The implementation of this projection has been divided into two pacakges, where the data collection happens when a tour is running and hence implemented in the existing \texttt{tourr} package.
-The diagnostics of the algorithm via plots and other auxiliary functions have been implemented in a new pacakge, \texttt{ferrn}.
-
-When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
-Additional messages during the optimisation can be displayed by argument \texttt{verbose\ =\ TRUE}.
-Notice that the tibble object allows the list-column \texttt{basis} to be printed out nicely with the dimension of the projection basis readily available.
-Below presents a sample of the data object.
-
-\begin{verbatim}
-## # A tibble: 5 x 8
-##      id basis             index_val info          tries  loop method       alpha
-##   <int> <list>                <dbl> <chr>         <dbl> <dbl> <chr>        <dbl>
-## 1     1 <dbl[,1] [5 x 1]>     0.749 new_basis         1     1 search_bett~   0.5
-## 2     2 <dbl[,1] [5 x 1]>     0.730 random_search     2     1 search_bett~   0.5
-## 3     3 <dbl[,1] [5 x 1]>     0.743 random_search     2     2 search_bett~   0.5
-## 4     4 <dbl[,1] [5 x 1]>     0.736 random_search     2     3 search_bett~   0.5
-## 5     5 <dbl[,1] [5 x 1]>     0.747 random_search     2     4 search_bett~   0.5
-\end{verbatim}
-
-Once the data object has been obtained, the package, \texttt{ferrn}, provides automatic diagnostic plots as described in Section \ref{vis-diag}.
+The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing \texttt{tourr} package while the optimiser diagnostics have been implemented in a new package, \texttt{ferrn}.
+When a guided tour is run, the users can choose if the data from optimisation should be collected via the \texttt{verbose} argument.
+Once the data object has been obtained, the package, \texttt{ferrn}, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
 The structure of package functionality has been listed below.
 
-\textbf{Main functions}
+\begin{itemize}
+\item
+  \texttt{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
+\item
+  \texttt{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+\item
+  \texttt{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument \texttt{animate\ =\ TRUE}
+\item
+  \texttt{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+\item
+  \texttt{get\_*()} extracts and manipulates certain components from the existing data object.
 
@@
   \item
-  \texttt{explore\_trace\_search()}: produce summary plots, as shown in Figure \ref{fig:toy-search}
+    \texttt{get\_best()}: extracts the best basis found in the data object
   \item
-  \texttt{explore\_trace\_interp()}: produce trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+    \texttt{get\_start()}: extracts the starting basis
   \item
-  \texttt{explore\_space\_pca()}: produce plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
-Animated version in Figure \ref{fig:toy-pca-aniamted} can be acquired via the argument \texttt{animate\ =\ TRUE}
+    \texttt{get\_interp()}: extracts the observations in the interpolation
   \item
-  \texttt{explore\_space\_tour()}: produce animated tour view on the full space of the projection bases, as shown in Figure \ref{toy-tour}
+    \texttt{get\_search\_count()}: produces the summary table of the number of observation in each iteration
+  \item
+    \texttt{get\_basis\_matrix()}: flattens all the bases into a matrix
   \end{itemize}
-
-\textbf{Auxiliary functions}
-
-\begin{itemize}
 \item
-  \texttt{get\_*()} extract and, in some cases, manipulate certain components from the existing data object.
+  \texttt{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
 
@@
   \item
-    \texttt{get\_best()}: extract the best basis found by the algorithm
+    \texttt{bind\_theoretical()}: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value.
   \item
-    \texttt{get\_start()}: extract the starting basis
-  \item
-    \texttt{get\_interp()}: extract the observations in the interpolation and prepare to be supplied to \texttt{explore\_trace\_interp()}.
-  \item
-    \texttt{get\_search\_count()}: produce the summary table of the number of observation in each iteration and prepare to be supplied to \texttt{explore\_trace\_search()}
-  \item
-    \texttt{get\_basis\_matrix()}: flattern all the bases into a row and stack them to form a large matrix for \texttt{explore\_space\_tour()}
+    \texttt{bind\_random()}: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
+\texttt{bind\_random\_matrix()} binds the points to the basis matrix.
   \end{itemize}
 \item
-  \texttt{bind\_*()} incorporate additional information that is outside the tour optimisation.
+  Color
 
@@
   \item
-    \texttt{bind\_theoretical()}: incorporate the best possible basis to the existing data object with the supply of the index function and original data for producing the index value.
+    \texttt{botanical\_palettes}: a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
   \item
-    \texttt{bind\_random()}: generates a large number of points on the surface of a high dimensional sphere and attaches it to the existing data object and output as a tibble object.
-\texttt{bind\_random\_matrix()} returns a matrix.
-  \end{itemize}
-\item
-  Color and relevel
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \texttt{botanical\_palettes}: a collection of color palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequantial palettes cotain fern and acacia.
-  \item
-    \texttt{botanical\_pal()}: a color interpolator useful for \texttt{scale\_color\_botanical()}
+    \texttt{botanical\_pal()}: a color interpolator
   \item
     \texttt{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.
-  \item
-    \texttt{relevel\_geo()}: manipulate the level in the \texttt{info} column in the data object when geodesic search is used
-  \item
-    \texttt{relevel\_better()}: manipulate the level in the \texttt{info} column in the data object when search\_better related search is used
   \end{itemize}
@@
 
-Four diagnostic plots have been introduced in the paper to investigate the progression and the projection space of an optimser, ranging from simple summary plot to advanced high dimensional animation.
-The implementation of these visualisations are designed to be easy-to-use with each plot can be produced with a simple supply of the data object and few additional arguments.
-More advanced users may decide to modify on top of the basic diagnostic pltos or even build their own.
+Four diagnostic plots have been introduced to investigate the progression and the projection space of an optimiser.
+The implementation of these visualisations is designed to be easy-to-use with each plot can be produced with a simple supply of the data object.
+More advanced users may decide to modify on top of the basic plots or even build their own.
 
-Most of the work in this project has been translated into code in two pacakges: the collection of the data object is implemented in the existing \texttt{tourr} package; manipulation and visualisation of the data object are implemented in the new \texttt{ferrn} package.
-Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions i.e.~scagnostics, association, and information index \citep{laa2020using} and understand how the optimisers behave for index functions with different structures.
+Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \texttt{tourr} package; manipulation and visualisation of the data object are implemented in the new \texttt{ferrn} package.
+Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.~scagnostics, association, and information index \citep{laa2020using} and understand how the optimisers behave for index functions with different structures.
 

### 2020-10-16 paper/paper.Rmd
Small changes in the introduction.

@@
 - R
-abstract: Friedman & Tukey commented on their initial paper on projection pursuit in 1974 that "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedures.
-In this paper, we illustrates the diagnostics of optimisers for projection pursuit guided tour via the construction of a data object and producing diagnostic plots.
-These diagnostics workflow can be applied generally on optimisers to assess its performance.
+abstract: Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
+The initial paper by Friedman & Tukey in 1974 stated that  "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
+In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
+These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
 An R package, ferrn, has been created to implement the diagnostics.
@@
 
-Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see information they would otherwise not see.
-This motivates our work of creating plots to diagnose optimisation algorithms with the aim to understand and compare features of different algorithms.
+Visualisation is widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of.
+This motivates our work of creating plots to diagnose optimisation algorithms: we aim to understand and compare the behaviour of different algorithms.
 
-In an optimization problem the goal is to find the best solution within the space of all feasible solutions, which typically is represented by a set of constraints.
-The problem consists in optimizing an objective function $f: S \rightarrow \mathbb{R}$ with $S \in \mathbb{R}^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.
+The goal of optimisation is to find the best solution within the space of all feasible solutions, typically represented by a set of constraints.
+The problem can be formulated as the maximisation (or minimisation) of an objective function $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.
 
-The problem context of the optimisation that has been worked in this paper is projection pursuit guided tour, which are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
-Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
+The focus of this paper is on optimisation in the projection pursuit guided tour, which is an exploratory data analysis tool that detects interesting structure in high-dimensional data visible in low-dimensional projections.
+The target of the optimisation is thus to identify the most interesting low-dimensional view, given by a corresponding projection matrix.
 
 The remainder of the paper is organised as follows. 
-Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods. 
-Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem and its features, and introduces three main existing algorithms.
-Section \ref{vis-diag} presents new visual diagnostics designs. Data object is first defined to pinned down the data structure, several diagnostic plots are then defined with sample examples given.
-Section \ref{application} shows the application of how the diagnostic plots designed previously can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
+Section \ref{optim} provides an overview of optimisation methods, specifically line search methods. 
+Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and introduces three existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and used in different types of diagnostic plots.
+Section \ref{application} shows applications of how these plots can be used to understand and compare different algorithms.
+We also discuss how these insights contribute to modifications that improve the algorithms.
+Finally, Section \ref{implementation} describes the R package: ferrn, that implements the visual diagnostics.
 

### 2020-10-17 paper/paper.Rmd
Changes in Section 2.

@@
 
-Given an optimisation problem, two broad approaches find the optimum based on different thinking and hence have its own strength and deficits.
-An analytical approach aims to find the optimal solution in a finite number of steps, but the closed-form solution may not be available when the problem starts to become complex. 
-An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution and researchers can decide when to stop if a desirable accuracy has been achieved.
-However, a potential issue is that it may end up finding a local optimum.
-In this paper, the main focus is on a particular iterative method:  *line search method* [@fletcher2013practical].
-In a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search conducts its search via an iterative up as in Equation \ref{eq:line-search}.
+For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
+An iterative approach finds the optimum by progressively improving the current solution.
+Researchers can then decide when to stop if a desirable accuracy has been achieved.
+One potential issue with this technique is that it may end up finding a local optimum.
+
+In this paper, the main focus is on the  *line search method* [@fletcher2013practical].
+In a simple one-dimensional problem, finding the value of $x$ that minimises the function $f(x)$, the line search takes iterative steps as in Equation \ref{eq:line-search},
 
 \begin{equation}
-x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)}
+x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)},
 \label{eq:line-search}
@@
 where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
-Theoretical results have demonstrated the global convergence of the algorithm when $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
-However, in practice, this is rarely implemented due to its computational demanding or sometimes the existence of such a minimisation[@fletcher2013practical].
-A modified approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$ and this approach turns out to be efficient in practical problems despite the lost of guarantee on global convergence.
+Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
+This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical].
+A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$.
+Similarly, for  minimisation we would require a decrease in function value at each step.
+This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
 

### 2020-10-18 paper/paper.Rmd
Starting to work on Section 3.

@@
 
+Modern development of line search methods focuses on improving the computation of the searching direction $d^{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
+Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
+
 # Projection pursuit guided tour {#tour}
 
-Modern development of the line search methods focuses on proposing different computation on the searching direction: $d^{(j)}$ and various approximations on the step size: $\alpha_j$ catered for practical optimisation problems.
-The specific problem context discussed in this paper is projection pursuit guided tour.
-Projection pursuit and guided tour are two separate methods in exploratory data analysis focusing on different aspects: Projection pursuit, coined by @friedman1974projection, detects interesting structures (i.e.
-clustering, outliers and skewness) in multivariate data via low dimensions projection whilst guided tour is one particular variation in a broader class of data visualisation method, tour.
+The projection pursuit guided tour combines two different methods in exploratory data analysis, focusing on different aspects.
+Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
+clustering, outliers and skewness) in multivariate data via low dimensions projection.
+The guided tour is using ideas from projection pursuit to define a particular variation in a broader class of data visualisation methods, building on the grand tour approach [@As85].
 
-Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis.
-Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection.
+To define projection pursuit, we first need to establish the notation used.
+Let $\mathbf{X}_{n \times p}$ be the data matrix, with $n$ observations in $p$ dimensions.
+A d-dimensional projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
+Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$.
+This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a projection.
+
 A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
 
-As a general visualisation method, A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
-Different tours choose the low dimensional planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly in the high dimensional space and manual tour[@cook1997manual] chooses a sequence of bases that gradually phase in and out one variable to view the contribution of that variable in the visualisation.
-Guided tour, the main interest of this paper, chooses its projection with the aid of projection pursuit.
-Given a random start, projection pursuit iteratively finds bases with higher index value and guided tour constructs geodesic interpolation between two consecutive to form a tour path.
+As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
+Different tour types choose these planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly and a manual tour[@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
+Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
+Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
+
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-Figure \@ref(fig:tour-path) shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, the interpolations between two blue frames.
+Figure \@ref(fig:tour-path) shows a sketch of the tour path.
+The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
 The tour method has been implemented in the `tourr` package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].

### 2020-10-22 paper/paper.Rmd
name

@@
 
-- name: Author 1
+- name: H.Sherry Zhang
   thanks: The authors gratefully acknowledge ...
-  affiliation: Department of YYY, University of XXX
+  affiliation: Department of Econometrics and Business Statistics, Monash University 
 - name: Author 2
@@
 - projection pursuit 
-- guided tourr
-- visual 
+- guided tour
+- visualisation
 - diagnostics 

### 2020-10-22 paper/paper.tex
name

@@
   \author{
-        Author 1 \thanks{The authors gratefully acknowledge \ldots{}} \\
-    Department of YYY, University of XXX\\
+        H.Sherry Zhang \thanks{The authors gratefully acknowledge \ldots{}} \\
+    Department of Econometrics and Business Statistics, Monash University\\
      and \\     Author 2 \\
@@
 \begin{abstract}
-Friedman \& Tukey commented on their initial paper on projection pursuit in 1974 that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedures.
-In this paper, we illustrates the diagnostics of optimisers for projection pursuit guided tour via the construction of a data object and producing diagnostic plots.
-These diagnostics workflow can be applied generally on optimisers to assess its performance.
+Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
+The initial paper by Friedman \& Tukey in 1974 stated that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
+In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
+These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
 An R package, ferrn, has been created to implement the diagnostics.
@@
 \noindent%
-{\it Keywords:} optimisation, projection pursuit, guided tourr, visual, diagnostics, R
+{\it Keywords:} optimisation, projection pursuit, guided tour, visualisation, diagnostics, R
 \vfill
@@
 
-Visualisation has been widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to see information they would otherwise not see.
-This motivates our work of creating plots to diagnose optimisation algorithms with the aim to understand and compare features of different algorithms.
+Visualisation is widely used in exploratory data analysis.
+Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of.
+This motivates our work of creating plots to diagnose optimisation algorithms: we aim to understand and compare the behaviour of different algorithms.
 
-In an optimization problem the goal is to find the best solution within the space of all feasible solutions, which typically is represented by a set of constraints.
-The problem consists in optimizing an objective function \(f: S \rightarrow \mathbb{R}\) with \(S \in \mathbb{R}^n\) in a reduced space given by the problem constraints to either minimize or maximize a function.
+The goal of optimisation is to find the best solution within the space of all feasible solutions, typically represented by a set of constraints.
+The problem can be formulated as the maximisation (or minimisation) of an objective function \(f: S \rightarrow \mathbb{R}\), with \(S \in \mathbb{R}^n\), in a reduced space given by the problem constraints on \(f\).
 
-The problem context of the optimisation that has been worked in this paper is projection pursuit guided tour, which are exploratory data analysis tools that detect interesting structure of high dimensional data through projection on low dimensional space.
-Optimisation is applied here to search for the low dimensional space that finds the most interesting projection.
+The focus of this paper is on optimisation in the projection pursuit guided tour, which is an exploratory data analysis tool that detects interesting structure in high-dimensional data visible in low-dimensional projections.
+The target of the optimisation is thus to identify the most interesting low-dimensional view, given by a corresponding projection matrix.
 
 The remainder of the paper is organised as follows.
-Section \ref{optim} provides a literature review of optimisation methods, specifically the line search methods.
-Section \ref{tour} reviews projection pursuit guided tour, forms the optimisation problem and its features, and introduces three main existing algorithms.
-Section \ref{vis-diag} presents new visual diagnostics designs. Data object is first defined to pinned down the data structure, several diagnostic plots are then defined with sample examples given.
-Section \ref{application} shows the application of how the diagnostic plots designed previously can be used to understand and compare different algorithms and how they contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package: ferrn, that implements all the visual diagnostics above.
+Section \ref{optim} provides an overview of optimisation methods, specifically line search methods.
+Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and introduces three existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and used in different types of diagnostic plots.
+Section \ref{application} shows applications of how these plots can be used to understand and compare different algorithms.
+We also discuss how these insights contribute to modifications that improve the algorithms.
+Finally, Section \ref{implementation} describes the R package: ferrn, that implements the visual diagnostics.
 
@@
 
-Given an optimisation problem, two broad approaches find the optimum based on different thinking and hence have its own strength and deficits.
-An analytical approach aims to find the optimal solution in a finite number of steps, but the closed-form solution may not be available when the problem starts to become complex.
-An iterative approach, on the other hand, finds the optimum based on the idea of making progressive improvement to the current solution and researchers can decide when to stop if a desirable accuracy has been achieved.
-However, a potential issue is that it may end up finding a local optimum.
-In this paper, the main focus is on a particular iterative method: \emph{line search method} \citep{fletcher2013practical}.
-In a simple one-dimensional problem of finding the \(x\) that minimises \(f(x)\), line search conducts its search via an iterative up as in Equation \ref{eq:line-search}.
+For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
+An iterative approach finds the optimum by progressively improving the current solution.
+Researchers can then decide when to stop if a desirable accuracy has been achieved.
+One potential issue with this technique is that it may end up finding a local optimum.
+
+In this paper, the main focus is on the \emph{line search method} \citep{fletcher2013practical}.
+In a simple one-dimensional problem, finding the value of \(x\) that minimises the function \(f(x)\), the line search takes iterative steps as in Equation \ref{eq:line-search},
 
 \begin{equation}
-x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)}
+x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)},
 \label{eq:line-search}
@@
 where \(d^{(j)}\) is the searching direction in iteration \(j\), and \(\alpha_j\) is the step-size.
-Theoretical results have demonstrated the global convergence of the algorithm when \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
-However, in practice, this is rarely implemented due to its computational demanding or sometimes the existence of such a minimisation\citep{fletcher2013practical}.
-A modified approach is to impose a mandatory decrease in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\) and this approach turns out to be efficient in practical problems despite the lost of guarantee on global convergence.
+Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
+This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist \citep{fletcher2013practical}.
+A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\).
+Similarly, for minimisation we would require a decrease in function value at each step.
+This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
+
+Modern development of line search methods focuses on improving the computation of the searching direction \(d^{(j)}\), and on approximations of the step size \(\alpha_j\), catering for practical optimisation problems.
+Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
 
@@
 
-Modern development of the line search methods focuses on proposing different computation on the searching direction: \(d^{(j)}\) and various approximations on the step size: \(\alpha_j\) catered for practical optimisation problems.
-The specific problem context discussed in this paper is projection pursuit guided tour.
-Projection pursuit and guided tour are two separate methods in exploratory data analysis focusing on different aspects: Projection pursuit, coined by \citet{friedman1974projection}, detects interesting structures (i.e.~clustering, outliers and skewness) in multivariate data via low dimensions projection whilst guided tour is one particular variation in a broader class of data visualisation method, tour.
+The projection pursuit guided tour combines two different methods in exploratory data analysis, focusing on different aspects.
+Projection pursuit, coined by \citet{friedman1974projection}, detects interesting structures (e.g.~clustering, outliers and skewness) in multivariate data via low dimensions projection.
+The guided tour is using ideas from projection pursuit to define a particular variation in a broader class of data visualisation methods, building on the grand tour approach \citep{As85}.
 
-Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection basis.
-Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projection basis \(\mathbf{A}\) onto an index value \(I\), this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the ``interestingness'' of a projection.
+To define projection pursuit, we first need to establish the notation used.
+Let \(\mathbf{X}_{n \times p}\) be the data matrix, with \(n\) observations in \(p\) dimensions.
+A d-dimensional projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\), and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection matrix.
+Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projected data \(\mathbf{Y}\) (corresponding to an associated projection matrix \(\mathbf{A}\)) onto an index value \(I\).
+This is commonly known as the projection pursuit index function, or just index function, and is used to measure the ``interestingness'' of a projection.
+
 A number of index functions have been proposed in the literature to detect different data structures, including Legendre index \citep{friedman1974projection}, Hermite index \citep{hall1989polynomial}, natural Hermite index \citep{cook1993projection}, chi-square index \citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA index \citep{lee2010projection}.
 
-As a general visualisation method, A tour produces animated visualisation of the high dimensional data via rotating low dimension planes.
-Different tours choose the low dimensional planes differently, for example, a grand tour\citep{cook2008grand} selects the planes randomly in the high dimensional space and manual tour\citep{cook1997manual} chooses a sequence of bases that gradually phase in and out one variable to view the contribution of that variable in the visualisation.
-Guided tour, the main interest of this paper, chooses its projection with the aid of projection pursuit.
-Given a random start, projection pursuit iteratively finds bases with higher index value and guided tour constructs geodesic interpolation between two consecutive to form a tour path.
+As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
+Different tour types choose these planes differently, for example, a grand tour\citep{cook2008grand} selects the planes randomly and a manual tour\citep{cook1997manual} gradually phases in and out one variable, to understand the contribution of that variable in the projection.
+Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
+Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
+
 Mathematical details of the geodesic interpolation can be found in \citet{buja2005computational}.
-Figure \ref{fig:tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, the interpolations between two blue frames.
+Figure \ref{fig:tour-path} shows a sketch of the tour path.
+The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
 The tour method has been implemented in the \texttt{tourr} package in R, available on the Comprehensive R Archive Network at \url{https://cran.r-project.org/web/packages/tourr/} \citep{wickham2011tourrpackage}.
@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-better-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-random-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
 

### 2020-10-28 paper/paper.Rmd
update paper.Rmd
removed name T for linear projection + minor corrections (up to section 5)

@@
 
-For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
+For a given problem we might rely on numerical methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
 An iterative approach finds the optimum by progressively improving the current solution.
@@
 Let $\mathbf{X}_{n \times p}$ be the data matrix, with $n$ observations in $p$ dimensions.
-A d-dimensional projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
+A d-dimensional projection can be seen as a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
 Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$.
@@
 As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
-Different tour types choose these planes differently, for example, a grand tour[@cook2008grand] selects the planes randomly and a manual tour[@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
+Different tour types choose these planes differently, for example, a grand tour [@cook2008grand] selects the planes randomly and a manual tour [@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
 Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
@@
 
-where $I_d$ is the d-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be an orthogonal matrices.
+where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthogonal matrices.
 
 Several features of this optimisation are worth noticing.
-First of all, this is a constraint optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
+First of all, this is a constrained optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
 It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
@@
 
-Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allow a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
+Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allows a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
 Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
@@
 This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when local maximum is presented.
+This feature is particularly useful when local maxima are present.
 The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
@@
 @cook1995grand used a gradient ascent algorithm on the space of the projection bases.
-In gradient ascent, one first find the direction for improvement via computing the gradient information.
+In gradient ascent, one first finds the direction for improvement via computing the gradient information.
 In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
@@
 In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
-Possible values for `search_better` and `search_better_random` includes `random_search`,  `new_basis`, and `interpolation`.
+Possible values for `search_better` and `search_better_random` include `random_search`,  `new_basis`, and `interpolation`.
 `search_geodesic` has a wider variety that includes `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
@@
 $t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
-$l$ is the counter for each search/interpolation, which provides the information of how many basis the algorithm has searched before finding one to return.
-There are other parameters of interest, depends on the particular problem content and they are denoted as *$V_{p}$*.
+$l$ is the counter for each search/interpolation, which provides the information of how many bases the algorithm has searched before finding one to return.
+There are other parameters of interest, which depend on the particular problem content and they are denoted as *$V_{p}$*.
 Two most common examples include $V_1 = \text{method}$, which tags the name of the algorithm used, and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
@@
 where $T^{\prime} = T + k_{J}+ l_{J+1}$.
-Note that there is no output in iteration $J + 1$ since the optimiser can't find a better basis and the algorithm terminates.
-In this notation, final basis found is $A_T$ with the highest index value $I_T$.
+Note that there is no output in iteration $J + 1$ since the fact that it is the last iteration means that the optimiser cannot find a better basis and the algorithm terminates.
+In this notation, the final basis found is $A_T$ with the highest index value $I_T$.
 
@@
 The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation forms a row and each variable forms a column.
-With tidy data structure, data wrangling and visualisation have been significantly simplified by the well-developed packages like dplyr[@dplyr] and ggplot2[@ggplot2].
+With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as dplyr[@dplyr] and ggplot2[@ggplot2].
 
-The construction of diagnostic plots uses the concept of grammar of graphic [@wickham2010layered] in ggplot2.
-In grammar of graphic, plots are not defined by its appearance (i.e.
+The construction of diagnostic plots uses the concept of grammar of graphics [@wickham2010layered] in ggplot2.
+In grammar of graphics, plots are not defined by their appearance (i.e.
 boxplot, histogram, scatter plot etc) but by stacked layers.
-In the construction of diagnostic plots, there are multiple elements one may wants to emphasize and there's no single plot name that would meet the need, the stacked layers concept, on the other hand, allow information to be overlaid on the each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
+In the construction of diagnostic plots, there are multiple elements one may want to emphasize and there is no single plot name that would meet the need.
+The stacked layers concept, on the other hand, allows information to be overlaid on each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
 
-## Check how hard the optimiser is working
+## Checking how hard the optimiser is working
 
@@
 Boxplot is a suitable candidate that can provide five points summary of each iteration.
-Other additional information not presented in the  boxplot and then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
+Other additional information not presented in the  boxplot can then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
 Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a `cutoff` parameter.
 
-Figure \@ref(fig:toy-search) shows a sample of the plot constructed using `search_better` with different value on the parameter `max_tries`.
+Figure \@ref(fig:toy-search) shows a sample of the plot constructed using `search_better` with different values of the parameter `max_tries`.
 Comparing the returned index value of each iteration shows that the termination at `max_tries = 25` is not sufficient for `search_better` to explore the parameter space and a value of 500 is preferred over 25 in this context.
@@
 
-(ref:toy-search) A comparison of `search_better` with different parameter value on `max_tries`.
-A six variable dataset `boa6` is used with `holes` index on a 2D problem.
-A `max_tries = 500` since it allows the optimiser to find better basis with higher index value at iteration six.
+(ref:toy-search) A comparison of `search_better` with different values of the parameter `max_tries`.
+A six-variable dataset `boa6` is used with `holes` index on a 2D problem.
+`max_tries = 500` allows the optimiser to find better basis with higher index value at iteration six.
 
@@
 The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
-The middle path has an increase-then-decreases pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
-The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost cases is a deliberate construction of `search_better_random` where an inferior basis can be accepted in a probability manner so as to avoid trapping in the local maximum.
+The middle path has an increase-then-decrease pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
+The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost case is a deliberate construction of `search_better_random` where an inferior basis can be accepted with some probability so as to avoid being trapped in a local maximum.
 
@@
 
-(ref:toy-pca) 1D projection on the 5 variable dataset `boa5` with two optimisers: `search_better` and `search_geodesic`.
+(ref:toy-pca) 1D projection on the 5-variable dataset `boa5` with two optimisers: `search_better` and `search_geodesic`.
 The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
@@
 The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in `search_geodesic` has been flipped positive to ensure that bases with same projected images but a sign difference are represented by the same point in the plot.
+They are initially simulated with the same starting points but all the bases in `search_geodesic` has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.
 
 Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go over two or three.
-While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (d > 3), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.
+Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ live on the surface of a $p \times d$ dimension sphere, where the dimension can easily go over two or three.
+While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space ($d > 3$), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.
 principal component analysis.
-To better perceive the the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
+To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
 
-Figure \@ref(fig:toy-pca) plots the first two principal components of two search paths, one using `search_better` and another uses `search_geodesic`.
-The search in the space reduced by PCA matches with the optimiser description before where the random sampling in `search_better` is broader, controlled by `alpha` parameter, which is default to 0.5 while the directional search in `search_geodesic`, controlled by `delta` with a default of 0.01, is so tiny that it can barely seen.
+Figure \@ref(fig:toy-pca) plots the first two principal components of two search paths, one using `search_better` and the other using `search_geodesic`.
+The search in the space reduced by PCA matches with the optimiser description before where the random sampling in `search_better` is broader, controlled by `alpha` parameter, which is default to 0.5 while the directional search in `search_geodesic`, controlled by `delta` with a default of 0.01, is so tiny that it can barely be seen.
 
@@
 (ref:toy-tour) A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row view the space from a direction that is close to the one in PCA plot.
+The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
 The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
 
-Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a more holistic stereoscopical view of the search.
+While viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a more holistic stereoscopic view of the search.
 To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compare to the PCA plot, the animated rotation (tour) display in Figure \@ref(fig:toy-tour) gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
-Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and grand tour produces a sequence of angle that view the search from different direction.
-As an evidence, the last frame in Figure \@ref(fig:toy-tour) is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to one in Figure \@ref(fig:toy-pca).
+Compared to the PCA plot, the animated rotation (tour) displayed in Figure \@ref(fig:toy-tour) gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
+As an evidence, the last frame in Figure \@ref(fig:toy-tour) is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \@ref(fig:toy-pca).
 
@@
 For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
-If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
+If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
 When multiple optima are presented, `search_better` may stuck at the local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.

### 2020-10-29 paper/paper.Rmd
update paper.Rmd
minor corrections from Section 5 to the end

@@
 If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are presented, `search_better` may stuck at the local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+When multiple optima are present, `search_better` may get stuck at a local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
 In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
@@
 
-Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Random variables with different structures have been simulated and the distribution of each is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 Variable `x1`, `x8`, `x9` and `x10` are normal distributed with zero mean and unit variance and `x2` to `x7` are mixtures of normal distributions with varied weights and locations.
@@
 
-## A problem of not monotonic {#monotonic}
+## A problem of non-monotonicity {#monotonic}
 
 In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
-This pattern is undesirable since the optimiser could have start the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
+This pattern is undesirable since the optimiser could have started the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
 This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \@ref(fig:interruption).
+After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \@ref(fig:interruption).
 
@@
 (ref:interruption) Two-D projection on `boa6` data with holes index optimised by `search_better`.
-A comparison of the trace plot with and without the interruption.
+A comparison of the trace plot with and without interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
@@
 
-Once the final basis has been found by an algorithm, one may want to push further to investigate whether there's an even better basis in the close neighbourhood.
+Once the final basis has been found by an algorithm, one may want to push further to investigate whether there is an even better basis in the close neighbourhood.
 This motivates the polish search where the final basis is supplied as the start of a new guided tour to search for any local breakthrough.
@@
 2) the percentage change of the index value need to be larger than 1e-5; and
-3) the alpha parameter on itself need to be larger than 0.01
+3) the alpha parameter on itself needs to be larger than 0.01
 
@@
 The grey dashed line represents the index value of the theoretical best basis.
-`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of theoretical best basis.
+`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of the theoretical best basis.
 
@@
 `norm_kol`, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{Y}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as $F_{.}(u)$ with the subscript indicating the projection or the random normal variable,  the `norm_kol` index is defined by
+Denote the ECDF function as $F_{.}(u)$ with the subscript indicating either the projection or the random normal variable,  the `norm_kol` index is defined by
 
@@
 Figure \@ref(fig:noisy-better-geo) compares the tracing plot of two optimisers: `search_geodesic` and `search_better`.
-This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely improvement of the index value.
-On the other hand, `search_better` is doing a relatively well.
-With theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] produces an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
+This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely any improvement of the index value.
+On the other hand, `search_better` is doing relatively well.
+With the theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] producing an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
 A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 3)`].
-At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumptions in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
@@
 (ref:kol-random) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
-`search_better_random` has a probabilistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the global maximum.
+`search_better_random` has a probabilistic acceptance implementation that would also accept a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.
 
@@
 
-The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is  presented.
+The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
 The dataset used is `boa6` where `x2` and `x7` are informative.
@@
 The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find `V7` because they are born in Rome.
-On the other end, the ones on the first two rows can only find `V2`, largely due to the fact that it doesn't even have chances to search near `V7`.
+On the other end, the ones on the first two rows can only find `V2`, largely due to the fact that they do not even have chances to search near `V7`.
 
@@
 This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \@ref(fig:kol-random)).
-The two cases that neither the algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \@ref(fig:kol-better) and position (1,5) and (1,3) in Figure \@ref(fig:kol-random)).
+The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \@ref(fig:kol-better) and position (1,5) and (1,3) in Figure \@ref(fig:kol-random)).
 This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near `V7` to be adequately explored.
 
-In Figure \ref{kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated basis and hence allows a wider search.
+In Figure \@ref{fig:kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated bases and hence allows a wider search.
 Seed 9982 ((2,4) in Figure \@ref(fig:kol-random-tuned)) and seed 2463 ((3,2) in Figure \@ref(fig:kol-random-tuned)) benefit from this and find the global maximum.
 A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
-In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, the allows a near `V2` bases to switch to a near `V7` bases and results these cases to find the global maximum.
-While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near `V7` basis to near `V2` basis and it doesn't come back and leave those simulations in the local maximum.
+In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, this allows a near `V2` bases to switch to a near `V7` bases and as a result these cases find the global maximum.
+While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near `V7` basis to near `V2` basis, it does not come back and leaves those simulations in the local maximum.
 
-The conclusion from this experiment is that the usage of `search_better_random` and increase the search space are methods that can avoid trapping in local maximum but the best solution will depends on the starting points of the simulation and the seed used.
+The conclusion from this experiment is that the usage of `search_better_random` and increasing the search space are methods that can avoid getting trapped in local maxima but the solution will still depend on the starting points of the simulation and the seed used.
 

### 2020-10-29 paper/paper.Rmd
Working version section 1-1

@@
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+  \usepackage{todonotes}
 output:
@@
 Visualisation is widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of.
-This motivates our work of creating plots to diagnose optimisation algorithms: we aim to understand and compare the behaviour of different algorithms.
+Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
+The work presented in this paper is at the intersection of data visualization and optimization methods and aims to bring visualization tools into optimization problems to assist practitioners to better understand the performance of their optimization methods in practice.
+This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms \todo[inline]{(THERE ARE TWO THINGS HERE --> VISUAL METHODS + THE IMPLEMENTATION OF FINE TUNNING METHODS RIGHT ?}.
+\todo[inline, color = green]{Add intro to fine tunning methods here too to set the scene}
+The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
+Broadly speaking optimization can be unconstrained or constrained.
+The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
+$min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with  certain properties defined in $f\in L^p$ .
+In this case typically solutions rely on gradient descent or ascent methods.
+While in the constrained problem additional restrictions  are introduced via a set of functions that can be convex or non-convex $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the optimization problem in this case can be written as
+$min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods includig linear and quadratic programming can be used.
+<!-- $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.  -->
 
-The goal of optimisation is to find the best solution within the space of all feasible solutions, typically represented by a set of constraints.
-The problem can be formulated as the maximisation (or minimisation) of an objective function $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.
-
-The focus of this paper is on optimisation in the projection pursuit guided tour, which is an exploratory data analysis tool that detects interesting structure in high-dimensional data visible in low-dimensional projections.
-The target of the optimisation is thus to identify the most interesting low-dimensional view, given by a corresponding projection matrix.
+The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
+The target of the optimisation is thus to identify the most interesting low-dimensional views of the data given by a corresponding projection matrix.
+The most interesting structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space.
+\todo[inline]{CHALLENGES here that can also be set as general optimization challenges.
+However this depends on the sucess of index optimization, right?}
+However, the optimization challenges encounter inthe projection pursuit guided tour problem are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
 
@@
 
-In this paper, the main focus is on the  *line search method* [@fletcher2013practical].
-In a simple one-dimensional problem, finding the value of $x$ that minimises the function $f(x)$, the line search takes iterative steps as in Equation \ref{eq:line-search},
+Optimization problems are present in many areas of study and even though in some cases, analytical solutions can be found, the majority of optimization problems rely on numerical methods to find the optimal solution to a problem.
+In optimization, numerical methods follow interative approaches that aims at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
 
-\begin{equation}
-x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)},
-\label{eq:line-search}
-\end{equation}
+In this paper we focus on unconstrained optimization as defined in section XX and assume that it is not possible an analytical solution in closed form.
+That is, our problem consists of finding the mininum of function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
+In this scenario, gradient descent methods have taken the stage for many decades.
+In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate.
+The step size is the key in the sucess of the optimization problem here as inadequate step
+sizes might lead to a failure in convergence.
+In order to solve this problem line minimization or line search methods are proposed (REF) where the step size is chosen in such a way that convergence is ensured.
+In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is stablished $S_k$ then the objective is to find
+the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear).
+Then move on to the next pont following $x_{k+1} = x_k + \nu_kS_k$
+\todo[inline]{lconnection with derivative free methods}
+ **From Fletcher's book next paragraph --> clean up** Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
+This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical].
+A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$.
+Similarly, for  minimisation we would require a decrease in function value at each step.
+This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
 
-where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
+Modern development of line search methods focuses on improving the computation of the searching direction $S_{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
+Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
+
+\todo[inline, color = pink]{Question for Sherry --> which implementaion is your search better? search better random? and search geodesic?}
+
+\todo[inline]{line search--> search better,  search geodesic}
+
+<!-- For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems. -->
+<!-- An iterative approach finds the optimum by progressively improving the current solution.
+Researchers can then decide when to stop if a desirable accuracy has been achieved.
+One potential issue with this technique is that it may end up finding a local optimum.
+-->
+
+<!-- In this paper, the main focus is on the  *line search method* [@fletcher2013practical].
+In a simple one-dimensional problem, finding the value of $x$ that minimises the function $f(x)$, the line search takes iterative steps as in Equation \ref{eq:line-search}, -->
+
+<!-- \begin{equation} -->
+<!-- x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)}, -->
+<!-- \label{eq:line-search} -->
+<!-- \end{equation} -->
+
+<!-- where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
 Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
@@
 This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
+-->
 
-Modern development of line search methods focuses on improving the computation of the searching direction $d^{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
+<!-- Modern development of line search methods focuses on improving the computation of the searching direction $d^{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
 Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
+-->
 

### 2020-10-29 paper/paper.tex
Working version section 1-1

@@
 %% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e} \usepackage{todonotes}
 
@@
 Visualisation is widely used in exploratory data analysis.
-Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of.
-This motivates our work of creating plots to diagnose optimisation algorithms: we aim to understand and compare the behaviour of different algorithms.
+Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
+The work presented in this paper is at the intersection of data visualization and optimization methods and aims to bring visualization tools into optimization problems to assist practitioners to better understand the performance of their optimization methods in practice.
+This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms \todo[inline]{(THERE ARE TWO THINGS HERE --> VISUAL METHODS + THE IMPLEMENTATION OF FINE TUNNING METHODS RIGHT ?}.
+\todo[inline, color = green]{Add intro to fine tunning methods here too to set the scene}
+The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
+Broadly speaking optimization can be unconstrained or constrained.
+The unconstrained problem can be formulated as a minimization (or maximization) problem such as
+\(min_{x} f(x)\) where \(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with certain properties defined in \(f\in L^p\) .
+In this case typically solutions rely on gradient descent or ascent methods.
+While in the constrained problem additional restrictions are introduced via a set of functions that can be convex or non-convex \(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and hence the optimization problem in this case can be written as
+\(min_{x} f(x)\) \emph{subject to} \(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex optimization methods includig linear and quadratic programming can be used.
 
-The goal of optimisation is to find the best solution within the space of all feasible solutions, typically represented by a set of constraints.
-The problem can be formulated as the maximisation (or minimisation) of an objective function \(f: S \rightarrow \mathbb{R}\), with \(S \in \mathbb{R}^n\), in a reduced space given by the problem constraints on \(f\).
-
-The focus of this paper is on optimisation in the projection pursuit guided tour, which is an exploratory data analysis tool that detects interesting structure in high-dimensional data visible in low-dimensional projections.
-The target of the optimisation is thus to identify the most interesting low-dimensional view, given by a corresponding projection matrix.
+The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect \emph{interesting structures} or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
+The target of the optimisation is thus to identify the most interesting low-dimensional views of the data given by a corresponding projection matrix.
+The most interesting structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space.
+\todo[inline]{CHALLENGES here that can also be set as general optimization challenges.
+However this depends on the sucess of index optimization, right?}
+However, the optimization challenges encounter inthe projection pursuit guided tour problem are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
 
@@
 
-For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
-An iterative approach finds the optimum by progressively improving the current solution.
-Researchers can then decide when to stop if a desirable accuracy has been achieved.
-One potential issue with this technique is that it may end up finding a local optimum.
+Optimization problems are present in many areas of study and even though in some cases, analytical solutions can be found, the majority of optimization problems rely on numerical methods to find the optimal solution to a problem.
+In optimization, numerical methods follow interative approaches that aims at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
 
-In this paper, the main focus is on the \emph{line search method} \citep{fletcher2013practical}.
-In a simple one-dimensional problem, finding the value of \(x\) that minimises the function \(f(x)\), the line search takes iterative steps as in Equation \ref{eq:line-search},
-
-\begin{equation}
-x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)},
-\label{eq:line-search}
-\end{equation}
-
-where \(d^{(j)}\) is the searching direction in iteration \(j\), and \(\alpha_j\) is the step-size.
-Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
+In this paper we focus on unconstrained optimization as defined in section XX and assume that it is not possible an analytical solution in closed form.
+That is, our problem consists of finding the mininum of function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
+In this scenario, gradient descent methods have taken the stage for many decades.
+In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate.
+The step size is the key in the sucess of the optimization problem here as inadequate step
+sizes might lead to a failure in convergence.
+In order to solve this problem line minimization or line search methods are proposed (REF) where the step size is chosen in such a way that convergence is ensured.
+In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is stablished \(S_k\) then the objective is to find
+the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear).
+Then move on to the next pont following \(x_{k+1} = x_k + \nu_kS_k\)
+\todo[inline]{lconnection with derivative free methods}
+\textbf{From Fletcher's book next paragraph --\textgreater{} clean up} Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
 This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist \citep{fletcher2013practical}.
@@
 
-Modern development of line search methods focuses on improving the computation of the searching direction \(d^{(j)}\), and on approximations of the step size \(\alpha_j\), catering for practical optimisation problems.
+Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(j)}\), and on approximations of the step size \(\alpha_j\), catering for practical optimisation problems.
 Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
+
+\todo[inline, color = pink]{Question for Sherry --> which implementaion is your search better? search better random? and search geodesic?}
+
+\todo[inline]{line search--> search better,  search geodesic}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/pmen0008/Desktop/paper-tour-vis/img/tour_path}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-better-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-1}
 
@@
 
-{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
+{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-tuned-1}
 

### 2020-11-02 paper/paper.Rmd
Update on Optmisation methods section

@@
 
-Visualisation is widely used in exploratory data analysis.
+Visualisation is widely used in exploratory data analysis (REF).
 Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-The work presented in this paper is at the intersection of data visualization and optimization methods and aims to bring visualization tools into optimization problems to assist practitioners to better understand the performance of their optimization methods in practice.
-This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms \todo[inline]{(THERE ARE TWO THINGS HERE --> VISUAL METHODS + THE IMPLEMENTATION OF FINE TUNNING METHODS RIGHT ?}.
-\todo[inline, color = green]{Add intro to fine tunning methods here too to set the scene}
+The work presented in this paper combines data visualization with optimization methods.
+It aims at bring visualization tools into optimization problems to assist users to better understand the performance of their optimization methods in practice.
+This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms.
+
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained.
+Broadly speaking optimization can be unconstrained or constrained (REF).
 The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
-$min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with  certain properties defined in $f\in L^p$ .
-In this case typically solutions rely on gradient descent or ascent methods.
-While in the constrained problem additional restrictions  are introduced via a set of functions that can be convex or non-convex $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the optimization problem in this case can be written as
-$min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods includig linear and quadratic programming can be used.
+$min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with  certain properties defined in an $L^p$ space.
+In this case, solutions rely on gradient descent or ascent methods.
+In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as
+$min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
 <!-- $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.  -->
@@
 The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
-The target of the optimisation is thus to identify the most interesting low-dimensional views of the data given by a corresponding projection matrix.
-The most interesting structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space.
-\todo[inline]{CHALLENGES here that can also be set as general optimization challenges.
-However this depends on the sucess of index optimization, right?}
-However, the optimization challenges encounter inthe projection pursuit guided tour problem are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
+The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
+The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (REF).
+
+The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
+Examples of those include the existance of multiple maxima (local and global), the tradeof between computational burden and proximity to the maxima,  dealing with noisy objective functions, that might be non smooth and non diferentiable (REF).
+Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
 
@@
 
-For a given problem we might rely on numerical methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
-An iterative approach finds the optimum by progressively improving the current solution.
+<!-- For a given problem we might rely on numerical methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
+-->
+<!-- An iterative approach finds the optimum by progressively improving the current solution.
 Researchers can then decide when to stop if a desirable accuracy has been achieved.
 One potential issue with this technique is that it may end up finding a local optimum.
+-->
 
-Optimization problems are present in many areas of study and even though in some cases, analytical solutions can be found, the majority of optimization problems rely on numerical methods to find the optimal solution to a problem.
-In optimization, numerical methods follow interative approaches that aims at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
+Optimization problems are ubiquitous in many areas of study.
+While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem.
+These numerical methods follow interative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the tradeoff between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-In this paper we focus on unconstrained optimization as defined in section XX and assume that it is not possible an analytical solution in closed form.
-That is, our problem consists of finding the mininum of function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
-In this scenario, gradient descent methods have taken the stage for many decades.
-In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate.
-The step size is the key in the sucess of the optimization problem here as inadequate step
+In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
+That is, our problem consists of finding the mininum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
+In this scenario, gradient descent methods have taken the stage for many decades now.
+In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
+The step size is the key in the sucess of the optimization problem as inadequate step
 sizes might lead to a failure in convergence.
-In order to solve this problem line minimization or line search methods are proposed (REF) where the step size is chosen in such a way that convergence is ensured.
+In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
 In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is stablished $S_k$ then the objective is to find

### 2020-11-02 paper/paper.tex
Update on Optmisation methods section

@@
 
-Visualisation is widely used in exploratory data analysis.
+Visualisation is widely used in exploratory data analysis (REF).
 Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-The work presented in this paper is at the intersection of data visualization and optimization methods and aims to bring visualization tools into optimization problems to assist practitioners to better understand the performance of their optimization methods in practice.
-This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms \todo[inline]{(THERE ARE TWO THINGS HERE --> VISUAL METHODS + THE IMPLEMENTATION OF FINE TUNNING METHODS RIGHT ?}.
-\todo[inline, color = green]{Add intro to fine tunning methods here too to set the scene}
+The work presented in this paper combines data visualization with optimization methods.
+It aims at bring visualization tools into optimization problems to assist users to better understand the performance of their optimization methods in practice.
+This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms.
+
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained.
+Broadly speaking optimization can be unconstrained or constrained (REF).
 The unconstrained problem can be formulated as a minimization (or maximization) problem such as
-\(min_{x} f(x)\) where \(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with certain properties defined in \(f\in L^p\) .
-In this case typically solutions rely on gradient descent or ascent methods.
-While in the constrained problem additional restrictions are introduced via a set of functions that can be convex or non-convex \(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and hence the optimization problem in this case can be written as
-\(min_{x} f(x)\) \emph{subject to} \(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex optimization methods includig linear and quadratic programming can be used.
+\(min_{x} f(x)\) where \(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with certain properties defined in an \(L^p\) space.
+In this case, solutions rely on gradient descent or ascent methods.
+In the constrained optimization problem additional restrictions are introduced via a set of functions that can be convex or non-convex: \(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and hence the problem can be written as
+\(min_{x} f(x)\) \emph{subject to} \(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
 
 The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect \emph{interesting structures} or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
-The target of the optimisation is thus to identify the most interesting low-dimensional views of the data given by a corresponding projection matrix.
-The most interesting structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space.
-\todo[inline]{CHALLENGES here that can also be set as general optimization challenges.
-However this depends on the sucess of index optimization, right?}
-However, the optimization challenges encounter inthe projection pursuit guided tour problem are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
+The target of the optimisation is to identify the most \emph{interesting} low-dimensional views of the data given by a corresponding projection matrix.
+The most \emph{interesting} structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (REF).
+
+The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
+Examples of those include the existance of multiple maxima (local and global), the tradeof between computational burden and proximity to the maxima, dealing with noisy objective functions, that might be non smooth and non diferentiable (REF).
+Those are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
 
@@
 
-Optimization problems are present in many areas of study and even though in some cases, analytical solutions can be found, the majority of optimization problems rely on numerical methods to find the optimal solution to a problem.
-In optimization, numerical methods follow interative approaches that aims at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
+Optimization problems are ubiquitous in many areas of study.
+While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem.
+These numerical methods follow interative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the tradeoff between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-In this paper we focus on unconstrained optimization as defined in section XX and assume that it is not possible an analytical solution in closed form.
-That is, our problem consists of finding the mininum of function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
-In this scenario, gradient descent methods have taken the stage for many decades.
-In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate.
-The step size is the key in the sucess of the optimization problem here as inadequate step
+In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
+That is, our problem consists of finding the mininum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
+In this scenario, gradient descent methods have taken the stage for many decades now.
+In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
+The step size is the key in the sucess of the optimization problem as inadequate step
 sizes might lead to a failure in convergence.
-In order to solve this problem line minimization or line search methods are proposed (REF) where the step size is chosen in such a way that convergence is ensured.
+In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
 In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is stablished \(S_k\) then the objective is to find
@@
 Let \(\mathbf{X}_{n \times p}\) be the data matrix, with \(n\) observations in \(p\) dimensions.
-A d-dimensional projection can be seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\), and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection matrix.
+A d-dimensional projection can be seen as a linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\), and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection matrix.
 Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projected data \(\mathbf{Y}\) (corresponding to an associated projection matrix \(\mathbf{A}\)) onto an index value \(I\).
@@
 As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
-Different tour types choose these planes differently, for example, a grand tour\citep{cook2008grand} selects the planes randomly and a manual tour\citep{cook1997manual} gradually phases in and out one variable, to understand the contribution of that variable in the projection.
+Different tour types choose these planes differently, for example, a grand tour \citep{cook2008grand} selects the planes randomly and a manual tour \citep{cook1997manual} gradually phases in and out one variable, to understand the contribution of that variable in the projection.
 Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
@@
 
-where \(I_d\) is the d-dimensional identity matrix and the constraint requires the projection bases \(\mathbf{A}\) to be an orthogonal matrices.
+where \(I_d\) is the \(d\)-dimensional identity matrix and the constraint requires the projection bases \(\mathbf{A}\) to be orthogonal matrices.
 
 Several features of this optimisation are worth noticing.
-First of all, this is a constraint optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
+First of all, this is a constrained optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
 It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
@@
 
-Simulated annealing (\texttt{search\_better\_random}) \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allow a probabilistic acceptance of a basis with lower index value based on the annealing \(T(l)\).
+Simulated annealing (\texttt{search\_better\_random}) \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allows a probabilistic acceptance of a basis with lower index value based on the annealing \(T(l)\).
 Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
@@
 \citet{cook1995grand} used a gradient ascent algorithm on the space of the projection bases.
-In gradient ascent, one first find the direction for improvement via computing the gradient information.
+In gradient ascent, one first finds the direction for improvement via computing the gradient information.
 In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
@@
 In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3) State: \(S\), which labels the observation with detailed stage in the optimisation.
-Possible values for \texttt{search\_better} and \texttt{search\_better\_random} includes \texttt{random\_search}, \texttt{new\_basis}, and \texttt{interpolation}.
+Possible values for \texttt{search\_better} and \texttt{search\_better\_random} include \texttt{random\_search}, \texttt{new\_basis}, and \texttt{interpolation}.
 \texttt{search\_geodesic} has a wider variety that includes \texttt{new\_basis}, \texttt{direction\_search}, \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and \texttt{interpolation}.
@@
 \(t\) is a unique identifier that prescribes the natural ordering of each observation; \(j\) is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
-\(l\) is the counter for each search/interpolation, which provides the information of how many basis the algorithm has searched before finding one to return.
-There are other parameters of interest, depends on the particular problem content and they are denoted as \emph{\(V_{p}\)}.
+\(l\) is the counter for each search/interpolation, which provides the information of how many bases the algorithm has searched before finding one to return.
+There are other parameters of interest, which depend on the particular problem content and they are denoted as \emph{\(V_{p}\)}.
 Two most common examples include \(V_1 = \text{method}\), which tags the name of the algorithm used, and \(V_2 = \text{alpha}\), the neighbourhood parameter that controls the size in sampling candidate bases.
@@
 where \(T^{\prime} = T + k_{J}+ l_{J+1}\).
-Note that there is no output in iteration \(J + 1\) since the optimiser can't find a better basis and the algorithm terminates.
-In this notation, final basis found is \(A_T\) with the highest index value \(I_T\).
+Note that there is no output in iteration \(J + 1\) since the fact that it is the last iteration means that the optimiser cannot find a better basis and the algorithm terminates.
+In this notation, the final basis found is \(A_T\) with the highest index value \(I_T\).
 
 The data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that requires each observation forms a row and each variable forms a column.
-With tidy data structure, data wrangling and visualisation have been significantly simplified by the well-developed packages like dplyr\citep{dplyr} and ggplot2\citep{ggplot2}.
+With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as dplyr\citep{dplyr} and ggplot2\citep{ggplot2}.
 
-The construction of diagnostic plots uses the concept of grammar of graphic \citep{wickham2010layered} in ggplot2.
-In grammar of graphic, plots are not defined by its appearance (i.e.~boxplot, histogram, scatter plot etc) but by stacked layers.
-In the construction of diagnostic plots, there are multiple elements one may wants to emphasize and there's no single plot name that would meet the need, the stacked layers concept, on the other hand, allow information to be overlaid on the each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
+The construction of diagnostic plots uses the concept of grammar of graphics \citep{wickham2010layered} in ggplot2.
+In grammar of graphics, plots are not defined by their appearance (i.e.~boxplot, histogram, scatter plot etc) but by stacked layers.
+In the construction of diagnostic plots, there are multiple elements one may want to emphasize and there is no single plot name that would meet the need.
+The stacked layers concept, on the other hand, allows information to be overlaid on each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
 
-\hypertarget{check-how-hard-the-optimiser-is-working}{%
-\subsection{Check how hard the optimiser is working}\label{check-how-hard-the-optimiser-is-working}}
+\hypertarget{checking-how-hard-the-optimiser-is-working}{%
+\subsection{Checking how hard the optimiser is working}\label{checking-how-hard-the-optimiser-is-working}}
 
@@
 Boxplot is a suitable candidate that can provide five points summary of each iteration.
-Other additional information not presented in the boxplot and then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
+Other additional information not presented in the boxplot can then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
 Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a \texttt{cutoff} parameter.
 
-Figure \ref{fig:toy-search} shows a sample of the plot constructed using \texttt{search\_better} with different value on the parameter \texttt{max\_tries}.
+Figure \ref{fig:toy-search} shows a sample of the plot constructed using \texttt{search\_better} with different values of the parameter \texttt{max\_tries}.
 Comparing the returned index value of each iteration shows that the termination at \texttt{max\_tries\ =\ 25} is not sufficient for \texttt{search\_better} to explore the parameter space and a value of 500 is preferred over 25 in this context.
@@
 
-\caption{A comparison of \texttt{search\_better} with different parameter value on \texttt{max\_tries}.
-A six variable dataset \texttt{boa6} is used with \texttt{holes} index on a 2D problem.
-A \texttt{max\_tries\ =\ 500} since it allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
+\caption{A comparison of \texttt{search\_better} with different values of the parameter \texttt{max\_tries}.
+A six-variable dataset \texttt{boa6} is used with \texttt{holes} index on a 2D problem.
+\texttt{max\_tries\ =\ 500} allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
 \end{figure}
@@
 The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
-The middle path has an increase-then-decreases pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
-The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost cases is a deliberate construction of \texttt{search\_better\_random} where an inferior basis can be accepted in a probability manner so as to avoid trapping in the local maximum.
+The middle path has an increase-then-decrease pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
+The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost case is a deliberate construction of \texttt{search\_better\_random} where an inferior basis can be accepted with some probability so as to avoid being trapped in a local maximum.
 
@@
 
-\caption{1D projection on the 5 variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
+\caption{1D projection on the 5-variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
 The yellow point corresponds to the theoretical best basis {[}0, 1, 0, 0, 0{]} with \texttt{V2} being the only non-normal variable in the dataset.
@@
 The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
+They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
 \end{figure}
@@
 Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go over two or three.
-While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (d \textgreater{} 3), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.~principal component analysis.
-To better perceive the the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
+Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) live on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go over two or three.
+While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (\(d > 3\)), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.~principal component analysis.
+To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
 
-Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using \texttt{search\_better} and another uses \texttt{search\_geodesic}.
-The search in the space reduced by PCA matches with the optimiser description before where the random sampling in \texttt{search\_better} is broader, controlled by \texttt{alpha} parameter, which is default to 0.5 while the directional search in \texttt{search\_geodesic}, controlled by \texttt{delta} with a default of 0.01, is so tiny that it can barely seen.
+Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using \texttt{search\_better} and the other using \texttt{search\_geodesic}.
+The search in the space reduced by PCA matches with the optimiser description before where the random sampling in \texttt{search\_better} is broader, controlled by \texttt{alpha} parameter, which is default to 0.5 while the directional search in \texttt{search\_geodesic}, controlled by \texttt{delta} with a default of 0.01, is so tiny that it can barely be seen.
 
@@
 \caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row view the space from a direction that is close to the one in PCA plot.
+The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
 The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}\label{fig:toy-tour}
@@
 
-Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a more holistic stereoscopical view of the search.
+While viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a more holistic stereoscopic view of the search.
 To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compare to the PCA plot, the animated rotation (tour) display in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
-Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and grand tour produces a sequence of angle that view the search from different direction.
-As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to one in Figure \ref{fig:toy-pca}.
+Compared to the PCA plot, the animated rotation (tour) displayed in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
+As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
 
@@
 For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
-If the index function is smooth and has single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are presented, \texttt{search\_better} may stuck at the local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
+If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
+When multiple optima are present, \texttt{search\_better} may get stuck at a local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
 In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
@@
 
-Random variables with different structures has been simulated and the distribution of each is presented in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Random variables with different structures have been simulated and the distribution of each is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are normal distributed with zero mean and unit variance and \texttt{x2} to \texttt{x7} are mixtures of normal distributions with varied weights and locations.
@@
 \hypertarget{monotonic}{%
-\subsection{A problem of not monotonic}\label{monotonic}}
+\subsection{A problem of non-monotonicity}\label{monotonic}}
 
 In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
-This pattern is undesirable since the optimiser could have start the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
+This pattern is undesirable since the optimiser could have started the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
 This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
+After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
 
@@
 \caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_better}.
-A comparison of the trace plot with and without the interruption.
+A comparison of the trace plot with and without interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
@@
 \item
-  the alpha parameter on itself need to be larger than 0.01
+  the alpha parameter on itself needs to be larger than 0.01
 \end{enumerate}
@@
 The grey dashed line represents the index value of the theoretical best basis.
-\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of theoretical best basis.}\label{fig:noisy-better-geo}
+\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of the theoretical best basis.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 \texttt{norm\_kol}, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
-Denotes the ECDF function as \(F_{.}(u)\) with the subscript indicating the projection or the random normal variable, the \texttt{norm\_kol} index is defined by
+Denote the ECDF function as \(F_{.}(u)\) with the subscript indicating either the projection or the random normal variable, the \texttt{norm\_kol} index is defined by
 
@@
 Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: \texttt{search\_geodesic} and \texttt{search\_better}.
-This time, the interpolated path is no longer smooth when using either algorithm and \texttt{search\_geodesic} fails to optimise this index with barely improvement of the index value.
-On the other hand, \texttt{search\_better} is doing a relatively well.
-With theoretical best basis {[}0, 1, 0, 0, 0{]} produces an index value of 0.182, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
+This time, the interpolated path is no longer smooth when using either algorithm and \texttt{search\_geodesic} fails to optimise this index with barely any improvement of the index value.
+On the other hand, \texttt{search\_better} is doing relatively well.
+With the theoretical best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.182, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
 A further polish step will give a marginal improvement of index value to 0.175 with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}.
-At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumption in Equation \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumptions in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
 
@@
 \caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
-\texttt{search\_better\_random} has a probabilistic acceptance implementation that would also a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more incidents find the global maximum.}\label{fig:kol-random}
+\texttt{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
 \end{figure}
@@
 
-The second experiment with the noisy index is to understand how the optimisers perform when the local maximum is presented.
+The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
 The dataset used is \texttt{boa6} where \texttt{x2} and \texttt{x7} are informative.
@@
 The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7} because they are born in Rome.
-On the other end, the ones on the first two rows can only find \texttt{V2}, largely due to the fact that it doesn't even have chances to search near \texttt{V7}.
+On the other end, the ones on the first two rows can only find \texttt{V2}, largely due to the fact that they do not even have chances to search near \texttt{V7}.
 
@@
 This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
-The two cases that neither the algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
+The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
 This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near \texttt{V7} to be adequately explored.
 
-In Figure \ref{kol-random-tuned}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated basis and hence allows a wider search.
+In Figure @ref\{fig:kol-random-tuned\}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
+Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated bases and hence allows a wider search.
 Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
 A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
-In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, the allows a near \texttt{V2} bases to switch to a near \texttt{V7} bases and results these cases to find the global maximum.
-While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near \texttt{V7} basis to near \texttt{V2} basis and it doesn't come back and leave those simulations in the local maximum.
+In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, this allows a near \texttt{V2} bases to switch to a near \texttt{V7} bases and as a result these cases find the global maximum.
+While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near \texttt{V7} basis to near \texttt{V2} basis, it does not come back and leaves those simulations in the local maximum.
 
-The conclusion from this experiment is that the usage of \texttt{search\_better\_random} and increase the search space are methods that can avoid trapping in local maximum but the best solution will depends on the starting points of the simulation and the seed used.
+The conclusion from this experiment is that the usage of \texttt{search\_better\_random} and increasing the search space are methods that can avoid getting trapped in local maxima but the solution will still depend on the starting points of the simulation and the seed used.
 

### 2020-11-15 paper/paper.Rmd
Editing second section

@@
 The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existance of multiple maxima (local and global), the tradeof between computational burden and proximity to the maxima,  dealing with noisy objective functions, that might be non smooth and non diferentiable (REF).
+Examples of those include the existance of multiple maxima (local and global), the tradeoff between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (REF).
 Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
@@
 While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem.
-These numerical methods follow interative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accurary is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the tradeoff between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
 In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
@@
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
-The step size is the key in the sucess of the optimization problem as inadequate step
-sizes might lead to a failure in convergence.
+In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
+The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
 In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is stablished $S_k$ then the objective is to find
-the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear).
+In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is established $S_k$ then the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear).
 Then move on to the next pont following $x_{k+1} = x_k + \nu_kS_k$
 \todo[inline]{lconnection with derivative free methods}
+
  **From Fletcher's book next paragraph --> clean up** Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].

### 2020-11-15 paper/paper.tex
Editing second section

@@
 The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existance of multiple maxima (local and global), the tradeof between computational burden and proximity to the maxima, dealing with noisy objective functions, that might be non smooth and non diferentiable (REF).
+Examples of those include the existance of multiple maxima (local and global), the tradeoff between computational burden and proximity to the maxima, dealing with noisy objective functions that might be non smooth and non diferentiable (REF).
 Those are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
@@
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
-The step size is the key in the sucess of the optimization problem as inadequate step
-sizes might lead to a failure in convergence.
+In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
+The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
 In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is stablished \(S_k\) then the objective is to find
-the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear).
+In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is established \(S_k\) then the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear).
 Then move on to the next pont following \(x_{k+1} = x_k + \nu_kS_k\)
 \todo[inline]{lconnection with derivative free methods}
+
 \textbf{From Fletcher's book next paragraph --\textgreater{} clean up} Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.

### 2020-11-15 paper/paper.Rmd
Updating section 2

@@
 In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
-That is, our problem consists of finding the mininum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
+That is, our problem consists of finding the minimum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
+In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
 In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is established $S_k$ then the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear).
-Then move on to the next pont following $x_{k+1} = x_k + \nu_kS_k$
-\todo[inline]{lconnection with derivative free methods}
-
- **From Fletcher's book next paragraph --> clean up** Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
-This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical].
-A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$.
-Similarly, for  minimisation we would require a decrease in function value at each step.
-This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
-
-Modern development of line search methods focuses on improving the computation of the searching direction $S_{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
+In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction $S_k$ is established then the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
+Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$.
+Modern development of line search methods focuses on improving the computation of the searching direction $S_{(k)}$, and on approximations of the step size $\alpha_k$, catering for practical optimisation problems.
 Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
 
-\todo[inline, color = pink]{Question for Sherry --> which implementaion is your search better? search better random? and search geodesic?}
+Derivative free --> adding
 
-\todo[inline]{line search--> search better,  search geodesic}
+ <!-- **From Fletcher's book next paragraph --> clean up** Theoretical results have -->
+ <!-- demonstrated the global convergence of the algorithm when the step-size -->
+ <!-- $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ -->
+ <!-- [@curry1944method]. This is however rarely implemented in practice, either -->
+ <!-- because of the high computational demand or because the solution does not exist -->
+ <!-- [@fletcher2013practical]. A modified approach, in the case of maximisation, is -->
+ <!-- to impose a mandatory increase in the objective function for each iteration: -->
+ <!-- $f^{(j+1)}> f^{(j)}$. Similarly, for  minimisation we would require a decrease -->
+ <!-- in function value at each step. This approach turns out to be efficient in -->
+ <!-- practical problems, but global convergence cannot be guaranteed in this case. -->
 

### 2020-11-15 paper/paper.tex
Updating section 2

@@
 In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
-That is, our problem consists of finding the mininum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
+That is, our problem consists of finding the minimum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
+In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
 In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is established \(S_k\) then the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear).
-Then move on to the next pont following \(x_{k+1} = x_k + \nu_kS_k\)
-\todo[inline]{lconnection with derivative free methods}
-
-\textbf{From Fletcher's book next paragraph --\textgreater{} clean up} Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}.
-This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist \citep{fletcher2013practical}.
-A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\).
-Similarly, for minimisation we would require a decrease in function value at each step.
-This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
-
-Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(j)}\), and on approximations of the step size \(\alpha_j\), catering for practical optimisation problems.
+In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction \(S_k\) is established then the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line search).
+Then move on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\).
+Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(k)}\), and on approximations of the step size \(\alpha_k\), catering for practical optimisation problems.
 Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
 
-\todo[inline, color = pink]{Question for Sherry --> which implementaion is your search better? search better random? and search geodesic?}
+Derivative free --\textgreater{} adding
 
-\todo[inline]{line search--> search better,  search geodesic}
+clean up** Theoretical results have --\textgreater{}
 

### 2020-11-17 paper/paper.Rmd
updating references

@@
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained (REF).
+Broadly speaking optimization can be unconstrained or constrained (@kelley1999iterative).
 The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
@@
 
-The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
+The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (@buja2005computational) which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (cook2008grand).
 The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
-The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (REF).
+The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (@cook1993projection).
 
 The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existance of multiple maxima (local and global), the tradeoff between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (REF).
+Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (@jones1998efficient).
 Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
@@
 These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
 In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
+In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find a solution in closed form.
 That is, our problem consists of finding the minimum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF).
+In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (@fletcher2013practical).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
-In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction $S_k$ is established then the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
-Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$.
+These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned (@trefethen1997numerical).
+In order to solve some of those problems, line minimization or line search methods were proposed (@shi2004convergence) where the step size is chosen in such a way that convergence is ensured.
+In line search methods users are required to provide an initial estimate $x_{1}$, a search direction $S_k$ for each iteration where the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
+Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and repeat the process until the desire convergence is reached.
 Modern development of line search methods focuses on improving the computation of the searching direction $S_{(k)}$, and on approximations of the step size $\alpha_k$, catering for practical optimisation problems.
-Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
-
-Derivative free --> adding
+Here we focus on line search methods for the optimisation used in the projection pursuit guided tour described in the next section.
 

### 2020-11-17 paper/paper.tex
updating references

@@
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained (REF).
+Broadly speaking optimization can be unconstrained or constrained (\citet{kelley1999iterative}).
 The unconstrained problem can be formulated as a minimization (or maximization) problem such as
@@
 
-The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (REF) which is an exploratory data analysis tool that is defined to detect \emph{interesting structures} or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (REF).
+The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (\citet{buja2005computational}) which is an exploratory data analysis tool that is defined to detect \emph{interesting structures} or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (cook2008grand).
 The target of the optimisation is to identify the most \emph{interesting} low-dimensional views of the data given by a corresponding projection matrix.
-The most \emph{interesting} structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (REF).
+The most \emph{interesting} structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (\citet{cook1993projection}).
 
 The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existance of multiple maxima (local and global), the tradeoff between computational burden and proximity to the maxima, dealing with noisy objective functions that might be non smooth and non diferentiable (REF).
+Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima, dealing with noisy objective functions that might be non smooth and non diferentiable (\citet{jones1998efficient}).
 Those are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
@@
 These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
 In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form.
+In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find a solution in closed form.
 That is, our problem consists of finding the minimum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
 In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF).
+In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (\citet{fletcher2013practical}).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
-In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction \(S_k\) is established then the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line search).
-Then move on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\).
+These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned (\citet{trefethen1997numerical}).
+In order to solve some of those problems, line minimization or line search methods were proposed (\citet{shi2004convergence}) where the step size is chosen in such a way that convergence is ensured.
+In line search methods users are required to provide an initial estimate \(x_{1}\), a search direction \(S_k\) for each iteration where the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line search).
+Then move on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and repeat the process until the desire convergence is reached.
 Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(k)}\), and on approximations of the step size \(\alpha_k\), catering for practical optimisation problems.
-Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
-
-Derivative free --\textgreater{} adding
+Here we focus on line search methods for the optimisation used in the projection pursuit guided tour described in the next section.
 

### 2020-11-17 paper/paper.Rmd
Minor typos

@@
 A d-dimensional projection can be seen as a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
-Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$.
-This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a projection.
+Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$ (QUESTION: Isn't f the index function?).
+This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
 
@@
 
-The optimisation problem is stated as follows.
-Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
+The optimisation problem in the tour context is stated as follows:
+....
+.Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
 

### 2020-11-17 paper/paper.tex
Minor typos

@@
 A d-dimensional projection can be seen as a linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\), and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection matrix.
-Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projected data \(\mathbf{Y}\) (corresponding to an associated projection matrix \(\mathbf{A}\)) onto an index value \(I\).
-This is commonly known as the projection pursuit index function, or just index function, and is used to measure the ``interestingness'' of a projection.
+Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projected data \(\mathbf{Y}\) (corresponding to an associated projection matrix \(\mathbf{A}\)) onto an index value \(I\) (QUESTION: Isn't f the index function?).
+This is commonly known as the projection pursuit index function, or just index function, and is used to measure the ``interestingness'' of a given projection.
 
@@
 
-The optimisation problem is stated as follows.
-Given a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit finds the final projection basis \(\mathbf{A}_T\) that satisfies the following optimisation problem:
+The optimisation problem in the tour context is stated as follows:
+\ldots.
+.Given a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit finds the final projection basis \(\mathbf{A}_T\) that satisfies the following optimisation problem:
 

### 2020-11-19 paper/paper.tex
added a hand-sketch on plots

@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/pmen0008/Desktop/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/dicook/students_PhD/Sherry/paper-tour-vis/img/tour_path}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-better-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/kol-better-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/kol-random-1}
 
@@
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-tuned-1}
+{\centering \includegraphics{/Users/dicook/students_PhD/Sherry/paper-tour-vis/figs/kol-random-tuned-1}
 

### 2020-12-03 paper/paper.Rmd
Updating optimization methods section

@@
 
-The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
+The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
 Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (@jones1998efficient).
@@
 Optimization problems are ubiquitous in many areas of study.
-While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem.
+While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
 These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
@@
 
-In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find a solution in closed form.
-That is, our problem consists of finding the minimum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.
-In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (@fletcher2013practical).
+Furthermore, optimization methods can be divided in various classes (REF).
+In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
+
+Our interest is on constrained optimization as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form solution.
+That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
+In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives.
+
+In gradient descent methods the goal is to find the minimum of an objective function while in gradient ascent methods the interest is in the maximum.
+The minimum is searched for by taken steps  proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (@fletcher2013practical).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
@@
 Modern development of line search methods focuses on improving the computation of the searching direction $S_{(k)}$, and on approximations of the step size $\alpha_k$, catering for practical optimisation problems.
-Here we focus on line search methods for the optimisation used in the projection pursuit guided tour described in the next section.
+In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
 
- <!-- **From Fletcher's book next paragraph --> clean up** Theoretical results have -->
+\textbf{Gradient descent/ascent}
+
+- Objective function
+- Gradient of the objective function
+- A search direction
+- Step length --> line search methods ()
+- R available packages
+
+Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and the properties of the searching space.
+However,  the assumption of having a sufficiently smooth objective function where derivatives exist, it is not adequate for all optimization problems.
+In addition, even in the case where derivatives exists sometimes the searching space is such that their computation is not feasible or it is computationally too expensive.
+Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is in finding, in most cases, a near optimal solution.
+
+\textbf{Derivative free methods}
+
+- Do not rely on derivatives of the objective function
+- Near optimal solutions
+- R packages?
+- polish?
+
+ <!-- **From Fletcher's book next paragraph -->
  <!-- demonstrated the global convergence of the algorithm when the step-size -->

### 2020-12-03 paper/paper.tex
Updating optimization methods section

--- 2020-11-17 paper/paper.tex
@@
 
-The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general.
+The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
 Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima, dealing with noisy objective functions that might be non smooth and non diferentiable (\citet{jones1998efficient}).
@@
 
-In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find a solution in closed form.
-That is, our problem consists of finding the minimum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space.
-In this scenario, gradient descent methods have taken the stage for many decades now.
-In gradient descent methods the goal is to find the minimum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (\citet{fletcher2013practical}).
+Furthermore, optimization methods can be divided in various classes (REF).
+In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
+
+Our interest is on constrained optimization as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form solution.
+That is, the problem consists of finding the minimum or maximum of a function \(f \in L^p\) in the constrained \(\mathbb{A}\) space.
+In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives.
+
+In gradient descent methods the goal is to find the minimum of an objective function while in gradient ascent methods the interest is in the maximum.
+The minimum is searched for by taken steps proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (\citet{fletcher2013practical}).
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
@@
 Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(k)}\), and on approximations of the step size \(\alpha_k\), catering for practical optimisation problems.
-Here we focus on line search methods for the optimisation used in the projection pursuit guided tour described in the next section.
+In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
 
-clean up** Theoretical results have --\textgreater{}
+\textbf{Gradient descent/ascent}
+
+\begin{itemize}
+\tightlist
+\item
+  Objective function
+\item
+  Gradient of the objective function
+\item
+  A search direction
+\item
+  Step length --\textgreater{} line search methods ()
+\item
+  R available packages
+\end{itemize}
+
+Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and the properties of the searching space.
+However, the assumption of having a sufficiently smooth objective function where derivatives exist, it is not adequate for all optimization problems.
+In addition, even in the case where derivatives exists sometimes the searching space is such that their computation is not feasible or it is computationally too expensive.
+Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is in finding, in most cases, a near optimal solution.
+
+\textbf{Derivative free methods}
+
+\begin{itemize}
+\tightlist
+\item
+  Do not rely on derivatives of the objective function
+\item
+  Near optimal solutions
+\item
+  R packages?
+\item
+  polish?
+\end{itemize}
 

### 2020-12-04 paper/paper.Rmd
Updating section 2 - having fun with Sherry

@@
 These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
 In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-Furthermore, optimization methods can be divided in various classes (REF).
+Our interest is on constrained optimization (REF) as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
+That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
+
+<!-- In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives or their approximations. -->
+
+Optimization methods can be divided in various classes (REF).
 In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
-
-Our interest is on constrained optimization as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form solution.
-That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
-In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives.
 
@@
 Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and repeat the process until the desire convergence is reached.
-Modern development of line search methods focuses on improving the computation of the searching direction $S_{(k)}$, and on approximations of the step size $\alpha_k$, catering for practical optimisation problems.
+Modern development of line search methods focuses on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational storing
+burden of the intermediate computations while catering for practical optimisation problems.
 In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
 
-\textbf{Gradient descent/ascent}
+Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and on the complexity of the constraints.
+Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is on finding, in most cases, a near optimal solution.
+Examples of those  include response surface methodology (REF), stochastic approximation (REF),
+random search (REF) and  heuristic methods like tabu search and genetic algorithms (REF).
+Later, we will present an optimization algorithm within the random search class, namely simulated annealing for optimization with the guided tour.
+Moreover, we will introduce an algorithm to further improve optimization solutions both in gradient descent methods and in derive free optimization.
 
-- Objective function
-- Gradient of the objective function
-- A search direction
-- Step length --> line search methods ()
-- R available packages
-
-Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and the properties of the searching space.
-However,  the assumption of having a sufficiently smooth objective function where derivatives exist, it is not adequate for all optimization problems.
-In addition, even in the case where derivatives exists sometimes the searching space is such that their computation is not feasible or it is computationally too expensive.
-Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is in finding, in most cases, a near optimal solution.
-
-\textbf{Derivative free methods}
-
-- Do not rely on derivatives of the objective function
-- Near optimal solutions
-- R packages?
-- polish?
-
- <!-- **From Fletcher's book next paragraph -->
- <!-- demonstrated the global convergence of the algorithm when the step-size -->
- <!-- $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ -->
- <!-- [@curry1944method]. This is however rarely implemented in practice, either -->
- <!-- because of the high computational demand or because the solution does not exist -->
- <!-- [@fletcher2013practical]. A modified approach, in the case of maximisation, is -->
- <!-- to impose a mandatory increase in the objective function for each iteration: -->
- <!-- $f^{(j+1)}> f^{(j)}$. Similarly, for  minimisation we would require a decrease -->
- <!-- in function value at each step. This approach turns out to be efficient in -->
- <!-- practical problems, but global convergence cannot be guaranteed in this case. -->
-
-<!-- For a given problem we might rely on numeric methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems. -->
-<!-- An iterative approach finds the optimum by progressively improving the current solution.
-Researchers can then decide when to stop if a desirable accuracy has been achieved.
-One potential issue with this technique is that it may end up finding a local optimum.
--->
-
-<!-- In this paper, the main focus is on the  *line search method* [@fletcher2013practical].
-In a simple one-dimensional problem, finding the value of $x$ that minimises the function $f(x)$, the line search takes iterative steps as in Equation \ref{eq:line-search}, -->
-
-<!-- \begin{equation} -->
-<!-- x^{(j + 1)} = x^{(j)} + \alpha_k* d^{(j)}, -->
-<!-- \label{eq:line-search} -->
-<!-- \end{equation} -->
-
-<!-- where $d^{(j)}$ is the searching direction in iteration $j$, and $\alpha_j$ is the step-size.
-Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method].
-This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical].
-A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$.
-Similarly, for  minimisation we would require a decrease in function value at each step.
-This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
--->
-
-<!-- Modern development of line search methods focuses on improving the computation of the searching direction $d^{(j)}$, and on approximations of the step size $\alpha_j$, catering for practical optimisation problems.
-Here we focus on optimisation methods used with the projection pursuit guided tour described in the next section.
--->
+Several R implementations address optimization problems with both general purpose as well as tasks specifics solvers.
+The most prominent one within the general solvers is *optim* in (REF-stats package) which have functions for gradient based and derivative free methods.
+Other packages having similar functionalities include *nloptr* (REF), XX.
 

### 2020-12-04 paper/paper.tex
Updating section 2 - having fun with Sherry

@@
 These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
 In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
 
-Furthermore, optimization methods can be divided in various classes (REF).
+Our interest is on constrained optimization (REF) as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
+That is, the problem consists of finding the minimum or maximum of a function \(f \in L^p\) in the constrained \(\mathbb{A}\) space.
+
+Optimization methods can be divided in various classes (REF).
 In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
-
-Our interest is on constrained optimization as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form solution.
-That is, the problem consists of finding the minimum or maximum of a function \(f \in L^p\) in the constrained \(\mathbb{A}\) space.
-In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives.
 
@@
 Then move on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and repeat the process until the desire convergence is reached.
-Modern development of line search methods focuses on improving the computation of the searching direction \(S_{(k)}\), and on approximations of the step size \(\alpha_k\), catering for practical optimisation problems.
+Modern development of line search methods focuses on improving the computation of the searching direction \(S_{k}\), and on approximations of the step size \(\alpha_k\) together with reducing the computational storing
+burden of the intermediate computations while catering for practical optimisation problems.
 In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
 
-\textbf{Gradient descent/ascent}
+Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and on the complexity of the constraints.
+Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is on finding, in most cases, a near optimal solution.
+Examples of those include response surface methodology (REF), stochastic approximation (REF),
+random search (REF) and heuristic methods like tabu search and genetic algorithms (REF).
+Later, we will present an optimization algorithm within the random search class, namely simulated annealing for optimization with the guided tour.
+Moreover, we will introduce an algorithm to further improve optimization solutions both in gradient descent methods and in derive free optimization.
 
-\begin{itemize}
-\tightlist
-\item
-  Objective function
-\item
-  Gradient of the objective function
-\item
-  A search direction
-\item
-  Step length --\textgreater{} line search methods ()
-\item
-  R available packages
-\end{itemize}
-
-Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and the properties of the searching space.
-However, the assumption of having a sufficiently smooth objective function where derivatives exist, it is not adequate for all optimization problems.
-In addition, even in the case where derivatives exists sometimes the searching space is such that their computation is not feasible or it is computationally too expensive.
-Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is in finding, in most cases, a near optimal solution.
-
-\textbf{Derivative free methods}
-
-\begin{itemize}
-\tightlist
-\item
-  Do not rely on derivatives of the objective function
-\item
-  Near optimal solutions
-\item
-  R packages?
-\item
-  polish?
-\end{itemize}
+Several R implementations address optimization problems with both general purpose as well as tasks specifics solvers.
+The most prominent one within the general solvers is \emph{optim} in (REF-stats package) which have functions for gradient based and derivative free methods.
+Other packages having similar functionalities include \emph{nloptr} (REF), XX.
 

### 2020-12-04 paper/paper.Rmd
update codes and RJournal layout, not finishing...

@@
 blinded: 0
-authors:
-
+author:
   - name: H.Sherry Zhang
-  thanks: The authors gratefully acknowledge ...
     affiliation: Department of Econometrics and Business Statistics, Monash University 
-- name: Author 2
-  affiliation: Department of ZZZ, University of WWW
+    email:  Huize.Zhang@monash.edu
+abstract: >
+  Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
+The initial paper by Friedman & Tukey in 1974 stated that  "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
+In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
+These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
+An R package, ferrn, has been created to implement the diagnostics.
 keywords:
@@
 - R
-abstract: Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
-The initial paper by Friedman & Tukey in 1974 stated that  "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
-In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
-These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
-An R package, ferrn, has been created to implement the diagnostics.
-bibliography: biblio.bib
 preamble: >
@@
   bookdown::pdf_book:
-    base_format: rticles::asa_article
+    base_format: rticles::rjournal_article
     fig_height: 3
@@
   chunk_output_type: console
+
 ---
@@
                       root.dir = here::here(), fig.path = here::here("figs/"), 
-                      fig.align = 'center')
+                      fig.align = 'center', cache = TRUE)
 ```
@@
 
-Figure \@ref(fig:toy-search) shows a sample of the plot constructed using `search_better` with different values of the parameter `max_tries`.
+Figure \ref{fig:toy-search} shows a sample of the plot constructed using `search_better` with different values of the parameter `max_tries`.
 Comparing the returned index value of each iteration shows that the termination at `max_tries = 25` is not sufficient for `search_better` to explore the parameter space and a value of 500 is preferred over 25 in this context.
 
-```{r toy-search, fig.cap = "(ref:toy-search)"}
+```{r toy-search, fig.cap="\\label{fig:toy-search} A comparison of search\\_better with different values of the parameter max\\_tries.
+A six-variable dataset boa6 is used with holes index on a 2D problem.
+max\\_tries = 500 allows the optimiser to find better basis with higher index value at iteration six."}
 ```
-
-(ref:toy-search) A comparison of `search_better` with different values of the parameter `max_tries`.
-A six-variable dataset `boa6` is used with `holes` index on a 2D problem.
-`max_tries = 500` allows the optimiser to find better basis with higher index value at iteration six.
 
@@
 Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
-Figure \@ref(fig:toy-interp) presents the interpolation of three different tour paths each with different curvature.
+Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths each with different curvature.
 The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
@@
 
-```{r noisy-better-geo,fig.asp=0.5,fig.cap = "(ref:noisy-better-geo)"}
-```
+<!-- ```{r noisy-better-geo,fig.asp=0.5,fig.cap = "(ref:noisy-better-geo)"} -->
+<!-- ``` -->
 
@@
 On the other hand, `search_better` is doing relatively well.
-With the theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] producing an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
+With the theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] producing an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(nrow(boa5)), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
 A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 3)`].
@@
 The dataset used is `boa6` where `x2` and `x7` are informative.
-The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
+The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
 Hence, the global maximum happens when variable `x7` is found.
@@
 \clearpage
+
+\bibliography{RJreferences}

### 2020-12-04 paper/paper.tex
update codes and RJournal layout, not finishing...

@@
-% !TeX program = pdfLaTeX
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf}
-\usepackage{enumerate}
-\usepackage{natbib}
-\usepackage{textcomp}
-\usepackage[hyphens]{url} % not crucial - just used below for the URL
-\usepackage{hyperref}
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% !TeX root = RJwrapper.tex
+\title{Visual diagnostics for constrained optimisation with application to guided tours}
+\author{by H.Sherry Zhang}
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+\maketitle
 
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-%% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e} \usepackage{todonotes}
-
-\begin{document}
-
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\if0\blind
-{
-  \title{\bf Visual diagnostics for constrained optimisation with application to guided tours}
-
-  \author{
-        H.Sherry Zhang \thanks{The authors gratefully acknowledge \ldots{}} \\
-    Department of Econometrics and Business Statistics, Monash University\\
-     and \\     Author 2 \\
-    Department of ZZZ, University of WWW\\
-      }
-  \maketitle
-} \fi
-
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Visual diagnostics for constrained optimisation with application to guided tours}
-  \end{center}
-  \medskip
-} \fi
-
-\bigskip
-\begin{abstract}
+\abstract{%
 Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
@@
 An R package, ferrn, has been created to implement the diagnostics.
-\end{abstract}
-
-\noindent%
-{\it Keywords:} optimisation, projection pursuit, guided tour, visualisation, diagnostics, R
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
+}
 
@@
 Mathematical details of the geodesic interpolation can be found in \citet{buja2005computational}.
-Figure \ref{fig:tour-path} shows a sketch of the tour path.
+Figure @ref(fig:tour-path) shows a sketch of the tour path.
 The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/pmen0008/Desktop/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
 
@@
 
-\caption{Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
+\caption[(ref:tour-path)]{(ref:tour-path)}(\#fig:tour-path)
 \end{figure}
+\end{Schunk}
+
+(ref:tour-path) Each square (frame) represents the projected data with a corresponding basis.
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
 
@@
 
-\caption{A comparison of \texttt{search\_better} with different values of the parameter \texttt{max\_tries}.
-A six-variable dataset \texttt{boa6} is used with \texttt{holes} index on a 2D problem.
-\texttt{max\_tries\ =\ 500} allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
+\caption{\label{fig:toy-search} A comparison of search\_better with different values of the parameter max\_tries.
+A six-variable dataset boa6 is used with holes index on a 2D problem.
+max\_tries = 500 allows the optimiser to find better basis with higher index value at iteration six.}(\#fig:toy-search)
 \end{figure}
+\end{Schunk}
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-\caption{The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
+\caption[(ref:toy-interp)]{(ref:toy-interp)}(\#fig:toy-interp)
+\end{figure}
+\end{Schunk}
+
+(ref:toy-interp) The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
 The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}\label{fig:toy-interp}
-\end{figure}
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-\caption{1D projection on the 5-variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
+\caption[(ref:toy-pca)]{(ref:toy-pca)}(\#fig:toy-pca)
+\end{figure}
+\end{Schunk}
+
+(ref:toy-pca) 1D projection on the 5-variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
 The yellow point corresponds to the theoretical best basis {[}0, 1, 0, 0, 0{]} with \texttt{V2} being the only non-normal variable in the dataset.
@@
 The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
-\end{figure}
+They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.
 
@@
 
-Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using \texttt{search\_better} and the other using \texttt{search\_geodesic}.
+Figure @ref(fig:toy-pca) plots the first two principal components of two search paths, one using \texttt{search\_better} and the other using \texttt{search\_geodesic}.
 The search in the space reduced by PCA matches with the optimiser description before where the random sampling in \texttt{search\_better} is broader, controlled by \texttt{alpha} parameter, which is default to 0.5 while the directional search in \texttt{search\_geodesic}, controlled by \texttt{delta} with a default of 0.01, is so tiny that it can barely be seen.
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-\caption{A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\caption[(ref:toy-pca-animated)]{(ref:toy-pca-animated)}(\#fig:toy-pca-animated)
 \end{figure}
+\end{Schunk}
+
+(ref:toy-pca-animated) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.
 
 Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
-Figure \ref{fig:toy-pca-animated} shows six frames of an animated version of Figure \ref{fig:toy-pca} and this time, it shows that \texttt{search\_better} finds the optimum quicker than \texttt{search\_geodesic}.
+Figure @ref(fig:toy-pca-animated) shows six frames of an animated version of Figure @ref(fig:toy-pca) and this time, it shows that \texttt{search\_better} finds the optimum quicker than \texttt{search\_geodesic}.
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-\caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+\caption[(ref:toy-tour)]{(ref:toy-tour)}(\#fig:toy-tour)
+\end{figure}
+\end{Schunk}
+
+(ref:toy-tour) A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
 The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}\label{fig:toy-tour}
-\end{figure}
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
 
@@
 To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compared to the PCA plot, the animated rotation (tour) displayed in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Compared to the PCA plot, the animated rotation (tour) displayed in Figure @ref(fig:toy-tour) gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
 Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
-As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
+As an evidence, the last frame in Figure @ref(fig:toy-tour) is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure @ref(fig:toy-pca).
 
@@
 This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
+After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure @ref(fig:interruption).
 
-\begin{figure}
-
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
-
-}
-
-\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_better}.
+(ref:interruption) Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_better}.
 A comparison of the trace plot with and without interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.}\label{fig:interruption}
-\end{figure}
+The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
+\caption[(ref:polish)]{(ref:polish)}(\#fig:polish)
+\end{figure}
+\end{Schunk}
+
+(ref:polish) Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
 The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.}\label{fig:polish}
-\end{figure}
+The separation of the clusters on the y axis becomes sharper after the polish.
 
@@
 
-Figure \ref{fig:polish} presents the final projections found before and after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
+Figure @ref(fig:polish) presents the final projections found before and after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
 Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables.
-In terms of the projected data as in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
+In terms of the projected data as in Figure @ref(fig:polish), polish works to sharpen the edges of each cluster.
 
@@
 
-\begin{figure}
-
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
-
-}
-
-\caption{One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
+(ref:noisy-better-geo) One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
 The grey dashed line represents the index value of the theoretical best basis.
-\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of the theoretical best basis.}\label{fig:noisy-better-geo}
-\end{figure}
+\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of the theoretical best basis.
 
@@
 
-Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: \texttt{search\_geodesic} and \texttt{search\_better}.
+Figure @ref(fig:noisy-better-geo) compares the tracing plot of two optimisers: \texttt{search\_geodesic} and \texttt{search\_better}.
 This time, the interpolated path is no longer smooth when using either algorithm and \texttt{search\_geodesic} fails to optimise this index with barely any improvement of the index value.
 On the other hand, \texttt{search\_better} is doing relatively well.
-With the theoretical best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.182, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
+With the theoretical best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.17, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
 A further polish step will give a marginal improvement of index value to 0.175 with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}.
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-better-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}
 
@@
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
+\caption[(ref:kol-better)]{(ref:kol-better)}(\#fig:kol-better)
+\end{figure}
+\end{Schunk}
+
+(ref:kol-better) One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
 Each point represents a basis in the original 6D space, reduced to 2D by PCA.
 The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
-The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
-\end{figure}
+The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}
 
@@
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
+\caption[(ref:kol-random)]{(ref:kol-random)}(\#fig:kol-random)
+\end{figure}
+\end{Schunk}
+
+(ref:kol-random) One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
 \texttt{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
-\end{figure}
+This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-tuned-1}
+{\centering \includegraphics{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
 
@@
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with an larger searching neighbourhood of 0.7.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.}\label{fig:kol-random-tuned}
+\caption[(ref:kol-random-tuned)]{(ref:kol-random-tuned)}(\#fig:kol-random-tuned)
 \end{figure}
+\end{Schunk}
+
+(ref:kol-random-tuned) One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with an larger searching neighbourhood of 0.7.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.
 
@@
 The dataset used is \texttt{boa6} where \texttt{x2} and \texttt{x7} are informative.
-The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235, respectively.
+The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.176 and 0.235, respectively.
 Hence, the global maximum happens when variable \texttt{x7} is found.
-Simulation is done with 20 randomly generated seeds on two optimisers: \texttt{search\_better} (Figure \ref{fig:kol-better}) and \texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
-Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned} further increases the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge the search space.
+Simulation is done with 20 randomly generated seeds on two optimisers: \texttt{search\_better} (Figure @ref(fig:kol-better)) and \texttt{search\_better\_random} (Figure @ref(fig:kol-random) and Figure @ref(fig:kol-random-tuned)).
+Compared to Figure @ref(fig:kol-better) and @ref(fig:kol-random), Figure @ref(fig:kol-random-tuned) further increases the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge the search space.
 The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
 
-In Figure \ref{fig:kol-better}, the starting point largely affects whether \texttt{search\_better} will find the global maximum.
+In Figure @ref(fig:kol-better), the starting point largely affects whether \texttt{search\_better} will find the global maximum.
 The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7} because they are born in Rome.
@@
 A comparison between \texttt{search\_better} and \texttt{search\_better\_random} shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
-Two iconic cases that shows how \texttt{search\_better\_random} improves \texttt{search\_better} are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}).
+Two iconic cases that shows how \texttt{search\_better\_random} improves \texttt{search\_better} are seeds 4761 (position (1,4) in Figure @ref(fig:kol-better) and (2,2) in Figure @ref(fig:kol-random)) and seed 1842 ((2,2) in Figure @ref(fig:kol-better) and (4,5) in Figure @ref(fig:kol-random)).
 In both cases, \texttt{search\_better\_random} accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
 However, this change of search direction may sometimes have an adverse effect on \texttt{search\_better\_random} and change the search to the neighbourhood of \texttt{V2}.
-This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
-The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
+This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure @ref(fig:kol-random)).
+The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure @ref(fig:kol-better) and position (1,5) and (1,3) in Figure @ref(fig:kol-random)).
 This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near \texttt{V7} to be adequately explored.
@@
 Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated bases and hence allows a wider search.
-Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
+Seed 9982 ((2,4) in Figure @ref(fig:kol-random-tuned)) and seed 2463 ((3,2) in Figure @ref(fig:kol-random-tuned)) benefit from this and find the global maximum.
 A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
@@
 \item
-  \texttt{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
+  \texttt{explore\_trace\_search()}: produces summary plots, as shown in Figure @ref(fig:toy-search)
 \item
-  \texttt{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+  \texttt{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure @ref(fig:toy-interp)
 \item
-  \texttt{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
-Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument \texttt{animate\ =\ TRUE}
+  \texttt{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure @ref(fig:toy-pca).
+Animated version in Figure @ref(fig:toy-pca-animated) can be turned on via the argument \texttt{animate\ =\ TRUE}
 \item
-  \texttt{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+  \texttt{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure @ref(fig:toy-tour).
 \item
@@
 
-\bibliographystyle{agsm}
-\bibliography{biblio.bib}
+\bibliography{RJreferences}
 
-\end{document}
+\address{%
+H.Sherry Zhang\\
+Department of Econometrics and Business Statistics, Monash University\\
+\\
+}
+\href{mailto:Huize.Zhang@monash.edu}{\nolinkurl{Huize.Zhang@monash.edu}}

### 2020-12-04 scripts/examples.R

@@ -10,7 +10,7 @@ p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() 
 p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
   theme(legend.position = "none")
 
-(p1 | p2) & scale_color_botanical(palette = "fern")
+#(p1 | p2) & scale_color_botanical(palette = "fern")
 
 ## ----polish
 # set.seed(123456)
@@ -85,12 +85,18 @@ wrap_plots(gl)
 # save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
 # save(kol_1d_better_polish, file = here::here("data", "kol_1d_better_polish.rda"))
 
-index <- tourr::norm_kol()
+index <- tourr::norm_kol(nrow(boa5))
 theo_best_index_val <- index(as.matrix(boa5) %*% matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1))
 
-bind_rows(kol_1d_geo, kol_1d_better) %>%
-  explore_trace_interp(col = method, group = method) +
-  facet_wrap(vars(method), ncol = 2) +
+
+p1 <- kol_1d_geo %>%
+  explore_trace_interp() +
   geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) +
-  scale_color_botanical() +
-  theme(legend.position = "none", aspect.ratio = 1)
+  scale_color_botanical(discrete = FALSE)
+
+p2 <- kol_1d_better %>%
+  explore_trace_interp() +
+  geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) +
+  scale_color_botanical(discrete = FALSE)
+
+p1 | p2

### 2020-12-04 scripts/kol_result.R

@@ -35,10 +35,10 @@ sim_pca <- function(clean_up){
 
   pca <- clean_up %>%
     bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                     index = tourr::norm_kol(),
+                     index = tourr::norm_kol(nrow(boa6)),
                      raw_data = boa6) %>%
     bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                     index = tourr::norm_kol(),
+                     index = tourr::norm_kol(nrow(boa6)),
                      raw_data = boa6) %>%
     mutate(group_id = paste0(sim, seed)) %>%
     compute_pca(group = group_id) %>%

### 2020-12-04 scripts/vis-diag.R

@@ -9,24 +9,36 @@ files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
 ## ---- toy-search
-bind_rows(holes_2d_better_max_tries %>% mutate(max_tries = 500),
-          holes_2d_better %>% mutate(max_tries = 25)) %>%
-  explore_trace_search(group = max_tries) +
-  scale_color_botanical(palette = "daisy") +
-  facet_wrap(vars(max_tries),labeller = "label_both", ncol = 2) +
-  xlab("iteration")
+p1 <- holes_2d_better_max_tries %>%
+  mutate(max_tries = 500) %>%
+  explore_trace_search() +
+  scale_color_botanical(palette = "daisy")
+
+p2 <- holes_2d_better %>%
+  mutate(max_tries = 25) %>%
+  explore_trace_search() +
+  scale_color_botanical(palette = "daisy")
+
+p1 | p2
 
 ## ---- toy-interp
-bind_rows(holes_2d_better_max_tries %>% mutate(group = "Algorithm 1"),
-          interrupt_no %>% mutate(group = "Algorithm 2"),
-          holes_2d_better_random %>% mutate(group = "Algorithm 3")) %>%
-  mutate(group = fct_relevel(group, c("Algorithm 1",
-                                      "Algorithm 2",
-                                      "Algorithm 3"))) %>%
-  explore_trace_interp(group = group) +
-  theme(legend.position = "none") +
-  scale_color_botanical(palette = "fern") +
-  facet_wrap(vars(group), ncol = 3, scales = "free_x")
+p1 <- holes_2d_better_max_tries %>%
+  mutate(group = "Algorithm 1") %>%
+  explore_trace_interp() +
+  scale_color_botanical(palette = "fern", discrete = FALSE)
+
+# p2 <- interrupt_no %>%
+#   mutate(group = "Algorithm 2") %>%
+#   explore_trace_interp() +
+#   scale_color_botanical(palette = "fern", discrete = FALSE)
+
+p3 <- holes_2d_better_random %>%
+  mutate(group = "Algorithm 3") %>%
+  explore_trace_interp() +
+  scale_color_botanical(palette = "fern", discrete = FALSE)
+
+# p1 | p2 | p3
+p1 | p3
 
 ## ---- toy-pca
 pca <- bind_rows(holes_1d_geo, holes_1d_better) %>%

### 2020-12-09 paper/paper.Rmd
edit on reference, plot layout & captions, acknowleedgements

@@
 title: "Visual diagnostics for constrained optimisation with application to guided tours"
-blinded: 0
 author:
@@
 An R package, ferrn, has been created to implement the diagnostics.
-keywords:
-- optimisation
-- projection pursuit 
-- guided tour
-- visualisation
-- diagnostics 
-- R
 preamble: >
@@
   \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
-  \usepackage{todonotes}
-output:
-  bookdown::pdf_book:
-    base_format: rticles::rjournal_article
-    fig_height: 3
-editor_options: 
-  chunk_output_type: console
-
+output: rticles::rjournal_article
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
-                      root.dir = here::here(), fig.path = here::here("figs/"), 
-                      fig.align = 'center', cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, 
+                      message = FALSE, 
+                      warning = FALSE,
+                      root.dir = here::here(), 
+                      fig.path = here::here("figs/"), 
+                      cache = TRUE, 
+                      fig.align = "center",
+                      fig.height = 3,
+                      out.width = "100%")
 ```
@@
 
-Visualisation is widely used in exploratory data analysis (REF).
+Visualisation is widely used in exploratory data analysis [@tukey1977exploratory].
 Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
@@
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained (@kelley1999iterative).
+Broadly speaking optimization can be unconstrained or constrained [@kelley1999iterative].
 The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
-$min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with  certain properties defined in an $L^p$ space.
+$\min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with  certain properties defined in an $L^p$ space.
 In this case, solutions rely on gradient descent or ascent methods.
 In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as
-$min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
+$\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
 <!-- $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.  -->
 
-The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (@buja2005computational) which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (cook2008grand).
+The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour [@buja2005computational] which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours [@cook2008grand].
 The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
-The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (@cook1993projection).
+The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space [@cook1993projection].
 
 The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (@jones1998efficient).
+Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable [@jones1998efficient].
 Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
@@
 In gradient descent methods the goal is to find the minimum of an objective function while in gradient ascent methods the interest is in the maximum.
-The minimum is searched for by taken steps  proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (@fletcher2013practical).
+The minimum is searched for by taken steps  proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
-These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned (@trefethen1997numerical).
-In order to solve some of those problems, line minimization or line search methods were proposed (@shi2004convergence) where the step size is chosen in such a way that convergence is ensured.
+These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned [@trefethen1997numerical].
+In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
 In line search methods users are required to provide an initial estimate $x_{1}$, a search direction $S_k$ for each iteration where the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
@@
 Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is on finding, in most cases, a near optimal solution.
-Examples of those  include response surface methodology (REF), stochastic approximation (REF),
-random search (REF) and  heuristic methods like tabu search and genetic algorithms (REF).
+Examples of those include response surface methodology [@box1951experimental], stochastic approximation [@robbins1951stochastic], random search [@fu2015handbook] and  heuristic methods [@sorensen2013metaheuristics].
 Later, we will present an optimization algorithm within the random search class, namely simulated annealing for optimization with the guided tour.
@@
 Several R implementations address optimization problems with both general purpose as well as tasks specifics solvers.
-The most prominent one within the general solvers is *optim* in (REF-stats package) which have functions for gradient based and derivative free methods.
-Other packages having similar functionalities include *nloptr* (REF), XX.
+The most prominent one within the general solvers is `optim()` in the package \CRANpkg{stats}[@stats] which have functions for gradient based and derivative free methods.
+Other packages having similar functionalities include \CRANpkg{nloptr}.
 
@@
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-Figure \@ref(fig:tour-path) shows a sketch of the tour path.
+Figure \ref{fig:tour-path} shows a sketch of the tour path.
 The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
-The tour method has been implemented in the `tourr` package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].
+The tour method has been implemented in the R package \CRANpkg{tourr} [@tourr], available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/)
 
-```{r tour-path, out.height="30%", out.width="60%", fig.cap="(ref:tour-path)"}
+```{r tour-path, out.height="25%",out.width="50%",fig.cap="Each square (frame) represents the projected data with a corresponding basis.
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
 include_graphics(here::here("img/tour_path.png"))
@@
 
-(ref:tour-path) Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.
-
 ## Optimisation in the tour {#tour-optim}
 
+pa
 The optimisation problem in the tour context is stated as follows:
@@
 
-where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthogonal matrices.
+\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthogonal matrices.
 
@@
 
-where $T^{\prime} = T + k_{J}+ l_{J+1}$.
+\noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
 Note that there is no output in iteration $J + 1$ since the fact that it is the last iteration means that the optimiser cannot find a better basis and the algorithm terminates.
@@
 The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation forms a row and each variable forms a column.
-With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as dplyr[@dplyr] and ggplot2[@ggplot2].
+With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
 
@@
 
-```{r toy-search, fig.cap="\\label{fig:toy-search} A comparison of search\\_better with different values of the parameter max\\_tries.
-A six-variable dataset boa6 is used with holes index on a 2D problem.
-max\\_tries = 500 allows the optimiser to find better basis with higher index value at iteration six."}
+```{r toy-search, fig.width = 5.5, fig.cap="A comparison of \\code{search\\_better} with different values of the parameter \\code{max\\_tries}.
+A six-variable dataset \\code{boa6} is used with holes index on a 2D problem.
+\\code{max\\_tries = 500} allows the optimiser to find better basis with higher index value at iteration six."}
 ```
@@
 
-```{r toy-interp,fig.cap = "(ref:toy-interp)"}
+```{r toy-interp,fig.width = 5.5, fig.cap = "The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
+The color represents the number of iteration.
+It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies."}
 ```
-
-(ref:toy-interp) The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
-The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.
 
@@
 
-```{r toy-pca, fig.cap = "(ref:toy-pca)"}
-```
-
-(ref:toy-pca) 1D projection on the 5-variable dataset `boa5` with two optimisers: `search_better` and `search_geodesic`.
+```{r toy-pca, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: \\code{search\\_better} and \\code{search\\_geodesic}.
 The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
@@
 The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in `search_geodesic` has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.
+They are initially simulated with the same starting points but all the bases in \\code{search\\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot."}
+```
 
@@
 principal component analysis.
-To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.
+To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \CRANpkg{geozoo} [@geozoo] and PCA is conducted on both the bases and the points on the surface of the sphere.
 
-Figure \@ref(fig:toy-pca) plots the first two principal components of two search paths, one using `search_better` and the other using `search_geodesic`.
+Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using `search_better` and the other using `search_geodesic`.
 The search in the space reduced by PCA matches with the optimiser description before where the random sampling in `search_better` is broader, controlled by `alpha` parameter, which is default to 0.5 while the directional search in `search_geodesic`, controlled by `delta` with a default of 0.01, is so tiny that it can barely be seen.
@@
 
-```{r toy-pca-animated, fig.cap = "(ref:toy-pca-animated)"}
+```{r toy-pca-animated, fig.cap = "A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm."}
 ```
 
-(ref:toy-pca-animated) A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.
-
 Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
-Figure \@ref(fig:toy-pca-animated) shows six frames of an animated version of Figure \@ref(fig:toy-pca) and this time, it shows that `search_better` finds the optimum quicker than `search_geodesic`.
+Figure \ref{fig:toy-pca-animated} shows six frames of an animated version of Figure \ref{fig:toy-pca} and this time, it shows that `search_better` finds the optimum quicker than `search_geodesic`.
 
@@
 
-```{r toy-tour, fig.cap="(ref:toy-tour)"}
+```{r toy-tour, fig.cap="A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles."}
 ```
-
-(ref:toy-tour) A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
 
@@
 To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compared to the PCA plot, the animated rotation (tour) displayed in Figure \@ref(fig:toy-tour) gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
+Compared to the PCA plot, the animated rotation (tour) displayed in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
 Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
-As an evidence, the last frame in Figure \@ref(fig:toy-tour) is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \@ref(fig:toy-pca).
+As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
 
@@
 This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \@ref(fig:interruption).
+After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
 
-```{r interruption, fig.asp=0.5, fig.cap = "(ref:interruption)"}
-```
-
-(ref:interruption) Two-D projection on `boa6` data with holes index optimised by `search_better`.
+```{r interruption, fig.asp=0.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by \\code{search\\_better}.
 A comparison of the trace plot with and without interruption.
 When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.
+The interruption forces the interpolation to finish at the highest interpolated basis on the tour path."}
+```
 
@@
 
-```{r polish, fig.height= 2.5, fig.cap = "(ref:polish)"}
-```
-
-(ref:polish) Two-D projection on `boa6` data with holes index optimised by `search_geodesic`.
+```{r polish, fig.height= 2.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by \\code{search\\_geodesic}.
 The left panel shows the final projected data before polish and the right panel shows the one after.
 The separation of the clusters on the y axis becomes sharper after the polish.
+"}
+```
 
@@
 
-Figure \@ref(fig:polish) presents the final projections found before and after applying `search_polish` on `search_geodesic`.
+Figure \ref{fig:polish} presents the final projections found before and after applying `search_polish` on `search_geodesic`.
 Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables.
-In terms of the projected data as in Figure \@ref(fig:polish), polish works to sharpen the edges of each cluster.
+In terms of the projected data as in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
 
@@
 
-<!-- ```{r noisy-better-geo,fig.asp=0.5,fig.cap = "(ref:noisy-better-geo)"} -->
+<!-- ```{r noisy-better-geo,fig.asp=0.5,fig.cap = "One-D projection on `boa5` data with noisy index `norm_kol` optimised by `search_geodesic` and `search_better`.
+The grey dashed line represents the index value of the theoretical best basis.
+`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of the theoretical best basis."} -->
 <!-- ``` -->
-
-(ref:noisy-better-geo) One-D projection on `boa5` data with noisy index `norm_kol` optimised by `search_geodesic` and `search_better`.
-The grey dashed line represents the index value of the theoretical best basis.
-`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of the theoretical best basis.
 
@@
 
-Figure \@ref(fig:noisy-better-geo) compares the tracing plot of two optimisers: `search_geodesic` and `search_better`.
+Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: `search_geodesic` and `search_better`.
 This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely any improvement of the index value.
@@
 
-```{r kol-better, fig.height = 6, fig.cap="(ref:kol-better)"}
+```{r kol-better, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better} with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced to 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
+The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point."}
 ```
 
-(ref:kol-better) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better` with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced to 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
-The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.
-
-```{r kol-random, fig.height = 6, fig.cap = "(ref:kol-random)"}
+```{r kol-random, fig.height = 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with the same seeds.
+\\code{search\\_better\\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum."}
 ```
 
-(ref:kol-random) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with the same seeds.
-`search_better_random` has a probabilistic acceptance implementation that would also accept a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.
-
-```{r kol-random-tuned, fig.height= 6, fig.cap = "(ref:kol-random-tuned)"}
+```{r kol-random-tuned, fig.height= 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with an larger searching neighbourhood of 0.7.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
+"}
 ```
 
-(ref:kol-random-tuned) One-D projection of `norm_kol` index on `boa6` data optimised by `search_better_random` with an larger searching neighbourhood of 0.7.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
-
 The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
-The dataset used is `boa6` where `x2` and `x7` are informative.
+The dataset used is \\code{boa6} where `x2` and `x7` are informative.
 The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
 Hence, the global maximum happens when variable `x7` is found.
-Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \@ref(fig:kol-better)) and `search_better_random` (Figure \@ref(fig:kol-random) and Figure \@ref(fig:kol-random-tuned)).
-Compared to Figure \@ref(fig:kol-better) and \@ref(fig:kol-random), Figure \@ref(fig:kol-random-tuned) further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.
+Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
+Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.
 The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
 
-In Figure \@ref(fig:kol-better), the starting point largely affects whether `search_better` will find the global maximum.
+In Figure \ref{fig:kol-better}, the starting point largely affects whether `search_better` will find the global maximum.
 The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find `V7` because they are born in Rome.
@@
 A comparison between `search_better` and `search_better_random` shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
-Two iconic cases that shows how `search_better_random` improves `search_better` are seeds 4761 (position (1,4) in Figure \@ref(fig:kol-better) and (2,2) in Figure \@ref(fig:kol-random)) and seed 1842 ((2,2) in Figure \@ref(fig:kol-better) and (4,5) in Figure \@ref(fig:kol-random)).
+Two iconic cases that shows how `search_better_random` improves `search_better` are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}).
 In both cases, `search_better_random` accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
 However, this change of search direction may sometimes have an adverse effect on `search_better_random` and change the search to the neighbourhood of `V2`.
-This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \@ref(fig:kol-random)).
-The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \@ref(fig:kol-better) and position (1,5) and (1,3) in Figure \@ref(fig:kol-random)).
+This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
+The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
 This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near `V7` to be adequately explored.
@@
 Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated bases and hence allows a wider search.
-Seed 9982 ((2,4) in Figure \@ref(fig:kol-random-tuned)) and seed 2463 ((3,2) in Figure \@ref(fig:kol-random-tuned)) benefit from this and find the global maximum.
+Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
 A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
@@
 
-- `explore_trace_search()`: produces summary plots, as shown in Figure \@ref(fig:toy-search)
-- `explore_trace_interp()`: produces trace plots for the interpolation points, as shown in Figure \@ref(fig:toy-interp)
-- `explore_space_pca() `: produces plots of projection basis on the reduced space by PCA, as shown in Figure \@ref(fig:toy-pca).
-Animated version in Figure \@ref(fig:toy-pca-animated) can be turned on via the argument `animate = TRUE`
-- `explore_space_tour()`: produces animated tour view on the full space of the projection bases, as shown in Figure \@ref(fig:toy-tour).
+- `explore_trace_search()`: produces summary plots, as shown in Figure \ref{fig:toy-search}
+- `explore_trace_interp()`: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+- `explore_space_pca() `: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
+- `explore_space_tour()`: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
 
@@
 <!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing `tourr` package; manipulation and visualisation of the data object are implemented in the new `ferrn` package.
+Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \CRANpkg{tourr}[@tourr] package; manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
 Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.
 scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.
+
+# Acknowledgements
+
+This article is created using \CRANpkg{knitr}[@knitr] and \CRANpkg{rmarkdown} [@rmarkdown]  in R.
+The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-tour-vis}.
 

### 2020-12-09 paper/paper.tex
edit on reference, plot layout & captions, acknowleedgements

--- 2020-12-04 paper/paper.tex
@@
-% !TeX program = pdfLaTeX
-\documentclass[12pt]{article}
-\usepackage{amsmath}
-\usepackage{graphicx,psfrag,epsf}
-\usepackage{enumerate}
-\usepackage{natbib}
-\usepackage{textcomp}
-\usepackage[hyphens]{url} % not crucial - just used below for the URL
-\usepackage{hyperref}
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+% !TeX root = RJwrapper.tex
+\title{Visual diagnostics for constrained optimisation with application to
+guided tours}
+\author{by H.Sherry Zhang}
 
-%\pdfminorversion=4
-% NOTE: To produce blinded version, replace "0" with "1" below.
-\newcommand{\blind}{0}
+\maketitle
 
-% DON'T change margins - should be 1 inch all around.
-\addtolength{\oddsidemargin}{-.5in}%
-\addtolength{\evensidemargin}{-.5in}%
-\addtolength{\textwidth}{1in}%
-\addtolength{\textheight}{1.3in}%
-\addtolength{\topmargin}{-.8in}%
-
-%% load any required packages here
-\usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs} \usepackage[ruled,vlined, linesnumbered]{algorithm2e} \usepackage{todonotes}
-
-\begin{document}
-
-\def\spacingset#1{\renewcommand{\baselinestretch}%
-{#1}\small\normalsize} \spacingset{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\if0\blind
-{
-  \title{\bf Visual diagnostics for constrained optimisation with application to guided tours}
-
-  \author{
-        H.Sherry Zhang \thanks{The authors gratefully acknowledge \ldots{}} \\
-    Department of Econometrics and Business Statistics, Monash University\\
-     and \\     Author 2 \\
-    Department of ZZZ, University of WWW\\
+\abstract{%
+Projection pursuit searches for interesting low-dimensional views of
+high-dimensional data via the optimisation of an index function. The
+initial paper by Friedman \& Tukey in 1974 stated that ``the technique
+used for maximising the projection index strongly influences both the
+statistical and the computational aspects of the procedure.'' However,
+while many projection pursuit indices have been proposed in the
+literature, less work has been done on the optimisation procedures. In
+this paper we introduce visual diagnostics for optimisation algorithms,
+in particular those available in the projection pursuit guided tour.
+These diagnostics and workflows can be applied to a broad class of
+optimisers, to assess their performance. An R package, ferrn, has been
+created to implement the diagnostics.
 }
-  \maketitle
-} \fi
-
-\if1\blind
-{
-  \bigskip
-  \bigskip
-  \bigskip
-  \begin{center}
-    {\LARGE\bf Visual diagnostics for constrained optimisation with application to guided tours}
-  \end{center}
-  \medskip
-} \fi
-
-\bigskip
-\begin{abstract}
-Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
-The initial paper by Friedman \& Tukey in 1974 stated that ``the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure.'' However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
-In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
-These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
-An R package, ferrn, has been created to implement the diagnostics.
-\end{abstract}
-
-\noindent%
-{\it Keywords:} optimisation, projection pursuit, guided tour, visualisation, diagnostics, R
-\vfill
-
-\newpage
-\spacingset{1.45} % DON'T change the spacing!
 
@@
 
-Visualisation is widely used in exploratory data analysis (REF).
-Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-The work presented in this paper combines data visualization with optimization methods.
-It aims at bring visualization tools into optimization problems to assist users to better understand the performance of their optimization methods in practice.
-This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms.
+Visualisation is widely used in exploratory data analysis
+\citep{tukey1977exploratory}. Presenting information in a graphical
+format often allows people to uncover information they would otherwise
+not be aware of and provides a more comprehensive understanding of the
+problem at hand. The work presented in this paper combines data
+visualization with optimization methods. It aims at bring visualization
+tools into optimization problems to assist users to better understand
+the performance of their optimization methods in practice. This
+motivates our work of creating plots to diagnose optimisation algorithms
+which in turn can be also used not only to understand optimization
+performance but also to understand and compare the behaviour of
+different algorithms.
 
-The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained (\citet{kelley1999iterative}).
-The unconstrained problem can be formulated as a minimization (or maximization) problem such as
-\(min_{x} f(x)\) where \(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with certain properties defined in an \(L^p\) space.
-In this case, solutions rely on gradient descent or ascent methods.
-In the constrained optimization problem additional restrictions are introduced via a set of functions that can be convex or non-convex: \(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and hence the problem can be written as
-\(min_{x} f(x)\) \emph{subject to} \(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
+The goal of continuous optimisation is to find the best solution within
+the space of all feasible solutions where typically the best solution is
+decided by an objective function. Broadly speaking optimization can be
+unconstrained or constrained \citep{kelley1999iterative}. The
+unconstrained problem can be formulated as a minimization (or
+maximization) problem such as \(\min_{x} f(x)\) where
+\(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with
+certain properties defined in an \(L^p\) space. In this case, solutions
+rely on gradient descent or ascent methods. In the constrained
+optimization problem additional restrictions are introduced via a set of
+functions that can be convex or non-convex:
+\(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and
+hence the problem can be written as \(\min_{x} f(x)\) \emph{subject to}
+\(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex
+optimization methods including linear and quadratic programming can be
+used.
 
-The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour (\citet{buja2005computational}) which is an exploratory data analysis tool that is defined to detect \emph{interesting structures} or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours (cook2008grand).
-The target of the optimisation is to identify the most \emph{interesting} low-dimensional views of the data given by a corresponding projection matrix.
-The most \emph{interesting} structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space (\citet{cook1993projection}).
+The motivation of this paper is the optimisation problem arising in the
+projection pursuit guided tour \citep{buja2005computational} which is an
+exploratory data analysis tool that is defined to detect
+\emph{interesting structures} or features in high-dimensional data
+through a set of lower-dimensional projections that cover the entire
+high dimensional space using interpolation methods called tours
+\citep{cook2008grand}. The target of the optimisation is to identify the
+most \emph{interesting} low-dimensional views of the data given by a
+corresponding projection matrix. The most \emph{interesting} structures
+are formally defined by a function of projections, called index function
+which is optimized to uncover the most reveling structures in a high
+dimensional space \citep{cook1993projection}.
 
-The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima, dealing with noisy objective functions that might be non smooth and non diferentiable (\citet{jones1998efficient}).
-Those are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
+The optimization challenges encountered in the projection pursuit guided
+tour problem are common to those of optimization in general. Examples of
+those include the existence of multiple maxima (local and global), the
+trade off between computational burden and proximity to the maxima,
+dealing with noisy objective functions that might be non smooth and non
+diferentiable \citep{jones1998efficient}. Those are not unique to this
+context and therefore the visualization tools and optimization methods
+presented in this paper can be easily apply to any other optimization
+problems.
 
-The remainder of the paper is organised as follows.
-Section \ref{optim} provides an overview of optimisation methods, specifically line search methods.
-Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and introduces three existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and used in different types of diagnostic plots.
-Section \ref{application} shows applications of how these plots can be used to understand and compare different algorithms.
-We also discuss how these insights contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package: ferrn, that implements the visual diagnostics.
+The remainder of the paper is organised as follows. Section \ref{optim}
+provides an overview of optimisation methods, specifically line search
+methods. Section \ref{tour} reviews projection pursuit guided tour,
+defines the optimisation problem and introduces three existing
+algorithms. Section \ref{vis-diag} presents the new visual diagnostics.
+A data structure is defined to capture information during the
+optimisation, and used in different types of diagnostic plots. Section
+\ref{application} shows applications of how these plots can be used to
+understand and compare different algorithms. We also discuss how these
+insights contribute to modifications that improve the algorithms.
+Finally, Section \ref{implementation} describes the R package: ferrn,
+that implements the visual diagnostics.
 
@@
 
-Optimization problems are ubiquitous in many areas of study.
-While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
-These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
-In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
+Optimization problems are ubiquitous in many areas of study. While in
+some cases, analytical solutions can be found, the majority of problems
+rely on numerical methods to find the optimal solution. These numerical
+methods follow interactive approaches that aim at finding the optimum by
+progressively improving the current solution until a desirable accuracy
+is achieved. Although that principle seems uncomplicated, optimization
+methods need to deal with many challenges such as the existence of
+multiple maxima (local and global), the trade off between desirable
+accuracy and computational burden, as well as dealing with noisy
+objective functions and possible constraints. In addition, optimization
+results might depend on the algorithm starting values affecting the
+consistency of results.
 
-Our interest is on constrained optimization (REF) as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
-That is, the problem consists of finding the minimum or maximum of a function \(f \in L^p\) in the constrained \(\mathbb{A}\) space.
+Our interest is on constrained optimization (REF) as defined in the
+introduction section and assume that it is not possible to find a
+solution to the problem in the way of a closed form. That is, the
+problem consists of finding the minimum or maximum of a function
+\(f \in L^p\) in the constrained \(\mathbb{A}\) space.
 
-Optimization methods can be divided in various classes (REF).
-In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
+Optimization methods can be divided in various classes (REF). In this
+paper, we focused on methods that rely on the objective function
+derivatives known as gradient methods, and those which are known as
+derivative free methods where optimization does not rely on the gradient
+or derivatives information.
 
-In gradient descent methods the goal is to find the minimum of an objective function while in gradient ascent methods the interest is in the maximum.
-The minimum is searched for by taken steps proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (\citet{fletcher2013practical}).
-The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
-These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned (\citet{trefethen1997numerical}).
-In order to solve some of those problems, line minimization or line search methods were proposed (\citet{shi2004convergence}) where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate \(x_{1}\), a search direction \(S_k\) for each iteration where the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line search).
-Then move on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and repeat the process until the desire convergence is reached.
-Modern development of line search methods focuses on improving the computation of the searching direction \(S_{k}\), and on approximations of the step size \(\alpha_k\) together with reducing the computational storing
-burden of the intermediate computations while catering for practical optimisation problems.
-In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
+In gradient descent methods the goal is to find the minimum of an
+objective function while in gradient ascent methods the interest is in
+the maximum. The minimum is searched for by taken steps proportional to
+the negative gradient or of the function at the evaluation point
+(analogously positive gradient for gradient ascent methods), that is
+\(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or
+the learning rate \citep{fletcher2013practical}. The step size is the
+key in the success of the optimization problem as inadequate step sizes
+might lead to a failure in convergence. These methods are known to be
+relative slow and can have convergence issues when problems are poorly
+conditioned \citep{trefethen1997numerical}. In order to solve some of
+those problems, line minimization or line search methods were proposed
+\citep{shi2004convergence} where the step size is chosen in such a way
+that convergence is ensured. In line search methods users are required
+to provide an initial estimate \(x_{1}\), a search direction \(S_k\) for
+each iteration where the objective is to find the value of
+\(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with
+respect to \(\alpha_k\) (this step is called the line search). Then move
+on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and
+repeat the process until the desire convergence is reached. Modern
+development of line search methods focuses on improving the computation
+of the searching direction \(S_{k}\), and on approximations of the step
+size \(\alpha_k\) together with reducing the computational storing
+burden of the intermediate computations while catering for practical
+optimisation problems. In the next section two line search methods for
+the optimisation used in the projection pursuit guided tour will be
+presented.
 
-Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and on the complexity of the constraints.
-Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is on finding, in most cases, a near optimal solution.
-Examples of those include response surface methodology (REF), stochastic approximation (REF),
-random search (REF) and heuristic methods like tabu search and genetic algorithms (REF).
-Later, we will present an optimization algorithm within the random search class, namely simulated annealing for optimization with the guided tour.
-Moreover, we will introduce an algorithm to further improve optimization solutions both in gradient descent methods and in derive free optimization.
+Although gradient optimization methods are very popular, they rely on
+the existence of the objective function derivatives and on the
+complexity of the constraints. Because of that, the second big class of
+optimization problems look at derivative free methods where the emphasis
+is on finding, in most cases, a near optimal solution. Examples of those
+include response surface methodology \citep{box1951experimental},
+stochastic approximation \citep{robbins1951stochastic}, random search
+\citep{fu2015handbook} and heuristic methods
+\citep{sorensen2013metaheuristics}. Later, we will present an
+optimization algorithm within the random search class, namely simulated
+annealing for optimization with the guided tour. Moreover, we will
+introduce an algorithm to further improve optimization solutions both in
+gradient descent methods and in derive free optimization.
 
-Several R implementations address optimization problems with both general purpose as well as tasks specifics solvers.
-The most prominent one within the general solvers is \emph{optim} in (REF-stats package) which have functions for gradient based and derivative free methods.
-Other packages having similar functionalities include \emph{nloptr} (REF), XX.
+Several R implementations address optimization problems with both
+general purpose as well as tasks specifics solvers. The most prominent
+one within the general solvers is \texttt{optim()} in the package
+\CRANpkg{stats}\citep{stats} which have functions for gradient based and
+derivative free methods. Other packages having similar functionalities
+include \CRANpkg{nloptr}.
 
@@
 
-The projection pursuit guided tour combines two different methods in exploratory data analysis, focusing on different aspects.
-Projection pursuit, coined by \citet{friedman1974projection}, detects interesting structures (e.g.~clustering, outliers and skewness) in multivariate data via low dimensions projection.
-The guided tour is using ideas from projection pursuit to define a particular variation in a broader class of data visualisation methods, building on the grand tour approach \citep{As85}.
+The projection pursuit guided tour combines two different methods in
+exploratory data analysis, focusing on different aspects. Projection
+pursuit, coined by \citet{friedman1974projection}, detects interesting
+structures (e.g.~clustering, outliers and skewness) in multivariate data
+via low dimensions projection. The guided tour is using ideas from
+projection pursuit to define a particular variation in a broader class
+of data visualisation methods, building on the grand tour approach
+\citep{As85}.
 
-To define projection pursuit, we first need to establish the notation used.
-Let \(\mathbf{X}_{n \times p}\) be the data matrix, with \(n\) observations in \(p\) dimensions.
-A d-dimensional projection can be seen as a linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\), and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where \(\mathbf{Y}_{n \times d}\) is the projected data and \(\mathbf{A}_{p\times d}\) is the projection matrix.
-Define \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index function that maps the projected data \(\mathbf{Y}\) (corresponding to an associated projection matrix \(\mathbf{A}\)) onto an index value \(I\) (QUESTION: Isn't f the index function?).
-This is commonly known as the projection pursuit index function, or just index function, and is used to measure the ``interestingness'' of a given projection.
+To define projection pursuit, we first need to establish the notation
+used. Let \(\mathbf{X}_{n \times p}\) be the data matrix, with \(n\)
+observations in \(p\) dimensions. A d-dimensional projection can be seen
+as a linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\),
+and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
+\(\mathbf{Y}_{n \times d}\) is the projected data and
+\(\mathbf{A}_{p\times d}\) is the projection matrix. Define
+\(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
+function that maps the projected data \(\mathbf{Y}\) (corresponding to
+an associated projection matrix \(\mathbf{A}\)) onto an index value
+\(I\) (QUESTION: Isn't f the index function?). This is commonly known as
+the projection pursuit index function, or just index function, and is
+used to measure the ``interestingness'' of a given projection.
 
-A number of index functions have been proposed in the literature to detect different data structures, including Legendre index \citep{friedman1974projection}, Hermite index \citep{hall1989polynomial}, natural Hermite index \citep{cook1993projection}, chi-square index \citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA index \citep{lee2010projection}.
+A number of index functions have been proposed in the literature to
+detect different data structures, including Legendre index
+\citep{friedman1974projection}, Hermite index
+\citep{hall1989polynomial}, natural Hermite index
+\citep{cook1993projection}, chi-square index
+\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
+index \citep{lee2010projection}.
 
-As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
-Different tour types choose these planes differently, for example, a grand tour \citep{cook2008grand} selects the planes randomly and a manual tour \citep{cook1997manual} gradually phases in and out one variable, to understand the contribution of that variable in the projection.
-Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
-Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
+As a general visualisation method, a tour produces animations of high
+dimensional data via rotations between low dimension planes. Different
+tour types choose these planes differently, for example, a grand tour
+\citep{cook2008grand} selects the planes randomly and a manual tour
+\citep{cook1997manual} gradually phases in and out one variable, to
+understand the contribution of that variable in the projection. Guided
+tour, the main interest of this paper, chooses planes with the aid of
+projection pursuit. Given a random start, projection pursuit iteratively
+finds bases with higher index values and the guided tour constructs the
+geodesic interpolation between these planes to form a tour path.
 
-Mathematical details of the geodesic interpolation can be found in \citet{buja2005computational}.
-Figure \ref{fig:tour-path} shows a sketch of the tour path.
-The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
-The tour method has been implemented in the \texttt{tourr} package in R, available on the Comprehensive R Archive Network at \url{https://cran.r-project.org/web/packages/tourr/} \citep{wickham2011tourrpackage}.
+Mathematical details of the geodesic interpolation can be found in
+\citet{buja2005computational}. Figure \ref{fig:tour-path} shows a sketch
+of the tour path. The blue frames are produced by the projection pursuit
+optimisation algorithm, and the white frames interpolate between them.
+The tour method has been implemented in the R package \CRANpkg{tourr}
+\citep{tourr}, available on the Comprehensive R Archive Network at
+\url{https://cran.r-project.org/web/packages/tourr/}
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.3\textheight]{/Users/pmen0008/Desktop/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
 
@@
 
-\caption{Each square (frame) represents the projected data with a corresponding basis.
+\caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
+\end{Schunk}
 
@@
 
-The optimisation problem in the tour context is stated as follows:
-\ldots.
-.Given a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit finds the final projection basis \(\mathbf{A}_T\) that satisfies the following optimisation problem:
+pa The optimisation problem in the tour context is stated as follows:
+\ldots. .Given a randomly generated starting basis \(\mathbf{A}_1\),
+projection pursuit finds the final projection basis \(\mathbf{A}_T\)
+that satisfies the following optimisation problem:
 
@@
 
-where \(I_d\) is the \(d\)-dimensional identity matrix and the constraint requires the projection bases \(\mathbf{A}\) to be orthogonal matrices.
+\noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
+constraint requires the projection bases \(\mathbf{A}\) to be orthogonal
+matrices.
 
-Several features of this optimisation are worth noticing.
-First of all, this is a constrained optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
-It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
-Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
-Lastly, there is also one computational consideration: the optimisation procedure needs to be fast to compute since the tour animation is played in real-time.
+Several features of this optimisation are worth noticing. First of all,
+this is a constrained optimisation problem as the decision variables
+form the entries of a projection basis, which is required to be
+orthonormal. It is also likely that the objective function may not be
+differentiable for a constructed index function and in these cases,
+gradient-based methods may not work well. Although finding the global
+maximum is the goal of an optimisation problem, it is also interesting
+to inspect local maximum in projection pursuit since it could present
+unexpected interesting projections. Lastly, there is also one
+computational consideration: the optimisation procedure needs to be fast
+to compute since the tour animation is played in real-time.
 
@@
 
-Below we introduce three possible algorithms: \texttt{search\_better}, \texttt{search\_better\_random}, and \texttt{search\_geodesic}.
-The first two are derivative free methods that sample candidate bases in the neighbourhood whilst \texttt{search\_geodesic} is an analogue of gradient ascent on the projection basis space.
+Below we introduce three possible algorithms: \texttt{search\_better},
+\texttt{search\_better\_random}, and \texttt{search\_geodesic}. The
+first two are derivative free methods that sample candidate bases in the
+neighbourhood whilst \texttt{search\_geodesic} is an analogue of
+gradient ascent on the projection basis space.
 
@@
 
-\texttt{search\_better} is a random search device that samples a candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current basis \(\mathbf{A}_{\text{cur}}\) by \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\) where \(\alpha\) controls the radius of the sampling neighbourhood and \(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the same dimension as \(\mathbf{A}_{\text{cur}}\).
-\(\mathbf{A}_{l}\) is then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis \(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis for guided tour to construct an interpolation path.
-The next iteration of search begins after adjusting \(\alpha\) by a cooling parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\).
-The termination condition is when the maximum number of iteration \(l_{\max}\) is reached.
-The algorithm of \texttt{search\_better} is summarised in Algorithm \ref{random-search}.
-A slightly different cooling scheme has been proposed by \citet{posse1995projection} to include a halving parameter \(c\).
-Rather than reducing the radius of the searching neighbourhood, \(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if the last search takes more than \(c\) times to find an accepted basis to avoid the searching space being reduced too fast.
+\texttt{search\_better} is a random search device that samples a
+candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
+basis \(\mathbf{A}_{\text{cur}}\) by
+\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
+where \(\alpha\) controls the radius of the sampling neighbourhood and
+\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
+same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
+then orthogonalised to ensure the orthonormal constraint is fulfilled.
+When a basis is found with index value higher than the current basis
+\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
+for guided tour to construct an interpolation path. The next iteration
+of search begins after adjusting \(\alpha\) by a cooling parameter:
+\(\alpha_{j+1} = \alpha_j * \text{cooling}\). The termination condition
+is when the maximum number of iteration \(l_{\max}\) is reached. The
+algorithm of \texttt{search\_better} is summarised in Algorithm
+\ref{random-search}. A slightly different cooling scheme has been
+proposed by \citet{posse1995projection} to include a halving parameter
+\(c\). Rather than reducing the radius of the searching neighbourhood,
+\(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if
+the last search takes more than \(c\) times to find an accepted basis to
+avoid the searching space being reduced too fast.
 
@@
 
-Simulated annealing (\texttt{search\_better\_random}) \citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses the same sampling process as \texttt{search\_better} but allows a probabilistic acceptance of a basis with lower index value based on the annealing \(T(l)\).
-Given an initial \(T_0\), the temperature at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\] where \(I_{(\cdot)}\) denotes the index value of a given basis.
-This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when local maxima are present.
-The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+Simulated annealing (\texttt{search\_better\_random})
+\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses
+the same sampling process as \texttt{search\_better} but allows a
+probabilistic acceptance of a basis with lower index value based on the
+annealing \(T(l)\). Given an initial \(T_0\), the temperature at
+iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
+candidate basis fails to have an index value larger than the current
+basis, simulated annealing gives it a second chance to be accepted with
+probability
+\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
+where \(I_{(\cdot)}\) denotes the index value of a given basis. This
+implementation allows the algorithm to jump out of a local maximum and
+enables a more holistic search of the whole parameter space. This
+feature is particularly useful when local maxima are present. The
+algorithm can be written as replacing line 5-8 of Algorithm
+\ref{random-search} with Algorithm \ref{simulated_annealing}.
 
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm on the space of the projection bases.
-In gradient ascent, one first finds the direction for improvement via computing the gradient information.
-In \texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter \(\delta\).
-A geodesic is then constructed using the current basis and the one in \(2n\) bases with the highest index value.
-If the neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and works as the searching direction.
-The next step in gradient ascent is to conduct a line search to find the best improvement along the gradient direction and in \texttt{search\_geodesic}, this is replaced by optimising the index value along the geodesic direction over an 90 degree angle from \(-\pi/4\) to \(\pi/4\).
-The optima \(\mathbf{A}_{**}\) is returned for the current iteration if it meets the termination condition on percentage improvement.
-The procedure will also terminate if \(l_{\max}\) is reached.
-Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
+\citet{cook1995grand} used a gradient ascent algorithm on the space of
+the projection bases. In gradient ascent, one first finds the direction
+for improvement via computing the gradient information. In
+\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
+neighbourhood of the current basis, controlled by the neighbourhood
+parameter \(\delta\). A geodesic is then constructed using the current
+basis and the one in \(2n\) bases with the highest index value. If the
+neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is
+an analogue of the gradient information in the curved space and works as
+the searching direction. The next step in gradient ascent is to conduct
+a line search to find the best improvement along the gradient direction
+and in \texttt{search\_geodesic}, this is replaced by optimising the
+index value along the geodesic direction over an 90 degree angle from
+\(-\pi/4\) to \(\pi/4\). The optima \(\mathbf{A}_{**}\) is returned for
+the current iteration if it meets the termination condition on
+percentage improvement. The procedure will also terminate if
+\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
+steps in geodesic search.
 
@@
 
-To be able to make diagnostics on the optimisers, the algorithms need to populate a data structure with key elements of the algorithm.
-When the algorithms run, key information regarding the decision variable, objective function and hyper-parameters, needs to be recorded and stored as a data object so that it is ready to be supplied to the plotting functions for diagnostics.
+To be able to make diagnostics on the optimisers, the algorithms need to
+populate a data structure with key elements of the algorithm. When the
+algorithms run, key information regarding the decision variable,
+objective function and hyper-parameters, needs to be recorded and stored
+as a data object so that it is ready to be supplied to the plotting
+functions for diagnostics.
 
 \hypertarget{data-structure-for-diagnostics}{%
-\subsection{Data structure for diagnostics}\label{data-structure-for-diagnostics}}
+\subsection{Data structure for
+diagnostics}\label{data-structure-for-diagnostics}}
 
-In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3) State: \(S\), which labels the observation with detailed stage in the optimisation.
-Possible values for \texttt{search\_better} and \texttt{search\_better\_random} include \texttt{random\_search}, \texttt{new\_basis}, and \texttt{interpolation}.
-\texttt{search\_geodesic} has a wider variety that includes \texttt{new\_basis}, \texttt{direction\_search}, \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and \texttt{interpolation}.
+In the optimisation algorithms for projection pursuit, the three main
+elements to record are 1) projection bases: \(\mathbf{A}\), 2) index
+values: \(I\), and 3) State: \(S\), which labels the observation with
+detailed stage in the optimisation. Possible values for
+\texttt{search\_better} and \texttt{search\_better\_random} include
+\texttt{random\_search}, \texttt{new\_basis}, and
+\texttt{interpolation}. \texttt{search\_geodesic} has a wider variety
+that includes \texttt{new\_basis}, \texttt{direction\_search},
+\texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
+\texttt{interpolation}.
 
-Multiple iterators are also needed to index the data collected at different levels.
-\(t\) is a unique identifier that prescribes the natural ordering of each observation; \(j\) is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
-\(l\) is the counter for each search/interpolation, which provides the information of how many bases the algorithm has searched before finding one to return.
-There are other parameters of interest, which depend on the particular problem content and they are denoted as \emph{\(V_{p}\)}.
-Two most common examples include \(V_1 = \text{method}\), which tags the name of the algorithm used, and \(V_2 = \text{alpha}\), the neighbourhood parameter that controls the size in sampling candidate bases.
-A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
+Multiple iterators are also needed to index the data collected at
+different levels. \(t\) is a unique identifier that prescribes the
+natural ordering of each observation; \(j\) is the counter for each
+search-and-interpolate iteration, which remains the same within one
+round and has an increment of one once a new round starts. \(l\) is the
+counter for each search/interpolation, which provides the information of
+how many bases the algorithm has searched before finding one to return.
+There are other parameters of interest, which depend on the particular
+problem content and they are denoted as \emph{\(V_{p}\)}. Two most
+common examples include \(V_1 = \text{method}\), which tags the name of
+the algorithm used, and \(V_2 = \text{alpha}\), the neighbourhood
+parameter that controls the size in sampling candidate bases. A matrix
+notation of the data structure is presented in Equation
+\ref{eq:data-structure}.
 
@@
 
-where \(T^{\prime} = T + k_{J}+ l_{J+1}\).
-Note that there is no output in iteration \(J + 1\) since the fact that it is the last iteration means that the optimiser cannot find a better basis and the algorithm terminates.
-In this notation, the final basis found is \(A_T\) with the highest index value \(I_T\).
+\noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
+no output in iteration \(J + 1\) since the fact that it is the last
+iteration means that the optimiser cannot find a better basis and the
+algorithm terminates. In this notation, the final basis found is \(A_T\)
+with the highest index value \(I_T\).
 
-The data structure constructed above meets the tidy data principle \citep{wickham2014tidy} that requires each observation forms a row and each variable forms a column.
-With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as dplyr\citep{dplyr} and ggplot2\citep{ggplot2}.
+The data structure constructed above meets the tidy data principle
+\citep{wickham2014tidy} that requires each observation forms a row and
+each variable forms a column. With tidy data structure, data wrangling
+and visualisation have been significantly simplified by well-developed
+packages such as \CRANpkg{dplyr} \citep{dplyr} and \CRANpkg{ggplot2}
+\citep{ggplot2}.
 
-The construction of diagnostic plots uses the concept of grammar of graphics \citep{wickham2010layered} in ggplot2.
-In grammar of graphics, plots are not defined by their appearance (i.e.~boxplot, histogram, scatter plot etc) but by stacked layers.
-In the construction of diagnostic plots, there are multiple elements one may want to emphasize and there is no single plot name that would meet the need.
-The stacked layers concept, on the other hand, allows information to be overlaid on each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
+The construction of diagnostic plots uses the concept of grammar of
+graphics \citep{wickham2010layered} in ggplot2. In grammar of graphics,
+plots are not defined by their appearance (i.e.~boxplot, histogram,
+scatter plot etc) but by stacked layers. In the construction of
+diagnostic plots, there are multiple elements one may want to emphasize
+and there is no single plot name that would meet the need. The stacked
+layers concept, on the other hand, allows information to be overlaid on
+each other and one can build the plot from scratch as long as the
+variables have been stored in a dataset.
 
 \hypertarget{checking-how-hard-the-optimiser-is-working}{%
-\subsection{Checking how hard the optimiser is working}\label{checking-how-hard-the-optimiser-is-working}}
+\subsection{Checking how hard the optimiser is
+working}\label{checking-how-hard-the-optimiser-is-working}}
 
-A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum.
-Directly plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes longer for an optimiser to find a better basis towards the end.
-Another option is to use summarisation for each iteration.
-Boxplot is a suitable candidate that can provide five points summary of each iteration.
-Other additional information not presented in the boxplot can then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
-Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a \texttt{cutoff} parameter.
+A primary interest of diagnosing an optimiser is to study how it
+progressively finds its optimum. Directly plotting the index value
+across its natural order will cause the graph to be disproportional to
+the iteration since it usually takes longer for an optimiser to find a
+better basis towards the end. Another option is to use summarisation for
+each iteration. Boxplot is a suitable candidate that can provide five
+points summary of each iteration. Other additional information not
+presented in the boxplot can then be added with new layers, for example
+text information on the number of points can be added at the bottom of
+each iteration and the position of basis returned by projection pursuit
+can be highlighted in point. Further, an option to switch between
+displaying points and boxplot geometry is helpful when the number of
+observation is small in one iteration and this is achieved via a
+\texttt{cutoff} parameter.
 
-Figure \ref{fig:toy-search} shows a sample of the plot constructed using \texttt{search\_better} with different values of the parameter \texttt{max\_tries}.
-Comparing the returned index value of each iteration shows that the termination at \texttt{max\_tries\ =\ 25} is not sufficient for \texttt{search\_better} to explore the parameter space and a value of 500 is preferred over 25 in this context.
+Figure \ref{fig:toy-search} shows a sample of the plot constructed using
+\texttt{search\_better} with different values of the parameter
+\texttt{max\_tries}. Comparing the returned index value of each
+iteration shows that the termination at \texttt{max\_tries\ =\ 25} is
+not sufficient for \texttt{search\_better} to explore the parameter
+space and a value of 500 is preferred over 25 in this context.
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
 
@@
 
-\caption{A comparison of \texttt{search\_better} with different values of the parameter \texttt{max\_tries}.
-A six-variable dataset \texttt{boa6} is used with \texttt{holes} index on a 2D problem.
-\texttt{max\_tries\ =\ 500} allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
+\caption{A comparison of \code{search\_better} with different values of the parameter \code{max\_tries}.
+A six-variable dataset \code{boa6} is used with holes index on a 2D problem.
+\code{max\_tries = 500} allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
 \end{figure}
+\end{Schunk}
 
@@
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-\caption{The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
+\caption[The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index]{The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
 The color represents the number of iteration.
@@
 \end{figure}
+\end{Schunk}
 
-Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
-Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths each with different curvature.
-The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
-The middle path has an increase-then-decrease pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
-The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost case is a deliberate construction of \texttt{search\_better\_random} where an inferior basis can be accepted with some probability so as to avoid being trapped in a local maximum.
+Points on the interpolation path are another interest in tour since the
+projection on these bases will be played by the tour animation. Figure
+\ref{fig:toy-interp} presents the interpolation of three different tour
+paths each with different curvature. The leftmost plot shows an
+interpolation where the index value increases progressively and
+monotonically. The middle path has an increase-then-decrease pattern in
+the last two iterations and the rightmost path shows even a decrease in
+the index value at iteration three. The middle situation can be avoid
+via a construction of the interruption, which will be detailed in
+section \ref{monotonic} and the rightmost case is a deliberate
+construction of \texttt{search\_better\_random} where an inferior basis
+can be accepted with some probability so as to avoid being trapped in a
+local maximum.
 
 \hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
-\subsection{Understanding the optimiser's coverage of the search space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
+\subsection{Understanding the optimiser's coverage of the search
+space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-\caption{1D projection on the 5-variable dataset \texttt{boa5} with two optimisers: \texttt{search\_better} and \texttt{search\_geodesic}.
-The yellow point corresponds to the theoretical best basis {[}0, 1, 0, 0, 0{]} with \texttt{V2} being the only non-normal variable in the dataset.
+\caption{1D projection on the 5-variable dataset \code{boa5} with two optimisers: \code{search\_better} and \code{search\_geodesic}.
+The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
 The underlying grey points are randomly generated on the 5D space and reduced to 2D via PCA along with all the search points presented.
 The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in \texttt{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
+They are initially simulated with the same starting points but all the bases in \code{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
 \end{figure}
+\end{Schunk}
 
-Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases \(\mathbf{A}_{p \times d}\) live on the surface of a \(p \times d\) dimension sphere, where the dimension can easily go over two or three.
-While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space (\(d > 3\)), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.~principal component analysis.
-To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \texttt{geozoo} and PCA is conducted on both the bases and the points on the surface of the sphere.
+Apart from checking the progression of an optimiser, another interesting
+aspect is to visualise how the search looks like in its parameter space.
+Given the orthonormality constraint, the projection bases
+\(\mathbf{A}_{p \times d}\) live on the surface of a \(p \times d\)
+dimension sphere, where the dimension can easily go over two or three.
+While visualising the search paths on the original high dimensional
+sphere would require skills for the viewers to perceive rotation of
+geometry in higher dimensional space (\(d > 3\)), an easier alternative
+is to view the reduced space via some dimension reduction methods,
+i.e.~principal component analysis. To better perceive the search path as
+an embedding of a hollow sphere, random points on the high dimensional
+sphere is generated using the package \CRANpkg{geozoo} \citep{geozoo}
+and PCA is conducted on both the bases and the points on the surface of
+the sphere.
 
-Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using \texttt{search\_better} and the other using \texttt{search\_geodesic}.
-The search in the space reduced by PCA matches with the optimiser description before where the random sampling in \texttt{search\_better} is broader, controlled by \texttt{alpha} parameter, which is default to 0.5 while the directional search in \texttt{search\_geodesic}, controlled by \texttt{delta} with a default of 0.01, is so tiny that it can barely be seen.
+Figure \ref{fig:toy-pca} plots the first two principal components of two
+search paths, one using \texttt{search\_better} and the other using
+\texttt{search\_geodesic}. The search in the space reduced by PCA
+matches with the optimiser description before where the random sampling
+in \texttt{search\_better} is broader, controlled by \texttt{alpha}
+parameter, which is default to 0.5 while the directional search in
+\texttt{search\_geodesic}, controlled by \texttt{delta} with a default
+of 0.01, is so tiny that it can barely be seen.
 
 \hypertarget{animating-the-diagnostic-plots}{%
-\subsection{Animating the diagnostic plots}\label{animating-the-diagnostic-plots}}
+\subsection{Animating the diagnostic
+plots}\label{animating-the-diagnostic-plots}}
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-\caption{A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\caption[A selected number of frames from the animated PCA plot]{A selected number of frames from the animated PCA plot.
+With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
 \end{figure}
+\end{Schunk}
 
-Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
-Figure \ref{fig:toy-pca-animated} shows six frames of an animated version of Figure \ref{fig:toy-pca} and this time, it shows that \texttt{search\_better} finds the optimum quicker than \texttt{search\_geodesic}.
+Animated plots can be informative in diagnostics, especially in the case
+of PCA plot when the starting and ending of the search is not clear.
+Figure \ref{fig:toy-pca-animated} shows six frames of an animated
+version of Figure \ref{fig:toy-pca} and this time, it shows that
+\texttt{search\_better} finds the optimum quicker than
+\texttt{search\_geodesic}.
 
 \hypertarget{the-tour-looking-at-itself}{%
-\subsection{The tour looking at itself}\label{the-tour-looking-at-itself}}
+\subsection{The tour looking at
+itself}\label{the-tour-looking-at-itself}}
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-\caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+\caption[A selected number of frames from the tour animation for viewing the 5D space of all the projection bases]{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
 The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
@@
 \end{figure}
+\end{Schunk}
 
-While viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original \(p \times d\) dimension enables a more holistic stereoscopic view of the search.
-To view a high dimensional (\(d \ge 3\)) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compared to the PCA plot, the animated rotation (tour) displayed in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
-Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
-As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
+While viewing the bases on the reduced space via PCA shed some lights on
+the space the optimisers have explored, the visualisation on the
+original \(p \times d\) dimension enables a more holistic stereoscopic
+view of the search. To view a high dimensional (\(d \ge 3\)) object on a
+screen, an approach is to play the rotation of the object in animation
+and this can be done via a regular grand tour. Compared to the PCA plot,
+the animated rotation (tour) displayed in Figure \ref{fig:toy-tour}
+gives a more well-rounded view of the search and one can view the curved
+region of the tour path from different angles, which may not be
+presented in the PCA plot. Also the grand tour animation encompasses the
+PCA projection since the rotation from PCA is just one angle that
+maximises the variance of the bases and the grand tour produces a
+sequence of angles that view the search from different directions. As an
+evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select
+from the tour animation that is close to the PCA angle and the
+projection looks similar to the one in Figure \ref{fig:toy-pca}.
 
@@
 
-For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
-If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are present, \texttt{search\_better} may get stuck at a local maximum and in the case where the index function is non-smooth, \texttt{search\_geodesic} may even fail to find the maximum.
-In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
+For a particular index function, the best algorithm to optimise relates
+to the character of the index and the data. If the index function is
+smooth and has a single maximum, all of the three algorithms introduced
+above can find the maximum. When multiple optima are present,
+\texttt{search\_better} may get stuck at a local maximum and in the case
+where the index function is non-smooth, \texttt{search\_geodesic} may
+even fail to find the maximum. In this section, examples will be
+presented to outline how the diagnostic plots can be used to compare the
+performance of optimisers in different scenarios.
 
@@
 
-Random variables with different structures have been simulated and the distribution of each is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and \texttt{x10} are normal distributed with zero mean and unit variance and \texttt{x2} to \texttt{x7} are mixtures of normal distributions with varied weights and locations.
-The mixture variables have been scaled to have an overall unit variance before running the projection pursuit.
+Random variables with different structures have been simulated and the
+distribution of each is presented in Equations \ref{eq:sim-norm} to
+\ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and
+\texttt{x10} are normal distributed with zero mean and unit variance and
+\texttt{x2} to \texttt{x7} are mixtures of normal distributions with
+varied weights and locations. The mixture variables have been scaled to
+have an overall unit variance before running the projection pursuit.
 
@@
 
-In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
-This pattern is undesirable since the optimiser could have started the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
-This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
+In section \ref{toy-interp}, an interpolation with
+increase-then-decrease pattern has been presented. This pattern is
+undesirable since the optimiser could have started the next iteration
+from the highest basis on the tour path, as annotated as the
+interpolated basis in the plot, but instead, it is forced to start from
+the target basis. This motivates the design of an interruption to check
+the index value on the tour path so that the interpolating bases is
+accepted only up to the one with the largest index value. After
+implementing this interruption, the search finds a higher final index
+value with fewer steps as shown in the right panel of Figure
+\ref{fig:interruption}.
 
+\hypertarget{close-but-not-close-enough}{%
+\subsection{Close but not close
+enough}\label{close-but-not-close-enough}}
+
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_better}.
-A comparison of the trace plot with and without interruption.
-When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the highest interpolated basis on the tour path.}\label{fig:interruption}
+\caption[Two-D projection on \code{boa6} data with holes index optimised by \code{search\_geodesic}]{Two-D projection on \code{boa6} data with holes index optimised by \code{search\_geodesic}.
+The left panel shows the final projected data before polish and the right panel shows the one after.
+The separation of the clusters on the y axis becomes sharper after the polish.
+}\label{fig:polish}
 \end{figure}
+\end{Schunk}
 
-\hypertarget{close-but-not-close-enough}{%
-\subsection{Close but not close enough}\label{close-but-not-close-enough}}
+Once the final basis has been found by an algorithm, one may want to
+push further to investigate whether there is an even better basis in the
+close neighbourhood. This motivates the polish search where the final
+basis is supplied as the start of a new guided tour to search for any
+local breakthrough.
 
-\begin{figure}
-
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
-
-}
-
-\caption{Two-D projection on \texttt{boa6} data with holes index optimised by \texttt{search\_geodesic}.
-The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.}\label{fig:polish}
-\end{figure}
-
-Once the final basis has been found by an algorithm, one may want to push further to investigate whether there is an even better basis in the close neighbourhood.
-This motivates the polish search where the final basis is supplied as the start of a new guided tour to search for any local breakthrough.
-
-Similar to \texttt{search\_better} as a stochastic random search, \texttt{search\_polish} has a different scheme of reducing the search neighbourhood.
-In each search-interpolation iteration, \texttt{search\_better} has a fixed neighbourhood parameter alpha and this alpha is reduced by the cooling parameter only after an iteration finishes.
-On the contrary, \texttt{search\_polish} allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
-Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original \texttt{max.tries} limit.
-These include:
+Similar to \texttt{search\_better} as a stochastic random search,
+\texttt{search\_polish} has a different scheme of reducing the search
+neighbourhood. In each search-interpolation iteration,
+\texttt{search\_better} has a fixed neighbourhood parameter alpha and
+this alpha is reduced by the cooling parameter only after an iteration
+finishes. On the contrary, \texttt{search\_polish} allows alpha to be
+reduced during each iteration to exploit the search in the
+neighbourhood. Further, to avoid the case where alpha becomes too small
+and the further search is meaningless, three more stopping criteria have
+been added, on top of the original \texttt{max.tries} limit. These
+include:
 
@@
 \item
-  the distance between the candidate basis and the current basis needs to be larger than 1e-3;
+  the distance between the candidate basis and the current basis needs
+  to be larger than 1e-3;
 \item
-  the percentage change of the index value need to be larger than 1e-5; and
+  the percentage change of the index value need to be larger than 1e-5;
+  and
 \item
@@
 
-Figure \ref{fig:polish} presents the final projections found before and after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
-Polish search improves the index value from 0.9618 to 0.9627 with reduction of weights on the non-informative variables.
-In terms of the projected data as in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
+Figure \ref{fig:polish} presents the final projections found before and
+after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
+Polish search improves the index value from 0.9618 to 0.9627 with
+reduction of weights on the non-informative variables. In terms of the
+projected data as in Figure \ref{fig:polish}, polish works to sharpen
+the edges of each cluster.
 
 \hypertarget{seeing-the-signal-in-the-noise}{%
-\subsection{Seeing the signal in the noise}\label{seeing-the-signal-in-the-noise}}
+\subsection{Seeing the signal in the
+noise}\label{seeing-the-signal-in-the-noise}}
 
+The index function, up until this point, are all smooth, while this is
+not the case for all the index functions. \texttt{norm\_kol}, a 1D
+projection function based on the Kolmogorov test, compares the
+difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\)
+and a randomly generated normal distribution, \(y_n\) based on the
+empirical cumulated distribution function (ECDF). Denote the ECDF
+function as \(F_{.}(u)\) with the subscript indicating either the
+projection or the random normal variable, the \texttt{norm\_kol} index
+is defined by
+
+\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
+
+Figure \ref{fig:noisy-better-geo} compares the tracing plot of two
+optimisers: \texttt{search\_geodesic} and \texttt{search\_better}. This
+time, the interpolated path is no longer smooth when using either
+algorithm and \texttt{search\_geodesic} fails to optimise this index
+with barely any improvement of the index value. On the other hand,
+\texttt{search\_better} is doing relatively well. With the theoretical
+best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.17,
+\texttt{search\_better} finds the final basis {[} 0.0376, -0.9916,
+-0.0581, -0.0831, 0.0716{]} with an index value of 0.165. A further
+polish step will give a marginal improvement of index value to 0.175
+with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}. At
+this stage, the difference between the theoretical best and what has
+been found is likely due to simulation error since the best possible
+basis for a simulated data will be slightly off the theoretical best
+basis, which is derived based on the distributional assumptions in
+Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}
 
@@
 
-\caption{One-D projection on \texttt{boa5} data with noisy index \texttt{norm\_kol} optimised by \texttt{search\_geodesic} and \texttt{search\_better}.
-The grey dashed line represents the index value of the theoretical best basis.
-\texttt{search\_geodesic} fails to optimise the noisy index while \texttt{search\_better} has made reasonable improvements to reach the index value close to the one of the theoretical best basis.}\label{fig:noisy-better-geo}
+\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds.
+Each point represents a basis in the original 6D space, reduced to 2D by PCA.
+The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
+The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
 \end{figure}
+\end{Schunk}
 
-The index function, up until this point, are all smooth, while this is not the case for all the index functions.
-\texttt{norm\_kol}, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\) and a randomly generated normal distribution, \(y_n\) based on the empirical cumulated distribution function (ECDF).
-Denote the ECDF function as \(F_{.}(u)\) with the subscript indicating either the projection or the random normal variable, the \texttt{norm\_kol} index is defined by
-
-\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
-
-Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: \texttt{search\_geodesic} and \texttt{search\_better}.
-This time, the interpolated path is no longer smooth when using either algorithm and \texttt{search\_geodesic} fails to optimise this index with barely any improvement of the index value.
-On the other hand, \texttt{search\_better} is doing relatively well.
-With the theoretical best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.182, \texttt{search\_better} finds the final basis {[} 0.0376, -0.9916, -0.0581, -0.0831, 0.0716{]} with an index value of 0.165.
-A further polish step will give a marginal improvement of index value to 0.175 with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}.
-At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumptions in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-better-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}
 
@@
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better} with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced to 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when \texttt{V2} is found and the global maximum where \texttt{V7} is found.
-The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
+\caption{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with the same seeds.
+\code{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
+This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
 \end{figure}
+\end{Schunk}
 
+\begin{Schunk}
 \begin{figure}
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
 
@@
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with the same seeds.
-\texttt{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
+\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7.
+Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
+}\label{fig:kol-random-tuned}
 \end{figure}
+\end{Schunk}
 
-\begin{figure}
+The second experiment with the noisy index is to understand how the
+optimisers perform when a local maximum is present. The dataset used is
+\textbackslash code\{boa6\} where \texttt{x2} and \texttt{x7} are
+informative. The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]}
+and {[}0, 0, 1, 0, 0, 0{]} with index value 0.176 and 0.235,
+respectively. Hence, the global maximum happens when variable
+\texttt{x7} is found. Simulation is done with 20 randomly generated
+seeds on two optimisers: \texttt{search\_better} (Figure
+\ref{fig:kol-better}) and \texttt{search\_better\_random} (Figure
+\ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}). Compared to
+Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure
+\ref{fig:kol-random-tuned}further increases the neighbourhood parameter
+\texttt{alpha} from 0.5 to 0.7 to enlarge the search space. The data
+object is collected for each simulation and computation is done to
+reduce the bases from its original 6D to a 2D points by principal
+component analysis so as to view the search paths.
 
-{\centering \includegraphics{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-random-tuned-1}
+In Figure \ref{fig:kol-better}, the starting point largely affects
+whether \texttt{search\_better} will find the global maximum. The three
+seeds on row three: seed 9145, seed 2511, and seed 9209, and three on
+row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7}
+because they are born in Rome. On the other end, the ones on the first
+two rows can only find \texttt{V2}, largely due to the fact that they do
+not even have chances to search near \texttt{V7}.
 
-}
+A comparison between \texttt{search\_better} and
+\texttt{search\_better\_random} shows all four combinations where both,
+neither, or one of the two algorithms find the global maximum. Two
+iconic cases that shows how \texttt{search\_better\_random} improves
+\texttt{search\_better} are seeds 4761 (position (1,4) in Figure
+\ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed
+1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure
+\ref{fig:kol-random}). In both cases, \texttt{search\_better\_random}
+accepts inferior points at the initial iteration and this acceptance has
+later been proved to change the direction of the search and hence allows
+the optimiser to explore near the global optimum and in the end, find
+the global optimum. However, this change of search direction may
+sometimes have an adverse effect on \texttt{search\_better\_random} and
+change the search to the neighbourhood of \texttt{V2}. This happens on
+seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure
+\ref{fig:kol-random}). The two cases that neither algorithm finds the
+global optimum happens on seed 6170 and seed 2757 (position (1,1) and
+(1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in
+Figure \ref{fig:kol-random}). This is largely due to the fact that the
+searching neighbourhood is not large enough to allow the space near
+\texttt{V7} to be adequately explored.
 
-\caption{One-D projection of \texttt{norm\_kol} index on \texttt{boa6} data optimised by \texttt{search\_better\_random} with an larger searching neighbourhood of 0.7.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7.}\label{fig:kol-random-tuned}
-\end{figure}
+In Figure @ref\{fig:kol-random-tuned\}, the neighbourhood parameter
+\texttt{alpha} is increased from a 0.5 default to 0.7 to solve the
+insufficiency of search neighbourhood. Remember that a candidate basis
+is generated via a linear combination of the current basis and a
+randomly generated basis on the surface of the sphere, an increase of
+\texttt{alpha} gives more weights on the randomly generated bases and
+hence allows a wider search. Seed 9982 ((2,4) in Figure
+\ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure
+\ref{fig:kol-random-tuned}) benefit from this and find the global
+maximum. A noticeable feature of this enlarged search space is that the
+interpolation paths start to jump around the space but its usefulness is
+arguable. In seed 2888 (position (3,1)), seed 2986 (position (2,3)),
+seed 9209 (position (4,4)), just to name a few, this allows a near
+\texttt{V2} bases to switch to a near \texttt{V7} bases and as a result
+these cases find the global maximum. While in the case of seed 6746
+(position (1,2)) and seed 2757 (position (1,5)), the jump happens from
+near \texttt{V7} basis to near \texttt{V2} basis, it does not come back
+and leaves those simulations in the local maximum.
 
-The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
-The dataset used is \texttt{boa6} where \texttt{x2} and \texttt{x7} are informative.
-The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0, 0{]} with index value 0.182 and 0.235, respectively.
-Hence, the global maximum happens when variable \texttt{x7} is found.
-Simulation is done with 20 randomly generated seeds on two optimisers: \texttt{search\_better} (Figure \ref{fig:kol-better}) and \texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
-Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned} further increases the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge the search space.
-The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
-
-In Figure \ref{fig:kol-better}, the starting point largely affects whether \texttt{search\_better} will find the global maximum.
-The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7} because they are born in Rome.
-On the other end, the ones on the first two rows can only find \texttt{V2}, largely due to the fact that they do not even have chances to search near \texttt{V7}.
-
-A comparison between \texttt{search\_better} and \texttt{search\_better\_random} shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
-Two iconic cases that shows how \texttt{search\_better\_random} improves \texttt{search\_better} are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}).
-In both cases, \texttt{search\_better\_random} accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
-However, this change of search direction may sometimes have an adverse effect on \texttt{search\_better\_random} and change the search to the neighbourhood of \texttt{V2}.
-This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
-The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
-This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near \texttt{V7} to be adequately explored.
-
-In Figure @ref\{fig:kol-random-tuned\}, the neighbourhood parameter \texttt{alpha} is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of \texttt{alpha} gives more weights on the randomly generated bases and hence allows a wider search.
-Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
-A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
-In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, this allows a near \texttt{V2} bases to switch to a near \texttt{V7} bases and as a result these cases find the global maximum.
-While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near \texttt{V7} basis to near \texttt{V2} basis, it does not come back and leaves those simulations in the local maximum.
-
-The conclusion from this experiment is that the usage of \texttt{search\_better\_random} and increasing the search space are methods that can avoid getting trapped in local maxima but the solution will still depend on the starting points of the simulation and the seed used.
+The conclusion from this experiment is that the usage of
+\texttt{search\_better\_random} and increasing the search space are
+methods that can avoid getting trapped in local maxima but the solution
+will still depend on the starting points of the simulation and the seed
+used.
 
@@
 
-The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing \texttt{tourr} package while the optimiser diagnostics have been implemented in a new package, \texttt{ferrn}.
-When a guided tour is run, the users can choose if the data from optimisation should be collected via the \texttt{verbose} argument.
-Once the data object has been obtained, the package, \texttt{ferrn}, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
-The structure of package functionality has been listed below.
+The implementation of this projection has been divided into two
+packages: the data collection object is implemented in the existing
+\texttt{tourr} package while the optimiser diagnostics have been
+implemented in a new package, \texttt{ferrn}. When a guided tour is run,
+the users can choose if the data from optimisation should be collected
+via the \texttt{verbose} argument. Once the data object has been
+obtained, the package, \texttt{ferrn}, can provide four diagnostic plots
+as shown in Section \ref{vis-diag}. The structure of package
+functionality has been listed below.
 
@@
 \item
-  \texttt{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
+  \texttt{explore\_trace\_search()}: produces summary plots, as shown in
+  Figure \ref{fig:toy-search}
 \item
-  \texttt{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+  \texttt{explore\_trace\_interp()}: produces trace plots for the
+  interpolation points, as shown in Figure \ref{fig:toy-interp}
 \item
-  \texttt{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
-Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument \texttt{animate\ =\ TRUE}
+  \texttt{explore\_space\_pca()}: produces plots of projection basis on
+  the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+  Animated version in Figure \ref{fig:toy-pca-animated} can be turned on
+  via the argument \texttt{animate\ =\ TRUE}
 \item
-  \texttt{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+  \texttt{explore\_space\_tour()}: produces animated tour view on the
+  full space of the projection bases, as shown in Figure
+  \ref{fig:toy-tour}.
 \item
-  \texttt{get\_*()} extracts and manipulates certain components from the existing data object.
+  \texttt{get\_*()} extracts and manipulates certain components from the
+  existing data object.
 
@@
   \item
-    \texttt{get\_best()}: extracts the best basis found in the data object
+    \texttt{get\_best()}: extracts the best basis found in the data
+    object
   \item
@@
   \item
-    \texttt{get\_interp()}: extracts the observations in the interpolation
+    \texttt{get\_interp()}: extracts the observations in the
+    interpolation
   \item
-    \texttt{get\_search\_count()}: produces the summary table of the number of observation in each iteration
+    \texttt{get\_search\_count()}: produces the summary table of the
+    number of observation in each iteration
   \item
@@
 \item
-  \texttt{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
+  \texttt{bind\_*()} incorporates additional information outside the
+  tour optimisation into the data object.
 
@@
   \item
-    \texttt{bind\_theoretical()}: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value.
+    \texttt{bind\_theoretical()}: incorporates the best possible basis
+    to the existing data object with the supply of the index function
+    and original data for producing the index value.
   \item
-    \texttt{bind\_random()}: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
-\texttt{bind\_random\_matrix()} binds the points to the basis matrix.
+    \texttt{bind\_random()}: generates 1000 points on the high
+    dimensional surface of a sphere and binds it to the existing data
+    object and output as a tibble object.
+    \texttt{bind\_random\_matrix()} binds the points to the basis
+    matrix.
   \end{itemize}
@@
   \item
-    \texttt{botanical\_palettes}: a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+    \texttt{botanical\_palettes}: a collection of color palettes from
+    Australian native plants. Quantitative palettes include daisy,
+    banksia and cherry and sequential palettes contain fern and acacia.
   \item
@@
   \item
-    \texttt{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.
+    \texttt{scale\_color\_botanical()}: a ggplot construction for using
+    botanical palettes.
   \end{itemize}
@@
 
-This paper has illustrated setting up a data object that can be used for diagnosing a complex optimisation procedure.
-The ideas were illustrated using the optimisers available for projection pursuit guided tour.
-Here the constraint is the orthornormality condition of the projection bases.
-The approach used here could be broadly applied to understand other constrained optimisers.
+This paper has illustrated setting up a data object that can be used for
+diagnosing a complex optimisation procedure. The ideas were illustrated
+using the optimisers available for projection pursuit guided tour. Here
+the constraint is the orthornormality condition of the projection bases.
+The approach used here could be broadly applied to understand other
+constrained optimisers.
 
-Four diagnostic plots have been introduced to investigate the progression and the projection space of an optimiser.
-The implementation of these visualisations is designed to be easy-to-use with each plot can be produced with a simple supply of the data object.
-More advanced users may decide to modify on top of the basic plots or even build their own.
+Four diagnostic plots have been introduced to investigate the
+progression and the projection space of an optimiser. The implementation
+of these visualisations is designed to be easy-to-use with each plot can
+be produced with a simple supply of the data object. More advanced users
+may decide to modify on top of the basic plots or even build their own.
 
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \texttt{tourr} package; manipulation and visualisation of the data object are implemented in the new \texttt{ferrn} package.
-Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.~scagnostics, association, and information index \citep{laa2020using} and understand how the optimisers behave for index functions with different structures.
+Most of the work in this project has been translated into code in two
+packages: the collection of the data object is implemented in the
+existing \CRANpkg{tourr}\citep{tourr} package; manipulation and
+visualisation of the data object are implemented in the new \pkg{ferrn}
+package. Equipped with handy tools to diagnose the performance of
+optimisers, future work can extend the diagnostics to a wider range of
+index functions, i.e.~scagnostics, association, and information index
+\citep{laa2020using} and understand how the optimisers behave for index
+functions with different structures.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+This article is created using \CRANpkg{knitr}\citep{knitr} and
+\CRANpkg{rmarkdown} \citep{rmarkdown} in R. The source code for
+reproducing this paper can be found at:
+\url{https://github.com/huizezhang-sherry/paper-tour-vis}.
 

### 2020-12-09 scripts/examples.R

@@ -10,7 +10,7 @@ p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() 
 p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
   theme(legend.position = "none")
 
-#(p1 | p2) & scale_color_botanical(palette = "fern")
+#(p1 | p2) & scale_color_botanical(palette = "fern", discrete = TRUE)
 
 ## ----polish
 # set.seed(123456)

### 2020-12-09 paper/paper.Rmd
more editing and reference in optim method section

@@
 abstract: >
-  Projection pursuit searches for interesting low-dimensional views of high-dimensional data via the optimisation of an index function.
-The initial paper by Friedman & Tukey in 1974 stated that  "the  technique  used  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." However, while many projection pursuit indices have been proposed in the literature, less work has been done on the optimisation procedures.
-In this paper we introduce visual diagnostics for optimisation algorithms, in particular those available in the projection pursuit guided tour.
+  Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
+The first paper of projection pursuit by @friedman1974projection stated that "the  technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While many indices have been proposed in the literature, less work has been done on evaluating the performance of the optimisers.
+In this paper, we implement a data collection object for the optimisation in the guided tour and introduce visual diagnostics based on the data object collected.
 These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
-An R package, ferrn, has been created to implement the diagnostics.
+An R package, \pkg{ferrn}, has been created to implement the diagnostics.
 preamble: >
@@
 
-Visualisation is widely used in exploratory data analysis [@tukey1977exploratory].
-Presenting information in a graphical format often allows people to uncover information they would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-The work presented in this paper combines data visualization with optimization methods.
-It aims at bring visualization tools into optimization problems to assist users to better understand the performance of their optimization methods in practice.
+Visualisation is widely used in exploratory data analysis [@tukey1977exploratory; @unwin2015graphical;  @healy2018data; @wilke2019fundamentals].
+Presenting information in graphics often unveils information that would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
+The work presented in this paper brings visualization tools into optimisation problems with an aim to better understand the performance of the optimisers in practice.
 This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms.
@@
 We also discuss how these insights contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package: ferrn, that implements the visual diagnostics.
+Finally, Section \ref{implementation} describes the R package: \pkg{ferrn}, that implements the visual diagnostics.
 
@@
 
-Our interest is on constrained optimization (REF) as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
+Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
 That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
@@
 
-Optimization methods can be divided in various classes (REF).
+Optimization methods can be divided in various classes (NOTE: do we want one more sentence on various class: global optim, convex optim, stochastic optim etc?) [@kelley1999iterative; @boyd2004convex; @fletcher2013practical].
 In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
@@
 
-Several R implementations address optimization problems with both general purpose as well as tasks specifics solvers.
-The most prominent one within the general solvers is `optim()` in the package \CRANpkg{stats}[@stats] which have functions for gradient based and derivative free methods.
-Other packages having similar functionalities include \CRANpkg{nloptr}.
+Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
+The most prominent one within the general solvers is `optim()` in the \CRANpkg{stats} [@stats] package, which have functions for both gradient based and derivative free methods.
+Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
+For more specific optimisation solvers, readers are recommended to svisite the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
 
@@
 The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
-The tour method has been implemented in the R package \CRANpkg{tourr} [@tourr], available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/)
+The tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
 
@@
 
-Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization, @bertsimas1993simulated] uses the same sampling process as `search_better` but allows a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
+Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as `search_better` but allows a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
 Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
@@
 The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
-The dataset used is \\code{boa6} where `x2` and `x7` are informative.
+The dataset used is \code{boa6} where `x2` and `x7` are informative.
 The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
@@
 
-In Figure \@ref{fig:kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
+In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
 Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated bases and hence allows a wider search.
@@
 
-The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing `tourr` package while the optimiser diagnostics have been implemented in a new package, `ferrn`.
+The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing CRAN package \CRANpkg{tourr} [@tourr] while the optimiser diagnostics have been implemented in a new package, \pkg{ferrn}.
 When a guided tour is run, the users can choose if the data from optimisation should be collected via the `verbose` argument.
-Once the data object has been obtained, the package, `ferrn`, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
+Once the data object has been obtained, the package, \pkg{ferrn}, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
 The structure of package functionality has been listed below.

### 2020-12-09 paper/paper.tex
more editing and reference in optim method section

@@
 \abstract{%
-Projection pursuit searches for interesting low-dimensional views of
-high-dimensional data via the optimisation of an index function. The
-initial paper by Friedman \& Tukey in 1974 stated that ``the technique
-used for maximising the projection index strongly influences both the
-statistical and the computational aspects of the procedure.'' However,
-while many projection pursuit indices have been proposed in the
-literature, less work has been done on the optimisation procedures. In
-this paper we introduce visual diagnostics for optimisation algorithms,
-in particular those available in the projection pursuit guided tour.
-These diagnostics and workflows can be applied to a broad class of
-optimisers, to assess their performance. An R package, ferrn, has been
-created to implement the diagnostics.
+Guided tour searches for interesting low-dimensional views of
+high-dimensional data via optimising a projection pursuit index
+function. The first paper of projection pursuit by
+\citet{friedman1974projection} stated that ``the technique used for
+maximising the projection index strongly influences both the statistical
+and the computational aspects of the procedure.'' While many indices
+have been proposed in the literature, less work has been done on
+evaluating the performance of the optimisers. In this paper, we
+implement a data collection object for the optimisation in the guided
+tour and introduce visual diagnostics based on the data object
+collected. These diagnostics and workflows can be applied to a broad
+class of optimisers, to assess their performance. An R package,
+\pkg{ferrn}, has been created to implement the diagnostics.
 }
@@
 Visualisation is widely used in exploratory data analysis
-\citep{tukey1977exploratory}. Presenting information in a graphical
-format often allows people to uncover information they would otherwise
-not be aware of and provides a more comprehensive understanding of the
-problem at hand. The work presented in this paper combines data
-visualization with optimization methods. It aims at bring visualization
-tools into optimization problems to assist users to better understand
-the performance of their optimization methods in practice. This
+\citep{tukey1977exploratory, unwin2015graphical, healy2018data, wilke2019fundamentals}.
+Presenting information in graphics often unveils information that would
+otherwise not be aware of and provides a more comprehensive
+understanding of the problem at hand. The work presented in this paper
+brings visualization tools into optimisation problems with an aim to
+better understand the performance of the optimisers in practice. This
 motivates our work of creating plots to diagnose optimisation algorithms
@@
 insights contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package: ferrn,
-that implements the visual diagnostics.
+Finally, Section \ref{implementation} describes the R package:
+\pkg{ferrn}, that implements the visual diagnostics.
 
@@
 
-Our interest is on constrained optimization (REF) as defined in the
-introduction section and assume that it is not possible to find a
-solution to the problem in the way of a closed form. That is, the
-problem consists of finding the minimum or maximum of a function
-\(f \in L^p\) in the constrained \(\mathbb{A}\) space.
-
-Optimization methods can be divided in various classes (REF). In this
-paper, we focused on methods that rely on the objective function
+Our interest is on constrained optimization
+\citep{bertsekas2014constrained} as defined in the introduction section
+and assume that it is not possible to find a solution to the problem in
+the way of a closed form. That is, the problem consists of finding the
+minimum or maximum of a function \(f \in L^p\) in the constrained
+\(\mathbb{A}\) space.
+
+Optimization methods can be divided in various classes (NOTE: do we want
+one more sentence on various class: global optim, convex optim,
+stochastic optim etc?)
+\citep{kelley1999iterative, boyd2004convex, fletcher2013practical}. In
+this paper, we focused on methods that rely on the objective function
 derivatives known as gradient methods, and those which are known as
@@
 Several R implementations address optimization problems with both
-general purpose as well as tasks specifics solvers. The most prominent
-one within the general solvers is \texttt{optim()} in the package
-\CRANpkg{stats}\citep{stats} which have functions for gradient based and
-derivative free methods. Other packages having similar functionalities
-include \CRANpkg{nloptr}.
+general purpose as well as task specifics solvers. The most prominent
+one within the general solvers is \texttt{optim()} in the
+\CRANpkg{stats} \citep{stats} package, which have functions for both
+gradient based and derivative free methods. Another general solver
+specialised in non-linear optimisation is \CRANpkg{nloptr}
+\citep{nloptr}. For more specific optimisation solvers, readers are
+recommended to svisite the relevant sections in CRAN task review on
+\emph{optimisation and mathematical programming}
+\citep{crantaskreviewoptim}.
 
@@
 The tour method has been implemented in the R package \CRANpkg{tourr}
-\citep{tourr}, available on the Comprehensive R Archive Network at
-\url{https://cran.r-project.org/web/packages/tourr/}
+\citep{tourr}.
 
@@
 Simulated annealing (\texttt{search\_better\_random})
-\citep[\citet{bertsimas1993simulated}]{kirkpatrick1983optimization} uses
-the same sampling process as \texttt{search\_better} but allows a
+\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
+same sampling process as \texttt{search\_better} but allows a
 probabilistic acceptance of a basis with lower index value based on the
@@
 optimisers perform when a local maximum is present. The dataset used is
-\textbackslash code\{boa6\} where \texttt{x2} and \texttt{x7} are
-informative. The two theoretical best bases are {[}0, 1, 0, 0, 0, 0{]}
-and {[}0, 0, 1, 0, 0, 0{]} with index value 0.176 and 0.235,
-respectively. Hence, the global maximum happens when variable
-\texttt{x7} is found. Simulation is done with 20 randomly generated
-seeds on two optimisers: \texttt{search\_better} (Figure
-\ref{fig:kol-better}) and \texttt{search\_better\_random} (Figure
-\ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}). Compared to
-Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure
-\ref{fig:kol-random-tuned}further increases the neighbourhood parameter
-\texttt{alpha} from 0.5 to 0.7 to enlarge the search space. The data
-object is collected for each simulation and computation is done to
-reduce the bases from its original 6D to a 2D points by principal
-component analysis so as to view the search paths.
+\code{boa6} where \texttt{x2} and \texttt{x7} are informative. The two
+theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0,
+0{]} with index value 0.176 and 0.235, respectively. Hence, the global
+maximum happens when variable \texttt{x7} is found. Simulation is done
+with 20 randomly generated seeds on two optimisers:
+\texttt{search\_better} (Figure \ref{fig:kol-better}) and
+\texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure
+\ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and
+\ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases
+the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge
+the search space. The data object is collected for each simulation and
+computation is done to reduce the bases from its original 6D to a 2D
+points by principal component analysis so as to view the search paths.
 
@@
 
-In Figure @ref\{fig:kol-random-tuned\}, the neighbourhood parameter
+In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter
 \texttt{alpha} is increased from a 0.5 default to 0.7 to solve the
@@
 The implementation of this projection has been divided into two
-packages: the data collection object is implemented in the existing
-\texttt{tourr} package while the optimiser diagnostics have been
-implemented in a new package, \texttt{ferrn}. When a guided tour is run,
-the users can choose if the data from optimisation should be collected
-via the \texttt{verbose} argument. Once the data object has been
-obtained, the package, \texttt{ferrn}, can provide four diagnostic plots
-as shown in Section \ref{vis-diag}. The structure of package
+packages: the data collection object is implemented in the existing CRAN
+package \CRANpkg{tourr} \citep{tourr} while the optimiser diagnostics
+have been implemented in a new package, \pkg{ferrn}. When a guided tour
+is run, the users can choose if the data from optimisation should be
+collected via the \texttt{verbose} argument. Once the data object has
+been obtained, the package, \pkg{ferrn}, can provide four diagnostic
+plots as shown in Section \ref{vis-diag}. The structure of package
 functionality has been listed below.

### 2020-12-10 paper/paper.Rmd
optim method section is good for proofreading

@@
 The work presented in this paper brings visualization tools into optimisation problems with an aim to better understand the performance of the optimisers in practice.
-This motivates our work of creating plots to diagnose optimisation algorithms which in turn can be also used not only to understand optimization performance but also to understand and compare the behaviour of different algorithms.
 
@@
 
-The motivation of this paper is the optimisation problem arising in the projection pursuit guided tour [@buja2005computational] which is an exploratory data analysis tool that is defined to detect *interesting structures* or features  in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours [@cook2008grand].
+The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational] which is an exploratory data analysis tool that is defined to detect *interesting structures* or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours [@cook2008grand].
 The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
@@
 
-Optimization methods can be divided in various classes (NOTE: do we want one more sentence on various class: global optim, convex optim, stochastic optim etc?) [@kelley1999iterative; @boyd2004convex; @fletcher2013practical].
+Optimization methods can be divided in various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex], stochastic optimisation [@nocedal2006numerical] .
 In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
 
-In gradient descent methods the goal is to find the minimum of an objective function while in gradient ascent methods the interest is in the maximum.
-The minimum is searched for by taken steps  proportional to the negative gradient or of the function at the evaluation point (analogously positive gradient for gradient ascent methods), that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
+In gradient ascent methods the goal is to find the maximum of an objective function while in gradient descent methods the interest is in the minimum.
+The maximum is searched for by taken steps  proportional to the gradient of the function at the evaluation point that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
 The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
@@
 In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate $x_{1}$, a search direction $S_k$ for each iteration where the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
+In line search methods users are required to provide an initial estimate $x_{1}$, a search direction $S_k$ for each iteration where the objective is to find the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
 Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and repeat the process until the desire convergence is reached.

### 2020-12-10 paper/paper.tex
optim method section is good for proofreading

@@
 brings visualization tools into optimisation problems with an aim to
-better understand the performance of the optimisers in practice. This
-motivates our work of creating plots to diagnose optimisation algorithms
-which in turn can be also used not only to understand optimization
-performance but also to understand and compare the behaviour of
-different algorithms.
+better understand the performance of the optimisers in practice.
 
@@
 
-The motivation of this paper is the optimisation problem arising in the
+The focus of this paper is on the optimisation problem arising in the
 projection pursuit guided tour \citep{buja2005computational} which is an
@@
 
-Optimization methods can be divided in various classes (NOTE: do we want
-one more sentence on various class: global optim, convex optim,
-stochastic optim etc?)
-\citep{kelley1999iterative, boyd2004convex, fletcher2013practical}. In
-this paper, we focused on methods that rely on the objective function
-derivatives known as gradient methods, and those which are known as
-derivative free methods where optimization does not rely on the gradient
-or derivatives information.
-
-In gradient descent methods the goal is to find the minimum of an
-objective function while in gradient ascent methods the interest is in
-the maximum. The minimum is searched for by taken steps proportional to
-the negative gradient or of the function at the evaluation point
-(analogously positive gradient for gradient ascent methods), that is
-\(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or
+Optimization methods can be divided in various classes, such as global
+optimisation \citep{kelley1999iterative, fletcher2013practical}, convex
+optimisation \citep{boyd2004convex}, stochastic optimisation
+\citep{nocedal2006numerical} . In this paper, we focused on methods that
+rely on the objective function derivatives known as gradient methods,
+and those which are known as derivative free methods where optimization
+does not rely on the gradient or derivatives information.
+
+In gradient ascent methods the goal is to find the maximum of an
+objective function while in gradient descent methods the interest is in
+the minimum. The maximum is searched for by taken steps proportional to
+the gradient of the function at the evaluation point that is
+\(x_{k+1} = x_k +\alpha_kS_k\) where \(\alpha_k\) is the step size or
 the learning rate \citep{fletcher2013practical}. The step size is the
@@
 each iteration where the objective is to find the value of
-\(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with
+\(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
 respect to \(\alpha_k\) (this step is called the line search). Then move

### 2020-12-10 paper/paper.Rmd
update of optimisation section

@@
 Optimization problems are ubiquitous in many areas of study.
-While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
-These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existence of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
-In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.
+While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
+These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
+Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the possible existence of multiple maxima (local and global), the trade-off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
+In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.
 
-Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section and assume that it is not possible to find a solution to the problem in the way of a closed form.
+Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
 That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
@@
 
-Optimization methods can be divided in various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex], stochastic optimisation [@nocedal2006numerical] .
-In this paper, we focused on methods that rely on the objective function derivatives known as gradient methods, and those which are known as derivative free methods where optimization does not rely on the gradient or derivatives information.
+Optimization methods can be divided into various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex] or stochastic optimisation [@nocedal2006numerical].
+In this paper, we consider methods that rely on the objective function derivatives known as gradient methods, as well as on methods known as derivative-free methods where optimization does not rely on the knowledge of the gradient or other derivatives.
 
 In gradient ascent methods the goal is to find the maximum of an objective function while in gradient descent methods the interest is in the minimum.
-The maximum is searched for by taken steps  proportional to the gradient of the function at the evaluation point that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
-The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence.
-These methods are known to be relative slow and can have convergence issues when problems are poorly conditioned [@trefethen1997numerical].
+The maximum is searched for by taking steps proportional to the gradient of the function at the evaluation point, that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
+One difficulty is the choice of step size, as inadequate step sizes might prevent the optimisation method to converge to an extremum.
+Moreover, these methods are known to be relative slow and can have convergence issues when problems are poorly conditioned [@trefethen1997numerical].
 In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
-In line search methods users are required to provide an initial estimate $x_{1}$, a search direction $S_k$ for each iteration where the objective is to find the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
-Then move on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and repeat the process until the desire convergence is reached.
-Modern development of line search methods focuses on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational storing
-burden of the intermediate computations while catering for practical optimisation problems.
-In the next section two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
+In line search methods, users are required to provide an initial estimate $x_{1}$ and a search direction $S_k$ for each iteration.
+Then, the objective is to find the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
+Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached.
+Modern developments of line search methods focus on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational burden of storing intermediate computations while catering for practical optimisation problems.
+In the next section, two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
 
-Although gradient optimization methods are very popular, they rely on the existence of the objective function derivatives and on the complexity of the constraints.
-Because of that, the second big class of optimization problems look at derivative free methods where the emphasis is on finding, in most cases, a near optimal solution.
+Although gradient optimization methods are very popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
+Because of that, a large class of derivative-free optimization methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
 Examples of those include response surface methodology [@box1951experimental], stochastic approximation [@robbins1951stochastic], random search [@fu2015handbook] and  heuristic methods [@sorensen2013metaheuristics].
-Later, we will present an optimization algorithm within the random search class, namely simulated annealing for optimization with the guided tour.
-Moreover, we will introduce an algorithm to further improve optimization solutions both in gradient descent methods and in derive free optimization.
+Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour.
+Moreover, we will introduce an algorithm to further improve optimization solutions with both gradient descent methods and derivative-free optimisation.
 
 Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
-The most prominent one within the general solvers is `optim()` in the \CRANpkg{stats} [@stats] package, which have functions for both gradient based and derivative free methods.
+The most prominent one within the general solvers is `optim()` in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions.
 Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
-For more specific optimisation solvers, readers are recommended to svisite the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
+For more specific optimisation solvers, readers are recommended to visit the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
 

### 2020-12-13 paper/paper.Rmd
modify as per Nicolas suggestion and author

@@
   - name: H.Sherry Zhang
-    affiliation: Department of Econometrics and Business Statistics, Monash University 
-    email:  Huize.Zhang@monash.edu
+    affiliation: Monash University 
+    address: Department of Econometrics and Business Statistics
+    email:  huize.zhang@monash.edu
+  - name: Dianne Cook
+    affiliation: Monash University 
+    address: Department of Econometrics and Business Statistics
+    email:  dicook@monash.edu
+  - name: Ursula Laa
+    affiliation: University of Natural Resources and Life Sciences
+    address: Institute of Statistics
+    email:  ursula.laa@boku.ac.at  
+  - name: Nicolas Langrené
+    affiliation: Data61 CSIRO 
+    address: Locked Bag 38004 Docklands VIC 8012 Australia
+    email: Nicolas.Langrene@data61.csiro.au
+  - name: Patricia Menéndez
+    affiliation: Monash University 
+    address: Department of Econometrics and Business Statistics
+    email:  patricia.menendez@monash.edu
+
 abstract: >
@@
 Optimization methods can be divided into various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex] or stochastic optimisation [@nocedal2006numerical].
-In this paper, we consider methods that rely on the objective function derivatives known as gradient methods, as well as on methods known as derivative-free methods where optimization does not rely on the knowledge of the gradient or other derivatives.
+A large class of methods utilises the gradient information in the objective function with the most notable one being the gradient ascent (descent) method.
+Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
+Because of that, we consider another class of method in this paper: derivative-free methods.
+Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
+Examples of those include response surface methodology [@box1951experimental], stochastic approximation [@robbins1951stochastic], random search [@fu2015handbook] and heuristic methods [@sorensen2013metaheuristics].
+Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour.
 
-In gradient ascent methods the goal is to find the maximum of an objective function while in gradient descent methods the interest is in the minimum.
-The maximum is searched for by taking steps proportional to the gradient of the function at the evaluation point, that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
+<!-- In this paper, we consider derivative free method without the knowledge of the gradient. Somd df minic the graident method, biggest class is line search, then SA -->
+
+<!--  In the case of gradient ascent, the maximum is searched by taking steps proportional to the search direction at the evaluation point, that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
 One difficulty is the choice of step size, as inadequate step sizes might prevent the optimisation method to converge to an extremum.
 Moreover, these methods are known to be relative slow and can have convergence issues when problems are poorly conditioned [@trefethen1997numerical].
-In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
-In line search methods, users are required to provide an initial estimate $x_{1}$ and a search direction $S_k$ for each iteration.
-Then, the objective is to find the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line search).
+-->
+
+A common search scheme utilised by both derivative-free methods and gradient methods is line search.
+In line search methods, users are required to provide an initial estimate $x_{1}$ and at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
 Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached.
-Modern developments of line search methods focus on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational burden of storing intermediate computations while catering for practical optimisation problems.
-In the next section, two line search methods for the optimisation used in the projection pursuit guided tour will be presented.
+While gradient-based methods choose the search direction by the gradient, derivative-free methods determine the search direction through local information of the objective function.
+The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an extremum.
+An ideal step size can be chosen via finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.
 
-Although gradient optimization methods are very popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
-Because of that, a large class of derivative-free optimization methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
-Examples of those include response surface methodology [@box1951experimental], stochastic approximation [@robbins1951stochastic], random search [@fu2015handbook] and  heuristic methods [@sorensen2013metaheuristics].
-Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour.
-Moreover, we will introduce an algorithm to further improve optimization solutions with both gradient descent methods and derivative-free optimisation.
+<!-- Modern developments of line search methods focus on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational burden of storing intermediate computations while catering for practical optimisation problems.
+-->
+
+<!-- methods that rely on the objective function derivatives known as gradient methods, as well as on methods known as derivative-free methods where optimization does not rely on the knowledge of the gradient or other derivatives.
+-->
+
+<!-- In gradient ascent methods the goal is to find the maximum of an objective function while in gradient descent methods the interest is in the minimum.  -->
+ <!-- In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
+-->
+
+ <!--  In the next section, two line search methods for the optimisation used in the projection pursuit guided tour will be presented. -->
 
@@
 Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
-For more specific optimisation solvers, readers are recommended to visit the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
+Specific solvers for simulated annealing includes `optim(..., method  "SANN")` and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
+For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
 

### 2020-12-13 paper/paper.tex
modify as per Nicolas suggestion and author

@@
 guided tours}
-\author{by H.Sherry Zhang}
+\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
 
@@
 Optimization problems are ubiquitous in many areas of study. While in
-some cases, analytical solutions can be found, the majority of problems
+some cases analytical solutions can be found, the majority of problems
 rely on numerical methods to find the optimal solution. These numerical
-methods follow interactive approaches that aim at finding the optimum by
+methods follow iterative approaches that aim at finding the optimum by
 progressively improving the current solution until a desirable accuracy
 is achieved. Although that principle seems uncomplicated, optimization
-methods need to deal with many challenges such as the existence of
-multiple maxima (local and global), the trade off between desirable
+methods need to deal with many challenges such as the possible existence
+of multiple maxima (local and global), the trade-off between desirable
 accuracy and computational burden, as well as dealing with noisy
-objective functions and possible constraints. In addition, optimization
-results might depend on the algorithm starting values affecting the
-consistency of results.
+objective functions and possible constraints. In addition, the
+optimization results might depend on the algorithm starting values,
+affecting the consistency of results.
 
 Our interest is on constrained optimization
-\citep{bertsekas2014constrained} as defined in the introduction section
-and assume that it is not possible to find a solution to the problem in
-the way of a closed form. That is, the problem consists of finding the
+\citep{bertsekas2014constrained} as defined in the introduction section,
+and assuming it is not possible to find a solution to the problem in the
+way of a closed-form. That is, the problem consists of finding the
 minimum or maximum of a function \(f \in L^p\) in the constrained
@@
 
-Optimization methods can be divided in various classes, such as global
+Optimization methods can be divided into various classes, such as global
 optimisation \citep{kelley1999iterative, fletcher2013practical}, convex
-optimisation \citep{boyd2004convex}, stochastic optimisation
-\citep{nocedal2006numerical} . In this paper, we focused on methods that
-rely on the objective function derivatives known as gradient methods,
-and those which are known as derivative free methods where optimization
-does not rely on the gradient or derivatives information.
-
-In gradient ascent methods the goal is to find the maximum of an
-objective function while in gradient descent methods the interest is in
-the minimum. The maximum is searched for by taken steps proportional to
-the gradient of the function at the evaluation point that is
-\(x_{k+1} = x_k +\alpha_kS_k\) where \(\alpha_k\) is the step size or
-the learning rate \citep{fletcher2013practical}. The step size is the
-key in the success of the optimization problem as inadequate step sizes
-might lead to a failure in convergence. These methods are known to be
-relative slow and can have convergence issues when problems are poorly
-conditioned \citep{trefethen1997numerical}. In order to solve some of
-those problems, line minimization or line search methods were proposed
-\citep{shi2004convergence} where the step size is chosen in such a way
-that convergence is ensured. In line search methods users are required
-to provide an initial estimate \(x_{1}\), a search direction \(S_k\) for
-each iteration where the objective is to find the value of
+optimisation \citep{boyd2004convex} or stochastic optimisation
+\citep{nocedal2006numerical}. A large class of methods utilises the
+gradient information in the objective function with the most notable one
+being the gradient ascent (descent) method. Although gradient
+optimization methods are popular, they rely on the availability of the
+objective function derivatives and on the complexity of the constraints.
+Because of that, we consider another class of method in this paper:
+derivative-free methods. Derivative-free methods have been developed
+over the years, where the emphasis is on finding, in most cases, a near
+optimal solution. Examples of those include response surface methodology
+\citep{box1951experimental}, stochastic approximation
+\citep{robbins1951stochastic}, random search \citep{fu2015handbook} and
+heuristic methods \citep{sorensen2013metaheuristics}. Later, we will
+present a simulated annealing optimisation algorithm, which belongs to
+the class of random search methods, for optimisation with the guided
+tour.
+
+A common search scheme utilised by both derivative-free methods and
+gradient methods is line search. In line search methods, users are
+required to provide an initial estimate \(x_{1}\) and at each iteration,
+a search direction \(S_k\) and a step size \(\alpha_k\) are generated.
+Then one moves on to the next point following
+\(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
+desire convergence is reached. While gradient-based methods choose the
+search direction by the gradient, derivative-free methods determine the
+search direction through local information of the objective function.
+The choice of step size also needs considerations, as inadequate step
+sizes might prevent the optimisation method to converge to an extremum.
+An ideal step size can be chosen via finding the value of
 \(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
-respect to \(\alpha_k\) (this step is called the line search). Then move
-on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and
-repeat the process until the desire convergence is reached. Modern
-development of line search methods focuses on improving the computation
-of the searching direction \(S_{k}\), and on approximations of the step
-size \(\alpha_k\) together with reducing the computational storing
-burden of the intermediate computations while catering for practical
-optimisation problems. In the next section two line search methods for
-the optimisation used in the projection pursuit guided tour will be
-presented.
-
-Although gradient optimization methods are very popular, they rely on
-the existence of the objective function derivatives and on the
-complexity of the constraints. Because of that, the second big class of
-optimization problems look at derivative free methods where the emphasis
-is on finding, in most cases, a near optimal solution. Examples of those
-include response surface methodology \citep{box1951experimental},
-stochastic approximation \citep{robbins1951stochastic}, random search
-\citep{fu2015handbook} and heuristic methods
-\citep{sorensen2013metaheuristics}. Later, we will present an
-optimization algorithm within the random search class, namely simulated
-annealing for optimization with the guided tour. Moreover, we will
-introduce an algorithm to further improve optimization solutions both in
-gradient descent methods and in derive free optimization.
+respect to \(\alpha_k\) at each iteration.
 
@@
 one within the general solvers is \texttt{optim()} in the
-\CRANpkg{stats} \citep{stats} package, which have functions for both
-gradient based and derivative free methods. Another general solver
-specialised in non-linear optimisation is \CRANpkg{nloptr}
-\citep{nloptr}. For more specific optimisation solvers, readers are
-recommended to svisite the relevant sections in CRAN task review on
-\emph{optimisation and mathematical programming}
-\citep{crantaskreviewoptim}.
+\CRANpkg{stats} \citep{stats} package, which provides both
+gradient-based and derivative-free optimisation functions. Another
+general solver specialised in non-linear optimisation is
+\CRANpkg{nloptr} \citep{nloptr}. Specific solvers for simulated
+annealing includes \texttt{optim(...,\ method\ \ "SANN")} and package
+\CRANpkg{GenSA} \citep{gensa} that deals with more complicated objective
+functions. For other task specific solvers, readers are recommended to
+visit the relevant sections in CRAN task review on \emph{optimisation
+and mathematical programming} \citep{crantaskreviewoptim}.
 
@@
 H.Sherry Zhang\\
-Department of Econometrics and Business Statistics, Monash University\\
-\\
-}
-\href{mailto:Huize.Zhang@monash.edu}{\nolinkurl{Huize.Zhang@monash.edu}}
+Monash University\\
+Department of Econometrics and Business Statistics\\
+}
+\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
+
+\address{%
+Dianne Cook\\
+Monash University\\
+Department of Econometrics and Business Statistics\\
+}
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+
+\address{%
+Ursula Laa\\
+University of Natural Resources and Life Sciences\\
+Institute of Statistics\\
+}
+\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
+
+\address{%
+Nicolas Langrené\\
+Data61 CSIRO\\
+Locked Bag 38004 Docklands VIC 8012 Australia\\
+}
+\href{mailto:Nicolas.Langrene@data61.csiro.au}{\nolinkurl{Nicolas.Langrene@data61.csiro.au}}
+
+\address{%
+Patricia Menéndez\\
+Monash University\\
+Department of Econometrics and Business Statistics\\
+}
+\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}

### 2020-12-17 paper/paper.Rmd
update paper.Rmd
polishing of the optimization section

@@
   - name: Nicolas Langrené
-    affiliation: Data61 CSIRO 
-    address: Locked Bag 38004 Docklands VIC 8012 Australia
-    email: Nicolas.Langrene@data61.csiro.au
+    affiliation: CSIRO Data61 
+    address: 34 Village Street, Docklands VIC 3008 Australia
+    email: nicolas.langrene@csiro.au
   - name: Patricia Menéndez
@@
 These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the possible existence of multiple maxima (local and global), the trade-off between desirable accuracy and computational burden, as well as dealing with noisy objective functions and possible constraints.
+Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden.
 In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.
-
-Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
 
@@
 Optimization methods can be divided into various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex] or stochastic optimisation [@nocedal2006numerical].
-A large class of methods utilises the gradient information in the objective function with the most notable one being the gradient ascent (descent) method.
+Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
+That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
+
+A large class of methods utilises the gradient information of the objective function to perform the optimisation iterations, with the most notable one being the gradient ascent (descent) method.
 Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
-Because of that, we consider another class of method in this paper: derivative-free methods.
+Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable.
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.

### 2021-01-02 paper/paper.Rmd
last go-through before the vis-diag section

@@
 The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
-Broadly speaking optimization can be unconstrained or constrained [@kelley1999iterative].
+Broadly speaking, optimization can be unconstrained or constrained [@kelley1999iterative].
 The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
@@
 $\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
-<!-- $f: S \rightarrow \mathbb{R}$, with $S \in \mathbb{R}^n$, in a reduced space given by the problem constraints on $f$.  -->
 
@@
 The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
-The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most reveling structures in a high dimensional space [@cook1993projection].
+The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most revelling structures in a high dimensional space [@cook1993projection].
 
 The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable [@jones1998efficient].
-Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.
+Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non differentiable [@jones1998efficient].
+Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily applied to any other optimization problems.
 
@@
 
-<!-- For a given problem we might rely on numerical methods to find the optimal solution, since an analytical approach, requiring a closed-form solution, is often not possible for complex problems.
--->
-<!-- An iterative approach finds the optimum by progressively improving the current solution.
-Researchers can then decide when to stop if a desirable accuracy has been achieved.
-One potential issue with this technique is that it may end up finding a local optimum.
--->
-
 Optimization problems are ubiquitous in many areas of study.
@@
 In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.
-
-<!-- In this scenario, gradient descent or gradient ascent methods) have taken the stage for many decades and rely in the usage of the objective function derivatives or their approximations. -->
 
@@
 
-<!-- In this paper, we consider derivative free method without the knowledge of the gradient. Somd df minic the graident method, biggest class is line search, then SA -->
-
-<!--  In the case of gradient ascent, the maximum is searched by taking steps proportional to the search direction at the evaluation point, that is $x_{k+1} = x_k +\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate [@fletcher2013practical].
-One difficulty is the choice of step size, as inadequate step sizes might prevent the optimisation method to converge to an extremum.
-Moreover, these methods are known to be relative slow and can have convergence issues when problems are poorly conditioned [@trefethen1997numerical].
--->
-
 A common search scheme utilised by both derivative-free methods and gradient methods is line search.
-In line search methods, users are required to provide an initial estimate $x_{1}$ and at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
+In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
 Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached.
-While gradient-based methods choose the search direction by the gradient, derivative-free methods determine the search direction through local information of the objective function.
-The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an extremum.
+While gradient-based methods choose the search direction by the gradient, derivative-free methods uses local information of the objective function to determine the search direction.
+The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an optimum.
 An ideal step size can be chosen via finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.
-
-<!-- Modern developments of line search methods focus on improving the computation of the searching direction $S_{k}$, and on approximations of the step size $\alpha_k$ together with reducing the computational burden of storing intermediate computations while catering for practical optimisation problems.
--->
-
-<!-- methods that rely on the objective function derivatives known as gradient methods, as well as on methods known as derivative-free methods where optimization does not rely on the knowledge of the gradient or other derivatives.
--->
-
-<!-- In gradient ascent methods the goal is to find the maximum of an objective function while in gradient descent methods the interest is in the minimum.  -->
- <!-- In order to solve some of those problems, line minimization or line search methods were proposed [@shi2004convergence] where the step size is chosen in such a way that convergence is ensured.
--->
-
- <!--  In the next section, two line search methods for the optimisation used in the projection pursuit guided tour will be presented. -->
 
@@
 
-A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].
+A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] for supervised classification problems, and PDA index [@lee2010projection], which is an extension of the LDA index.
 
 As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
-Different tour types choose these planes differently, for example, a grand tour [@cook2008grand] selects the planes randomly and a manual tour [@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
-Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit.
+Different tour types choose these planes differently, for example, a grand tour [@cook2008grand] selects the planes randomly to provide a general overview and a manual tour [@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
+Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit to gradually reveal the most interesting projection in the low dimension space.
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
-
+Intuitively, Figure \ref{fig:tour-path} shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-Figure \ref{fig:tour-path} shows a sketch of the tour path.
-The blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
 The tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
@@
 
-pa
-The optimisation problem in the tour context is stated as follows:
-....
-.Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
+The optimisation problem in the tour context is stated as follows: Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
 
@@
 
-Below we introduce three possible algorithms: `search_better`, `search_better_random`, and `search_geodesic`.
-The first two are derivative free methods that sample candidate bases in the neighbourhood whilst `search_geodesic` is an analogue of gradient ascent on the projection basis space.
+Below we introduce three line search algorithms in tour: `search_better`, `search_better_random`, and `search_geodesic`.
+The first two are simulated annealing algorithms whilst `search_geodesic` evaluates neighbourhood basis to determine the search direction.
 
@@
 \SetAlgoLined
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
-    \Output{$\mathbf{A}_{l}$}
-  initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    \If{$I_{l} > I_{\text{cur}}$}{
-      \KwRet{$\mathbf{A}_{l}$} \;
+  \SetKwInOut{input}{input}
+  \SetKwInOut{output}{output}
+    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
+    \output{$\mathbf{A}_{l}$}
+    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
+   set $l = 1$\;
+  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
+    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    update $l = l + 1$\;
   }
-    $l = l + 1$\;
+  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
 }
@@
 The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
-A slightly different cooling scheme has been proposed by @posse1995projection to include a halving parameter $c$.
-Rather than reducing the radius of the searching neighbourhood, $\alpha$, at each iteration, Posse's design only adjust $\alpha$ if the last search takes more than $c$ times to find an accepted basis to avoid the searching space being reduced too fast.
+A slightly different cooling scheme has been proposed by @posse1995projection to avoid the search space being reduced too fast.
+A halving parameter $c$ is introduced and $\alpha$ is only adjusted if the last search takes more than $c$ times to find an accepted basis.
 
@@
 \SetAlgoLined
-    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
-      \eIf{$I_{l} > I_{\text{cur}}$}{
-        \KwRet{$\mathbf{A}_{l}$} \;
-      }{
-        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-        \If{$P > U$}{
-           \KwRet{$\mathbf{A}_{l}$} \;
-        }
+\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
+    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+    draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+    update $l = l + 1$\;
   }
@@
 This feature is particularly useful when local maxima are present.
-The algorithm can be written as replacing line 5-8 of Algorithm \ref{random-search} with Algorithm \ref{simulated_annealing}.
+The algorithm \ref{simulated_annealing} highlights how simulated annealing differs from random search in the inner loop.
 
@@
 \SetAlgoLined
-\SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n = 5$, $\delta$}
-    \Output{$\mathbf{A}_{**}$}
-  initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $2n$ bases in a small neighbourhood, $\delta$, of $\mathbf{A}_{\text{cur}}$ and ensure orthogonality \;
-    Find the one with the largest index value: $\mathbf{A}_{*}$\;
-    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
-    Optimise the index value on the geodesic $\mathcal{G}$ over a 90 degree window to produce the optima $\mathbf{A}_{**}$  \;
-    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-      \If{$p_{\text{diff}} > 0.001$}{
-         \KwRet{$\mathbf{A}_{**}$} \;
-      }
-    $l = l + 1$\;
+\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
+  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
+  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
+  compute the corresponding index value for each candidate bases\;
+  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
+  determine the step size via optimising the index value on the search direction over a 90 degree window\;
+  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
+  update $l = l + 1$\;
 }
@@
 
-@cook1995grand used a gradient ascent algorithm on the space of the projection bases.
-In gradient ascent, one first finds the direction for improvement via computing the gradient information.
-In `search_geodesic`, $2n$ bases are first generated in a tiny neighbourhood of the current basis, controlled by the neighbourhood parameter $\delta$.
-A geodesic is then constructed using the current basis and the one in $2n$ bases with the highest index value.
-If the neighbourhood parameter $\delta$ is tiny, the geodesic constructed is an analogue of the gradient information in the curved space and works as the searching direction.
-The next step in gradient ascent is to conduct a line search to find the best improvement along the gradient direction and in `search_geodesic`, this is replaced by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$.
-The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the termination condition on percentage improvement.
-The procedure will also terminate if $l_{\max}$ is reached.
-Algorithm \ref{search-geodesic} summarises the steps in geodesic search.
+@cook1995grand used a line search algorithm on the space of the projection bases.
+In `search_geodesic`, the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
+The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen.
+The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the percentage improve condition or when $l_{\max}$ is reached.
+Algorithm \ref{search-geodesic} summarises the inner loop in geodesic search.
 

### 2021-01-02 paper/paper.tex
last go-through before the vis-diag section

@@
 the space of all feasible solutions where typically the best solution is
-decided by an objective function. Broadly speaking optimization can be
+decided by an objective function. Broadly speaking, optimization can be
 unconstrained or constrained \citep{kelley1999iterative}. The
@@
 are formally defined by a function of projections, called index function
-which is optimized to uncover the most reveling structures in a high
+which is optimized to uncover the most revelling structures in a high
 dimensional space \citep{cook1993projection}.
@@
 dealing with noisy objective functions that might be non smooth and non
-diferentiable \citep{jones1998efficient}. Those are not unique to this
+differentiable \citep{jones1998efficient}. Those are not unique to this
 context and therefore the visualization tools and optimization methods
-presented in this paper can be easily apply to any other optimization
+presented in this paper can be easily applied to any other optimization
 problems.
@@
 progressively improving the current solution until a desirable accuracy
-is achieved. Although that principle seems uncomplicated, optimization
-methods need to deal with many challenges such as the possible existence
-of multiple maxima (local and global), the trade-off between desirable
-accuracy and computational burden, as well as dealing with noisy
-objective functions and possible constraints. In addition, the
-optimization results might depend on the algorithm starting values,
-affecting the consistency of results.
-
-Our interest is on constrained optimization
-\citep{bertsekas2014constrained} as defined in the introduction section,
-and assuming it is not possible to find a solution to the problem in the
-way of a closed-form. That is, the problem consists of finding the
-minimum or maximum of a function \(f \in L^p\) in the constrained
-\(\mathbb{A}\) space.
+is achieved. Although this principle seems uncomplicated, a number of
+challenges arise such as the possible existence of multiple maxima
+(local and global), constraints and noisy objective function, and the
+trade-off between desirable accuracy and computational burden. In
+addition, the optimization results might depend on the algorithm
+starting values, affecting the consistency of results.
 
@@
 optimisation \citep{boyd2004convex} or stochastic optimisation
-\citep{nocedal2006numerical}. A large class of methods utilises the
-gradient information in the objective function with the most notable one
-being the gradient ascent (descent) method. Although gradient
-optimization methods are popular, they rely on the availability of the
-objective function derivatives and on the complexity of the constraints.
-Because of that, we consider another class of method in this paper:
-derivative-free methods. Derivative-free methods have been developed
-over the years, where the emphasis is on finding, in most cases, a near
-optimal solution. Examples of those include response surface methodology
-\citep{box1951experimental}, stochastic approximation
-\citep{robbins1951stochastic}, random search \citep{fu2015handbook} and
-heuristic methods \citep{sorensen2013metaheuristics}. Later, we will
-present a simulated annealing optimisation algorithm, which belongs to
-the class of random search methods, for optimisation with the guided
-tour.
+\citep{nocedal2006numerical}. Our interest is on constrained
+optimization \citep{bertsekas2014constrained} as defined in the
+introduction section, and assuming it is not possible to find a solution
+to the problem in the way of a closed-form. That is, the problem
+consists of finding the minimum or maximum of a function \(f \in L^p\)
+in the constrained \(\mathbb{A}\) space.
+
+A large class of methods utilises the gradient information of the
+objective function to perform the optimisation iterations, with the most
+notable one being the gradient ascent (descent) method. Although
+gradient optimization methods are popular, they rely on the availability
+of the objective function derivatives and on the complexity of the
+constraints. Derivative-free methods, which do not rely on the knowledge
+of the gradient, are more generally applicable. Derivative-free methods
+have been developed over the years, where the emphasis is on finding, in
+most cases, a near optimal solution. Examples of those include response
+surface methodology \citep{box1951experimental}, stochastic
+approximation \citep{robbins1951stochastic}, random search
+\citep{fu2015handbook} and heuristic methods
+\citep{sorensen2013metaheuristics}. Later, we will present a simulated
+annealing optimisation algorithm, which belongs to the class of random
+search methods, for optimisation with the guided tour.
 
@@
 gradient methods is line search. In line search methods, users are
-required to provide an initial estimate \(x_{1}\) and at each iteration,
-a search direction \(S_k\) and a step size \(\alpha_k\) are generated.
-Then one moves on to the next point following
+required to provide an initial estimate \(x_{1}\) and, at each
+iteration, a search direction \(S_k\) and a step size \(\alpha_k\) are
+generated. Then one moves on to the next point following
 \(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
 desire convergence is reached. While gradient-based methods choose the
-search direction by the gradient, derivative-free methods determine the
-search direction through local information of the objective function.
+search direction by the gradient, derivative-free methods uses local
+information of the objective function to determine the search direction.
 The choice of step size also needs considerations, as inadequate step
-sizes might prevent the optimisation method to converge to an extremum.
+sizes might prevent the optimisation method to converge to an optimum.
 An ideal step size can be chosen via finding the value of
@@
 \citep{cook1993projection}, chi-square index
-\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
-index \citep{lee2010projection}.
+\citep{posse1995projection}, LDA index \citep{lee2005projection} for
+supervised classification problems, and PDA index
+\citep{lee2010projection}, which is an extension of the LDA index.
 
@@
 tour types choose these planes differently, for example, a grand tour
-\citep{cook2008grand} selects the planes randomly and a manual tour
-\citep{cook1997manual} gradually phases in and out one variable, to
-understand the contribution of that variable in the projection. Guided
-tour, the main interest of this paper, chooses planes with the aid of
-projection pursuit. Given a random start, projection pursuit iteratively
-finds bases with higher index values and the guided tour constructs the
-geodesic interpolation between these planes to form a tour path.
-
-Mathematical details of the geodesic interpolation can be found in
-\citet{buja2005computational}. Figure \ref{fig:tour-path} shows a sketch
-of the tour path. The blue frames are produced by the projection pursuit
-optimisation algorithm, and the white frames interpolate between them.
-The tour method has been implemented in the R package \CRANpkg{tourr}
+\citep{cook2008grand} selects the planes randomly to provide a general
+overview and a manual tour \citep{cook1997manual} gradually phases in
+and out one variable, to understand the contribution of that variable in
+the projection. Guided tour, the main interest of this paper, chooses
+planes with the aid of projection pursuit to gradually reveal the most
+interesting projection in the low dimension space. Given a random start,
+projection pursuit iteratively finds bases with higher index values and
+the guided tour constructs the geodesic interpolation between these
+planes to form a tour path. Intuitively, Figure \ref{fig:tour-path}
+shows a sketch of the tour path where the blue frames are produced by
+the projection pursuit optimisation algorithm, and the white frames
+interpolate between them. Mathematical details of the geodesic
+interpolation can be found in \citet{buja2005computational}. The tour
+method has been implemented in the R package \CRANpkg{tourr}
 \citep{tourr}.
@@
 
-pa The optimisation problem in the tour context is stated as follows:
-\ldots. .Given a randomly generated starting basis \(\mathbf{A}_1\),
-projection pursuit finds the final projection basis \(\mathbf{A}_T\)
-that satisfies the following optimisation problem:
+The optimisation problem in the tour context is stated as follows: Given
+a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit
+finds the final projection basis \(\mathbf{A}_T\) that satisfies the
+following optimisation problem:
 
@@
 
-Below we introduce three possible algorithms: \texttt{search\_better},
-\texttt{search\_better\_random}, and \texttt{search\_geodesic}. The
-first two are derivative free methods that sample candidate bases in the
-neighbourhood whilst \texttt{search\_geodesic} is an analogue of
-gradient ascent on the projection basis space.
+Below we introduce three line search algorithms in tour:
+\texttt{search\_better}, \texttt{search\_better\_random}, and
+\texttt{search\_geodesic}. The first two are simulated annealing
+algorithms whilst \texttt{search\_geodesic} evaluates neighbourhood
+basis to determine the search direction.
 
@@
 \SetAlgoLined
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{output}
-    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
-    \Output{$\mathbf{A}_{l}$}
-  initialisation\;
-  Set $l = 1$\;
-  \While{$l < l_{\max}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    \If{$I_{l} > I_{\text{cur}}$}{
-      \KwRet{$\mathbf{A}_{l}$} \;
-      }
-    $l = l + 1$\;
+  \SetKwInOut{input}{input}
+  \SetKwInOut{output}{output}
+    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
+    \output{$\mathbf{A}_{l}$}
+    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
+   set $l = 1$\;
+  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
+    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    update $l = l + 1$\;
+  }
+  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
 }
@@
 \ref{random-search}. A slightly different cooling scheme has been
-proposed by \citet{posse1995projection} to include a halving parameter
-\(c\). Rather than reducing the radius of the searching neighbourhood,
-\(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if
-the last search takes more than \(c\) times to find an accepted basis to
-avoid the searching space being reduced too fast.
+proposed by \citet{posse1995projection} to avoid the search space being
+reduced too fast. A halving parameter \(c\) is introduced and \(\alpha\)
+is only adjusted if the last search takes more than \(c\) times to find
+an accepted basis.
 
@@
 feature is particularly useful when local maxima are present. The
-algorithm can be written as replacing line 5-8 of Algorithm
-\ref{random-search} with Algorithm \ref{simulated_annealing}.
+algorithm \ref{simulated_annealing} highlights how simulated annealing
+differs from random search in the inner loop.
 
@@
 
-\citet{cook1995grand} used a gradient ascent algorithm on the space of
-the projection bases. In gradient ascent, one first finds the direction
-for improvement via computing the gradient information. In
-\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
-neighbourhood of the current basis, controlled by the neighbourhood
-parameter \(\delta\). A geodesic is then constructed using the current
-basis and the one in \(2n\) bases with the highest index value. If the
-neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is
-an analogue of the gradient information in the curved space and works as
-the searching direction. The next step in gradient ascent is to conduct
-a line search to find the best improvement along the gradient direction
-and in \texttt{search\_geodesic}, this is replaced by optimising the
-index value along the geodesic direction over an 90 degree angle from
-\(-\pi/4\) to \(\pi/4\). The optima \(\mathbf{A}_{**}\) is returned for
-the current iteration if it meets the termination condition on
-percentage improvement. The procedure will also terminate if
+\citet{cook1995grand} used a line search algorithm on the space of the
+projection bases. In \texttt{search\_geodesic}, the search direction is
+computed using the most prominent direction that deviating an tiny angle
+of \(\delta\) from the current basis. The step size is chosen by
+optimising the index value along the geodesic direction over an 90
+degree angle from \(-\pi/4\) to \(\pi/4\) along the search direction
+chosen. The optima \(\mathbf{A}_{**}\) is returned for the current
+iteration if it meets the percentage improve condition or when
 \(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
-steps in geodesic search.
+inner loop in geodesic search.
 
@@
 Nicolas Langrené\\
-Data61 CSIRO\\
-Locked Bag 38004 Docklands VIC 8012 Australia\\
-}
-\href{mailto:Nicolas.Langrene@data61.csiro.au}{\nolinkurl{Nicolas.Langrene@data61.csiro.au}}
+CSIRO Data61\\
+34 Village Street, Docklands VIC 3008 Australia\\
+}
+\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
 

### 2021-01-03 paper/paper.Rmd
add literature on past incident of using vis to understand optim - distill article; changes in vis diag

@@
 Presenting information in graphics often unveils information that would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
+Task specific tools presented by @li2020visualizing show how visualisation can be used to understand the behaviour of neural network on classification models, but no general visualisation tool available for diagnosing optimisation procedure.
 The work presented in this paper brings visualization tools into optimisation problems with an aim to better understand the performance of the optimisers in practice.
@@
   \caption{simulated annealing}
-  \label{simulated_annealing}
+  \label{simulated-annealing}
 \end{algorithm}
@@
 This feature is particularly useful when local maxima are present.
-The algorithm \ref{simulated_annealing} highlights how simulated annealing differs from random search in the inner loop.
+The algorithm \ref{simulated-annealing} highlights how simulated annealing differs from random search in the inner loop.
 
@@
 To be able to make diagnostics on the optimisers, the algorithms need to populate a data structure with key elements of the algorithm.
-When the algorithms run, key information regarding the decision variable, objective function and hyper-parameters, needs to be recorded and stored as a data object so that it is ready to be supplied to the plotting functions for diagnostics.
+When the algorithms run, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object so that it is ready to be supplied to the plotting functions for diagnostics.
 
@@
 
-Multiple iterators are also  needed to index the data collected at different levels.
-$t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate iteration, which remains the same within one round and has an increment of one once a new round starts.
-$l$ is the counter for each search/interpolation, which provides the information of how many bases the algorithm has searched before finding one to return.
-There are other parameters of interest, which depend on the particular problem content and they are denoted as *$V_{p}$*.
-Two most common examples include $V_1 = \text{method}$, which tags the name of the algorithm used, and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.
+Multiple iterators are also  needed to index the data collected at different levels: $t$ being a unique identifier that prescribes the natural ordering of each observation while $j$ and $l$ being the counter of the outer and inner loop, respectively, in Algorithm \ref{random-search}, \ref{simulated-annealing} and \ref{search-geodesic} above.
+Other parameters of interest recorded include $V_1 = \text{method}$ that tags the name of the optimiser, and $V_2 = \text{alpha}$ that indicates the sampling neighbourhood size.
 A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
 
-<!-- # ```{r iterators, out.width = "100%", out.height = "25%", fig.cap = "\\label{iterators} A sketch of the design of iterators in iterative algorithms."} -->
-<!-- # knitr::include_graphics(here::here("figures", "iterators.png")) -->
-<!-- # ``` -->
-
 \begin{equation}
+\renewcommand\arraystretch{2}  % default value: 1.0
 \left[
@@
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
-Note that there is no output in iteration $J + 1$ since the fact that it is the last iteration means that the optimiser cannot find a better basis and the algorithm terminates.
-In this notation, the final basis found is $A_T$ with the highest index value $I_T$.
+Note that there is no output in iteration $J + 1$ since the optimiser cannot find a better basis in the last iteration and the algorithm terminates.
+The final basis found is $A_T$ with the highest index value $I_T$.
 
-<!-- ```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."} -->
-<!-- knitr::include_graphics(path = here::here("figures/global_obj.png")) -->
-<!-- ``` -->
+The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation to form a row and each variable to form a column.
+With tidy data structure, data wrangling and visualisation can ben significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
 
-The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation forms a row and each variable forms a column.
-With tidy data structure, data wrangling and visualisation have been significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
-
-The construction of diagnostic plots uses the concept of grammar of graphics [@wickham2010layered] in ggplot2.
-In grammar of graphics, plots are not defined by their appearance (i.e.
-boxplot, histogram, scatter plot etc) but by stacked layers.
-In the construction of diagnostic plots, there are multiple elements one may want to emphasize and there is no single plot name that would meet the need.
-The stacked layers concept, on the other hand, allows information to be overlaid on each other and one can build the plot from scratch as long as the variables have been stored in a dataset.
+The construction of diagnostic plots adopts the core concept in ggplot2:  grammar of graphics [@wickham2010layered].
+In grammar of graphics, plots are not produced via calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but via the concept of stacked layers.
+Seeing plots as stacked layers empowers us to composite diagnostic plots with an emphasis on any variable in the data object without the redundancy of creating different commands for the same type of plots that highlights on different variables.
 
@@
 A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum.
-Directly plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes longer for an optimiser to find a better basis towards the end.
-Another option is to use summarisation for each iteration.
-Boxplot is a suitable candidate that can provide five points summary of each iteration.
-Other additional information not presented in the  boxplot can then be added with new layers, for example text information on the number of points can be added at the bottom of each iteration and the position of basis returned by projection pursuit can be highlighted in point.
-Further, an option to switch between displaying points and boxplot geometry is helpful when the number of observation is small in one iteration and this is achieved via a `cutoff` parameter.
+A simple treatment of plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes much longer for an optimiser to find a better basis towards the end.
+Another option is to use summarisation in each iteration.
+Boxplot is a suitable candidate that can provide five points summary of each iteration and additional information can be separately added with new layers, for example, text information on the number of points searched in each iteration is added at the bottom of each iteration, the bases returned for interpolation are highlighted in larger size and linked, and information regarding the last iteration where no basis is returned is turned to grey scale.
+Furthermore, an option to switch from boxplot back to point geometry is helpful when the number of observation is small in one iteration and can be achieved via the `cutoff` argument.
 

### 2021-01-03 paper/paper.tex
add literature on past incident of using vis to understand optim - distill article; changes in vis diag

@@
 otherwise not be aware of and provides a more comprehensive
-understanding of the problem at hand. The work presented in this paper
-brings visualization tools into optimisation problems with an aim to
-better understand the performance of the optimisers in practice.
+understanding of the problem at hand. Task specific tools presented by
+\citet{li2020visualizing} show how visualisation can be used to
+understand the behaviour of neural network on classification models, but
+no general visualisation tool available for diagnosing optimisation
+procedure. The work presented in this paper brings visualization tools
+into optimisation problems with an aim to better understand the
+performance of the optimisers in practice.
 
@@
 feature is particularly useful when local maxima are present. The
-algorithm \ref{simulated_annealing} highlights how simulated annealing
+algorithm \ref{simulated-annealing} highlights how simulated annealing
 differs from random search in the inner loop.
@@
 algorithms run, key information regarding the decision variable,
-objective function and hyper-parameters, needs to be recorded and stored
+objective function and hyper-parameters needs to be recorded and stored
 as a data object so that it is ready to be supplied to the plotting
@@
 Multiple iterators are also needed to index the data collected at
-different levels. \(t\) is a unique identifier that prescribes the
-natural ordering of each observation; \(j\) is the counter for each
-search-and-interpolate iteration, which remains the same within one
-round and has an increment of one once a new round starts. \(l\) is the
-counter for each search/interpolation, which provides the information of
-how many bases the algorithm has searched before finding one to return.
-There are other parameters of interest, which depend on the particular
-problem content and they are denoted as \emph{\(V_{p}\)}. Two most
-common examples include \(V_1 = \text{method}\), which tags the name of
-the algorithm used, and \(V_2 = \text{alpha}\), the neighbourhood
-parameter that controls the size in sampling candidate bases. A matrix
-notation of the data structure is presented in Equation
+different levels: \(t\) being a unique identifier that prescribes the
+natural ordering of each observation while \(j\) and \(l\) being the
+counter of the outer and inner loop, respectively, in Algorithm
+\ref{random-search}, \ref{simulated-annealing} and \ref{search-geodesic}
+above. Other parameters of interest recorded include
+\(V_1 = \text{method}\) that tags the name of the optimiser, and
+\(V_2 = \text{alpha}\) that indicates the sampling neighbourhood size. A
+matrix notation of the data structure is presented in Equation
 \ref{eq:data-structure}.
@@
 \begin{equation}
+\renewcommand\arraystretch{2}  % default value: 1.0
 \left[
@@
 \noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
-no output in iteration \(J + 1\) since the fact that it is the last
-iteration means that the optimiser cannot find a better basis and the
-algorithm terminates. In this notation, the final basis found is \(A_T\)
-with the highest index value \(I_T\).
+no output in iteration \(J + 1\) since the optimiser cannot find a
+better basis in the last iteration and the algorithm terminates. The
+final basis found is \(A_T\) with the highest index value \(I_T\).
 
 The data structure constructed above meets the tidy data principle
-\citep{wickham2014tidy} that requires each observation forms a row and
-each variable forms a column. With tidy data structure, data wrangling
-and visualisation have been significantly simplified by well-developed
+\citep{wickham2014tidy} that requires each observation to form a row and
+each variable to form a column. With tidy data structure, data wrangling
+and visualisation can ben significantly simplified by well-developed
 packages such as \CRANpkg{dplyr} \citep{dplyr} and \CRANpkg{ggplot2}
@@
 
-The construction of diagnostic plots uses the concept of grammar of
-graphics \citep{wickham2010layered} in ggplot2. In grammar of graphics,
-plots are not defined by their appearance (i.e.~boxplot, histogram,
-scatter plot etc) but by stacked layers. In the construction of
-diagnostic plots, there are multiple elements one may want to emphasize
-and there is no single plot name that would meet the need. The stacked
-layers concept, on the other hand, allows information to be overlaid on
-each other and one can build the plot from scratch as long as the
-variables have been stored in a dataset.
+The construction of diagnostic plots adopts the core concept in ggplot2:
+grammar of graphics \citep{wickham2010layered}. In grammar of graphics,
+plots are not produced via calling the commands, named by the appearance
+of the plot, i.e., boxplot and histogram, but via the concept of stacked
+layers. Seeing plots as stacked layers empowers us to composite
+diagnostic plots with an emphasis on any variable in the data object
+without the redundancy of creating different commands for the same type
+of plots that highlights on different variables.
 
@@
 A primary interest of diagnosing an optimiser is to study how it
-progressively finds its optimum. Directly plotting the index value
-across its natural order will cause the graph to be disproportional to
-the iteration since it usually takes longer for an optimiser to find a
-better basis towards the end. Another option is to use summarisation for
-each iteration. Boxplot is a suitable candidate that can provide five
-points summary of each iteration. Other additional information not
-presented in the boxplot can then be added with new layers, for example
-text information on the number of points can be added at the bottom of
-each iteration and the position of basis returned by projection pursuit
-can be highlighted in point. Further, an option to switch between
-displaying points and boxplot geometry is helpful when the number of
-observation is small in one iteration and this is achieved via a
-\texttt{cutoff} parameter.
+progressively finds its optimum. A simple treatment of plotting the
+index value across its natural order will cause the graph to be
+disproportional to the iteration since it usually takes much longer for
+an optimiser to find a better basis towards the end. Another option is
+to use summarisation in each iteration. Boxplot is a suitable candidate
+that can provide five points summary of each iteration and additional
+information can be separately added with new layers, for example, text
+information on the number of points searched in each iteration is added
+at the bottom of each iteration, the bases returned for interpolation
+are highlighted in larger size and linked, and information regarding the
+last iteration where no basis is returned is turned to grey scale.
+Furthermore, an option to switch from boxplot back to point geometry is
+helpful when the number of observation is small in one iteration and can
+be achieved via the \texttt{cutoff} argument.
 

### 2021-01-03 paper/paper.Rmd
update gridlines for explore_trace_interp()

@@
 
-```{r toy-search, fig.width = 5.5, fig.cap="A comparison of \\code{search\\_better} with different values of the parameter \\code{max\\_tries}.
+```{r toy-search, fig.cap="A comparison of \\code{search\\_better} with different values of the parameter \\code{max\\_tries}.
 A six-variable dataset \\code{boa6} is used with holes index on a 2D problem.
@@
 
-```{r toy-interp,fig.width = 5.5, fig.cap = "The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
+```{r toy-interp, fig.cap = "The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
 The color represents the number of iteration.

### 2021-01-03 scripts/examples.R

@@ -2,13 +2,13 @@
 p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>%
   mutate(anno = c("current basis", "interpolated basis", "target basis"))
 
-p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") +
-  geom_point(data = p1_anno) +
-  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) +
-  theme(legend.position = "none")
-
-p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
-  theme(legend.position = "none")
+# p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") +
+#   geom_point(data = p1_anno) +
+#   geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) +
+#   theme(legend.position = "none")
+#
+# p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
+#   theme(legend.position = "none")
 
 #(p1 | p2) & scale_color_botanical(palette = "fern", discrete = TRUE)
 

### 2021-01-03 scripts/vis-diag.R

@@ -19,12 +19,12 @@ p2 <- holes_2d_better %>%
   explore_trace_search() +
   scale_color_botanical(palette = "daisy")
 
-p1 | p2
+(p1 | p2) & ylim(0.76, 0.95)
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
   mutate(group = "Algorithm 1") %>%
-  explore_trace_interp() +
+  explore_trace_interp(accuracy_x = 4) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
 # p2 <- interrupt_no %>%
@@ -34,7 +34,7 @@ p1 <- holes_2d_better_max_tries %>%
 
 p3 <- holes_2d_better_random %>%
   mutate(group = "Algorithm 3") %>%
-  explore_trace_interp() +
+  explore_trace_interp(accuracy_x = 14) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
 # p1 | p2 | p3

### 2021-01-04 paper/paper.Rmd
changes on explore_trace_search/interp() plots and paragraphs

@@
 
-```{r toy-interp, fig.cap = "The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
-The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies."}
+```{r toy-interp, fig.cap = "Trace plots of two optimisation procedures with color representing each iteration.
+The first optimiser progresses monotonically towards the opitmum while the second has experienced some reduction in some iterations."}
 ```
@@
 Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
-Figure \ref{fig:toy-interp} presents the interpolation of three different tour paths each with different curvature.
-The leftmost plot shows an interpolation where the index value increases progressively and monotonically.
-The middle path has an increase-then-decrease pattern in the last two iterations and the rightmost path shows even a decrease in the index value at iteration three.
-The middle situation can be avoid via a construction of the interruption, which will be detailed in section \ref{monotonic} and the rightmost case is a deliberate construction of `search_better_random` where an inferior basis can be accepted with some probability so as to avoid being trapped in a local maximum.
+Figure \ref{fig:toy-interp} presents the interpolation of two different tour paths.
+The interpolation on the left has index value increase monotonically while the right one has the index value decreases in several iteration.
+Compare to the Figure \ref{fig:toy-search}, Figure \ref{fig:toy-interp} provides additional information on how the interpolation gradually reaches the target basis in each iteration: either monotonically or experiencing a turning point, as seen on the right of Figure \ref{fig:toy-interp}.
 

### 2021-01-04 paper/paper.tex
changes on explore_trace_search/interp() plots and paragraphs

@@
 
-\caption[The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index]{The resulting trace plot on the interpolated points has been plotted when using three different algorithms to optimise the index.
-The color represents the number of iteration.
-It can be observed that each algorithm differs in length in the optimisation and the curvature of the improvement for each algorithm also varies.}\label{fig:toy-interp}
+\caption[Trace plots of two optimisation procedures with color representing each iteration]{Trace plots of two optimisation procedures with color representing each iteration.
+The first optimiser progresses monotonically towards the opitmum while the second has experienced some reduction in some iterations.}\label{fig:toy-interp}
 \end{figure}
@@
 projection on these bases will be played by the tour animation. Figure
-\ref{fig:toy-interp} presents the interpolation of three different tour
-paths each with different curvature. The leftmost plot shows an
-interpolation where the index value increases progressively and
-monotonically. The middle path has an increase-then-decrease pattern in
-the last two iterations and the rightmost path shows even a decrease in
-the index value at iteration three. The middle situation can be avoid
-via a construction of the interruption, which will be detailed in
-section \ref{monotonic} and the rightmost case is a deliberate
-construction of \texttt{search\_better\_random} where an inferior basis
-can be accepted with some probability so as to avoid being trapped in a
-local maximum.
+\ref{fig:toy-interp} presents the interpolation of two different tour
+paths. The interpolation on the left has index value increase
+monotonically while the right one has the index value decreases in
+several iteration. Compare to the Figure \ref{fig:toy-search}, Figure
+\ref{fig:toy-interp} provides additional information on how the
+interpolation gradually reaches the target basis in each iteration:
+either monotonically or experiencing a turning point, as seen on the
+right of Figure \ref{fig:toy-interp}.
 

### 2021-01-04 scripts/vis-diag.R

@@ -11,15 +11,15 @@ purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 ## ---- toy-search
 p1 <- holes_2d_better_max_tries %>%
   mutate(max_tries = 500) %>%
-  explore_trace_search() +
+  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
   scale_color_botanical(palette = "daisy")
 
-p2 <- holes_2d_better %>%
+p2 <- holes_2d_better_random %>%
   mutate(max_tries = 25) %>%
-  explore_trace_search() +
+  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
   scale_color_botanical(palette = "daisy")
 
-(p1 | p2) & ylim(0.76, 0.95)
+(p1 | p2)
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
@@ -27,18 +27,12 @@ p1 <- holes_2d_better_max_tries %>%
   explore_trace_interp(accuracy_x = 4) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
-# p2 <- interrupt_no %>%
-#   mutate(group = "Algorithm 2") %>%
-#   explore_trace_interp() +
-#   scale_color_botanical(palette = "fern", discrete = FALSE)
-
-p3 <- holes_2d_better_random %>%
+p2 <- holes_2d_better_random %>%
   mutate(group = "Algorithm 3") %>%
   explore_trace_interp(accuracy_x = 14) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
-# p1 | p2 | p3
-p1 | p3
+p1 | p2
 
 ## ---- toy-pca
 pca <- bind_rows(holes_1d_geo, holes_1d_better) %>%

### 2021-01-04 paper/paper.Rmd
update pca plot

@@
 
-```{r toy-pca, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: \\code{search\\_better} and \\code{search\\_geodesic}.
+```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: \\code{search\\_better} and \\code{search\\_geodesic}.
 The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.

### 2021-01-04 scripts/vis-diag.R

@@ -35,20 +35,11 @@ p2 <- holes_2d_better_random %>%
 p1 | p2
 
 ## ---- toy-pca
-pca <- bind_rows(holes_1d_geo, holes_1d_better) %>%
+bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-                   index = tourr::holes(), raw_data = boa5) %>%
-  compute_pca(group = method)
-pca$aug %>%
-  ggplot(aes(x = PC1, y = PC2)) +
-  geom_point(data = pca$aug %>% filter(info == "randomly_generated"), col = "grey", size = 0.1) +
-  #geom_path(data =  pca$aug %>% filter(info == "interpolation"), aes(col = method), size = 2) +
-  geom_point(data =  pca$aug %>% filter(info != "randomly_generated"), aes(col = method)) +
-  geom_point(data = pca$aug %>% filter(info == "theoretical"), aes(col = method), size = 10) +
-  geom_point(data = get_start(pca$aug), aes(col = method), size = 5) +
-  scale_color_botanical(palette = "cherry") +
-  theme_void() +
-  theme(aspect.ratio = 1, legend.position =  "bottom")
+                   index = tourr::holes(), raw_data = boa5)  %>%
+  explore_space_pca(group = method, rand_size = 10) +
+  scale_color_botanical(palette = "fern")
 
 ## ----toy-pca-animated
 # ani <- pca$aug %>%

### 2021-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
modify as per R journal author instruction

--- 2021-01-04 paper/paper.Rmd
@@
 ---
-title: "Visual diagnostics for constrained optimisation with application to guided tours"
+title: "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours"
 author:
@@
 
-# Optimisation Methods {#optim}
+# Optimisation methods {#optim}
 
@@
 Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
-The most prominent one within the general solvers is `optim()` in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions.
+The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions.
 Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
-Specific solvers for simulated annealing includes `optim(..., method  "SANN")` and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
-For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].
+Specific solvers for simulated annealing includes \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
+For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].
 
@@
 The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation to form a row and each variable to form a column.
-With tidy data structure, data wrangling and visualisation can ben significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
+With tidy data structure, data wrangling and visualisation can be significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
 
@@
 
-- `explore_trace_search()`: produces summary plots, as shown in Figure \ref{fig:toy-search}
-- `explore_trace_interp()`: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
-- `explore_space_pca() `: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+- \code{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
+- \code{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+- \code{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
 Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
-- `explore_space_tour()`: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+- \code{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
 
-- `get_*()` extracts and manipulates certain components from the existing data object.
-  - `get_best()`: extracts the best basis found in the data object
-  - `get_start()`: extracts the starting basis
-  - `get_interp()`: extracts the observations in the interpolation
-  - `get_search_count()`: produces the summary table of the number of observation in each iteration
-  - `get_basis_matrix()`: flattens all the bases into a matrix
+- \code{get\_*()} extracts and manipulates certain components from the existing data object.
+  - \code{get\_best()}: extracts the best basis found in the data object
+  - \code{get\_start()}: extracts the starting basis
+  - \code{get\_interp()}: extracts the observations in the interpolation
+  - \code{get\_search\_count()}: produces the summary table of the number of observation in each iteration
+  - \code{get\_basis\_matrix()}: flattens all the bases into a matrix
 
-- `bind_*()` incorporates additional information outside the tour optimisation into the data object.
-  - `bind_theoretical()`: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
-  - `bind_random()`: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
+- \code{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
+  - \code{bind\_theoretical()}: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
+  - \code{bind\_random()}: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
 `bind_random_matrix()` binds the points to the basis matrix.
@@
 - Color
-  - `botanical_palettes`:  a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
-  - `botanical_pal()`: a color interpolator 
-  - `scale_color_botanical()`: a ggplot construction for using botanical palettes.
+  - \code{botanical\_palettes}: a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+  - \code{botanical\_pal()}: a color interpolator 
+  - \code{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.
 
@@
 
-\clearpage
-
-\bibliography{RJreferences}
+\bibliography{zhang-cook-laa-langrene-menendez}

### 2021-01-04 paper/zhang-cook-laa-langrene-menendez.tex
modify as per R journal author instruction

--- 2021-01-04 paper/paper.tex
@@
 % !TeX root = RJwrapper.tex
-\title{Visual diagnostics for constrained optimisation with application to
-guided tours}
+\title{Visual Diagnostics for Constrained Optimisation with Application to
+Guided Tours}
 \author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
@@
 \hypertarget{optim}{%
-\section{Optimisation Methods}\label{optim}}
+\section{Optimisation methods}\label{optim}}
 
@@
 general purpose as well as task specifics solvers. The most prominent
-one within the general solvers is \texttt{optim()} in the
-\CRANpkg{stats} \citep{stats} package, which provides both
-gradient-based and derivative-free optimisation functions. Another
-general solver specialised in non-linear optimisation is
-\CRANpkg{nloptr} \citep{nloptr}. Specific solvers for simulated
-annealing includes \texttt{optim(...,\ method\ \ "SANN")} and package
-\CRANpkg{GenSA} \citep{gensa} that deals with more complicated objective
-functions. For other task specific solvers, readers are recommended to
-visit the relevant sections in CRAN task review on \emph{optimisation
-and mathematical programming} \citep{crantaskreviewoptim}.
+one within the general solvers is \code{optim()} in the \CRANpkg{stats}
+\citep{stats} package, which provides both gradient-based and
+derivative-free optimisation functions. Another general solver
+specialised in non-linear optimisation is \CRANpkg{nloptr}
+\citep{nloptr}. Specific solvers for simulated annealing includes
+\code{optim(..., method  "SANN")} and package \CRANpkg{GenSA}
+\citep{gensa} that deals with more complicated objective functions. For
+other task specific solvers, readers are recommended to visit the
+relevant sections in CRAN task review on
+\ctv{optimisation and mathematical programming}
+\citep{crantaskreviewoptim}.
 
@@
 each variable to form a column. With tidy data structure, data wrangling
-and visualisation can ben significantly simplified by well-developed
+and visualisation can be significantly simplified by well-developed
 packages such as \CRANpkg{dplyr} \citep{dplyr} and \CRANpkg{ggplot2}
@@
 \item
-  \texttt{explore\_trace\_search()}: produces summary plots, as shown in
+  \code{explore\_trace\_search()}: produces summary plots, as shown in
   Figure \ref{fig:toy-search}
 \item
-  \texttt{explore\_trace\_interp()}: produces trace plots for the
+  \code{explore\_trace\_interp()}: produces trace plots for the
   interpolation points, as shown in Figure \ref{fig:toy-interp}
 \item
-  \texttt{explore\_space\_pca()}: produces plots of projection basis on
+  \code{explore\_space\_pca()}: produces plots of projection basis on
   the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
@@
 \item
-  \texttt{explore\_space\_tour()}: produces animated tour view on the
-  full space of the projection bases, as shown in Figure
-  \ref{fig:toy-tour}.
-\item
-  \texttt{get\_*()} extracts and manipulates certain components from the
+  \code{explore\_space\_tour()}: produces animated tour view on the full
+  space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+\item
+  \code{get\_*()} extracts and manipulates certain components from the
   existing data object.
@@
   \item
-    \texttt{get\_best()}: extracts the best basis found in the data
-    object
-  \item
-    \texttt{get\_start()}: extracts the starting basis
-  \item
-    \texttt{get\_interp()}: extracts the observations in the
-    interpolation
-  \item
-    \texttt{get\_search\_count()}: produces the summary table of the
+    \code{get\_best()}: extracts the best basis found in the data object
+  \item
+    \code{get\_start()}: extracts the starting basis
+  \item
+    \code{get\_interp()}: extracts the observations in the interpolation
+  \item
+    \code{get\_search\_count()}: produces the summary table of the
     number of observation in each iteration
   \item
-    \texttt{get\_basis\_matrix()}: flattens all the bases into a matrix
+    \code{get\_basis\_matrix()}: flattens all the bases into a matrix
   \end{itemize}
 \item
-  \texttt{bind\_*()} incorporates additional information outside the
-  tour optimisation into the data object.
+  \code{bind\_*()} incorporates additional information outside the tour
+  optimisation into the data object.
 
@@
   \item
-    \texttt{bind\_theoretical()}: incorporates the best possible basis
-    to the existing data object with the supply of the index function
-    and original data for producing the index value.
-  \item
-    \texttt{bind\_random()}: generates 1000 points on the high
-    dimensional surface of a sphere and binds it to the existing data
-    object and output as a tibble object.
-    \texttt{bind\_random\_matrix()} binds the points to the basis
-    matrix.
+    \code{bind\_theoretical()}: incorporates the best possible basis to
+    the existing data object with the supply of the index function and
+    original data for producing the index value.
+  \item
+    \code{bind\_random()}: generates 1000 points on the high dimensional
+    surface of a sphere and binds it to the existing data object and
+    output as a tibble object. \texttt{bind\_random\_matrix()} binds the
+    points to the basis matrix.
   \end{itemize}
@@
   \item
-    \texttt{botanical\_palettes}: a collection of color palettes from
+    \code{botanical\_palettes}: a collection of color palettes from
     Australian native plants. Quantitative palettes include daisy,
@@
   \item
-    \texttt{botanical\_pal()}: a color interpolator
-  \item
-    \texttt{scale\_color\_botanical()}: a ggplot construction for using
+    \code{botanical\_pal()}: a color interpolator
+  \item
+    \code{scale\_color\_botanical()}: a ggplot construction for using
     botanical palettes.
@@
 
-\clearpage
-
-\bibliography{RJreferences}
+\bibliography{zhang-cook-laa-langrene-menendez}
 

### 2021-01-14 paper/zhang-cook-laa-langrene-menendez.Rmd
writing for explore_trace_search()

@@
 
-Figure \ref{fig:toy-search} shows a sample of the plot constructed using `search_better` with different values of the parameter `max_tries`.
-Comparing the returned index value of each iteration shows that the termination at `max_tries = 25` is not sufficient for `search_better` to explore the parameter space and a value of 500 is preferred over 25 in this context.
+Figure \ref{fig:toy-search} shows the searches of two different optimisers: `search_better` and `search_better_random`.
+Both optimisers progress quickly at the first few iterations, take longer to find better basis in the later iterations, and finish off when hitting the maximum number of try, 499 and 999, respectively.
+Also, the target basis found in each iteration of `search_better` always has an increased index value while this is not the case for `search_better_random`.
+This explains why in this example, `search_better_random` takes longer to find the final basis, but in more complicated scenarios, the feature of probabilistic acceptance allows a more holistic search of the basis space and is more likely to find the global optimum.
 
-```{r toy-search, fig.cap="A comparison of \\code{search\\_better} with different values of the parameter \\code{max\\_tries}.
-A six-variable dataset \\code{boa6} is used with holes index on a 2D problem.
-\\code{max\\_tries = 500} allows the optimiser to find better basis with higher index value at iteration six."}
+```{r toy-search, fig.cap="A comparison of search by two otimisers: \\code{search\\_better} and \\code{search\\_better\\_random} on a six-variable dataset \\code{boa6} with holes index on a 2D problem.
+Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
+\\code{search\\_better\\_random} takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem."}
 ```

### 2021-01-14 paper/zhang-cook-laa-langrene-menendez.tex
writing for explore_trace_search()

@@
 
-Figure \ref{fig:toy-search} shows a sample of the plot constructed using
-\texttt{search\_better} with different values of the parameter
-\texttt{max\_tries}. Comparing the returned index value of each
-iteration shows that the termination at \texttt{max\_tries\ =\ 25} is
-not sufficient for \texttt{search\_better} to explore the parameter
-space and a value of 500 is preferred over 25 in this context.
+Figure \ref{fig:toy-search} shows the searches of two different
+optimisers: \texttt{search\_better} and \texttt{search\_better\_random}.
+Both optimisers progress quickly at the first few iterations, take
+longer to find better basis in the later iterations, and finish off when
+hitting the maximum number of try, 499 and 999, respectively. Also, the
+target basis found in each iteration of \texttt{search\_better} always
+has an increased index value while this is not the case for
+\texttt{search\_better\_random}. This explains why in this example,
+\texttt{search\_better\_random} takes longer to find the final basis,
+but in more complicated scenarios, the feature of probabilistic
+acceptance allows a more holistic search of the basis space and is more
+likely to find the global optimum.
 
@@
 
-\caption{A comparison of \code{search\_better} with different values of the parameter \code{max\_tries}.
-A six-variable dataset \code{boa6} is used with holes index on a 2D problem.
-\code{max\_tries = 500} allows the optimiser to find better basis with higher index value at iteration six.}\label{fig:toy-search}
+\caption{A comparison of search by two otimisers: \code{search\_better} and \code{search\_better\_random} on a six-variable dataset \code{boa6} with holes index on a 2D problem.
+Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
+\code{search\_better\_random} takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem.}\label{fig:toy-search}
 \end{figure}

### 2021-01-14 scripts/vis-diag.R

@@ -38,7 +38,7 @@ p1 | p2
 bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
-  explore_space_pca(group = method, rand_size = 10) +
+  explore_space_pca(group = method, interp_size = 1) +
   scale_color_botanical(palette = "fern")
 
 ## ----toy-pca-animated

### 2021-01-14 paper/zhang-cook-laa-langrene-menendez.Rmd
writing on explore_trace_interp()

@@
 
-```{r toy-interp, fig.cap = "Trace plots of two optimisation procedures with color representing each iteration.
-The first optimiser progresses monotonically towards the opitmum while the second has experienced some reduction in some iterations."}
+```{r toy-interp, fig.cap = "Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
+The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers."}
 ```
 
-Points on the interpolation path are another interest in tour since the projection on these bases will be played by the tour animation.
-Figure \ref{fig:toy-interp} presents the interpolation of two different tour paths.
-The interpolation on the left has index value increase monotonically while the right one has the index value decreases in several iteration.
-Compare to the Figure \ref{fig:toy-search}, Figure \ref{fig:toy-interp} provides additional information on how the interpolation gradually reaches the target basis in each iteration: either monotonically or experiencing a turning point, as seen on the right of Figure \ref{fig:toy-interp}.
+Viewing the index value of points on the interpolation path is another interest to the analysts since the projection on these bases will be played by the tour animation and it provides further information on how the index value changes when moving from one target basis to another.
+Figure \ref{fig:toy-interp} presents the index value against time on the interpolation path for the two optimisers with same configuration as \ref{fig:toy-search}.
+From Figure \ref{fig:toy-interp}, we further know that to reach a target basis with lower index value, the optimiser plotted on the right  first passes some bases with higher index value during the interpolation.
+If the bases on the interpolation path can be incorporated into the optimisation, we may be able to find the optimal basis quicker.
 

### 2021-01-14 paper/zhang-cook-laa-langrene-menendez.tex
writing on explore_trace_interp()

@@
 
-\caption[Trace plots of two optimisation procedures with color representing each iteration]{Trace plots of two optimisation procedures with color representing each iteration.
-The first optimiser progresses monotonically towards the opitmum while the second has experienced some reduction in some iterations.}\label{fig:toy-interp}
+\caption[Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure]{Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
+The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers.}\label{fig:toy-interp}
 \end{figure}
@@
 
-Points on the interpolation path are another interest in tour since the
-projection on these bases will be played by the tour animation. Figure
-\ref{fig:toy-interp} presents the interpolation of two different tour
-paths. The interpolation on the left has index value increase
-monotonically while the right one has the index value decreases in
-several iteration. Compare to the Figure \ref{fig:toy-search}, Figure
-\ref{fig:toy-interp} provides additional information on how the
-interpolation gradually reaches the target basis in each iteration:
-either monotonically or experiencing a turning point, as seen on the
-right of Figure \ref{fig:toy-interp}.
+Viewing the index value of points on the interpolation path is another
+interest to the analysts since the projection on these bases will be
+played by the tour animation and it provides further information on how
+the index value changes when moving from one target basis to another.
+Figure \ref{fig:toy-interp} presents the index value against time on the
+interpolation path for the two optimisers with same configuration as
+\ref{fig:toy-search}. From Figure \ref{fig:toy-interp}, we further know
+that to reach a target basis with lower index value, the optimiser
+plotted on the right first passes some bases with higher index value
+during the interpolation. If the bases on the interpolation path can be
+incorporated into the optimisation, we may be able to find the optimal
+basis quicker.
 

### 2021-01-15 paper/zhang-cook-laa-langrene-menendez.Rmd
update pca plot and writing on the plot design

@@
 
-```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: \\code{search\\_better} and \\code{search\\_geodesic}.
-The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
-The underlying grey points are randomly generated on the 5D space and reduced to 2D via PCA along with all the search points presented.
-The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in \\code{search\\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot."}
+```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: pseudo derivative and simulated annealing.
+All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
 ```
 
-Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ live on the surface of a $p \times d$ dimension sphere, where the dimension can easily go over two or three.
-While visualising the search paths on the original high dimensional sphere would require skills for the viewers to perceive rotation of geometry in higher dimensional space ($d > 3$), an easier alternative is to view the reduced space via some dimension reduction methods, i.e.
-principal component analysis.
-To better perceive the search path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using the package \CRANpkg{geozoo} [@geozoo] and PCA is conducted on both the bases and the points on the surface of the sphere.
+Apart from checking the progression of an optimiser, another interesting aspect of the diagnostics is to visualise how the visited bases looks like in its parameter space.
+Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and even with 1D projection, the basis space is high-dimensional.
+Dimension reduction methods, i.e.
+principal component analysis can be used to present the bases on 2D space with appropriate annotation to highlight the different key components of the optimisation.
 
-Figure \ref{fig:toy-pca} plots the first two principal components of two search paths, one using `search_better` and the other using `search_geodesic`.
-The search in the space reduced by PCA matches with the optimiser description before where the random sampling in `search_better` is broader, controlled by `alpha` parameter, which is default to 0.5 while the directional search in `search_geodesic`, controlled by `delta` with a default of 0.01, is so tiny that it can barely be seen.
+In a projection pursuit guided tour optimisation, there are 7 different components that are worth-noticing in the exploration: 1) The spherical space of all the basses; 2) The starting basis of an optimisation routine; 3) The anchor points where a new target basis is found; 4) The search basis that an optimiser has evaluated; 5) The interpolation path that constructed by the guided tour; 6) The annotation of the interrupted path and 7) The theoretical best basis, if applicable.
+Despite all have to be displayed on the diagnostic plot, these components differ in their importance with the most interesting ones being the starting point, anchor points and the interpolation path.
+
+Various other components can be down-played via size, transparency, linetype and proper approximation.
+The basis space is generated via random bases on the original high dimensional space via the CRAN package \CRANpkg{geozoo} [@geozoo] and projected down to 2D.
+Displaying the search space with hundreds of randomly generated dots can be dizzy, especially when the more important start, anchor and search bases are also displayed with point geometry.
+Given the fact that the basis space is a circle when projected to 2D, we can estimate the center and radius with all the bases recorded in the data object.
+Comparing to the start and anchor points, search points are less important as they fail to improve on the index value during the optimisation and hence an adjusted size and transparency are applied to them.
+A potential scenario that could happen is that two optimisers can start at the same basis but finishes with bases in opposite signs.
+This sign difference would not make a difference in the projected data but the PCA projection of the bases will display the finish points symmetric to each other, which can be disturbing for comparing the path close to the finish.
+One solution is to flip the sign of one optimiser so that they start at symmetric position but finish close to each other.
+To remind the analyst of this symmetry in the PCA space, a light dashed line is drawn to connect the starting points of two paths.
+The theoretical best basis is available if the data is simulated and can serve as a guide on how the optimisers progress to the final basis.
+Also, the search paths also have an increasing alpha hue from start to finish to indicate the direction of the optimsation when the theoretical guide is not available.
+To allow flexible annotation in different scenarios, a range of arguments are made available to adjust in the main plotting function \code{explore\_space\_pca()}.
+
+Figure \ref{fig:toy-pca} shows an example of the PCA plot on the same optimisers on a 1D projection problem with 5 variables.
 

### 2021-01-15 paper/zhang-cook-laa-langrene-menendez.tex
update pca plot and writing on the plot design

@@
 
-\caption{1D projection on the 5-variable dataset \code{boa5} with two optimisers: \code{search\_better} and \code{search\_geodesic}.
-The yellow point corresponds to the theoretical best basis [0, 1, 0, 0, 0] with `V2` being the only non-normal variable in the dataset.
-The underlying grey points are randomly generated on the 5D space and reduced to 2D via PCA along with all the search points presented.
-The enlarged color points that colored as the interpolation are the starting points of each algorithm.
-They are initially simulated with the same starting points but all the bases in \code{search\_geodesic} has been flipped positive to ensure that bases with the same projected images but a sign difference are represented by the same point in the plot.}\label{fig:toy-pca}
+\caption[1D projection on the 5-variable dataset \code{boa5} with two optimisers]{1D projection on the 5-variable dataset \code{boa5} with two optimisers: pseudo derivative and simulated annealing.
+All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
 \end{figure}
@@
 Apart from checking the progression of an optimiser, another interesting
-aspect is to visualise how the search looks like in its parameter space.
-Given the orthonormality constraint, the projection bases
-\(\mathbf{A}_{p \times d}\) live on the surface of a \(p \times d\)
-dimension sphere, where the dimension can easily go over two or three.
-While visualising the search paths on the original high dimensional
-sphere would require skills for the viewers to perceive rotation of
-geometry in higher dimensional space (\(d > 3\)), an easier alternative
-is to view the reduced space via some dimension reduction methods,
-i.e.~principal component analysis. To better perceive the search path as
-an embedding of a hollow sphere, random points on the high dimensional
-sphere is generated using the package \CRANpkg{geozoo} \citep{geozoo}
-and PCA is conducted on both the bases and the points on the surface of
-the sphere.
-
-Figure \ref{fig:toy-pca} plots the first two principal components of two
-search paths, one using \texttt{search\_better} and the other using
-\texttt{search\_geodesic}. The search in the space reduced by PCA
-matches with the optimiser description before where the random sampling
-in \texttt{search\_better} is broader, controlled by \texttt{alpha}
-parameter, which is default to 0.5 while the directional search in
-\texttt{search\_geodesic}, controlled by \texttt{delta} with a default
-of 0.01, is so tiny that it can barely be seen.
+aspect of the diagnostics is to visualise how the visited bases looks
+like in its parameter space. Given the orthonormality constraint, the
+space of projection bases \(\mathbf{A}_{p \times d}\) is a
+\(p \times d\) dimension sphere and even with 1D projection, the basis
+space is high-dimensional. Dimension reduction methods, i.e.~principal
+component analysis can be used to present the bases on 2D space with
+appropriate annotation to highlight the different key components of the
+optimisation.
+
+In a projection pursuit guided tour optimisation, there are 7 different
+components that are worth-noticing in the exploration: 1) The spherical
+space of all the basses; 2) The starting basis of an optimisation
+routine; 3) The anchor points where a new target basis is found; 4) The
+search basis that an optimiser has evaluated; 5) The interpolation path
+that constructed by the guided tour; 6) The annotation of the
+interrupted path and 7) The theoretical best basis, if applicable.
+Despite all have to be displayed on the diagnostic plot, these
+components differ in their importance with the most interesting ones
+being the starting point, anchor points and the interpolation path.
+
+Various other components can be down-played via size, transparency,
+linetype and proper approximation. The basis space is generated via
+random bases on the original high dimensional space via the CRAN package
+\CRANpkg{geozoo} \citep{geozoo} and projected down to 2D. Displaying the
+search space with hundreds of randomly generated dots can be dizzy,
+especially when the more important start, anchor and search bases are
+also displayed with point geometry. Given the fact that the basis space
+is a circle when projected to 2D, we can estimate the center and radius
+with all the bases recorded in the data object. Comparing to the start
+and anchor points, search points are less important as they fail to
+improve on the index value during the optimisation and hence an adjusted
+size and transparency are applied to them. A potential scenario that
+could happen is that two optimisers can start at the same basis but
+finishes with bases in opposite signs. This sign difference would not
+make a difference in the projected data but the PCA projection of the
+bases will display the finish points symmetric to each other, which can
+be disturbing for comparing the path close to the finish. One solution
+is to flip the sign of one optimiser so that they start at symmetric
+position but finish close to each other. To remind the analyst of this
+symmetry in the PCA space, a light dashed line is drawn to connect the
+starting points of two paths. The theoretical best basis is available if
+the data is simulated and can serve as a guide on how the optimisers
+progress to the final basis. Also, the search paths also have an
+increasing alpha hue from start to finish to indicate the direction of
+the optimsation when the theoretical guide is not available. To allow
+flexible annotation in different scenarios, a range of arguments are
+made available to adjust in the main plotting function
+\code{explore\_space\_pca()}.
+
+Figure \ref{fig:toy-pca} shows an example of the PCA plot on the same
+optimisers on a 1D projection problem with 5 variables.
 

### 2021-01-15 scripts/vis-diag.R

@@ -38,7 +38,7 @@ p1 | p2
 bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
-  explore_space_pca(group = method, interp_size = 1) +
+  explore_space_pca(group = method, interp_size = 1, anchor_size = 2, finish_size = 2) +
   scale_color_botanical(palette = "fern")
 
 ## ----toy-pca-animated

### 2021-01-15 paper/zhang-cook-laa-langrene-menendez.Rmd
list functions in ferrn in  implementation section

@@
 
-- \code{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
-- \code{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
-- \code{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+- Main plotting functions:
+
+  - \code{explore\_trace\_search()} produces summary plots, as shown in Figure \ref{fig:toy-search}
+  - \code{explore\_trace\_interp()}produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
+  - \code{explore\_space\_pca()} produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
 Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
-- \code{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+  - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
 
 - \code{get\_*()} extracts and manipulates certain components from the existing data object.
-  - \code{get\_best()}: extracts the best basis found in the data object
-  - \code{get\_start()}: extracts the starting basis
-  - \code{get\_interp()}: extracts the observations in the interpolation
-  - \code{get\_search\_count()}: produces the summary table of the number of observation in each iteration
+  - \code{get\_best()} extracts the best basis found in the data object
+  - \code{get\_start()} extracts the starting basis
+  - \code{get\_interp()} extracts the observations in the interpolation
+  - \code{get\_interp\_last()} extracts the end observations of the interpolation in each iteration
+  - \code{get\_anchor()} extracts the target observations found by the optimiser
+  - \code{get\_search()} extracts the search observations evaluated by the optimiser
+  - \code{get\_search\_count()} produces the summary table of the number of observation in each iteration
+  - \code{get\_center()} produces the center point of the basis space estimated by the starting points
+  - \code{get\_space\_param()} produces the coordinates of the center and radius of the basis space
+  - \code{get\_theo()} extracts the theoretical observations from the data object
+  - \code{get\_interrupt()} extract the end point of the interpolation and the target point when an interruption happens
   - \code{get\_basis\_matrix()}: flattens all the bases into a matrix
@@
 - \code{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
-  - \code{bind\_theoretical()}: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
-  - \code{bind\_random()}: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object.
-`bind_random_matrix()` binds the points to the basis matrix.
+  - \code{bind\_theoretical()} incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
+  - \code{bind\_random()} generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object. 
+  - \code{bind\_random\_matrix()} binds the points to the basis matrix.
 
-- Color
-  - \code{botanical\_palettes}: a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
-  - \code{botanical\_pal()}: a color interpolator 
-  - \code{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.
+- Utilities
+  - \code{add\_*()} are internal wrapper functions that facilitate the composition of PCA plot
+  - \code{theme\_fern()} and \code{format\_label()} for better display of the grid lines and axis formatting
+  - \code{clean\_method()} for clean up the name of the optimisers
+  - \code{botanical\_palettes} is a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+  - \code{botanical\_pal()} as the color interpolator 
+  - \code{scale\_color\_botanical()} is a ggplot2 scale for botanical palettes.
 

### 2021-01-15 paper/zhang-cook-laa-langrene-menendez.tex
list functions in ferrn in  implementation section

@@
 \item
-  \code{explore\_trace\_search()}: produces summary plots, as shown in
+  Main plotting functions:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \code{explore\_trace\_search()} produces summary plots, as shown in
     Figure \ref{fig:toy-search}
   \item
-  \code{explore\_trace\_interp()}: produces trace plots for the
+    \code{explore\_trace\_interp()}produces trace plots for the
     interpolation points, as shown in Figure \ref{fig:toy-interp}
   \item
-  \code{explore\_space\_pca()}: produces plots of projection basis on
+    \code{explore\_space\_pca()} produces plots of projection basis on
     the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
-  Animated version in Figure \ref{fig:toy-pca-animated} can be turned on
-  via the argument \texttt{animate\ =\ TRUE}
-\item
-  \code{explore\_space\_tour()}: produces animated tour view on the full
-  space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+    Animated version in Figure \ref{fig:toy-pca-animated} can be turned
+    on via the argument \texttt{animate\ =\ TRUE}
+  \item
+    \code{explore\_space\_tour()} produces animated tour view on the
+    full space of the projection bases, as shown in Figure
+    \ref{fig:toy-tour}.
+  \end{itemize}
 \item
@@
   \item
-    \code{get\_best()}: extracts the best basis found in the data object
-  \item
-    \code{get\_start()}: extracts the starting basis
-  \item
-    \code{get\_interp()}: extracts the observations in the interpolation
-  \item
-    \code{get\_search\_count()}: produces the summary table of the
-    number of observation in each iteration
+    \code{get\_best()} extracts the best basis found in the data object
+  \item
+    \code{get\_start()} extracts the starting basis
+  \item
+    \code{get\_interp()} extracts the observations in the interpolation
+  \item
+    \code{get\_interp\_last()} extracts the end observations of the
+    interpolation in each iteration
+  \item
+    \code{get\_anchor()} extracts the target observations found by the
+    optimiser
+  \item
+    \code{get\_search()} extracts the search observations evaluated by
+    the optimiser
+  \item
+    \code{get\_search\_count()} produces the summary table of the number
+    of observation in each iteration
+  \item
+    \code{get\_center()} produces the center point of the basis space
+    estimated by the starting points
+  \item
+    \code{get\_space\_param()} produces the coordinates of the center
+    and radius of the basis space
+  \item
+    \code{get\_theo()} extracts the theoretical observations from the
+    data object
+  \item
+    \code{get\_interrupt()} extract the end point of the interpolation
+    and the target point when an interruption happens
   \item
@@
   \item
-    \code{bind\_theoretical()}: incorporates the best possible basis to
+    \code{bind\_theoretical()} incorporates the best possible basis to
     the existing data object with the supply of the index function and
@@
   \item
-    \code{bind\_random()}: generates 1000 points on the high dimensional
+    \code{bind\_random()} generates 1000 points on the high dimensional
     surface of a sphere and binds it to the existing data object and
-    output as a tibble object. \texttt{bind\_random\_matrix()} binds the
-    points to the basis matrix.
+    output as a tibble object.
+  \item
+    \code{bind\_random\_matrix()} binds the points to the basis matrix.
   \end{itemize}
 \item
-  Color
+  Utilities
 
@@
   \item
-    \code{botanical\_palettes}: a collection of color palettes from
+    \code{add\_*()} are internal wrapper functions that facilitate the
+    composition of PCA plot
+  \item
+    \code{theme\_fern()} and \code{format\_label()} for better display
+    of the grid lines and axis formatting
+  \item
+    \code{clean\_method()} for clean up the name of the optimisers
+  \item
+    \code{botanical\_palettes} is a collection of color palettes from
     Australian native plants. Quantitative palettes include daisy,
@@
   \item
-    \code{botanical\_pal()}: a color interpolator
-  \item
-    \code{scale\_color\_botanical()}: a ggplot construction for using
-    botanical palettes.
+    \code{botanical\_pal()} as the color interpolator
+  \item
+    \code{scale\_color\_botanical()} is a ggplot2 scale for botanical
+    palettes.
   \end{itemize}

### 2021-01-21 paper/zhang-cook-laa-langrene-menendez.Rmd
more work on noisy index simulation, WIP

@@
 
-```{r kol-better, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better} with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced to 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
-The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point."}
-```
-
-```{r kol-random, fig.height = 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with the same seeds.
-\\code{search\\_better\\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum."}
-```
-
-```{r kol-random-tuned, fig.height= 6, fig.cap = "One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better\\_random} with an larger searching neighbourhood of 0.7.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
-"}
+```{r kol-result, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better}, starting from 20 different seeds. "}
 ```
@@
 Hence, the global maximum happens when variable `x7` is found.
-Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
+
+<!-- Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
 Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.
 The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
-
-In Figure \ref{fig:kol-better}, the starting point largely affects whether `search_better` will find the global maximum.
-The three seeds on row three: seed 9145, seed 2511, and seed 9209, and three on row four: seed 2888, seed 9334, and seed 9819 easily find `V7` because they are born in Rome.
-On the other end, the ones on the first two rows can only find `V2`, largely due to the fact that they do not even have chances to search near `V7`.
-
-A comparison between `search_better` and `search_better_random` shows all four combinations where both, neither, or one of the two algorithms find the global maximum.
-Two iconic cases that shows how `search_better_random` improves `search_better` are seeds 4761 (position (1,4) in Figure \ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed 1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure \ref{fig:kol-random}).
-In both cases, `search_better_random` accepts inferior points at the initial iteration and this acceptance has later been proved to change the direction of the search and hence allows the optimiser to explore near the global optimum and in the end, find the global optimum.
-However, this change of search direction may sometimes have an adverse effect on `search_better_random` and change the search to the neighbourhood of `V2`.
-This happens on seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure \ref{fig:kol-random}).
-The two cases that neither algorithm finds the global optimum happens on seed 6170 and seed 2757 (position (1,1) and (1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in Figure \ref{fig:kol-random}).
-This is largely due to the fact that the searching neighbourhood is not large enough to allow the space near `V7` to be adequately explored.
-
-In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter `alpha` is increased from a 0.5 default to 0.7 to solve the insufficiency of search neighbourhood.
-Remember that a candidate basis is generated via a linear combination of the current basis and a randomly generated basis on the surface of the sphere, an increase of `alpha` gives more weights on the randomly generated bases and hence allows a wider search.
-Seed 9982 ((2,4) in Figure \ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure \ref{fig:kol-random-tuned}) benefit from this and find the global maximum.
-A noticeable feature of this enlarged search space is that the interpolation paths start to jump around the space but its usefulness is arguable.
-In seed 2888 (position (3,1)), seed 2986 (position (2,3)), seed 9209 (position (4,4)), just to name a few, this allows a near `V2` bases to switch to a near `V7` bases and as a result these cases find the global maximum.
-While in the case of seed 6746 (position (1,2)) and seed 2757 (position (1,5)), the jump happens from near `V7` basis to near `V2` basis, it does not come back and leaves those simulations in the local maximum.
+-->
 

### 2021-01-21 paper/zhang-cook-laa-langrene-menendez.tex
more work on noisy index simulation, WIP

@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-better-1}
-
-}
-
-\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better} with 20 randomly generated seeds.
-Each point represents a basis in the original 6D space, reduced to 2D by PCA.
-The grey points are random bases generated from 6-D hollow sphere and the two yellow points represent the local maximum corresponds to basis when `V2` is found and the global maximum where `V7` is found.
-The points produced by the search algorithm is colored by whether the global or local maximum is found with the interpolated bases highlighted as a path and the final basis as an amplified point.}\label{fig:kol-better}
-\end{figure}
-\end{Schunk}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-1}
-
-}
-
-\caption{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with the same seeds.
-\code{search\_better\_random} has a probabilistic acceptance implementation that would also accept a basis with lower index value.
-This design allows the optimiser to jump out of the local maximum and hence more instances find the global maximum.}\label{fig:kol-random}
-\end{figure}
-\end{Schunk}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-random-tuned-1}
-
-}
-
-\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better\_random} with an larger searching neighbourhood of 0.7.
-Further tuning of the parameter has been taken place to adjust the neighbourhood parameter `alpha` from 0.5 to 0.7.
-}\label{fig:kol-random-tuned}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+
+}
+
+\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds.
+}\label{fig:kol-result}
 \end{figure}
@@
 0{]} with index value 0.176 and 0.235, respectively. Hence, the global
-maximum happens when variable \texttt{x7} is found. Simulation is done
-with 20 randomly generated seeds on two optimisers:
-\texttt{search\_better} (Figure \ref{fig:kol-better}) and
-\texttt{search\_better\_random} (Figure \ref{fig:kol-random} and Figure
-\ref{fig:kol-random-tuned}). Compared to Figure \ref{fig:kol-better} and
-\ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases
-the neighbourhood parameter \texttt{alpha} from 0.5 to 0.7 to enlarge
-the search space. The data object is collected for each simulation and
-computation is done to reduce the bases from its original 6D to a 2D
-points by principal component analysis so as to view the search paths.
-
-In Figure \ref{fig:kol-better}, the starting point largely affects
-whether \texttt{search\_better} will find the global maximum. The three
-seeds on row three: seed 9145, seed 2511, and seed 9209, and three on
-row four: seed 2888, seed 9334, and seed 9819 easily find \texttt{V7}
-because they are born in Rome. On the other end, the ones on the first
-two rows can only find \texttt{V2}, largely due to the fact that they do
-not even have chances to search near \texttt{V7}.
-
-A comparison between \texttt{search\_better} and
-\texttt{search\_better\_random} shows all four combinations where both,
-neither, or one of the two algorithms find the global maximum. Two
-iconic cases that shows how \texttt{search\_better\_random} improves
-\texttt{search\_better} are seeds 4761 (position (1,4) in Figure
-\ref{fig:kol-better} and (2,2) in Figure \ref{fig:kol-random}) and seed
-1842 ((2,2) in Figure \ref{fig:kol-better} and (4,5) in Figure
-\ref{fig:kol-random}). In both cases, \texttt{search\_better\_random}
-accepts inferior points at the initial iteration and this acceptance has
-later been proved to change the direction of the search and hence allows
-the optimiser to explore near the global optimum and in the end, find
-the global optimum. However, this change of search direction may
-sometimes have an adverse effect on \texttt{search\_better\_random} and
-change the search to the neighbourhood of \texttt{V2}. This happens on
-seed 9209 and seed 9982 (position (1,1) and (1,2) in Figure
-\ref{fig:kol-random}). The two cases that neither algorithm finds the
-global optimum happens on seed 6170 and seed 2757 (position (1,1) and
-(1,2) in Figure \ref{fig:kol-better} and position (1,5) and (1,3) in
-Figure \ref{fig:kol-random}). This is largely due to the fact that the
-searching neighbourhood is not large enough to allow the space near
-\texttt{V7} to be adequately explored.
-
-In Figure \ref{fig:kol-random-tuned}, the neighbourhood parameter
-\texttt{alpha} is increased from a 0.5 default to 0.7 to solve the
-insufficiency of search neighbourhood. Remember that a candidate basis
-is generated via a linear combination of the current basis and a
-randomly generated basis on the surface of the sphere, an increase of
-\texttt{alpha} gives more weights on the randomly generated bases and
-hence allows a wider search. Seed 9982 ((2,4) in Figure
-\ref{fig:kol-random-tuned}) and seed 2463 ((3,2) in Figure
-\ref{fig:kol-random-tuned}) benefit from this and find the global
-maximum. A noticeable feature of this enlarged search space is that the
-interpolation paths start to jump around the space but its usefulness is
-arguable. In seed 2888 (position (3,1)), seed 2986 (position (2,3)),
-seed 9209 (position (4,4)), just to name a few, this allows a near
-\texttt{V2} bases to switch to a near \texttt{V7} bases and as a result
-these cases find the global maximum. While in the case of seed 6746
-(position (1,2)) and seed 2757 (position (1,5)), the jump happens from
-near \texttt{V7} basis to near \texttt{V2} basis, it does not come back
-and leaves those simulations in the local maximum.
+maximum happens when variable \texttt{x7} is found.
 

### 2021-01-21 scripts/kol_result.R

@@ -1,52 +1,25 @@
-## ----kol-better
-bind_boa6_theo <- function(dt){
-  dt %>%
-    bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                     index = tourr::norm_kol(nrow(boa6)),
-                     raw_data = boa6) %>%
-    bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                     index = tourr::norm_kol(nrow(boa6)),
-                     raw_data = boa6)
-}
-
-
+## ----kol-result
 clean %>%
-  filter(sim == "kol_better",method != "search_polish") %>%
-  bind_boa6_theo() %>%
-  explore_space_pca(group = seed, color = var_found,
+  filter(method != "search_polish") %>%
+  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
+                   index = tourr::norm_kol(nrow(boa6)),
+                   raw_data = boa6) %>%
+  bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
+                   index = tourr::norm_kol(nrow(boa6)),
+                   raw_data = boa6) %>%
+  mutate(seed_sim = paste0(seed, sim),
+         alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
+         search_name = ifelse(str_detect(sim, "random"), "search_better_random", "search_better")) %>%
+  compute_pca(group = seed_sim) %>%
+  pluck("aug") %>%
+  explore_space_pca(details = FALSE, pca = FALSE, group = seed_sim, color = var_found,
                     precision = 0.001, theo_size = 15,
-                    start_size = 7) +
+                    start_size = 1, end_size = 3,
+                    interp_size = 0.5) +
   scale_color_botanical() +
-  facet_wrap(vars(fct_reorder(seed, index_val_best)))
+  facet_grid(alpha ~ search_name, labeller = label_both)
 
-# kol_better_tuned
-kol_better_tuned <- clean %>%
-  filter(sim == "kol_better_tuned", method != "search_polish") %>%
-  bind_boa6_theo() %>%
-  explore_space_pca(group = seed, color = var_found,
-                    precision = 0.001, theo_size = 15,
-                    start_size = 7) +
-  scale_color_botanical() +
-  facet_wrap(vars(fct_reorder(seed, index_val_best)))
 
-## ----kol-random
[…]

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.Rmd
update the two plots with noisy index and some writing

@@
 
-The index function, up until this point, are all smooth, while this is not the case for all the index functions.
-`norm_kol`, a 1D projection function based on the Kolmogorov test, compares the difference between the 1D projected data, $\mathbf{Y}_{n \times 1}$ and a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Denote the ECDF function as $F_{.}(u)$ with the subscript indicating either the projection or the random normal variable,  the `norm_kol` index is defined by
+The index function, up until this point, are all smooth, while this is not the case for all.
+A 1D projection function based on the Kolmogorov test, `norm_kol`, compares the projected data, $\mathbf{Y}_{n \times 1}$ to a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
+Let $F_{.}(u)$ be the ECDF function with the subscript $P$ and $u$ indicating the projected and randomly generated ECDF respectively,  the index is defined as:
 
 $$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
+With a more complex index function, it is interesting to understand 1) if all the three optimisers can reach the optimum, 2) when a local optimum is presented, whether the simulated annealing algorithm can escape that local optimum and find the global one.
 
-Figure \ref{fig:noisy-better-geo} compares the tracing plot of two optimisers: `search_geodesic` and `search_better`.
-This time, the interpolated path is no longer smooth when using either algorithm and  `search_geodesic` fails to optimise this index with barely any improvement of the index value.
-On the other hand, `search_better` is doing relatively well.
-With the theoretical best basis [`r matrix(c(0, 1, 0, 0, 0))`] producing an index value of `r kol_1d_better %>% bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1), index = tourr::norm_kol(nrow(boa5)), raw_data = boa5) %>% tail(1) %>% pull(index_val)`, `search_better` finds the final basis [`r get_best(kol_1d_better)$basis %>% .[[1]] %>% format(digits = 3)`] with an index value of `r get_best(kol_1d_better)$index_val %>% format(digits = 4)`.
-A further polish step will give a marginal improvement of index value to `r get_best(kol_1d_better_polish)$index_val` with a basis of [`r get_best(kol_1d_better_polish)$basis %>% .[[1]] %>% format(digits = 3)`].
-At this stage, the difference between the theoretical best and what has been found is likely due to simulation error since the best possible basis for a simulated data will be slightly off the theoretical best basis, which is derived based on the distributional assumptions in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+```{r noisy-better-geo, fig.env = "widefigure", fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
+The trace plot shows the failure of pseudo derivative on non-smooth index.
+The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index."}
+```
+
+Figure \ref{fig:noisy-better-geo} presents the tracing plots of two optimisers: pseudo derivative and simulated annealing and as expected, the interpolated path is no longer smooth in either case.
+There is barely any improvement on the index value in the pseudo derivative algorithm while the simulated annealing algorithm has managed to get close to the index value of the theoretical best, indicated by the horizontal dashed line.
+In the PCA plot, simulated annealing is able to move gradually towards the theoretical best while the pseudo derivative method breaks after the initial few attempts.
 

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.tex
update the two plots with noisy index and some writing

@@
 The index function, up until this point, are all smooth, while this is
-not the case for all the index functions. \texttt{norm\_kol}, a 1D
-projection function based on the Kolmogorov test, compares the
-difference between the 1D projected data, \(\mathbf{Y}_{n \times 1}\)
-and a randomly generated normal distribution, \(y_n\) based on the
-empirical cumulated distribution function (ECDF). Denote the ECDF
-function as \(F_{.}(u)\) with the subscript indicating either the
-projection or the random normal variable, the \texttt{norm\_kol} index
-is defined by
-
-\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\]
-
-Figure \ref{fig:noisy-better-geo} compares the tracing plot of two
-optimisers: \texttt{search\_geodesic} and \texttt{search\_better}. This
-time, the interpolated path is no longer smooth when using either
-algorithm and \texttt{search\_geodesic} fails to optimise this index
-with barely any improvement of the index value. On the other hand,
-\texttt{search\_better} is doing relatively well. With the theoretical
-best basis {[}0, 1, 0, 0, 0{]} producing an index value of 0.17,
-\texttt{search\_better} finds the final basis {[} 0.0376, -0.9916,
--0.0581, -0.0831, 0.0716{]} with an index value of 0.165. A further
-polish step will give a marginal improvement of index value to 0.175
-with a basis of {[} 0.0223, -0.9965, -0.0352, -0.0591, 0.0418{]}. At
-this stage, the difference between the theoretical best and what has
-been found is likely due to simulation error since the best possible
-basis for a simulated data will be slightly off the theoretical best
-basis, which is derived based on the distributional assumptions in
-Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+not the case for all. A 1D projection function based on the Kolmogorov
+test, \texttt{norm\_kol}, compares the projected data,
+\(\mathbf{Y}_{n \times 1}\) to a randomly generated normal distribution,
+\(y_n\) based on the empirical cumulated distribution function (ECDF).
+Let \(F_{.}(u)\) be the ECDF function with the subscript \(P\) and \(u\)
+indicating the projected and randomly generated ECDF respectively, the
+index is defined as:
+
+\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\] With a more complex
+index function, it is interesting to understand 1) if all the three
+optimisers can reach the optimum, 2) when a local optimum is presented,
+whether the simulated annealing algorithm can escape that local optimum
+and find the global one.
+
+\begin{Schunk}
+\begin{widefigure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+
+}
+
+\caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index.
+The trace plot shows the failure of pseudo derivative on non-smooth index.
+The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
+\end{widefigure}
+\end{Schunk}
+
+Figure \ref{fig:noisy-better-geo} presents the tracing plots of two
+optimisers: pseudo derivative and simulated annealing and as expected,
+the interpolated path is no longer smooth in either case. There is
+barely any improvement on the index value in the pseudo derivative
+algorithm while the simulated annealing algorithm has managed to get
+close to the index value of the theoretical best, indicated by the
+horizontal dashed line. In the PCA plot, simulated annealing is able to
+move gradually towards the theoretical best while the pseudo derivative
+method breaks after the initial few attempts.
 

### 2021-01-24 scripts/examples.R

@@ -60,23 +60,24 @@ gl <-  lapply(list(before, after), grid::rasterGrob)
 wrap_plots(gl)
 
 ## ---- noisy-better-geo
+## code for generating kol_1d_geo, kol_1d_better and kol_1d_better_polish
 # set.seed(123456)
 # kol_1d_geo <-
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
-#                                              search_f =  search_geodesic, max.tries = 100),
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+#                                              search_f =  search_geodesic, max.tries = 200),
 #                rescale = FALSE, verbose = TRUE)
 #
 # set.seed(123456)
 # kol_1d_better <-
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
-#                                              search_f =  search_better, max.tries = 100),
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+#                                              search_f =  search_better, max.tries = 200),
 #                rescale = FALSE, verbose = TRUE)
 #
 # last <- get_best(kol_1d_better)$basis %>% .[[1]]
 #
 # set.seed(123456)
 # kol_1d_better_polish <-
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(), d = 1,
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
 #                                              search_f =  search_polish, max.tries = 100),
 #                rescale = FALSE, verbose = TRUE, start = last)
 #
@@ -85,18 +86,27 @@ wrap_plots(gl)
 # save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
 # save(kol_1d_better_polish, file = here::here("data", "kol_1d_better_polish.rda"))
 
+
 index <- tourr::norm_kol(nrow(boa5))
 theo_best_index_val <- index(as.matrix(boa5) %*% matrix(c(0, 1, 0, 0, 0), nrow = 5, ncol = 1))
 
+dt <- dplyr::bind_rows(kol_1d_geo, kol_1d_better) %>%
[…]

### 2021-01-24 scripts/kol_result.R

@@ -1,5 +1,5 @@
 ## ----kol-result
-clean %>%
+clean_pca <- clean %>%
   filter(method != "search_polish") %>%
   bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
                    index = tourr::norm_kol(nrow(boa6)),
@@ -11,14 +11,27 @@ clean %>%
          alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
          search_name = ifelse(str_detect(sim, "random"), "search_better_random", "search_better")) %>%
   compute_pca(group = seed_sim) %>%
-  pluck("aug") %>%
+  pluck("aug")
+
+global_max <- clean_pca %>%
+  filter(info == "theoretical") %>%
+  filter(index_val == max(index_val)) %>%
+  select(-alpha, -search_name)
+
+local_max <- clean_pca %>%
+  filter(info == "theoretical") %>%
+  filter(index_val == min(index_val)) %>%
+  select(-alpha, -search_name)
+
+clean_pca %>% drop_na() %>%
   explore_space_pca(details = FALSE, pca = FALSE, group = seed_sim, color = var_found,
-                    precision = 0.001, theo_size = 15,
-                    start_size = 1, end_size = 3,
+                    precision = 0.001,
+                    start_size = 1, end_size = 2.5,
                     interp_size = 0.5) +
+  add_theo(dt = global_max) +
+  add_theo(dt = local_max, theo_label = "x", theo_size = 12) +
   scale_color_botanical() +
   facet_grid(alpha ~ search_name, labeller = label_both)
-
 
 
 

### 2021-01-24 scripts/vis-diag.R

@@ -38,7 +38,7 @@ p1 | p2
 bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
-  explore_space_pca(group = method, interp_size = 1, anchor_size = 2, finish_size = 2) +
+  explore_space_pca(group = method, interp_size = 1, anchor_size = 2, finish_size = 2, ratio = 4.5) +
   scale_color_botanical(palette = "fern")
 
 ## ----toy-pca-animated

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.Rmd
more changes on the plot for the last simulation and writing

@@
 
-```{r noisy-better-geo, fig.env = "widefigure", fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
+```{r noisy-better-geo, fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
 The trace plot shows the failure of pseudo derivative on non-smooth index.
@@
 
-```{r kol-result, fig.height = 6, fig.cap="One-D projection of \\code{norm\\_kol} index on \\code{boa6} data optimised by \\code{search\\_better}, starting from 20 different seeds. "}
+```{r kol-result, fig.height = 6, fig.cap="One-D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by \\code{search\\_better} and \\code{search\\_better\\_random}.
+Two different neighbourhood size of 0.5 and 0.7 is used.
+The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing."}
 ```
 
-The second experiment with the noisy index is to understand how the optimisers perform when a local maximum is present.
-The dataset used is \code{boa6} where `x2` and `x7` are informative.
-The two theoretical best bases are [`r matrix(c(0, 1, 0, 0, 0, 0), nrow = 6)`] and [`r matrix(c(0, 0, 1, 0, 0, 0), nrow = 6)`] with index value `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 1, 0, 0, 0, 0), nrow = 6); index(proj)` and `r index <- tourr::norm_kol(nrow(boa6)); proj <- as.matrix(boa6) %*% matrix(c(0, 0, 1, 0, 0, 0), nrow = 6); index(proj)`, respectively.
-Hence, the global maximum happens when variable `x7` is found.
-
-<!-- Simulation is done with 20 randomly generated seeds on two optimisers: `search_better` (Figure \ref{fig:kol-better}) and `search_better_random` (Figure \ref{fig:kol-random} and Figure \ref{fig:kol-random-tuned}).
-Compared to Figure \ref{fig:kol-better} and \ref{fig:kol-random}, Figure \ref{fig:kol-random-tuned}further increases the neighbourhood parameter `alpha` from 0.5 to 0.7 to enlarge the search space.
-The data object is collected for each simulation and computation is done to reduce the bases from its original 6D to a 2D points by principal component analysis so as to view the search paths.
--->
-
-The conclusion from this experiment is that the usage of `search_better_random` and increasing the search space are methods that can avoid getting trapped in local maxima but the solution will still depend on the starting points of the simulation and the seed used.
+The next experiment compares two optimisers: `search_better` and `search_better_random` for 20 simulation paths.
+Two search neighbourhood sizes, `alpha`, of 0.5 and 0.7 are also compared to understand the effect of neighbourhood size where a larger value indicating a wider search.
+Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected interpolation paths faceted by optimiser and alpha.
+Several results can be drawn through the comparison: 1) With `search_better` and `alpha = 0.5`, despite being the simplest and fastest, the optimiser fails to optimise for three instances where they finish neither near the local nor the global optimum.
+2) With the same optimiser but a larger neighbourhood size `alpha = 0.7`, more seeds have found the global optimum.
+3) Comparing different optimisers `search_better` and `search_better_random` with the same search size of 0.5, `search_better_random` finishes closer to the theoretical for the ones that find the local optimum.
+This indicates the optimiser is working harder to examine if there is other optimum in the space but a small search neighbourhood size makes it hard to conduct a holistic search of the whole space.
+4) With `search_better_random` and a larger neighbourhood size of 0.7, more seeds have found the global optimum.
+Also some points are getting closer to the position of the theoretical best, as comparing with `alpha = 0.5`, and this would indicate fewer work for the subsequent polish.
 

### 2021-01-24 paper/zhang-cook-laa-langrene-menendez.tex
more changes on the plot for the last simulation and writing

@@
 \begin{Schunk}
-\begin{widefigure}
+\begin{figure}
 
@@
 The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
-\end{widefigure}
+\end{figure}
 \end{Schunk}
@@
 
-\caption[One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds]{One-D projection of \code{norm\_kol} index on \code{boa6} data optimised by \code{search\_better}, starting from 20 different seeds.
-}\label{fig:kol-result}
+\caption[One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}]{One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}.
+Two different neighbourhood size of 0.5 and 0.7 is used.
+The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing.}\label{fig:kol-result}
 \end{figure}
@@
 
-The second experiment with the noisy index is to understand how the
-optimisers perform when a local maximum is present. The dataset used is
-\code{boa6} where \texttt{x2} and \texttt{x7} are informative. The two
-theoretical best bases are {[}0, 1, 0, 0, 0, 0{]} and {[}0, 0, 1, 0, 0,
-0{]} with index value 0.176 and 0.235, respectively. Hence, the global
-maximum happens when variable \texttt{x7} is found.
-
-The conclusion from this experiment is that the usage of
-\texttt{search\_better\_random} and increasing the search space are
-methods that can avoid getting trapped in local maxima but the solution
-will still depend on the starting points of the simulation and the seed
-used.
+The next experiment compares two optimisers: \texttt{search\_better} and
+\texttt{search\_better\_random} for 20 simulation paths. Two search
+neighbourhood sizes, \texttt{alpha}, of 0.5 and 0.7 are also compared to
+understand the effect of neighbourhood size where a larger value
+indicating a wider search. Figure \ref{fig:kol-result} shows 80 (20 * 4)
+PCA projected interpolation paths faceted by optimiser and alpha.
+Several results can be drawn through the comparison: 1) With
+\texttt{search\_better} and \texttt{alpha\ =\ 0.5}, despite being the
+simplest and fastest, the optimiser fails to optimise for three
+instances where they finish neither near the local nor the global
+optimum. 2) With the same optimiser but a larger neighbourhood size
+\texttt{alpha\ =\ 0.7}, more seeds have found the global optimum. 3)
+Comparing different optimisers \texttt{search\_better} and
+\texttt{search\_better\_random} with the same search size of 0.5,
+\texttt{search\_better\_random} finishes closer to the theoretical for
+the ones that find the local optimum. This indicates the optimiser is
+working harder to examine if there is other optimum in the space but a
+small search neighbourhood size makes it hard to conduct a holistic
+search of the whole space. 4) With \texttt{search\_better\_random} and a
+larger neighbourhood size of 0.7, more seeds have found the global
+optimum. Also some points are getting closer to the position of the
+theoretical best, as comparing with \texttt{alpha\ =\ 0.5}, and this
+would indicate fewer work for the subsequent polish.
 

### 2021-01-24 scripts/kol_result.R

@@ -9,7 +9,9 @@ clean_pca <- clean %>%
                    raw_data = boa6) %>%
   mutate(seed_sim = paste0(seed, sim),
          alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
-         search_name = ifelse(str_detect(sim, "random"), "search_better_random", "search_better")) %>%
+         search_name = ifelse(str_detect(sim, "random"), "search_better_random", "search_better"),
+         var_found = as.factor(ifelse(var_found == "V7", "global", "local")),
+         var_found = fct_relevel(var_found, "global", after = 1)) %>%
   compute_pca(group = seed_sim) %>%
   pluck("aug")
 
@@ -25,11 +27,10 @@ local_max <- clean_pca %>%
 
 clean_pca %>% drop_na() %>%
   explore_space_pca(details = FALSE, pca = FALSE, group = seed_sim, color = var_found,
-                    precision = 0.001,
-                    start_size = 1, end_size = 2.5,
+                    start_size = 1, end_size = 2,
                     interp_size = 0.5) +
-  add_theo(dt = global_max) +
-  add_theo(dt = local_max, theo_label = "x", theo_size = 12) +
+  add_theo(dt = global_max, theo_size = 20) +
+  add_theo(dt = local_max, theo_label = "x", theo_size = 10) +
   scale_color_botanical() +
   facet_grid(alpha ~ search_name, labeller = label_both)
 

### 2021-01-24 scripts/vis-diag.R

@@ -12,24 +12,24 @@ purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 p1 <- holes_2d_better_max_tries %>%
   mutate(max_tries = 500) %>%
   explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
-  scale_color_botanical(palette = "daisy")
+  scale_color_botanical(palette = "fern")
 
 p2 <- holes_2d_better_random %>%
   mutate(max_tries = 25) %>%
   explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
-  scale_color_botanical(palette = "daisy")
+  scale_color_botanical(palette = "fern")
 
 (p1 | p2)
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
   mutate(group = "Algorithm 1") %>%
-  explore_trace_interp(accuracy_x = 4) +
+  explore_trace_interp(iter = id, color = tries, accuracy_x = 4) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
 p2 <- holes_2d_better_random %>%
   mutate(group = "Algorithm 3") %>%
-  explore_trace_interp(accuracy_x = 14) +
+  explore_trace_interp(iter = id, color = tries, accuracy_x = 14) +
   scale_color_botanical(palette = "fern", discrete = FALSE)
 
 p1 | p2

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.Rmd
clean up interruption and writing

@@
 library(ferrn)
-read_chunk(here::here('scripts/vis-diag.R'))
-read_chunk(here::here('scripts/examples.R'))
+read_chunk(here::here('scripts/toy.R'))
+read_chunk(here::here('scripts/diagnose.R'))
 read_chunk(here::here('scripts/kol_result.R'))
@@
 
-For a particular index function, the best algorithm to optimise relates to the character of the index and the data.
+For a particular index function, the best algorithm to optimise depends on the character of the index and the data.
 If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are present, `search_better` may get stuck at a local maximum and in the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
-In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers in different scenarios.
+When multiple optima are present, `search_better` may get stuck at a local maximum.
+In the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers.
 
@@
 
-Random variables with different structures have been simulated and the distribution of each is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable `x1`, `x8`, `x9` and `x10` are normal distributed with zero mean and unit variance and `x2` to `x7` are mixtures of normal distributions with varied weights and locations.
-The mixture variables have been scaled to have an overall unit variance before running the projection pursuit.
+Random variables with different distributions have been simulated and the distributional form of each variable is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
+All the variables have been scaled to have an overall unit variance before running the projection pursuit.
 
@@
 
-In section \ref{toy-interp}, an interpolation with increase-then-decrease pattern has been presented.
-This pattern is undesirable since the optimiser could have started the next iteration from the highest basis on the tour path, as annotated as the interpolated basis in the plot, but instead, it is forced to start from the target basis.
-This motivates the design of an interruption to check the index value on the tour path so that the interpolating bases is accepted only up to the one with the largest index value.
-After implementing this interruption, the search finds a higher final index value with fewer steps as shown in the right panel of Figure \ref{fig:interruption}.
+A non-monotonic interpolation path has already been given in Figure \ref{fig:toy-interp}: to get to the target basis from the current one, the interpolator may have passed a basis with even higher index value.
+This motivates the design of an interruption to check whether higher index value can be reached during the interpolation.
+If such a basis can be found, the optimiser will start the next iteration from that basis instead of the target one.
+Figure \ref{fig:interruption} contrasts the trace with and without the interruption for the simulated annealing optimiser: `search_better`.
+The first two interpolation are identical while in the third interpolation, a better basis has been found than the target one.
+Rather than starting the next iteration from the "target basis" on Time 62, as shown in the left panel, the interruption, on the right panel, starts the next iteration from the "interpolation basis" on Time 61.
+The interruption has also detected better basis on the fourth interpolation.
+With the interruption, the optimiser is able to find the final basis with a higher index value (0.865 vs.
+0.879).
 
-```{r interruption, fig.asp=0.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by \\code{search\\_better}.
-A comparison of the trace plot with and without interruption.
-When the interruption does not take place, the index value of the target basis can be smaller than the interpolated basis.
-The interruption forces the interpolation to finish at the highest interpolated basis on the tour path."}
+```{r interruption, fig.asp=0.5,fig.env = "widefigure", fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index optimised by \\code{search\\_better}.
+The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
+With this implementation, the optimiser is able to find a better final basis."}
 ```

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.tex
clean up interruption and writing

@@
 
-For a particular index function, the best algorithm to optimise relates
-to the character of the index and the data. If the index function is
+For a particular index function, the best algorithm to optimise depends
+on the character of the index and the data. If the index function is
 smooth and has a single maximum, all of the three algorithms introduced
 above can find the maximum. When multiple optima are present,
-\texttt{search\_better} may get stuck at a local maximum and in the case
+\texttt{search\_better} may get stuck at a local maximum. In the case
 where the index function is non-smooth, \texttt{search\_geodesic} may
@@
 presented to outline how the diagnostic plots can be used to compare the
-performance of optimisers in different scenarios.
+performance of optimisers.
 
@@
 
-Random variables with different structures have been simulated and the
-distribution of each is presented in Equations \ref{eq:sim-norm} to
-\ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8}, \texttt{x9} and
-\texttt{x10} are normal distributed with zero mean and unit variance and
-\texttt{x2} to \texttt{x7} are mixtures of normal distributions with
-varied weights and locations. The mixture variables have been scaled to
-have an overall unit variance before running the projection pursuit.
+Random variables with different distributions have been simulated and
+the distributional form of each variable is presented in Equations
+\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8},
+\texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
+variance and \texttt{x2} to \texttt{x7} are normal mixtures with varied
+weights and locations. All the variables have been scaled to have an
+overall unit variance before running the projection pursuit.
 
@@
 
-In section \ref{toy-interp}, an interpolation with
-increase-then-decrease pattern has been presented. This pattern is
-undesirable since the optimiser could have started the next iteration
-from the highest basis on the tour path, as annotated as the
-interpolated basis in the plot, but instead, it is forced to start from
-the target basis. This motivates the design of an interruption to check
-the index value on the tour path so that the interpolating bases is
-accepted only up to the one with the largest index value. After
-implementing this interruption, the search finds a higher final index
-value with fewer steps as shown in the right panel of Figure
-\ref{fig:interruption}.
+A non-monotonic interpolation path has already been given in Figure
+\ref{fig:toy-interp}: to get to the target basis from the current one,
+the interpolator may have passed a basis with even higher index value.
+This motivates the design of an interruption to check whether higher
+index value can be reached during the interpolation. If such a basis can
+be found, the optimiser will start the next iteration from that basis
+instead of the target one. Figure \ref{fig:interruption} contrasts the
+trace with and without the interruption for the simulated annealing
+optimiser: \texttt{search\_better}. The first two interpolation are
+identical while in the third interpolation, a better basis has been
+found than the target one. Rather than starting the next iteration from
+the ``target basis'' on Time 62, as shown in the left panel, the
+interruption, on the right panel, starts the next iteration from the
+``interpolation basis'' on Time 61. The interruption has also detected
+better basis on the fourth interpolation. With the interruption, the
+optimiser is able to find the final basis with a higher index value
+(0.865 vs.~0.879).
+
+\begin{Schunk}
+\begin{widefigure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+
+}
+
+\caption[Trace plot of 2D projection on \code{boa6} data with holes index optimised by \code{search\_better}]{Trace plot of 2D projection on \code{boa6} data with holes index optimised by \code{search\_better}.
+The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
+With this implementation, the optimiser is able to find a better final basis.}\label{fig:interruption}
+\end{widefigure}
+\end{Schunk}
 

### 2021-01-25 scripts/clean_interrupt.R

@@ -1,0 +1,27 @@
+origin <- interrupt_no
+
+before <- origin %>%
+  select(basis, index_val, info, method, tries, loop, id) %>%
+  mutate( id = row_number(),
+          loop = ifelse(id == 1, 1, loop),
+          method = ifelse(id == 1, "search_better", method),
+          tries = ifelse(id == 1, tries, tries + 1)) %>%
+  group_by(tries, info) %>%
+  mutate(loop = ifelse(is.na(loop), row_number(), loop)) %>%
+  ungroup()
+
+origin2 <- interrupt_yes
+
+after <- origin2 %>%
+  select(basis, index_val, info, method, tries, loop, id) %>%
+  mutate( id = row_number(),
+          loop = ifelse(id == 1, 1, loop),
+          method = ifelse(id == 1, "search_better", method),
+          tries = ifelse(id == 1, tries, tries + 1)) %>%
+  group_by(tries, info) %>%
+  mutate(loop = ifelse(is.na(loop), row_number(), loop)) %>%
+  ungroup()
+
+
+save(before, file = here::here("data", "before.rda"))
+save(after, file = here::here("data", "after.rda"))

### 2021-01-25 scripts/diagnose.R

@@ -1,16 +1,28 @@
 ## ----interruption
-p1_anno <- interrupt_no %>% filter(info == "interpolation") %>% mutate(id = row_number()) %>% filter(id %in% c(44, 60, 62)) %>%
-  mutate(anno = c("current basis", "interpolated basis", "target basis"))
+interp <- before %>% get_interp()
 
-# p1 <- interrupt_no %>% mutate(id = row_number() - 1) %>% explore_trace_interp() + ggtitle("without interruption") +
-#   geom_point(data = p1_anno) +
-#   geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5) + ylim(0.8, 0.9) + xlim(0, 80) +
-#   theme(legend.position = "none")
-#
-# p2 <- interrupt_yes %>% explore_trace_interp() + ggtitle("with interruption") + ylim(0.8, 0.9) + xlim(0, 80) +
-#   theme(legend.position = "none")
+p1_anno <- bind_rows(
+  interp %>% filter(tries == 4, info == "interpolation") %>%
+    filter(index_val == max(index_val)) %>%
+    mutate(anno = "interpolation basis"),
+  interp %>% get_interp_last %>% filter(tries %in% c(3, 4)) %>%
+    mutate(anno = c("current basis", "target basis"))
+) %>% arrange(id)
 
-#(p1 | p2) & scale_color_botanical(palette = "fern", discrete = TRUE)
+p1 <- before %>%
+  explore_trace_interp(iter = id, color = tries, accuracy_y = 0.001) +
+  scale_color_botanical(discrete = FALSE) +
+  ggtitle("without interruption") +
+  geom_point(data = p1_anno, size = 3) +
+  geom_point(data = get_best(after) %>% mutate(id = 78), size = 0, color = "white") +
+  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
+
+p2 <- after %>%
+  explore_trace_interp(iter = id, color = tries, accuracy_y = 0.001) +
+  ggtitle("with interruption") +
+  scale_color_botanical(discrete = FALSE)
+
+(p1 | p2)
 
 ## ----polish
 # set.seed(123456)

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.Rmd
a coarse go-through of search_better -> SA etc, writing on polish, widefigure for trace plots

@@
 
-Below we introduce three line search algorithms in tour: `search_better`, `search_better_random`, and `search_geodesic`.
-The first two are simulated annealing algorithms whilst `search_geodesic` evaluates neighbourhood basis to determine the search direction.
+Below we introduce three line search algorithms in tour: simulated annealing (SA), simulated annealing with jump out (SAJO), and pseudo derivative (PD).
 
@@
 
-`search_better` is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
+Simulated annealing (SA) is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
 $\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled.
@@
 The termination condition is when the maximum number of iteration $l_{\max}$ is reached.
-The algorithm of `search_better` is summarised in Algorithm \ref{random-search}.
+The algorithm of simulated annealing is summarised in Algorithm \ref{random-search}.
 A slightly different cooling scheme has been proposed by @posse1995projection to avoid the search space being reduced too fast.
@@
 
-Simulated annealing (`search_better_random`) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as `search_better` but allows a probabilistic acceptance of a basis with lower index value based on the annealing $T(l)$.
+Simulated annealing with jump out (SAJO) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process but allows a probabilistic acceptance of a inferior basis with lower index value based on the annealing $T(l)$.
 Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
@@
 This feature is particularly useful when local maxima are present.
-The algorithm \ref{simulated-annealing} highlights how simulated annealing differs from random search in the inner loop.
+The algorithm \ref{simulated-annealing} highlights how SAJO differs from SA in the inner loop.
 
@@
 
-@cook1995grand used a line search algorithm on the space of the projection bases.
-In `search_geodesic`, the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
+In pseudo derivative [@cook1995grand] method, the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
 The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen.
 The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the percentage improve condition or when $l_{\max}$ is reached.
-Algorithm \ref{search-geodesic} summarises the inner loop in geodesic search.
+Algorithm \ref{search-geodesic} summarises the inner loop of the pseudo derivative search.
 
@@
 
-In the optimisation algorithms for projection pursuit, the three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
-Possible values for `search_better` and `search_better_random` include `random_search`,  `new_basis`, and `interpolation`.
-`search_geodesic` has a wider variety that includes `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
+Three main elements are recorded for the projection pursuit optimisers: 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
+For optimiser SA and SAJO, possible values of the state include `random_search`,  `new_basis`, and `interpolation`.
+pseudo derivative search has a wider variety of state including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
 
@@
 
-Figure \ref{fig:toy-search} shows the searches of two different optimisers: `search_better` and `search_better_random`.
+Figure \ref{fig:toy-search} shows the searches of two different optimisers: SA (left) and SAJO (right).
 Both optimisers progress quickly at the first few iterations, take longer to find better basis in the later iterations, and finish off when hitting the maximum number of try, 499 and 999, respectively.
-Also, the target basis found in each iteration of `search_better` always has an increased index value while this is not the case for `search_better_random`.
-This explains why in this example, `search_better_random` takes longer to find the final basis, but in more complicated scenarios, the feature of probabilistic acceptance allows a more holistic search of the basis space and is more likely to find the global optimum.
+Also, the target basis found on the left panel always has an increased index value in each iteration while this is not the case for the plot on the right.
+This explains why in this example, the optimiser SAJO takes longer to find the final basis, but in more complicated scenarios, the feature of probabilistic acceptance allows a more holistic search of the basis space and is more likely to find the global optimum.
 
-```{r toy-search, fig.cap="A comparison of search by two otimisers: \\code{search\\_better} and \\code{search\\_better\\_random} on a six-variable dataset \\code{boa6} with holes index on a 2D problem.
+```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \\code{boa6} with holes index on a 2D problem.
 Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
-\\code{search\\_better\\_random} takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem."}
+Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem."}
 ```
@@
 
-```{r toy-interp, fig.cap = "Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
+```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
 The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers."}
@@
 
-```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: pseudo derivative and simulated annealing.
+```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
 All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
@@
 If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are present, `search_better` may get stuck at a local maximum.
-In the case where the index function is non-smooth, `search_geodesic` may even fail to find the maximum.
+When multiple optima are present, simulated annealing (SA) may get stuck at a local maximum.
+In the case where the index function is non-smooth, pseudo derivative search may even fail to find the maximum.
 In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers.
@@
 If such a basis can be found, the optimiser will start the next iteration from that basis instead of the target one.
-Figure \ref{fig:interruption} contrasts the trace with and without the interruption for the simulated annealing optimiser: `search_better`.
-The first two interpolation are identical while in the third interpolation, a better basis has been found than the target one.
+Figure \ref{fig:interruption} contrasts the trace with and without the interruption for the simulated annealing optimiser and it can be observed that while the first two interpolation are identical, a better basis has been found than the target one in the third interpolation.
 Rather than starting the next iteration from the "target basis" on Time 62, as shown in the left panel, the interruption, on the right panel, starts the next iteration from the "interpolation basis" on Time 61.
@@
 
-```{r interruption, fig.asp=0.5,fig.env = "widefigure", fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index optimised by \\code{search\\_better}.
+```{r interruption, fig.env = "widefigure", fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index optimised by optimiser SA.
 The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
@@
 
-```{r polish, fig.height= 2.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by \\code{search\\_geodesic}.
+```{r polish, fig.height= 2.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by pseudo derivative.
 The left panel shows the final projected data before polish and the right panel shows the one after.
@@
 
-Once the final basis has been found by an algorithm, one may want to push further to investigate whether there is an even better basis in the close neighbourhood.
-This motivates the polish search where the final basis is supplied as the start of a new guided tour to search for any local breakthrough.
+Once the final basis has been found by an optimiser, one may want to push further in the close neighbourhood to find an even better basis.
+A polish search takes the final basis as the start of a new guided tour to search for local breakthrough.
 
-Similar to `search_better` as a stochastic random search, `search_polish` has a different scheme of reducing the search neighbourhood.
-In each search-interpolation iteration, `search_better` has a fixed neighbourhood parameter alpha and this alpha is reduced by the cooling parameter only after an iteration finishes.
-On the contrary, `search_polish` allows alpha to be reduced during each iteration to exploit the search in the neighbourhood.
-Further, to avoid the case where alpha becomes too small and the further search is meaningless, three more stopping criteria have been added, on top of the original `max.tries` limit.
-These include:
+The polish algorithm is similar to the simulated annealing one with three main distinctions on 1) the number of candidate basis generated each time in the inner loop; 2) the search neighbourhood, `alpha`, and 3) the termination conditions.
+In simulated annealing, only one candidate basis $\mathbf{A}_l$ is generated every time while the polish algorithm allows for a specified number of basis to be evaluated for each repetition in the inner loop via an argument `n_sample`.
+The search neighbourhood in simulated annealing is reduced in the outer loop, that is, the alpha for one iteration of search remains the same, while in polish search, `alpha` is reduced once a set of candidate bases fails to make an improvement.
+To avoid the case the neighbourhood size becoming too small to generate meaningful search, more termination criteria have been added, apart from the original `max.tries` condition and  these include:
 
-1) the distance between the candidate basis and the current basis needs to be larger than 1e-3;
+1) the distance between the basis found and the current basis needs to be larger than 1e-3;
 2) the percentage change of the index value need to be larger than 1e-5; and
@@
 
-Figure \ref{fig:polish} presents the final projections found before and after applying `search_polish` on `search_geodesic`.
-Polish search improves the index value from `r get_best(holes_2d_geo)$index_val %>% format(digits = 4)`  to `r get_best(holes_2d_geo_polish)$index_val %>% format(digits = 4)` with reduction of weights on the non-informative variables.
-In terms of the projected data as in Figure \ref{fig:polish}, polish works to sharpen the edges of each cluster.
+Figure \ref{fig:polish} presents the final projections found before (left) and after(right) applying the polish using the pseudo derivative optimiser.
+Before polishing, the four clusters are clearly shown while the edges of each cluster is relatively unshaved.
+The polish search start with this basis and on the right panel, the edges of the clusters are better trimmed.
 
 ## Seeing the signal in the noise
-
-<!-- ```{r noisy-better-geo,fig.asp=0.5,fig.cap = "One-D projection on `boa5` data with noisy index `norm_kol` optimised by `search_geodesic` and `search_better`.
-The grey dashed line represents the index value of the theoretical best basis.
-`search_geodesic` fails to optimise the noisy index while `search_better` has made reasonable improvements to reach the index value close to the one of the theoretical best basis."} -->
-<!-- ``` -->
 
@@
 
-$$\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]$$
+$$\max \left[F_{P}(u) - F_{y}(u)\right]$$
 With a more complex index function, it is interesting to understand 1) if all the three optimisers can reach the optimum, 2) when a local optimum is presented, whether the simulated annealing algorithm can escape that local optimum and find the global one.
 
-```{r noisy-better-geo, fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
+```{r noisy-better-geo, fig.env = "widefigure", fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
 The trace plot shows the failure of pseudo derivative on non-smooth index.
@@
 
-```{r kol-result, fig.height = 6, fig.cap="One-D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by \\code{search\\_better} and \\code{search\\_better\\_random}.
+```{r kol-result, fig.height = 6, fig.cap="1D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by SA and SAJO.
 Two different neighbourhood size of 0.5 and 0.7 is used.
@@
 
-The next experiment compares two optimisers: `search_better` and `search_better_random` for 20 simulation paths.
+The next experiment compares the two simulated annealing optimisers, SA and SAJO,  for 20 simulation paths.
 Two search neighbourhood sizes, `alpha`, of 0.5 and 0.7 are also compared to understand the effect of neighbourhood size where a larger value indicating a wider search.
 Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected interpolation paths faceted by optimiser and alpha.
-Several results can be drawn through the comparison: 1) With `search_better` and `alpha = 0.5`, despite being the simplest and fastest, the optimiser fails to optimise for three instances where they finish neither near the local nor the global optimum.
-2) With the same optimiser but a larger neighbourhood size `alpha = 0.7`, more seeds have found the global optimum.
-3) Comparing different optimisers `search_better` and `search_better_random` with the same search size of 0.5, `search_better_random` finishes closer to the theoretical for the ones that find the local optimum.
-This indicates the optimiser is working harder to examine if there is other optimum in the space but a small search neighbourhood size makes it hard to conduct a holistic search of the whole space.
-4) With `search_better_random` and a larger neighbourhood size of 0.7, more seeds have found the global optimum.
-Also some points are getting closer to the position of the theoretical best, as comparing with `alpha = 0.5`, and this would indicate fewer work for the subsequent polish.
+Several results can be drawn through the comparison: 1) With optimiser SA and a 0.5 neighbourhood search size, despite being the simplest and fastest, the optimiser fails to optimise for three instances where they finish neither near the local nor the global optimum.
+2) With the same optimiser but a larger neighbourhood size of 0.7, more seeds have found the global optimum.
+3) Comparing between the optimiser SA and SAJO with the same search size of 0.5, SAJO finishes closer to the theoretical for the ones that find the local optimum.
+This indicates that it is working harder to examine if there is other optimum in the space but a small search neighbourhood size makes it hard to conduct a holistic search of the whole space.
+4) With optimiser SAJO and a larger neighbourhood size of 0.7, more seeds have found the global optimum.
+Also some finish points are getting closer to the position of the theoretical best, as comparing with the 0.5 neighbourhood size, and this would indicate fewer work for the subsequent polish.
 

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.tex
a coarse go-through of search_better -> SA etc, writing on polish, widefigure for trace plots

@@
 
-Below we introduce three line search algorithms in tour:
-\texttt{search\_better}, \texttt{search\_better\_random}, and
-\texttt{search\_geodesic}. The first two are simulated annealing
-algorithms whilst \texttt{search\_geodesic} evaluates neighbourhood
-basis to determine the search direction.
+Below we introduce three line search algorithms in tour: simulated
+annealing (SA), simulated annealing with jump out (SAJO), and pseudo
+derivative (PD).
 
@@
 
-\texttt{search\_better} is a random search device that samples a
+Simulated annealing (SA) is a random search device that samples a
 candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
@@
 is when the maximum number of iteration \(l_{\max}\) is reached. The
-algorithm of \texttt{search\_better} is summarised in Algorithm
+algorithm of simulated annealing is summarised in Algorithm
 \ref{random-search}. A slightly different cooling scheme has been
@@
 
-Simulated annealing (\texttt{search\_better\_random})
+Simulated annealing with jump out (SAJO)
 \citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
-same sampling process as \texttt{search\_better} but allows a
-probabilistic acceptance of a basis with lower index value based on the
-annealing \(T(l)\). Given an initial \(T_0\), the temperature at
-iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
-candidate basis fails to have an index value larger than the current
-basis, simulated annealing gives it a second chance to be accepted with
-probability
+same sampling process but allows a probabilistic acceptance of a
+inferior basis with lower index value based on the annealing \(T(l)\).
+Given an initial \(T_0\), the temperature at iteration \(l\) is defined
+as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to
+have an index value larger than the current basis, simulated annealing
+gives it a second chance to be accepted with probability
 \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
@@
 feature is particularly useful when local maxima are present. The
-algorithm \ref{simulated-annealing} highlights how simulated annealing
-differs from random search in the inner loop.
+algorithm \ref{simulated-annealing} highlights how SAJO differs from SA
+in the inner loop.
 
@@
 
-\citet{cook1995grand} used a line search algorithm on the space of the
-projection bases. In \texttt{search\_geodesic}, the search direction is
-computed using the most prominent direction that deviating an tiny angle
-of \(\delta\) from the current basis. The step size is chosen by
+In pseudo derivative \citep{cook1995grand} method, the search direction
+is computed using the most prominent direction that deviating an tiny
+angle of \(\delta\) from the current basis. The step size is chosen by
 optimising the index value along the geodesic direction over an 90
@@
 \(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
-inner loop in geodesic search.
+inner loop of the pseudo derivative search.
 
@@
 
-In the optimisation algorithms for projection pursuit, the three main
-elements to record are 1) projection bases: \(\mathbf{A}\), 2) index
-values: \(I\), and 3) State: \(S\), which labels the observation with
-detailed stage in the optimisation. Possible values for
-\texttt{search\_better} and \texttt{search\_better\_random} include
-\texttt{random\_search}, \texttt{new\_basis}, and
-\texttt{interpolation}. \texttt{search\_geodesic} has a wider variety
-that includes \texttt{new\_basis}, \texttt{direction\_search},
+Three main elements are recorded for the projection pursuit optimisers:
+1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3)
+State: \(S\), which labels the observation with detailed stage in the
+optimisation. For optimiser SA and SAJO, possible values of the state
+include \texttt{random\_search}, \texttt{new\_basis}, and
+\texttt{interpolation}. pseudo derivative search has a wider variety of
+state including \texttt{new\_basis}, \texttt{direction\_search},
 \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
@@
 Figure \ref{fig:toy-search} shows the searches of two different
-optimisers: \texttt{search\_better} and \texttt{search\_better\_random}.
-Both optimisers progress quickly at the first few iterations, take
-longer to find better basis in the later iterations, and finish off when
-hitting the maximum number of try, 499 and 999, respectively. Also, the
-target basis found in each iteration of \texttt{search\_better} always
-has an increased index value while this is not the case for
-\texttt{search\_better\_random}. This explains why in this example,
-\texttt{search\_better\_random} takes longer to find the final basis,
-but in more complicated scenarios, the feature of probabilistic
-acceptance allows a more holistic search of the basis space and is more
-likely to find the global optimum.
+optimisers: SA (left) and SAJO (right). Both optimisers progress quickly
+at the first few iterations, take longer to find better basis in the
+later iterations, and finish off when hitting the maximum number of try,
+499 and 999, respectively. Also, the target basis found on the left
+panel always has an increased index value in each iteration while this
+is not the case for the plot on the right. This explains why in this
+example, the optimiser SAJO takes longer to find the final basis, but in
+more complicated scenarios, the feature of probabilistic acceptance
+allows a more holistic search of the basis space and is more likely to
+find the global optimum.
 
 \begin{Schunk}
-\begin{figure}
+\begin{widefigure}
 
@@
 
-\caption{A comparison of search by two otimisers: \code{search\_better} and \code{search\_better\_random} on a six-variable dataset \code{boa6} with holes index on a 2D problem.
+\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \code{boa6} with holes index on a 2D problem.
 Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
-\code{search\_better\_random} takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem.}\label{fig:toy-search}
-\end{figure}
+Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem.}\label{fig:toy-search}
+\end{widefigure}
 \end{Schunk}
@@
 \begin{Schunk}
-\begin{figure}
+\begin{widefigure}
 
@@
 The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers.}\label{fig:toy-interp}
-\end{figure}
+\end{widefigure}
 \end{Schunk}
@@
 
-\caption[1D projection on the 5-variable dataset \code{boa5} with two optimisers]{1D projection on the 5-variable dataset \code{boa5} with two optimisers: pseudo derivative and simulated annealing.
+\caption[1D projection on the 5-variable dataset \code{boa5} with two optimisers]{1D projection on the 5-variable dataset \code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
 All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
@@
 smooth and has a single maximum, all of the three algorithms introduced
-above can find the maximum. When multiple optima are present,
-\texttt{search\_better} may get stuck at a local maximum. In the case
-where the index function is non-smooth, \texttt{search\_geodesic} may
-even fail to find the maximum. In this section, examples will be
-presented to outline how the diagnostic plots can be used to compare the
-performance of optimisers.
+above can find the maximum. When multiple optima are present, simulated
+annealing (SA) may get stuck at a local maximum. In the case where the
+index function is non-smooth, pseudo derivative search may even fail to
+find the maximum. In this section, examples will be presented to outline
+how the diagnostic plots can be used to compare the performance of
+optimisers.
 
@@
 trace with and without the interruption for the simulated annealing
-optimiser: \texttt{search\_better}. The first two interpolation are
-identical while in the third interpolation, a better basis has been
-found than the target one. Rather than starting the next iteration from
-the ``target basis'' on Time 62, as shown in the left panel, the
+optimiser and it can be observed that while the first two interpolation
+are identical, a better basis has been found than the target one in the
+third interpolation. Rather than starting the next iteration from the
+``target basis'' on Time 62, as shown in the left panel, the
 interruption, on the right panel, starts the next iteration from the
@@
 
-\caption[Trace plot of 2D projection on \code{boa6} data with holes index optimised by \code{search\_better}]{Trace plot of 2D projection on \code{boa6} data with holes index optimised by \code{search\_better}.
+\caption[Trace plot of 2D projection on \code{boa6} data with holes index optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index optimised by optimiser SA.
 The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
@@
 
-\caption[Two-D projection on \code{boa6} data with holes index optimised by \code{search\_geodesic}]{Two-D projection on \code{boa6} data with holes index optimised by \code{search\_geodesic}.
+\caption[Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative]{Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative.
 The left panel shows the final projected data before polish and the right panel shows the one after.
@@
 
-Once the final basis has been found by an algorithm, one may want to
-push further to investigate whether there is an even better basis in the
-close neighbourhood. This motivates the polish search where the final
-basis is supplied as the start of a new guided tour to search for any
-local breakthrough.
-
-Similar to \texttt{search\_better} as a stochastic random search,
-\texttt{search\_polish} has a different scheme of reducing the search
-neighbourhood. In each search-interpolation iteration,
-\texttt{search\_better} has a fixed neighbourhood parameter alpha and
-this alpha is reduced by the cooling parameter only after an iteration
-finishes. On the contrary, \texttt{search\_polish} allows alpha to be
-reduced during each iteration to exploit the search in the
-neighbourhood. Further, to avoid the case where alpha becomes too small
-and the further search is meaningless, three more stopping criteria have
-been added, on top of the original \texttt{max.tries} limit. These
-include:
+Once the final basis has been found by an optimiser, one may want to
+push further in the close neighbourhood to find an even better basis. A
+polish search takes the final basis as the start of a new guided tour to
+search for local breakthrough.
+
+The polish algorithm is similar to the simulated annealing one with
+three main distinctions on 1) the number of candidate basis generated
+each time in the inner loop; 2) the search neighbourhood,
+\texttt{alpha}, and 3) the termination conditions. In simulated
+annealing, only one candidate basis \(\mathbf{A}_l\) is generated every
+time while the polish algorithm allows for a specified number of basis
+to be evaluated for each repetition in the inner loop via an argument
+\texttt{n\_sample}. The search neighbourhood in simulated annealing is
+reduced in the outer loop, that is, the alpha for one iteration of
+search remains the same, while in polish search, \texttt{alpha} is
+reduced once a set of candidate bases fails to make an improvement. To
+avoid the case the neighbourhood size becoming too small to generate
+meaningful search, more termination criteria have been added, apart from
+the original \texttt{max.tries} condition and these include:
 
@@
 \item
-  the distance between the candidate basis and the current basis needs
-  to be larger than 1e-3;
+  the distance between the basis found and the current basis needs to be
+  larger than 1e-3;
 \item
@@
 
-Figure \ref{fig:polish} presents the final projections found before and
-after applying \texttt{search\_polish} on \texttt{search\_geodesic}.
-Polish search improves the index value from 0.9618 to 0.9627 with
-reduction of weights on the non-informative variables. In terms of the
-projected data as in Figure \ref{fig:polish}, polish works to sharpen
-the edges of each cluster.
+Figure \ref{fig:polish} presents the final projections found before
+(left) and after(right) applying the polish using the pseudo derivative
+optimiser. Before polishing, the four clusters are clearly shown while
+the edges of each cluster is relatively unshaved. The polish search
+start with this basis and on the right panel, the edges of the clusters
+are better trimmed.
 
@@
 
-\[\max \left[F_{\mathbf{P}}(u) - F_{y}(u)\right]\] With a more complex
-index function, it is interesting to understand 1) if all the three
-optimisers can reach the optimum, 2) when a local optimum is presented,
-whether the simulated annealing algorithm can escape that local optimum
-and find the global one.
+\[\max \left[F_{P}(u) - F_{y}(u)\right]\] With a more complex index
+function, it is interesting to understand 1) if all the three optimisers
+can reach the optimum, 2) when a local optimum is presented, whether the
+simulated annealing algorithm can escape that local optimum and find the
+global one.
 
 \begin{Schunk}
-\begin{figure}
+\begin{widefigure}
 
@@
 The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
-\end{figure}
+\end{widefigure}
 \end{Schunk}
@@
 
-\caption[One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}]{One-D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by \code{search\_better} and \code{search\_better\_random}.
+\caption[1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO]{1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO.
 Two different neighbourhood size of 0.5 and 0.7 is used.
@@
 
-The next experiment compares two optimisers: \texttt{search\_better} and
-\texttt{search\_better\_random} for 20 simulation paths. Two search
-neighbourhood sizes, \texttt{alpha}, of 0.5 and 0.7 are also compared to
-understand the effect of neighbourhood size where a larger value
-indicating a wider search. Figure \ref{fig:kol-result} shows 80 (20 * 4)
-PCA projected interpolation paths faceted by optimiser and alpha.
-Several results can be drawn through the comparison: 1) With
-\texttt{search\_better} and \texttt{alpha\ =\ 0.5}, despite being the
-simplest and fastest, the optimiser fails to optimise for three
-instances where they finish neither near the local nor the global
-optimum. 2) With the same optimiser but a larger neighbourhood size
-\texttt{alpha\ =\ 0.7}, more seeds have found the global optimum. 3)
-Comparing different optimisers \texttt{search\_better} and
-\texttt{search\_better\_random} with the same search size of 0.5,
-\texttt{search\_better\_random} finishes closer to the theoretical for
-the ones that find the local optimum. This indicates the optimiser is
-working harder to examine if there is other optimum in the space but a
-small search neighbourhood size makes it hard to conduct a holistic
-search of the whole space. 4) With \texttt{search\_better\_random} and a
-larger neighbourhood size of 0.7, more seeds have found the global
-optimum. Also some points are getting closer to the position of the
-theoretical best, as comparing with \texttt{alpha\ =\ 0.5}, and this
-would indicate fewer work for the subsequent polish.
+The next experiment compares the two simulated annealing optimisers, SA
+and SAJO, for 20 simulation paths. Two search neighbourhood sizes,
+\texttt{alpha}, of 0.5 and 0.7 are also compared to understand the
+effect of neighbourhood size where a larger value indicating a wider
+search. Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected
+interpolation paths faceted by optimiser and alpha. Several results can
+be drawn through the comparison: 1) With optimiser SA and a 0.5
+neighbourhood search size, despite being the simplest and fastest, the
+optimiser fails to optimise for three instances where they finish
+neither near the local nor the global optimum. 2) With the same
+optimiser but a larger neighbourhood size of 0.7, more seeds have found
+the global optimum. 3) Comparing between the optimiser SA and SAJO with
+the same search size of 0.5, SAJO finishes closer to the theoretical for
+the ones that find the local optimum. This indicates that it is working
+harder to examine if there is other optimum in the space but a small
+search neighbourhood size makes it hard to conduct a holistic search of
+the whole space. 4) With optimiser SAJO and a larger neighbourhood size
+of 0.7, more seeds have found the global optimum. Also some finish
+points are getting closer to the position of the theoretical best, as
+comparing with the 0.5 neighbourhood size, and this would indicate fewer
+work for the subsequent polish.
 

### 2021-01-25 scripts/diagnose.R

@@ -14,7 +14,7 @@ p1 <- before %>%
   scale_color_botanical(discrete = FALSE) +
   ggtitle("without interruption") +
   geom_point(data = p1_anno, size = 3) +
-  geom_point(data = get_best(after) %>% mutate(id = 78), size = 0, color = "white") +
+  geom_hline(data = get_best(after) %>% mutate(id = 78), aes(yintercept = index_val), color = "grey90") +
   geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
 
 p2 <- after %>%

### 2021-01-25 scripts/toy.R

@@ -19,7 +19,7 @@ p2 <- holes_2d_better_random %>%
   explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
   scale_color_botanical(palette = "fern")
 
-(p1 | p2)
+(p1 | p2) & theme_bw() + theme(legend.position = "none")
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.Rmd
code, gif and vimeo link for animated pca plot

@@
 
-```{r toy-pca-animated, fig.cap = "A selected number of frames from the animated PCA plot. With animation, it is easier to track the progression from the start to finish in each algorithm."}
+```{r toy-pca-animated, fig.cap = "A selected number of frames from the animated PCA plots.
+With animation, it is easier to track the progression of each optimiser from the start to finish.
+A full video of the aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
 ```

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.tex
code, gif and vimeo link for animated pca plot

@@
 
-\caption[A selected number of frames from the animated PCA plot]{A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\caption{A selected number of frames from the animated PCA plots.
+With animation, it is easier to track the progression of each optimiser from the start to finish.
+A full video of the aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
 \end{figure}
@@
 \begin{Schunk}
+\begin{Soutput}
+#> Called from: explore_trace_interp(., color = method, iter = id, group = method)
+#> debug: iter <- dplyr::enexpr(iter)
+#> debug: color <- dplyr::enexpr(color)
+#> debug: group <- dplyr::enexpr(group)
+#> debug: if ("search_geodesic" %in% dt$method) {
+#>     dt <- dt %>% clean_method()
+#> }
+#> debug: dt <- dt %>% clean_method()
+#> debug: dt_interp <- get_interp(dt, group = !!group)
+#> debug: interp_last <- dplyr::bind_rows(get_start(dt), get_interp_last(dt, 
+#>     group = !!group))
+#> debug: tick_x <- format_label(interp_last %>% dplyr::pull(!!iter), accuracy = accuracy_x)
+#> debug: tick_y <- format_label(interp_last$index_val, accuracy = accuracy_y)
+#> debug: a <- dt_interp %>% group_by(group = !!group) %>% dplyr::summarise(row = dplyr::n(), 
+#>     diff = max(!!iter) - min(!!iter) + 1)
+#> debug: if (!all(a$row == a$diff, TRUE)) {
+#>     stop("there is gap(s) in the variable iter!")
+#> }
+#> debug: p <- dt_interp %>% ggplot2::ggplot(ggplot2::aes(x = !!iter, y = .data$index_val, 
+#>     group = !!group)) + ggplot2::geom_line() + ggplot2::geom_point(data = interp_last, 
+#>     ggplot2::aes(col = !!color), size = 3) + ggplot2::geom_vline(data = interp_last, 
+#>     ggplot2::aes(xintercept = !!iter), lty = "dashed", alpha = 0.3) + 
+#>     ggplot2::scale_x_continuous(breaks = tick_x) + ggplot2::scale_y_continuous(breaks = tick_y, 
+#>     labels = scales::label_number(accuracy = accuracy_y)) + theme_fern() + 
+#>     ggplot2::theme(legend.position = "none") + ggplot2::ylab("Index value") + 
+#>     ggplot2::xlab("Time")
+#> debug: if (nrow(dt_interp) < cutoff) {
+#>     p <- p + ggplot2::geom_point(ggplot2::aes(x = !!iter, y = .data$index_val, 
+#>         col = !!color))
+#> }
+#> debug: p
+\end{Soutput}
 \begin{widefigure}

### 2021-01-25 scripts/toy.R

@@ -42,25 +42,26 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
   scale_color_botanical(palette = "fern")
 
 ## ----toy-pca-animated
-# ani <- pca$aug %>%
-#   ggplot(aes(x = PC1, y = PC2)) +
-#   geom_point(data = pca$aug %>% filter(info == "randomly_generated"), col = "grey", size = 0.5) +
-#   #geom_path(data =  pca$aug %>% filter(info == "interpolation"), aes(col = method), size = 2) +
-#   geom_point(data =  pca$aug %>% filter(info != "randomly_generated"), aes(col = method), size = 5) +
-#   geom_point(data = pca$aug %>% filter(info == "theoretical"), aes(col = method), size = 15) +
-#   geom_point(data = get_start(pca$aug), aes(col = method), size = 10) +
-#   scale_color_botanical(palette = "cherry") +
-#   theme_void() +
-#   theme(aspect.ratio = 1, legend.position =  "none",
-#         panel.border = element_rect(colour = "black", fill=NA, size=1)) +
-#   gganimate::transition_states(id) +
-#   gganimate::shadow_mark()
-#
-# animate(ani, nframes = 350, device = "png",
+# dt <- dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                    index = tourr::holes(), raw_data = boa5)
+# ani <- dt %>%
+#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
+#                     theo_size = 45, start_size = 10, end_size = 20) +
+#   theme(legend.position = "none") +
+#   scale_color_botanical(palette = "fern")
+# animate(ani, nframes = 100, device = "png",
 #         renderer = file_renderer("anim/pca/",
 #                                  prefix = "pca", overwrite = TRUE))
 
-frames <- c("0002", "0058", "0078", "0133", "0172", "0350")
+# vimeo version with legend
+# ani <- dt %>%
+#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
+#                     theo_size = 45, start_size = 10, end_size = 20) +
+#   scale_color_botanical(palette = "fern")
+# anim_save(ani, filename = here::here("anim/toy-pca-animated.mp4"))
+
[…]

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.Rmd
clean up the toy tour view and vimeo

@@
 With animation, it is easier to track the progression of each optimiser from the start to finish.
-A full video of the aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
+A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
 ```
@@
 ```{r toy-tour, fig.cap="A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles."}
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
+A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
 ```

### 2021-01-25 paper/zhang-cook-laa-langrene-menendez.tex
clean up the toy tour view and vimeo

@@
 With animation, it is easier to track the progression of each optimiser from the start to finish.
-A full video of the aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
+A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
 \end{figure}
@@
 
-\caption[A selected number of frames from the tour animation for viewing the 5D space of all the projection bases]{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The second frame on the top row views the space from a direction that is close to the one in the PCA plot.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.}\label{fig:toy-tour}
+\caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
+The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
+A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
 \end{figure}

### 2021-01-25 scripts/toy.R

@@ -68,48 +68,39 @@ gl <-  lapply(rl, grid::rasterGrob)
 wrap_plots(gl)
 
 ## ----toy-tour
-# set.seed(1)
-# sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
-# path1 <- holes_1d_better$basis %>%
-#   flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
-# path2 <- holes_1d_geo$basis %>%
-#   flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE) %>% -.
-# theoretical <- matrix(c(0, 1, 0, 0, 0), nrow = 1, byrow = TRUE)
-# start <- get_start(holes_1d_better) %>% pull(basis) %>% .[[1]] %>% matrix(nrow = 1)
-# dt <- rbind(sphere, path1, path2, theoretical, start, -start)
-# colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
-# pal <- c("#D3D3D3",c("#524340",  #orchre
-#                      "#B4B754",  # green
-#                      "#F3B422" # yellow
-# ))
-# color <- c(rep(pal[1], nrow(sphere)),
-#            rep(pal[2], nrow(path1)),
-#            rep(pal[3], nrow(path2)),
-#            rep(pal[4], 1), # theoretical
-#            pal[2], # for start
-#            pal[3] # for start
-# )
-# cex <- c(rep(1, nrow(sphere)),
-#          rep(3, nrow(path1)),
-#          rep(3, nrow(path2)),
-#          10, # theoretical
-#          rep(5, 2) # for start
-# )
-#
-# animate_xy(dt, col = color, cex = cex, tour_path = grand_tour()) # trial
+# prep <- prep_space_tour(dplyr::bind_rows(holes_1d_better, holes_1d_geo) %>%
+#                      bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                                       index = tourr::holes(), raw_data = boa5),
+#                    group = method, palette = botanical_palettes$fern[c(1,6)], axes = "bottomleft",
+#                    point_size = 3, end_size = 8, theo_size = 8)
 #
 # set.seed(123)
 # render(
-#   dt,
+#   prep$basis,
 #   tour_path = grand_tour(),
 #   dev = "png",
-#   display = display_xy(col=color,cex = cex, axes = "bottomleft"),
+#   display = display_xy(col = prep$col, cex = prep$cex, pch = prep$pch,
[…]

### 2021-01-27 paper/zhang-cook-laa-langrene-menendez.tex
update code for scale color as per the latest ferrn

--- 2021-01-25 paper/zhang-cook-laa-langrene-menendez.tex
@@
 
-\caption[A selected number of frames from the animated PCA plot]{A selected number of frames from the animated PCA plot.
-With animation, it is easier to track the progression from the start to finish in each algorithm.}\label{fig:toy-pca-animated}
+\caption{A selected number of frames from the animated PCA plots.
+With animation, it is easier to track the progression of each optimiser from the start to finish.
+A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
 \end{figure}

### 2021-01-27 scripts/diagnose.R

@@ -10,17 +10,17 @@ p1_anno <- bind_rows(
 ) %>% arrange(id)
 
 p1 <- before %>%
-  explore_trace_interp(iter = id, color = tries, accuracy_y = 0.001) +
-  scale_color_botanical(discrete = FALSE) +
+  explore_trace_interp(accuracy_y = 0.001) +
+  scale_color_continuous_botanical() +
   ggtitle("without interruption") +
   geom_point(data = p1_anno, size = 3) +
   geom_hline(data = get_best(after) %>% mutate(id = 78), aes(yintercept = index_val), color = "grey90") +
   geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
 
 p2 <- after %>%
-  explore_trace_interp(iter = id, color = tries, accuracy_y = 0.001) +
+  explore_trace_interp(accuracy_y = 0.001) +
   ggtitle("with interruption") +
-  scale_color_botanical(discrete = FALSE)
+  scale_color_continuous_botanical()
 
 (p1 | p2)
 
@@ -103,9 +103,9 @@ dt <- dplyr::bind_rows(kol_1d_geo, kol_1d_better) %>%
   bind_theoretical(matrix = matrix(c(0, -1, 0, 0, 0), ncol = 1), tourr::norm_kol(nrow(boa5)), raw_data = boa5)
 
 p1 <- dt %>%
-  explore_trace_interp(color = method, iter = id,group = method) +
+  explore_trace_interp(group = method) +
   geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) +
-  scale_color_botanical() +
+  scale_color_continuous_botanical() +
   facet_wrap(vars(method), scales = "free_x")
 
 pca <- dt %>%
@@ -114,7 +114,7 @@ pca <- dt %>%
 
 p2 <- pca %>%
   explore_space_pca(group = method, pca = FALSE, details = FALSE, start_size = 1) +
-  scale_color_botanical() +
+  scale_color_discrete_botanical() +
   theme(legend.position = "none")
 
 (p1 | p2)

### 2021-01-27 scripts/kol_result.R

@@ -31,6 +31,6 @@ clean_pca %>% drop_na() %>%
                     interp_size = 0.5) +
   add_theo(dt = global_max, theo_size = 20) +
   add_theo(dt = local_max, theo_label = "x", theo_size = 10) +
-  scale_color_botanical() +
+  scale_color_discrete_botanical() +
   facet_grid(alpha ~ search_name, labeller = label_both)
 

### 2021-01-27 scripts/toy.R

@@ -12,25 +12,25 @@ purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 p1 <- holes_2d_better_max_tries %>%
   mutate(max_tries = 500) %>%
   explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
-  scale_color_botanical(palette = "fern")
+  scale_color_continuous_botanical()
 
 p2 <- holes_2d_better_random %>%
   mutate(max_tries = 25) %>%
   explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
-  scale_color_botanical(palette = "fern")
+  scale_color_continuous_botanical()
 
 (p1 | p2) & theme_bw() + theme(legend.position = "none")
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
   mutate(group = "Algorithm 1") %>%
-  explore_trace_interp(iter = id, color = tries, accuracy_x = 4) +
-  scale_color_botanical(palette = "fern", discrete = FALSE)
+  explore_trace_interp(accuracy_x = 4) +
+  scale_color_continuous_botanical()
 
 p2 <- holes_2d_better_random %>%
   mutate(group = "Algorithm 3") %>%
-  explore_trace_interp(iter = id, color = tries, accuracy_x = 14) +
-  scale_color_botanical(palette = "fern", discrete = FALSE)
+  explore_trace_interp( accuracy_x = 14) +
+  scale_color_continuous_botanical()
 
 p1 | p2
 
@@ -40,7 +40,7 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
                    index = tourr::holes(), raw_data = boa5)  %>%
   explore_space_pca(group = method, details = TRUE,
                     interp_size = 1, anchor_size = 2, finish_size = 2, ratio = 4.5) +
-  scale_color_botanical(palette = "fern")
+  scale_color_discrete_botanical()
 
 ## ----toy-pca-animated
 # dt <- dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.Rmd
text editing on search, trace and pca plot WIP

@@
 }
-  \caption{random search}
+  \caption{Simulated annealing (SA)}
   \label{random-search}
@@
   }
-  \caption{simulated annealing}
+  \caption{Simulated annealing with jump out (SAJO)}
   \label{simulated-annealing}
@@
 }
-\caption{search geodesic}
+\caption{Pseudo derivative (PD)}
 \label{search-geodesic}
@@
 
-In pseudo derivative [@cook1995grand] method, the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
+In pseudo derivative search [@cook1995grand], the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
 The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen.
@@
 
-A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum.
-A simple treatment of plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes much longer for an optimiser to find a better basis towards the end.
-Another option is to use summarisation in each iteration.
-Boxplot is a suitable candidate that can provide five points summary of each iteration and additional information can be separately added with new layers, for example, text information on the number of points searched in each iteration is added at the bottom of each iteration, the bases returned for interpolation are highlighted in larger size and linked, and information regarding the last iteration where no basis is returned is turned to grey scale.
-Furthermore, an option to switch from boxplot back to point geometry is helpful when the number of observation is small in one iteration and can be achieved via the `cutoff` argument.
+A starting point of diagnosing an optimiser is to understand how many search the optimiser has conducted.
+One may want to simply plot the index value of the search points across its natural order.
+A point geometry may work well if each iteration has a similar number of points,  however, the plot will be disproportionate if some iterations have considerably more points than other - it over-emphasizes the iterations that have more search points!
+An alternative is to summarise the search in each iteration using boxplot where each iteration will be spaced out equally.
+Occasionally, one may still want to switch back to point geometry if the number of points is small and this can be achieved via the `cutoff` argument.
+Additional annotation can be added with new layers, for example, the number of points searched in each iteration can be added as text label at the bottom of each iteration; the anchor bases to interpolate are connected and highlighted in a larger size; and the color of the last iteration is in a grey scale to indicate no better basis found in this iteration.
 
 Figure \ref{fig:toy-search} shows the searches of two different optimisers: SA (left) and SAJO (right).
-Both optimisers progress quickly at the first few iterations, take longer to find better basis in the later iterations, and finish off when hitting the maximum number of try, 499 and 999, respectively.
-Also, the target basis found on the left panel always has an increased index value in each iteration while this is not the case for the plot on the right.
-This explains why in this example, the optimiser SAJO takes longer to find the final basis, but in more complicated scenarios, the feature of probabilistic acceptance allows a more holistic search of the basis space and is more likely to find the global optimum.
+Both optimisers quickly find better bases at the first few iterations,  then take longer in the later iterations, and finally finish when the maximum number of evaluation is reached.
+The anchor bases, the ones with the highest index value in each iteration, always have an increased index value in the optimiser SA while this is not the case for SAJO.
+This feature gives SA an advantage in simple example, like the one presented while the jump out component in SAJO allows for a more holistic search in more complicated examples.
 
-```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \\code{boa6} with holes index on a 2D problem.
-Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
-Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem."}
+```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using holes index.
+Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
 ```
@@
 
-```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
-The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers."}
+```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
+Both traces are smooth while the change of index value in each iteration may not be monotinic."}
 ```
 
-Viewing the index value of points on the interpolation path is another interest to the analysts since the projection on these bases will be played by the tour animation and it provides further information on how the index value changes when moving from one target basis to another.
-Figure \ref{fig:toy-interp} presents the index value against time on the interpolation path for the two optimisers with same configuration as \ref{fig:toy-search}.
-From Figure \ref{fig:toy-interp}, we further know that to reach a target basis with lower index value, the optimiser plotted on the right  first passes some bases with higher index value during the interpolation.
-If the bases on the interpolation path can be incorporated into the optimisation, we may be able to find the optimal basis quicker.
+Another interest of diagnosing is to examine how the index value progresses for the points on the interpolation path since the projection on these bases are played by the tour animation.
+Trace plots are created with plotting the index value of the interpolation points against time.
+Figure \ref{fig:toy-interp} presents the trace plot of the two optimisation routines as \ref{fig:toy-search}.
+The trace plot is smooth in both plots and this is a character of the holes index function used.
+An interesting discovery is that with optimiser SAJO, the interpolator, in iteration 6-8, first passes through some bases with higher index value than the anchor bases.
 
@@
 
-```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
+```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
 All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
@@
 
-Apart from checking the progression of an optimiser, another interesting aspect of the diagnostics is to visualise how the visited bases looks like in its parameter space.
-Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and even with 1D projection, the basis space is high-dimensional.
-Dimension reduction methods, i.e.
-principal component analysis can be used to present the bases on 2D space with appropriate annotation to highlight the different key components of the optimisation.
+Apart from checking the progression of an optimiser, looking at how the points evaluated by the optimisers look like in the space is another interest.
+Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and  dimension reduction methods, i.e.
+principal component analysis is applied first to project all the bases onto a 2D space.
+In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The starting basis; 2) The anchor bases that have the highest index value in each iteration; 3) The search bases that an optimiser evaluated to produce the anchor bases; 4) The interpolating bases on the interpolation path; and finally 5) the ending basis.
+The importance of these basis differs with the most important ones being the start, interpolating and ending bases.
+Anchor and search bases can be turned on with argument `details = TRUE`.
 
-In a projection pursuit guided tour optimisation, there are 7 different components that are worth-noticing in the exploration: 1) The spherical space of all the basses; 2) The starting basis of an optimisation routine; 3) The anchor points where a new target basis is found; 4) The search basis that an optimiser has evaluated; 5) The interpolation path that constructed by the guided tour; 6) The annotation of the interrupted path and 7) The theoretical best basis, if applicable.
-Despite all have to be displayed on the diagnostic plot, these components differ in their importance with the most interesting ones being the starting point, anchor points and the interpolation path.
+Several annotations have been made to facilitate better reading of this plot.
 

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.tex
text editing on search, trace and pca plot WIP

@@
 
-In pseudo derivative \citep{cook1995grand} method, the search direction
+In pseudo derivative search \citep{cook1995grand}, the search direction
 is computed using the most prominent direction that deviating an tiny
@@
 
-A primary interest of diagnosing an optimiser is to study how it
-progressively finds its optimum. A simple treatment of plotting the
-index value across its natural order will cause the graph to be
-disproportional to the iteration since it usually takes much longer for
-an optimiser to find a better basis towards the end. Another option is
-to use summarisation in each iteration. Boxplot is a suitable candidate
-that can provide five points summary of each iteration and additional
-information can be separately added with new layers, for example, text
-information on the number of points searched in each iteration is added
-at the bottom of each iteration, the bases returned for interpolation
-are highlighted in larger size and linked, and information regarding the
-last iteration where no basis is returned is turned to grey scale.
-Furthermore, an option to switch from boxplot back to point geometry is
-helpful when the number of observation is small in one iteration and can
-be achieved via the \texttt{cutoff} argument.
+A starting point of diagnosing an optimiser is to understand how many
+search the optimiser has conducted. One may want to simply plot the
+index value of the search points across its natural order. A point
+geometry may work well if each iteration has a similar number of points,
+however, the plot will be disproportionate if some iterations have
+considerably more points than other - it over-emphasizes the iterations
+that have more search points! An alternative is to summarise the search
+in each iteration using boxplot where each iteration will be spaced out
+equally. Occasionally, one may still want to switch back to point
+geometry if the number of points is small and this can be achieved via
+the \texttt{cutoff} argument. Additional annotation can be added with
+new layers, for example, the number of points searched in each iteration
+can be added as text label at the bottom of each iteration; the anchor
+bases to interpolate are connected and highlighted in a larger size; and
+the color of the last iteration is in a grey scale to indicate no better
+basis found in this iteration.
 
 Figure \ref{fig:toy-search} shows the searches of two different
-optimisers: SA (left) and SAJO (right). Both optimisers progress quickly
-at the first few iterations, take longer to find better basis in the
-later iterations, and finish off when hitting the maximum number of try,
-499 and 999, respectively. Also, the target basis found on the left
-panel always has an increased index value in each iteration while this
-is not the case for the plot on the right. This explains why in this
-example, the optimiser SAJO takes longer to find the final basis, but in
-more complicated scenarios, the feature of probabilistic acceptance
-allows a more holistic search of the basis space and is more likely to
-find the global optimum.
+optimisers: SA (left) and SAJO (right). Both optimisers quickly find
+better bases at the first few iterations, then take longer in the later
+iterations, and finally finish when the maximum number of evaluation is
+reached. The anchor bases, the ones with the highest index value in each
+iteration, always have an increased index value in the optimiser SA
+while this is not the case for SAJO. This feature gives SA an advantage
+in simple example, like the one presented while the jump out component
+in SAJO allows for a more holistic search in more complicated examples.
 
@@
 
-\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \code{boa6} with holes index on a 2D problem.
-Both optimisers finish when the maximum number of try is reached and attain similar index value in the end.
-Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem.}\label{fig:toy-search}
+\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using holes index.
+Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
 \end{widefigure}
@@
 
-\caption[Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure]{Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure.
-The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers.}\label{fig:toy-interp}
+\caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
+Both traces are smooth while the change of index value in each iteration may not be monotinic.}\label{fig:toy-interp}
 \end{widefigure}
@@
 
-Viewing the index value of points on the interpolation path is another
-interest to the analysts since the projection on these bases will be
-played by the tour animation and it provides further information on how
-the index value changes when moving from one target basis to another.
-Figure \ref{fig:toy-interp} presents the index value against time on the
-interpolation path for the two optimisers with same configuration as
-\ref{fig:toy-search}. From Figure \ref{fig:toy-interp}, we further know
-that to reach a target basis with lower index value, the optimiser
-plotted on the right first passes some bases with higher index value
-during the interpolation. If the bases on the interpolation path can be
-incorporated into the optimisation, we may be able to find the optimal
-basis quicker.
+Another interest of diagnosing is to examine how the index value
+progresses for the points on the interpolation path since the projection
+on these bases are played by the tour animation. Trace plots are created
+with plotting the index value of the interpolation points against time.
+Figure \ref{fig:toy-interp} presents the trace plot of the two
+optimisation routines as \ref{fig:toy-search}. The trace plot is smooth
+in both plots and this is a character of the holes index function used.
+An interesting discovery is that with optimiser SAJO, the interpolator,
+in iteration 6-8, first passes through some bases with higher index
+value than the anchor bases.
 
@@
 
-\caption[1D projection on the 5-variable dataset \code{boa5} with two optimisers]{1D projection on the 5-variable dataset \code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
+\caption[1D projection on the 5-variable dataset \code{boa5}5) with two optimisers]{1D projection on the 5-variable dataset \code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
 All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
@@
 
-Apart from checking the progression of an optimiser, another interesting
-aspect of the diagnostics is to visualise how the visited bases looks
-like in its parameter space. Given the orthonormality constraint, the
-space of projection bases \(\mathbf{A}_{p \times d}\) is a
-\(p \times d\) dimension sphere and even with 1D projection, the basis
-space is high-dimensional. Dimension reduction methods, i.e.~principal
-component analysis can be used to present the bases on 2D space with
-appropriate annotation to highlight the different key components of the
-optimisation.
-
-In a projection pursuit guided tour optimisation, there are 7 different
-components that are worth-noticing in the exploration: 1) The spherical
-space of all the basses; 2) The starting basis of an optimisation
-routine; 3) The anchor points where a new target basis is found; 4) The
-search basis that an optimiser has evaluated; 5) The interpolation path
-that constructed by the guided tour; 6) The annotation of the
-interrupted path and 7) The theoretical best basis, if applicable.
-Despite all have to be displayed on the diagnostic plot, these
-components differ in their importance with the most interesting ones
-being the starting point, anchor points and the interpolation path.
+Apart from checking the progression of an optimiser, looking at how the
+points evaluated by the optimisers look like in the space is another
+interest. Given the orthonormality constraint, the space of projection
+bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\) dimension sphere
+and dimension reduction methods, i.e.~principal component analysis is
+applied first to project all the bases onto a 2D space. In a projection
+pursuit guided tour optimisation, there are various type of bases
+involved: 1) The starting basis; 2) The anchor bases that have the
+highest index value in each iteration; 3) The search bases that an
+optimiser evaluated to produce the anchor bases; 4) The interpolating
+bases on the interpolation path; and finally 5) the ending basis. The
+importance of these basis differs with the most important ones being the
+start, interpolating and ending bases. Anchor and search bases can be
+turned on with argument \texttt{details\ =\ TRUE}.
+
+Several annotations have been made to facilitate better reading of this
+plot.
 

### 2021-01-28 scripts/diagnose.R

@@ -15,7 +15,7 @@ p1 <- before %>%
   ggtitle("without interruption") +
   geom_point(data = p1_anno, size = 3) +
   geom_hline(data = get_best(after) %>% mutate(id = 78), aes(yintercept = index_val), color = "grey90") +
-  geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
+  ggrepel::geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
 
 p2 <- after %>%
   explore_trace_interp(accuracy_y = 0.001) +

### 2021-01-28 scripts/toy.R

@@ -1,10 +1,7 @@
 ## ---- load-pkg
 library(ferrn)
 library(tidyverse)
-library(tourr)
-library(gganimate)
 library(patchwork)
-library(ggrepel)
 files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
@@ -39,7 +36,7 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
   explore_space_pca(group = method, details = TRUE,
-                    interp_size = 1, anchor_size = 2, finish_size = 2, ratio = 4.5) +
+                    interp_size = 1, ratio = 7, dir_alpha = 0.5) +
   scale_color_discrete_botanical()
 
 ## ----toy-pca-animated

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.Rmd
text editing till PCA plot

@@
 
-```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using holes index.
+```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
@@
 
-```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
-All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
+```{r toy-pca, fig.height=6, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes indexwith two optimisers: SA and PD.
+All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases."}
 ```
@@
 principal component analysis is applied first to project all the bases onto a 2D space.
-In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The starting basis; 2) The anchor bases that have the highest index value in each iteration; 3) The search bases that an optimiser evaluated to produce the anchor bases; 4) The interpolating bases on the interpolation path; and finally 5) the ending basis.
-The importance of these basis differs with the most important ones being the start, interpolating and ending bases.
+In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The start basis; 2) The anchor bases that have the highest index value in each iteration; 3) The search bases that an optimiser evaluated to produce the anchor bases; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
+The importance of these basis differs with the most important ones being the start, interpolating and end bases.
 Anchor and search bases can be turned on with argument `details = TRUE`.
+Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
+This happens because the projection is invariate to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing end bases.
+A flip sign device is implemented to flip the sign of all the bases in one optimisation if the situation described above happens and to ensure different routines finish close to each other.
 
-Several annotations have been made to facilitate better reading of this plot.
+Several annotations have been made to help understanding this plot.
+The original space of the bases is a high dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
+Along with the bases recorded during the optimisation and a zero basis, PCA is performed to get the first two PC coordinates of all the bases.
+The search space in the 2D space is a circle with the origin being the PC coordinates of the zero matrix, and radius estimated as the largest distance between the origin and all the bases.
+The theoretical best basis is known with simulated data and can be labelled for comparing the how close the ending basis to the theoretical one for different optimisers.
+Various aesthetics, i.e.
+size, alpha and color, are applicable to each component to adjust for the presentation.
 
-Various other components can be down-played via size, transparency, linetype and proper approximation.
-The basis space is generated via random bases on the original high dimensional space via the CRAN package \CRANpkg{geozoo} [@geozoo] and projected down to 2D.
-Displaying the search space with hundreds of randomly generated dots can be dizzy, especially when the more important start, anchor and search bases are also displayed with point geometry.
-Given the fact that the basis space is a circle when projected to 2D, we can estimate the center and radius with all the bases recorded in the data object.
-Comparing to the start and anchor points, search points are less important as they fail to improve on the index value during the optimisation and hence an adjusted size and transparency are applied to them.
-A potential scenario that could happen is that two optimisers can start at the same basis but finishes with bases in opposite signs.
-This sign difference would not make a difference in the projected data but the PCA projection of the bases will display the finish points symmetric to each other, which can be disturbing for comparing the path close to the finish.
-One solution is to flip the sign of one optimiser so that they start at symmetric position but finish close to each other.
-To remind the analyst of this symmetry in the PCA space, a light dashed line is drawn to connect the starting points of two paths.
-The theoretical best basis is available if the data is simulated and can serve as a guide on how the optimisers progress to the final basis.
-Also, the search paths also have an increasing alpha hue from start to finish to indicate the direction of the optimsation when the theoretical guide is not available.
-To allow flexible annotation in different scenarios, a range of arguments are made available to adjust in the main plotting function \code{explore\_space\_pca()}.
-
-Figure \ref{fig:toy-pca} shows an example of the PCA plot on the same optimisers on a 1D projection problem with 5 variables.
+Figure \ref{fig:toy-pca} presents the PCA plot of a 1D projection of a 5-variable dataset, \code{boa5}, using the holes index with two optimisers, PD and SA, being compared.
+Both optimisers get close to the theoretical bases, annotated by "*" and in this particular example pseudo derivative search gets closer.
+One can also appreciate the nature of the two different searches: PD first evaluates points in a small neighbourhood for a promising direction, while SA evaluates points randomly in the search space to search for the next target.
+There are dashed lines annotated for SA and it describes the interruption of the interpolation, which has been briefly mentioned in Figure \ref{fig:toy-interp} and will be discussed in details in Section \ref{monotonic}.
 
@@
 As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
+
+## Forming a torus
 
@@
 The polish search start with this basis and on the right panel, the edges of the clusters are better trimmed.
+
+## fliping the sign in the bases
 

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.tex
text editing till PCA plot

@@
 
-\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using holes index.
+\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
@@
 
-\caption[1D projection on the 5-variable dataset \code{boa5}5) with two optimisers]{1D projection on the 5-variable dataset \code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD).
-All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
+\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
+All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
 \end{figure}
@@
 pursuit guided tour optimisation, there are various type of bases
-involved: 1) The starting basis; 2) The anchor bases that have the
-highest index value in each iteration; 3) The search bases that an
-optimiser evaluated to produce the anchor bases; 4) The interpolating
-bases on the interpolation path; and finally 5) the ending basis. The
-importance of these basis differs with the most important ones being the
-start, interpolating and ending bases. Anchor and search bases can be
-turned on with argument \texttt{details\ =\ TRUE}.
-
-Several annotations have been made to facilitate better reading of this
-plot.
-
-Various other components can be down-played via size, transparency,
-linetype and proper approximation. The basis space is generated via
-random bases on the original high dimensional space via the CRAN package
-\CRANpkg{geozoo} \citep{geozoo} and projected down to 2D. Displaying the
-search space with hundreds of randomly generated dots can be dizzy,
-especially when the more important start, anchor and search bases are
-also displayed with point geometry. Given the fact that the basis space
-is a circle when projected to 2D, we can estimate the center and radius
-with all the bases recorded in the data object. Comparing to the start
-and anchor points, search points are less important as they fail to
-improve on the index value during the optimisation and hence an adjusted
-size and transparency are applied to them. A potential scenario that
-could happen is that two optimisers can start at the same basis but
-finishes with bases in opposite signs. This sign difference would not
-make a difference in the projected data but the PCA projection of the
-bases will display the finish points symmetric to each other, which can
-be disturbing for comparing the path close to the finish. One solution
-is to flip the sign of one optimiser so that they start at symmetric
-position but finish close to each other. To remind the analyst of this
-symmetry in the PCA space, a light dashed line is drawn to connect the
-starting points of two paths. The theoretical best basis is available if
-the data is simulated and can serve as a guide on how the optimisers
-progress to the final basis. Also, the search paths also have an
-increasing alpha hue from start to finish to indicate the direction of
-the optimsation when the theoretical guide is not available. To allow
-flexible annotation in different scenarios, a range of arguments are
-made available to adjust in the main plotting function
-\code{explore\_space\_pca()}.
-
-Figure \ref{fig:toy-pca} shows an example of the PCA plot on the same
-optimisers on a 1D projection problem with 5 variables.
+involved: 1) The start basis; 2) The anchor bases that have the highest
+index value in each iteration; 3) The search bases that an optimiser
+evaluated to produce the anchor bases; 4) The interpolating bases on the
+interpolation path; and finally 5) the end basis. The importance of
+these basis differs with the most important ones being the start,
+interpolating and end bases. Anchor and search bases can be turned on
+with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers can
+start with the same basis but finish with bases of opposite sign. This
+happens because the projection is invariate to the sign difference of
+the bases and so does the index value, however, this creates
+difficulties for comparing end bases. A flip sign device is implemented
+to flip the sign of all the bases in one optimisation if the situation
+described above happens and to ensure different routines finish close to
+each other.
+
+Several annotations have been made to help understanding this plot. The
+original space of the bases is a high dimensional sphere and random
+bases on the sphere can be generated via the CRAN package
+\CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded during
+the optimisation and a zero basis, PCA is performed to get the first two
+PC coordinates of all the bases. The search space in the 2D space is a
+circle with the origin being the PC coordinates of the zero matrix, and
+radius estimated as the largest distance between the origin and all the
+bases. The theoretical best basis is known with simulated data and can
+be labelled for comparing the how close the ending basis to the
+theoretical one for different optimisers. Various aesthetics, i.e.~size,
+alpha and color, are applicable to each component to adjust for the
+presentation.
+
+Figure \ref{fig:toy-pca} presents the PCA plot of a 1D projection of a
+5-variable dataset, \code{boa5}, using the holes index with two
+optimisers, PD and SA, being compared. Both optimisers get close to the
+theoretical bases, annotated by "*" and in this particular example
+pseudo derivative search gets closer. One can also appreciate the nature
+of the two different searches: PD first evaluates points in a small
+neighbourhood for a promising direction, while SA evaluates points
+randomly in the search space to search for the next target. There are
+dashed lines annotated for SA and it describes the interruption of the
+interpolation, which has been briefly mentioned in Figure
+\ref{fig:toy-interp} and will be discussed in details in Section
+\ref{monotonic}.
 
@@
 projection looks similar to the one in Figure \ref{fig:toy-pca}.
+
+\hypertarget{forming-a-torus}{%
+\subsection{Forming a torus}\label{forming-a-torus}}
 
@@
 are better trimmed.
+
+\hypertarget{fliping-the-sign-in-the-bases}{%
+\subsection{fliping the sign in the
+bases}\label{fliping-the-sign-in-the-bases}}
 

### 2021-01-28 scripts/toy.R

@@ -36,7 +36,7 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
   explore_space_pca(group = method, details = TRUE,
-                    interp_size = 1, ratio = 7, dir_alpha = 0.5) +
+                    interp_size = 1, ratio = 5) +
   scale_color_discrete_botanical()
 
 ## ----toy-pca-animated

### 2021-01-29 paper/zhang-cook-laa-langrene-menendez.Rmd
writing on toy examples are now done

@@
 
-## Understanding the optimiser's coverage of the search space
+## Understanding the optimiser's coverage of the search space {#toy-pca}
 
@@
 Various aesthetics, i.e.
-size, alpha and color, are applicable to each component to adjust for the presentation.
+size, alpha and color, are applicable to emphasize the crucial elements and adjust for the presentation.
+For example, anchor points and search points are less important and a smaller size and alpha is used; The interpolation paths are more informative if alpha increases from start to finish (that is, the path is more opaque towards the end).
 
@@
 
-```{r toy-pca-animated, fig.cap = "A selected number of frames from the animated PCA plots.
-With animation, it is easier to track the progression of each optimiser from the start to finish.
+```{r toy-pca-animated, fig.cap = "Six selected frames from the animated PCA plots in the previous figure.
+With animation, the start and end position of each optimisation is easier to identify.
 A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
@@
 
-Animated plots can be informative in diagnostics, especially in the case of PCA plot when the starting and ending of the search is not clear.
-Figure \ref{fig:toy-pca-animated} shows six frames of an animated version of Figure \ref{fig:toy-pca} and this time, it shows that `search_better` finds the optimum quicker than `search_geodesic`.
+Animation is another display to show how the search porgresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation.
+Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
+An additional piece of information one can learn from this animation is that the SA search finds its end basis quicker than the PD search since SA has finished in the 5th frame while PD is still making more progression.
 
@@
 
-```{r toy-tour, fig.cap="A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
+```{r toy-tour, fig.cap="Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+The tour animation allows for an appreciation of the search bases in the original high dimensional space.
 A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
@@
 
-While viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a more holistic stereoscopic view of the search.
-To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation and this can be done via a regular grand tour.
-Compared to the PCA plot, the animated rotation (tour) displayed in Figure \ref{fig:toy-tour} gives a more well-rounded view of the search and one can view the curved region of the tour path from different angles, which may not be presented in the PCA plot.
-Also the grand tour animation encompasses the PCA projection since the rotation from PCA is just one angle that maximises the variance of the bases and the grand tour produces a sequence of angles that view the search from different directions.
-As an evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select from the tour animation that is close to the PCA angle and the projection looks similar to the one in Figure \ref{fig:toy-pca}.
+As mentioned in Section \ref{toy-pca}, the original $p \times d$ dimension space can be simulated via randomly generated bases via \CRANpkg{geozoo} [@geozoo] package.
+A tour plot rotates these random bases and the ones collected by the optimisation in the original space and hence provides a stereoscopic view of the search.
+While the PCA plot projects the bases from the direction that maximises the variance, the tour plot display of the original high dimensional space from various directions using animation.
+Figure \ref{fig:toy-tour} shows the tour plot of the same two optimisations in the 5D space.
 

### 2021-01-29 paper/zhang-cook-laa-langrene-menendez.tex
writing on toy examples are now done

@@
 
-\hypertarget{understanding-the-optimisers-coverage-of-the-search-space}{%
+\hypertarget{toy-pca}{%
 \subsection{Understanding the optimiser's coverage of the search
-space}\label{understanding-the-optimisers-coverage-of-the-search-space}}
+space}\label{toy-pca}}
 
@@
 theoretical one for different optimisers. Various aesthetics, i.e.~size,
-alpha and color, are applicable to each component to adjust for the
-presentation.
+alpha and color, are applicable to emphasize the crucial elements and
+adjust for the presentation. For example, anchor points and search
+points are less important and a smaller size and alpha is used; The
+interpolation paths are more informative if alpha increases from start
+to finish (that is, the path is more opaque towards the end).
 
@@
 
-\caption{A selected number of frames from the animated PCA plots.
-With animation, it is easier to track the progression of each optimiser from the start to finish.
+\caption{Six selected frames from the animated PCA plots in the previous figure.
+With animation, the start and end position of each optimisation is easier to identify.
 A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
@@
 
-Animated plots can be informative in diagnostics, especially in the case
-of PCA plot when the starting and ending of the search is not clear.
-Figure \ref{fig:toy-pca-animated} shows six frames of an animated
-version of Figure \ref{fig:toy-pca} and this time, it shows that
-\texttt{search\_better} finds the optimum quicker than
-\texttt{search\_geodesic}.
+Animation is another display to show how the search porgresses from
+start to finish in the space and an \texttt{animate\ =\ TRUE} argument
+is used to enable the animation. Figure \ref{fig:toy-pca-animated} shows
+six frames from the animation of the PCA plot in Figure
+\ref{fig:toy-pca}. An additional piece of information one can learn from
+this animation is that the SA search finds its end basis quicker than
+the PD search since SA has finished in the 5th frame while PD is still
+making more progression.
 
@@
 
-\caption{A selected number of frames from the tour animation for viewing the 5D space of all the projection bases.
-The tour animation allows for a more holistic view of the full space in high dimensions from different angles.
+\caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+The tour animation allows for an appreciation of the search bases in the original high dimensional space.
 A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
@@
 
-While viewing the bases on the reduced space via PCA shed some lights on
-the space the optimisers have explored, the visualisation on the
-original \(p \times d\) dimension enables a more holistic stereoscopic
-view of the search. To view a high dimensional (\(d \ge 3\)) object on a
-screen, an approach is to play the rotation of the object in animation
-and this can be done via a regular grand tour. Compared to the PCA plot,
-the animated rotation (tour) displayed in Figure \ref{fig:toy-tour}
-gives a more well-rounded view of the search and one can view the curved
-region of the tour path from different angles, which may not be
-presented in the PCA plot. Also the grand tour animation encompasses the
-PCA projection since the rotation from PCA is just one angle that
-maximises the variance of the bases and the grand tour produces a
-sequence of angles that view the search from different directions. As an
-evidence, the last frame in Figure \ref{fig:toy-tour} is a frame select
-from the tour animation that is close to the PCA angle and the
-projection looks similar to the one in Figure \ref{fig:toy-pca}.
+As mentioned in Section \ref{toy-pca}, the original \(p \times d\)
+dimension space can be simulated via randomly generated bases via
+\CRANpkg{geozoo} \citep{geozoo} package. A tour plot rotates these
+random bases and the ones collected by the optimisation in the original
+space and hence provides a stereoscopic view of the search. While the
+PCA plot projects the bases from the direction that maximises the
+variance, the tour plot display of the original high dimensional space
+from various directions using animation. Figure \ref{fig:toy-tour} shows
+the tour plot of the same two optimisations in the 5D space.
 

### 2021-01-29 scripts/toy.R

@@ -2,6 +2,7 @@
 library(ferrn)
 library(tidyverse)
 library(patchwork)
+library(tourr)
 files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
@@ -59,7 +60,7 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
 #   scale_color_botanical(palette = "fern")
 # anim_save(ani, filename = here::here("anim/toy-pca-animated.mp4"))
 
-frames <- c("0001", "0038", "0046", "0079", "0096", "0100")
+frames <- c("0001", "0038", "0046", "0079", "0086", "0100")
 ani <- paste0(here::here("anim/"), "pca/", "pca", frames, ".png")
 rl <- lapply(ani, png::readPNG)
 gl <-  lapply(rl, grid::rasterGrob)

### 2021-01-30 paper/zhang-cook-laa-langrene-menendez.Rmd
section monotonicity

@@
                       cache = TRUE,
-                      fig.align = "center",
+                      fig.align = "centre",
                       fig.height = 3,
@@
 
+```{r }
+# spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
+```
+
 # Introduction
@@
 In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as
-$\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as Langrange multipliers and convex optimization methods including linear and quadratic programming can be used.
+$\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as multipliers and convex optimization methods including linear and quadratic programming can be used.
 
@@
 Occasionally, one may still want to switch back to point geometry if the number of points is small and this can be achieved via the `cutoff` argument.
-Additional annotation can be added with new layers, for example, the number of points searched in each iteration can be added as text label at the bottom of each iteration; the anchor bases to interpolate are connected and highlighted in a larger size; and the color of the last iteration is in a grey scale to indicate no better basis found in this iteration.
+Additional annotation can be added with new layers, for example, the number of points searched in each iteration can be added as text label at the bottom of each iteration; the anchor bases to interpolate are connected and highlighted in a larger size; and the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
 
@@
 
-```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
+```{r toy-search, fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
@@
 
-```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
+```{r toy-interp, fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
 Both traces are smooth while the change of index value in each iteration may not be monotinic."}
@@
 
-```{r toy-pca, fig.height=6, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes indexwith two optimisers: SA and PD.
+```{r toy-pca, fig.height=5, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes indexwith two optimisers: SA and PD.
 All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases."}
@@
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariate to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing end bases.
+This happens because the projection is invariant to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing end bases.
 A flip sign device is implemented to flip the sign of all the bases in one optimisation if the situation described above happens and to ensure different routines finish close to each other.
@@
 Various aesthetics, i.e.
-size, alpha and color, are applicable to emphasize the crucial elements and adjust for the presentation.
+size, alpha and colour, are applicable to emphasize the crucial elements and adjust for the presentation.
 For example, anchor points and search points are less important and a smaller size and alpha is used; The interpolation paths are more informative if alpha increases from start to finish (that is, the path is more opaque towards the end).
@@
 
-Animation is another display to show how the search porgresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation.
+Animation is another display to show how the search progresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation.
 Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
@@
 
+Something to fill here ...
+
 # Diagnosing an optimiser {#application}
 
-For a particular index function, the best algorithm to optimise depends on the character of the index and the data.
-If the index function is smooth and has a single maximum, all of the three algorithms introduced above can find the maximum.
-When multiple optima are present, simulated annealing (SA) may get stuck at a local maximum.
-In the case where the index function is non-smooth, pseudo derivative search may even fail to find the maximum.
-In this section, examples will be presented to outline how the diagnostic plots can be used to compare the performance of optimisers.
+In this section, several examples will be presented to show how the diagnostic plots discover something unexpected in the optimisation algorithm, and guide the implementation of some new features.
 
@@
 Random variables with different distributions have been simulated and the distributional form of each variable is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
-All the variables have been scaled to have an overall unit variance before running the projection pursuit.
+Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance, and `x2` to `x7` are normal mixtures with varied weights and locations.
+All the variables have been scaled to have an overall unit variance before projection pursuit.
 
@@
 
-A non-monotonic interpolation path has already been given in Figure \ref{fig:toy-interp}: to get to the target basis from the current one, the interpolator may have passed a basis with even higher index value.
-This motivates the design of an interruption to check whether higher index value can be reached during the interpolation.
-If such a basis can be found, the optimiser will start the next iteration from that basis instead of the target one.
-Figure \ref{fig:interruption} contrasts the trace with and without the interruption for the simulated annealing optimiser and it can be observed that while the first two interpolation are identical, a better basis has been found than the target one in the third interpolation.
-Rather than starting the next iteration from the "target basis" on Time 62, as shown in the left panel, the interruption, on the right panel, starts the next iteration from the "interpolation basis" on Time 61.
-The interruption has also detected better basis on the fourth interpolation.
-With the interruption, the optimiser is able to find the final basis with a higher index value (0.865 vs.
-0.879).
+An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with higher index value than the target one.
+With optimiser SAJO, a non-monotonic interpolation is justified since the target bases do not necessarily have a higher index value than the current one, while this is not the case for SA.
+The original trace plot for a 2D projection, optimised by SA, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+Hence, an interruption is implemented to stop the interpolation at the basis with the highest index value.
+The right panel of Figure \ref{fig:interruption} presents the trace plot after implementing such interruption and it can be observed that while the first two interpolations are identical, in the third interpolation the basis at time 61 has a higher index value than the target.
+Rather than starting the next iteration from the target basis on Time 65 as shown in the left panel, the optimiser, on the right panel, starts the next iteration from the interpolation basis on Time 61.
+With the interruption, the optimiser is able to find the final basis with a index value higher than the most promising basis when the interruption is not implemented.
 
-```{r interruption, fig.env = "widefigure", fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index optimised by optimiser SA.
-The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
-With this implementation, the optimiser is able to find a better final basis."}
+```{r interruption, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised by optimiser SA."}
 ```
@@
 
-## fliping the sign in the bases
+## Flipping the sign in the bases
 
@@
 
-```{r kol-result, fig.height = 6, fig.cap="1D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by SA and SAJO.
+```{r kol-result, fig.height = 5, fig.cap="1D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by SA and SAJO.
 Two different neighbourhood size of 0.5 and 0.7 is used.
-The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing."}
+The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The colour of the path is based on whether the optimiser finds the global or local optimum, after polishing."}
 ```
@@
   - \code{get\_search\_count()} produces the summary table of the number of observation in each iteration
-  - \code{get\_center()} produces the center point of the basis space estimated by the starting points
-  - \code{get\_space\_param()} produces the coordinates of the center and radius of the basis space
+  - \code{get\_centre()} produces the centre point of the basis space estimated by the starting points
+  - \code{get\_space\_param()} produces the coordinates of the centre and radius of the basis space
   - \code{get\_theo()} extracts the theoretical observations from the data object
@@
   - \code{clean\_method()} for clean up the name of the optimisers
-  - \code{botanical\_palettes} is a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
-  - \code{botanical\_pal()} as the color interpolator 
-  - \code{scale\_color\_botanical()} is a ggplot2 scale for botanical palettes.
+  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants.
+Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+  - \code{botanical\_pal()} as the colour interpolator
 
@@
 The ideas were illustrated using the optimisers available for projection pursuit guided tour.
-Here the constraint is the orthornormality condition of the projection bases.
+Here the constraint is the orthonormality condition of the projection bases.
 The approach used here could be broadly applied to understand other constrained optimisers.

### 2021-01-30 paper/zhang-cook-laa-langrene-menendez.tex
section monotonicity

@@
 hence the problem can be written as \(\min_{x} f(x)\) \emph{subject to}
-\(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex
+\(g_i(x) \leq 0\). Here methods such as multipliers and convex
 optimization methods including linear and quadratic programming can be
@@
 \begin{figure}
-
-{\centering \includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
-
-}
-
-\caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
+\includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path} \caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
@@
 bases to interpolate are connected and highlighted in a larger size; and
-the color of the last iteration is in a grey scale to indicate no better
-basis found in this iteration.
+the colour of the last iteration is in a grey scale to indicate no
+better basis found in this iteration.
 
@@
 \begin{Schunk}
-\begin{widefigure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
-
-}
-
-\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1} \caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
-\end{widefigure}
+\end{figure}
 \end{Schunk}
@@
 \begin{Schunk}
-\begin{widefigure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
-
-}
-
-\caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1} \caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
 Both traces are smooth while the change of index value in each iteration may not be monotinic.}\label{fig:toy-interp}
-\end{widefigure}
+\end{figure}
 \end{Schunk}
@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
-
-}
-
-\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1} \caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
 All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
@@
 start with the same basis but finish with bases of opposite sign. This
-happens because the projection is invariate to the sign difference of
+happens because the projection is invariant to the sign difference of
 the bases and so does the index value, however, this creates
@@
 theoretical one for different optimisers. Various aesthetics, i.e.~size,
-alpha and color, are applicable to emphasize the crucial elements and
+alpha and colour, are applicable to emphasize the crucial elements and
 adjust for the presentation. For example, anchor points and search
@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
-
-}
-
-\caption{Six selected frames from the animated PCA plots in the previous figure.
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1} \caption{Six selected frames from the animated PCA plots in the previous figure.
 With animation, the start and end position of each optimisation is easier to identify.
@@
 
-Animation is another display to show how the search porgresses from
+Animation is another display to show how the search progresses from
 start to finish in the space and an \texttt{animate\ =\ TRUE} argument
@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
-
-}
-
-\caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1} \caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
 The tour animation allows for an appreciation of the search bases in the original high dimensional space.
@@
 
+Something to fill here \ldots{}
+
 \hypertarget{application}{%
@@
 
-For a particular index function, the best algorithm to optimise depends
-on the character of the index and the data. If the index function is
-smooth and has a single maximum, all of the three algorithms introduced
-above can find the maximum. When multiple optima are present, simulated
-annealing (SA) may get stuck at a local maximum. In the case where the
-index function is non-smooth, pseudo derivative search may even fail to
-find the maximum. In this section, examples will be presented to outline
-how the diagnostic plots can be used to compare the performance of
-optimisers.
+In this section, several examples will be presented to show how the
+diagnostic plots discover something unexpected in the optimisation
+algorithm, and guide the implementation of some new features.
 
@@
 \texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
-variance and \texttt{x2} to \texttt{x7} are normal mixtures with varied
+variance, and \texttt{x2} to \texttt{x7} are normal mixtures with varied
 weights and locations. All the variables have been scaled to have an
-overall unit variance before running the projection pursuit.
+overall unit variance before projection pursuit.
 
@@
 
-A non-monotonic interpolation path has already been given in Figure
-\ref{fig:toy-interp}: to get to the target basis from the current one,
-the interpolator may have passed a basis with even higher index value.
-This motivates the design of an interruption to check whether higher
-index value can be reached during the interpolation. If such a basis can
-be found, the optimiser will start the next iteration from that basis
-instead of the target one. Figure \ref{fig:interruption} contrasts the
-trace with and without the interruption for the simulated annealing
-optimiser and it can be observed that while the first two interpolation
-are identical, a better basis has been found than the target one in the
-third interpolation. Rather than starting the next iteration from the
-``target basis'' on Time 62, as shown in the left panel, the
-interruption, on the right panel, starts the next iteration from the
-``interpolation basis'' on Time 61. The interruption has also detected
-better basis on the fourth interpolation. With the interruption, the
-optimiser is able to find the final basis with a higher index value
-(0.865 vs.~0.879).
+An example of non-monotonic interpolation has been given in Figure
+\ref{fig:toy-interp}: a path that passes bases with higher index value
+than the target one. With optimiser SAJO, a non-monotonic interpolation
+is justified since the target bases do not necessarily have a higher
+index value than the current one, while this is not the case for SA. The
+original trace plot for a 2D projection, optimised by SA, is shown on
+the left panel of Figure \ref{fig:interruption} and one can observe
+clearly that the non-monotonic interpolation has undermined the
+optimiser to realise its full potential. Hence, an interruption is
+implemented to stop the interpolation at the basis with the highest
+index value. The right panel of Figure \ref{fig:interruption} presents
+the trace plot after implementing such interruption and it can be
+observed that while the first two interpolations are identical, in the
+third interpolation the basis at time 61 has a higher index value than
+the target. Rather than starting the next iteration from the target
+basis on Time 65 as shown in the left panel, the optimiser, on the right
+panel, starts the next iteration from the interpolation basis on Time
+61. With the interruption, the optimiser is able to find the final basis
+with a index value higher than the most promising basis when the
+interruption is not implemented.
 
 \begin{Schunk}
-\begin{widefigure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
-
-}
-
-\caption[Trace plot of 2D projection on \code{boa6} data with holes index optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index optimised by optimiser SA.
-The purpose of the plot is to compare the effect of implementing the interruption that avoid fully interpolating to the target basis if a better one can be found during the interpolation.
-With this implementation, the optimiser is able to find a better final basis.}\label{fig:interruption}
-\end{widefigure}
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1} \caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA.}\label{fig:interruption}
+\end{figure}
 \end{Schunk}
@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
-
-}
-
-\caption[Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative]{Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative.
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1} \caption[Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative]{Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative.
 The left panel shows the final projected data before polish and the right panel shows the one after.
@@
 
-\hypertarget{fliping-the-sign-in-the-bases}{%
-\subsection{fliping the sign in the
-bases}\label{fliping-the-sign-in-the-bases}}
+\hypertarget{flipping-the-sign-in-the-bases}{%
+\subsection{Flipping the sign in the
+bases}\label{flipping-the-sign-in-the-bases}}
 
@@
 \begin{Schunk}
-\begin{widefigure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
-
-}
-
-\caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index.
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1} \caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index.
 The trace plot shows the failure of pseudo derivative on non-smooth index.
 The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
-\end{widefigure}
+\end{figure}
 \end{Schunk}
@@
 \begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
-
-}
-
-\caption[1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO]{1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO.
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1} \caption[1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO]{1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO.
 Two different neighbourhood size of 0.5 and 0.7 is used.
-The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The color of the path is based on whether the optimiser finds the global or local optimum, after polishing.}\label{fig:kol-result}
+The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The colour of the path is based on whether the optimiser finds the global or local optimum, after polishing.}\label{fig:kol-result}
 \end{figure}
@@
   \item
-    \code{get\_center()} produces the center point of the basis space
+    \code{get\_centre()} produces the centre point of the basis space
     estimated by the starting points
   \item
-    \code{get\_space\_param()} produces the coordinates of the center
+    \code{get\_space\_param()} produces the coordinates of the centre
     and radius of the basis space
@@
   \item
-    \code{botanical\_palettes} is a collection of color palettes from
+    \code{botanical\_palettes} is a collection of colour palettes from
     Australian native plants. Quantitative palettes include daisy,
@@
   \item
-    \code{botanical\_pal()} as the color interpolator
-  \item
-    \code{scale\_color\_botanical()} is a ggplot2 scale for botanical
-    palettes.
+    \code{botanical\_pal()} as the colour interpolator
   \end{itemize}
@@
 using the optimisers available for projection pursuit guided tour. Here
-the constraint is the orthornormality condition of the projection bases.
+the constraint is the orthonormality condition of the projection bases.
 The approach used here could be broadly applied to understand other

### 2021-01-30 scripts/clean_interrupt.R

@@ -1,6 +1,6 @@
 origin <- interrupt_no
 
-before <- origin %>%
+clean_before <- origin %>%
   select(basis, index_val, info, method, tries, loop, id) %>%
   mutate( id = row_number(),
           loop = ifelse(id == 1, 1, loop),
@@ -9,6 +9,17 @@ before <- origin %>%
   group_by(tries, info) %>%
   mutate(loop = ifelse(is.na(loop), row_number(), loop)) %>%
   ungroup()
+
+last <- clean_before %>% filter(info == "new_basis") %>% mutate(info = "interpolation")
+
+before <- bind_rows(clean_before, last) %>%
+  arrange(tries) %>%
+  mutate(search = ifelse(info == "interpolation", "interp", "search")) %>%
+  group_by(tries, search) %>%
+  mutate(loop = row_number()) %>%
+  ungroup() %>%
+  mutate(id = row_number()) %>%
+  dplyr::select(-search)
 
 origin2 <- interrupt_yes
 
@@ -17,10 +28,12 @@ after <- origin2 %>%
   mutate( id = row_number(),
           loop = ifelse(id == 1, 1, loop),
           method = ifelse(id == 1, "search_better", method),
-          tries = ifelse(id == 1, tries, tries + 1)) %>%
-  group_by(tries, info) %>%
-  mutate(loop = ifelse(is.na(loop), row_number(), loop)) %>%
-  ungroup()
+          tries = ifelse(id == 1, tries, tries + 1),
+          search = ifelse(info == "interpolation", "interp", "search")) %>%
+  group_by(tries, search) %>%
+  mutate(loop = row_number()) %>%
+  ungroup() %>%
+  dplyr::select(-search)
 
 
 save(before, file = here::here("data", "before.rda"))

### 2021-01-30 scripts/diagnose.R

@@ -12,13 +12,13 @@ p1_anno <- bind_rows(
 p1 <- before %>%
   explore_trace_interp(accuracy_y = 0.001) +
   scale_color_continuous_botanical() +
-  ggtitle("without interruption") +
-  geom_point(data = p1_anno, size = 3) +
   geom_hline(data = get_best(after) %>% mutate(id = 78), aes(yintercept = index_val), color = "grey90") +
-  ggrepel::geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5)
+  ggrepel::geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5) +
+  ggtitle("without interruption")
 
 p2 <- after %>%
   explore_trace_interp(accuracy_y = 0.001) +
+
   ggtitle("with interruption") +
   scale_color_continuous_botanical()
 

### 2021-01-31 paper/zhang-cook-laa-langrene-menendez.Rmd
polish section done

@@
 
-```{r polish, fig.height= 2.5, fig.cap = "Two-D projection on \\code{boa6} data with holes index optimised by pseudo derivative.
-The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.
-"}
+```{r polish, fig.height= 5, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised first by SA and then by polish search.
+The polish step results in clearer cut on the edges of the four clusters in the data structure."}
 ```
 
-Once the final basis has been found by an optimiser, one may want to push further in the close neighbourhood to find an even better basis.
-A polish search takes the final basis as the start of a new guided tour to search for local breakthrough.
-
-The polish algorithm is similar to the simulated annealing one with three main distinctions on 1) the number of candidate basis generated each time in the inner loop; 2) the search neighbourhood, `alpha`, and 3) the termination conditions.
-In simulated annealing, only one candidate basis $\mathbf{A}_l$ is generated every time while the polish algorithm allows for a specified number of basis to be evaluated for each repetition in the inner loop via an argument `n_sample`.
-The search neighbourhood in simulated annealing is reduced in the outer loop, that is, the alpha for one iteration of search remains the same, while in polish search, `alpha` is reduced once a set of candidate bases fails to make an improvement.
-To avoid the case the neighbourhood size becoming too small to generate meaningful search, more termination criteria have been added, apart from the original `max.tries` condition and  these include:
+Once the final basis has been found by an optimiser, one may want to push further in the close neighbourhood to see if an even better basis can be found.
+A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local breakthrough.
+The polish algorithm is similar to the SA with three main distinctions: 1) 100 rather than one candidate bases are generated each time in the inner loop; 2) the search neighbourhood, `alpha`, is reduced in the inner loop, rather than in the outer loop, once a set of candidate bases fails to make an improvement; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current in terms of the distance in the space,  percentage change in the index value, and neighbourhood:
 
@@
 
-Figure \ref{fig:polish} presents the final projections found before (left) and after(right) applying the polish using the pseudo derivative optimiser.
-Before polishing, the four clusters are clearly shown while the edges of each cluster is relatively unshaved.
-The polish search start with this basis and on the right panel, the edges of the clusters are better trimmed.
+Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by SA and followed by the polish search, along with the projection on the final basis of each algorithm.
+The end basis found by SA reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
+Polish works with this end basis and further pushes the index value to produce even clear structure in the data.
+One can observe that the projection after polishing better distinguishes the four clusters with more trimmed edges, espeically along the vertical axis.
 

### 2021-01-31 paper/zhang-cook-laa-langrene-menendez.tex
polish section done

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1} \caption[Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative]{Two-D projection on \code{boa6} data with holes index optimised by pseudo derivative.
-The left panel shows the final projected data before polish and the right panel shows the one after.
-The separation of the clusters on the y axis becomes sharper after the polish.
-}\label{fig:polish}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1} \caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search.
+The polish step results in clearer cut on the edges of the four clusters in the data structure.}\label{fig:polish}
 \end{figure}
@@
 Once the final basis has been found by an optimiser, one may want to
-push further in the close neighbourhood to find an even better basis. A
-polish search takes the final basis as the start of a new guided tour to
-search for local breakthrough.
-
-The polish algorithm is similar to the simulated annealing one with
-three main distinctions on 1) the number of candidate basis generated
-each time in the inner loop; 2) the search neighbourhood,
-\texttt{alpha}, and 3) the termination conditions. In simulated
-annealing, only one candidate basis \(\mathbf{A}_l\) is generated every
-time while the polish algorithm allows for a specified number of basis
-to be evaluated for each repetition in the inner loop via an argument
-\texttt{n\_sample}. The search neighbourhood in simulated annealing is
-reduced in the outer loop, that is, the alpha for one iteration of
-search remains the same, while in polish search, \texttt{alpha} is
-reduced once a set of candidate bases fails to make an improvement. To
-avoid the case the neighbourhood size becoming too small to generate
-meaningful search, more termination criteria have been added, apart from
-the original \texttt{max.tries} condition and these include:
+push further in the close neighbourhood to see if an even better basis
+can be found. A polish search takes the final basis of an optimiser as
+the start of a new guided tour to search for local breakthrough. The
+polish algorithm is similar to the SA with three main distinctions: 1)
+100 rather than one candidate bases are generated each time in the inner
+loop; 2) the search neighbourhood, \texttt{alpha}, is reduced in the
+inner loop, rather than in the outer loop, once a set of candidate bases
+fails to make an improvement; and 3) three more termination conditions
+have been added to ensure the new basis generated is distinguishable
+from the current in terms of the distance in the space, percentage
+change in the index value, and neighbourhood:
 
@@
 
-Figure \ref{fig:polish} presents the final projections found before
-(left) and after(right) applying the polish using the pseudo derivative
-optimiser. Before polishing, the four clusters are clearly shown while
-the edges of each cluster is relatively unshaved. The polish search
-start with this basis and on the right panel, the edges of the clusters
-are better trimmed.
+Figure \ref{fig:polish} presents the trace plot of a 2D projection,
+optimised by SA and followed by the polish search, along with the
+projection on the final basis of each algorithm. The end basis found by
+SA reveals the four clusters in the data, but the edges of each cluster
+are not clean-cut. Polish works with this end basis and further pushes
+the index value to produce even clear structure in the data. One can
+observe that the projection after polishing better distinguishes the
+four clusters with more trimmed edges, espeically along the vertical
+axis.
 

### 2021-01-31 scripts/diagnose.R

@@ -28,7 +28,7 @@ p2 <- after %>%
 # set.seed(123456)
 # holes_2d_better <-
 #   animate_xy(boa6, tour_path = guided_tour(holes(), d = 2,
-#                                            search_f =  search_better, max.tries = 500),
+#                                            search_f =  search_better, max.tries = 400),
 #              rescale = FALSE)
 # last_basis <- get_best(holes_2d_better)$basis %>% .[[1]]
 # set.seed(123456)
@@ -37,12 +37,15 @@ p2 <- after %>%
 #                                            search_f =  search_polish),
 #              rescale = FALSE, start = last_basis)
 #
+# save(holes_2d_better, file = here::here("data", "holes_2d_better.rda"))
+# save(holes_2d_better_polish, file = here::here("data", "holes_2d_better_polish.rda"))
+#
 # set.seed(123456)
 # render(
 #   boa6,
-#   tour_path = guided_tour(holes(), d = 2, search_f = search_better, max.tries = 500),
+#   tour_path = guided_tour(holes(), d = 2, search_f = search_better, max.tries = 400),
 #   dev = "png",
-#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[6]]),
 #   rescale = FALSE,
 #   frames = 500,
 #   file = here::here("anim","polish", "before%03d.png")
@@ -55,16 +58,28 @@ p2 <- after %>%
 #   boa6,
 #   tour_path =  guided_tour(holes(), d = 2, search_f = search_polish),
 #   dev = "png",
-#   display = display_xy(axes = "bottomleft", verbose = TRUE),
+#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[1]]),
 #   start = last_basis,
 #   rescale = FALSE,
 #   frames = 100,
 #   file = here::here("anim","polish", "after%03d.png")
 # )
-before <- png::readPNG(here::here("anim","polish", "before079.png"))
-after <- png::readPNG(here::here("anim","polish", "after005.png"))
-gl <-  lapply(list(before, after), grid::rasterGrob)
-patchwork::wrap_plots(gl)
+
+p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
+  clean_method() %>%
+  mutate(method = factor(method, levels = c("SA", "polish"))) %>%
[…]

### 2021-01-31 paper/zhang-cook-laa-langrene-menendez.tex
add the plot for the orientation section

@@
 
-\hypertarget{flipping-the-sign-in-the-bases}{%
-\subsection{Flipping the sign in the
-bases}\label{flipping-the-sign-in-the-bases}}
+\hypertarget{reconciling-the-orientation}{%
+\subsection{Reconciling the
+orientation}\label{reconciling-the-orientation}}
+
+Something to fill here \ldots{}
+
+\begin{Schunk}
+\begin{figure}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1} \caption[PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463]{PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463.}\label{fig:flip-sign}
+\end{figure}
+\end{Schunk}
 

### 2021-01-31 scripts/diagnose.R

@@ -81,6 +81,32 @@ p1 +
   inset_element(grid::rasterGrob(before), left = 0.77, bottom = 0.45, right = 0.89, top = 0.67, align_to = "full") +
   inset_element(grid::rasterGrob(after), left = 0.87, bottom = 0.7, right = 1, top = 1, align_to = "full")
 
+## ---- flip-sign
+# set.seed(2463)
+# orientation_corrected <-
+#   animate_dist(boa6,
+#                tour_path = guided_tour(norm_kol(nrow(boa6)), d = 1,
+#                                        search_f = search_better, max.tries = 200, 0.7),
+#                rescale = FALSE)
+#
+# use ggobi/tourr version 0.6.0
+# set.seed(2463)
+# orientation_different <-
+#   animate_dist(boa6,
+#                tour_path = guided_tour(norm_kol(nrow(boa6)), d = 1,
+#                                        search_f = search_better, max.tries = 200, 0.7),
+#                rescale = FALSE)
+#
+# save(orientation_corrected, file = here::here("data", "orientation_corrected.rda"))
+# save(orientation_different, file = here::here("data", "orientation_different.rda"))
+
+bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
+          orientation_different %>% mutate(sign = "original")) %>%
+  explore_space_pca(group = sign, flip = FALSE, start_size = 3) +
+  scale_color_discrete_botanical() +
+  facet_wrap(vars(fct_relevel(sign, c("original", "flipped")))) +
+  theme(legend.position = "none")
+
 ## ---- noisy-better-geo
 ## code for generating kol_1d_geo, kol_1d_better and kol_1d_better_polish
 # set.seed(123456)

### 2021-01-31 paper/zhang-cook-laa-langrene-menendez.Rmd
add the plot for the orientation section
writing on comparing the three for kol_norm

@@
 
-## Flipping the sign in the bases
+## Reconciling the orientation
+
+Something to fill here ...
+
+```{r flip-sign, fig.cap = "PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463."}
+```
 
@@
 
-The index function, up until this point, are all smooth, while this is not the case for all.
-A 1D projection function based on the Kolmogorov test, `norm_kol`, compares the projected data, $\mathbf{Y}_{n \times 1}$ to a randomly generated normal distribution, $y_n$ based on the empirical cumulated distribution function (ECDF).
-Let $F_{.}(u)$ be the ECDF function with the subscript $P$ and $u$ indicating the projected and randomly generated ECDF respectively,  the index is defined as:
+The index function before this section are all smooth, while this is not the case for all.
+A 1D projection function based on the Kolmogorov test, $I^{nk}(n)$, compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the empirical cumulated distribution function (ECDF).
+Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov, $I^{nk}(n)$, index is defined as:
 
-$$\max \left[F_{P}(u) - F_{y}(u)\right]$$
-With a more complex index function, it is interesting to understand 1) if all the three optimisers can reach the optimum, 2) when a local optimum is presented, whether the simulated annealing algorithm can escape that local optimum and find the global one.
+$$I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]$$
+With a non-smooth index function, two research questions are raised to understand
 
-```{r noisy-better-geo, fig.cap = "1D projection on 5\\-variable dataset \\code{boa5} with \\code{norm\\_kol} index.
-The trace plot shows the failure of pseudo derivative on non-smooth index.
-The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index."}
+1) if any optimiser fails to optimise a non-smooth index; and 
+2) whether the optimisers can find the global optimum despite the presence of a local optimum
+
+```{r noisy-better-geo, fig.height = 6, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimiers.
+Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
 ```
 
-Figure \ref{fig:noisy-better-geo} presents the tracing plots of two optimisers: pseudo derivative and simulated annealing and as expected, the interpolated path is no longer smooth in either case.
-There is barely any improvement on the index value in the pseudo derivative algorithm while the simulated annealing algorithm has managed to get close to the index value of the theoretical best, indicated by the horizontal dashed line.
-In the PCA plot, simulated annealing is able to move gradually towards the theoretical best while the pseudo derivative method breaks after the initial few attempts.
+Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers.
+As expected, none of the interpolated path is smooth.
+There is barely any improvement made by PD, indicating its failure in optimising non-smooth index.
+While SA and SAJO have managed to get close to the index value of the theoretical best, trace plot shows that it takes SAJO much longer to interpolate towards the final basis.
+This long interpolation path is partially due to the fluctuation in the early iterations, where the SAJO tends to generously accept inferior basis before concentrating near the optimum.
+The PCA plot shows the interpolation path and search points excluding the last termination iteration.
+[A sentence for PD].
+Comparing the amount of random search done by SA and SAJO, it is surprising that although SAJO doesn't actually carry that many of search as SA.
+Combine the insight from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement will be made, while SAJO first broadly accepts basis in the space and only starts the extensive search in a narrowed subspace.
+The specific decision of which optimiser to use will depends on the index curve in the basis space but if the basis space is highly non-smooth and spike, accepting inferior bases at first, as SAJO, can lead to a more efficient search, in terms of the overall number of points evaluated.
 

### 2021-01-31 paper/zhang-cook-laa-langrene-menendez.tex
writing on comparing the three for kol_norm

@@
 
-The index function, up until this point, are all smooth, while this is
-not the case for all. A 1D projection function based on the Kolmogorov
-test, \texttt{norm\_kol}, compares the projected data,
-\(\mathbf{Y}_{n \times 1}\) to a randomly generated normal distribution,
-\(y_n\) based on the empirical cumulated distribution function (ECDF).
-Let \(F_{.}(u)\) be the ECDF function with the subscript \(P\) and \(u\)
-indicating the projected and randomly generated ECDF respectively, the
-index is defined as:
-
-\[\max \left[F_{P}(u) - F_{y}(u)\right]\] With a more complex index
-function, it is interesting to understand 1) if all the three optimisers
-can reach the optimum, 2) when a local optimum is presented, whether the
-simulated annealing algorithm can escape that local optimum and find the
-global one.
+The index function before this section are all smooth, while this is not
+the case for all. A 1D projection function based on the Kolmogorov test,
+\(I^{nk}(n)\), compares the projected data, \(\mathbf{Y}_{n \times 1}\),
+to a randomly generated normal distribution,
+\(\mathcal{N}_{n \times 1}\), based on the empirical cumulated
+distribution function (ECDF). Let \(F_{.}(n)\) be the ECDF function,
+with two possible subscripts \(Y\) and \(\mathcal{N}\) representing the
+projected and randomly generated data, and \(n\) denoting the number of
+observation, the Normal Kolmogorov, \(I^{nk}(n)\), index is defined as:
+
+\[I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
+non-smooth index function, two research questions are raised to
+understand
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\tightlist
+\item
+  if any optimiser fails to optimise a non-smooth index; and
+\item
+  whether the optimisers can find the global optimum despite the
+  presence of a local optimum
+\end{enumerate}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1} \caption[1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index]{1D projection on 5\-variable dataset \code{boa5} with \code{norm\_kol} index.
-The trace plot shows the failure of pseudo derivative on non-smooth index.
-The PCA plot also shows that simulated annealing is searching the area close to the theoretical best while pseudo derivative can't couple with the non-continuity in the index.}\label{fig:noisy-better-geo}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1} \caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers.
+Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 
-Figure \ref{fig:noisy-better-geo} presents the tracing plots of two
-optimisers: pseudo derivative and simulated annealing and as expected,
-the interpolated path is no longer smooth in either case. There is
-barely any improvement on the index value in the pseudo derivative
-algorithm while the simulated annealing algorithm has managed to get
-close to the index value of the theoretical best, indicated by the
-horizontal dashed line. In the PCA plot, simulated annealing is able to
-move gradually towards the theoretical best while the pseudo derivative
-method breaks after the initial few attempts.
+Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
+all three optimisers. As expected, none of the interpolated path is
+smooth. There is barely any improvement made by PD, indicating its
+failure in optimising non-smooth index. While SA and SAJO have managed
+to get close to the index value of the theoretical best, trace plot
+shows that it takes SAJO much longer to interpolate towards the final
+basis. This long interpolation path is partially due to the fluctuation
+in the early iterations, where the SAJO tends to generously accept
+inferior basis before concentrating near the optimum. The PCA plot shows
+the interpolation path and search points excluding the last termination
+iteration. {[}A sentence for PD{]}. Comparing the amount of random
+search done by SA and SAJO, it is surprising that although SAJO doesn't
+actually carry that many of search as SA. Combine the insight from both
+the trace and PCA plot, one can learn the two different search
+strategies by SA and SAJO: SA frequently samples in the space and only
+make a move when an improvement will be made, while SAJO first broadly
+accepts basis in the space and only starts the extensive search in a
+narrowed subspace. The specific decision of which optimiser to use will
+depends on the index curve in the basis space but if the basis space is
+highly non-smooth and spike, accepting inferior bases at first, as SAJO,
+can lead to a more efficient search, in terms of the overall number of
+points evaluated.
 

### 2021-01-31 scripts/diagnose.R

@@ -113,38 +113,51 @@ bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
 # kol_1d_geo <-
 #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
 #                                              search_f =  search_geodesic, max.tries = 200),
-#                rescale = FALSE, verbose = TRUE)
+#                rescale = FALSE)
 #
 # set.seed(123456)
 # kol_1d_better <-
 #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
 #                                              search_f =  search_better, max.tries = 200),
-#                rescale = FALSE, verbose = TRUE)
+#                rescale = FALSE)
 #
-# last <- get_best(kol_1d_better)$basis %>% .[[1]]
+# last_basis <- get_best(kol_1d_better)$basis %>% .[[1]]
 #
 # set.seed(123456)
 # kol_1d_better_polish <-
 #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
-#                                              search_f =  search_polish, max.tries = 100),
-#                rescale = FALSE, verbose = TRUE, start = last)
+#                                              search_f = search_polish, max.tries = 200,
+#                                              start = last_basis),
+#                rescale = FALSE)
+# set.seed(123456)
+# kol_1d_better_random <-
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+#                                              search_f =  search_better_random, max.tries = 200),
+#                rescale = FALSE)
 #
+# last_basis <- get_best(kol_1d_better_random)$basis %>% .[[1]]
 #
+# set.seed(123456)
+# kol_1d_better_random_polish <-
+#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+#                                              search_f =  search_polish, max.tries = 200,
+#                                              start = last_basis),
+#                rescale = FALSE)
 # save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
 # save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
[…]

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.Rmd
writing on the simulation example

@@
 
-```{r kol-result, fig.height = 5, fig.cap="1D projection of data \\code{boa6} on \\code{norm\\_kol} index,  optimised by SA and SAJO.
-Two different neighbourhood size of 0.5 and 0.7 is used.
-The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The colour of the path is based on whether the optimiser finds the global or local optimum, after polishing."}
+```{r kol-result, fig.height = 5, fig.cap="PCA plot of 1D projection on the 6\\-variable data, \\code{boa6}, with normal kolmogorov index.
+Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
+Local optimum (x), is presented in this experiment, along with the global optimum (*).
+The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing."}
 ```
 
-The next experiment compares the two simulated annealing optimisers, SA and SAJO,  for 20 simulation paths.
-Two search neighbourhood sizes, `alpha`, of 0.5 and 0.7 are also compared to understand the effect of neighbourhood size where a larger value indicating a wider search.
-Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected interpolation paths faceted by optimiser and alpha.
-Several results can be drawn through the comparison: 1) With optimiser SA and a 0.5 neighbourhood search size, despite being the simplest and fastest, the optimiser fails to optimise for three instances where they finish neither near the local nor the global optimum.
-2) With the same optimiser but a larger neighbourhood size of 0.7, more seeds have found the global optimum.
-3) Comparing between the optimiser SA and SAJO with the same search size of 0.5, SAJO finishes closer to the theoretical for the ones that find the local optimum.
-This indicates that it is working harder to examine if there is other optimum in the space but a small search neighbourhood size makes it hard to conduct a holistic search of the whole space.
-4) With optimiser SAJO and a larger neighbourhood size of 0.7, more seeds have found the global optimum.
-Also some finish points are getting closer to the position of the theoretical best, as comparing with the 0.5 neighbourhood size, and this would indicate fewer work for the subsequent polish.
+The next experiment compares the performance of SA and SAJO when a local maximum exists.
+Two search neighbourhood sizes: 0.5 and 0.7, are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
+Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by  optimiser and search size.
+With SA and a search size of 0.5, despite being the simplest and fastest, SA fails to optimise in three instances where they finish neither near the local nor the global optimum.
+With a larger search size of 0.7, none of the seed fail and more seeds find the global maximum.
+Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to improve the final basis found, despite having longer interpolation paths.
+However, the denser path near the local maximum is an indicator that SAJO is working hard to examine if there is any other optimum in the space but the relative small search size has diminished its ability to reach out to the true global maximum.
+With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
+This indicates that SAJO, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the underlying true optimum.
 

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.tex
writing on the simulation example

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1} \caption[1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO]{1D projection of data \code{boa6} on \code{norm\_kol} index,  optimised by SA and SAJO.
-Two different neighbourhood size of 0.5 and 0.7 is used.
-The finish point of each path is larger than the start point and optimums are separately labelled with * representing the global optimum and x the local optimum The colour of the path is based on whether the optimiser finds the global or local optimum, after polishing.}\label{fig:kol-result}
+\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1} \caption[PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index]{PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index.
+Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
+Local optimum (x), is presented in this experiment, along with the global optimum (*).
+The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
 \end{figure}
@@
 
-The next experiment compares the two simulated annealing optimisers, SA
-and SAJO, for 20 simulation paths. Two search neighbourhood sizes,
-\texttt{alpha}, of 0.5 and 0.7 are also compared to understand the
-effect of neighbourhood size where a larger value indicating a wider
-search. Figure \ref{fig:kol-result} shows 80 (20 * 4) PCA projected
-interpolation paths faceted by optimiser and alpha. Several results can
-be drawn through the comparison: 1) With optimiser SA and a 0.5
-neighbourhood search size, despite being the simplest and fastest, the
-optimiser fails to optimise for three instances where they finish
-neither near the local nor the global optimum. 2) With the same
-optimiser but a larger neighbourhood size of 0.7, more seeds have found
-the global optimum. 3) Comparing between the optimiser SA and SAJO with
-the same search size of 0.5, SAJO finishes closer to the theoretical for
-the ones that find the local optimum. This indicates that it is working
-harder to examine if there is other optimum in the space but a small
-search neighbourhood size makes it hard to conduct a holistic search of
-the whole space. 4) With optimiser SAJO and a larger neighbourhood size
-of 0.7, more seeds have found the global optimum. Also some finish
-points are getting closer to the position of the theoretical best, as
-comparing with the 0.5 neighbourhood size, and this would indicate fewer
-work for the subsequent polish.
+The next experiment compares the performance of SA and SAJO when a local
+maximum exists. Two search neighbourhood sizes: 0.5 and 0.7, are
+compared to understand how a large search neighbourhood would affect the
+final basis and the length of the search. Figure \ref{fig:kol-result}
+shows 80 paths simulated using 20 seeds in the PCA plot, faceted by
+optimiser and search size. With SA and a search size of 0.5, despite
+being the simplest and fastest, SA fails to optimise in three instances
+where they finish neither near the local nor the global optimum. With a
+larger search size of 0.7, none of the seed fail and more seeds find the
+global maximum. Comparing SA and SAJO for a search size of 0.5, SAJO
+doesn't seem to improve the final basis found, despite having longer
+interpolation paths. However, the denser path near the local maximum is
+an indicator that SAJO is working hard to examine if there is any other
+optimum in the space but the relative small search size has diminished
+its ability to reach out to the true global maximum. With a larger
+search size, almost all the seeds (16 out of 20) have found the global
+maximum and some final bases are much closer to the theoretical best, as
+compared to the three other cases. This indicates that SAJO, with a
+reasonable large search size, is able to overcome the local optimum and
+optimise close towards the underlying true optimum.
 

### 2021-02-01 scripts/kol_result.R

@@ -9,7 +9,7 @@ clean_pca <- clean %>%
                    raw_data = boa6) %>%
   mutate(seed_sim = paste0(seed, sim),
          alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
-         search_name = ifelse(str_detect(sim, "random"), "search_better_random", "search_better"),
+         optimiser = ifelse(str_detect(sim, "random"), "SAJO", "SA"),
          var_found = as.factor(ifelse(var_found == "V7", "global", "local")),
          var_found = fct_relevel(var_found, "global", after = 1)) %>%
   compute_pca(group = seed_sim) %>%
@@ -18,19 +18,19 @@ clean_pca <- clean %>%
 global_max <- clean_pca %>%
   filter(info == "theoretical") %>%
   filter(index_val == max(index_val)) %>%
-  select(-alpha, -search_name)
+  select(-alpha, -optimiser)
 
 local_max <- clean_pca %>%
   filter(info == "theoretical") %>%
   filter(index_val == min(index_val)) %>%
-  select(-alpha, -search_name)
+  select(-alpha, -optimiser)
 
-clean_pca %>% drop_na() %>%
+clean_pca %>% drop_na() %>% rename(`search size` = alpha) %>%
   explore_space_pca(pca = FALSE, group = seed_sim, color = var_found,
-                    start_size = 1, end_size = 2,
+                    start_size = 1, start_alpha = 0.5, end_size = 2,
                     interp_size = 0.5) +
   add_theo(dt = global_max, theo_size = 20) +
   add_theo(dt = local_max, theo_label = "x", theo_size = 10) +
   scale_color_discrete_botanical() +
-  facet_grid(alpha ~ search_name, labeller = label_both)
+  facet_grid(`search size` ~ optimiser, labeller = label_both)
 

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.Rmd
read through everything before section vis diag

@@
   Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
-The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While many indices have been proposed in the literature, less work has been done on evaluating the performance of the optimisers.
-In this paper, we implement a data collection object for the optimisation in the guided tour and introduce visual diagnostics based on the data object collected.
+The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers.
+In this paper, we implement a data collection object in the optimisation of projeciton pursuit guided tour and introduce visual diagnostics based on the data object collected.
 These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
@@
                       cache = TRUE,
-                      fig.align = "centre",
+                      fig.align = "center",
                       fig.height = 3,
@@
 Presenting information in graphics often unveils information that would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-Task specific tools presented by @li2020visualizing show how visualisation can be used to understand the behaviour of neural network on classification models, but no general visualisation tool available for diagnosing optimisation procedure.
+Task specific tools such as @li2020visualizing show how visualisation can be used to understand the behaviour of neural network on classification models, but no general visualisation tool is available for diagnosing optimisation procedures.
 The work presented in this paper brings visualization tools into optimisation problems with an aim to better understand the performance of the optimisers in practice.
@@
 
-The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational] which is an exploratory data analysis tool that is defined to detect *interesting structures* or features in high-dimensional data through a set of lower-dimensional projections that cover the entire high dimensional space using interpolation methods called tours [@cook2008grand].
-The target of the optimisation is to identify the most *interesting* low-dimensional views of the data given by a corresponding projection matrix.
-The most *interesting* structures are formally defined by a function of projections, called index function which is optimized to uncover the most revelling structures in a high dimensional space [@cook1993projection].
+The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand].
+The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view.
+The interestingness of the structure is defined by the index function, a function of the projection basis.
 
 The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existence of multiple maxima (local and global), the trade off between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non differentiable [@jones1998efficient].
+Examples of those include the existence of multiple optima (local and global), the trade off between computational burden and proximity to the optima,  dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient].
 Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily applied to any other optimization problems.
@@
 Section \ref{optim} provides an overview of optimisation methods, specifically line search methods. 
-Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and introduces three existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and used in different types of diagnostic plots.
-Section \ref{application} shows applications of how these plots can be used to understand and compare different algorithms.
-We also discuss how these insights contribute to modifications that improve the algorithms.
+Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and outlines three existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics.
+A data structure is defined to capture information during the optimisation, and different diagnostic plots are designed with the data collected.
+Section \ref{application} shows applications of how these diagnostic plots can be used to discover interesting aspects of different optimisers and guide improvement to the existing algorithms.
 Finally, Section \ref{implementation} describes the R package: \pkg{ferrn}, that implements the visual diagnostics.
@@
 
-The projection pursuit guided tour combines two different methods in exploratory data analysis, focusing on different aspects.
+Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) in exploratory data analysis.
 Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
 clustering, outliers and skewness) in multivariate data via low dimensions projection.
-The guided tour is using ideas from projection pursuit to define a particular variation in a broader class of data visualisation methods, building on the grand tour approach [@As85].
+Guided tour is one variation of a broader class of data visualisation methods, tour, which visualises high-dimensional data through a series of animated projection.
 
-To define projection pursuit, we first need to establish the notation used.
-Let $\mathbf{X}_{n \times p}$ be the data matrix, with $n$ observations in $p$ dimensions.
-A d-dimensional projection can be seen as a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
-Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ (corresponding to an associated projection matrix $\mathbf{A}$) onto an index value $I$ (QUESTION: Isn't f the index function?).
+Notation is first established before defining projection pursuit.
+Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions.
+A d-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto an scalar.
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
+A number of index functions have been proposed in the literature to detect different data structures [friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
 
-A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] for supervised classification problems, and PDA index [@lee2010projection], which is an extension of the LDA index.
-
-As a general visualisation method, a tour produces animations of high dimensional data via rotations between low dimension planes.
-Different tour types choose these planes differently, for example, a grand tour [@cook2008grand] selects the planes randomly to provide a general overview and a manual tour [@cook1997manual] gradually phases in and out one variable, to understand the contribution of that variable in the projection.
-Guided tour, the main interest of this paper, chooses planes with the aid of projection pursuit to gradually reveal the most interesting projection in the low dimension space.
+As a general visualisation method, tour produces animations of high dimensional data via rotations low dimension planes.
+Different tour types choose these planes differently: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
+Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection.
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
-Intuitively, Figure \ref{fig:tour-path} shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation algorithm, and the white frames interpolate between them.
+Figure \ref{fig:tour-path} shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations  between the blue frames.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
@@
 
-```{r tour-path, out.height="25%",out.width="50%",fig.cap="Each square (frame) represents the projected data with a corresponding basis.
+```{r tour-path, out.height="25%",out.width="50%", fig.cap="Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
@@
 
-The optimisation problem in the tour context is stated as follows: Given a randomly generated starting basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
+The optimisation problem in projection pursuit is stated as follows: Given a randomly start basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
 
@@
 
-\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthogonal matrices.
+\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orhtonormal.
 
-Several features of this optimisation are worth noticing.
-First of all, this is a constrained optimisation problem as the decision variables form the entries of a projection basis, which is required to be orthonormal.
-It is also likely that the objective function may not be differentiable for a constructed index function and in these cases, gradient-based methods may not work well.
-Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
-Lastly, there is also one computational consideration: the optimisation procedure needs to be fast to compute since the tour animation is played in real-time.
+Several features of this optimisation are worth-noticing.
+1) The optimisation is constrained and the space of the bases is a d-dimensional sphere.
+2) The objective function may not be differentiable for a constructed index function.
+3) Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
+4) The optimisation procedure needs to be easy to compute since the tour animation needs to be played in real-time during the optimisation.
 
@@
 
-Below we introduce three line search algorithms in tour: simulated annealing (SA), simulated annealing with jump out (SAJO), and pseudo derivative (PD).
+Three optimisers have been implemented in the \CRANpkg{tourr} [@tourr] package: simulated annealing (SA), simulated annealing with jump out (SAJO), and pseudo derivative (PD).
+Simulated annealing (SA) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
+$\mathbf{A}_{l}$ is then orthonormalised to fulfil the basis constraint.
+If $\mathbf{A}_{l}$ has an index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the optimiser outputs $\mathbf{A}_{l}$ for guided tour to construct an interpolation path.
+The neighbourhood parameter $\alpha$ is adjusted by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$ before the next iteration starts.
+The optimiser terminates when the maximum number of iteration $l_{\max}$ is reached before a better basis can be found.
+The algorithm of SA is summarised in Algorithm \ref{random-search}.
+@posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
+In his proposal, the $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
 
@@
 
-Simulated annealing (SA) is a random search device that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is a randomly generated matrix with the same dimension as $\mathbf{A}_{\text{cur}}$.
-$\mathbf{A}_{l}$ is then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the search terminates and outputs the basis for guided tour to construct an interpolation path.
-The next iteration of search begins after adjusting $\alpha$ by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$.
-The termination condition is when the maximum number of iteration $l_{\max}$ is reached.
-The algorithm of simulated annealing is summarised in Algorithm \ref{random-search}.
-A slightly different cooling scheme has been proposed by @posse1995projection to avoid the search space being reduced too fast.
-A halving parameter $c$ is introduced and $\alpha$ is only adjusted if the last search takes more than $c$ times to find an accepted basis.
+Simulated annealing with jump out (SAJO) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as SA but allows a probabilistic acceptance of a basis with lower index value than the current.
+Given an initial value of $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, SAJO gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+This implementation allows the optimiser to make a move and explore the basis space even the landing basis does have a higher index value and hence enable the optimiser to jump out of a local optimum.
+The algorithm \ref{simulated-annealing} highlights how SAJO differs from SA in the inner loop.
 
@@
 
-Simulated annealing with jump out (SAJO) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process but allows a probabilistic acceptance of a inferior basis with lower index value based on the annealing $T(l)$.
-Given an initial $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, simulated annealing gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
-This implementation allows the algorithm to jump out of a local maximum and enables a more holistic search of the whole parameter space.
-This feature is particularly useful when local maxima are present.
-The algorithm \ref{simulated-annealing} highlights how SAJO differs from SA in the inner loop.
+Pseudo derivative (PD) search [@cook1995grand] uses a different strategy than SA and SAJO.
+Rather than randomly sample the basis space, PD first computes a search direction by evaluating vases close to the current basis.
+A step size is then chosen along the geodesic direction by another optimisation over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.
+Algorithm \ref{search-geodesic} summarises the inner loop of the PD.
 
@@
 \end{algorithm}
-
-In pseudo derivative search [@cook1995grand], the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis.
-The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen.
-The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the percentage improve condition or when $l_{\max}$ is reached.
-Algorithm \ref{search-geodesic} summarises the inner loop of the pseudo derivative search.
 
@@
 
-  - \code{explore\_trace\_search()} produces summary plots, as shown in Figure \ref{fig:toy-search}
-  - \code{explore\_trace\_interp()}produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
-  - \code{explore\_space\_pca()} produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
+  - \code{explore\_trace\_search()} produces summary plots in Figure \ref{fig:toy-search}
+  - \code{explore\_trace\_interp()}produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}
+  - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
 Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
-  - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
+  - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
 
 - \code{get\_*()} extracts and manipulates certain components from the existing data object.
-  - \code{get\_best()} extracts the best basis found in the data object
-  - \code{get\_start()} extracts the starting basis
-  - \code{get\_interp()} extracts the observations in the interpolation
-  - \code{get\_interp\_last()} extracts the end observations of the interpolation in each iteration
-  - \code{get\_anchor()} extracts the target observations found by the optimiser
-  - \code{get\_search()} extracts the search observations evaluated by the optimiser
-  - \code{get\_search\_count()} produces the summary table of the number of observation in each iteration
-  - \code{get\_centre()} produces the centre point of the basis space estimated by the starting points
+  - \code{get\_anchor()} extracts target observations 
+  - \code{get\_basis\_matrix()} flattens all the bases into a matrix
+  - \code{get\_best()} extracts the observation with the highest index value in the data object
+  - \code{get\_dir\_search()} extracts directional search observations for PD search
+  - \code{get\_interp()} extracts interpolated observations
+  - \code{get\_interp\_last()} extracts the ending interpolated observations in each iteration
+  - \code{get\_interrupt()} extracts the ending interpolated observations and the target observations if  the interpolation is interrupted
+  - \code{get\_search()} extracts search observations 
+  - \code{get\_search\_count()} extracts the count of search observations 
   - \code{get\_space\_param()} produces the coordinates of the centre and radius of the basis space
-  - \code{get\_theo()} extracts the theoretical observations from the data object
-  - \code{get\_interrupt()} extract the end point of the interpolation and the target point when an interruption happens
-  - \code{get\_basis\_matrix()}: flattens all the bases into a matrix
+  - \code{get\_start()} extracts the starting observation
+  - \code{get\_theo()} extracts the theoretical best observations, if presented
 
 - \code{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
-  - \code{bind\_theoretical()} incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
-  - \code{bind\_random()} generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object. 
-  - \code{bind\_random\_matrix()} binds the points to the basis matrix.
+  - \code{bind\_theoretical()} binds the theoretical best observation in simulated experiment
+  - \code{bind\_random()} binds randomly generated bases in the projection bases space to the data object
+  - \code{bind\_random\_matrix()} binds randomly generated bases and outputs in a matrix format
 
@@
   - \code{clean\_method()} for clean up the name of the optimisers
-  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants.
-Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
+  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia
   - \code{botanical\_pal()} as the colour interpolator 
+  - \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the colour and fill of the plot
 

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.tex
read through everything before section vis diag

@@
 maximising the projection index strongly influences both the statistical
-and the computational aspects of the procedure.'' While many indices
-have been proposed in the literature, less work has been done on
+and the computational aspects of the procedure.'' While much work has
+been done in proposing indices in the literature, less has been done on
 evaluating the performance of the optimisers. In this paper, we
-implement a data collection object for the optimisation in the guided
-tour and introduce visual diagnostics based on the data object
-collected. These diagnostics and workflows can be applied to a broad
-class of optimisers, to assess their performance. An R package,
+implement a data collection object in the optimisation of projeciton
+pursuit guided tour and introduce visual diagnostics based on the data
+object collected. These diagnostics and workflows can be applied to a
+broad class of optimisers, to assess their performance. An R package,
 \pkg{ferrn}, has been created to implement the diagnostics.
@@
 otherwise not be aware of and provides a more comprehensive
-understanding of the problem at hand. Task specific tools presented by
+understanding of the problem at hand. Task specific tools such as
 \citet{li2020visualizing} show how visualisation can be used to
 understand the behaviour of neural network on classification models, but
-no general visualisation tool available for diagnosing optimisation
-procedure. The work presented in this paper brings visualization tools
+no general visualisation tool is available for diagnosing optimisation
+procedures. The work presented in this paper brings visualization tools
 into optimisation problems with an aim to better understand the
@@
 The focus of this paper is on the optimisation problem arising in the
-projection pursuit guided tour \citep{buja2005computational} which is an
-exploratory data analysis tool that is defined to detect
-\emph{interesting structures} or features in high-dimensional data
-through a set of lower-dimensional projections that cover the entire
-high dimensional space using interpolation methods called tours
-\citep{cook2008grand}. The target of the optimisation is to identify the
-most \emph{interesting} low-dimensional views of the data given by a
-corresponding projection matrix. The most \emph{interesting} structures
-are formally defined by a function of projections, called index function
-which is optimized to uncover the most revelling structures in a high
-dimensional space \citep{cook1993projection}.
+projection pursuit guided tour \citep{buja2005computational}, an
+exploratory data analysis tool used for detecting interesting structures
+in high-dimensional data through a set of lower-dimensional projections
+\citep{cook2008grand}. The goal of the optimisation is to identify the
+projection, characterised by the projection basis, that gives the most
+interesting low-dimensional view. The interestingness of the structure
+is defined by the index function, a function of the projection basis.
 
@@
 tour problem are common to those of optimization in general. Examples of
-those include the existence of multiple maxima (local and global), the
-trade off between computational burden and proximity to the maxima,
-dealing with noisy objective functions that might be non smooth and non
-differentiable \citep{jones1998efficient}. Those are not unique to this
-context and therefore the visualization tools and optimization methods
-presented in this paper can be easily applied to any other optimization
-problems.
+those include the existence of multiple optima (local and global), the
+trade off between computational burden and proximity to the optima,
+dealing with noisy objective functions that might be non-smooth and
+non-differentiable \citep{jones1998efficient}. Those are not unique to
+this context and therefore the visualization tools and optimization
+methods presented in this paper can be easily applied to any other
+optimization problems.
 
@@
 methods. Section \ref{tour} reviews projection pursuit guided tour,
-defines the optimisation problem and introduces three existing
-algorithms. Section \ref{vis-diag} presents the new visual diagnostics.
-A data structure is defined to capture information during the
-optimisation, and used in different types of diagnostic plots. Section
-\ref{application} shows applications of how these plots can be used to
-understand and compare different algorithms. We also discuss how these
-insights contribute to modifications that improve the algorithms.
-Finally, Section \ref{implementation} describes the R package:
-\pkg{ferrn}, that implements the visual diagnostics.
+defines the optimisation problem and outlines three existing algorithms.
+Section \ref{vis-diag} presents the new visual diagnostics. A data
+structure is defined to capture information during the optimisation, and
+different diagnostic plots are designed with the data collected. Section
+\ref{application} shows applications of how these diagnostic plots can
+be used to discover interesting aspects of different optimisers and
+guide improvement to the existing algorithms. Finally, Section
+\ref{implementation} describes the R package: \pkg{ferrn}, that
+implements the visual diagnostics.
 
@@
 
-The projection pursuit guided tour combines two different methods in
-exploratory data analysis, focusing on different aspects. Projection
-pursuit, coined by \citet{friedman1974projection}, detects interesting
-structures (e.g.~clustering, outliers and skewness) in multivariate data
-via low dimensions projection. The guided tour is using ideas from
-projection pursuit to define a particular variation in a broader class
-of data visualisation methods, building on the grand tour approach
-\citep{As85}.
-
-To define projection pursuit, we first need to establish the notation
-used. Let \(\mathbf{X}_{n \times p}\) be the data matrix, with \(n\)
-observations in \(p\) dimensions. A d-dimensional projection can be seen
-as a linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\),
-and defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
+Projection pursuit guided tour combines two different methods
+(projection pursuit and guided tour) in exploratory data analysis.
+Projection pursuit, coined by \citet{friedman1974projection}, detects
+interesting structures (e.g.~clustering, outliers and skewness) in
+multivariate data via low dimensions projection. Guided tour is one
+variation of a broader class of data visualisation methods, tour, which
+visualises high-dimensional data through a series of animated
+projection.
+
+Notation is first established before defining projection pursuit. Let
+\(\mathbf{X}_{n \times p}\) be the data matrix with \(n\) observations
+in \(p\) dimensions. A d-dimensional projection is a linear
+transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\), and defined
+as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{Y}_{n \times d}\) is the projected data and
-\(\mathbf{A}_{p\times d}\) is the projection matrix. Define
+\(\mathbf{A}_{p\times d}\) is the projection basis Define
 \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
-function that maps the projected data \(\mathbf{Y}\) (corresponding to
-an associated projection matrix \(\mathbf{A}\)) onto an index value
-\(I\) (QUESTION: Isn't f the index function?). This is commonly known as
-the projection pursuit index function, or just index function, and is
-used to measure the ``interestingness'' of a given projection.
-
-A number of index functions have been proposed in the literature to
-detect different data structures, including Legendre index
-\citep{friedman1974projection}, Hermite index
-\citep{hall1989polynomial}, natural Hermite index
-\citep{cook1993projection}, chi-square index
-\citep{posse1995projection}, LDA index \citep{lee2005projection} for
-supervised classification problems, and PDA index
-\citep{lee2010projection}, which is an extension of the LDA index.
-
-As a general visualisation method, a tour produces animations of high
-dimensional data via rotations between low dimension planes. Different
-tour types choose these planes differently, for example, a grand tour
-\citep{cook2008grand} selects the planes randomly to provide a general
-overview and a manual tour \citep{cook1997manual} gradually phases in
-and out one variable, to understand the contribution of that variable in
-the projection. Guided tour, the main interest of this paper, chooses
-planes with the aid of projection pursuit to gradually reveal the most
-interesting projection in the low dimension space. Given a random start,
-projection pursuit iteratively finds bases with higher index values and
-the guided tour constructs the geodesic interpolation between these
-planes to form a tour path. Intuitively, Figure \ref{fig:tour-path}
-shows a sketch of the tour path where the blue frames are produced by
-the projection pursuit optimisation algorithm, and the white frames
-interpolate between them. Mathematical details of the geodesic
-interpolation can be found in \citet{buja2005computational}. The tour
-method has been implemented in the R package \CRANpkg{tourr}
-\citep{tourr}.
+function that maps the projected data \(\mathbf{Y}\) onto an scalar.
+This is commonly known as the projection pursuit index function, or just
+index function, and is used to measure the ``interestingness'' of a
+given projection. A number of index functions have been proposed in the
+literature to detect different data structures
+\citep[friedman1974projection;][]{hall1989polynomial, cook1993projection, posse1995projection}
+and to classify groups \citep{lee2005projection, lee2010projection}.
+
+As a general visualisation method, tour produces animations of high
+dimensional data via rotations low dimension planes. Different tour
+types choose these planes differently: grand tour \citep{cook2008grand}
+selects the planes randomly to provide a general overview; manual tour
+\citep{cook1997manual} gradually phases in and out one variable to
+understand the contribution of that variable in the projection. Guided
+tour, the main interest of this paper, chooses the planes with the aid
+of projection pursuit, to gradually reveal the most interesting
+projection. Given a random start, projection pursuit iteratively finds
+bases with higher index values and the guided tour constructs the
+geodesic interpolation between these planes to form a tour path. Figure
+\ref{fig:tour-path} shows a sketch of the tour path where the blue
+frames are produced by the projection pursuit optimisation and the white
+frames are interpolations between the blue frames. Mathematical details
+of the geodesic interpolation can be found in
+\citet{buja2005computational}. The tour method has been implemented in
+the R package \CRANpkg{tourr} \citep{tourr}.
 
@@
 \begin{figure}
-\includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path} \caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
+
+{\centering \includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
+
+}
+
+\caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
@@
 
-The optimisation problem in the tour context is stated as follows: Given
-a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit
-finds the final projection basis \(\mathbf{A}_T\) that satisfies the
-following optimisation problem:
+The optimisation problem in projection pursuit is stated as follows:
+Given a randomly start basis \(\mathbf{A}_1\), projection pursuit finds
+the final projection basis \(\mathbf{A}_T\) that satisfies the following
+optimisation problem:
 
@@
 \noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
-constraint requires the projection bases \(\mathbf{A}\) to be orthogonal
-matrices.
-
-Several features of this optimisation are worth noticing. First of all,
-this is a constrained optimisation problem as the decision variables
-form the entries of a projection basis, which is required to be
-orthonormal. It is also likely that the objective function may not be
-differentiable for a constructed index function and in these cases,
-gradient-based methods may not work well. Although finding the global
-maximum is the goal of an optimisation problem, it is also interesting
-to inspect local maximum in projection pursuit since it could present
-unexpected interesting projections. Lastly, there is also one
-computational consideration: the optimisation procedure needs to be fast
-to compute since the tour animation is played in real-time.
+constraint requires the projection bases \(\mathbf{A}\) to be
+orhtonormal.
+
+Several features of this optimisation are worth-noticing. 1) The
+optimisation is constrained and the space of the bases is a
+d-dimensional sphere. 2) The objective function may not be
+differentiable for a constructed index function. 3) Although finding the
+global maximum is the goal of an optimisation problem, it is also
+interesting to inspect local maximum in projection pursuit since it
+could present unexpected interesting projections. 4) The optimisation
+procedure needs to be easy to compute since the tour animation needs to
+be played in real-time during the optimisation.
 
@@
 
-Below we introduce three line search algorithms in tour: simulated
-annealing (SA), simulated annealing with jump out (SAJO), and pseudo
-derivative (PD).
+Three optimisers have been implemented in the \CRANpkg{tourr}
+\citep{tourr} package: simulated annealing (SA), simulated annealing
+with jump out (SAJO), and pseudo derivative (PD). Simulated annealing
+(SA) is a random search optimiser that samples a candidate basis
+\(\mathbf{A}_{l}\) in the neighbourhood of the current basis
+\(\mathbf{A}_{\text{cur}}\) by
+\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
+where \(\alpha\) controls the radius of the sampling neighbourhood and
+\(\mathbf{A}_{\text{rand}}\) is generated randomly. \(\mathbf{A}_{l}\)
+is then orthonormalised to fulfil the basis constraint. If
+\(\mathbf{A}_{l}\) has an index value higher than the current basis
+\(\mathbf{A}_{\text{cur}}\), the optimiser outputs \(\mathbf{A}_{l}\)
+for guided tour to construct an interpolation path. The neighbourhood
+parameter \(\alpha\) is adjusted by a cooling parameter:
+\(\alpha_{j+1} = \alpha_j * \text{cooling}\) before the next iteration
+starts. The optimiser terminates when the maximum number of iteration
+\(l_{\max}\) is reached before a better basis can be found. The
+algorithm of SA is summarised in Algorithm \ref{random-search}.
+\citet{posse1995projection} has proposed a slightly different cooling
+scheme by introducing a halving parameter \(c\). In his proposal, the
+\(\alpha\) is only adjusted if the last iteration takes more than \(c\)
+times to find a better basis.
 
@@
 
-Simulated annealing (SA) is a random search device that samples a
-candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
-basis \(\mathbf{A}_{\text{cur}}\) by
-\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
-where \(\alpha\) controls the radius of the sampling neighbourhood and
-\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
-same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
-then orthogonalised to ensure the orthonormal constraint is fulfilled.
-When a basis is found with index value higher than the current basis
-\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
-for guided tour to construct an interpolation path. The next iteration
-of search begins after adjusting \(\alpha\) by a cooling parameter:
-\(\alpha_{j+1} = \alpha_j * \text{cooling}\). The termination condition
-is when the maximum number of iteration \(l_{\max}\) is reached. The
-algorithm of simulated annealing is summarised in Algorithm
-\ref{random-search}. A slightly different cooling scheme has been
-proposed by \citet{posse1995projection} to avoid the search space being
-reduced too fast. A halving parameter \(c\) is introduced and \(\alpha\)
-is only adjusted if the last search takes more than \(c\) times to find
-an accepted basis.
+Simulated annealing with jump out (SAJO)
+\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
+same sampling process as SA but allows a probabilistic acceptance of a
+basis with lower index value than the current. Given an initial value of
+\(T_0\), the temperature at iteration \(l\) is defined as
+\(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to have
+an index value larger than the current basis, SAJO gives it a second
+chance to be accepted with probability
+\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
+where \(I_{(\cdot)}\) denotes the index value of a given basis. This
+implementation allows the optimiser to make a move and explore the basis
+space even the landing basis does have a higher index value and hence
+enable the optimiser to jump out of a local optimum. The algorithm
+\ref{simulated-annealing} highlights how SAJO differs from SA in the
+inner loop.
 
@@
 
-Simulated annealing with jump out (SAJO)
-\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
-same sampling process but allows a probabilistic acceptance of a
-inferior basis with lower index value based on the annealing \(T(l)\).
-Given an initial \(T_0\), the temperature at iteration \(l\) is defined
-as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to
-have an index value larger than the current basis, simulated annealing
-gives it a second chance to be accepted with probability
-\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
-where \(I_{(\cdot)}\) denotes the index value of a given basis. This
-implementation allows the algorithm to jump out of a local maximum and
-enables a more holistic search of the whole parameter space. This
-feature is particularly useful when local maxima are present. The
-algorithm \ref{simulated-annealing} highlights how SAJO differs from SA
-in the inner loop.
+Pseudo derivative (PD) search \citep{cook1995grand} uses a different
+strategy than SA and SAJO. Rather than randomly sample the basis space,
+PD first computes a search direction by evaluating vases close to the
+current basis. A step size is then chosen along the geodesic direction
+by another optimisation over an 90 degree angle from \(-\pi/4\) to
+\(\pi/4\). The resulting candidate basis \(\mathbf{A}_{**}\) is returned
+for the current iteration if it has a higher index value than the
+current one. Algorithm \ref{search-geodesic} summarises the inner loop
+of the PD.
 
@@
 \end{algorithm}
-
-In pseudo derivative search \citep{cook1995grand}, the search direction
-is computed using the most prominent direction that deviating an tiny
-angle of \(\delta\) from the current basis. The step size is chosen by
-optimising the index value along the geodesic direction over an 90
-degree angle from \(-\pi/4\) to \(\pi/4\) along the search direction
-chosen. The optima \(\mathbf{A}_{**}\) is returned for the current
-iteration if it meets the percentage improve condition or when
-\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
-inner loop of the pseudo derivative search.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1} \caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+
+}
+
+\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1} \caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+
+}
+
+\caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
 Both traces are smooth while the change of index value in each iteration may not be monotinic.}\label{fig:toy-interp}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1} \caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+
+}
+
+\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
 All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1} \caption{Six selected frames from the animated PCA plots in the previous figure.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+
+}
+
+\caption{Six selected frames from the animated PCA plots in the previous figure.
 With animation, the start and end position of each optimisation is easier to identify.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1} \caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+
+}
+
+\caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
 The tour animation allows for an appreciation of the search bases in the original high dimensional space.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1} \caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA.}\label{fig:interruption}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+
+}
+
+\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA.}\label{fig:interruption}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1} \caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+
+}
+
+\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search.
 The polish step results in clearer cut on the edges of the four clusters in the data structure.}\label{fig:polish}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1} \caption[PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463]{PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463.}\label{fig:flip-sign}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+
+}
+
+\caption[PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463]{PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463.}\label{fig:flip-sign}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1} \caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+
+}
+
+\caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers.
 Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1} \caption[PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index]{PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index.
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+
+}
+
+\caption[PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index]{PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index.
 Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
@@
   \item
-    \code{explore\_trace\_search()} produces summary plots, as shown in
-    Figure \ref{fig:toy-search}
+    \code{explore\_trace\_search()} produces summary plots in Figure
+    \ref{fig:toy-search}
   \item
     \code{explore\_trace\_interp()}produces trace plots for the
-    interpolation points, as shown in Figure \ref{fig:toy-interp}
-  \item
-    \code{explore\_space\_pca()} produces plots of projection basis on
-    the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
-    Animated version in Figure \ref{fig:toy-pca-animated} can be turned
-    on via the argument \texttt{animate\ =\ TRUE}
+    interpolation points in Figure \ref{fig:toy-interp}
+  \item
+    \code{explore\_space\_pca()} produces the PCA plot of projection
+    bases on the reduced space in Figure \ref{fig:toy-pca}. Animated
+    version in Figure \ref{fig:toy-pca-animated} can be turned on via
+    the argument \texttt{animate\ =\ TRUE}
   \item
     \code{explore\_space\_tour()} produces animated tour view on the
-    full space of the projection bases, as shown in Figure
-    \ref{fig:toy-tour}.
+    full space of the projection bases in Figure \ref{fig:toy-tour}.
   \end{itemize}
@@
   \item
-    \code{get\_best()} extracts the best basis found in the data object
-  \item
-    \code{get\_start()} extracts the starting basis
-  \item
-    \code{get\_interp()} extracts the observations in the interpolation
-  \item
-    \code{get\_interp\_last()} extracts the end observations of the
-    interpolation in each iteration
-  \item
-    \code{get\_anchor()} extracts the target observations found by the
-    optimiser
-  \item
-    \code{get\_search()} extracts the search observations evaluated by
-    the optimiser
-  \item
-    \code{get\_search\_count()} produces the summary table of the number
-    of observation in each iteration
-  \item
-    \code{get\_centre()} produces the centre point of the basis space
-    estimated by the starting points
+    \code{get\_anchor()} extracts target observations
+  \item
+    \code{get\_basis\_matrix()} flattens all the bases into a matrix
+  \item
+    \code{get\_best()} extracts the observation with the highest index
+    value in the data object
+  \item
+    \code{get\_dir\_search()} extracts directional search observations
+    for PD search
+  \item
+    \code{get\_interp()} extracts interpolated observations
+  \item
+    \code{get\_interp\_last()} extracts the ending interpolated
+    observations in each iteration
+  \item
+    \code{get\_interrupt()} extracts the ending interpolated
+    observations and the target observations if the interpolation is
+    interrupted
+  \item
+    \code{get\_search()} extracts search observations
+  \item
+    \code{get\_search\_count()} extracts the count of search
+    observations
   \item
@@
   \item
-    \code{get\_theo()} extracts the theoretical observations from the
-    data object
-  \item
-    \code{get\_interrupt()} extract the end point of the interpolation
-    and the target point when an interruption happens
-  \item
-    \code{get\_basis\_matrix()}: flattens all the bases into a matrix
+    \code{get\_start()} extracts the starting observation
+  \item
+    \code{get\_theo()} extracts the theoretical best observations, if
+    presented
   \end{itemize}
@@
   \item
-    \code{bind\_theoretical()} incorporates the best possible basis to
-    the existing data object with the supply of the index function and
-    original data for producing the index value.
-  \item
-    \code{bind\_random()} generates 1000 points on the high dimensional
-    surface of a sphere and binds it to the existing data object and
-    output as a tibble object.
-  \item
-    \code{bind\_random\_matrix()} binds the points to the basis matrix.
+    \code{bind\_theoretical()} binds the theoretical best observation in
+    simulated experiment
+  \item
+    \code{bind\_random()} binds randomly generated bases in the
+    projection bases space to the data object
+  \item
+    \code{bind\_random\_matrix()} binds randomly generated bases and
+    outputs in a matrix format
   \end{itemize}
@@
     \code{botanical\_palettes} is a collection of colour palettes from
-    Australian native plants. Quantitative palettes include daisy,
-    banksia and cherry and sequential palettes contain fern and acacia.
+    Australian native plants Quantitative palettes include daisy,
+    banksia and cherry and sequential palettes contain fern and acacia
   \item
     \code{botanical\_pal()} as the colour interpolator
+  \item
+    \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the
+    colour and fill of the plot
   \end{itemize}

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.Rmd
finish proofreading

@@
 
-```{r tour-path, out.height="25%",out.width="50%", fig.cap="Each square (frame) represents the projected data with a corresponding basis.
+```{r tour-path, out.height="20%",out.width="50%", fig.cap="Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
@@
 \noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orhtonormal.
-
 Several features of this optimisation are worth-noticing.
@@
 
-To be able to make diagnostics on the optimisers, the algorithms need to populate a data structure with key elements of the algorithm.
-When the algorithms run, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object so that it is ready to be supplied to the plotting functions for diagnostics.
+Before making diagnostic plots, information generated by the optimisation need to be collected and a pre-defined data structure solidifies the information being collected and streamlines the subsequent plot-making.
 
@@
 
-Three main elements are recorded for the projection pursuit optimisers: 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation.
-For optimiser SA and SAJO, possible values of the state include `random_search`,  `new_basis`, and `interpolation`.
-pseudo derivative search has a wider variety of state including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
-
-Multiple iterators are also  needed to index the data collected at different levels: $t$ being a unique identifier that prescribes the natural ordering of each observation while $j$ and $l$ being the counter of the outer and inner loop, respectively, in Algorithm \ref{random-search}, \ref{simulated-annealing} and \ref{search-geodesic} above.
-Other parameters of interest recorded include $V_1 = \text{method}$ that tags the name of the optimiser, and $V_2 = \text{alpha}$ that indicates the sampling neighbourhood size.
+Three main pieces of information are recorded for the projection pursuit optimisers: 1) projection bases $\mathbf{A}$, 2) index values $I$, and 3) state $S$.
+For SA and SAJO, possible state include `random_search`,  `new_basis`, and `interpolation`.
+Pseudo derivative (PD) has a wider variety of state including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
+Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation;  $j$ and $l$ are the counter of the outer and inner loop respectively.
+Other parameters of interest recorded include `method` that tags the name of the optimiser, and `alpha` that indicates the sampling neighbourhood size for searching observations.
 A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
@@
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
-Note that there is no output in iteration $J + 1$ since the optimiser cannot find a better basis in the last iteration and the algorithm terminates.
-The final basis found is $A_T$ with the highest index value $I_T$.
+Note that there is no output in iteration $J + 1$ since the optimiser doesn't find a better basis in the last iteration and terminates.
+The final basis found is $A_T$ with index value $I_T$.
 
@@
 
-The construction of diagnostic plots adopts the core concept in ggplot2:  grammar of graphics [@wickham2010layered].
-In grammar of graphics, plots are not produced via calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but via the concept of stacked layers.
-Seeing plots as stacked layers empowers us to composite diagnostic plots with an emphasis on any variable in the data object without the redundancy of creating different commands for the same type of plots that highlights on different variables.
+The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2.
+In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers.
+Seeing plots as stacked layers gives analysts the freedom to composite plots with multiple elements on hands.
 
@@
 
-A starting point of diagnosing an optimiser is to understand how many search the optimiser has conducted.
-One may want to simply plot the index value of the search points across its natural order.
-A point geometry may work well if each iteration has a similar number of points,  however, the plot will be disproportionate if some iterations have considerably more points than other - it over-emphasizes the iterations that have more search points!
-An alternative is to summarise the search in each iteration using boxplot where each iteration will be spaced out equally.
-Occasionally, one may still want to switch back to point geometry if the number of points is small and this can be achieved via the `cutoff` argument.
-Additional annotation can be added with new layers, for example, the number of points searched in each iteration can be added as text label at the bottom of each iteration; the anchor bases to interpolate are connected and highlighted in a larger size; and the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
+A starting point of diagnosing an optimiser is to understand how many search an optimiser has conducted.
+One may want to simply plot the index value of the search points across its natural order, but a point geometry may work well if each iteration has a similar number of points.
+When some iterations have considerably more points than others, using a point geometry over-emphasizes the iterations that have more search points since these iterations will occupy the mast majority of the plot space.
+An alternative is to summarise the search in each iteration using boxplot and each iteration will then be spaced out equally.
+Occasionally, one may still want to switch back to a point geometry if the number of points is small in a particular iteration and this can be achieved via the `cutoff` argument in the search plot function.
+Additional annotations are added to facilitate better reading of the plot and these includes
+1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
 
-Figure \ref{fig:toy-search} shows the searches of two different optimisers: SA (left) and SAJO (right).
-Both optimisers quickly find better bases at the first few iterations,  then take longer in the later iterations, and finally finish when the maximum number of evaluation is reached.
-The anchor bases, the ones with the highest index value in each iteration, always have an increased index value in the optimiser SA while this is not the case for SAJO.
-This feature gives SA an advantage in simple example, like the one presented while the jump out component in SAJO allows for a more holistic search in more complicated examples.
+Figure \ref{fig:toy-search} shows an example of the search plot for SA (left) and SAJO (right).
+Both optimisers quickly find better bases in the first few iterations and then take longer to find one in the later iterations.
+The anchor bases, the ones found with the highest index value in each iteration, always have an increased index value in the optimiser SA while this is not the case for SAJO.
+This feature gives SA an advantage in this simple example to quickly find the optimum.
 
@@
 
-Another interest of diagnosing is to examine how the index value progresses for the points on the interpolation path since the projection on these bases are played by the tour animation.
+Another interest of diagnostic is to examine how the index value improves for interpolating bases since the projection on these bases is played by the tour animation.
 Trace plots are created with plotting the index value of the interpolation points against time.
-Figure \ref{fig:toy-interp} presents the trace plot of the two optimisation routines as \ref{fig:toy-search}.
-The trace plot is smooth in both plots and this is a character of the holes index function used.
-An interesting discovery is that with optimiser SAJO, the interpolator, in iteration 6-8, first passes through some bases with higher index value than the anchor bases.
+Figure \ref{fig:toy-interp} presents the trace plot of the same optimisation routine as Figure \ref{fig:toy-search}.
+The trace plot is smooth and an interesting discovery is that in SAJO, the interpolation in iteration 6-8 is not monotonic.
+Sometimes, this non-monotonicity is acceptable since the target basis does not necessarily have a higher index value than the current one, while in some cases, the optimisation will be better off if these intermediate basis can be used as the start of the next iteration (see section \ref{monotonic} for a detailed discussion on this).
 
@@
 
-```{r toy-pca, fig.height=5, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes indexwith two optimisers: SA and PD.
+```{r toy-pca, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes index with two optimisers: SA and PD.
 All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases."}
@@
 
-Apart from checking the progression of an optimiser, looking at how the points evaluated by the optimisers look like in the space is another interest.
+Apart from checking the search and progression of an optimiser, looking at where the bases are positioned in the basis space is another interest.
 Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and  dimension reduction methods, i.e.
-principal component analysis is applied first to project all the bases onto a 2D space.
-In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The start basis; 2) The anchor bases that have the highest index value in each iteration; 3) The search bases that an optimiser evaluated to produce the anchor bases; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
-The importance of these basis differs with the most important ones being the start, interpolating and end bases.
+principal component analysis is applied to first project all the bases onto a 2D space.
+In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The start basis; 2) The search bases that an optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
+The importance of these basis differs and the most important ones are the start, interpolating and end bases.
 Anchor and search bases can be turned on with argument `details = TRUE`.
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing end bases.
-A flip sign device is implemented to flip the sign of all the bases in one optimisation if the situation described above happens and to ensure different routines finish close to each other.
+This happens because the projection is invariant to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing the optimisers since the end bases will be symmetric with respect to the origin.
+A sign flipping step is conducted to flip the signs of all the bases in one routine if different optimisations finish at opposite places.
 
 Several annotations have been made to help understanding this plot.
-The original space of the bases is a high dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
-Along with the bases recorded during the optimisation and a zero basis, PCA is performed to get the first two PC coordinates of all the bases.
-The search space in the 2D space is a circle with the origin being the PC coordinates of the zero matrix, and radius estimated as the largest distance between the origin and all the bases.
-The theoretical best basis is known with simulated data and can be labelled for comparing the how close the ending basis to the theoretical one for different optimisers.
+As mentioned previously, the original basis space is a high dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
+Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated.
+PCA is performed to get the first two PC coordinates of all the bases and the centre of the circle is the PC coordinates of the zero matrix, and radius is estimated as the largest distance from the centre to the basis.
+The theoretical best basis is known for simulated data and can be labelled to compare the how close the end basis found by the optimisers is to the theoretical best.
 Various aesthetics, i.e.
-size, alpha and colour, are applicable to emphasize the crucial elements and adjust for the presentation.
-For example, anchor points and search points are less important and a smaller size and alpha is used; The interpolation paths are more informative if alpha increases from start to finish (that is, the path is more opaque towards the end).
+size, alpha and colour, are applicable to emphasize critical elements and adjust for the presentation.
+For example, anchor points and search points are less important and hence a smaller size and alpha is used; Alpha can also be applied on the interpolation paths to show the start to finish from transparent to opaque.
 
-Figure \ref{fig:toy-pca} presents the PCA plot of a 1D projection of a 5-variable dataset, \code{boa5}, using the holes index with two optimisers, PD and SA, being compared.
-Both optimisers get close to the theoretical bases, annotated by "*" and in this particular example pseudo derivative search gets closer.
-One can also appreciate the nature of the two different searches: PD first evaluates points in a small neighbourhood for a promising direction, while SA evaluates points randomly in the search space to search for the next target.
-There are dashed lines annotated for SA and it describes the interruption of the interpolation, which has been briefly mentioned in Figure \ref{fig:toy-interp} and will be discussed in details in Section \ref{monotonic}.
+Figure \ref{fig:toy-pca} shows the PCA plot of SA and PD for a 1D projection problem.
+Both optimisers find the optimum but PD gets closer.
+With PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while SA evaluates points randomly in the search space to search for the next target.
+There are dashed lines annotated for SA and it describes the interruption of the interpolation, which will be discussed in details in Section \ref{monotonic}.
 
@@
 
-Animation is another display to show how the search progresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation.
+Animation is another display to show how the search progresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation in the PCA plot.
 Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
-An additional piece of information one can learn from this animation is that the SA search finds its end basis quicker than the PD search since SA has finished in the 5th frame while PD is still making more progression.
+An additional piece of information one can learn from this animation is that SA finds its end basis quicker than PD since SA finishes its search in the 5th frame while PD is still making more progression.
 
@@
 
-As mentioned in Section \ref{toy-pca}, the original $p \times d$ dimension space can be simulated via randomly generated bases via \CRANpkg{geozoo} [@geozoo] package.
-A tour plot rotates these random bases and the ones collected by the optimisation in the original space and hence provides a stereoscopic view of the search.
+As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in \CRANpkg{geozoo} [@geozoo] package.
 While the PCA plot projects the bases from the direction that maximises the variance, the tour plot display of the original high dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows the tour plot of the same two optimisations in the 5D space.
+Figure \ref{fig:toy-tour} shows frames from the tour plot of the same two optimisations in its original 5D space.
 
@@
 
-In this section, several examples will be presented to show how the diagnostic plots discover something unexpected in the optimisation algorithm, and guide the implementation of some new features.
+In this section, several examples will be presented to show how the diagnostic plots discover something unexpected in projection pursuit optimisation, and guide the implementation of new features.
 
@@
 Random variables with different distributions have been simulated and the distributional form of each variable is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance, and `x2` to `x7` are normal mixtures with varied weights and locations.
+Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
 All the variables have been scaled to have an overall unit variance before projection pursuit.
@@
 An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with higher index value than the target one.
-With optimiser SAJO, a non-monotonic interpolation is justified since the target bases do not necessarily have a higher index value than the current one, while this is not the case for SA.
-The original trace plot for a 2D projection, optimised by SA, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
-Hence, an interruption is implemented to stop the interpolation at the basis with the highest index value.
-The right panel of Figure \ref{fig:interruption} presents the trace plot after implementing such interruption and it can be observed that while the first two interpolations are identical, in the third interpolation the basis at time 61 has a higher index value than the target.
-Rather than starting the next iteration from the target basis on Time 65 as shown in the left panel, the optimiser, on the right panel, starts the next iteration from the interpolation basis on Time 61.
-With the interruption, the optimiser is able to find the final basis with a index value higher than the most promising basis when the interruption is not implemented.
+For SAJO, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for SA.
+The original trace plot for a 2D projection problem, optimised by SA, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+Hence, an interruption is implemented to stop at the best basis found in the interpolation.
+The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at Time 61 has a higher index value than the target in the third interpolations.
+Rather than starting the next iteration from the target basis on Time 65, SA starts the next iteration at Time 61 on the right panel and reaches a better final basis.
 
@@
 A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local breakthrough.
-The polish algorithm is similar to the SA with three main distinctions: 1) 100 rather than one candidate bases are generated each time in the inner loop; 2) the search neighbourhood, `alpha`, is reduced in the inner loop, rather than in the outer loop, once a set of candidate bases fails to make an improvement; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current in terms of the distance in the space,  percentage change in the index value, and neighbourhood:
+The polish algorithm is similar to the SA but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space,  percentage change in the index value, and neighbourhood size:
 
@@
 
-Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by SA and followed by the polish search, along with the projection on the final basis of each algorithm.
+Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by SA and followed by the polish.
+THe projection by the final basis of each algorithm is also shown.
 The end basis found by SA reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
-Polish works with this end basis and further pushes the index value to produce even clear structure in the data.
-One can observe that the projection after polishing better distinguishes the four clusters with more trimmed edges, espeically along the vertical axis.
+Polish works with this end basis and further pushes the index value to produce much clear edges of the cluster, especially along the vertical axis.
 
@@
 
-The index function before this section are all smooth, while this is not the case for all.
-A 1D projection function based on the Kolmogorov test, $I^{nk}(n)$, compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the empirical cumulated distribution function (ECDF).
-Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov, $I^{nk}(n)$, index is defined as:
+The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indices.
+A 1D projection function, $I^{nk}(n)$, compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the Kolmogorov test.
+Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index, $I^{nk}(n)$, is defined as:
 
 $$I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]$$
-With a non-smooth index function, two research questions are raised to understand
+With a non-smooth index function, two research questions are raised:
 
-1) if any optimiser fails to optimise a non-smooth index; and 
+1) whether any optimiser fails to optimise this non-smooth index; and 
 2) whether the optimisers can find the global optimum despite the presence of a local optimum
 
-```{r noisy-better-geo, fig.height = 6, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimiers.
+```{r noisy-better-geo, fig.height = 4, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimiers.
 Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
@@
 
-Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers.
-As expected, none of the interpolated path is smooth.
+Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated path is smooth.
 There is barely any improvement made by PD, indicating its failure in optimising non-smooth index.
 While SA and SAJO have managed to get close to the index value of the theoretical best, trace plot shows that it takes SAJO much longer to interpolate towards the final basis.
-This long interpolation path is partially due to the fluctuation in the early iterations, where the SAJO tends to generously accept inferior basis before concentrating near the optimum.
+This long interpolation path is partially due to the fluctuation in the early iterations, where the SAJO tends to generously accept inferior bases before concentrating near the optimum.
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
 [A sentence for PD].
-Comparing the amount of random search done by SA and SAJO, it is surprising that although SAJO doesn't actually carry that many of search as SA.
-Combine the insight from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement will be made, while SAJO first broadly accepts basis in the space and only starts the extensive search in a narrowed subspace.
-The specific decision of which optimiser to use will depends on the index curve in the basis space but if the basis space is highly non-smooth and spike, accepting inferior bases at first, as SAJO, can lead to a more efficient search, in terms of the overall number of points evaluated.
+Comparing the amount of random search done by SA and SAJO, it is surprising that SAJO doesn't carry that many of sampling as SA.
+Combine the insight from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SAJO first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
+The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as what SAJO has done, can lead to a more efficient search, in terms of the overall number of points evaluated.
 
-```{r kol-result, fig.height = 5, fig.cap="PCA plot of 1D projection on the 6\\-variable data, \\code{boa6}, with normal kolmogorov index.
+```{r kol-result, fig.height = 4, fig.cap="PCA plot of 1D projection on the 6\\-variable data, \\code{boa6}, with normal kolmogorov index.
 Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
@@
 The next experiment compares the performance of SA and SAJO when a local maximum exists.
-Two search neighbourhood sizes: 0.5 and 0.7, are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
+Two search neighbourhood sizes: 0.5 and 0.7 are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
 Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by  optimiser and search size.
-With SA and a search size of 0.5, despite being the simplest and fastest, SA fails to optimise in three instances where they finish neither near the local nor the global optimum.
-With a larger search size of 0.7, none of the seed fail and more seeds find the global maximum.
+With SA and a search size of 0.5, despite being the simplest and fastest, the optimiser fails in three instances where the final basis lands neither near the local nor the global optimum.
+With a larger search size of 0.7, more seeds have found the global maximum.
 Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to improve the final basis found, despite having longer interpolation paths.
-However, the denser path near the local maximum is an indicator that SAJO is working hard to examine if there is any other optimum in the space but the relative small search size has diminished its ability to reach out to the true global maximum.
+However, the denser paths near the local maximum is an indicator that SAJO is working hard to examine if there is any other optimum in the basis space but the relative small search size has diminished its ability to reach the global maximum.
 With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
-This indicates that SAJO, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the underlying true optimum.
+This indicates that SAJO, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the global optimum.
 
@@
 The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing CRAN package \CRANpkg{tourr} [@tourr] while the optimiser diagnostics have been implemented in a new package, \pkg{ferrn}.
-When a guided tour is run, the users can choose if the data from optimisation should be collected via the `verbose` argument.
-Once the data object has been obtained, the package, \pkg{ferrn}, can provide four diagnostic plots as shown in Section \ref{vis-diag}.
-The structure of package functionality has been listed below.
+When a guided tour is run, the users can choose to collect the data by binding `animate_*()` to a variable.
+Once the data object has been obtained, the \pkg{ferrn} package provides diagnostic plots shown in Section \ref{vis-diag}.
+The main functions of the \pkg{ferrn} package functionality is listed below.
 
@@
   - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
-Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
+Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
   - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
@@
 
+- \code{add\_*()} create ggprotos for different components for the PCA plot
+  - \code{add\_anchor()} is a wrapper for plotting anchor bases
+  - \code{add\_anno()} is a wrapper for annotating the symmetry of start bases
+  - \code{add\_dir\_search()} is a wrapper for plotting the directional search bases with magnified distance
+  - \code{add\_end()} is a wrapper for plotting end bases
+  - \code{add\_interp()} is a wrapper for plotting the interpolation path
+  - \code{add\_interp\_last()} is a wrapper for plotting the last interpolation bases for comparing with target bases when interruption is carried
+  - \code{add\_interrupt()} is a wrapper for linking the last interpolation bases with target ones when interruption is carried
+  - \code{add\_search()} is a wrapper for plotting search bases
+  - \code{add\_space()} is a wrapper for plotting the circular space 
+  - \code{add\_start()} is a wrapper for plotting start bases
+  - \code{add\_theo()} is a wrapper for plotting theoretical best bases, if applicable
+
 - Utilities
-  - \code{add\_*()} are internal wrapper functions that facilitate the composition of PCA plot
   - \code{theme\_fern()} and \code{format\_label()} for better display of the grid lines and axis formatting

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.tex
finish proofreading

@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
 
@@
 constraint requires the projection bases \(\mathbf{A}\) to be
-orhtonormal.
-
-Several features of this optimisation are worth-noticing. 1) The
-optimisation is constrained and the space of the bases is a
+orhtonormal. Several features of this optimisation are worth-noticing.
+1) The optimisation is constrained and the space of the bases is a
 d-dimensional sphere. 2) The objective function may not be
@@
 
-To be able to make diagnostics on the optimisers, the algorithms need to
-populate a data structure with key elements of the algorithm. When the
-algorithms run, key information regarding the decision variable,
-objective function and hyper-parameters needs to be recorded and stored
-as a data object so that it is ready to be supplied to the plotting
-functions for diagnostics.
+Before making diagnostic plots, information generated by the
+optimisation need to be collected and a pre-defined data structure
+solidifies the information being collected and streamlines the
+subsequent plot-making.
 
@@
 
-Three main elements are recorded for the projection pursuit optimisers:
-1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3)
-State: \(S\), which labels the observation with detailed stage in the
-optimisation. For optimiser SA and SAJO, possible values of the state
-include \texttt{random\_search}, \texttt{new\_basis}, and
-\texttt{interpolation}. pseudo derivative search has a wider variety of
+Three main pieces of information are recorded for the projection pursuit
+optimisers: 1) projection bases \(\mathbf{A}\), 2) index values \(I\),
+and 3) state \(S\). For SA and SAJO, possible state include
+\texttt{random\_search}, \texttt{new\_basis}, and
+\texttt{interpolation}. Pseudo derivative (PD) has a wider variety of
 state including \texttt{new\_basis}, \texttt{direction\_search},
 \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
-\texttt{interpolation}.
-
-Multiple iterators are also needed to index the data collected at
-different levels: \(t\) being a unique identifier that prescribes the
-natural ordering of each observation while \(j\) and \(l\) being the
-counter of the outer and inner loop, respectively, in Algorithm
-\ref{random-search}, \ref{simulated-annealing} and \ref{search-geodesic}
-above. Other parameters of interest recorded include
-\(V_1 = \text{method}\) that tags the name of the optimiser, and
-\(V_2 = \text{alpha}\) that indicates the sampling neighbourhood size. A
-matrix notation of the data structure is presented in Equation
-\ref{eq:data-structure}.
+\texttt{interpolation}. Multiple iterators index the information
+collected at different levels: \(t\) is a unique identifier prescribing
+the natural ordering of each observation; \(j\) and \(l\) are the
+counter of the outer and inner loop respectively. Other parameters of
+interest recorded include \texttt{method} that tags the name of the
+optimiser, and \texttt{alpha} that indicates the sampling neighbourhood
+size for searching observations. A matrix notation of the data structure
+is presented in Equation \ref{eq:data-structure}.
 
@@
 \noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
-no output in iteration \(J + 1\) since the optimiser cannot find a
-better basis in the last iteration and the algorithm terminates. The
-final basis found is \(A_T\) with the highest index value \(I_T\).
+no output in iteration \(J + 1\) since the optimiser doesn't find a
+better basis in the last iteration and terminates. The final basis found
+is \(A_T\) with index value \(I_T\).
 
@@
 
-The construction of diagnostic plots adopts the core concept in ggplot2:
-grammar of graphics \citep{wickham2010layered}. In grammar of graphics,
-plots are not produced via calling the commands, named by the appearance
-of the plot, i.e., boxplot and histogram, but via the concept of stacked
-layers. Seeing plots as stacked layers empowers us to composite
-diagnostic plots with an emphasis on any variable in the data object
-without the redundancy of creating different commands for the same type
-of plots that highlights on different variables.
+The construction of diagnostic plots adopts the core concept of grammar
+of graphics \citep{wickham2010layered} in ggplot2. In grammar of
+graphics, plots are not produced by calling the commands, named by the
+appearance of the plot, i.e., boxplot and histogram, but by the concept
+of stacked layers. Seeing plots as stacked layers gives analysts the
+freedom to composite plots with multiple elements on hands.
 
@@
 A starting point of diagnosing an optimiser is to understand how many
-search the optimiser has conducted. One may want to simply plot the
-index value of the search points across its natural order. A point
-geometry may work well if each iteration has a similar number of points,
-however, the plot will be disproportionate if some iterations have
-considerably more points than other - it over-emphasizes the iterations
-that have more search points! An alternative is to summarise the search
-in each iteration using boxplot where each iteration will be spaced out
-equally. Occasionally, one may still want to switch back to point
-geometry if the number of points is small and this can be achieved via
-the \texttt{cutoff} argument. Additional annotation can be added with
-new layers, for example, the number of points searched in each iteration
-can be added as text label at the bottom of each iteration; the anchor
-bases to interpolate are connected and highlighted in a larger size; and
-the colour of the last iteration is in a grey scale to indicate no
-better basis found in this iteration.
-
-Figure \ref{fig:toy-search} shows the searches of two different
-optimisers: SA (left) and SAJO (right). Both optimisers quickly find
-better bases at the first few iterations, then take longer in the later
-iterations, and finally finish when the maximum number of evaluation is
-reached. The anchor bases, the ones with the highest index value in each
-iteration, always have an increased index value in the optimiser SA
-while this is not the case for SAJO. This feature gives SA an advantage
-in simple example, like the one presented while the jump out component
-in SAJO allows for a more holistic search in more complicated examples.
+search an optimiser has conducted. One may want to simply plot the index
+value of the search points across its natural order, but a point
+geometry may work well if each iteration has a similar number of points.
+When some iterations have considerably more points than others, using a
+point geometry over-emphasizes the iterations that have more search
+points since these iterations will occupy the mast majority of the plot
+space. An alternative is to summarise the search in each iteration using
+boxplot and each iteration will then be spaced out equally.
+Occasionally, one may still want to switch back to a point geometry if
+the number of points is small in a particular iteration and this can be
+achieved via the \texttt{cutoff} argument in the search plot function.
+Additional annotations are added to facilitate better reading of the
+plot and these includes\\
+1) the number of points searched in each iteration can be added as text
+label at the bottom of each iteration; 2) the anchor bases to
+interpolate are connected and highlighted in a larger size; and 3) the
+colour of the last iteration is in a grey scale to indicate no better
+basis found in this iteration.
+
+Figure \ref{fig:toy-search} shows an example of the search plot for SA
+(left) and SAJO (right). Both optimisers quickly find better bases in
+the first few iterations and then take longer to find one in the later
+iterations. The anchor bases, the ones found with the highest index
+value in each iteration, always have an increased index value in the
+optimiser SA while this is not the case for SAJO. This feature gives SA
+an advantage in this simple example to quickly find the optimum.
 
@@
 
-Another interest of diagnosing is to examine how the index value
-progresses for the points on the interpolation path since the projection
-on these bases are played by the tour animation. Trace plots are created
-with plotting the index value of the interpolation points against time.
-Figure \ref{fig:toy-interp} presents the trace plot of the two
-optimisation routines as \ref{fig:toy-search}. The trace plot is smooth
-in both plots and this is a character of the holes index function used.
-An interesting discovery is that with optimiser SAJO, the interpolator,
-in iteration 6-8, first passes through some bases with higher index
-value than the anchor bases.
+Another interest of diagnostic is to examine how the index value
+improves for interpolating bases since the projection on these bases is
+played by the tour animation. Trace plots are created with plotting the
+index value of the interpolation points against time. Figure
+\ref{fig:toy-interp} presents the trace plot of the same optimisation
+routine as Figure \ref{fig:toy-search}. The trace plot is smooth and an
+interesting discovery is that in SAJO, the interpolation in iteration
+6-8 is not monotonic. Sometimes, this non-monotonicity is acceptable
+since the target basis does not necessarily have a higher index value
+than the current one, while in some cases, the optimisation will be
+better off if these intermediate basis can be used as the start of the
+next iteration (see section \ref{monotonic} for a detailed discussion on
+this).
 
@@
 
-\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD.
+\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD.
 All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
@@
 
-Apart from checking the progression of an optimiser, looking at how the
-points evaluated by the optimisers look like in the space is another
+Apart from checking the search and progression of an optimiser, looking
+at where the bases are positioned in the basis space is another
 interest. Given the orthonormality constraint, the space of projection
@@
 and dimension reduction methods, i.e.~principal component analysis is
-applied first to project all the bases onto a 2D space. In a projection
+applied to first project all the bases onto a 2D space. In a projection
 pursuit guided tour optimisation, there are various type of bases
-involved: 1) The start basis; 2) The anchor bases that have the highest
-index value in each iteration; 3) The search bases that an optimiser
-evaluated to produce the anchor bases; 4) The interpolating bases on the
+involved: 1) The start basis; 2) The search bases that an optimiser
+evaluated to produce the anchor bases; 3) The anchor bases that have the
+highest index value in each iteration; 4) The interpolating bases on the
 interpolation path; and finally 5) the end basis. The importance of
-these basis differs with the most important ones being the start,
+these basis differs and the most important ones are the start,
 interpolating and end bases. Anchor and search bases can be turned on
@@
 the bases and so does the index value, however, this creates
-difficulties for comparing end bases. A flip sign device is implemented
-to flip the sign of all the bases in one optimisation if the situation
-described above happens and to ensure different routines finish close to
-each other.
-
-Several annotations have been made to help understanding this plot. The
-original space of the bases is a high dimensional sphere and random
-bases on the sphere can be generated via the CRAN package
-\CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded during
-the optimisation and a zero basis, PCA is performed to get the first two
-PC coordinates of all the bases. The search space in the 2D space is a
-circle with the origin being the PC coordinates of the zero matrix, and
-radius estimated as the largest distance between the origin and all the
-bases. The theoretical best basis is known with simulated data and can
-be labelled for comparing the how close the ending basis to the
-theoretical one for different optimisers. Various aesthetics, i.e.~size,
-alpha and colour, are applicable to emphasize the crucial elements and
-adjust for the presentation. For example, anchor points and search
-points are less important and a smaller size and alpha is used; The
-interpolation paths are more informative if alpha increases from start
-to finish (that is, the path is more opaque towards the end).
-
-Figure \ref{fig:toy-pca} presents the PCA plot of a 1D projection of a
-5-variable dataset, \code{boa5}, using the holes index with two
-optimisers, PD and SA, being compared. Both optimisers get close to the
-theoretical bases, annotated by "*" and in this particular example
-pseudo derivative search gets closer. One can also appreciate the nature
-of the two different searches: PD first evaluates points in a small
-neighbourhood for a promising direction, while SA evaluates points
-randomly in the search space to search for the next target. There are
-dashed lines annotated for SA and it describes the interruption of the
-interpolation, which has been briefly mentioned in Figure
-\ref{fig:toy-interp} and will be discussed in details in Section
-\ref{monotonic}.
+difficulties for comparing the optimisers since the end bases will be
+symmetric with respect to the origin. A sign flipping step is conducted
+to flip the signs of all the bases in one routine if different
+optimisations finish at opposite places.
+
+Several annotations have been made to help understanding this plot. As
+mentioned previously, the original basis space is a high dimensional
+sphere and random bases on the sphere can be generated via the CRAN
+package \CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded
+during the optimisation and a zero basis, the parameters (centre and
+radius) of the 2D space can be estimated. PCA is performed to get the
+first two PC coordinates of all the bases and the centre of the circle
+is the PC coordinates of the zero matrix, and radius is estimated as the
+largest distance from the centre to the basis. The theoretical best
+basis is known for simulated data and can be labelled to compare the how
+close the end basis found by the optimisers is to the theoretical best.
+Various aesthetics, i.e.~size, alpha and colour, are applicable to
+emphasize critical elements and adjust for the presentation. For
+example, anchor points and search points are less important and hence a
+smaller size and alpha is used; Alpha can also be applied on the
+interpolation paths to show the start to finish from transparent to
+opaque.
+
+Figure \ref{fig:toy-pca} shows the PCA plot of SA and PD for a 1D
+projection problem. Both optimisers find the optimum but PD gets closer.
+With PCA plot, one can visually appreciate the nature of these two
+optimisers: PD first evaluates points in a small neighbourhood for a
+promising direction, while SA evaluates points randomly in the search
+space to search for the next target. There are dashed lines annotated
+for SA and it describes the interruption of the interpolation, which
+will be discussed in details in Section \ref{monotonic}.
 
@@
 start to finish in the space and an \texttt{animate\ =\ TRUE} argument
-is used to enable the animation. Figure \ref{fig:toy-pca-animated} shows
-six frames from the animation of the PCA plot in Figure
-\ref{fig:toy-pca}. An additional piece of information one can learn from
-this animation is that the SA search finds its end basis quicker than
-the PD search since SA has finished in the 5th frame while PD is still
+is used to enable the animation in the PCA plot. Figure
+\ref{fig:toy-pca-animated} shows six frames from the animation of the
+PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
+one can learn from this animation is that SA finds its end basis quicker
+than PD since SA finishes its search in the 5th frame while PD is still
 making more progression.
@@
 
-As mentioned in Section \ref{toy-pca}, the original \(p \times d\)
-dimension space can be simulated via randomly generated bases via
-\CRANpkg{geozoo} \citep{geozoo} package. A tour plot rotates these
-random bases and the ones collected by the optimisation in the original
-space and hence provides a stereoscopic view of the search. While the
-PCA plot projects the bases from the direction that maximises the
-variance, the tour plot display of the original high dimensional space
-from various directions using animation. Figure \ref{fig:toy-tour} shows
-the tour plot of the same two optimisations in the 5D space.
+As mentioned previously, the original \(p \times d\) dimension space can
+be simulated via randomly generated bases in \CRANpkg{geozoo}
+\citep{geozoo} package. While the PCA plot projects the bases from the
+direction that maximises the variance, the tour plot display of the
+original high dimensional space from various directions using animation.
+Figure \ref{fig:toy-tour} shows frames from the tour plot of the same
+two optimisations in its original 5D space.
 
@@
 In this section, several examples will be presented to show how the
-diagnostic plots discover something unexpected in the optimisation
-algorithm, and guide the implementation of some new features.
+diagnostic plots discover something unexpected in projection pursuit
+optimisation, and guide the implementation of new features.
 
@@
 \texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
-variance, and \texttt{x2} to \texttt{x7} are normal mixtures with varied
+variance and \texttt{x2} to \texttt{x7} are normal mixtures with varied
 weights and locations. All the variables have been scaled to have an
@@
 \ref{fig:toy-interp}: a path that passes bases with higher index value
-than the target one. With optimiser SAJO, a non-monotonic interpolation
-is justified since the target bases do not necessarily have a higher
-index value than the current one, while this is not the case for SA. The
-original trace plot for a 2D projection, optimised by SA, is shown on
-the left panel of Figure \ref{fig:interruption} and one can observe
-clearly that the non-monotonic interpolation has undermined the
+than the target one. For SAJO, a non-monotonic interpolation is
+justified since target bases do not necessarily have a higher index
+value than the current one, while this is not the case for SA. The
+original trace plot for a 2D projection problem, optimised by SA, is
+shown on the left panel of Figure \ref{fig:interruption} and one can
+observe clearly that the non-monotonic interpolation has undermined the
 optimiser to realise its full potential. Hence, an interruption is
-implemented to stop the interpolation at the basis with the highest
-index value. The right panel of Figure \ref{fig:interruption} presents
-the trace plot after implementing such interruption and it can be
-observed that while the first two interpolations are identical, in the
-third interpolation the basis at time 61 has a higher index value than
-the target. Rather than starting the next iteration from the target
-basis on Time 65 as shown in the left panel, the optimiser, on the right
-panel, starts the next iteration from the interpolation basis on Time
-61. With the interruption, the optimiser is able to find the final basis
-with a index value higher than the most promising basis when the
-interruption is not implemented.
+implemented to stop at the best basis found in the interpolation. The
+right panel of Figure \ref{fig:interruption} shows the trace plot after
+implementing the interruption and while the first two interpolations are
+identical, the basis at Time 61 has a higher index value than the target
+in the third interpolations. Rather than starting the next iteration
+from the target basis on Time 65, SA starts the next iteration at Time
+61 on the right panel and reaches a better final basis.
 
@@
 the start of a new guided tour to search for local breakthrough. The
-polish algorithm is similar to the SA with three main distinctions: 1)
-100 rather than one candidate bases are generated each time in the inner
-loop; 2) the search neighbourhood, \texttt{alpha}, is reduced in the
-inner loop, rather than in the outer loop, once a set of candidate bases
-fails to make an improvement; and 3) three more termination conditions
+polish algorithm is similar to the SA but with three distinctions: 1) a
+hundred rather than one candidate bases are generated each time in the
+inner loop; 2) the neighbourhood size is reduced in the inner loop,
+rather than in the outer loop; and 3) three more termination conditions
 have been added to ensure the new basis generated is distinguishable
-from the current in terms of the distance in the space, percentage
-change in the index value, and neighbourhood:
+from the current one in terms of the distance in the space, percentage
+change in the index value, and neighbourhood size:
 
@@
 Figure \ref{fig:polish} presents the trace plot of a 2D projection,
-optimised by SA and followed by the polish search, along with the
-projection on the final basis of each algorithm. The end basis found by
-SA reveals the four clusters in the data, but the edges of each cluster
-are not clean-cut. Polish works with this end basis and further pushes
-the index value to produce even clear structure in the data. One can
-observe that the projection after polishing better distinguishes the
-four clusters with more trimmed edges, espeically along the vertical
-axis.
+optimised by SA and followed by the polish. THe projection by the final
+basis of each algorithm is also shown. The end basis found by SA reveals
+the four clusters in the data, but the edges of each cluster are not
+clean-cut. Polish works with this end basis and further pushes the index
+value to produce much clear edges of the cluster, especially along the
+vertical axis.
 
@@
 
-The index function before this section are all smooth, while this is not
-the case for all. A 1D projection function based on the Kolmogorov test,
-\(I^{nk}(n)\), compares the projected data, \(\mathbf{Y}_{n \times 1}\),
-to a randomly generated normal distribution,
-\(\mathcal{N}_{n \times 1}\), based on the empirical cumulated
-distribution function (ECDF). Let \(F_{.}(n)\) be the ECDF function,
-with two possible subscripts \(Y\) and \(\mathcal{N}\) representing the
-projected and randomly generated data, and \(n\) denoting the number of
-observation, the Normal Kolmogorov, \(I^{nk}(n)\), index is defined as:
+The holes index function used for all the examples before this section
+produces a smooth interpolation, while this is not the case for all the
+indices. A 1D projection function, \(I^{nk}(n)\), compares the projected
+data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated normal
+distribution, \(\mathcal{N}_{n \times 1}\), based on the Kolmogorov
+test. Let \(F_{.}(n)\) be the ECDF function, with two possible
+subscripts \(Y\) and \(\mathcal{N}\) representing the projected and
+randomly generated data, and \(n\) denoting the number of observation,
+the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
 
 \[I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
-non-smooth index function, two research questions are raised to
-understand
+non-smooth index function, two research questions are raised:
 
@@
 \item
-  if any optimiser fails to optimise a non-smooth index; and
+  whether any optimiser fails to optimise this non-smooth index; and
 \item
@@
 Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
-all three optimisers. As expected, none of the interpolated path is
+all three optimisers and as expected, none of the interpolated path is
 smooth. There is barely any improvement made by PD, indicating its
@@
 in the early iterations, where the SAJO tends to generously accept
-inferior basis before concentrating near the optimum. The PCA plot shows
+inferior bases before concentrating near the optimum. The PCA plot shows
 the interpolation path and search points excluding the last termination
 iteration. {[}A sentence for PD{]}. Comparing the amount of random
-search done by SA and SAJO, it is surprising that although SAJO doesn't
-actually carry that many of search as SA. Combine the insight from both
-the trace and PCA plot, one can learn the two different search
-strategies by SA and SAJO: SA frequently samples in the space and only
-make a move when an improvement will be made, while SAJO first broadly
-accepts basis in the space and only starts the extensive search in a
-narrowed subspace. The specific decision of which optimiser to use will
-depends on the index curve in the basis space but if the basis space is
-highly non-smooth and spike, accepting inferior bases at first, as SAJO,
-can lead to a more efficient search, in terms of the overall number of
-points evaluated.
+search done by SA and SAJO, it is surprising that SAJO doesn't carry
+that many of sampling as SA. Combine the insight from both the trace and
+PCA plot, one can learn the two different search strategies by SA and
+SAJO: SA frequently samples in the space and only make a move when an
+improvement is guaranteed to be made, while SAJO first broadly accepts
+bases in the space and then starts the extensive sampling in a narrowed
+subspace. The specific decision of which optimiser to use will depend on
+the index curve in the basis space but if the basis space is non-smooth,
+accepting inferior bases at first, as what SAJO has done, can lead to a
+more efficient search, in terms of the overall number of points
+evaluated.
 
@@
 The next experiment compares the performance of SA and SAJO when a local
-maximum exists. Two search neighbourhood sizes: 0.5 and 0.7, are
-compared to understand how a large search neighbourhood would affect the
-final basis and the length of the search. Figure \ref{fig:kol-result}
-shows 80 paths simulated using 20 seeds in the PCA plot, faceted by
-optimiser and search size. With SA and a search size of 0.5, despite
-being the simplest and fastest, SA fails to optimise in three instances
-where they finish neither near the local nor the global optimum. With a
-larger search size of 0.7, none of the seed fail and more seeds find the
-global maximum. Comparing SA and SAJO for a search size of 0.5, SAJO
-doesn't seem to improve the final basis found, despite having longer
-interpolation paths. However, the denser path near the local maximum is
-an indicator that SAJO is working hard to examine if there is any other
-optimum in the space but the relative small search size has diminished
-its ability to reach out to the true global maximum. With a larger
-search size, almost all the seeds (16 out of 20) have found the global
-maximum and some final bases are much closer to the theoretical best, as
-compared to the three other cases. This indicates that SAJO, with a
-reasonable large search size, is able to overcome the local optimum and
-optimise close towards the underlying true optimum.
+maximum exists. Two search neighbourhood sizes: 0.5 and 0.7 are compared
+to understand how a large search neighbourhood would affect the final
+basis and the length of the search. Figure \ref{fig:kol-result} shows 80
+paths simulated using 20 seeds in the PCA plot, faceted by optimiser and
+search size. With SA and a search size of 0.5, despite being the
+simplest and fastest, the optimiser fails in three instances where the
+final basis lands neither near the local nor the global optimum. With a
+larger search size of 0.7, more seeds have found the global maximum.
+Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to
+improve the final basis found, despite having longer interpolation
+paths. However, the denser paths near the local maximum is an indicator
+that SAJO is working hard to examine if there is any other optimum in
+the basis space but the relative small search size has diminished its
+ability to reach the global maximum. With a larger search size, almost
+all the seeds (16 out of 20) have found the global maximum and some
+final bases are much closer to the theoretical best, as compared to the
+three other cases. This indicates that SAJO, with a reasonable large
+search size, is able to overcome the local optimum and optimise close
+towards the global optimum.
 
@@
 have been implemented in a new package, \pkg{ferrn}. When a guided tour
-is run, the users can choose if the data from optimisation should be
-collected via the \texttt{verbose} argument. Once the data object has
-been obtained, the package, \pkg{ferrn}, can provide four diagnostic
-plots as shown in Section \ref{vis-diag}. The structure of package
-functionality has been listed below.
+is run, the users can choose to collect the data by binding
+\texttt{animate\_*()} to a variable. Once the data object has been
+obtained, the \pkg{ferrn} package provides diagnostic plots shown in
+Section \ref{vis-diag}. The main functions of the \pkg{ferrn} package
+functionality is listed below.
 
@@
     version in Figure \ref{fig:toy-pca-animated} can be turned on via
-    the argument \texttt{animate\ =\ TRUE}
+    the argument \texttt{animate\ =\ TRUE}.
   \item
@@
 \item
-  Utilities
+  \code{add\_*()} create ggprotos for different components for the PCA
+  plot
 
@@
   \item
-    \code{add\_*()} are internal wrapper functions that facilitate the
-    composition of PCA plot
+    \code{add\_anchor()} is a wrapper for plotting anchor bases
+  \item
+    \code{add\_anno()} is a wrapper for annotating the symmetry of start
+    bases
+  \item
+    \code{add\_dir\_search()} is a wrapper for plotting the directional
+    search bases with magnified distance
+  \item
+    \code{add\_end()} is a wrapper for plotting end bases
+  \item
+    \code{add\_interp()} is a wrapper for plotting the interpolation
+    path
+  \item
+    \code{add\_interp\_last()} is a wrapper for plotting the last
+    interpolation bases for comparing with target bases when
+    interruption is carried
+  \item
+    \code{add\_interrupt()} is a wrapper for linking the last
+    interpolation bases with target ones when interruption is carried
+  \item
+    \code{add\_search()} is a wrapper for plotting search bases
+  \item
+    \code{add\_space()} is a wrapper for plotting the circular space
+  \item
+    \code{add\_start()} is a wrapper for plotting start bases
+  \item
+    \code{add\_theo()} is a wrapper for plotting theoretical best bases,
+    if applicable
+  \end{itemize}
+\item
+  Utilities
+
+  \begin{itemize}
+  \tightlist
   \item

### 2021-02-01 scripts/toy.R

@@ -37,7 +37,7 @@ bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
   explore_space_pca(group = method, details = TRUE,
-                    interp_size = 1, ratio = 5) +
+                    interp_size = 1, ratio = 9) +
   scale_color_discrete_botanical()
 
 ## ----toy-pca-animated

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
run spell check

@@
 
-\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orhtonormal.
+\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
 Several features of this optimisation are worth-noticing.
@@
 Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by SA and followed by the polish.
-THe projection by the final basis of each algorithm is also shown.
+The projection by the final basis of each algorithm is also shown.
 The end basis found by SA reveals the four clusters in the data, but the edges of each cluster are not clean-cut.

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.tex
run spell check

@@
 constraint requires the projection bases \(\mathbf{A}\) to be
-orhtonormal. Several features of this optimisation are worth-noticing.
+orthonormal. Several features of this optimisation are worth-noticing.
 1) The optimisation is constrained and the space of the bases is a
@@
 Figure \ref{fig:polish} presents the trace plot of a 2D projection,
-optimised by SA and followed by the polish. THe projection by the final
+optimised by SA and followed by the polish. The projection by the final
 basis of each algorithm is also shown. The end basis found by SA reveals

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
underscore to dash in tour-path.png

@@
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
-include_graphics(here::here("img/tour_path.png"))
+include_graphics(here::here("img/tour-path.png"))
 ```

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
Updated text in the first 2 sections

@@
 Visualisation is widely used in exploratory data analysis [@tukey1977exploratory; @unwin2015graphical;  @healy2018data; @wilke2019fundamentals].
-Presenting information in graphics often unveils information that would otherwise not be aware of and provides a more comprehensive understanding of the problem at hand.
-Task specific tools such as @li2020visualizing show how visualisation can be used to understand the behaviour of neural network on classification models, but no general visualisation tool is available for diagnosing optimisation procedures.
-The work presented in this paper brings visualization tools into optimisation problems with an aim to better understand the performance of the optimisers in practice.
+Presenting information in graphics often unveils information that would otherwise not be discovered and provides a more comprehensive understanding of the problem at hand.
+Task specific tools such as @li2020visualizing show how visualisation can be used to understand, for instance, the behaviour of neural network on classification models.
+However, no general visualisation tool is available for diagnosing optimisation procedures.
+The work presented in this paper brings visualization tools into optimisation problems with the aim to better understand the performance of optimisers in practice.
 
@@
 clustering, outliers and skewness) in multivariate data via low dimensions projection.
-Guided tour is one variation of a broader class of data visualisation methods, tour, which visualises high-dimensional data through a series of animated projection.
+Guided tour is one variation of a broader class of data visualisation methods, tour, which displays high-dimensional data through a series of animated projections.
 
-Notation is first established before defining projection pursuit.
+<!-- Notation is first established before defining projection pursuit. -->
 Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions.
-A d-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$, and defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis Define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto an scalar.
+A d-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix.
+We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto an scalar value.
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
@@
 
-As a general visualisation method, tour produces animations of high dimensional data via rotations low dimension planes.
-Different tour types choose these planes differently: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
+As a general visualisation method, tour produces animations of high dimensional data via rotations of low dimension planes.
+There are different tour implementations depending on how the high dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
 Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection.
-Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs the geodesic interpolation between these planes to form a tour path.
+Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
 Figure \ref{fig:tour-path} shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations  between the blue frames.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-The tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
+The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
 

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.tex
underscore to dash in tour-path.png
Updated text in the first 2 sections

@@
 Presenting information in graphics often unveils information that would
-otherwise not be aware of and provides a more comprehensive
+otherwise not be discovered and provides a more comprehensive
 understanding of the problem at hand. Task specific tools such as
 \citet{li2020visualizing} show how visualisation can be used to
-understand the behaviour of neural network on classification models, but
-no general visualisation tool is available for diagnosing optimisation
-procedures. The work presented in this paper brings visualization tools
-into optimisation problems with an aim to better understand the
-performance of the optimisers in practice.
+understand, for instance, the behaviour of neural network on
+classification models. However, no general visualisation tool is
+available for diagnosing optimisation procedures. The work presented in
+this paper brings visualization tools into optimisation problems with
+the aim to better understand the performance of optimisers in practice.
 
@@
 variation of a broader class of data visualisation methods, tour, which
-visualises high-dimensional data through a series of animated
-projection.
-
-Notation is first established before defining projection pursuit. Let
-\(\mathbf{X}_{n \times p}\) be the data matrix with \(n\) observations
-in \(p\) dimensions. A d-dimensional projection is a linear
-transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\), and defined
-as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
+displays high-dimensional data through a series of animated projections.
+
+Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
+observations in \(p\) dimensions. A d-dimensional projection is a linear
+transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\) defined as
+\(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{Y}_{n \times d}\) is the projected data and
-\(\mathbf{A}_{p\times d}\) is the projection basis Define
+\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
 \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
-function that maps the projected data \(\mathbf{Y}\) onto an scalar.
-This is commonly known as the projection pursuit index function, or just
-index function, and is used to measure the ``interestingness'' of a
-given projection. A number of index functions have been proposed in the
-literature to detect different data structures
+function that maps the projected data \(\mathbf{Y}\) onto an scalar
+value. This is commonly known as the projection pursuit index function,
+or just index function, and is used to measure the ``interestingness''
+of a given projection. A number of index functions have been proposed in
+the literature to detect different data structures
 \citep[friedman1974projection;][]{hall1989polynomial, cook1993projection, posse1995projection}
@@
 As a general visualisation method, tour produces animations of high
-dimensional data via rotations low dimension planes. Different tour
-types choose these planes differently: grand tour \citep{cook2008grand}
-selects the planes randomly to provide a general overview; manual tour
+dimensional data via rotations of low dimension planes. There are
+different tour implementations depending on how the high dimensional
+space is investigated: grand tour \citep{cook2008grand} selects the
+planes randomly to provide a general overview; manual tour
 \citep{cook1997manual} gradually phases in and out one variable to
@@
 projection. Given a random start, projection pursuit iteratively finds
-bases with higher index values and the guided tour constructs the
-geodesic interpolation between these planes to form a tour path. Figure
+bases with higher index values and the guided tour constructs a geodesic
+interpolation between these planes to form a tour path. Figure
 \ref{fig:tour-path} shows a sketch of the tour path where the blue
@@
 of the geodesic interpolation can be found in
-\citet{buja2005computational}. The tour method has been implemented in
-the R package \CRANpkg{tourr} \citep{tourr}.
+\citet{buja2005computational}. The aforementioned tour method has been
+implemented in the R package \CRANpkg{tourr} \citep{tourr}.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/pmen0008/Documents/paper-tour-vis/img/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/flip-sign-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/kol-result-1}
 

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
Add comments in Optimisation in the tour section

@@
 
-The optimisation problem in projection pursuit is stated as follows: Given a randomly start basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
+In projection pursuit the optimisation problem (aims at ...
+describe it in simple words.) That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
 
@@
 
-\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
+\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in section XX, and  $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
 Several features of this optimisation are worth-noticing.

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.tex
Add comments in Optimisation in the tour section

@@
 
-The optimisation problem in projection pursuit is stated as follows:
-Given a randomly start basis \(\mathbf{A}_1\), projection pursuit finds
-the final projection basis \(\mathbf{A}_T\) that satisfies the following
-optimisation problem:
+In projection pursuit the optimisation problem (aims at \ldots{}
+describe it in simple words.) That is, it starts with a given randomly
+selected basis \(\mathbf{A}_1\) and aims at finding an optimal final
+projection basis \(\mathbf{A}_T\) in each iteration that satisfies the
+following optimisation problem:
 
@@
 
-\noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
-constraint requires the projection bases \(\mathbf{A}\) to be
+\noindent where \(f\), \(\mathcal{A}\), and \(\mathbf{X}\) are defined
+as in section XX, and \(I_d\) is the \(d\)-dimensional identity matrix
+and the constraint requires the projection bases \(\mathbf{A}\) to be
 orthonormal. Several features of this optimisation are worth-noticing.

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
update zhang-cook-laa-langrene-menendez.Rmd
minor corrections/typos

--- 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
-Specific solvers for simulated annealing includes \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
+Specific solvers for simulated annealing include \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
 For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].
@@
 Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
-clustering, outliers and skewness) in multivariate data via low dimensions projection.
+clustering, outliers and skewness) in multivariate data via low-dimensional projections.
 Guided tour is one variation of a broader class of data visualisation methods, tour, which displays high-dimensional data through a series of animated projections.
@@
 Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions.
-A d-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix.
-We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto an scalar value.
+A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix.
+We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value.
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-A number of index functions have been proposed in the literature to detect different data structures [friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
+A number of index functions have been proposed in the literature to detect different data structures [@friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
 
-As a general visualisation method, tour produces animations of high dimensional data via rotations of low dimension planes.
+As a general visualisation method, tour produces animations of high dimensional data via rotations of low-dimensional planes.
 There are different tour implementations depending on how the high dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
@@
 
-The optimisation problem in projection pursuit is stated as follows: Given a randomly start basis $\mathbf{A}_1$, projection pursuit finds the final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
+In projection pursuit the optimisation problem \textcolor{red}{(aims at ...
+describe it in simple words.)} That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
 
@@
 
-\noindent where $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
+\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in section \textcolor{red}{XX}, and  $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
 Several features of this optimisation are worth-noticing.
-1) The optimisation is constrained and the space of the bases is a d-dimensional sphere.
+1) The optimisation is constrained and the space of the bases is a $d$-dimensional sphere.
 2) The objective function may not be differentiable for a constructed index function.
@@
 \Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
-  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
-  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
-  compute the corresponding index value for each candidate bases\;
-  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
-  determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
+  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
+  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
+  Compute the corresponding index value for each candidate bases\;
+  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
+  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
+  find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
   update $l = l + 1$\;
@@
 Three main pieces of information are recorded for the projection pursuit optimisers: 1) projection bases $\mathbf{A}$, 2) index values $I$, and 3) state $S$.
-For SA and SAJO, possible state include `random_search`,  `new_basis`, and `interpolation`.
-Pseudo derivative (PD) has a wider variety of state including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
+For SA and SAJO, possible states include `random_search`,  `new_basis`, and `interpolation`.
+Pseudo derivative (PD) has a wider variety of states including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
 Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation;  $j$ and $l$ are the counter of the outer and inner loop respectively.
 Other parameters of interest recorded include `method` that tags the name of the optimiser, and `alpha` that indicates the sampling neighbourhood size for searching observations.
-A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.
+A matrix notation of the data structure is presented in equation \ref{eq:data-structure}.
 
@@
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
-Note that there is no output in iteration $J + 1$ since the optimiser doesn't find a better basis in the last iteration and terminates.
+Note that there is no output in iteration $J + 1$ since the optimiser does not find a better basis in the last iteration and terminates.
 The final basis found is $A_T$ with index value $I_T$.
@@
 
-A starting point of diagnosing an optimiser is to understand how many search an optimiser has conducted.
+A starting point of diagnosing an optimiser is to understand how many searches an optimiser has conducted.
 One may want to simply plot the index value of the search points across its natural order, but a point geometry may work well if each iteration has a similar number of points.
 When some iterations have considerably more points than others, using a point geometry over-emphasizes the iterations that have more search points since these iterations will occupy the mast majority of the plot space.
-An alternative is to summarise the search in each iteration using boxplot and each iteration will then be spaced out equally.
+An alternative is to summarise the search in each iteration using boxplots and each iteration will then be spaced out equally.
 Occasionally, one may still want to switch back to a point geometry if the number of points is small in a particular iteration and this can be achieved via the `cutoff` argument in the search plot function.
-Additional annotations are added to facilitate better reading of the plot and these includes
+Additional annotations are added to facilitate better reading of the plot and these include
 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
@@
 
-```{r toy-search, fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
+```{r toy-search, fig.cap="A comparison of search by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
+Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
 ```
@@
 ```{r toy-interp, fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
-Both traces are smooth while the change of index value in each iteration may not be monotinic."}
+Both traces are smooth while the change of index value in each iteration may not be monotonic."}
 ```
 
-Another interest of diagnostic is to examine how the index value improves for interpolating bases since the projection on these bases is played by the tour animation.
+Another interesting diagnostic is to examine how the index value improves between interpolating bases since the projection on these bases is played by the tour animation.
 Trace plots are created with plotting the index value of the interpolation points against time.

### 2021-02-03 paper/zhang-cook-laa-langrene-menendez.Rmd
minor fixes

@@
 Apart from checking the search and progression of an optimiser, looking at where the bases are positioned in the basis space is another interest.
-Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and  dimension reduction methods, i.e.
+Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$-dimensional sphere.
+A dimensionality reduction method, e.g.
 principal component analysis is applied to first project all the bases onto a 2D space.
-In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The start basis; 2) The search bases that an optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
-The importance of these basis differs and the most important ones are the start, interpolating and end bases.
-Anchor and search bases can be turned on with argument `details = TRUE`.
+In a projection pursuit guided tour optimisation, there are various types of bases involved: 1) The starting basis; 2) The search bases that the optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
+The importance of these bases differs and the most important ones are the starting, interpolating and end bases.
+The anchor and search bases can be turned on with argument `details = TRUE`.
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
@@
 Several annotations have been made to help understanding this plot.
-As mentioned previously, the original basis space is a high dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
+As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
 Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated.
 PCA is performed to get the first two PC coordinates of all the bases and the centre of the circle is the PC coordinates of the zero matrix, and radius is estimated as the largest distance from the centre to the basis.
-The theoretical best basis is known for simulated data and can be labelled to compare the how close the end basis found by the optimisers is to the theoretical best.
+The theoretical best basis is known for simulated data and can be labelled to compare how close the end basis found by the optimisers is to the theoretical best.
 Various aesthetics, i.e.
 size, alpha and colour, are applicable to emphasize critical elements and adjust for the presentation.
-For example, anchor points and search points are less important and hence a smaller size and alpha is used; Alpha can also be applied on the interpolation paths to show the start to finish from transparent to opaque.
+For example, anchor points and search points are less important and hence a smaller size and alpha is used.
+Alpha can also be applied on the interpolation paths to show the start to finish from transparent to opaque.
 
@@
 
-Animation is another display to show how the search progresses from start to finish in the space and an  `animate = TRUE` argument is used to enable the animation in the PCA plot.
+Animation is another type of display to show how the search progresses from start to finish in the space.
+An  `animate = TRUE` argument is used to enable the animation in the PCA plot.
 Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
-An additional piece of information one can learn from this animation is that SA finds its end basis quicker than PD since SA finishes its search in the 5th frame while PD is still making more progression.
+An additional piece of information one can learn from this animation is that SA finds its end basis quicker than PD since SA finishes its search in the 5th frame while PD is still making more progress.
 
@@
 
-```{r toy-tour, fig.cap="Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+```{r toy-tour, fig.cap="Six selected frames from the tour animation for viewing the same two optimisations as previous two figures.
 The tour animation allows for an appreciation of the search bases in the original high dimensional space.
@@
 
-As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in \CRANpkg{geozoo} [@geozoo] package.
-While the PCA plot projects the bases from the direction that maximises the variance, the tour plot display of the original high dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows frames from the tour plot of the same two optimisations in its original 5D space.
+As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \CRANpkg{geozoo} [@geozoo] package.
+While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.
+Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original 5D space.
 
@@
 
-Something to fill here ...
+\textcolor{red}{Something to fill here ...}
 
@@
 
-Random variables with different distributions have been simulated and the distributional form of each variable is presented in Equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variable `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
+Random variables with different distributions have been simulated and the distributional form of each variable is presented in equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Variables `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
 All the variables have been scaled to have an overall unit variance before projection pursuit.
@@
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
-The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at Time 61 has a higher index value than the target in the third interpolations.
+The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at Time 61 has a higher index value than the target in the third interpolation.
 Rather than starting the next iteration from the target basis on Time 65, SA starts the next iteration at Time 61 on the right panel and reaches a better final basis.
@@
 
-```{r noisy-better-geo, fig.height = 4, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimiers.
-Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
+```{r noisy-better-geo, fig.height = 4, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimisers.
+Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
 ```
@@
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
-[A sentence for PD].
-Comparing the amount of random search done by SA and SAJO, it is surprising that SAJO doesn't carry that many of sampling as SA.
-Combine the insight from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SAJO first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
+\textcolor{red}{[A sentence for PD]}.
+Comparing the amount of random search done by SA and SAJO, it is surprising that SAJO does not carry as many samples as SA.
+Combining the insights from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SAJO first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
 The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as what SAJO has done, can lead to a more efficient search, in terms of the overall number of points evaluated.
@@
 Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
-Local optimum (x), is presented in this experiment, along with the global optimum (*).
+The local optimum (x), is presented in this experiment, along with the global optimum (*).
 The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing."}
@@
 With a larger search size of 0.7, more seeds have found the global maximum.
-Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to improve the final basis found, despite having longer interpolation paths.
+Comparing SA and SAJO for a search size of 0.5, SAJO does not seem to improve the final basis found, despite having longer interpolation paths.
 However, the denser paths near the local maximum is an indicator that SAJO is working hard to examine if there is any other optimum in the basis space but the relative small search size has diminished its ability to reach the global maximum.
@@
   - \code{theme\_fern()} and \code{format\_label()} for better display of the grid lines and axis formatting
-  - \code{clean\_method()} for clean up the name of the optimisers
-  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia
+  - \code{clean\_method()} to clean up the name of the optimisers
+  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia
   - \code{botanical\_pal()} as the colour interpolator 
@@
 <!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \CRANpkg{tourr}[@tourr] package; manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
+Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \CRANpkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
 Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.
@@
 
-This article is created using \CRANpkg{knitr}[@knitr] and \CRANpkg{rmarkdown} [@rmarkdown]  in R.
+This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown]  in R.
 The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-tour-vis}.

### 2021-02-04 paper/zhang-cook-laa-langrene-menendez.Rmd
changes on some highlighted places and a minor tweak of search plot

@@
 
-In projection pursuit the optimisation problem \textcolor{red}{(aims at ...
-describe it in simple words.)} That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
+In projection pursuit the optimisation problem aims at finding the global optimum that gives the most interesting projection according to an index function.
+That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
 
@@
 
-\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in section \textcolor{red}{XX}, and  $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
+\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in the previous section, and  $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
 Several features of this optimisation are worth-noticing.

### 2021-02-04 paper/zhang-cook-laa-langrene-menendez.tex
changes on some highlighted places and a minor tweak of search plot

--- 2021-02-03 paper/zhang-cook-laa-langrene-menendez.tex
@@
 specialised in non-linear optimisation is \CRANpkg{nloptr}
-\citep{nloptr}. Specific solvers for simulated annealing includes
+\citep{nloptr}. Specific solvers for simulated annealing include
 \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA}
@@
 interesting structures (e.g.~clustering, outliers and skewness) in
-multivariate data via low dimensions projection. Guided tour is one
+multivariate data via low-dimensional projections. Guided tour is one
 variation of a broader class of data visualisation methods, tour, which
@@
 Let \(\mathbf{X}_{n \times p}\) be the data matrix with \(n\)
-observations in \(p\) dimensions. A d-dimensional projection is a linear
-transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\) defined as
-\(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
+observations in \(p\) dimensions. A \(d\)-dimensional projection is a
+linear transformation from \(\mathbb{R}^p\) into \(\mathbb{R}^d\)
+defined as \(\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}\), where
 \(\mathbf{Y}_{n \times d}\) is the projected data and
@@
 \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
-function that maps the projected data \(\mathbf{Y}\) onto an scalar
+function that maps the projected data \(\mathbf{Y}\) onto a scalar
 value. This is commonly known as the projection pursuit index function,
@@
 the literature to detect different data structures
-\citep[friedman1974projection;][]{hall1989polynomial, cook1993projection, posse1995projection}
+\citep{friedman1974projection, hall1989polynomial, cook1993projection, posse1995projection}
 and to classify groups \citep{lee2005projection, lee2010projection}.
@@
 As a general visualisation method, tour produces animations of high
-dimensional data via rotations of low dimension planes. There are
+dimensional data via rotations of low-dimensional planes. There are
 different tour implementations depending on how the high dimensional
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/pmen0008/Documents/paper-tour-vis/img/tour-path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour-path}
 
@@
 
-The optimisation problem in projection pursuit is stated as follows:
-Given a randomly start basis \(\mathbf{A}_1\), projection pursuit finds
-the final projection basis \(\mathbf{A}_T\) that satisfies the following
+In projection pursuit the optimisation problem aims at finding the
+global optimum that gives the most interesting projection according to
+an index function. That is, it starts with a given randomly selected
+basis \(\mathbf{A}_1\) and aims at finding an optimal final projection
+basis \(\mathbf{A}_T\) in each iteration that satisfies the following
 optimisation problem:
@@
 
-\noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
-constraint requires the projection bases \(\mathbf{A}\) to be
-orthonormal. Several features of this optimisation are worth-noticing.
-1) The optimisation is constrained and the space of the bases is a
-d-dimensional sphere. 2) The objective function may not be
-differentiable for a constructed index function. 3) Although finding the
-global maximum is the goal of an optimisation problem, it is also
-interesting to inspect local maximum in projection pursuit since it
+\noindent where \(f\), \(\mathcal{A}\), and \(\mathbf{X}\) are defined
+as in the previous section, and \(I_d\) is the \(d\)-dimensional
+identity matrix and the constraint requires the projection bases
+\(\mathbf{A}\) to be orthonormal. Several features of this optimisation
+are worth-noticing. 1) The optimisation is constrained and the space of
+the bases is a \(d\)-dimensional sphere. 2) The objective function may
+not be differentiable for a constructed index function. 3) Although
+finding the global maximum is the goal of an optimisation problem, it is
+also interesting to inspect local maximum in projection pursuit since it
 could present unexpected interesting projections. 4) The optimisation
@@
 optimisers: 1) projection bases \(\mathbf{A}\), 2) index values \(I\),
-and 3) state \(S\). For SA and SAJO, possible state include
+and 3) state \(S\). For SA and SAJO, possible states include
 \texttt{random\_search}, \texttt{new\_basis}, and
 \texttt{interpolation}. Pseudo derivative (PD) has a wider variety of
-state including \texttt{new\_basis}, \texttt{direction\_search},
+states including \texttt{new\_basis}, \texttt{direction\_search},
 \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
@@
 size for searching observations. A matrix notation of the data structure
-is presented in Equation \ref{eq:data-structure}.
+is presented in equation \ref{eq:data-structure}.
 
@@
 \noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
-no output in iteration \(J + 1\) since the optimiser doesn't find a
+no output in iteration \(J + 1\) since the optimiser does not find a
 better basis in the last iteration and terminates. The final basis found
@@
 A starting point of diagnosing an optimiser is to understand how many
-search an optimiser has conducted. One may want to simply plot the index
-value of the search points across its natural order, but a point
+searches an optimiser has conducted. One may want to simply plot the
+index value of the search points across its natural order, but a point
 geometry may work well if each iteration has a similar number of points.
@@
 space. An alternative is to summarise the search in each iteration using
-boxplot and each iteration will then be spaced out equally.
+boxplots and each iteration will then be spaced out equally.
 Occasionally, one may still want to switch back to a point geometry if
@@
 Additional annotations are added to facilitate better reading of the
-plot and these includes\\
+plot and these include\\
 1) the number of points searched in each iteration can be added as text
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-search-1}
-
-}
-
-\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+
+}
+
+\caption[A comparison of search by two optimisers]{A comparison of search by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
+Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 \caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
-Both traces are smooth while the change of index value in each iteration may not be monotinic.}\label{fig:toy-interp}
+Both traces are smooth while the change of index value in each iteration may not be monotonic.}\label{fig:toy-interp}
 \end{figure}
@@
 
-Another interest of diagnostic is to examine how the index value
-improves for interpolating bases since the projection on these bases is
-played by the tour animation. Trace plots are created with plotting the
-index value of the interpolation points against time. Figure
+Another interesting diagnostic is to examine how the index value
+improves between interpolating bases since the projection on these bases
+is played by the tour animation. Trace plots are created with plotting
+the index value of the interpolation points against time. Figure
 \ref{fig:toy-interp} presents the trace plot of the same optimisation
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
 
@@
 interest. Given the orthonormality constraint, the space of projection
-bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\) dimension sphere
-and dimension reduction methods, i.e.~principal component analysis is
-applied to first project all the bases onto a 2D space. In a projection
-pursuit guided tour optimisation, there are various type of bases
-involved: 1) The start basis; 2) The search bases that an optimiser
-evaluated to produce the anchor bases; 3) The anchor bases that have the
-highest index value in each iteration; 4) The interpolating bases on the
-interpolation path; and finally 5) the end basis. The importance of
-these basis differs and the most important ones are the start,
-interpolating and end bases. Anchor and search bases can be turned on
-with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers can
-start with the same basis but finish with bases of opposite sign. This
-happens because the projection is invariant to the sign difference of
-the bases and so does the index value, however, this creates
-difficulties for comparing the optimisers since the end bases will be
-symmetric with respect to the origin. A sign flipping step is conducted
-to flip the signs of all the bases in one routine if different
-optimisations finish at opposite places.
+bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\)-dimensional
+sphere. A dimensionality reduction method, e.g.~principal component
+analysis is applied to first project all the bases onto a 2D space. In a
+projection pursuit guided tour optimisation, there are various types of
+bases involved: 1) The starting basis; 2) The search bases that the
+optimiser evaluated to produce the anchor bases; 3) The anchor bases
+that have the highest index value in each iteration; 4) The
+interpolating bases on the interpolation path; and finally 5) the end
+basis. The importance of these bases differs and the most important ones
+are the starting, interpolating and end bases. The anchor and search
+bases can be turned on with argument \texttt{details\ =\ TRUE}.
+Sometimes, two optimisers can start with the same basis but finish with
+bases of opposite sign. This happens because the projection is invariant
+to the sign difference of the bases and so does the index value,
+however, this creates difficulties for comparing the optimisers since
+the end bases will be symmetric with respect to the origin. A sign
+flipping step is conducted to flip the signs of all the bases in one
+routine if different optimisations finish at opposite places.
 
 Several annotations have been made to help understanding this plot. As
-mentioned previously, the original basis space is a high dimensional
+mentioned previously, the original basis space is a high-dimensional
 sphere and random bases on the sphere can be generated via the CRAN
@@
 largest distance from the centre to the basis. The theoretical best
-basis is known for simulated data and can be labelled to compare the how
+basis is known for simulated data and can be labelled to compare how
 close the end basis found by the optimisers is to the theoretical best.
@@
 example, anchor points and search points are less important and hence a
-smaller size and alpha is used; Alpha can also be applied on the
+smaller size and alpha is used. Alpha can also be applied on the
 interpolation paths to show the start to finish from transparent to
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-Animation is another display to show how the search progresses from
-start to finish in the space and an \texttt{animate\ =\ TRUE} argument
+Animation is another type of display to show how the search progresses
+from start to finish in the space. An \texttt{animate\ =\ TRUE} argument
 is used to enable the animation in the PCA plot. Figure
@@
 than PD since SA finishes its search in the 5th frame while PD is still
-making more progression.
+making more progress.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/toy-tour-1}
-
-}
-
-\caption{Six selected frames from the tour animation for viewing the same two optmisations as previous two figures.
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+
+}
+
+\caption{Six selected frames from the tour animation for viewing the same two optimisations as previous two figures.
 The tour animation allows for an appreciation of the search bases in the original high dimensional space.
@@
 As mentioned previously, the original \(p \times d\) dimension space can
-be simulated via randomly generated bases in \CRANpkg{geozoo}
+be simulated via randomly generated bases in the \CRANpkg{geozoo}
 \citep{geozoo} package. While the PCA plot projects the bases from the
-direction that maximises the variance, the tour plot display of the
-original high dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows frames from the tour plot of the same
-two optimisations in its original 5D space.
+direction that maximises the variance, the tour plot displays the
+original high-dimensional space from various directions using animation.
+Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
+same two optimisations in its original 5D space.
 
@@
 
-Something to fill here \ldots{}
+\textcolor{red}{Something to fill here ...}
 
@@
 Random variables with different distributions have been simulated and
-the distributional form of each variable is presented in Equations
-\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8},
-\texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
-variance and \texttt{x2} to \texttt{x7} are normal mixtures with varied
-weights and locations. All the variables have been scaled to have an
-overall unit variance before projection pursuit.
+the distributional form of each variable is presented in equations
+\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1},
+\texttt{x8}, \texttt{x9} and \texttt{x10} are normal noise with zero
+mean and unit variance and \texttt{x2} to \texttt{x7} are normal
+mixtures with varied weights and locations. All the variables have been
+scaled to have an overall unit variance before projection pursuit.
 
@@
 identical, the basis at Time 61 has a higher index value than the target
-in the third interpolations. Rather than starting the next iteration
-from the target basis on Time 65, SA starts the next iteration at Time
-61 on the right panel and reaches a better final basis.
+in the third interpolation. Rather than starting the next iteration from
+the target basis on Time 65, SA starts the next iteration at Time 61 on
+the right panel and reaches a better final basis.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/noisy-better-geo-1}
-
-}
-
-\caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimiers.
-Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO success in optimise the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+
+}
+
+\caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers.
+Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 the interpolation path and search points excluding the last termination
-iteration. {[}A sentence for PD{]}. Comparing the amount of random
-search done by SA and SAJO, it is surprising that SAJO doesn't carry
-that many of sampling as SA. Combine the insight from both the trace and
-PCA plot, one can learn the two different search strategies by SA and
-SAJO: SA frequently samples in the space and only make a move when an
-improvement is guaranteed to be made, while SAJO first broadly accepts
-bases in the space and then starts the extensive sampling in a narrowed
-subspace. The specific decision of which optimiser to use will depend on
-the index curve in the basis space but if the basis space is non-smooth,
-accepting inferior bases at first, as what SAJO has done, can lead to a
-more efficient search, in terms of the overall number of points
-evaluated.
+iteration. \textcolor{red}{[A sentence for PD]}. Comparing the amount of
+random search done by SA and SAJO, it is surprising that SAJO does not
+carry as many samples as SA. Combining the insights from both the trace
+and PCA plot, one can learn the two different search strategies by SA
+and SAJO: SA frequently samples in the space and only make a move when
+an improvement is guaranteed to be made, while SAJO first broadly
+accepts bases in the space and then starts the extensive sampling in a
+narrowed subspace. The specific decision of which optimiser to use will
+depend on the index curve in the basis space but if the basis space is
+non-smooth, accepting inferior bases at first, as what SAJO has done,
+can lead to a more efficient search, in terms of the overall number of
+points evaluated.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Documents/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
 
@@
 Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
-Local optimum (x), is presented in this experiment, along with the global optimum (*).
+The local optimum (x), is presented in this experiment, along with the global optimum (*).
 The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
@@
 larger search size of 0.7, more seeds have found the global maximum.
-Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to
+Comparing SA and SAJO for a search size of 0.5, SAJO does not seem to
 improve the final basis found, despite having longer interpolation
@@
   \item
-    \code{clean\_method()} for clean up the name of the optimisers
+    \code{clean\_method()} to clean up the name of the optimisers
   \item
     \code{botanical\_palettes} is a collection of colour palettes from
-    Australian native plants Quantitative palettes include daisy,
+    Australian native plants. Quantitative palettes include daisy,
     banksia and cherry and sequential palettes contain fern and acacia
@@
 packages: the collection of the data object is implemented in the
-existing \CRANpkg{tourr}\citep{tourr} package; manipulation and
-visualisation of the data object are implemented in the new \pkg{ferrn}
-package. Equipped with handy tools to diagnose the performance of
-optimisers, future work can extend the diagnostics to a wider range of
-index functions, i.e.~scagnostics, association, and information index
-\citep{laa2020using} and understand how the optimisers behave for index
-functions with different structures.
+existing \CRANpkg{tourr} \citep{tourr} package; while the manipulation
+and visualisation of the data object are implemented in the new
+\pkg{ferrn} package. Equipped with handy tools to diagnose the
+performance of optimisers, future work can extend the diagnostics to a
+wider range of index functions, i.e.~scagnostics, association, and
+information index \citep{laa2020using} and understand how the optimisers
+behave for index functions with different structures.
 
@@
 
-This article is created using \CRANpkg{knitr}\citep{knitr} and
+This article is created using \CRANpkg{knitr} \citep{knitr} and
 \CRANpkg{rmarkdown} \citep{rmarkdown} in R. The source code for

### 2021-02-05 paper/zhang-cook-laa-langrene-menendez.Rmd
minor fixes: keywords,interestingness, optim prob and captions

@@
     email:  patricia.menendez@monash.edu 
-
-abstract: >
+abstract: |
   Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
@@
 An R package, \pkg{ferrn}, has been created to implement the diagnostics.
+
+  \noindent \textit{keywords}: graphics, multivariate, optimization
 preamble: >
@@
 
-<!-- Notation is first established before defining projection pursuit. -->
 Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions.
@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-A number of index functions have been proposed in the literature to detect different data structures [@friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
+In the context of projection pursuit, interestingness means departure from normality [@diaconis1984asymptotics; @huber1985projection].
+\textcolor{red}{only include a couple of indices, using e.g.}A number of index functions have been proposed in the literature to detect different data structures [@friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
 
@@
 
-```{r tour-path, out.height="20%",out.width="50%", fig.cap="Each square (frame) represents the projected data with a corresponding basis.
+```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the steps in a tour path.
+Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
@@
 
-In projection pursuit the optimisation problem aims at finding the global optimum that gives the most interesting projection according to an index function.
+In projection pursuit the optimisation problem aims at finding the global and local maxima that give interesting projections according to an index function.
 That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
@@
 \begin{align}
-&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
-&s.t.  \mathbf{A}^{\prime} \mathbf{A} = I_d
+\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A})  ~~~ s.t. ~~~ \mathbf{A}^{\prime} \mathbf{A} = I_d
 \end{align}
@@
 
-```{r toy-search, fig.cap="A comparison of search by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
+```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
 Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
@@
 
-```{r toy-interp, fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
-Both traces are smooth while the change of index value in each iteration may not be monotonic."}
+```{r toy-interp, fig.cap = "Examing the index values as the optimisation progress for two optimisers using trace plot.
+The lines indicate the interpolation and the dots indicate a new target is generated."}
 ```
+
+<!-- The right hand side has the jump out, which has a lower index value - also explain this in the text. -->
+
+<!-- Both traces are smooth while the change of index value in each iteration may not be monotonic. -->
 

### 2021-02-05 paper/zhang-cook-laa-langrene-menendez.tex
minor fixes: keywords,interestingness, optim prob and captions

@@
 \pkg{ferrn}, has been created to implement the diagnostics.
+
+\noindent \textit{keywords}: graphics, multivariate, optimization
 }
@@
 or just index function, and is used to measure the ``interestingness''
-of a given projection. A number of index functions have been proposed in
-the literature to detect different data structures
+of a given projection. In the context of projection pursuit,
+interestingness means departure from normality
+\citep{diaconis1984asymptotics, huber1985projection}.
+\textcolor{red}{only include a couple of indices, using e.g.}A number of
+index functions have been proposed in the literature to detect different
+data structures
 \citep{friedman1974projection, hall1989polynomial, cook1993projection, posse1995projection}
@@
 
-\caption[Each square (frame) represents the projected data with a corresponding basis]{Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
+\caption[An illustration for demonstrating the steps in a tour path]{An illustration for demonstrating the steps in a tour path.
+Each square (frame) represents the projected data with a corresponding basis.
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}(\#fig:tour-path)
 \end{figure}
@@
 In projection pursuit the optimisation problem aims at finding the
-global optimum that gives the most interesting projection according to
+global and local maxima that give interesting projections according to
 an index function. That is, it starts with a given randomly selected
@@
 
-\caption[A comparison of search by two optimisers]{A comparison of search by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
+\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
+Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}(\#fig:toy-search)
 \end{figure}
@@
 
-\caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure.
-Both traces are smooth while the change of index value in each iteration may not be monotonic.}\label{fig:toy-interp}
+\caption[Examing the index values as the optimisation progress for two optimisers using trace plot]{Examing the index values as the optimisation progress for two optimisers using trace plot.
+The lines indicate the interpolation and the dots indicate a new target is generated.}(\#fig:toy-interp)
 \end{figure}
@@
 \caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD.
-All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
+All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}(\#fig:toy-pca)
 \end{figure}
@@
 With animation, the start and end position of each optimisation is easier to identify.
-A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
+A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}(\#fig:toy-pca-animated)
 \end{figure}

### 2021-02-06 paper/zhang-cook-laa-langrene-menendez.Rmd
orientation check section and small tweaks

@@
     \output{$\mathbf{A}_{l}$}
-    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
   \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   set $l = 1$\;
+   Set $l = 1$\;
   \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    update $l = l + 1$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    Update $l = l + 1$\;
   }
-  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
+  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
 }
@@
 \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
-    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-    draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-    update $l = l + 1$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+    Update $l = l + 1$\;
   }
@@
   Determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-  update $l = l + 1$\;
+  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
+  Update $l = l + 1$\;
 }
@@
 
-## Reconciling the orientation
-
-Something to fill here ...
-
-```{r flip-sign, fig.cap = "PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463."}
-```
-
 ## Seeing the signal in the noise
@@
 This indicates that SAJO, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the global optimum.
+
+## Reconciling the orientation
+
+```{r flip-sign, fig.cap = "Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463.
+On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory.
+After implementing an orientation check on the determinant, this phenomenon disappears on the right panel."}
+```
+
+One interesting situation observed in the previous examples is that, for some simulations as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
+As mentioned previously, bases with opposite signs do not affect the projection and index value, but clearly, we prefer the target to have the same orientation as the current basis.
+The orientation of two bases can be checked via calculating the determinant and a negative value indicates opposite direction.
+Hence an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
+The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exist.
 

### 2021-02-06 paper/zhang-cook-laa-langrene-menendez.tex
orientation check section and small tweaks

@@
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}(\#fig:tour-path)
+Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
@@
     \output{$\mathbf{A}_{l}$}
-    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
   \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   set $l = 1$\;
+   Set $l = 1$\;
   \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    update $l = l + 1$\;
-  }
-  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    Update $l = l + 1$\;
+  }
+  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
 }
@@
 \caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}(\#fig:toy-search)
+Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
 \end{figure}
@@
 \caption[Examing the index values as the optimisation progress for two optimisers using trace plot]{Examing the index values as the optimisation progress for two optimisers using trace plot.
-The lines indicate the interpolation and the dots indicate a new target is generated.}(\#fig:toy-interp)
+The lines indicate the interpolation and the dots indicate a new target is generated.}\label{fig:toy-interp}
 \end{figure}
@@
 \caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD.
-All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}(\#fig:toy-pca)
+All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
 \end{figure}
@@
 With animation, the start and end position of each optimisation is easier to identify.
-A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}(\#fig:toy-pca-animated)
+A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
 \end{figure}
@@
 vertical axis.
-
-\hypertarget{reconciling-the-orientation}{%
-\subsection{Reconciling the
-orientation}\label{reconciling-the-orientation}}
-
-Something to fill here \ldots{}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
-
-}
-
-\caption[PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463]{PCA plot of 1D projection of boa6 data, optimised by SA, with seed 2463.}\label{fig:flip-sign}
-\end{figure}
-\end{Schunk}
 
@@
 
+\hypertarget{reconciling-the-orientation}{%
+\subsection{Reconciling the
+orientation}\label{reconciling-the-orientation}}
+
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+
+}
+
+\caption[Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463]{Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463.
+On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory.
+After implementing an orientation check on the determinant, this phenomenon disappears on the right panel.}\label{fig:flip-sign}
+\end{figure}
+\end{Schunk}
+
+One interesting situation observed in the previous examples is that, for
+some simulations as shown on the left panel of Figure
+\ref{fig:flip-sign}, the target basis is generated on the other half of
+the basis space and the interpolator seems to draw a straight line to
+interpolate. As mentioned previously, bases with opposite signs do not
+affect the projection and index value, but clearly, we prefer the target
+to have the same orientation as the current basis. The orientation of
+two bases can be checked via calculating the determinant and a negative
+value indicates opposite direction. Hence an orientation check is
+carried out once a new target basis is generated and the sign in the
+target basis will be flipped if a negative determinant is obtained. The
+interpolation after implementing the orientation check is shown on the
+right panel of Figure \ref{fig:flip-sign} where the unsatisfactory
+interpolation no longer exist.
+
 \hypertarget{implementation}{%

### 2021-02-06 scripts/diagnose.R

@@ -69,7 +69,7 @@ p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
   clean_method() %>%
   mutate(method = factor(method, levels = c("SA", "polish"))) %>%
   get_interp() %>%
-  explore_trace_interp(color = method, cutoff = 100) +
+  explore_trace_interp(color = method, cutoff = 100, target_size = 2, interp_size = 2) +
   scale_color_discrete_botanical(breaks = c("SA", "polish"), label = c("SA", "polish")) +
   theme(legend.position = "bottom",
         plot.margin = margin(2.5, 2, 0, 0, "cm"))
@@ -100,9 +100,14 @@ p1 +
 # save(orientation_corrected, file = here::here("data", "orientation_corrected.rda"))
 # save(orientation_different, file = here::here("data", "orientation_different.rda"))
 
-bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
+dt <- bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
           orientation_different %>% mutate(sign = "original")) %>%
-  explore_space_pca(group = sign, flip = FALSE, start_size = 3) +
+  compute_pca() %>%
+  pluck("aug")
+
+dt %>%
+  explore_space_pca(group = sign, pca = FALSE ,flip = FALSE, start_size = 3) +
+  add_anchor(dt = get_interp_last(dt), anchor_color = sign) +
   scale_color_discrete_botanical() +
   facet_wrap(vars(fct_relevel(sign, c("original", "flipped")))) +
   theme(legend.position = "none")

### 2021-02-07 paper/zhang-cook-laa-langrene-menendez.Rmd
fix captions and regenerate plots

@@
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
+Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation."}
 include_graphics(here::here("img/tour-path.png"))
@@
 
-```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
+```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6}, using the holes index.
+Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
 ```
@@
 
-```{r toy-interp, fig.cap = "Examing the index values as the optimisation progress for two optimisers using trace plot.
-The lines indicate the interpolation and the dots indicate a new target is generated."}
+```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
+The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
+The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
 ```
@@
 
-```{r toy-pca, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes index with two optimisers: SA and PD.
-All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases."}
+```{r toy-pca, fig.cap = "The PCA-projected basis space with two search paths optimised by SA and PD.
+The holes index is optimised for the 1D projection problem on the five-variable dataset, \\code{boa5}.
+All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
 ```
@@
 
-```{r toy-pca-animated, fig.cap = "Six selected frames from the animated PCA plots in the previous figure.
-With animation, the start and end position of each optimisation is easier to identify.
-A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
+```{r toy-pca-animated, fig.cap = "Six frames selected from the animation of search paths in the PCA-projected basis space.
+With animation, the start and end position of each optimisation routine is clearly identified.
+The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
 ```
@@
 
-```{r toy-tour, fig.cap="Six selected frames from the tour animation for viewing the same two optimisations as previous two figures.
-The tour animation allows for an appreciation of the search bases in the original high dimensional space.
-A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
+```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
+The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
 ```
@@
 
-```{r interruption, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised by optimiser SA."}
+```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data with the holes index, optimised by SA.
+On the left panel, basis with higher index value is found during the interpolation but not used.
+The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
 ```
@@
 
-```{r polish, fig.height= 5, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised first by SA and then by polish search.
-The polish step results in clearer cut on the edges of the four clusters in the data structure."}
+```{r polish, fig.height= 5, fig.cap = "The interpolation of a 2D projection problem on \\code{boa6} data with the holes index being first optimised by SA and then then polished.
+The projection by the final baiss of each algorithm is also presented.
+Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge.."}
 ```
@@
 
-```{r noisy-better-geo, fig.height = 4, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimisers.
-Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
+```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \\code{boa5}.
+Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best."}
 ```
@@
 
-```{r kol-result, fig.height = 4, fig.cap="PCA plot of 1D projection on the 6\\-variable data, \\code{boa6}, with normal kolmogorov index.
-Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
-The local optimum (x), is presented in this experiment, along with the global optimum (*).
-The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing."}
+```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+The optimisation is done on the six-variable data, \\code{boa6} with 1D projecton index $I^{nk}(n)$.
+A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).
+The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing."}
 ```
@@
 
-```{r flip-sign, fig.cap = "Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463.
-On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory.
-After implementing an orientation check on the determinant, this phenomenon disappears on the right panel."}
+```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis.
+Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463.
+The dots represent the target basis in each iteration and the path shows the interpolation.
+On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable.
+An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis.
+The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears."}
 ```

### 2021-02-07 paper/zhang-cook-laa-langrene-menendez.tex
fix captions and regenerate plots

@@
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
+Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
@@
 
-\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index.
-Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
+\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6}, using the holes index.
+Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
 \end{figure}
@@
 
-\caption[Examing the index values as the optimisation progress for two optimisers using trace plot]{Examing the index values as the optimisation progress for two optimisers using trace plot.
-The lines indicate the interpolation and the dots indicate a new target is generated.}\label{fig:toy-interp}
+\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
+The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
+The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
 \end{figure}
@@
 
-\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD.
-All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
+\caption[The PCA-projected basis space with two search paths optimised by SA and PD]{The PCA-projected basis space with two search paths optimised by SA and PD.
+The holes index is optimised for the 1D projection problem on the five-variable dataset, \code{boa5}.
+All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
 \end{figure}
@@
 
-\caption{Six selected frames from the animated PCA plots in the previous figure.
-With animation, the start and end position of each optimisation is easier to identify.
-A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
+\caption{Six frames selected from the animation of search paths in the PCA-projected basis space.
+With animation, the start and end position of each optimisation routine is clearly identified.
+The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
 \end{figure}
@@
 
-\caption{Six selected frames from the tour animation for viewing the same two optimisations as previous two figures.
-The tour animation allows for an appreciation of the search bases in the original high dimensional space.
-A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
+\caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
 \end{figure}
@@
 
-\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA.}\label{fig:interruption}
+\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA.
+On the left panel, basis with higher index value is found during the interpolation but not used.
+The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
 \end{figure}
@@
 
-\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search.
-The polish step results in clearer cut on the edges of the four clusters in the data structure.}\label{fig:polish}
+\caption[The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished]{The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished.
+The projection by the final baiss of each algorithm is also presented.
+Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge..}\label{fig:polish}
 \end{figure}
@@
 
-\caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers.
-Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
+\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}.
+Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 
-\caption[PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index]{PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index.
-Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates.
-The local optimum (x), is presented in this experiment, along with the global optimum (*).
-The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
+\caption[Comparing 20 search paths in the PCA-projected basis space with two optimisers]{Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+The optimisation is done on the six-variable data, \code{boa6} with 1D projecton index $I^{nk}(n)$.
+A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).
+The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
 \end{figure}
@@
 
-\caption[Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463]{Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463.
-On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory.
-After implementing an orientation check on the determinant, this phenomenon disappears on the right panel.}\label{fig:flip-sign}
+\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis.
+Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463.
+The dots represent the target basis in each iteration and the path shows the interpolation.
+On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable.
+An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis.
+The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
 \end{figure}

### 2021-02-07 scripts/diagnose.R

@@ -160,7 +160,7 @@ dt <- dplyr::bind_rows(kol_1d_geo, kol_1d_better, kol_1d_better_random) %>%
   bind_theoretical(matrix = matrix(c(0, -1, 0, 0, 0), ncol = 1), tourr::norm_kol(nrow(boa5)), raw_data = boa5)
 
 p1 <- dt %>%
-  explore_trace_interp(group = method, color = method) +
+  explore_trace_interp(group = method, color = method, accuracy_y = 0.02) +
   geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) +
   scale_color_discrete_botanical() +
   facet_wrap(vars(method), scales = "free_x")

### 2021-02-09 paper/zhang-cook-laa-langrene-menendez.Rmd
changes after talking to patricia

@@
 In projection pursuit the optimisation problem aims at finding the global and local maxima that give interesting projections according to an index function.
-That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ in each iteration that satisfies the following optimisation problem:
+That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
 
@@
 
-\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in the previous section, and  $I_d$ is the $d$-dimensional identity matrix and the constraint requires the projection bases $\mathbf{A}$ to be orthonormal.
-Several features of this optimisation are worth-noticing.
-1) The optimisation is constrained and the space of the bases is a $d$-dimensional sphere.
-2) The objective function may not be differentiable for a constructed index function.
-3) Although finding the global maximum is the goal of an optimisation problem, it is also interesting to inspect local maximum in projection pursuit since it could present unexpected interesting projections.
-4) The optimisation procedure needs to be easy to compute since the tour animation needs to be played in real-time during the optimisation.
+\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in the previous section,  $I_d$ is the $d$-dimensional identity matrix and the constraint ensures the projection bases $\mathbf{A}$ to be orthonormal.
+It is worth noticing the following: 1) The optimisation is constrained and the basis space is a $d$-dimensional sphere.
+2) There may be index functions for which the objective function might not be differentiable.
+3) While finding the global optimum is the goal of the optimisation problem, interesting projections may also appear in the local optimum.
+4) The optimisation should be computationally agile since the tour animation is viewed by the users during the optimisation.
 
@@
 
-Before making diagnostic plots, information generated by the optimisation need to be collected and a pre-defined data structure solidifies the information being collected and streamlines the subsequent plot-making.
+A data structure for diagnosing optimisers in projection pursuit guided tour is first defined.
+With a fixed data structure, the diagnostic plots presented later in this paper can be easily applied to diagnose other optimisation procedures.
 
@@
 The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
-The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
+The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
 ```
 
-<!-- The right hand side has the jump out, which has a lower index value - also explain this in the text. -->
-
-<!-- Both traces are smooth while the change of index value in each iteration may not be monotonic. -->
-
 Another interesting diagnostic is to examine how the index value improves between interpolating bases since the projection on these bases is played by the tour animation.
-Trace plots are created with plotting the index value of the interpolation points against time.
-Figure \ref{fig:toy-interp} presents the trace plot of the same optimisation routine as Figure \ref{fig:toy-search}.
-The trace plot is smooth and an interesting discovery is that in SAJO, the interpolation in iteration 6-8 is not monotonic.
-Sometimes, this non-monotonicity is acceptable since the target basis does not necessarily have a higher index value than the current one, while in some cases, the optimisation will be better off if these intermediate basis can be used as the start of the next iteration (see section \ref{monotonic} for a detailed discussion on this).
+Trace plots are created by plotting the index value against time.
+Figure \ref{fig:toy-interp} presents the trace plot of the same optimisers as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
+It may seem bizarre at the first sight that the interpolation in the left panel first passes some bases with higher index value before decreases to a lower target.
+This happens because on the one hand, the probabilistic acceptance in SAJO implies that some worse bases will be accepted by the optimiser.
+On the other hand, guided tour interpolate between the current and target basis to provide a smooth transition between projections.
+These two facts indicate that a non-monotonic interpolation can't be avoided for SAJO.
+Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for SA.
 
@@
 With animation, the start and end position of each optimisation routine is clearly identified.
-The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
+The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}."}
 ```
@@
 ```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
-The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
+The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}."}
 ```

### 2021-02-09 paper/zhang-cook-laa-langrene-menendez.tex
changes after talking to patricia

@@
 basis \(\mathbf{A}_1\) and aims at finding an optimal final projection
-basis \(\mathbf{A}_T\) in each iteration that satisfies the following
-optimisation problem:
+basis \(\mathbf{A}_T\) that satisfies the following optimisation
+problem:
 
@@
 \noindent where \(f\), \(\mathcal{A}\), and \(\mathbf{X}\) are defined
-as in the previous section, and \(I_d\) is the \(d\)-dimensional
-identity matrix and the constraint requires the projection bases
-\(\mathbf{A}\) to be orthonormal. Several features of this optimisation
-are worth-noticing. 1) The optimisation is constrained and the space of
-the bases is a \(d\)-dimensional sphere. 2) The objective function may
-not be differentiable for a constructed index function. 3) Although
-finding the global maximum is the goal of an optimisation problem, it is
-also interesting to inspect local maximum in projection pursuit since it
-could present unexpected interesting projections. 4) The optimisation
-procedure needs to be easy to compute since the tour animation needs to
-be played in real-time during the optimisation.
+as in the previous section, \(I_d\) is the \(d\)-dimensional identity
+matrix and the constraint ensures the projection bases \(\mathbf{A}\) to
+be orthonormal. It is worth noticing the following: 1) The optimisation
+is constrained and the basis space is a \(d\)-dimensional sphere. 2)
+There may be index functions for which the objective function might not
+be differentiable. 3) While finding the global optimum is the goal of
+the optimisation problem, interesting projections may also appear in the
+local optimum. 4) The optimisation should be computationally agile since
+the tour animation is viewed by the users during the optimisation.
 
@@
 
-Before making diagnostic plots, information generated by the
-optimisation need to be collected and a pre-defined data structure
-solidifies the information being collected and streamlines the
-subsequent plot-making.
+A data structure for diagnosing optimisers in projection pursuit guided
+tour is first defined. With a fixed data structure, the diagnostic plots
+presented later in this paper can be easily applied to diagnose other
+optimisation procedures.
 
@@
 The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
-The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
+The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
 \end{figure}
@@
 improves between interpolating bases since the projection on these bases
-is played by the tour animation. Trace plots are created with plotting
-the index value of the interpolation points against time. Figure
-\ref{fig:toy-interp} presents the trace plot of the same optimisation
-routine as Figure \ref{fig:toy-search}. The trace plot is smooth and an
-interesting discovery is that in SAJO, the interpolation in iteration
-6-8 is not monotonic. Sometimes, this non-monotonicity is acceptable
-since the target basis does not necessarily have a higher index value
-than the current one, while in some cases, the optimisation will be
-better off if these intermediate basis can be used as the start of the
-next iteration (see section \ref{monotonic} for a detailed discussion on
-this).
+is played by the tour animation. Trace plots are created by plotting the
+index value against time. Figure \ref{fig:toy-interp} presents the trace
+plot of the same optimisers as Figure \ref{fig:toy-search} and one can
+observe that the trace is smooth in both cases. It may seem bizarre at
+the first sight that the interpolation in the left panel first passes
+some bases with higher index value before decreases to a lower target.
+This happens because on the one hand, the probabilistic acceptance in
+SAJO implies that some worse bases will be accepted by the optimiser. On
+the other hand, guided tour interpolate between the current and target
+basis to provide a smooth transition between projections. These two
+facts indicate that a non-monotonic interpolation can't be avoided for
+SAJO. Later, in section \ref{monotonic}, there will be a discussion on
+improving the non-monotonic interpolation for SA.
 
@@
 With animation, the start and end position of each optimisation routine is clearly identified.
-The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
+The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}.}\label{fig:toy-pca-animated}
 \end{figure}
@@
 \caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
-The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
+The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}.}\label{fig:toy-tour}
 \end{figure}

### 2021-02-09 scripts/toy.R

@@ -67,13 +67,11 @@ gl <-  lapply(rl, grid::rasterGrob)
 wrap_plots(gl)
 
 ## ----toy-tour
-# prep <- prep_space_tour(dplyr::bind_rows(holes_1d_better, holes_1d_geo) %>%
-#                      bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-#                                       index = tourr::holes(), raw_data = boa5),
-#                    group = method, palette = botanical_palettes$fern[c(1,6)], axes = "bottomleft",
-#                    point_size = 3, end_size = 8, theo_size = 8)
+# prep <- prep_space_tour(dplyr::bind_rows(holes_1d_better, holes_1d_geo), flip = TRUE,
+#                         group = method, palette = botanical_palettes$fern[c(1,6)], axes = "bottomleft",
+#                         point_size = 3, end_size = 8)
 #
-# set.seed(123)
+# set.seed(123456)
 # render(
 #   prep$basis,
 #   tour_path = grand_tour(),
@@ -87,7 +85,7 @@ wrap_plots(gl)
 # )
 #
 # # render gif
-# set.seed(123)
+# set.seed(123456)
 # render_gif(
 #   prep$basis,
 #   tour_path = grand_tour(),
@@ -99,7 +97,7 @@ wrap_plots(gl)
 #   gif_file = here::here("anim","tour.gif")
 # )
 
-frames <- c("001", "038", "046", "079", "096", "100")
+frames <- c("001", "021", "056", "072", "079", "088")
 ani <- paste0(here::here("anim/"), "tour/", "tour", frames, ".png")
 rl <- lapply(ani, png::readPNG)
 gl <-  lapply(rl, grid::rasterGrob)

### 2021-02-09 paper/zhang-cook-laa-langrene-menendez.Rmd
tour path on torus

@@
 
+```{r toy-torus, fig.cap="Six frames selected from animating 5000 randomly generated 2D projection bases.
+\\textcolor{red}{one more sentence here.} The full video of the animation can be found at \\url{https://vimeo.com/user133215810/review/510046144/49e902d8ef}.
+"}
+```
+
+The orthogonality of the axes in the basis corresponds to the two circles being perpendicular to each other in the torus.
+
 \textcolor{red}{Something to fill here ...}

### 2021-02-09 paper/zhang-cook-laa-langrene-menendez.tex
tour path on torus

@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+
+}
+
+\caption{Six frames selected from animating 5000 randomly generated 2D projection bases.
+\textcolor{red}{one more sentence here.} The full video of the animation can be found at \url{https://vimeo.com/user133215810/review/510046144/49e902d8ef}.
+}\label{fig:toy-torus}
+\end{figure}
+\end{Schunk}
+
+The orthogonality of the axes in the basis corresponds to the two
+circles being perpendicular to each other in the torus.
+
 \textcolor{red}{Something to fill here ...}

### 2021-02-09 scripts/toy.R

@@ -103,3 +103,90 @@ rl <- lapply(ani, png::readPNG)
 gl <-  lapply(rl, grid::rasterGrob)
 wrap_plots(gl)
 
+## ----toy-torus
+# proj_d <-  2 # 2D because then you have 2 orthogonal circles
+# d <- 6
+# n_point <- 5000
+# index <- holes()
+# set.seed(123456)
+# random <- map(1:n_point, ~basis_random(n = d/proj_d,  d=proj_d)) %>%
+#   purrr::flatten_dbl() %>% matrix(ncol = d, byrow = TRUE) %>% as_tibble()
+#
+# set.seed(123456)
+# holes_2d_geo_3var <-
+#   animate_xy(boa6[, 1:3], tour_path = guided_tour(holes(), d = 2,
+#                                            search_f =  search_geodesic),
+#              rescale = FALSE)
+#
+# last <- get_best(holes_2d_geo_3var)$basis %>% .[[1]]
+#
+# set.seed(123456)
+# holes_2d_geo_3var_polish <-
+#   animate_xy(boa6[, 1:3], tour_path = guided_tour(holes(), d = 2,
+#                                                   search_f =  search_polish,
+#                                                   start = last),
+#              rescale = FALSE)
+#
+# set.seed(123456)
+# holes_2d_better_3var <-
+#   animate_xy(boa6[, 1:3], tour_path = guided_tour(holes(), d = 2,
+#                                                   search_f =  search_better),
+#              rescale = FALSE)
+#
+# last <- get_best(holes_2d_better_3var)$basis %>% .[[1]]
+#
+# set.seed(123456)
+# holes_2d_better_3var_polish <-
+#   animate_xy(boa6[, 1:3], tour_path = guided_tour(holes(), d = 2,
+#                                                   search_f =  search_polish),
+#              rescale = FALSE, start = last)
+#
+# save(holes_2d_better_3var, file = here::here("data", "holes_2d_better_3var.rda"))
+# save(holes_2d_geo_3var, file = here::here("data", "holes_2d_geo_3var.rda"))
+#
+#
+# path_geo <- bind_rows(holes_2d_geo_3var) %>% get_interp() %>% get_basis_matrix() %>% as_tibble()
+# path_better <- bind_rows(holes_2d_better_3var) %>% get_interp() %>% get_basis_matrix() %>% as_tibble()
[…]

### 2021-02-11 scripts/diagnose.R
fix kol-geo-pca plot and writing

@@ -126,28 +126,28 @@ dt %>%
 #                                              search_f =  search_better, max.tries = 200),
 #                rescale = FALSE)
 #
-# last_basis <- get_best(kol_1d_better)$basis %>% .[[1]]
+# # last_basis <- get_best(kol_1d_better)$basis %>% .[[1]]
+# #
+# # set.seed(123456)
+# # kol_1d_better_polish <-
+# #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+# #                                              search_f = search_polish, max.tries = 200),
+# #                rescale = FALSE, start = last_basis)
 #
-# set.seed(123456)
-# kol_1d_better_polish <-
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
-#                                              search_f = search_polish, max.tries = 200,
-#                                              start = last_basis),
-#                rescale = FALSE)
 # set.seed(123456)
 # kol_1d_better_random <-
 #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
 #                                              search_f =  search_better_random, max.tries = 200),
 #                rescale = FALSE)
 #
-# last_basis <- get_best(kol_1d_better_random)$basis %>% .[[1]]
+# # last_basis <- get_best(kol_1d_better_random)$basis %>% .[[1]]
+# #
+# # set.seed(123456)
+# # kol_1d_better_random_polish <-
+# #   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
+# #                                              search_f =  search_polish, max.tries = 200),
+# #                rescale = FALSE, start = last_basis)
 #
-# set.seed(123456)
-# kol_1d_better_random_polish <-
-#   animate_dist(boa5, tour_path = guided_tour(norm_kol(nrow(boa5)), d = 1,
-#                                              search_f =  search_polish, max.tries = 200,
-#                                              start = last_basis),
-#                rescale = FALSE)
 # save(kol_1d_geo, file = here::here("data", "kol_1d_geo.rda"))
 # save(kol_1d_better, file = here::here("data", "kol_1d_better.rda"))
[…]

### 2021-02-16 paper/zhang-cook-laa-langrene-menendez.Rmd
fix kol-geo-pca plot and writing
write up torus section and update  vimeo links

--- 2021-02-09 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 With animation, the start and end position of each optimisation routine is clearly identified.
-The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}."}
+The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
 ```
@@
 ```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
-The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}."}
+The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
 ```
@@
 
-\textcolor{red}{Something to fill here ...}
+While the previous few examples have looked at the space of 1D bases in a unit sphere, this section visualises the space of 2D bases.
+Recall that the columns in a 2D basis are orthogonal to each other, the space of $p \times 2$ bases is a $p$-D torus.
+If take $p = 3$, one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
+The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality of the columns in the bases.
+Two paths optimised by SA and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous few 1D space plots.
+The final basis found by PD and SA are shown in a larger shape  and printed below respectively:
+
+```{r}
+holes_2d_geo_3var %>% get_best() %>% pull(basis) %>% .[[1]]
+```
+
+```{r}
+holes_2d_better_3var %>% get_best() %>% pull(basis) %>% .[[1]]
+```
+
+Both optimisers have found the third variable in the first column, but the second variable in the second column differs by a sign.One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and be symmetric when projected onto the other (due the the sign difference in the second column).
+In Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
+
+```{r toy-torus, fig.cap="Six frames selected from the torus view of 2D basis space.
+The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index.
+The grey points show the space of the 2D basis as a torus.
+The full video of the animation can be found at \\url{https://vimeo.com/512882784}."}
+```
 
@@
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
-\textcolor{red}{[A sentence for PD]}.
+Pseudo Derivative (PD) quickly gets stuck near the starting position.
 Comparing the amount of random search done by SA and SAJO, it is surprising that SAJO does not carry as many samples as SA.

### 2021-02-16 paper/zhang-cook-laa-langrene-menendez.tex
fix kol-geo-pca plot and writing
write up torus section and update  vimeo links

--- 2021-02-09 paper/zhang-cook-laa-langrene-menendez.tex
@@
 With animation, the start and end position of each optimisation routine is clearly identified.
-The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}.}\label{fig:toy-pca-animated}
+The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
 \end{figure}
@@
 \caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
-The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}.}\label{fig:toy-tour}
+The full video of the animation can be found at \url{https://vimeo.com/512885379}.}\label{fig:toy-tour}
 \end{figure}
@@
 
-\textcolor{red}{Something to fill here ...}
+While the previous few examples have looked at the space of 1D bases in
+a unit sphere, this section visualises the space of 2D bases. Recall
+that the columns in a 2D basis are orthogonal to each other, the space
+of \(p \times 2\) bases is a \(p\)-D torus. If take \(p = 3\), one would
+see a classical 3D torus shape as shown by the grey points in Figure
+\ref{fig:toy-torus}. The two circles of the torus can be observed to be
+perpendicular to each other and this can be linked back to the
+orthogonality of the columns in the bases. Two paths optimised by SA and
+PD are plotted on top of the torus and coloured in green and brown
+respectively to match the previous few 1D space plots. The final basis
+found by PD and SA are shown in a larger shape and printed below
+respectively:
+
+\begin{Schunk}
+\begin{Soutput}
+#>              [,1]        [,2]
+#> [1,]  0.001196285  0.03273881
+#> [2,] -0.242432715  0.96965761
+#> [3,] -0.970167484 -0.24226493
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Soutput}
+#>             [,1]         [,2]
+#> [1,]  0.05707994 -0.007220138
+#> [2,] -0.40196202 -0.915510160
+#> [3,] -0.91387549  0.402230054
+\end{Soutput}
+\end{Schunk}
+
+Both optimisers have found the third variable in the first column, but
+the second variable in the second column differs by a sign.One would
+expect to see this in the torus plot as the final bases match each other
+when projected onto one torus circle (due to the same sign in the first
+column) and be symmetric when projected onto the other (due the the sign
+difference in the second column). In Figure \ref{fig:toy-torus}, this
+can be seen most clearly in frame 5 where the two circles are rotated
+into a line from our view.
+
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+
+}
+
+\caption{Six frames selected from the torus view of 2D basis space.
+The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index.
+The grey points show the space of the 2D basis as a torus.
+The full video of the animation can be found at \url{https://vimeo.com/512882784}.}\label{fig:toy-torus}
+\end{figure}
+\end{Schunk}
 
@@
 the interpolation path and search points excluding the last termination
-iteration. \textcolor{red}{[A sentence for PD]}. Comparing the amount of
-random search done by SA and SAJO, it is surprising that SAJO does not
-carry as many samples as SA. Combining the insights from both the trace
-and PCA plot, one can learn the two different search strategies by SA
-and SAJO: SA frequently samples in the space and only make a move when
-an improvement is guaranteed to be made, while SAJO first broadly
-accepts bases in the space and then starts the extensive sampling in a
-narrowed subspace. The specific decision of which optimiser to use will
-depend on the index curve in the basis space but if the basis space is
-non-smooth, accepting inferior bases at first, as what SAJO has done,
-can lead to a more efficient search, in terms of the overall number of
-points evaluated.
+iteration. Pseudo Derivative (PD) quickly gets stuck near the starting
+position. Comparing the amount of random search done by SA and SAJO, it
+is surprising that SAJO does not carry as many samples as SA. Combining
+the insights from both the trace and PCA plot, one can learn the two
+different search strategies by SA and SAJO: SA frequently samples in the
+space and only make a move when an improvement is guaranteed to be made,
+while SAJO first broadly accepts bases in the space and then starts the
+extensive sampling in a narrowed subspace. The specific decision of
+which optimiser to use will depend on the index curve in the basis space
+but if the basis space is non-smooth, accepting inferior bases at first,
+as what SAJO has done, can lead to a more efficient search, in terms of
+the overall number of points evaluated.
 

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.Rmd
better examples of projection pursuit indices

@@
 The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers.
-In this paper, we implement a data collection object in the optimisation of projeciton pursuit guided tour and introduce visual diagnostics based on the data object collected.
-These diagnostics and workflows can be applied to a broad class of optimisers, to assess their performance.
+In this paper, we implement a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics based on the data object collected.
+These diagnostics and workflow can be applied to a broad class of optimisers, to assess their performance.
 An R package, \pkg{ferrn}, has been created to implement the diagnostics.
@@
 ```{r }
-# spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
+#spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
 ```
@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-In the context of projection pursuit, interestingness means departure from normality [@diaconis1984asymptotics; @huber1985projection].
-\textcolor{red}{only include a couple of indices, using e.g.}A number of index functions have been proposed in the literature to detect different data structures [@friedman1974projection; @hall1989polynomial; @cook1993projection; @posse1995projection] and to classify groups [@lee2005projection; @lee2010projection].
-
+In projection pursuit, higher index values are associated with interesting projections that depart from normality [@diaconis1984asymptotics; @huber1985projection], e.g.
+Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial] and natural Hermite index [@cook1993projection] for measuring the distance between the density of projected data and a standard normal distribution, LDA index [@lee2005projection] and PDA index [@lee2010projection] for finding clustering structure in the data, and skewness-based index [@Loperfido2018] and kurtosis-based index [@Loperfido2020] for outlier detection.
 As a general visualisation method, tour produces animations of high dimensional data via rotations of low-dimensional planes.

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.tex
better examples of projection pursuit indices

@@
 evaluating the performance of the optimisers. In this paper, we
-implement a data collection object in the optimisation of projeciton
+implement a data collection object in the optimisation of projection
 pursuit guided tour and introduce visual diagnostics based on the data
-object collected. These diagnostics and workflows can be applied to a
+object collected. These diagnostics and workflow can be applied to a
 broad class of optimisers, to assess their performance. An R package,
@@
 or just index function, and is used to measure the ``interestingness''
-of a given projection. In the context of projection pursuit,
-interestingness means departure from normality
-\citep{diaconis1984asymptotics, huber1985projection}.
-\textcolor{red}{only include a couple of indices, using e.g.}A number of
-index functions have been proposed in the literature to detect different
-data structures
-\citep{friedman1974projection, hall1989polynomial, cook1993projection, posse1995projection}
-and to classify groups \citep{lee2005projection, lee2010projection}.
-
-As a general visualisation method, tour produces animations of high
-dimensional data via rotations of low-dimensional planes. There are
-different tour implementations depending on how the high dimensional
-space is investigated: grand tour \citep{cook2008grand} selects the
-planes randomly to provide a general overview; manual tour
+of a given projection. In projection pursuit, higher index values are
+associated with interesting projections that depart from normality
+\citep{diaconis1984asymptotics, huber1985projection}, e.g.~Legendre
+index \citep{friedman1974projection}, Hermite index
+\citep{hall1989polynomial} and natural Hermite index
+\citep{cook1993projection} for measuring the distance between the
+density of projected data and a standard normal distribution, LDA index
+\citep{lee2005projection} and PDA index \citep{lee2010projection} for
+finding clustering structure in the data, and skewness-based index
+\citep{Loperfido2018} and kurtosis-based index \citep{Loperfido2020} for
+outlier detection. As a general visualisation method, tour produces
+animations of high dimensional data via rotations of low-dimensional
+planes. There are different tour implementations depending on how the
+high dimensional space is investigated: grand tour \citep{cook2008grand}
+selects the planes randomly to provide a general overview; manual tour
 \citep{cook1997manual} gradually phases in and out one variable to

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.Rmd
clean up captions

@@
 
-```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the steps in a tour path.
+```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation."}
+Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."}
 include_graphics(here::here("img/tour-path.png"))
@@
 
-```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6}, using the holes index.
+```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index.
 Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
@@
 
-```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
-The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
-The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
+```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right).
+The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
+Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
 ```
@@
 
-```{r toy-pca, fig.cap = "The PCA-projected basis space with two search paths optimised by SA and PD.
-The holes index is optimised for the 1D projection problem on the five-variable dataset, \\code{boa5}.
-All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
+```{r toy-pca, fig.cap = "Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
 ```
@@
 
-```{r toy-pca-animated, fig.cap = "Six frames selected from the animation of search paths in the PCA-projected basis space.
-With animation, the start and end position of each optimisation routine is clearly identified.
+```{r toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot.
+With animation, the progression of the search paths from start to finish are better identified.
+SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame.
 The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
@@
 
-```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
+```{r toy-tour, fig.cap="Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
+The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \\CRANpkg{geozoo}.
 The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
@@
 
-```{r toy-torus, fig.cap="Six frames selected from the torus view of 2D basis space.
-The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index.
-The grey points show the space of the 2D basis as a torus.
+```{r toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green).
+The projection problem is a 2D projection with three variables using holes index.
+The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
 The full video of the animation can be found at \\url{https://vimeo.com/512882784}."}
@@
 
-```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data with the holes index, optimised by SA.
+```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by SA.
 On the left panel, basis with higher index value is found during the interpolation but not used.
-The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
 ```
@@
 
-```{r polish, fig.height= 5, fig.cap = "The interpolation of a 2D projection problem on \\code{boa6} data with the holes index being first optimised by SA and then then polished.
-The projection by the final baiss of each algorithm is also presented.
-Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge.."}
+```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index.
+Optimiser SA is used before the polish.
+The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
 ```
@@
 
-```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \\code{boa5}.
-Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best."}
+```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
+Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best."}
 ```
@@
 
-```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
-The optimisation is done on the six-variable data, \\code{boa6} with 1D projecton index $I^{nk}(n)$.
-A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).
-The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing."}
+```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
 ```
@@
 
-```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis.
-Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463.
+```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
+Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463.
 The dots represent the target basis in each iteration and the path shows the interpolation.
-On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable.
-An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis.
-The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears."}
+On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
+The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears."}
 ```

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.tex
clean up captions

@@
 
-\caption[An illustration for demonstrating the steps in a tour path]{An illustration for demonstrating the steps in a tour path.
+\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
+Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
@@
 
-\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6}, using the holes index.
+\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index.
 Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
@@
 
-\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
-The lines indicate the interpolation and the dots indicate new targets generated by the optimisers.
-The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
+\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right).
+The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
+Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 
-\caption[The PCA-projected basis space with two search paths optimised by SA and PD]{The PCA-projected basis space with two search paths optimised by SA and PD.
-The holes index is optimised for the 1D projection problem on the five-variable dataset, \code{boa5}.
-All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
+\caption[Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
+The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
 \end{figure}
@@
 
-\caption{Six frames selected from the animation of search paths in the PCA-projected basis space.
-With animation, the start and end position of each optimisation routine is clearly identified.
+\caption{Six frames selected from the animated version of the previous plot.
+With animation, the progression of the search paths from start to finish are better identified.
+SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame.
 The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
@@
 
-\caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+\caption{Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \CRANpkg{geozoo}.
 The full video of the animation can be found at \url{https://vimeo.com/512885379}.}\label{fig:toy-tour}
@@
 
-\caption{Six frames selected from the torus view of 2D basis space.
-The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index.
-The grey points show the space of the 2D basis as a torus.
+\caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green).
+The projection problem is a 2D projection with three variables using holes index.
+The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
 The full video of the animation can be found at \url{https://vimeo.com/512882784}.}\label{fig:toy-torus}
@@
 
-\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA.
+\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA.
 On the left panel, basis with higher index value is found during the interpolation but not used.
-The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
 \end{figure}
@@
 
-\caption[The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished]{The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished.
-The projection by the final baiss of each algorithm is also presented.
-Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge..}\label{fig:polish}
+\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index.
+Optimiser SA is used before the polish.
+The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
 \end{figure}
@@
 
-\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}.
-Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
+\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}.
+Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 
-\caption[Comparing 20 search paths in the PCA-projected basis space with two optimisers]{Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
-The optimisation is done on the six-variable data, \code{boa6} with 1D projecton index $I^{nk}(n)$.
-A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).
-The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
+\caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
 \end{figure}
@@
 
-\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis.
-Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463.
+\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
+Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463.
 The dots represent the target basis in each iteration and the path shows the interpolation.
-On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable.
-An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis.
-The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
+On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
+The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
 \end{figure}

### 2021-03-01 paper/zhang-cook-laa-langrene-menendez.Rmd
SA -> CRS, SAJO -> SA

@@
 library(ferrn)
-read_chunk(here::here('scripts/toy.R'))
-read_chunk(here::here('scripts/diagnose.R'))
-read_chunk(here::here('scripts/kol_result.R'))
-read_chunk(here::here('scripts/implementation.R'))
+
+knitr::read_chunk(
+  here::here("scripts/toy.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/diagnose.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/implementation.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/kol_result.R")
+)
+
 ```
@@
 
-Three optimisers have been implemented in the \CRANpkg{tourr} [@tourr] package: simulated annealing (SA), simulated annealing with jump out (SAJO), and pseudo derivative (PD).
-Simulated annealing (SA) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
+Three optimisers have been implemented in the \CRANpkg{tourr} [@tourr] package: Creeping random search (CRS), simulated annealing (SA), and pseudo derivative (PD).
+Creeping random search (CRS) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
 $\mathbf{A}_{l}$ is then orthonormalised to fulfil the basis constraint.
@@
 }
-  \caption{Simulated annealing (SA)}
+  \caption{Creeping random search (CRS)}
   \label{random-search}
@@
 
-Simulated annealing with jump out (SAJO) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as SA but allows a probabilistic acceptance of a basis with lower index value than the current.
+Simulated annealing (SA) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current.
 Given an initial value of $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
@@
   }
-  \caption{Simulated annealing with jump out (SAJO)}
+  \caption{Simulated annealing (SA)}
   \label{simulated-annealing}
@@
 Three main pieces of information are recorded for the projection pursuit optimisers: 1) projection bases $\mathbf{A}$, 2) index values $I$, and 3) state $S$.
-For SA and SAJO, possible states include `random_search`,  `new_basis`, and `interpolation`.
+For CRS and SA, possible states include `random_search`,  `new_basis`, and `interpolation`.
 Pseudo derivative (PD) has a wider variety of states including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
@@
 
-Figure \ref{fig:toy-search} shows an example of the search plot for SA (left) and SAJO (right).
+Figure \ref{fig:toy-search} shows an example of the search plot for CRS (left) and SA (right).
 Both optimisers quickly find better bases in the first few iterations and then take longer to find one in the later iterations.
-The anchor bases, the ones found with the highest index value in each iteration, always have an increased index value in the optimiser SA while this is not the case for SAJO.
+The anchor bases, the ones found with the highest index value in each iteration, always have an increased index value in the optimiser CRS while this is not the case for SA.
 This feature gives SA an advantage in this simple example to quickly find the optimum.
 
-```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index.
-Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
+```{r toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index.
+Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis."}
 ```
@@
 
-```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right).
+```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
 The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
 ```
@@
 It may seem bizarre at the first sight that the interpolation in the left panel first passes some bases with higher index value before decreases to a lower target.
-This happens because on the one hand, the probabilistic acceptance in SAJO implies that some worse bases will be accepted by the optimiser.
+This happens because on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
 On the other hand, guided tour interpolate between the current and target basis to provide a smooth transition between projections.
-These two facts indicate that a non-monotonic interpolation can't be avoided for SAJO.
-Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for SA.
+These two facts indicate that a non-monotonic interpolation can't be avoided for SA.
+Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for CRS.
 
@@
 
-```{r toy-pca, fig.cap = "Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 
-Figure \ref{fig:toy-pca} shows the PCA plot of SA and PD for a 1D projection problem.
+Figure \ref{fig:toy-pca} shows the PCA plot of CRS and PD for a 1D projection problem.
 Both optimisers find the optimum but PD gets closer.
-With PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while SA evaluates points randomly in the search space to search for the next target.
-There are dashed lines annotated for SA and it describes the interruption of the interpolation, which will be discussed in details in Section \ref{monotonic}.
+With PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
+There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in details in Section \ref{monotonic}.
 
@@
 With animation, the progression of the search paths from start to finish are better identified.
-SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame.
+CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
 The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
@@
 Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
-An additional piece of information one can learn from this animation is that SA finds its end basis quicker than PD since SA finishes its search in the 5th frame while PD is still making more progress.
+An additional piece of information one can learn from this animation is that CRS finds its end basis quicker than PD since CRS finishes its search in the 5th frame while PD is still making more progress.
 
@@
 The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality of the columns in the bases.
-Two paths optimised by SA and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous few 1D space plots.
-The final basis found by PD and SA are shown in a larger shape  and printed below respectively:
+Two paths optimised by CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous few 1D space plots.
+The final basis found by PD and CRS are shown in a larger shape  and printed below respectively:
 
@@
 
-```{r toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green).
+```{r toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using holes index.
@@
 An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with higher index value than the target one.
-For SAJO, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for SA.
-The original trace plot for a 2D projection problem, optimised by SA, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+For SA, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for CRS.
+The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
 The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at Time 61 has a higher index value than the target in the third interpolation.
-Rather than starting the next iteration from the target basis on Time 65, SA starts the next iteration at Time 61 on the right panel and reaches a better final basis.
+Rather than starting the next iteration from the target basis on Time 65, CRS starts the next iteration at Time 61 on the right panel and reaches a better final basis.
 
-```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by SA.
+```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS.
 On the left panel, basis with higher index value is found during the interpolation but not used.
@@
 ```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index.
-Optimiser SA is used before the polish.
-The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
+Optimiser CRS is used before the polish.
+The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
 ```
@@
 A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local breakthrough.
-The polish algorithm is similar to the SA but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space,  percentage change in the index value, and neighbourhood size:
+The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space,  percentage change in the index value, and neighbourhood size:
 
@@
 
-Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by SA and followed by the polish.
+Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by CRS and followed by the polish.
 The projection by the final basis of each algorithm is also shown.
-The end basis found by SA reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
+The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
 Polish works with this end basis and further pushes the index value to produce much clear edges of the cluster, especially along the vertical axis.
@@
 ```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
-Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index.
-Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best."}
+Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best."}
 ```
@@
 There is barely any improvement made by PD, indicating its failure in optimising non-smooth index.
-While SA and SAJO have managed to get close to the index value of the theoretical best, trace plot shows that it takes SAJO much longer to interpolate towards the final basis.
-This long interpolation path is partially due to the fluctuation in the early iterations, where the SAJO tends to generously accept inferior bases before concentrating near the optimum.
+While CRS and SA have managed to get close to the index value of the theoretical best, trace plot shows that it takes SA much longer to interpolate towards the final basis.
+This long interpolation path is partially due to the fluctuation in the early iterations, where the SA tends to generously accept inferior bases before concentrating near the optimum.
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
 Pseudo Derivative (PD) quickly gets stuck near the starting position.
-Comparing the amount of random search done by SA and SAJO, it is surprising that SAJO does not carry as many samples as SA.
-Combining the insights from both the trace and PCA plot, one can learn the two different search strategies by SA and SAJO: SA frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SAJO first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
-The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as what SAJO has done, can lead to a more efficient search, in terms of the overall number of points evaluated.
+Comparing the amount of random search done by CRS and SA, it is surprising that SA does not carry as many samples as CRS.
+Combining the insights from both the trace and PCA plot, one can learn the two different search strategies by CRS and SA: CRS frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SA first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
+The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as what SA has done, can lead to a more efficient search, in terms of the overall number of points evaluated.
 
-```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
 The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
@@
 
-The next experiment compares the performance of SA and SAJO when a local maximum exists.
+The next experiment compares the performance of CRS and SA when a local maximum exists.
 Two search neighbourhood sizes: 0.5 and 0.7 are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
 Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by  optimiser and search size.
-With SA and a search size of 0.5, despite being the simplest and fastest, the optimiser fails in three instances where the final basis lands neither near the local nor the global optimum.
+With CRS and a search size of 0.5, despite being the simplest and fastest, the optimiser fails in three instances where the final basis lands neither near the local nor the global optimum.
 With a larger search size of 0.7, more seeds have found the global maximum.
-Comparing SA and SAJO for a search size of 0.5, SAJO does not seem to improve the final basis found, despite having longer interpolation paths.
-However, the denser paths near the local maximum is an indicator that SAJO is working hard to examine if there is any other optimum in the basis space but the relative small search size has diminished its ability to reach the global maximum.
+Comparing CRS and SA for a search size of 0.5, SA does not seem to improve the final basis found, despite having longer interpolation paths.
+However, the denser paths near the local maximum is an indicator that SA is working hard to examine if there is any other optimum in the basis space but the relative small search size has diminished its ability to reach the global maximum.
 With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
-This indicates that SAJO, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the global optimum.
+This indicates that SA, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the global optimum.
 
@@
 ```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
-Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463.
+Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
 The dots represent the target basis in each iteration and the path shows the interpolation.

### 2021-03-01 paper/zhang-cook-laa-langrene-menendez.tex
SA -> CRS, SAJO -> SA

@@
 Guided Tours}
-\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}
+\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, and Patricia Menéndez}
 
@@
 Three optimisers have been implemented in the \CRANpkg{tourr}
-\citep{tourr} package: simulated annealing (SA), simulated annealing
-with jump out (SAJO), and pseudo derivative (PD). Simulated annealing
-(SA) is a random search optimiser that samples a candidate basis
+\citep{tourr} package: Creeping random search (CRS), simulated annealing
+(SA), and pseudo derivative (PD). Creeping random search (CRS) is a
+random search optimiser that samples a candidate basis
 \(\mathbf{A}_{l}\) in the neighbourhood of the current basis
@@
 
-Simulated annealing with jump out (SAJO)
+Simulated annealing (SA)
 \citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
-same sampling process as SA but allows a probabilistic acceptance of a
+same sampling process as CRS but allows a probabilistic acceptance of a
 basis with lower index value than the current. Given an initial value of
@@
 optimisers: 1) projection bases \(\mathbf{A}\), 2) index values \(I\),
-and 3) state \(S\). For SA and SAJO, possible states include
+and 3) state \(S\). For CRS and SA, possible states include
 \texttt{random\_search}, \texttt{new\_basis}, and
@@
 
-Figure \ref{fig:toy-search} shows an example of the search plot for SA
-(left) and SAJO (right). Both optimisers quickly find better bases in
-the first few iterations and then take longer to find one in the later
+Figure \ref{fig:toy-search} shows an example of the search plot for CRS
+(left) and SA (right). Both optimisers quickly find better bases in the
+first few iterations and then take longer to find one in the later
 iterations. The anchor bases, the ones found with the highest index
 value in each iteration, always have an increased index value in the
-optimiser SA while this is not the case for SAJO. This feature gives SA
+optimiser CRS while this is not the case for SA. This feature gives SA
 an advantage in this simple example to quickly find the optimum.
@@
 
-\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index.
-Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
+\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index.
+Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.}\label{fig:toy-search}
 \end{figure}
@@
 
-\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right).
+\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
 The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 some bases with higher index value before decreases to a lower target.
-This happens because on the one hand, the probabilistic acceptance in
-SAJO implies that some worse bases will be accepted by the optimiser. On
-the other hand, guided tour interpolate between the current and target
-basis to provide a smooth transition between projections. These two
-facts indicate that a non-monotonic interpolation can't be avoided for
-SAJO. Later, in section \ref{monotonic}, there will be a discussion on
-improving the non-monotonic interpolation for SA.
+This happens because on the one hand, the probabilistic acceptance in SA
+implies that some worse bases will be accepted by the optimiser. On the
+other hand, guided tour interpolate between the current and target basis
+to provide a smooth transition between projections. These two facts
+indicate that a non-monotonic interpolation can't be avoided for SA.
+Later, in section \ref{monotonic}, there will be a discussion on
+improving the non-monotonic interpolation for CRS.
 
@@
 
-\caption[Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
+\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 
-Figure \ref{fig:toy-pca} shows the PCA plot of SA and PD for a 1D
+Figure \ref{fig:toy-pca} shows the PCA plot of CRS and PD for a 1D
 projection problem. Both optimisers find the optimum but PD gets closer.
@@
 optimisers: PD first evaluates points in a small neighbourhood for a
-promising direction, while SA evaluates points randomly in the search
+promising direction, while CRS evaluates points randomly in the search
 space to search for the next target. There are dashed lines annotated
-for SA and it describes the interruption of the interpolation, which
+for CRS and it describes the interruption of the interpolation, which
 will be discussed in details in Section \ref{monotonic}.
@@
 With animation, the progression of the search paths from start to finish are better identified.
-SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame.
+CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
 The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
@@
 PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
-one can learn from this animation is that SA finds its end basis quicker
-than PD since SA finishes its search in the 5th frame while PD is still
-making more progress.
+one can learn from this animation is that CRS finds its end basis
+quicker than PD since CRS finishes its search in the 5th frame while PD
+is still making more progress.
 
@@
 perpendicular to each other and this can be linked back to the
-orthogonality of the columns in the bases. Two paths optimised by SA and
-PD are plotted on top of the torus and coloured in green and brown
+orthogonality of the columns in the bases. Two paths optimised by CRS
+and PD are plotted on top of the torus and coloured in green and brown
 respectively to match the previous few 1D space plots. The final basis
-found by PD and SA are shown in a larger shape and printed below
+found by PD and CRS are shown in a larger shape and printed below
 respectively:
@@
 
-\caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green).
+\caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using holes index.
@@
 \ref{fig:toy-interp}: a path that passes bases with higher index value
-than the target one. For SAJO, a non-monotonic interpolation is
-justified since target bases do not necessarily have a higher index
-value than the current one, while this is not the case for SA. The
-original trace plot for a 2D projection problem, optimised by SA, is
-shown on the left panel of Figure \ref{fig:interruption} and one can
-observe clearly that the non-monotonic interpolation has undermined the
-optimiser to realise its full potential. Hence, an interruption is
-implemented to stop at the best basis found in the interpolation. The
-right panel of Figure \ref{fig:interruption} shows the trace plot after
-implementing the interruption and while the first two interpolations are
-identical, the basis at Time 61 has a higher index value than the target
-in the third interpolation. Rather than starting the next iteration from
-the target basis on Time 65, SA starts the next iteration at Time 61 on
-the right panel and reaches a better final basis.
+than the target one. For SA, a non-monotonic interpolation is justified
+since target bases do not necessarily have a higher index value than the
+current one, while this is not the case for CRS. The original trace plot
+for a 2D projection problem, optimised by CRS, is shown on the left
+panel of Figure \ref{fig:interruption} and one can observe clearly that
+the non-monotonic interpolation has undermined the optimiser to realise
+its full potential. Hence, an interruption is implemented to stop at the
+best basis found in the interpolation. The right panel of Figure
+\ref{fig:interruption} shows the trace plot after implementing the
+interruption and while the first two interpolations are identical, the
+basis at Time 61 has a higher index value than the target in the third
+interpolation. Rather than starting the next iteration from the target
+basis on Time 65, CRS starts the next iteration at Time 61 on the right
+panel and reaches a better final basis.
 
@@
 
-\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA.
+\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS.
 On the left panel, basis with higher index value is found during the interpolation but not used.
@@
 \caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index.
-Optimiser SA is used before the polish.
-The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
+Optimiser CRS is used before the polish.
+The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
 \end{figure}
@@
 the start of a new guided tour to search for local breakthrough. The
-polish algorithm is similar to the SA but with three distinctions: 1) a
+polish algorithm is similar to the CRS but with three distinctions: 1) a
 hundred rather than one candidate bases are generated each time in the
@@
 Figure \ref{fig:polish} presents the trace plot of a 2D projection,
-optimised by SA and followed by the polish. The projection by the final
-basis of each algorithm is also shown. The end basis found by SA reveals
-the four clusters in the data, but the edges of each cluster are not
-clean-cut. Polish works with this end basis and further pushes the index
-value to produce much clear edges of the cluster, especially along the
-vertical axis.
+optimised by CRS and followed by the polish. The projection by the final
+basis of each algorithm is also shown. The end basis found by CRS
+reveals the four clusters in the data, but the edges of each cluster are
+not clean-cut. Polish works with this end basis and further pushes the
+index value to produce much clear edges of the cluster, especially along
+the vertical axis.
 
@@
 \caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}.
-Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index.
-Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
+Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 smooth. There is barely any improvement made by PD, indicating its
-failure in optimising non-smooth index. While SA and SAJO have managed
-to get close to the index value of the theoretical best, trace plot
-shows that it takes SAJO much longer to interpolate towards the final
-basis. This long interpolation path is partially due to the fluctuation
-in the early iterations, where the SAJO tends to generously accept
-inferior bases before concentrating near the optimum. The PCA plot shows
-the interpolation path and search points excluding the last termination
+failure in optimising non-smooth index. While CRS and SA have managed to
+get close to the index value of the theoretical best, trace plot shows
+that it takes SA much longer to interpolate towards the final basis.
+This long interpolation path is partially due to the fluctuation in the
+early iterations, where the SA tends to generously accept inferior bases
+before concentrating near the optimum. The PCA plot shows the
+interpolation path and search points excluding the last termination
 iteration. Pseudo Derivative (PD) quickly gets stuck near the starting
-position. Comparing the amount of random search done by SA and SAJO, it
-is surprising that SAJO does not carry as many samples as SA. Combining
+position. Comparing the amount of random search done by CRS and SA, it
+is surprising that SA does not carry as many samples as CRS. Combining
 the insights from both the trace and PCA plot, one can learn the two
-different search strategies by SA and SAJO: SA frequently samples in the
+different search strategies by CRS and SA: CRS frequently samples in the
 space and only make a move when an improvement is guaranteed to be made,
-while SAJO first broadly accepts bases in the space and then starts the
+while SA first broadly accepts bases in the space and then starts the
 extensive sampling in a narrowed subspace. The specific decision of
@@
 but if the basis space is non-smooth, accepting inferior bases at first,
-as what SAJO has done, can lead to a more efficient search, in terms of
+as what SA has done, can lead to a more efficient search, in terms of
 the overall number of points evaluated.
@@
 
-\caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7.
+\caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
 The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
@@
 
-The next experiment compares the performance of SA and SAJO when a local
+The next experiment compares the performance of CRS and SA when a local
 maximum exists. Two search neighbourhood sizes: 0.5 and 0.7 are compared
@@
 paths simulated using 20 seeds in the PCA plot, faceted by optimiser and
-search size. With SA and a search size of 0.5, despite being the
+search size. With CRS and a search size of 0.5, despite being the
 simplest and fastest, the optimiser fails in three instances where the
@@
 larger search size of 0.7, more seeds have found the global maximum.
-Comparing SA and SAJO for a search size of 0.5, SAJO does not seem to
+Comparing CRS and SA for a search size of 0.5, SA does not seem to
 improve the final basis found, despite having longer interpolation
 paths. However, the denser paths near the local maximum is an indicator
-that SAJO is working hard to examine if there is any other optimum in
-the basis space but the relative small search size has diminished its
+that SA is working hard to examine if there is any other optimum in the
+basis space but the relative small search size has diminished its
 ability to reach the global maximum. With a larger search size, almost
@@
 final bases are much closer to the theoretical best, as compared to the
-three other cases. This indicates that SAJO, with a reasonable large
+three other cases. This indicates that SA, with a reasonable large
 search size, is able to overcome the local optimum and optimise close
@@
 \caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
-Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463.
+Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
 The dots represent the target basis in each iteration and the path shows the interpolation.
@@
 H.Sherry Zhang\\
-Monash University\\
+Monash University\\%
 Department of Econometrics and Business Statistics\\
-}
-\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
+%
+%
+%
+\\\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
+}
 
@@
 Dianne Cook\\
-Monash University\\
+Monash University\\%
 Department of Econometrics and Business Statistics\\
-}
-\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+%
+%
+%
+\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+}
 
@@
 Ursula Laa\\
-University of Natural Resources and Life Sciences\\
+University of Natural Resources and Life Sciences\\%
 Institute of Statistics\\
-}
-\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
+%
+%
+%
+\\\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
+}
 
@@
 Nicolas Langrené\\
-CSIRO Data61\\
+CSIRO Data61\\%
 34 Village Street, Docklands VIC 3008 Australia\\
-}
-\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
+%
+%
+%
+\\\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
+}
 
@@
 Patricia Menéndez\\
-Monash University\\
+Monash University\\%
 Department of Econometrics and Business Statistics\\
-}
-\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}
+%
+%
+%
+\\\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}
+}

### 2021-03-01 paper/zhang-cook-laa-langrene-menendez.R

@@ -1,0 +1,89 @@
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, 
+                      message = FALSE, 
+                      warning = FALSE,
+                      root.dir = here::here(), 
+                      fig.path = here::here("figs/"), 
+                      cache = TRUE,
+                      fig.align = "center",
+                      fig.height = 3,
+                      out.width = "100%")
+
+
+## ----external, include = FALSE, cache = FALSE---------------------------------
+library(knitr)
+library(tidyverse)
+library(ferrn)
+
+knitr::read_chunk(
+  here::here("scripts/toy.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/diagnose.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/implementation.R")
+)
+
+knitr::read_chunk(
+  here::here("scripts/kol_result.R")
+)
+
+
+
+
+
+
+## ----load-pkg-----------------------------------------------------------------
+
+
+## -----------------------------------------------------------------------------
+#spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
+
+
+## ----tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
+include_graphics(here::here("img/tour-path.png"))
+
+
+## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis."----
+
+
+## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."----
+
+
[…]

### 2021-03-01 scripts/diagnose.R

@@ -67,10 +67,10 @@ p2 <- after %>%
 
 p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
   clean_method() %>%
-  mutate(method = factor(method, levels = c("SA", "polish"))) %>%
+  mutate(method = factor(method, levels = c("CRS", "polish"))) %>%
   get_interp() %>%
   explore_trace_interp(color = method, cutoff = 100, target_size = 2, interp_size = 2) +
-  scale_color_discrete_botanical(breaks = c("SA", "polish"), label = c("SA", "polish")) +
+  scale_color_discrete_botanical(breaks = c("CFRS", "polish"), label = c("CRS", "polish")) +
   theme(legend.position = "bottom",
         plot.margin = margin(2.5, 2, 0, 0, "cm"))
 
@@ -163,7 +163,7 @@ p1 <- dt %>%
   explore_trace_interp(group = method, color = method, accuracy_y = 0.02) +
   geom_hline(yintercept = theo_best_index_val, alpha = 0.5, linetype = 2) +
   scale_color_discrete_botanical() +
-  facet_wrap(vars(method), scales = "free_x")
+  facet_wrap(vars(fct_relevel(method, c("PD", "CRS", "SA"))), scales = "free_x")
 
 pca <- dt %>%
   compute_pca(group = method) %>%
@@ -177,7 +177,7 @@ p2 <- pca %>%
                  dir_col = method, dir_alpha = 0.2) +
   scale_color_discrete_botanical() +
   theme(legend.position = "none") +
-  facet_wrap(vars(method))
+  facet_wrap(vars(fct_relevel(method, c("PD", "CRS", "SA"))))
 
 (p1 / p2)
 

### 2021-03-02 paper/zhang-cook-laa-langrene-menendez.Rmd
change color for polish example

@@
 The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indices.
-A 1D projection function, $I^{nk}(n)$, compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the Kolmogorov test.
+A 1D projection function, $I^{K}(n)$, compar‚es the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the Kolmogorov test.
 Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index, $I^{nk}(n)$, is defined as:
 
-$$I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]$$
+$$I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]$$
 With a non-smooth index function, two research questions are raised:

### 2021-03-02 paper/zhang-cook-laa-langrene-menendez.tex
change color for polish example

@@
 produces a smooth interpolation, while this is not the case for all the
-indices. A 1D projection function, \(I^{nk}(n)\), compares the projected
+indices. A 1D projection function, \(I^{K}(n)\), compar‚es the projected
 data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated normal
@@
 
-\[I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
+\[I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
 non-smooth index function, two research questions are raised:

### 2021-03-02 scripts/diagnose.R

@@ -45,12 +45,11 @@ p2 <- after %>%
 #   boa6,
 #   tour_path = guided_tour(holes(), d = 2, search_f = search_better, max.tries = 400),
 #   dev = "png",
-#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[6]]),
+#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[1]]),
 #   rescale = FALSE,
-#   frames = 500,
+#   frames = 00,
 #   file = here::here("anim","polish", "before%03d.png")
 # )
-#
 # last_basis <- get_best(holes_2d_better)$basis %>% .[[1]]
 #
 # set.seed(123456)
@@ -58,7 +57,7 @@ p2 <- after %>%
 #   boa6,
 #   tour_path =  guided_tour(holes(), d = 2, search_f = search_polish),
 #   dev = "png",
-#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[1]]),
+#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[6]]),
 #   start = last_basis,
 #   rescale = FALSE,
 #   frames = 100,
@@ -70,16 +69,19 @@ p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
   mutate(method = factor(method, levels = c("CRS", "polish"))) %>%
   get_interp() %>%
   explore_trace_interp(color = method, cutoff = 100, target_size = 2, interp_size = 2) +
-  scale_color_discrete_botanical(breaks = c("CFRS", "polish"), label = c("CRS", "polish")) +
-  theme(legend.position = "bottom",
-        plot.margin = margin(2.5, 2, 0, 0, "cm"))
+  scale_color_discrete_botanical(breaks = c("CRS", "polish"), label = c("CRS", "polish")) +
+  theme(legend.position = "bottom")
+
+wrap <- function(path) png::readPNG(path) %>% grid::rasterGrob() %>% wrap_plots()
 
 before <- png::readPNG(here::here("anim","polish", "before073.png"))
 after <- png::readPNG(here::here("anim","polish", "after006.png"))
 
-p1 +
-  inset_element(grid::rasterGrob(before), left = 0.77, bottom = 0.45, right = 0.89, top = 0.67, align_to = "full") +
[…]

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.Rmd
small changes flagged by Gael in the confirmation

@@
 In this paper, we implement a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics based on the data object collected.
-These diagnostics and workflow can be applied to a broad class of optimisers, to assess their performance.
+These diagnostics and this workflow can be applied to a broad class of optimisers, to assess their performance.
 An R package, \pkg{ferrn}, has been created to implement the diagnostics.
@@
 
-Optimization problems are ubiquitous in many areas of study.
+Optimisation problems are ubiquitous in many areas of study.
 While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130."}
 ```
@@
 Optimiser CRS is used before the polish.
-The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
+The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis."}
 ```

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.tex
small changes flagged by Gael in the confirmation

@@
 pursuit guided tour and introduce visual diagnostics based on the data
-object collected. These diagnostics and workflow can be applied to a
-broad class of optimisers, to assess their performance. An R package,
+object collected. These diagnostics and this workflow can be applied to
+a broad class of optimisers, to assess their performance. An R package,
 \pkg{ferrn}, has been created to implement the diagnostics.
@@
 
-Optimization problems are ubiquitous in many areas of study. While in
+Optimisation problems are ubiquitous in many areas of study. While in
 some cases analytical solutions can be found, the majority of problems
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 Optimiser CRS is used before the polish.
-The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
+The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis.}\label{fig:polish}
 \end{figure}

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.Rmd
more changes

@@
 Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial] and natural Hermite index [@cook1993projection] for measuring the distance between the density of projected data and a standard normal distribution, LDA index [@lee2005projection] and PDA index [@lee2010projection] for finding clustering structure in the data, and skewness-based index [@Loperfido2018] and kurtosis-based index [@Loperfido2020] for outlier detection.
+
 As a general visualisation method, tour produces animations of high dimensional data via rotations of low-dimensional planes.
@@
 
-\noindent where $f$, $\mathcal{A}$, and $\mathbf{X}$ are defined as in the previous section,  $I_d$ is the $d$-dimensional identity matrix and the constraint ensures the projection bases $\mathbf{A}$ to be orthonormal.
-It is worth noticing the following: 1) The optimisation is constrained and the basis space is a $d$-dimensional sphere.
+\noindent where $f$ and $\mathbf{X}$ are defined as in the previous section, $\mathcal{A}$ is the set of all $p$-dimensional projection bases, $I_d$ is the $d$-dimensional identity matrix, and the constraint ensures the projection bases, $\mathbf{A}$, to be orthonormal.
+It is worth noticing the following: 1) The optimisation is constrained and the orthonormality constraint imposes a geometrical structure on the bases space: it forms a Stiefel manifold.
 2) There may be index functions for which the objective function might not be differentiable.

### 2021-03-08 paper/zhang-cook-laa-langrene-menendez.tex
more changes

@@
 
-\noindent where \(f\), \(\mathcal{A}\), and \(\mathbf{X}\) are defined
-as in the previous section, \(I_d\) is the \(d\)-dimensional identity
-matrix and the constraint ensures the projection bases \(\mathbf{A}\) to
-be orthonormal. It is worth noticing the following: 1) The optimisation
-is constrained and the basis space is a \(d\)-dimensional sphere. 2)
-There may be index functions for which the objective function might not
-be differentiable. 3) While finding the global optimum is the goal of
-the optimisation problem, interesting projections may also appear in the
+\noindent where \(f\) and \(\mathbf{X}\) are defined as in the previous
+section, \(\mathcal{A}\) is the set of all \(p\)-dimensional projection
+bases, \(I_d\) is the \(d\)-dimensional identity matrix, and the
+constraint ensures the projection bases, \(\mathbf{A}\), to be
+orthonormal. It is worth noticing the following: 1) The optimisation is
+constrained and the orthonormality constraint imposes a geometrical
+structure on the bases space: it forms a Stiefel manifold. 2) There may
+be index functions for which the objective function might not be
+differentiable. 3) While finding the global optimum is the goal of the
+optimisation problem, interesting projections may also appear in the
 local optimum. 4) The optimisation should be computationally agile since

### 2021-03-10 paper/zhang-cook-laa-langrene-menendez.Rmd
render graphs

@@
 
-```{r toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, fig.asp=1, fig.height=2.5, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.

### 2021-03-10 paper/zhang-cook-laa-langrene-menendez.R

@@ -6,7 +6,7 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.path = here::here("figs/"), 
                       cache = TRUE,
                       fig.align = "center",
-                      fig.height = 3,
+                      fig.height = 2.5,
                       out.width = "100%")
 
 
@@ -50,10 +50,10 @@ include_graphics(here::here("img/tour-path.png"))
 ## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis."----
 
 
-## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."----
+## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130."----
 
 
-## ----toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
+## ----toy-pca, fig.height = 3, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
 
 
[…]

### 2021-03-10 scripts/toy.R

@@ -32,12 +32,12 @@ p2 <- holes_2d_better_random %>%
 
 p1 | p2
 
-## ---- toy-pca
+## ----toy-pca
 bind_rows(holes_1d_geo, holes_1d_better) %>%
   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
                    index = tourr::holes(), raw_data = boa5)  %>%
   explore_space_pca(group = method, details = TRUE,
-                    interp_size = 1, ratio = 9) +
+                    interp_size = 1) +
   scale_color_discrete_botanical()
 
 ## ----toy-pca-animated

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.Rmd
writing in optim section, algo to appendix, small tweaks in vis diag section, a tweak of the fig 9 (WIP)

@@
 The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view.
-The interestingness of the structure is defined by the index function, a function of the projection basis.
+A view is interesting if it can show some unexpected structure of the data, e.g., bi-modality, clustering and outliers.
 
@@
 
-Optimization methods can be divided into various classes, such as global optimisation [@kelley1999iterative; @fletcher2013practical], convex optimisation [@boyd2004convex] or stochastic optimisation [@nocedal2006numerical].
-Our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathbb{A}$ space.
+The optimisation problem of our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
+That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathcal{A}$ space.
 
-A large class of methods utilises the gradient information of the objective function to perform the optimisation iterations, with the most notable one being the gradient ascent (descent) method.
+A large class of methods utilises the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method.
 Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
+As will be shown in the next section, the independent variable, in our optimisation problem, is a matrix of projection basis and the computational time required to perform differentation on a matrix could impede the render of tour animation.
+Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods.
 Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable.
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
-Examples of those include response surface methodology [@box1951experimental], stochastic approximation [@robbins1951stochastic], random search [@fu2015handbook] and heuristic methods [@sorensen2013metaheuristics].
-Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour.
+Below will introduce two random search methods: creeping random search and simulated annealing, and a pseudo derivative search.
+
+<!-- Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour. -->
 
@@
 The optimiser terminates when the maximum number of iteration $l_{\max}$ is reached before a better basis can be found.
-The algorithm of SA is summarised in Algorithm \ref{random-search}.
+The algorithm of SA is summarised in Algorithm \ref{random-search} in the appendix.
 @posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
 In his proposal, the $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
-
-\begin{algorithm}
-\SetAlgoLined
-  \SetKwInOut{input}{input}
-  \SetKwInOut{output}{output}
-    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
-    \output{$\mathbf{A}_{l}$}
-    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
-  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   Set $l = 1$\;
-  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    Update $l = l + 1$\;
-  }
-  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
-}
-  \caption{Creeping random search (CRS)}
-  \label{random-search}
-\end{algorithm}
 
@@
 This implementation allows the optimiser to make a move and explore the basis space even the landing basis does have a higher index value and hence enable the optimiser to jump out of a local optimum.
-The algorithm \ref{simulated-annealing} highlights how SAJO differs from SA in the inner loop.
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-    Update $l = l + 1$\;
-  }
-  \caption{Simulated annealing (SA)}
-  \label{simulated-annealing}
-\end{algorithm}
+The algorithm \ref{simulated-annealing} in the appendix highlights how SAJO differs from SA in the inner loop.
 
@@
 The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.
-Algorithm \ref{search-geodesic} summarises the inner loop of the PD.
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
-  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
-  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
-  Compute the corresponding index value for each candidate bases\;
-  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
-  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-  Update $l = l + 1$\;
-}
-\caption{Pseudo derivative (PD)}
-\label{search-geodesic}
-\end{algorithm}
+Algorithm \ref{search-geodesic} in the appendix summarises the inner loop of the PD.
 
@@
 A data structure for diagnosing optimisers in projection pursuit guided tour is first defined.
-With a fixed data structure, the diagnostic plots presented later in this paper can be easily applied to diagnose other optimisation procedures.
+With this data structure, four types of diagnostic plots are presented.
+In the section, examples will be given to show how these plot can be used to reveal issues in the projection pursuit optimisation.
 
@@
 2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
-3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{31}  & V_{32}\\
 \vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
@@
 \vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
 \vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
@@
 \text{search} \\
-\text{search} \\
 \vdots \\
@@
 \hline
-\text{interpolate} \\
 \text{interpolate} \\
@@
 In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers.
-Seeing plots as stacked layers gives analysts the freedom to composite plots with multiple elements on hands.
+Seeing plots as stacked layers gives analysts the freedom to compose plots with multiple elements on hand.
 
-## Checking how hard the optimiser is working
+## Diagnostic 1: Checking how hard the optimiser is working
 
@@
 
-## Examining the optimisation progress {#toy-interp}
+## Diagnostic 2: Examining the optimisation progress {#toy-interp}
 
@@
 
-## Understanding the optimiser's coverage of the search space {#toy-pca}
+## Diagnostic 3a: Understanding the optimiser's coverage of the search space {#toy-pca}
 
@@
 
-## Animating the diagnostic plots
+## Diagnostic 3b: Animating the diagnostic plots
 
@@
 
-## The tour looking at itself
+## Diagnostic 4a: The tour looking at itself
 
@@
 
-## Forming a torus
+## Diagnostic 4b: Forming a torus
 
@@
 \bibliography{zhang-cook-laa-langrene-menendez}
+
+\newpage
+
+# Appendix
+
+Three algorithms (Creeping random search, simulated annealing, and pseudo derivative) used in projection pursuit guided tour optimisation.
+
+\begin{algorithm}
+\SetAlgoLined
+  \SetKwInOut{input}{input}
+  \SetKwInOut{output}{output}
+    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
+    \output{$\mathbf{A}_{l}$}
+    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
+   Set $l = 1$\;
+  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    Update $l = l + 1$\;
+  }
+  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
+}
+  \caption{Creeping random search (CRS)}
+  \label{random-search}
+\end{algorithm}
+
+\begin{algorithm}
+\SetAlgoLined
+\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+    Update $l = l + 1$\;
+  }
+  \caption{Simulated annealing (SA)}
+  \label{simulated-annealing}
+\end{algorithm}
+
+\begin{algorithm}
+\SetAlgoLined
+\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
+  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
+  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
+  Compute the corresponding index value for each candidate bases\;
+  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
+  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
+  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
+  Update $l = l + 1$\;
+}
+\caption{Pseudo derivative (PD)}
+\label{search-geodesic}
+\end{algorithm}

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.tex
render graphs
writing in optim section, algo to appendix, small tweaks in vis diag section, a tweak of the fig 9 (WIP)

@@
 projection, characterised by the projection basis, that gives the most
-interesting low-dimensional view. The interestingness of the structure
-is defined by the index function, a function of the projection basis.
+interesting low-dimensional view. A view is interesting if it can show
+some unexpected structure of the data, e.g., bi-modality, clustering and
+outliers.
 
@@
 
-Optimization methods can be divided into various classes, such as global
-optimisation \citep{kelley1999iterative, fletcher2013practical}, convex
-optimisation \citep{boyd2004convex} or stochastic optimisation
-\citep{nocedal2006numerical}. Our interest is on constrained
-optimization \citep{bertsekas2014constrained} as defined in the
-introduction section, and assuming it is not possible to find a solution
-to the problem in the way of a closed-form. That is, the problem
-consists of finding the minimum or maximum of a function \(f \in L^p\)
-in the constrained \(\mathbb{A}\) space.
+The optimisation problem of our interest is on constrained optimization
+\citep{bertsekas2014constrained} as defined in the introduction section,
+and assuming it is not possible to find a solution to the problem in the
+way of a closed-form. That is, the problem consists of finding the
+minimum or maximum of a function \(f \in L^p\) in the constrained
+\(\mathcal{A}\) space.
 
 A large class of methods utilises the gradient information of the
-objective function to perform the optimisation iterations, with the most
-notable one being the gradient ascent (descent) method. Although
-gradient optimization methods are popular, they rely on the availability
-of the objective function derivatives and on the complexity of the
-constraints. Derivative-free methods, which do not rely on the knowledge
-of the gradient, are more generally applicable. Derivative-free methods
-have been developed over the years, where the emphasis is on finding, in
-most cases, a near optimal solution. Examples of those include response
-surface methodology \citep{box1951experimental}, stochastic
-approximation \citep{robbins1951stochastic}, random search
-\citep{fu2015handbook} and heuristic methods
-\citep{sorensen2013metaheuristics}. Later, we will present a simulated
-annealing optimisation algorithm, which belongs to the class of random
-search methods, for optimisation with the guided tour.
+objective function, with the most notable one being the gradient ascent
+(descent) method. Although gradient optimization methods are popular,
+they rely on the availability of the objective function derivatives and
+on the complexity of the constraints. As will be shown in the next
+section, the independent variable, in our optimisation problem, is a
+matrix of projection basis and the computational time required to
+perform differentation on a matrix could impede the render of tour
+animation. Hence, gradient-based methods are not the focus of this paper
+and considerations will be given to derivative-free methods.
+Derivative-free methods, which do not rely on the knowledge of the
+gradient, are more generally applicable. Derivative-free methods have
+been developed over the years, where the emphasis is on finding, in most
+cases, a near optimal solution. Below will introduce two random search
+methods: creeping random search and simulated annealing, and a pseudo
+derivative search.
 
@@
 \citep{Loperfido2018} and kurtosis-based index \citep{Loperfido2020} for
-outlier detection. As a general visualisation method, tour produces
-animations of high dimensional data via rotations of low-dimensional
-planes. There are different tour implementations depending on how the
-high dimensional space is investigated: grand tour \citep{cook2008grand}
-selects the planes randomly to provide a general overview; manual tour
+outlier detection.
+
+As a general visualisation method, tour produces animations of high
+dimensional data via rotations of low-dimensional planes. There are
+different tour implementations depending on how the high dimensional
+space is investigated: grand tour \citep{cook2008grand} selects the
+planes randomly to provide a general overview; manual tour
 \citep{cook1997manual} gradually phases in and out one variable to
@@
 \(l_{\max}\) is reached before a better basis can be found. The
-algorithm of SA is summarised in Algorithm \ref{random-search}.
-\citet{posse1995projection} has proposed a slightly different cooling
-scheme by introducing a halving parameter \(c\). In his proposal, the
-\(\alpha\) is only adjusted if the last iteration takes more than \(c\)
-times to find a better basis.
-
-\begin{algorithm}
-\SetAlgoLined
-  \SetKwInOut{input}{input}
-  \SetKwInOut{output}{output}
-    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
-    \output{$\mathbf{A}_{l}$}
-    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
-  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   Set $l = 1$\;
-  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    Update $l = l + 1$\;
-  }
-  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
-}
-  \caption{Creeping random search (CRS)}
-  \label{random-search}
-\end{algorithm}
+algorithm of SA is summarised in Algorithm \ref{random-search} in the
+appendix. \citet{posse1995projection} has proposed a slightly different
+cooling scheme by introducing a halving parameter \(c\). In his
+proposal, the \(\alpha\) is only adjusted if the last iteration takes
+more than \(c\) times to find a better basis.
 
@@
 enable the optimiser to jump out of a local optimum. The algorithm
-\ref{simulated-annealing} highlights how SAJO differs from SA in the
-inner loop.
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-    Update $l = l + 1$\;
-  }
-  \caption{Simulated annealing (SA)}
-  \label{simulated-annealing}
-\end{algorithm}
+\ref{simulated-annealing} in the appendix highlights how SAJO differs
+from SA in the inner loop.
 
@@
 for the current iteration if it has a higher index value than the
-current one. Algorithm \ref{search-geodesic} summarises the inner loop
-of the PD.
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
-  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
-  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
-  Compute the corresponding index value for each candidate bases\;
-  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
-  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-  Update $l = l + 1$\;
-}
-\caption{Pseudo derivative (PD)}
-\label{search-geodesic}
-\end{algorithm}
+current one. Algorithm \ref{search-geodesic} in the appendix summarises
+the inner loop of the PD.
 
@@
 A data structure for diagnosing optimisers in projection pursuit guided
-tour is first defined. With a fixed data structure, the diagnostic plots
-presented later in this paper can be easily applied to diagnose other
-optimisation procedures.
+tour is first defined. With this data structure, four types of
+diagnostic plots are presented. In the section, examples will be given
+to show how these plot can be used to reveal issues in the projection
+pursuit optimisation.
 
@@
 of stacked layers. Seeing plots as stacked layers gives analysts the
-freedom to composite plots with multiple elements on hands.
-
-\hypertarget{checking-how-hard-the-optimiser-is-working}{%
-\subsection{Checking how hard the optimiser is
-working}\label{checking-how-hard-the-optimiser-is-working}}
+freedom to compose plots with multiple elements on hand.
+
+\hypertarget{diagnostic-1-checking-how-hard-the-optimiser-is-working}{%
+\subsection{Diagnostic 1: Checking how hard the optimiser is
+working}\label{diagnostic-1-checking-how-hard-the-optimiser-is-working}}
 
@@
 \hypertarget{toy-interp}{%
-\subsection{Examining the optimisation progress}\label{toy-interp}}
+\subsection{Diagnostic 2: Examining the optimisation
+progress}\label{toy-interp}}
 
@@
 \hypertarget{toy-pca}{%
-\subsection{Understanding the optimiser's coverage of the search
-space}\label{toy-pca}}
+\subsection{Diagnostic 3a: Understanding the optimiser's coverage of the
+search space}\label{toy-pca}}
 
@@
 
-\hypertarget{animating-the-diagnostic-plots}{%
-\subsection{Animating the diagnostic
-plots}\label{animating-the-diagnostic-plots}}
+\hypertarget{diagnostic-3b-animating-the-diagnostic-plots}{%
+\subsection{Diagnostic 3b: Animating the diagnostic
+plots}\label{diagnostic-3b-animating-the-diagnostic-plots}}
 
@@
 
-\hypertarget{the-tour-looking-at-itself}{%
-\subsection{The tour looking at
-itself}\label{the-tour-looking-at-itself}}
+\hypertarget{diagnostic-4a-the-tour-looking-at-itself}{%
+\subsection{Diagnostic 4a: The tour looking at
+itself}\label{diagnostic-4a-the-tour-looking-at-itself}}
 
@@
 
-\hypertarget{forming-a-torus}{%
-\subsection{Forming a torus}\label{forming-a-torus}}
+\hypertarget{diagnostic-4b-forming-a-torus}{%
+\subsection{Diagnostic 4b: Forming a
+torus}\label{diagnostic-4b-forming-a-torus}}
 
@@
 
+\newpage
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+Three algorithms (Creeping random search, simulated annealing, and
+pseudo derivative) used in projection pursuit guided tour optimisation.
+
+\begin{algorithm}
+\SetAlgoLined
+  \SetKwInOut{input}{input}
+  \SetKwInOut{output}{output}
+    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
+    \output{$\mathbf{A}_{l}$}
+    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
+  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
+   Set $l = 1$\;
+  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
+    Update $l = l + 1$\;
+  }
+  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
+  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
+  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
+}
+  \caption{Creeping random search (CRS)}
+  \label{random-search}
+\end{algorithm}
+
+\begin{algorithm}
+\SetAlgoLined
+\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
+    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
+    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
+    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
+    Update $l = l + 1$\;
+  }
+  \caption{Simulated annealing (SA)}
+  \label{simulated-annealing}
+\end{algorithm}
+
+\begin{algorithm}
+\SetAlgoLined
+\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
+  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
+  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
+  Compute the corresponding index value for each candidate bases\;
+  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
+  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
+  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
+  Update $l = l + 1$\;
+}
+\caption{Pseudo derivative (PD)}
+\label{search-geodesic}
+\end{algorithm}
+
 \address{%

### 2021-03-16 scripts/diagnose.R

@@ -74,10 +74,13 @@ p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
 
 wrap <- function(path) png::readPNG(path) %>% grid::rasterGrob() %>% wrap_plots()
 
-before <- png::readPNG(here::here("anim","polish", "before073.png"))
-after <- png::readPNG(here::here("anim","polish", "after006.png"))
-
-(wrap_plots(grid::rasterGrob(before)) |wrap_plots(grid::rasterGrob(after))) / p1
+first <- here::here("anim","polish", "before001.png")
+before <- here::here("anim","polish", "before073.png")
+after <- here::here("anim","polish", "after006.png")
+file <- c(first, before, after)
+rl <- lapply(file, png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+wrap_plots(gl) / p1 + plot_annotation(tag_levels = "A")
 
 # p1 +
 #   inset_element(grid::rasterGrob(before), left = 0.77, bottom = 0.45, right = 0.89, top = 0.67, align_to = "full") +

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.Rmd
fix various plots and writing on interestingness of the index

@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-In projection pursuit, higher index values are associated with interesting projections that depart from normality [@diaconis1984asymptotics; @huber1985projection], e.g.
-Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial] and natural Hermite index [@cook1993projection] for measuring the distance between the density of projected data and a standard normal distribution, LDA index [@lee2005projection] and PDA index [@lee2010projection] for finding clustering structure in the data, and skewness-based index [@Loperfido2018] and kurtosis-based index [@Loperfido2020] for outlier detection.
+An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal as $n$ and $p$ approaching infinity under certain conditions.
+@huber1985projection illustrates that, for indices based on entropy measures, [...].
+Index functions  proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution.(Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]); chi-square index [@posse1995projection] for detecting spiral structure; kurtosis [@Loperfido2020] and skewness [@Loperfido2018] indices for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 
@@
 
-```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index.
-Optimiser CRS is used before the polish.
-The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis."}
+```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index.
+Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis."}
 ```
@@
 
-Figure \ref{fig:polish} presents the trace plot of a 2D projection, optimised by CRS and followed by the polish.
-The projection by the final basis of each algorithm is also shown.
+Figure \ref{fig:polish} presents the projectd data and trace plot of a 2D projection, optimised by CRS and followed by the polish.
+In the top row shows the intial projection, the final projection after CRS, and the final projection after polish, repsectively.
 The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
@@
 
-```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
+```{r noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
 Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index.

### 2021-03-16 paper/zhang-cook-laa-langrene-menendez.tex
fix various plots and writing on interestingness of the index

@@
 or just index function, and is used to measure the ``interestingness''
-of a given projection. In projection pursuit, higher index values are
-associated with interesting projections that depart from normality
-\citep{diaconis1984asymptotics, huber1985projection}, e.g.~Legendre
-index \citep{friedman1974projection}, Hermite index
-\citep{hall1989polynomial} and natural Hermite index
-\citep{cook1993projection} for measuring the distance between the
-density of projected data and a standard normal distribution, LDA index
-\citep{lee2005projection} and PDA index \citep{lee2010projection} for
-finding clustering structure in the data, and skewness-based index
-\citep{Loperfido2018} and kurtosis-based index \citep{Loperfido2020} for
-outlier detection.
+of a given projection. An interesting projection shows structures that
+are non-normal and this has been argued for two reasons: Theoretical
+proofs from \citet{diaconis1984asymptotics} show that projections tend
+to be normal as \(n\) and \(p\) approaching infinity under certain
+conditions. \citet{huber1985projection} illustrates that, for indices
+based on entropy measures, {[}\ldots{}{]}. Index functions proposed in
+the litearture include early indices that can be categorised by
+measuring the \(L^2\) distance between the projection and a normal
+distribution.(Legendre index \citep{friedman1974projection}, Hermite
+index \citep{hall1989polynomial}, and natural Hermite index
+\citep{cook1993projection}); chi-square index
+\citep{posse1995projection} for detecting spiral structure; kurtosis
+\citep{Loperfido2020} and skewness \citep{Loperfido2018} indices for
+detecting outliers in financial time series; and most recently,
+scagnostic indices \citep{laa2020using} for summarising structures in
+scatterplot matrices based on eight scagnostic measures.
 
@@
 
-\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index.
-Optimiser CRS is used before the polish.
-The clustering structure in the data is detected by CRS but the polish step improves the index value and produces clearer boundaries of the clusters, especially along the vertial axis.}\label{fig:polish}
+\caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
+Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis.}\label{fig:polish}
 \end{figure}
@@
 
-Figure \ref{fig:polish} presents the trace plot of a 2D projection,
-optimised by CRS and followed by the polish. The projection by the final
-basis of each algorithm is also shown. The end basis found by CRS
+Figure \ref{fig:polish} presents the projectd data and trace plot of a
+2D projection, optimised by CRS and followed by the polish. In the top
+row shows the intial projection, the final projection after CRS, and the
+final projection after polish, repsectively. The end basis found by CRS
 reveals the four clusters in the data, but the edges of each cluster are

### 2021-03-16 scripts/diagnose.R

@@ -71,7 +71,6 @@ p1 <- bind_rows(holes_2d_better, holes_2d_better_polish) %>%
   explore_trace_interp(color = method, cutoff = 100, target_size = 2, interp_size = 2) +
   scale_color_discrete_botanical(breaks = c("CRS", "polish"), label = c("CRS", "polish")) +
   theme(legend.position = "bottom")
-
 wrap <- function(path) png::readPNG(path) %>% grid::rasterGrob() %>% wrap_plots()
 
 first <- here::here("anim","polish", "before001.png")
@@ -80,12 +79,11 @@ after <- here::here("anim","polish", "after006.png")
 file <- c(first, before, after)
 rl <- lapply(file, png::readPNG)
 gl <-  lapply(rl, grid::rasterGrob)
-wrap_plots(gl) / p1 + plot_annotation(tag_levels = "A")
 
-# p1 +
-#   inset_element(grid::rasterGrob(before), left = 0.77, bottom = 0.45, right = 0.89, top = 0.67, align_to = "full") +
-#   inset_element(grid::rasterGrob(after), left = 0.87, bottom = 0.7, right = 1, top = 1, align_to = "full")
+lay <- rbind(c(1,2,3),
+             c(4,4,4))
 
+gridExtra::grid.arrange(gl[[1]], gl[[2]], gl[[3]], p1, layout_matrix = lay)
 ## ---- flip-sign
 # set.seed(2463)
 # orientation_corrected <-
@@ -108,14 +106,16 @@ wrap_plots(gl) / p1 + plot_annotation(tag_levels = "A")
 dt <- bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
           orientation_different %>% mutate(sign = "original")) %>%
   compute_pca() %>%
-  pluck("aug")
+  pluck("aug") %>%
+  mutate(sign = fct_relevel(as.factor(sign), c("original", "flipped")))
 
 dt %>%
   explore_space_pca(group = sign, pca = FALSE ,flip = FALSE, start_size = 3) +
   add_anchor(dt = get_interp_last(dt), anchor_color = sign) +
   scale_color_discrete_botanical() +
-  facet_wrap(vars(fct_relevel(sign, c("original", "flipped")))) +
-  theme(legend.position = "none")
+  facet_wrap(vars(sign)) +
[…]

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.Rmd
some changes on optim method section

@@
 
-Optimisation problems are ubiquitous in many areas of study.
+<!-- Optimisation problems are ubiquitous in many areas of study.
 While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
@@
 In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.
+-->
 
@@
 Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods.
+
 Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable.
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
-Below will introduce two random search methods: creeping random search and simulated annealing, and a pseudo derivative search.
+Below will introduce three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo derivative search.
 
-<!-- Later, we will present a simulated annealing optimisation algorithm, which belongs to the class of random search methods, for optimisation with the guided tour. -->
+Random search [in practice, non-convex, non-smooth].
+The original (pure) random search, Brooks, randomly evaluate points in the space regardless of the current position.
+Although gaurantee global convergence, as the dimensional increases, blindly sample points in the entire space would cost a long time for convergence to occur [spall, 6].
+Various algorithms improve on the pure random search by either concentrating on a narrower sampling space or a different updating machanism.
+Creeping random search, [R.C White], is such a variation where a candidate point is generated within a neighbourhood of the current point and the current point is only updated if the candidate point makes an improvment on the function value.
+[no convergence guarantee].
+A different random search algorithm, simulated annealing, takes a different updating procedure than creeping random search.
+Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where nonimproving candidates can still have a chance to be accepted.
+The convergence of simulated annealing algorithms has been widely researched [reference, reference, reference] and a global convergence is guaranteed [under what condition].
 
-A common search scheme utilised by both derivative-free methods and gradient methods is line search.
+Pseudo derivative search uses a common search scheme in optimistaion: line search.
 In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
 Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached.
-While gradient-based methods choose the search direction by the gradient, derivative-free methods uses local information of the objective function to determine the search direction.
+In derivative-free methods, local information of the objective function are used to determine the search direction.
 The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an optimum.
@@
 
-Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
+<!-- Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
 The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions.
@@
 For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].
+-->
 

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.tex
some changes on optim method section

@@
 
-Optimisation problems are ubiquitous in many areas of study. While in
-some cases analytical solutions can be found, the majority of problems
-rely on numerical methods to find the optimal solution. These numerical
-methods follow iterative approaches that aim at finding the optimum by
-progressively improving the current solution until a desirable accuracy
-is achieved. Although this principle seems uncomplicated, a number of
-challenges arise such as the possible existence of multiple maxima
-(local and global), constraints and noisy objective function, and the
-trade-off between desirable accuracy and computational burden. In
-addition, the optimization results might depend on the algorithm
-starting values, affecting the consistency of results.
-
 The optimisation problem of our interest is on constrained optimization
@@
 and considerations will be given to derivative-free methods.
+
 Derivative-free methods, which do not rely on the knowledge of the
@@
 been developed over the years, where the emphasis is on finding, in most
-cases, a near optimal solution. Below will introduce two random search
-methods: creeping random search and simulated annealing, and a pseudo
-derivative search.
-
-A common search scheme utilised by both derivative-free methods and
-gradient methods is line search. In line search methods, users are
-required to provide an initial estimate \(x_{1}\) and, at each
-iteration, a search direction \(S_k\) and a step size \(\alpha_k\) are
-generated. Then one moves on to the next point following
-\(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
-desire convergence is reached. While gradient-based methods choose the
-search direction by the gradient, derivative-free methods uses local
-information of the objective function to determine the search direction.
-The choice of step size also needs considerations, as inadequate step
-sizes might prevent the optimisation method to converge to an optimum.
-An ideal step size can be chosen via finding the value of
-\(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
-respect to \(\alpha_k\) at each iteration.
-
-Several R implementations address optimization problems with both
-general purpose as well as task specifics solvers. The most prominent
-one within the general solvers is \code{optim()} in the \CRANpkg{stats}
-\citep{stats} package, which provides both gradient-based and
-derivative-free optimisation functions. Another general solver
-specialised in non-linear optimisation is \CRANpkg{nloptr}
-\citep{nloptr}. Specific solvers for simulated annealing include
-\code{optim(..., method  "SANN")} and package \CRANpkg{GenSA}
-\citep{gensa} that deals with more complicated objective functions. For
-other task specific solvers, readers are recommended to visit the
-relevant sections in CRAN task review on
-\ctv{optimisation and mathematical programming}
-\citep{crantaskreviewoptim}.
+cases, a near optimal solution. Below will introduce three
+derivative-free methods, two of which are random search methods:
+creeping random search and simulated annealing, and the other one is
+pseudo derivative search.
+
+Random search {[}in practice, non-convex, non-smooth{]}. The original
+(pure) random search, Brooks, randomly evaluate points in the space
+regardless of the current position. Although gaurantee global
+convergence, as the dimensional increases, blindly sample points in the
+entire space would cost a long time for convergence to occur {[}spall,
+6{]}. Various algorithms improve on the pure random search by either
+concentrating on a narrower sampling space or a different updating
+machanism. Creeping random search, {[}R.C White{]}, is such a variation
+where a candidate point is generated within a neighbourhood of the
+current point and the current point is only updated if the candidate
+point makes an improvment on the function value. {[}no convergence
+guarantee{]}. A different random search algorithm, simulated annealing,
+takes a different updating procedure than creeping random search. Rather
+than only updating the current point when an improvement is made,
+simulated annealing uses a Metropolis acceptance criterion, where
+nonimproving candidates can still have a chance to be accepted. The
+convergence of simulated annealing algorithms has been widely researched
+{[}reference, reference, reference{]} and a global convergence is
+guaranteed {[}under what condition{]}.
+
+Pseudo derivative search uses a common search scheme in optimistaion:
+line search. In line search methods, users are required to provide an
+initial estimate \(x_{1}\) and, at each iteration, a search direction
+\(S_k\) and a step size \(\alpha_k\) are generated. Then one moves on to
+the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and the process
+is repeated until the desire convergence is reached. In derivative-free
+methods, local information of the objective function are used to
+determine the search direction. The choice of step size also needs
+considerations, as inadequate step sizes might prevent the optimisation
+method to converge to an optimum. An ideal step size can be chosen via
+finding the value of \(\alpha_k \in \mathbb{R}\) that maximises
+\(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) at each iteration.
 

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.Rmd
changes on optim methods done

@@
 
-Random search [in practice, non-convex, non-smooth].
-The original (pure) random search, Brooks, randomly evaluate points in the space regardless of the current position.
-Although gaurantee global convergence, as the dimensional increases, blindly sample points in the entire space would cost a long time for convergence to occur [spall, 6].
-Various algorithms improve on the pure random search by either concentrating on a narrower sampling space or a different updating machanism.
-Creeping random search, [R.C White], is such a variation where a candidate point is generated within a neighbourhood of the current point and the current point is only updated if the candidate point makes an improvment on the function value.
-[no convergence guarantee].
-A different random search algorithm, simulated annealing, takes a different updating procedure than creeping random search.
+Random search mehtods [@zabinsky2013stochastic] contain a random sampling component in its algorithm and has shown its ability to optimise non-convex and non-smooth functions.
+The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
+As the dimension of the space becomes larger, sampling from the entire space would indicate a long time for convergence to occur despite a guaranteed global convergence [@spall2005introduction].
+Various algorithms have then been developed to improve pure random search by either concentrating on a narrower sampling space or using a different updating machanism.
+Creeping random search [@White1971-ci] is such a variation where a candidate point is generated within a neighbourhood of the current point.
+This gives creeping random search an advantage of being fast to compute, however, global convergence is not longer guaranteed.
+On the other hand, simulated annealing [@kirkpatrick1983optimization; @bertsimas1993simulated], takes a different updating machanism from the pure random search.
 Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where nonimproving candidates can still have a chance to be accepted.
-The convergence of simulated annealing algorithms has been widely researched [reference, reference, reference] and a global convergence is guaranteed [under what condition].
+The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @belisle1992convergence; @granville1994simulated] and the global optimum can be attained under mild conditions.
 

### 2021-03-17 paper/zhang-cook-laa-langrene-menendez.tex
changes on optim methods done

@@
 
-Random search {[}in practice, non-convex, non-smooth{]}. The original
-(pure) random search, Brooks, randomly evaluate points in the space
-regardless of the current position. Although gaurantee global
-convergence, as the dimensional increases, blindly sample points in the
-entire space would cost a long time for convergence to occur {[}spall,
-6{]}. Various algorithms improve on the pure random search by either
-concentrating on a narrower sampling space or a different updating
-machanism. Creeping random search, {[}R.C White{]}, is such a variation
-where a candidate point is generated within a neighbourhood of the
-current point and the current point is only updated if the candidate
-point makes an improvment on the function value. {[}no convergence
-guarantee{]}. A different random search algorithm, simulated annealing,
-takes a different updating procedure than creeping random search. Rather
-than only updating the current point when an improvement is made,
-simulated annealing uses a Metropolis acceptance criterion, where
-nonimproving candidates can still have a chance to be accepted. The
-convergence of simulated annealing algorithms has been widely researched
-{[}reference, reference, reference{]} and a global convergence is
-guaranteed {[}under what condition{]}.
+Random search mehtods \citep{zabinsky2013stochastic} contain a random
+sampling component in its algorithm and has shown its ability to
+optimise non-convex and non-smooth functions. The initial random search
+algorithm, pure random search \citep{Brooks1958-zk}, draws candidate
+points from the entire space without using any information of the
+current position and updates the current position when an improvement on
+the objective function is made. As the dimension of the space becomes
+larger, sampling from the entire space would indicate a long time for
+convergence to occur despite a guaranteed global convergence
+\citep{spall2005introduction}. Various algorithms have then been
+developed to improve pure random search by either concentrating on a
+narrower sampling space or using a different updating machanism.
+Creeping random search \citep{White1971-ci} is such a variation where a
+candidate point is generated within a neighbourhood of the current
+point. This gives creeping random search an advantage of being fast to
+compute, however, global convergence is not longer guaranteed. On the
+other hand, simulated annealing
+\citep{kirkpatrick1983optimization, bertsimas1993simulated}, takes a
+different updating machanism from the pure random search. Rather than
+only updating the current point when an improvement is made, simulated
+annealing uses a Metropolis acceptance criterion, where nonimproving
+candidates can still have a chance to be accepted. The convergence of
+simulated annealing algorithms has been widely researched
+\citep{mitra1986convergence, belisle1992convergence, granville1994simulated}
+and the global optimum can be attained under mild conditions.
 

### 2021-03-18 paper/zhang-cook-laa-langrene-menendez.Rmd
interesting index polished

@@
 The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some unexpected structure of the data, e.g., bi-modality, clustering and outliers.
+A view is interesting if it can show some structures of the data, e.g., bi-modality, clustering and outliers, other than normal distribution.
 
@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal as $n$ and $p$ approaching infinity under certain conditions.
-@huber1985projection illustrates that, for indices based on entropy measures, [...].
-Index functions  proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution.(Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]); chi-square index [@posse1995projection] for detecting spiral structure; kurtosis [@Loperfido2020] and skewness [@Loperfido2018] indices for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal, as $n$ and $p$ approaching infinity under certain conditions.
+Index functions proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 

### 2021-03-18 paper/zhang-cook-laa-langrene-menendez.tex
interesting index polished

@@
 interesting low-dimensional view. A view is interesting if it can show
-some unexpected structure of the data, e.g., bi-modality, clustering and
-outliers.
+some structures of the data, e.g., bi-modality, clustering and outliers,
+other than normal distribution.
 
@@
 proofs from \citet{diaconis1984asymptotics} show that projections tend
-to be normal as \(n\) and \(p\) approaching infinity under certain
-conditions. \citet{huber1985projection} illustrates that, for indices
-based on entropy measures, {[}\ldots{}{]}. Index functions proposed in
-the litearture include early indices that can be categorised by
-measuring the \(L^2\) distance between the projection and a normal
-distribution.(Legendre index \citep{friedman1974projection}, Hermite
-index \citep{hall1989polynomial}, and natural Hermite index
-\citep{cook1993projection}); chi-square index
-\citep{posse1995projection} for detecting spiral structure; kurtosis
-\citep{Loperfido2020} and skewness \citep{Loperfido2018} indices for
-detecting outliers in financial time series; and most recently,
-scagnostic indices \citep{laa2020using} for summarising structures in
-scatterplot matrices based on eight scagnostic measures.
+to be normal, as \(n\) and \(p\) approaching infinity under certain
+conditions. Index functions proposed in the litearture include early
+indices that can be categorised by measuring the \(L^2\) distance
+between the projection and a normal distribution: Legendre index
+\citep{friedman1974projection}, Hermite index
+\citep{hall1989polynomial}, and natural Hermite index
+\citep{cook1993projection}; chi-square index \citep{posse1995projection}
+for detecting spiral structure; kurtosis index \citep{Loperfido2020} and
+skewness index \citep{Loperfido2018} for detecting outliers in financial
+time series; and most recently, scagnostic indices \citep{laa2020using}
+for summarising structures in scatterplot matrices based on eight
+scagnostic measures.
 

### 2021-03-22 paper/zhang-cook-laa-langrene-menendez.Rmd
Small changes in the first few sections.

@@
 Visualisation is widely used in exploratory data analysis [@tukey1977exploratory; @unwin2015graphical;  @healy2018data; @wilke2019fundamentals].
-Presenting information in graphics often unveils information that would otherwise not be discovered and provides a more comprehensive understanding of the problem at hand.
-Task specific tools such as @li2020visualizing show how visualisation can be used to understand, for instance, the behaviour of neural network on classification models.
+Presenting information in graphics often unveils insights that would otherwise not be discovered and provides a more comprehensive understanding of the problem at hand.
+Task specific tools such as @li2020visualizing show how visualisation can be used to understand, for instance, the behaviour of the optimisation for the example of neural network classification models.
 However, no general visualisation tool is available for diagnosing optimisation procedures.
@@
 
-The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is decided by an objective function.
+The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is defined via an objective function.
 Broadly speaking, optimization can be unconstrained or constrained [@kelley1999iterative].
@@
 The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand].
-The goal of the optimisation is to identify the projection, characterised by the projection basis, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some structures of the data, e.g., bi-modality, clustering and outliers, other than normal distribution.
+The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view.
+A view is interesting if it can show some structures of the data other than a normal distribution, e.g., bi-modality, clustering and outliers.
 
 The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general.
-Examples of those include the existence of multiple optima (local and global), the trade off between computational burden and proximity to the optima,  dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient].
-Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily applied to any other optimization problems.
+Examples include the existence of multiple optima (local and global), the trade-off between computational burden and proximity to the optima, or dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient].
+The visualization tools and optimization methods presented in this paper can therefore easily be applied to other optimization problems.
 
@@
 A data structure is defined to capture information during the optimisation, and different diagnostic plots are designed with the data collected.
-Section \ref{application} shows applications of how these diagnostic plots can be used to discover interesting aspects of different optimisers and guide improvement to the existing algorithms.
+Section \ref{application} shows applications of how these diagnostic plots can be used to discover interesting aspects of different optimisers and guide improvements to the existing algorithms.
 Finally, Section \ref{implementation} describes the R package: \pkg{ferrn}, that implements the visual diagnostics.
@@
 
-A large class of methods utilises the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method.
-Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the complexity of the constraints.
-As will be shown in the next section, the independent variable, in our optimisation problem, is a matrix of projection basis and the computational time required to perform differentation on a matrix could impede the render of tour animation.
+A large class of methods uses the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method.
+Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the limited complexity of the constraints.
+As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the render of tour animation.
 Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods.
@@
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
-Below will introduce three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo derivative search.
+Here we consider three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo derivative search.
 
-Random search mehtods [@zabinsky2013stochastic] contain a random sampling component in its algorithm and has shown its ability to optimise non-convex and non-smooth functions.
+Random search methods [@zabinsky2013stochastic] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
 The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
-As the dimension of the space becomes larger, sampling from the entire space would indicate a long time for convergence to occur despite a guaranteed global convergence [@spall2005introduction].
-Various algorithms have then been developed to improve pure random search by either concentrating on a narrower sampling space or using a different updating machanism.
-Creeping random search [@White1971-ci] is such a variation where a candidate point is generated within a neighbourhood of the current point.
-This gives creeping random search an advantage of being fast to compute, however, global convergence is not longer guaranteed.
-On the other hand, simulated annealing [@kirkpatrick1983optimization; @bertsimas1993simulated], takes a different updating machanism from the pure random search.
-Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where nonimproving candidates can still have a chance to be accepted.
+As the dimension of the space becomes larger, sufficient sampling from the entire space would require a long time for convergence to occur, despite a guaranteed global convergence [@spall2005introduction].
+Various algorithms have thus been developed to improve pure random search by either concentrating on a narrower sampling space or using a different updating machanism.
+Creeping random search [@White1971-ci] is such a variation, where a candidate point is generated within a neighbourhood of the current point.
+This makes creeping random search faster to compute, however, global convergence is not longer guaranteed.
+On the other hand, simulated annealing [@kirkpatrick1983optimization; @bertsimas1993simulated], introduces a different updating mechanism.
+Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where worse candidates still have a chance to be accepted.
 The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @belisle1992convergence; @granville1994simulated] and the global optimum can be attained under mild conditions.
@@
 In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
-Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desire convergence is reached.
-In derivative-free methods, local information of the objective function are used to determine the search direction.
-The choice of step size also needs considerations, as inadequate step sizes might prevent the optimisation method to converge to an optimum.
-An ideal step size can be chosen via finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.
+Then one moves on to the next point following $x_{k+1} = x_k + \alpha_kS_k$ and the process is repeated until the desired convergence is reached.
+In derivative-free methods, local information of the objective function is used to determine the search direction.
+The choice of step size also needs consideration, as inadequate step sizes might prevent the optimisation method to converge to an optimum.
+An ideal step size can be chosen by finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.
 
@@
 Let $\mathbf{X}_{n \times p}$ be the data matrix with $n$ observations in $p$ dimensions.
-A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis matrix.
+A $d$-dimensional projection is a linear transformation from $\mathbb{R}^p$ into $\mathbb{R}^d$ defined as $\mathbf{Y} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{Y}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection matrix.
 We define $f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}$ to be an index function that maps the projected data $\mathbf{Y}$ onto a scalar value.
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal, as $n$ and $p$ approaching infinity under certain conditions.
+An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
 Index functions proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
@@
 As a general visualisation method, tour produces animations of high dimensional data via rotations of low-dimensional planes.
-There are different tour implementations depending on how the high dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
+There are different versions depending on how the high dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
 Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection.
@@
 
-In projection pursuit the optimisation problem aims at finding the global and local maxima that give interesting projections according to an index function.
+In projection pursuit the optimisation aims at finding the global and local maxima that give interesting projections according to an index function.
 That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
@@
 3) While finding the global optimum is the goal of the optimisation problem, interesting projections may also appear in the local optimum.
-4) The optimisation should be computationally agile since the tour animation is viewed by the users during the optimisation.
+4) The optimisation should be fast to compute since the tour animation is viewed by the users during the optimisation.
 
@@
 The optimiser terminates when the maximum number of iteration $l_{\max}$ is reached before a better basis can be found.
-The algorithm of SA is summarised in Algorithm \ref{random-search} in the appendix.
+The algorithm of CRS is summarised in Algorithm \ref{random-search} in the appendix.
 @posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
@@
 Given an initial value of $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, SAJO gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
-This implementation allows the optimiser to make a move and explore the basis space even the landing basis does have a higher index value and hence enable the optimiser to jump out of a local optimum.
-The algorithm \ref{simulated-annealing} in the appendix highlights how SAJO differs from SA in the inner loop.
+When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enable the optimiser to jump out of a local optimum.
+The algorithm \ref{simulated-annealing} in the appendix highlights how SA differs from CRS in the inner loop.
 
-Pseudo derivative (PD) search [@cook1995grand] uses a different strategy than SA and SAJO.
-Rather than randomly sample the basis space, PD first computes a search direction by evaluating vases close to the current basis.
-A step size is then chosen along the geodesic direction by another optimisation over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+Pseudo derivative (PD) search [@cook1995grand] uses a different strategy than CRS and SA.
+Rather than randomly sample the basis space, PD first computes a search direction by evaluating bases close to the current basis.
+A step size is then chosen along the corresponding geodesic by another optimisation over an 90 degree angle from $-\pi/4$ to $\pi/4$.
 The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.Rmd
Some changes in visual diagnostics section.

@@
 
-A starting point of diagnosing an optimiser is to understand how many searches an optimiser has conducted.
-One may want to simply plot the index value of the search points across its natural order, but a point geometry may work well if each iteration has a similar number of points.
-When some iterations have considerably more points than others, using a point geometry over-emphasizes the iterations that have more search points since these iterations will occupy the mast majority of the plot space.
-An alternative is to summarise the search in each iteration using boxplots and each iteration will then be spaced out equally.
+A starting point of diagnosing an optimiser is to understand how many searches it has conducted, i.e.
+we to summarise how the index is increasing over iterations and how many tries need to be sampled at each iteration.
+This is achieved using the search plot function: a boxplot shows the distribution of index values for each try, where the selected basis (corresponding to the highest index value) is always shown as a point.
+In case of few tries at a given iteration showing the data points directly may be preferred over the boxplot, this is controlled via the `cutoff` argument.
+Additional annotations are added to facilitate better reading of the plot and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
+
+<!-- One may want to simply plot the index value of the search points across its natural order.
+While such a point geometry may work well if each iteration has a similar number of points, when some iterations have considerably more points than others, it will over-emphasizes the iterations that have more search points and these iterations will occupy the vast majority of the plot space.
+An alternative is to summarise the search in each iteration using boxplots, with each iteration spaced out equally.
 Occasionally, one may still want to switch back to a point geometry if the number of points is small in a particular iteration and this can be achieved via the `cutoff` argument in the search plot function.
@@
 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
+-->
 
@@
 The anchor bases, the ones found with the highest index value in each iteration, always have an increased index value in the optimiser CRS while this is not the case for SA.
-This feature gives SA an advantage in this simple example to quickly find the optimum.
+This feature gives CRS an advantage in this simple example to quickly find the optimum.
 
@@
 
-Another interesting diagnostic is to examine how the index value improves between interpolating bases since the projection on these bases is played by the tour animation.
+Another interesting diagnostic is to examine how the index value improves between interpolating bases since the projection on these bases is shown in the tour animation.
 Trace plots are created by plotting the index value against time.
-Figure \ref{fig:toy-interp} presents the trace plot of the same optimisers as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
-It may seem bizarre at the first sight that the interpolation in the left panel first passes some bases with higher index value before decreases to a lower target.
+Figure \ref{fig:toy-interp} presents the trace plot of the same optimisations as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
+It may seem bizarre at the first sight that the interpolation sometimes passes bases with higher index value before it decreases to a lower target.
 This happens because on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
-On the other hand, guided tour interpolate between the current and target basis to provide a smooth transition between projections.
-These two facts indicate that a non-monotonic interpolation can't be avoided for SA.
+In addition, the guided tour interpolates between the current and target basis to provide a smooth transition between projections, and sometimes a higher index value will be observed along the interpolation path.
+This indicates that a non-monotonic interpolation can't be avoided, even for CRS.
 Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for CRS.
@@
 
-Apart from checking the search and progression of an optimiser, looking at where the bases are positioned in the basis space is another interest.
-Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$-dimensional sphere.
+Apart from checking the search and progression of an optimiser, looking at where the bases are positioned in the basis space is also of interest.
+Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a Stiefel manifold.
+For one-dimensional projections this forms a $p$-dimensional sphere.
 A dimensionality reduction method, e.g.
@@
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to the sign difference of the bases and so does the index value, however, this creates difficulties for comparing the optimisers since the end bases will be symmetric with respect to the origin.
+This happens because the projection is invariant to orientation of the basis and so is the index value.
+However, this creates difficulties for comparing the optimisers since the end bases will be symmetric with respect to the origin.
 A sign flipping step is conducted to flip the signs of all the bases in one routine if different optimisations finish at opposite places.
@@
 Both optimisers find the optimum but PD gets closer.
-With PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
+With the PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
 There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in details in Section \ref{monotonic}.
@@
 While the previous few examples have looked at the space of 1D bases in a unit sphere, this section visualises the space of 2D bases.
-Recall that the columns in a 2D basis are orthogonal to each other, the space of $p \times 2$ bases is a $p$-D torus.
-If take $p = 3$, one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
-The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality of the columns in the bases.
-Two paths optimised by CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous few 1D space plots.
+Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a $p$-D torus.
+For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
+The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality.
+Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots.
 The final basis found by PD and CRS are shown in a larger shape  and printed below respectively:
@@
 
-Both optimisers have found the third variable in the first column, but the second variable in the second column differs by a sign.One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and be symmetric when projected onto the other (due the the sign difference in the second column).
+Both optimisers have found the third variable in the first direction, and the second variable in the second direction.
+Note however the different orientation of the basis, following from the different sign in the second column.
+One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and be symmetric when projected onto the other (due the the sign difference in the second column).
 In Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.Rmd
proper citation of ferrn

@@
 These diagnostics and this workflow can be applied to a broad class of optimisers, to assess their performance.
-An R package, \pkg{ferrn}, has been created to implement the diagnostics.
+An R package, \CRANpkg{ferrn} [@ferrn], has been created to implement the diagnostics.
 
@@
 Section \ref{application} shows applications of how these diagnostic plots can be used to discover interesting aspects of different optimisers and guide improvements to the existing algorithms.
-Finally, Section \ref{implementation} describes the R package: \pkg{ferrn}, that implements the visual diagnostics.
+Finally, Section \ref{implementation} describes the R package: \CRANpkg{ferrn} [@ferrn], that implements the visual diagnostics.
 
@@
 
-The implementation of this projection has been divided into two packages: the data collection object is implemented in the existing CRAN package \CRANpkg{tourr} [@tourr] while the optimiser diagnostics have been implemented in a new package, \pkg{ferrn}.
-When a guided tour is run, the users can choose to collect the data by binding `animate_*()` to a variable.
-Once the data object has been obtained, the \pkg{ferrn} package provides diagnostic plots shown in Section \ref{vis-diag}.
-The main functions of the \pkg{ferrn} package functionality is listed below.
+This project contributes to the software development in two packages: a data collection object is implemented in \CRANpkg{tourr} [@tourr], while the visual diagnostics of the optimisers is implemented in \CRANpkg{ferrn} [@ferrn].
+Below list functions in the \CRANpkg{ferrn} [@ferrn] package:
 

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.tex
proper citation of ferrn

@@
 % !TeX root = RJwrapper.tex
-\title{Visual Diagnostics for Constrained Optimisation with Application to
-Guided Tours}
-\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, and Patricia Menéndez}
+\title{Visual Diagnostics for Constrained Optimisation with Application
+to Guided Tours}
+\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
+Langrené, and Patricia Menéndez}
 
@@
 a broad class of optimisers, to assess their performance. An R package,
-\pkg{ferrn}, has been created to implement the diagnostics.
+\CRANpkg{ferrn} \citep{ferrn}, has been created to implement the
+diagnostics.
 
@@
 \citep{tukey1977exploratory, unwin2015graphical, healy2018data, wilke2019fundamentals}.
-Presenting information in graphics often unveils information that would
+Presenting information in graphics often unveils insights that would
 otherwise not be discovered and provides a more comprehensive
@@
 \citet{li2020visualizing} show how visualisation can be used to
-understand, for instance, the behaviour of neural network on
-classification models. However, no general visualisation tool is
-available for diagnosing optimisation procedures. The work presented in
-this paper brings visualization tools into optimisation problems with
-the aim to better understand the performance of optimisers in practice.
+understand, for instance, the behaviour of the optimisation for the
+example of neural network classification models. However, no general
+visualisation tool is available for diagnosing optimisation procedures.
+The work presented in this paper brings visualization tools into
+optimisation problems with the aim to better understand the performance
+of optimisers in practice.
 
@@
 the space of all feasible solutions where typically the best solution is
-decided by an objective function. Broadly speaking, optimization can be
+defined via an objective function. Broadly speaking, optimization can be
 unconstrained or constrained \citep{kelley1999iterative}. The
@@
 \citep{cook2008grand}. The goal of the optimisation is to identify the
-projection, characterised by the projection basis, that gives the most
+projection, represented by the projection matrix, that gives the most
 interesting low-dimensional view. A view is interesting if it can show
-some structures of the data, e.g., bi-modality, clustering and outliers,
-other than normal distribution.
+some structures of the data other than a normal distribution, e.g.,
+bi-modality, clustering and outliers.
 
 The optimization challenges encountered in the projection pursuit guided
-tour problem are common to those of optimization in general. Examples of
-those include the existence of multiple optima (local and global), the
-trade off between computational burden and proximity to the optima,
+tour problem are common to those of optimization in general. Examples
+include the existence of multiple optima (local and global), the
+trade-off between computational burden and proximity to the optima, or
 dealing with noisy objective functions that might be non-smooth and
-non-differentiable \citep{jones1998efficient}. Those are not unique to
-this context and therefore the visualization tools and optimization
-methods presented in this paper can be easily applied to any other
-optimization problems.
+non-differentiable \citep{jones1998efficient}. The visualization tools
+and optimization methods presented in this paper can therefore easily be
+applied to other optimization problems.
 
@@
 be used to discover interesting aspects of different optimisers and
-guide improvement to the existing algorithms. Finally, Section
-\ref{implementation} describes the R package: \pkg{ferrn}, that
-implements the visual diagnostics.
+guide improvements to the existing algorithms. Finally, Section
+\ref{implementation} describes the R package: \CRANpkg{ferrn}
+\citep{ferrn}, that implements the visual diagnostics.
 
@@
 
-A large class of methods utilises the gradient information of the
-objective function, with the most notable one being the gradient ascent
-(descent) method. Although gradient optimization methods are popular,
-they rely on the availability of the objective function derivatives and
-on the complexity of the constraints. As will be shown in the next
-section, the independent variable, in our optimisation problem, is a
-matrix of projection basis and the computational time required to
-perform differentation on a matrix could impede the render of tour
+A large class of methods uses the gradient information of the objective
+function, with the most notable one being the gradient ascent (descent)
+method. Although gradient optimization methods are popular, they rely on
+the availability of the objective function derivatives and on the
+limited complexity of the constraints. As will be shown in the next
+section, the independent variables in our optimisation problem are the
+entries of a projection matrix and the computational time required to
+perform differentiation on a matrix could impede the render of tour
 animation. Hence, gradient-based methods are not the focus of this paper
@@
 been developed over the years, where the emphasis is on finding, in most
-cases, a near optimal solution. Below will introduce three
-derivative-free methods, two of which are random search methods:
-creeping random search and simulated annealing, and the other one is
-pseudo derivative search.
-
-Random search mehtods \citep{zabinsky2013stochastic} contain a random
-sampling component in its algorithm and has shown its ability to
-optimise non-convex and non-smooth functions. The initial random search
-algorithm, pure random search \citep{Brooks1958-zk}, draws candidate
-points from the entire space without using any information of the
-current position and updates the current position when an improvement on
-the objective function is made. As the dimension of the space becomes
-larger, sampling from the entire space would indicate a long time for
-convergence to occur despite a guaranteed global convergence
-\citep{spall2005introduction}. Various algorithms have then been
-developed to improve pure random search by either concentrating on a
-narrower sampling space or using a different updating machanism.
-Creeping random search \citep{White1971-ci} is such a variation where a
-candidate point is generated within a neighbourhood of the current
-point. This gives creeping random search an advantage of being fast to
-compute, however, global convergence is not longer guaranteed. On the
-other hand, simulated annealing
-\citep{kirkpatrick1983optimization, bertsimas1993simulated}, takes a
-different updating machanism from the pure random search. Rather than
-only updating the current point when an improvement is made, simulated
-annealing uses a Metropolis acceptance criterion, where nonimproving
-candidates can still have a chance to be accepted. The convergence of
-simulated annealing algorithms has been widely researched
+cases, a near optimal solution. Here we consider three derivative-free
+methods, two of which are random search methods: creeping random search
+and simulated annealing, and the other one is pseudo derivative search.
+
+Random search methods \citep{zabinsky2013stochastic} have a random
+sampling component as part of their algorithm and have been shown to
+have the ability to optimise non-convex and non-smooth functions. The
+initial random search algorithm, pure random search
+\citep{Brooks1958-zk}, draws candidate points from the entire space
+without using any information of the current position and updates the
+current position when an improvement on the objective function is made.
+As the dimension of the space becomes larger, sufficient sampling from
+the entire space would require a long time for convergence to occur,
+despite a guaranteed global convergence \citep{spall2005introduction}.
+Various algorithms have thus been developed to improve pure random
+search by either concentrating on a narrower sampling space or using a
+different updating machanism. Creeping random search
+\citep{White1971-ci} is such a variation, where a candidate point is
+generated within a neighbourhood of the current point. This makes
+creeping random search faster to compute, however, global convergence is
+not longer guaranteed. On the other hand, simulated annealing
+\citep{kirkpatrick1983optimization, bertsimas1993simulated}, introduces
+a different updating mechanism. Rather than only updating the current
+point when an improvement is made, simulated annealing uses a Metropolis
+acceptance criterion, where worse candidates still have a chance to be
+accepted. The convergence of simulated annealing algorithms has been
+widely researched
 \citep{mitra1986convergence, belisle1992convergence, granville1994simulated}
@@
 the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and the process
-is repeated until the desire convergence is reached. In derivative-free
-methods, local information of the objective function are used to
+is repeated until the desired convergence is reached. In derivative-free
+methods, local information of the objective function is used to
 determine the search direction. The choice of step size also needs
-considerations, as inadequate step sizes might prevent the optimisation
-method to converge to an optimum. An ideal step size can be chosen via
+consideration, as inadequate step sizes might prevent the optimisation
+method to converge to an optimum. An ideal step size can be chosen by
 finding the value of \(\alpha_k \in \mathbb{R}\) that maximises
@@
 \(\mathbf{Y}_{n \times d}\) is the projected data and
-\(\mathbf{A}_{p\times d}\) is the projection basis matrix. We define
+\(\mathbf{A}_{p\times d}\) is the projection matrix. We define
 \(f: \mathbb{R}^{n \times d} \mapsto \mathbb{R}\) to be an index
@@
 proofs from \citet{diaconis1984asymptotics} show that projections tend
-to be normal, as \(n\) and \(p\) approaching infinity under certain
+to be normal, as \(n\) and \(p\) approach infinity under certain
 conditions. Index functions proposed in the litearture include early
@@
 dimensional data via rotations of low-dimensional planes. There are
-different tour implementations depending on how the high dimensional
-space is investigated: grand tour \citep{cook2008grand} selects the
-planes randomly to provide a general overview; manual tour
+different versions depending on how the high dimensional space is
+investigated: grand tour \citep{cook2008grand} selects the planes
+randomly to provide a general overview; manual tour
 \citep{cook1997manual} gradually phases in and out one variable to
@@
 
-In projection pursuit the optimisation problem aims at finding the
-global and local maxima that give interesting projections according to
-an index function. That is, it starts with a given randomly selected
-basis \(\mathbf{A}_1\) and aims at finding an optimal final projection
-basis \(\mathbf{A}_T\) that satisfies the following optimisation
-problem:
+In projection pursuit the optimisation aims at finding the global and
+local maxima that give interesting projections according to an index
+function. That is, it starts with a given randomly selected basis
+\(\mathbf{A}_1\) and aims at finding an optimal final projection basis
+\(\mathbf{A}_T\) that satisfies the following optimisation problem:
 
@@
 optimisation problem, interesting projections may also appear in the
-local optimum. 4) The optimisation should be computationally agile since
-the tour animation is viewed by the users during the optimisation.
+local optimum. 4) The optimisation should be fast to compute since the
+tour animation is viewed by the users during the optimisation.
 
@@
 \(l_{\max}\) is reached before a better basis can be found. The
-algorithm of SA is summarised in Algorithm \ref{random-search} in the
+algorithm of CRS is summarised in Algorithm \ref{random-search} in the
 appendix. \citet{posse1995projection} has proposed a slightly different
@@
 \(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to have
-an index value larger than the current basis, SAJO gives it a second
+an index value larger than the current basis, SA gives it a second
 chance to be accepted with probability
@@
 implementation allows the optimiser to make a move and explore the basis
-space even the landing basis does have a higher index value and hence
-enable the optimiser to jump out of a local optimum. The algorithm
-\ref{simulated-annealing} in the appendix highlights how SAJO differs
-from SA in the inner loop.
+space even if the candidate basis does not have a higher index value and
+hence enable the optimiser to jump out of a local optimum. The algorithm
+\ref{simulated-annealing} in the appendix highlights how SA differs from
+CRS in the inner loop.
 
 Pseudo derivative (PD) search \citep{cook1995grand} uses a different
-strategy than SA and SAJO. Rather than randomly sample the basis space,
-PD first computes a search direction by evaluating vases close to the
-current basis. A step size is then chosen along the geodesic direction
-by another optimisation over an 90 degree angle from \(-\pi/4\) to
-\(\pi/4\). The resulting candidate basis \(\mathbf{A}_{**}\) is returned
-for the current iteration if it has a higher index value than the
-current one. Algorithm \ref{search-geodesic} in the appendix summarises
-the inner loop of the PD.
+strategy than CRS and SA. Rather than randomly sample the basis space,
+PD first computes a search direction by evaluating bases close to the
+current basis. A step size is then chosen along the corresponding
+geodesic by another optimisation over an 90 degree angle from \(-\pi/4\)
+to \(\pi/4\). The resulting candidate basis \(\mathbf{A}_{**}\) is
+returned for the current iteration if it has a higher index value than
+the current one. Algorithm \ref{search-geodesic} in the appendix
+summarises the inner loop of the PD.
 
@@
 A starting point of diagnosing an optimiser is to understand how many
-searches an optimiser has conducted. One may want to simply plot the
-index value of the search points across its natural order, but a point
-geometry may work well if each iteration has a similar number of points.
-When some iterations have considerably more points than others, using a
-point geometry over-emphasizes the iterations that have more search
-points since these iterations will occupy the mast majority of the plot
-space. An alternative is to summarise the search in each iteration using
-boxplots and each iteration will then be spaced out equally.
-Occasionally, one may still want to switch back to a point geometry if
-the number of points is small in a particular iteration and this can be
-achieved via the \texttt{cutoff} argument in the search plot function.
-Additional annotations are added to facilitate better reading of the
-plot and these include\\
-1) the number of points searched in each iteration can be added as text
-label at the bottom of each iteration; 2) the anchor bases to
-interpolate are connected and highlighted in a larger size; and 3) the
-colour of the last iteration is in a grey scale to indicate no better
-basis found in this iteration.
+searches it has conducted, i.e.~we to summarise how the index is
+increasing over iterations and how many tries need to be sampled at each
+iteration. This is achieved using the search plot function: a boxplot
+shows the distribution of index values for each try, where the selected
+basis (corresponding to the highest index value) is always shown as a
+point. In case of few tries at a given iteration showing the data points
+directly may be preferred over the boxplot, this is controlled via the
+\texttt{cutoff} argument. Additional annotations are added to facilitate
+better reading of the plot and these include 1) the number of points
+searched in each iteration can be added as text label at the bottom of
+each iteration; 2) the anchor bases to interpolate are connected and
+highlighted in a larger size; and 3) the colour of the last iteration is
+in a grey scale to indicate no better basis found in this iteration.
 
@@
 value in each iteration, always have an increased index value in the
-optimiser CRS while this is not the case for SA. This feature gives SA
+optimiser CRS while this is not the case for SA. This feature gives CRS
 an advantage in this simple example to quickly find the optimum.
@@
 improves between interpolating bases since the projection on these bases
-is played by the tour animation. Trace plots are created by plotting the
+is shown in the tour animation. Trace plots are created by plotting the
 index value against time. Figure \ref{fig:toy-interp} presents the trace
-plot of the same optimisers as Figure \ref{fig:toy-search} and one can
-observe that the trace is smooth in both cases. It may seem bizarre at
-the first sight that the interpolation in the left panel first passes
-some bases with higher index value before decreases to a lower target.
-This happens because on the one hand, the probabilistic acceptance in SA
-implies that some worse bases will be accepted by the optimiser. On the
-other hand, guided tour interpolate between the current and target basis
-to provide a smooth transition between projections. These two facts
-indicate that a non-monotonic interpolation can't be avoided for SA.
-Later, in section \ref{monotonic}, there will be a discussion on
-improving the non-monotonic interpolation for CRS.
+plot of the same optimisations as Figure \ref{fig:toy-search} and one
+can observe that the trace is smooth in both cases. It may seem bizarre
+at the first sight that the interpolation sometimes passes bases with
+higher index value before it decreases to a lower target. This happens
+because on the one hand, the probabilistic acceptance in SA implies that
+some worse bases will be accepted by the optimiser. In addition, the
+guided tour interpolates between the current and target basis to provide
+a smooth transition between projections, and sometimes a higher index
+value will be observed along the interpolation path. This indicates that
+a non-monotonic interpolation can't be avoided, even for CRS. Later, in
+section \ref{monotonic}, there will be a discussion on improving the
+non-monotonic interpolation for CRS.
 
@@
 Apart from checking the search and progression of an optimiser, looking
-at where the bases are positioned in the basis space is another
+at where the bases are positioned in the basis space is also of
 interest. Given the orthonormality constraint, the space of projection
-bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\)-dimensional
-sphere. A dimensionality reduction method, e.g.~principal component
-analysis is applied to first project all the bases onto a 2D space. In a
-projection pursuit guided tour optimisation, there are various types of
-bases involved: 1) The starting basis; 2) The search bases that the
-optimiser evaluated to produce the anchor bases; 3) The anchor bases
-that have the highest index value in each iteration; 4) The
-interpolating bases on the interpolation path; and finally 5) the end
-basis. The importance of these bases differs and the most important ones
-are the starting, interpolating and end bases. The anchor and search
-bases can be turned on with argument \texttt{details\ =\ TRUE}.
-Sometimes, two optimisers can start with the same basis but finish with
-bases of opposite sign. This happens because the projection is invariant
-to the sign difference of the bases and so does the index value,
-however, this creates difficulties for comparing the optimisers since
-the end bases will be symmetric with respect to the origin. A sign
-flipping step is conducted to flip the signs of all the bases in one
-routine if different optimisations finish at opposite places.
+bases \(\mathbf{A}_{p \times d}\) is a Stiefel manifold. For
+one-dimensional projections this forms a \(p\)-dimensional sphere. A
+dimensionality reduction method, e.g.~principal component analysis is
+applied to first project all the bases onto a 2D space. In a projection
+pursuit guided tour optimisation, there are various types of bases
+involved: 1) The starting basis; 2) The search bases that the optimiser
+evaluated to produce the anchor bases; 3) The anchor bases that have the
+highest index value in each iteration; 4) The interpolating bases on the
+interpolation path; and finally 5) the end basis. The importance of
+these bases differs and the most important ones are the starting,
+interpolating and end bases. The anchor and search bases can be turned
+on with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers
+can start with the same basis but finish with bases of opposite sign.
+This happens because the projection is invariant to orientation of the
+basis and so is the index value. However, this creates difficulties for
+comparing the optimisers since the end bases will be symmetric with
+respect to the origin. A sign flipping step is conducted to flip the
+signs of all the bases in one routine if different optimisations finish
+at opposite places.
 
@@
 projection problem. Both optimisers find the optimum but PD gets closer.
-With PCA plot, one can visually appreciate the nature of these two
+With the PCA plot, one can visually appreciate the nature of these two
 optimisers: PD first evaluates points in a small neighbourhood for a
@@
 a unit sphere, this section visualises the space of 2D bases. Recall
-that the columns in a 2D basis are orthogonal to each other, the space
-of \(p \times 2\) bases is a \(p\)-D torus. If take \(p = 3\), one would
-see a classical 3D torus shape as shown by the grey points in Figure
-\ref{fig:toy-torus}. The two circles of the torus can be observed to be
-perpendicular to each other and this can be linked back to the
-orthogonality of the columns in the bases. Two paths optimised by CRS
-and PD are plotted on top of the torus and coloured in green and brown
-respectively to match the previous few 1D space plots. The final basis
-found by PD and CRS are shown in a larger shape and printed below
-respectively:
+that the columns in a 2D basis are orthogonal to each other, so the
+space of \(p \times 2\) bases is a \(p\)-D torus. For \(p = 3\) one
+would see a classical 3D torus shape as shown by the grey points in
+Figure \ref{fig:toy-torus}. The two circles of the torus can be observed
+to be perpendicular to each other and this can be linked back to the
+orthogonality. Two paths from CRS and PD are plotted on top of the torus
+and coloured in green and brown respectively to match the previous
+plots. The final basis found by PD and CRS are shown in a larger shape
+and printed below respectively:
 
@@
 
-Both optimisers have found the third variable in the first column, but
-the second variable in the second column differs by a sign.One would
-expect to see this in the torus plot as the final bases match each other
-when projected onto one torus circle (due to the same sign in the first
-column) and be symmetric when projected onto the other (due the the sign
-difference in the second column). In Figure \ref{fig:toy-torus}, this
-can be seen most clearly in frame 5 where the two circles are rotated
-into a line from our view.
+Both optimisers have found the third variable in the first direction,
+and the second variable in the second direction. Note however the
+different orientation of the basis, following from the different sign in
+the second column. One would expect to see this in the torus plot as the
+final bases match each other when projected onto one torus circle (due
+to the same sign in the first column) and be symmetric when projected
+onto the other (due the the sign difference in the second column). In
+Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5
+where the two circles are rotated into a line from our view.
 
@@
 
-The implementation of this projection has been divided into two
-packages: the data collection object is implemented in the existing CRAN
-package \CRANpkg{tourr} \citep{tourr} while the optimiser diagnostics
-have been implemented in a new package, \pkg{ferrn}. When a guided tour
-is run, the users can choose to collect the data by binding
-\texttt{animate\_*()} to a variable. Once the data object has been
-obtained, the \pkg{ferrn} package provides diagnostic plots shown in
-Section \ref{vis-diag}. The main functions of the \pkg{ferrn} package
-functionality is listed below.
+This project contributes to the software development in two packages: a
+data collection object is implemented in \CRANpkg{tourr} \citep{tourr},
+while the visual diagnostics of the optimisers is implemented in
+\CRANpkg{ferrn} \citep{ferrn}. Below list functions in the
+\CRANpkg{ferrn} \citep{ferrn} package:
 

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.Rmd
Small changes in application section.

--- 2021-03-23 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 For SA, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for CRS.
-The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure \ref{fig:interruption} and one can observe clearly that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure \ref{fig:interruption} and one can observe that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
@@
 Once the final basis has been found by an optimiser, one may want to push further in the close neighbourhood to see if an even better basis can be found.
-A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local breakthrough.
+A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local improvements.
 The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space,  percentage change in the index value, and neighbourhood size:
@@
 
-Figure \ref{fig:polish} presents the projectd data and trace plot of a 2D projection, optimised by CRS and followed by the polish.
-In the top row shows the intial projection, the final projection after CRS, and the final projection after polish, repsectively.
+Figure \ref{fig:polish} presents the projectd data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
+The top row shows the intial projection, the final projection after CRS, and the final projection after polish, repsectively.
 The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
@@
 The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indices.
-A 1D projection function, $I^{K}(n)$, compar‚es the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, based on the Kolmogorov test.
+An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
 Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index, $I^{nk}(n)$, is defined as:
@@
 
-Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated path is smooth.
-There is barely any improvement made by PD, indicating its failure in optimising non-smooth index.
-While CRS and SA have managed to get close to the index value of the theoretical best, trace plot shows that it takes SA much longer to interpolate towards the final basis.
+Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated paths are smooth.
+There is barely any improvement made by PD, indicating its failure in optimising non-smooth index functions.
+While CRS and SA have managed to get close to the index value of the theoretical best, the trace plot shows that it takes SA much longer to find the final basis.
 This long interpolation path is partially due to the fluctuation in the early iterations, where the SA tends to generously accept inferior bases before concentrating near the optimum.
@@
 Combining the insights from both the trace and PCA plot, one can learn the two different search strategies by CRS and SA: CRS frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SA first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
-The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as what SA has done, can lead to a more efficient search, in terms of the overall number of points evaluated.
+The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as SA has done here, can lead to a more efficient search, in terms of the overall number of points evaluated.
 
@@
 With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
-This indicates that SA, with a reasonable large search size, is able to overcome the local optimum and optimise close towards the global optimum.
+This indicates that SA, with a reasonable large search window, is able to overcome the local optimum and optimise close towards the global optimum.
 

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.tex
Small changes in application section.

@@
 a broad class of optimisers, to assess their performance. An R package,
-\CRANpkg{ferrn} \citep{ferrn}, has been created to implement the
-diagnostics.
+\pkg{ferrn}, has been created to implement the diagnostics.
 
@@
 guide improvements to the existing algorithms. Finally, Section
-\ref{implementation} describes the R package: \CRANpkg{ferrn}
-\citep{ferrn}, that implements the visual diagnostics.
+\ref{implementation} describes the R package: \pkg{ferrn}, that
+implements the visual diagnostics.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour-path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/ursulalaa/papers/paper-tour-vis/img/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-torus-1}
 
@@
 for a 2D projection problem, optimised by CRS, is shown on the left
-panel of Figure \ref{fig:interruption} and one can observe clearly that
-the non-monotonic interpolation has undermined the optimiser to realise
-its full potential. Hence, an interruption is implemented to stop at the
+panel of Figure \ref{fig:interruption} and one can observe that the
+non-monotonic interpolation has undermined the optimiser to realise its
+full potential. Hence, an interruption is implemented to stop at the
 best basis found in the interpolation. The right panel of Figure
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/polish-1}
 
@@
 can be found. A polish search takes the final basis of an optimiser as
-the start of a new guided tour to search for local breakthrough. The
+the start of a new guided tour to search for local improvements. The
 polish algorithm is similar to the CRS but with three distinctions: 1) a
@@
 Figure \ref{fig:polish} presents the projectd data and trace plot of a
-2D projection, optimised by CRS and followed by the polish. In the top
+2D projection, optimised by CRS and followed by the polish step. The top
 row shows the intial projection, the final projection after CRS, and the
@@
 produces a smooth interpolation, while this is not the case for all the
-indices. A 1D projection function, \(I^{K}(n)\), compar‚es the projected
-data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated normal
-distribution, \(\mathcal{N}_{n \times 1}\), based on the Kolmogorov
-test. Let \(F_{.}(n)\) be the ECDF function, with two possible
-subscripts \(Y\) and \(\mathcal{N}\) representing the projected and
-randomly generated data, and \(n\) denoting the number of observation,
-the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
+indices. An example of a noisy index function for 1D projections
+compares the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly
+generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
+Kolmogorov test. Let \(F_{.}(n)\) be the ECDF function, with two
+possible subscripts \(Y\) and \(\mathcal{N}\) representing the projected
+and randomly generated data, and \(n\) denoting the number of
+observation, the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
-all three optimisers and as expected, none of the interpolated path is
+all three optimisers and as expected, none of the interpolated paths are
 smooth. There is barely any improvement made by PD, indicating its
-failure in optimising non-smooth index. While CRS and SA have managed to
-get close to the index value of the theoretical best, trace plot shows
-that it takes SA much longer to interpolate towards the final basis.
+failure in optimising non-smooth index functions. While CRS and SA have
+managed to get close to the index value of the theoretical best, the
+trace plot shows that it takes SA much longer to find the final basis.
 This long interpolation path is partially due to the fluctuation in the
@@
 but if the basis space is non-smooth, accepting inferior bases at first,
-as what SA has done, can lead to a more efficient search, in terms of
+as SA has done here, can lead to a more efficient search, in terms of
 the overall number of points evaluated.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/flip-sign-1}
 
@@
 
-This project contributes to the software development in two packages: a
-data collection object is implemented in \CRANpkg{tourr} \citep{tourr},
-while the visual diagnostics of the optimisers is implemented in
-\CRANpkg{ferrn} \citep{ferrn}. Below list functions in the
-\CRANpkg{ferrn} \citep{ferrn} package:
+The implementation of this projection has been divided into two
+packages: the data collection object is implemented in the existing CRAN
+package \CRANpkg{tourr} \citep{tourr} while the optimiser diagnostics
+have been implemented in a new package, \pkg{ferrn}. When a guided tour
+is run, the users can choose to collect the data by binding
+\texttt{animate\_*()} to a variable. Once the data object has been
+obtained, the \pkg{ferrn} package provides diagnostic plots shown in
+Section \ref{vis-diag}. The main functions of the \pkg{ferrn} package
+functionality is listed below.
 

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.tex

@@
 guide improvements to the existing algorithms. Finally, Section
-\ref{implementation} describes the R package: \pkg{ferrn}, that
-implements the visual diagnostics.
+\ref{implementation} describes the R package: \CRANpkg{ferrn}
+\citep{ferrn}, that implements the visual diagnostics.
 
@@
 three other cases. This indicates that SA, with a reasonable large
-search size, is able to overcome the local optimum and optimise close
+search window, is able to overcome the local optimum and optimise close
 towards the global optimum.

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.Rmd
Final few changes.

@@
   - \code{get\_start()} extracts the starting observation
-  - \code{get\_theo()} extracts the theoretical best observations, if presented
+  - \code{get\_theo()} extracts the theoretical best observations, if given
 
@@
 
-- \code{add\_*()} create ggprotos for different components for the PCA plot
-  - \code{add\_anchor()} is a wrapper for plotting anchor bases
-  - \code{add\_anno()} is a wrapper for annotating the symmetry of start bases
-  - \code{add\_dir\_search()} is a wrapper for plotting the directional search bases with magnified distance
-  - \code{add\_end()} is a wrapper for plotting end bases
-  - \code{add\_interp()} is a wrapper for plotting the interpolation path
-  - \code{add\_interp\_last()} is a wrapper for plotting the last interpolation bases for comparing with target bases when interruption is carried
-  - \code{add\_interrupt()} is a wrapper for linking the last interpolation bases with target ones when interruption is carried
-  - \code{add\_search()} is a wrapper for plotting search bases
-  - \code{add\_space()} is a wrapper for plotting the circular space 
-  - \code{add\_start()} is a wrapper for plotting start bases
-  - \code{add\_theo()} is a wrapper for plotting theoretical best bases, if applicable
+- \code{add\_*()} provides wrapper functions to create ggprotos for different components for the PCA plot
+  - \code{add\_anchor()} for plotting anchor bases
+  - \code{add\_anno()} for annotating the symmetry of start bases
+  - \code{add\_dir\_search()} for plotting the directional search bases with magnified distance
+  - \code{add\_end()} for plotting end bases
+  - \code{add\_interp()} for plotting the interpolation path
+  - \code{add\_interp\_last()} for plotting the last interpolation bases for comparing with target bases when interruption is used
+  - \code{add\_interrupt()} for linking the last interpolation bases with target ones when interruption is used
+  - \code{add\_search()} for plotting search bases
+  - \code{add\_space()} for plotting the circular space 
+  - \code{add\_start()} for plotting start bases
+  - \code{add\_theo()} for plotting theoretical best bases, if applicable
 
@@
 
-This paper has illustrated setting up a data object that can be used for diagnosing a complex optimisation procedure.
+This paper has set up a data object that can be used for diagnosing a complex optimisation procedure.
 The ideas were illustrated using the optimisers available for projection pursuit guided tour.
 Here the constraint is the orthonormality condition of the projection bases.
-The approach used here could be broadly applied to understand other constrained optimisers.
+The approach used broadly applies to other constrained optimisers.
 
 <!-- something about the diagnostic plots  -->
-Four diagnostic plots have been introduced to investigate the progression and the projection space of an optimiser.
-The implementation of these visualisations is designed to be easy-to-use with each plot can be produced with a simple supply of the data object.
-More advanced users may decide to modify on top of the basic plots or even build their own.
+Four diagnostic plots have been introduced to investigate the progression of the optimiser and the coverage in search space.
+The implementation of these visualisations is designed to be easy-to-use, each plot can be produced by supplying the data object to the corresponding function.
+More advanced users may decide to modify these basic graphs using the provided wrapper functions, or even build their own.
 

### 2021-03-23 paper/zhang-cook-laa-langrene-menendez.tex
Final few changes.

@@
     \code{get\_theo()} extracts the theoretical best observations, if
-    presented
+    given
   \end{itemize}
@@
 \item
-  \code{add\_*()} create ggprotos for different components for the PCA
-  plot
+  \code{add\_*()} provides wrapper functions to create ggprotos for
+  different components for the PCA plot
 
@@
   \item
-    \code{add\_anchor()} is a wrapper for plotting anchor bases
-  \item
-    \code{add\_anno()} is a wrapper for annotating the symmetry of start
-    bases
-  \item
-    \code{add\_dir\_search()} is a wrapper for plotting the directional
-    search bases with magnified distance
-  \item
-    \code{add\_end()} is a wrapper for plotting end bases
-  \item
-    \code{add\_interp()} is a wrapper for plotting the interpolation
-    path
-  \item
-    \code{add\_interp\_last()} is a wrapper for plotting the last
-    interpolation bases for comparing with target bases when
-    interruption is carried
-  \item
-    \code{add\_interrupt()} is a wrapper for linking the last
-    interpolation bases with target ones when interruption is carried
-  \item
-    \code{add\_search()} is a wrapper for plotting search bases
-  \item
-    \code{add\_space()} is a wrapper for plotting the circular space
-  \item
-    \code{add\_start()} is a wrapper for plotting start bases
-  \item
-    \code{add\_theo()} is a wrapper for plotting theoretical best bases,
-    if applicable
+    \code{add\_anchor()} for plotting anchor bases
+  \item
+    \code{add\_anno()} for annotating the symmetry of start bases
+  \item
+    \code{add\_dir\_search()} for plotting the directional search bases
+    with magnified distance
+  \item
+    \code{add\_end()} for plotting end bases
+  \item
+    \code{add\_interp()} for plotting the interpolation path
+  \item
+    \code{add\_interp\_last()} for plotting the last interpolation bases
+    for comparing with target bases when interruption is used
+  \item
+    \code{add\_interrupt()} for linking the last interpolation bases
+    with target ones when interruption is used
+  \item
+    \code{add\_search()} for plotting search bases
+  \item
+    \code{add\_space()} for plotting the circular space
+  \item
+    \code{add\_start()} for plotting start bases
+  \item
+    \code{add\_theo()} for plotting theoretical best bases, if
+    applicable
   \end{itemize}
@@
 
-This paper has illustrated setting up a data object that can be used for
-diagnosing a complex optimisation procedure. The ideas were illustrated
-using the optimisers available for projection pursuit guided tour. Here
-the constraint is the orthonormality condition of the projection bases.
-The approach used here could be broadly applied to understand other
-constrained optimisers.
+This paper has set up a data object that can be used for diagnosing a
+complex optimisation procedure. The ideas were illustrated using the
+optimisers available for projection pursuit guided tour. Here the
+constraint is the orthonormality condition of the projection bases. The
+approach used broadly applies to other constrained optimisers.
 
 Four diagnostic plots have been introduced to investigate the
-progression and the projection space of an optimiser. The implementation
-of these visualisations is designed to be easy-to-use with each plot can
-be produced with a simple supply of the data object. More advanced users
-may decide to modify on top of the basic plots or even build their own.
+progression of the optimiser and the coverage in search space. The
+implementation of these visualisations is designed to be easy-to-use,
+each plot can be produced by supplying the data object to the
+corresponding function. More advanced users may decide to modify these
+basic graphs using the provided wrapper functions, or even build their
+own.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
minor tweaks

@@
 The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some structures of the data other than a normal distribution, e.g., bi-modality, clustering and outliers.
+A view is interesting if it can show some structures of the data other than a normal distribution, e.g., bi-modality, clustering, and outliers.
 
@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-An interesting projection shows structures that are non-normal and this has been argued for two reasons: Theoretical proofs from @diaconis1984asymptotics show that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
+An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
 Index functions proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
@@
 A starting point of diagnosing an optimiser is to understand how many searches it has conducted, i.e.
-we to summarise how the index is increasing over iterations and how many tries need to be sampled at each iteration.
+we want to summarise how the index is increasing over iterations and how many tries need to be sampled at each iteration.
 This is achieved using the search plot function: a boxplot shows the distribution of index values for each try, where the selected basis (corresponding to the highest index value) is always shown as a point.
@@
 principal component analysis is applied to first project all the bases onto a 2D space.
-In a projection pursuit guided tour optimisation, there are various types of bases involved: 1) The starting basis; 2) The search bases that the optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally 5) the end basis.
+In a projection pursuit guided tour optimisation, there are various types of bases involved: 1) The starting basis; 2) The search bases that the optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally, 5) the end basis.
 The importance of these bases differs and the most important ones are the starting, interpolating and end bases.
@@
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to orientation of the basis and so is the index value.
+This happens because the projection is invariant to the orientation of the basis and so is the index value.
 However, this creates difficulties for comparing the optimisers since the end bases will be symmetric with respect to the origin.
@@
 For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
-The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality.
+The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition.
 Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots.
@@
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
-The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at Time 61 has a higher index value than the target in the third interpolation.
-Rather than starting the next iteration from the target basis on Time 65, CRS starts the next iteration at Time 61 on the right panel and reaches a better final basis.
+The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at time 61 has a higher index value than the target in the third interpolation.
+Rather than starting the next iteration from the target basis on time 65, CRS starts the next iteration at time 61 on the right panel and reaches a better final basis.
 
@@
 While CRS and SA have managed to get close to the index value of the theoretical best, the trace plot shows that it takes SA much longer to find the final basis.
-This long interpolation path is partially due to the fluctuation in the early iterations, where the SA tends to generously accept inferior bases before concentrating near the optimum.
+This long interpolation path is partially due to the fluctuation in the early iterations, where SA tends to generously accept inferior bases before concentrating near the optimum.
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
@@
 <!-- something about the diagnostic plots  -->
-Four diagnostic plots have been introduced to investigate the progression of the optimiser and the coverage in search space.
+Four diagnostic plots have been introduced to investigate the progression of the optimiser and its coverage in search space.
 The implementation of these visualisations is designed to be easy-to-use, each plot can be produced by supplying the data object to the corresponding function.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
minor tweaks

@@
 some structures of the data other than a normal distribution, e.g.,
-bi-modality, clustering and outliers.
+bi-modality, clustering, and outliers.
 
@@
 of a given projection. An interesting projection shows structures that
-are non-normal and this has been argued for two reasons: Theoretical
-proofs from \citet{diaconis1984asymptotics} show that projections tend
-to be normal, as \(n\) and \(p\) approach infinity under certain
-conditions. Index functions proposed in the litearture include early
-indices that can be categorised by measuring the \(L^2\) distance
-between the projection and a normal distribution: Legendre index
+are non-normal since theoretical proofs from
+\citet{diaconis1984asymptotics} have shown that projections tend to be
+normal, as \(n\) and \(p\) approach infinity under certain conditions.
+Index functions proposed in the litearture include early indices that
+can be categorised by measuring the \(L^2\) distance between the
+projection and a normal distribution: Legendre index
 \citep{friedman1974projection}, Hermite index
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/ursulalaa/papers/paper-tour-vis/img/tour-path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour-path}
 
@@
 A starting point of diagnosing an optimiser is to understand how many
-searches it has conducted, i.e.~we to summarise how the index is
+searches it has conducted, i.e.~we want to summarise how the index is
 increasing over iterations and how many tries need to be sampled at each
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
 
@@
 highest index value in each iteration; 4) The interpolating bases on the
-interpolation path; and finally 5) the end basis. The importance of
+interpolation path; and finally, 5) the end basis. The importance of
 these bases differs and the most important ones are the starting,
@@
 can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to orientation of the
-basis and so is the index value. However, this creates difficulties for
-comparing the optimisers since the end bases will be symmetric with
+This happens because the projection is invariant to the orientation of
+the basis and so is the index value. However, this creates difficulties
+for comparing the optimisers since the end bases will be symmetric with
 respect to the origin. A sign flipping step is conducted to flip the
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 to be perpendicular to each other and this can be linked back to the
-orthogonality. Two paths from CRS and PD are plotted on top of the torus
-and coloured in green and brown respectively to match the previous
-plots. The final basis found by PD and CRS are shown in a larger shape
-and printed below respectively:
+orthogonality condition. Two paths from CRS and PD are plotted on top of
+the torus and coloured in green and brown respectively to match the
+previous plots. The final basis found by PD and CRS are shown in a
+larger shape and printed below respectively:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
 
@@
 interruption and while the first two interpolations are identical, the
-basis at Time 61 has a higher index value than the target in the third
+basis at time 61 has a higher index value than the target in the third
 interpolation. Rather than starting the next iteration from the target
-basis on Time 65, CRS starts the next iteration at Time 61 on the right
+basis on time 65, CRS starts the next iteration at time 61 on the right
 panel and reaches a better final basis.
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 This long interpolation path is partially due to the fluctuation in the
-early iterations, where the SA tends to generously accept inferior bases
+early iterations, where SA tends to generously accept inferior bases
 before concentrating near the optimum. The PCA plot shows the
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
 
@@
 Four diagnostic plots have been introduced to investigate the
-progression of the optimiser and the coverage in search space. The
+progression of the optimiser and its coverage in search space. The
 implementation of these visualisations is designed to be easy-to-use,

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
changes according to ursula's ntoes

@@
 
-The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is defined via an objective function.
+<!-- The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is defined via an objective function.
 Broadly speaking, optimization can be unconstrained or constrained [@kelley1999iterative].
-The unconstrained problem can be formulated as a minimization (or maximization)  problem such as
-$\min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with certain properties defined in an $L^p$ space.
+The unconstrained problem can be formulated as a minimization (or maximization)  problem such as  -->
+<!-- $\min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with certain properties defined in an $L^p$ space.
 In this case, solutions rely on gradient descent or ascent methods.
-In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as
-$\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as multipliers and convex optimization methods including linear and quadratic programming can be used.
+In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as -->
+<!-- $\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as multipliers and convex optimization methods including linear and quadratic programming can be used. -->
 
@@
 ```{r toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index.
-Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis."}
+Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
+In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and boxplot is used to summarise the distribution of the index values.
+There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey."}
 ```
@@
 All the variables have been scaled to have an overall unit variance before projection pursuit.
+Holes index [@cook2008grand], used for detecting binormality of the variables, is used throughout the examples unless otherwise specified.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
changes according to ursula's ntoes

@@
 of optimisers in practice.
-
-The goal of continuous optimisation is to find the best solution within
-the space of all feasible solutions where typically the best solution is
-defined via an objective function. Broadly speaking, optimization can be
-unconstrained or constrained \citep{kelley1999iterative}. The
-unconstrained problem can be formulated as a minimization (or
-maximization) problem such as \(\min_{x} f(x)\) where
-\(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with
-certain properties defined in an \(L^p\) space. In this case, solutions
-rely on gradient descent or ascent methods. In the constrained
-optimization problem additional restrictions are introduced via a set of
-functions that can be convex or non-convex:
-\(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and
-hence the problem can be written as \(\min_{x} f(x)\) \emph{subject to}
-\(g_i(x) \leq 0\). Here methods such as multipliers and convex
-optimization methods including linear and quadratic programming can be
-used.
 
@@
 \caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index.
-Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.}\label{fig:toy-search}
+Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
+In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and boxplot is used to summarise the distribution of the index values.
+There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.}\label{fig:toy-search}
 \end{figure}
@@
 mixtures with varied weights and locations. All the variables have been
-scaled to have an overall unit variance before projection pursuit.
+scaled to have an overall unit variance before projection pursuit. Holes
+index \citep{cook2008grand}, used for detecting binormality of the
+variables, is used throughout the examples unless otherwise specified.
 

### 2021-03-25 scripts/kol_result.R

@@ -9,7 +9,7 @@ clean_pca <- clean %>%
                    raw_data = boa6) %>%
   mutate(seed_sim = paste0(seed, sim),
          alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
-         optimiser = ifelse(str_detect(sim, "random"), "SAJO", "SA"),
+         optimiser = ifelse(str_detect(sim, "random"), "SA", "CRS"),
          var_found = as.factor(ifelse(var_found == "V7", "global", "local")),
          var_found = fct_relevel(var_found, "global", after = 1)) %>%
   compute_pca(group = seed_sim) %>%

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
some changes based on patricia's comment s

@@
 abstract: |
-  Guided tour searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
+  Guided tour [@cook1995grand] searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
 The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers.
-In this paper, we implement a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics based on the data object collected.
-These diagnostics and this workflow can be applied to a broad class of optimisers, to assess their performance.
+In this paper, we propose a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics for examining the performance of the optimisation procedure.
+These diagnostics and this workflow can be applied to a broad class of optimisers to assess their performance.
 An R package, \CRANpkg{ferrn} [@ferrn], has been created to implement the diagnostics.
@@
 A data structure is defined to capture information during the optimisation, and different diagnostic plots are designed with the data collected.
-Section \ref{application} shows applications of how these diagnostic plots can be used to discover interesting aspects of different optimisers and guide improvements to the existing algorithms.
+Section \ref{application} shows applications of how these diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms.
 Finally, Section \ref{implementation} describes the R package: \CRANpkg{ferrn} [@ferrn], that implements the visual diagnostics.
@@
 
-The optimisation problem of our interest is on constrained optimization [@bertsekas2014constrained] as defined in the introduction section, and assuming it is not possible to find a solution to the problem in the way of a closed-form.
+The optimisation problem considered in this paper is on constrained optimization [@bertsekas2014constrained] and assuming it is not possible to find a solution to the problem in the way of a closed-form.
 That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathcal{A}$ space.
 
-A large class of methods uses the gradient information of the objective function, with the most notable one being the gradient ascent (descent) method.
-Although gradient optimization methods are popular, they rely on the availability of the objective function derivatives and on the limited complexity of the constraints.
+Gradient-based methods are commonly used to optimise an objective function, with the most notable one being the gradient ascent (descent) method.
+Although these methods are popular, they rely on the availability of the objective function derivatives.
 As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the render of tour animation.
@@
 
-Random search methods [@zabinsky2013stochastic] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
+Random search methods [@zabinsky2013stochastic; @Romeijn2009] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
 The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
@@
 Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where worse candidates still have a chance to be accepted.
-The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @belisle1992convergence; @granville1994simulated] and the global optimum can be attained under mild conditions.
+The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @belisle1992convergence; @granville1994simulated] and the global optimum can be attained under mild regularity conditions.
 
@@
 
-Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) in exploratory data analysis.
+Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high dimensional space.
 Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
 clustering, outliers and skewness) in multivariate data via low-dimensional projections.
-Guided tour is one variation of a broader class of data visualisation methods, tour, which displays high-dimensional data through a series of animated projections.
+Guided tour [@cook2008grand; @cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.
 
@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-Index functions proposed in the litearture include early indices that can be categorised by measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+There has been plentiful index functions proposed in the literature and here list a few: early indices that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 
@@
 
-A data structure for diagnosing optimisers in projection pursuit guided tour is first defined.
-With this data structure, four types of diagnostic plots are presented.
-In the section, examples will be given to show how these plot can be used to reveal issues in the projection pursuit optimisation.
+A data structure for diagnosing optimisers in projection pursuit guided tour is first defined. With this data structure, four types of diagnostic plots are presented.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
some changes based on patricia's comment s

@@
 \abstract{%
-Guided tour searches for interesting low-dimensional views of
-high-dimensional data via optimising a projection pursuit index
-function. The first paper of projection pursuit by
-\citet{friedman1974projection} stated that ``the technique used for
+Guided tour \citep{cook1995grand} searches for interesting
+low-dimensional views of high-dimensional data via optimising a
+projection pursuit index function. The first paper of projection pursuit
+by \citet{friedman1974projection} stated that ``the technique used for
 maximising the projection index strongly influences both the statistical
@@
 been done in proposing indices in the literature, less has been done on
-evaluating the performance of the optimisers. In this paper, we
-implement a data collection object in the optimisation of projection
-pursuit guided tour and introduce visual diagnostics based on the data
-object collected. These diagnostics and this workflow can be applied to
-a broad class of optimisers, to assess their performance. An R package,
-\CRANpkg{ferrn} \citep{ferrn}, has been created to implement the
-diagnostics.
+evaluating the performance of the optimisers. In this paper, we propose
+a data collection object in the optimisation of projection pursuit
+guided tour and introduce visual diagnostics for examining the
+performance of the optimisation procedure. These diagnostics and this
+workflow can be applied to a broad class of optimisers to assess their
+performance. An R package, \CRANpkg{ferrn} \citep{ferrn}, has been
+created to implement the diagnostics.
 
@@
 \ref{application} shows applications of how these diagnostic plots can
-be used to discover interesting aspects of different optimisers and
-guide improvements to the existing algorithms. Finally, Section
+be used to examine the performance of different optimisers and guide
+improvements to the existing algorithms. Finally, Section
 \ref{implementation} describes the R package: \CRANpkg{ferrn}
@@
 
-The optimisation problem of our interest is on constrained optimization
-\citep{bertsekas2014constrained} as defined in the introduction section,
-and assuming it is not possible to find a solution to the problem in the
-way of a closed-form. That is, the problem consists of finding the
-minimum or maximum of a function \(f \in L^p\) in the constrained
-\(\mathcal{A}\) space.
-
-A large class of methods uses the gradient information of the objective
+The optimisation problem considered in this paper is on constrained
+optimization \citep{bertsekas2014constrained} and assuming it is not
+possible to find a solution to the problem in the way of a closed-form.
+That is, the problem consists of finding the minimum or maximum of a
+function \(f \in L^p\) in the constrained \(\mathcal{A}\) space.
+
+Gradient-based methods are commonly used to optimise an objective
 function, with the most notable one being the gradient ascent (descent)
-method. Although gradient optimization methods are popular, they rely on
-the availability of the objective function derivatives and on the
-limited complexity of the constraints. As will be shown in the next
-section, the independent variables in our optimisation problem are the
-entries of a projection matrix and the computational time required to
-perform differentiation on a matrix could impede the render of tour
-animation. Hence, gradient-based methods are not the focus of this paper
-and considerations will be given to derivative-free methods.
+method. Although these methods are popular, they rely on the
+availability of the objective function derivatives. As will be shown in
+the next section, the independent variables in our optimisation problem
+are the entries of a projection matrix and the computational time
+required to perform differentiation on a matrix could impede the render
+of tour animation. Hence, gradient-based methods are not the focus of
+this paper and considerations will be given to derivative-free methods.
 
@@
 
-Random search methods \citep{zabinsky2013stochastic} have a random
-sampling component as part of their algorithm and have been shown to
-have the ability to optimise non-convex and non-smooth functions. The
+Random search methods \citep{zabinsky2013stochastic, Romeijn2009} have a
+random sampling component as part of their algorithm and have been shown
+to have the ability to optimise non-convex and non-smooth functions. The
 initial random search algorithm, pure random search
@@
 \citep{mitra1986convergence, belisle1992convergence, granville1994simulated}
-and the global optimum can be attained under mild conditions.
+and the global optimum can be attained under mild regularity conditions.
 
@@
 Projection pursuit guided tour combines two different methods
-(projection pursuit and guided tour) in exploratory data analysis.
-Projection pursuit, coined by \citet{friedman1974projection}, detects
-interesting structures (e.g.~clustering, outliers and skewness) in
-multivariate data via low-dimensional projections. Guided tour is one
-variation of a broader class of data visualisation methods, tour, which
+(projection pursuit and guided tour) to explore interesting features in
+a high dimensional space. Projection pursuit, coined by
+\citet{friedman1974projection}, detects interesting structures
+(e.g.~clustering, outliers and skewness) in multivariate data via
+low-dimensional projections. Guided tour
+\citep{cook2008grand, cook1995grand} is one variation of a broader class
+of data visualisation methods, tour \citep{buja2005computational}, which
 displays high-dimensional data through a series of animated projections.
@@
 normal, as \(n\) and \(p\) approach infinity under certain conditions.
-Index functions proposed in the litearture include early indices that
-can be categorised by measuring the \(L^2\) distance between the
-projection and a normal distribution: Legendre index
-\citep{friedman1974projection}, Hermite index
+There has been plentiful index functions proposed in the literature and
+here list a few: early indices that can be categorised as measuring the
+\(L^2\) distance between the projection and a normal distribution:
+Legendre index \citep{friedman1974projection}, Hermite index
 \citep{hall1989polynomial}, and natural Hermite index
@@
 tour is first defined. With this data structure, four types of
-diagnostic plots are presented. In the section, examples will be given
-to show how these plot can be used to reveal issues in the projection
-pursuit optimisation.
+diagnostic plots are presented.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
finish editing on patricia's comments; remove references no longer used in .bib file

@@
 
-Derivative-free methods, which do not rely on the knowledge of the gradient, are more generally applicable.
+Derivative-free methods [@rios2013derivative; @conn2009introduction], which do not rely on the knowledge of the gradient, are more generally applicable.
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
@@
 
-Random search methods [@zabinsky2013stochastic; @Romeijn2009] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
+Random search methods [@zabinsky2013stochastic; @Romeijn2009; @andradottir2015review] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
 The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
@@
 
-Three main pieces of information are recorded for the projection pursuit optimisers: 1) projection bases $\mathbf{A}$, 2) index values $I$, and 3) state $S$.
+Three main pieces of information are recorded during the projection pursuit optimisation: 1) projection bases $\mathbf{A}$, 2) index values $I$, and 3) state $S$.
 For CRS and SA, possible states include `random_search`,  `new_basis`, and `interpolation`.
@@
 A starting point of diagnosing an optimiser is to understand how many searches it has conducted, i.e.
-we want to summarise how the index is increasing over iterations and how many tries need to be sampled at each iteration.
-This is achieved using the search plot function: a boxplot shows the distribution of index values for each try, where the selected basis (corresponding to the highest index value) is always shown as a point.
-In case of few tries at a given iteration showing the data points directly may be preferred over the boxplot, this is controlled via the `cutoff` argument.
+we want to summarise how the index is increasing over iterations and how many basis need to be sampled at each iteration.
+This is achieved using the function `explore_trace_search()`: a boxplot shows the distribution of index values for each try, where the accepted basis (corresponding to the highest index value) is always shown as a point.
+In the case of few tries at a given iteration, showing the data points directly may be preferred over the boxplot, this is controlled via the `cutoff` argument.
 Additional annotations are added to facilitate better reading of the plot and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
@@
 
-Another interesting diagnostic is to examine how the index value improves between interpolating bases since the projection on these bases is shown in the tour animation.
+Another interesting feature to examine is the changes on the index value between interpolating bases since the projection on these bases is shown in the tour animation.
 Trace plots are created by plotting the index value against time.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
finish editing on patricia's comments; remove references no longer used in .bib file

@@
 
-Derivative-free methods, which do not rely on the knowledge of the
-gradient, are more generally applicable. Derivative-free methods have
-been developed over the years, where the emphasis is on finding, in most
-cases, a near optimal solution. Here we consider three derivative-free
-methods, two of which are random search methods: creeping random search
-and simulated annealing, and the other one is pseudo derivative search.
-
-Random search methods \citep{zabinsky2013stochastic, Romeijn2009} have a
-random sampling component as part of their algorithm and have been shown
-to have the ability to optimise non-convex and non-smooth functions. The
-initial random search algorithm, pure random search
+Derivative-free methods
+\citep{rios2013derivative, conn2009introduction}, which do not rely on
+the knowledge of the gradient, are more generally applicable.
+Derivative-free methods have been developed over the years, where the
+emphasis is on finding, in most cases, a near optimal solution. Here we
+consider three derivative-free methods, two of which are random search
+methods: creeping random search and simulated annealing, and the other
+one is pseudo derivative search.
+
+Random search methods
+\citep{zabinsky2013stochastic, Romeijn2009, andradottir2015review} have
+a random sampling component as part of their algorithm and have been
+shown to have the ability to optimise non-convex and non-smooth
+functions. The initial random search algorithm, pure random search
 \citep{Brooks1958-zk}, draws candidate points from the entire space
@@
 
-Three main pieces of information are recorded for the projection pursuit
-optimisers: 1) projection bases \(\mathbf{A}\), 2) index values \(I\),
-and 3) state \(S\). For CRS and SA, possible states include
-\texttt{random\_search}, \texttt{new\_basis}, and
+Three main pieces of information are recorded during the projection
+pursuit optimisation: 1) projection bases \(\mathbf{A}\), 2) index
+values \(I\), and 3) state \(S\). For CRS and SA, possible states
+include \texttt{random\_search}, \texttt{new\_basis}, and
 \texttt{interpolation}. Pseudo derivative (PD) has a wider variety of
@@
 searches it has conducted, i.e.~we want to summarise how the index is
-increasing over iterations and how many tries need to be sampled at each
-iteration. This is achieved using the search plot function: a boxplot
-shows the distribution of index values for each try, where the selected
-basis (corresponding to the highest index value) is always shown as a
-point. In case of few tries at a given iteration showing the data points
-directly may be preferred over the boxplot, this is controlled via the
-\texttt{cutoff} argument. Additional annotations are added to facilitate
-better reading of the plot and these include 1) the number of points
-searched in each iteration can be added as text label at the bottom of
-each iteration; 2) the anchor bases to interpolate are connected and
-highlighted in a larger size; and 3) the colour of the last iteration is
-in a grey scale to indicate no better basis found in this iteration.
+increasing over iterations and how many basis need to be sampled at each
+iteration. This is achieved using the function
+\texttt{explore\_trace\_search()}: a boxplot shows the distribution of
+index values for each try, where the accepted basis (corresponding to
+the highest index value) is always shown as a point. In the case of few
+tries at a given iteration, showing the data points directly may be
+preferred over the boxplot, this is controlled via the \texttt{cutoff}
+argument. Additional annotations are added to facilitate better reading
+of the plot and these include 1) the number of points searched in each
+iteration can be added as text label at the bottom of each iteration; 2)
+the anchor bases to interpolate are connected and highlighted in a
+larger size; and 3) the colour of the last iteration is in a grey scale
+to indicate no better basis found in this iteration.
 
@@
 
-Another interesting diagnostic is to examine how the index value
-improves between interpolating bases since the projection on these bases
-is shown in the tour animation. Trace plots are created by plotting the
-index value against time. Figure \ref{fig:toy-interp} presents the trace
-plot of the same optimisations as Figure \ref{fig:toy-search} and one
-can observe that the trace is smooth in both cases. It may seem bizarre
-at the first sight that the interpolation sometimes passes bases with
+Another interesting feature to examine is the changes on the index value
+between interpolating bases since the projection on these bases is shown
+in the tour animation. Trace plots are created by plotting the index
+value against time. Figure \ref{fig:toy-interp} presents the trace plot
+of the same optimisations as Figure \ref{fig:toy-search} and one can
+observe that the trace is smooth in both cases. It may seem bizarre at
+the first sight that the interpolation sometimes passes bases with
 higher index value before it decreases to a lower target. This happens

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
abstract changed and use of CRANpkg corrected

@@
 abstract: |
-  Guided tour [@cook1995grand] searches for interesting low-dimensional views of high-dimensional data via optimising a projection pursuit index function.
-The first paper of projection pursuit by @friedman1974projection stated that "the technique used for maximising the projection index strongly influences both the statistical and the computational aspects of the procedure." While much work has been done in proposing  indices in the literature, less has been done on evaluating the performance of the optimisers.
-In this paper, we propose a data collection object in the optimisation of projection pursuit guided tour and introduce visual diagnostics for examining the performance of the optimisation procedure.
-These diagnostics and this workflow can be applied to a broad class of optimisers to assess their performance.
-An R package, \CRANpkg{ferrn} [@ferrn], has been created to implement the diagnostics.
+  A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimization path.
+Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
+While much work has been done in developing new indexes in the literature, less has been done on understanding the optimization.
+Index functions can be noisy, they might have multiple local maxima as well as an optimal maximum, and it is constrained to generate orthonormal projection frames, which complicates the optimization.
+In addition, projection pursuit is primarily used for exploratory data analysis, and finding the local maxima is also useful.
+The guided tour is especially useful for exploration, because it conducts geodesic interpolation connecting steps in the optimisation and shows how the projected data changes as a maxima is approached.
+This work provides new visual diagnostics for examining a choice of optimisation procedure, based on the provision of a new data object which collects information throughout the optimisation.
+It has helped to diagnose and fix several problems with projection pursuit guided tour.
+This work might be useful more broadly for diagnosing optimisers, and comparing their performance.
+The diagnostics are implemented in the R package, \pkg{ferrn}.
 
@@
 
-Three optimisers have been implemented in the \CRANpkg{tourr} [@tourr] package: Creeping random search (CRS), simulated annealing (SA), and pseudo derivative (PD).
+Three optimisers have been implemented in the \pkg{tourr} [@tourr] package: Creeping random search (CRS), simulated annealing (SA), and pseudo derivative (PD).
 Creeping random search (CRS) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
@@
 
-This project contributes to the software development in two packages: a data collection object is implemented in \CRANpkg{tourr} [@tourr], while the visual diagnostics of the optimisers is implemented in \CRANpkg{ferrn} [@ferrn].
-Below list functions in the \CRANpkg{ferrn} [@ferrn] package:
+This project contributes to the software development in two packages: a data collection object is implemented in \pkg{tourr} [@tourr], while the visual diagnostics of the optimisers is implemented in \pkg{ferrn} [@ferrn].
+Below list functions in the \pkg{ferrn} [@ferrn] package:
 
@@
 <!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \CRANpkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
+Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \pkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
 Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
abstract changed and use of CRANpkg corrected

@@
 \abstract{%
-Guided tour \citep{cook1995grand} searches for interesting
-low-dimensional views of high-dimensional data via optimising a
-projection pursuit index function. The first paper of projection pursuit
-by \citet{friedman1974projection} stated that ``the technique used for
-maximising the projection index strongly influences both the statistical
-and the computational aspects of the procedure.'' While much work has
-been done in proposing indices in the literature, less has been done on
-evaluating the performance of the optimisers. In this paper, we propose
-a data collection object in the optimisation of projection pursuit
-guided tour and introduce visual diagnostics for examining the
-performance of the optimisation procedure. These diagnostics and this
-workflow can be applied to a broad class of optimisers to assess their
-performance. An R package, \CRANpkg{ferrn} \citep{ferrn}, has been
-created to implement the diagnostics.
+A guided tour helps to visualise high-dimensional data by showing the
+low-dimensional projections along a projection pursuit optimization
+path. Projection pursuit is a generalisation of principal component
+analysis, in the sense that different indexes are used to define the
+interestingness of the projected data. While much work has been done in
+developing new indexes in the literature, less has been done on
+understanding the optimization. Index functions can be noisy, they might
+have multiple local maxima as well as an optimal maximum, and it is
+constrained to generate orthonormal projection frames, which complicates
+the optimization. In addition, projection pursuit is primarily used for
+exploratory data analysis, and finding the local maxima is also useful.
+The guided tour is especially useful for exploration, because it
+conducts geodesic interpolation connecting steps in the optimisation and
+shows how the projected data changes as a maxima is approached. This
+work provides new visual diagnostics for examining a choice of
+optimisation procedure, based on the provision of a new data object
+which collects information throughout the optimisation. It has helped to
+diagnose and fix several problems with projection pursuit guided tour.
+This work might be useful more broadly for diagnosing optimisers, and
+comparing their performance. The diagnostics are implemented in the R
+package, \pkg{ferrn}.
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour-path}
+{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/img/tour-path}
 
@@
 
-Three optimisers have been implemented in the \CRANpkg{tourr}
-\citep{tourr} package: Creeping random search (CRS), simulated annealing
-(SA), and pseudo derivative (PD). Creeping random search (CRS) is a
-random search optimiser that samples a candidate basis
-\(\mathbf{A}_{l}\) in the neighbourhood of the current basis
-\(\mathbf{A}_{\text{cur}}\) by
+Three optimisers have been implemented in the \pkg{tourr} \citep{tourr}
+package: Creeping random search (CRS), simulated annealing (SA), and
+pseudo derivative (PD). Creeping random search (CRS) is a random search
+optimiser that samples a candidate basis \(\mathbf{A}_{l}\) in the
+neighbourhood of the current basis \(\mathbf{A}_{\text{cur}}\) by
 \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-pca-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-tour-vis/figs/flip-sign-1}
 
@@
 This project contributes to the software development in two packages: a
-data collection object is implemented in \CRANpkg{tourr} \citep{tourr},
+data collection object is implemented in \pkg{tourr} \citep{tourr},
 while the visual diagnostics of the optimisers is implemented in
-\CRANpkg{ferrn} \citep{ferrn}. Below list functions in the
-\CRANpkg{ferrn} \citep{ferrn} package:
+\pkg{ferrn} \citep{ferrn}. Below list functions in the \pkg{ferrn}
+\citep{ferrn} package:
 
@@
 packages: the collection of the data object is implemented in the
-existing \CRANpkg{tourr} \citep{tourr} package; while the manipulation
-and visualisation of the data object are implemented in the new
-\pkg{ferrn} package. Equipped with handy tools to diagnose the
-performance of optimisers, future work can extend the diagnostics to a
-wider range of index functions, i.e.~scagnostics, association, and
-information index \citep{laa2020using} and understand how the optimisers
-behave for index functions with different structures.
+existing \pkg{tourr} \citep{tourr} package; while the manipulation and
+visualisation of the data object are implemented in the new \pkg{ferrn}
+package. Equipped with handy tools to diagnose the performance of
+optimisers, future work can extend the diagnostics to a wider range of
+index functions, i.e.~scagnostics, association, and information index
+\citep{laa2020using} and understand how the optimisers behave for index
+functions with different structures.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
a missing action

--- 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
@@
 
-The construction of diagnostic plots adopts the core concept of grammar
-of graphics \citep{wickham2010layered} in ggplot2. In grammar of
-graphics, plots are not produced by calling the commands, named by the
-appearance of the plot, i.e., boxplot and histogram, but by the concept
-of stacked layers. Seeing plots as stacked layers gives analysts the
-freedom to compose plots with multiple elements on hand.
-
 \hypertarget{diagnostic-1-checking-how-hard-the-optimiser-is-working}{%

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
a missing action
introduction modified

@@
 
-The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand].
+The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis technique used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand].
 The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some structures of the data other than a normal distribution, e.g., bi-modality, clustering, and outliers.
+A view is interesting if it can show some structures of the data that depart from normalality, such as bimodality, clustering, or outliers.
 
@@
 Examples include the existence of multiple optima (local and global), the trade-off between computational burden and proximity to the optima, or dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient].
-The visualization tools and optimization methods presented in this paper can therefore easily be applied to other optimization problems.
+The visualization tools, optimization methods and conceptual framework presented in this paper can therefore be applied to other optimization problems.
 
 The remainder of the paper is organised as follows. 
-Section \ref{optim} provides an overview of optimisation methods, specifically line search methods. 
-Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and outlines three existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics.
-A data structure is defined to capture information during the optimisation, and different diagnostic plots are designed with the data collected.
-Section \ref{application} shows applications of how these diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms.
-Finally, Section \ref{implementation} describes the R package: \CRANpkg{ferrn} [@ferrn], that implements the visual diagnostics.
+The next section provides an overview of optimisation methods, specifically line search methods.
+A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
+The third section presents the new visual diagnostics, including the the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
+An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms applications is shown using simulated data.
+Finally, an explanation of the \ref{implementation} in the R package, \CRANpkg{ferrn} [@ferrn], is provided.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
introduction modified

--- 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
@@
 projection pursuit guided tour \citep{buja2005computational}, an
-exploratory data analysis tool used for detecting interesting structures
-in high-dimensional data through a set of lower-dimensional projections
-\citep{cook2008grand}. The goal of the optimisation is to identify the
-projection, represented by the projection matrix, that gives the most
-interesting low-dimensional view. A view is interesting if it can show
-some structures of the data other than a normal distribution, e.g.,
-bi-modality, clustering, and outliers.
+exploratory data analysis technique used for detecting interesting
+structures in high-dimensional data through a set of lower-dimensional
+projections \citep{cook2008grand}. The goal of the optimisation is to
+identify the projection, represented by the projection matrix, that
+gives the most interesting low-dimensional view. A view is interesting
+if it can show some structures of the data that depart from normalality,
+such as bimodality, clustering, or outliers.
 
@@
 dealing with noisy objective functions that might be non-smooth and
-non-differentiable \citep{jones1998efficient}. The visualization tools
-and optimization methods presented in this paper can therefore easily be
-applied to other optimization problems.
-
-The remainder of the paper is organised as follows. Section \ref{optim}
+non-differentiable \citep{jones1998efficient}. The visualization tools,
+optimization methods and conceptual framework presented in this paper
+can therefore be applied to other optimization problems.
+
+The remainder of the paper is organised as follows. The next section
 provides an overview of optimisation methods, specifically line search
-methods. Section \ref{tour} reviews projection pursuit guided tour,
-defines the optimisation problem and outlines three existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics. A data
-structure is defined to capture information during the optimisation, and
-different diagnostic plots are designed with the data collected. Section
-\ref{application} shows applications of how these diagnostic plots can
-be used to examine the performance of different optimisers and guide
-improvements to the existing algorithms. Finally, Section
-\ref{implementation} describes the R package: \CRANpkg{ferrn}
-\citep{ferrn}, that implements the visual diagnostics.
+methods. A review of the projection pursuit guided tour, an overview of
+the optimisation problem and outlines of three existing algorithms
+follows. The third section presents the new visual diagnostics,
+including the the design of a data structure to capture information
+during the optimisation, from which several diagnostic plots are
+created. An illustration of how the diagnostic plots can be used to
+examine the performance of different optimisers and guide improvements
+to the existing algorithms applications is shown using simulated data.
+Finally, an explanation of the \ref{implementation} in the R package,
+\CRANpkg{ferrn} \citep{ferrn}, is provided.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
introduction modified
conclusions

@@
                       fig.height = 3,
+                      fig.retina = 5,
                       out.width = "100%")
@@
 An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms applications is shown using simulated data.
-Finally, an explanation of the \ref{implementation} in the R package, \CRANpkg{ferrn} [@ferrn], is provided.
+Finally, an explanation of the implementation in the R package, \CRANpkg{ferrn} [@ferrn], is provided.
 
@@
 
-This paper has set up a data object that can be used for diagnosing a complex optimisation procedure.
-The ideas were illustrated using the optimisers available for projection pursuit guided tour.
+This paper has provided several visual diagnostics that can be used for diagnosing a complex optimisation procedure.
+The methods were illustrated using the optimisers available for projection pursuit guided tour.
 Here the constraint is the orthonormality condition of the projection bases.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
introduction modified
conclusions

@@
 to the existing algorithms applications is shown using simulated data.
-Finally, an explanation of the \ref{implementation} in the R package,
+Finally, an explanation of the implementation in the R package,
 \CRANpkg{ferrn} \citep{ferrn}, is provided.
@@
 
-This paper has set up a data object that can be used for diagnosing a
-complex optimisation procedure. The ideas were illustrated using the
-optimisers available for projection pursuit guided tour. Here the
-constraint is the orthonormality condition of the projection bases. The
-approach used broadly applies to other constrained optimisers.
+This paper has provided several visual diagnostics that can be used for
+diagnosing a complex optimisation procedure. The methods were
+illustrated using the optimisers available for projection pursuit guided
+tour. Here the constraint is the orthonormality condition of the
+projection bases. The approach used broadly applies to other constrained
+optimisers.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd

--- 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 
-The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2.
+<!-- The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2.
 In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers.
 Seeing plots as stacked layers gives analysts the freedom to compose plots with multiple elements on hand.
+-->
 
@@
 
-This paper has set up a data object that can be used for diagnosing a complex optimisation procedure.
-The ideas were illustrated using the optimisers available for projection pursuit guided tour.
-Here the constraint is the orthonormality condition of the projection bases.
-The approach used broadly applies to other constrained optimisers.
+This paper has provided several visual diagnostics that can be used for diagnosing a complex optimisation procedure.
+The methods were illustrated using the optimisers available for projection pursuit guided tour.
+The methods are implemented in the package \pkg{ferrn}, and changes were made in the \pkg{tourr} package to generate the necessary data.
+This work complements with the work of @laa2020using and will be useful for researchers when designing new projection pursuit indexes.
 
-<!-- something about the diagnostic plots  -->
-Four diagnostic plots have been introduced to investigate the progression of the optimiser and its coverage in search space.
-The implementation of these visualisations is designed to be easy-to-use, each plot can be produced by supplying the data object to the corresponding function.
-More advanced users may decide to modify these basic graphs using the provided wrapper functions, or even build their own.
-
-<!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \pkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
-Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.
-scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.
+This work might be considered an effort to bring transparency into algorithms.
+Although little attention is paid by algorithm developers to providing ways to output information during intermediate steps, this is an important component for enabling others to understand and diagnose the performance.
+Algorithms are an essential component of artificial intelligence that is used to make daily life easier.
+Interpretability of algorithms is important to guard against aspects like bias and inappropriate use.
+The data object underlying the visual diagnostics here is an example of what might be useful in algorithm development generally.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex

--- 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
@@
 
-This paper has set up a data object that can be used for diagnosing a
-complex optimisation procedure. The ideas were illustrated using the
-optimisers available for projection pursuit guided tour. Here the
-constraint is the orthonormality condition of the projection bases. The
-approach used broadly applies to other constrained optimisers.
-
-Four diagnostic plots have been introduced to investigate the
-progression of the optimiser and its coverage in search space. The
-implementation of these visualisations is designed to be easy-to-use,
-each plot can be produced by supplying the data object to the
-corresponding function. More advanced users may decide to modify these
-basic graphs using the provided wrapper functions, or even build their
-own.
-
-Most of the work in this project has been translated into code in two
-packages: the collection of the data object is implemented in the
-existing \CRANpkg{tourr} \citep{tourr} package; while the manipulation
-and visualisation of the data object are implemented in the new
-\pkg{ferrn} package. Equipped with handy tools to diagnose the
-performance of optimisers, future work can extend the diagnostics to a
-wider range of index functions, i.e.~scagnostics, association, and
-information index \citep{laa2020using} and understand how the optimisers
-behave for index functions with different structures.
+This paper has provided several visual diagnostics that can be used for
+diagnosing a complex optimisation procedure. The methods were
+illustrated using the optimisers available for projection pursuit guided
+tour. The methods are implemented in the package \pkg{ferrn}, and
+changes were made in the \pkg{tourr} package to generate the necessary
+data. This work complements with the work of \citet{laa2020using} and
+will be useful for researchers when designing new projection pursuit
+indexes.
+
+This work might be considered an effort to bring transparency into
+algorithms. Although little attention is paid by algorithm developers to
+providing ways to output information during intermediate steps, this is
+an important component for enabling others to understand and diagnose
+the performance. Algorithms are an essential component of artificial
+intelligence that is used to make daily life easier. Interpretability of
+algorithms is important to guard against aspects like bias and
+inappropriate use. The data object underlying the visual diagnostics
+here is an example of what might be useful in algorithm development
+generally.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
clean up figures

--- 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
   here::here("scripts/diagnose.R")
-)
-
-knitr::read_chunk(
-  here::here("scripts/implementation.R")
 )
@@
 
-```{r toy-pca, fig.asp=1, fig.height=2.5, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, fig.height = 5, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 
-<!-- something about the diagnostic plots  -->
-Four diagnostic plots have been introduced to investigate the progression of the optimiser and its coverage in search space.
-The implementation of these visualisations is designed to be easy-to-use, each plot can be produced by supplying the data object to the corresponding function.
-More advanced users may decide to modify these basic graphs using the provided wrapper functions, or even build their own.
-
-<!-- summarise what we have done and what might be done in the future. -->
-Most of the work in this project has been translated into code in two packages: the collection of the data object is implemented in the existing \pkg{tourr} [@tourr] package; while the manipulation and visualisation of the data object are implemented in the new \pkg{ferrn} package.
-Equipped with handy tools to diagnose the performance of optimisers, future work can extend the diagnostics to a wider range of index functions, i.e.
-scagnostics, association, and information index [@laa2020using] and understand how the optimisers behave for index functions with different structures.
+This work might be considered an effort to bring transparency into algorithms.
+Although little attention is paid by algorithm developers to providing ways to output information during intermediate steps, this is an important component for enabling others to understand and diagnose the performance.
+Algorithms are an essential component of artificial intelligence that is used to make daily life easier.
+Interpretability of algorithms is important to guard against aspects like bias and inappropriate use.
+The data object underlying the visual diagnostics here is an example of what might be useful in algorithm development generally.
 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
clean up figures

@@
 dealing with noisy objective functions that might be non-smooth and
-non-differentiable \citep{jones1998efficient}. The visualization tools
-and optimization methods presented in this paper can therefore easily be
-applied to other optimization problems.
-
-The remainder of the paper is organised as follows. Section \ref{optim}
+non-differentiable \citep{jones1998efficient}. The visualization tools,
+optimization methods and conceptual framework presented in this paper
+can therefore be applied to other optimization problems.
+
+The remainder of the paper is organised as follows. The next section
 provides an overview of optimisation methods, specifically line search
-methods. Section \ref{tour} reviews projection pursuit guided tour,
-defines the optimisation problem and outlines three existing algorithms.
-Section \ref{vis-diag} presents the new visual diagnostics. A data
-structure is defined to capture information during the optimisation, and
-different diagnostic plots are designed with the data collected. Section
-\ref{application} shows applications of how these diagnostic plots can
-be used to examine the performance of different optimisers and guide
-improvements to the existing algorithms. Finally, Section
-\ref{implementation} describes the R package: \CRANpkg{ferrn}
-\citep{ferrn}, that implements the visual diagnostics.
+methods. A review of the projection pursuit guided tour, an overview of
+the optimisation problem and outlines of three existing algorithms
+follows. The third section presents the new visual diagnostics,
+including the the design of a data structure to capture information
+during the optimisation, from which several diagnostic plots are
+created. An illustration of how the diagnostic plots can be used to
+examine the performance of different optimisers and guide improvements
+to the existing algorithms applications is shown using simulated data.
+Finally, an explanation of the implementation in the R package,
+\CRANpkg{ferrn} \citep{ferrn}, is provided.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca}
 
@@
 illustrated using the optimisers available for projection pursuit guided
-tour. The methods are implemented in the package \pkg{ferrn}, and
-changes were made in the \pkg{tourr} package to generate the necessary
-data. This work complements with the work of \citet{laa2020using} and
-will be useful for researchers when designing new projection pursuit
-indexes.
+tour. Here the constraint is the orthonormality condition of the
+projection bases. The approach used broadly applies to other constrained
+optimisers.
 

### 2021-03-25 scripts/toy.R

@@ -33,12 +33,14 @@ p2 <- holes_2d_better_random %>%
 p1 | p2
 
 ## ----toy-pca
-bind_rows(holes_1d_geo, holes_1d_better) %>%
-  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-                   index = tourr::holes(), raw_data = boa5)  %>%
-  explore_space_pca(group = method, details = TRUE,
-                    interp_size = 1) +
-  scale_color_discrete_botanical()
+# bind_rows(holes_1d_geo, holes_1d_better) %>%
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                    index = tourr::holes(), raw_data = boa5)  %>%
+#   explore_space_pca(group = method, details = TRUE,
+#                     interp_size = 1) +
+#   scale_color_discrete_botanical()
+knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
+
 
 ## ----toy-pca-animated
 # dt <- dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
changes based on checklist WIP

@@
     email:  patricia.menendez@monash.edu 
-abstract: |
-  A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimization path.
+abstract: A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimization path.
 Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
@@
 The diagnostics are implemented in the R package, \pkg{ferrn}.
-
-  \noindent \textit{keywords}: graphics, multivariate, optimization
+keywords: graphics, multivariate, optimization
 preamble: >
@@
 
-Derivative-free methods [@rios2013derivative; @conn2009introduction], which do not rely on the knowledge of the gradient, are more generally applicable.
+Derivative-free methods [@conn2009introduction; @rios2013derivative], which do not rely on the knowledge of the gradient, are more generally applicable.
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
@@
 
-Random search methods [@zabinsky2013stochastic; @Romeijn2009; @andradottir2015review] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
+Random search methods [@Romeijn2009; @zabinsky2013stochastic; @andradottir2015review] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
 The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
@@
 Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where worse candidates still have a chance to be accepted.
-The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @belisle1992convergence; @granville1994simulated] and the global optimum can be attained under mild regularity conditions.
+The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @granville1994simulated] and the global optimum can be attained under mild regularity conditions.
 
@@
 clustering, outliers and skewness) in multivariate data via low-dimensional projections.
-Guided tour [@cook2008grand; @cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.
+Guided tour [@cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.
 
@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There has been plentiful index functions proposed in the literature and here list a few: early indices that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+There has been plentiful index functions proposed in the literature and here list a few: early indices that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index[@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 
@@
 
-Simulated annealing (SA) [@kirkpatrick1983optimization; @bertsimas1993simulated] uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current.
+Simulated annealing (SA) uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current.
 Given an initial value of $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
@@
 
-Pseudo derivative (PD) search [@cook1995grand] uses a different strategy than CRS and SA.
+Pseudo derivative (PD) search uses a different strategy than CRS and SA.
 Rather than randomly sample the basis space, PD first computes a search direction by evaluating bases close to the current basis.
@@
 \renewcommand\arraystretch{2}  % default value: 1.0
-\left[
-\begin{array}{c|ccc|cc|cc}
+\begin{pmatrix}
 t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
@@
 T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
-\end{array}
-\right]
+\end{pmatrix}
 = 
-\left[
-\begin{array}{c}
+\begin{pmatrix}
 \text{column name} \\
@@
 \text{search (no output)} \\
-\end{array}
-\right]
+\end{pmatrix}
 \label{eq:data-structure}
@@
 With tidy data structure, data wrangling and visualisation can be significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].
-
-<!-- The construction of diagnostic plots adopts the core concept of grammar of graphics [@wickham2010layered] in ggplot2.
-In grammar of graphics, plots are not produced by calling the commands, named by the appearance of the plot, i.e., boxplot and histogram, but by the concept of stacked layers.
-Seeing plots as stacked layers gives analysts the freedom to compose plots with multiple elements on hand.
--->
 
@@
 Several annotations have been made to help understanding this plot.
-As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the CRAN package \CRANpkg{geozoo} [@geozoo].
+As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via \CRANpkg{geozoo} [@geozoo] package.
 Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated.
@@
 
-As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \CRANpkg{geozoo} [@geozoo] package.
+As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \pkg{geozoo} [@geozoo] package.
 While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
changes based on checklist WIP

@@
 package, \pkg{ferrn}.
-
-\noindent \textit{keywords}: graphics, multivariate, optimization
 }
@@
 Derivative-free methods
-\citep{rios2013derivative, conn2009introduction}, which do not rely on
+\citep{conn2009introduction, rios2013derivative}, which do not rely on
 the knowledge of the gradient, are more generally applicable.
@@
 Random search methods
-\citep{zabinsky2013stochastic, Romeijn2009, andradottir2015review} have
+\citep{Romeijn2009, zabinsky2013stochastic, andradottir2015review} have
 a random sampling component as part of their algorithm and have been
@@
 accepted. The convergence of simulated annealing algorithms has been
-widely researched
-\citep{mitra1986convergence, belisle1992convergence, granville1994simulated}
+widely researched \citep{mitra1986convergence, granville1994simulated}
 and the global optimum can be attained under mild regularity conditions.
@@
 (e.g.~clustering, outliers and skewness) in multivariate data via
-low-dimensional projections. Guided tour
-\citep{cook2008grand, cook1995grand} is one variation of a broader class
-of data visualisation methods, tour \citep{buja2005computational}, which
-displays high-dimensional data through a series of animated projections.
+low-dimensional projections. Guided tour \citep{cook1995grand} is one
+variation of a broader class of data visualisation methods, tour
+\citep{buja2005computational}, which displays high-dimensional data
+through a series of animated projections.
 
@@
 \citep{cook1993projection}; chi-square index \citep{posse1995projection}
-for detecting spiral structure; kurtosis index \citep{Loperfido2020} and
-skewness index \citep{Loperfido2018} for detecting outliers in financial
-time series; and most recently, scagnostic indices \citep{laa2020using}
-for summarising structures in scatterplot matrices based on eight
-scagnostic measures.
+for detecting spiral structure; LDA index\citep{lee2005projection} and
+PDA \citep{lee2010projection} index for supervised classification;
+kurtosis index \citep{Loperfido2020} and skewness index
+\citep{Loperfido2018} for detecting outliers in financial time series;
+and most recently, scagnostic indices \citep{laa2020using} for
+summarising structures in scatterplot matrices based on eight scagnostic
+measures.
 
@@
 
-Simulated annealing (SA)
-\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
-same sampling process as CRS but allows a probabilistic acceptance of a
-basis with lower index value than the current. Given an initial value of
-\(T_0\), the temperature at iteration \(l\) is defined as
-\(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to have
-an index value larger than the current basis, SA gives it a second
-chance to be accepted with probability
+Simulated annealing (SA) uses the same sampling process as CRS but
+allows a probabilistic acceptance of a basis with lower index value than
+the current. Given an initial value of \(T_0\), the temperature at
+iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
+candidate basis fails to have an index value larger than the current
+basis, SA gives it a second chance to be accepted with probability
 \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
@@
 
-Pseudo derivative (PD) search \citep{cook1995grand} uses a different
-strategy than CRS and SA. Rather than randomly sample the basis space,
-PD first computes a search direction by evaluating bases close to the
-current basis. A step size is then chosen along the corresponding
-geodesic by another optimisation over an 90 degree angle from \(-\pi/4\)
-to \(\pi/4\). The resulting candidate basis \(\mathbf{A}_{**}\) is
-returned for the current iteration if it has a higher index value than
-the current one. Algorithm \ref{search-geodesic} in the appendix
-summarises the inner loop of the PD.
+Pseudo derivative (PD) search uses a different strategy than CRS and SA.
+Rather than randomly sample the basis space, PD first computes a search
+direction by evaluating bases close to the current basis. A step size is
+then chosen along the corresponding geodesic by another optimisation
+over an 90 degree angle from \(-\pi/4\) to \(\pi/4\). The resulting
+candidate basis \(\mathbf{A}_{**}\) is returned for the current
+iteration if it has a higher index value than the current one. Algorithm
+\ref{search-geodesic} in the appendix summarises the inner loop of the
+PD.
 
@@
 mentioned previously, the original basis space is a high-dimensional
-sphere and random bases on the sphere can be generated via the CRAN
-package \CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded
+sphere and random bases on the sphere can be generated via
+\CRANpkg{geozoo} \citep{geozoo} package. Along with the bases recorded
 during the optimisation and a zero basis, the parameters (centre and
@@
 As mentioned previously, the original \(p \times d\) dimension space can
-be simulated via randomly generated bases in the \CRANpkg{geozoo}
+be simulated via randomly generated bases in the \pkg{geozoo}
 \citep{geozoo} package. While the PCA plot projects the bases from the

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
indices -> index, australian english, some gramma issues

@@
     email:  patricia.menendez@monash.edu 
-abstract: A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimization path.
+abstract: A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimisation path.
 Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
-While much work has been done in developing new indexes in the literature, less has been done on understanding the optimization.
+While much work has been done in developing new indexes in the literature, less has been done on understanding the optimisation.
 Index functions can be noisy, they might have multiple local maxima as well as an optimal maximum, and it is constrained to generate orthonormal projection frames, which complicates the optimization.
@@
 The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some structures of the data that depart from normalality, such as bimodality, clustering, or outliers.
+A view is interesting if it can show some structures of the data that depart from normality, such as bimodality, clustering, or outliers.
 
@@
 As the dimension of the space becomes larger, sufficient sampling from the entire space would require a long time for convergence to occur, despite a guaranteed global convergence [@spall2005introduction].
-Various algorithms have thus been developed to improve pure random search by either concentrating on a narrower sampling space or using a different updating machanism.
+Various algorithms have thus been developed to improve pure random search by either concentrating on a narrower sampling space or using a different updating mechanism.
 Creeping random search [@White1971-ci] is such a variation, where a candidate point is generated within a neighbourhood of the current point.
@@
 
-Pseudo derivative search uses a common search scheme in optimistaion: line search.
+Pseudo derivative search uses a common search scheme in optimisation: line search.
 In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There has been plentiful index functions proposed in the literature and here list a few: early indices that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index[@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indices [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+There has been plentiful index functions proposed in the literature and here list a few: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index[@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 
@@
 When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
-This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enable the optimiser to jump out of a local optimum.
+This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enables the optimiser to jump out of a local optimum.
 The algorithm \ref{simulated-annealing} in the appendix highlights how SA differs from CRS in the inner loop.
@@
 In the case of few tries at a given iteration, showing the data points directly may be preferred over the boxplot, this is controlled via the `cutoff` argument.
-Additional annotations are added to facilitate better reading of the plot and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
+Additional annotations are added to facilitate better reading of the plot and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a greyscale to indicate no better basis found in this iteration.
 
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130."}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reaching the target bases in time 76-130."}
 ```
 
-Another interesting feature to examine is the changes on the index value between interpolating bases since the projection on these bases is shown in the tour animation.
+Another interesting feature to examine is the changes in the index value between interpolating bases since the projection on these bases is shown in the tour animation.
 Trace plots are created by plotting the index value against time.
@@
 It may seem bizarre at the first sight that the interpolation sometimes passes bases with higher index value before it decreases to a lower target.
-This happens because on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
+This happens because, on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
 In addition, the guided tour interpolates between the current and target basis to provide a smooth transition between projections, and sometimes a higher index value will be observed along the interpolation path.
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
 ```
@@
 Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a Stiefel manifold.
-For one-dimensional projections this forms a $p$-dimensional sphere.
+For one-dimensional projections, this forms a $p$-dimensional sphere.
 A dimensionality reduction method, e.g.
@@
 In a projection pursuit guided tour optimisation, there are various types of bases involved: 1) The starting basis; 2) The search bases that the optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally, 5) the end basis.
-The importance of these bases differs and the most important ones are the starting, interpolating and end bases.
+The importance of these bases differs and the most important ones are the starting, interpolating, and end bases.
 The anchor and search bases can be turned on with argument `details = TRUE`.
@@
 This happens because the projection is invariant to the orientation of the basis and so is the index value.
-However, this creates difficulties for comparing the optimisers since the end bases will be symmetric with respect to the origin.
+However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin.
 A sign flipping step is conducted to flip the signs of all the bases in one routine if different optimisations finish at opposite places.
@@
 Various aesthetics, i.e.
-size, alpha and colour, are applicable to emphasize critical elements and adjust for the presentation.
-For example, anchor points and search points are less important and hence a smaller size and alpha is used.
+size, alpha, and colour, are applicable to emphasize critical elements and adjust for the presentation.
+For example, anchor points and search points are less important and hence a smaller size and alpha are used.
 Alpha can also be applied on the interpolation paths to show the start to finish from transparent to opaque.
@@
 With the PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
-There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in details in Section \ref{monotonic}.
+There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in detail in Section \ref{monotonic}.
 
@@
 ```{r toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot.
-With animation, the progression of the search paths from start to finish are better identified.
+With animation, the progression of the search paths from start to finish is better identified.
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
@@
 Note however the different orientation of the basis, following from the different sign in the second column.
-One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and be symmetric when projected onto the other (due the the sign difference in the second column).
+One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and are symmetric when projected onto the other (due the sign difference in the second column).
 In Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
@@
 ```{r toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
-The projection problem is a 2D projection with three variables using holes index.
+The projection problem is a 2D projection with three variables using the holes index.
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
@@
 Random variables with different distributions have been simulated and the distributional form of each variable is presented in equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variables `x1`, `x8`, `x9` and `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
+Variables `x1`, `x8` to `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
 All the variables have been scaled to have an overall unit variance before projection pursuit.
-Holes index [@cook2008grand], used for detecting binormality of the variables, is used throughout the examples unless otherwise specified.
+Holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
 
@@
 ```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index.
-Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis."}
+The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis."}
 ```
@@
 A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local improvements.
-The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space,  percentage change in the index value, and neighbourhood size:
+The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space, the percentage change in the index value, and neighbourhood size:
 
 1) the distance between the basis found and the current basis needs to be larger than 1e-3;
-2) the percentage change of the index value need to be larger than 1e-5; and
+2) the percentage change of the index value needs to be larger than 1e-5; and
 3) the alpha parameter on itself needs to be larger than 0.01
 
-Figure \ref{fig:polish} presents the projectd data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
-The top row shows the intial projection, the final projection after CRS, and the final projection after polish, repsectively.
-The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
-Polish works with this end basis and further pushes the index value to produce much clear edges of the cluster, especially along the vertical axis.
+Figure \ref{fig:polish} presents the projected data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
+The top row shows the initial projection, the final projection after CRS, and the final projection after polish, respectively The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
+Polish works with this end basis and further pushes the index value to produce clearer edges of the cluster, especially along the vertical axis.
 
@@
 
-The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indices.
-An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
+The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indexes An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
 Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index, $I^{nk}(n)$, is defined as:
@@
 ```{r noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
-Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
 Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best."}
@@
 ```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
 ```
@@
 Two search neighbourhood sizes: 0.5 and 0.7 are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
-Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by  optimiser and search size.
+Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the optimiser and search size.
 With CRS and a search size of 0.5, despite being the simplest and fastest, the optimiser fails in three instances where the final basis lands neither near the local nor the global optimum.
@@
 Comparing CRS and SA for a search size of 0.5, SA does not seem to improve the final basis found, despite having longer interpolation paths.
-However, the denser paths near the local maximum is an indicator that SA is working hard to examine if there is any other optimum in the basis space but the relative small search size has diminished its ability to reach the global maximum.
+However, the denser paths near the local maximum is an indicator that SA is working hard to examine if there is any other optimum in the basis space but the relatively small search size has diminished its ability to reach the global maximum.
 With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
@@
 
-One interesting situation observed in the previous examples is that, for some simulations as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
+One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
 As mentioned previously, bases with opposite signs do not affect the projection and index value, but clearly, we prefer the target to have the same orientation as the current basis.
@@
 Hence an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
-The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exist.
+The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exists.
 
@@
   - \code{clean\_method()} to clean up the name of the optimisers
-  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia
+  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants.
+Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
   - \code{botanical\_pal()} as the colour interpolator 

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
indices -> index, australian english, some gramma issues

@@
 A guided tour helps to visualise high-dimensional data by showing the
-low-dimensional projections along a projection pursuit optimization
+low-dimensional projections along a projection pursuit optimisation
 path. Projection pursuit is a generalisation of principal component
@@
 developing new indexes in the literature, less has been done on
-understanding the optimization. Index functions can be noisy, they might
+understanding the optimisation. Index functions can be noisy, they might
 have multiple local maxima as well as an optimal maximum, and it is
@@
 gives the most interesting low-dimensional view. A view is interesting
-if it can show some structures of the data that depart from normalality,
+if it can show some structures of the data that depart from normality,
 such as bimodality, clustering, or outliers.
@@
 search by either concentrating on a narrower sampling space or using a
-different updating machanism. Creeping random search
+different updating mechanism. Creeping random search
 \citep{White1971-ci} is such a variation, where a candidate point is
@@
 
-Pseudo derivative search uses a common search scheme in optimistaion:
+Pseudo derivative search uses a common search scheme in optimisation:
 line search. In line search methods, users are required to provide an
@@
 There has been plentiful index functions proposed in the literature and
-here list a few: early indices that can be categorised as measuring the
+here list a few: early indexes that can be categorised as measuring the
 \(L^2\) distance between the projection and a normal distribution:
@@
 \citep{Loperfido2018} for detecting outliers in financial time series;
-and most recently, scagnostic indices \citep{laa2020using} for
+and most recently, scagnostic indexes \citep{laa2020using} for
 summarising structures in scatterplot matrices based on eight scagnostic
@@
 space even if the candidate basis does not have a higher index value and
-hence enable the optimiser to jump out of a local optimum. The algorithm
-\ref{simulated-annealing} in the appendix highlights how SA differs from
-CRS in the inner loop.
+hence enables the optimiser to jump out of a local optimum. The
+algorithm \ref{simulated-annealing} in the appendix highlights how SA
+differs from CRS in the inner loop.
 
@@
 the anchor bases to interpolate are connected and highlighted in a
-larger size; and 3) the colour of the last iteration is in a grey scale
+larger size; and 3) the colour of the last iteration is in a greyscale
 to indicate no better basis found in this iteration.
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reach the target bases in time 76-130.}\label{fig:toy-interp}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reaching the target bases in time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 
-Another interesting feature to examine is the changes on the index value
+Another interesting feature to examine is the changes in the index value
 between interpolating bases since the projection on these bases is shown
@@
 higher index value before it decreases to a lower target. This happens
-because on the one hand, the probabilistic acceptance in SA implies that
-some worse bases will be accepted by the optimiser. In addition, the
-guided tour interpolates between the current and target basis to provide
-a smooth transition between projections, and sometimes a higher index
-value will be observed along the interpolation path. This indicates that
-a non-monotonic interpolation can't be avoided, even for CRS. Later, in
-section \ref{monotonic}, there will be a discussion on improving the
-non-monotonic interpolation for CRS.
+because, on the one hand, the probabilistic acceptance in SA implies
+that some worse bases will be accepted by the optimiser. In addition,
+the guided tour interpolates between the current and target basis to
+provide a smooth transition between projections, and sometimes a higher
+index value will be observed along the interpolation path. This
+indicates that a non-monotonic interpolation can't be avoided, even for
+CRS. Later, in section \ref{monotonic}, there will be a discussion on
+improving the non-monotonic interpolation for CRS.
 
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
 \end{figure}
@@
 bases \(\mathbf{A}_{p \times d}\) is a Stiefel manifold. For
-one-dimensional projections this forms a \(p\)-dimensional sphere. A
+one-dimensional projections, this forms a \(p\)-dimensional sphere. A
 dimensionality reduction method, e.g.~principal component analysis is
@@
 these bases differs and the most important ones are the starting,
-interpolating and end bases. The anchor and search bases can be turned
+interpolating, and end bases. The anchor and search bases can be turned
 on with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers
@@
 the basis and so is the index value. However, this creates difficulties
-for comparing the optimisers since the end bases will be symmetric with
-respect to the origin. A sign flipping step is conducted to flip the
-signs of all the bases in one routine if different optimisations finish
-at opposite places.
+for comparing the optimisers since the end bases will be symmetric to
+the origin. A sign flipping step is conducted to flip the signs of all
+the bases in one routine if different optimisations finish at opposite
+places.
 
@@
 close the end basis found by the optimisers is to the theoretical best.
-Various aesthetics, i.e.~size, alpha and colour, are applicable to
+Various aesthetics, i.e.~size, alpha, and colour, are applicable to
 emphasize critical elements and adjust for the presentation. For
 example, anchor points and search points are less important and hence a
-smaller size and alpha is used. Alpha can also be applied on the
+smaller size and alpha are used. Alpha can also be applied on the
 interpolation paths to show the start to finish from transparent to
@@
 for CRS and it describes the interruption of the interpolation, which
-will be discussed in details in Section \ref{monotonic}.
+will be discussed in detail in Section \ref{monotonic}.
 
@@
 \caption{Six frames selected from the animated version of the previous plot.
-With animation, the progression of the search paths from start to finish are better identified.
+With animation, the progression of the search paths from start to finish is better identified.
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
@@
 final bases match each other when projected onto one torus circle (due
-to the same sign in the first column) and be symmetric when projected
-onto the other (due the the sign difference in the second column). In
-Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5
-where the two circles are rotated into a line from our view.
+to the same sign in the first column) and are symmetric when projected
+onto the other (due the sign difference in the second column). In Figure
+\ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the
+two circles are rotated into a line from our view.
 
@@
 \caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
-The projection problem is a 2D projection with three variables using holes index.
+The projection problem is a 2D projection with three variables using the holes index.
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
@@
 the distributional form of each variable is presented in equations
-\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1},
-\texttt{x8}, \texttt{x9} and \texttt{x10} are normal noise with zero
-mean and unit variance and \texttt{x2} to \texttt{x7} are normal
-mixtures with varied weights and locations. All the variables have been
-scaled to have an overall unit variance before projection pursuit. Holes
-index \citep{cook2008grand}, used for detecting binormality of the
-variables, is used throughout the examples unless otherwise specified.
+\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1}, \texttt{x8}
+to \texttt{x10} are normal noise with zero mean and unit variance and
+\texttt{x2} to \texttt{x7} are normal mixtures with varied weights and
+locations. All the variables have been scaled to have an overall unit
+variance before projection pursuit. Holes index \citep{cook2008grand},
+used for detecting bimodality of the variables, is used throughout the
+examples unless otherwise specified.
 
@@
 \caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
-Top row shows the intial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertial axis.}\label{fig:polish}
+The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.}\label{fig:polish}
 \end{figure}
@@
 have been added to ensure the new basis generated is distinguishable
-from the current one in terms of the distance in the space, percentage
-change in the index value, and neighbourhood size:
+from the current one in terms of the distance in the space, the
+percentage change in the index value, and neighbourhood size:
 
@@
 \item
-  the percentage change of the index value need to be larger than 1e-5;
+  the percentage change of the index value needs to be larger than 1e-5;
   and
@@
 
-Figure \ref{fig:polish} presents the projectd data and trace plot of a
+Figure \ref{fig:polish} presents the projected data and trace plot of a
 2D projection, optimised by CRS and followed by the polish step. The top
-row shows the intial projection, the final projection after CRS, and the
-final projection after polish, repsectively. The end basis found by CRS
-reveals the four clusters in the data, but the edges of each cluster are
-not clean-cut. Polish works with this end basis and further pushes the
-index value to produce much clear edges of the cluster, especially along
-the vertical axis.
+row shows the initial projection, the final projection after CRS, and
+the final projection after polish, respectively The end basis found by
+CRS reveals the four clusters in the data, but the edges of each cluster
+are not clean-cut. Polish works with this end basis and further pushes
+the index value to produce clearer edges of the cluster, especially
+along the vertical axis.
 
@@
 produces a smooth interpolation, while this is not the case for all the
-indices. An example of a noisy index function for 1D projections
-compares the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly
-generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
-Kolmogorov test. Let \(F_{.}(n)\) be the ECDF function, with two
-possible subscripts \(Y\) and \(\mathcal{N}\) representing the projected
-and randomly generated data, and \(n\) denoting the number of
-observation, the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
+indexes An example of a noisy index function for 1D projections compares
+the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated
+normal distribution, \(\mathcal{N}_{n \times 1}\), using the Kolmogorov
+test. Let \(F_{.}(n)\) be the ECDF function, with two possible
+subscripts \(Y\) and \(\mathcal{N}\) representing the projected and
+randomly generated data, and \(n\) denoting the number of observation,
+the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
 
@@
 \caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}.
-Both CRS and SA succeeds in the optimisation, PD fails to optimise this non-smooth index.
+Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
 Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
@@
 \caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
 \end{figure}
@@
 basis and the length of the search. Figure \ref{fig:kol-result} shows 80
-paths simulated using 20 seeds in the PCA plot, faceted by optimiser and
-search size. With CRS and a search size of 0.5, despite being the
+paths simulated using 20 seeds in the PCA plot, faceted by the optimiser
+and search size. With CRS and a search size of 0.5, despite being the
 simplest and fastest, the optimiser fails in three instances where the
@@
 that SA is working hard to examine if there is any other optimum in the
-basis space but the relative small search size has diminished its
+basis space but the relatively small search size has diminished its
 ability to reach the global maximum. With a larger search size, almost
@@
 One interesting situation observed in the previous examples is that, for
-some simulations as shown on the left panel of Figure
+some simulations, as shown on the left panel of Figure
 \ref{fig:flip-sign}, the target basis is generated on the other half of
@@
 right panel of Figure \ref{fig:flip-sign} where the unsatisfactory
-interpolation no longer exist.
+interpolation no longer exists.
 
@@
     Australian native plants. Quantitative palettes include daisy,
-    banksia and cherry and sequential palettes contain fern and acacia
+    banksia, and cherry and sequential palettes contain fern and acacia
   \item

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.Rmd
proofreading checklist done

@@
     address: Department of Econometrics and Business Statistics 
+    url: Melbourne, Australia
     email:  huize.zhang@monash.edu
@@
     address: Department of Econometrics and Business Statistics
+    url: Melbourne, Australia
     email:  dicook@monash.edu
@@
     address: Institute of Statistics
+    url: Vienna, Austria
     email:  ursula.laa@boku.ac.at  
@@
     address: 34 Village Street, Docklands VIC 3008 Australia
+    url: Melbourne, Australia
     email: nicolas.langrene@csiro.au
@@
     address: Department of Econometrics and Business Statistics
+    url: Melbourne, Australia
     email:  patricia.menendez@monash.edu 
@@
 preamble: >
-  \usepackage{amssymb, amsmath, mathtools, dsfont, bbm, array, booktabs}
-  \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
+  \usepackage{mathtools}
+  \usepackage[ruled, linesnumbered]{algorithm2e}
 output: rticles::rjournal_article
@@
 
-```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the frames in a tour path.
+```{r tour-path, out.width="70%", fig.cap="An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
 Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."}
-include_graphics(here::here("img/tour-path.png"))
 ```

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.R

@@ -6,7 +6,8 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.path = here::here("figs/"), 
                       cache = TRUE,
                       fig.align = "center",
-                      fig.height = 2.5,
+                      fig.height = 3,
+                      fig.retina = 5,
                       out.width = "100%")
 
 
@@ -24,15 +25,8 @@ knitr::read_chunk(
 )
 
 knitr::read_chunk(
-  here::here("scripts/implementation.R")
-)
-
-knitr::read_chunk(
   here::here("scripts/kol_result.R")
 )
-
-
-
 
 
 
@@ -43,20 +37,19 @@ knitr::read_chunk(
 #spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
 
 
-## ----tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
-include_graphics(here::here("img/tour-path.png"))
+## ----tour-path, out.width="70%", fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
 
 
-## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis."----
+## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey."----
 
 
[…]

### 2021-03-26 scripts/toy.R

@@ -6,6 +6,9 @@ library(tourr)
 files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
+## ----tour-path
+include_graphics(here::here("figs/tour-path.png"))
+
 ## ---- toy-search
 p1 <- holes_2d_better_max_tries %>%
   mutate(max_tries = 500) %>%

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.Rmd
minor fixes to optimisation section

@@
 
-The optimisation problem considered in this paper is on constrained optimization [@bertsekas2014constrained] and assuming it is not possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists of finding the minimum or maximum of a function $f \in L^p$ in the constrained $\mathcal{A}$ space.
+The type of optimisation problem considered in this paper is constrained optimization [@bertsekas2014constrained], assuming it is not possible to find a solution to the problem in the way of a closed-form.
+That is, the problem consists in finding the minimum or maximum of a function $f \in L^p$ in the constrained space $\mathcal{A}$.
 
@@
 Although these methods are popular, they rely on the availability of the objective function derivatives.
-As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the render of tour animation.
-Hence, gradient-based methods are not the focus of this paper and considerations will be given to derivative-free methods.
+As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the rendering of tour animation.
+Hence, gradient-based methods are not the focus of this paper and consideration will be given to derivative-free methods.
 
@@
 Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
-Here we consider three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo derivative search.
+Here we consider three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo-derivative search.
 
@@
 Creeping random search [@White1971-ci] is such a variation, where a candidate point is generated within a neighbourhood of the current point.
-This makes creeping random search faster to compute, however, global convergence is not longer guaranteed.
+This makes creeping random search faster to compute, however, global convergence is no longer guaranteed.
 On the other hand, simulated annealing [@kirkpatrick1983optimization; @bertsimas1993simulated], introduces a different updating mechanism.
@@
 
-Pseudo derivative search uses a common search scheme in optimisation: line search.
+Pseudo-derivative search uses a common search scheme in optimisation: line search.
 In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.Rmd
several minor corrections

@@
     email:  patricia.menendez@monash.edu 
-abstract: A guided tour helps to visualise high-dimensional data by showing the low-dimensional projections along a projection pursuit optimisation path.
+abstract: A guided tour helps to visualise high-dimensional data by showing low-dimensional projections along a projection pursuit optimisation path.
 Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
 While much work has been done in developing new indexes in the literature, less has been done on understanding the optimisation.
-Index functions can be noisy, they might have multiple local maxima as well as an optimal maximum, and it is constrained to generate orthonormal projection frames, which complicates the optimization.
+Index functions can be noisy, might have multiple local maxima as well as an optimal maximum, and are constrained to generate orthonormal projection frames, which complicates the optimization.
 In addition, projection pursuit is primarily used for exploratory data analysis, and finding the local maxima is also useful.
@@
 This work might be useful more broadly for diagnosing optimisers, and comparing their performance.
-The diagnostics are implemented in the R package, \pkg{ferrn}.
+The diagnostics are implemented in the R package \pkg{ferrn}.
 keywords: graphics, multivariate, optimization
@@
 The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view.
-A view is interesting if it can show some structures of the data that depart from normality, such as bimodality, clustering, or outliers.
+A view is said to be interesting if it can show some structures of the data that depart from normality, such as bimodality, clustering, or outliers.
 
@@
 The remainder of the paper is organised as follows. 
-The next section provides an overview of optimisation methods, specifically line search methods.
+The next section provides an overview of optimisation methods, specifically random search and line search methods.
 A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
-The third section presents the new visual diagnostics, including the the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
+The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
 An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms applications is shown using simulated data.
@@
 
-Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high dimensional space.
+Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space.
 Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There has been plentiful index functions proposed in the literature and here list a few: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index[@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
 
-As a general visualisation method, tour produces animations of high dimensional data via rotations of low-dimensional planes.
-There are different versions depending on how the high dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
+As a general visualisation method, tour produces animations of high-dimensional data via rotations of low-dimensional planes.
+There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview;  manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
 Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection.
@@
 
-Three optimisers have been implemented in the \pkg{tourr} [@tourr] package: Creeping random search (CRS), simulated annealing (SA), and pseudo derivative (PD).
+Three optimisers have been implemented in the \pkg{tourr} [@tourr] package: creeping random search (CRS), simulated annealing (SA), and pseudo-derivative (PD).
 Creeping random search (CRS) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
@@
 
-Simulated annealing (SA) uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current.
-Given an initial value of $T_0$, the temperature at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+Simulated annealing (SA) uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current one.
+Given an initial value of $T_0$, the ``temperature'' at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
 When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
@@
 
-Pseudo derivative (PD) search uses a different strategy than CRS and SA.
+Pseudo-derivative (PD) search uses a different strategy than CRS and SA.
 Rather than randomly sample the basis space, PD first computes a search direction by evaluating bases close to the current basis.
@@
 For CRS and SA, possible states include `random_search`,  `new_basis`, and `interpolation`.
-Pseudo derivative (PD) has a wider variety of states including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
+Pseudo-derivative (PD) has a wider variety of states including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
 Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation;  $j$ and $l$ are the counter of the outer and inner loop respectively.
@@
 Figure \ref{fig:toy-interp} presents the trace plot of the same optimisations as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
-It may seem bizarre at the first sight that the interpolation sometimes passes bases with higher index value before it decreases to a lower target.
+It may seem bizarre at first sight that the interpolation sometimes passes bases with higher index value before it decreases to a lower target.
 This happens because, on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
 In addition, the guided tour interpolates between the current and target basis to provide a smooth transition between projections, and sometimes a higher index value will be observed along the interpolation path.
-This indicates that a non-monotonic interpolation can't be avoided, even for CRS.
+This indicates that a non-monotonic interpolation cannot be avoided, even for CRS.
 Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for CRS.
@@
 Several annotations have been made to help understanding this plot.
-As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via \CRANpkg{geozoo} [@geozoo] package.
+As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
 Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated.
-PCA is performed to get the first two PC coordinates of all the bases and the centre of the circle is the PC coordinates of the zero matrix, and radius is estimated as the largest distance from the centre to the basis.
+A PCA is performed to get the first two PC coordinates of all the bases and the centre of the circle is the PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
 The theoretical best basis is known for simulated data and can be labelled to compare how close the end basis found by the optimisers is to the theoretical best.
 Various aesthetics, i.e.
-size, alpha, and colour, are applicable to emphasize critical elements and adjust for the presentation.
+size, alpha (transparency), and colour, are applicable to emphasize critical elements and adjust for the presentation.
 For example, anchor points and search points are less important and hence a smaller size and alpha are used.
@@
 
-Animation is another type of display to show how the search progresses from start to finish in the space.
+An animation is another type of display to show how the search progresses from start to finish in the space.
 An  `animate = TRUE` argument is used to enable the animation in the PCA plot.
@@
 
-```{r toy-tour, fig.cap="Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
+```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \\CRANpkg{geozoo}.
@@
 All the variables have been scaled to have an overall unit variance before projection pursuit.
-Holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
+The holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
 
@@
 Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best."}
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best."}
 ```
@@
 The PCA plot shows the interpolation path and search points excluding the last termination iteration.
-Pseudo Derivative (PD) quickly gets stuck near the starting position.
+Pseudo-Derivative (PD) quickly gets stuck near the starting position.
 Comparing the amount of random search done by CRS and SA, it is surprising that SA does not carry as many samples as CRS.
@@
 
-```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
+```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
 The optimisation is on the 1D projection index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
@@
 This project contributes to the software development in two packages: a data collection object is implemented in \pkg{tourr} [@tourr], while the visual diagnostics of the optimisers is implemented in \pkg{ferrn} [@ferrn].
-Below list functions in the \pkg{ferrn} [@ferrn] package:
+The functions in the \pkg{ferrn} [@ferrn] package are listed below:
 
@@
 
-  - \code{explore\_trace\_search()} produces summary plots in Figure \ref{fig:toy-search}
-  - \code{explore\_trace\_interp()}produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}
+  - \code{explore\_trace\_search()} produces summary plots in Figure \ref{fig:toy-search}.
+  - \code{explore\_trace\_interp()} produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}.
   - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
-Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
+The animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
   - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
@@
 
-Three algorithms (Creeping random search, simulated annealing, and pseudo derivative) used in projection pursuit guided tour optimisation.
+Three algorithms (creeping random search, simulated annealing, and pseudo-derivative) used in projection pursuit guided tour optimisation.
 
@@
 }
-\caption{Pseudo derivative (PD)}
+\caption{Pseudo-derivative (PD)}
 \label{search-geodesic}

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.Rmd
compile pdf

@@
 Simulated annealing (SA) uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current one.
-Given an initial value of $T_0$, the ``temperature'' at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+Given an initial value of $T_0$, the "temperature" at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
 When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
@@
 As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
-Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated.
-A PCA is performed to get the first two PC coordinates of all the bases and the centre of the circle is the PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
+Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated by PCA.
+The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
 The theoretical best basis is known for simulated data and can be labelled to compare how close the end basis found by the optimisers is to the theoretical best.

### 2021-03-26 paper/zhang-cook-laa-langrene-menendez.tex
proofreading checklist done
compile pdf

@@
 \abstract{%
-A guided tour helps to visualise high-dimensional data by showing the
+A guided tour helps to visualise high-dimensional data by showing
 low-dimensional projections along a projection pursuit optimisation
@@
 developing new indexes in the literature, less has been done on
-understanding the optimisation. Index functions can be noisy, they might
-have multiple local maxima as well as an optimal maximum, and it is
-constrained to generate orthonormal projection frames, which complicates
-the optimization. In addition, projection pursuit is primarily used for
+understanding the optimisation. Index functions can be noisy, might have
+multiple local maxima as well as an optimal maximum, and are constrained
+to generate orthonormal projection frames, which complicates the
+optimization. In addition, projection pursuit is primarily used for
 exploratory data analysis, and finding the local maxima is also useful.
@@
 comparing their performance. The diagnostics are implemented in the R
-package, \pkg{ferrn}.
+package \pkg{ferrn}.
 }
@@
 identify the projection, represented by the projection matrix, that
-gives the most interesting low-dimensional view. A view is interesting
-if it can show some structures of the data that depart from normality,
-such as bimodality, clustering, or outliers.
+gives the most interesting low-dimensional view. A view is said to be
+interesting if it can show some structures of the data that depart from
+normality, such as bimodality, clustering, or outliers.
 
@@
 The remainder of the paper is organised as follows. The next section
-provides an overview of optimisation methods, specifically line search
-methods. A review of the projection pursuit guided tour, an overview of
-the optimisation problem and outlines of three existing algorithms
-follows. The third section presents the new visual diagnostics,
-including the the design of a data structure to capture information
-during the optimisation, from which several diagnostic plots are
-created. An illustration of how the diagnostic plots can be used to
+provides an overview of optimisation methods, specifically random search
+and line search methods. A review of the projection pursuit guided tour,
+an overview of the optimisation problem and outlines of three existing
+algorithms follows. The third section presents the new visual
+diagnostics, including the design of a data structure to capture
+information during the optimisation, from which several diagnostic plots
+are created. An illustration of how the diagnostic plots can be used to
 examine the performance of different optimisers and guide improvements
@@
 
-The optimisation problem considered in this paper is on constrained
-optimization \citep{bertsekas2014constrained} and assuming it is not
+The type of optimisation problem considered in this paper is constrained
+optimization \citep{bertsekas2014constrained}, assuming it is not
 possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists of finding the minimum or maximum of a
-function \(f \in L^p\) in the constrained \(\mathcal{A}\) space.
+That is, the problem consists in finding the minimum or maximum of a
+function \(f \in L^p\) in the constrained space \(\mathcal{A}\).
 
@@
 are the entries of a projection matrix and the computational time
-required to perform differentiation on a matrix could impede the render
-of tour animation. Hence, gradient-based methods are not the focus of
-this paper and considerations will be given to derivative-free methods.
+required to perform differentiation on a matrix could impede the
+rendering of tour animation. Hence, gradient-based methods are not the
+focus of this paper and consideration will be given to derivative-free
+methods.
 
@@
 methods: creeping random search and simulated annealing, and the other
-one is pseudo derivative search.
+one is pseudo-derivative search.
 
@@
 creeping random search faster to compute, however, global convergence is
-not longer guaranteed. On the other hand, simulated annealing
+no longer guaranteed. On the other hand, simulated annealing
 \citep{kirkpatrick1983optimization, bertsimas1993simulated}, introduces
@@
 
-Pseudo derivative search uses a common search scheme in optimisation:
+Pseudo-derivative search uses a common search scheme in optimisation:
 line search. In line search methods, users are required to provide an
@@
 (projection pursuit and guided tour) to explore interesting features in
-a high dimensional space. Projection pursuit, coined by
+a high-dimensional space. Projection pursuit, coined by
 \citet{friedman1974projection}, detects interesting structures
@@
 normal, as \(n\) and \(p\) approach infinity under certain conditions.
-There has been plentiful index functions proposed in the literature and
-here list a few: early indexes that can be categorised as measuring the
-\(L^2\) distance between the projection and a normal distribution:
+There have been many index functions proposed in the literature, here
+are a few examples: early indexes that can be categorised as measuring
+the \(L^2\) distance between the projection and a normal distribution:
 Legendre index \citep{friedman1974projection}, Hermite index
@@
 \citep{cook1993projection}; chi-square index \citep{posse1995projection}
-for detecting spiral structure; LDA index\citep{lee2005projection} and
+for detecting spiral structure; LDA index \citep{lee2005projection} and
 PDA \citep{lee2010projection} index for supervised classification;
@@
 
-As a general visualisation method, tour produces animations of high
-dimensional data via rotations of low-dimensional planes. There are
-different versions depending on how the high dimensional space is
+As a general visualisation method, tour produces animations of
+high-dimensional data via rotations of low-dimensional planes. There are
+different versions depending on how the high-dimensional space is
 investigated: grand tour \citep{cook2008grand} selects the planes
@@
 
-{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/tour-path}
 
@@
 Three optimisers have been implemented in the \pkg{tourr} \citep{tourr}
-package: Creeping random search (CRS), simulated annealing (SA), and
-pseudo derivative (PD). Creeping random search (CRS) is a random search
+package: creeping random search (CRS), simulated annealing (SA), and
+pseudo-derivative (PD). Creeping random search (CRS) is a random search
 optimiser that samples a candidate basis \(\mathbf{A}_{l}\) in the
@@
 allows a probabilistic acceptance of a basis with lower index value than
-the current. Given an initial value of \(T_0\), the temperature at
-iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
-candidate basis fails to have an index value larger than the current
-basis, SA gives it a second chance to be accepted with probability
+the current one. Given an initial value of \(T_0\), the ``temperature''
+at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
+When a candidate basis fails to have an index value larger than the
+current basis, SA gives it a second chance to be accepted with
+probability
 \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
@@
 
-Pseudo derivative (PD) search uses a different strategy than CRS and SA.
+Pseudo-derivative (PD) search uses a different strategy than CRS and SA.
 Rather than randomly sample the basis space, PD first computes a search
@@
 include \texttt{random\_search}, \texttt{new\_basis}, and
-\texttt{interpolation}. Pseudo derivative (PD) has a wider variety of
+\texttt{interpolation}. Pseudo-derivative (PD) has a wider variety of
 states including \texttt{new\_basis}, \texttt{direction\_search},
@@
 observe that the trace is smooth in both cases. It may seem bizarre at
-the first sight that the interpolation sometimes passes bases with
-higher index value before it decreases to a lower target. This happens
-because, on the one hand, the probabilistic acceptance in SA implies
-that some worse bases will be accepted by the optimiser. In addition,
-the guided tour interpolates between the current and target basis to
-provide a smooth transition between projections, and sometimes a higher
-index value will be observed along the interpolation path. This
-indicates that a non-monotonic interpolation can't be avoided, even for
-CRS. Later, in section \ref{monotonic}, there will be a discussion on
-improving the non-monotonic interpolation for CRS.
+first sight that the interpolation sometimes passes bases with higher
+index value before it decreases to a lower target. This happens because,
+on the one hand, the probabilistic acceptance in SA implies that some
+worse bases will be accepted by the optimiser. In addition, the guided
+tour interpolates between the current and target basis to provide a
+smooth transition between projections, and sometimes a higher index
+value will be observed along the interpolation path. This indicates that
+a non-monotonic interpolation cannot be avoided, even for CRS. Later, in
+section \ref{monotonic}, there will be a discussion on improving the
+non-monotonic interpolation for CRS.
 
@@
 mentioned previously, the original basis space is a high-dimensional
-sphere and random bases on the sphere can be generated via
+sphere and random bases on the sphere can be generated via the
 \CRANpkg{geozoo} \citep{geozoo} package. Along with the bases recorded
 during the optimisation and a zero basis, the parameters (centre and
-radius) of the 2D space can be estimated. PCA is performed to get the
-first two PC coordinates of all the bases and the centre of the circle
-is the PC coordinates of the zero matrix, and radius is estimated as the
-largest distance from the centre to the basis. The theoretical best
-basis is known for simulated data and can be labelled to compare how
-close the end basis found by the optimisers is to the theoretical best.
-Various aesthetics, i.e.~size, alpha, and colour, are applicable to
-emphasize critical elements and adjust for the presentation. For
-example, anchor points and search points are less important and hence a
-smaller size and alpha are used. Alpha can also be applied on the
-interpolation paths to show the start to finish from transparent to
-opaque.
+radius) of the 2D space can be estimated by PCA. The centre of the
+circle is the first two PC coordinates of the zero matrix, and the
+radius is estimated as the largest distance from the centre to the
+basis. The theoretical best basis is known for simulated data and can be
+labelled to compare how close the end basis found by the optimisers is
+to the theoretical best. Various aesthetics, i.e.~size, alpha
+(transparency), and colour, are applicable to emphasize critical
+elements and adjust for the presentation. For example, anchor points and
+search points are less important and hence a smaller size and alpha are
+used. Alpha can also be applied on the interpolation paths to show the
+start to finish from transparent to opaque.
 
@@
 
-Animation is another type of display to show how the search progresses
-from start to finish in the space. An \texttt{animate\ =\ TRUE} argument
-is used to enable the animation in the PCA plot. Figure
-\ref{fig:toy-pca-animated} shows six frames from the animation of the
-PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
-one can learn from this animation is that CRS finds its end basis
-quicker than PD since CRS finishes its search in the 5th frame while PD
-is still making more progress.
+An animation is another type of display to show how the search
+progresses from start to finish in the space. An
+\texttt{animate\ =\ TRUE} argument is used to enable the animation in
+the PCA plot. Figure \ref{fig:toy-pca-animated} shows six frames from
+the animation of the PCA plot in Figure \ref{fig:toy-pca}. An additional
+piece of information one can learn from this animation is that CRS finds
+its end basis quicker than PD since CRS finishes its search in the 5th
+frame while PD is still making more progress.
 
@@
 
-\caption{Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \CRANpkg{geozoo}.
@@
 locations. All the variables have been scaled to have an overall unit
-variance before projection pursuit. Holes index \citep{cook2008grand},
-used for detecting bimodality of the variables, is used throughout the
-examples unless otherwise specified.
+variance before projection pursuit. The holes index
+\citep{cook2008grand}, used for detecting bimodality of the variables,
+is used throughout the examples unless otherwise specified.
 
@@
 Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 interpolation path and search points excluding the last termination
-iteration. Pseudo Derivative (PD) quickly gets stuck near the starting
+iteration. Pseudo-Derivative (PD) quickly gets stuck near the starting
 position. Comparing the amount of random search done by CRS and SA, it
@@
 
-\caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
+\caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
 The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
@@
 while the visual diagnostics of the optimisers is implemented in
-\pkg{ferrn} \citep{ferrn}. Below list functions in the \pkg{ferrn}
-\citep{ferrn} package:
+\pkg{ferrn} \citep{ferrn}. The functions in the \pkg{ferrn}
+\citep{ferrn} package are listed below:
 
@@
     \code{explore\_trace\_search()} produces summary plots in Figure
-    \ref{fig:toy-search}
-  \item
-    \code{explore\_trace\_interp()}produces trace plots for the
-    interpolation points in Figure \ref{fig:toy-interp}
+    \ref{fig:toy-search}.
+  \item
+    \code{explore\_trace\_interp()} produces trace plots for the
+    interpolation points in Figure \ref{fig:toy-interp}.
   \item
     \code{explore\_space\_pca()} produces the PCA plot of projection
-    bases on the reduced space in Figure \ref{fig:toy-pca}. Animated
+    bases on the reduced space in Figure \ref{fig:toy-pca}. The animated
     version in Figure \ref{fig:toy-pca-animated} can be turned on via
@@
 
-Three algorithms (Creeping random search, simulated annealing, and
-pseudo derivative) used in projection pursuit guided tour optimisation.
+Three algorithms (creeping random search, simulated annealing, and
+pseudo-derivative) used in projection pursuit guided tour optimisation.
 
@@
 %
-%
+Melbourne, Australia%
 %
@@
 %
-%
+Vienna, Austria%
 %

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.Rmd
settle down all the issues w. Ursula; into one script

@@
 ```{r external, include = FALSE, cache = FALSE}
-library(knitr)
-library(tidyverse)
-library(ferrn)
-
 knitr::read_chunk(
-  here::here("scripts/toy.R")
+  here::here("paper", "zhang-cook-laa-langrene-menendez.R")
 )
-
-knitr::read_chunk(
-  here::here("scripts/diagnose.R")
-)
-
-knitr::read_chunk(
-  here::here("scripts/kol_result.R")
-)
-
 ```
@@
 As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the rendering of tour animation.
+In addition, some objective functions rely on the empirical distribution of the data, which makes it in general not possible to get the gradient.
 Hence, gradient-based methods are not the focus of this paper and consideration will be given to derivative-free methods.
@@
 
-```{r toy-pca, fig.height = 5, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 A polish search takes the final basis of an optimiser as the start of a new guided tour to search for local improvements.
-The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space, the percentage change in the index value, and neighbourhood size:
+The polish algorithm is similar to the CRS but with three distinctions: 1) a hundred rather than one candidate bases are generated each time in the inner loop; 2) the neighbourhood size is reduced in the inner loop, rather than in the outer loop; and 3) three more termination conditions have been added to ensure the new basis generated is distinguishable from the current one in terms of the distance in the space, the relative change in the index value, and neighbourhood size:
 
 1) the distance between the basis found and the current basis needs to be larger than 1e-3;
-2) the percentage change of the index value needs to be larger than 1e-5; and
+2) the relative change of the index value needs to be larger than 1e-5; and
 3) the alpha parameter on itself needs to be larger than 0.01

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.tex
settle down all the issues w. Ursula; into one script

@@
 required to perform differentiation on a matrix could impede the
-rendering of tour animation. Hence, gradient-based methods are not the
+rendering of tour animation. In addition, some objective functions rely
+on the empirical distribution of the data, which makes it in general not
+possible to get the gradient. Hence, gradient-based methods are not the
 focus of this paper and consideration will be given to derivative-free
@@
 have been added to ensure the new basis generated is distinguishable
-from the current one in terms of the distance in the space, the
-percentage change in the index value, and neighbourhood size:
+from the current one in terms of the distance in the space, the relative
+change in the index value, and neighbourhood size:
 
@@
 \item
-  the percentage change of the index value needs to be larger than 1e-5;
+  the relative change of the index value needs to be larger than 1e-5;
   and

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.R

@@ -1,82 +1,543 @@
-## ----setup, include=FALSE-----------------------------------------------------
-knitr::opts_chunk$set(echo = FALSE, 
-                      message = FALSE, 
-                      warning = FALSE,
-                      root.dir = here::here(), 
-                      fig.path = here::here("figs/"), 
-                      cache = TRUE,
-                      fig.align = "center",
-                      fig.height = 3,
-                      fig.retina = 5,
-                      out.width = "100%")
-
-
-## ----external, include = FALSE, cache = FALSE---------------------------------
-library(knitr)
-library(tidyverse)
+## ---- load-pkg
 library(ferrn)
+library(tidyverse)
+library(patchwork)
+library(tourr)
+files <- paste0("data/", list.files(here::here("data")))
+purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
-knitr::read_chunk(
-  here::here("scripts/toy.R")
-)
+## ----tour-path
+knitr::include_graphics(here::here("figs/tour-path.png"))
 
-knitr::read_chunk(
-  here::here("scripts/diagnose.R")
-)
+## ---- toy-search
+p1 <- holes_2d_better_max_tries %>%
+  mutate(max_tries = 500) %>%
+  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
+  scale_color_continuous_botanical()
 
-knitr::read_chunk(
-  here::here("scripts/kol_result.R")
-)
+p2 <- holes_2d_better_random %>%
+  mutate(max_tries = 25) %>%
+  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
+  scale_color_continuous_botanical()
 
+(p1 | p2) & theme_bw() + theme(legend.position = "none")
 
+## ---- toy-interp
+p1 <- holes_2d_better_max_tries %>%
+  mutate(group = "Algorithm 1") %>%
+  explore_trace_interp(accuracy_x = 4) +
+  scale_color_continuous_botanical()
 
-## ----load-pkg-----------------------------------------------------------------
+p2 <- holes_2d_better_random %>%
+  mutate(group = "Algorithm 3") %>%
+  explore_trace_interp( accuracy_x = 14) +
+  scale_color_continuous_botanical()
 
+p1 | p2
 
-## -----------------------------------------------------------------------------
[…]

### 2021-03-27 scripts/clean_interrupt.R

@@ -1,40 +1,0 @@
-origin <- interrupt_no
-
-clean_before <- origin %>%
-  select(basis, index_val, info, method, tries, loop, id) %>%
-  mutate( id = row_number(),
-          loop = ifelse(id == 1, 1, loop),
-          method = ifelse(id == 1, "search_better", method),
-          tries = ifelse(id == 1, tries, tries + 1)) %>%
-  group_by(tries, info) %>%
-  mutate(loop = ifelse(is.na(loop), row_number(), loop)) %>%
-  ungroup()
-
-last <- clean_before %>% filter(info == "new_basis") %>% mutate(info = "interpolation")
-
-before <- bind_rows(clean_before, last) %>%
-  arrange(tries) %>%
-  mutate(search = ifelse(info == "interpolation", "interp", "search")) %>%
-  group_by(tries, search) %>%
-  mutate(loop = row_number()) %>%
-  ungroup() %>%
-  mutate(id = row_number()) %>%
-  dplyr::select(-search)
-
-origin2 <- interrupt_yes
-
-after <- origin2 %>%
-  select(basis, index_val, info, method, tries, loop, id) %>%
-  mutate( id = row_number(),
-          loop = ifelse(id == 1, 1, loop),
-          method = ifelse(id == 1, "search_better", method),
-          tries = ifelse(id == 1, tries, tries + 1),
-          search = ifelse(info == "interpolation", "interp", "search")) %>%
-  group_by(tries, search) %>%
-  mutate(loop = row_number()) %>%
-  ungroup() %>%
-  dplyr::select(-search)
-
-
-save(before, file = here::here("data", "before.rda"))
-save(after, file = here::here("data", "after.rda"))

### 2021-03-27 scripts/diagnose.R

@@ -1,189 +1,0 @@
-## ----interruption
-interp <- before %>% get_interp()
-
-p1_anno <- bind_rows(
-  interp %>% filter(tries == 4, info == "interpolation") %>%
-    filter(index_val == max(index_val)) %>%
-    mutate(anno = "interpolation basis"),
-  interp %>% get_interp_last %>% filter(tries %in% c(3, 4)) %>%
-    mutate(anno = c("current basis", "target basis"))
-) %>% arrange(id)
-
-p1 <- before %>%
-  explore_trace_interp(accuracy_y = 0.001) +
-  scale_color_continuous_botanical() +
-  geom_hline(data = get_best(after) %>% mutate(id = 78), aes(yintercept = index_val), color = "grey90") +
-  ggrepel::geom_label_repel(data = p1_anno, aes(label = anno), box.padding = 0.5, alpha = 0.5) +
-  ggtitle("without interruption")
-
-p2 <- after %>%
-  explore_trace_interp(accuracy_y = 0.001) +
-
-  ggtitle("with interruption") +
-  scale_color_continuous_botanical()
-
-(p1 | p2)
-
-## ----polish
-# set.seed(123456)
-# holes_2d_better <-
-#   animate_xy(boa6, tour_path = guided_tour(holes(), d = 2,
-#                                            search_f =  search_better, max.tries = 400),
-#              rescale = FALSE)
-# last_basis <- get_best(holes_2d_better)$basis %>% .[[1]]
-# set.seed(123456)
-# holes_2d_better_polish <-
-#   animate_xy(boa6, tour_path = guided_tour(holes(), d = 2,
-#                                            search_f =  search_polish),
-#              rescale = FALSE, start = last_basis)
-#
-# save(holes_2d_better, file = here::here("data", "holes_2d_better.rda"))
-# save(holes_2d_better_polish, file = here::here("data", "holes_2d_better_polish.rda"))
-#
-# set.seed(123456)
-# render(
-#   boa6,
-#   tour_path = guided_tour(holes(), d = 2, search_f = search_better, max.tries = 400),
-#   dev = "png",
-#   display = display_xy(axes = "off", verbose = TRUE, col = botanical_palettes$fern[[1]]),
-#   rescale = FALSE,
-#   frames = 00,
-#   file = here::here("anim","polish", "before%03d.png")
-# )
-# last_basis <- get_best(holes_2d_better)$basis %>% .[[1]]
-#
[…]

### 2021-03-27 scripts/kol_result.R

@@ -1,36 +1,0 @@
-## ----kol-result
-clean_pca <- clean %>%
-  filter(method != "search_polish") %>%
-  bind_theoretical(matrix = matrix(c(0, 1, 0, 0, 0, 0), nrow = 6),
-                   index = tourr::norm_kol(nrow(boa6)),
-                   raw_data = boa6) %>%
-  bind_theoretical(matrix = matrix(c(0, 0, 1, 0, 0, 0), nrow = 6),
-                   index = tourr::norm_kol(nrow(boa6)),
-                   raw_data = boa6) %>%
-  mutate(seed_sim = paste0(seed, sim),
-         alpha = ifelse(str_detect(sim, "tuned"), 0.7, 0.5),
-         optimiser = ifelse(str_detect(sim, "random"), "SA", "CRS"),
-         var_found = as.factor(ifelse(var_found == "V7", "global", "local")),
-         var_found = fct_relevel(var_found, "global", after = 1)) %>%
-  compute_pca(group = seed_sim) %>%
-  pluck("aug")
-
-global_max <- clean_pca %>%
-  filter(info == "theoretical") %>%
-  filter(index_val == max(index_val)) %>%
-  select(-alpha, -optimiser)
-
-local_max <- clean_pca %>%
-  filter(info == "theoretical") %>%
-  filter(index_val == min(index_val)) %>%
-  select(-alpha, -optimiser)
-
-clean_pca %>% drop_na() %>% rename(`search size` = alpha) %>%
-  explore_space_pca(pca = FALSE, group = seed_sim, color = var_found,
-                    start_size = 1, start_alpha = 0.5, end_size = 2,
-                    interp_size = 0.5) +
-  add_theo(dt = global_max, theo_size = 20) +
-  add_theo(dt = local_max, theo_label = "x", theo_size = 10) +
-  scale_color_discrete_botanical() +
-  facet_grid(`search size` ~ optimiser, labeller = label_both)
-

### 2021-03-27 scripts/kol_sim.R

@@ -1,76 +1,0 @@
-compute_kol_sim <- function(optim_data, polish_data, search_f, alpha = 0.5, max.tries = 200){
-  #browser()
-  set.seed(123)
-  seed <- sample.int(10000, 20)
-  optim_data <- list()
-  polish_data <- list()
-
-  for (i in 1: length(seed)){
-    cat("i = ", i, "seed = ", seed[i], "\n")
-    set.seed(seed[i])
-    optim_data[[i]] <-
-      animate_dist(boa6,
-                   tour_path = guided_tour(norm_kol(nrow(boa6)), d = 1,
-                                           search_f = search_f, max.tries = max.tries, alpha),
-                   rescale = FALSE, verbose = TRUE) %>%
-      mutate(seed = seed[i])
-
-    last_basis <- get_best(optim_data[[i]])$basis %>% .[[1]]
-
-    set.seed(seed[i])
-    polish_data[[i]] <-
-      animate_dist(boa6,
-                   tour_path = guided_tour(norm_kol(nrow(boa6)), d= 1,
-                                           search_f = search_polish, max.tries = 500),
-                   rescale = FALSE, verbose = TRUE,
-                   start = last_basis) %>%
-      mutate(seed = seed[i])
-  }
-
-  list(optim_data = optim_data, polish_data = polish_data)
-}
-
-kol_better <- compute_kol_sim(kol_better_optim, kol_better_polish, search_f = search_better)
-kol_better_tuned <- compute_kol_sim(kol_better_optim_tuned, kol_better_polish_tuned, search_f = search_better, alpha = 0.7)
-kol_random <- compute_kol_sim(kol_random_optim, kol_random_polish, search_f = search_better_random)
-kol_random_tuned <- compute_kol_sim(kol_random_optim_tuned, kol_random_polish_tuned, search_f = search_better_random, alpha = 0.7)
-
-# clean up the four results
-sim_clean_up <- function(sim_data) {
-  sim_name <- rep(names(sim_data), each = 20)
-
-  optim_tidy <- purrr::map(sim_data,  ~.x$optim_data) %>%
-    unlist(recursive = FALSE) %>%
-    map2(sim_name, ~.x %>% mutate(sim = .y)) %>%
-    bind_rows()
-
-  polish_tidy <- purrr::map(sim_data,  ~.x$polish_data) %>%
-    unlist(recursive = FALSE) %>%
-    map2(sim_name, ~.x %>% mutate(sim = .y)) %>%
[…]

### 2021-03-27 scripts/toy.R

@@ -1,197 +1,0 @@
-## ---- load-pkg
-library(ferrn)
-library(tidyverse)
-library(patchwork)
-library(tourr)
-files <- paste0("data/", list.files(here::here("data")))
-purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
-
-## ----tour-path
-include_graphics(here::here("figs/tour-path.png"))
-
-## ---- toy-search
-p1 <- holes_2d_better_max_tries %>%
-  mutate(max_tries = 500) %>%
-  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
-  scale_color_continuous_botanical()
-
-p2 <- holes_2d_better_random %>%
-  mutate(max_tries = 25) %>%
-  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
-  scale_color_continuous_botanical()
-
-(p1 | p2) & theme_bw() + theme(legend.position = "none")
-
-## ---- toy-interp
-p1 <- holes_2d_better_max_tries %>%
-  mutate(group = "Algorithm 1") %>%
-  explore_trace_interp(accuracy_x = 4) +
-  scale_color_continuous_botanical()
-
-p2 <- holes_2d_better_random %>%
-  mutate(group = "Algorithm 3") %>%
-  explore_trace_interp( accuracy_x = 14) +
-  scale_color_continuous_botanical()
-
-p1 | p2
-
-## ----toy-pca
-# bind_rows(holes_1d_geo, holes_1d_better) %>%
-#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-#                    index = tourr::holes(), raw_data = boa5)  %>%
-#   explore_space_pca(group = method, details = TRUE,
-#                     interp_size = 1) +
-#   scale_color_discrete_botanical()
-knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
-
-
-## ----toy-pca-animated
-# dt <- dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%
-#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-#                    index = tourr::holes(), raw_data = boa5)
-# ani <- dt %>%
-#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
-#                     theo_size = 45, start_size = 10, end_size = 20) +
[…]

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.Rmd
Minor fixes in manuscript

@@
 Index functions can be noisy, might have multiple local maxima as well as an optimal maximum, and are constrained to generate orthonormal projection frames, which complicates the optimization.
-In addition, projection pursuit is primarily used for exploratory data analysis, and finding the local maxima is also useful.
+In addition, projection pursuit is primarily used for exploratory data analysis.
 The guided tour is especially useful for exploration, because it conducts geodesic interpolation connecting steps in the optimisation and shows how the projected data changes as a maxima is approached.
@@
 A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
-The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
-An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms applications is shown using simulated data.
+The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation process, from which several diagnostic plots are created.
+An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to existing algorithms is shown using simulated data.
 Finally, an explanation of the implementation in the R package, \CRANpkg{ferrn} [@ferrn], is provided.
@@
 Although these methods are popular, they rely on the availability of the objective function derivatives.
-As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix and the computational time required to perform differentiation on a matrix could impede the rendering of tour animation.
+As will be shown in the next section, the independent variables in our optimisation problem are the entries of a projection matrix, and the computational time required to perform differentiation on a matrix could impede the rendering of tour animation.
 In addition, some objective functions rely on the empirical distribution of the data, which makes it in general not possible to get the gradient.
@@
 3) While finding the global optimum is the goal of the optimisation problem, interesting projections may also appear in the local optimum.
-4) The optimisation should be fast to compute since the tour animation is viewed by the users during the optimisation.
+4) The optimisation should be fast to compute since the tour animation is viewed by the users during the optimisation process.
 
@@
 Three optimisers have been implemented in the \pkg{tourr} [@tourr] package: creeping random search (CRS), simulated annealing (SA), and pseudo-derivative (PD).
-Creeping random search (CRS) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
+Creeping random search (CRS) is a random search optimiser that samples a candidate basis $\mathbf{A}_{l}$ in the neighbourhood of the current basis $\mathbf{A}_{\text{cur}}$ by $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ where $\alpha \in [0,1]$ controls the radius of the sampling neighbourhood and $\mathbf{A}_{\text{rand}}$ is generated randomly.
 $\mathbf{A}_{l}$ is then orthonormalised to fulfil the basis constraint.
@@
 @posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
-In his proposal, the $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
+In his proposal $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
 
 Simulated annealing (SA) uses the same sampling process as CRS but allows a probabilistic acceptance of a basis with lower index value than the current one.
-Given an initial value of $T_0$, the "temperature" at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)}$ denotes the index value of a given basis.
+Given an initial value of $T_0 \in \mathbb{R^{+}}$, the "temperature" at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
+When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)} \in \mathbb{R}$ denotes the index value of a given basis.
 This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enables the optimiser to jump out of a local optimum.
@@
 Rather than randomly sample the basis space, PD first computes a search direction by evaluating bases close to the current basis.
-A step size is then chosen along the corresponding geodesic by another optimisation over an 90 degree angle from $-\pi/4$ to $\pi/4$.
+A step size is then chosen along the corresponding geodesic by another optimisation over a 90 degree angle from $-\pi/4$ to $\pi/4$.
 The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.tex
Minor fixes in manuscript

@@
 % !TeX root = RJwrapper.tex
-\title{Visual Diagnostics for Constrained Optimisation with Application
-to Guided Tours}
-\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
-Langrené, and Patricia Menéndez}
+\title{Visual Diagnostics for Constrained Optimisation with Application to
+Guided Tours}
+\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, and Patricia Menéndez}
 
@@
 optimization. In addition, projection pursuit is primarily used for
-exploratory data analysis, and finding the local maxima is also useful.
-The guided tour is especially useful for exploration, because it
-conducts geodesic interpolation connecting steps in the optimisation and
-shows how the projected data changes as a maxima is approached. This
-work provides new visual diagnostics for examining a choice of
-optimisation procedure, based on the provision of a new data object
-which collects information throughout the optimisation. It has helped to
-diagnose and fix several problems with projection pursuit guided tour.
-This work might be useful more broadly for diagnosing optimisers, and
-comparing their performance. The diagnostics are implemented in the R
-package \pkg{ferrn}.
+exploratory data analysis. The guided tour is especially useful for
+exploration, because it conducts geodesic interpolation connecting steps
+in the optimisation and shows how the projected data changes as a maxima
+is approached. This work provides new visual diagnostics for examining a
+choice of optimisation procedure, based on the provision of a new data
+object which collects information throughout the optimisation. It has
+helped to diagnose and fix several problems with projection pursuit
+guided tour. This work might be useful more broadly for diagnosing
+optimisers, and comparing their performance. The diagnostics are
+implemented in the R package \pkg{ferrn}.
 }
@@
 diagnostics, including the design of a data structure to capture
-information during the optimisation, from which several diagnostic plots
-are created. An illustration of how the diagnostic plots can be used to
-examine the performance of different optimisers and guide improvements
-to the existing algorithms applications is shown using simulated data.
+information during the optimisation process, from which several
+diagnostic plots are created. An illustration of how the diagnostic
+plots can be used to examine the performance of different optimisers and
+guide improvements to existing algorithms is shown using simulated data.
 Finally, an explanation of the implementation in the R package,
@@
 the next section, the independent variables in our optimisation problem
-are the entries of a projection matrix and the computational time
+are the entries of a projection matrix, and the computational time
 required to perform differentiation on a matrix could impede the
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/tour-path}
 
@@
 local optimum. 4) The optimisation should be fast to compute since the
-tour animation is viewed by the users during the optimisation.
+tour animation is viewed by the users during the optimisation process.
 
@@
 \(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
-where \(\alpha\) controls the radius of the sampling neighbourhood and
-\(\mathbf{A}_{\text{rand}}\) is generated randomly. \(\mathbf{A}_{l}\)
-is then orthonormalised to fulfil the basis constraint. If
-\(\mathbf{A}_{l}\) has an index value higher than the current basis
-\(\mathbf{A}_{\text{cur}}\), the optimiser outputs \(\mathbf{A}_{l}\)
-for guided tour to construct an interpolation path. The neighbourhood
-parameter \(\alpha\) is adjusted by a cooling parameter:
-\(\alpha_{j+1} = \alpha_j * \text{cooling}\) before the next iteration
-starts. The optimiser terminates when the maximum number of iteration
-\(l_{\max}\) is reached before a better basis can be found. The
-algorithm of CRS is summarised in Algorithm \ref{random-search} in the
-appendix. \citet{posse1995projection} has proposed a slightly different
-cooling scheme by introducing a halving parameter \(c\). In his
-proposal, the \(\alpha\) is only adjusted if the last iteration takes
+where \(\alpha \in [0,1]\) controls the radius of the sampling
+neighbourhood and \(\mathbf{A}_{\text{rand}}\) is generated randomly.
+\(\mathbf{A}_{l}\) is then orthonormalised to fulfil the basis
+constraint. If \(\mathbf{A}_{l}\) has an index value higher than the
+current basis \(\mathbf{A}_{\text{cur}}\), the optimiser outputs
+\(\mathbf{A}_{l}\) for guided tour to construct an interpolation path.
+The neighbourhood parameter \(\alpha\) is adjusted by a cooling
+parameter: \(\alpha_{j+1} = \alpha_j * \text{cooling}\) before the next
+iteration starts. The optimiser terminates when the maximum number of
+iteration \(l_{\max}\) is reached before a better basis can be found.
+The algorithm of CRS is summarised in Algorithm \ref{random-search} in
+the appendix. \citet{posse1995projection} has proposed a slightly
+different cooling scheme by introducing a halving parameter \(c\). In
+his proposal \(\alpha\) is only adjusted if the last iteration takes
 more than \(c\) times to find a better basis.
@@
 allows a probabilistic acceptance of a basis with lower index value than
-the current one. Given an initial value of \(T_0\), the ``temperature''
-at iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\).
-When a candidate basis fails to have an index value larger than the
-current basis, SA gives it a second chance to be accepted with
-probability
+the current one. Given an initial value of \(T_0 \in \mathbb{R^{+}}\),
+the ``temperature'' at iteration \(l\) is defined as
+\(T(l) = \frac{T_0}{\log(l + 1)}\). When a candidate basis fails to have
+an index value larger than the current basis, SA gives it a second
+chance to be accepted with probability
 \[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
-where \(I_{(\cdot)}\) denotes the index value of a given basis. This
-implementation allows the optimiser to make a move and explore the basis
-space even if the candidate basis does not have a higher index value and
-hence enables the optimiser to jump out of a local optimum. The
-algorithm \ref{simulated-annealing} in the appendix highlights how SA
-differs from CRS in the inner loop.
+where \(I_{(\cdot)} \in \mathbb{R}\) denotes the index value of a given
+basis. This implementation allows the optimiser to make a move and
+explore the basis space even if the candidate basis does not have a
+higher index value and hence enables the optimiser to jump out of a
+local optimum. The algorithm \ref{simulated-annealing} in the appendix
+highlights how SA differs from CRS in the inner loop.
 
@@
 then chosen along the corresponding geodesic by another optimisation
-over an 90 degree angle from \(-\pi/4\) to \(\pi/4\). The resulting
+over a 90 degree angle from \(-\pi/4\) to \(\pi/4\). The resulting
 candidate basis \(\mathbf{A}_{**}\) is returned for the current
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/flip-sign-1}
 

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.Rmd
Final touch done

@@
 2) the relative change of the index value needs to be larger than 1e-5; and
-3) the alpha parameter on itself needs to be larger than 0.01
+3) the alpha parameter needs to be larger than 0.01
 
@@
 The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indexes An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
-Let $F_{.}(n)$ be the ECDF function, with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index, $I^{nk}(n)$, is defined as:
+Let $F_{.}(n)$ be the ECDF function (empirical cummulative distribution function), with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index $I^{nk}(n)$, is defined as:
 
-$$I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]$$
+$$I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right].$$
 With a non-smooth index function, two research questions are raised:

### 2021-03-27 paper/zhang-cook-laa-langrene-menendez.tex
Final touch done

@@
 \item
-  the alpha parameter on itself needs to be larger than 0.01
+  the alpha parameter needs to be larger than 0.01
 \end{enumerate}
@@
 normal distribution, \(\mathcal{N}_{n \times 1}\), using the Kolmogorov
-test. Let \(F_{.}(n)\) be the ECDF function, with two possible
-subscripts \(Y\) and \(\mathcal{N}\) representing the projected and
-randomly generated data, and \(n\) denoting the number of observation,
-the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:
-
-\[I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
+test. Let \(F_{.}(n)\) be the ECDF function (empirical cummulative
+distribution function), with two possible subscripts \(Y\) and
+\(\mathcal{N}\) representing the projected and randomly generated data,
+and \(n\) denoting the number of observation, the Normal Kolmogorov
+index \(I^{nk}(n)\), is defined as:
+
+\[I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right].\] With a
 non-smooth index function, two research questions are raised:

### 2021-03-28 cover-letter/cover-letter.Rmd
cover letter and last change

---
from: H. Sherry Zhang
fromqualifications: BComm (Honours)
fromtitle: PhD student
fromWWW: huizezhangsh.netlify.app
fromemail: Huize.Zhang@monash.edu
fromphone: +61 4 6854 1397
fromsignature: sig.png
toaddress: R Journal
opening: Dear Editor
closing: Regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our manuscript "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours" for publication in the R Journal.

The manuscript uses visual diagnostic plots to discover deficiencies in the optimisation algorithms of projection pursuit guided tour. We introduce a workflow of collecting data from the optimisation algorithms and using it to produce visual diagnostic plots with a new R package "ferrn".

In addition, our diagnostic plots provide visual interpretation to the algorithms diagnosed and allow others to understand how different components in the algorithms contribute to the final optimum found. This could inspire software developers to draw more attention to algorithm transparency when designing and implementing new algorithms.

We believe that this manuscript is a good fit to the R Journal.

Thank you for considering this manuscript.

### 2021-03-28 paper/zhang-cook-laa-langrene-menendez.Rmd
cover letter and last change

@@
 Index functions can be noisy, might have multiple local maxima as well as an optimal maximum, and are constrained to generate orthonormal projection frames, which complicates the optimization.
-In addition, projection pursuit is primarily used for exploratory data analysis.
+In addition, projection pursuit is primarily used for exploratory data analysis, and finding the local maxima is also useful.
 The guided tour is especially useful for exploration, because it conducts geodesic interpolation connecting steps in the optimisation and shows how the projected data changes as a maxima is approached.
@@
 A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
-The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation process, from which several diagnostic plots are created.
+The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
 An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to existing algorithms is shown using simulated data.
@@
 3) While finding the global optimum is the goal of the optimisation problem, interesting projections may also appear in the local optimum.
-4) The optimisation should be fast to compute since the tour animation is viewed by the users during the optimisation process.
+4) The optimisation should be fast to compute since the tour animation is viewed by the users during the optimisation.
 

### 2021-03-28 paper/zhang-cook-laa-langrene-menendez.tex
cover letter and last change

@@
 optimization. In addition, projection pursuit is primarily used for
-exploratory data analysis. The guided tour is especially useful for
-exploration, because it conducts geodesic interpolation connecting steps
-in the optimisation and shows how the projected data changes as a maxima
-is approached. This work provides new visual diagnostics for examining a
-choice of optimisation procedure, based on the provision of a new data
-object which collects information throughout the optimisation. It has
-helped to diagnose and fix several problems with projection pursuit
-guided tour. This work might be useful more broadly for diagnosing
-optimisers, and comparing their performance. The diagnostics are
-implemented in the R package \pkg{ferrn}.
+exploratory data analysis, and finding the local maxima is also useful.
+The guided tour is especially useful for exploration, because it
+conducts geodesic interpolation connecting steps in the optimisation and
+shows how the projected data changes as a maxima is approached. This
+work provides new visual diagnostics for examining a choice of
+optimisation procedure, based on the provision of a new data object
+which collects information throughout the optimisation. It has helped to
+diagnose and fix several problems with projection pursuit guided tour.
+This work might be useful more broadly for diagnosing optimisers, and
+comparing their performance. The diagnostics are implemented in the R
+package \pkg{ferrn}.
 }
@@
 diagnostics, including the design of a data structure to capture
-information during the optimisation process, from which several
-diagnostic plots are created. An illustration of how the diagnostic
-plots can be used to examine the performance of different optimisers and
-guide improvements to existing algorithms is shown using simulated data.
-Finally, an explanation of the implementation in the R package,
-\CRANpkg{ferrn} \citep{ferrn}, is provided.
+information during the optimisation, from which several diagnostic plots
+are created. An illustration of how the diagnostic plots can be used to
+examine the performance of different optimisers and guide improvements
+to existing algorithms is shown using simulated data. Finally, an
+explanation of the implementation in the R package, \CRANpkg{ferrn}
+\citep{ferrn}, is provided.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/tour-path}
 
@@
 local optimum. 4) The optimisation should be fast to compute since the
-tour animation is viewed by the users during the optimisation process.
+tour animation is viewed by the users during the optimisation.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Desktop/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
 

### 2021-08-22 review/response.Rmd
Draft response to reviewers

---
title: "Response"
author: "H. Sherry Zhang"
date: "22/08/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reviewer 1

## The Review

This work presents visual diagnostic tools for different methods of projection pursuit optimization. They use line plots, PCA and the tour method to visually compare the behavior of different optimizers. Finally, they showcased the analysis of three optimizers using their tool.

The tool is likely to be of interest to the R community.  Most of my comments are about the technical aspect of the content. The writing is overall good, but please see my comment under the "writing" regarding to the clarity of the text.

### Major comment

- "Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign...However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin. A sign flipping step is conducted..." and the "Reconciling the orientation" section
  - For 1D projections, flipping the sign is sufficient. For p-D projections of $\mathbb{R}^n$​, one may need a little more. To my knowledge most index functions are invariant to any rotations that happens within the p-subspace. This makes me wonder that in general, one may want to align two bases so that their difference does not contain any "within-subspace" components. In other words, one may want to compare representations in the Grassmannian manifold  (space of k-subspaces) instead of in the Stiefel manifold (space of orthonormal k-frames).

\textcolor{red}{Is this a comment or anything to change?}

### Minor comment

- The software seems to be specific to projection pursuit. I wish the authors could comment more on its potential generality to other optimizations. For example, can one generalizes the tour method to visualize any parameter trace of an optimizer?

**It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated. In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package. **

- Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
  - This is related to the major comment. In the figure, PD flipped in sign but stopped at a basis representation similar to the destination of the CRS. Flipping the sign back, does it mean that PD and CRS diverged but stopped at the similar subspace with opposite basis representations? Also, given that PD and CRS have the same starting bases and one is flipped, why wouldn't the dashed line pass the origin (gray dot) in the figure?

\textcolor{red}{The grey dot is the the zero matrix after projecting by the PCA. }

- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus"
  - It is not clear to me why "so". For p=2, the space of 2x2 bases is a pair of 1-D tori instead of just one (because you have two orientations for a pair of orthogonal vectors), right? For p=3, the space of 3x2 bases is SO(3), which is not a 3-D torus? Ref: https://en.wikipedia.org/wiki/Stiefel_manifold

\textcolor{red}{TODO}

### Writing

- "Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated by PCA. The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."
  - It took me multiple passes to get what this means. Perhaps instead starting with something like: "We use PCA to project and visualize the parameters/bases in 2D. The centre of the 2D view is..."?

**Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis. The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."**

- "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
  - I think this is the first mention of "5D space"/5D data in the main text (apart from the caption in Figure 4), so it may be worth clarifying what the original 5D space is.

\textcolor{red}{The description of data is in the subsection simulation setup in the next section.}

- "While the previous few examples have looked at the space of 1D bases ...  2D bases"
  - I don't think the terms "1D bases" and "2D bases" are standard. Perhaps (orthonormal) "1-frames" and "2-frames"?

**Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."**

# Reviewer 2

These experiments are interesting and so is the methodology.
The conclusions could do with beefing up
What can we learn about how to optimise the holes index for 1 and 2d?
Does this apply to other smooth indices? What about non-smooth indices?
What about indices in other dimensions?

\textcolor{red}{Is the reviewer suggesting adding more details in the first paragraph of the conclusion?}

1. Should refer to earlier work on scagnostic indices, page 3

\textcolor{red}{my understanding is that although scagnostics were introduced much earlier, using it as an index function for tourr is a recent implemntation.}

2. figure 2 caption "using the holes index"  "and a boxplot". What is the colour scale for the points? The text related to figure 2 in the bottom of page 5 should mention the dataset and the index, giving references if needed.

**Fixed language. Added mention the color scale in the figure 2 caption:  "The color scale is from the customised botanical palette in the ferrn package."**

3. "with higher index value"  -> "with a higher index value". The word "the" is missing in a few other places in the paper.

**corrected.**

4. In Figure 3 would be better if the y-axes scales were the same for the two graphs.

**fixed**

5. What is the star in figures 4 and 5? This is not mentioned in the text or caption

**Added "The black star is the theoretical best basis the optimisers aiming to find." in the caption of figure 4.**

6. In the text you say things like "details=TRUE" and "animate=TRUE" but what functions are you referring to?

**Mention of "details=TRUE" and "animate=TRUE" is removed from subsection diagnostic 3a and 3b to the implementation section: `explore_space_pca()` produces the PCA plot of projection bases on the reduced space. Figure 4 includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`. The animated version in Figure 5 is produced with argument `animate = TRUE`.**

7. In Figure 9, it seems CRS is doing better here than in Figure 8-- maybe this could be mentioned.

**Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing. Cross comparing the two experiments is less meaningful.**

8. "not the case for all the indexes" missing a "." page 11

**added**

9. For the Kolmogorov index, why are you using randomly generated data instead of normal quantiles?

\textcolor{red}{TODO}

10. Fix references with "et al"

**fixed**

### 2021-08-22 paper/zhang-cook-laa-langrene-menendez.Rmd
changes made as per the reviewers

@@
 
-```{r toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index.
+```{r toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index.
 Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
-In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and boxplot is used to summarise the distribution of the index values.
-There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey."}
+In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values.
+There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
+The color scale is from the customised botanical palette in the ferrn package."}
 ```
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reaching the target bases in time 76-130."}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."}
 ```
@@
 Figure \ref{fig:toy-interp} presents the trace plot of the same optimisations as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
-It may seem bizarre at first sight that the interpolation sometimes passes bases with higher index value before it decreases to a lower target.
+It may seem bizarre at first sight that the interpolation sometimes passes bases with higher index values before it decreases to a lower target.
 This happens because, on the one hand, the probabilistic acceptance in SA implies that some worse bases will be accepted by the optimiser.
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+The black star is the theoretical best basis the optimisers aiming to find.
 All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
@@
 The importance of these bases differs and the most important ones are the starting, interpolating, and end bases.
-The anchor and search bases can be turned on with argument `details = TRUE`.
 Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
@@
 As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
-Along with the bases recorded during the optimisation and a zero basis, the parameters (centre and radius) of the 2D space can be estimated by PCA.
-The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
+We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
+The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
 The theoretical best basis is known for simulated data and can be labelled to compare how close the end basis found by the optimisers is to the theoretical best.
@@
 An animation is another type of display to show how the search progresses from start to finish in the space.
-An  `animate = TRUE` argument is used to enable the animation in the PCA plot.
 Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
@@
 
-While the previous few examples have looked at the space of 1D bases in a unit sphere, this section visualises the space of 2D bases.
+While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames.
 Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a $p$-D torus.
@@
 
-An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with higher index value than the target one.
+An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with a higher index value than the target one.
 For SA, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for CRS.
@@
 ```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS.
-On the left panel, basis with higher index value is found during the interpolation but not used.
-On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
+On the left panel, basis with a higher index value is found during the interpolation but not used.
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel."}
 ```
@@
 
-The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indexes An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
+The holes index function used for all the examples before this section produces a smooth interpolation, while this is not the case for all the indexes.
+An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
 Let $F_{.}(n)$ be the ECDF function (empirical cummulative distribution function), with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index $I^{nk}(n)$, is defined as:
@@
 The animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
+   - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space.
+Figure \ref{fig:toy-pca} includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
+The animated version in Figure \ref{fig:toy-pca-animated} is produced with argument `animate = TRUE`.
   - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.

### 2021-08-22 paper/zhang-cook-laa-langrene-menendez.tex
changes made as per the reviewers

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-search-1}
-
-}
-
-\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index.
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-search-1}
+
+}
+
+\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using the holes index.
 Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
-In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and boxplot is used to summarise the distribution of the index values.
-There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.}\label{fig:toy-search}
+In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values.
+There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
+The color scale is from the customised botanical palette in the ferrn package.}\label{fig:toy-search}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-interp-1}
 
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index value before reaching the target bases in time 76-130.}\label{fig:toy-interp}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 first sight that the interpolation sometimes passes bases with higher
-index value before it decreases to a lower target. This happens because,
-on the one hand, the probabilistic acceptance in SA implies that some
-worse bases will be accepted by the optimiser. In addition, the guided
-tour interpolates between the current and target basis to provide a
-smooth transition between projections, and sometimes a higher index
-value will be observed along the interpolation path. This indicates that
-a non-monotonic interpolation cannot be avoided, even for CRS. Later, in
-section \ref{monotonic}, there will be a discussion on improving the
-non-monotonic interpolation for CRS.
+index values before it decreases to a lower target. This happens
+because, on the one hand, the probabilistic acceptance in SA implies
+that some worse bases will be accepted by the optimiser. In addition,
+the guided tour interpolates between the current and target basis to
+provide a smooth transition between projections, and sometimes a higher
+index value will be observed along the interpolation path. This
+indicates that a non-monotonic interpolation cannot be avoided, even for
+CRS. Later, in section \ref{monotonic}, there will be a discussion on
+improving the non-monotonic interpolation for CRS.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca}
 
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+The black star is the theoretical best basis the optimisers aiming to find.
 All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
@@
 these bases differs and the most important ones are the starting,
-interpolating, and end bases. The anchor and search bases can be turned
-on with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers
-can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to the orientation of
-the basis and so is the index value. However, this creates difficulties
-for comparing the optimisers since the end bases will be symmetric to
-the origin. A sign flipping step is conducted to flip the signs of all
-the bases in one routine if different optimisations finish at opposite
-places.
+interpolating, and end bases. Sometimes, two optimisers can start with
+the same basis but finish with bases of opposite sign. This happens
+because the projection is invariant to the orientation of the basis and
+so is the index value. However, this creates difficulties for comparing
+the optimisers since the end bases will be symmetric to the origin. A
+sign flipping step is conducted to flip the signs of all the bases in
+one routine if different optimisations finish at opposite places.
 
@@
 sphere and random bases on the sphere can be generated via the
-\CRANpkg{geozoo} \citep{geozoo} package. Along with the bases recorded
-during the optimisation and a zero basis, the parameters (centre and
-radius) of the 2D space can be estimated by PCA. The centre of the
-circle is the first two PC coordinates of the zero matrix, and the
-radius is estimated as the largest distance from the centre to the
-basis. The theoretical best basis is known for simulated data and can be
-labelled to compare how close the end basis found by the optimisers is
-to the theoretical best. Various aesthetics, i.e.~size, alpha
-(transparency), and colour, are applicable to emphasize critical
+\CRANpkg{geozoo} \citep{geozoo} package. We use PCA to estimate the 2D
+space that contains all the bases recorded and the zero basis. The
+centre of the 2D view is the first two PC coordinates of the zero
+matrix, and the radius is estimated as the largest distance from the
+centre to the basis. The theoretical best basis is known for simulated
+data and can be labelled to compare how close the end basis found by the
+optimisers is to the theoretical best. Various aesthetics, i.e.~size,
+alpha (transparency), and colour, are applicable to emphasize critical
 elements and adjust for the presentation. For example, anchor points and
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 An animation is another type of display to show how the search
-progresses from start to finish in the space. An
-\texttt{animate\ =\ TRUE} argument is used to enable the animation in
-the PCA plot. Figure \ref{fig:toy-pca-animated} shows six frames from
-the animation of the PCA plot in Figure \ref{fig:toy-pca}. An additional
-piece of information one can learn from this animation is that CRS finds
-its end basis quicker than PD since CRS finishes its search in the 5th
-frame while PD is still making more progress.
+progresses from start to finish in the space. Figure
+\ref{fig:toy-pca-animated} shows six frames from the animation of the
+PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
+one can learn from this animation is that CRS finds its end basis
+quicker than PD since CRS finishes its search in the 5th frame while PD
+is still making more progress.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-While the previous few examples have looked at the space of 1D bases in
-a unit sphere, this section visualises the space of 2D bases. Recall
+While the previous few examples have looked at the space of 1-frames in
+a unit sphere, this section visualises the space of 2-frames. Recall
 that the columns in a 2D basis are orthogonal to each other, so the
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-torus-1}
 
@@
 An example of non-monotonic interpolation has been given in Figure
-\ref{fig:toy-interp}: a path that passes bases with higher index value
+\ref{fig:toy-interp}: a path that passes bases with a higher index value
 than the target one. For SA, a non-monotonic interpolation is justified
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/interruption-1}
 
@@
 \caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS.
-On the left panel, basis with higher index value is found during the interpolation but not used.
-On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
+On the left panel, basis with a higher index value is found during the interpolation but not used.
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/polish-1}
 
@@
 produces a smooth interpolation, while this is not the case for all the
-indexes An example of a noisy index function for 1D projections compares
-the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated
-normal distribution, \(\mathcal{N}_{n \times 1}\), using the Kolmogorov
-test. Let \(F_{.}(n)\) be the ECDF function (empirical cummulative
-distribution function), with two possible subscripts \(Y\) and
-\(\mathcal{N}\) representing the projected and randomly generated data,
-and \(n\) denoting the number of observation, the Normal Kolmogorov
-index \(I^{nk}(n)\), is defined as:
+indexes. An example of a noisy index function for 1D projections
+compares the projected data, \(\mathbf{Y}_{n \times 1}\), to a randomly
+generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
+Kolmogorov test. Let \(F_{.}(n)\) be the ECDF function (empirical
+cummulative distribution function), with two possible subscripts \(Y\)
+and \(\mathcal{N}\) representing the projected and randomly generated
+data, and \(n\) denoting the number of observation, the Normal
+Kolmogorov index \(I^{nk}(n)\), is defined as:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/flip-sign-1}
 
@@
   \item
+    \code{explore\_space\_pca()} produces the PCA plot of projection
+    bases on the reduced space. Figure \ref{fig:toy-pca} includes the
+    additional details of anchor and search bases, which can be turned
+    on by the argument \texttt{details\ =\ TRUE}. The animated version
+    in Figure \ref{fig:toy-pca-animated} is produced with argument
+    \texttt{animate\ =\ TRUE}.
+  \item
     \code{explore\_space\_tour()} produces animated tour view on the
@@
 %
-Melbourne, Australia%
+Melbourne, Australia\\%
 %
-\\\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
+\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}%
 }
@@
 %
-Melbourne, Australia%
+Melbourne, Australia\\%
 %
-\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
+\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
 }
@@
 %
-Vienna, Austria%
+Vienna, Austria\\%
 %
-\\\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
+\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}%
 }
@@
 %
-Melbourne, Australia%
+Melbourne, Australia\\%
 %
-\\\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
+\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}%
 }
@@
 %
-Melbourne, Australia%
+Melbourne, Australia\\%
 %
-\\\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}
-}
+\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}%
+}

### 2021-08-22 paper/zhang-cook-laa-langrene-menendez.R

@@ -1,543 +1,69 @@
-## ---- load-pkg
-library(ferrn)
-library(tidyverse)
-library(patchwork)
-library(tourr)
-files <- paste0("data/", list.files(here::here("data")))
-purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
-
-## ----tour-path
-knitr::include_graphics(here::here("figs/tour-path.png"))
-
-## ---- toy-search
-p1 <- holes_2d_better_max_tries %>%
-  mutate(max_tries = 500) %>%
-  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
-  scale_color_continuous_botanical()
-
-p2 <- holes_2d_better_random %>%
-  mutate(max_tries = 25) %>%
-  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
-  scale_color_continuous_botanical()
-
-(p1 | p2) & theme_bw() + theme(legend.position = "none")
-
-## ---- toy-interp
-p1 <- holes_2d_better_max_tries %>%
-  mutate(group = "Algorithm 1") %>%
-  explore_trace_interp(accuracy_x = 4) +
-  scale_color_continuous_botanical()
-
-p2 <- holes_2d_better_random %>%
-  mutate(group = "Algorithm 3") %>%
-  explore_trace_interp( accuracy_x = 14) +
-  scale_color_continuous_botanical()
-
-p1 | p2
-
-## ----toy-pca
-# p <- bind_rows(holes_1d_geo, holes_1d_better) %>%
-#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
-#                    index = tourr::holes(), raw_data = boa5)  %>%
-#   explore_space_pca(group = method, details = TRUE,
-#                     interp_size = 1) +
-#   scale_color_discrete_botanical()
-#
-# ggsave(p, filename = "toy-pca.pdf",
-#        path = here::here("figs"),
-#        width = 4, height = 4)
-knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
+## ----setup, include=FALSE-----------------------------------------------------
+knitr::opts_chunk$set(echo = FALSE, 
+                      message = FALSE, 
+                      warning = FALSE,
+                      root.dir = here::here(), 
+                      fig.path = here::here("figs/"), 
[…]

### 2021-08-25 review/response.Rmd
Thoughts on major comment

@@
 \textcolor{red}{Is this a comment or anything to change?}
+I think the reviewer want us to comment on the sign flipping for higher order dimensions.
+It seems that the reviewer is suggesting that instead of comparing the bases on the space of the orthonormal k-frames it might be good to compare their representation in the Grassmannian manifold.
+
+Grassmannian manifold: GR9k,V) is a space that parametrize all k-dimension a linear subspaces of the n-dimensional vector sace V.(Wikpedia source)
 

### 2021-08-25 review/response.Rmd
Adding thought about reviewers comments

@@
 
+Probably here the reviewer would like to see a kind of "recipie" about how to use the framework of the paper for other optimisers?
+
 - Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
@@
 
+It might be worth to mentioned about the 5D in the simulation sep up subsection so that it is introduced explicitly for the readers.
+
 - "While the previous few examples have looked at the space of 1D bases ...  2D bases"
@@
 **Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."**
+
+And the frames are orthonormal, right?
 
@@
 \textcolor{red}{Is the reviewer suggesting adding more details in the first paragraph of the conclusion?}
+
+I think the reviewer here wants to see a bit more detail on the conclusions and have provided a number of questions that could be answered in order to add some interesting points on that section.
 
@@
 
+"The black star **represents** the theoretical ...
+
 6. In the text you say things like "details=TRUE" and "animate=TRUE" but what functions are you referring to?
@@
 
+The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data.
+I am not sure what would be the reasoning for using normal quantiles in this index.
+That would imply you are assuming normality on the theoretial pdf.
+
 10. Fix references with "et al"

### 2021-09-02 review/response.Rmd
more changes as per the reviewers

@@
 
+**ANS: With p-D bases, the correction of orientation is a bit more complicated.
+Take 2D as an example, two bases can be aligned in the first axis but not in the second.
+This would means there are $2^p$ possibility to check.
+**
+
 ### Minor comment
@@
 
-\textcolor{red}{The grey dot is the the zero matrix after projecting by the PCA. }
+**Yes the reviewer is giving a decent explanation in describing the effect of flipping.
+The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
+This has now been corrected.
+**
 
-- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus"
+- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus".
+
   - It is not clear to me why "so".
@@
 
-\textcolor{red}{TODO}
+- We change the wording of the sentence to "**a torus in the $p$-D space**"
 
@@
 
-\textcolor{red}{The description of data is in the subsection simulation setup in the next section.}
+**The explanation of data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
+To avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".
+**
 
@@
 
-\textcolor{red}{TODO}
-
-The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data.
+**The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data.
 I am not sure what would be the reasoning for using normal quantiles in this index.
-That would imply you are assuming normality on the theoretial pdf.
+That would imply you are assuming normality on the theoretial pdf.**
 

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.Rmd
more changes as per the reviewers

@@
 knitr::read_chunk(
-  here::here("paper", "zhang-cook-laa-langrene-menendez.R")
+  here::here("paper", "script.R")
 )
@@
 While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original 5D space.
+Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original space.
 
@@
 While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames.
-Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a $p$-D torus.
+Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space.
 For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.tex
more changes as per the reviewers

@@
 that the columns in a 2D basis are orthogonal to each other, so the
-space of \(p \times 2\) bases is a \(p\)-D torus. For \(p = 3\) one
-would see a classical 3D torus shape as shown by the grey points in
-Figure \ref{fig:toy-torus}. The two circles of the torus can be observed
-to be perpendicular to each other and this can be linked back to the
-orthogonality condition. Two paths from CRS and PD are plotted on top of
-the torus and coloured in green and brown respectively to match the
-previous plots. The final basis found by PD and CRS are shown in a
-larger shape and printed below respectively:
+space of \(p \times 2\) bases is a torus in the \(p\)-D space. For
+\(p = 3\) one would see a classical 3D torus shape as shown by the grey
+points in Figure \ref{fig:toy-torus}. The two circles of the torus can
+be observed to be perpendicular to each other and this can be linked
+back to the orthogonality condition. Two paths from CRS and PD are
+plotted on top of the torus and coloured in green and brown respectively
+to match the previous plots. The final basis found by PD and CRS are
+shown in a larger shape and printed below respectively:
 

### 2021-09-02 paper/script.R

@@ -1,0 +1,543 @@
+## ---- load-pkg
+library(ferrn)
+library(tidyverse)
+library(patchwork)
+library(tourr)
+files <- paste0("data/", list.files(here::here("data")))
+purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
+
+## ----tour-path
+knitr::include_graphics(here::here("figs/tour-path.png"))
+
+## ---- toy-search
+p1 <- holes_2d_better_max_tries %>%
+  mutate(max_tries = 500) %>%
+  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
+  scale_color_continuous_botanical()
+
+p2 <- holes_2d_better_random %>%
+  mutate(max_tries = 25) %>%
+  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
+  scale_color_continuous_botanical()
+
+(p1 | p2) & theme_bw() + theme(legend.position = "none")
+
+## ---- toy-interp
+p1 <- holes_2d_better_max_tries %>%
+  mutate(group = "Algorithm 1") %>%
+  explore_trace_interp(accuracy_x = 4) +
+  scale_color_continuous_botanical()
+
+p2 <- holes_2d_better_random %>%
+  mutate(group = "Algorithm 3") %>%
+  explore_trace_interp( accuracy_x = 14) +
+  scale_color_continuous_botanical()
+
+p1 | p2
+
+## ----toy-pca
+# p <- bind_rows(holes_1d_geo, holes_1d_better) %>%
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                    index = tourr::holes(), raw_data = boa5)  %>%
+#   explore_space_pca(group = method, details = TRUE,
+#                     interp_size = 1) +
+#   scale_color_discrete_botanical()
+#
+# ggsave(p, filename = "toy-pca.pdf",
+#        path = here::here("figs"),
+#        width = 4, height = 4)
+knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
+
+
+## ----toy-pca-animated
+# dt <- dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%
+#   bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
+#                    index = tourr::holes(), raw_data = boa5)
[…]

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.R

@@ -13,7 +13,7 @@ knitr::opts_chunk$set(echo = FALSE,
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------
 knitr::read_chunk(
-  here::here("paper", "zhang-cook-laa-langrene-menendez.R")
+  here::here("paper", "script.R")
 )
 
 

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.Rmd
Added scagnostics reference.

@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures.
+There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag,@WW08].
 

### 2021-09-02 paper/zhang-cook-laa-langrene-menendez.tex
Added scagnostics reference.

@@
 summarising structures in scatterplot matrices based on eight scagnostic
-measures.
+measures \citep[\citet{WW08}]{scag}.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-search-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-tour-1}
 
@@
 Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
-same two optimisations in its original 5D space.
+same two optimisations in its original space.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/noisy-better-geo-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/kol-result-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/hzha400/Documents/PhD/research/zzz-done/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/flip-sign-1}
 
@@
 %
-Melbourne, Australia\\%
+Melbourne, Australia%
 %
-\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}%
+\\\href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}
 }
@@
 %
-Melbourne, Australia\\%
+Melbourne, Australia%
 %
-\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}%
+\\\href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}
 }
@@
 %
-Vienna, Austria\\%
+Vienna, Austria%
 %
-\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}%
+\\\href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}
 }
@@
 %
-Melbourne, Australia\\%
+Melbourne, Australia%
 %
-\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}%
+\\\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
 }
@@
 %
-Melbourne, Australia\\%
+Melbourne, Australia%
 %
-\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}%
-}
+\\\href{mailto:patricia.menendez@monash.edu}{\nolinkurl{patricia.menendez@monash.edu}}
+}

### 2021-09-03 paper/zhang-cook-laa-langrene-menendez.tex
update plot aesthetics

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-search-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-search}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/noisy-better-geo-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/kol-result-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/kol-result}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/ursulalaa/papers/paper-tour-vis/figs/flip-sign-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/flip-sign}
 

### 2021-09-03 paper/script.R

@@ -10,17 +10,23 @@ purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 knitr::include_graphics(here::here("figs/tour-path.png"))
 
 ## ---- toy-search
-p1 <- holes_2d_better_max_tries %>%
-  mutate(max_tries = 500) %>%
-  explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.93) +
-  scale_color_continuous_botanical()
+# p1 <- holes_2d_better_max_tries %>%
+#   mutate(max_tries = 500) %>%
+#   explore_trace_search(label.size = 0.03,label.padding = 0.1, segment.size = 0,  extend_lower = 0.99) +
+#   scale_color_continuous_botanical()
+#
+# p2 <- holes_2d_better_random %>%
+#   mutate(max_tries = 25) %>%
+#   explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.99) +
+#   scale_color_continuous_botanical()
+#
+# p <- (p1 | p2) + plot_layout(widths = c(1.5,2)) & theme_bw() & theme(legend.position = "none")
+#
+# ggsave(p, filename = "toy-search.pdf",
+#        path = here::here("figs"), device = "pdf",
+#        width = 10, height = 5)
 
-p2 <- holes_2d_better_random %>%
-  mutate(max_tries = 25) %>%
-  explore_trace_search(label.size = 0.01, label.padding = 0.1, segment.size = 0, extend_lower = 0.93) +
-  scale_color_continuous_botanical()
-
-(p1 | p2) & theme_bw() + theme(legend.position = "none")
+knitr::include_graphics(here::here("figs/toy-search.pdf"))
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
@@ -33,7 +39,7 @@ p2 <- holes_2d_better_random %>%
   explore_trace_interp( accuracy_x = 14) +
   scale_color_continuous_botanical()
 
-p1 | p2
+(p1 | p2) & ylim(0.8, 0.96)
 
 ## ----toy-pca
 # p <- bind_rows(holes_1d_geo, holes_1d_better) %>%
@@ -41,11 +47,11 @@ p1 | p2
 #                    index = tourr::holes(), raw_data = boa5)  %>%
 #   explore_space_pca(group = method, details = TRUE,
 #                     interp_size = 1) +
[…]

### 2021-09-03 paper/zhang-cook-laa-langrene-menendez.Rmd
try new web article format

@@
     email:  patricia.menendez@monash.edu 
-abstract: A guided tour helps to visualise high-dimensional data by showing low-dimensional projections along a projection pursuit optimisation path.
+abstract: >
+  A guided tour helps to visualise high-dimensional data by showing low-dimensional projections along a projection pursuit optimisation path.
 Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
@@
   \usepackage[ruled, linesnumbered]{algorithm2e}
-output: rticles::rjournal_article
+output: 
+  rjtools::rjournal_web_article
+bibliography: zhang-cook-laa-langrene-menendez.bib
 ---
@@
 
-```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
-The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
-Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."}
-```
-
 Another interesting feature to examine is the changes in the index value between interpolating bases since the projection on these bases is shown in the tour animation.
@@
 
+```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
+The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
+Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."}
+```
+
 ## Diagnostic 3a: Understanding the optimiser's coverage of the search space {#toy-pca}
-
-```{r toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
-The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star is the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
-```
 
@@
 
+```{r toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+The black star is the theoretical best basis the optimisers aiming to find.
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
+```
+
 ## Diagnostic 3b: Animating the diagnostic plots
+
+An animation is another type of display to show how the search progresses from start to finish in the space.
+Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-pca-animated-interactive}', '\\ref{fig:toy-pca-animated}'))` shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
+An additional piece of information one can learn from this animation is that CRS finds its end basis quicker than PD since CRS finishes its search in the 5th frame while PD is still making more progress.
 
@@
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
-The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
+The full video of the animation can be found at \\url{https://vimeo.com/504242845}.", eval=knitr::is_latex_output()}
 ```
 
-An animation is another type of display to show how the search progresses from start to finish in the space.
-Figure \ref{fig:toy-pca-animated} shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
-An additional piece of information one can learn from this animation is that CRS finds its end basis quicker than PD since CRS finishes its search in the 5th frame while PD is still making more progress.
+```{r toy-pca-animated-interactive, fig.height = 7, fig.width = 7, fig.cap = "this is the interactive version", eval=knitr::is_html_output()}
+```
 
@@
 
+As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \pkg{geozoo} [@geozoo] package.
+While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.
+Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-torus-interactive}', '\\ref{fig:toy-torus}'))` shows some frames from the tour plot of the same two optimisations in its original space.
+
 ```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
-The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \\CRANpkg{geozoo}.
+The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
 The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
 ```
-
-As mentioned previously, the original $p \times d$ dimension space can be simulated via randomly generated bases in the \pkg{geozoo} [@geozoo] package.
-While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows some frames from the tour plot of the same two optimisations in its original space.
 
@@
 One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and are symmetric when projected onto the other (due the sign difference in the second column).
-In Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
+In Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-torus-interactive}', '\\ref{fig:toy-torus}'))`, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
 
@@
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found at \\url{https://vimeo.com/512882784}."}
+The full video of the animation can be found at \\url{https://vimeo.com/512882784}.", eval = knitr::is_latex_output()}
+```
+
+```{r toy-torus-interactive, out.height = "10%", out.width = "10%", fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
+The projection problem is a 2D projection with three variables using the holes index.
+The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
+The full video of the animation can be found at \\url{https://vimeo.com/512882784}.", eval = knitr::is_html_output()}
 ```
@@
 
-```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index.
-The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis."}
-```
-
 Once the final basis has been found by an optimiser, one may want to push further in the close neighbourhood to see if an even better basis can be found.
@@
 Polish works with this end basis and further pushes the index value to produce clearer edges of the cluster, especially along the vertical axis.
+
+```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index.
+The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis."}
+```
 
@@
 
-```{r noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
-Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best."}
-```
-
 Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated paths are smooth.
@@
 
-```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projection index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
+```{r noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}.
+Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best."}
 ```
@@
 
+```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
+```
+
 ## Reconciling the orientation
+
+One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
+As mentioned previously, bases with opposite signs do not affect the projection and index value, but clearly, we prefer the target to have the same orientation as the current basis.
+The orientation of two bases can be checked via calculating the determinant and a negative value indicates opposite direction.
+Hence an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
+The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exists.
 
@@
 ```
-
-One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
-As mentioned previously, bases with opposite signs do not affect the projection and index value, but clearly, we prefer the target to have the same orientation as the current basis.
-The orientation of two bases can be checked via calculating the determinant and a negative value indicates opposite direction.
-Hence an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
-The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exists.
 
@@
 
-\bibliography{zhang-cook-laa-langrene-menendez}
-
 \newpage

### 2021-09-03 paper/script.R

@@ -22,11 +22,11 @@ knitr::include_graphics(here::here("figs/tour-path.png"))
 #
 # p <- (p1 | p2) + plot_layout(widths = c(1.5,2)) & theme_bw() & theme(legend.position = "none")
 #
-# ggsave(p, filename = "toy-search.pdf",
-#        path = here::here("figs"), device = "pdf",
+# ggsave(p, filename = "toy-search.svg",
+#        path = here::here("figs"), device = "svg",
 #        width = 10, height = 5)
 
-knitr::include_graphics(here::here("figs/toy-search.pdf"))
+knitr::include_graphics(here::here("figs/toy-search.svg"))
 
 ## ---- toy-interp
 p1 <- holes_2d_better_max_tries %>%
@@ -50,9 +50,9 @@ p2 <- holes_2d_better_random %>%
 #   scale_color_discrete_botanical() +
 #   theme(legend.text = element_text(size = "10pt"))
 #
-# ggsave(p, filename = "toy-pca.pdf",
+# ggsave(p, filename = "toy-pca.svg",
 #        path = here::here("figs"))
-knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
+knitr::include_graphics(here::here("figs", "toy-pca.svg"))
 
 
 ## ----toy-pca-animated
@@ -68,18 +68,21 @@ knitr::include_graphics(here::here("figs", "toy-pca.pdf"))
 #         renderer = file_renderer("anim/pca/",
 #                                  prefix = "pca", overwrite = TRUE))
 
-# vimeo version with legend
-# ani <- dt %>%
-#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
-#                     theo_size = 45, start_size = 10, end_size = 20) +
-#   scale_color_botanical(palette = "fern")
-# anim_save(ani, filename = here::here("anim/toy-pca-animated.mp4"))
-
 frames <- c("0001", "0038", "0046", "0079", "0086", "0100")
 ani <- paste0(here::here("anim/"), "pca/", "pca", frames, ".png")
 rl <- lapply(ani, png::readPNG)
 gl <-  lapply(rl, grid::rasterGrob)
 wrap_plots(gl)
+
+
+## ----toy-pca-animated-interactive
+# ani <- dt %>%
+#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
[…]

### 2021-09-05 review/response.Rmd
finish response to reviewer

--- 2021-08-25 review/response.Rmd
@@
 
-\textcolor{red}{Is this a comment or anything to change?}
-I think the reviewer want us to comment on the sign flipping for higher order dimensions.
-It seems that the reviewer is suggesting that instead of comparing the bases on the space of the orthonormal k-frames it might be good to compare their representation in the Grassmannian manifold.
-
-Grassmannian manifold: GR9k,V) is a space that parametrize all k-dimension a linear subspaces of the n-dimensional vector sace V.(Wikpedia source)
+\textcolor{red}{With p-D bases, the correction of orientation is a bit more complicated.
+Take 2D as an example, two bases can be aligned in the first axis but not in the second.
+This would means there are $2^p$ possibility to check.}
 
@@
 
-**It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated.
+\textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated.
 In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package.
-**
+For other optimisers, reserachers need first to create a data collection object in a tibble that track the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.}
 
@@
 
-\textcolor{red}{The grey dot is the the zero matrix after projecting by the PCA. }
+\textcolor{red}{Yes the reviewer is giving a decent explanation in describing the effect of flipping.
+The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
+This has now been corrected.}
 
-- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus"
+- "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus".
+
   - It is not clear to me why "so".
@@
 
-\textcolor{red}{TODO}
+\textcolor{red}{We change the wording of the sentence to "**a torus in the $p$-D space**"}
 
@@
 
-**Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
-The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."**
+\textcolor{red}{Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
+The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}
 
@@
 
-\textcolor{red}{The description of data is in the subsection simulation setup in the next section.}
+\textcolor{red}{The explanation of data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
+To avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
 
@@
 
-**Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."**
+\textcolor{red}{Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
 
@@
 
-\textcolor{red}{Is the reviewer suggesting adding more details in the first paragraph of the conclusion?}
+\textcolor{red}{A paragraph summarising the finding is added as the second paragram in the conclusion}
 
@@
 
-\textcolor{red}{my understanding is that although scagnostics were introduced much earlier, using it as an index function for tourr is a recent implemntation.}
+\textcolor{red}{Added reference of (Wilkinson et al., 2005) and  (Wilkinson and Wills, 2008):}
+
+\textcolor{red}{L. Wilkinson and G. Wills. Scagnostics distributions. Journal of Computational and Graphical Statistics, 17 (2):473–491, 2008}
+
+\textcolor{red}{L. Wilkinson, A. Anand, and R. Grossman. Graph-theoretic scagnostics. In IEEE Symposium on Information Visualization, 2005. INFOVIS 2005., pages 157–164, Oct 2005.}
 
@@
 
-**Fixed language. Added mention the color scale in the figure 2 caption:  "The color scale is from the customised botanical palette in the ferrn package."**
+\textcolor{red}{Fixed language. Added mention the color scale in the figure 2 caption:  "The color scale is from the customised botanical palette in the ferrn package."}
 
@@
 
-**corrected.**
+\textcolor{red}{corrected.}
 
@@
 
-**fixed**
+\textcolor{red}{fixed}
 
@@
 
-**Added "The black star is the theoretical best basis the optimisers aiming to find." in the caption of figure 4.**
+\textcolor{red}{Added "The black star represents the theoretical best basis the optimisers aiming to find." in the caption of figure 4.}
 
@@
 
-**Mention of "details=TRUE" and "animate=TRUE" is removed from subsection diagnostic 3a and 3b to the implementation section: `explore_space_pca()` produces the PCA plot of projection bases on the reduced space.
+\textcolor{red}{Mention of "details=TRUE" and "animate=TRUE" is removed from subsection diagnostic 3a and 3b to the implementation section: `explore\_space\_pca()` produces the PCA plot of projection bases on the reduced space.
 Figure 4 includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
-The animated version in Figure 5 is produced with argument `animate = TRUE`.**
+The animated version in Figure 5 is produced with argument `animate = TRUE`.}
 
@@
 
-**Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing.
-Cross comparing the two experiments is less meaningful.**
+\textcolor{red}{Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing.
+Cross comparing the two experiments is less meaningful.}
 
@@
 
-**added**
+\textcolor{red}{added}
 
@@
 
-\textcolor{red}{TODO}
+\textcolor{red}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data.
+I am not sure what would be the reasoning for using normal quantiles in this index.
+That would imply you are assuming normality on the theoretial pdf.}
 
@@
 
-**fixed**
+\textcolor{red}{fixed}

### 2021-09-05 paper/zhang-cook-laa-langrene-menendez.Rmd
tweak gif for html rendering

@@
 This work might be useful more broadly for diagnosing optimisers, and comparing their performance.
-The diagnostics are implemented in the R package \pkg{ferrn}.
+The diagnostics are implemented in the R package [ferrn](https://github.com/huizezhang-sherry/ferrn).
 keywords: graphics, multivariate, optimization
@@
 An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag,@WW08].
+There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag; @WW08].
 
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star is the theoretical best basis the optimisers aiming to find.
+The black star represents the theoretical best basis the optimisers aiming to find.
 All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
@@
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
-The full video of the animation can be found at \\url{https://vimeo.com/504242845}.", eval=knitr::is_latex_output()}
+The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()}
 ```
 
-```{r toy-pca-animated-interactive, fig.height = 7, fig.width = 7, fig.cap = "this is the interactive version", eval=knitr::is_html_output()}
+```{r toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\ref{fig:toy-pca}.
+With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
 ```
@@
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
-The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
+The full animation can be seen in the html version of the paper."}
+```
+
+```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated} in the full basis space.
+The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo."}
 ```
@@
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found at \\url{https://vimeo.com/512882784}.", eval = knitr::is_latex_output()}
+The full video of the animation can be found in the html version of the paper.", eval = knitr::is_latex_output()}
 ```
 
-```{r toy-torus-interactive, out.height = "10%", out.width = "10%", fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
+```{r toy-torus-interactive, fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using the holes index.
-The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found at \\url{https://vimeo.com/512882784}.", eval = knitr::is_html_output()}
+The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.", eval = knitr::is_html_output()}
 ```
@@
 
-  - \code{explore\_trace\_search()} produces summary plots in Figure \ref{fig:toy-search}.
-  - \code{explore\_trace\_interp()} produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}.
-  - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
+  - `explore_trace_search()` produces summary plots in Figure \ref{fig:toy-search}.
+  - `explore_trace_interp()` produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}.
+  - `explore_space_pca()` produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
 The animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
-   - \code{explore\_space\_pca()} produces the PCA plot of projection bases on the reduced space.
+   - `explore_space_pca()` produces the PCA plot of projection bases on the reduced space.
 Figure \ref{fig:toy-pca} includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
 The animated version in Figure \ref{fig:toy-pca-animated} is produced with argument `animate = TRUE`.
-  - \code{explore\_space\_tour()} produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
+  - `explore_space_tour()` produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
 
-- \code{get\_*()} extracts and manipulates certain components from the existing data object.
-  - \code{get\_anchor()} extracts target observations 
-  - \code{get\_basis\_matrix()} flattens all the bases into a matrix
-  - \code{get\_best()} extracts the observation with the highest index value in the data object
-  - \code{get\_dir\_search()} extracts directional search observations for PD search
-  - \code{get\_interp()} extracts interpolated observations
-  - \code{get\_interp\_last()} extracts the ending interpolated observations in each iteration
-  - \code{get\_interrupt()} extracts the ending interpolated observations and the target observations if  the interpolation is interrupted
-  - \code{get\_search()} extracts search observations 
-  - \code{get\_search\_count()} extracts the count of search observations 
-  - \code{get\_space\_param()} produces the coordinates of the centre and radius of the basis space
-  - \code{get\_start()} extracts the starting observation
-  - \code{get\_theo()} extracts the theoretical best observations, if given
+- `get_*()` extracts and manipulates certain components from the existing data object.
+  - `get_anchor()` extracts target observations 
+  - `get_basis_matrix()` flattens all the bases into a matrix
+  - `get_best()` extracts the observation with the highest index value in the data object
+  - `get_dir_search()` extracts directional search observations for PD search
+  - `get_interp()` extracts interpolated observations
+  - `get_interp_last()` extracts the ending interpolated observations in each iteration
+  - `get_interrupt()` extracts the ending interpolated observations and the target observations if  the interpolation is interrupted
+  - `get_search()` extracts search observations 
+  - `get_search_count()` extracts the count of search observations 
+  - `get_space_param()` produces the coordinates of the centre and radius of the basis space
+  - `get_start()` extracts the starting observation
+  - `get_theo()` extracts the theoretical best observations, if given
 
-- \code{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
-  - \code{bind\_theoretical()} binds the theoretical best observation in simulated experiment
-  - \code{bind\_random()} binds randomly generated bases in the projection bases space to the data object
-  - \code{bind\_random\_matrix()} binds randomly generated bases and outputs in a matrix format
+- `bind_*()` incorporates additional information outside the tour optimisation into the data object.
+  - `bind_theoretical()` binds the theoretical best observation in simulated experiment
+  - `bind_random()` binds randomly generated bases in the projection bases space to the data object
+  - `bind_random_matrix()` binds randomly generated bases and outputs in a matrix format
 
-- \code{add\_*()} provides wrapper functions to create ggprotos for different components for the PCA plot
-  - \code{add\_anchor()} for plotting anchor bases
-  - \code{add\_anno()} for annotating the symmetry of start bases
-  - \code{add\_dir\_search()} for plotting the directional search bases with magnified distance
-  - \code{add\_end()} for plotting end bases
-  - \code{add\_interp()} for plotting the interpolation path
-  - \code{add\_interp\_last()} for plotting the last interpolation bases for comparing with target bases when interruption is used
-  - \code{add\_interrupt()} for linking the last interpolation bases with target ones when interruption is used
-  - \code{add\_search()} for plotting search bases
-  - \code{add\_space()} for plotting the circular space 
-  - \code{add\_start()} for plotting start bases
-  - \code{add\_theo()} for plotting theoretical best bases, if applicable
+- `add_*()` provides wrapper functions to create ggprotos for different components for the PCA plot
+  - `add_anchor()` for plotting anchor bases
+  - `add_anno()` for annotating the symmetry of start bases
+  - `add_dir_search()` for plotting the directional search bases with magnified distance
+  - `add_end()` for plotting end bases
+  - `add_interp()` for plotting the interpolation path
+  - `add_interp_last()` for plotting the last interpolation bases for comparing with target bases when interruption is used
+  - `add_interrupt()` for linking the last interpolation bases with target ones when interruption is used
+  - `add_search()` for plotting search bases
+  - `add_space()` for plotting the circular space 
+  - `add_start()` for plotting start bases
+  - `add_theo()` for plotting theoretical best bases, if applicable
 
 - Utilities
-  - \code{theme\_fern()} and \code{format\_label()} for better display of the grid lines and axis formatting
-  - \code{clean\_method()} to clean up the name of the optimisers
-  - \code{botanical\_palettes} is a collection of colour palettes from Australian native plants.
-Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
-  - \code{botanical\_pal()} as the colour interpolator 
-  - \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the colour and fill of the plot
+  - `theme_fern()` and `format_label()` for better display of the grid lines and axis formatting
+  - `clean_method()` to clean up the name of the optimisers
+  - `botanical_palettes()` is a collection of colour palettes from Australian native plants. Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
+  - `botanical_pal()` as the colour interpolator 
+  - `scale_color_*()` and `scale_fill_*()` for scaling the colour and fill of the plot
 
@@
 The approach used broadly applies to other constrained optimisers.
+
+Experiments in this paper have shown that random search algorithms tend to require larger search size for 2D indexes as compared to 1D indexes.
+All the three optimisers tend to work well for smooth indexes while PD fails generally fails for non-smooth indexes.
+This is due to the fact that derivative based methods are in general not suitable for optimising non-smooth functions.
 

### 2021-09-05 paper/script.R

@@ -63,7 +63,7 @@ knitr::include_graphics(here::here("figs", "toy-pca.svg"))
 #   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
 #                     theo_size = 45, start_size = 10, end_size = 20) +
 #   theme(legend.position = "none") +
-#   scale_color_botanical(palette = "fern")
+#   scale_color_continuous_botanical(palette = "fern")
 # animate(ani, nframes = 100, device = "png",
 #         renderer = file_renderer("anim/pca/",
 #                                  prefix = "pca", overwrite = TRUE))
@@ -77,10 +77,12 @@ wrap_plots(gl)
 
 ## ----toy-pca-animated-interactive
 # ani <- dt %>%
-#   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
-#                     theo_size = 45, start_size = 10, end_size = 20) +
-#   scale_color_botanical(palette = "fern")
-# anim_save(ani, filename = here::here("anim/toy-pca-animated.mp4"))
+#   explore_space_pca(group = method, animate = TRUE, interp_size = 1,
+#                     theo_size = 20, start_size = 3, end_size = 5) +
+#   scale_color_discrete_botanical(palette = "fern")
+# anim_save(ani,
+#           filename = here::here("anim/toy-pca-animated.gif"),
+#           renderer = gifski_renderer(width = 500, height = 500))
 knitr::include_graphics(here::here("anim/toy-pca-animated.gif"))
 
 
@@ -102,7 +104,17 @@ knitr::include_graphics(here::here("anim/toy-pca-animated.gif"))
 #   here::here("anim","tour", "tour%03d.png")
 # )
 #
-# # render gif
+
+frames <- c("001", "021", "056", "072", "079", "088")
+ani <- paste0(here::here("anim/"), "tour/", "tour", frames, ".png")
+rl <- lapply(ani, png::readPNG)
+gl <-  lapply(rl, grid::rasterGrob)
+wrap_plots(gl)
+
+## ----toy-tour-interactive
+# prep <- prep_space_tour(dplyr::bind_rows(holes_1d_better, holes_1d_geo), flip = TRUE,
[…]

### 2021-09-05 paper/zhang-cook-laa-langrene-menendez.Rmd
update rendering

@@
 ```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated} in the full basis space.
-The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo."}
+The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.", eval = knitr::is_html_output()}
 ```

### 2021-09-05 paper/zhang-cook-laa-langrene-menendez.tex
update rendering

@@
 comparing their performance. The diagnostics are implemented in the R
-package \pkg{ferrn}.
+package \href{https://github.com/huizezhang-sherry/ferrn}{ferrn}.
 }
@@
 summarising structures in scatterplot matrices based on eight scagnostic
-measures \citep[\citet{WW08}]{scag}.
+measures \citep{scag, WW08}.
 
@@
 
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
-
-}
-
-\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
-The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
-Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
-\end{figure}
-\end{Schunk}
-
 Another interesting feature to examine is the changes in the index value
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
+
+}
+
+\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right).
+The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
+Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{toy-pca}{%
@@
 search space}\label{toy-pca}}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca}
-
-}
-
-\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
-The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star is the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
-\end{figure}
-\end{Schunk}
 
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca}
+
+}
+
+\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
+The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
+The black star represents the theoretical best basis the optimisers aiming to find.
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{diagnostic-3b-animating-the-diagnostic-plots}{%
@@
 plots}\label{diagnostic-3b-animating-the-diagnostic-plots}}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
-
-}
-
-\caption{Six frames selected from the animated version of the previous plot.
-With animation, the progression of the search paths from start to finish is better identified.
-CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
-The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
-\end{figure}
-\end{Schunk}
 
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
+
+}
+
+\caption[Six frames selected from the animated version of the previous plot]{Six frames selected from the animated version of the previous plot.
+With animation, the progression of the search paths from start to finish is better identified.
+CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
+The full video of the animation can be found in the html version of the paper.}\label{fig:toy-pca-animated}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{diagnostic-4a-the-tour-looking-at-itself}{%
@@
 itself}\label{diagnostic-4a-the-tour-looking-at-itself}}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
-
-}
-
-\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
-The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \CRANpkg{geozoo}.
-The full video of the animation can be found at \url{https://vimeo.com/512885379}.}\label{fig:toy-tour}
-\end{figure}
-\end{Schunk}
 
@@
 original high-dimensional space from various directions using animation.
-Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
+Figure \ref{fig:toy-torus} shows some frames from the tour plot of the
 same two optimisations in its original space.
+
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
+
+}
+
+\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
+The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
+\end{figure}
+\end{Schunk}
 
@@
 
-\caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
+\caption[Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green)]{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using the holes index.
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found at \url{https://vimeo.com/512882784}.}\label{fig:toy-torus}
+The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
 \end{figure}
@@
 
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/polish-1}
-
-}
-
-\caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
-The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.}\label{fig:polish}
-\end{figure}
-\end{Schunk}
-
 Once the final basis has been found by an optimiser, one may want to
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/polish-1}
+
+}
+
+\caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
+The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.}\label{fig:polish}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{seeing-the-signal-in-the-noise}{%
@@
 \end{enumerate}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
-
-}
-
-\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}.
-Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
-\end{figure}
-\end{Schunk}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/kol-result}
-
-}
-
-\caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
+
+}
+
+\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}.
+Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/kol-result}
+
+}
+
+\caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{reconciling-the-orientation}{%
@@
 orientation}\label{reconciling-the-orientation}}
-
-\begin{Schunk}
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/flip-sign}
-
-}
-
-\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
-Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
-The dots represent the target basis in each iteration and the path shows the interpolation.
-On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
-The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
-\end{figure}
-\end{Schunk}
 
@@
 
+\begin{Schunk}
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/flip-sign}
+
+}
+
+\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis.
+Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
+The dots represent the target basis in each iteration and the path shows the interpolation.
+On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
+The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
+\end{figure}
+\end{Schunk}
+
 \hypertarget{implementation}{%
@@
   \item
-    \code{explore\_trace\_search()} produces summary plots in Figure
+    \texttt{explore\_trace\_search()} produces summary plots in Figure
     \ref{fig:toy-search}.
   \item
-    \code{explore\_trace\_interp()} produces trace plots for the
+    \texttt{explore\_trace\_interp()} produces trace plots for the
     interpolation points in Figure \ref{fig:toy-interp}.
   \item
-    \code{explore\_space\_pca()} produces the PCA plot of projection
+    \texttt{explore\_space\_pca()} produces the PCA plot of projection
     bases on the reduced space in Figure \ref{fig:toy-pca}. The animated
@@
   \item
-    \code{explore\_space\_pca()} produces the PCA plot of projection
+    \texttt{explore\_space\_pca()} produces the PCA plot of projection
     bases on the reduced space. Figure \ref{fig:toy-pca} includes the
@@
   \item
-    \code{explore\_space\_tour()} produces animated tour view on the
+    \texttt{explore\_space\_tour()} produces animated tour view on the
     full space of the projection bases in Figure \ref{fig:toy-tour}.
@@
 \item
-  \code{get\_*()} extracts and manipulates certain components from the
+  \texttt{get\_*()} extracts and manipulates certain components from the
   existing data object.
@@
   \item
-    \code{get\_anchor()} extracts target observations
-  \item
-    \code{get\_basis\_matrix()} flattens all the bases into a matrix
-  \item
-    \code{get\_best()} extracts the observation with the highest index
+    \texttt{get\_anchor()} extracts target observations
+  \item
+    \texttt{get\_basis\_matrix()} flattens all the bases into a matrix
+  \item
+    \texttt{get\_best()} extracts the observation with the highest index
     value in the data object
   \item
-    \code{get\_dir\_search()} extracts directional search observations
+    \texttt{get\_dir\_search()} extracts directional search observations
     for PD search
   \item
-    \code{get\_interp()} extracts interpolated observations
-  \item
-    \code{get\_interp\_last()} extracts the ending interpolated
+    \texttt{get\_interp()} extracts interpolated observations
+  \item
+    \texttt{get\_interp\_last()} extracts the ending interpolated
     observations in each iteration
   \item
-    \code{get\_interrupt()} extracts the ending interpolated
+    \texttt{get\_interrupt()} extracts the ending interpolated
     observations and the target observations if the interpolation is
@@
   \item
-    \code{get\_search()} extracts search observations
-  \item
-    \code{get\_search\_count()} extracts the count of search
+    \texttt{get\_search()} extracts search observations
+  \item
+    \texttt{get\_search\_count()} extracts the count of search
     observations
   \item
-    \code{get\_space\_param()} produces the coordinates of the centre
+    \texttt{get\_space\_param()} produces the coordinates of the centre
     and radius of the basis space
   \item
-    \code{get\_start()} extracts the starting observation
-  \item
-    \code{get\_theo()} extracts the theoretical best observations, if
+    \texttt{get\_start()} extracts the starting observation
+  \item
+    \texttt{get\_theo()} extracts the theoretical best observations, if
     given
@@
 \item
-  \code{bind\_*()} incorporates additional information outside the tour
-  optimisation into the data object.
+  \texttt{bind\_*()} incorporates additional information outside the
+  tour optimisation into the data object.
 
@@
   \item
-    \code{bind\_theoretical()} binds the theoretical best observation in
-    simulated experiment
-  \item
-    \code{bind\_random()} binds randomly generated bases in the
+    \texttt{bind\_theoretical()} binds the theoretical best observation
+    in simulated experiment
+  \item
+    \texttt{bind\_random()} binds randomly generated bases in the
     projection bases space to the data object
   \item
-    \code{bind\_random\_matrix()} binds randomly generated bases and
+    \texttt{bind\_random\_matrix()} binds randomly generated bases and
     outputs in a matrix format
@@
 \item
-  \code{add\_*()} provides wrapper functions to create ggprotos for
+  \texttt{add\_*()} provides wrapper functions to create ggprotos for
   different components for the PCA plot
@@
   \item
-    \code{add\_anchor()} for plotting anchor bases
-  \item
-    \code{add\_anno()} for annotating the symmetry of start bases
-  \item
-    \code{add\_dir\_search()} for plotting the directional search bases
-    with magnified distance
-  \item
-    \code{add\_end()} for plotting end bases
-  \item
-    \code{add\_interp()} for plotting the interpolation path
-  \item
-    \code{add\_interp\_last()} for plotting the last interpolation bases
-    for comparing with target bases when interruption is used
-  \item
-    \code{add\_interrupt()} for linking the last interpolation bases
+    \texttt{add\_anchor()} for plotting anchor bases
+  \item
+    \texttt{add\_anno()} for annotating the symmetry of start bases
+  \item
+    \texttt{add\_dir\_search()} for plotting the directional search
+    bases with magnified distance
+  \item
+    \texttt{add\_end()} for plotting end bases
+  \item
+    \texttt{add\_interp()} for plotting the interpolation path
+  \item
+    \texttt{add\_interp\_last()} for plotting the last interpolation
+    bases for comparing with target bases when interruption is used
+  \item
+    \texttt{add\_interrupt()} for linking the last interpolation bases
     with target ones when interruption is used
   \item
-    \code{add\_search()} for plotting search bases
-  \item
-    \code{add\_space()} for plotting the circular space
-  \item
-    \code{add\_start()} for plotting start bases
-  \item
-    \code{add\_theo()} for plotting theoretical best bases, if
+    \texttt{add\_search()} for plotting search bases
+  \item
+    \texttt{add\_space()} for plotting the circular space
+  \item
+    \texttt{add\_start()} for plotting start bases
+  \item
+    \texttt{add\_theo()} for plotting theoretical best bases, if
     applicable
@@
   \item
-    \code{theme\_fern()} and \code{format\_label()} for better display
-    of the grid lines and axis formatting
-  \item
-    \code{clean\_method()} to clean up the name of the optimisers
-  \item
-    \code{botanical\_palettes} is a collection of colour palettes from
-    Australian native plants. Quantitative palettes include daisy,
+    \texttt{theme\_fern()} and \texttt{format\_label()} for better
+    display of the grid lines and axis formatting
+  \item
+    \texttt{clean\_method()} to clean up the name of the optimisers
+  \item
+    \texttt{botanical\_palettes()} is a collection of colour palettes
+    from Australian native plants. Quantitative palettes include daisy,
     banksia, and cherry and sequential palettes contain fern and acacia
   \item
-    \code{botanical\_pal()} as the colour interpolator
-  \item
-    \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the
-    colour and fill of the plot
+    \texttt{botanical\_pal()} as the colour interpolator
+  \item
+    \texttt{scale\_color\_*()} and \texttt{scale\_fill\_*()} for scaling
+    the colour and fill of the plot
   \end{itemize}
@@
 optimisers.
+
+Experiments in this paper have shown that random search algorithms tend
+to require larger search size for 2D indexes as compared to 1D indexes.
+All the three optimisers tend to work well for smooth indexes while PD
+fails generally fails for non-smooth indexes. This is due to the fact
+that derivative based methods are in general not suitable for optimising
+non-smooth functions.
 
@@
 \url{https://github.com/huizezhang-sherry/paper-tour-vis}.
-
-\bibliography{zhang-cook-laa-langrene-menendez}
 
@@
 
+\bibliography{zhang-cook-laa-langrene-menendez.bib}
+
 \address{%

### 2021-09-05 paper/zhang-cook-laa-langrene-menendez.R

@@ -33,13 +33,21 @@ knitr::read_chunk(
 ## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."----
 
 
-## ----toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star is the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
+## ----toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
 
 
-## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found at \\url{https://vimeo.com/504242845}."----
+## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
 
 
[…]

### 2021-09-06 review/response.Rmd
Updating answers to reviewers

@@
 ```
+\textcolor{green}{Many thanks for reviewing the manuscript and for your comments and suggestions. Please see below the answers to your review comments.}
 
@@
 
+\textcolor{green}{providing --> given that the ...?}
+
 - Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
@@
 
+\textcolor{green}{Thanks for sharing your interpretation of the effects of flipping, that's right!...}
+
 - "Recall that the columns in a 2D basis are orthogonal to each other, so the space of *p* × 2 bases is a *p*-D torus".
@@
 
-\textcolor{red}{We change the wording of the sentence to "**a torus in the $p$-D space**"}
+\textcolor{red}{We changed the wording of the sentence to "**a torus in the $p$-D space**"}
 
@@
 
-\textcolor{red}{Changed to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
+\textcolor{red}{The paragraph has now been updated to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
 The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}
+
+\textcolor{green}{Change --> The paragraph has now been updated to...}
 
@@
 
-\textcolor{red}{The explanation of data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
-To avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
+\textcolor{red}{The explanation of that data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
+However, to avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
+
+\textcolor{green}{The data set is first introduced in subsection ....}
+\textcolor{green}{However, to avoid ambiguity --> However, to avoid ambiguity}
 
@@
 
-\textcolor{red}{Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
+\textcolor{red}{Those terms have now been changed and the sentence now reads: "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
+
+\textcolor{green}{I changed to -->Those terms have now been changed and the sentence now reads:}
 
@@
 
-\textcolor{red}{A paragraph summarising the finding is added as the second paragram in the conclusion}
+\textcolor{red}{A paragraph further summarising the findings has been added in the second paragram in the conclusion}
 
 1. Should refer to earlier work on scagnostic indices, page 3
+
+\textcolor{green}{New references have been added as follows:}
 
@@
 
+\textcolor{green}{The caption has now been re-written and color scale explanation added}
+
 3. "with higher index value"  -> "with a higher index value". The word "the" is missing in a few other places in the paper.
 
-\textcolor{red}{corrected.}
+\textcolor{green}{That has now been corrected.}
 
@@
 
-\textcolor{red}{fixed}
+\textcolor{green}{It is now fixed}
 
@@
 
-\textcolor{red}{Added "The black star represents the theoretical best basis the optimisers aiming to find." in the caption of figure 4.}
+\textcolor{red}{The caption has now been ammended to: "The black star represents the theoretical best basis the optimisers aiming to find." in the caption of figure 4.}
 
@@
 
-\textcolor{red}{added}
+\textcolor{red}{It has now been added}
 
@@
 
-\textcolor{red}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on data.
-I am not sure what would be the reasoning for using normal quantiles in this index.
-That would imply you are assuming normality on the theoretial pdf.}
+\textcolor{green}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on the data.
+In our case we are not assuming normality on the theoretical pdf and that is the reason why we are not using normal quantiles.}
 
@@
 
-\textcolor{red}{fixed}
+\textcolor{red}{Fixed!}

### 2021-09-06 review/response.Rmd
Note on Kolmogorov Index notation
Fixing typo

@@
 
-\textcolor{green}{Thanks for sharing your interpretation of the effects of flipping, that's right!...}
+\textcolor{green}{Thanks for sharing your interpretation on the effects of flipping, that's right!...}
 
@@
 \textcolor{green}{The data set is first introduced in subsection ....}
-\textcolor{green}{However, to avoid ambiguity --> However, to avoid ambiguity}
+\textcolor{green}{However, to avoid ambiguity --> However, to avoid confusion}
 
@@
 
+\textcolor{purple}{I think the reviewer is confused here because in the text it is written Normal Kolmogorov.
+However, the definition is just the standard Kolmogorov Index, right?
+Shall we just changed to standard Kolmogorov Index instead?}
+
 10. Fix references with "et al"

### 2021-09-06 review/response.Rmd
Added comments to the response

@@
 
+\textcolor{magenta}{Once you go to 2D you have to take into account rotation in addition to flips, so the thing to check would be if the two frames are spanning the same subspace, and aligning them will be a bit more work.
+Might be a good idea to add 1-2 sentences in the paper and rephrase the reply.}
+
 ### Minor comment
@@
 \textcolor{green}{providing --> given that the ...?}
+
+\textcolor{magenta}{Here could reiterate points about constrained optimisation and that we simulate from the allowed parameter space, in those cases using the tour method to visualize the trace is especially interesting.
+Is there any documentation of how the data structure needs to look like for ferrn input?
+It might be a good idea to point to that.}
 
@@
 
+\textcolor{magenta}{I think Di mentioned a reference to cite when you first mention the torus in the text, it should be added to the paper and mentioned here in the reply.}
+
 ### Writing
@@
 \textcolor{green}{Change --> The paragraph has now been updated to...}
+
+\textcolor{magenta}{You could be a bit more carful with the phrasing still, the PCA finds the 2D space that is best suited to visualize the bases (while any 2D projection will contain them).}
 
@@
 
+\textcolor{magenta}{The new paragraph should be improved a bit further.
+Can you connect how the method was useful to how other people might want to use it?
+I think Fig 10 is a nice example of how we understand the optimizers both in terms of how well they climb towards higher index values, how well they cover the search space and also how direct the path towards the optimum is.
+This lead to insights for the guided tour example, and would be useful generally when comparing optimizers on difficult problems.}
+
 1. Should refer to earlier work on scagnostic indices, page 3
@@
 \textcolor{green}{The caption has now been re-written and color scale explanation added}
+
+\textcolor{magenta}{You should mention what color is indicating here, is it just a way of distinguishing iterations or does it carry any additional information?}
 
@@
 
+\textcolor{magenta}{It would still be interesting to mention what was different, i.e.
+were all the settings the same and the better performance is just random (by different random seed), or did you change some setting of the optimizer that might have helped improve the performance?}
+
 8. "not the case for all the indexes" missing a "." page 11

### 2021-09-08 review/response.Rmd
response to patricia & urtricia & ursula's comments

--- 2021-09-06 review/response.Rmd
@@
 ```
-\textcolor{green}{Many thanks for reviewing the manuscript and for your comments and suggestions. Please see below the answers to your review comments.}
+
+Many thanks for reviewing the manuscript and for your comments and suggestions. Please see below the answers to your review comments.
 
@@
 A sign flipping step is conducted..." and the "Reconciling the orientation" section
+
     -   For 1D projections, flipping the sign is sufficient.
@@
 
-\textcolor{red}{With p-D bases, the correction of orientation is a bit more complicated.
-Take 2D as an example, two bases can be aligned in the first axis but not in the second.
-This would means there are $2^p$ possibility to check.}
+\textcolor{red}{With p-D bases, the correction of orientation is a bit more complicated since both rotation and flip need to be taken into account.
+For 2D bases, one need to check if the two frames are spanning the same subspace and pointing at the same direction.}
+
+\textcolor{magenta}{Once you go to 2D you have to take into account rotation in addition to flips, so the thing to check would be if the two frames are spanning the same subspace, and aligning them will be a bit more work.
+Might be a good idea to add 1-2 sentences in the paper and rephrase the reply.}
 
@@
 
-\textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, providing the parameter space can be simulated.
-In our examples, the parameter to trace (projection base ($p \times d$) ) lives in an p-sphere, which can be simulated by the `geozoo` package.
-For other optimisers, reserachers need first to create a data collection object in a tibble that track the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.}
-
-\textcolor{green}{providing --> given that the ...?}
+\textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, given that the parameter space can be simulated.
+In our examples, the orthonormality constraint regulates the parameter space of the projection basis ($p \times d$) to be an p-sphere, which can be simulated by the `geozoo` package.
+For other optimisers, reserachers need first to create a data collection object in a tibble that tracks the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.
+Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the `expore\_\*` functions.}
 
 -   Figure 4 "...All the bases in PD have been flipped for easier comparison of the final bases..."
+
     -   This is related to the major comment.
@@
 
-\textcolor{red}{Yes the reviewer is giving a decent explanation in describing the effect of flipping.
+\textcolor{red}{Thanks for sharing your interpretation on the effects of flipping, that's right!
+The reviewer is giving a decent explanation in describing the effect of flipping.
 The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
 This has now been corrected.}
-
-\textcolor{green}{Thanks for sharing your interpretation of the effects of flipping, that's right!...}
 
@@
 For p=3, the space of 3x2 bases is SO(3), which is not a 3-D torus?
-Ref: https://en.wikipedia.org/wiki/Stiefel_manifold
+Ref: <https://en.wikipedia.org/wiki/Stiefel_manifold>
 
-\textcolor{red}{We changed the wording of the sentence to "**a torus in the $p$-D space**"}
+\textcolor{red}{We changed the wording of the sentence to: "a torus in the $p$-D space"}
+
+\textcolor{magenta}{I think Di mentioned a reference to cite when you first mention the torus in the text, it should be added to the paper and mentioned here in the reply.}
 
@@
 The centre of the circle is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."
-  - It took me multiple passes to get what this means. Perhaps instead starting with something like: "We use PCA to project and visualize the parameters/bases in 2D. The centre of the 2D view is..."?
 
-\textcolor{red}{The paragraph has now been updated to "We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
+    -   It took me multiple passes to get what this means.
+Perhaps instead starting with something like: "We use PCA to project and visualize the parameters/bases in 2D.
+The centre of the 2D view is..."?
+
+\textcolor{red}{Changed to "We use PCA to estimate the 2D space that is best suited to visualise the bases.
 The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}
 
-\textcolor{green}{Change --> The paragraph has now been updated to...}
+-   "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
 
-- "Figure 6 shows some frames from the tour plot of the same two optimisations in its original 5D space."
     -   I think this is the first mention of "5D space"/5D data in the main text (apart from the caption in Figure 4), so it may be worth clarifying what the original 5D space is.
 
-\textcolor{red}{The explanation of that data set is introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
-However, to avoid ambiguity, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
-
-\textcolor{green}{The data set is first introduced in subsection ....}
-\textcolor{green}{However, to avoid ambiguity --> However, to avoid ambiguity}
+\textcolor{red}{The data set is first introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
+However, to avoid confusion, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
 
 -   "While the previous few examples have looked at the space of 1D bases ... 2D bases"
+
     -   I don't think the terms "1D bases" and "2D bases" are standard. Perhaps (orthonormal) "1-frames" and "2-frames"?
 
-\textcolor{red}{Those terms have now been changed and the sentence now reads: "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
-
-\textcolor{green}{I changed to -->Those terms have now been changed and the sentence now reads:}
+\textcolor{red}{Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
 
@@
 These experiments are interesting and so is the methodology.
-The conclusions could do with beefing up
-What can we learn about how to optimise the holes index for 1 and 2d?
-Does this apply to other smooth indices? What about non-smooth indices?
+The conclusions could do with beefing up What can we learn about how to optimise the holes index for 1 and 2d?
+Does this apply to other smooth indices?
+What about non-smooth indices?
 What about indices in other dimensions?
@@
 
-\textcolor{green}{New references have been added as follows:}
+\textcolor{red}{New references have been added as follows:}
 
@@
 
-\textcolor{red}{Fixed language. Added mention the color scale in the figure 2 caption:  "The color scale is from the customised botanical palette in the ferrn package."}
+\textcolor{red}{The caption has now been re-written and color scale explanation added: "The color scale is from the customised botanical palette in the ferrn package."}
 
-\textcolor{green}{The caption has now been re-written and color scale explanation added}
+3.  "with higher index value" -\> "with a higher index value". The word "the" is missing in a few other places in the paper.
 
-3. "with higher index value"  -> "with a higher index value". The word "the" is missing in a few other places in the paper.
-
-\textcolor{green}{That has now been corrected.}
+\textcolor{red}{That has now been corrected.}
 
@@
 
-\textcolor{green}{It is now fixed}
+\textcolor{red}{It is now fixed}
 
@@
 \textcolor{red}{Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing.
+In the experiment of Figure 9, the parameter `max.tries` is set to 400 to allow guided tour to do its best instead of a default 25 in Figure 8.
 Cross comparing the two experiments is less meaningful.}
@@
 
-\textcolor{green}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on the data.
-In our case we are not assuming normality on the theoretical pdf and that is the reason why we are not using normal quantiles.}
+\textcolor{red}{The Kolmogorov Smirnoff index here looks at the difference between a theoretical pdf and an empirical one based on the data.
+In our case we are not assuming normality on the theoretical pdf and that is the reason why we are not using normal quantiles.
+To make it clearer, we change the wording from "Normal Kolmogorov Index" to "Kolmogorove Index".}
 

### 2021-09-08 paper/zhang-cook-laa-langrene-menendez.Rmd
response to patricia & urtricia & ursula's comments

@@
 output: 
-  rjtools::rjournal_web_article
+  rjtools::rjournal_pdf_article
 bibliography: zhang-cook-laa-langrene-menendez.bib
@@
 The unconstrained problem can be formulated as a minimization (or maximization)  problem such as  -->
+
 <!-- $\min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with certain properties defined in an $L^p$ space.
@@
 In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as -->
+
 <!-- $\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as multipliers and convex optimization methods including linear and quadratic programming can be used. -->
@@
 
+```{=tex}
 \begin{align}
@@
 \end{align}
-
+```
 \noindent where $f$ and $\mathbf{X}$ are defined as in the previous section, $\mathcal{A}$ is the set of all $p$-dimensional projection bases, $I_d$ is the $d$-dimensional identity matrix, and the constraint ensures the projection bases, $\mathbf{A}$, to be orthonormal.
@@
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
@@
 
+```{=html}
 <!-- One may want to simply plot the index value of the search points across its natural order.
@@
 -->
-
+```
 Figure \ref{fig:toy-search} shows an example of the search plot for CRS (left) and SA (right).
@@
 As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
-We use PCA to estimate the 2D space that contains all the bases recorded and the zero basis.
+We use PCA to estimate the 2D space that is best suited to visualise the bases.
 The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
@@
 \end{align}
-
+```
 ## A problem of non-monotonicity {#monotonic}
@@
 An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
-Let $F_{.}(n)$ be the ECDF function (empirical cummulative distribution function), with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Normal Kolmogorov index $I^{nk}(n)$, is defined as:
+Let $F_{.}(n)$ be the ECDF function (empirical cummulative distribution function), with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Kolmogorov index $I^{nk}(n)$, is defined as:
 
-$$I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right].$$
-With a non-smooth index function, two research questions are raised:
+$$I^{K}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right].$$ With a non-smooth index function, two research questions are raised:
 
@@
 -   `get_*()` extracts and manipulates certain components from the existing data object.
+
     -   `get_anchor()` extracts target observations
@@
 -   `bind_*()` incorporates additional information outside the tour optimisation into the data object.
+
     -   `bind_theoretical()` binds the theoretical best observation in simulated experiment
@@
 -   `add_*()` provides wrapper functions to create ggprotos for different components for the PCA plot
+
     -   `add_anchor()` for plotting anchor bases
@@
 -   Utilities
+
     -   `theme_fern()` and `format_label()` for better display of the grid lines and axis formatting
     -   `clean_method()` to clean up the name of the optimisers
-  - `botanical_palettes()` is a collection of colour palettes from Australian native plants. Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
+    -   `botanical_palettes()` is a collection of colour palettes from Australian native plants.
+Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
     -   `botanical_pal()` as the colour interpolator
@@
 Experiments in this paper have shown that random search algorithms tend to require larger search size for 2D indexes as compared to 1D indexes.
-All the three optimisers tend to work well for smooth indexes while PD fails generally fails for non-smooth indexes.
+All the three optimisers tend to work well for smooth indexes while PD generally fails for non-smooth indexes.
 This is due to the fact that derivative based methods are in general not suitable for optimising non-smooth functions.
+
+For other researchers, the tools implemented in the ferrn package can track the progression of the target function and monitor parameter coverage of the search space in both reduced 2D space and the full space.
+These visualisation can lead to insights for evaluating the performance of multiple optimisers in a specific problem, as the example shown for comparing the three optimisers when optimising the kolmogorov index in the paper.
 
@@
 
+```{=tex}
 \begin{algorithm}
@@
 \end{algorithm}
-
+```
+```{=tex}
 \begin{algorithm}
@@
 \end{algorithm}
+```

### 2021-09-08 paper/zhang-cook-laa-langrene-menendez.tex
response to patricia & urtricia & ursula's comments

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-search}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-search}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-interp-1}
 
@@
 \CRANpkg{geozoo} \citep{geozoo} package. We use PCA to estimate the 2D
-space that contains all the bases recorded and the zero basis. The
-centre of the 2D view is the first two PC coordinates of the zero
-matrix, and the radius is estimated as the largest distance from the
-centre to the basis. The theoretical best basis is known for simulated
-data and can be labelled to compare how close the end basis found by the
-optimisers is to the theoretical best. Various aesthetics, i.e.~size,
-alpha (transparency), and colour, are applicable to emphasize critical
-elements and adjust for the presentation. For example, anchor points and
-search points are less important and hence a smaller size and alpha are
-used. Alpha can also be applied on the interpolation paths to show the
-start to finish from transparent to opaque.
+space that is best suited to visualise the bases. The centre of the 2D
+view is the first two PC coordinates of the zero matrix, and the radius
+is estimated as the largest distance from the centre to the basis. The
+theoretical best basis is known for simulated data and can be labelled
+to compare how close the end basis found by the optimisers is to the
+theoretical best. Various aesthetics, i.e.~size, alpha (transparency),
+and colour, are applicable to emphasize critical elements and adjust for
+the presentation. For example, anchor points and search points are less
+important and hence a smaller size and alpha are used. Alpha can also be
+applied on the interpolation paths to show the start to finish from
+transparent to opaque.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/polish-1}
 
@@
 and \(\mathcal{N}\) representing the projected and randomly generated
-data, and \(n\) denoting the number of observation, the Normal
-Kolmogorov index \(I^{nk}(n)\), is defined as:
+data, and \(n\) denoting the number of observation, the Kolmogorov index
+\(I^{nk}(n)\), is defined as:
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/noisy-better-geo}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/kol-result}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/kol-result}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/huizezhang/Documents/research/paper-tour-vis/figs/flip-sign}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/flip-sign}
 
@@
 All the three optimisers tend to work well for smooth indexes while PD
-fails generally fails for non-smooth indexes. This is due to the fact
-that derivative based methods are in general not suitable for optimising
+generally fails for non-smooth indexes. This is due to the fact that
+derivative based methods are in general not suitable for optimising
 non-smooth functions.
+
+For other researchers, the tools implemented in the ferrn package can
+track the progression of the target function and monitor parameter
+coverage of the search space in both reduced 2D space and the full
+space. These visualisation can lead to insights for evaluating the
+performance of multiple optimisers in a specific problem, as the example
+shown for comparing the three optimisers when optimising the kolmogorov
+index in the paper.
 

### 2021-09-08 review/response.Rmd
Small changes to conclusions and reply

@@
 
-\textcolor{magenta}{Once you go to 2D you have to take into account rotation in addition to flips, so the thing to check would be if the two frames are spanning the same subspace, and aligning them will be a bit more work.
-Might be a good idea to add 1-2 sentences in the paper and rephrase the reply.}
+\textcolor{magenta}{We agree that most index functions should be rotation invariant (this is in fact a desirable property, but there are exceptions where indexes can still be very useful but not rotation invariant, see discussion in Laa, Cook (2020)).
+Thus when comparing frames defining projections to d>=2 dimensions, we may want to introduce a rotation that will align the bases as much as possible within the common subspace, while potentially quantifying e.g.
+the angle between two planes.
+This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.
+XXX do you want to mention something about this in the paper?}
 

### 2021-09-08 paper/zhang-cook-laa-langrene-menendez.Rmd
Small changes to conclusions and reply

@@
 
-Experiments in this paper have shown that random search algorithms tend to require larger search size for 2D indexes as compared to 1D indexes.
-All the three optimisers tend to work well for smooth indexes while PD generally fails for non-smooth indexes.
-This is due to the fact that derivative based methods are in general not suitable for optimising non-smooth functions.
-
-For other researchers, the tools implemented in the ferrn package can track the progression of the target function and monitor parameter coverage of the search space in both reduced 2D space and the full space.
-These visualisation can lead to insights for evaluating the performance of multiple optimisers in a specific problem, as the example shown for comparing the three optimisers when optimising the kolmogorov index in the paper.
+The tools were implemented in the ferrn package and can track the progression of the target function and monitor parameter coverage of the search space in both reduced 2D space and the full space.
+These visualisation can lead to insights for evaluating and comparing the performance of multiple optimisers in a specific problem.
+They can provide better understanding of existing methods or motivate the development of new approaches.
+For example, we have compared how three optimisers perform when maximising a non-smooth index function and illustrated how the pseudo-derivative search fails in this setting.
+The observations from our experiments have also been translated into improved optimisation methods for the guided tour, e.g.
+we introduced the option to interrupt the search if a better basis is found along the path.
 

### 2021-09-10 review/response.Rmd
add reference of torus in p-D space and more explanation on flipping

@@
 
-\textcolor{red}{With p-D bases, the correction of orientation is a bit more complicated since both rotation and flip need to be taken into account.
-For 2D bases, one need to check if the two frames are spanning the same subspace and pointing at the same direction.}
-
-\textcolor{magenta}{We agree that most index functions should be rotation invariant (this is in fact a desirable property, but there are exceptions where indexes can still be very useful but not rotation invariant, see discussion in Laa, Cook (2020)).
+\textcolor{red}{We agree that most index functions should be rotation invariant (this is in fact a desirable property, but there are exceptions where indexes can still be very useful but not rotation invariant, see discussion in Laa, Cook (2020)).
 Thus when comparing frames defining projections to d>=2 dimensions, we may want to introduce a rotation that will align the bases as much as possible within the common subspace, while potentially quantifying e.g.
 the angle between two planes.
-This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.
-XXX do you want to mention something about this in the paper?}
+This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.}
 
@@
 \textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, given that the parameter space can be simulated.
-In our examples, the orthonormality constraint regulates the parameter space of the projection basis ($p \times d$) to be an p-sphere, which can be simulated by the `geozoo` package.
+In our examples, the orthonormality constraint regulates the parameter space of the projection basis ($p \times d$) to be an p-sphere, which can be simulated by the geozoo package.
 For other optimisers, reserachers need first to create a data collection object in a tibble that tracks the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.
-Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the `expore\_\*` functions.}
+Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the expore\_\* functions.}
 
@@
 
-\textcolor{red}{We changed the wording of the sentence to: "a torus in the $p$-D space"}
-
-\textcolor{magenta}{I think Di mentioned a reference to cite when you first mention the torus in the text, it should be added to the paper and mentioned here in the reply.}
+\textcolor{red}{We changed the wording of the sentence to: "a torus in the $p$-D space" and provide a reference for this statement: (Buja and Asimov, 1986).}
 

### 2021-09-10 paper/zhang-cook-laa-langrene-menendez.Rmd
add reference of torus in p-D space and more explanation on flipping

@@
 While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames.
-Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space.
+Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space [@Buja1986-sw].
 For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
@@
 One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
-As mentioned previously, bases with opposite signs do not affect the projection and index value, but clearly, we prefer the target to have the same orientation as the current basis.
-The orientation of two bases can be checked via calculating the determinant and a negative value indicates opposite direction.
-Hence an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
+Bases with opposite signs do not affect the projection and index value, but we would prefer the target to have the same orientation as the current basis.
+The orientation of two bases can be checked via calculating the determinant and a negative value suggests the two bases having different orientation.
+For 1D bases, this can be corrected by flipping the sign in one of the basis while the correction for a general p-D bases is more complex and beyond the scope of this paper.
+Here an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
 The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exists.

### 2021-09-10 paper/zhang-cook-laa-langrene-menendez.tex
add reference of torus in p-D space and more explanation on flipping

@@
 that the columns in a 2D basis are orthogonal to each other, so the
-space of \(p \times 2\) bases is a torus in the \(p\)-D space. For
-\(p = 3\) one would see a classical 3D torus shape as shown by the grey
-points in Figure \ref{fig:toy-torus}. The two circles of the torus can
-be observed to be perpendicular to each other and this can be linked
-back to the orthogonality condition. Two paths from CRS and PD are
-plotted on top of the torus and coloured in green and brown respectively
-to match the previous plots. The final basis found by PD and CRS are
-shown in a larger shape and printed below respectively:
+space of \(p \times 2\) bases is a torus in the \(p\)-D space
+\citep{Buja1986-sw}. For \(p = 3\) one would see a classical 3D torus
+shape as shown by the grey points in Figure \ref{fig:toy-torus}. The two
+circles of the torus can be observed to be perpendicular to each other
+and this can be linked back to the orthogonality condition. Two paths
+from CRS and PD are plotted on top of the torus and coloured in green
+and brown respectively to match the previous plots. The final basis
+found by PD and CRS are shown in a larger shape and printed below
+respectively:
 
@@
 the basis space and the interpolator seems to draw a straight line to
-interpolate. As mentioned previously, bases with opposite signs do not
-affect the projection and index value, but clearly, we prefer the target
-to have the same orientation as the current basis. The orientation of
-two bases can be checked via calculating the determinant and a negative
-value indicates opposite direction. Hence an orientation check is
-carried out once a new target basis is generated and the sign in the
-target basis will be flipped if a negative determinant is obtained. The
-interpolation after implementing the orientation check is shown on the
-right panel of Figure \ref{fig:flip-sign} where the unsatisfactory
-interpolation no longer exists.
+interpolate. Bases with opposite signs do not affect the projection and
+index value, but we would prefer the target to have the same orientation
+as the current basis. The orientation of two bases can be checked via
+calculating the determinant and a negative value suggests the two bases
+having different orientation. For 1D bases, this can be corrected by
+flipping the sign in one of the basis while the correction for a general
+p-D bases is more complex and beyond the scope of this paper. Here an
+orientation check is carried out once a new target basis is generated
+and the sign in the target basis will be flipped if a negative
+determinant is obtained. The interpolation after implementing the
+orientation check is shown on the right panel of Figure
+\ref{fig:flip-sign} where the unsatisfactory interpolation no longer
+exists.
 
@@
 
-Experiments in this paper have shown that random search algorithms tend
-to require larger search size for 2D indexes as compared to 1D indexes.
-All the three optimisers tend to work well for smooth indexes while PD
-generally fails for non-smooth indexes. This is due to the fact that
-derivative based methods are in general not suitable for optimising
-non-smooth functions.
-
-For other researchers, the tools implemented in the ferrn package can
-track the progression of the target function and monitor parameter
-coverage of the search space in both reduced 2D space and the full
-space. These visualisation can lead to insights for evaluating the
-performance of multiple optimisers in a specific problem, as the example
-shown for comparing the three optimisers when optimising the kolmogorov
-index in the paper.
+The tools were implemented in the ferrn package and can track the
+progression of the target function and monitor parameter coverage of the
+search space in both reduced 2D space and the full space. These
+visualisation can lead to insights for evaluating and comparing the
+performance of multiple optimisers in a specific problem. They can
+provide better understanding of existing methods or motivate the
+development of new approaches. For example, we have compared how three
+optimisers perform when maximising a non-smooth index function and
+illustrated how the pseudo-derivative search fails in this setting. The
+observations from our experiments have also been translated into
+improved optimisation methods for the guided tour, e.g.~we introduced
+the option to interrupt the search if a better basis is found along the
+path.
 

### 2021-09-16 review/response.Rmd
changes made during the meeting

@@
 
-Many thanks for reviewing the manuscript and for your comments and suggestions. Please see below the answers to your review comments.
+\textcolor{red}{We really appreciate the helpful comments and feedback.
+Below is a point-by-point explanation of how we have changed the paper.
+The reviewer comments are in black and our responses are in red.}
+
+\textcolor{red}{In our revision, we have used the new format for the journal and produced both a pdf format and an html format. The html format includes interactive graphics whenever appropriate.}
 
@@
 the angle between two planes.
-This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.}
+This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.
+The text in section "Reconciling the orientation" has been updated to reflect this explanation.}
 
@@
 For other optimisers, reserachers need first to create a data collection object in a tibble that tracks the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.
-Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the expore\_\* functions.}
+Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the explore_* functions.
+we have added a sentence in the conclusion to point to this possibility.
+}
 
@@
 \textcolor{red}{Thanks for sharing your interpretation on the effects of flipping, that's right!
-The reviewer is giving a decent explanation in describing the effect of flipping.
 The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
@@
 
-\textcolor{red}{Added reference of (Wilkinson et al., 2005) and  (Wilkinson and Wills, 2008):}
-
 \textcolor{red}{L. Wilkinson and G. Wills. Scagnostics distributions. Journal of Computational and Graphical Statistics, 17 (2):473–491, 2008}

### 2021-09-16 paper/zhang-cook-laa-langrene-menendez.Rmd
update web article
changes made during the meeting

@@
 output: 
-  rjtools::rjournal_pdf_article
+  rjtools::rjournal_web_article
 bibliography: zhang-cook-laa-langrene-menendez.bib
@@
 
-```{r toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\ref{fig:toy-pca}.
+```{r toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\\\ref{fig:toy-pca}.
 With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
@@
 The orientation of two bases can be checked via calculating the determinant and a negative value suggests the two bases having different orientation.
-For 1D bases, this can be corrected by flipping the sign in one of the basis while the correction for a general p-D bases is more complex and beyond the scope of this paper.
+For 1D bases, this can be corrected by flipping the sign in one of the basis while the correction for a general p-D bases is more complex and beyond the scope of this work.
 Here an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
@@
 
-This paper has provided several visual diagnostics that can be used for diagnosing a complex optimisation procedure.
+This paper has provided several visual diagnostics that can be used for understanding a complex optimisation procedure and are implemented in the ferrn package.
 The methods were illustrated using the optimisers available for projection pursuit guided tour.
 Here the constraint is the orthonormality condition of the projection bases.
-The approach used broadly applies to other constrained optimisers.
+The approach used broadly applies to other constrained optimisers, by saving the path of the optimiser in a form required to input into the ferrn package.
+These high-dimensional data objects can then be viewed using the tour.
 
-The tools were implemented in the ferrn package and can track the progression of the target function and monitor parameter coverage of the search space in both reduced 2D space and the full space.
+The progressive optimisation of a target function and it's coverage of the search space can be viewed in both reduced 2D space and the full space.
 These visualisation can lead to insights for evaluating and comparing the performance of multiple optimisers in a specific problem.

### 2021-09-16 paper/zhang-cook-laa-langrene-menendez.tex
changes made during the meeting

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-search}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-search}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/interruption-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/polish-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/kol-result}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/kol-result}
 
@@
 flipping the sign in one of the basis while the correction for a general
-p-D bases is more complex and beyond the scope of this paper. Here an
+p-D bases is more complex and beyond the scope of this work. Here an
 orientation check is carried out once a new target basis is generated
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/researh/paper-tour-vis/figs/flip-sign}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/flip-sign}
 
@@
 This paper has provided several visual diagnostics that can be used for
-diagnosing a complex optimisation procedure. The methods were
-illustrated using the optimisers available for projection pursuit guided
-tour. Here the constraint is the orthonormality condition of the
-projection bases. The approach used broadly applies to other constrained
-optimisers.
-
-The tools were implemented in the ferrn package and can track the
-progression of the target function and monitor parameter coverage of the
-search space in both reduced 2D space and the full space. These
-visualisation can lead to insights for evaluating and comparing the
-performance of multiple optimisers in a specific problem. They can
-provide better understanding of existing methods or motivate the
-development of new approaches. For example, we have compared how three
-optimisers perform when maximising a non-smooth index function and
+understanding a complex optimisation procedure and are implemented in
+the ferrn package. The methods were illustrated using the optimisers
+available for projection pursuit guided tour. Here the constraint is the
+orthonormality condition of the projection bases. The approach used
+broadly applies to other constrained optimisers, by saving the path of
+the optimiser in a form required to input into the ferrn package. These
+high-dimensional data objects can then be viewed using the tour.
+
+The progressive optimisation of a target function and it's coverage of
+the search space can be viewed in both reduced 2D space and the full
+space. These visualisation can lead to insights for evaluating and
+comparing the performance of multiple optimisers in a specific problem.
+They can provide better understanding of existing methods or motivate
+the development of new approaches. For example, we have compared how
+three optimisers perform when maximising a non-smooth index function and
 illustrated how the pseudo-derivative search fails in this setting. The

### 2021-09-16 paper/zhang-cook-laa-langrene-menendez.R

@@ -39,7 +39,7 @@ knitr::read_chunk(
 ## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
 
 
-## ----toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\ref{fig:toy-pca}. With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()----
+## ----toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\\\ref{fig:toy-pca}. With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()----
 #> NA
 
 

### 2021-09-18 review/response.Rmd
go through response again

@@
 For other optimisers, reserachers need first to create a data collection object in a tibble that tracks the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.
-Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the explore_* functions.
-we have added a sentence in the conclusion to point to this possibility.
-}
+Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the explore functions.
+we have added a sentence in the conclusion to point to this possibility.}
 
@@
 The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
-This has now been corrected.}
+This has now been corrected in the package and Figure 4 has been updated accordingly.}
 
@@
 
-\textcolor{red}{We changed the wording of the sentence to: "a torus in the $p$-D space" and provide a reference for this statement: (Buja and Asimov, 1986).}
+\textcolor{red}{We changed the wording of the sentence to: "a torus in the $p$-D space" and provide a reference for this statement.}
 
@@
 
-\textcolor{red}{Changed to "We use PCA to estimate the 2D space that is best suited to visualise the bases.
+\textcolor{red}{The sentence has been re-phrased into: "We use PCA to estimate the 2D space that is best suited to visualise the bases.
 The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}
@@
 
-\textcolor{red}{The data set is first introduced in the subsection *Simulation setup* under *Diagnosing an optimiser*.
-However, to avoid confusion, we change the "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
+\textcolor{red}{The data set is first introduced in the subsection Simulation setup under Diagnosing an optimiser.
+To avoid confusion, we change the wording from "original 5D space" to "original space" and give more details about the original space in the figure caption: "The basis space in this example is a 5D unit sphere ...".}
 
@@
 
-\textcolor{red}{A paragraph further summarising the findings has been added in the second paragram in the conclusion}
+\textcolor{red}{A paragraph further summarising the findings has been added in the second paragram in the conclusion.}
 
@@
 
-\textcolor{red}{New references have been added as follows:}
-
-\textcolor{red}{L. Wilkinson and G. Wills. Scagnostics distributions. Journal of Computational and Graphical Statistics, 17 (2):473–491, 2008}
-
-\textcolor{red}{L. Wilkinson, A. Anand, and R. Grossman. Graph-theoretic scagnostics. In IEEE Symposium on Information Visualization, 2005. INFOVIS 2005., pages 157–164, Oct 2005.}
+\textcolor{red}{Two references of Wilkinson work on scagnostics have been added. }
 
@@
 
-\textcolor{red}{The caption has now been re-written and color scale explanation added: "The color scale is from the customised botanical palette in the ferrn package."}
+\textcolor{red}{The caption has now been updated and explanation for the color scale is added: "The color scale is from the customised botanical palette in the ferrn package."}
 
@@
 
-\textcolor{red}{That has now been corrected.}
+\textcolor{red}{Corrected.}
 
@@
 
-\textcolor{red}{It is now fixed}
+\textcolor{red}{Fixed.}
 
@@
 
-\textcolor{red}{The caption has now been ammended to: "The black star represents the theoretical best basis the optimisers aiming to find." in the caption of figure 4.}
+\textcolor{red}{The black star represents the theoretical best basis the optimisers aiming to find. This has been added in the caption of Figure 4.}
 
@@
 
-\textcolor{red}{Mention of "details=TRUE" and "animate=TRUE" is removed from subsection diagnostic 3a and 3b to the implementation section: `explore\_space\_pca()` produces the PCA plot of projection bases on the reduced space.
-Figure 4 includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
-The animated version in Figure 5 is produced with argument `animate = TRUE`.}
+\textcolor{red}{"details=TRUE" and "animate=TRUE" refers to the function "explore space pca".
+These two arguments are now mentioned in the impelmentation section under function "explore space pca", instead of subsection Diagnostic 3a and 3b.}
 
@@
 \textcolor{red}{Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing.
-In the experiment of Figure 9, the parameter `max.tries` is set to 400 to allow guided tour to do its best instead of a default 25 in Figure 8.
+In the experiment of Figure 9, the parameter "max.tries" is set to 400 to allow guided tour to do its best instead of a default 25 as in Figure 8.
 Cross comparing the two experiments is less meaningful.}
@@
 
-\textcolor{red}{It has now been added}
+\textcolor{red}{Added.}
 

### 2021-09-18 review/response.Rmd
add using max.tries = 400 in Figure 9 caption

@@
 
-\textcolor{red}{Figure 8 and 9 focuses on different aspects: Figure 8 compares the effect of implementing an interruption while Figure 9 compares the effect of polishing.
-In the experiment of Figure 9, the parameter "max.tries" is set to 400 to allow guided tour to do its best instead of a default 25 as in Figure 8.
-Cross comparing the two experiments is less meaningful.}
+\textcolor{red}{In the experiment of Figure 9, the parameter "max.tries" is set to 400 to allow guided tour to do its best instead of a default 25 as in Figure 8.
+A sentence in the Figure 9 captions is added to reflect this.}
 

### 2021-09-18 paper/zhang-cook-laa-langrene-menendez.Rmd
add using max.tries = 400 in Figure 9 caption

@@
 The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis."}
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
+Note that the parameter \\code{max.tries} is set to 400 in this experiment for CRS to do its best.
+"}
 ```

### 2021-09-18 paper/zhang-cook-laa-langrene-menendez.tex
add using max.tries = 400 in Figure 9 caption

@@
 
-\caption[Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
+\caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
 The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.}\label{fig:polish}
+The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
+Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best.
+}\label{fig:polish}
 \end{figure}

### 2021-09-18 paper/zhang-cook-laa-langrene-menendez.R

@@ -68,7 +68,7 @@ holes_2d_better_3var %>% get_best() %>% pull(basis) %>% .[[1]]
 ## ----interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS. On the left panel, basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel."----
 
 
-## ----polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis."----
+## ----polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index. The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value. The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis. Note that the parameter \\code{max.tries} is set to 400 in this experiment for CRS to do its best. "----
 
 
 ## ----noisy-better-geo, fig.height = 5, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}. Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index. Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best."----

### 2021-09-21 paper/zhang-cook-laa-langrene-menendez.Rmd
response to review revised

@@
 As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
-We use PCA to estimate the 2D space that is best suited to visualise the bases.
-The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis.
-The theoretical best basis is known for simulated data and can be labelled to compare how close the end basis found by the optimisers is to the theoretical best.
+We use PCA to project and visualize the parameters/bases in 2D.
+The centre of the 2D view is the first two PCs of the data matrix.
+It theoretically should be a circle, but may have some irregular edges due to finite sampling.
+Thus the edge is smoothed by using a radius estimated as the largest distance from the centre to any basis.
+In the simulation, the theoretical best basis is known and can be labelled to compare how close to this that the optimisers stopped.
 Various aesthetics, i.e.
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star represents the theoretical best basis the optimisers aiming to find.
+The theoretical best basis is indicated by the black star.
 All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
@@
 Bases with opposite signs do not affect the projection and index value, but we would prefer the target to have the same orientation as the current basis.
-The orientation of two bases can be checked via calculating the determinant and a negative value suggests the two bases having different orientation.
-For 1D bases, this can be corrected by flipping the sign in one of the basis while the correction for a general p-D bases is more complex and beyond the scope of this work.
+The orientation of two bases can be computationally checked by calculating the determinant -- a negative value suggests the two bases having different orientation.
+For 1D bases, this can be corrected by flipping the sign in one basis.
+For higher dimensions, it can be a bit more difficult because the orthonormality of the basis needs to be maintained also when an individual vector is flipped.
 Here an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
@@
 The methods were illustrated using the optimisers available for projection pursuit guided tour.
-Here the constraint is the orthonormality condition of the projection bases.
-The approach used broadly applies to other constrained optimisers, by saving the path of the optimiser in a form required to input into the ferrn package.
+Here the constraint is the orthonormality condition of the projection bases, which corresponds to optimisation over spheres and torii.
+The approach described broadly applies to other constrained optimisers.
+Although the manifold in $p$-space might be different the diagnostic techniques are the same.
+A researcher would begin by saving the path of the optimiser in a form required to input into the ferrn package, as described in this paper.
+One might generally make more samples from the constrained space upon which to assess and compare the optimisation paths.
 These high-dimensional data objects can then be viewed using the tour.
@@
 The progressive optimisation of a target function and it's coverage of the search space can be viewed in both reduced 2D space and the full space.
-These visualisation can lead to insights for evaluating and comparing the performance of multiple optimisers in a specific problem.
+These visualisations can lead to insights for evaluating and comparing the performance of multiple optimisers operating on the same task.
 They can provide better understanding of existing methods or motivate the development of new approaches.

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.Rmd
misc fix as per issue 18

--- 2021-09-18 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
   \usepackage[ruled, linesnumbered]{algorithm2e}
+type: package
+csl: "biostats.csl"
+creative_commons: CC BY
 output: 
@@
                       fig.retina = 5,
-                      out.width = "100%")
+                      out.width = "100%",
+                      fig.cap = TRUE)
 ```
@@
 The type of optimisation problem considered in this paper is constrained optimization [@bertsekas2014constrained], assuming it is not possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists in finding the minimum or maximum of a function $f \in L^p$ in the constrained space $\mathcal{A}$.
+That is, the problem consists in finding the minimum or maximum of a function $f \in L^p$ in the constrained space $\mathcal{A}$, where $L^p$ defines the vector space of function $f$, whose $p$th power are integrable.
 
@@
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
-Figure \ref{fig:tour-path} shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames.
+Figure \@ref(fig:tour-path) shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
@@
 Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation; $j$ and $l$ are the counter of the outer and inner loop respectively.
-Other parameters of interest recorded include `method` that tags the name of the optimiser, and `alpha` that indicates the sampling neighbourhood size for searching observations.
-A matrix notation of the data structure is presented in equation \ref{eq:data-structure}.
+Other parameters of interest recorded, $V$, include `method` that tags the name of the optimiser, and `alpha` that indicates the sampling neighbourhood size for searching observations.
+A matrix notation of the data structure is presented in the table below.
 
-```{=tex}
-\begin{equation}
-\renewcommand\arraystretch{2}  % default value: 1.0
-\begin{pmatrix}
-t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
-\hline
-1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
-\hline
-2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
-\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
-\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
-\hline
-\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
-\hline
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-\hline
-\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
-\hline
-\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
-\hline
-\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
-\end{pmatrix}
-= 
-\begin{pmatrix}
-\text{column name} \\
-\hline
-\text{search (start basis)} \\
-\hline
-\text{search} \\
-\vdots \\
-\text{search (accepted basis)} \\
-\hline
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\vdots \\
-\hline
-\text{search} \\
-\vdots \\
-\text{search (final basis)} \\
-\hline
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\text{search (no output)} \\
-\vdots \\
-\text{search (no output)} \\
-\end{pmatrix}
-\label{eq:data-structure}
-\end{equation}
-```
+| $t$| $\mathbf{A}$| $I$| $S$| $j$| $l$| $V_{1}$| $V_{2}$| Notes|
+|---|---|---|---|---|---|---|---|-------|
+| $1$ | $\mathbf{A}_1$| $I_1$| $S_1$| 1| 1| $V_{11}$| $V_{12}$| start basis|
+| $2$ | $\mathbf{A}_2$| $I_2$| $S_2$| 2| 1| $V_{21}$| $V_{22}$| search|
+| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
+| ... | ... | ...| ...| 2     | $l_2$   | ...| ...| search (accepted) |
+| ... | ... | ...| ...| 2     | 1       | ...| ...| interpolation|
+| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
+| ... | ... | ...| ...| 2     | $k_2$   | ...| ...| interpolation|
+| ... | ... | ...| ...| ...   | ...     | ...| ...| ...|
+| ... | ... | ...| ...| $J$   | 1       | ...| ...| search|
+| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
+| $T$ | $\mathbf{A}_T$ | $I_T$| $S_T$| $J$  | $l_J$| $V_{T1}$| $V_{T2}$| search (final)|
+| ... | ... | ...| ...| $J$   | 1| ...| ...| interpolation|
+| ... | ... | ...| ...| ...   | ...| ...| ...| ...|
+| ... | ... | ...| ...| $J$   | $k_J$| ...| ...| interpolation|
+| ... | ... | ...| ...| $J+1$   | 1| ...| ...| search (last round)|
+| ... | ... | ...| ...| ...   | ...| ...| ...| ...    |
+| $T^{\prime}$ | $\mathbf{A}_{T^{\prime}}$  | $I_{T^{\prime}}$ | $S_{T^{\prime}}$ | $J+1$ | $l_{J+1}$ | $V_{{T}^{\prime}1}$ | $V_{{T}^{\prime}2}$ | search (last round)   |
+
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
@@
 ```
-Figure \ref{fig:toy-search} shows an example of the search plot for CRS (left) and SA (right).
+Figure \@ref(fig:toy-search) shows an example of the search plot for CRS (left) and SA (right).
 Both optimisers quickly find better bases in the first few iterations and then take longer to find one in the later iterations.
@@
 
-```{r toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 
-```{r toy-pca-animated-interactive,  out.wdith= "100px", fig.cap = "Animated version of Figure \\\\ref{fig:toy-pca}.
+```{r toy-pca-animated-interactive, fig.cap = "Animated version of Figure 4.
 With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
@@
 
-```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
+```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\@ref(fig:toy-pca) and \\@ref(fig:toy-pca-animated).
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
-The full animation can be seen in the html version of the paper."}
+The full animation can be seen in the html version of the paper.", eval = knitr::is_latex_output()}
 ```
 
-```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated} in the full basis space.
+```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure 4 and 5 in the full basis space.
 The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.", eval = knitr::is_html_output()}
@@
 
-    -   `explore_trace_search()` produces summary plots in Figure \ref{fig:toy-search}.
+    -   `explore_trace_search()` produces summary plots in Figure \@ref(fig:toy-search).
     -   `explore_trace_interp()` produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}.
@@
 This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R.
-The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-tour-vis}.
+The source code for reproducing this paper can be found at: <https://github.com/huizezhang-sherry/paper-tour-vis>.
 
 \newpage
-
-# Appendix
-
-Three algorithms (creeping random search, simulated annealing, and pseudo-derivative) used in projection pursuit guided tour optimisation.
-
-```{=tex}
-\begin{algorithm}
-\SetAlgoLined
-  \SetKwInOut{input}{input}
-  \SetKwInOut{output}{output}
-    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
-    \output{$\mathbf{A}_{l}$}
-    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
-  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   Set $l = 1$\;
-  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    Update $l = l + 1$\;
-  }
-  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
-}
-  \caption{Creeping random search (CRS)}
-  \label{random-search}
-\end{algorithm}
-```
-```{=tex}
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-    Update $l = l + 1$\;
-  }
-  \caption{Simulated annealing (SA)}
-  \label{simulated-annealing}
-\end{algorithm}
-```
-```{=tex}
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
-  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
-  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
-  Compute the corresponding index value for each candidate bases\;
-  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
-  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-  Update $l = l + 1$\;
-}
-\caption{Pseudo-derivative (PD)}
-\label{search-geodesic}
-\end{algorithm}
-```

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.tex
misc fix as per issue 18

@@
 That is, the problem consists in finding the minimum or maximum of a
-function \(f \in L^p\) in the constrained space \(\mathcal{A}\).
+function \(f \in L^p\) in the constrained space \(\mathcal{A}\), where
+\(L^p\) defines the vector space of function \(f\), whose \(p\)th power
+are integrable.
 
@@
 interpolation between these planes to form a tour path. Figure
-\ref{fig:tour-path} shows a sketch of the tour path where the blue
+@ref(fig:tour-path) shows a sketch of the tour path where the blue
 frames are produced by the projection pursuit optimisation and the white
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/tour-path}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}
 
@@
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
+Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}(\#fig:tour-path)
 \end{figure}
@@
 counter of the outer and inner loop respectively. Other parameters of
-interest recorded include \texttt{method} that tags the name of the
-optimiser, and \texttt{alpha} that indicates the sampling neighbourhood
-size for searching observations. A matrix notation of the data structure
-is presented in equation \ref{eq:data-structure}.
-
-\begin{equation}
-\renewcommand\arraystretch{2}  % default value: 1.0
-\begin{pmatrix}
-t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
-\hline
-1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
-\hline
-2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
-\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
-\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
-\hline
-\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
-\hline
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-\hline
-\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
-\hline
-\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
-\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
-\hline
-\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
-\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
-T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
-\end{pmatrix}
-= 
-\begin{pmatrix}
-\text{column name} \\
-\hline
-\text{search (start basis)} \\
-\hline
-\text{search} \\
-\vdots \\
-\text{search (accepted basis)} \\
-\hline
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\vdots \\
-\hline
-\text{search} \\
-\vdots \\
-\text{search (final basis)} \\
-\hline
-\text{interpolate} \\
-\vdots \\
-\text{interpolate} \\
-\hline
-\text{search (no output)} \\
-\vdots \\
-\text{search (no output)} \\
-\end{pmatrix}
-\label{eq:data-structure}
-\end{equation}
+interest recorded, \(V\), include \texttt{method} that tags the name of
+the optimiser, and \texttt{alpha} that indicates the sampling
+neighbourhood size for searching observations. A matrix notation of the
+data structure is presented in the table below.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.23}}@{}}
+\toprule
+\(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
+\(V_{2}\) & Notes \\
+\midrule
+\endhead
+\(1\) & \(\mathbf{A}_1\) & \(I_1\) & \(S_1\) & 1 & 1 & \(V_{11}\) &
+\(V_{12}\) & start basis \\
+\(2\) & \(\mathbf{A}_2\) & \(I_2\) & \(S_2\) & 2 & 1 & \(V_{21}\) &
+\(V_{22}\) & search \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(l_2\) & \ldots{} &
+\ldots{} & search (accepted) \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & 1 & \ldots{} & \ldots{}
+& interpolation \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & 2 & \(k_2\) & \ldots{} &
+\ldots{} & interpolation \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
+\ldots{} & search \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\(T\) & \(\mathbf{A}_T\) & \(I_T\) & \(S_T\) & \(J\) & \(l_J\) &
+\(V_{T1}\) & \(V_{T2}\) & search (final) \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & 1 & \ldots{} &
+\ldots{} & interpolation \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J\) & \(k_J\) & \ldots{} &
+\ldots{} & interpolation \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \(J+1\) & 1 & \ldots{} &
+\ldots{} & search (last round) \\
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{} & \ldots{} \\
+\(T^{\prime}\) & \(\mathbf{A}_{T^{\prime}}\) & \(I_{T^{\prime}}\) &
+\(S_{T^{\prime}}\) & \(J+1\) & \(l_{J+1}\) & \(V_{{T}^{\prime}1}\) &
+\(V_{{T}^{\prime}2}\) & search (last round) \\
+\bottomrule
+\end{longtable}
 
@@
 
-Figure \ref{fig:toy-search} shows an example of the search plot for CRS
+Figure @ref(fig:toy-search) shows an example of the search plot for CRS
 (left) and SA (right). Both optimisers quickly find better bases in the
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-search}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-search}
 
@@
 There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
-The color scale is from the customised botanical palette in the ferrn package.}\label{fig:toy-search}
+The color scale is from the customised botanical palette in the ferrn package.}(\#fig:toy-search)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp-1}
 
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}(\#fig:toy-interp)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}
 
@@
 The black star represents the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}(\#fig:toy-pca)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca-animated-1}
 
@@
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
-The full video of the animation can be found in the html version of the paper.}\label{fig:toy-pca-animated}
+The full video of the animation can be found in the html version of the paper.}(\#fig:toy-pca-animated)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-tour-1}
-
-}
-
-\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-tour-1}
+
+}
+
+\caption[Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated)]{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated).
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
-The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
+The full animation can be seen in the html version of the paper.}(\#fig:toy-tour)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-torus-1}
 
@@
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
+The full video of the animation can be found in the html version of the paper.}(\#fig:toy-torus)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interruption-1}
 
@@
 On the left panel, basis with a higher index value is found during the interpolation but not used.
-On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}(\#fig:interruption)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish-1}
 
@@
 Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best.
-}\label{fig:polish}
+}(\#fig:polish)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/noisy-better-geo}
 
@@
 Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}(\#fig:noisy-better-geo)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/kol-result}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/kol-result}
 
@@
 \caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}(\#fig:kol-result)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-tour-vis/figs/flip-sign}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/flip-sign}
 
@@
 On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
-The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
+The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}(\#fig:flip-sign)
 \end{figure}
@@
     \texttt{explore\_trace\_search()} produces summary plots in Figure
-    \ref{fig:toy-search}.
+    @ref(fig:toy-search).
   \item
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-Three algorithms (creeping random search, simulated annealing, and
-pseudo-derivative) used in projection pursuit guided tour optimisation.
-
-\begin{algorithm}
-\SetAlgoLined
-  \SetKwInOut{input}{input}
-  \SetKwInOut{output}{output}
-    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
-    \output{$\mathbf{A}_{l}$}
-    Generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
-  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
-   Set $l = 1$\;
-  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
-    Update $l = l + 1$\;
-  }
-  Update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
-  Construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
-  Update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
-}
-  \caption{Creeping random search (CRS)}
-  \label{random-search}
-\end{algorithm}
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
-    Generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
-    Compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
-    Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
-    Update $l = l + 1$\;
-  }
-  \caption{Simulated annealing (SA)}
-  \label{simulated-annealing}
-\end{algorithm}
-
-\begin{algorithm}
-\SetAlgoLined
-\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
-  Generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
-  Compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensuring orthogonality\;
-  Compute the corresponding index value for each candidate bases\;
-  Determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
-  Determine the step size via optimising the index value on the search direction over a 90 degree window\;
-  Find the optimum $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
-  Update $l = l + 1$\;
-}
-\caption{Pseudo-derivative (PD)}
-\label{search-geodesic}
-\end{algorithm}
+\begin{verbatim}
+\end{verbatim}
 

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.R

@@ -8,7 +8,8 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.align = "center",
                       fig.height = 3,
                       fig.retina = 5,
-                      out.width = "100%")
+                      out.width = "100%",
+                      fig.cap = TRUE)
 
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------
@@ -33,20 +34,20 @@ knitr::read_chunk(
 ## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."----
 
 
-## ----toy-pca, fig.height = 4, fig.asp = 1, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
+## ----toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers aiming to find. All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
 
 
 ## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
 
 
[…]

### 2021-09-28 review/response.Rmd

@@
 
-\textcolor{red}{We agree that most index functions should be rotation invariant (this is in fact a desirable property, but there are exceptions where indexes can still be very useful but not rotation invariant, see discussion in Laa, Cook (2020)).
-Thus when comparing frames defining projections to d>=2 dimensions, we may want to introduce a rotation that will align the bases as much as possible within the common subspace, while potentially quantifying e.g.
-the angle between two planes.
-This is an interesting problem, however the implementation of this general approach is beyond the scope of the current paper.
+\textcolor{red}{Your comment involves two different, and quite separate aspects, of projection pursuit.
+One is that index functions should be rotation invariant, that within the same plane any rotation of the index should have same value.
+There are exceptions where rotation variant indexes can still be very useful, see discussion in Laa and Cook (2020).
+And yes, operating on a Stiefel manifold can help - actually this is a work in progress to add to the tourr package.
+The same optimisation diagnostics reported in this paper should apply to Stiefel manifolds in addition to Grassmann manifolds.}
+
+\textcolor{red}{The second element of your comment refers to the direction of projection vectors.
+The index value will be the same regardless of direction, but the representation of a vector as a dot on the surface of a sphere means that it may be located on either side depending on the direction.
+Technically it is simply that the signs of the coefficient are opposite.
+You encounter this in many different analyses, including PCA.
+In PCA, there is a convention applied to force the first element of the project vector to be positive.
+That is analogous to what we are doing with the flipping operation.
 The text in section "Reconciling the orientation" has been updated to reflect this explanation.}
@@
 \textcolor{red}{It is feasible to generalise the tour method for diagnosing other optimisers, given that the parameter space can be simulated.
-In our examples, the orthonormality constraint regulates the parameter space of the projection basis ($p \times d$) to be an p-sphere, which can be simulated by the geozoo package.
-For other optimisers, reserachers need first to create a data collection object in a tibble that tracks the parameter during the optimisation, and then to use ferrn to produce the diagnostic plots as shown in the paper.
-Some examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the explore functions.
-we have added a sentence in the conclusion to point to this possibility.}
+In our examples, the orthonormality constraint regulates the parameter space of the projection basis ($p \times d$) to be a sphere, or torus, which can be simulated by the geozoo package.
+In other problems the parameter space may take different forms in the $p$-dimensional space.
+Nevertheless, this can be visualised and the ferrn package should be possible to use.
+Researchers need first to create a data collection object in a tibble structure, analgous to the one described in our paper, that tracks the parameter during the optimisation.
+Then the ferrn diagnostic plots could be used.
+Some additional examples of the data collection objects that we created are included in the ferrn package and should be referred to as the format for inputs into the explore functions.
+We have added several sentences in the conclusion explaining this.}
 
@@
     -   This is related to the major comment.
-In the figure, PD flipped in sign but stopped at a basis representation similar to the destination of the CRS.
+In the Figure, PD flipped in sign but stopped at a basis representation similar to the destination of the CRS.
 Flipping the sign back, does it mean that PD and CRS diverged but stopped at the similar subspace with opposite basis representations?
@@
 \textcolor{red}{Thanks for sharing your interpretation on the effects of flipping, that's right!
-The gray dot should land on the dashed line and this is a mistake due to the data not being updated after flipping in the code.
+The gray dot should land on the dashed line and this was a mistake in the code.
 This has now been corrected in the package and Figure 4 has been updated accordingly.}
@@
 
-\textcolor{red}{The sentence has been re-phrased into: "We use PCA to estimate the 2D space that is best suited to visualise the bases.
-The centre of the 2D view is the first two PC coordinates of the zero matrix, and the radius is estimated as the largest distance from the centre to the basis."}
+\textcolor{red}{The sentence has been re-phrased as suggested.}
 
@@
 
-\textcolor{red}{A paragraph further summarising the findings has been added in the second paragram in the conclusion.}
+\textcolor{red}{The conclusion has been expanded.}
 
@@
 
-\textcolor{red}{The caption has now been updated and explanation for the color scale is added: "The color scale is from the customised botanical palette in the ferrn package."}
+\textcolor{red}{The caption has now been updated and explanation for the color scale is added."}
 
@@
 
-\textcolor{red}{The black star represents the theoretical best basis the optimisers aiming to find. This has been added in the caption of Figure 4.}
+\textcolor{red}{The black star represents the theoretical best basis the optimisers aiming to find.
+This has been added in the caption of Figure 4, but not Figure 5 as this refers back to Figure 4 in the caption.}
 
@@
 \textcolor{red}{"details=TRUE" and "animate=TRUE" refers to the function "explore space pca".
-These two arguments are now mentioned in the impelmentation section under function "explore space pca", instead of subsection Diagnostic 3a and 3b.}
+These two arguments are now mentioned in the implementation section under function "explore space pca", instead of subsection Diagnostic 3a and 3b.}
 
-7.  In Figure 9, it seems CRS is doing better here than in Figure 8-- maybe this could be mentioned.
+7.  In Figure 9, it seems CRS is doing better here than in Figure 8 -- maybe this could be mentioned.
 
@@
 In our case we are not assuming normality on the theoretical pdf and that is the reason why we are not using normal quantiles.
-To make it clearer, we change the wording from "Normal Kolmogorov Index" to "Kolmogorove Index".}
+To make it clearer, we change the wording from "Normal Kolmogorov Index" to "Kolmogorov Index".}
 

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.Rmd
add css to fix bullet point line spacing

@@
 preamble: >
-  \usepackage{mathtools}
+  \usepackage{mathtools, longtable}
   \usepackage[ruled, linesnumbered]{algorithm2e}
@@
 output: 
-  rjtools::rjournal_web_article
+  rjtools::rjournal_web_article:
+    css: "style.css"
 bibliography: zhang-cook-laa-langrene-menendez.bib
@@
                       fig.retina = 5,
-                      out.width = "100%",
-                      fig.cap = TRUE)
+                      out.width = "100%")
 ```

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.tex
add css to fix bullet point line spacing

@@
 sphere and random bases on the sphere can be generated via the
-\CRANpkg{geozoo} \citep{geozoo} package. We use PCA to estimate the 2D
-space that is best suited to visualise the bases. The centre of the 2D
-view is the first two PC coordinates of the zero matrix, and the radius
-is estimated as the largest distance from the centre to the basis. The
-theoretical best basis is known for simulated data and can be labelled
-to compare how close the end basis found by the optimisers is to the
-theoretical best. Various aesthetics, i.e.~size, alpha (transparency),
+\CRANpkg{geozoo} \citep{geozoo} package. We use PCA to project and
+visualize the parameters/bases in 2D. The centre of the 2D view is the
+first two PCs of the data matrix. It theoretically should be a circle,
+but may have some irregular edges due to finite sampling. Thus the edge
+is smoothed by using a radius estimated as the largest distance from the
+centre to any basis. In the simulation, the theoretical best basis is
+known and can be labelled to compare how close to this that the
+optimisers stopped. Various aesthetics, i.e.~size, alpha (transparency),
 and colour, are applicable to emphasize critical elements and adjust for
@@
 index value, but we would prefer the target to have the same orientation
-as the current basis. The orientation of two bases can be checked via
-calculating the determinant and a negative value suggests the two bases
-having different orientation. For 1D bases, this can be corrected by
-flipping the sign in one of the basis while the correction for a general
-p-D bases is more complex and beyond the scope of this work. Here an
-orientation check is carried out once a new target basis is generated
-and the sign in the target basis will be flipped if a negative
-determinant is obtained. The interpolation after implementing the
-orientation check is shown on the right panel of Figure
+as the current basis. The orientation of two bases can be
+computationally checked by calculating the determinant -- a negative
+value suggests the two bases having different orientation. For 1D bases,
+this can be corrected by flipping the sign in one basis. For higher
+dimensions, it can be a bit more difficult because the orthonormality of
+the basis needs to be maintained also when an individual vector is
+flipped. Here an orientation check is carried out once a new target
+basis is generated and the sign in the target basis will be flipped if a
+negative determinant is obtained. The interpolation after implementing
+the orientation check is shown on the right panel of Figure
 \ref{fig:flip-sign} where the unsatisfactory interpolation no longer
@@
 available for projection pursuit guided tour. Here the constraint is the
-orthonormality condition of the projection bases. The approach used
-broadly applies to other constrained optimisers, by saving the path of
-the optimiser in a form required to input into the ferrn package. These
+orthonormality condition of the projection bases, which corresponds to
+optimisation over spheres and torii. The approach described broadly
+applies to other constrained optimisers. Although the manifold in
+\(p\)-space might be different the diagnostic techniques are the same. A
+researcher would begin by saving the path of the optimiser in a form
+required to input into the ferrn package, as described in this paper.
+One might generally make more samples from the constrained space upon
+which to assess and compare the optimisation paths. These
 high-dimensional data objects can then be viewed using the tour.
@@
 the search space can be viewed in both reduced 2D space and the full
-space. These visualisation can lead to insights for evaluating and
-comparing the performance of multiple optimisers in a specific problem.
-They can provide better understanding of existing methods or motivate
-the development of new approaches. For example, we have compared how
-three optimisers perform when maximising a non-smooth index function and
-illustrated how the pseudo-derivative search fails in this setting. The
-observations from our experiments have also been translated into
-improved optimisation methods for the guided tour, e.g.~we introduced
-the option to interrupt the search if a better basis is found along the
-path.
+space. These visualisations can lead to insights for evaluating and
+comparing the performance of multiple optimisers operating on the same
+task. They can provide better understanding of existing methods or
+motivate the development of new approaches. For example, we have
+compared how three optimisers perform when maximising a non-smooth index
+function and illustrated how the pseudo-derivative search fails in this
+setting. The observations from our experiments have also been translated
+into improved optimisation methods for the guided tour, e.g.~we
+introduced the option to interrupt the search if a better basis is found
+along the path.
 

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.R

@@ -8,8 +8,7 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.align = "center",
                       fig.height = 3,
                       fig.retina = 5,
-                      out.width = "100%",
-                      fig.cap = TRUE)
+                      out.width = "100%")
 
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.Rmd
control gif size with css

@@
     css: "style.css"
+    fig_caption: yes
 bibliography: zhang-cook-laa-langrene-menendez.bib
@@
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
-Figure \@ref(fig:tour-path) shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
@@
 
-```{r tour-path, out.width="70%", fig.cap="An illustration for demonstrating the frames in a tour path.
+```{r tour-path, out.width="70%", fig.cap="\\label{fig:tour-path} An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
@@
 A matrix notation of the data structure is presented in the table below.
+
+\newpage
 
@@
 
-```{r toy-pca-animated-interactive, fig.cap = "Animated version of Figure 4.
+```{r toy-pca-animated-interactive, out.extra="id = 'gif'", fig.cap = "Animated version of Figure 4.
 With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
@@
 
-```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure 4 and 5 in the full basis space.
+```{r toy-tour-interactive, out.extra="id = 'tour-gif'",fig.cap="Animating the two paths from Figure 4 and 5 in the full basis space.
 The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.", eval = knitr::is_html_output()}
@@
 
-```{r toy-torus-interactive, fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
+```{r toy-torus-interactive, out.extra="id = 'tour-gif'", fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using the holes index.
@@
 
-1)  the distance between the basis found and the current basis needs to be larger than 1e-3;
+1)  the distance between the basis found and the current needs to be larger than 1e-3;
 2)  the relative change of the index value needs to be larger than 1e-5; and

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.tex
control gif size with css

@@
 interpolation between these planes to form a tour path. Figure
-@ref(fig:tour-path) shows a sketch of the tour path where the blue
+\ref{fig:tour-path} shows a sketch of the tour path where the blue
 frames are produced by the projection pursuit optimisation and the white
@@
 
-\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path.
+\caption{\label{fig:tour-path} An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
@@
 data structure is presented in the table below.
+
+\newpage
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interruption-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interrupt}
 
@@
 \item
-  the distance between the basis found and the current basis needs to be
+  the distance between the basis found and the current needs to be
   larger than 1e-3;
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish}
 

### 2021-09-28 paper/zhang-cook-laa-langrene-menendez.R

@@ -8,7 +8,8 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.align = "center",
                       fig.height = 3,
                       fig.retina = 5,
-                      out.width = "100%")
+                      out.width = "100%",
+                      fig.cap = TRUE)
 
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------
@@ -24,7 +25,7 @@ knitr::read_chunk(
 #spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
 
 
-## ----tour-path, out.width="70%", fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
+## ----tour-path, out.width="70%", fig.cap="\\label{fig:tour-path} An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
 
 
 ## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the ferrn package."----
@@ -39,14 +40,14 @@ knitr::read_chunk(
 ## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
 
 
[…]

### 2021-09-30 paper/zhang-cook-laa-langrene-menendez.Rmd
final clean up regarding the dual output formats

--- 2021-09-28 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 
-```{r tour-path, out.width="70%", fig.cap="An illustration for demonstrating the frames in a tour path.
+```{r tour-path, fig.cap="An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
@@
 The optimiser terminates when the maximum number of iteration $l_{\max}$ is reached before a better basis can be found.
-The algorithm of CRS is summarised in Algorithm \ref{random-search} in the appendix.
+The algorithm of CRS can be found in the appendix.
 @posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
@@
 This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enables the optimiser to jump out of a local optimum.
-The algorithm \ref{simulated-annealing} in the appendix highlights how SA differs from CRS in the inner loop.
+The second algorithm in the appendix highlights how SA differs from CRS in the inner loop.
 
@@
 The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.
-Algorithm \ref{search-geodesic} in the appendix summarises the inner loop of the PD.
+The third algorithm in the appendix summarises the inner loop of the PD.
 
@@
 ```
-Figure \@ref(fig:toy-search) shows an example of the search plot for CRS (left) and SA (right).
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))` shows an example of the search plot for CRS (left) and SA (right).
 Both optimisers quickly find better bases in the first few iterations and then take longer to find one in the later iterations.
@@
 Trace plots are created by plotting the index value against time.
-Figure \ref{fig:toy-interp} presents the trace plot of the same optimisations as Figure \ref{fig:toy-search} and one can observe that the trace is smooth in both cases.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-interp}','\\@ref(fig:toy-interp)'))` presents the trace plot of the same optimisations as Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))` and one can observe that the trace is smooth in both cases.
 It may seem bizarre at first sight that the interpolation sometimes passes bases with higher index values before it decreases to a lower target.
@@
 This indicates that a non-monotonic interpolation cannot be avoided, even for CRS.
-Later, in section \ref{monotonic}, there will be a discussion on improving the non-monotonic interpolation for CRS.
+Later, in Section *A problem of non-monotonicity*, there will be a discussion on improving the non-monotonic interpolation for CRS.
 
@@
 As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
-We use PCA to project and visualize the parameters/bases in 2D.
+We use PCA to project and visualize the parameters/ bases in 2D.
 The centre of the 2D view is the first two PCs of the data matrix.
@@
 
-Figure \ref{fig:toy-pca} shows the PCA plot of CRS and PD for a 1D projection problem.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca}','\\@ref(fig:toy-pca)'))` shows the PCA plot of CRS and PD for a 1D projection problem.
 Both optimisers find the optimum but PD gets closer.
 With the PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
-There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in detail in Section \ref{monotonic}.
+There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in detail in Section *A problem of non-monotonicity*.
 
-```{r toy-pca, fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, out.width = "70%", fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 An animation is another type of display to show how the search progresses from start to finish in the space.
-Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-pca-animated-interactive}', '\\ref{fig:toy-pca-animated}'))` shows six frames from the animation of the PCA plot in Figure \ref{fig:toy-pca}.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca-animated}','\\@ref(fig:toy-pca-animated-interactive)'))` shows the animated version (six frames from the animation if viewed in pdf) of the PCA plot in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca}','\\@ref(fig:toy-pca)'))`.
 An additional piece of information one can learn from this animation is that CRS finds its end basis quicker than PD since CRS finishes its search in the 5th frame while PD is still making more progress.
@@
 
-```{r toy-pca-animated-interactive, fig.cap = "Animated version of Figure 4.
+```{r toy-pca-animated-interactive, out.extra="class = 'gif'", fig.cap = "Animated version of Figure 4.
 With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
@@
 While the PCA plot projects the bases from the direction that maximises the variance, the tour plot displays the original high-dimensional space from various directions using animation.
-Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-torus-interactive}', '\\ref{fig:toy-torus}'))` shows some frames from the tour plot of the same two optimisations in its original space.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-tour}','\\@ref(fig:toy-tour-interactive)'))` shows some frames from the tour plot of the same two optimisations in its original space.
 
-```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\@ref(fig:toy-pca) and \\@ref(fig:toy-pca-animated).
+```{r toy-tour, fig.cap="Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}.
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
@@
 
-```{r toy-tour-interactive, fig.cap="Animating the two paths from Figure 4 and 5 in the full basis space.
+```{r toy-tour-interactive, out.extra="class = 'tour-gif'",fig.cap="Animating the two paths from Figure 4 and 5 in the full basis space.
 The full basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.", eval = knitr::is_html_output()}
@@
 Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space [@Buja1986-sw].
-For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure \ref{fig:toy-torus}.
+For $p = 3$ one would see a classical 3D torus shape as shown by the grey points in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-torus}','\\@ref(fig:toy-torus-interactive)'))`.
 The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition.
@@
 One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and are symmetric when projected onto the other (due the sign difference in the second column).
-In Figure `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\ref{fig:toy-torus-interactive}', '\\ref{fig:toy-torus}'))`, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
+In Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-torus}','\\@ref(fig:toy-torus-interactive)'))`, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.
 
@@
 
-```{r toy-torus-interactive, fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
+```{r toy-torus-interactive, out.extra="class = 'tour-gif'", fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green).
 The projection problem is a 2D projection with three variables using the holes index.
@@
 
-An example of non-monotonic interpolation has been given in Figure \ref{fig:toy-interp}: a path that passes bases with a higher index value than the target one.
+An example of non-monotonic interpolation has been given in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-interp}','\\@ref(fig:toy-interp)'))`: a path that passes bases with a higher index value than the target one.
 For SA, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for CRS.
-The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure \ref{fig:interruption} and one can observe that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))` and one can observe that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
-The right panel of Figure \ref{fig:interruption} shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at time 61 has a higher index value than the target in the third interpolation.
+The right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))` shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at time 61 has a higher index value than the target in the third interpolation.
 Rather than starting the next iteration from the target basis on time 65, CRS starts the next iteration at time 61 on the right panel and reaches a better final basis.
@@
 
-Figure \ref{fig:polish} presents the projected data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:polish}','\\@ref(fig:polish)'))` presents the projected data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
 The top row shows the initial projection, the final projection after CRS, and the final projection after polish, respectively The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
@@
 
-Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated paths are smooth.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:noisy-better-geo}','\\@ref(fig:noisy-better-geo)'))` presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated paths are smooth.
 There is barely any improvement made by PD, indicating its failure in optimising non-smooth index functions.
@@
 Two search neighbourhood sizes: 0.5 and 0.7 are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
-Figure \ref{fig:kol-result} shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the optimiser and search size.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:kol-result}','\\@ref(fig:kol-result)'))` shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the optimiser and search size.
 With CRS and a search size of 0.5, despite being the simplest and fastest, the optimiser fails in three instances where the final basis lands neither near the local nor the global optimum.
@@
 
-One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure \ref{fig:flip-sign}, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
+One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))`, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
 Bases with opposite signs do not affect the projection and index value, but we would prefer the target to have the same orientation as the current basis.
@@
 Here an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
-The interpolation after implementing the orientation check is shown on the right panel of Figure \ref{fig:flip-sign} where the unsatisfactory interpolation no longer exists.
+The interpolation after implementing the orientation check is shown on the right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))` where the unsatisfactory interpolation no longer exists.
 
@@
 
-    -   `explore_trace_search()` produces summary plots in Figure \@ref(fig:toy-search).
-    -   `explore_trace_interp()` produces trace plots for the interpolation points in Figure \ref{fig:toy-interp}.
-    -   `explore_space_pca()` produces the PCA plot of projection bases on the reduced space in Figure \ref{fig:toy-pca}.
-The animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`.
+    -   `explore_trace_search()` produces summary plots in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))`.
+    -   `explore_trace_interp()` produces trace plots for the interpolation points in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-interp}','\\@ref(fig:toy-interp)'))`.
     -   `explore_space_pca()` produces the PCA plot of projection bases on the reduced space.
-Figure \ref{fig:toy-pca} includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
-The animated version in Figure \ref{fig:toy-pca-animated} is produced with argument `animate = TRUE`.
-    -   `explore_space_tour()` produces animated tour view on the full space of the projection bases in Figure \ref{fig:toy-tour}.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca}','\\@ref(fig:toy-pca)'))` includes the additional details of anchor and search bases, which can be turned on by the argument `details = TRUE`.
+The animated version in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca-animated}','\\@ref(fig:toy-pca-animated-interactive)'))` is produced with argument `animate = TRUE`.
+    -   `explore_space_tour()` produces animated tour view on the full space of the projection bases in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-tour}','\\@ref(fig:toy-tour-interactive)'))`.
 

### 2021-09-30 paper/zhang-cook-laa-langrene-menendez.tex
final clean up regarding the dual output formats

@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}
-
-}
-
-\caption{\label{fig:tour-path} An illustration for demonstrating the frames in a tour path.
+{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}
+
+}
+
+\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path.
 Each square (frame) represents the projected data with a corresponding basis.
-Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}(\#fig:tour-path)
+Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
 \end{figure}
@@
 iteration \(l_{\max}\) is reached before a better basis can be found.
-The algorithm of CRS is summarised in Algorithm \ref{random-search} in
-the appendix. \citet{posse1995projection} has proposed a slightly
-different cooling scheme by introducing a halving parameter \(c\). In
-his proposal \(\alpha\) is only adjusted if the last iteration takes
-more than \(c\) times to find a better basis.
+The algorithm of CRS can be found in the appendix.
+\citet{posse1995projection} has proposed a slightly different cooling
+scheme by introducing a halving parameter \(c\). In his proposal
+\(\alpha\) is only adjusted if the last iteration takes more than \(c\)
+times to find a better basis.
 
@@
 higher index value and hence enables the optimiser to jump out of a
-local optimum. The algorithm \ref{simulated-annealing} in the appendix
-highlights how SA differs from CRS in the inner loop.
+local optimum. The second algorithm in the appendix highlights how SA
+differs from CRS in the inner loop.
 
@@
 candidate basis \(\mathbf{A}_{**}\) is returned for the current
-iteration if it has a higher index value than the current one. Algorithm
-\ref{search-geodesic} in the appendix summarises the inner loop of the
-PD.
+iteration if it has a higher index value than the current one. The third
+algorithm in the appendix summarises the inner loop of the PD.
 
@@
 
-Figure @ref(fig:toy-search) shows an example of the search plot for CRS
+Figure \ref{fig:toy-search} shows an example of the search plot for CRS
 (left) and SA (right). Both optimisers quickly find better bases in the
@@
 There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
-The color scale is from the customised botanical palette in the ferrn package.}(\#fig:toy-search)
+The color scale is from the customised botanical palette in the ferrn package.}\label{fig:toy-search}
 \end{figure}
@@
 indicates that a non-monotonic interpolation cannot be avoided, even for
-CRS. Later, in section \ref{monotonic}, there will be a discussion on
-improving the non-monotonic interpolation for CRS.
+CRS. Later, in Section \emph{A problem of non-monotonicity}, there will
+be a discussion on improving the non-monotonic interpolation for CRS.
 
@@
 Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}(\#fig:toy-interp)
+Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 \CRANpkg{geozoo} \citep{geozoo} package. We use PCA to project and
-visualize the parameters/bases in 2D. The centre of the 2D view is the
+visualize the parameters/ bases in 2D. The centre of the 2D view is the
 first two PCs of the data matrix. It theoretically should be a circle,
@@
 for CRS and it describes the interruption of the interpolation, which
-will be discussed in detail in Section \ref{monotonic}.
+will be discussed in detail in Section \emph{A problem of
+non-monotonicity}.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}
 
@@
 The black star represents the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}(\#fig:toy-pca)
+All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
 \end{figure}
@@
 progresses from start to finish in the space. Figure
-\ref{fig:toy-pca-animated} shows six frames from the animation of the
-PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
-one can learn from this animation is that CRS finds its end basis
-quicker than PD since CRS finishes its search in the 5th frame while PD
-is still making more progress.
+\ref{fig:toy-pca-animated} shows the animated version (six frames from
+the animation if viewed in pdf) of the PCA plot in Figure
+\ref{fig:toy-pca}. An additional piece of information one can learn from
+this animation is that CRS finds its end basis quicker than PD since CRS
+finishes its search in the 5th frame while PD is still making more
+progress.
 
@@
 CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
-The full video of the animation can be found in the html version of the paper.}(\#fig:toy-pca-animated)
+The full video of the animation can be found in the html version of the paper.}\label{fig:toy-pca-animated}
 \end{figure}
@@
 original high-dimensional space from various directions using animation.
-Figure \ref{fig:toy-torus} shows some frames from the tour plot of the
+Figure \ref{fig:toy-tour} shows some frames from the tour plot of the
 same two optimisations in its original space.
@@
 
-\caption[Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated)]{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \@ref(fig:toy-pca) and \@ref(fig:toy-pca-animated).
+\caption{Six frames selected from rotating the high-dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}.
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
-The full animation can be seen in the html version of the paper.}(\#fig:toy-tour)
+The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
 \end{figure}
@@
 The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
-The full video of the animation can be found in the html version of the paper.}(\#fig:toy-torus)
+The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
 \end{figure}
@@
 On the left panel, basis with a higher index value is found during the interpolation but not used.
-On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}(\#fig:interruption)
+On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
 \end{figure}
@@
 Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best.
-}(\#fig:polish)
+}\label{fig:polish}
 \end{figure}
@@
 Both CRS and SA succeed in the optimisation, PD fails to optimise this non-smooth index.
-Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}(\#fig:noisy-better-geo)
+Further, SA takes  much longer than CRS to finish the optimisation, but finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
 \end{figure}
@@
 \caption[Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space faceted by two optimisers: CRS and SA, and two search sizes: 0.5 and 0.7.
-The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}(\#fig:kol-result)
+The optimisation is on the 1D projection index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
 \end{figure}
@@
 On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
-The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}(\#fig:flip-sign)
+The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
 \end{figure}
@@
     \texttt{explore\_trace\_search()} produces summary plots in Figure
-    @ref(fig:toy-search).
+    \ref{fig:toy-search}.
   \item
@@
     interpolation points in Figure \ref{fig:toy-interp}.
-  \item
-    \texttt{explore\_space\_pca()} produces the PCA plot of projection
-    bases on the reduced space in Figure \ref{fig:toy-pca}. The animated
-    version in Figure \ref{fig:toy-pca-animated} can be turned on via
-    the argument \texttt{animate\ =\ TRUE}.
   \item
@@
 
-\begin{verbatim}
-\end{verbatim}
-
 \bibliography{zhang-cook-laa-langrene-menendez.bib}

### 2021-09-30 paper/script.R

@@ -577,7 +577,7 @@ knitr::include_graphics(here::here("figs", "noisy-better-geo.svg"))
 #         legend.text = element_text(size = "10pt"))
 #
 # ggsave(p, filename = "kol-result.svg",
-#        path = here::here("figs"))
+#        path = here::here("figs"), width = 10, height = 10)
 knitr::include_graphics(here::here("figs", "kol-result.svg"))
 
 ## ---- flip-sign
@@ -598,7 +598,7 @@ knitr::include_graphics(here::here("figs", "kol-result.svg"))
 #
 # save(orientation_corrected, file = here::here("data", "orientation_corrected.rda"))
 # save(orientation_different, file = here::here("data", "orientation_different.rda"))
-
+#
 # dt <- bind_rows(orientation_corrected %>% mutate(sign = "flipped"),
 #                 orientation_different %>% mutate(sign = "original")) %>%
 #   compute_pca() %>%
@@ -614,5 +614,5 @@ knitr::include_graphics(here::here("figs", "kol-result.svg"))
 #         legend.text = element_text(size = "10pt"))
 #
 # ggsave(p, filename = "flip-sign.svg",
-#        path = here::here("figs"), width = 10, height = 5)
+#        path = here::here("figs"))
 knitr::include_graphics(here::here("figs", "flip-sign.svg"))

### 2021-09-30 paper/zhang-cook-laa-langrene-menendez.R

@@ -8,8 +8,7 @@ knitr::opts_chunk$set(echo = FALSE,
                       fig.align = "center",
                       fig.height = 3,
                       fig.retina = 5,
-                      out.width = "100%",
-                      fig.cap = TRUE)
+                      out.width = "100%")
 
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------
@@ -25,7 +24,7 @@ knitr::read_chunk(
 #spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
 
 
-## ----tour-path, out.width="70%", fig.cap="\\label{fig:tour-path} An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
+## ----tour-path, fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
 
 
 ## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the ferrn package."----
@@ -34,20 +33,20 @@ knitr::read_chunk(
 ## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."----
 
 
[…]

### 2021-10-01 review/response.Rmd
last change to paper and response to reviewer

@@
 ---
-title: "Response"
-author: "H. Sherry Zhang"
-date: "22/08/2021"
+title: "Response to reviewer"
+author: "Article ID: 2021-69"
 output: pdf_document
@@
 
-\textcolor{red}{Changed to "While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames."}
+\textcolor{red}{We keep the term "1D bases" and "2D basis" in the paper since 1) "frame" is conventinally "basis", and 2) we use this term also in the software.
+To make this clearer, a sentence is added at the end of section "Projection pursuit guided tour" to note this.}
 
@@
 
-3.  "with higher index value" -\> "with a higher index value". The word "the" is missing in a few other places in the paper.
+3.  "with higher index value" -\> "with a higher index value". The word "the" is missing in a few other places in the paper. \textcolor{red}{Corrected.}
 
-\textcolor{red}{Corrected.}
-
-4.  In Figure 3 would be better if the y-axes scales were the same for the two graphs.
-
-\textcolor{red}{Fixed.}
+4.  In Figure 3 would be better if the y-axes scales were the same for the two graphs. \textcolor{red}{Fixed.}
 
@@
 
-8.  "not the case for all the indexes" missing a "." page 11
-
-\textcolor{red}{Added.}
+8.  "not the case for all the indexes" missing a "." page 11 \textcolor{red}{Added.}
 
@@
 
-10. Fix references with "et al"
-
-\textcolor{red}{Fixed!}
+10. Fix references with "et al" \textcolor{red}{Fixed!}

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.Rmd
last change to paper and response to reviewer

@@
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
+Notice that the term *frame* used in Buja's paper is conventionally called *basis*, which is used in this paper and the associated software.
 The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
@@
 
-| $t$| $\mathbf{A}$| $I$| $S$| $j$| $l$| $V_{1}$| $V_{2}$| Notes|
+| $t$| $\mathbf{A}$| $I$| $S$| $j$| $l$| $V_{1}$| $V_{2}$| Description|
 |---|---|---|---|---|---|---|---|-------|
@@
 
-While the previous few examples have looked at the space of 1-frames in a unit sphere, this section visualises the space of 2-frames.
+While the previous few examples have looked at the space of 1D basis in a unit sphere, this section visualises the space of 2D basis.
 Recall that the columns in a 2D basis are orthogonal to each other, so the space of $p \times 2$ bases is a torus in the $p$-D space [@Buja1986-sw].

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.tex
last change to paper and response to reviewer

@@
 of the geodesic interpolation can be found in
-\citet{buja2005computational}. The aforementioned tour method has been
-implemented in the R package \CRANpkg{tourr} \citep{tourr}.
+\citet{buja2005computational}. Notice that the term \emph{frame} used in
+Buja's paper is conventionally called \emph{basis}, which is used in
+this paper and the associated software. The aforementioned tour method
+has been implemented in the R package \CRANpkg{tourr} \citep{tourr}.
 
@@
 \(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
-\(V_{2}\) & Notes \\
+\(V_{2}\) & Description \\
 \midrule
@@
 
-While the previous few examples have looked at the space of 1-frames in
-a unit sphere, this section visualises the space of 2-frames. Recall
+While the previous few examples have looked at the space of 1D basis in
+a unit sphere, this section visualises the space of 2D basis. Recall
 that the columns in a 2D basis are orthogonal to each other, so the

### 2021-10-01 review/response.Rmd
last tweaks

@@
 
-\textcolor{red}{We keep the term "1D bases" and "2D basis" in the paper since 1) "frame" is conventinally "basis", and 2) we use this term also in the software.
-To make this clearer, a sentence is added at the end of section "Projection pursuit guided tour" to note this.}
+\textcolor{red}{You are right that we could use frame instead of basis, and both are used conventionally.
+We have elected to stay with the use of basis, and have  some sentences in the section titled "Projection pursuit guided tour" to explain the terminology.}
 

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.Rmd
last tweaks

@@
 
-# Projection pursuit guided tour {#tour}
+# Projection pursuit guided tour
 
-Projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space.
+A projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space.
 Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
@@
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
-Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue frames are produced by the projection pursuit optimisation and the white frames are interpolations between the blue frames.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation and the white squares represent planes in the geodesic interpolation between targets.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-Notice that the term *frame* used in Buja's paper is conventionally called *basis*, which is used in this paper and the associated software.
-The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
+(Note that the term *frame* used in Buja's paper to refer to a particular set of orthonormal vectors defining a plane.
+This is also conventionally referred to as a basis, which is used in this paper and the associated software.) The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
 

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.tex
last tweaks

@@
 
-\hypertarget{tour}{%
-\section{Projection pursuit guided tour}\label{tour}}
-
-Projection pursuit guided tour combines two different methods
+\hypertarget{projection-pursuit-guided-tour}{%
+\section{Projection pursuit guided
+tour}\label{projection-pursuit-guided-tour}}
+
+A projection pursuit guided tour combines two different methods
 (projection pursuit and guided tour) to explore interesting features in
@@
 \ref{fig:tour-path} shows a sketch of the tour path where the blue
-frames are produced by the projection pursuit optimisation and the white
-frames are interpolations between the blue frames. Mathematical details
-of the geodesic interpolation can be found in
-\citet{buja2005computational}. Notice that the term \emph{frame} used in
-Buja's paper is conventionally called \emph{basis}, which is used in
-this paper and the associated software. The aforementioned tour method
-has been implemented in the R package \CRANpkg{tourr} \citep{tourr}.
+squares represent planes (targets) selected by the projection pursuit
+optimisation and the white squares represent planes in the geodesic
+interpolation between targets. Mathematical details of the geodesic
+interpolation can be found in \citet{buja2005computational}. (Note that
+the term \emph{frame} used in Buja's paper to refer to a particular set
+of orthonormal vectors defining a plane. This is also conventionally
+referred to as a basis, which is used in this paper and the associated
+software.) The aforementioned tour method has been implemented in the R
+package \CRANpkg{tourr} \citep{tourr}.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-search}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-search}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-interp}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}
+{\centering \includegraphics[width=0.7\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca-animated-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-pca-animated-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-tour-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-tour-1}
 
@@
 The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package geozoo.
-The full animation can be seen in the html version of the paper.}\label{fig:toy-tour}
+The full animation can be seen in the html version of the paper.}(\#fig:toy-tour)
 \end{figure}
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-torus-1}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/toy-torus-1}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interrupt}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/interrupt}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/polish}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/noisy-better-geo}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/kol-result}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/kol-result}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/flip-sign}
+{\centering \includegraphics[width=1\linewidth]{/Users/cookd/students_PhD/Sherry/paper-ferrn/figs/flip-sign}
 
@@
 
+\begin{verbatim}
+\end{verbatim}
+
 \bibliography{zhang-cook-laa-langrene-menendez.bib}

### 2021-10-01 paper/zhang-cook-laa-langrene-menendez.R

@@ -79,3 +79,5 @@ holes_2d_better_3var %>% get_best() %>% pull(basis) %>% .[[1]]
 
 ## ----flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears."----
 
+```{.r .distill-force-highlighting-css}
+```

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
copy edit changes & misc: yaml

@@
 author:
-  - name: H.Sherry Zhang
+  - name: H. Sherry Zhang
     affiliation: Monash University 
-    address: Department of Econometrics and Business Statistics 
-    url: Melbourne, Australia
+    address: 
+      - Department of Econometrics and Business Statistics
+      - Melbourne, Australia
     email: huize.zhang@monash.edu
@@
     affiliation: Monash University 
-    address: Department of Econometrics and Business Statistics
-    url: Melbourne, Australia
+    address: 
+      - Department of Econometrics and Business Statistics
+      - Melbourne, Australia
     email: dicook@monash.edu
@@
     affiliation: University of Natural Resources and Life Sciences
-    address: Institute of Statistics
-    url: Vienna, Austria
+    address: 
+      - Institute of Statistics
+      - Vienna, Austria
     email: ursula.laa@boku.ac.at  
@@
     affiliation: CSIRO Data61 
-    address: 34 Village Street, Docklands VIC 3008 Australia
-    url: Melbourne, Australia
+    address: 
+       - 34 Village Street, Docklands VIC 3008 Australia
+       - Melbourne, Australia
     email: nicolas.langrene@csiro.au
@@
     affiliation: Monash University 
-    address: Department of Econometrics and Business Statistics
-    url: Melbourne, Australia
+    address: 
+      - Department of Econometrics and Business Statistics
+      - Melbourne, Australia
     email: patricia.menendez@monash.edu 
@@
   A guided tour helps to visualise high-dimensional data by showing low-dimensional projections along a projection pursuit optimisation path.
-Projection pursuit is a generalisation of principal component analysis, in the sense that different indexes are used to define the interestingness of the projected data.
+Projection pursuit is a generalisation of principal component analysis in the sense that different indexes are used to define the interestingness of the projected data.
 While much work has been done in developing new indexes in the literature, less has been done on understanding the optimisation.
@@
 In addition, projection pursuit is primarily used for exploratory data analysis, and finding the local maxima is also useful.
-The guided tour is especially useful for exploration, because it conducts geodesic interpolation connecting steps in the optimisation and shows how the projected data changes as a maxima is approached.
-This work provides new visual diagnostics for examining a choice of optimisation procedure, based on the provision of a new data object which collects information throughout the optimisation.
+The guided tour is especially useful for exploration because it conducts geodesic interpolation connecting steps in the optimisation and shows how the projected data changes as a maxima is approached.
+This work provides new visual diagnostics for examining a choice of optimisation procedure based on the provision of a new data object which collects information throughout the optimisation.
 It has helped to diagnose and fix several problems with projection pursuit guided tour.
-This work might be useful more broadly for diagnosing optimisers, and comparing their performance.
+This work might be useful more broadly for diagnosing optimisers and comparing their performance.
 The diagnostics are implemented in the R package [ferrn](https://github.com/huizezhang-sherry/ferrn).
@@
 type: package
-csl: "biostats.csl"
-creative_commons: CC BY
 output: 
-  rjtools::rjournal_web_article:
+  rjtools::rjournal_web_article
 css: "style.css"
@@
 Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
-Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:tour-path}','\\@ref(fig:tour-path)'))` shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation and the white squares represent planes in the geodesic interpolation between targets.
+Figure \@ref(fig:tour-path)  shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation and the white squares represent planes in the geodesic interpolation between targets.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
copy edit changes & misc: yaml

--- 2021-10-01 paper/zhang-cook-laa-langrene-menendez.tex
@@
 to Guided Tours}
-\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
+\author{by H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas
 Langrené, and Patricia Menéndez}
@@
 path. Projection pursuit is a generalisation of principal component
-analysis, in the sense that different indexes are used to define the
+analysis in the sense that different indexes are used to define the
 interestingness of the projected data. While much work has been done in
@@
 exploratory data analysis, and finding the local maxima is also useful.
-The guided tour is especially useful for exploration, because it
-conducts geodesic interpolation connecting steps in the optimisation and
-shows how the projected data changes as a maxima is approached. This
-work provides new visual diagnostics for examining a choice of
-optimisation procedure, based on the provision of a new data object
-which collects information throughout the optimisation. It has helped to
-diagnose and fix several problems with projection pursuit guided tour.
-This work might be useful more broadly for diagnosing optimisers, and
-comparing their performance. The diagnostics are implemented in the R
-package \href{https://github.com/huizezhang-sherry/ferrn}{ferrn}.
+The guided tour is especially useful for exploration because it conducts
+geodesic interpolation connecting steps in the optimisation and shows
+how the projected data changes as a maxima is approached. This work
+provides new visual diagnostics for examining a choice of optimisation
+procedure based on the provision of a new data object which collects
+information throughout the optimisation. It has helped to diagnose and
+fix several problems with projection pursuit guided tour. This work
+might be useful more broadly for diagnosing optimisers and comparing
+their performance. The diagnostics are implemented in the R package
+\href{https://github.com/huizezhang-sherry/ferrn}{ferrn}.
 }
@@
 interpolation between these planes to form a tour path. Figure
-\ref{fig:tour-path} shows a sketch of the tour path where the blue
-frames are produced by the projection pursuit optimisation and the white
-frames are interpolations between the blue frames. Mathematical details
-of the geodesic interpolation can be found in
-\citet{buja2005computational}. Notice that the term \emph{frame} used in
-Buja's paper is conventionally called \emph{basis}, which is used in
-this paper and the associated software. The aforementioned tour method
-has been implemented in the R package \CRANpkg{tourr} \citep{tourr}.
+@ref(fig:tour-path) shows a sketch of the tour path where the blue
+squares represent planes (targets) selected by the projection pursuit
+optimisation and the white squares represent planes in the geodesic
+interpolation between targets. Mathematical details of the geodesic
+interpolation can be found in \citet{buja2005computational}. (Note that
+the term \emph{frame} used in Buja's paper to refer to a particular set
+of orthonormal vectors defining a plane. This is also conventionally
+referred to as a basis, which is used in this paper and the associated
+software.) The aforementioned tour method has been implemented in the R
+package \CRANpkg{tourr} \citep{tourr}.
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.10}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.23}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.0968}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.2258}}@{}}
 \toprule
-\(t\) & \(\mathbf{A}\) & \(I\) & \(S\) & \(j\) & \(l\) & \(V_{1}\) &
-\(V_{2}\) & Description \\
+\begin{minipage}[b]{\linewidth}\raggedright
+\(t\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(\mathbf{A}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(I\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(S\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(j\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(l\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(V_{1}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(V_{2}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Description
+\end{minipage} \\
 \midrule
@@
 \address{%
-H.Sherry Zhang\\
+H. Sherry Zhang\\
 Monash University\\%
-Department of Econometrics and Business Statistics\\
-%
-Melbourne, Australia\\%
+Department of Econometrics and Business Statistics\\ Melbourne,
+Australia\\
+%
+%
 %
@@
 Monash University\\%
-Department of Econometrics and Business Statistics\\
-%
-Melbourne, Australia\\%
+Department of Econometrics and Business Statistics\\ Melbourne,
+Australia\\
+%
+%
 %
@@
 University of Natural Resources and Life Sciences\\%
-Institute of Statistics\\
-%
-Vienna, Austria\\%
+Institute of Statistics\\ Vienna, Austria\\
+%
+%
 %
@@
 CSIRO Data61\\%
-34 Village Street, Docklands VIC 3008 Australia\\
-%
-Melbourne, Australia\\%
+34 Village Street, Docklands VIC 3008 Australia\\ Melbourne, Australia\\
+%
+%
 %

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.R

@@ -79,5 +79,3 @@ holes_2d_better_3var %>% get_best() %>% pull(basis) %>% .[[1]]
 
 ## ----flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears."----
 
-```{.r .distill-force-highlighting-css}
-```

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
before diagnostic 1

@@
   \usepackage[ruled, linesnumbered]{algorithm2e}
+  \usepackage{longtable}
 type: package
@@
 
-```{r }
-#spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
-```
-
 # Introduction
@@
 Presenting information in graphics often unveils insights that would otherwise not be discovered and provides a more comprehensive understanding of the problem at hand.
-Task specific tools such as @li2020visualizing show how visualisation can be used to understand, for instance, the behaviour of the optimisation for the example of neural network classification models.
+Task-specific tools such as @li2020visualizing show how visualisation can be used to understand, for instance, the behaviour of the optimisation for the example of neural network classification models.
 However, no general visualisation tool is available for diagnosing optimisation procedures.
 The work presented in this paper brings visualization tools into optimisation problems with the aim to better understand the performance of optimisers in practice.
-
-<!-- The goal of continuous optimisation is to find the best solution within the space of all feasible solutions where typically the best solution is defined via an objective function.
-Broadly speaking, optimization can be unconstrained or constrained [@kelley1999iterative].
-The unconstrained problem can be formulated as a minimization (or maximization)  problem such as  -->
-
-<!-- $\min_{x} f(x)$ where $f:\mathbb{R}^n \rightarrow \mathbb{R}$ is an objective function with certain properties defined in an $L^p$ space.
-In this case, solutions rely on gradient descent or ascent methods.
-In the constrained optimization problem additional restrictions  are introduced via a set of functions that can be convex or non-convex: $g_i:\mathbb{R}^n \rightarrow \mathbb{R}$ for $i = 1, \ldots k$ and hence the problem can be written as -->
-
-<!-- $\min_{x} f(x)$ *subject to* $g_i(x) \leq 0$. Here methods such as multipliers and convex optimization methods including linear and quadratic programming can be used. -->
 
@@
 Examples include the existence of multiple optima (local and global), the trade-off between computational burden and proximity to the optima, or dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient].
-The visualization tools, optimization methods and conceptual framework presented in this paper can therefore be applied to other optimization problems.
+The visualization tools, optimization methods, and conceptual framework presented in this paper can therefore be applied to other optimization problems.
 
@@
 The next section provides an overview of optimisation methods, specifically random search and line search methods.
-A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
+A review of the projection pursuit guided tour, an overview of the optimisation problem and, outlines of three existing algorithms follows.
 The third section presents the new visual diagnostics, including the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
@@
 
-<!-- Optimisation problems are ubiquitous in many areas of study.
-While in some cases analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution.
-These numerical methods follow iterative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved.
-Although this principle seems uncomplicated, a number of challenges arise such as the possible existence of multiple maxima (local and global), constraints and noisy objective function, and the trade-off between desirable accuracy and computational burden.
-In addition, the optimization results might depend on the algorithm starting values, affecting the consistency of results.
--->
-
 The type of optimisation problem considered in this paper is constrained optimization [@bertsekas2014constrained], assuming it is not possible to find a solution to the problem in the way of a closed-form.
-That is, the problem consists in finding the minimum or maximum of a function $f \in L^p$ in the constrained space $\mathcal{A}$, where $L^p$ defines the vector space of function $f$, whose $p$th power are integrable.
+That is, the problem consists in finding the minimum or maximum of a function $f \in L^p$ in the constrained space $\mathcal{A}$, where $L^p$ defines the vector space of function $f$, whose $p$th power is integrable.
 
@@
 In addition, some objective functions rely on the empirical distribution of the data, which makes it in general not possible to get the gradient.
-Hence, gradient-based methods are not the focus of this paper and consideration will be given to derivative-free methods.
+Hence, gradient-based methods are not the focus of this paper, and consideration will be given to derivative-free methods.
 
 Derivative-free methods [@conn2009introduction; @rios2013derivative], which do not rely on the knowledge of the gradient, are more generally applicable.
-Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near optimal solution.
+Derivative-free methods have been developed over the years, where the emphasis is on finding, in most cases, a near-optimal solution.
 Here we consider three derivative-free methods, two of which are random search methods: creeping random search and simulated annealing, and the other one is pseudo-derivative search.
 
-Random search methods [@Romeijn2009; @zabinsky2013stochastic; @andradottir2015review] have a random sampling component as part of their algorithm and have been shown to have the ability to optimise non-convex and non-smooth functions.
+Random search methods [@Romeijn2009; @zabinsky2013stochastic; @andradottir2015review] have a random sampling component as part of their algorithms and have been shown to have the ability to optimise non-convex and non-smooth functions.
 The initial random search algorithm, pure random search [@Brooks1958-zk], draws candidate points from the entire space without using any information of the current position and updates the current position when an improvement on the objective function is made.
@@
 Creeping random search [@White1971-ci] is such a variation, where a candidate point is generated within a neighbourhood of the current point.
-This makes creeping random search faster to compute, however, global convergence is no longer guaranteed.
+This makes creeping random search faster to compute but global convergence is no longer guaranteed.
 On the other hand, simulated annealing [@kirkpatrick1983optimization; @bertsimas1993simulated], introduces a different updating mechanism.
-Rather than only updating the current point when an improvement is made, simulated annealing uses a Metropolis acceptance criterion, where worse candidates still have a chance to be accepted.
+Rather than only updating the current point when an improvement is made, it uses a Metropolis acceptance criterion, where worse candidates still have a chance to be accepted.
 The convergence of simulated annealing algorithms has been widely researched [@mitra1986convergence; @granville1994simulated] and the global optimum can be attained under mild regularity conditions.
 
-Pseudo-derivative search uses a common search scheme in optimisation: line search.
+The pseudo-derivative search uses a common search scheme in optimisation: line search.
 In line search methods, users are required to provide an initial estimate $x_{1}$ and, at each iteration, a search direction $S_k$ and a step size $\alpha_k$ are generated.
@@
 In derivative-free methods, local information of the objective function is used to determine the search direction.
-The choice of step size also needs consideration, as inadequate step sizes might prevent the optimisation method to converge to an optimum.
+The choice of step size also needs consideration, as inadequate step sizes might prevent the optimisation method from converging to an optimum.
 An ideal step size can be chosen by finding the value of $\alpha_k \in \mathbb{R}$ that maximises $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ at each iteration.
-
-<!-- Several R implementations address optimization problems with both general purpose as well as task specifics solvers.
-The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions.
-Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr].
-Specific solvers for simulated annealing include \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions.
-For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].
--->
 
@@
 A projection pursuit guided tour combines two different methods (projection pursuit and guided tour) to explore interesting features in a high-dimensional space.
-Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g.
-clustering, outliers and skewness) in multivariate data via low-dimensional projections.
+Projection pursuit, coined by @friedman1974projection, detects interesting structures (e.g., clustering, outliers, and skewness) in multivariate data via low-dimensional projections.
 Guided tour [@cook1995grand] is one variation of a broader class of data visualisation methods, tour [@buja2005computational], which displays high-dimensional data through a series of animated projections.
@@
 This is commonly known as the projection pursuit index function, or just index function, and is used to measure the "interestingness" of a given projection.
-An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal, as $n$ and $p$ approach infinity under certain conditions.
-There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], and natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag; @WW08].
+An interesting projection shows structures that are non-normal since theoretical proofs from @diaconis1984asymptotics have shown that projections tend to be normal as $n$ and $p$ approach infinity under certain conditions.
+There have been many index functions proposed in the literature, here are a few examples: early indexes that can be categorised as measuring the $L^2$ distance between the projection and a normal distribution: Legendre index [@friedman1974projection]; Hermite index [@hall1989polynomial]; natural Hermite index [@cook1993projection]; chi-square index [@posse1995projection] for detecting spiral structure; LDA index [@lee2005projection] and PDA [@lee2010projection] index for supervised classification; kurtosis index [@Loperfido2020] and skewness index [@Loperfido2018] for detecting outliers in financial time series; and most recently, scagnostic indexes [@laa2020using] for summarising structures in scatterplot matrices based on eight scagnostic measures [@scag; @WW08].
 
@@
 There are different versions depending on how the high-dimensional space is investigated: grand tour [@cook2008grand] selects the planes randomly to provide a general overview; manual tour [@cook1997manual] gradually phases in and out one variable to understand the contribution of that variable in the projection.
-Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit, to gradually reveal the most interesting projection.
-Given a random start, projection pursuit iteratively finds bases with higher index values and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
-Figure \@ref(fig:tour-path)  shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation and the white squares represent planes in the geodesic interpolation between targets.
+Guided tour, the main interest of this paper, chooses the planes with the aid of projection pursuit to gradually reveal the most interesting projection.
+Given a random start, projection pursuit iteratively finds bases with higher index values, and the guided tour constructs a geodesic interpolation between these planes to form a tour path.
+Figure \@ref(fig:tour-path)  shows a sketch of the tour path where the blue squares represent planes (targets) selected by the projection pursuit optimisation, and the white squares represent planes in the geodesic interpolation between targets.
 Mathematical details of the geodesic interpolation can be found in @buja2005computational.
-(Note that the term *frame* used in Buja's paper to refer to a particular set of orthonormal vectors defining a plane.
+(Note that the term *frame* used in Buja's paper refers to a particular set of orthonormal vectors defining a plane.
 This is also conventionally referred to as a basis, which is used in this paper and the associated software.) The aforementioned tour method has been implemented in the R package \CRANpkg{tourr} [@tourr].
@@
 
-In projection pursuit the optimisation aims at finding the global and local maxima that give interesting projections according to an index function.
+In projection pursuit, the optimisation aims at finding the global and local maxima that give interesting projections according to an index function.
 That is, it starts with a given randomly selected basis $\mathbf{A}_1$ and aims at finding an optimal final projection basis $\mathbf{A}_T$ that satisfies the following optimisation problem:
@@
 \begin{align}
-\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A})  ~~~ s.t. ~~~ \mathbf{A}^{\prime} \mathbf{A} = I_d
+\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A})  ~~~ s.t. ~~~ \mathbf{A}^{\prime} \mathbf{A} = I_d \,,
 \end{align}
@@
 \noindent where $f$ and $\mathbf{X}$ are defined as in the previous section, $\mathcal{A}$ is the set of all $p$-dimensional projection bases, $I_d$ is the $d$-dimensional identity matrix, and the constraint ensures the projection bases, $\mathbf{A}$, to be orthonormal.
-It is worth noticing the following: 1) The optimisation is constrained and the orthonormality constraint imposes a geometrical structure on the bases space: it forms a Stiefel manifold.
+It is worth noticing the following: 1) The optimisation is constrained, and the orthonormality constraint imposes a geometrical structure on the bases space: it forms a Stiefel manifold.
 2) There may be index functions for which the objective function might not be differentiable.
@@
 $\mathbf{A}_{l}$ is then orthonormalised to fulfil the basis constraint.
-If $\mathbf{A}_{l}$ has an index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the optimiser outputs $\mathbf{A}_{l}$ for guided tour to construct an interpolation path.
+If $\mathbf{A}_{l}$ has an index value higher than the current basis $\mathbf{A}_{\text{cur}}$, the optimiser outputs $\mathbf{A}_{l}$ for a guided tour to construct an interpolation path.
 The neighbourhood parameter $\alpha$ is adjusted by a cooling parameter: $\alpha_{j+1} = \alpha_j * \text{cooling}$ before the next iteration starts.
@@
 @posse1995projection has proposed a slightly different cooling scheme by introducing a halving parameter $c$.
-In his proposal $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
+In his proposal, $\alpha$ is only adjusted if the last iteration takes more than $c$ times to find a better basis.
 
@@
 Given an initial value of $T_0 \in \mathbb{R^{+}}$, the "temperature" at iteration $l$ is defined as $T(l) = \frac{T_0}{\log(l + 1)}$.
-When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}$$ where $I_{(\cdot)} \in \mathbb{R}$ denotes the index value of a given basis.
-This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value and hence enables the optimiser to jump out of a local optimum.
+When a candidate basis fails to have an index value larger than the current basis, SA gives it a second chance to be accepted with probability $$P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\} \,,$$ where $I_{(\cdot)} \in \mathbb{R}$ denotes the index value of a given basis.
+This implementation allows the optimiser to make a move and explore the basis space even if the candidate basis does not have a higher index value.
+Hence it enables the optimiser to jump out of a local optimum.
 The second algorithm in the appendix highlights how SA differs from CRS in the inner loop.
@@
 Rather than randomly sample the basis space, PD first computes a search direction by evaluating bases close to the current basis.
-A step size is then chosen along the corresponding geodesic by another optimisation over a 90 degree angle from $-\pi/4$ to $\pi/4$.
+The step size is then chosen along the corresponding geodesic by another optimisation over a 90 degree angle from $-\pi/4$ to $\pi/4$.
 The resulting candidate basis $\mathbf{A}_{**}$ is returned for the current iteration if it has a higher index value than the current one.
@@
 For CRS and SA, possible states include `random_search`, `new_basis`, and `interpolation`.
-Pseudo-derivative (PD) has a wider variety of states including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
-Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation; $j$ and $l$ are the counter of the outer and inner loop respectively.
+Pseudo-derivative (PD) has a wider variety of states, including `new_basis`, `direction_search`, `best_direction_search`, `best_line_search`, and `interpolation`.
+Multiple iterators index the information collected at different levels: $t$ is a unique identifier prescribing the natural ordering of each observation; $j$ and $l$ are the counter of the outer and inner loop, respectively.
 Other parameters of interest recorded, $V$, include `method` that tags the name of the optimiser, and `alpha` that indicates the sampling neighbourhood size for searching observations.
@@
 
+Table: Data structure of the information recorded for visual diagnostics.
+
 \noindent where $T^{\prime} = T + k_{J}+ l_{J+1}$.
@@
 
-```{=html}
-<!-- One may want to simply plot the index value of the search points across its natural order.
-While such a point geometry may work well if each iteration has a similar number of points, when some iterations have considerably more points than others, it will over-emphasizes the iterations that have more search points and these iterations will occupy the vast majority of the plot space.
-An alternative is to summarise the search in each iteration using boxplots, with each iteration spaced out equally.
-Occasionally, one may still want to switch back to a point geometry if the number of points is small in a particular iteration and this can be achieved via the `cutoff` argument in the search plot function.
-Additional annotations are added to facilitate better reading of the plot and these include
-1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a grey scale to indicate no better basis found in this iteration.
--->
-```
 Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))` shows an example of the search plot for CRS (left) and SA (right).

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
before diagnostic 1

@@
 otherwise not be discovered and provides a more comprehensive
-understanding of the problem at hand. Task specific tools such as
+understanding of the problem at hand. Task-specific tools such as
 \citet{li2020visualizing} show how visualisation can be used to
@@
 non-differentiable \citep{jones1998efficient}. The visualization tools,
-optimization methods and conceptual framework presented in this paper
+optimization methods, and conceptual framework presented in this paper
 can therefore be applied to other optimization problems.
@@
 and line search methods. A review of the projection pursuit guided tour,
-an overview of the optimisation problem and outlines of three existing
+an overview of the optimisation problem and, outlines of three existing
 algorithms follows. The third section presents the new visual
@@
 \(L^p\) defines the vector space of function \(f\), whose \(p\)th power
-are integrable.
+is integrable.
 
@@
 possible to get the gradient. Hence, gradient-based methods are not the
-focus of this paper and consideration will be given to derivative-free
+focus of this paper, and consideration will be given to derivative-free
 methods.
@@
 Derivative-free methods have been developed over the years, where the
-emphasis is on finding, in most cases, a near optimal solution. Here we
+emphasis is on finding, in most cases, a near-optimal solution. Here we
 consider three derivative-free methods, two of which are random search
@@
 \citep{Romeijn2009, zabinsky2013stochastic, andradottir2015review} have
-a random sampling component as part of their algorithm and have been
+a random sampling component as part of their algorithms and have been
 shown to have the ability to optimise non-convex and non-smooth
@@
 generated within a neighbourhood of the current point. This makes
-creeping random search faster to compute, however, global convergence is
-no longer guaranteed. On the other hand, simulated annealing
+creeping random search faster to compute but global convergence is no
+longer guaranteed. On the other hand, simulated annealing
 \citep{kirkpatrick1983optimization, bertsimas1993simulated}, introduces
 a different updating mechanism. Rather than only updating the current
-point when an improvement is made, simulated annealing uses a Metropolis
-acceptance criterion, where worse candidates still have a chance to be
-accepted. The convergence of simulated annealing algorithms has been
-widely researched \citep{mitra1986convergence, granville1994simulated}
-and the global optimum can be attained under mild regularity conditions.
-
-Pseudo-derivative search uses a common search scheme in optimisation:
-line search. In line search methods, users are required to provide an
-initial estimate \(x_{1}\) and, at each iteration, a search direction
-\(S_k\) and a step size \(\alpha_k\) are generated. Then one moves on to
-the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and the process
-is repeated until the desired convergence is reached. In derivative-free
-methods, local information of the objective function is used to
-determine the search direction. The choice of step size also needs
-consideration, as inadequate step sizes might prevent the optimisation
-method to converge to an optimum. An ideal step size can be chosen by
-finding the value of \(\alpha_k \in \mathbb{R}\) that maximises
-\(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) at each iteration.
+point when an improvement is made, it uses a Metropolis acceptance
+criterion, where worse candidates still have a chance to be accepted.
+The convergence of simulated annealing algorithms has been widely
+researched \citep{mitra1986convergence, granville1994simulated} and the
+global optimum can be attained under mild regularity conditions.
+
+The pseudo-derivative search uses a common search scheme in
+optimisation: line search. In line search methods, users are required to
+provide an initial estimate \(x_{1}\) and, at each iteration, a search
+direction \(S_k\) and a step size \(\alpha_k\) are generated. Then one
+moves on to the next point following \(x_{k+1} = x_k + \alpha_kS_k\) and
+the process is repeated until the desired convergence is reached. In
+derivative-free methods, local information of the objective function is
+used to determine the search direction. The choice of step size also
+needs consideration, as inadequate step sizes might prevent the
+optimisation method from converging to an optimum. An ideal step size
+can be chosen by finding the value of \(\alpha_k \in \mathbb{R}\) that
+maximises \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) at each
+iteration.
 
@@
 a high-dimensional space. Projection pursuit, coined by
-\citet{friedman1974projection}, detects interesting structures
-(e.g.~clustering, outliers and skewness) in multivariate data via
+\citet{friedman1974projection}, detects interesting structures (e.g.,
+clustering, outliers, and skewness) in multivariate data via
 low-dimensional projections. Guided tour \citep{cook1995grand} is one
@@
 \citet{diaconis1984asymptotics} have shown that projections tend to be
-normal, as \(n\) and \(p\) approach infinity under certain conditions.
+normal as \(n\) and \(p\) approach infinity under certain conditions.
 There have been many index functions proposed in the literature, here
@@
 the \(L^2\) distance between the projection and a normal distribution:
-Legendre index \citep{friedman1974projection}, Hermite index
-\citep{hall1989polynomial}, and natural Hermite index
+Legendre index \citep{friedman1974projection}; Hermite index
+\citep{hall1989polynomial}; natural Hermite index
 \citep{cook1993projection}; chi-square index \citep{posse1995projection}
@@
 tour, the main interest of this paper, chooses the planes with the aid
-of projection pursuit, to gradually reveal the most interesting
+of projection pursuit to gradually reveal the most interesting
 projection. Given a random start, projection pursuit iteratively finds
-bases with higher index values and the guided tour constructs a geodesic
-interpolation between these planes to form a tour path. Figure
+bases with higher index values, and the guided tour constructs a
+geodesic interpolation between these planes to form a tour path. Figure
 @ref(fig:tour-path) shows a sketch of the tour path where the blue
 squares represent planes (targets) selected by the projection pursuit
-optimisation and the white squares represent planes in the geodesic
+optimisation, and the white squares represent planes in the geodesic
 interpolation between targets. Mathematical details of the geodesic
 interpolation can be found in \citet{buja2005computational}. (Note that
-the term \emph{frame} used in Buja's paper to refer to a particular set
-of orthonormal vectors defining a plane. This is also conventionally
+the term \emph{frame} used in Buja's paper refers to a particular set of
+orthonormal vectors defining a plane. This is also conventionally
 referred to as a basis, which is used in this paper and the associated
@@
 
-In projection pursuit the optimisation aims at finding the global and
+In projection pursuit, the optimisation aims at finding the global and
 local maxima that give interesting projections according to an index
@@
 orthonormal. It is worth noticing the following: 1) The optimisation is
-constrained and the orthonormality constraint imposes a geometrical
+constrained, and the orthonormality constraint imposes a geometrical
 structure on the bases space: it forms a Stiefel manifold. 2) There may
@@
 current basis \(\mathbf{A}_{\text{cur}}\), the optimiser outputs
-\(\mathbf{A}_{l}\) for guided tour to construct an interpolation path.
+\(\mathbf{A}_{l}\) for a guided tour to construct an interpolation path.
 The neighbourhood parameter \(\alpha\) is adjusted by a cooling
@@
 \citet{posse1995projection} has proposed a slightly different cooling
-scheme by introducing a halving parameter \(c\). In his proposal
+scheme by introducing a halving parameter \(c\). In his proposal,
 \(\alpha\) is only adjusted if the last iteration takes more than \(c\)
@@
 chance to be accepted with probability
-\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
+\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\} \,,\]
 where \(I_{(\cdot)} \in \mathbb{R}\) denotes the index value of a given
@@
 explore the basis space even if the candidate basis does not have a
-higher index value and hence enables the optimiser to jump out of a
+higher index value. Hence it enables the optimiser to jump out of a
 local optimum. The second algorithm in the appendix highlights how SA
@@
 Rather than randomly sample the basis space, PD first computes a search
-direction by evaluating bases close to the current basis. A step size is
-then chosen along the corresponding geodesic by another optimisation
+direction by evaluating bases close to the current basis. The step size
+is then chosen along the corresponding geodesic by another optimisation
 over a 90 degree angle from \(-\pi/4\) to \(\pi/4\). The resulting
@@
 \texttt{interpolation}. Pseudo-derivative (PD) has a wider variety of
-states including \texttt{new\_basis}, \texttt{direction\_search},
+states, including \texttt{new\_basis}, \texttt{direction\_search},
 \texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
@@
 the natural ordering of each observation; \(j\) and \(l\) are the
-counter of the outer and inner loop respectively. Other parameters of
+counter of the outer and inner loop, respectively. Other parameters of
 interest recorded, \(V\), include \texttt{method} that tags the name of
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 16\tabcolsep) * \real{0.2258}}@{}}
+\caption{Data structure of the information recorded for visual
+diagnostics.}\tabularnewline
+\toprule
+\begin{minipage}[b]{\linewidth}\raggedright
+\(t\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(\mathbf{A}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(I\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(S\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(j\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(l\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(V_{1}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\(V_{2}\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+Description
+\end{minipage} \\
+\midrule
+\endfirsthead
 \toprule

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.R

@@ -20,10 +20,6 @@ knitr::read_chunk(
 ## ----load-pkg-----------------------------------------------------------------
 
 
-## -----------------------------------------------------------------------------
-#spelling::spell_check_files(here::here("paper/zhang-cook-laa-langrene-menendez.Rmd"), lang = "en-GB", ignore = spelling::get_wordlist())
-
-
 ## ----tour-path, fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
 
 

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
section vis diagnose done: )

@@
 
-A starting point of diagnosing an optimiser is to understand how many searches it has conducted, i.e.
-we want to summarise how the index is increasing over iterations and how many basis need to be sampled at each iteration.
+A starting point of diagnosing an optimiser is to understand how many searches it has conducted, i.e., we want to summarise how the index is increasing over iterations and how many basis need to be sampled at each iteration.
 This is achieved using the function `explore_trace_search()`: a boxplot shows the distribution of index values for each try, where the accepted basis (corresponding to the highest index value) is always shown as a point.
-In the case of few tries at a given iteration, showing the data points directly may be preferred over the boxplot, this is controlled via the `cutoff` argument.
-Additional annotations are added to facilitate better reading of the plot and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; and 3) the colour of the last iteration is in a greyscale to indicate no better basis found in this iteration.
+When there are only few tries at a given iteration, showing the data points directly is preferred over the boxplot and this is controlled via the `cutoff` argument.
+Additional annotations are added to facilitate better reading of the plot, and these include 1) the number of points searched in each iteration can be added as text label at the bottom of each iteration; 2) the anchor bases to interpolate are connected and highlighted in a larger size; 3) the colour of the last iteration is in greyscale to indicate no better basis found in this iteration.
 
@@
 ```{r toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index.
-Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
-In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values.
+Both optimisers reach the final basis with a similar index value, while it takes SA longer to find the final basis.
+In the earlier iterations, optimisers quickly find a better basis to proceed, while in the later iterations, most sampled bases fail to make an improvement on the index value, and a boxplot is used to summarise the distribution of the index values.
 There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
-The color scale is from the customised botanical palette in the ferrn package."}
+The color scale is from the customised botanical palette in the **ferrn** package."}
 ```
@@
 Trace plots are created by plotting the index value against time.
-Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-interp}','\\@ref(fig:toy-interp)'))` presents the trace plot of the same optimisations as Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))` and one can observe that the trace is smooth in both cases.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-interp}','\\@ref(fig:toy-interp)'))` presents the trace plot of the same optimisations as Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-search}','\\@ref(fig:toy-search)'))`, and one can observe that the trace is smooth in both cases.
 It may seem bizarre at first sight that the interpolation sometimes passes bases with higher index values before it decreases to a lower target.
@@
 The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}.
-Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."}
+Lines indicate the interpolation, and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth, while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."}
 ```
@@
 For one-dimensional projections, this forms a $p$-dimensional sphere.
-A dimensionality reduction method, e.g.
-principal component analysis is applied to first project all the bases onto a 2D space.
+A dimensionality reduction method, e.g., principal component analysis, is applied to first project all the bases onto a 2D space.
 In a projection pursuit guided tour optimisation, there are various types of bases involved: 1) The starting basis; 2) The search bases that the optimiser evaluated to produce the anchor bases; 3) The anchor bases that have the highest index value in each iteration; 4) The interpolating bases on the interpolation path; and finally, 5) the end basis.
-The importance of these bases differs and the most important ones are the starting, interpolating, and end bases.
-Sometimes, two optimisers can start with the same basis but finish with bases of opposite sign.
-This happens because the projection is invariant to the orientation of the basis and so is the index value.
+The importance of these bases differs but the most important ones are the starting, interpolating, and end bases.
+Sometimes, two optimisers can start with the same basis but finish with bases of opposite signs.
+This happens because the projection is invariant to the orientation of the basis, and so is the index value.
 However, this creates difficulties for comparing the optimisers since the end bases will be symmetric to the origin.
@@
 
-Several annotations have been made to help understanding this plot.
-As mentioned previously, the original basis space is a high-dimensional sphere and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
+Several annotations have been made to help understand this plot.
+As mentioned previously, the original basis space is a high-dimensional sphere, and random bases on the sphere can be generated via the \CRANpkg{geozoo} [@geozoo] package.
 We use PCA to project and visualize the parameters/ bases in 2D.
 The centre of the 2D view is the first two PCs of the data matrix.
-It theoretically should be a circle, but may have some irregular edges due to finite sampling.
+It theoretically should be a circle but may have some irregular edges due to finite sampling.
 Thus the edge is smoothed by using a radius estimated as the largest distance from the centre to any basis.
 In the simulation, the theoretical best basis is known and can be labelled to compare how close to this that the optimisers stopped.
-Various aesthetics, i.e.
-size, alpha (transparency), and colour, are applicable to emphasize critical elements and adjust for the presentation.
-For example, anchor points and search points are less important and hence a smaller size and alpha are used.
-Alpha can also be applied on the interpolation paths to show the start to finish from transparent to opaque.
+Various aesthetics, i.e., size, alpha (transparency), and colour, are applicable to emphasize critical elements and adjust for the presentation.
+For example, anchor points and search points are less important, and hence a smaller size and alpha are used.
+Alpha can also be applied on the interpolation paths to show start to finish from transparent to opaque.
 
 Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca}','\\@ref(fig:toy-pca)'))` shows the PCA plot of CRS and PD for a 1D projection problem.
-Both optimisers find the optimum but PD gets closer.
+Both optimisers find the optimum, but PD gets closer.
 With the PCA plot, one can visually appreciate the nature of these two optimisers: PD first evaluates points in a small neighbourhood for a promising direction, while CRS evaluates points randomly in the search space to search for the next target.
-There are dashed lines annotated for CRS and it describes the interruption of the interpolation, which will be discussed in detail in Section *A problem of non-monotonicity*.
+There are dashed lines annotated for CRS, and it describes the interruption of the interpolation, which will be discussed in detail in Section *A problem of non-monotonicity*.
 
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star represents the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
+The black star represents the theoretical best basis the optimisers are aiming to find.
+All the bases in PD have been flipped for easier comparison of the final bases, and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
 ```
@@
 
-An animation is another type of display to show how the search progresses from start to finish in the space.
+Animation is another type of display to show how the search progresses from start to finish in the space.
 Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca-animated}','\\@ref(fig:toy-pca-animated-interactive)'))` shows the animated version (six frames from the animation if viewed in pdf) of the PCA plot in Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-pca}','\\@ref(fig:toy-pca)'))`.
@@
 The two circles of the torus can be observed to be perpendicular to each other and this can be linked back to the orthogonality condition.
-Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown respectively to match the previous plots.
-The final basis found by PD and CRS are shown in a larger shape and printed below respectively:
+Two paths from CRS and PD are plotted on top of the torus and coloured in green and brown, respectively, to match the previous plots.
+The final basis found by PD and CRS are shown in a larger shape and printed below, respectively:
 
@@
 
-Both optimisers have found the third variable in the first direction, and the second variable in the second direction.
-Note however the different orientation of the basis, following from the different sign in the second column.
-One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and are symmetric when projected onto the other (due the sign difference in the second column).
+Both optimisers have found the third variable in the first direction and the second variable in the second direction.
+Note, however, the different orientation of the basis, following from the different sign in the second column.
+One would expect to see this in the torus plot as the final bases match each other when projected onto one torus circle (due to the same sign in the first column) and are symmetric when projected onto the other (due to the different sign in the second column).
 In Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:toy-torus}','\\@ref(fig:toy-torus-interactive)'))`, this can be seen most clearly in frame 5 where the two circles are rotated into a line from our view.

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
section vis diagnose done: )

@@
 A starting point of diagnosing an optimiser is to understand how many
-searches it has conducted, i.e.~we want to summarise how the index is
+searches it has conducted, i.e., we want to summarise how the index is
 increasing over iterations and how many basis need to be sampled at each
@@
 index values for each try, where the accepted basis (corresponding to
-the highest index value) is always shown as a point. In the case of few
-tries at a given iteration, showing the data points directly may be
-preferred over the boxplot, this is controlled via the \texttt{cutoff}
-argument. Additional annotations are added to facilitate better reading
-of the plot and these include 1) the number of points searched in each
-iteration can be added as text label at the bottom of each iteration; 2)
-the anchor bases to interpolate are connected and highlighted in a
-larger size; and 3) the colour of the last iteration is in a greyscale
-to indicate no better basis found in this iteration.
+the highest index value) is always shown as a point. When there are only
+few tries at a given iteration, showing the data points directly is
+preferred over the boxplot and this is controlled via the
+\texttt{cutoff} argument. Additional annotations are added to facilitate
+better reading of the plot, and these include 1) the number of points
+searched in each iteration can be added as text label at the bottom of
+each iteration; 2) the anchor bases to interpolate are connected and
+highlighted in a larger size; 3) the colour of the last iteration is in
+greyscale to indicate no better basis found in this iteration.
 
@@
 \caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using the holes index.
-Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis.
-In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values.
+Both optimisers reach the final basis with a similar index value, while it takes SA longer to find the final basis.
+In the earlier iterations, optimisers quickly find a better basis to proceed, while in the later iterations, most sampled bases fail to make an improvement on the index value, and a boxplot is used to summarise the distribution of the index values.
 There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey.
-The color scale is from the customised botanical palette in the ferrn package.}\label{fig:toy-search}
+The color scale is from the customised botanical palette in the **ferrn** package.}\label{fig:toy-search}
 \end{figure}
@@
 value against time. Figure \ref{fig:toy-interp} presents the trace plot
-of the same optimisations as Figure \ref{fig:toy-search} and one can
+of the same optimisations as Figure \ref{fig:toy-search}, and one can
 observe that the trace is smooth in both cases. It may seem bizarre at
@@
 The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}.
-Lines indicate the interpolation and dots indicate new target bases generated by the optimisers.
-Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
+Lines indicate the interpolation, and dots indicate new target bases generated by the optimisers.
+Interpolation in both optimisation is smooth, while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130.}\label{fig:toy-interp}
 \end{figure}
@@
 one-dimensional projections, this forms a \(p\)-dimensional sphere. A
-dimensionality reduction method, e.g.~principal component analysis is
+dimensionality reduction method, e.g., principal component analysis, is
 applied to first project all the bases onto a 2D space. In a projection
@@
 interpolation path; and finally, 5) the end basis. The importance of
-these bases differs and the most important ones are the starting,
+these bases differs but the most important ones are the starting,
 interpolating, and end bases. Sometimes, two optimisers can start with
-the same basis but finish with bases of opposite sign. This happens
-because the projection is invariant to the orientation of the basis and
+the same basis but finish with bases of opposite signs. This happens
+because the projection is invariant to the orientation of the basis, and
 so is the index value. However, this creates difficulties for comparing
@@
 
-Several annotations have been made to help understanding this plot. As
+Several annotations have been made to help understand this plot. As
 mentioned previously, the original basis space is a high-dimensional
-sphere and random bases on the sphere can be generated via the
+sphere, and random bases on the sphere can be generated via the
 \CRANpkg{geozoo} \citep{geozoo} package. We use PCA to project and
 visualize the parameters/ bases in 2D. The centre of the 2D view is the
-first two PCs of the data matrix. It theoretically should be a circle,
+first two PCs of the data matrix. It theoretically should be a circle
 but may have some irregular edges due to finite sampling. Thus the edge
@@
 known and can be labelled to compare how close to this that the
-optimisers stopped. Various aesthetics, i.e.~size, alpha (transparency),
-and colour, are applicable to emphasize critical elements and adjust for
-the presentation. For example, anchor points and search points are less
-important and hence a smaller size and alpha are used. Alpha can also be
-applied on the interpolation paths to show the start to finish from
-transparent to opaque.
+optimisers stopped. Various aesthetics, i.e., size, alpha
+(transparency), and colour, are applicable to emphasize critical
+elements and adjust for the presentation. For example, anchor points and
+search points are less important, and hence a smaller size and alpha are
+used. Alpha can also be applied on the interpolation paths to show start
+to finish from transparent to opaque.
 
 Figure \ref{fig:toy-pca} shows the PCA plot of CRS and PD for a 1D
-projection problem. Both optimisers find the optimum but PD gets closer.
-With the PCA plot, one can visually appreciate the nature of these two
-optimisers: PD first evaluates points in a small neighbourhood for a
-promising direction, while CRS evaluates points randomly in the search
-space to search for the next target. There are dashed lines annotated
-for CRS and it describes the interruption of the interpolation, which
-will be discussed in detail in Section \emph{A problem of
-non-monotonicity}.
+projection problem. Both optimisers find the optimum, but PD gets
+closer. With the PCA plot, one can visually appreciate the nature of
+these two optimisers: PD first evaluates points in a small neighbourhood
+for a promising direction, while CRS evaluates points randomly in the
+search space to search for the next target. There are dashed lines
+annotated for CRS, and it describes the interruption of the
+interpolation, which will be discussed in detail in Section \emph{A
+problem of non-monotonicity}.
 
@@
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
-The black star represents the theoretical best basis the optimisers aiming to find.
-All the bases in PD have been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
+The black star represents the theoretical best basis the optimisers are aiming to find.
+All the bases in PD have been flipped for easier comparison of the final bases, and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
 \end{figure}
@@
 
-An animation is another type of display to show how the search
-progresses from start to finish in the space. Figure
-\ref{fig:toy-pca-animated} shows the animated version (six frames from
-the animation if viewed in pdf) of the PCA plot in Figure
-\ref{fig:toy-pca}. An additional piece of information one can learn from
-this animation is that CRS finds its end basis quicker than PD since CRS
-finishes its search in the 5th frame while PD is still making more
-progress.
+Animation is another type of display to show how the search progresses
+from start to finish in the space. Figure \ref{fig:toy-pca-animated}
+shows the animated version (six frames from the animation if viewed in
+pdf) of the PCA plot in Figure \ref{fig:toy-pca}. An additional piece of
+information one can learn from this animation is that CRS finds its end
+basis quicker than PD since CRS finishes its search in the 5th frame
+while PD is still making more progress.
 
@@
 from CRS and PD are plotted on top of the torus and coloured in green
-and brown respectively to match the previous plots. The final basis
-found by PD and CRS are shown in a larger shape and printed below
+and brown, respectively, to match the previous plots. The final basis
+found by PD and CRS are shown in a larger shape and printed below,
 respectively:
@@
 
-Both optimisers have found the third variable in the first direction,
-and the second variable in the second direction. Note however the
+Both optimisers have found the third variable in the first direction and
+the second variable in the second direction. Note, however, the
 different orientation of the basis, following from the different sign in
@@
 to the same sign in the first column) and are symmetric when projected
-onto the other (due the sign difference in the second column). In Figure
-\ref{fig:toy-torus}, this can be seen most clearly in frame 5 where the
-two circles are rotated into a line from our view.
+onto the other (due to the different sign in the second column). In
+Figure \ref{fig:toy-torus}, this can be seen most clearly in frame 5
+where the two circles are rotated into a line from our view.
 

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.R

@@ -23,13 +23,13 @@ knitr::read_chunk(
 ## ----tour-path, fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."----
 
 
-## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed while in the later iterations, most sampled bases fail to make an improvement on the index value and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the ferrn package."----
+## ----toy-search, fig.cap="A comparison of the searches by two optimisers: CRS (left) and SA (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using the holes index. Both optimisers reach the final basis with a similar index value, while it takes SA longer to find the final basis. In the earlier iterations, optimisers quickly find a better basis to proceed, while in the later iterations, most sampled bases fail to make an improvement on the index value, and a boxplot is used to summarise the distribution of the index values. There is no better basis found in the last iteration, 9 (left) and 15 (right), before reaching the maximum number of try and hence it is colored grey. The color scale is from the customised botanical palette in the **ferrn** package."----
 
 
-## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."----
[…]

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
copy editing towards the end

@@
 The projection problem is a 2D projection with three variables using the holes index.
-The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
+The grey points are randomly generated 2D projection bases in the space, and it can be observed that these points form a torus.
 The full video of the animation can be found in the html version of the paper.", eval = knitr::is_latex_output()}
@@
 
-Random variables with different distributions have been simulated and the distributional form of each variable is presented in equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variables `x1`, `x8` to `x10` are normal noise with zero mean and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
-All the variables have been scaled to have an overall unit variance before projection pursuit.
+Random variables with different distributions have been simulated, and the distributional form of each variable is presented in equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
+Variables `x1`, `x8` to `x10` are normal noise with zero mean, and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
+All the variables have been scaled to have overall unit variance before projection pursuit.
 The holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
 
-```{=tex}
 \begin{align}
@@
 \end{align}
-```
+
 ## A problem of non-monotonicity {#monotonic}
@@
 For SA, a non-monotonic interpolation is justified since target bases do not necessarily have a higher index value than the current one, while this is not the case for CRS.
-The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))` and one can observe that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
+The original trace plot for a 2D projection problem, optimised by CRS, is shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))`, and one can observe that the non-monotonic interpolation has undermined the optimiser to realise its full potential.
 Hence, an interruption is implemented to stop at the best basis found in the interpolation.
-The right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))` shows the trace plot after implementing the interruption and while the first two interpolations are identical, the basis at time 61 has a higher index value than the target in the third interpolation.
+The right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:interruption}','\\@ref(fig:interruption)'))` shows the trace plot after implementing the interruption, and while the first two interpolations are identical, the basis at time 61 has a higher index value than the target in the third interpolation.
 Rather than starting the next iteration from the target basis on time 65, CRS starts the next iteration at time 61 on the right panel and reaches a better final basis.
@@
 ```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS.
-On the left panel, basis with a higher index value is found during the interpolation but not used.
+On the left panel, the basis with a higher index value is found during the interpolation but not used.
 On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel."}
@@
 2)  the relative change of the index value needs to be larger than 1e-5; and
-3)  the alpha parameter needs to be larger than 0.01
+3)  the alpha parameter needs to be larger than 0.01.
 
 Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:polish}','\\@ref(fig:polish)'))` presents the projected data and trace plot of a 2D projection, optimised by CRS and followed by the polish step.
-The top row shows the initial projection, the final projection after CRS, and the final projection after polish, respectively The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
+The top row shows the initial projection, the final projection after CRS, and the final projection after polish, respectively.
+The end basis found by CRS reveals the four clusters in the data, but the edges of each cluster are not clean-cut.
 Polish works with this end basis and further pushes the index value to produce clearer edges of the cluster, especially along the vertical axis.
@@
 ```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing for a 2D projection problem on \\code{boa6} data using holes index.
-The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
+The top row shows the initial projected data and the final views after CRS and polish search, and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel), but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
 Note that the parameter \\code{max.tries} is set to 400 in this experiment for CRS to do its best.
@@
 An example of a noisy index function for 1D projections compares the projected data, $\mathbf{Y}_{n \times 1}$, to a randomly generated normal distribution, $\mathcal{N}_{n \times 1}$, using the Kolmogorov test.
-Let $F_{.}(n)$ be the ECDF function (empirical cummulative distribution function), with two possible subscripts $Y$ and $\mathcal{N}$ representing the projected and randomly generated data, and $n$ denoting the number of observation, the Kolmogorov index $I^{nk}(n)$, is defined as:
+Let $F_{.}(n)$ be the empirical cumulative distribution function (ECDF) with two possible subscripts, $Y$ and $\mathcal{N}$, representing the projected and randomly generated data, and $n$ denoting the number of observations, the Kolmogorov index $I^{nk}(n)$, is defined as:
 
@@
 1)  whether any optimiser fails to optimise this non-smooth index; and
-2)  whether the optimisers can find the global optimum despite the presence of a local optimum
+2)  whether the optimisers can find the global optimum despite the presence of a local optimum.
 
-Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:noisy-better-geo}','\\@ref(fig:noisy-better-geo)'))` presents the trace and PCA plots of all three optimisers and as expected, none of the interpolated paths are smooth.
+Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:noisy-better-geo}','\\@ref(fig:noisy-better-geo)'))` presents the trace and PCA plots of all three optimisers, and as expected, none of the interpolated paths are smooth.
 There is barely any improvement made by PD, indicating its failure in optimising non-smooth index functions.
@@
 This long interpolation path is partially due to the fluctuation in the early iterations, where SA tends to generously accept inferior bases before concentrating near the optimum.
-The PCA plot shows the interpolation path and search points excluding the last termination iteration.
+The PCA plot shows the interpolation path and search points, excluding the last termination iteration.
 Pseudo-Derivative (PD) quickly gets stuck near the starting position.
@@
 Combining the insights from both the trace and PCA plot, one can learn the two different search strategies by CRS and SA: CRS frequently samples in the space and only make a move when an improvement is guaranteed to be made, while SA first broadly accepts bases in the space and then starts the extensive sampling in a narrowed subspace.
-The specific decision of which optimiser to use will depend on the index curve in the basis space but if the basis space is non-smooth, accepting inferior bases at first, as SA has done here, can lead to a more efficient search, in terms of the overall number of points evaluated.
+The specific decision of which optimiser to use will depend on the index curve in the basis space, but if the basis space is non-smooth, accepting inferior bases at first, as SA has done here, can lead to a more efficient search in terms of the overall number of points evaluated.
 
@@
 The next experiment compares the performance of CRS and SA when a local maximum exists.
-Two search neighbourhood sizes: 0.5 and 0.7 are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
+Two search neighbourhood sizes, 0.5 and 0.7, are compared to understand how a large search neighbourhood would affect the final basis and the length of the search.
 Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:kol-result}','\\@ref(fig:kol-result)'))` shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the optimiser and search size.
@@
 Comparing CRS and SA for a search size of 0.5, SA does not seem to improve the final basis found, despite having longer interpolation paths.
-However, the denser paths near the local maximum is an indicator that SA is working hard to examine if there is any other optimum in the basis space but the relatively small search size has diminished its ability to reach the global maximum.
-With a larger search size, almost all the seeds (16 out of 20) have found the global maximum and some final bases are much closer to the theoretical best, as compared to the three other cases.
+However, the denser paths near the local maximum are an indicator that SA is working hard to examine if there is any other optimum in the basis space, but the relatively small search size has diminished its ability to reach the global maximum.
+With a larger search size, almost all the seeds (16 out of 20) have found the global maximum, and some final bases are much closer to the theoretical best, as compared to the three other cases.
 This indicates that SA, with a reasonable large search window, is able to overcome the local optimum and optimise close towards the global optimum.
@@
 
-One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))`, the target basis is generated on the other half of the basis space and the interpolator seems to draw a straight line to interpolate.
+One interesting situation observed in the previous examples is that, for some simulations, as shown on the left panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))`, the target basis is generated on the other half of the basis space, and the interpolator seems to draw a straight line to interpolate.
 Bases with opposite signs do not affect the projection and index value, but we would prefer the target to have the same orientation as the current basis.
-The orientation of two bases can be computationally checked by calculating the determinant -- a negative value suggests the two bases having different orientation.
-For 1D bases, this can be corrected by flipping the sign in one basis.
-For higher dimensions, it can be a bit more difficult because the orthonormality of the basis needs to be maintained also when an individual vector is flipped.
-Here an orientation check is carried out once a new target basis is generated and the sign in the target basis will be flipped if a negative determinant is obtained.
-The interpolation after implementing the orientation check is shown on the right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))` where the unsatisfactory interpolation no longer exists.
+The orientation of two bases can be computationally checked by calculating the determinant -- a negative value suggests the two bases have a different orientation.
+For 1D bases, this can be corrected by flipping the sign on one basis.
+For higher dimensions, it can be a bit more difficult because the orthonormality of the basis needs to be also maintained when an individual vector is flipped.
+Here, an orientation check is carried out once a new target basis is generated, and the sign in the target basis will be flipped if a negative determinant is obtained.
+The interpolation after implementing the orientation check is shown on the right panel of Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:flip-sign}','\\@ref(fig:flip-sign)'))`, where the unsatisfactory interpolation no longer exists.
 
@@
 Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
-The dots represent the target basis in each iteration and the path shows the interpolation.
-On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
-The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears."}
+The dots represent the target basis in each iteration, and the path shows the interpolation.
+On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space), and the interpolator crosses the origin to perform the interpolation.
+The right panel shows the same interpolation after implementing an orientation check, and the undesirable interpolation disappears."}
 ```
@@
 
-    -   `get_anchor()` extracts target observations
-    -   `get_basis_matrix()` flattens all the bases into a matrix
-    -   `get_best()` extracts the observation with the highest index value in the data object
-    -   `get_dir_search()` extracts directional search observations for PD search
-    -   `get_interp()` extracts interpolated observations
-    -   `get_interp_last()` extracts the ending interpolated observations in each iteration
-    -   `get_interrupt()` extracts the ending interpolated observations and the target observations if the interpolation is interrupted
-    -   `get_search()` extracts search observations
-    -   `get_search_count()` extracts the count of search observations
-    -   `get_space_param()` produces the coordinates of the centre and radius of the basis space
-    -   `get_start()` extracts the starting observation
-    -   `get_theo()` extracts the theoretical best observations, if given
+    -   `get_anchor()` extracts target observations.
+    -   `get_basis_matrix()` flattens all the bases into a matrix.
+    -   `get_best()` extracts the observation with the highest index value in the data object.
+    -   `get_dir_search()` extracts directional search observations for PD search.
+    -   `get_interp()` extracts interpolated observations.
+    -   `get_interp_last()` extracts the ending interpolated observations in each iteration.
+    -   `get_interrupt()` extracts the ending interpolated observations and the target observations if the interpolation is .interrupted
+    -   `get_search()` extracts search observations.
+    -   `get_search_count()` extracts the count of search observations.
+    -   `get_space_param()` produces the coordinates of the centre and radius of the basis space.
+    -   `get_start()` extracts the starting observation.
+    -   `get_theo()` extracts the theoretical best observations, if given.
 
@@
 
-    -   `bind_theoretical()` binds the theoretical best observation in simulated experiment
-    -   `bind_random()` binds randomly generated bases in the projection bases space to the data object
-    -   `bind_random_matrix()` binds randomly generated bases and outputs in a matrix format
+    -   `bind_theoretical()` binds the theoretical best observation in simulated experiment.
+    -   `bind_random()` binds randomly generated bases in the projection bases space to the data object.
+    -   `bind_random_matrix()` binds randomly generated bases and outputs in a matrix format.
 
@@
 
-    -   `add_anchor()` for plotting anchor bases
-    -   `add_anno()` for annotating the symmetry of start bases
-    -   `add_dir_search()` for plotting the directional search bases with magnified distance
-    -   `add_end()` for plotting end bases
-    -   `add_interp()` for plotting the interpolation path
-    -   `add_interp_last()` for plotting the last interpolation bases for comparing with target bases when interruption is used
-    -   `add_interrupt()` for linking the last interpolation bases with target ones when interruption is used
-    -   `add_search()` for plotting search bases
-    -   `add_space()` for plotting the circular space
-    -   `add_start()` for plotting start bases
-    -   `add_theo()` for plotting theoretical best bases, if applicable
+    -   `add_anchor()` for plotting anchor bases.
+    -   `add_anno()` for annotating the symmetry of start bases.
+    -   `add_dir_search()` for plotting the directional search bases with magnified distance.
+    -   `add_end()` for plotting end bases.
+    -   `add_interp()` for plotting the interpolation path.
+    -   `add_interp_last()` for plotting the last interpolation bases for comparing with target bases when interruption is used.
+    -   `add_interrupt()` for linking the last interpolation bases with target ones when interruption is used.
+    -   `add_search()` for plotting search bases.
+    -   `add_space()` for plotting the circular space.
+    -   `add_start()` for plotting start bases.
+    -   `add_theo()` for plotting theoretical best bases, if applicable.
 
@@
 
-    -   `theme_fern()` and `format_label()` for better display of the grid lines and axis formatting
-    -   `clean_method()` to clean up the name of the optimisers
+    -   `theme_fern()` and `format_label()` for better display of the grid lines and axis formatting.
+    -   `clean_method()` to clean up the name of the optimisers.
     -   `botanical_palettes()` is a collection of colour palettes from Australian native plants.
-Quantitative palettes include daisy, banksia, and cherry and sequential palettes contain fern and acacia
-    -   `botanical_pal()` as the colour interpolator
-    -   `scale_color_*()` and `scale_fill_*()` for scaling the colour and fill of the plot
+Quantitative palettes include daisy, banksia, and cherry, and sequential palettes contain fern and acacia.
+    -   `botanical_pal()` as the colour interpolator.
+    -   `scale_color_*()` and `scale_fill_*()` for scaling the colour and fill of the plot.
 
@@
 
-This paper has provided several visual diagnostics that can be used for understanding a complex optimisation procedure and are implemented in the ferrn package.
+This paper has provided several visual diagnostics that can be used for understanding a complex optimisation procedure and are implemented in the **ferrn** package.
 The methods were illustrated using the optimisers available for projection pursuit guided tour.
@@
 
-The progressive optimisation of a target function and it's coverage of the search space can be viewed in both reduced 2D space and the full space.
+The progressive optimisation of a target function and its coverage of the search space can be viewed in both reduced 2D space and the full space.
 These visualisations can lead to insights for evaluating and comparing the performance of multiple optimisers operating on the same task.
-They can provide better understanding of existing methods or motivate the development of new approaches.
-For example, we have compared how three optimisers perform when maximising a non-smooth index function and illustrated how the pseudo-derivative search fails in this setting.
-The observations from our experiments have also been translated into improved optimisation methods for the guided tour, e.g.
-we introduced the option to interrupt the search if a better basis is found along the path.
+They can provide a better understanding of existing methods or motivate the development of new approaches.
+For example, we have compared how three optimisers perform when maximising a non-smooth index function and have illustrated how the pseudo-derivative search fails in this setting.
+The observations from our experiments have also been translated into improved optimisation methods for the guided tour, e.g., we introduced the option to interrupt the search if a better basis is found along the path.
 
@@
 This article is created using \CRANpkg{knitr} [@knitr] and \CRANpkg{rmarkdown} [@rmarkdown] in R.
-The source code for reproducing this paper can be found at: <https://github.com/huizezhang-sherry/paper-tour-vis>.
-
-\newpage
+The source code for reproducing this paper can be found at: <https://github.com/huizezhang-sherry/paper-ferrn>.

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
copy editing towards the end

@@
 The projection problem is a 2D projection with three variables using the holes index.
-The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.
+The grey points are randomly generated 2D projection bases in the space, and it can be observed that these points form a torus.
 The full video of the animation can be found in the html version of the paper.}\label{fig:toy-torus}
@@
 
-Random variables with different distributions have been simulated and
+Random variables with different distributions have been simulated, and
 the distributional form of each variable is presented in equations
 \ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1}, \texttt{x8}
-to \texttt{x10} are normal noise with zero mean and unit variance and
+to \texttt{x10} are normal noise with zero mean, and unit variance and
 \texttt{x2} to \texttt{x7} are normal mixtures with varied weights and
-locations. All the variables have been scaled to have an overall unit
+locations. All the variables have been scaled to have overall unit
 variance before projection pursuit. The holes index
@@
 for a 2D projection problem, optimised by CRS, is shown on the left
-panel of Figure \ref{fig:interruption} and one can observe that the
+panel of Figure \ref{fig:interruption}, and one can observe that the
 non-monotonic interpolation has undermined the optimiser to realise its
@@
 \ref{fig:interruption} shows the trace plot after implementing the
-interruption and while the first two interpolations are identical, the
+interruption, and while the first two interpolations are identical, the
 basis at time 61 has a higher index value than the target in the third
@@
 \caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by CRS.
-On the left panel, basis with a higher index value is found during the interpolation but not used.
+On the left panel, the basis with a higher index value is found during the interpolation but not used.
 On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel.}\label{fig:interruption}
@@
 \item
-  the alpha parameter needs to be larger than 0.01
+  the alpha parameter needs to be larger than 0.01.
 \end{enumerate}
@@
 row shows the initial projection, the final projection after CRS, and
-the final projection after polish, respectively The end basis found by
+the final projection after polish, respectively. The end basis found by
 CRS reveals the four clusters in the data, but the edges of each cluster
@@
 \caption{Comparison of the projected data before and after using polishing for a 2D projection problem on \code{boa6} data using holes index.
-The top row shows the initial projected data and the final views after CRS and polish search and the second row traces the index value.
-The clustering structure in the data is detected by CRS (top middle panel) but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
+The top row shows the initial projected data and the final views after CRS and polish search, and the second row traces the index value.
+The clustering structure in the data is detected by CRS (top middle panel), but the polish step improves the index value and produces clearer boundaries of the clusters (top right panel), especially along the vertical axis.
 Note that the parameter \code{max.tries} is set to 400 in this experiment for CRS to do its best.
@@
 generated normal distribution, \(\mathcal{N}_{n \times 1}\), using the
-Kolmogorov test. Let \(F_{.}(n)\) be the ECDF function (empirical
-cummulative distribution function), with two possible subscripts \(Y\)
-and \(\mathcal{N}\) representing the projected and randomly generated
-data, and \(n\) denoting the number of observation, the Kolmogorov index
+Kolmogorov test. Let \(F_{.}(n)\) be the empirical cumulative
+distribution function (ECDF) with two possible subscripts, \(Y\) and
+\(\mathcal{N}\), representing the projected and randomly generated data,
+and \(n\) denoting the number of observations, the Kolmogorov index
 \(I^{nk}(n)\), is defined as:
@@
   whether the optimisers can find the global optimum despite the
-  presence of a local optimum
+  presence of a local optimum.
 \end{enumerate}
@@
 Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
-all three optimisers and as expected, none of the interpolated paths are
-smooth. There is barely any improvement made by PD, indicating its
+all three optimisers, and as expected, none of the interpolated paths
+are smooth. There is barely any improvement made by PD, indicating its
 failure in optimising non-smooth index functions. While CRS and SA have
@@
 before concentrating near the optimum. The PCA plot shows the
-interpolation path and search points excluding the last termination
+interpolation path and search points, excluding the last termination
 iteration. Pseudo-Derivative (PD) quickly gets stuck near the starting
@@
 extensive sampling in a narrowed subspace. The specific decision of
-which optimiser to use will depend on the index curve in the basis space
-but if the basis space is non-smooth, accepting inferior bases at first,
-as SA has done here, can lead to a more efficient search, in terms of
-the overall number of points evaluated.
+which optimiser to use will depend on the index curve in the basis
+space, but if the basis space is non-smooth, accepting inferior bases at
+first, as SA has done here, can lead to a more efficient search in terms
+of the overall number of points evaluated.
 
@@
 The next experiment compares the performance of CRS and SA when a local
-maximum exists. Two search neighbourhood sizes: 0.5 and 0.7 are compared
-to understand how a large search neighbourhood would affect the final
-basis and the length of the search. Figure \ref{fig:kol-result} shows 80
-paths simulated using 20 seeds in the PCA plot, faceted by the optimiser
-and search size. With CRS and a search size of 0.5, despite being the
-simplest and fastest, the optimiser fails in three instances where the
-final basis lands neither near the local nor the global optimum. With a
-larger search size of 0.7, more seeds have found the global maximum.
-Comparing CRS and SA for a search size of 0.5, SA does not seem to
-improve the final basis found, despite having longer interpolation
-paths. However, the denser paths near the local maximum is an indicator
-that SA is working hard to examine if there is any other optimum in the
-basis space but the relatively small search size has diminished its
-ability to reach the global maximum. With a larger search size, almost
-all the seeds (16 out of 20) have found the global maximum and some
-final bases are much closer to the theoretical best, as compared to the
-three other cases. This indicates that SA, with a reasonable large
-search window, is able to overcome the local optimum and optimise close
-towards the global optimum.
+maximum exists. Two search neighbourhood sizes, 0.5 and 0.7, are
+compared to understand how a large search neighbourhood would affect the
+final basis and the length of the search. Figure \ref{fig:kol-result}
+shows 80 paths simulated using 20 seeds in the PCA plot, faceted by the
+optimiser and search size. With CRS and a search size of 0.5, despite
+being the simplest and fastest, the optimiser fails in three instances
+where the final basis lands neither near the local nor the global
+optimum. With a larger search size of 0.7, more seeds have found the
+global maximum. Comparing CRS and SA for a search size of 0.5, SA does
+not seem to improve the final basis found, despite having longer
+interpolation paths. However, the denser paths near the local maximum
+are an indicator that SA is working hard to examine if there is any
+other optimum in the basis space, but the relatively small search size
+has diminished its ability to reach the global maximum. With a larger
+search size, almost all the seeds (16 out of 20) have found the global
+maximum, and some final bases are much closer to the theoretical best,
+as compared to the three other cases. This indicates that SA, with a
+reasonable large search window, is able to overcome the local optimum
+and optimise close towards the global optimum.
 
@@
 \ref{fig:flip-sign}, the target basis is generated on the other half of
-the basis space and the interpolator seems to draw a straight line to
+the basis space, and the interpolator seems to draw a straight line to
 interpolate. Bases with opposite signs do not affect the projection and
@@
 computationally checked by calculating the determinant -- a negative
-value suggests the two bases having different orientation. For 1D bases,
-this can be corrected by flipping the sign in one basis. For higher
+value suggests the two bases have a different orientation. For 1D bases,
+this can be corrected by flipping the sign on one basis. For higher
 dimensions, it can be a bit more difficult because the orthonormality of
-the basis needs to be maintained also when an individual vector is
-flipped. Here an orientation check is carried out once a new target
-basis is generated and the sign in the target basis will be flipped if a
-negative determinant is obtained. The interpolation after implementing
+the basis needs to be also maintained when an individual vector is
+flipped. Here, an orientation check is carried out once a new target
+basis is generated, and the sign in the target basis will be flipped if
+a negative determinant is obtained. The interpolation after implementing
 the orientation check is shown on the right panel of Figure
-\ref{fig:flip-sign} where the unsatisfactory interpolation no longer
+\ref{fig:flip-sign}, where the unsatisfactory interpolation no longer
 exists.
@@
 Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using CRS with seed 2463.
-The dots represent the target basis in each iteration and the path shows the interpolation.
-On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation.
-The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
+The dots represent the target basis in each iteration, and the path shows the interpolation.
+On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space), and the interpolator crosses the origin to perform the interpolation.
+The right panel shows the same interpolation after implementing an orientation check, and the undesirable interpolation disappears.}\label{fig:flip-sign}
 \end{figure}
@@
   \item
-    \texttt{get\_anchor()} extracts target observations
-  \item
-    \texttt{get\_basis\_matrix()} flattens all the bases into a matrix
+    \texttt{get\_anchor()} extracts target observations.
+  \item
+    \texttt{get\_basis\_matrix()} flattens all the bases into a matrix.
   \item
     \texttt{get\_best()} extracts the observation with the highest index
-    value in the data object
+    value in the data object.
   \item
     \texttt{get\_dir\_search()} extracts directional search observations
-    for PD search
-  \item
-    \texttt{get\_interp()} extracts interpolated observations
+    for PD search.
+  \item
+    \texttt{get\_interp()} extracts interpolated observations.
   \item
     \texttt{get\_interp\_last()} extracts the ending interpolated
-    observations in each iteration
+    observations in each iteration.
   \item
@@
     observations and the target observations if the interpolation is
-    interrupted
-  \item
-    \texttt{get\_search()} extracts search observations
+    .interrupted
+  \item
+    \texttt{get\_search()} extracts search observations.
   \item
     \texttt{get\_search\_count()} extracts the count of search
-    observations
+    observations.
   \item
     \texttt{get\_space\_param()} produces the coordinates of the centre
-    and radius of the basis space
-  \item
-    \texttt{get\_start()} extracts the starting observation
+    and radius of the basis space.
+  \item
+    \texttt{get\_start()} extracts the starting observation.
   \item
     \texttt{get\_theo()} extracts the theoretical best observations, if
-    given
+    given.
   \end{itemize}
@@
     \texttt{bind\_theoretical()} binds the theoretical best observation
-    in simulated experiment
+    in simulated experiment.
   \item
     \texttt{bind\_random()} binds randomly generated bases in the
-    projection bases space to the data object
+    projection bases space to the data object.
   \item
     \texttt{bind\_random\_matrix()} binds randomly generated bases and
-    outputs in a matrix format
+    outputs in a matrix format.
   \end{itemize}
@@
   \item
-    \texttt{add\_anchor()} for plotting anchor bases
-  \item
-    \texttt{add\_anno()} for annotating the symmetry of start bases
+    \texttt{add\_anchor()} for plotting anchor bases.
+  \item
+    \texttt{add\_anno()} for annotating the symmetry of start bases.
   \item
     \texttt{add\_dir\_search()} for plotting the directional search
-    bases with magnified distance
-  \item
-    \texttt{add\_end()} for plotting end bases
-  \item
-    \texttt{add\_interp()} for plotting the interpolation path
+    bases with magnified distance.
+  \item
+    \texttt{add\_end()} for plotting end bases.
+  \item
+    \texttt{add\_interp()} for plotting the interpolation path.
   \item
     \texttt{add\_interp\_last()} for plotting the last interpolation
-    bases for comparing with target bases when interruption is used
+    bases for comparing with target bases when interruption is used.
   \item
     \texttt{add\_interrupt()} for linking the last interpolation bases
-    with target ones when interruption is used
-  \item
-    \texttt{add\_search()} for plotting search bases
-  \item
-    \texttt{add\_space()} for plotting the circular space
-  \item
-    \texttt{add\_start()} for plotting start bases
+    with target ones when interruption is used.
+  \item
+    \texttt{add\_search()} for plotting search bases.
+  \item
+    \texttt{add\_space()} for plotting the circular space.
+  \item
+    \texttt{add\_start()} for plotting start bases.
   \item
     \texttt{add\_theo()} for plotting theoretical best bases, if
-    applicable
+    applicable.
   \end{itemize}
@@
     \texttt{theme\_fern()} and \texttt{format\_label()} for better
-    display of the grid lines and axis formatting
-  \item
-    \texttt{clean\_method()} to clean up the name of the optimisers
+    display of the grid lines and axis formatting.
+  \item
+    \texttt{clean\_method()} to clean up the name of the optimisers.
   \item
@@
     from Australian native plants. Quantitative palettes include daisy,
-    banksia, and cherry and sequential palettes contain fern and acacia
-  \item
-    \texttt{botanical\_pal()} as the colour interpolator
+    banksia, and cherry, and sequential palettes contain fern and
+    acacia.
+  \item
+    \texttt{botanical\_pal()} as the colour interpolator.
   \item
     \texttt{scale\_color\_*()} and \texttt{scale\_fill\_*()} for scaling
-    the colour and fill of the plot
+    the colour and fill of the plot.
   \end{itemize}
@@
 understanding a complex optimisation procedure and are implemented in
-the ferrn package. The methods were illustrated using the optimisers
-available for projection pursuit guided tour. Here the constraint is the
-orthonormality condition of the projection bases, which corresponds to
-optimisation over spheres and torii. The approach described broadly
-applies to other constrained optimisers. Although the manifold in
-\(p\)-space might be different the diagnostic techniques are the same. A
-researcher would begin by saving the path of the optimiser in a form
-required to input into the ferrn package, as described in this paper.
-One might generally make more samples from the constrained space upon
-which to assess and compare the optimisation paths. These
+the \textbf{ferrn} package. The methods were illustrated using the
+optimisers available for projection pursuit guided tour. Here the
+constraint is the orthonormality condition of the projection bases,
+which corresponds to optimisation over spheres and torii. The approach
+described broadly applies to other constrained optimisers. Although the
+manifold in \(p\)-space might be different the diagnostic techniques are
+the same. A researcher would begin by saving the path of the optimiser
+in a form required to input into the ferrn package, as described in this
+paper. One might generally make more samples from the constrained space
+upon which to assess and compare the optimisation paths. These
 high-dimensional data objects can then be viewed using the tour.
 
-The progressive optimisation of a target function and it's coverage of
+The progressive optimisation of a target function and its coverage of
 the search space can be viewed in both reduced 2D space and the full
@@
 comparing the performance of multiple optimisers operating on the same
-task. They can provide better understanding of existing methods or
+task. They can provide a better understanding of existing methods or
 motivate the development of new approaches. For example, we have
 compared how three optimisers perform when maximising a non-smooth index
-function and illustrated how the pseudo-derivative search fails in this
-setting. The observations from our experiments have also been translated
-into improved optimisation methods for the guided tour, e.g.~we
-introduced the option to interrupt the search if a better basis is found
-along the path.
+function and have illustrated how the pseudo-derivative search fails in
+this setting. The observations from our experiments have also been
+translated into improved optimisation methods for the guided tour, e.g.,
+we introduced the option to interrupt the search if a better basis is
+found along the path.
 
@@
 reproducing this paper can be found at:
-\url{https://github.com/huizezhang-sherry/paper-tour-vis}.
-
-\newpage
+\url{https://github.com/huizezhang-sherry/paper-ferrn}.
 

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.R

@@ -54,17 +54,17 @@ holes_2d_geo_3var %>% get_best() %>% pull(basis) %>% .[[1]]
 holes_2d_better_3var %>% get_best() %>% pull(basis) %>% .[[1]]
 
 
-## ----toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.", eval = knitr::is_latex_output()----
+## ----toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space, and it can be observed that these points form a torus. The full video of the animation can be found in the html version of the paper.", eval = knitr::is_latex_output()----
 
 
 ## ----toy-torus-interactive, out.extra="class = 'tour-gif'", fig.cap="Animation of rotating the 2D basis space along with two search paths optimised by PD (brown) and CRS (green). The projection problem is a 2D projection with three variables using the holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus.", eval = knitr::is_html_output()----
 #> NA
 
 
-## ----interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS. On the left panel, basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel."----
+## ----interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by CRS. On the left panel, the basis with a higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with a higher index value, as shown on the right panel."----
 
 
[…]

### 2022-01-04 cover-letter/cover-letter.Rmd
some clean up

@@
 opening: Dear Editor
-closing: Regards
+closing: Regards,
 spacing: 1.4
-output: MonashEBSTemplates::letter
+output: monash::letter
 ---

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
some clean up

@@
 
-```{=tex}
 \begin{align}
@@
 \end{align}
-```
+
 \noindent where $f$ and $\mathbf{X}$ are defined as in the previous section, $\mathcal{A}$ is the set of all $p$-dimensional projection bases, $I_d$ is the $d$-dimensional identity matrix, and the constraint ensures the projection bases, $\mathbf{A}$, to be orthonormal.
@@
 
-Random variables with different distributions have been simulated, and the distributional form of each variable is presented in equations \ref{eq:sim-norm} to \ref{eq:sim-x7}.
-Variables `x1`, `x8` to `x10` are normal noise with zero mean, and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
-All the variables have been scaled to have overall unit variance before projection pursuit.
-The holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
+Random variables with different distributions have been simulated as follows:
 
 \begin{align}
-x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \label{eq:sim-norm} \\
-x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\label{eq:sim-x2}\\
+x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \\
+x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\\
 \Pr(x_3) &= 
@@
 0 & \text{otherwise}
-\end{cases}\label{eq:sim-x3}\\
-x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1) \label{eq:sim-x4}\\
-x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\label{eq:sim-x5}\\
-x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\label{eq:sim-x6}\\
+\end{cases}\\
+x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1)\\
+x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\\
+x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\\
 x_7 &\sim 0.5 \mathcal{N}(-5, 1) + 0.5 \mathcal{N}(5, 1) 
-\label{eq:sim-x7}
 \end{align}
+
+\noindent Variables `x1`, `x8` to `x10` are normal noise with zero mean, and unit variance and `x2` to `x7` are normal mixtures with varied weights and locations.
+All the variables have been scaled to have overall unit variance before projection pursuit.
+The holes index [@cook2008grand], used for detecting bimodality of the variables, is used throughout the examples unless otherwise specified.
 

### 2022-01-04 paper/zhang-cook-laa-langrene-menendez.tex
some clean up

@@
 
-Random variables with different distributions have been simulated, and
-the distributional form of each variable is presented in equations
-\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variables \texttt{x1}, \texttt{x8}
-to \texttt{x10} are normal noise with zero mean, and unit variance and
-\texttt{x2} to \texttt{x7} are normal mixtures with varied weights and
-locations. All the variables have been scaled to have overall unit
-variance before projection pursuit. The holes index
-\citep{cook2008grand}, used for detecting bimodality of the variables,
-is used throughout the examples unless otherwise specified.
+Random variables with different distributions have been simulated as
+follows:
 
 \begin{align}
-x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \label{eq:sim-norm} \\
-x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\label{eq:sim-x2}\\
+x_1 \overset{d}{=} x_8 \overset{d}{=} x_9 \overset{d}{=} x_{10}& \sim \mathcal{N}(0, 1) \\
+x_2 &\sim 0.5 \mathcal{N}(-3, 1) + 0.5 \mathcal{N}(3, 1)\\
 \Pr(x_3) &= 
@@
 0 & \text{otherwise}
-\end{cases}\label{eq:sim-x3}\\
-x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1) \label{eq:sim-x4}\\
-x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\label{eq:sim-x5}\\
-x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\label{eq:sim-x6}\\
+\end{cases}\\
+x_4 &\sim 0.25 \mathcal{N}(-3, 1) + 0.75 \mathcal{N}(3, 1)\\
+x_5 &\sim \frac{1}{3} \mathcal{N}(-5, 1) + \frac{1}{3} \mathcal{N}(0, 1) + \frac{1}{3} \mathcal{N}(5, 1)\\
+x_6 &\sim 0.45 \mathcal{N}(-5, 1) + 0.1 \mathcal{N}(0, 1) + 0.45 \mathcal{N}(5, 1)\\
 x_7 &\sim 0.5 \mathcal{N}(-5, 1) + 0.5 \mathcal{N}(5, 1) 
-\label{eq:sim-x7}
 \end{align}
+
+Variables \texttt{x1}, \texttt{x8} to \texttt{x10} are normal noise with
+zero mean, and unit variance and \texttt{x2} to \texttt{x7} are normal
+mixtures with varied weights and locations. All the variables have been
+scaled to have overall unit variance before projection pursuit. The
+holes index \citep{cook2008grand}, used for detecting bimodality of the
+variables, is used throughout the examples unless otherwise specified.
 

### 2022-01-05 paper/zhang-cook-laa-langrene-menendez.tex
final clean up

@@
 
-Variables \texttt{x1}, \texttt{x8} to \texttt{x10} are normal noise with
-zero mean, and unit variance and \texttt{x2} to \texttt{x7} are normal
-mixtures with varied weights and locations. All the variables have been
-scaled to have overall unit variance before projection pursuit. The
-holes index \citep{cook2008grand}, used for detecting bimodality of the
-variables, is used throughout the examples unless otherwise specified.
+\noindent Variables \texttt{x1}, \texttt{x8} to \texttt{x10} are normal
+noise with zero mean, and unit variance and \texttt{x2} to \texttt{x7}
+are normal mixtures with varied weights and locations. All the variables
+have been scaled to have overall unit variance before projection
+pursuit. The holes index \citep{cook2008grand}, used for detecting
+bimodality of the variables, is used throughout the examples unless
+otherwise specified.
 

### 2022-01-05 zhang-cook-laa-langrene-menendez.tex
move paper to the top level directory

--- 2022-01-05 paper/zhang-cook-laa-langrene-menendez.tex
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/tour-path}
+{\centering \includegraphics[width=1\linewidth]{figs/tour-path}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-search}
+{\centering \includegraphics[width=1\linewidth]{figs/toy-search}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-interp}
+{\centering \includegraphics[width=1\linewidth]{figs/toy-interp}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/toy-pca}
+{\centering \includegraphics[width=0.7\linewidth]{figs/toy-pca}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/interrupt}
+{\centering \includegraphics[width=1\linewidth]{figs/interrupt}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/polish}
+{\centering \includegraphics[width=1\linewidth]{figs/polish}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/noisy-better-geo}
+{\centering \includegraphics[width=1\linewidth]{figs/noisy-better-geo}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/kol-result}
+{\centering \includegraphics[width=1\linewidth]{figs/kol-result}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-ferrn/figs/flip-sign}
+{\centering \includegraphics[width=1\linewidth]{figs/flip-sign}
 

### 2022-01-05 script.R

@@ -7,7 +7,7 @@ files <- paste0("data/", list.files(here::here("data")))
 purrr::walk(.x = files, ~load(here::here(.x), env = globalenv()))
 
 ## ----tour-path
-knitr::include_graphics(here::here("figs/tour-path.png"))
+knitr::include_graphics("figs/tour-path.png")
 
 ## ---- toy-search
 # p1 <- holes_2d_better_max_tries %>%
@@ -26,7 +26,7 @@ knitr::include_graphics(here::here("figs/tour-path.png"))
 #        path = here::here("figs"), device = "svg",
 #        width = 10, height = 5)
 
-knitr::include_graphics(here::here("figs/toy-search.svg"))
+knitr::include_graphics("figs/toy-search.svg")
 
 ## ---- toy-interp
 # p1 <- holes_2d_better_max_tries %>%
@@ -45,7 +45,7 @@ knitr::include_graphics(here::here("figs/toy-search.svg"))
 #        path = here::here("figs"), device = "svg",
 #        width = 10, height = 5)
 
-knitr::include_graphics(here::here("figs/toy-interp.svg"))
+knitr::include_graphics("figs/toy-interp.svg")
 
 ## ----toy-pca
 # p <- bind_rows(holes_1d_geo, holes_1d_better) %>%
@@ -58,7 +58,7 @@ knitr::include_graphics(here::here("figs/toy-interp.svg"))
 #
 # ggsave(p, filename = "toy-pca.svg",
 #        path = here::here("figs"))
-knitr::include_graphics(here::here("figs", "toy-pca.svg"))
+knitr::include_graphics("figs/toy-pca.svg")
 
 
 ## ----toy-pca-animated
@@ -89,7 +89,7 @@ wrap_plots(gl)
 # anim_save(ani,
 #           filename = here::here("anim/toy-pca-animated.gif"),
 #           renderer = gifski_renderer(width = 500, height = 500))
-knitr::include_graphics(here::here("anim/toy-pca-animated.gif"))
+knitr::include_graphics("anim/toy-pca-animated.gif")
 
 
 ## ----toy-tour
@@ -132,7 +132,7 @@ wrap_plots(gl)
 #   frames = 100,
 #   gif_file = here::here("anim","tour.gif")
 # )
-knitr::include_graphics(here::here("anim/tour.gif"))
+knitr::include_graphics("anim/tour.gif")
 
 ## ----toy-torus
 # generate 2D paths
@@ -242,7 +242,7 @@ wrap_plots(gl)
[…]

### 2022-01-05 zhang-cook-laa-langrene-menendez.R

@@ -13,7 +13,7 @@ knitr::opts_chunk$set(echo = FALSE,
 
 ## ----external, include = FALSE, cache = FALSE---------------------------------
 knitr::read_chunk(
-  here::here("paper", "script.R")
+  here::here("script.R")
 )
 
 

### 2022-01-13 zhang-cook-laa-langrene-menendez.Rmd
move paper to the top level directory
fix PCA plot colour and caption

--- 2022-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
@@
 knitr::read_chunk(
-  here::here("paper", "script.R")
+  here::here("script.R")
 )
@@
 
-```{r toy-pca, out.width = "70%", fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
+```{r toy-pca, out.width = "70%", fig.cap = "Search paths of CRS (brown) and PD (green) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 With animation, the progression of the search paths from start to finish is better identified.
-CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
+CRS (brown) finishes the optimisation quicker than PD (green) since there is no further movement for CRS in the sixth frame.
 The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()}
@@
 ```{r toy-pca-animated-interactive, out.extra="class = 'gif'", fig.cap = "Animated version of Figure 4.
-With animation, the progression of the search paths from start to finish can better identified and CRS (green) finishes the optimisation quicker than PD (brown).", eval=knitr::is_html_output()}
+With animation, the progression of the search paths from start to finish can better identified and CRS (brown) finishes the optimisation quicker than PD (green).", eval=knitr::is_html_output()}
 ```

### 2022-01-13 zhang-cook-laa-langrene-menendez.tex
fix PCA plot colour and caption

@@
 
-\caption[Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
+\caption[Search paths of CRS (brown) and PD (green) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of CRS (brown) and PD (green) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index.
 The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA.
@@
 With animation, the progression of the search paths from start to finish is better identified.
-CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame.
+CRS (brown) finishes the optimisation quicker than PD (green) since there is no further movement for CRS in the sixth frame.
 The full video of the animation can be found in the html version of the paper.}\label{fig:toy-pca-animated}

### 2022-01-13 script.R

@@ -69,9 +69,9 @@ knitr::include_graphics("figs/toy-pca.svg")
 #   explore_space_pca(group = method, animate = TRUE, interp_size = 3,
 #                     theo_size = 45, start_size = 10, end_size = 20) +
 #   theme(legend.position = "none") +
-#   scale_color_continuous_botanical(palette = "fern")
-# animate(ani, nframes = 100, device = "png",
-#         renderer = file_renderer("anim/pca/",
+#   scale_color_discrete_botanical(palette = "fern")
+# gganimate::animate(ani, nframes = 100, device = "png",
+#         renderer = gganimate::file_renderer("anim/pca/",
 #                                  prefix = "pca", overwrite = TRUE))
 
 frames <- c("0001", "0038", "0046", "0079", "0086", "0100")

### 2022-01-13 zhang-cook-laa-langrene-menendez.R

@@ -29,13 +29,13 @@ knitr::read_chunk(
 ## ----toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: CRS (left) and SA (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation, and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth, while SA is observed to first pass by some bases with higher index values before reaching the target bases in time 76-130."----
 
 
-## ----toy-pca, out.width = "70%", fig.cap = "Search paths of CRS (green) and PD (brown) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers are aiming to find. All the bases in PD have been flipped for easier comparison of the final bases, and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
+## ----toy-pca, out.width = "70%", fig.cap = "Search paths of CRS (brown) and PD (green) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. The black star represents the theoretical best basis the optimisers are aiming to find. All the bases in PD have been flipped for easier comparison of the final bases, and a grey dashed line has been annotated to indicate the symmetry of the two start bases."----
 
 
-## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (green) finishes the optimisation quicker than PD (brown) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
+## ----toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish is better identified. CRS (brown) finishes the optimisation quicker than PD (green) since there is no further movement for CRS in the sixth frame. The full video of the animation can be found in the html version of the paper.", eval=knitr::is_latex_output()----
 
 
[…]
